### 2017-02-16 paper-LIN-PQ.tex
added Chao's files

%\documentclass[a4paper,UKenglish]{lipics}
\documentclass{llncs}
%\usepackage{mathbbold}

%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{conjecture}[theorem]{Conjecture}
%\newtheorem{corollary}[theorem]{Corollary}
%\newtheorem{proposition}[theorem]{Proposition}
%\newtheorem{lemma}[theorem]{Lemma}
%\newdef{definition}[theorem]{Definition}
%\newdef{remark}[theorem]{Remark}
%\newdef{example}[theorem]{Example}

\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{color}
%\usepackage{txfonts}
%\usepackage{pxfonts}
%\usepackage{mathabx}
%\usepackage{makeidx}  % allows for indexgeneration
\usepackage{verbatim}
\usepackage{url}
%\usepackage{hyperref}
\usepackage{times}
\usepackage{ulem}    %(\emph{....} in the texfile) is underlined instead of italic
\normalem
\usepackage{enumerate}
%\usepackage[square, comma, sort&compress, numbers]{natbib}

%added by Yi Lv%
%\usepackage[linesnumbered,ruled,procnumbered,noend]{algorithm2e}
%\usepackage[linesnumbered,ruled,procnumbered,vlined]{algorithm2e}
\usepackage[linesnumbered,ruled,procnumbered,noend]{algorithm2e}
\usepackage{stmaryrd}
%\usepackage{hyperref}
%added by Wang Chao%
%\usepackage{mathrsfs}
%\usepackage{extarrows}
\usepackage{thmtools}
\usepackage{thm-restate}

%add for TIkz
\usepackage[version=0.96]{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,snakes,automata,backgrounds,petri}
\usepackage[latin1]{inputenc}
\usepackage{placeins}
\usepackage[bookmarks=false]{hyperref}

%\declaretheorem[name=Claim]{clm}
%\declaretheorem[name=Lemma]{lem}
%\declaretheorem[name=Theorem]{thm}

\DeclareSymbolFont{largesymbolsA}{U}{txexa}{m}{n}
\SetSymbolFont{largesymbolsA}{bold}{U}{txexa}{bx}{n}
\DeclareFontSubstitution{U}{txexa}{m}{n}
\DeclareMathSymbol{\bigsqcupplus}{\mathop}{largesymbolsA}{"02}

% "Import" \bigsqcupplus from txfonts without loading txfonts (since
% it changes the default font and replaces many symbols)

% Commands by Catuscia

%\DeclareMathOperator*{\minarg}{{\rm minarg}}

%Previous notation for the probabilistic choice
%\newcommand{\probsum}{\bigoplus}
%\newcommand{\smallprobsum}[1]{\mkern4mu{\textstyle\circ\mkern-15.5mu\sum_{#1}\:}}
%\newcommand{\bigprobsum}[1]{{\;\displaystyle\odot\mkern-21mu\sum_{#1}}}
%\newcommand{\probsum}[1]{\mathchoice{\bigprobsum{#1}}{\smallprobsum{#1}}{\smallprobsum{#1}}{\smallprobsum{#1}}}
%For the uniform distribution
%\newcommand{\usmallprobsum}[1]{\mkern4mu{\textstyle\circ\mkern-15.5mu\sum^\mathcal{U}_{#1}\:}}
%\newcommand{\ubigprobsum}[1]{{\;\displaystyle\odot\mkern-21mu\sum^\mathcal{U}_{#1}}}
%\newcommand{\uprobsum}[1]{\mathchoice{\ubigprobsum{#1}}{\usmallprobsum{#1}}{\usmallprobsum{#1}}{\usmallprobsum{#1}}}

%\newcommand{\nondsum}{\bigbox}
\newcommand{\nondsum}{\bigsqcupplus}
\newcommand{\probplus}[1]{\oplus_{#1}}
%\newcommand{\nondplus}{\square}
\newcommand{\bang}{!\,}
\newcommand{\nondplus}{{\textstyle\bigsqcupplus}}
\newcommand{\partmap}{\rightharpoonup}
\newcommand{\map}{\rightarrow}
%\newcommand{\exc}{\phi}
\newcommand{\exc}{\alpha}
\newcommand{\exec}{\mathit{exec}}
\newcommand{\execp}{\mathit{execp}}
\newcommand{\Act}{\mathit{Act}}
\newcommand{\Sec}{\mathit{Sec}}
\newcommand{\Obs}{\mathit{Obs}}
\newcommand{\etree}{\mathit{etree}}
\newcommand{\lstate}{\mathit{lst}}
\newcommand{\fstate}{\mathit{fst}}
%\newcommand{\STATE}{\mathcal{P}r}  $ original marked
%\newcommand{\STATE}{\mathcal{P}}   $ marked by me
%\newcommand{\st}{P}
\newcommand{\trans}{\mathcal{T}}
\newcommand{\Aut}{\mathcal{M}}
\newcommand{\init}{\mathit{init}}
\newcommand{\perr}{\mathcal{P}}

\newcommand{\calo}{\mathcal{O}}
\newcommand{\cals}{\mathcal{S}}
\newcommand{\sseq}{\vec s}
\newcommand{\oseq}{\vec o}
\newcommand{\ccsp}{CCS$_p$}

\newcommand{\bigfrac}[2]{\frac{\raisebox{1ex}{$#1$}}{\raisebox{-1.5ex}{$#2$}}}
\newcommand{\nondarr}[1]{\overset{#1}{\longrightarrow}}
\newcommand{\Nondarr}[1]{\overset{#1}{\Longrightarrow}}
\newcommand{\vectorArrow}[1]{\stackrel{\longrightarrow}{\mbox{#1}}}
\newcommand{\probarr}[1]{\overset{#1}{\dashrightarrow}}
\newcommand{\paral}{\,|\,}
\newcommand{\outp}[1]{\overline{#1}}
\renewcommand{\Pr}{{\rm Pr}}

%Commands by Chao Wang

%memory models%
\newcommand{\TSO}{\textrm{TSO}}
\newcommand{\PSO}{\textrm{PSO}}

%correctness conditions%
\newcommand{\lin}{\textrm{linearizability}}
\newcommand{\slin}{\textrm{static linearizability}}
\newcommand{\qlin}{\textrm{quasi linearizability}}
\newcommand{\TTlin}{\textrm{TSO-to-TSO linearizability}}

\newcommand{\pair}[2]{\langle #1 , #2 \rangle}% pairs
\newcommand{\setof}[2]{\{ \, #1 \mid #2 \, \}}% Sets
\newcommand{\set}[1]{\{ {#1}  \}  }
%\newcommand{\map}[3]{{#1} \colon {#2} \longmapsto {#3}} %functions
\newcommand{\den}[1]{[\![#1]\!]}% Denotation of
\newcommand{\mean}[1]{|\!|#1|\!|}
\newcommand{\forget}[1]{}

%%%%%%%%%%%%%GENERAL%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\itbox}[1]{{\it #1\/}}
\newcommand{\un}[1]{\uline{#1}}%\underline{#1}}
\newcommand{\ov}[1]{\overline{#1}}
\newcommand{\smallspace}{\vspace{10mm}}
\newcommand{\is}{\mbox{$\Longleftarrow\ $}}
\newcommand{\pright}[1]{\hfill{#1}}
\newcommand{\bnfor}{\;\;\mid\;\;}

\newcommand{\ar}[1]{\stackrel{\scriptstyle #1}{\longrightarrow}}

%%%%%%%%%italics in math mode
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\newcommand{\true}{{\it true}}
%\newcommand{\false}{{\it false}}
\newcommand{\calB}{{\cal B}}
\newcommand{\calF}{{\cal F}}
\newcommand{\calP}{{\cal P}}
\newcommand{\order}{{\cal O}}
\newcommand{\size}[1]{|#1|}

%other notations%
\newcommand{\LTS}{\textit{LTS}}
\newcommand{\bedt}[1]{{\color{blue}#1}}
\newcommand{\redt}[1]{{\color{red}#1}}

%\pagestyle{empty}
%\pagestyle{plain}

%\def\lastname{Xu,Palamidessi}

\title{Reducing Linearizability of Priority Queues and More Data Structures to State Reachability}

\author {Chao Wang \and Ahmed Bouajjani \and Constantin Enea}
\institute{Institut de Recherche en Informatique Fondamentale, \\Univ. Paris Diderot (Paris 7)
}

\begin{document}

\maketitle

\begin{abstract}
Abstract
\end{abstract}

\forget{
\noindent Keywords: weak memory model, $\textit{linearizability}$,
$\textit{TSO-to-TSO linearizability}$
}

\section{Introduction}
\label{sec:introduction}
Introduction

\smallskip

\noindent {\bf Related work}
%\label{sec:related work}
\noindent Related work

\section{Inductive Rules for Priority Queues}
\label{sec:inductive rules for priority queues}

Given a sequence $l$ of priority queue operations, we say that $l$ is data-differentiated, if for each item $d \in \mathcal{D}$, $\textit{put}(d,\_)$ appears at most once in $l$. Note that a data-differentiated sequence $l$ may contains more than one items with a same priority. The subset of data-differentiated operation sequences of a set $S$ is denoted by $S_{\neq}$. For priority queue, the renaming functions rename only the items and keep the priorities unchanged. For priority queue, we modify the notion of data-independence \cite{Wolper:1986} and use data-differentiated instead of differentiated:

%\vspace{-6pt}
\begin{definition}[data-independence of priority queue]\label{def:priority-value data-independence}
The set $S$ of sequences of priority queue operations (resp., histories)  is data independent, if:
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] for all $l \in S$, there exists $l' \in S_{\neq}$, and a renaming $r$, such that $l=r(l')$,

\item[-] for all $l \in S$ and for all renaming $r$, $r(l) \in S$.
\end{itemize}
\end{definition}

Similar as in \cite{Bouajjani:2015}, we consider only completed executions, where each call action has a corresponding return action. This simplification is sound when implementation methods can always make progress in isolation.

For the inductive rules of priority queue, our rules need to concern the cost of operations. Formally, the inductive rules for priority queues is as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $R_{\textit{pr0}} \equiv \epsilon \in \textit{PQueue}$

\item[-] $R_{\textit{pr1}} \equiv (u \cdot v \cdot w \in \textit{PQueue}) \wedge (u \cdot v \cdot w \in (\textit{put},\textit{rm})^*) \wedge (\textit{pri} >$ priorities of $u \cdot v \cdot w ) \wedge (\textit{matched}(u \cdot v, \textit{put}, \textit{rm}) ) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w \in \textit{PQueue})$

\item[-] $R_{\textit{pr2}} \equiv (u \cdot v \cdot w \in \textit{PQueue}) \wedge (u \cdot v \cdot w \in (\textit{put},\textit{rm})^*) \wedge (\textit{pri} =$ the maximal priority of $u \cdot v \cdot w ) \wedge (\textit{pri} >$ priorities of unmatched items in $u \cdot v \cdot w) \wedge (\textit{matched}(u \cdot v, \textit{put}, \textit{rm}) ) \wedge ($all $\textit{put}$ operation with priority $\textit{pri}$ is in $u) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w \in \textit{PQueue})$

\item[-] $R_{\textit{pr3}} \equiv (u \cdot v \in \textit{PQueue}) \wedge ( u \cdot v \in (\textit{put},\textit{rm})^*) \wedge (\textit{pri} =$ the maximal priority of items in $u \cdot v) \wedge (\textit{pri} >$ priorities of matched items in $u \cdot v \cdot w ) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{PQueue})$

\item[-] $R_{\textit{pr4}} \equiv (u \cdot v \in \textit{PQueue}) \wedge ( u \cdot v \in (\textit{put},\textit{rm})^*) \wedge (\textit{pri} =$ the maximal priority of items in $u \cdot v) \wedge (\textit{pri} =$ the maximal priority of matched items in $u \cdot v \cdot w ) \wedge ($all matched $\textit{put}$ with priority $\textit{pri}$ is in $u) \wedge ($all unmatched $\textit{put}$ with priority $\textit{pri}$ is in $u) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{PQueue})$

\item[-] $R_{\textit{pr5}} \equiv (u \cdot v \in \textit{PQueue}) \wedge (\textit{matched}(u,\textit{put},\textit{rm}) ) \Rightarrow (u \cdot \textit{rmEmpty} \cdot v \in \textit{PQueue})$
\end{itemize}

Given a sequence $l$ of possibly $\textit{put}$, $\textit{rm}$ and $\textit{rmEmpty}$ operations, we say that $l$ contains unmatched $\textit{rm}$ operations, if the number of $\textit{rm}(d)$ operations is larger than the number of $\textit{put}(d,\_)$ operations for some $d$ in $l$. By data-independence, we can detect the existence of unmatched $\textit{rm}$ operations by automata. A automaton $\mathcal{A}_{\textit{unmRm1}}$ is given in \figurename~\ref{fig:automata for unmatched rm1}, where $c_1=\textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm),\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$, $c_2=c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c_1 + \textit{cal}(\textit{put},b,\textit{anyPri}) + \textit{ret}(\textit{rm},b)$. Here $\textit{anyPri}$ is a specific predicates that is satisfied by any priority. $\mathcal{A}_{\textit{unmRm1}}$ is used to detect the existence of a item that has been removed but has never been inserted.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_RM1.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{unmRm1}}$}
  \label{fig:automata for unmatched rm1}
\end{figure}

An automaton $\mathcal{A}_{\textit{unmRm2}}$ is given in \figurename~\ref{fig:automata for unmatched rm2}, where $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm),$ $\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$, and $c_1 = c_1 + \textit{ret}(\textit{rm},b)$.  $\mathcal{A}_{\textit{unmRm1}}$ is used to detect the existence of a item that has been removed more than once but has been inserted once.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_RM2.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{unmRm2}}$}
  \label{fig:automata for unmatched rm2}
\end{figure}

The following lemma states that we can detect the existence of unmatched $\textit{rm}$ by $\mathcal{A}_{\textit{unmRm1}}$ and $\mathcal{A}_{\textit{unmRm2}}$.

\begin{restatable}{lemma}{DetectUnmatchedRmByAutomata}
\label{lemma:detect unmatched rm by automata}

Given a data-independent $\mathcal{I}$, there is some unmatched $\textit{rm}$ in some sequence of $\mathcal{I}$, if and only if $( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm1}} \neq \emptyset) \vee ( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm2}} \neq \emptyset )$.
\end{restatable}

\begin {proof}

For the $\textit{if}$ direction, assume that $( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm1}} \neq \emptyset) \vee ( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm2}} \neq \emptyset )$. Then, there exists a sequence $l \in \mathcal{I}$ and $l$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$ or $\mathcal{A}_{\textit{unmRm2}}$. If $l$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$, then $l$ contains at least one $\textit{rm}(b)$ operation and no $\textit{put}(b,\_)$ operation; Else, if $l$ is accepted by $\mathcal{A}_{\textit{unmRm2}}$, then $l$ contains at least two $\textit{rm}(b)$ operations and one $\textit{put}(b,\_)$ operation. In both cases, $l$ contains unmatched $\textit{rm}$ operations.

To prove the $\textit{only if}$ direction assume that there exists a sequence $l \in \mathcal{I}$ that contains unmatched $\textit{rm}$ operations. By data-independence, there exists data-differentiated sequence $l'$ and renaming function $r$, such that $l=r(l')$. We need to prove that in $l'$, one of the following cases holds: (1) $l'$ contains $\textit{rm}(d_1)$ for some $d_1$ and does not contain any $\textit{put}(d_1,\_)$, and (2) $l'$ contains one $\textit{put}(d_1,\_)$ operation and more than one $\textit{rm}(d_1)$ operations. We prove this by contradiction. Assume by contradiction that both cases do not hold. Then, for each value $d$ of $l'$, there is either one $\textit{put}(d,\_)$ and one $\textit{rm}(d)$ in $l'$, or one $\textit{put}(d,\_)$ and no $\textit{rm}(d)$ in $l'$. It is obvious that such $l'$ does not contain unmatched $\textit{rm}$ operation, and renaming a sequence that does not contain unmatched $\textit{rm}$ we can only obtain another sequence that does not contain unmatched $\textit{rm}$. This contradicts that $l=r(l')$ has unmatched $\textit{rm}$.

Let $l''$ be obtained from $ll$ by renaming $d_1$ to $b$ and renaming other value to $a$. If the first case holds, $l''$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$; Otherwise, the second case holds, and $l''$ is accepted by $\mathcal{A}_{\textit{unmRm2}}$. \qed
\end {proof}

Therefore, from now on, we can safely assume that there are no unmatched $\textit{rm}$ operations. The notion of $\textit{last}$ is defined as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $l$ contains $\textit{rmEmpty}$ operations, then $\textit{last}(l) = R_{\textit{pr5}}$.
\item[-] Else, $l$ does not contain any $\textit{rmEmpty}$ operation.
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If the maximal priority of $l$ is an unmatched $\textit{put}$ operation $o$, and the priority of $o$ equals the priority of some matched $\textit{put}$ operations in $l$, then $\textit{last}(l) = R_{\textit{pr4}}$.

    \item[-] If the maximal priority of $l$ is an unmatched $\textit{put}$ operation $o$, and the priority of $o$ is larger than the priority of all matched $\textit{put}$ operations in $l$, then $\textit{last}(l) = R_{\textit{pr3}}$.

    \item[-] If the maximal priority of $l$ is a matched $\textit{put}$ operation $o$, the priority of $o$ equals the priority of some matched $\textit{put}$ operations in $l$, and the priority of $o$ is larger than the priority of all unmatched $\textit{put}$ operations in $l$, then $\textit{last}(l) = R_{\textit{pr2}}$.

    \item[-] If the maximal priority of $l$ is an matched $\textit{put}$ operation $o$, and the priority of $o$ is larger than the priority of all $\textit{put}$ operations in $l$, then $\textit{last}(l) = R_{\textit{pr1}}$.

    \item[-] Otherwise, $l = \epsilon$, and $\textit{last}(l) = R_{\textit{pr0}}$.
    \end{itemize}
\end{itemize}

%The following lemma states that the inductive rules for priority queues are sound and complete.

\section{Step-by-Step Linearizability of Priority Queues}
\label{sec:step-by-step linearizability of priority queues}

The following lemma states that $\textit{PQueue}$ is closed under projection. This lemma holds since the predicates appears in inductive rules of $\textit{PQueue}$ are closed under projection.

\begin{restatable}{lemma}{PQisClosedUnderProjection}
\label{lemma:PQ is closed under projection}

$\textit{PQueue}$ is closed under projection.

\end{restatable}

%When the context is clear, we shorten priority-value data-independence as data-independence.
For priority queue, the notion of step-by-step linearizability is modified as follows:

%\vspace{-6pt}
\begin{definition}[step-by-step linearizability for priority queue]\label{def:step-by-step linearizability of priority queue}
A data structure $S=R_1, \ldots, R_n$ is said be to step-by-step linearizable, if for any data-differentiated sequence $e$, if $e$ is linearizable w.r.t. $\textit{MR}_{\textit{pri}}$ ($1 \leq i \leq 4$) with witness $x$, we have:

$e \setminus x \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket \Rightarrow e \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$

\noindent if $e$ is linearizable w.r.t. $\textit{MR}_{\textit{pr5}}$ with witness $x$, we have:

$e \setminus x \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket \Rightarrow e \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$
\end{definition}

Let $\textit{proj}(s)$ denote the projection of $s$ obtained by erasing all the operations of several items from $s$. Here we assume that each $\textit{rm}(\textit{empty})$ is differentiated from each other, and we can consider each $\textit{rm}(\textit{empty})$ as having a unique item.

\begin{restatable}{lemma}{PQasMultiInMRpriforSequence}
\label{lemma:PQ as multi in MRpri for sequence}
$s \in \textit{PQueue}$, if and only if, $\forall s'$, if $s' \in \textit{proj}(s)$, then $s' \in \textit{MR}_{\textit{pri}}$ for some $0 \leq i \leq 5$.
\end{restatable}

\begin {proof}

The \textit{only if} direction is a direct consequence of Lemma \ref{lemma:PQ is closed under projection}.

To prove the $\textit{if}$ direction. The $\textit{if}$ direction holds trivially when $s = \epsilon$. When $s \neq \epsilon$. since $s \in \textit{proj}(s)$, we generate sequence $s_1$ from $s$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $s$ has $\textit{rm}(\textit{empty})$ operations: $s_1$ is generated from $s$ by erasing some $\textit{rm}(\textit{empty})$ operation.

\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in $S_1$ and unmatched $\textit{put}$ operations of some items in $S_2$: $s_1$ is generated from $s$ by erasing the unmatched $\textit{put}$ operation of items last putted in $S_2$.

\item[-] If the maximal priority in $s$ is unmatched $\textit{put}$ operations of some items in $S_2$ and $\vert S_2 \vert =1$: $s_1$ is generated from $s$ by erasing this unmatched $\textit{put}$ operation.

\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert \geq 2$: $s_1$ is generated from $s$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item which is last putted in $S$.

\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert = 1$: $s_1$ is generated from $s$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item in $S$.
\end{itemize}

Similarly, we obtain $s_2$ from $s_1$, obtain $s_3$ from $s_2$, $\ldots$, until we obtain some $s_m = \epsilon$. It is obvious that $s_m \in \textit{PQueue}$, and by the five rules of $\textit{PQueue}$, it is easy to see that $s_{\textit{m-1}},\ldots,s_1,s \in \textit{PQueue}$. \qed
\end {proof}

\begin{restatable}{lemma}{PQasMultiInMRpriforHistory}
\label{lemma:PQ as multi in MRpri for history}
Given a data-differentiated history $h$, assume that $\textit{PQueue}$ is step-by-step linearizability. $h \sqsubseteq \textit{PQueue}$, if and only if, $\forall h'$, if $h' \in \textit{proj}(h)$, then $h' \sqsubseteq \textit{MR}_{\textit{pri}}$, where $i = \textit{last}(h')$.
\end{restatable}

\begin {proof}

The \textit{only if} direction is a direct consequence of Lemma \ref{lemma:PQ is closed under projection}.

To prove the $\textit{if}$ direction. Since $h \sqsubseteq \textit{PQueue}$, we can see that $h \sqsubseteq l \in \textit{PQueue}$. Each projection $h'$ of $h$ can be linearizable to some projection $l'$ of $l$. By Lemma \ref{lemma:PQ as multi in MRpri for sequence}, $l' \in \textit{MR}_{\textit{prj}}$ for some $0 \leq j \leq 5$. Since $\textit{last}$ of a sequence is uniquely determined, we can see that $l' \in \textit{MR}_{\textit{pri}}$, where $i = \textit{last}(h')$.

The $\textit{if}$ direction holds trivially when $h = \epsilon$. When $h \neq \epsilon$. since $h \in \textit{proj}(h)$, we generate sequence $h_1$ from $h$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $h$ has $\textit{rm}(\textit{empty})$ operations: Then $\textit{last}(h) = R_{\textit{pr5}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr5}}$. $h_1$ is generated from $h$ by erasing some $\textit{rm}(\textit{empty})$ operation.

\item[-] If the maximal priority in $h$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in $S_1$ and unmatched $\textit{put}$ operations of some items in $S_2$: Then $\textit{last}(h) = R_{\textit{pr4}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr4}}$. $h_1$ is generated from $h$ by erasing a unmatched $\textit{put}(x,\_)$, such that $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$ for each $y \in S_2$.

\item[-] If the maximal priority in $s$ is unmatched $\textit{put}$ operations of some items in $S_2$ and $\vert S_2 \vert =1$: Then $\textit{last}(h) = R_{\textit{pr3}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr3}}$. $h_1$ is generated from $h$ by erasing this unmatched $\textit{put}$ operation.

\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert \geq 2$: Then $\textit{last}(h) = R_{\textit{pr2}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr2}}$. $h_1$ is generated from $h$ by erasing $\textit{put}(x,\_)$ and $\textit{rm}(x)$, such that for each $y \in S$, $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$, and $\textit{rm}(x)$ does not happen before $\textit{rm}(y)$.

\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert = 1$: Then $\textit{last}(h) = R_{\textit{pr1}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr1}}$. $h_1$ is generated from $h$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item in $S$.
\end{itemize}

Similarly, we obtain $h_2$ from $h_1$, obtain $h_3$ from $h_2$, $\ldots$, until we obtain some $h_m = \epsilon$. Assume that for each $i$, when generating $h_i$ from $h_{\textit{i-1}}$, we erase operations of item $d_i$. It is obvious that $h_m = h_{\textit{m-1}} \setminus d_m \sqsubseteq \textit{PQueue}$, and since $h_{\textit{m-1}} \sqsubseteq \textit{MR}_{\textit{prj}_{\textit{m-1}}}$, where $\textit{last}(h_{\textit{m-1}}) = R_{\textit{prj}_{\textit{m-1}}}$, we can see that $h_{\textit{m-1}} \sqsubseteq \textit{PQueue}$. Similarly, we can prove that $h_{\textit{m-1}},\ldots,h_1,h \sqsubseteq \textit{PQueue}$. This completes the proof of the $\textit{if}$ direction. \qed
\end {proof}

To prove that $\textit{PQueue}$ is step-by-step linearizability, we use five lemmas for five rules of $\textit{PQueue}$.

\begin{restatable}{lemma}{Rpr1isStepByStepLinearizability}
\label{lemma:Rpr1 is step-by-step linearizability}

$R_{\textit{pr1}}$ is step-by-step linearizability.

\end{restatable}

\begin {proof}

Assume that $l$ matches $R_{\textit{pr1}}$ with witness $x$. Assume that  Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$. Then, let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.

According to $R_{\textit{pr1}}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where the $\textit{put}$ operations and $\textit{rm}$ operations in $u \cdot v$ are matched, and $\textit{pri}_x$ is larger than any priority in $u \cdot v \cdot w$. Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$. Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$ operations, we can see that $D_L \cap D_R = \emptyset$. Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$. Let $S_1$ be the set of all operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_{L1}$ be the shortest prefix of $l'_L$ that contains all operations of $S_1$, and let $l'_{L2}$ be the sequence such that $l'_L = l'_{L1} \cdot l'_{L2}$. It is obvious that if $\textit{put}(x,\textit{pri}_x)$ happens before an operation $o$ in $h$, then $o$ is in $l'_{L2}$. Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.

Since priority queue is closed under projection (Lemma \ref{lemma:PQ is closed under projection}) and $u \cdot v$ contain matched $\textit{put}$ and $\textit{rm}$ operations, we know that $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$. Since $\textit{pri}_x$ is larger than any priority in $h$, it is easy to see that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$. Since the priority queue is empty after executing $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$, we know that $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R \in \textit{PQueue}$.

It remains to prove that $h \sqsubseteq l''$. To prove this, we define a graph $G$ whose nodes are the operations of $h$ and there is an edge from operation $o_1$ to $o_2$, if one of the following case holds

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $o_1$ happens-before $o_2$ in h,

\item[-] the method event corresponding to $o_1$ in $l''$ is before the one corresponding to $o_2$.
\end{itemize}

Assume there is a cycle in $G$. According the the property of interval order and the fact that the order of $l''$ is total, we know that there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$. Then, we can see

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $o_1,o_2 \in l'_{L1}$, or $o_2 \in l'_{L1} \wedge o_1 \in l'_{L2}$, or $o_1,o_2 \in l'_{L2}$, or $o_1,o_2 \in L'_R$, then $l'$ contradicts with happen before relation of $h$.

\item[-] If $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_{L1}$, then since $\textit{put}(x,\textit{pri}_x)$ happens before $o_2$, we know that $o_2 \in l'_{L2}$, which contradicts that $o_2 \in l'_{L1}$.

\item[-] If $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L1}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L1}$, or $o_1 = \textit{rm}(x) \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 \in L'_R \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 = \textit{rm}(x)$, then $l$ contradicts with happen before relation of $h$.

\item[-] If $o_1 \in l'_{L2} \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, then since $o_1$ happens before $\textit{put}(x,\textit{pri}_x)$, we know that $o_1 \in S_1$ and then $o_1 \in l'_{L1}$, which contradicts that $o_1 \in l'_{L2}$.
\end{itemize}

Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed

\end {proof}

\begin{restatable}{lemma}{Rpr2isStepByStepLinearizability}
\label{lemma:Rpr2 is step-by-step linearizability}

$R_{\textit{pr2}}$ is step-by-step linearizability.

\end{restatable}

\begin {proof}

Assume that $l$ matches $R_{\textit{pr2}}$ with witness $x$. Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$. Let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.

Since $h' \sqsubseteq l'$, there exists a sequence $h_{l'}$ of call, return and operations, such that (1) the projection of $h_{l'}$ into call and return action is $h$, (2) the projection of $h_{l'}$ into operations is $l'$, and (3) for each matched pair of call action $c$ and return action $r$ of $h$, there exists an operation $o$, such that $a$ before $o$, $o$ before $r$, and $o$ is the operation of $c$ and $r$. Intuitively, $h_{l'}$ is obtained from $h$ by adding linearization points.

According to $R_{\textit{pr2}}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where (1) the $\textit{put}$ operations and $\textit{rm}$ operations in $u \cdot v$ are matched, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v \cdot w$, (3) there exists matched $\textit{put}$ and $\textit{rm}$ operations in $u \cdot v \cdot w$ that has priority $\textit{pri}_x$, (4) $\textit{pri}_x$ is larger than priorities of any unmatched put in $u \cdot v \cdot w$, and (5) all $\textit{put}$ operation with priority $\textit{pri}_x$ is in $u$. Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$. Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$ operations, we can see that $D_L \cap D_R = \emptyset$. Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.

Let $l'_{\textit{L0}}$ be the shortest prefix of $l'_L$, such that every operation in $l'_{\textit{L0}}$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $h_{l'}$. Let $\textit{OP}_x = \{ \textit{put}(d,\textit{pri}_x) \vert \textit{put}(d,\textit{pri}_x)$ is an operation of $h\}$. Let $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t$, where $l_t$ is the projection of $l'$ into operations of $\textit{OP}_x$ that are not in $l'_{\textit{L0}}$. Let $\textit{OP}'_x$ be the set of operations in $l_t$. Let $l'_{\textit{L2}}$ be the projection of $l'_L$ into operations of $l'_L$ that are not in $l'_{\textit{L1}}$. Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.

As in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, we know that $l'_L \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$. It is obvious that $l'_{\textit{L0}} \in \textit{PQueue}$, and then $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t \in \textit{PQueue}$. Since there is no $\textit{rm}(\textit{Empty})$ in $h$ and $\textit{pri}_x$ is the maximal priority of $h$, $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$. Since all $\textit{put}$ operations with priority $\textit{pri}$ is in $l'_{L1}$, we know that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$. As in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, we know that $l'' \in \textit{PQueue}$.

To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}. Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$. Then, we can see

\begin{itemize}
\setlength{\itemsep}{0.5pt}

\item[-] $o_1 \in L'_R \wedge (o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2} \vee o_2 = \textit{rm}(x))$, or $o_1 = \textit{rm}(x) \wedge (o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2})$: Then $l$ contradicts with happen before relation of $h$.

\item[-] $o_1, o_2 \in l'_R$ or $o_1, o_2 \in l'_{L2}$: Then $l'$ contradicts with happen before relation of $h$.

\item[-] $o_1, o_2 \in l'_{L1}$: If $(o_1, o_2 \in l'_{L0}) \vee (o_1,o_2 \in \textit{OP}'_x)$, then $l'$ contradicts with happen before relation of $h$.

    Else, it is the case that $o_1 \in \textit{OP}'_x \wedge o_2 \in l'_{L0}$. We can see that $o_1$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and the linearization point of $o_1$ is after the return action of $\textit{put}(x,\textit{pri}_x)$. Since $o_1 <_{\textit{hb}} o_2$, it is easy to see that $\textit{put}(x,\textit{pri}_x) <_{\textit{hb}} o_2$, and then $o_2 \notin l'_{L0}$, which contradicts that $o_2 \in l'_{L0}$.

\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_{L1}$: Since $o_1 <_{\textit{hb}} o_2$, it is obvious that $o_2 \notin l'_{L1}$, contradicts that $o_2 \in l'_{L1}$.

\item[-] $o_1 \in l'_{L2} \wedge o_2 \in l'_{L1}$: If $o_2 \in l'_{L0}$, then $l'$ contradicts with happen before relation of $h$.

    Else, it is the case that $o_2 \in \textit{OP}'_x$. $o_2$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and the linearization point of $o_2$ is after the return action of $\textit{put}(x,\textit{pri}_x)$. Since $o_1 <_{\textit{hb}} o_2$, the return action of $o_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$. Then it is easy to see that $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.

\item[-] $o_1 \in l'_{L2} \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: Then $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.
\end{itemize}

Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
\end {proof}

\begin{restatable}{lemma}{Rpr3isStepByStepLinearizability}
\label{lemma:Rpr3 is step-by-step linearizability}

$R_{\textit{pr3}}$ is step-by-step linearizability.

\end{restatable}

\begin {proof}

Assume that $l$ matches $R_{\textit{pr3}}$ with witness $x$. Assume that  Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$. Then, let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.

According to $R_{\textit{pr3}}$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, where $\textit{pri}_x$ is the maximal priority in $u \cdot v$, and $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ operations in $u \cdot v$ (may equal priority of some unmatched $\textit{put}$ operations in $u \cdot v$). Let $S_1$ be the set of all operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_1$ be the shortest prefix of $l'$ that contains all operations of $S_1$, and let $l'_2$ be the sequence such that $l' = l'_1 \cdot l'_2$. Let sequence $l'' = l'_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_2$.

Since $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ operations in $u \cdot v$, it is easy to see that $l'' \in \textit{PQueue}$. It remains to prove that $h \sqsubseteq l''$. To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}. Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$. Then, we can see

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $o_1,o_2 \in l'_1$, or $o_1,o_2 \in l'_2$, or $o_1 \in l'_2 \wedge o_2 \in l'_1$: Then $l'$ contradicts with happen before relation of $h$.

\item[-] $o_1 \in l'_2 \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: $o_1 \in S_1$, and then $o_1 \in l'_1$, which contradicts that $o_1 \in l'_2$.

\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_1$: We can see that in $l'$, $o_2$ must after every operation in $S_1$, and then $o_2 \notin l'_1$, contradicts that $o_2 \in l'_1$.
\end{itemize}

Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
\end {proof}

The following lemma states that, from linearization of sub-histories, we can merge them and obtain a linearization (regardless of whether it belongs to sequential specification) of the whole history.

\begin{restatable}{lemma}{MergeTwoLinearization}
\label{lemma:merge two linearization}

Given a history $h$ and operation set $S_1$, $S_2$, let $h_1 = h \vert{S_1}$ and $h_2 = h \vert{S_2}$. Assume that $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, and $S_1 \cup S_2$ contains all operations of $h$. Then, there exists a sequence $l$, such that $h \sqsubseteq l$, $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$.
\end{restatable}

\begin {proof}

Given a history $h$ and a operation $o \in S_2$, let $\textit{MB}(o) = \{ o' \vert o' \in S_1$ and $o'$ happens before $o$ in $h \}$, let $\textit{SBI}(o) = \textit{min}\{ i \vert l_1[0,i]$ contains all elements of $\textit{MB}(o) \}$.

Let $l = s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[n] \cdot s_{\textit{n+1}}$ be generated as follows, where $n = \vert l_2 \vert$:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $s_1 = l_1[0, \textit{SBI}(l_2[1])]$,

\item[-] If $s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[i]$ already contains $l_1(\textit{SBI}(l_2[\textit{i+1}]))$, then $s_{\textit{i+1}} = \epsilon$. Otherwise, $s_{\textit{i+1}}$ is a subsequence of $l_1$, which starts from the next of last elements of $s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[i]$ in $l_1$ and ends in $l_1(\textit{SBI}(l_2[\textit{i+1}]))$.
\end{itemize}

It is obvious that $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$, and it remains to prove that $h \sqsubseteq l$. We prove this by contradiction. Assume that $h$ is not linearizable with respect to $l$. Then there must be two operations of $h$, such that $o_1 <_{hb} o_2$ in $h$ but $o_2$ before $0_1$ in $l$. Since $l \vert{S_1} = l_1$, $l \vert{S_2} = l_2$, and $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, it is easy to see that it is impossible that $o_1,o_2 \in S_1$ or $o_1,o_2 \in S_2$. There are only two possibilities:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $o_1 \in S_1 \wedge o_2 \in S_2$. Then we can see that $o_2=l_2[i]$ and $o_1 \in s_j$ for some $i < j$. Since $o_1 <_{hb} o_2$, we know that $o_1 \in \textit{SBI}(o_2)$. By the construction of $l$, we know that $o_1$ must be in $s_k$ for some $k \leq i$, contradicts that $o_1 \in s_j$ with $i < j$.

\item[-] $o_1 \in S_2 \wedge o_2 \in S_1$. Then we can see that $o_2 \in s_i$ and $o_1 = l_2[j]$ for some $i \leq j$. It is easy to see that this leads to contradiction when $i = j$. For the case of $i \neq j$, we need to satisfy the following requirements: (1) $o_1$ ($l_2[j]$) does not happen before $l_2[i]$, (2) $l_2[i]$ does not happen before $o_2$, (3) $o_2$ is either overlap or happens before $o' \in \textit{MB}(l_2[i])$, and (4) $o' <_{hb} l_2[i]$. By enumeration we can see that it is impossible that above four conditions be satisfied while $o_1 <_{hb} o_2$.
\end{itemize}

This completes the proof of this lemma. \qed
\end {proof}

\begin{restatable}{lemma}{Rpr4isStepByStepLinearizability}
\label{lemma:Rpr4 is step-by-step linearizability}

$R_{\textit{pr4}}$ is step-by-step linearizability.

\end{restatable}

\begin {proof}

Assume that $l$ matches $R_{\textit{pr4}}$ with witness $x$. Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$. Let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.

According to $R_{\textit{pr4}}$, there exist sequences $u$ and $v$, such that (1) $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v$, (3) $\textit{pri}_x$ equals the maximal priority of matched $\textit{put}$ operations in $u \cdot v$, (4) all matched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$ and (5) all the other unmatched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$.

Since $h' \sqsubseteq l'$, there exists a sequence $h_{l'}$ of call, return and operations, such that (1) the projection of $h$ into call and return action is $h$, (2) the projection of $h_{l'}$ into operations is $l'$, and (3) for each matched pair of call action $c$ and return action $r$ of $h$, there exists an operation $o$, such that $a$ before $o$, $o$ before $r$, and $o$ is the operation of $c$ and $r$. Intuitively, $h_{l'}$ is obtained from $h$ by adding linearization points.

According to $R_{\textit{pr4}}$, there exist sequences $u$ and $v$, such that (1) $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v$, (3) $\textit{pri}_x$ equals the maximal priority of matched $\textit{put}$ operations in $u \cdot v$, (4) all matched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$ and (5) all the other unmatched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$.

Let $l'_1$ be the shortest prefix of $l'$, such that every operation in $l'_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $h_{l'}$. Let $\textit{OP}_x$ be the set of $\textit{put}$ operations with priority $\textit{pri}_x$ and their matched $\textit{rm}$ operations in $h$, let $\textit{OP}'_x$ be the set of $\textit{put}$ operations in $\textit{OP}_x$ that do not appear in $l'_1$, and let $\textit{OP}''_x$ be the set of $\textit{rm}$ operations in $\textit{OP}_x$ that do not appear in $l'_1$. Let $l_x''=l'_1 \cdot l'_2 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_3$, where $l'_2$ is the projection of $l'$ into operations of $\textit{OP}'_x$, and $l'_3$ is the projection of $l'$ into operations of $\textit{OP}''_x$.

Let $h_x$ be the projection of $h$ into $\textit{OP}_x$. Let us prove that $h_x \sqsubseteq l''_x$ by contradiction. Assume that there exists operations $o_1,o_2$, such that $o_1$ happens before $o_2$ in $h_x$, and $o_2$ is before $o_1$ in $l''_x$. Then, we can see

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $(o_1 \in l'_3) \wedge (o_2 \in l'_1 \vee o_2 \in l'_2 \vee o'_2 \in l'_3)$, or $(o_1 \in l'_2) \wedge (o_2 \in l'_1 \vee o_2 \in l'_2)$, or $o_1,o_2 \in l'_1$: Then $l'$ contradicts with happen before relation of $h$.

\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_1$: Since the linearization point of $o_2$ is before return action of $o_1$, $o_2$ either overlaps with $o_1$, or happens before $o_1$, which contradicts that $o_1$ happens before $o_2$.

\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_2$: Since $l'_2$ only contains $\textit{put}$ operations with priority $\textit{pri}_x$ and are in $\textit{OP}'_x$, it is not hard to see that each operation in $l'_2$ must overlap with $\textit{put}(x,\textit{pri}_x)$, which contradicts that $o_1$ happens before $o_2$.

\item[-] $o_1 \in l'_3 \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: According to the construction of $l''$, we can see that $1_2 \in l'_1$, which contradicts that $1_2 \in l'_3$.
\end{itemize}

Therefore, we know that $G$ is acyclic, and then we know that $h_x \sqsubseteq l''_x$.

Let $\textit{OP}_{\textit{-x}}$ be the set of $\textit{put}$ operations with non-$\textit{pri}_x$ priority, and their matched $\textit{rm}$ operations. Let $h_{\textit{-x}}$ be the projection of $h$ into $\textit{OP}_{\textit{-x}}$. Let $l''_{\textit{-x}}$ be the projection of $l'$ into$\textit{OP}_{\textit{-x}}$. It is obvious that $h_{\textit{-x}} \sqsubseteq l''_{\textit{-x}}$. By Lemma \ref{lemma:merge two linearization}, there exists a sequence $l'''$, such that $h \sqsubseteq l'''$, $l''' \vert{\textit{OP}_x} = l_x''$ and $l''' \vert{\textit{OP}_{\textit{-x}}} = l''_{\textit{-x}}$.

It is obvious that $l''_{\textit{-x}} \in \textit{PQueue}$. Let $l'_x$ be the projection of $l'$ into $\textit{OP}_x - \{ \textit{put}(x,\textit{pri}_x) \}$. It is obvious that $l'_x \in \textit{PQueue}$ and $l'_1 \in \textit{PQueue}$. Since (1) $l'_1 \cdot l'_2 \cdot l'_3$ is obtained from $l'_x$ by making $\textit{put}$ operations in $l'_2$ to happen earlier, and (2) $l'_3$ contains only $\textit{rm}$ operations, we can see that $l'_1 \cdot l'_2 \cdot l'_3 \in \textit{PQueue}$. Since $l'_3$ contains only $\textit{rm}$ operations, it is easy to see that $l_x'' \in \textit{PQueue}$. Since $\textit{pri}_x$ is the maximal priority of $h$ and $h$ does not contain $\textit{rm}(\textit{Empty})$ operations, the operations in $l_x''$ does not influence the correctness of $l''_{\textit{-x}} \in \textit{PQueue}$. Therefore, we can see that $l''' \in \textit{PQueue}$. \qed
\end {proof}

\begin{restatable}{lemma}{Rpr5isStepByStepLinearizability}
\label{lemma:Rpr5 is step-by-step linearizability}

$R_{\textit{pr5}}$ is step-by-step linearizability.

\end{restatable}

\begin {proof}

Assume that $l$ matches $R_{\textit{pr5}}$ with operation $o=\textit{rmEmpty}$. Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$. Let $h'=h \setminus o$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$. Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$ such that $h' \sqsubseteq l'$.

According to $R_{\textit{pr5}}$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{rmEmpty} \cdot v$, where the $\textit{put}$ operations and $\textit{rm}$ operations in $u$ are matched.

Let $D_L$ be the set of items in $u$ and $D_R$ be the set of items in $v$. Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$. Let sequence $l'' = l'_L \cdot o \cdot L'_R$. Since priority queue is closed under projection (Lemma \ref{lemma:PQ is closed under projection}) and $u$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $l'_L \in \textit{PQueue}$ and the the priority queue is empty after executing $l'_L$. Then we know that $l'_L \cdot o \in \textit{PQueue}$. Since $l'_R$ is obtained from $l'$ by discarding pairs of matched $\textit{put}$ and $\textit{rm}$ operations, it is easy to see that $L'_R \in \textit{PQueue}$, and then we know that $l'' = l'_L \cdot o \cdot L'_R \in \textit{PQueue}$.

It remains to prove that $h \sqsubseteq l''$. To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}. Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$. Then, we can see

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $o_1,o_2 \in l'_L$, or $o_1,o_2 \in l'_R$: Then $l'$ contradicts with happen before relation of $h$.

\item[-] If $o_1=o \wedge o_2 \in l'_L$, or $o_1 \in l'_R \wedge o_2 \in l'_L$, or $o_1 \in l'_R \wedge o_2 = 0$, then $l$ contradicts with happen before relation of $h$.
\end{itemize}

Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$. \qed
\end {proof}

The following lemma states that $\textit{PQueue}$ is step-by-step linearizability, it is a direct consequence of Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, Lemma \ref{lemma:Rpr2 is step-by-step linearizability}, Lemma \ref{lemma:Rpr3 is step-by-step linearizability}, Lemma \ref{lemma:Rpr4 is step-by-step linearizability} and Lemma \ref{lemma:Rpr5 is step-by-step linearizability}.

\begin{restatable}{lemma}{PQueueisStepByStepLinearizability}
\label{lemma:PQueue is step-by-step linearizability}

$\textit{PQueue}$ is step-by-step linearizability.

\end{restatable}

\section{Co-Regular of Priority Queues}
\label{sec:co-regular of priority queues}

\subsection{Co-Regular of $R_{\textit{pr1}}$}
\label{subsec:co-regular of Rpr1}

Let us introduce the notion of gap for matched $\textit{put}$ and $\textit{rm}$ operations.

\begin{definition}[gap for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:gap for matched put and rm operations}

Let $h$ be a data-differentiated history and $o_1 = \textit{put}(x,\textit{pri}_x)$, $o_2 = \textit{rm}(x)$ be a pair of matched $\textit{put}$ and $\textit{rm}$ operation of $h$. We say that $h$ has a gap on operations $o_1$ and $o_2$, if there is a partition of the operations of $h$ into $L$, $M$ and $R$ satisfying:
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $L \cup M$ has matched $\textit{put}$ and $\textit{rm}$ operations.

\item[-] no operation of $R$ happens before an operation of $L$, $o_1$, $M$ or $o_2$.

\item[-] no operation of $o_2$ happens before an operation of $L$, $o_1$ or $M$.

\item[-] no operation of $M$ happens before an operation of $L$ or $o_1$.

\item[-] no operation of $o_1$ happens before an operation of $L$.

\item[-] no operation in $L \cup M \cup R$ has priority $\textit{pri}_x$.
\end{itemize}
\end{definition}

Given a history $h$ and set $O$ of some operations in $h$. It is easy to see that there exists a sequence $l$ of operations of $O$, such that $l$ does not contradicts the happen before relation of $h$. Therefore, from now on, we can use this fact to directly generate such sequence from such $O$.

The following lemma reduce linearizable with respect to $\textit{MR}_{\textit{pr1}}$ into gap of matched $\textit{put}$ and $\textit{rm}$ operations.

\begin{restatable}{lemma}{GapEqualsLinforRpr1}
\label{lemma:Gap Equals Lin for Rpr1}
A data-differentiated history $h$ has a projection $h'$, such that $\textit{last}(h') = R_{\textit{pr1}}$ and $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$, if and only if there exists a projection $h''$ of $h$, such that $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is in $h''$, $\textit{pri}_x$ is the maximal priority of $h''$, and one the the following case holds:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] case $1$: $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$,
\item[-] case $2$: there is no gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
\end{itemize}
\end{restatable}

\begin {proof}

Let us prove the $\textit{only if}$ direction by contradiction. Assume that there exists such $h'$, while case $1$ and case $2$ do not hold for any projection of $h$. Since case $1$ does not hold for any projection of $h$, we know that $\forall y \in D$, $\textit{rm}(y)$ does not happens before $\textit{put}(y,\_)$ in $h$. Since case $2$ does not hold for any projection of $h$ and $\textit{last}(h') = R_{\textit{pr1}}$, we can see that $h'$ has gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$, while $\textit{pri}_x$ is the maximal priority in $h'$. Since $\textit{last}(h') = R_{\textit{pr1}}$, we can see that $\textit{pri}_x$ is larger than priority of all other $\textit{put}$ operations of $h$.

By the properties of the gap, there is partitions $L$, $M$ and $R$. Let $u$, $v$ and $w$ contain the operations of $L$, $M$ and $R$, respectively. It is easy to see that $h'$ is linearizable with respect to $l''=u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$ and $l'' \in \textit{MR}_{\textit{pr1}}$. This contradicts the fact that $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$.

Let us begin to prove the $\textit{if}$ direction.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If case $1$ holds for $h''$, then $h \vert_{ \{ x \} }$ is the $h'$ as required.

\item[-] If case $2$ holds, assume by contradiction that, for every projection $h'$ of $h$, if $\textit{last}(h') = R_{\textit{pr1}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr1}}$. Since $h''$ is also a projection of $h$, we know that there exist sequences $u$, $v$ and $w$, such that $h'' \sqsubseteq u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, $u \cdot v$ contains matched $\textit{put}$ and $\textit{rm}$ operations, and $\textit{pri}_x$ is larger than priority of $u \cdot v \cdot w$. Let $L$, $M$ and $R$ be the operations of $u$, $v$ and $w$, respectively. Then, the partitions $L$, $M$ and $R$ form a gap on operation $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$. This contradicts the fact that there is no gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
\end{itemize}

This completes the proof of this lemma. \qed
\end {proof}

\begin{definition}[left-right constraint for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:left-right constraint for matched put and rm operations}
Given a data-differentiated history $h$, and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$. The left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is the graph $G$ where:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] the nodes are the items of $h$, to which we add a node,

\item[-] there is an edge from item $d_1$ to $x$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$ and $\textit{pri}_1 \neq \textit{pri}_x$,

\item[-] there is an edge from $x$ to item $d_1$, if $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$, and the priority of $d_1$ is not $\textit{pri}_x$,

\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{rm}(d_2,\textit{pri}_2)$, $\textit{pri}_1 \neq \textit{pri}_x$, and $\textit{pri}_2 \neq \textit{pri}_x$.
\end{itemize}
\end{definition}

Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$. Let $\textit{LMSet}_1(h,x) = \{ o \vert$ $o$ is not an operation of $x$, the priority of $o$ does not equal the priority of $x$, and either it happens before $\textit{put}(x)$ or $\textit{rm}(x)$ in $h$, or there is an operation $o'$ with the same item of $o$, such that $o'$ happens before $\textit{put}(x)$ or $\textit{rm}(x)$ in $h$ $\}$. For each $i \geq 1$, let $\textit{LMSet}_{\textit{i+1}}(h,x) = \{ o \vert$ $o$ is not an operation of $x$ or items in $\textit{LMSet}_k(h,x)$ for each $k \leq i$, the priority of $o$ does not equal the priority of $x$, and either it happens before some operation $o' \in \textit{LMSet}_i(h,x)$ in $h$, or there is an operation $o''$ with the same item of $o$ and $o''$ happens before some operation $o' \in \textit{LMSet}_i(h,x)$ in $h$ $\}$. %We can see that for each $i$, $\textit{LMSet}_i(h,x)$ contains matched $\textit{put}$ and $\textit{rm}$ operations, and
We can see that $\textit{LMSet}_i(h,x) \cap \textit{LMSet}_j(h,x) = \emptyset$ for any $i \neq j$. Let $\textit{LMSet}(h,x) = \textit{LMSet}_1(h,x) \cup \textit{LMSet}_2(h,x) \cup \ldots$. %{\color {red} We say that a history $h$ of priority queue is matched before $x$, if (1) $h$ contains no $\textit{rmEmpty}$ operations, (2) $ h \sqsubseteq l_1 \cdot \textit{put}(x,\_) \cdot l_2 \cdot \textit{rm}(x) \cdot l_3$ for some $l_1,l_2,l_3$, and $l_1 \cdot l_2$ contains matched $\textit{put}$ and $\textit{rm}$ operations.}

The following is an example that shows that, given a data-differentiated history $h$, the set $\textit{LMSet}_i(h,x)$ can be non-empty for each $1 \leq i \leq 3$. In \figurename~\ref{fig:his nobound of LMSet} , $\textit{LMSet}_1(h,x) = \{ \textit{put}(a,1),\textit{rm}(a) \}$, $\textit{LMSet}_2(h,x) = \{ \textit{put}(b,1),\textit{rm}(b) \}$ and $\textit{LMSet}_3(h,x) = \{ \textit{put}(a,c),\textit{rm}(c) \}$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5 \textwidth]{PIC_HIS_NOBOUNDOF_LMSET.pdf}
%\vspace{-10pt}
  \caption{An example for non-emptiness of $\textit{LMSet}_1(h,x)$, $\textit{LMSet}_2(h,x)$ and $\textit{LMSet}_3(h,x)$}
  \label{fig:his nobound of LMSet}
\end{figure}

\begin{restatable}{lemma}{RmxDoesNotHappenBeforeLMSetForRpr1}
\label{lemma:Rmx does not happen before LMSet for Rpr1}
Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$. %, such that $\textit{pri}_x$ is the maximal priority of $h$, and $\textit{last}(h) = R_{\textit{pr1}}$.
Assume that $\textit{rm}(y)$ does not happen before $\textit{put}(y,\_)$ for each item $y$ in $h$. Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$. Assume that $G$ has no cycle going through $x$. Then, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}(h,x)$.
\end{restatable}

\begin {proof}

We prove this lemma by induction, and prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x)$, in $\textit{LMSet}_2(h,x)$, $\ldots$. Note that, according to the definition of $\textit{LMSet}(h,x)$, $\textit{pri}_x$ is different to priority of all items in $\textit{LMSet}(h,x)$.

\noindent (1) Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x)$ by contradiction. Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{LMSet}_1(h,x)$ is an operation of item $d$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d$ does not equals $\textit{pri}_x$).

We use a triple $(t_1,t_2,t_3)$ to represent related information. $t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$. $t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation. $t_2$ and $t_3$ is used for the reason of $o \in \textit{LMSet}_1(h,x)$: $o \in \textit{LMSet}_1(h,x)$, since an operation (of kind $t_2$) of $d$ happens before an operation (of kind $t_3$) of $x$. Let us consider all the possible cases of $(t_1,t_2,t_3)$:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.

\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{rm}(x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{rm}(x)$.

\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x) )$. By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(x,\textit{pri}_x) ) \vee ( \textit{rm}(d) <_{hb} \textit{put}(d,\_) )$, which is impossible.

\item[-] $(\textit{put},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{rm}(x) )$. We can see that $\textit{rm}(d) <_{hb} \textit{rm}(x) <_{hb} \textit{put}(d,\_)$, which contradicts that $\textit{rm}(d)$ does not happen before $\textit{put}(d,\_)$.

\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x) )$. We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.

\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{rm}(x) )$. We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.

\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.

\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{rm}(x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{rm}(x)$.
\end{itemize}

This completes the proof for $\textit{LMSet}_1(h,x)$.

\noindent (2) Assume we already prove that for some $j \geq 1$, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$. Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_{\textit{j+1}}(h,x)$ by contradiction. Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$ is an operation of item $d_{\textit{j+1}}$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d_{\textit{j+1}}$ does not equals $\textit{pri}_x$). We use a triple $(t_1,t_2,t_3)$ to represent related information. $t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$. $t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation. $t_2$ and $t_3$ is used for the reason of $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$: $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$, since an operation (of kind $t_2$) of $d_{\textit{j+1}}$ happens before an operation (of kind $t_3$) of $d_j$, where $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{LMSet}_j(h,x)$. Let us consider all the possible cases of $(t_1,t_2,t_3)$:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_)$. We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.

\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j,\_)$. We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j,\_) ) \wedge ( \textit{rm}(d_j) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.

\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$. By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \vee ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_{\textit{j+1}},\_) )$, which is impossible.

\item[-] $(\textit{put},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$. By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \vee ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_{\textit{j+1}},\_) )$, which is impossible.

\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) )$. Let us consider the reason of $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{LMSet}_j(h,x)$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $( j > 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(_{\textit{j-1}},\_), \textit{rm}(_{\textit{j-1}}) \in \textit{LMSet}_{\textit{j-1}}(h,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, we can see that $\textit{put}(d_{\textit{j+1}},\_) <_{hb} o''$, and then operations of $d_{\textit{j+1}}$ is in $\textit{LMSet}_j(h,x)$, contradicts that operations of $d_{\textit{j+1}}$ is in $\textit{LMSet}_{\textit{j+1}}(h,x)$.

    \item[-] If $( j = 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of $x$: Similar to above case.

    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{LMSet}_{\textit{j-1}}(h,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, we can see that $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} o'' ) \vee ( \textit{rm}(d_j) <_{hb} \textit{put}(d_j,\_) )$, which is impossible.

    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of $x$: Similar to above case.
    \end{itemize}

\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Let $T_{\textit{ind}}$ be the set of sentences $\{ \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}), \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j),\ldots, \textit{put}(d_{\textit{ind+1}},\_) <_{hb} \textit{rm}(d_{\textit{ind}}) \}$. Here each $d_i$ is a item of some operation in $\textit{LMSet}_i(h,x)$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d_i$ does not equals $\textit{pri}_x$ for each $i$). Let us prove that from $T_j$ we can obtain contradiction by induction:

    {\bf Base case $1$}: From $T_1$ we can obtain contradiction.

    Let us prove base case $1$:

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $\textit{put}(d_1,\_)$ happens $o$, and $o$ is an operation of $x$. Then there is a cycle $x \rightarrow d_{\textit{j+1}} \rightarrow \ldots \rightarrow d_1 \rightarrow x$ in $G$, contradicts that $G$ has no cycle going through $x$.

    \item[-] If $\textit{rm}(d_1)$ happens before $o$, and $o$ is an operation of $x$. Then since $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1) <_{hb} o$, we can see that $\textit{put}(d_2,\_) <_{hb} o$, and then $\textit{put}(d_2,\_) \in \textit{LMSet}_1(h,x)$, contradicts that $\textit{put}(d_2,\_) \in \textit{LMSet}_2(h,x)$.
    \end{itemize}

    {\bf Base case $2$}: From $T_2$ we can obtain contradiction.

    Let us prove base case $2$: If $\textit{rm}(d_2) <_{hb} o$, and $o$ is an operation of $d_1$, then since $( \textit{put}(d_3,\_) <_{hb} \textit{rm}(d_2) ) \wedge ( \textit{rm}(d_2) <_{hb} o )$, we know that $\textit{put}(d_3,\_) <_{hb} o$. This implies that $\textit{put}(d_3,\_) \in \textit{LMSet}_2(h,x)$, contradicts that $\textit{rm}(d_3,\_) \in \textit{LMSet}_3(h,x)$. Therefore, it is only possible that $\textit{put}(d_2,\_)$ happens before an operation of $d_1$.

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{put}(d_1,\_)$ and $\textit{put}(d_1,\_)$ happens before operations of $x$, then we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, which is impossible.

    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{put}(d_1,\_)$ and $\textit{rm}(d_1)$ happens before operations of $x$, then by interval order, we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, or $\textit{rm}(d_1) <_{hb} \textit{put}(d_1,\_)$, which is impossible.

    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{put}(d_1,\_)$ happens before operations of $x$, then $x \rightarrow d_{\textit{j+1}} \rightarrow \ldots \rightarrow d_1 \rightarrow x$ in $G$, contradicts that $G$ has no cycle going through $x$.

    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1)$ happens before operations of $x$, then we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, which is impossible.
    \end{itemize}

    {\bf induction step}: Given $\textit{ind} \geq 3$, if from $T_{\textit{ind-1}}$ we can obtain contradiction, then from $T_{\textit{ind}}$ we can also contain contradiction.

    Prove of the induction step: Similarly as base case $2$, we can prove that it is only possible that $\textit{put}(d_{\textit{ind}},\_)$ happens before operations of $d_{\textit{ind-1}}$.

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$ and $\textit{put}(d_{\textit{ind-1}},\_)$ happens before operations of $d_{\textit{ind-2}}$, then we know that $\textit{put}(d_{\textit{ind}})$ happens before operation of $d_{\textit{ind-2}}$, which is impossible.

    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$ and $\textit{rm}(d_{\textit{ind-1}})$ happens before operations of $d_{\textit{ind-2}}$, then by interval order, we know that $\textit{put}(d_{\textit{ind}},\_)$ happens before operation of $d_{\textit{ind-2}}$, or $\textit{rm}(d_{\textit{ind-1}}) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$, which is impossible.

    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{rm}(d_{\textit{ind-1}})$, then we obtain $T_{\textit{ind-1}}$, which already contain contradiction.
    \end{itemize}

    By base case $1$, base case $2$ and the induction step, it is easy to see that $T_j$ contains contradiction.

\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$. We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.

\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$. We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \wedge ( \textit{rm}(d_j) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
\end{itemize}

This completes the proof for $\textit{LMSet}_{\textit{j+1}}(h,x)$. Therefore, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}(h,x) = \textit{LMSet}_1(h,x) \cup \textit{LMSet}_2(h,x) \cup \ldots$. \qed
\end {proof}

\begin{restatable}{lemma}{LMSetHasMatchedPutandRmOperations}
\label{lemma:LMSet has matched put and rm operations}
Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$. Assume that $\textit{rm}(y)$ does not happen before $\textit{put}(y,\_)$ for each item $y$ in $h$. Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$. Assume that $G$ has no cycle going through $x$. Then, $\textit{LMSet}(h,x)$ contains only matched $\textit{put}$ and $\textit{rm}$ operations.
\end{restatable}
\begin {proof}

We prove this lemma by contradiction. Assume that there exists a value, such that $\textit{LMSet}(h,x)$ contains only its $\textit{put}$ operation and does not contain its $\textit{rm}$ operation. Then we can see that there exists $d_1,\ldots,d_j$. Intuitively, $d_1,\ldots,d_j$ are items in $\textit{LMSet}_1(h,x), \ldots, \textit{LMSet}_i(h,x)$, respectively. $\textit{LMSet}(h,x)$ contains $\textit{put}(d_j,\_)$ and does not contain $\textit{rm}(d_j)$. And each $d_i$ is the reason of $d_{\textit{i+1}} \in \textit{LMSet}_{\textit{i+1}}(h,x)$. Formally, we require that

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] For each $1 \leq i \leq j$, operations of $d_i$ belongs to $\textit{LMSet}_i(h,x)$. %For each $1 \leq i \leq j$, the priority of $d_i$ is not $\textit{pri}_x$.

\item[-] For each $i \neq j$, $\textit{put}(d_i,\_),\textit{rm}(d_i) \in \textit{LMSet}_i(h,x)$. $\textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x)$, and $h$ does not contain $\textit{rm}(d_j)$.

\item[-] An operation of $d_1$ happens before an operation of $x$. For each $1 < i \leq j$, an operation of $d_j$ happens an operation of $d_{\textit{i-1}}$.

\item[-] For each $k$ and $\textit{ind}$, if $k > \textit{ind+1}$, then no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.
\end{itemize}

According to the definition of $\textit{LMSet}(h,x)$, it is easy to see that such $d_1,\ldots,d_j$ exists. Let us prove the following fact:

\noindent {\bf $\textit{fact}_1$}: Given $1 \leq i < j$, it can not be the case that $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap.

Proof of $\textit{fact}_1$: We prove $\textit{fact}_1$ by contradiction. Assume that for some $i \neq j$, $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap. Since $\textit{put}(d_i,\_), \textit{rm}(d_i) \in \textit{LMSet}_i(h,x)$, we know that an operation $o_i$ of $d_i$ happens before operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$. Moreover, since $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap, it is not hard to see that the call action of $\textit{put}(d_i,\_)$ and the call action of $\textit{rm}(d_i)$ is before the call action of $o_{\textit{i-1}}$. Since $\textit{put}(d_{\textit{i+1}},\_), \textit{rm}(d_{\textit{i+1}}) \in \textit{LMSet}_{\textit{i+1}}(h,x)$, we know that an operation $o'_{\textit{i+1}}$ of $d_{\textit{i+1}}$ happens before operation $o'_i$ of $d_i$. Then, it is not hard to see that $o'_{\textit{i+1}}$ also happens before $o_{\textit{i-1}}$, which contradicts that For each $k > \textit{ind+1}$, no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.

We already know that an operation of $d_1$ happens before an operation of $x$. By $\textit{fact}_1$, we can ensure that $\textit{put}(d_1,\_)$ happens before an operation of $x$, and then $d_1 \rightarrow x$ in $G$. For each $1 < i < j$, we know that an operation $o_i$ of $d_i$ happens before an operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$. By $\textit{fact}_1$, we can ensure that $o_i=\textit{put}(d_i,\_)$ and $o_{\textit{i-1}}=\textit{rm}(d_{\textit{i-1}})$, and then $d_i \rightarrow d_{\textit{i-1}}$ in $G$. Since $h$ contains $\textit{put}(d_j)$ and does not contain $\textit{rm}(d_j)$, we know that $x \rightarrow d_j$ in $G$. Then $G$ has a cycle going through $x$, contradicts that $G$ has no cycle going through $x$. \qed
\end {proof}

\begin{restatable}{lemma}{GapEqualsConstraintforRpr1}
\label{lemma:Gap Equals Constraint for Rpr1}
Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$. %, where $\textit{pri}_x$ has the maximal priority of $h$, and $\textit{last}(h) = R_{\textit{pr1}}$.
Assume that for each item in $h$, its $\textit{rm}$ operation does not happen before its $\textit{put}$ operation. Let $G$ be the graph representing the left-right constraint of $\textit{put}(x)$ and $\textit{rm}(x)$. There is a gap on $\textit{put}(x)$ and $\textit{rm}(x)$, if and only if $G$ has no cycle going through $x$.
\end{restatable}

\begin {proof}

To prove the $\textit{only if}$ direction, assume that there is a gap on $\textit{put}(x)$ and $\textit{rm}(x)$, and let it be $L$, $M$ and $R$. Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ in $G$. It is obvious that the priority of each $d_i$ is not $\textit{pri}_x$. Then our proof proceeds as follows:

According to the definition of left-right constraint, there are two possibilities. The first possibility is that, $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$. It is obvious that $\textit{rm}(d_1) \in R$, and then since $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_1),\textit{rm}(d_1) \in R$. Then,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Since $d_1 \rightarrow d_2$, by definition of $G$, we know that $\textit{put}(d_1)$ happens before $\textit{rm}(d_2)$. Since $\textit{put}(d_1) \in R$ and $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_2),\textit{rm}(d_2) \in R$. Similarly, for each $1 \leq i \leq m$, we know that $\textit{put}(d_i),\textit{rm}(d_i) \in R$.

\item[-] Since $d_m \rightarrow x$,
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] if $\textit{put}(d_m)$ happens before $\textit{put}(x)$, then we can see that $\textit{put}(d_m) \in L$, which contradicts that $\textit{put}(d_m) \in R$.

    \item[-] if $\textit{put}(d_m)$ happens before $\textit{rm}(x)$, then we can see that $\textit{put}(d_m) \in L \cup M$, which contradicts that $\textit{put}(d_m) \in R$.
    \end{itemize}
\end{itemize}

The second possibility is that, $h$ contains one $\textit{put}(d_1,\_)$ operation and no $\textit{rm}(d_1)$ operation. Note that for each $j > 1$, $h$ contains $\textit{put}(d_j,\_)$ and $\textit{rm}(d_j)$. Since $d_m \rightarrow x$, is is obvious that $\textit{put}(d_m) \in L \cup M$. Since $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_m),\textit{rm}(d_m) \in R$. Then,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Since $d_{\textit{m-1}} \rightarrow d_m$, by definition of $G$, we know that $\textit{put}(d_{\textit{m-1}})$ happens before $\textit{rm}(d_m)$. Since $\textit{rm}(d_m) \in L \cup M$ and $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_{\textit{m-1}}),\textit{rm}(d_{\textit{m-1}}) \in L \cup M$. Similarly, for each $1 < i \leq m$, we know that $\textit{put}(d_i),\textit{rm}(d_i) \in L \cup M$, and $\textit{put}(d_1)\in L \cup M$.

\item[-] Since $d_1 \rightarrow d_2$ in $G$, $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2)$. Since $\textit{put}(d_2),\textit{rm}(d_2) \in L \cup M$, we can see that $\textit{put}(d_1) \in L \cup M$. However, there is one $\textit{put}(d_1,\_)$ operation and no $\textit{rm}(d_1)$ operation in $h$, contradicts that $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
\end{itemize}

This completes the proof of the $\textit{only if}$ direction.

To prove the $\textit{if}$ direction, assume that $G$ has no cycle going through $x$. Let $O_L$ be the set of operations that happen before $\textit{put}(x)$ in $h$. It is easy to see that $O_L \subseteq \textit{LMSet}(h,x)$. Let $O_M = \textit{LMSet}(h,x) \setminus O_L$. Let $O_h$ be the set of operations of $h$, and let $O_R = O_h \setminus \textit{LMSet}(h,x)$. Let $L$, $M$ and $R$ be $O_L$, $O_M$ and $O_R$, respectively.

By Lemma \ref{lemma:LMSet has matched put and rm operations}, we can see that $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations. It remains to prove that $L$, $M$ and $R$ is a gap of $\textit{put}(x)$ and $\textit{rm}(x)$. We prove this by showing that all the following cases are impossible:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Case $1$: Some operation $o_r \in R$ happens before $\textit{rm}(x)$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.

\item[-] Case $2$: Some operation $o_r \in R$ happens before some operation $o_{lm} \in L \cup M$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.

\item[-] Case $3$: Some operation $o_r \in R$ happens before $\textit{put}(x)$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.

\item[-] Case $4$: $\textit{rm}(x)$ happens before some $o_{lm} \in \textit{LMSet}(h,x) = L \cup M$. By Lemma \ref{lemma:Rmx does not happen before LMSet for Rpr1} we know that this is impossible.

\item[-] Case $5$: $\textit{rm}(x)$ happens before $\textit{put}(x)$. This contradicts that for each item in $h$, its $\textit{rm}$ operation does not happen before its $\textit{put}$ operation.

\item[-] Case $6$: Some operation $o_m \in M$ happens before $\textit{put}(x)$. Then we know that $o_m \in L$, which contradicts that $o_m \in M$.

\item[-] Case $7$: Some operation $o_m \in M$ happens before some operation $o_l \in L$. Then we know that $o_m \in L$, which contradicts that $o_m \in M$.

\item[-] Case $8$: $\textit{put}(x)$ happens before some operation $o_l \in L$. This is impossible.
\end{itemize}

This completes the proof of the $\textit{if}$ direction.

\qed
\end {proof}

A automaton $\mathcal{A}_{\textit{Rpr1-1}}$ is given in \figurename~\ref{fig:automata for Rpro-1}, where $c=\textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm,a),$ $\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$. $\mathcal{A}_{\textit{Rpr1-1}}$ is used to detect the existence of a item whose $\textit{rm}$ happens before its $\textit{put}$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-1.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-1}}$}
  \label{fig:automata for Rpro-1}
\end{figure}

Let us begin to represent several automata that is used for capture the case that, in a sub-history $h'$ of a history $h$, $\textit{last}(h')=R_{\textit{pr1}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$, and the reason is that there is no gap on $\textit{put}(y,\textit{pri}_y)$ and $\textit{rm}(y)$ in $h'$, where $y$ is the item that has maximal priority in $h'$. Such case is dealt with by four automata, since there are four possibilities of $h' \vert_{y}$.

The first automaton is $\mathcal{A}_{\textit{Rpr1-pprr}}$, which is given in \figurename~\ref{fig:automata for Rpro-pprr}. In \figurename~\ref{fig:automata for Rpro-pprr}, $c_1= \textit{cal}(\textit{put},c,$ $\textit{anyPri}),\textit{ret}(\textit{put},c), \textit{cal}(\textit{rm},c), \textit{ret}(\textit{rm},c)$, $c_2 = c_1 + \textit{cal}(\textit{put},a,\textit{les}_p)$, and $c_3 = c_2 + \textit{ret}(\textit{rm},a)$. Here $p$ is a variable for priority, and $\textit{les}_p$ is a predicate that accepts all priorities that are less than the value of $p$. Here $\textit{pprr}$ represents that, $h' \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$ (we consider $b$ as $y$ and $p$ as $\textit{pri}_y$). In \figurename~\ref{fig:automata for Rpro-pprr}, the three paths that (1) from $q_{\textit{init}}$ to $q_2$ to $q_3$, (2) from $q_{\textit{init}}$ to $q_2$ to $q_{11}$, and (3) from $q_{\textit{init}}$ to $q_{10}$ to $q_{11}$ represents three possible positions of $\textit{ret}(\textit{put},a)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-pprr.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-pprr}}$}
  \label{fig:automata for Rpro-pprr}
\end{figure}

The second automaton is $\mathcal{A}_{\textit{Rpr1-prpr}}$, which is given in \figurename~\ref{fig:automata for Rpro-prpr}. In $\mathcal{A}_{\textit{Rpr1-prpr}}$, $c_1, c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$. Here $\textit{pprr}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{put},b,p) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-prpr.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-prpr}}$}
  \label{fig:automata for Rpro-prpr}
\end{figure}

The third automaton is $\mathcal{A}_{\textit{Rpr1-rppr}}$, which is given in \figurename~\ref{fig:automata for Rpro-rppr}. In $\mathcal{A}_{\textit{Rpr1-rppr}}$, $c_1, c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$. Here $\textit{rppr}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-rppr.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-rppr}}$}
  \label{fig:automata for Rpro-rppr}
\end{figure}

The forth automaton is $\mathcal{A}_{\textit{Rpr1-rprp}}$, which is given in \figurename~\ref{fig:automata for Rpro-rprp}. In $\mathcal{A}_{\textit{Rpr1-rprp}}$, $c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$, and $c_4=c_1 + \textit{ret}(put,b)$. Here $\textit{rprp}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b) \cdot \textit{ret}(\textit{put})$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-rprp.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-rprp}}$}
  \label{fig:automata for Rpro-rprp}
\end{figure}

Given a data-differentiated history $h$, we say that a item $x$ with priority $\textit{pri}_x$ in $h$ is covered by items $d_1,\ldots,d_m$ in $h$, if the priorities of $d_1,\ldots,d_m$ is not $\textit{pri}_x$, and

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{put}(d_m,\_)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$,

\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,

\item[-] $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$, or $\textit{rm}(d_1)$ does not exists in $h$
\end{itemize}

According to the definition of left-right constraint, in a data-differentiated history $h$, there is a cycle going through $x$, if and only if there exists items $d_1,\ldots,d_m$, such that $x$ is covered by $d_1,\ldots,d_m$.

\begin{restatable}{lemma}{Rpr1IsCoRegular}
\label{lemma:Rpr1 is co-regular}
$R_{\textit{pr1}}$ is co-regular.
\end{restatable}

\begin {proof}

Let $\textit{AutS-R}_{\textit{pr1}} = \{ \mathcal{A}_{\textit{Rpr1-1}} , \mathcal{A}_{\textit{Rpr1-pprr}}, \mathcal{A}_{\textit{Rpr1-prpr}}, \mathcal{A}_{\textit{Rpr1-rppr}}, \mathcal{A}_{\textit{Rpr1-rprp}} \}$. We need to prove that:

\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr1}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr1}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr1}}$.

By Lemma \ref{lemma:Gap Equals Lin for Rpr1} and Lemma \ref{lemma:Gap Equals Constraint for Rpr1}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:

\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr1}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h''$ is a projection of $h$, such that $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is in $h''$, $\textit{pri}_x$ is the maximal priority of $h''$, and one the the following case holds:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] case $1$: $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$,
\item[-] case $2$: $x$ is covered by some items $d_1,\ldots,d_m$ in $h''$.
\end{itemize}

\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\textit{AutS-R}_{\textit{pr1}}$. By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-1}}$: Let $d_1$ a item in $h_2$ such that $r(d_1)=b$ and let $h'' = h_2 \vert_{ \{ d_1 \} }$. $d_1$ is the item with maximal priority in $h''$. It is obvious that $\textit{last}(h'') = R_{\textit{pr1}}$ and $\textit{rm}(d_1)$ happens before $\textit{put}(d_1,\_)$ in $h''$.

\item[-] If $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-pprr}}$, $\mathcal{A}_{\textit{Rpr1-prpr}}$, $\mathcal{A}_{\textit{Rpr1-rppr}}$, or $\mathcal{A}_{\textit{Rpr1-rprp}}$: Let $d_0$ be a item in $h_2$ such that $r(d_0)=b$, let $d_1,\ldots,d_m$ be the items in $h_2$ such that $r(d_i)=a$ for each $1 \leq i \leq m$. Let $h'' = h_1 \vert_{ \{ d_0, d_1, \ldots, d_m \} }$. It is obvious that the priority of $d_0$ is larger than the priorities of other operations in $h''$, and then $\textit{last}(h'') = R_{\textit{pr1}}$. It is easy to see that $d_0$ is covered by $d_1,\ldots,d_m$.
\end{itemize}

\noindent The $\textit{if}$ direction: Assume that there exists such $h$, $h''$, $x$ and $\textit{pri}_x$. Since $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{pri}_x$ is larger than the priorities of other operations in $h''$. Then,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If case $1$ holds: Then, $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$ in $h''$. Let $h_1$ be obtained from $h$ by renaming $x$ to $b$ and renaming all other items to $a$. By data-independence, $h_1 \in \mathcal{I}$, and it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-1}}$.
\item[-] case $2$: Then, $x$ is covered by some items $d_1,\ldots,d_m$ in $h''$. There are four possibilities of $h'' \vert_{ \{ x \} }$: (1) $\textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{rm},x)$, (2) $\textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},x)$, (3) $\textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},x)$ and (4) $\textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{rm},x) \cdot \textit{ret}(\textit{put})$.

    Let $h_1$ be obtained from $h$ by renaming $x$ into $b$, renaming $d_1,\ldots,d_m$ into $a$ and renaming other items into $c$. We can prove that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-pprr}}$, $\mathcal{A}_{\textit{Rpr1-prpr}}$, $\mathcal{A}_{\textit{Rpr1-rppr}}$ or $\mathcal{A}_{\textit{Rpr1-rprp}}$.

    For example, if $h'' \vert_{ \{ x \} } = \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{rm},x)$, and $\textit{ret}(\textit{rm},d_m)$ is before $\textit{call}(\textit{put},x,\textit{pri}_x)$ and is after $\textit{ret}(\textit{put},x)$ in $h''$. Then in  $\mathcal{A}_{\textit{Rpr1-pprr}}$, $h_1$ is accepted by transitions from $q_{\textit{init}}$ to $q_2$ to $q_{11}$ and finally to $q_7$ (if $\textit{rm}(d_1)$ does not exists in $h''$) or $q_9$ (if $\textit{rm}(d_1)$ exists in $h''$).
\end{itemize}

This completes the proof of this lemma. \qed
\end {proof}

\subsection{Co-Regular of $R_{\textit{pr2}}$}
\label{subsec:co-regular of Rpr2}

An automaton $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$ is given in \figurename~\ref{fig:automata for FIFO of single priority}. In \figurename~\ref{fig:automata for FIFO of single priority}, let $c_1= \textit{cal}(\textit{put},c,\textit{anyPri}),\textit{ret}(\textit{put},c)$, $\textit{cal}(\textit{rm},c),\textit{ret}(\textit{rm},c)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, and $c_3 = c_2 + \textit{cal}(\textit{rm},a) + \textit{ret}(\textit{rm},a)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_QUEUE_FOR_SINGLE_PRI.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$}
  \label{fig:automata for FIFO of single priority}
\end{figure}

According to \cite{Bouajjani:2015}, given a data-independent set $S$ of queue executions, whether every sequence $s$ in $S$ is linearizable to queue can be reduced into checking emptiness of intersection of $S$ and several finite automata. Let $\textit{transToQueue}(h)$ be a sequence $h'$ that is generated from $h$ by transforming $\textit{put}$ and $\textit{rm}$ into $\textit{enq}$ and $\textit{deq}$, respectively, and then discarding priorities. A priority queue history is called single-priority, if all of its items has only one priority. The following lemma states that we can also use finite automata to check ``first in first out property'' for single-priority sub-sequence of priority queue history.

\begin{restatable}{lemma}{AutoForPQwithSignlePri}
\label{lemma:automata for priority queue with single priority}

Given a data-independent implementations $\mathcal{I}$ of priority queue, let $\mathcal{A} = \{ \mathcal{A}_{\textit{unmRm1}}, \mathcal{A}_{\textit{Rpr0-1}}, \mathcal{A}_{\textit{unmRm2}}, \mathcal{A}_{\textit{FIFO-Sig-Pri}} \}$, then,

$\mathcal{I} \cap \mathcal{A} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority, and $\textit{transToQueue}(e')$ does not linearizable to queue.
\end{restatable}

\begin {proof}

According to \cite{Bouajjani:2015}, checking linearizable with respect to queue can be reduced into checking the following three conditions for data-differentiated histories:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{deq}(a)$ happens before $\textit{enq}(a)$ (or $\textit{enq}(a)$ does not exist) for some item $a$,

\item[-] there are two or more $\textit{deq}(a)$ for some item $a$,

\item[-] $\textit{enq}(a)$ happens before $\textit{enq}(b)$, and $\textit{deq}(b)$ happens before $\textit{deq}(a)$ (or $\textit{deq}(a)$ does not exists) for some items $a,b$.
\end{itemize}

It is not hard to see that (1) the first condition is dealt with by $\mathcal{A}_{\textit{unmRm1}}$ (if $\textit{put}(a,\_)$ does not exist) and $\mathcal{A}_{\textit{Rpr0-1}}$ (if $\textit{put}(a,\_)$ exists), (2) the second condition is dealt with by $\mathcal{A}_{\textit{unmRm2}}$, and (3) the third condition is dealt with by $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$. \qed
\end {proof}

By Lemma \ref{lemma:automata for priority queue with single priority}, from now on, it is safe for us to assume that, for each data-differentiated priority queue history $h$ and each $h'$ of projection of $h$, if $h'$ is single-priority, then $\textit{transToQueue}(h')$ is linearizable to queue. One consequence of this assumption is that, in $h'$, there does not exists an item $a$, such that $\textit{rm}(a)$ happens before $\textit{put}(a)$.

By the proof of Lemma \ref{lemma:Rpr1 is co-regular}, with the help of $\mathcal{A}_{\textit{Rpr1-pprr}}, \mathcal{A}_{\textit{Rpr1-prpr}}, \mathcal{A}_{\textit{Rpr1-rppr}}$ and $\mathcal{A}_{\textit{Rpr1-rprp}}$, from now on, we can safely assume that given a data-differentiated history $h$ and operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ in $h$, there does not exists any items $d_1,\ldots,d_m$, such that $x$ is not covered by items $d_1,\ldots,d_m$ and the priority of $d_1,\ldots,d_m$ is less that $x$.

However, above two assumptions are not enough for imply the linearizability to priority queue. One example is shown in \figurename~\ref{fig:history introduct ob order1}. The history $h$ of \figurename~\ref{fig:history introduct ob order1} is not linearizable, even if $h \vert_{1}$ and $h \vert_{2}$ are both linearizable, and either $a$ or $b$ is covered by other items. One explanation is that, since $\textit{put}(b,1)$ happens before $\textit{put}(a,1)$, in choosing linearization points, we should make $\textit{rm}(b)$ before $\textit{rm}(a)$. However, to satisfy the requirements of $R_{\textit{pr2}}$, the linearization points of $\textit{rm}(a)$ can only locates in the interval from $\textit{ret}(\textit{rm},d)$ to $\textit{cal}(\textit{put},e,2)$. But, $\textit{rm}(b)$ is after this entire interval, and we can not locate the linearization point of $\textit{rm}(b)$ in this interval. When there is no such phenomenon, such as the history in \figurename~\ref{fig:history introduct ob order2}, the history becomes linearizable.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER1.pdf}
%\vspace{-10pt}
  \caption{above assumptions do not implies linearizability (1)}
  \label{fig:history introduct ob order1}
\end{figure}

%\begin{figure}[htbp]
%  \centering
%  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER3.pdf}
%\vspace{-10pt}
%  \caption{above assumptions do not implies linearizability (2)}
%  \label{fig:history introduct ob order1}
%\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER2.pdf}
%\vspace{-10pt}
  \caption{above assumptions do not implies linearizability (2)}
  \label{fig:history introduct ob order2}
\end{figure}

In the following part of this subsection, we use $<_{\textit{ob}}$ order for the case when we can infer that an item is inserted earlier than another item, and we use gap-point to represent the intervals when it is possible to locate linearizaton points of $\textit{rm}(a)$.

\begin{definition}[gap-point for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:gap-point for matched put and rm operations}

Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x),\textit{rm}(x)$ of $h$, we say $o$ is a gap-point of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$, if we can obtain a sequence $\textit{gapPSeq}(h)$ from $h$ by adding $o$ to some locations between actions of $h$, and satisfies that,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] In $\textit{gapPSeq}(h)$, $o$ is after $\textit{call}(\textit{put},x,\textit{pri}_x)$ and $\textit{call}(\textit{rm},x)$, and is before $\textit{ret}(\textit{rm},x)$.
\item[-] For each item $y$ with non-$\textit{pri}_x$ priority in $\textit{gapPSeq}(h)$. If $\textit{put}(y,\_)$ and $\textit{rm}(y)$ is in $h$, then it can not be the case that $\textit{ret}(\textit{put},y,\_)$ is before $o$ and $\textit{call}(\textit{rm},x)$ is after $o$; Else if $\textit{put}(y,\_)$ is in $h$ and $\textit{rm}(y)$ is not in $h$, then it can not be the case that $\textit{ret}(\textit{put},y,\_)$ is before $o$;
\end{itemize}

We say that $o$ is a gap-point of priority $\textit{pri}_x$, if we can obtain a sequence $\textit{gapPSeq}(h)$ from $h$ by adding $o$ to some locations between actions of $h$, and satisfies only the latter requirement.
\end{definition}

\begin{restatable}{lemma}{GapEqualsGapPoint}
\label{lemma:gap eqals the existence of gap point}
Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$. There is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, if and only if there is a gap-point $o$ of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
\end{restatable}

\begin {proof}

To prove the $\textit{only if}$ direction, we need to prove that: if there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, then there is no gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.

We prove this by contradiction. Assume that there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, but there is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$. Since there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, in the time interval starts from $\textit{cal}(\textit{put},x,\textit{pri}_x)$ and $\textit{cal}(\textit{rm},x)$ and ends at $\textit{ret}(\textit{rm},x)$, there is always some item with non-$\textit{pri}_x$ priority in priority queue. Therefore, it is not hard to prove that there is a cycle going through the left-right constraint of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, and by Lemma \ref{lemma:Gap Equals Constraint for Rpr0}, there is no gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, which contradicts that there is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.

To prove the $\textit{if}$ direction, assume that there is a gap-point $o$ of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$ and let $h_1$ be the sequence in definition of gap-point, which is generated from $h$ by adding $o$ into $h$. Let sequence $h_2$ be a sequence of call, return and operations, such that the projection of $h_2$ into call and return actions is $h$. $h_2$ is generated as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] First, from $h_1$, transform the point $o$ into $\textit{rm}(x)$.

\item[-] Then for all the other items $y$, (1) adding an operation $\textit{put}(y,p)$ at some position between $\textit{call}(\textit{put},y,\_)$ and $\textit{ret}(\textit{put},y)$, and (2) adding an operation $\textit{rm}(y)$ at some position between $\textit{call}(\textit{rm},y)$ and $\textit{ret}(\textit{rm},y)$.

\item[-] For all items $z$ with the same priority as $x$, the positions for adding operations of $z$ into $h$ is arbitrary (but need to satisfy above restriction. In the following we implicitly require this).

\item[-] For all items $z$ whose priority is different from that of $x$,

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}

    \item[-] If $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, and $\textit{ret}(\textit{put},z,\_)$ is before $o$ in $h_1$: Then in $h_2$, $\textit{rm}(z,\_)$ is just after $\textit{call}(\textit{rm},z)$.

    \item[-] If $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, and $o$ is before $\textit{ret}(\textit{put},z,\_)$ $h_1$: Then in $h_2$, $\textit{put}(z,\_)$ is just before $\textit{ret}(\textit{put},z)$.

    \item[-] Otherwise, we can see that $\textit{put}(z,\_)$ and $\textit{rm}(z)$ overlaps in $h$. Then we add $o_1=\textit{put}(z,\_)$ and $o_2=\textit{rm}(z)$ in the time interval of interleaving of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ overlaps in $h$, make sure $o_2$ next to $o_1$, and make sure either both of them is before $\textit{rm}(x)$ or both of them is after $\textit{rm}(x)$ in $h_2$.
    \end{itemize}
\end{itemize}

According to the definition of gap-point, it is not hard to see that we can construct such $h_2$. Let $l$ be the projection of $h_2$ into operations, and $l'$ be the projection of $l$ into $x$ and all items that has priorities not equals that of $x$. Let $l'= u \cdot \textit{put}(x,\_) \cdot v \cdot \textit{rm}(x) \cdot w$. It is easy to see that $h \sqsubseteq l$, and $u \cdot v$ contains matched $\textit{put}$ and $\textit{rm}$ operations. Let $L$, $M$ and $R$ be the set of operations in $u$, $v$ and $w$, respectively. Then it is obvious that $L$, $M$ and $R$ is a gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$. \qed
\end {proof}

Given a data-differentiated history $h$ and two items $a,b$, we say that $a <_{\textit{ob}} b$, if the priority of $a$ and $b$ is the same, and one of the following cases holds:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Type $A$: $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$ in $h$,

\item[-] Type $B$: $\textit{rm}(a)$ happens before $\textit{rm}(b)$ in $h$,

\item[-] Type $C$: $\textit{rm}(a)$ happens before $\textit{put}(b,\_)$ in $h$,
\end{itemize}

Let $<_{\textit{ob}}^*$ be the transitive closure of $<_{\textit{ob}}$. Intuitively, $a <_{\textit{ob}} b$ means that $a$ should be inserted earlier than $b$, and $a <_{\textit{ob}}^* b$ means that we can infer that $a$ should be inserted earlier than $b$ with the help of other items. We can explicitly write $a <_{\textit{ob}}^A b$ to mean that $a <_{\textit{ob}} b$ and the reason is that $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$. Similarly, we can explicitly write $a <_{\textit{ob}}^B b$ and $a <_{\textit{ob}}^C b$ to mean that $a <_{\textit{ob}} b$ and the reason is that $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$ and $\textit{rm}(a)$ happens before $\textit{put}(b,\_)$, respectively.

According to the definition of $<_{\textit{ob}}^*$, if $a <_{\textit{ob}}^* b$, then there exists $c_1,\ldots,c_n$, such that $a <_{\textit{ob}} c_1 <_{\textit{ob}} \ldots <_{\textit{ob}} b$. The following lemma states that, the number of intermediate items $c_i$ is in fact bounded.

\begin{restatable}{lemma}{OBOrderHasBoundedLength}
\label{lemma:ob order has bounded length}
Given a data-differentiated history $h$. Assume that $a <_{\textit{ob}} a_1 <_{\textit{ob}} \ldots <_{\textit{ob}} a_m <_{\textit{ob}} b$, then either $a <_{\textit{ob}} b$, or $a <_{\textit{ob}} a_i <_{\textit{ob}} b$ for some $i$, or $a <_{\textit{ob}} a_i <_{\textit{ob}} a_j <_{\textit{ob}} b$ for some $i,j$.
\end{restatable}

\begin {proof}

Our proof proceed as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^A$,$<_{\textit{ob}}^B \cdot <_{\textit{ob}}^B$ and $<_{\textit{ob}}^C \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^A c_2 <_{\textit{ob}}^A c_1$, then $\textit{put}(c_3,\_)$ happens before $\textit{put}(c_2,\_)$, and $\textit{put}(c_2,\_)$ happens before $\textit{put}(c_1,\_)$. Therefore, it is obvious hat $\textit{put}(c_3,\_)$ happens before $\textit{put}(c_1,\_)$ and $c_3 <_{\textit{ob}}^A c_1$.

    Similarly, if $c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^B c_1$, then $c_3 <_{\textit{ob}}^B c_1$.

    If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^C c_1$: Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$. Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$. Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$. Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.

    Therefore, when we meet successive $<_{\textit{ob}}^A$, it is safe to leave only the first and the last elements and ignore intermediate elements. Similar cases hold for $<_{\textit{ob}}^B$ and $<_{\textit{ob}}^C$.

\item[-] $<_{\textit{ob}}^A$ and $<_{\textit{ob}}^C$:

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^A c_2 <_{\textit{ob}}^C c_1$. Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$. Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$. Since $c_3 <_{\textit{ob}}^A c_2$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$. Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^A c_1$.

    \item[-] ($<_{\textit{ob}}^C \cdot <_{\textit{ob}}^A$): If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^A c_1$. Since $c_2 <_{\textit{ob}}^A c_1$, $\textit{put}(c_2,\_)$ happens before $\textit{put}(c_1,\_)$. Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{rm}(c_3)$ happens before $\textit{put}(c_2,\_)$. Therefore, $\textit{rm}(c_3)$ happens before $\textit{put}(c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
    \end{itemize}

\item[-] $<_{\textit{ob}}^B$ and $<_{\textit{ob}}^C$:

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] ($<_{\textit{ob}}^B \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^C c_1$. Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$. It is obvious that $\textit{cal}(\textit{rm},c_2)$ is before $\textit{ret}(\textit{rm},c_2)$. Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$. Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.

    \item[-] ($<_{\textit{ob}}^C \cdot <_{\textit{ob}}^B$): If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^B c_1$. Since $c_2 <_{\textit{ob}}^B c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{rm},c_1)$. Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$. Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$. Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_1)$, and $c_3 <_{\textit{ob}}^B c_1$.
    \end{itemize}

\item[-]  ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^B \cdot <_{\textit{ob}}^A$): If $c_4 <_{\textit{ob}}^A c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^A c_1$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$: Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$. Then $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$. This implies that $c_4 <_{\textit{ob}}^A c_3 <_{\textit{ob}}^C c_1$. According to the fact for $<_{\textit{ob}}^A \cdot <_{\textit{ob}}^C$, we know that $c_4  <_{\textit{ob}}^A c_1$.

    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is after $\textit{cal}(\textit{put},c_1,\_)$: Since $c_2 <_{\textit{ob}}^A c_1$, $\textit{ret}(\textit{put},c_2,\_)$ is before $\textit{cal}(\textit{put},c_1,\_)$. Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{rm}(c_3)$ happens before $\textit{rm}(c_2)$, and then we know that $\textit{put}(c_2,\_)$ can not happen before $\textit{put}(c_3,\_)$. Since $\textit{put}(c_2,\_)$ does not happen before $\textit{put}(c_3,\_)$, $\textit{cal}(\textit{put},c_3,\_)$ is before $\textit{ret}(\textit{put},c_2,\_)$. Since $c_4 <_{\textit{ob}}^A c_3$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_3,\_)$. Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_4 <_{\textit{ob}}^A c_1$.
    \end{itemize}

\end{itemize}

Based on above results, given $a <_{\textit{ob}}^{b_1} a_1 <_{\textit{ob}} \ldots <_{\textit{ob}}^{b_m} a_m <_{\textit{ob}}^{b_{\textit{m+1}}} b$, where each $b_i$ is in $\{ A,B,C \}$, we can merge relations, until we got one of the following facts:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $a <_{\textit{ob}}^A b$, $a <_{\textit{ob}}^B b$ or $a <_{\textit{ob}}^C b$,

\item[-] $a <_{\textit{ob}}^A a_i <_{\textit{ob}}^B b$, or $a <_{\textit{ob}}^B a_i <_{\textit{ob}}^A b$, for some $i$,

\item[-] $a <_{\textit{ob}}^B a_i <_{\textit{ob}}^A a_j <_{\textit{ob}}^B b$, for some $i$ and $j$,
\end{itemize}

This completes the proof of this lemma. \qed
\end {proof}

Given a differentiated history $h$ and operations $\textit{put}(x,\_)$ and $\textit{rm}(x)$ of $h$, the time interval from the time point when both $\textit{call}(\textit{put},x,\_)$ and $\textit{cal}(\textit{rm},x)$ happen and ends at $\textit{ret}(\textit{rm},x)$ is called the effective interval of $x$ in $h$.

\begin{restatable}{lemma}{Rpr2AsGPandOBOrder}
\label{lemma:Rpr2 as gap-point and ob order}
Given a data-differentiated history $h$. There exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr2}}$, and $h'$ does not linearizable to $\textit{MR}_{\textit{pr2}}$, if and only if there exists $h'' \in \textit{proj}(h)$, $\textit{last}(h'') = R_{\textit{pr2}}$, $x$ and $y$ has maximal priority in $h''$, $y <_{\textit{ob}} x$ in $h''$, and the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$ in $h''$.
\end{restatable}

\begin {proof}

To prove the $\textit{if}$ direction, let $h_1$ be the history that is obtained from $h''$ by erasing all operations of items that has same priority as $x$, except for operations of $x$ and $y$. It is obvious that $\textit{last}(h_1) = R_{\textit{pr2}}$.
According to $y <_{\textit{ob}} x$,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $y <_{\textit{ob}}^A x$, then $\textit{put}(y,\_)$ happens before $\textit{put}(x,\_)$. Since in $R_{\textit{pr2}}$, $\textit{put}(\textit{itm},\_)$ is after $\textit{put}$ operations of items with maximal priority (except for $\textit{itm}$), we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
\item[-] If $y <_{\textit{ob}}^B x$, then $\textit{rm}(y)$ happens before $\textit{rm}(x)$. Since in $R_{\textit{pr2}}$, $\textit{rm}(\textit{itm})$ is the last $\textit{rm}$ operation of items with maximal priority, we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.

\item[-] If $y <_{\textit{ob}}^C x$, then $\textit{rm}(y)$ happens before $\textit{put}(x,\_)$. Since in $R_{\textit{pr2}}$, $\textit{rm}(\textit{itm})$ is the last $\textit{rm}$ operation of items with maximal priority, we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
\end{itemize}

Similarly as the proof in Lemma \ref{lemma:gap eqals the existence of gap point}, we can see that the only possible position for locating linearizaton point of $\textit{rm}(x)$ is at gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$. Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, then there exists unmatched operations before $\textit{rm}(x)$.

Since the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},x)$ in $h''$ (and also in $h_1$), if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, then $\textit{rm}(x)$ will be before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},x)$.

Therefore, for every sequence $l = u \cdot \textit{put}(x,\_) \cdot v \cdot \textit{rm}(x) \cdot w$, if $h_1 \sqsubseteq l$, then either $u \cdot v$ contains some unmatched operations of priority less than priority of $x$, or $w$ contains $\textit{put}(y,\_)$ or $\textit{rm}(y)$. In both cases, $l \notin \textit{MR}_{\textit{pr2}}$.

To prove the $\textit{only if}$ direction, we prove its contrapositive. Assume we already know that for each $h'' \in \textit{proj}(h)$ with $\textit{last}(h'') = R_{\textit{pr2}}$, for each $x$ and $y$ has maximal priority in $h''$, if $y <_{\textit{ob}} x$ in $h''$, then the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},x)$ in $h''$. We need to prove that for each $h' \in \textit{proj}(h)$, if $\textit{last}(h') = R_{\textit{pr2}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr2}}$.

Let $\textit{pri}$ be the maximal priority of $h'$. Recall that we already assume that each single-priority history (if we consider $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$) can be ``linearizable to queue'', and item with larger priority is not covered by items with smaller priority. Our proof proceed as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Let $h_{\textit{pri}}$ be the projection of $h$ into operations of priority $\textit{pri}$. Since each single-priority history can be ``linearizable to queue'', there exists sequence $l_{\textit{pri}}$, such that $h_{\textit{pri}} \sqsubseteq l_{\textit{pri}}$, and when we treat $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$, $l_{\textit{pri}}$ belongs to queue.

\item[-] Step $1$: Let $a_1$ be the last inserted item of $l_{\textit{pri}}$.

$\textit{Cond1}$: If for each item $b$ with priority $\textit{pri}$ in $h'$, the rightmost gap-point $o$ of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$ does not before any $\textit{cal}(\textit{put},b,\textit{pri})$ or $\textit{cal}(\textit{rm},b)$ in $h'$, then we can

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] Locate the linearization point of $\textit{rm}(a_1)$ at $o$,
    \item[-] If $\textit{put}(a_1,\textit{pri})$ overlaps with $\textit{rm}(a_1)$, then locate the linearization point of $\textit{put}(a_1,\_)$ just before the linearization point of $\textit{rm}(a_1)$. Otherwise, $\textit{put}(a_1,\textit{pri}) <_{\textit{hb}} \textit{rm}(a_1)$, and we locate the linearization point of $\textit{put}(a_1,\_)$ just before its return action.
    \item[-] Locate linearization points of operations of items of priority $\textit{pri}$ (except for $a_1$) just after the call action of the operation.
    \item[-] For item $z$ with priority smaller than $\textit{pri}$. If both $\textit{cal}(\textit{put},z,\_)$ and $\textit{cal}(\textit{rm},z)$ is before $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just after their call actions. If both $\textit{ret}(\textit{put},z)$ and $\textit{ret}(\textit{rm},z)$ is after $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just before their return actions.

        Otherwise, either (1) $\textit{rm}(z) <_{\textit{hb}} \textit{put}(z,\_)$, or (2) $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, $\textit{ret}(\textit{put},z,\_)$ before $o$, and $o$ before $\textit{call}(\textit{rm},z)$. The first case violates that each single-priority history can be ``linearizable to queue'', and the second case violates that $o$ is a gap-point.
    \end{itemize}

    In this way we construct a sequence $l$, and it is easy to see that $h' \sqsubseteq l$ and $l$ satisfies the requirement of $R_{\textit{pr2}}$.

\item[-] Otherwise, there exists $a_2$ with priority $\textit{pri}$, such that the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$ in $h'$. We can see that the rightmost gap-point of $\textit{put}(a_2,\textit{pri})$ and $\textit{rm}(a_2)$ is after the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$. By assumption we know that $a_2$ does not $<_{\textit{ob}}$ to $a_1$.

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If for each item $b$ with priority $\textit{pri}$ in $h'$, $a_2$ does not $<_{\textit{ob}}$ to $b$. Then we go to step $1$ and treat $a_2$ as $a_1$ (When $\textit{Cond1}$ of step $1$ holds, we break the loop).
    \item[-] Otherwise, there exists $a_3$ with priority $\textit{pri}$ such that $a_2 <_{\textit{ob}} a_3$. By assumption,we know that the rightmost gap-point of $\textit{put}(a_3,\textit{pri})$ and $\textit{rm}(a_3)$ is not before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$. Then we can see that the rightmost gap-point of $\textit{put}(a_3,\textit{pri})$ and $\textit{rm}(a_3)$ is after the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$. Then we go to step $1$ and treat $a_3$ as $a_1$ (When $\textit{Cond1}$ of step $1$ holds, we break the loop).
    \end{itemize}
\end{itemize}

It is not hard to see that, if $b_i$ is the $a_1$ in the $\textit{i-th}$ loop of our proof, $b_j$ is the $a_1$ in the $\textit{j-th}$ loop of our proof, and $i<j$, then the rightmost gap-point of $\textit{put}(b_j,\textit{pri})$ and $\textit{rm}(b_j)$ is after the rightmost gap-point of $\textit{put}(b_i,\textit{pri})$ and $\textit{rm}(b_i)$. Thus, we finally stop at some $d$. $d$ satisfies that, for each item $c$ with priority $\textit{pri}$, $d$ does not $<_{\textit{ob}}$ to $c$. $d$ also satisfies that, for each item $c$ with priority $\textit{pri}$ in $h'$, the rightmost gap-point $o'$ of $\textit{put}(d,\textit{pri})$ and $\textit{rm}(d)$ does not before any $\textit{cal}(\textit{put},c,\textit{pri})$ or $\textit{cal}(\textit{rm},c)$. Then, we can

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Locate the linearization point of $\textit{rm}(d)$ at the rightmost gap-point $o'$ of $\textit{put}(d,\textit{pri})$ and $\textit{rm}(d)$,

\item[-] Since for each item $c$ with priority $\textit{pri}$ in $h'$, $o'$ is after any $\textit{cal}(\textit{put},c,\textit{pri})$ and $\textit{cal}(\textit{rm},c)$. For each $c$, one of the following cases hold: (1) $\textit{cal}(\textit{rm},c)$ is before $o'$ and $o'$ is before $\textit{ret}(\textit{rm},c)$, (2) $\textit{cal}(\textit{rm},c)$ and $\textit{ret}(\textit{rm},c)$ are before $o'$. We locate the linearization point of $\textit{rm}(c)$ just after its call action. Such linearization point is obviously before $o'$.

\item[-] If $o'$ is before $\textit{ret}(\textit{put},d,\textit{pri})$, then locate the linearization point of $\textit{put}(d,\textit{pri})$ just before $o'$. Else, $\textit{ret}(\textit{put},d,\textit{pri})$ is after $o'$, and we locate the linearization point of $\textit{put}(d,\textit{pri})$ just before the return action of it.

\item[-] If $o'$ is before $\textit{ret}(\textit{put},d,\textit{pri})$, we can see that, for each item $c$ with priority $\textit{pri}$ in $h'$, $\textit{cal}(\textit{put},c,\textit{pri})$ is before $o'$. Then, and we locate the linearization point of $\textit{put}(c,\textit{pri})$ just before $o'$ and the linearization point of $\textit{put}(d,\textit{pri})$.

    Otherwise, $\textit{ret}(\textit{put},d,\textit{pri})$ is before $o'$. Since $d$ does not $<_{\textit{ob}}$ to $c$, $\textit{put}(d,\textit{pri})$ does not happen before $\textit{put}(c,\textit{pri})$, or we can say, $\textit{ret}(\textit{put},d,\textit{pri})$ is after $\textit{cal}(\textit{put},c,\textit{pri})$. Then, we locate the linearization point of $\textit{put}(c,\textit{pri})$ just after its call action.
\end{itemize}

In this way we construct a sequence $l_d$, and it is easy to see that $h' \sqsubseteq l_d$ and $l_d$ satisfies the requirement of $R_{\textit{pr2}}$ with $d$ acts as $\textit{itm}$. This completes the proof of $\textit{if}$ direction. \qed
\end {proof}

Let us begin to represent several automata that is used for capture the case that, in a sub-history $h''$ of $h$, $\textit{last}(h'') = R_{\textit{pr2}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr2}}$, and the reason is that there exists items $a$ and $b$ with maximal priority in $h''$, $a <_{\textit{ob}} b$ in $h''$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.

According to Lemma \ref{lemma:ob order has bounded length}, if $a <_{\textit{ob}} b$, then one of the following cases hold: (1) $a <_{\textit{ob}}^A b$, (2) $a <_{\textit{ob}}^B b$, (3) $a <_{\textit{ob}}^C b$, (4) $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$ for some $a_1$, (5) $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$ for some $a_1$, or (6) $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$ for some $a_1$ and $a_2$. Although there are many possible enumeration of $a$, $b$, $a_1$ and $b_1$, the following lemma states that we need only to consider five enumeration of operations of $a$, $b$, $a_1$ and $b_1$.

\begin{restatable}{lemma}{FiveEnmuerationisEnoughForRpr2}
\label{lemma:five enumeration is enough for Rpr2}
Given a data-differentiated history $h$, $\textit{last}(h) = R_{\textit{pr2}}$, and $h$ does not linearizable to $\textit{MR}_{\textit{pr2}}$. Assume that there exists items $a$ and $b$ with maximal priority in $h''$, $a <_{\textit{ob}} b$, and the rightmost gap-point $o$ of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$. Let $a_1$ and $a_2$ be the value when possibly $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$, or $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$. Then, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists) and $a_2$ (if exists).
\end{restatable}
\begin {proof}

Let us prove by consider all the possible reason of $a <_{\textit{ob}}^A b$. According to Lemma \ref{lemma:ob order has bounded length}, we need to consider six reasons: %There may exists items with smaller priority of $a$ and covers some time interval, and assume that we rename them into $d$:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Reason $1$, $a <_{\textit{ob}}^A b$:

    Since $a <_{\textit{ob}}^A b$, $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$. Since $o$ is after $\textit{cal}(\textit{put},a,\_)$, we can see that $o$ is before $\textit{cal}(\textit{rm},b)$.

    Since single-priority must satisfy the first in firs out property, $\textit{rm}(b)$ does not happen before $\textit{rm}(a)$, and $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$. If $\textit{ret}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$, since we already assume that there exists gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$, this gap-point is also a gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$, and is after $o$, which contradicts that $o$ is the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$. Therefore, $\textit{ret}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$.

    According to above discussion, there are two possible enumeration of operations of $a$ and $b$, as shown in \figurename~\ref{fig:history enumeration 1 for Rpr2} and \figurename~\ref{fig:history enumeration 2 for Rpr2}. Here we explicitly draw the leftmost gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$ as $o'$. Since the position of $\textit{ret}(\textit{put},b)$ does not influence the correctness, we can simply ignore it.

\item[-] Reason $2$, $a <_{\textit{ob}}^B b$:

    Since $a <_{\textit{ob}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$. Since $o$ is after $\textit{cal}(\textit{rm},b)$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$. Then, we can see that $o$ is before $\textit{cal}(\textit{put},a,\_)$. This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\_)$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible. Therefore, we can safely ignore this reason.

\item[-] Reason $3$, $a <_{\textit{ob}}^C b$:

    Since $a <_{\textit{ob}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},b)$. Since $o$ is after $\textit{cal}(\textit{put},b)$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$. Then, we can see that $o$ is before $\textit{cal}(\textit{put},a,\_)$. This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\_)$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible. Therefore, we can safely ignore this reason.

\item[-] Reason $4$, $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$:

    Since $a_1 <_{\textit{ob}}^B b$, $\textit{rm}(a_1) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_1)$ is before $\textit{cal}(\textit{rm},b)$. Since $\textit{rm}(a_1)$ does not happen before $\textit{put}(a_1)$, $\textit{cal}(\textit{put},a_1,\_)$ is before $\textit{ret}(\textit{rm},a_1)$. Since $a <_{\textit{ob}}^A a_1$, $\textit{ret}(\textit{put},a,\_)$ is before $\textit{cal}(\textit{put},a_1,\_)$. Therefore, $\textit{ret}(\textit{put},a,\_)$ is before $\textit{cal}(\textit{rm},b)$. Since $\textit{cal}(\textit{rm},b)$ is before $o$, we can see that $\textit{ret}(\textit{put},a,\_)$ is before $o$. Then, we can see that $o$ is before $\textit{cal}(\textit{rm},a)$.

    If $\textit{cal}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$, then $\textit{rm}(b) <_{\textit{hb}} \textit{rm}(a)$. Let $h' = h \vert_{ \{ a,a_1,b \} }$. It is not hard to see that $h'$ can not linearizable to a sequence that satisfies the first in first our property, and this contradicts to our assumption that each single-priority history satisfies the first in first out property. Therefore, $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$. Similarly as the case of reason $1$, we can see that $\textit{ret}(\textit{rm},b)$ is before $\textit{ret}(\textit{rm},a)$.

    According to above discussion, there are three possible enumeration of operations of $a$, $a_1$ and $b$, as shown in \figurename~\ref{fig:history enumeration 3 for Rpr2}, \figurename~\ref{fig:history enumeration 4 for Rpr2} and \figurename~\ref{fig:history enumeration 5 for Rpr2}. Here we explicitly draw the leftmost gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$ as $o'$. Since the position of $\textit{cal}(\textit{put},b,\_)$, $\textit{ret}(\textit{put},b)$, $\textit{ret}(\textit{put},a_1,\_)$ and $\textit{cal}(\textit{put},a,\_)$ do not influence the correctness, we can simply ignore it.

\item[-] Reason $5$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$:

    Since $a_1 <_{\textit{ob}}^A b$, $\textit{ret}(\textit{put},a_1,\_)$ is before $\textit{call}(\textit{put},b,\_)$. Since $\textit{call}(\textit{put},b,\_)$ is before $o$, we can see that $\textit{ret}(\textit{put},a_1,\_)$ is before $o$.

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o$ is before $\textit{cal}(\textit{rm},a)$: Then $o$ is obviously before $\textit{ret}(\textit{rm},a)$. Since $a <_{\textit{ob}}^B a_1$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},a_1)$. Then we can see that, $o$ is before $\textit{cal}(\textit{rm},a_1)$, and remember that $a_1 <_{\textit{ob}}^A b$. Then we can goto the case of reason $1$ and treat $a_1$ as $a$. Therefore, we can safely ignore this.

    \item[-] If $o$ is before $\textit{cal}(\textit{put},a,\_)$: Since $\textit{rm}(a)$ does not happen before $\textit{put}(a,\_)$, we can see that $\textit{cal}(\textit{put},a,\_)$ is before $\textit{ret}(\textit{rm},a)$, and then $o$ is before $\textit{ret}(\textit{rm},a)$. Then similarly as above case, we can see that $o$ is before $\textit{cal}(\textit{rm},a_1)$, and $a_1 <_{\textit{ob}}^A b$. Then we can goto the case of reason $1$ and treat $a_1$ as $a$. Therefore, we can safely ignore this.
    \end{itemize}

\item[-] Reason $6$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$:

    Since $a_2 <_{\textit{ob}}^B b$, $\textit{rm}(a_2) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_2)$ is before $\textit{cal}(\textit{rm},b)$. %Since $\textit{rm}(a_2)$ does not happen before $\textit{put}(a_2,\_)$, $\textit{cal}(\textit{put},a_2,\_)$ is before $\textit{ret}(\textit{rm},a_2)$. %Since $a_1 <_{\textit{ob}}^A a_2$, $\textit{ret}(\textit{put},a_1,\_)$ is before $\textit{cal}(\textit{put},a_2,\_)$.
    Since $a_1 <_{\textit{ob}}^A a_2$, $\textit{put}(a_1,\_) <_{\textit{hb}} \textit{put}(a_2,\_)$, and $\textit{rm}(a_2)$ does not happen before $\textit{rm}(a_1)$, and $\textit{cal}(\textit{rm},a_1)$ is before $\textit{ret}(\textit{rm},a_2)$. Since $a <_{\textit{ob}}^B a_1$, $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(a_1)$, and $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},a_1)$.

    Therefore, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$. Since $\textit{cal}(\textit{rm},b)$ is before $o$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$. Then it is easy to see that $o$ is before $\textit{cal}(\textit{put},a,\_)$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\_)$, and $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a,\_)$. This is impossible, and we can safely ignore this reason.
\end{itemize}

This completes the proof of this lemma. \qed
\end {proof}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-1.pdf}
%\vspace{-10pt}
  \caption{The first possible enumeration.}
  \label{fig:history enumeration 1 for Rpr2}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-2.pdf}
%\vspace{-10pt}
  \caption{The second possible enumeration.}
  \label{fig:history enumeration 2 for Rpr2}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-3.pdf}
%\vspace{-10pt}
  \caption{The third possible enumeration.}
  \label{fig:history enumeration 3 for Rpr2}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-4.pdf}
%\vspace{-10pt}
  \caption{The forth possible enumeration.}
  \label{fig:history enumeration 4 for Rpr2}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-5.pdf}
%\vspace{-10pt}
  \caption{The fifth possible enumeration.}
  \label{fig:history enumeration 5 for Rpr2}
\end{figure}

Given a data-differentiated history $h$, two actions $\textit{act}_1$, $\textit{act}_2$ of items of priority $\textit{pri}$, and assume that $\textit{act}_1$ is before $\textit{act}_2$ in $h$. We say that $\textit{act}_1$, $\textit{act}_2$ is covered by items $d_1,\ldots,d_m$ in $h$, if the priorities of $d_1,\ldots,d_m$ is not $\textit{pri}$, and

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{ret}(\textit{put},d_m,\_)$ is before $\textit{act}_1$,

\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,

\item[-] $\textit{act}_2$ is before $\textit{cal}(\textit{rm},d_1)$.
\end{itemize}

According to the definition of gap-point, in \figurename~\ref{fig:history enumeration 1 for Rpr2}, \figurename~\ref{fig:history enumeration 2 for Rpr2}, \figurename~\ref{fig:history enumeration 3 for Rpr2}, \figurename~\ref{fig:history enumeration 4 for Rpr2} or \figurename~\ref{fig:history enumeration 5 for Rpr2}, the location of $o$ and $o'$ is equivalent to that, there exists items $d_1,\ldots,d_m$ in $h$ in $h$, such that $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ is covered by $d_1,\ldots,d_m$, $\textit{cal}(\textit{rm},b)$ is before $\textit{ret}(\textit{put},d_m,\_)$ in $h$, and $\textit{cal}(\textit{rm},d_1)$ is before $\textit{ret}(\textit{rm},a)$ in $h$. We say that such $d_1,\ldots,d_m$ constitute the rightmost gap of $\textit{put}(b,\_)$ and $\textit{rm}(b)$.

An automaton $\mathcal{A}_{\textit{Rpr2-1}}$ is given in \figurename~\ref{fig:automata for first enumeration of Rpr2}, and it is constructed for the first enumeration in \figurename~\ref{fig:history enumeration 1 for Rpr2}. In \figurename~\ref{fig:automata for first enumeration of Rpr2}, we renames the items that covers $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ into $d$, and rename the renaming items into $e$. Here let $c = \textit{cal}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{cal}(\textit{rm},e), \textit{ret}(\textit{rm},e)$, $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_1.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{Rpr2-1}}$}
  \label{fig:automata for first enumeration of Rpr2}
\end{figure}

An automaton $\mathcal{A}_{\textit{Rpr2-2}}$ is given in \figurename~\ref{fig:automata for second enumeration of Rpr2}, and it is constructed for the second enumeration in \figurename~\ref{fig:history enumeration 2 for Rpr2}. In \figurename~\ref{fig:automata for second enumeration of Rpr2}, $c_1$, $c_2$, $c_3$ and $c_4$ is same as that in \figurename~\ref{fig:automata for first enumeration of Rpr2}.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_2.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{Rpr2-2}}$}
  \label{fig:automata for second enumeration of Rpr2}
\end{figure}

For the third enumeration in \figurename~\ref{fig:history enumeration 3 for Rpr2}. Since we want to ensure that $a$ and $b$ are putted only once, we need to explicitly record the positions of $\textit{cal}(\textit{put},a,p)$ and $\textit{cal}(\textit{put},b,p)$. Since the positions of $\textit{cal}(\textit{put},a,p)$ and $\textit{cal}(\textit{put},b,p)$ are not fixed, there are finite possible cases to consider, as shown below:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{rm},b)$ and before $\textit{cal}(\textit{rm},a)$: There are two possible positions of $\textit{cal}(\textit{put},a,p)$: (1) before $\textit{cal}(\textit{rm},a_1)$, and (2) after $\textit{cal}(\textit{rm},a_1)$, and before $\textit{ret}(\textit{put},a)$.

\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$ and before $\textit{cal}(\textit{rm},b)$: same as above case.

\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{put},a_1,\_)$ and before $\textit{ret}(\textit{rm},a_1)$: same as above case.

\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{put},a)$ and before $\textit{cal}(\textit{put},a_1,p)$: same as above case.

\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$: There are three possible positions of $\textit{cal}(\textit{put},a,p)$: (1) after $\textit{cal}(\textit{put},b,p)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{cal}(\textit{put}b,p)$, and (3) before $\textit{call}(\textit{rm},a_1)$.

\item[-] If $\textit{cal}(\textit{put},b,p)$ is before $\textit{cal}(\textit{rm},a_1)$: There are three possible positions of $\textit{cal}(\textit{put},a,p)$: (1) after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{cal}(\textit{put},b,p)$ and before $\textit{cal}(\textit{rm},a_1)$, and (3) before $\textit{cal}(\textit{put},b,p)$.
\end{itemize}

Therefore, there are fourteen possible cases that satisfy the third enumeration in \figurename~\ref{fig:history enumeration 3 for Rpr2}. For each case, we construct an finite automaton. Let $\textit{Aut}_{\textit{Rpr2-3}}$ be the set of finite automata that is constructed for above fourteen cases. For example, for the case $\textit{ca}_1$ when $\textit{cal}(\textit{put},a,p)$ is before $\textit{cal}(\textit{rm},a_1)$, $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$, and $\textit{cal}(\textit{put},b,p)$ is before $\textit{cal}(\textit{rm},b)$, we construct a finite automaton in \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}. In \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}, let $c$ and $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$ the same as that in \figurename~\ref{fig:automata for first enumeration of Rpr2}. Let $c_2 = c_1 + \textit{ret}(\textit{put},a_1)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$, $c_4 = c_3 + \textit{ret}(\textit{rm},d)$, and $c_5 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{put},a_1) + \textit{ret}(\textit{rm},d)$. The other finite automata in $\textit{Aut}_{\textit{Rpr2-3}}$ can be similarly constructed.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_3.pdf}
%\vspace{-10pt}
  \caption{Automaton for $\textit{ca}_1$}
  \label{fig:automata for ca1 of third enumeration of Rpr2}
\end{figure}

Similarly, we construct sets $\textit{Aut}_{\textit{Rpr2-4}}$ and $\textit{Aut}_{\textit{Rpr2-5}}$ of finite automata for the forth enumeration in \figurename~\ref{fig:history enumeration 4 for Rpr2} and the fifth enumeration in \figurename~\ref{fig:history enumeration 5 for Rpr2}, respectively.

\begin{restatable}{lemma}{Rpr2IsCoRegular}
\label{lemma:Rpr2 is co-regular}
$R_{\textit{pr2}}$ is co-regular.
\end{restatable}

\begin {proof}

Let $\textit{AutS-R}_{\textit{pr2}} = \{ \mathcal{A}_{\textit{Rpr2-1}}, \mathcal{A}_{\textit{Rpr2-2}} \} \cup \textit{Aut}_{\textit{Rpr2-3}} \cup \textit{Aut}_{\textit{Rpr2-4}} \cup \textit{Aut}_{\textit{Rpr2-5}}$. We need to prove that:

\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr2}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr2}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr2}}$.

By Lemma \ref{lemma:Rpr2 as gap-point and ob order}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:

\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr0}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, such that $\textit{last}(h') = R_{\textit{pr2}}$, $a$ and $b$ has maximal priority in $h'$, $a <_{\textit{ob}} b$ in $h'$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.

\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\textit{AutS-R}_{\textit{pr2}}$. By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$. Since $h_1$ is accepted by $\textit{AutS-R}_{\textit{pr2}}$, let $x$ and $y$ be the items that are renamed into $b$ and $a$ by $r$, respectively, and let $d_1,\ldots,d_m$ be the items that are renamed into $d$ by $r$.

let $h'' = h_2 \vert_{ \{ x,y,d_1,\ldots,d_m \} }$. It is obvious that $\textit{last}(h'') = R_{\textit{pr2}}$. According to our construction of automata in $\textit{AutS-R}_{\textit{pr2}}$, it is not hard to see that $x$ and $y$ has maximal priority in $h_2$, $y <_{\textit{ob}} x$ in $h''$, and the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$ in $h''$. %By Lemma \ref{lemma:Rpr2 as gap-point and ob order}, we know that $h''$ is not linearizable to $\textit{MR}_{\textit{pr2}}$.

\noindent The $\textit{if}$ direction: Assume that there exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr2}}$, $a'$ and $b'$ has maximal priority in $h'$, $a' <_{\textit{ob}} b'$ in $h'$, and the rightmost gap-point of $\textit{put}(b',\_)$ and $\textit{rm}(b')$ is before $\textit{cal}(\textit{put},a',\_)$ or $\textit{cal}(\textit{rm},a')$ in $h''$. By data-independence, we can obtain history $h_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the items $d_1,\ldots,d_m$ that constitute the rightmost gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, we rename them into $d$, and (3) rename the other items into $e$. It is easy to see that $\textit{last}(h_1) = R_{\textit{pr2}}$, $a$ and $b$ has maximal priority in $h_1$, $a <_{\textit{ob}} b$ in $h_1$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$. By Lemma \ref{lemma:five enumeration is enough for Rpr2}, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists) and $a_2$ (if exists). Then

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $a <_{\textit{ob}} b$ because of the first enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr2-1}}$.

\item[-] If $a <_{\textit{ob}} b$ because of the second enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr2-2}}$.

\item[-] If $a <_{\textit{ob}} b$ because of the third enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-3}}$.

\item[-] If $a <_{\textit{ob}} b$ because of the forth enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-4}}$.

\item[-] If $a <_{\textit{ob}} b$ because of the fifth enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-5}}$.
\end{itemize}

This completes the proof of this lemma. \qed
\end {proof}

\subsection{Co-Regular of $R_{\textit{pr3}}$}
\label{subsec:co-regular of Rpr3}

\begin{restatable}{lemma}{Rpr3IsAlwaysCoRegular}
\label{lemma:Rpr3 is always co-regular}
Given a data-differentiated history $h$, if $\textit{last}(h) = R_{\textit{pr3}}$, then $h \sqsubseteq \textit{MR}_{\textit{pr3}}$.
\end{restatable}

\begin {proof}

Since $\textit{last}(h) = R_{\textit{pr3}}$, there exists item $x$ with $\textit{pri}_x$ in $h$, such that $h \vert_{ \{ x \}} = \textit{put}(x,\textit{pri}_x)$, and $\textit{pri}_x$ is larger than any priority of $h$.

For each operation, we locate its linearization point just after its call action. In such way we obtain a sequence $l$ of operations, and it is easy to see that $h \sqsubseteq l$. It is obvious that $l$ contains $\textit{put}(x,\textit{pri})$, and then $l \in \textit{MR}_{\textit{pr3}}$. This completes the proof of this lemma. \qed
\end {proof}

\subsection{Co-Regular of $R_{\textit{pr4}}$}
\label{subsec:co-regular of Rpr4}

\begin{restatable}{lemma}{Rp4AsHappenBefore}
\label{lemma:Rpr4 as happen before}
Given a data-differentiated history $h$. There exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr4}}$, and $h'$ does not linearizable to $\textit{MR}_{\textit{pr4}}$, if and only if there exists $h'' \in \textit{proj}(h)$, $\textit{last}(h'') = R_{\textit{pr4}}$, $x$ and $y$ has maximal priority in $h''$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,\_) <_{\textit{hb}} \textit{put}(y,\_)$.
\end{restatable}

\begin {proof}

To prove the $\textit{if}$ direction, let $h_1 = h \vert_{ \{ x,y \} }$. It is easy to see that $\textit{last}(h') = R_{\textit{pr4}}$ and $h_1$ does not linearizable to $\textit{MR}_{\textit{pr4}}$.

To prove the $\textit{only if}$ direction, we prove its contrapositive. Assume that for each $h'' \in \textit{proj}(h)$, such that $\textit{last}(h'') = R_{\textit{pr4}}$, for each $x$ and $y$ has maximal priority in $h''$, if $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, then $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$. We need to prove that, for each $h' \in \textit{proj}(h)$, if $\textit{last}(h') = R_{\textit{pr4}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr4}}$.

Given such $h'$, let $x_1,\ldots,x_m$ be the set of items with maximal priority and has unmatched $\textit{put}$ operations, let $y_1,\ldots,y_n$ be the set of items with maximal priority and has matched $\textit{put}$ and $\textit{rm}$ operations. It is easy to see that, for each $x_i$ and $y_j$, $\textit{cal}(\textit{put},y_j)$ is before $\textit{ret}(\textit{put},x_i,\_)$. Then we locate the linearization points of $h'$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] For each $x_i$, locate the linearization point of $\textit{put}(x_i,\_)$ just before its return action.

\item[-] For each $y_j$, locate the lineariztion point of $\textit{put}(y_j,\_)$ jest after its call action.

\item[-] For other operations, locate their linearization points at an arbitrary location after its call action and before its return action.
\end{itemize}

Let $l$ be the sequence of linearization points. It is easy to see that $h \sqsubseteq l$, and $l \in \textit{MR}_{\textit{pr4}}$. Here the $\textit{itm}$ in $R_{\textit{pr4}}$ is chosen to be $\textit{ret}(\textit{put},x_{\textit{ind}},\_)$, where the return action of $x_{\textit{ind}}$ is later than return actions of other $x_i$. This completes the proof of this lemma. \qed
\end {proof}

\begin{restatable}{lemma}{Rpr4IsAlwaysCoRegular}
\label{lemma:Rpr4 is always co-regular}
Given a data-differentiated history $h$, if $\textit{last}(h) = R_{\textit{pr4}}$, then $h \sqsubseteq \textit{MR}_{\textit{pr4}}$.
\end{restatable}

\begin {proof}

According to Lemma \ref{lemma:Rpr4 as happen before}, if $\textit{last}(h) = R_{\textit{pr4}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr4}}$, then there exists $x$ and $y$ with maximal priority in $h$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,\_) <_{\textit{hb}} \textit{put}(y,\_)$. Let $h_1 = h \vert_{ \{ x,y \} }$. It is obvious that $\textit{transToQueue}(h_1)$ does not linearizable to queue. This contradicts the assumption that every single-priority history can be ``linearizable to queue'', and thus, we can safely ignore this case. \qed
\end {proof}

\subsection{Co-Regular of $R_{\textit{pr5}}$}
\label{subsec:co-regular of Rpr5}

$R_{\textit{pr5}} \equiv (u \cdot v \in \textit{PQueue}) \wedge (\textit{matched}(u,\textit{put},\textit{rm}) ) \Rightarrow (u \cdot \textit{rmEmpty} \cdot v \in \textit{PQueue})$

Similar as the notion of gap in \cite{Bouajjani:2015}, we propose the notion of gap of $\textit{em}(\textit{empty})$ operations.

\begin{definition}[gap for $\textit{em}(\textit{empty})$ operations]\label{def:gap for rmEmpty operations}

Let $h$ be a data-differentiated history and $o = \textit{rm}(\textit{empty})$ of $h$. We say that $h$ has a gap on operation $o$, if there is a partition of the operations of $h$ into $L$ and $R$ satisfying:
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $L$ has matched $\textit{put}$ and $\textit{rm}$ operations.

\item[-] no operation of $R$ happens before an operation of $L$ or $o$.

\item[-] $o$ does not happen before operations of $L$
\end{itemize}
\end{definition}

Compared to the gap on $\textit{put}$ and $\textit{rm}$ operations, here we do not consider the priority of operations in $L$ or $R$.

\begin{restatable}{lemma}{GapEqualsLinforRpr5}
\label{lemma:Gap Equals Lin for Rpr5}
A data-differentiated history $h$ has a projection $h'$, such that $\textit{last}(h') = R_{\textit{pr5}}$ and $h'$ does not linearizable to $\textit{MR}_{\textit{pr5}}$, if and only if there exists a projection $h''$ of $h$, such that $\textit{last}(h'') = R_{\textit{pr5}}$, there exists $o = \textit{rm}(\textit{empty})$ in $h''$, and there is no gap of $o$ in $h''$.
\end{restatable}

\begin {proof}

Let us prove the $\textit{only if}$ direction by contradiction. Assume that there exists such $h'$, and for each $\textit{rm}(\textit{empty})$ in $h'$, there is a gap of $o$ in $h''$. Since $\textit{last}(h') = R_{\textit{pr5}}$, there is exists at least $\textit{rm}(\textit{empty})$ operation in $h$, and let $o$ be arbitrary one of them. Since there is a gap of $o$ in $h$, let $L$ and $R$ be the sets of the gap. Let $u$ be a sequence of operations of $L$, and $u$ is generated from $h$ by locating linearization points of each operation in $L$ just after its call action. Let $v$ be a sequence of operations of $R$, and $v$ is generated from $h$ by locating linearization points of each operation in $R$ just before its return action. It is easy to see that $h \sqsubseteq l = u \cdot o \cdot v$, and $l \in \textit{MR}_{\textit{pr5}}$. This contradicts that $h'$ does not linearizable to $\textit{MR}_{\textit{pr5}}$.

To prove the $\textit{if}$ direction, let $h_1$ be a history generate from $h''$ by discarding all other $\textit{rm}(\textit{Empty})$ operations. By assumption, it is obvious that there is no gap of $o$ in $h_1$. Assume by contradiction that there exists $l'$, such that $h_1 \sqsubseteq l' \in \textit{MR}_{\textit{pr5}}$. It is easy to see that $l' = u' \cdot o \cdot v'$, such that $u'$ contains matched $\textit{put}$ and $\textit{rm}$ operations. Let $L$ be the set of operations of $u'$, and $R$ be the set of operations in $v'$. Since $h_1 \sqsubseteq l'$, it is easy to see that $L$ and $R$ is a gap of $o$ in $h_1$, contradicts that there is no gap of $o$ in $h_1$. \qed
\end {proof}

Similar as the notion of gap in \cite{Bouajjani:2015}, we propose the notion left-right constraint of $\textit{rm}(\textit{empty})$ operation.

\begin{definition}[left-right constraint for $\textit{rm}(\textit{empty})$ operation]\label{def:left-right constraint for rmEmpty operation}
Given a data-differentiated history $h$, and $o = \textit{rm}(\textit{empty})$ of $h$. The left-right constraint of $o$ is the graph $G$ where:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] the nodes are the items of $h$ or $o$, to which we add a node,

\item[-] there is an edge from item $d_1$ to $o$, if $\textit{put}(d_1,\_)$ happens before $o$,

\item[-] there is an edge from $o$ to item $d_1$, if $o$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$,

\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$.
\end{itemize}
\end{definition}

Compared to the left-right constraint of $\textit{put}$ and $\textit{rm}$ operations, here we do not consider the priority. We also do not consider other $\textit{rm}(\textit{empty})$ operations in $h$.

Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$. Let $\textit{LMSet}_1(h,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some item, either $\textit{op}$ happens before $o$ in $h$, or there is an operation $\textit{op}'$ with the same item of $\textit{op}$, such that $\textit{op}'$ happens before $o$ in $h$ $\}$. For each $i \geq 1$, let $\textit{LMSet}_{\textit{i+1}}(h,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some item, $\textit{op}$ is not an operation of items in $\textit{LMSet}_k(h,o)$ for each $k \leq i$, either $\textit{op}$ happens before some operation $o' \in \textit{LMSet}_i(h,o)$ in $h$, or there is an operation $\textit{op}''$ with the same item of $o$ and $\textit{op}''$ happens before some operation $o' \in \textit{LMSet}_i(h,o)$ in $h$ $\}$. We can see that $\textit{LMSet}_i(h,o) \cap \textit{LMSet}_j(h,o) = \emptyset$ for any $i \neq j$. Let $\textit{LMSet}(h,o) = \textit{LMSet}_1(h,o) \cup \textit{LMSet}_2(h,o) \cup \ldots$.

\begin{restatable}{lemma}{RmEmptyDoesNotHappenBeforeLMSetForRpr5}
\label{lemma:RmEmpty does not happen before LMSet for Rpr5}
Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$. Let $G$ be the graph representing the left-right constraint of $o$. Assume that $G$ has no cycle going through $o$. Then, $o$ does not happen before any operation in $\textit{LMSet}(h,o)$.
\end{restatable}

This Lemma can be similarly proved as Lemma \ref{lemma:Rmx does not happen before LMSet for Rpr1}.

\begin{restatable}{lemma}{LMSetHasMatchedPutandRmOperationsForRpr5}
\label{lemma:LMSet has matched put and rm operations for Rpr5}
Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$. Let $G$ be the graph representing the left-right constraint of $o$. Assume that $G$ has no cycle going through $o$. Then, $\textit{LMSet}(h,o)$ contains only matched $\textit{put}$ and $\textit{rm}$ operations.
\end{restatable}

This Lemma can be similarly proved as Lemma \ref{lemma:LMSet has matched put and rm operations}.

\begin{restatable}{lemma}{GapEqualsConstraintforRpr5}
\label{lemma:Gap Equals Constraint for Rpr5}
Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$. Let $G$ be the graph representing the left-right constraint of $o$. There is a gap on $o$, if and only if $G$ has no cycle going through $o$.
\end{restatable}

\begin {proof}

To prove the $\textit{only if}$ direction, assume that there is a gap on $o$, and let it be $L$ and $R$. Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow o \rightarrow d_1$ in $G$. Since $o \rightarrow d_1$, either $o$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$. Since $L$ contains matched $\textit{put}$ and $\textit{rm}$ operations, it is easy to see that operations of $d_1$ is in $R$. Since $d_1 \rightarrow d_2$, we can see that $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$. Since operations of $d_1$ is in $R$ and $L$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we can see that operations of $d_2$ is in $R$. Similarly, we can prove that operations of $d_3,\ldots,d_m$ is in $R$. Since $d_m \rightarrow o$, $\textit{put}(d_m,\_)$ happens before $o$, and $\textit{put}(d_m,\_) \in L$. This contradicts that operations of $d_m$ is in $R$. This completes the proof of the $\textit{only if}$ direction.

To prove the $\textit{if}$ direction, assume that $G$ has no cycle going through $o$. Let $O_h$ be the set of operations of $h$. Let $O_L = \textit{LMSet}(h,o)$, $O_R = O_h \setminus O_L$.

By Lemma \ref{lemma:LMSet has matched put and rm operations for Rpr5}, we can see that $O_L = \textit{LMSet}(h,o)$ contains only matched $\textit{put}$ and $\textit{rm}$ operations. It remains to prove that $L$ and $R$ is a gap of $o$. We prove this by showing that all the following cases are impossible:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Case $1$: Some operation $o_r \in R$ happens before $o$. Then we know that $o_r \in \textit{LMSet}(h,o)$, which contradicts that $o_r \in R$.

\item[-] Case $2$: Some operation $o_r \in R$ happens before some operation $o_l$. Then we know that $o_r \in \textit{LMSet}(h,o)$, which contradicts that $o_r \in R$.

\item[-] Case $4$: $o$ happens before some $o_l \in \textit{LMSet}(h,o) = L$. By Lemma \ref{lemma:RmEmpty does not happen before LMSet for Rpr5}, we know that this is impossible.
\end{itemize}

This completes the proof of the $\textit{if}$ direction.

\qed
\end {proof}

Let us begin to represent an automaton that is used for capture the case that, in a sub-history $h'$ of a history $h$, $\textit{last}(h')=R_{\textit{pr5}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr5}}$, and the reason is that there is no gap on a $\textit{rm}(\textit{empty})$ operation in $h'$. The automaton is $\mathcal{A}_{\textit{Rpr5}}$, which is given in \figurename~\ref{fig:automata for Rpr5}. In \figurename~\ref{fig:automata for Rpr5}, let $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d), \textit{ret}(\textit{rm},d)$, $c_1 = c + \textit{cal}(\textit{put},b,\textit{anPri})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c + \textit{ret}(\textit{rm},b)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr5.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{Rpr5}}$}
  \label{fig:automata for Rpr5}
\end{figure}

Given a data-differentiated history $h$, we say that $o = \textit{rm}(\textit{empty})$ in $h$ is covered by items $d_1,\ldots,d_m$ in $h$, if

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{put}(d_m,\_)$ happens before $o$,

\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,

\item[-] $o$ happens before $\textit{rm}(d_1)$, or $\textit{rm}(d_1)$ does not exists in $h$
\end{itemize}

According to the definition of left-right constraint for $o$, in a data-differentiated history $h$, there is a cycle going through $o$, if and only if there exists items $d_1,\ldots,d_m$, such that $o$ is covered by $d_1,\ldots,d_m$.

\begin{restatable}{lemma}{Rpr5IsCoRegular}
\label{lemma:Rpr1 5s co-regular}
$R_{\textit{pr5}}$ is co-regular.
\end{restatable}

\begin {proof}

We need to prove that:

\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\mathcal{A}_{\textit{Rpr5}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr5}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr5}}$.

By Lemma \ref{lemma:Gap Equals Lin for Rpr5} and Lemma \ref{lemma:Gap Equals Constraint for Rpr5}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:

\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\mathcal{A}_{\textit{Rpr5}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h''$ is a projection of $h$, such that $\textit{last}(h'') = R_{\textit{pr5}}$, $o = \textit{rm}(\textit{empty})$ is in $h''$, and $o$ is covered by some items $d_1,\ldots,d_m$ in $h''$.

\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\mathcal{A}_{\textit{Rpr5}}$. By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$. Let $d_1,\ldots,d_m$ be the items in $h_2$ such that $r(d_i)=d$ for each $1 \leq i \leq m$. Let $h'' = h_2 \vert_{ \{ o, d_0, d_1, \ldots, d_m \} }$. It is obvious that $\textit{last}(h'') = R_{\textit{pr5}}$. It is easy to see that $o$ is covered by $d_1,\ldots,d_m$.

\noindent The $\textit{if}$ direction: Assume that there exists such $h$, $h''$, $o$ and $d_1,\ldots,d_m$. Then, let $h_1$ be obtained from $h$ by renaming $d_1,\ldots,d_m$ into $b$ and renaming other items into $d$. We can prove that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr5}}$.

This completes the proof of this lemma. \qed
\end {proof}

\section{Relate Queue and Stack with Priority Queue}
\label{sec:relate queue and stack with priority queue}

\subsection{Relate Queue with Priority Queue}
\label{subsec:relate queue with priority queue}

Given a history $h_q$ of queue operations, let $\textit{UniPriTrans}(h_q)$ be a history of priority queue, which is generated as follows: transform each $\textit{call}(\textit{enq},a)$, $\textit{ret}(\textit{enq},a)$, $\textit{call}(\textit{deq},a)$ and $\textit{ret}(\textit{deq},a)$ of $h_q$ into $\textit{call}(\textit{put},a,1)$, $\textit{ret}(\textit{put},a)$, $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$.

\begin{restatable}{lemma}{RelateQueuewithPQ}
\label{lemma:relate queue with priority queue}
Given a history $h_q$ of queue operations, $h_q$ is linearizable to queue, if and only if $\textit{UniPriTrans}(h_q) \sqsubseteq \textit{PQueue}$.
\end{restatable}

\begin {proof}

This is obvious, since we already ensure that for each single-priority history $h$, $\textit{transToQueue}(h)$ is linearizable to queue. \qed
\end {proof}

\subsection{Relate Stack with Priority Queue}
\label{subsec:relate stack with priority queue}

\section{Conclusion and Future Work}
\label{sec:conclusion and future work}

Conclusion

%\FloatBarrier

%\bibliography{short}
%\bibliographystyle{plain}
%\bibliographystyle{splncs}
%\bibliography{biblio_cat.bib}

\begin{thebibliography}{50}

\bibitem{Bouajjani:2015}
Bouajjani, A., Emmi, M., Enea, C., Hamza, J.:
\newblock On reducing linearizability to state reachability.
\newblock In: Halld{\'{o}}rsson, M.M. et al. (eds.) ICALP 2015, Part II, pp. 95--107. Springer (2015)

\bibitem{Wolper:1986}
Wolper, P.:
\newblock Expressing interesting properties of programs in propositional
  temporal logic.
\newblock In {\em POPL'86}. pp.184--193. (1986)

\end{thebibliography}
\newpage

\appendix

\end{document}

### 2017-02-16 paper-LIN-PQ.tex
trivial testing

@@
 
-Given a sequence $l$ of priority queue operations, we say that $l$ is data-differentiated, if for each item $d \in \mathcal{D}$, $\textit{put}(d,\_)$ appears at most once in $l$.
+given a sequence $l$ of priority queue operations, we say that $l$ is data-differentiated, if for each item $d \in \mathcal{D}$, $\textit{put}(d,\_)$ appears at most once in $l$.
 Note that a data-differentiated sequence $l$ may contains more than one items with a same priority.

### 2017-02-20 abstract.tex
preliminary,data-independence,rules

--- 2017-02-16 paper-LIN-PQ.tex
@@
-%\documentclass[a4paper,UKenglish]{lipics}
-\documentclass{llncs}
-%\usepackage{mathbbold}
-
-%\newtheorem{theorem}{Theorem}[section]
-%\newtheorem{conjecture}[theorem]{Conjecture}
-%\newtheorem{corollary}[theorem]{Corollary}
-%\newtheorem{proposition}[theorem]{Proposition}
-%\newtheorem{lemma}[theorem]{Lemma}
-%\newdef{definition}[theorem]{Definition}
-%\newdef{remark}[theorem]{Remark}
-%\newdef{example}[theorem]{Example}
-
-\usepackage{epsfig}
-\usepackage{amsmath}
-\usepackage{color}
-%\usepackage{txfonts}
-%\usepackage{pxfonts}
-%\usepackage{mathabx}
-%\usepackage{makeidx}  % allows for indexgeneration
-\usepackage{verbatim}
-\usepackage{url}
-%\usepackage{hyperref}
-\usepackage{times}
-\usepackage{ulem}    %(\emph{....} in the texfile) is underlined instead of italic
-\normalem
-\usepackage{enumerate}
-%\usepackage[square, comma, sort&compress, numbers]{natbib}
-
-%added by Yi Lv%
-%\usepackage[linesnumbered,ruled,procnumbered,noend]{algorithm2e}
-%\usepackage[linesnumbered,ruled,procnumbered,vlined]{algorithm2e}
-\usepackage[linesnumbered,ruled,procnumbered,noend]{algorithm2e}
-\usepackage{stmaryrd}
-%\usepackage{hyperref}
-%added by Wang Chao%
-%\usepackage{mathrsfs}
-%\usepackage{extarrows}
-\usepackage{thmtools}
-\usepackage{thm-restate}
-
-%add for TIkz
-\usepackage[version=0.96]{pgf}
-\usepackage{tikz}
-\usetikzlibrary{arrows,shapes,snakes,automata,backgrounds,petri}
-\usepackage[latin1]{inputenc}
-\usepackage{placeins}
-\usepackage[bookmarks=false]{hyperref}
-
-%\declaretheorem[name=Claim]{clm}
-%\declaretheorem[name=Lemma]{lem}
-%\declaretheorem[name=Theorem]{thm}
-
-\DeclareSymbolFont{largesymbolsA}{U}{txexa}{m}{n}
-\SetSymbolFont{largesymbolsA}{bold}{U}{txexa}{bx}{n}
-\DeclareFontSubstitution{U}{txexa}{m}{n}
-\DeclareMathSymbol{\bigsqcupplus}{\mathop}{largesymbolsA}{"02}
-
-% "Import" \bigsqcupplus from txfonts without loading txfonts (since
-% it changes the default font and replaces many symbols)
-
-% Commands by Catuscia
-
-%\DeclareMathOperator*{\minarg}{{\rm minarg}}
-
-%Previous notation for the probabilistic choice
-%\newcommand{\probsum}{\bigoplus}
-%\newcommand{\smallprobsum}[1]{\mkern4mu{\textstyle\circ\mkern-15.5mu\sum_{#1}\:}}
-%\newcommand{\bigprobsum}[1]{{\;\displaystyle\odot\mkern-21mu\sum_{#1}}}
-%\newcommand{\probsum}[1]{\mathchoice{\bigprobsum{#1}}{\smallprobsum{#1}}{\smallprobsum{#1}}{\smallprobsum{#1}}}
-%For the uniform distribution
-%\newcommand{\usmallprobsum}[1]{\mkern4mu{\textstyle\circ\mkern-15.5mu\sum^\mathcal{U}_{#1}\:}}
-%\newcommand{\ubigprobsum}[1]{{\;\displaystyle\odot\mkern-21mu\sum^\mathcal{U}_{#1}}}
-%\newcommand{\uprobsum}[1]{\mathchoice{\ubigprobsum{#1}}{\usmallprobsum{#1}}{\usmallprobsum{#1}}{\usmallprobsum{#1}}}
-
-%\newcommand{\nondsum}{\bigbox}
-\newcommand{\nondsum}{\bigsqcupplus}
-\newcommand{\probplus}[1]{\oplus_{#1}}
-%\newcommand{\nondplus}{\square}
-\newcommand{\bang}{!\,}
-\newcommand{\nondplus}{{\textstyle\bigsqcupplus}}
-\newcommand{\partmap}{\rightharpoonup}
-\newcommand{\map}{\rightarrow}
-%\newcommand{\exc}{\phi}
-\newcommand{\exc}{\alpha}
-\newcommand{\exec}{\mathit{exec}}
-\newcommand{\execp}{\mathit{execp}}
-\newcommand{\Act}{\mathit{Act}}
-\newcommand{\Sec}{\mathit{Sec}}
-\newcommand{\Obs}{\mathit{Obs}}
-\newcommand{\etree}{\mathit{etree}}
-\newcommand{\lstate}{\mathit{lst}}
-\newcommand{\fstate}{\mathit{fst}}
-%\newcommand{\STATE}{\mathcal{P}r}  $ original marked
-%\newcommand{\STATE}{\mathcal{P}}   $ marked by me
-%\newcommand{\st}{P}
-\newcommand{\trans}{\mathcal{T}}
-\newcommand{\Aut}{\mathcal{M}}
-\newcommand{\init}{\mathit{init}}
-\newcommand{\perr}{\mathcal{P}}
-
-\newcommand{\calo}{\mathcal{O}}
-\newcommand{\cals}{\mathcal{S}}
-\newcommand{\sseq}{\vec s}
-\newcommand{\oseq}{\vec o}
-\newcommand{\ccsp}{CCS$_p$}
-
-\newcommand{\bigfrac}[2]{\frac{\raisebox{1ex}{$#1$}}{\raisebox{-1.5ex}{$#2$}}}
-\newcommand{\nondarr}[1]{\overset{#1}{\longrightarrow}}
-\newcommand{\Nondarr}[1]{\overset{#1}{\Longrightarrow}}
-\newcommand{\vectorArrow}[1]{\stackrel{\longrightarrow}{\mbox{#1}}}
-\newcommand{\probarr}[1]{\overset{#1}{\dashrightarrow}}
-\newcommand{\paral}{\,|\,}
-\newcommand{\outp}[1]{\overline{#1}}
-\renewcommand{\Pr}{{\rm Pr}}
-
-%Commands by Chao Wang
-
-%memory models%
-\newcommand{\TSO}{\textrm{TSO}}
-\newcommand{\PSO}{\textrm{PSO}}
-
-%correctness conditions%
-\newcommand{\lin}{\textrm{linearizability}}
-\newcommand{\slin}{\textrm{static linearizability}}
-\newcommand{\qlin}{\textrm{quasi linearizability}}
-\newcommand{\TTlin}{\textrm{TSO-to-TSO linearizability}}
-
-\newcommand{\pair}[2]{\langle #1 , #2 \rangle}% pairs
-\newcommand{\setof}[2]{\{ \, #1 \mid #2 \, \}}% Sets
-\newcommand{\set}[1]{\{ {#1}  \}  }
-%\newcommand{\map}[3]{{#1} \colon {#2} \longmapsto {#3}} %functions
-\newcommand{\den}[1]{[\![#1]\!]}% Denotation of
-\newcommand{\mean}[1]{|\!|#1|\!|}
-\newcommand{\forget}[1]{}
-
-%%%%%%%%%%%%%GENERAL%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\itbox}[1]{{\it #1\/}}
-\newcommand{\un}[1]{\uline{#1}}%\underline{#1}}
-\newcommand{\ov}[1]{\overline{#1}}
-\newcommand{\smallspace}{\vspace{10mm}}
-\newcommand{\is}{\mbox{$\Longleftarrow\ $}}
-\newcommand{\pright}[1]{\hfill{#1}}
-\newcommand{\bnfor}{\;\;\mid\;\;}
-
-\newcommand{\ar}[1]{\stackrel{\scriptstyle #1}{\longrightarrow}}
-
-%%%%%%%%%italics in math mode
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\newcommand{\true}{{\it true}}
-%\newcommand{\false}{{\it false}}
-\newcommand{\calB}{{\cal B}}
-\newcommand{\calF}{{\cal F}}
-\newcommand{\calP}{{\cal P}}
-\newcommand{\order}{{\cal O}}
-\newcommand{\size}[1]{|#1|}
-
-%other notations%
-\newcommand{\LTS}{\textit{LTS}}
-\newcommand{\bedt}[1]{{\color{blue}#1}}
-\newcommand{\redt}[1]{{\color{red}#1}}
-
-%\pagestyle{empty}
-%\pagestyle{plain}
-
-%\def\lastname{Xu,Palamidessi}
-
-\title{Reducing Linearizability of Priority Queues and More Data Structures to State Reachability}
-
-\author {Chao Wang \and Ahmed Bouajjani \and Constantin Enea}
-\institute{Institut de Recherche en Informatique Fondamentale, \\Univ. Paris Diderot (Paris 7)
-}
-
-\begin{document}
-
-\maketitle
-
 \begin{abstract}
-Abstract
+I am abstract
 \end{abstract}
@@
 }
-
-\section{Introduction}
-\label{sec:introduction}
-Introduction
-
-\smallskip
-
-\noindent {\bf Related work}
-%\label{sec:related work}
-\noindent Related work
-
-\section{Inductive Rules for Priority Queues}
-\label{sec:inductive rules for priority queues}
-
-given a sequence $l$ of priority queue operations, we say that $l$ is data-differentiated, if for each item $d \in \mathcal{D}$, $\textit{put}(d,\_)$ appears at most once in $l$.
-Note that a data-differentiated sequence $l$ may contains more than one items with a same priority.
-The subset of data-differentiated operation sequences of a set $S$ is denoted by $S_{\neq}$.
-For priority queue, the renaming functions rename only the items and keep the priorities unchanged.
-For priority queue, we modify the notion of data-independence \cite{Wolper:1986} and use data-differentiated instead of differentiated:
-
-%\vspace{-6pt}
-\begin{definition}[data-independence of priority queue]\label{def:priority-value data-independence}
-The set $S$ of sequences of priority queue operations (resp., histories)  is data independent, if:
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] for all $l \in S$, there exists $l' \in S_{\neq}$, and a renaming $r$, such that $l=r(l')$,
-
-\item[-] for all $l \in S$ and for all renaming $r$, $r(l) \in S$.
-\end{itemize}
-\end{definition}
-
-Similar as in \cite{Bouajjani:2015}, we consider only completed executions, where each call action has a corresponding return action.
-This simplification is sound when implementation methods can always make progress in isolation.
-
-For the inductive rules of priority queue, our rules need to concern the cost of operations. Formally, the inductive rules for priority queues is as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $R_{\textit{pr0}} \equiv \epsilon \in \textit{PQueue}$
-
-\item[-] $R_{\textit{pr1}} \equiv (u \cdot v \cdot w \in \textit{PQueue}) \wedge (u \cdot v \cdot w \in (\textit{put},\textit{rm})^*) \wedge (\textit{pri} >$ priorities of $u \cdot v \cdot w ) \wedge (\textit{matched}(u \cdot v, \textit{put}, \textit{rm}) ) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w \in \textit{PQueue})$
-
-\item[-] $R_{\textit{pr2}} \equiv (u \cdot v \cdot w \in \textit{PQueue}) \wedge (u \cdot v \cdot w \in (\textit{put},\textit{rm})^*) \wedge (\textit{pri} =$ the maximal priority of $u \cdot v \cdot w ) \wedge (\textit{pri} >$ priorities of unmatched items in $u \cdot v \cdot w) \wedge (\textit{matched}(u \cdot v, \textit{put}, \textit{rm}) ) \wedge ($all $\textit{put}$ operation with priority $\textit{pri}$ is in $u) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w \in \textit{PQueue})$
-
-\item[-] $R_{\textit{pr3}} \equiv (u \cdot v \in \textit{PQueue}) \wedge ( u \cdot v \in (\textit{put},\textit{rm})^*) \wedge (\textit{pri} =$ the maximal priority of items in $u \cdot v) \wedge (\textit{pri} >$ priorities of matched items in $u \cdot v \cdot w ) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{PQueue})$
-
-\item[-] $R_{\textit{pr4}} \equiv (u \cdot v \in \textit{PQueue}) \wedge ( u \cdot v \in (\textit{put},\textit{rm})^*) \wedge (\textit{pri} =$ the maximal priority of items in $u \cdot v) \wedge (\textit{pri} =$ the maximal priority of matched items in $u \cdot v \cdot w ) \wedge ($all matched $\textit{put}$ with priority $\textit{pri}$ is in $u) \wedge ($all unmatched $\textit{put}$ with priority $\textit{pri}$ is in $u) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{PQueue})$
-
-\item[-] $R_{\textit{pr5}} \equiv (u \cdot v \in \textit{PQueue}) \wedge (\textit{matched}(u,\textit{put},\textit{rm}) ) \Rightarrow (u \cdot \textit{rmEmpty} \cdot v \in \textit{PQueue})$
-\end{itemize}
-
-Given a sequence $l$ of possibly $\textit{put}$, $\textit{rm}$ and $\textit{rmEmpty}$ operations, we say that $l$ contains unmatched $\textit{rm}$ operations, if the number of $\textit{rm}(d)$ operations is larger than the number of $\textit{put}(d,\_)$ operations for some $d$ in $l$.
-By data-independence, we can detect the existence of unmatched $\textit{rm}$ operations by automata.
-A automaton $\mathcal{A}_{\textit{unmRm1}}$ is given in \figurename~\ref{fig:automata for unmatched rm1}, where $c_1=\textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm),\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$, $c_2=c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c_1 + \textit{cal}(\textit{put},b,\textit{anyPri}) + \textit{ret}(\textit{rm},b)$.
-Here $\textit{anyPri}$ is a specific predicates that is satisfied by any priority.
-$\mathcal{A}_{\textit{unmRm1}}$ is used to detect the existence of a item that has been removed but has never been inserted.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_RM1.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{unmRm1}}$}
-  \label{fig:automata for unmatched rm1}
-\end{figure}
-
-An automaton $\mathcal{A}_{\textit{unmRm2}}$ is given in \figurename~\ref{fig:automata for unmatched rm2}, where $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm),$ $\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$, and $c_1 = c_1 + \textit{ret}(\textit{rm},b)$.
-$\mathcal{A}_{\textit{unmRm1}}$ is used to detect the existence of a item that has been removed more than once but has been inserted once.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_RM2.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{unmRm2}}$}
-  \label{fig:automata for unmatched rm2}
-\end{figure}
-
-The following lemma states that we can detect the existence of unmatched $\textit{rm}$ by $\mathcal{A}_{\textit{unmRm1}}$ and $\mathcal{A}_{\textit{unmRm2}}$.
-
-\begin{restatable}{lemma}{DetectUnmatchedRmByAutomata}
-\label{lemma:detect unmatched rm by automata}
-
-Given a data-independent $\mathcal{I}$, there is some unmatched $\textit{rm}$ in some sequence of $\mathcal{I}$, if and only if $( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm1}} \neq \emptyset) \vee ( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm2}} \neq \emptyset )$.
-\end{restatable}
-
-\begin {proof}
-
-For the $\textit{if}$ direction, assume that $( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm1}} \neq \emptyset) \vee ( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm2}} \neq \emptyset )$.
-Then, there exists a sequence $l \in \mathcal{I}$ and $l$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$ or $\mathcal{A}_{\textit{unmRm2}}$.
-If $l$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$, then $l$ contains at least one $\textit{rm}(b)$ operation and no $\textit{put}(b,\_)$ operation; Else, if $l$ is accepted by $\mathcal{A}_{\textit{unmRm2}}$, then $l$ contains at least two $\textit{rm}(b)$ operations and one $\textit{put}(b,\_)$ operation.
-In both cases, $l$ contains unmatched $\textit{rm}$ operations.
-
-To prove the $\textit{only if}$ direction assume that there exists a sequence $l \in \mathcal{I}$ that contains unmatched $\textit{rm}$ operations.
-By data-independence, there exists data-differentiated sequence $l'$ and renaming function $r$, such that $l=r(l')$.
-We need to prove that in $l'$, one of the following cases holds: (1) $l'$ contains $\textit{rm}(d_1)$ for some $d_1$ and does not contain any $\textit{put}(d_1,\_)$, and (2) $l'$ contains one $\textit{put}(d_1,\_)$ operation and more than one $\textit{rm}(d_1)$ operations.
-We prove this by contradiction.
-Assume by contradiction that both cases do not hold.
-Then, for each value $d$ of $l'$, there is either one $\textit{put}(d,\_)$ and one $\textit{rm}(d)$ in $l'$, or one $\textit{put}(d,\_)$ and no $\textit{rm}(d)$ in $l'$.
-It is obvious that such $l'$ does not contain unmatched $\textit{rm}$ operation, and renaming a sequence that does not contain unmatched $\textit{rm}$ we can only obtain another sequence that does not contain unmatched $\textit{rm}$.
-This contradicts that $l=r(l')$ has unmatched $\textit{rm}$.
-
-Let $l''$ be obtained from $ll$ by renaming $d_1$ to $b$ and renaming other value to $a$.
-If the first case holds, $l''$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$; Otherwise, the second case holds, and $l''$ is accepted by $\mathcal{A}_{\textit{unmRm2}}$.
-\qed
-\end {proof}
-
-Therefore, from now on, we can safely assume that there are no unmatched $\textit{rm}$ operations. The notion of $\textit{last}$ is defined as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $l$ contains $\textit{rmEmpty}$ operations, then $\textit{last}(l) = R_{\textit{pr5}}$.
-\item[-] Else, $l$ does not contain any $\textit{rmEmpty}$ operation.
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If the maximal priority of $l$ is an unmatched $\textit{put}$ operation $o$, and the priority of $o$ equals the priority of some matched $\textit{put}$ operations in $l$, then $\textit{last}(l) = R_{\textit{pr4}}$.
-
-    \item[-] If the maximal priority of $l$ is an unmatched $\textit{put}$ operation $o$, and the priority of $o$ is larger than the priority of all matched $\textit{put}$ operations in $l$, then $\textit{last}(l) = R_{\textit{pr3}}$.
-
-    \item[-] If the maximal priority of $l$ is a matched $\textit{put}$ operation $o$, the priority of $o$ equals the priority of some matched $\textit{put}$ operations in $l$, and the priority of $o$ is larger than the priority of all unmatched $\textit{put}$ operations in $l$, then $\textit{last}(l) = R_{\textit{pr2}}$.
-
-    \item[-] If the maximal priority of $l$ is an matched $\textit{put}$ operation $o$, and the priority of $o$ is larger than the priority of all $\textit{put}$ operations in $l$, then $\textit{last}(l) = R_{\textit{pr1}}$.
-
-    \item[-] Otherwise, $l = \epsilon$, and $\textit{last}(l) = R_{\textit{pr0}}$.
-    \end{itemize}
-\end{itemize}
-
-%The following lemma states that the inductive rules for priority queues are sound and complete.
-
-\section{Step-by-Step Linearizability of Priority Queues}
-\label{sec:step-by-step linearizability of priority queues}
-
-The following lemma states that $\textit{PQueue}$ is closed under projection. This lemma holds since the predicates appears in inductive rules of $\textit{PQueue}$ are closed under projection.
-
-\begin{restatable}{lemma}{PQisClosedUnderProjection}
-\label{lemma:PQ is closed under projection}
-
-$\textit{PQueue}$ is closed under projection.
-
-\end{restatable}
-
-%When the context is clear, we shorten priority-value data-independence as data-independence.
-For priority queue, the notion of step-by-step linearizability is modified as follows:
-
-%\vspace{-6pt}
-\begin{definition}[step-by-step linearizability for priority queue]\label{def:step-by-step linearizability of priority queue}
-A data structure $S=R_1, \ldots, R_n$ is said be to step-by-step linearizable, if for any data-differentiated sequence $e$, if $e$ is linearizable w.r.t.
-$\textit{MR}_{\textit{pri}}$ ($1 \leq i \leq 4$) with witness $x$, we have:
-
-$e \setminus x \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket \Rightarrow e \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$
-
-\noindent if $e$ is linearizable w.r.t. $\textit{MR}_{\textit{pr5}}$ with witness $x$, we have:
-
-$e \setminus x \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket \Rightarrow e \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$
-\end{definition}
-
-Let $\textit{proj}(s)$ denote the projection of $s$ obtained by erasing all the operations of several items from $s$.
-Here we assume that each $\textit{rm}(\textit{empty})$ is differentiated from each other, and we can consider each $\textit{rm}(\textit{empty})$ as having a unique item.
-
-\begin{restatable}{lemma}{PQasMultiInMRpriforSequence}
-\label{lemma:PQ as multi in MRpri for sequence}
-$s \in \textit{PQueue}$, if and only if, $\forall s'$, if $s' \in \textit{proj}(s)$, then $s' \in \textit{MR}_{\textit{pri}}$ for some $0 \leq i \leq 5$.
-\end{restatable}
-
-\begin {proof}
-
-The \textit{only if} direction is a direct consequence of Lemma \ref{lemma:PQ is closed under projection}.
-
-To prove the $\textit{if}$ direction.
-The $\textit{if}$ direction holds trivially when $s = \epsilon$.
-When $s \neq \epsilon$.
-since $s \in \textit{proj}(s)$, we generate sequence $s_1$ from $s$ as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $s$ has $\textit{rm}(\textit{empty})$ operations: $s_1$ is generated from $s$ by erasing some $\textit{rm}(\textit{empty})$ operation.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in $S_1$ and unmatched $\textit{put}$ operations of some items in $S_2$: $s_1$ is generated from $s$ by erasing the unmatched $\textit{put}$ operation of items last putted in $S_2$.
-
-\item[-] If the maximal priority in $s$ is unmatched $\textit{put}$ operations of some items in $S_2$ and $\vert S_2 \vert =1$: $s_1$ is generated from $s$ by erasing this unmatched $\textit{put}$ operation.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert \geq 2$: $s_1$ is generated from $s$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item which is last putted in $S$.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert = 1$: $s_1$ is generated from $s$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item in $S$.
-\end{itemize}
-
-Similarly, we obtain $s_2$ from $s_1$, obtain $s_3$ from $s_2$, $\ldots$, until we obtain some $s_m = \epsilon$.
-It is obvious that $s_m \in \textit{PQueue}$, and by the five rules of $\textit{PQueue}$, it is easy to see that $s_{\textit{m-1}},\ldots,s_1,s \in \textit{PQueue}$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{PQasMultiInMRpriforHistory}
-\label{lemma:PQ as multi in MRpri for history}
-Given a data-differentiated history $h$, assume that $\textit{PQueue}$ is step-by-step linearizability.
-$h \sqsubseteq \textit{PQueue}$, if and only if, $\forall h'$, if $h' \in \textit{proj}(h)$, then $h' \sqsubseteq \textit{MR}_{\textit{pri}}$, where $i = \textit{last}(h')$.
-\end{restatable}
-
-\begin {proof}
-
-The \textit{only if} direction is a direct consequence of Lemma \ref{lemma:PQ is closed under projection}.
-
-To prove the $\textit{if}$ direction.
-Since $h \sqsubseteq \textit{PQueue}$, we can see that $h \sqsubseteq l \in \textit{PQueue}$.
-Each projection $h'$ of $h$ can be linearizable to some projection $l'$ of $l$.
-By Lemma \ref{lemma:PQ as multi in MRpri for sequence}, $l' \in \textit{MR}_{\textit{prj}}$ for some $0 \leq j \leq 5$.
-Since $\textit{last}$ of a sequence is uniquely determined, we can see that $l' \in \textit{MR}_{\textit{pri}}$, where $i = \textit{last}(h')$.
-
-The $\textit{if}$ direction holds trivially when $h = \epsilon$. When $h \neq \epsilon$. since $h \in \textit{proj}(h)$, we generate sequence $h_1$ from $h$ as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $h$ has $\textit{rm}(\textit{empty})$ operations: Then $\textit{last}(h) = R_{\textit{pr5}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr5}}$.
-$h_1$ is generated from $h$ by erasing some $\textit{rm}(\textit{empty})$ operation.
-
-\item[-] If the maximal priority in $h$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in $S_1$ and unmatched $\textit{put}$ operations of some items in $S_2$: Then $\textit{last}(h) = R_{\textit{pr4}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr4}}$.
-$h_1$ is generated from $h$ by erasing a unmatched $\textit{put}(x,\_)$, such that $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$ for each $y \in S_2$.
-
-\item[-] If the maximal priority in $s$ is unmatched $\textit{put}$ operations of some items in $S_2$ and $\vert S_2 \vert =1$: Then $\textit{last}(h) = R_{\textit{pr3}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr3}}$.
-$h_1$ is generated from $h$ by erasing this unmatched $\textit{put}$ operation.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert \geq 2$: Then $\textit{last}(h) = R_{\textit{pr2}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr2}}$.
-$h_1$ is generated from $h$ by erasing $\textit{put}(x,\_)$ and $\textit{rm}(x)$, such that for each $y \in S$, $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$, and $\textit{rm}(x)$ does not happen before $\textit{rm}(y)$.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert = 1$: Then $\textit{last}(h) = R_{\textit{pr1}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr1}}$.
-$h_1$ is generated from $h$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item in $S$.
-\end{itemize}
-
-Similarly, we obtain $h_2$ from $h_1$, obtain $h_3$ from $h_2$, $\ldots$, until we obtain some $h_m = \epsilon$.
-Assume that for each $i$, when generating $h_i$ from $h_{\textit{i-1}}$, we erase operations of item $d_i$.
-It is obvious that $h_m = h_{\textit{m-1}} \setminus d_m \sqsubseteq \textit{PQueue}$, and since $h_{\textit{m-1}} \sqsubseteq \textit{MR}_{\textit{prj}_{\textit{m-1}}}$, where $\textit{last}(h_{\textit{m-1}}) = R_{\textit{prj}_{\textit{m-1}}}$, we can see that $h_{\textit{m-1}} \sqsubseteq \textit{PQueue}$.
-Similarly, we can prove that $h_{\textit{m-1}},\ldots,h_1,h \sqsubseteq \textit{PQueue}$.
-This completes the proof of the $\textit{if}$ direction.
-\qed
-\end {proof}
-
-To prove that $\textit{PQueue}$ is step-by-step linearizability, we use five lemmas for five rules of $\textit{PQueue}$.
-
-\begin{restatable}{lemma}{Rpr1isStepByStepLinearizability}
-\label{lemma:Rpr1 is step-by-step linearizability}
-
-$R_{\textit{pr1}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr1}}$ with witness $x$.
-Assume that  Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Then, let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-According to $R_{\textit{pr1}}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where the $\textit{put}$ operations and $\textit{rm}$ operations in $u \cdot v$ are matched, and $\textit{pri}_x$ is larger than any priority in $u \cdot v \cdot w$.
-Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$.
-Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$ operations, we can see that $D_L \cap D_R = \emptyset$.
-Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
-Let $S_1$ be the set of all operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_{L1}$ be the shortest prefix of $l'_L$ that contains all operations of $S_1$, and let $l'_{L2}$ be the sequence such that $l'_L = l'_{L1} \cdot l'_{L2}$.
-It is obvious that if $\textit{put}(x,\textit{pri}_x)$ happens before an operation $o$ in $h$, then $o$ is in $l'_{L2}$.
-Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.
-
-Since priority queue is closed under projection (Lemma \ref{lemma:PQ is closed under projection}) and $u \cdot v$ contain matched $\textit{put}$ and $\textit{rm}$ operations, we know that $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$.
-Since $\textit{pri}_x$ is larger than any priority in $h$, it is easy to see that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$.
-Since the priority queue is empty after executing $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$, we know that $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R \in \textit{PQueue}$.
-
-It remains to prove that $h \sqsubseteq l''$.
-To prove this, we define a graph $G$ whose nodes are the operations of $h$ and there is an edge from operation $o_1$ to $o_2$, if one of the following case holds
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o_1$ happens-before $o_2$ in h,
-
-\item[-] the method event corresponding to $o_1$ in $l''$ is before the one corresponding to $o_2$.
-\end{itemize}
-
-Assume there is a cycle in $G$.
-According the the property of interval order and the fact that the order of $l''$ is total, we know that there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $o_1,o_2 \in l'_{L1}$, or $o_2 \in l'_{L1} \wedge o_1 \in l'_{L2}$, or $o_1,o_2 \in l'_{L2}$, or $o_1,o_2 \in L'_R$, then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] If $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_{L1}$, then since $\textit{put}(x,\textit{pri}_x)$ happens before $o_2$, we know that $o_2 \in l'_{L2}$, which contradicts that $o_2 \in l'_{L1}$.
-
-\item[-] If $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L1}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L1}$, or $o_1 = \textit{rm}(x) \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 \in L'_R \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 = \textit{rm}(x)$, then $l$ contradicts with happen before relation of $h$.
-
-\item[-] If $o_1 \in l'_{L2} \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, then since $o_1$ happens before $\textit{put}(x,\textit{pri}_x)$, we know that $o_1 \in S_1$ and then $o_1 \in l'_{L1}$, which contradicts that $o_1 \in l'_{L2}$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
-
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr2isStepByStepLinearizability}
-\label{lemma:Rpr2 is step-by-step linearizability}
-
-$R_{\textit{pr2}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr2}}$ with witness $x$.
-Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-Since $h' \sqsubseteq l'$, there exists a sequence $h_{l'}$ of call, return and operations, such that (1) the projection of $h_{l'}$ into call and return action is $h$, (2) the projection of $h_{l'}$ into operations is $l'$, and (3) for each matched pair of call action $c$ and return action $r$ of $h$, there exists an operation $o$, such that $a$ before $o$, $o$ before $r$, and $o$ is the operation of $c$ and $r$.
-Intuitively, $h_{l'}$ is obtained from $h$ by adding linearization points.
-
-According to $R_{\textit{pr2}}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where (1) the $\textit{put}$ operations and $\textit{rm}$ operations in $u \cdot v$ are matched, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v \cdot w$, (3) there exists matched $\textit{put}$ and $\textit{rm}$ operations in $u \cdot v \cdot w$ that has priority $\textit{pri}_x$, (4) $\textit{pri}_x$ is larger than priorities of any unmatched put in $u \cdot v \cdot w$, and (5) all $\textit{put}$ operation with priority $\textit{pri}_x$ is in $u$.
-Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$.
-Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$ operations, we can see that $D_L \cap D_R = \emptyset$.
-Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
-
-Let $l'_{\textit{L0}}$ be the shortest prefix of $l'_L$, such that every operation in $l'_{\textit{L0}}$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $h_{l'}$.
-Let $\textit{OP}_x = \{ \textit{put}(d,\textit{pri}_x) \vert \textit{put}(d,\textit{pri}_x)$ is an operation of $h\}$.
-Let $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t$, where $l_t$ is the projection of $l'$ into operations of $\textit{OP}_x$ that are not in $l'_{\textit{L0}}$.
-Let $\textit{OP}'_x$ be the set of operations in $l_t$.
-Let $l'_{\textit{L2}}$ be the projection of $l'_L$ into operations of $l'_L$ that are not in $l'_{\textit{L1}}$.
-Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.
-
-As in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, we know that $l'_L \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$.
-It is obvious that $l'_{\textit{L0}} \in \textit{PQueue}$, and then $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t \in \textit{PQueue}$.
-Since there is no $\textit{rm}(\textit{Empty})$ in $h$ and $\textit{pri}_x$ is the maximal priority of $h$, $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$.
-Since all $\textit{put}$ operations with priority $\textit{pri}$ is in $l'_{L1}$, we know that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$.
-As in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, we know that $l'' \in \textit{PQueue}$.
-
-To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}.
-Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-
-\item[-] $o_1 \in L'_R \wedge (o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2} \vee o_2 = \textit{rm}(x))$, or $o_1 = \textit{rm}(x) \wedge (o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2})$: Then $l$ contradicts with happen before relation of $h$.
-
-\item[-] $o_1, o_2 \in l'_R$ or $o_1, o_2 \in l'_{L2}$: Then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] $o_1, o_2 \in l'_{L1}$: If $(o_1, o_2 \in l'_{L0}) \vee (o_1,o_2 \in \textit{OP}'_x)$, then $l'$ contradicts with happen before relation of $h$.
-
-    Else, it is the case that $o_1 \in \textit{OP}'_x \wedge o_2 \in l'_{L0}$.
-We can see that $o_1$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and the linearization point of $o_1$ is after the return action of $\textit{put}(x,\textit{pri}_x)$.
-Since $o_1 <_{\textit{hb}} o_2$, it is easy to see that $\textit{put}(x,\textit{pri}_x) <_{\textit{hb}} o_2$, and then $o_2 \notin l'_{L0}$, which contradicts that $o_2 \in l'_{L0}$.
-
-\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_{L1}$: Since $o_1 <_{\textit{hb}} o_2$, it is obvious that $o_2 \notin l'_{L1}$, contradicts that $o_2 \in l'_{L1}$.
-
-\item[-] $o_1 \in l'_{L2} \wedge o_2 \in l'_{L1}$: If $o_2 \in l'_{L0}$, then $l'$ contradicts with happen before relation of $h$.
-
-    Else, it is the case that $o_2 \in \textit{OP}'_x$.
-$o_2$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and the linearization point of $o_2$ is after the return action of $\textit{put}(x,\textit{pri}_x)$.
-Since $o_1 <_{\textit{hb}} o_2$, the return action of $o_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$.
-Then it is easy to see that $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.
-
-\item[-] $o_1 \in l'_{L2} \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: Then $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr3isStepByStepLinearizability}
-\label{lemma:Rpr3 is step-by-step linearizability}
-
-$R_{\textit{pr3}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr3}}$ with witness $x$.
-Assume that  Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Then, let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-According to $R_{\textit{pr3}}$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, where $\textit{pri}_x$ is the maximal priority in $u \cdot v$, and $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ operations in $u \cdot v$ (may equal priority of some unmatched $\textit{put}$ operations in $u \cdot v$).
-Let $S_1$ be the set of all operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_1$ be the shortest prefix of $l'$ that contains all operations of $S_1$, and let $l'_2$ be the sequence such that $l' = l'_1 \cdot l'_2$.
-Let sequence $l'' = l'_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_2$.
-
-Since $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ operations in $u \cdot v$, it is easy to see that $l'' \in \textit{PQueue}$.
-It remains to prove that $h \sqsubseteq l''$.
-To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}.
-Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o_1,o_2 \in l'_1$, or $o_1,o_2 \in l'_2$, or $o_1 \in l'_2 \wedge o_2 \in l'_1$: Then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] $o_1 \in l'_2 \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: $o_1 \in S_1$, and then $o_1 \in l'_1$, which contradicts that $o_1 \in l'_2$.
-
-\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_1$: We can see that in $l'$, $o_2$ must after every operation in $S_1$, and then $o_2 \notin l'_1$, contradicts that $o_2 \in l'_1$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
-\end {proof}
-
-The following lemma states that, from linearization of sub-histories, we can merge them and obtain a linearization (regardless of whether it belongs to sequential specification) of the whole history.
-
-\begin{restatable}{lemma}{MergeTwoLinearization}
-\label{lemma:merge two linearization}
-
-Given a history $h$ and operation set $S_1$, $S_2$, let $h_1 = h \vert{S_1}$ and $h_2 = h \vert{S_2}$.
-Assume that $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, and $S_1 \cup S_2$ contains all operations of $h$.
-Then, there exists a sequence $l$, such that $h \sqsubseteq l$, $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$.
-\end{restatable}
-
-\begin {proof}
-
-Given a history $h$ and a operation $o \in S_2$, let $\textit{MB}(o) = \{ o' \vert o' \in S_1$ and $o'$ happens before $o$ in $h \}$, let $\textit{SBI}(o) = \textit{min}\{ i \vert l_1[0,i]$ contains all elements of $\textit{MB}(o) \}$.
-
-Let $l = s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[n] \cdot s_{\textit{n+1}}$ be generated as follows, where $n = \vert l_2 \vert$:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $s_1 = l_1[0, \textit{SBI}(l_2[1])]$,
-
-\item[-] If $s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[i]$ already contains $l_1(\textit{SBI}(l_2[\textit{i+1}]))$, then $s_{\textit{i+1}} = \epsilon$.
-Otherwise, $s_{\textit{i+1}}$ is a subsequence of $l_1$, which starts from the next of last elements of $s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[i]$ in $l_1$ and ends in $l_1(\textit{SBI}(l_2[\textit{i+1}]))$.
-\end{itemize}
-
-It is obvious that $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$, and it remains to prove that $h \sqsubseteq l$.
-We prove this by contradiction.
-Assume that $h$ is not linearizable with respect to $l$.
-Then there must be two operations of $h$, such that $o_1 <_{hb} o_2$ in $h$ but $o_2$ before $0_1$ in $l$.
-Since $l \vert{S_1} = l_1$, $l \vert{S_2} = l_2$, and $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, it is easy to see that it is impossible that $o_1,o_2 \in S_1$ or $o_1,o_2 \in S_2$.
-There are only two possibilities:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o_1 \in S_1 \wedge o_2 \in S_2$.
-Then we can see that $o_2=l_2[i]$ and $o_1 \in s_j$ for some $i < j$.
-Since $o_1 <_{hb} o_2$, we know that $o_1 \in \textit{SBI}(o_2)$.
-By the construction of $l$, we know that $o_1$ must be in $s_k$ for some $k \leq i$, contradicts that $o_1 \in s_j$ with $i < j$.
-
-\item[-] $o_1 \in S_2 \wedge o_2 \in S_1$.
-Then we can see that $o_2 \in s_i$ and $o_1 = l_2[j]$ for some $i \leq j$.
-It is easy to see that this leads to contradiction when $i = j$.
-For the case of $i \neq j$, we need to satisfy the following requirements: (1) $o_1$ ($l_2[j]$) does not happen before $l_2[i]$, (2) $l_2[i]$ does not happen before $o_2$, (3) $o_2$ is either overlap or happens before $o' \in \textit{MB}(l_2[i])$, and (4) $o' <_{hb} l_2[i]$.
-By enumeration we can see that it is impossible that above four conditions be satisfied while $o_1 <_{hb} o_2$.
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr4isStepByStepLinearizability}
-\label{lemma:Rpr4 is step-by-step linearizability}
-
-$R_{\textit{pr4}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr4}}$ with witness $x$.
-Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-According to $R_{\textit{pr4}}$, there exist sequences $u$ and $v$, such that (1) $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v$, (3) $\textit{pri}_x$ equals the maximal priority of matched $\textit{put}$ operations in $u \cdot v$, (4) all matched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$ and (5) all the other unmatched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$.
-
-Since $h' \sqsubseteq l'$, there exists a sequence $h_{l'}$ of call, return and operations, such that (1) the projection of $h$ into call and return action is $h$, (2) the projection of $h_{l'}$ into operations is $l'$, and (3) for each matched pair of call action $c$ and return action $r$ of $h$, there exists an operation $o$, such that $a$ before $o$, $o$ before $r$, and $o$ is the operation of $c$ and $r$.
-Intuitively, $h_{l'}$ is obtained from $h$ by adding linearization points.
-
-According to $R_{\textit{pr4}}$, there exist sequences $u$ and $v$, such that (1) $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v$, (3) $\textit{pri}_x$ equals the maximal priority of matched $\textit{put}$ operations in $u \cdot v$, (4) all matched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$ and (5) all the other unmatched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$.
-
-Let $l'_1$ be the shortest prefix of $l'$, such that every operation in $l'_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $h_{l'}$.
-Let $\textit{OP}_x$ be the set of $\textit{put}$ operations with priority $\textit{pri}_x$ and their matched $\textit{rm}$ operations in $h$, let $\textit{OP}'_x$ be the set of $\textit{put}$ operations in $\textit{OP}_x$ that do not appear in $l'_1$, and let $\textit{OP}''_x$ be the set of $\textit{rm}$ operations in $\textit{OP}_x$ that do not appear in $l'_1$.
-Let $l_x''=l'_1 \cdot l'_2 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_3$, where $l'_2$ is the projection of $l'$ into operations of $\textit{OP}'_x$, and $l'_3$ is the projection of $l'$ into operations of $\textit{OP}''_x$.
-
-Let $h_x$ be the projection of $h$ into $\textit{OP}_x$.
-Let us prove that $h_x \sqsubseteq l''_x$ by contradiction.
-Assume that there exists operations $o_1,o_2$, such that $o_1$ happens before $o_2$ in $h_x$, and $o_2$ is before $o_1$ in $l''_x$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $(o_1 \in l'_3) \wedge (o_2 \in l'_1 \vee o_2 \in l'_2 \vee o'_2 \in l'_3)$, or $(o_1 \in l'_2) \wedge (o_2 \in l'_1 \vee o_2 \in l'_2)$, or $o_1,o_2 \in l'_1$: Then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_1$: Since the linearization point of $o_2$ is before return action of $o_1$, $o_2$ either overlaps with $o_1$, or happens before $o_1$, which contradicts that $o_1$ happens before $o_2$.
-
-\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_2$: Since $l'_2$ only contains $\textit{put}$ operations with priority $\textit{pri}_x$ and are in $\textit{OP}'_x$, it is not hard to see that each operation in $l'_2$ must overlap with $\textit{put}(x,\textit{pri}_x)$, which contradicts that $o_1$ happens before $o_2$.
-
-\item[-] $o_1 \in l'_3 \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: According to the construction of $l''$, we can see that $1_2 \in l'_1$, which contradicts that $1_2 \in l'_3$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h_x \sqsubseteq l''_x$.
-
-Let $\textit{OP}_{\textit{-x}}$ be the set of $\textit{put}$ operations with non-$\textit{pri}_x$ priority, and their matched $\textit{rm}$ operations.
-Let $h_{\textit{-x}}$ be the projection of $h$ into $\textit{OP}_{\textit{-x}}$.
-Let $l''_{\textit{-x}}$ be the projection of $l'$ into$\textit{OP}_{\textit{-x}}$.
-It is obvious that $h_{\textit{-x}} \sqsubseteq l''_{\textit{-x}}$.
-By Lemma \ref{lemma:merge two linearization}, there exists a sequence $l'''$, such that $h \sqsubseteq l'''$, $l''' \vert{\textit{OP}_x} = l_x''$ and $l''' \vert{\textit{OP}_{\textit{-x}}} = l''_{\textit{-x}}$.
-
-It is obvious that $l''_{\textit{-x}} \in \textit{PQueue}$.
-Let $l'_x$ be the projection of $l'$ into $\textit{OP}_x - \{ \textit{put}(x,\textit{pri}_x) \}$.
-It is obvious that $l'_x \in \textit{PQueue}$ and $l'_1 \in \textit{PQueue}$.
-Since (1) $l'_1 \cdot l'_2 \cdot l'_3$ is obtained from $l'_x$ by making $\textit{put}$ operations in $l'_2$ to happen earlier, and (2) $l'_3$ contains only $\textit{rm}$ operations, we can see that $l'_1 \cdot l'_2 \cdot l'_3 \in \textit{PQueue}$.
-Since $l'_3$ contains only $\textit{rm}$ operations, it is easy to see that $l_x'' \in \textit{PQueue}$.
-Since $\textit{pri}_x$ is the maximal priority of $h$ and $h$ does not contain $\textit{rm}(\textit{Empty})$ operations, the operations in $l_x''$ does not influence the correctness of $l''_{\textit{-x}} \in \textit{PQueue}$.
-Therefore, we can see that $l''' \in \textit{PQueue}$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr5isStepByStepLinearizability}
-\label{lemma:Rpr5 is step-by-step linearizability}
-
-$R_{\textit{pr5}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr5}}$ with operation $o=\textit{rmEmpty}$.
-Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Let $h'=h \setminus o$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-According to $R_{\textit{pr5}}$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{rmEmpty} \cdot v$, where the $\textit{put}$ operations and $\textit{rm}$ operations in $u$ are matched.
-
-Let $D_L$ be the set of items in $u$ and $D_R$ be the set of items in $v$.
-Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
-Let sequence $l'' = l'_L \cdot o \cdot L'_R$.
-Since priority queue is closed under projection (Lemma \ref{lemma:PQ is closed under projection}) and $u$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $l'_L \in \textit{PQueue}$ and the the priority queue is empty after executing $l'_L$.
-Then we know that $l'_L \cdot o \in \textit{PQueue}$.
-Since $l'_R$ is obtained from $l'$ by discarding pairs of matched $\textit{put}$ and $\textit{rm}$ operations, it is easy to see that $L'_R \in \textit{PQueue}$, and then we know that $l'' = l'_L \cdot o \cdot L'_R \in \textit{PQueue}$.
-
-It remains to prove that $h \sqsubseteq l''$.
-To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}.
-Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o_1,o_2 \in l'_L$, or $o_1,o_2 \in l'_R$: Then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] If $o_1=o \wedge o_2 \in l'_L$, or $o_1 \in l'_R \wedge o_2 \in l'_L$, or $o_1 \in l'_R \wedge o_2 = 0$, then $l$ contradicts with happen before relation of $h$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$. \qed
-\end {proof}
-
-The following lemma states that $\textit{PQueue}$ is step-by-step linearizability, it is a direct consequence of Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, Lemma \ref{lemma:Rpr2 is step-by-step linearizability}, Lemma \ref{lemma:Rpr3 is step-by-step linearizability}, Lemma \ref{lemma:Rpr4 is step-by-step linearizability} and Lemma \ref{lemma:Rpr5 is step-by-step linearizability}.
-
-\begin{restatable}{lemma}{PQueueisStepByStepLinearizability}
-\label{lemma:PQueue is step-by-step linearizability}
-
-$\textit{PQueue}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\section{Co-Regular of Priority Queues}
-\label{sec:co-regular of priority queues}
-
-\subsection{Co-Regular of $R_{\textit{pr1}}$}
-\label{subsec:co-regular of Rpr1}
-
-Let us introduce the notion of gap for matched $\textit{put}$ and $\textit{rm}$ operations.
-
-\begin{definition}[gap for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:gap for matched put and rm operations}
-
-Let $h$ be a data-differentiated history and $o_1 = \textit{put}(x,\textit{pri}_x)$, $o_2 = \textit{rm}(x)$ be a pair of matched $\textit{put}$ and $\textit{rm}$ operation of $h$.
-We say that $h$ has a gap on operations $o_1$ and $o_2$, if there is a partition of the operations of $h$ into $L$, $M$ and $R$ satisfying:
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $L \cup M$ has matched $\textit{put}$ and $\textit{rm}$ operations.
-
-\item[-] no operation of $R$ happens before an operation of $L$, $o_1$, $M$ or $o_2$.
-
-\item[-] no operation of $o_2$ happens before an operation of $L$, $o_1$ or $M$.
-
-\item[-] no operation of $M$ happens before an operation of $L$ or $o_1$.
-
-\item[-] no operation of $o_1$ happens before an operation of $L$.
-
-\item[-] no operation in $L \cup M \cup R$ has priority $\textit{pri}_x$.
-\end{itemize}
-\end{definition}
-
-Given a history $h$ and set $O$ of some operations in $h$.
-It is easy to see that there exists a sequence $l$ of operations of $O$, such that $l$ does not contradicts the happen before relation of $h$.
-Therefore, from now on, we can use this fact to directly generate such sequence from such $O$.
-
-The following lemma reduce linearizable with respect to $\textit{MR}_{\textit{pr1}}$ into gap of matched $\textit{put}$ and $\textit{rm}$ operations.
-
-\begin{restatable}{lemma}{GapEqualsLinforRpr1}
-\label{lemma:Gap Equals Lin for Rpr1}
-A data-differentiated history $h$ has a projection $h'$, such that $\textit{last}(h') = R_{\textit{pr1}}$ and $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$, if and only if there exists a projection $h''$ of $h$, such that $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is in $h''$, $\textit{pri}_x$ is the maximal priority of $h''$, and one the the following case holds:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] case $1$: $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$,
-\item[-] case $2$: there is no gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-\end{itemize}
-\end{restatable}
-
-\begin {proof}
-
-Let us prove the $\textit{only if}$ direction by contradiction.
-Assume that there exists such $h'$, while case $1$ and case $2$ do not hold for any projection of $h$.
-Since case $1$ does not hold for any projection of $h$, we know that $\forall y \in D$, $\textit{rm}(y)$ does not happens before $\textit{put}(y,\_)$ in $h$.
-Since case $2$ does not hold for any projection of $h$ and $\textit{last}(h') = R_{\textit{pr1}}$, we can see that $h'$ has gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$, while $\textit{pri}_x$ is the maximal priority in $h'$.
-Since $\textit{last}(h') = R_{\textit{pr1}}$, we can see that $\textit{pri}_x$ is larger than priority of all other $\textit{put}$ operations of $h$.
-
-By the properties of the gap, there is partitions $L$, $M$ and $R$.
-Let $u$, $v$ and $w$ contain the operations of $L$, $M$ and $R$, respectively.
-It is easy to see that $h'$ is linearizable with respect to $l''=u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$ and $l'' \in \textit{MR}_{\textit{pr1}}$.
-This contradicts the fact that $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$.
-
-Let us begin to prove the $\textit{if}$ direction.
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If case $1$ holds for $h''$, then $h \vert_{ \{ x \} }$ is the $h'$ as required.
-
-\item[-] If case $2$ holds, assume by contradiction that, for every projection $h'$ of $h$, if $\textit{last}(h') = R_{\textit{pr1}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr1}}$.
-Since $h''$ is also a projection of $h$, we know that there exist sequences $u$, $v$ and $w$, such that $h'' \sqsubseteq u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, $u \cdot v$ contains matched $\textit{put}$ and $\textit{rm}$ operations, and $\textit{pri}_x$ is larger than priority of $u \cdot v \cdot w$.
-Let $L$, $M$ and $R$ be the operations of $u$, $v$ and $w$, respectively.
-Then, the partitions $L$, $M$ and $R$ form a gap on operation $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-This contradicts the fact that there is no gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\begin{definition}[left-right constraint for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:left-right constraint for matched put and rm operations}
-Given a data-differentiated history $h$, and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-The left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is the graph $G$ where:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] the nodes are the items of $h$, to which we add a node,
-
-\item[-] there is an edge from item $d_1$ to $x$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$ and $\textit{pri}_1 \neq \textit{pri}_x$,
-
-\item[-] there is an edge from $x$ to item $d_1$, if $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$, and the priority of $d_1$ is not $\textit{pri}_x$,
-
-\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{rm}(d_2,\textit{pri}_2)$, $\textit{pri}_1 \neq \textit{pri}_x$, and $\textit{pri}_2 \neq \textit{pri}_x$.
-\end{itemize}
-\end{definition}
-
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-Let $\textit{LMSet}_1(h,x) = \{ o \vert$ $o$ is not an operation of $x$, the priority of $o$ does not equal the priority of $x$, and either it happens before $\textit{put}(x)$ or $\textit{rm}(x)$ in $h$, or there is an operation $o'$ with the same item of $o$, such that $o'$ happens before $\textit{put}(x)$ or $\textit{rm}(x)$ in $h$ $\}$.
-For each $i \geq 1$, let $\textit{LMSet}_{\textit{i+1}}(h,x) = \{ o \vert$ $o$ is not an operation of $x$ or items in $\textit{LMSet}_k(h,x)$ for each $k \leq i$, the priority of $o$ does not equal the priority of $x$, and either it happens before some operation $o' \in \textit{LMSet}_i(h,x)$ in $h$, or there is an operation $o''$ with the same item of $o$ and $o''$ happens before some operation $o' \in \textit{LMSet}_i(h,x)$ in $h$ $\}$.
-%We can see that for each $i$, $\textit{LMSet}_i(h,x)$ contains matched $\textit{put}$ and $\textit{rm}$ operations, and
-We can see that $\textit{LMSet}_i(h,x) \cap \textit{LMSet}_j(h,x) = \emptyset$ for any $i \neq j$.
-Let $\textit{LMSet}(h,x) = \textit{LMSet}_1(h,x) \cup \textit{LMSet}_2(h,x) \cup \ldots$.
-%{\color {red} We say that a history $h$ of priority queue is matched before $x$, if (1) $h$ contains no $\textit{rmEmpty}$ operations, (2) $ h \sqsubseteq l_1 \cdot \textit{put}(x,\_) \cdot l_2 \cdot \textit{rm}(x) \cdot l_3$ for some $l_1,l_2,l_3$, and $l_1 \cdot l_2$ contains matched $\textit{put}$ and $\textit{rm}$ operations.}
-
-The following is an example that shows that, given a data-differentiated history $h$, the set $\textit{LMSet}_i(h,x)$ can be non-empty for each $1 \leq i \leq 3$.
-In \figurename~\ref{fig:his nobound of LMSet} , $\textit{LMSet}_1(h,x) = \{ \textit{put}(a,1),\textit{rm}(a) \}$, $\textit{LMSet}_2(h,x) = \{ \textit{put}(b,1),\textit{rm}(b) \}$ and $\textit{LMSet}_3(h,x) = \{ \textit{put}(a,c),\textit{rm}(c) \}$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{PIC_HIS_NOBOUNDOF_LMSET.pdf}
-%\vspace{-10pt}
-  \caption{An example for non-emptiness of $\textit{LMSet}_1(h,x)$, $\textit{LMSet}_2(h,x)$ and $\textit{LMSet}_3(h,x)$}
-  \label{fig:his nobound of LMSet}
-\end{figure}
-
-\begin{restatable}{lemma}{RmxDoesNotHappenBeforeLMSetForRpr1}
-\label{lemma:Rmx does not happen before LMSet for Rpr1}
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-%, such that $\textit{pri}_x$ is the maximal priority of $h$, and $\textit{last}(h) = R_{\textit{pr1}}$.
-Assume that $\textit{rm}(y)$ does not happen before $\textit{put}(y,\_)$ for each item $y$ in $h$.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-Assume that $G$ has no cycle going through $x$.
-Then, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}(h,x)$.
-\end{restatable}
-
-\begin {proof}
-
-We prove this lemma by induction, and prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x)$, in $\textit{LMSet}_2(h,x)$, $\ldots$.
-Note that, according to the definition of $\textit{LMSet}(h,x)$, $\textit{pri}_x$ is different to priority of all items in $\textit{LMSet}(h,x)$.
-
-\noindent (1) Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x)$ by contradiction.
-Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{LMSet}_1(h,x)$ is an operation of item $d$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d$ does not equals $\textit{pri}_x$).
-
-We use a triple $(t_1,t_2,t_3)$ to represent related information.
-$t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$.
-$t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation.
-$t_2$ and $t_3$ is used for the reason of $o \in \textit{LMSet}_1(h,x)$: $o \in \textit{LMSet}_1(h,x)$, since an operation (of kind $t_2$) of $d$ happens before an operation (of kind $t_3$) of $x$.
-Let us consider all the possible cases of $(t_1,t_2,t_3)$:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.
-
-\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{rm}(x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{rm}(x)$.
-
-\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x) )$.
-By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(x,\textit{pri}_x) ) \vee ( \textit{rm}(d) <_{hb} \textit{put}(d,\_) )$, which is impossible.
-
-\item[-] $(\textit{put},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{rm}(x) )$.
-We can see that $\textit{rm}(d) <_{hb} \textit{rm}(x) <_{hb} \textit{put}(d,\_)$, which contradicts that $\textit{rm}(d)$ does not happen before $\textit{put}(d,\_)$.
-
-\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x) )$.
-We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.
-
-\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{rm}(x) )$.
-We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.
-
-\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.
-
-\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{rm}(x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{rm}(x)$.
-\end{itemize}
-
-This completes the proof for $\textit{LMSet}_1(h,x)$.
-
-\noindent (2) Assume we already prove that for some $j \geq 1$, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_{\textit{j+1}}(h,x)$ by contradiction.
-Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$ is an operation of item $d_{\textit{j+1}}$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d_{\textit{j+1}}$ does not equals $\textit{pri}_x$).
-We use a triple $(t_1,t_2,t_3)$ to represent related information.
-$t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$.
-$t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation.
-$t_2$ and $t_3$ is used for the reason of $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$: $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$, since an operation (of kind $t_2$) of $d_{\textit{j+1}}$ happens before an operation (of kind $t_3$) of $d_j$, where $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{LMSet}_j(h,x)$.
-Let us consider all the possible cases of $(t_1,t_2,t_3)$:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_)$.
-We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-
-\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j,\_)$.
-We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j,\_) ) \wedge ( \textit{rm}(d_j) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-
-\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$.
-By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \vee ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_{\textit{j+1}},\_) )$, which is impossible.
-
-\item[-] $(\textit{put},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$.
-By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \vee ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_{\textit{j+1}},\_) )$, which is impossible.
-
-\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) )$.
-Let us consider the reason of $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{LMSet}_j(h,x)$:
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $( j > 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(_{\textit{j-1}},\_), \textit{rm}(_{\textit{j-1}}) \in \textit{LMSet}_{\textit{j-1}}(h,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, we can see that $\textit{put}(d_{\textit{j+1}},\_) <_{hb} o''$, and then operations of $d_{\textit{j+1}}$ is in $\textit{LMSet}_j(h,x)$, contradicts that operations of $d_{\textit{j+1}}$ is in $\textit{LMSet}_{\textit{j+1}}(h,x)$.
-
-    \item[-] If $( j = 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of $x$: Similar to above case.
-
-    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{LMSet}_{\textit{j-1}}(h,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, we can see that $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} o'' ) \vee ( \textit{rm}(d_j) <_{hb} \textit{put}(d_j,\_) )$, which is impossible.
-
-    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of $x$: Similar to above case.
-    \end{itemize}
-
-\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Let $T_{\textit{ind}}$ be the set of sentences $\{ \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}), \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j),\ldots, \textit{put}(d_{\textit{ind+1}},\_) <_{hb} \textit{rm}(d_{\textit{ind}}) \}$.
-Here each $d_i$ is a item of some operation in $\textit{LMSet}_i(h,x)$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d_i$ does not equals $\textit{pri}_x$ for each $i$).
-Let us prove that from $T_j$ we can obtain contradiction by induction:
-
-    {\bf Base case $1$}: From $T_1$ we can obtain contradiction.
-
-    Let us prove base case $1$:
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{put}(d_1,\_)$ happens $o$, and $o$ is an operation of $x$.
-Then there is a cycle $x \rightarrow d_{\textit{j+1}} \rightarrow \ldots \rightarrow d_1 \rightarrow x$ in $G$, contradicts that $G$ has no cycle going through $x$.
-
-    \item[-] If $\textit{rm}(d_1)$ happens before $o$, and $o$ is an operation of $x$.
-Then since $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1) <_{hb} o$, we can see that $\textit{put}(d_2,\_) <_{hb} o$, and then $\textit{put}(d_2,\_) \in \textit{LMSet}_1(h,x)$, contradicts that $\textit{put}(d_2,\_) \in \textit{LMSet}_2(h,x)$.
-    \end{itemize}
-
-    {\bf Base case $2$}: From $T_2$ we can obtain contradiction.
-
-    Let us prove base case $2$: If $\textit{rm}(d_2) <_{hb} o$, and $o$ is an operation of $d_1$, then since $( \textit{put}(d_3,\_) <_{hb} \textit{rm}(d_2) ) \wedge ( \textit{rm}(d_2) <_{hb} o )$, we know that $\textit{put}(d_3,\_) <_{hb} o$.
-This implies that $\textit{put}(d_3,\_) \in \textit{LMSet}_2(h,x)$, contradicts that $\textit{rm}(d_3,\_) \in \textit{LMSet}_3(h,x)$.
-Therefore, it is only possible that $\textit{put}(d_2,\_)$ happens before an operation of $d_1$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{put}(d_1,\_)$ and $\textit{put}(d_1,\_)$ happens before operations of $x$, then we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, which is impossible.
-
-    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{put}(d_1,\_)$ and $\textit{rm}(d_1)$ happens before operations of $x$, then by interval order, we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, or $\textit{rm}(d_1) <_{hb} \textit{put}(d_1,\_)$, which is impossible.
-
-    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{put}(d_1,\_)$ happens before operations of $x$, then $x \rightarrow d_{\textit{j+1}} \rightarrow \ldots \rightarrow d_1 \rightarrow x$ in $G$, contradicts that $G$ has no cycle going through $x$.
-
-    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1)$ happens before operations of $x$, then we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, which is impossible.
-    \end{itemize}
-
-    {\bf induction step}: Given $\textit{ind} \geq 3$, if from $T_{\textit{ind-1}}$ we can obtain contradiction, then from $T_{\textit{ind}}$ we can also contain contradiction.
-
-    Prove of the induction step: Similarly as base case $2$, we can prove that it is only possible that $\textit{put}(d_{\textit{ind}},\_)$ happens before operations of $d_{\textit{ind-1}}$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$ and $\textit{put}(d_{\textit{ind-1}},\_)$ happens before operations of $d_{\textit{ind-2}}$, then we know that $\textit{put}(d_{\textit{ind}})$ happens before operation of $d_{\textit{ind-2}}$, which is impossible.
-
-    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$ and $\textit{rm}(d_{\textit{ind-1}})$ happens before operations of $d_{\textit{ind-2}}$, then by interval order, we know that $\textit{put}(d_{\textit{ind}},\_)$ happens before operation of $d_{\textit{ind-2}}$, or $\textit{rm}(d_{\textit{ind-1}}) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$, which is impossible.
-
-    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{rm}(d_{\textit{ind-1}})$, then we obtain $T_{\textit{ind-1}}$, which already contain contradiction.
-    \end{itemize}
-
-    By base case $1$, base case $2$ and the induction step, it is easy to see that $T_j$ contains contradiction.
-
-\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$.
-We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-
-\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$.
-We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \wedge ( \textit{rm}(d_j) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-\end{itemize}
-
-This completes the proof for $\textit{LMSet}_{\textit{j+1}}(h,x)$.
-Therefore, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}(h,x) = \textit{LMSet}_1(h,x) \cup \textit{LMSet}_2(h,x) \cup \ldots$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{LMSetHasMatchedPutandRmOperations}
-\label{lemma:LMSet has matched put and rm operations}
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-Assume that $\textit{rm}(y)$ does not happen before $\textit{put}(y,\_)$ for each item $y$ in $h$.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-Assume that $G$ has no cycle going through $x$.
-Then, $\textit{LMSet}(h,x)$ contains only matched $\textit{put}$ and $\textit{rm}$ operations.
-\end{restatable}
-\begin {proof}
-
-We prove this lemma by contradiction.
-Assume that there exists a value, such that $\textit{LMSet}(h,x)$ contains only its $\textit{put}$ operation and does not contain its $\textit{rm}$ operation.
-Then we can see that there exists $d_1,\ldots,d_j$.
-Intuitively, $d_1,\ldots,d_j$ are items in $\textit{LMSet}_1(h,x), \ldots, \textit{LMSet}_i(h,x)$, respectively.
-$\textit{LMSet}(h,x)$ contains $\textit{put}(d_j,\_)$ and does not contain $\textit{rm}(d_j)$.
-And each $d_i$ is the reason of $d_{\textit{i+1}} \in \textit{LMSet}_{\textit{i+1}}(h,x)$.
-Formally, we require that
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] For each $1 \leq i \leq j$, operations of $d_i$ belongs to $\textit{LMSet}_i(h,x)$. %For each $1 \leq i \leq j$, the priority of $d_i$ is not $\textit{pri}_x$.
-
-\item[-] For each $i \neq j$, $\textit{put}(d_i,\_),\textit{rm}(d_i) \in \textit{LMSet}_i(h,x)$. $\textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x)$, and $h$ does not contain $\textit{rm}(d_j)$.
-
-\item[-] An operation of $d_1$ happens before an operation of $x$. For each $1 < i \leq j$, an operation of $d_j$ happens an operation of $d_{\textit{i-1}}$.
-
-\item[-] For each $k$ and $\textit{ind}$, if $k > \textit{ind+1}$, then no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.
-\end{itemize}
-
-According to the definition of $\textit{LMSet}(h,x)$, it is easy to see that such $d_1,\ldots,d_j$ exists. Let us prove the following fact:
-
-\noindent {\bf $\textit{fact}_1$}: Given $1 \leq i < j$, it can not be the case that $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap.
-
-Proof of $\textit{fact}_1$: We prove $\textit{fact}_1$ by contradiction.
-Assume that for some $i \neq j$, $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap.
-Since $\textit{put}(d_i,\_), \textit{rm}(d_i) \in \textit{LMSet}_i(h,x)$, we know that an operation $o_i$ of $d_i$ happens before operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$.
-Moreover, since $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap, it is not hard to see that the call action of $\textit{put}(d_i,\_)$ and the call action of $\textit{rm}(d_i)$ is before the call action of $o_{\textit{i-1}}$.
-Since $\textit{put}(d_{\textit{i+1}},\_), \textit{rm}(d_{\textit{i+1}}) \in \textit{LMSet}_{\textit{i+1}}(h,x)$, we know that an operation $o'_{\textit{i+1}}$ of $d_{\textit{i+1}}$ happens before operation $o'_i$ of $d_i$.
-Then, it is not hard to see that $o'_{\textit{i+1}}$ also happens before $o_{\textit{i-1}}$, which contradicts that For each $k > \textit{ind+1}$, no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.
-
-We already know that an operation of $d_1$ happens before an operation of $x$.
-By $\textit{fact}_1$, we can ensure that $\textit{put}(d_1,\_)$ happens before an operation of $x$, and then $d_1 \rightarrow x$ in $G$.
-For each $1 < i < j$, we know that an operation $o_i$ of $d_i$ happens before an operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$.
-By $\textit{fact}_1$, we can ensure that $o_i=\textit{put}(d_i,\_)$ and $o_{\textit{i-1}}=\textit{rm}(d_{\textit{i-1}})$, and then $d_i \rightarrow d_{\textit{i-1}}$ in $G$.
-Since $h$ contains $\textit{put}(d_j)$ and does not contain $\textit{rm}(d_j)$, we know that $x \rightarrow d_j$ in $G$.
-Then $G$ has a cycle going through $x$, contradicts that $G$ has no cycle going through $x$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{GapEqualsConstraintforRpr1}
-\label{lemma:Gap Equals Constraint for Rpr1}
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-%, where $\textit{pri}_x$ has the maximal priority of $h$, and $\textit{last}(h) = R_{\textit{pr1}}$.
-Assume that for each item in $h$, its $\textit{rm}$ operation does not happen before its $\textit{put}$ operation.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x)$ and $\textit{rm}(x)$.
-There is a gap on $\textit{put}(x)$ and $\textit{rm}(x)$, if and only if $G$ has no cycle going through $x$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{only if}$ direction, assume that there is a gap on $\textit{put}(x)$ and $\textit{rm}(x)$, and let it be $L$, $M$ and $R$.
-Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ in $G$.
-It is obvious that the priority of each $d_i$ is not $\textit{pri}_x$.
-Then our proof proceeds as follows:
-
-According to the definition of left-right constraint, there are two possibilities.
-The first possibility is that, $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$.
-It is obvious that $\textit{rm}(d_1) \in R$, and then since $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_1),\textit{rm}(d_1) \in R$.
-Then,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Since $d_1 \rightarrow d_2$, by definition of $G$, we know that $\textit{put}(d_1)$ happens before $\textit{rm}(d_2)$.
-Since $\textit{put}(d_1) \in R$ and $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_2),\textit{rm}(d_2) \in R$.
-Similarly, for each $1 \leq i \leq m$, we know that $\textit{put}(d_i),\textit{rm}(d_i) \in R$.
-
-\item[-] Since $d_m \rightarrow x$,
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] if $\textit{put}(d_m)$ happens before $\textit{put}(x)$, then we can see that $\textit{put}(d_m) \in L$, which contradicts that $\textit{put}(d_m) \in R$.
-
-    \item[-] if $\textit{put}(d_m)$ happens before $\textit{rm}(x)$, then we can see that $\textit{put}(d_m) \in L \cup M$, which contradicts that $\textit{put}(d_m) \in R$.
-    \end{itemize}
-\end{itemize}
-
-The second possibility is that, $h$ contains one $\textit{put}(d_1,\_)$ operation and no $\textit{rm}(d_1)$ operation.
-Note that for each $j > 1$, $h$ contains $\textit{put}(d_j,\_)$ and $\textit{rm}(d_j)$.
-Since $d_m \rightarrow x$, is is obvious that $\textit{put}(d_m) \in L \cup M$.
-Since $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_m),\textit{rm}(d_m) \in R$.
-Then,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Since $d_{\textit{m-1}} \rightarrow d_m$, by definition of $G$, we know that $\textit{put}(d_{\textit{m-1}})$ happens before $\textit{rm}(d_m)$.
-Since $\textit{rm}(d_m) \in L \cup M$ and $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_{\textit{m-1}}),\textit{rm}(d_{\textit{m-1}}) \in L \cup M$.
-Similarly, for each $1 < i \leq m$, we know that $\textit{put}(d_i),\textit{rm}(d_i) \in L \cup M$, and $\textit{put}(d_1)\in L \cup M$.
-
-\item[-] Since $d_1 \rightarrow d_2$ in $G$, $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2)$.
-Since $\textit{put}(d_2),\textit{rm}(d_2) \in L \cup M$, we can see that $\textit{put}(d_1) \in L \cup M$.
-However, there is one $\textit{put}(d_1,\_)$ operation and no $\textit{rm}(d_1)$ operation in $h$, contradicts that $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
-\end{itemize}
-
-This completes the proof of the $\textit{only if}$ direction.
-
-To prove the $\textit{if}$ direction, assume that $G$ has no cycle going through $x$.
-Let $O_L$ be the set of operations that happen before $\textit{put}(x)$ in $h$.
-It is easy to see that $O_L \subseteq \textit{LMSet}(h,x)$.
-Let $O_M = \textit{LMSet}(h,x) \setminus O_L$.
-Let $O_h$ be the set of operations of $h$, and let $O_R = O_h \setminus \textit{LMSet}(h,x)$.
-Let $L$, $M$ and $R$ be $O_L$, $O_M$ and $O_R$, respectively.
-
-By Lemma \ref{lemma:LMSet has matched put and rm operations}, we can see that $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
-It remains to prove that $L$, $M$ and $R$ is a gap of $\textit{put}(x)$ and $\textit{rm}(x)$.
-We prove this by showing that all the following cases are impossible:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Case $1$: Some operation $o_r \in R$ happens before $\textit{rm}(x)$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $2$: Some operation $o_r \in R$ happens before some operation $o_{lm} \in L \cup M$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $3$: Some operation $o_r \in R$ happens before $\textit{put}(x)$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $4$: $\textit{rm}(x)$ happens before some $o_{lm} \in \textit{LMSet}(h,x) = L \cup M$. By Lemma \ref{lemma:Rmx does not happen before LMSet for Rpr1} we know that this is impossible.
-
-\item[-] Case $5$: $\textit{rm}(x)$ happens before $\textit{put}(x)$. This contradicts that for each item in $h$, its $\textit{rm}$ operation does not happen before its $\textit{put}$ operation.
-
-\item[-] Case $6$: Some operation $o_m \in M$ happens before $\textit{put}(x)$. Then we know that $o_m \in L$, which contradicts that $o_m \in M$.
-
-\item[-] Case $7$: Some operation $o_m \in M$ happens before some operation $o_l \in L$. Then we know that $o_m \in L$, which contradicts that $o_m \in M$.
-
-\item[-] Case $8$: $\textit{put}(x)$ happens before some operation $o_l \in L$. This is impossible.
-\end{itemize}
-
-This completes the proof of the $\textit{if}$ direction.
-
-\qed
-\end {proof}
-
-A automaton $\mathcal{A}_{\textit{Rpr1-1}}$ is given in \figurename~\ref{fig:automata for Rpro-1}, where $c=\textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm,a),$ $\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$.
-$\mathcal{A}_{\textit{Rpr1-1}}$ is used to detect the existence of a item whose $\textit{rm}$ happens before its $\textit{put}$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-1.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-1}}$}
-  \label{fig:automata for Rpro-1}
-\end{figure}
-
-Let us begin to represent several automata that is used for capture the case that, in a sub-history $h'$ of a history $h$, $\textit{last}(h')=R_{\textit{pr1}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$, and the reason is that there is no gap on $\textit{put}(y,\textit{pri}_y)$ and $\textit{rm}(y)$ in $h'$, where $y$ is the item that has maximal priority in $h'$.
-Such case is dealt with by four automata, since there are four possibilities of $h' \vert_{y}$.
-
-The first automaton is $\mathcal{A}_{\textit{Rpr1-pprr}}$, which is given in \figurename~\ref{fig:automata for Rpro-pprr}.
-In \figurename~\ref{fig:automata for Rpro-pprr}, $c_1= \textit{cal}(\textit{put},c,$ $\textit{anyPri}),\textit{ret}(\textit{put},c), \textit{cal}(\textit{rm},c), \textit{ret}(\textit{rm},c)$, $c_2 = c_1 + \textit{cal}(\textit{put},a,\textit{les}_p)$, and $c_3 = c_2 + \textit{ret}(\textit{rm},a)$.
-Here $p$ is a variable for priority, and $\textit{les}_p$ is a predicate that accepts all priorities that are less than the value of $p$.
-Here $\textit{pprr}$ represents that, $h' \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$ (we consider $b$ as $y$ and $p$ as $\textit{pri}_y$).
-In \figurename~\ref{fig:automata for Rpro-pprr}, the three paths that (1) from $q_{\textit{init}}$ to $q_2$ to $q_3$, (2) from $q_{\textit{init}}$ to $q_2$ to $q_{11}$, and (3) from $q_{\textit{init}}$ to $q_{10}$ to $q_{11}$ represents three possible positions of $\textit{ret}(\textit{put},a)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-pprr.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-pprr}}$}
-  \label{fig:automata for Rpro-pprr}
-\end{figure}
-
-The second automaton is $\mathcal{A}_{\textit{Rpr1-prpr}}$, which is given in \figurename~\ref{fig:automata for Rpro-prpr}.
-In $\mathcal{A}_{\textit{Rpr1-prpr}}$, $c_1, c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$.
-Here $\textit{pprr}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{put},b,p) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-prpr.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-prpr}}$}
-  \label{fig:automata for Rpro-prpr}
-\end{figure}
-
-The third automaton is $\mathcal{A}_{\textit{Rpr1-rppr}}$, which is given in \figurename~\ref{fig:automata for Rpro-rppr}.
-In $\mathcal{A}_{\textit{Rpr1-rppr}}$, $c_1, c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$.
-Here $\textit{rppr}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-rppr.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-rppr}}$}
-  \label{fig:automata for Rpro-rppr}
-\end{figure}
-
-The forth automaton is $\mathcal{A}_{\textit{Rpr1-rprp}}$, which is given in \figurename~\ref{fig:automata for Rpro-rprp}.
-In $\mathcal{A}_{\textit{Rpr1-rprp}}$, $c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$, and $c_4=c_1 + \textit{ret}(put,b)$.
-Here $\textit{rprp}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b) \cdot \textit{ret}(\textit{put})$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-rprp.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-rprp}}$}
-  \label{fig:automata for Rpro-rprp}
-\end{figure}
-
-Given a data-differentiated history $h$, we say that a item $x$ with priority $\textit{pri}_x$ in $h$ is covered by items $d_1,\ldots,d_m$ in $h$, if the priorities of $d_1,\ldots,d_m$ is not $\textit{pri}_x$, and
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{put}(d_m,\_)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$,
-
-\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,
-
-\item[-] $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$, or $\textit{rm}(d_1)$ does not exists in $h$
-\end{itemize}
-
-According to the definition of left-right constraint, in a data-differentiated history $h$, there is a cycle going through $x$, if and only if there exists items $d_1,\ldots,d_m$, such that $x$ is covered by $d_1,\ldots,d_m$.
-
-\begin{restatable}{lemma}{Rpr1IsCoRegular}
-\label{lemma:Rpr1 is co-regular}
-$R_{\textit{pr1}}$ is co-regular.
-\end{restatable}
-
-\begin {proof}
-
-Let $\textit{AutS-R}_{\textit{pr1}} = \{ \mathcal{A}_{\textit{Rpr1-1}} , \mathcal{A}_{\textit{Rpr1-pprr}}, \mathcal{A}_{\textit{Rpr1-prpr}}, \mathcal{A}_{\textit{Rpr1-rppr}}, \mathcal{A}_{\textit{Rpr1-rprp}} \}$.
-We need to prove that:
-
-\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr1}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr1}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr1}}$.
-
-By Lemma \ref{lemma:Gap Equals Lin for Rpr1} and Lemma \ref{lemma:Gap Equals Constraint for Rpr1}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:
-
-\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr1}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h''$ is a projection of $h$, such that $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is in $h''$, $\textit{pri}_x$ is the maximal priority of $h''$, and one the the following case holds:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] case $1$: $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$,
-\item[-] case $2$: $x$ is covered by some items $d_1,\ldots,d_m$ in $h''$.
-\end{itemize}
-
-\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\textit{AutS-R}_{\textit{pr1}}$.
-By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$.
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-1}}$: Let $d_1$ a item in $h_2$ such that $r(d_1)=b$ and let $h'' = h_2 \vert_{ \{ d_1 \} }$.
-$d_1$ is the item with maximal priority in $h''$.
-It is obvious that $\textit{last}(h'') = R_{\textit{pr1}}$ and $\textit{rm}(d_1)$ happens before $\textit{put}(d_1,\_)$ in $h''$.
-
-\item[-] If $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-pprr}}$, $\mathcal{A}_{\textit{Rpr1-prpr}}$, $\mathcal{A}_{\textit{Rpr1-rppr}}$, or $\mathcal{A}_{\textit{Rpr1-rprp}}$: Let $d_0$ be a item in $h_2$ such that $r(d_0)=b$, let $d_1,\ldots,d_m$ be the items in $h_2$ such that $r(d_i)=a$ for each $1 \leq i \leq m$.
-Let $h'' = h_1 \vert_{ \{ d_0, d_1, \ldots, d_m \} }$.
-It is obvious that the priority of $d_0$ is larger than the priorities of other operations in $h''$, and then $\textit{last}(h'') = R_{\textit{pr1}}$.
-It is easy to see that $d_0$ is covered by $d_1,\ldots,d_m$.
-\end{itemize}
-
-\noindent The $\textit{if}$ direction: Assume that there exists such $h$, $h''$, $x$ and $\textit{pri}_x$.
-Since $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{pri}_x$ is larger than the priorities of other operations in $h''$.
-Then,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If case $1$ holds: Then, $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$ in $h''$.
-Let $h_1$ be obtained from $h$ by renaming $x$ to $b$ and renaming all other items to $a$.
-By data-independence, $h_1 \in \mathcal{I}$, and it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-1}}$.
-\item[-] case $2$: Then, $x$ is covered by some items $d_1,\ldots,d_m$ in $h''$.
-There are four possibilities of $h'' \vert_{ \{ x \} }$: (1) $\textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{rm},x)$, (2) $\textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},x)$, (3) $\textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},x)$ and (4) $\textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{rm},x) \cdot \textit{ret}(\textit{put})$.
-
-    Let $h_1$ be obtained from $h$ by renaming $x$ into $b$, renaming $d_1,\ldots,d_m$ into $a$ and renaming other items into $c$.
-We can prove that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-pprr}}$, $\mathcal{A}_{\textit{Rpr1-prpr}}$, $\mathcal{A}_{\textit{Rpr1-rppr}}$ or $\mathcal{A}_{\textit{Rpr1-rprp}}$.
-
-    For example, if $h'' \vert_{ \{ x \} } = \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{rm},x)$, and $\textit{ret}(\textit{rm},d_m)$ is before $\textit{call}(\textit{put},x,\textit{pri}_x)$ and is after $\textit{ret}(\textit{put},x)$ in $h''$.
-Then in  $\mathcal{A}_{\textit{Rpr1-pprr}}$, $h_1$ is accepted by transitions from $q_{\textit{init}}$ to $q_2$ to $q_{11}$ and finally to $q_7$ (if $\textit{rm}(d_1)$ does not exists in $h''$) or $q_9$ (if $\textit{rm}(d_1)$ exists in $h''$).
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\subsection{Co-Regular of $R_{\textit{pr2}}$}
-\label{subsec:co-regular of Rpr2}
-
-An automaton $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$ is given in \figurename~\ref{fig:automata for FIFO of single priority}.
-In \figurename~\ref{fig:automata for FIFO of single priority}, let $c_1= \textit{cal}(\textit{put},c,\textit{anyPri}),\textit{ret}(\textit{put},c)$, $\textit{cal}(\textit{rm},c),\textit{ret}(\textit{rm},c)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, and $c_3 = c_2 + \textit{cal}(\textit{rm},a) + \textit{ret}(\textit{rm},a)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_QUEUE_FOR_SINGLE_PRI.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$}
-  \label{fig:automata for FIFO of single priority}
-\end{figure}
-
-According to \cite{Bouajjani:2015}, given a data-independent set $S$ of queue executions, whether every sequence $s$ in $S$ is linearizable to queue can be reduced into checking emptiness of intersection of $S$ and several finite automata.
-Let $\textit{transToQueue}(h)$ be a sequence $h'$ that is generated from $h$ by transforming $\textit{put}$ and $\textit{rm}$ into $\textit{enq}$ and $\textit{deq}$, respectively, and then discarding priorities.
-A priority queue history is called single-priority, if all of its items has only one priority.
-The following lemma states that we can also use finite automata to check ``first in first out property'' for single-priority sub-sequence of priority queue history.
-
-\begin{restatable}{lemma}{AutoForPQwithSignlePri}
-\label{lemma:automata for priority queue with single priority}
-
-Given a data-independent implementations $\mathcal{I}$ of priority queue, let $\mathcal{A} = \{ \mathcal{A}_{\textit{unmRm1}}, \mathcal{A}_{\textit{Rpr0-1}}, \mathcal{A}_{\textit{unmRm2}}, \mathcal{A}_{\textit{FIFO-Sig-Pri}} \}$, then,
-
-$\mathcal{I} \cap \mathcal{A} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority, and $\textit{transToQueue}(e')$ does not linearizable to queue.
-\end{restatable}
-
-\begin {proof}
-
-According to \cite{Bouajjani:2015}, checking linearizable with respect to queue can be reduced into checking the following three conditions for data-differentiated histories:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{deq}(a)$ happens before $\textit{enq}(a)$ (or $\textit{enq}(a)$ does not exist) for some item $a$,
-
-\item[-] there are two or more $\textit{deq}(a)$ for some item $a$,
-
-\item[-] $\textit{enq}(a)$ happens before $\textit{enq}(b)$, and $\textit{deq}(b)$ happens before $\textit{deq}(a)$ (or $\textit{deq}(a)$ does not exists) for some items $a,b$.
-\end{itemize}
-
-It is not hard to see that (1) the first condition is dealt with by $\mathcal{A}_{\textit{unmRm1}}$ (if $\textit{put}(a,\_)$ does not exist) and $\mathcal{A}_{\textit{Rpr0-1}}$ (if $\textit{put}(a,\_)$ exists), (2) the second condition is dealt with by $\mathcal{A}_{\textit{unmRm2}}$, and (3) the third condition is dealt with by $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$.
-\qed
-\end {proof}
-
-By Lemma \ref{lemma:automata for priority queue with single priority}, from now on, it is safe for us to assume that, for each data-differentiated priority queue history $h$ and each $h'$ of projection of $h$, if $h'$ is single-priority, then $\textit{transToQueue}(h')$ is linearizable to queue.
-One consequence of this assumption is that, in $h'$, there does not exists an item $a$, such that $\textit{rm}(a)$ happens before $\textit{put}(a)$.
-
-By the proof of Lemma \ref{lemma:Rpr1 is co-regular}, with the help of $\mathcal{A}_{\textit{Rpr1-pprr}}, \mathcal{A}_{\textit{Rpr1-prpr}}, \mathcal{A}_{\textit{Rpr1-rppr}}$ and $\mathcal{A}_{\textit{Rpr1-rprp}}$, from now on, we can safely assume that given a data-differentiated history $h$ and operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ in $h$, there does not exists any items $d_1,\ldots,d_m$, such that $x$ is not covered by items $d_1,\ldots,d_m$ and the priority of $d_1,\ldots,d_m$ is less that $x$.
-
-However, above two assumptions are not enough for imply the linearizability to priority queue.
-One example is shown in \figurename~\ref{fig:history introduct ob order1}.
-The history $h$ of \figurename~\ref{fig:history introduct ob order1} is not linearizable, even if $h \vert_{1}$ and $h \vert_{2}$ are both linearizable, and either $a$ or $b$ is covered by other items.
-One explanation is that, since $\textit{put}(b,1)$ happens before $\textit{put}(a,1)$, in choosing linearization points, we should make $\textit{rm}(b)$ before $\textit{rm}(a)$.
-However, to satisfy the requirements of $R_{\textit{pr2}}$, the linearization points of $\textit{rm}(a)$ can only locates in the interval from $\textit{ret}(\textit{rm},d)$ to $\textit{cal}(\textit{put},e,2)$.
-But, $\textit{rm}(b)$ is after this entire interval, and we can not locate the linearization point of $\textit{rm}(b)$ in this interval.
-When there is no such phenomenon, such as the history in \figurename~\ref{fig:history introduct ob order2}, the history becomes linearizable.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER1.pdf}
-%\vspace{-10pt}
-  \caption{above assumptions do not implies linearizability (1)}
-  \label{fig:history introduct ob order1}
-\end{figure}
-
-%\begin{figure}[htbp]
-%  \centering
-%  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER3.pdf}
-%\vspace{-10pt}
-%  \caption{above assumptions do not implies linearizability (2)}
-%  \label{fig:history introduct ob order1}
-%\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER2.pdf}
-%\vspace{-10pt}
-  \caption{above assumptions do not implies linearizability (2)}
-  \label{fig:history introduct ob order2}
-\end{figure}
-
-In the following part of this subsection, we use $<_{\textit{ob}}$ order for the case when we can infer that an item is inserted earlier than another item, and we use gap-point to represent the intervals when it is possible to locate linearizaton points of $\textit{rm}(a)$.
-
-\begin{definition}[gap-point for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:gap-point for matched put and rm operations}
-
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x),\textit{rm}(x)$ of $h$, we say $o$ is a gap-point of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$, if we can obtain a sequence $\textit{gapPSeq}(h)$ from $h$ by adding $o$ to some locations between actions of $h$, and satisfies that,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] In $\textit{gapPSeq}(h)$, $o$ is after $\textit{call}(\textit{put},x,\textit{pri}_x)$ and $\textit{call}(\textit{rm},x)$, and is before $\textit{ret}(\textit{rm},x)$.
-\item[-] For each item $y$ with non-$\textit{pri}_x$ priority in $\textit{gapPSeq}(h)$.
-If $\textit{put}(y,\_)$ and $\textit{rm}(y)$ is in $h$, then it can not be the case that $\textit{ret}(\textit{put},y,\_)$ is before $o$ and $\textit{call}(\textit{rm},x)$ is after $o$; Else if $\textit{put}(y,\_)$ is in $h$ and $\textit{rm}(y)$ is not in $h$, then it can not be the case that $\textit{ret}(\textit{put},y,\_)$ is before $o$;
-\end{itemize}
-
-We say that $o$ is a gap-point of priority $\textit{pri}_x$, if we can obtain a sequence $\textit{gapPSeq}(h)$ from $h$ by adding $o$ to some locations between actions of $h$, and satisfies only the latter requirement.
-\end{definition}
-
-\begin{restatable}{lemma}{GapEqualsGapPoint}
-\label{lemma:gap eqals the existence of gap point}
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-There is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, if and only if there is a gap-point $o$ of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{only if}$ direction, we need to prove that: if there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, then there is no gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-
-We prove this by contradiction.
-Assume that there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, but there is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-Since there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, in the time interval starts from $\textit{cal}(\textit{put},x,\textit{pri}_x)$ and $\textit{cal}(\textit{rm},x)$ and ends at $\textit{ret}(\textit{rm},x)$, there is always some item with non-$\textit{pri}_x$ priority in priority queue.
-Therefore, it is not hard to prove that there is a cycle going through the left-right constraint of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, and by Lemma \ref{lemma:Gap Equals Constraint for Rpr0}, there is no gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, which contradicts that there is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-
-To prove the $\textit{if}$ direction, assume that there is a gap-point $o$ of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$ and let $h_1$ be the sequence in definition of gap-point, which is generated from $h$ by adding $o$ into $h$.
-Let sequence $h_2$ be a sequence of call, return and operations, such that the projection of $h_2$ into call and return actions is $h$.
-$h_2$ is generated as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] First, from $h_1$, transform the point $o$ into $\textit{rm}(x)$.
-
-\item[-] Then for all the other items $y$, (1) adding an operation $\textit{put}(y,p)$ at some position between $\textit{call}(\textit{put},y,\_)$ and $\textit{ret}(\textit{put},y)$, and (2) adding an operation $\textit{rm}(y)$ at some position between $\textit{call}(\textit{rm},y)$ and $\textit{ret}(\textit{rm},y)$.
-
-\item[-] For all items $z$ with the same priority as $x$, the positions for adding operations of $z$ into $h$ is arbitrary (but need to satisfy above restriction.
-In the following we implicitly require this).
-
-\item[-] For all items $z$ whose priority is different from that of $x$,
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-
-    \item[-] If $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, and $\textit{ret}(\textit{put},z,\_)$ is before $o$ in $h_1$: Then in $h_2$, $\textit{rm}(z,\_)$ is just after $\textit{call}(\textit{rm},z)$.
-
-    \item[-] If $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, and $o$ is before $\textit{ret}(\textit{put},z,\_)$ $h_1$: Then in $h_2$, $\textit{put}(z,\_)$ is just before $\textit{ret}(\textit{put},z)$.
-
-    \item[-] Otherwise, we can see that $\textit{put}(z,\_)$ and $\textit{rm}(z)$ overlaps in $h$.
-Then we add $o_1=\textit{put}(z,\_)$ and $o_2=\textit{rm}(z)$ in the time interval of interleaving of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ overlaps in $h$, make sure $o_2$ next to $o_1$, and make sure either both of them is before $\textit{rm}(x)$ or both of them is after $\textit{rm}(x)$ in $h_2$.
-    \end{itemize}
-\end{itemize}
-
-According to the definition of gap-point, it is not hard to see that we can construct such $h_2$.
-Let $l$ be the projection of $h_2$ into operations, and $l'$ be the projection of $l$ into $x$ and all items that has priorities not equals that of $x$.
-Let $l'= u \cdot \textit{put}(x,\_) \cdot v \cdot \textit{rm}(x) \cdot w$.
-It is easy to see that $h \sqsubseteq l$, and $u \cdot v$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
-Let $L$, $M$ and $R$ be the set of operations in $u$, $v$ and $w$, respectively.
-Then it is obvious that $L$, $M$ and $R$ is a gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-\qed
-\end {proof}
-
-Given a data-differentiated history $h$ and two items $a,b$, we say that $a <_{\textit{ob}} b$, if the priority of $a$ and $b$ is the same, and one of the following cases holds:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Type $A$: $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$ in $h$,
-
-\item[-] Type $B$: $\textit{rm}(a)$ happens before $\textit{rm}(b)$ in $h$,
-
-\item[-] Type $C$: $\textit{rm}(a)$ happens before $\textit{put}(b,\_)$ in $h$,
-\end{itemize}
-
-Let $<_{\textit{ob}}^*$ be the transitive closure of $<_{\textit{ob}}$.
-Intuitively, $a <_{\textit{ob}} b$ means that $a$ should be inserted earlier than $b$, and $a <_{\textit{ob}}^* b$ means that we can infer that $a$ should be inserted earlier than $b$ with the help of other items.
-We can explicitly write $a <_{\textit{ob}}^A b$ to mean that $a <_{\textit{ob}} b$ and the reason is that $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$.
-Similarly, we can explicitly write $a <_{\textit{ob}}^B b$ and $a <_{\textit{ob}}^C b$ to mean that $a <_{\textit{ob}} b$ and the reason is that $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$ and $\textit{rm}(a)$ happens before $\textit{put}(b,\_)$, respectively.
-
-According to the definition of $<_{\textit{ob}}^*$, if $a <_{\textit{ob}}^* b$, then there exists $c_1,\ldots,c_n$, such that $a <_{\textit{ob}} c_1 <_{\textit{ob}} \ldots <_{\textit{ob}} b$.
-The following lemma states that, the number of intermediate items $c_i$ is in fact bounded.
-
-\begin{restatable}{lemma}{OBOrderHasBoundedLength}
-\label{lemma:ob order has bounded length}
-Given a data-differentiated history $h$.
-Assume that $a <_{\textit{ob}} a_1 <_{\textit{ob}} \ldots <_{\textit{ob}} a_m <_{\textit{ob}} b$, then either $a <_{\textit{ob}} b$, or $a <_{\textit{ob}} a_i <_{\textit{ob}} b$ for some $i$, or $a <_{\textit{ob}} a_i <_{\textit{ob}} a_j <_{\textit{ob}} b$ for some $i,j$.
-\end{restatable}
-
-\begin {proof}
-
-Our proof proceed as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^A$,$<_{\textit{ob}}^B \cdot <_{\textit{ob}}^B$ and $<_{\textit{ob}}^C \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^A c_2 <_{\textit{ob}}^A c_1$, then $\textit{put}(c_3,\_)$ happens before $\textit{put}(c_2,\_)$, and $\textit{put}(c_2,\_)$ happens before $\textit{put}(c_1,\_)$.
-Therefore, it is obvious hat $\textit{put}(c_3,\_)$ happens before $\textit{put}(c_1,\_)$ and $c_3 <_{\textit{ob}}^A c_1$.
-
-    Similarly, if $c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^B c_1$, then $c_3 <_{\textit{ob}}^B c_1$.
-
-    If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^C c_1$: Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
-Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
-
-    Therefore, when we meet successive $<_{\textit{ob}}^A$, it is safe to leave only the first and the last elements and ignore intermediate elements.
-Similar cases hold for $<_{\textit{ob}}^B$ and $<_{\textit{ob}}^C$.
-
-\item[-] $<_{\textit{ob}}^A$ and $<_{\textit{ob}}^C$:
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^A c_2 <_{\textit{ob}}^C c_1$.
-Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{ob}}^A c_2$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
-Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^A c_1$.
-
-    \item[-] ($<_{\textit{ob}}^C \cdot <_{\textit{ob}}^A$): If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^A c_1$.
-Since $c_2 <_{\textit{ob}}^A c_1$, $\textit{put}(c_2,\_)$ happens before $\textit{put}(c_1,\_)$.
-Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{rm}(c_3)$ happens before $\textit{put}(c_2,\_)$.
-Therefore, $\textit{rm}(c_3)$ happens before $\textit{put}(c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
-    \end{itemize}
-
-\item[-] $<_{\textit{ob}}^B$ and $<_{\textit{ob}}^C$:
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] ($<_{\textit{ob}}^B \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^C c_1$.
-Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-It is obvious that $\textit{cal}(\textit{rm},c_2)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$.
-Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
-
-    \item[-] ($<_{\textit{ob}}^C \cdot <_{\textit{ob}}^B$): If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^B c_1$.
-Since $c_2 <_{\textit{ob}}^B c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{rm},c_1)$.
-Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
-Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_1)$, and $c_3 <_{\textit{ob}}^B c_1$.
-    \end{itemize}
-
-\item[-]  ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^B \cdot <_{\textit{ob}}^A$): If $c_4 <_{\textit{ob}}^A c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^A c_1$:
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$: Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$.
-Then $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
-This implies that $c_4 <_{\textit{ob}}^A c_3 <_{\textit{ob}}^C c_1$.
-According to the fact for $<_{\textit{ob}}^A \cdot <_{\textit{ob}}^C$, we know that $c_4  <_{\textit{ob}}^A c_1$.
-
-    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is after $\textit{cal}(\textit{put},c_1,\_)$: Since $c_2 <_{\textit{ob}}^A c_1$, $\textit{ret}(\textit{put},c_2,\_)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{rm}(c_3)$ happens before $\textit{rm}(c_2)$, and then we know that $\textit{put}(c_2,\_)$ can not happen before $\textit{put}(c_3,\_)$.
-Since $\textit{put}(c_2,\_)$ does not happen before $\textit{put}(c_3,\_)$, $\textit{cal}(\textit{put},c_3,\_)$ is before $\textit{ret}(\textit{put},c_2,\_)$.
-Since $c_4 <_{\textit{ob}}^A c_3$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_3,\_)$.
-Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_4 <_{\textit{ob}}^A c_1$.
-    \end{itemize}
-
-\end{itemize}
-
-Based on above results, given $a <_{\textit{ob}}^{b_1} a_1 <_{\textit{ob}} \ldots <_{\textit{ob}}^{b_m} a_m <_{\textit{ob}}^{b_{\textit{m+1}}} b$, where each $b_i$ is in $\{ A,B,C \}$, we can merge relations, until we got one of the following facts:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $a <_{\textit{ob}}^A b$, $a <_{\textit{ob}}^B b$ or $a <_{\textit{ob}}^C b$,
-
-\item[-] $a <_{\textit{ob}}^A a_i <_{\textit{ob}}^B b$, or $a <_{\textit{ob}}^B a_i <_{\textit{ob}}^A b$, for some $i$,
-
-\item[-] $a <_{\textit{ob}}^B a_i <_{\textit{ob}}^A a_j <_{\textit{ob}}^B b$, for some $i$ and $j$,
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-Given a differentiated history $h$ and operations $\textit{put}(x,\_)$ and $\textit{rm}(x)$ of $h$, the time interval from the time point when both $\textit{call}(\textit{put},x,\_)$ and $\textit{cal}(\textit{rm},x)$ happen and ends at $\textit{ret}(\textit{rm},x)$ is called the effective interval of $x$ in $h$.
-
-\begin{restatable}{lemma}{Rpr2AsGPandOBOrder}
-\label{lemma:Rpr2 as gap-point and ob order}
-Given a data-differentiated history $h$.
-There exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr2}}$, and $h'$ does not linearizable to $\textit{MR}_{\textit{pr2}}$, if and only if there exists $h'' \in \textit{proj}(h)$, $\textit{last}(h'') = R_{\textit{pr2}}$, $x$ and $y$ has maximal priority in $h''$, $y <_{\textit{ob}} x$ in $h''$, and the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$ in $h''$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{if}$ direction, let $h_1$ be the history that is obtained from $h''$ by erasing all operations of items that has same priority as $x$, except for operations of $x$ and $y$.
-It is obvious that $\textit{last}(h_1) = R_{\textit{pr2}}$.
-According to $y <_{\textit{ob}} x$,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $y <_{\textit{ob}}^A x$, then $\textit{put}(y,\_)$ happens before $\textit{put}(x,\_)$.
-Since in $R_{\textit{pr2}}$, $\textit{put}(\textit{itm},\_)$ is after $\textit{put}$ operations of items with maximal priority (except for $\textit{itm}$), we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
-\item[-] If $y <_{\textit{ob}}^B x$, then $\textit{rm}(y)$ happens before $\textit{rm}(x)$.
-Since in $R_{\textit{pr2}}$, $\textit{rm}(\textit{itm})$ is the last $\textit{rm}$ operation of items with maximal priority, we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
-
-\item[-] If $y <_{\textit{ob}}^C x$, then $\textit{rm}(y)$ happens before $\textit{put}(x,\_)$.
-Since in $R_{\textit{pr2}}$, $\textit{rm}(\textit{itm})$ is the last $\textit{rm}$ operation of items with maximal priority, we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
-\end{itemize}
-
-Similarly as the proof in Lemma \ref{lemma:gap eqals the existence of gap point}, we can see that the only possible position for locating linearizaton point of $\textit{rm}(x)$ is at gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$.
-Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, then there exists unmatched operations before $\textit{rm}(x)$.
-
-Since the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},x)$ in $h''$ (and also in $h_1$), if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, then $\textit{rm}(x)$ will be before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},x)$.
-
-Therefore, for every sequence $l = u \cdot \textit{put}(x,\_) \cdot v \cdot \textit{rm}(x) \cdot w$, if $h_1 \sqsubseteq l$, then either $u \cdot v$ contains some unmatched operations of priority less than priority of $x$, or $w$ contains $\textit{put}(y,\_)$ or $\textit{rm}(y)$.
-In both cases, $l \notin \textit{MR}_{\textit{pr2}}$.
-
-To prove the $\textit{only if}$ direction, we prove its contrapositive.
-Assume we already know that for each $h'' \in \textit{proj}(h)$ with $\textit{last}(h'') = R_{\textit{pr2}}$, for each $x$ and $y$ has maximal priority in $h''$, if $y <_{\textit{ob}} x$ in $h''$, then the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},x)$ in $h''$.
-We need to prove that for each $h' \in \textit{proj}(h)$, if $\textit{last}(h') = R_{\textit{pr2}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr2}}$.
-
-Let $\textit{pri}$ be the maximal priority of $h'$.
-Recall that we already assume that each single-priority history (if we consider $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$) can be ``linearizable to queue'', and item with larger priority is not covered by items with smaller priority.
-Our proof proceed as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Let $h_{\textit{pri}}$ be the projection of $h$ into operations of priority $\textit{pri}$.
-Since each single-priority history can be ``linearizable to queue'', there exists sequence $l_{\textit{pri}}$, such that $h_{\textit{pri}} \sqsubseteq l_{\textit{pri}}$, and when we treat $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$, $l_{\textit{pri}}$ belongs to queue.
-
-\item[-] Step $1$: Let $a_1$ be the last inserted item of $l_{\textit{pri}}$.
-
-$\textit{Cond1}$: If for each item $b$ with priority $\textit{pri}$ in $h'$, the rightmost gap-point $o$ of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$ does not before any $\textit{cal}(\textit{put},b,\textit{pri})$ or $\textit{cal}(\textit{rm},b)$ in $h'$, then we can
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] Locate the linearization point of $\textit{rm}(a_1)$ at $o$,
-    \item[-] If $\textit{put}(a_1,\textit{pri})$ overlaps with $\textit{rm}(a_1)$, then locate the linearization point of $\textit{put}(a_1,\_)$ just before the linearization point of $\textit{rm}(a_1)$.
-Otherwise, $\textit{put}(a_1,\textit{pri}) <_{\textit{hb}} \textit{rm}(a_1)$, and we locate the linearization point of $\textit{put}(a_1,\_)$ just before its return action.
-    \item[-] Locate linearization points of operations of items of priority $\textit{pri}$ (except for $a_1$) just after the call action of the operation.
-    \item[-] For item $z$ with priority smaller than $\textit{pri}$.
-If both $\textit{cal}(\textit{put},z,\_)$ and $\textit{cal}(\textit{rm},z)$ is before $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just after their call actions.
-If both $\textit{ret}(\textit{put},z)$ and $\textit{ret}(\textit{rm},z)$ is after $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just before their return actions.
-
-        Otherwise, either (1) $\textit{rm}(z) <_{\textit{hb}} \textit{put}(z,\_)$, or (2) $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, $\textit{ret}(\textit{put},z,\_)$ before $o$, and $o$ before $\textit{call}(\textit{rm},z)$.
-The first case violates that each single-priority history can be ``linearizable to queue'', and the second case violates that $o$ is a gap-point.
-    \end{itemize}
-
-    In this way we construct a sequence $l$, and it is easy to see that $h' \sqsubseteq l$ and $l$ satisfies the requirement of $R_{\textit{pr2}}$.
-
-\item[-] Otherwise, there exists $a_2$ with priority $\textit{pri}$, such that the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$ in $h'$.
-We can see that the rightmost gap-point of $\textit{put}(a_2,\textit{pri})$ and $\textit{rm}(a_2)$ is after the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$.
-By assumption we know that $a_2$ does not $<_{\textit{ob}}$ to $a_1$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If for each item $b$ with priority $\textit{pri}$ in $h'$, $a_2$ does not $<_{\textit{ob}}$ to $b$.
-Then we go to step $1$ and treat $a_2$ as $a_1$ (When $\textit{Cond1}$ of step $1$ holds, we break the loop).
-    \item[-] Otherwise, there exists $a_3$ with priority $\textit{pri}$ such that $a_2 <_{\textit{ob}} a_3$.
-By assumption,we know that the rightmost gap-point of $\textit{put}(a_3,\textit{pri})$ and $\textit{rm}(a_3)$ is not before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$.
-Then we can see that the rightmost gap-point of $\textit{put}(a_3,\textit{pri})$ and $\textit{rm}(a_3)$ is after the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$.
-Then we go to step $1$ and treat $a_3$ as $a_1$ (When $\textit{Cond1}$ of step $1$ holds, we break the loop).
-    \end{itemize}
-\end{itemize}
-
-It is not hard to see that, if $b_i$ is the $a_1$ in the $\textit{i-th}$ loop of our proof, $b_j$ is the $a_1$ in the $\textit{j-th}$ loop of our proof, and $i<j$, then the rightmost gap-point of $\textit{put}(b_j,\textit{pri})$ and $\textit{rm}(b_j)$ is after the rightmost gap-point of $\textit{put}(b_i,\textit{pri})$ and $\textit{rm}(b_i)$.
-Thus, we finally stop at some $d$.
-$d$ satisfies that, for each item $c$ with priority $\textit{pri}$, $d$ does not $<_{\textit{ob}}$ to $c$.
-$d$ also satisfies that, for each item $c$ with priority $\textit{pri}$ in $h'$, the rightmost gap-point $o'$ of $\textit{put}(d,\textit{pri})$ and $\textit{rm}(d)$ does not before any $\textit{cal}(\textit{put},c,\textit{pri})$ or $\textit{cal}(\textit{rm},c)$.
-Then, we can
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Locate the linearization point of $\textit{rm}(d)$ at the rightmost gap-point $o'$ of $\textit{put}(d,\textit{pri})$ and $\textit{rm}(d)$,
-
-\item[-] Since for each item $c$ with priority $\textit{pri}$ in $h'$, $o'$ is after any $\textit{cal}(\textit{put},c,\textit{pri})$ and $\textit{cal}(\textit{rm},c)$.
-For each $c$, one of the following cases hold: (1) $\textit{cal}(\textit{rm},c)$ is before $o'$ and $o'$ is before $\textit{ret}(\textit{rm},c)$, (2) $\textit{cal}(\textit{rm},c)$ and $\textit{ret}(\textit{rm},c)$ are before $o'$.
-We locate the linearization point of $\textit{rm}(c)$ just after its call action.
-Such linearization point is obviously before $o'$.
-
-\item[-] If $o'$ is before $\textit{ret}(\textit{put},d,\textit{pri})$, then locate the linearization point of $\textit{put}(d,\textit{pri})$ just before $o'$.
-Else, $\textit{ret}(\textit{put},d,\textit{pri})$ is after $o'$, and we locate the linearization point of $\textit{put}(d,\textit{pri})$ just before the return action of it.
-
-\item[-] If $o'$ is before $\textit{ret}(\textit{put},d,\textit{pri})$, we can see that, for each item $c$ with priority $\textit{pri}$ in $h'$, $\textit{cal}(\textit{put},c,\textit{pri})$ is before $o'$.
-Then, and we locate the linearization point of $\textit{put}(c,\textit{pri})$ just before $o'$ and the linearization point of $\textit{put}(d,\textit{pri})$.
-
-    Otherwise, $\textit{ret}(\textit{put},d,\textit{pri})$ is before $o'$.
-Since $d$ does not $<_{\textit{ob}}$ to $c$, $\textit{put}(d,\textit{pri})$ does not happen before $\textit{put}(c,\textit{pri})$, or we can say, $\textit{ret}(\textit{put},d,\textit{pri})$ is after $\textit{cal}(\textit{put},c,\textit{pri})$.
-Then, we locate the linearization point of $\textit{put}(c,\textit{pri})$ just after its call action.
-\end{itemize}
-
-In this way we construct a sequence $l_d$, and it is easy to see that $h' \sqsubseteq l_d$ and $l_d$ satisfies the requirement of $R_{\textit{pr2}}$ with $d$ acts as $\textit{itm}$.
-This completes the proof of $\textit{if}$ direction.
-\qed
-\end {proof}
-
-Let us begin to represent several automata that is used for capture the case that, in a sub-history $h''$ of $h$, $\textit{last}(h'') = R_{\textit{pr2}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr2}}$, and the reason is that there exists items $a$ and $b$ with maximal priority in $h''$, $a <_{\textit{ob}} b$ in $h''$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
-
-According to Lemma \ref{lemma:ob order has bounded length}, if $a <_{\textit{ob}} b$, then one of the following cases hold: (1) $a <_{\textit{ob}}^A b$, (2) $a <_{\textit{ob}}^B b$, (3) $a <_{\textit{ob}}^C b$, (4) $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$ for some $a_1$, (5) $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$ for some $a_1$, or (6) $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$ for some $a_1$ and $a_2$.
-Although there are many possible enumeration of $a$, $b$, $a_1$ and $b_1$, the following lemma states that we need only to consider five enumeration of operations of $a$, $b$, $a_1$ and $b_1$.
-
-\begin{restatable}{lemma}{FiveEnmuerationisEnoughForRpr2}
-\label{lemma:five enumeration is enough for Rpr2}
-Given a data-differentiated history $h$, $\textit{last}(h) = R_{\textit{pr2}}$, and $h$ does not linearizable to $\textit{MR}_{\textit{pr2}}$.
-Assume that there exists items $a$ and $b$ with maximal priority in $h''$, $a <_{\textit{ob}} b$, and the rightmost gap-point $o$ of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
-Let $a_1$ and $a_2$ be the value when possibly $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$, or $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$.
-Then, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists) and $a_2$ (if exists).
-\end{restatable}
-\begin {proof}
-
-Let us prove by consider all the possible reason of $a <_{\textit{ob}}^A b$.
-According to Lemma \ref{lemma:ob order has bounded length}, we need to consider six reasons: %There may exists items with smaller priority of $a$ and covers some time interval, and assume that we rename them into $d$:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Reason $1$, $a <_{\textit{ob}}^A b$:
-
-    Since $a <_{\textit{ob}}^A b$, $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$.
-Since $o$ is after $\textit{cal}(\textit{put},a,\_)$, we can see that $o$ is before $\textit{cal}(\textit{rm},b)$.
-
-    Since single-priority must satisfy the first in firs out property, $\textit{rm}(b)$ does not happen before $\textit{rm}(a)$, and $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
-If $\textit{ret}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$, since we already assume that there exists gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$, this gap-point is also a gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$, and is after $o$, which contradicts that $o$ is the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$.
-Therefore, $\textit{ret}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$.
-
-    According to above discussion, there are two possible enumeration of operations of $a$ and $b$, as shown in \figurename~\ref{fig:history enumeration 1 for Rpr2} and \figurename~\ref{fig:history enumeration 2 for Rpr2}.
-Here we explicitly draw the leftmost gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$ as $o'$.
-Since the position of $\textit{ret}(\textit{put},b)$ does not influence the correctness, we can simply ignore it.
-
-\item[-] Reason $2$, $a <_{\textit{ob}}^B b$:
-
-    Since $a <_{\textit{ob}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$.
-Since $o$ is after $\textit{cal}(\textit{rm},b)$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$.
-Then, we can see that $o$ is before $\textit{cal}(\textit{put},a,\_)$.
-This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\_)$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible.
-Therefore, we can safely ignore this reason.
-
-\item[-] Reason $3$, $a <_{\textit{ob}}^C b$:
-
-    Since $a <_{\textit{ob}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},b)$.
-Since $o$ is after $\textit{cal}(\textit{put},b)$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$.
-Then, we can see that $o$ is before $\textit{cal}(\textit{put},a,\_)$.
-This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\_)$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible.
-Therefore, we can safely ignore this reason.
-
-\item[-] Reason $4$, $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$:
-
-    Since $a_1 <_{\textit{ob}}^B b$, $\textit{rm}(a_1) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_1)$ is before $\textit{cal}(\textit{rm},b)$.
-Since $\textit{rm}(a_1)$ does not happen before $\textit{put}(a_1)$, $\textit{cal}(\textit{put},a_1,\_)$ is before $\textit{ret}(\textit{rm},a_1)$.
-Since $a <_{\textit{ob}}^A a_1$, $\textit{ret}(\textit{put},a,\_)$ is before $\textit{cal}(\textit{put},a_1,\_)$.
-Therefore, $\textit{ret}(\textit{put},a,\_)$ is before $\textit{cal}(\textit{rm},b)$.
-Since $\textit{cal}(\textit{rm},b)$ is before $o$, we can see that $\textit{ret}(\textit{put},a,\_)$ is before $o$.
-Then, we can see that $o$ is before $\textit{cal}(\textit{rm},a)$.
-
-    If $\textit{cal}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$, then $\textit{rm}(b) <_{\textit{hb}} \textit{rm}(a)$.
-Let $h' = h \vert_{ \{ a,a_1,b \} }$.
-It is not hard to see that $h'$ can not linearizable to a sequence that satisfies the first in first our property, and this contradicts to our assumption that each single-priority history satisfies the first in first out property.
-Therefore, $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
-Similarly as the case of reason $1$, we can see that $\textit{ret}(\textit{rm},b)$ is before $\textit{ret}(\textit{rm},a)$.
-
-    According to above discussion, there are three possible enumeration of operations of $a$, $a_1$ and $b$, as shown in \figurename~\ref{fig:history enumeration 3 for Rpr2}, \figurename~\ref{fig:history enumeration 4 for Rpr2} and \figurename~\ref{fig:history enumeration 5 for Rpr2}.
-Here we explicitly draw the leftmost gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$ as $o'$.
-Since the position of $\textit{cal}(\textit{put},b,\_)$, $\textit{ret}(\textit{put},b)$, $\textit{ret}(\textit{put},a_1,\_)$ and $\textit{cal}(\textit{put},a,\_)$ do not influence the correctness, we can simply ignore it.
-
-\item[-] Reason $5$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$:
-
-    Since $a_1 <_{\textit{ob}}^A b$, $\textit{ret}(\textit{put},a_1,\_)$ is before $\textit{call}(\textit{put},b,\_)$.
-Since $\textit{call}(\textit{put},b,\_)$ is before $o$, we can see that $\textit{ret}(\textit{put},a_1,\_)$ is before $o$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $o$ is before $\textit{cal}(\textit{rm},a)$: Then $o$ is obviously before $\textit{ret}(\textit{rm},a)$.
-Since $a <_{\textit{ob}}^B a_1$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},a_1)$.
-Then we can see that, $o$ is before $\textit{cal}(\textit{rm},a_1)$, and remember that $a_1 <_{\textit{ob}}^A b$.
-Then we can goto the case of reason $1$ and treat $a_1$ as $a$.
-Therefore, we can safely ignore this.
-
-    \item[-] If $o$ is before $\textit{cal}(\textit{put},a,\_)$: Since $\textit{rm}(a)$ does not happen before $\textit{put}(a,\_)$, we can see that $\textit{cal}(\textit{put},a,\_)$ is before $\textit{ret}(\textit{rm},a)$, and then $o$ is before $\textit{ret}(\textit{rm},a)$.
-Then similarly as above case, we can see that $o$ is before $\textit{cal}(\textit{rm},a_1)$, and $a_1 <_{\textit{ob}}^A b$.
-Then we can goto the case of reason $1$ and treat $a_1$ as $a$.
-Therefore, we can safely ignore this.
-    \end{itemize}
-
-\item[-] Reason $6$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$:
-
-    Since $a_2 <_{\textit{ob}}^B b$, $\textit{rm}(a_2) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_2)$ is before $\textit{cal}(\textit{rm},b)$.
-%Since $\textit{rm}(a_2)$ does not happen before $\textit{put}(a_2,\_)$, $\textit{cal}(\textit{put},a_2,\_)$ is before $\textit{ret}(\textit{rm},a_2)$.
-%Since $a_1 <_{\textit{ob}}^A a_2$, $\textit{ret}(\textit{put},a_1,\_)$ is before $\textit{cal}(\textit{put},a_2,\_)$.
-    Since $a_1 <_{\textit{ob}}^A a_2$, $\textit{put}(a_1,\_) <_{\textit{hb}} \textit{put}(a_2,\_)$, and $\textit{rm}(a_2)$ does not happen before $\textit{rm}(a_1)$, and $\textit{cal}(\textit{rm},a_1)$ is before $\textit{ret}(\textit{rm},a_2)$.
-Since $a <_{\textit{ob}}^B a_1$, $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(a_1)$, and $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},a_1)$.
-
-    Therefore, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$.
-Since $\textit{cal}(\textit{rm},b)$ is before $o$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$.
-Then it is easy to see that $o$ is before $\textit{cal}(\textit{put},a,\_)$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\_)$, and $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a,\_)$.
-This is impossible, and we can safely ignore this reason.
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-1.pdf}
-%\vspace{-10pt}
-  \caption{The first possible enumeration.}
-  \label{fig:history enumeration 1 for Rpr2}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-2.pdf}
-%\vspace{-10pt}
-  \caption{The second possible enumeration.}
-  \label{fig:history enumeration 2 for Rpr2}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-3.pdf}
-%\vspace{-10pt}
-  \caption{The third possible enumeration.}
-  \label{fig:history enumeration 3 for Rpr2}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-4.pdf}
-%\vspace{-10pt}
-  \caption{The forth possible enumeration.}
-  \label{fig:history enumeration 4 for Rpr2}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-5.pdf}
-%\vspace{-10pt}
-  \caption{The fifth possible enumeration.}
-  \label{fig:history enumeration 5 for Rpr2}
-\end{figure}
-
-Given a data-differentiated history $h$, two actions $\textit{act}_1$, $\textit{act}_2$ of items of priority $\textit{pri}$, and assume that $\textit{act}_1$ is before $\textit{act}_2$ in $h$.
-We say that $\textit{act}_1$, $\textit{act}_2$ is covered by items $d_1,\ldots,d_m$ in $h$, if the priorities of $d_1,\ldots,d_m$ is not $\textit{pri}$, and
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{ret}(\textit{put},d_m,\_)$ is before $\textit{act}_1$,
-
-\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,
-
-\item[-] $\textit{act}_2$ is before $\textit{cal}(\textit{rm},d_1)$.
-\end{itemize}
-
-According to the definition of gap-point, in \figurename~\ref{fig:history enumeration 1 for Rpr2}, \figurename~\ref{fig:history enumeration 2 for Rpr2}, \figurename~\ref{fig:history enumeration 3 for Rpr2}, \figurename~\ref{fig:history enumeration 4 for Rpr2} or \figurename~\ref{fig:history enumeration 5 for Rpr2}, the location of $o$ and $o'$ is equivalent to that, there exists items $d_1,\ldots,d_m$ in $h$ in $h$, such that $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ is covered by $d_1,\ldots,d_m$, $\textit{cal}(\textit{rm},b)$ is before $\textit{ret}(\textit{put},d_m,\_)$ in $h$, and $\textit{cal}(\textit{rm},d_1)$ is before $\textit{ret}(\textit{rm},a)$ in $h$.
-We say that such $d_1,\ldots,d_m$ constitute the rightmost gap of $\textit{put}(b,\_)$ and $\textit{rm}(b)$.
-
-An automaton $\mathcal{A}_{\textit{Rpr2-1}}$ is given in \figurename~\ref{fig:automata for first enumeration of Rpr2}, and it is constructed for the first enumeration in \figurename~\ref{fig:history enumeration 1 for Rpr2}.
-In \figurename~\ref{fig:automata for first enumeration of Rpr2}, we renames the items that covers $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ into $d$, and rename the renaming items into $e$.
-Here let $c = \textit{cal}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{cal}(\textit{rm},e), \textit{ret}(\textit{rm},e)$, $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_1.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr2-1}}$}
-  \label{fig:automata for first enumeration of Rpr2}
-\end{figure}
-
-An automaton $\mathcal{A}_{\textit{Rpr2-2}}$ is given in \figurename~\ref{fig:automata for second enumeration of Rpr2}, and it is constructed for the second enumeration in \figurename~\ref{fig:history enumeration 2 for Rpr2}.
-In \figurename~\ref{fig:automata for second enumeration of Rpr2}, $c_1$, $c_2$, $c_3$ and $c_4$ is same as that in \figurename~\ref{fig:automata for first enumeration of Rpr2}.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_2.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr2-2}}$}
-  \label{fig:automata for second enumeration of Rpr2}
-\end{figure}
-
-For the third enumeration in \figurename~\ref{fig:history enumeration 3 for Rpr2}.
-Since we want to ensure that $a$ and $b$ are putted only once, we need to explicitly record the positions of $\textit{cal}(\textit{put},a,p)$ and $\textit{cal}(\textit{put},b,p)$.
-Since the positions of $\textit{cal}(\textit{put},a,p)$ and $\textit{cal}(\textit{put},b,p)$ are not fixed, there are finite possible cases to consider, as shown below:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{rm},b)$ and before $\textit{cal}(\textit{rm},a)$: There are two possible positions of $\textit{cal}(\textit{put},a,p)$: (1) before $\textit{cal}(\textit{rm},a_1)$, and (2) after $\textit{cal}(\textit{rm},a_1)$, and before $\textit{ret}(\textit{put},a)$.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$ and before $\textit{cal}(\textit{rm},b)$: same as above case.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{put},a_1,\_)$ and before $\textit{ret}(\textit{rm},a_1)$: same as above case.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{put},a)$ and before $\textit{cal}(\textit{put},a_1,p)$: same as above case.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$: There are three possible positions of $\textit{cal}(\textit{put},a,p)$: (1) after $\textit{cal}(\textit{put},b,p)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{cal}(\textit{put}b,p)$, and (3) before $\textit{call}(\textit{rm},a_1)$.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is before $\textit{cal}(\textit{rm},a_1)$: There are three possible positions of $\textit{cal}(\textit{put},a,p)$: (1) after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{cal}(\textit{put},b,p)$ and before $\textit{cal}(\textit{rm},a_1)$, and (3) before $\textit{cal}(\textit{put},b,p)$.
-\end{itemize}
-
-Therefore, there are fourteen possible cases that satisfy the third enumeration in \figurename~\ref{fig:history enumeration 3 for Rpr2}.
-For each case, we construct an finite automaton.
-Let $\textit{Aut}_{\textit{Rpr2-3}}$ be the set of finite automata that is constructed for above fourteen cases.
-For example, for the case $\textit{ca}_1$ when $\textit{cal}(\textit{put},a,p)$ is before $\textit{cal}(\textit{rm},a_1)$, $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$, and $\textit{cal}(\textit{put},b,p)$ is before $\textit{cal}(\textit{rm},b)$, we construct a finite automaton in \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}.
-In \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}, let $c$ and $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$ the same as that in \figurename~\ref{fig:automata for first enumeration of Rpr2}.
-Let $c_2 = c_1 + \textit{ret}(\textit{put},a_1)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$, $c_4 = c_3 + \textit{ret}(\textit{rm},d)$, and $c_5 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{put},a_1) + \textit{ret}(\textit{rm},d)$.
-The other finite automata in $\textit{Aut}_{\textit{Rpr2-3}}$ can be similarly constructed.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_3.pdf}
-%\vspace{-10pt}
-  \caption{Automaton for $\textit{ca}_1$}
-  \label{fig:automata for ca1 of third enumeration of Rpr2}
-\end{figure}
-
-Similarly, we construct sets $\textit{Aut}_{\textit{Rpr2-4}}$ and $\textit{Aut}_{\textit{Rpr2-5}}$ of finite automata for the forth enumeration in \figurename~\ref{fig:history enumeration 4 for Rpr2} and the fifth enumeration in \figurename~\ref{fig:history enumeration 5 for Rpr2}, respectively.
-
-\begin{restatable}{lemma}{Rpr2IsCoRegular}
-\label{lemma:Rpr2 is co-regular}
-$R_{\textit{pr2}}$ is co-regular.
-\end{restatable}
-
-\begin {proof}
-
-Let $\textit{AutS-R}_{\textit{pr2}} = \{ \mathcal{A}_{\textit{Rpr2-1}}, \mathcal{A}_{\textit{Rpr2-2}} \} \cup \textit{Aut}_{\textit{Rpr2-3}} \cup \textit{Aut}_{\textit{Rpr2-4}} \cup \textit{Aut}_{\textit{Rpr2-5}}$.
-We need to prove that:
-
-\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr2}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr2}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr2}}$.
-
-By Lemma \ref{lemma:Rpr2 as gap-point and ob order}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:
-
-\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr0}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, such that $\textit{last}(h') = R_{\textit{pr2}}$, $a$ and $b$ has maximal priority in $h'$, $a <_{\textit{ob}} b$ in $h'$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
-
-\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\textit{AutS-R}_{\textit{pr2}}$.
-By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$.
-Since $h_1$ is accepted by $\textit{AutS-R}_{\textit{pr2}}$, let $x$ and $y$ be the items that are renamed into $b$ and $a$ by $r$, respectively, and let $d_1,\ldots,d_m$ be the items that are renamed into $d$ by $r$.
-
-let $h'' = h_2 \vert_{ \{ x,y,d_1,\ldots,d_m \} }$.
-It is obvious that $\textit{last}(h'') = R_{\textit{pr2}}$.
-According to our construction of automata in $\textit{AutS-R}_{\textit{pr2}}$, it is not hard to see that $x$ and $y$ has maximal priority in $h_2$, $y <_{\textit{ob}} x$ in $h''$, and the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$ in $h''$.
-%By Lemma \ref{lemma:Rpr2 as gap-point and ob order}, we know that $h''$ is not linearizable to $\textit{MR}_{\textit{pr2}}$.
-
-\noindent The $\textit{if}$ direction: Assume that there exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr2}}$, $a'$ and $b'$ has maximal priority in $h'$, $a' <_{\textit{ob}} b'$ in $h'$, and the rightmost gap-point of $\textit{put}(b',\_)$ and $\textit{rm}(b')$ is before $\textit{cal}(\textit{put},a',\_)$ or $\textit{cal}(\textit{rm},a')$ in $h''$.
-By data-independence, we can obtain history $h_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the items $d_1,\ldots,d_m$ that constitute the rightmost gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, we rename them into $d$, and (3) rename the other items into $e$.
-It is easy to see that $\textit{last}(h_1) = R_{\textit{pr2}}$, $a$ and $b$ has maximal priority in $h_1$, $a <_{\textit{ob}} b$ in $h_1$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
-By Lemma \ref{lemma:five enumeration is enough for Rpr2}, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists) and $a_2$ (if exists).
-Then
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $a <_{\textit{ob}} b$ because of the first enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr2-1}}$.
-
-\item[-] If $a <_{\textit{ob}} b$ because of the second enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr2-2}}$.
-
-\item[-] If $a <_{\textit{ob}} b$ because of the third enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-3}}$.
-
-\item[-] If $a <_{\textit{ob}} b$ because of the forth enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-4}}$.
-
-\item[-] If $a <_{\textit{ob}} b$ because of the fifth enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-5}}$.
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\subsection{Co-Regular of $R_{\textit{pr3}}$}
-\label{subsec:co-regular of Rpr3}
-
-\begin{restatable}{lemma}{Rpr3IsAlwaysCoRegular}
-\label{lemma:Rpr3 is always co-regular}
-Given a data-differentiated history $h$, if $\textit{last}(h) = R_{\textit{pr3}}$, then $h \sqsubseteq \textit{MR}_{\textit{pr3}}$.
-\end{restatable}
-
-\begin {proof}
-
-Since $\textit{last}(h) = R_{\textit{pr3}}$, there exists item $x$ with $\textit{pri}_x$ in $h$, such that $h \vert_{ \{ x \}} = \textit{put}(x,\textit{pri}_x)$, and $\textit{pri}_x$ is larger than any priority of $h$.
-
-For each operation, we locate its linearization point just after its call action.
-In such way we obtain a sequence $l$ of operations, and it is easy to see that $h \sqsubseteq l$.
-It is obvious that $l$ contains $\textit{put}(x,\textit{pri})$, and then $l \in \textit{MR}_{\textit{pr3}}$.
-This completes the proof of this lemma.
-\qed
-\end {proof}
-
-\subsection{Co-Regular of $R_{\textit{pr4}}$}
-\label{subsec:co-regular of Rpr4}
-
-\begin{restatable}{lemma}{Rp4AsHappenBefore}
-\label{lemma:Rpr4 as happen before}
-Given a data-differentiated history $h$.
-There exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr4}}$, and $h'$ does not linearizable to $\textit{MR}_{\textit{pr4}}$, if and only if there exists $h'' \in \textit{proj}(h)$, $\textit{last}(h'') = R_{\textit{pr4}}$, $x$ and $y$ has maximal priority in $h''$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,\_) <_{\textit{hb}} \textit{put}(y,\_)$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{if}$ direction, let $h_1 = h \vert_{ \{ x,y \} }$. It is easy to see that $\textit{last}(h') = R_{\textit{pr4}}$ and $h_1$ does not linearizable to $\textit{MR}_{\textit{pr4}}$.
-
-To prove the $\textit{only if}$ direction, we prove its contrapositive.
-Assume that for each $h'' \in \textit{proj}(h)$, such that $\textit{last}(h'') = R_{\textit{pr4}}$, for each $x$ and $y$ has maximal priority in $h''$, if $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, then $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$.
-We need to prove that, for each $h' \in \textit{proj}(h)$, if $\textit{last}(h') = R_{\textit{pr4}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr4}}$.
-
-Given such $h'$, let $x_1,\ldots,x_m$ be the set of items with maximal priority and has unmatched $\textit{put}$ operations, let $y_1,\ldots,y_n$ be the set of items with maximal priority and has matched $\textit{put}$ and $\textit{rm}$ operations.
-It is easy to see that, for each $x_i$ and $y_j$, $\textit{cal}(\textit{put},y_j)$ is before $\textit{ret}(\textit{put},x_i,\_)$.
-Then we locate the linearization points of $h'$ as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] For each $x_i$, locate the linearization point of $\textit{put}(x_i,\_)$ just before its return action.
-
-\item[-] For each $y_j$, locate the lineariztion point of $\textit{put}(y_j,\_)$ jest after its call action.
-
-\item[-] For other operations, locate their linearization points at an arbitrary location after its call action and before its return action.
-\end{itemize}
-
-Let $l$ be the sequence of linearization points.
-It is easy to see that $h \sqsubseteq l$, and $l \in \textit{MR}_{\textit{pr4}}$.
-Here the $\textit{itm}$ in $R_{\textit{pr4}}$ is chosen to be $\textit{ret}(\textit{put},x_{\textit{ind}},\_)$, where the return action of $x_{\textit{ind}}$ is later than return actions of other $x_i$.
-This completes the proof of this lemma.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr4IsAlwaysCoRegular}
-\label{lemma:Rpr4 is always co-regular}
-Given a data-differentiated history $h$, if $\textit{last}(h) = R_{\textit{pr4}}$, then $h \sqsubseteq \textit{MR}_{\textit{pr4}}$.
-\end{restatable}
-
-\begin {proof}
-
-According to Lemma \ref{lemma:Rpr4 as happen before}, if $\textit{last}(h) = R_{\textit{pr4}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr4}}$, then there exists $x$ and $y$ with maximal priority in $h$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,\_) <_{\textit{hb}} \textit{put}(y,\_)$.
-Let $h_1 = h \vert_{ \{ x,y \} }$.
-It is obvious that $\textit{transToQueue}(h_1)$ does not linearizable to queue.
-This contradicts the assumption that every single-priority history can be ``linearizable to queue'', and thus, we can safely ignore this case.
-\qed
-\end {proof}
-
-\subsection{Co-Regular of $R_{\textit{pr5}}$}
-\label{subsec:co-regular of Rpr5}
-
-$R_{\textit{pr5}} \equiv (u \cdot v \in \textit{PQueue}) \wedge (\textit{matched}(u,\textit{put},\textit{rm}) ) \Rightarrow (u \cdot \textit{rmEmpty} \cdot v \in \textit{PQueue})$
-
-Similar as the notion of gap in \cite{Bouajjani:2015}, we propose the notion of gap of $\textit{em}(\textit{empty})$ operations.
-
-\begin{definition}[gap for $\textit{em}(\textit{empty})$ operations]\label{def:gap for rmEmpty operations}
-
-Let $h$ be a data-differentiated history and $o = \textit{rm}(\textit{empty})$ of $h$.
-We say that $h$ has a gap on operation $o$, if there is a partition of the operations of $h$ into $L$ and $R$ satisfying:
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $L$ has matched $\textit{put}$ and $\textit{rm}$ operations.
-
-\item[-] no operation of $R$ happens before an operation of $L$ or $o$.
-
-\item[-] $o$ does not happen before operations of $L$
-\end{itemize}
-\end{definition}
-
-Compared to the gap on $\textit{put}$ and $\textit{rm}$ operations, here we do not consider the priority of operations in $L$ or $R$.
-
-\begin{restatable}{lemma}{GapEqualsLinforRpr5}
-\label{lemma:Gap Equals Lin for Rpr5}
-A data-differentiated history $h$ has a projection $h'$, such that $\textit{last}(h') = R_{\textit{pr5}}$ and $h'$ does not linearizable to $\textit{MR}_{\textit{pr5}}$, if and only if there exists a projection $h''$ of $h$, such that $\textit{last}(h'') = R_{\textit{pr5}}$, there exists $o = \textit{rm}(\textit{empty})$ in $h''$, and there is no gap of $o$ in $h''$.
-\end{restatable}
-
-\begin {proof}
-
-Let us prove the $\textit{only if}$ direction by contradiction.
-Assume that there exists such $h'$, and for each $\textit{rm}(\textit{empty})$ in $h'$, there is a gap of $o$ in $h''$.
-Since $\textit{last}(h') = R_{\textit{pr5}}$, there is exists at least $\textit{rm}(\textit{empty})$ operation in $h$, and let $o$ be arbitrary one of them.
-Since there is a gap of $o$ in $h$, let $L$ and $R$ be the sets of the gap.
-Let $u$ be a sequence of operations of $L$, and $u$ is generated from $h$ by locating linearization points of each operation in $L$ just after its call action.
-Let $v$ be a sequence of operations of $R$, and $v$ is generated from $h$ by locating linearization points of each operation in $R$ just before its return action.
-It is easy to see that $h \sqsubseteq l = u \cdot o \cdot v$, and $l \in \textit{MR}_{\textit{pr5}}$.
-This contradicts that $h'$ does not linearizable to $\textit{MR}_{\textit{pr5}}$.
-
-To prove the $\textit{if}$ direction, let $h_1$ be a history generate from $h''$ by discarding all other $\textit{rm}(\textit{Empty})$ operations.
-By assumption, it is obvious that there is no gap of $o$ in $h_1$.
-Assume by contradiction that there exists $l'$, such that $h_1 \sqsubseteq l' \in \textit{MR}_{\textit{pr5}}$.
-It is easy to see that $l' = u' \cdot o \cdot v'$, such that $u'$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
-Let $L$ be the set of operations of $u'$, and $R$ be the set of operations in $v'$.
-Since $h_1 \sqsubseteq l'$, it is easy to see that $L$ and $R$ is a gap of $o$ in $h_1$, contradicts that there is no gap of $o$ in $h_1$.
-\qed
-\end {proof}
-
-Similar as the notion of gap in \cite{Bouajjani:2015}, we propose the notion left-right constraint of $\textit{rm}(\textit{empty})$ operation.
-
-\begin{definition}[left-right constraint for $\textit{rm}(\textit{empty})$ operation]\label{def:left-right constraint for rmEmpty operation}
-Given a data-differentiated history $h$, and $o = \textit{rm}(\textit{empty})$ of $h$. The left-right constraint of $o$ is the graph $G$ where:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] the nodes are the items of $h$ or $o$, to which we add a node,
-
-\item[-] there is an edge from item $d_1$ to $o$, if $\textit{put}(d_1,\_)$ happens before $o$,
-
-\item[-] there is an edge from $o$ to item $d_1$, if $o$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$,
-
-\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$.
-\end{itemize}
-\end{definition}
-
-Compared to the left-right constraint of $\textit{put}$ and $\textit{rm}$ operations, here we do not consider the priority.
-We also do not consider other $\textit{rm}(\textit{empty})$ operations in $h$.
-
-Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$.
-Let $\textit{LMSet}_1(h,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some item, either $\textit{op}$ happens before $o$ in $h$, or there is an operation $\textit{op}'$ with the same item of $\textit{op}$, such that $\textit{op}'$ happens before $o$ in $h$ $\}$.
-For each $i \geq 1$, let $\textit{LMSet}_{\textit{i+1}}(h,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some item, $\textit{op}$ is not an operation of items in $\textit{LMSet}_k(h,o)$ for each $k \leq i$, either $\textit{op}$ happens before some operation $o' \in \textit{LMSet}_i(h,o)$ in $h$, or there is an operation $\textit{op}''$ with the same item of $o$ and $\textit{op}''$ happens before some operation $o' \in \textit{LMSet}_i(h,o)$ in $h$ $\}$.
-We can see that $\textit{LMSet}_i(h,o) \cap \textit{LMSet}_j(h,o) = \emptyset$ for any $i \neq j$.
-Let $\textit{LMSet}(h,o) = \textit{LMSet}_1(h,o) \cup \textit{LMSet}_2(h,o) \cup \ldots$.
-
-\begin{restatable}{lemma}{RmEmptyDoesNotHappenBeforeLMSetForRpr5}
-\label{lemma:RmEmpty does not happen before LMSet for Rpr5}
-Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$.
-Let $G$ be the graph representing the left-right constraint of $o$.
-Assume that $G$ has no cycle going through $o$.
-Then, $o$ does not happen before any operation in $\textit{LMSet}(h,o)$.
-\end{restatable}
-
-This Lemma can be similarly proved as Lemma \ref{lemma:Rmx does not happen before LMSet for Rpr1}.
-
-\begin{restatable}{lemma}{LMSetHasMatchedPutandRmOperationsForRpr5}
-\label{lemma:LMSet has matched put and rm operations for Rpr5}
-Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$.
-Let $G$ be the graph representing the left-right constraint of $o$.
-Assume that $G$ has no cycle going through $o$.
-Then, $\textit{LMSet}(h,o)$ contains only matched $\textit{put}$ and $\textit{rm}$ operations.
-\end{restatable}
-
-This Lemma can be similarly proved as Lemma \ref{lemma:LMSet has matched put and rm operations}.
-
-\begin{restatable}{lemma}{GapEqualsConstraintforRpr5}
-\label{lemma:Gap Equals Constraint for Rpr5}
-Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$.
-Let $G$ be the graph representing the left-right constraint of $o$.
-There is a gap on $o$, if and only if $G$ has no cycle going through $o$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{only if}$ direction, assume that there is a gap on $o$, and let it be $L$ and $R$.
-Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow o \rightarrow d_1$ in $G$.
-Since $o \rightarrow d_1$, either $o$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$.
-Since $L$ contains matched $\textit{put}$ and $\textit{rm}$ operations, it is easy to see that operations of $d_1$ is in $R$.
-Since $d_1 \rightarrow d_2$, we can see that $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$.
-Since operations of $d_1$ is in $R$ and $L$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we can see that operations of $d_2$ is in $R$.
-Similarly, we can prove that operations of $d_3,\ldots,d_m$ is in $R$.
-Since $d_m \rightarrow o$, $\textit{put}(d_m,\_)$ happens before $o$, and $\textit{put}(d_m,\_) \in L$.
-This contradicts that operations of $d_m$ is in $R$.
-This completes the proof of the $\textit{only if}$ direction.
-
-To prove the $\textit{if}$ direction, assume that $G$ has no cycle going through $o$. Let $O_h$ be the set of operations of $h$. Let $O_L = \textit{LMSet}(h,o)$, $O_R = O_h \setminus O_L$.
-
-By Lemma \ref{lemma:LMSet has matched put and rm operations for Rpr5}, we can see that $O_L = \textit{LMSet}(h,o)$ contains only matched $\textit{put}$ and $\textit{rm}$ operations.
-It remains to prove that $L$ and $R$ is a gap of $o$.
-We prove this by showing that all the following cases are impossible:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Case $1$: Some operation $o_r \in R$ happens before $o$. Then we know that $o_r \in \textit{LMSet}(h,o)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $2$: Some operation $o_r \in R$ happens before some operation $o_l$. Then we know that $o_r \in \textit{LMSet}(h,o)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $4$: $o$ happens before some $o_l \in \textit{LMSet}(h,o) = L$. By Lemma \ref{lemma:RmEmpty does not happen before LMSet for Rpr5}, we know that this is impossible.
-\end{itemize}
-
-This completes the proof of the $\textit{if}$ direction.
-
-\qed
-\end {proof}
-
-Let us begin to represent an automaton that is used for capture the case that, in a sub-history $h'$ of a history $h$, $\textit{last}(h')=R_{\textit{pr5}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr5}}$, and the reason is that there is no gap on a $\textit{rm}(\textit{empty})$ operation in $h'$.
-The automaton is $\mathcal{A}_{\textit{Rpr5}}$, which is given in \figurename~\ref{fig:automata for Rpr5}.
-In \figurename~\ref{fig:automata for Rpr5}, let $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d), \textit{ret}(\textit{rm},d)$, $c_1 = c + \textit{cal}(\textit{put},b,\textit{anPri})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c + \textit{ret}(\textit{rm},b)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr5.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr5}}$}
-  \label{fig:automata for Rpr5}
-\end{figure}
-
-Given a data-differentiated history $h$, we say that $o = \textit{rm}(\textit{empty})$ in $h$ is covered by items $d_1,\ldots,d_m$ in $h$, if
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{put}(d_m,\_)$ happens before $o$,
-
-\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,
-
-\item[-] $o$ happens before $\textit{rm}(d_1)$, or $\textit{rm}(d_1)$ does not exists in $h$
-\end{itemize}
-
-According to the definition of left-right constraint for $o$, in a data-differentiated history $h$, there is a cycle going through $o$, if and only if there exists items $d_1,\ldots,d_m$, such that $o$ is covered by $d_1,\ldots,d_m$.
-
-\begin{restatable}{lemma}{Rpr5IsCoRegular}
-\label{lemma:Rpr1 5s co-regular}
-$R_{\textit{pr5}}$ is co-regular.
-\end{restatable}
-
-\begin {proof}
-
-We need to prove that:
-
-\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\mathcal{A}_{\textit{Rpr5}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr5}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr5}}$.
-
-By Lemma \ref{lemma:Gap Equals Lin for Rpr5} and Lemma \ref{lemma:Gap Equals Constraint for Rpr5}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:
-
-\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\mathcal{A}_{\textit{Rpr5}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h''$ is a projection of $h$, such that $\textit{last}(h'') = R_{\textit{pr5}}$, $o = \textit{rm}(\textit{empty})$ is in $h''$, and $o$ is covered by some items $d_1,\ldots,d_m$ in $h''$.
-
-\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\mathcal{A}_{\textit{Rpr5}}$.
-By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$.
-Let $d_1,\ldots,d_m$ be the items in $h_2$ such that $r(d_i)=d$ for each $1 \leq i \leq m$.
-Let $h'' = h_2 \vert_{ \{ o, d_0, d_1, \ldots, d_m \} }$.
-It is obvious that $\textit{last}(h'') = R_{\textit{pr5}}$.
-It is easy to see that $o$ is covered by $d_1,\ldots,d_m$.
-
-\noindent The $\textit{if}$ direction: Assume that there exists such $h$, $h''$, $o$ and $d_1,\ldots,d_m$.
-Then, let $h_1$ be obtained from $h$ by renaming $d_1,\ldots,d_m$ into $b$ and renaming other items into $d$.
-We can prove that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr5}}$.
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\section{Relate Queue and Stack with Priority Queue}
-\label{sec:relate queue and stack with priority queue}
-
-\subsection{Relate Queue with Priority Queue}
-\label{subsec:relate queue with priority queue}
-
-Given a history $h_q$ of queue operations, let $\textit{UniPriTrans}(h_q)$ be a history of priority queue, which is generated as follows: transform each $\textit{call}(\textit{enq},a)$, $\textit{ret}(\textit{enq},a)$, $\textit{call}(\textit{deq},a)$ and $\textit{ret}(\textit{deq},a)$ of $h_q$ into $\textit{call}(\textit{put},a,1)$, $\textit{ret}(\textit{put},a)$, $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$.
-
-\begin{restatable}{lemma}{RelateQueuewithPQ}
-\label{lemma:relate queue with priority queue}
-Given a history $h_q$ of queue operations, $h_q$ is linearizable to queue, if and only if $\textit{UniPriTrans}(h_q) \sqsubseteq \textit{PQueue}$.
-\end{restatable}
-
-\begin {proof}
-
-This is obvious, since we already ensure that for each single-priority history $h$, $\textit{transToQueue}(h)$ is linearizable to queue. \qed
-\end {proof}
-
-\subsection{Relate Stack with Priority Queue}
-\label{subsec:relate stack with priority queue}
-
-\section{Conclusion and Future Work}
-\label{sec:conclusion and future work}
-
-Conclusion
-
-%\FloatBarrier
-
-%\bibliography{short}
-%\bibliographystyle{plain}
-%\bibliographystyle{splncs}
-%\bibliography{biblio_cat.bib}
-
-\begin{thebibliography}{50}
-
-\bibitem{Bouajjani:2015}
-Bouajjani, A., Emmi, M., Enea, C., Hamza, J.:
-\newblock On reducing linearizability to state reachability.
-\newblock In: Halld{\'{o}}rsson, M.M. et al. (eds.) ICALP 2015, Part II, pp. 95--107. Springer (2015)
-
-\bibitem{Wolper:1986}
-Wolper, P.:
-\newblock Expressing interesting properties of programs in propositional
-  temporal logic.
-\newblock In {\em POPL'86}. pp.184--193. (1986)
-
-\end{thebibliography}
-\newpage
-
-\appendix
-
-\end{document}

### 2017-02-20 appendix.tex
preliminary,data-independence,rules

\section{Proof in Section \ref{sec:data-independence of priority queue}}
\label{sec:appendix in section data-independence of PQ}

\DataDifferentiatedisEnoughforPQ

\begin {proof}

To prove the $\textit{only if}$ direction, given a data-differentiated execution $e \in \mathcal{I}_{\neq}$. By assumption, it is linearizable with respect to a sequential execution $l \in S$, and the bijection between the operations of $e$ and the method events of $l$, ensures that $l$ is differentiated and belongs to $S_{\neq}$.

To prove the $\textit{if}$ direction, given an execution $e \in \mathcal{I}$. By data independence of $\mathcal{I}$, we know that there exists $e' \in \mathcal{I}_{\neq}$ and a renaming function $r$, such that $r(e') = e$. By assumption, $e'$ is linearizable with respect to a sequential execution $l' \in S_{\neq}$. Let $l=r(l')$. By data independence of $S$ it is easy to see that $l \in S$, and it is easy to see that $e \sqsubseteq l$  using the same bijection used for $e' \sqsubseteq l'$. \qed
\end {proof}

### 2017-02-20 data-independence-PQ.tex
preliminary,data-independence,rules

\section{Data-Independence of Priority Queue}
\label{sec:data-independence of priority queue}

A priority queue contains two method: $\textit{put}$ and $\textit{rm}$. A $\textit{put}$ method has two arguments, while the first argument is an item and the second argument is priority. A $\textit{put}$ method is used to put an item into the priority queue with certain priority. Here we assume that the item is chosen from a specific (possibly infinite) data domain $\mathbb{D}$ and priority is chosen from the set $\mathbb{N}$ of natural number. A $\textit{rm}$ method intends to remove the item with maximal priority in priority queue and then returns it. If the priority queue is empty, then $\textit{rm}$ returns $\textit{empty}$. If there are multiple items in priority queue with a same maximal priority, then $\textit{rm}$ will return the item that is putted earliest among such items. We say that $\textit{put}(a,p)$ matches $\textit{rm}(b)$, if $a = b$.

Let us slightly modify the notion of differentiated and data-independence in \cite{Wolper:1986} and propose data-differentiated sequence and data-independence for priority queues.

Let $\_$ denote an element, of which the value is irrelevant. A sequential execution $e$ of priority is said to be data-differentiated if, for all $d \in \mathbb{D}$, there is at most one method event $\textit{put}(d,\_)$ in $e$. Note that a data-differentiated sequential execution $e$ may contains more than one items with a same priority. The subset of data-differentiated sequential executions of a set $S$ is denoted by $S_{\neq}$. The definition extends to (sets of) executions and histories. For instance, an execution is data-differentiated if, for all $d \in \mathbb{D}$, there is at most one $\textit{cal}_{\_}(\textit{put},d,\_)$.

\begin{example}\label{example:data-differentiated}
$\textit{cal}_{o_1}(\textit{put},a,7) \cdot \textit{ret}_{o_1}(\textit{put},a) \cdot \textit{cal}_{o_2}(\textit{put},a,8) \cdot \textit{ret}_{o_2}(\textit{put},a)$ is not data-differentiated, since there are two $\textit{put}$ methods with the same item.
\end{example}

A renaming function $r$ for priority queue is a function from $\mathbb{D}$ to $\mathbb{D}$. Given a sequential execution (resp., execution or history) $e$, we denote by $r(e)$ the sequential execution (resp., execution or history) obtained from $e$ by replacing every item $x$ by $r(x)$. Note that here the renaming functions rename only the items and keep the priorities unchanged.

%\vspace{-6pt}
\begin{definition}\label{def:priority-value data-independence}
The set of sequential executions (resp., executions or histories) $S$ is data-independent, if:
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] for all $e \in S$, there exists $e' \in S'$, and a renaming function $r$, such that $e=r(e')$,

\item[-] for all $e \in S$ and for all renaming $r$, $r(e) \in S$.
\end{itemize}
\end{definition}

The following lemma states that, when checking that a data-independent implementation $\mathcal{I}$ is linearizable with respect to a data-independent specification, it is enough to do so for data-differentiated executions, similar as that in \cite{Abdulla:2013}, where the notion of data-independence and differentiated in \cite{Wolper:1986} is used. Thus, in the remainder of the paper, we focus on characterizing linearizability for data-differentiated executions, rather than arbitrary ones. The proof of this lemma can be found in Appendix \ref{sec:appendix in section data-independence of PQ}.

\begin{restatable}{lemma}{DataDifferentiatedisEnoughforPQ}
\label{lemma:data differentiated is enough for PQ}
A data-independent implementation $\mathcal{I}$ is linearizable with respect to a data-independent specification $S$, if and only if $\mathcal{I}_{\neq}$ is linearizable with respect to $S_{\neq}$.
\end{restatable}

### 2017-02-20 introduction.tex
preliminary,data-independence,rules

--- 2017-02-16 paper-LIN-PQ.tex
@@
-%\documentclass[a4paper,UKenglish]{lipics}
-\documentclass{llncs}
-%\usepackage{mathbbold}
-
-%\newtheorem{theorem}{Theorem}[section]
-%\newtheorem{conjecture}[theorem]{Conjecture}
-%\newtheorem{corollary}[theorem]{Corollary}
-%\newtheorem{proposition}[theorem]{Proposition}
-%\newtheorem{lemma}[theorem]{Lemma}
-%\newdef{definition}[theorem]{Definition}
-%\newdef{remark}[theorem]{Remark}
-%\newdef{example}[theorem]{Example}
-
-\usepackage{epsfig}
-\usepackage{amsmath}
-\usepackage{color}
-%\usepackage{txfonts}
-%\usepackage{pxfonts}
-%\usepackage{mathabx}
-%\usepackage{makeidx}  % allows for indexgeneration
-\usepackage{verbatim}
-\usepackage{url}
-%\usepackage{hyperref}
-\usepackage{times}
-\usepackage{ulem}    %(\emph{....} in the texfile) is underlined instead of italic
-\normalem
-\usepackage{enumerate}
-%\usepackage[square, comma, sort&compress, numbers]{natbib}
-
-%added by Yi Lv%
-%\usepackage[linesnumbered,ruled,procnumbered,noend]{algorithm2e}
-%\usepackage[linesnumbered,ruled,procnumbered,vlined]{algorithm2e}
-\usepackage[linesnumbered,ruled,procnumbered,noend]{algorithm2e}
-\usepackage{stmaryrd}
-%\usepackage{hyperref}
-%added by Wang Chao%
-%\usepackage{mathrsfs}
-%\usepackage{extarrows}
-\usepackage{thmtools}
-\usepackage{thm-restate}
-
-%add for TIkz
-\usepackage[version=0.96]{pgf}
-\usepackage{tikz}
-\usetikzlibrary{arrows,shapes,snakes,automata,backgrounds,petri}
-\usepackage[latin1]{inputenc}
-\usepackage{placeins}
-\usepackage[bookmarks=false]{hyperref}
-
-%\declaretheorem[name=Claim]{clm}
-%\declaretheorem[name=Lemma]{lem}
-%\declaretheorem[name=Theorem]{thm}
-
-\DeclareSymbolFont{largesymbolsA}{U}{txexa}{m}{n}
-\SetSymbolFont{largesymbolsA}{bold}{U}{txexa}{bx}{n}
-\DeclareFontSubstitution{U}{txexa}{m}{n}
-\DeclareMathSymbol{\bigsqcupplus}{\mathop}{largesymbolsA}{"02}
-
-% "Import" \bigsqcupplus from txfonts without loading txfonts (since
-% it changes the default font and replaces many symbols)
-
-% Commands by Catuscia
-
-%\DeclareMathOperator*{\minarg}{{\rm minarg}}
-
-%Previous notation for the probabilistic choice
-%\newcommand{\probsum}{\bigoplus}
-%\newcommand{\smallprobsum}[1]{\mkern4mu{\textstyle\circ\mkern-15.5mu\sum_{#1}\:}}
-%\newcommand{\bigprobsum}[1]{{\;\displaystyle\odot\mkern-21mu\sum_{#1}}}
-%\newcommand{\probsum}[1]{\mathchoice{\bigprobsum{#1}}{\smallprobsum{#1}}{\smallprobsum{#1}}{\smallprobsum{#1}}}
-%For the uniform distribution
-%\newcommand{\usmallprobsum}[1]{\mkern4mu{\textstyle\circ\mkern-15.5mu\sum^\mathcal{U}_{#1}\:}}
-%\newcommand{\ubigprobsum}[1]{{\;\displaystyle\odot\mkern-21mu\sum^\mathcal{U}_{#1}}}
-%\newcommand{\uprobsum}[1]{\mathchoice{\ubigprobsum{#1}}{\usmallprobsum{#1}}{\usmallprobsum{#1}}{\usmallprobsum{#1}}}
-
-%\newcommand{\nondsum}{\bigbox}
-\newcommand{\nondsum}{\bigsqcupplus}
-\newcommand{\probplus}[1]{\oplus_{#1}}
-%\newcommand{\nondplus}{\square}
-\newcommand{\bang}{!\,}
-\newcommand{\nondplus}{{\textstyle\bigsqcupplus}}
-\newcommand{\partmap}{\rightharpoonup}
-\newcommand{\map}{\rightarrow}
-%\newcommand{\exc}{\phi}
-\newcommand{\exc}{\alpha}
-\newcommand{\exec}{\mathit{exec}}
-\newcommand{\execp}{\mathit{execp}}
-\newcommand{\Act}{\mathit{Act}}
-\newcommand{\Sec}{\mathit{Sec}}
-\newcommand{\Obs}{\mathit{Obs}}
-\newcommand{\etree}{\mathit{etree}}
-\newcommand{\lstate}{\mathit{lst}}
-\newcommand{\fstate}{\mathit{fst}}
-%\newcommand{\STATE}{\mathcal{P}r}  $ original marked
-%\newcommand{\STATE}{\mathcal{P}}   $ marked by me
-%\newcommand{\st}{P}
-\newcommand{\trans}{\mathcal{T}}
-\newcommand{\Aut}{\mathcal{M}}
-\newcommand{\init}{\mathit{init}}
-\newcommand{\perr}{\mathcal{P}}
-
-\newcommand{\calo}{\mathcal{O}}
-\newcommand{\cals}{\mathcal{S}}
-\newcommand{\sseq}{\vec s}
-\newcommand{\oseq}{\vec o}
-\newcommand{\ccsp}{CCS$_p$}
-
-\newcommand{\bigfrac}[2]{\frac{\raisebox{1ex}{$#1$}}{\raisebox{-1.5ex}{$#2$}}}
-\newcommand{\nondarr}[1]{\overset{#1}{\longrightarrow}}
-\newcommand{\Nondarr}[1]{\overset{#1}{\Longrightarrow}}
-\newcommand{\vectorArrow}[1]{\stackrel{\longrightarrow}{\mbox{#1}}}
-\newcommand{\probarr}[1]{\overset{#1}{\dashrightarrow}}
-\newcommand{\paral}{\,|\,}
-\newcommand{\outp}[1]{\overline{#1}}
-\renewcommand{\Pr}{{\rm Pr}}
-
-%Commands by Chao Wang
-
-%memory models%
-\newcommand{\TSO}{\textrm{TSO}}
-\newcommand{\PSO}{\textrm{PSO}}
-
-%correctness conditions%
-\newcommand{\lin}{\textrm{linearizability}}
-\newcommand{\slin}{\textrm{static linearizability}}
-\newcommand{\qlin}{\textrm{quasi linearizability}}
-\newcommand{\TTlin}{\textrm{TSO-to-TSO linearizability}}
-
-\newcommand{\pair}[2]{\langle #1 , #2 \rangle}% pairs
-\newcommand{\setof}[2]{\{ \, #1 \mid #2 \, \}}% Sets
-\newcommand{\set}[1]{\{ {#1}  \}  }
-%\newcommand{\map}[3]{{#1} \colon {#2} \longmapsto {#3}} %functions
-\newcommand{\den}[1]{[\![#1]\!]}% Denotation of
-\newcommand{\mean}[1]{|\!|#1|\!|}
-\newcommand{\forget}[1]{}
-
-%%%%%%%%%%%%%GENERAL%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\itbox}[1]{{\it #1\/}}
-\newcommand{\un}[1]{\uline{#1}}%\underline{#1}}
-\newcommand{\ov}[1]{\overline{#1}}
-\newcommand{\smallspace}{\vspace{10mm}}
-\newcommand{\is}{\mbox{$\Longleftarrow\ $}}
-\newcommand{\pright}[1]{\hfill{#1}}
-\newcommand{\bnfor}{\;\;\mid\;\;}
-
-\newcommand{\ar}[1]{\stackrel{\scriptstyle #1}{\longrightarrow}}
-
-%%%%%%%%%italics in math mode
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\newcommand{\true}{{\it true}}
-%\newcommand{\false}{{\it false}}
-\newcommand{\calB}{{\cal B}}
-\newcommand{\calF}{{\cal F}}
-\newcommand{\calP}{{\cal P}}
-\newcommand{\order}{{\cal O}}
-\newcommand{\size}[1]{|#1|}
-
-%other notations%
-\newcommand{\LTS}{\textit{LTS}}
-\newcommand{\bedt}[1]{{\color{blue}#1}}
-\newcommand{\redt}[1]{{\color{red}#1}}
-
-%\pagestyle{empty}
-%\pagestyle{plain}
-
-%\def\lastname{Xu,Palamidessi}
-
-\title{Reducing Linearizability of Priority Queues and More Data Structures to State Reachability}
-
-\author {Chao Wang \and Ahmed Bouajjani \and Constantin Enea}
-\institute{Institut de Recherche en Informatique Fondamentale, \\Univ. Paris Diderot (Paris 7)
-}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-Abstract
-\end{abstract}
-
-\forget{
-\noindent Keywords: weak memory model, $\textit{linearizability}$,
-$\textit{TSO-to-TSO linearizability}$
-}
-
 \section{Introduction}
 \label{sec:introduction}
-Introduction
+I am introduction
 
@@
 \noindent Related work
-
-\section{Inductive Rules for Priority Queues}
-\label{sec:inductive rules for priority queues}
-
-given a sequence $l$ of priority queue operations, we say that $l$ is data-differentiated, if for each item $d \in \mathcal{D}$, $\textit{put}(d,\_)$ appears at most once in $l$.
-Note that a data-differentiated sequence $l$ may contains more than one items with a same priority.
-The subset of data-differentiated operation sequences of a set $S$ is denoted by $S_{\neq}$.
-For priority queue, the renaming functions rename only the items and keep the priorities unchanged.
-For priority queue, we modify the notion of data-independence \cite{Wolper:1986} and use data-differentiated instead of differentiated:
-
-%\vspace{-6pt}
-\begin{definition}[data-independence of priority queue]\label{def:priority-value data-independence}
-The set $S$ of sequences of priority queue operations (resp., histories)  is data independent, if:
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] for all $l \in S$, there exists $l' \in S_{\neq}$, and a renaming $r$, such that $l=r(l')$,
-
-\item[-] for all $l \in S$ and for all renaming $r$, $r(l) \in S$.
-\end{itemize}
-\end{definition}
-
-Similar as in \cite{Bouajjani:2015}, we consider only completed executions, where each call action has a corresponding return action.
-This simplification is sound when implementation methods can always make progress in isolation.
-
-For the inductive rules of priority queue, our rules need to concern the cost of operations. Formally, the inductive rules for priority queues is as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $R_{\textit{pr0}} \equiv \epsilon \in \textit{PQueue}$
-
-\item[-] $R_{\textit{pr1}} \equiv (u \cdot v \cdot w \in \textit{PQueue}) \wedge (u \cdot v \cdot w \in (\textit{put},\textit{rm})^*) \wedge (\textit{pri} >$ priorities of $u \cdot v \cdot w ) \wedge (\textit{matched}(u \cdot v, \textit{put}, \textit{rm}) ) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w \in \textit{PQueue})$
-
-\item[-] $R_{\textit{pr2}} \equiv (u \cdot v \cdot w \in \textit{PQueue}) \wedge (u \cdot v \cdot w \in (\textit{put},\textit{rm})^*) \wedge (\textit{pri} =$ the maximal priority of $u \cdot v \cdot w ) \wedge (\textit{pri} >$ priorities of unmatched items in $u \cdot v \cdot w) \wedge (\textit{matched}(u \cdot v, \textit{put}, \textit{rm}) ) \wedge ($all $\textit{put}$ operation with priority $\textit{pri}$ is in $u) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w \in \textit{PQueue})$
-
-\item[-] $R_{\textit{pr3}} \equiv (u \cdot v \in \textit{PQueue}) \wedge ( u \cdot v \in (\textit{put},\textit{rm})^*) \wedge (\textit{pri} =$ the maximal priority of items in $u \cdot v) \wedge (\textit{pri} >$ priorities of matched items in $u \cdot v \cdot w ) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{PQueue})$
-
-\item[-] $R_{\textit{pr4}} \equiv (u \cdot v \in \textit{PQueue}) \wedge ( u \cdot v \in (\textit{put},\textit{rm})^*) \wedge (\textit{pri} =$ the maximal priority of items in $u \cdot v) \wedge (\textit{pri} =$ the maximal priority of matched items in $u \cdot v \cdot w ) \wedge ($all matched $\textit{put}$ with priority $\textit{pri}$ is in $u) \wedge ($all unmatched $\textit{put}$ with priority $\textit{pri}$ is in $u) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{PQueue})$
-
-\item[-] $R_{\textit{pr5}} \equiv (u \cdot v \in \textit{PQueue}) \wedge (\textit{matched}(u,\textit{put},\textit{rm}) ) \Rightarrow (u \cdot \textit{rmEmpty} \cdot v \in \textit{PQueue})$
-\end{itemize}
-
-Given a sequence $l$ of possibly $\textit{put}$, $\textit{rm}$ and $\textit{rmEmpty}$ operations, we say that $l$ contains unmatched $\textit{rm}$ operations, if the number of $\textit{rm}(d)$ operations is larger than the number of $\textit{put}(d,\_)$ operations for some $d$ in $l$.
-By data-independence, we can detect the existence of unmatched $\textit{rm}$ operations by automata.
-A automaton $\mathcal{A}_{\textit{unmRm1}}$ is given in \figurename~\ref{fig:automata for unmatched rm1}, where $c_1=\textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm),\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$, $c_2=c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c_1 + \textit{cal}(\textit{put},b,\textit{anyPri}) + \textit{ret}(\textit{rm},b)$.
-Here $\textit{anyPri}$ is a specific predicates that is satisfied by any priority.
-$\mathcal{A}_{\textit{unmRm1}}$ is used to detect the existence of a item that has been removed but has never been inserted.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_RM1.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{unmRm1}}$}
-  \label{fig:automata for unmatched rm1}
-\end{figure}
-
-An automaton $\mathcal{A}_{\textit{unmRm2}}$ is given in \figurename~\ref{fig:automata for unmatched rm2}, where $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm),$ $\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$, and $c_1 = c_1 + \textit{ret}(\textit{rm},b)$.
-$\mathcal{A}_{\textit{unmRm1}}$ is used to detect the existence of a item that has been removed more than once but has been inserted once.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_RM2.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{unmRm2}}$}
-  \label{fig:automata for unmatched rm2}
-\end{figure}
-
-The following lemma states that we can detect the existence of unmatched $\textit{rm}$ by $\mathcal{A}_{\textit{unmRm1}}$ and $\mathcal{A}_{\textit{unmRm2}}$.
-
-\begin{restatable}{lemma}{DetectUnmatchedRmByAutomata}
-\label{lemma:detect unmatched rm by automata}
-
-Given a data-independent $\mathcal{I}$, there is some unmatched $\textit{rm}$ in some sequence of $\mathcal{I}$, if and only if $( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm1}} \neq \emptyset) \vee ( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm2}} \neq \emptyset )$.
-\end{restatable}
-
-\begin {proof}
-
-For the $\textit{if}$ direction, assume that $( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm1}} \neq \emptyset) \vee ( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm2}} \neq \emptyset )$.
-Then, there exists a sequence $l \in \mathcal{I}$ and $l$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$ or $\mathcal{A}_{\textit{unmRm2}}$.
-If $l$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$, then $l$ contains at least one $\textit{rm}(b)$ operation and no $\textit{put}(b,\_)$ operation; Else, if $l$ is accepted by $\mathcal{A}_{\textit{unmRm2}}$, then $l$ contains at least two $\textit{rm}(b)$ operations and one $\textit{put}(b,\_)$ operation.
-In both cases, $l$ contains unmatched $\textit{rm}$ operations.
-
-To prove the $\textit{only if}$ direction assume that there exists a sequence $l \in \mathcal{I}$ that contains unmatched $\textit{rm}$ operations.
-By data-independence, there exists data-differentiated sequence $l'$ and renaming function $r$, such that $l=r(l')$.
-We need to prove that in $l'$, one of the following cases holds: (1) $l'$ contains $\textit{rm}(d_1)$ for some $d_1$ and does not contain any $\textit{put}(d_1,\_)$, and (2) $l'$ contains one $\textit{put}(d_1,\_)$ operation and more than one $\textit{rm}(d_1)$ operations.
-We prove this by contradiction.
-Assume by contradiction that both cases do not hold.
-Then, for each value $d$ of $l'$, there is either one $\textit{put}(d,\_)$ and one $\textit{rm}(d)$ in $l'$, or one $\textit{put}(d,\_)$ and no $\textit{rm}(d)$ in $l'$.
-It is obvious that such $l'$ does not contain unmatched $\textit{rm}$ operation, and renaming a sequence that does not contain unmatched $\textit{rm}$ we can only obtain another sequence that does not contain unmatched $\textit{rm}$.
-This contradicts that $l=r(l')$ has unmatched $\textit{rm}$.
-
-Let $l''$ be obtained from $ll$ by renaming $d_1$ to $b$ and renaming other value to $a$.
-If the first case holds, $l''$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$; Otherwise, the second case holds, and $l''$ is accepted by $\mathcal{A}_{\textit{unmRm2}}$.
-\qed
-\end {proof}
-
-Therefore, from now on, we can safely assume that there are no unmatched $\textit{rm}$ operations. The notion of $\textit{last}$ is defined as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $l$ contains $\textit{rmEmpty}$ operations, then $\textit{last}(l) = R_{\textit{pr5}}$.
-\item[-] Else, $l$ does not contain any $\textit{rmEmpty}$ operation.
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If the maximal priority of $l$ is an unmatched $\textit{put}$ operation $o$, and the priority of $o$ equals the priority of some matched $\textit{put}$ operations in $l$, then $\textit{last}(l) = R_{\textit{pr4}}$.
-
-    \item[-] If the maximal priority of $l$ is an unmatched $\textit{put}$ operation $o$, and the priority of $o$ is larger than the priority of all matched $\textit{put}$ operations in $l$, then $\textit{last}(l) = R_{\textit{pr3}}$.
-
-    \item[-] If the maximal priority of $l$ is a matched $\textit{put}$ operation $o$, the priority of $o$ equals the priority of some matched $\textit{put}$ operations in $l$, and the priority of $o$ is larger than the priority of all unmatched $\textit{put}$ operations in $l$, then $\textit{last}(l) = R_{\textit{pr2}}$.
-
-    \item[-] If the maximal priority of $l$ is an matched $\textit{put}$ operation $o$, and the priority of $o$ is larger than the priority of all $\textit{put}$ operations in $l$, then $\textit{last}(l) = R_{\textit{pr1}}$.
-
-    \item[-] Otherwise, $l = \epsilon$, and $\textit{last}(l) = R_{\textit{pr0}}$.
-    \end{itemize}
-\end{itemize}
-
-%The following lemma states that the inductive rules for priority queues are sound and complete.
-
-\section{Step-by-Step Linearizability of Priority Queues}
-\label{sec:step-by-step linearizability of priority queues}
-
-The following lemma states that $\textit{PQueue}$ is closed under projection. This lemma holds since the predicates appears in inductive rules of $\textit{PQueue}$ are closed under projection.
-
-\begin{restatable}{lemma}{PQisClosedUnderProjection}
-\label{lemma:PQ is closed under projection}
-
-$\textit{PQueue}$ is closed under projection.
-
-\end{restatable}
-
-%When the context is clear, we shorten priority-value data-independence as data-independence.
-For priority queue, the notion of step-by-step linearizability is modified as follows:
-
-%\vspace{-6pt}
-\begin{definition}[step-by-step linearizability for priority queue]\label{def:step-by-step linearizability of priority queue}
-A data structure $S=R_1, \ldots, R_n$ is said be to step-by-step linearizable, if for any data-differentiated sequence $e$, if $e$ is linearizable w.r.t.
-$\textit{MR}_{\textit{pri}}$ ($1 \leq i \leq 4$) with witness $x$, we have:
-
-$e \setminus x \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket \Rightarrow e \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$
-
-\noindent if $e$ is linearizable w.r.t. $\textit{MR}_{\textit{pr5}}$ with witness $x$, we have:
-
-$e \setminus x \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket \Rightarrow e \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$
-\end{definition}
-
-Let $\textit{proj}(s)$ denote the projection of $s$ obtained by erasing all the operations of several items from $s$.
-Here we assume that each $\textit{rm}(\textit{empty})$ is differentiated from each other, and we can consider each $\textit{rm}(\textit{empty})$ as having a unique item.
-
-\begin{restatable}{lemma}{PQasMultiInMRpriforSequence}
-\label{lemma:PQ as multi in MRpri for sequence}
-$s \in \textit{PQueue}$, if and only if, $\forall s'$, if $s' \in \textit{proj}(s)$, then $s' \in \textit{MR}_{\textit{pri}}$ for some $0 \leq i \leq 5$.
-\end{restatable}
-
-\begin {proof}
-
-The \textit{only if} direction is a direct consequence of Lemma \ref{lemma:PQ is closed under projection}.
-
-To prove the $\textit{if}$ direction.
-The $\textit{if}$ direction holds trivially when $s = \epsilon$.
-When $s \neq \epsilon$.
-since $s \in \textit{proj}(s)$, we generate sequence $s_1$ from $s$ as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $s$ has $\textit{rm}(\textit{empty})$ operations: $s_1$ is generated from $s$ by erasing some $\textit{rm}(\textit{empty})$ operation.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in $S_1$ and unmatched $\textit{put}$ operations of some items in $S_2$: $s_1$ is generated from $s$ by erasing the unmatched $\textit{put}$ operation of items last putted in $S_2$.
-
-\item[-] If the maximal priority in $s$ is unmatched $\textit{put}$ operations of some items in $S_2$ and $\vert S_2 \vert =1$: $s_1$ is generated from $s$ by erasing this unmatched $\textit{put}$ operation.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert \geq 2$: $s_1$ is generated from $s$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item which is last putted in $S$.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert = 1$: $s_1$ is generated from $s$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item in $S$.
-\end{itemize}
-
-Similarly, we obtain $s_2$ from $s_1$, obtain $s_3$ from $s_2$, $\ldots$, until we obtain some $s_m = \epsilon$.
-It is obvious that $s_m \in \textit{PQueue}$, and by the five rules of $\textit{PQueue}$, it is easy to see that $s_{\textit{m-1}},\ldots,s_1,s \in \textit{PQueue}$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{PQasMultiInMRpriforHistory}
-\label{lemma:PQ as multi in MRpri for history}
-Given a data-differentiated history $h$, assume that $\textit{PQueue}$ is step-by-step linearizability.
-$h \sqsubseteq \textit{PQueue}$, if and only if, $\forall h'$, if $h' \in \textit{proj}(h)$, then $h' \sqsubseteq \textit{MR}_{\textit{pri}}$, where $i = \textit{last}(h')$.
-\end{restatable}
-
-\begin {proof}
-
-The \textit{only if} direction is a direct consequence of Lemma \ref{lemma:PQ is closed under projection}.
-
-To prove the $\textit{if}$ direction.
-Since $h \sqsubseteq \textit{PQueue}$, we can see that $h \sqsubseteq l \in \textit{PQueue}$.
-Each projection $h'$ of $h$ can be linearizable to some projection $l'$ of $l$.
-By Lemma \ref{lemma:PQ as multi in MRpri for sequence}, $l' \in \textit{MR}_{\textit{prj}}$ for some $0 \leq j \leq 5$.
-Since $\textit{last}$ of a sequence is uniquely determined, we can see that $l' \in \textit{MR}_{\textit{pri}}$, where $i = \textit{last}(h')$.
-
-The $\textit{if}$ direction holds trivially when $h = \epsilon$. When $h \neq \epsilon$. since $h \in \textit{proj}(h)$, we generate sequence $h_1$ from $h$ as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $h$ has $\textit{rm}(\textit{empty})$ operations: Then $\textit{last}(h) = R_{\textit{pr5}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr5}}$.
-$h_1$ is generated from $h$ by erasing some $\textit{rm}(\textit{empty})$ operation.
-
-\item[-] If the maximal priority in $h$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in $S_1$ and unmatched $\textit{put}$ operations of some items in $S_2$: Then $\textit{last}(h) = R_{\textit{pr4}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr4}}$.
-$h_1$ is generated from $h$ by erasing a unmatched $\textit{put}(x,\_)$, such that $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$ for each $y \in S_2$.
-
-\item[-] If the maximal priority in $s$ is unmatched $\textit{put}$ operations of some items in $S_2$ and $\vert S_2 \vert =1$: Then $\textit{last}(h) = R_{\textit{pr3}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr3}}$.
-$h_1$ is generated from $h$ by erasing this unmatched $\textit{put}$ operation.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert \geq 2$: Then $\textit{last}(h) = R_{\textit{pr2}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr2}}$.
-$h_1$ is generated from $h$ by erasing $\textit{put}(x,\_)$ and $\textit{rm}(x)$, such that for each $y \in S$, $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$, and $\textit{rm}(x)$ does not happen before $\textit{rm}(y)$.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert = 1$: Then $\textit{last}(h) = R_{\textit{pr1}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr1}}$.
-$h_1$ is generated from $h$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item in $S$.
-\end{itemize}
-
-Similarly, we obtain $h_2$ from $h_1$, obtain $h_3$ from $h_2$, $\ldots$, until we obtain some $h_m = \epsilon$.
-Assume that for each $i$, when generating $h_i$ from $h_{\textit{i-1}}$, we erase operations of item $d_i$.
-It is obvious that $h_m = h_{\textit{m-1}} \setminus d_m \sqsubseteq \textit{PQueue}$, and since $h_{\textit{m-1}} \sqsubseteq \textit{MR}_{\textit{prj}_{\textit{m-1}}}$, where $\textit{last}(h_{\textit{m-1}}) = R_{\textit{prj}_{\textit{m-1}}}$, we can see that $h_{\textit{m-1}} \sqsubseteq \textit{PQueue}$.
-Similarly, we can prove that $h_{\textit{m-1}},\ldots,h_1,h \sqsubseteq \textit{PQueue}$.
-This completes the proof of the $\textit{if}$ direction.
-\qed
-\end {proof}
-
-To prove that $\textit{PQueue}$ is step-by-step linearizability, we use five lemmas for five rules of $\textit{PQueue}$.
-
-\begin{restatable}{lemma}{Rpr1isStepByStepLinearizability}
-\label{lemma:Rpr1 is step-by-step linearizability}
-
-$R_{\textit{pr1}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr1}}$ with witness $x$.
-Assume that  Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Then, let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-According to $R_{\textit{pr1}}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where the $\textit{put}$ operations and $\textit{rm}$ operations in $u \cdot v$ are matched, and $\textit{pri}_x$ is larger than any priority in $u \cdot v \cdot w$.
-Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$.
-Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$ operations, we can see that $D_L \cap D_R = \emptyset$.
-Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
-Let $S_1$ be the set of all operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_{L1}$ be the shortest prefix of $l'_L$ that contains all operations of $S_1$, and let $l'_{L2}$ be the sequence such that $l'_L = l'_{L1} \cdot l'_{L2}$.
-It is obvious that if $\textit{put}(x,\textit{pri}_x)$ happens before an operation $o$ in $h$, then $o$ is in $l'_{L2}$.
-Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.
-
-Since priority queue is closed under projection (Lemma \ref{lemma:PQ is closed under projection}) and $u \cdot v$ contain matched $\textit{put}$ and $\textit{rm}$ operations, we know that $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$.
-Since $\textit{pri}_x$ is larger than any priority in $h$, it is easy to see that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$.
-Since the priority queue is empty after executing $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$, we know that $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R \in \textit{PQueue}$.
-
-It remains to prove that $h \sqsubseteq l''$.
-To prove this, we define a graph $G$ whose nodes are the operations of $h$ and there is an edge from operation $o_1$ to $o_2$, if one of the following case holds
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o_1$ happens-before $o_2$ in h,
-
-\item[-] the method event corresponding to $o_1$ in $l''$ is before the one corresponding to $o_2$.
-\end{itemize}
-
-Assume there is a cycle in $G$.
-According the the property of interval order and the fact that the order of $l''$ is total, we know that there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $o_1,o_2 \in l'_{L1}$, or $o_2 \in l'_{L1} \wedge o_1 \in l'_{L2}$, or $o_1,o_2 \in l'_{L2}$, or $o_1,o_2 \in L'_R$, then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] If $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_{L1}$, then since $\textit{put}(x,\textit{pri}_x)$ happens before $o_2$, we know that $o_2 \in l'_{L2}$, which contradicts that $o_2 \in l'_{L1}$.
-
-\item[-] If $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L1}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L1}$, or $o_1 = \textit{rm}(x) \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 \in L'_R \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 = \textit{rm}(x)$, then $l$ contradicts with happen before relation of $h$.
-
-\item[-] If $o_1 \in l'_{L2} \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, then since $o_1$ happens before $\textit{put}(x,\textit{pri}_x)$, we know that $o_1 \in S_1$ and then $o_1 \in l'_{L1}$, which contradicts that $o_1 \in l'_{L2}$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
-
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr2isStepByStepLinearizability}
-\label{lemma:Rpr2 is step-by-step linearizability}
-
-$R_{\textit{pr2}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr2}}$ with witness $x$.
-Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-Since $h' \sqsubseteq l'$, there exists a sequence $h_{l'}$ of call, return and operations, such that (1) the projection of $h_{l'}$ into call and return action is $h$, (2) the projection of $h_{l'}$ into operations is $l'$, and (3) for each matched pair of call action $c$ and return action $r$ of $h$, there exists an operation $o$, such that $a$ before $o$, $o$ before $r$, and $o$ is the operation of $c$ and $r$.
-Intuitively, $h_{l'}$ is obtained from $h$ by adding linearization points.
-
-According to $R_{\textit{pr2}}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where (1) the $\textit{put}$ operations and $\textit{rm}$ operations in $u \cdot v$ are matched, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v \cdot w$, (3) there exists matched $\textit{put}$ and $\textit{rm}$ operations in $u \cdot v \cdot w$ that has priority $\textit{pri}_x$, (4) $\textit{pri}_x$ is larger than priorities of any unmatched put in $u \cdot v \cdot w$, and (5) all $\textit{put}$ operation with priority $\textit{pri}_x$ is in $u$.
-Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$.
-Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$ operations, we can see that $D_L \cap D_R = \emptyset$.
-Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
-
-Let $l'_{\textit{L0}}$ be the shortest prefix of $l'_L$, such that every operation in $l'_{\textit{L0}}$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $h_{l'}$.
-Let $\textit{OP}_x = \{ \textit{put}(d,\textit{pri}_x) \vert \textit{put}(d,\textit{pri}_x)$ is an operation of $h\}$.
-Let $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t$, where $l_t$ is the projection of $l'$ into operations of $\textit{OP}_x$ that are not in $l'_{\textit{L0}}$.
-Let $\textit{OP}'_x$ be the set of operations in $l_t$.
-Let $l'_{\textit{L2}}$ be the projection of $l'_L$ into operations of $l'_L$ that are not in $l'_{\textit{L1}}$.
-Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.
-
-As in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, we know that $l'_L \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$.
-It is obvious that $l'_{\textit{L0}} \in \textit{PQueue}$, and then $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t \in \textit{PQueue}$.
-Since there is no $\textit{rm}(\textit{Empty})$ in $h$ and $\textit{pri}_x$ is the maximal priority of $h$, $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$.
-Since all $\textit{put}$ operations with priority $\textit{pri}$ is in $l'_{L1}$, we know that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$.
-As in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, we know that $l'' \in \textit{PQueue}$.
-
-To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}.
-Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-
-\item[-] $o_1 \in L'_R \wedge (o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2} \vee o_2 = \textit{rm}(x))$, or $o_1 = \textit{rm}(x) \wedge (o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2})$: Then $l$ contradicts with happen before relation of $h$.
-
-\item[-] $o_1, o_2 \in l'_R$ or $o_1, o_2 \in l'_{L2}$: Then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] $o_1, o_2 \in l'_{L1}$: If $(o_1, o_2 \in l'_{L0}) \vee (o_1,o_2 \in \textit{OP}'_x)$, then $l'$ contradicts with happen before relation of $h$.
-
-    Else, it is the case that $o_1 \in \textit{OP}'_x \wedge o_2 \in l'_{L0}$.
-We can see that $o_1$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and the linearization point of $o_1$ is after the return action of $\textit{put}(x,\textit{pri}_x)$.
-Since $o_1 <_{\textit{hb}} o_2$, it is easy to see that $\textit{put}(x,\textit{pri}_x) <_{\textit{hb}} o_2$, and then $o_2 \notin l'_{L0}$, which contradicts that $o_2 \in l'_{L0}$.
-
-\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_{L1}$: Since $o_1 <_{\textit{hb}} o_2$, it is obvious that $o_2 \notin l'_{L1}$, contradicts that $o_2 \in l'_{L1}$.
-
-\item[-] $o_1 \in l'_{L2} \wedge o_2 \in l'_{L1}$: If $o_2 \in l'_{L0}$, then $l'$ contradicts with happen before relation of $h$.
-
-    Else, it is the case that $o_2 \in \textit{OP}'_x$.
-$o_2$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and the linearization point of $o_2$ is after the return action of $\textit{put}(x,\textit{pri}_x)$.
-Since $o_1 <_{\textit{hb}} o_2$, the return action of $o_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$.
-Then it is easy to see that $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.
-
-\item[-] $o_1 \in l'_{L2} \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: Then $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr3isStepByStepLinearizability}
-\label{lemma:Rpr3 is step-by-step linearizability}
-
-$R_{\textit{pr3}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr3}}$ with witness $x$.
-Assume that  Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Then, let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-According to $R_{\textit{pr3}}$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, where $\textit{pri}_x$ is the maximal priority in $u \cdot v$, and $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ operations in $u \cdot v$ (may equal priority of some unmatched $\textit{put}$ operations in $u \cdot v$).
-Let $S_1$ be the set of all operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_1$ be the shortest prefix of $l'$ that contains all operations of $S_1$, and let $l'_2$ be the sequence such that $l' = l'_1 \cdot l'_2$.
-Let sequence $l'' = l'_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_2$.
-
-Since $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ operations in $u \cdot v$, it is easy to see that $l'' \in \textit{PQueue}$.
-It remains to prove that $h \sqsubseteq l''$.
-To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}.
-Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o_1,o_2 \in l'_1$, or $o_1,o_2 \in l'_2$, or $o_1 \in l'_2 \wedge o_2 \in l'_1$: Then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] $o_1 \in l'_2 \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: $o_1 \in S_1$, and then $o_1 \in l'_1$, which contradicts that $o_1 \in l'_2$.
-
-\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_1$: We can see that in $l'$, $o_2$ must after every operation in $S_1$, and then $o_2 \notin l'_1$, contradicts that $o_2 \in l'_1$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
-\end {proof}
-
-The following lemma states that, from linearization of sub-histories, we can merge them and obtain a linearization (regardless of whether it belongs to sequential specification) of the whole history.
-
-\begin{restatable}{lemma}{MergeTwoLinearization}
-\label{lemma:merge two linearization}
-
-Given a history $h$ and operation set $S_1$, $S_2$, let $h_1 = h \vert{S_1}$ and $h_2 = h \vert{S_2}$.
-Assume that $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, and $S_1 \cup S_2$ contains all operations of $h$.
-Then, there exists a sequence $l$, such that $h \sqsubseteq l$, $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$.
-\end{restatable}
-
-\begin {proof}
-
-Given a history $h$ and a operation $o \in S_2$, let $\textit{MB}(o) = \{ o' \vert o' \in S_1$ and $o'$ happens before $o$ in $h \}$, let $\textit{SBI}(o) = \textit{min}\{ i \vert l_1[0,i]$ contains all elements of $\textit{MB}(o) \}$.
-
-Let $l = s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[n] \cdot s_{\textit{n+1}}$ be generated as follows, where $n = \vert l_2 \vert$:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $s_1 = l_1[0, \textit{SBI}(l_2[1])]$,
-
-\item[-] If $s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[i]$ already contains $l_1(\textit{SBI}(l_2[\textit{i+1}]))$, then $s_{\textit{i+1}} = \epsilon$.
-Otherwise, $s_{\textit{i+1}}$ is a subsequence of $l_1$, which starts from the next of last elements of $s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[i]$ in $l_1$ and ends in $l_1(\textit{SBI}(l_2[\textit{i+1}]))$.
-\end{itemize}
-
-It is obvious that $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$, and it remains to prove that $h \sqsubseteq l$.
-We prove this by contradiction.
-Assume that $h$ is not linearizable with respect to $l$.
-Then there must be two operations of $h$, such that $o_1 <_{hb} o_2$ in $h$ but $o_2$ before $0_1$ in $l$.
-Since $l \vert{S_1} = l_1$, $l \vert{S_2} = l_2$, and $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, it is easy to see that it is impossible that $o_1,o_2 \in S_1$ or $o_1,o_2 \in S_2$.
-There are only two possibilities:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o_1 \in S_1 \wedge o_2 \in S_2$.
-Then we can see that $o_2=l_2[i]$ and $o_1 \in s_j$ for some $i < j$.
-Since $o_1 <_{hb} o_2$, we know that $o_1 \in \textit{SBI}(o_2)$.
-By the construction of $l$, we know that $o_1$ must be in $s_k$ for some $k \leq i$, contradicts that $o_1 \in s_j$ with $i < j$.
-
-\item[-] $o_1 \in S_2 \wedge o_2 \in S_1$.
-Then we can see that $o_2 \in s_i$ and $o_1 = l_2[j]$ for some $i \leq j$.
-It is easy to see that this leads to contradiction when $i = j$.
-For the case of $i \neq j$, we need to satisfy the following requirements: (1) $o_1$ ($l_2[j]$) does not happen before $l_2[i]$, (2) $l_2[i]$ does not happen before $o_2$, (3) $o_2$ is either overlap or happens before $o' \in \textit{MB}(l_2[i])$, and (4) $o' <_{hb} l_2[i]$.
-By enumeration we can see that it is impossible that above four conditions be satisfied while $o_1 <_{hb} o_2$.
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr4isStepByStepLinearizability}
-\label{lemma:Rpr4 is step-by-step linearizability}
-
-$R_{\textit{pr4}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr4}}$ with witness $x$.
-Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-According to $R_{\textit{pr4}}$, there exist sequences $u$ and $v$, such that (1) $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v$, (3) $\textit{pri}_x$ equals the maximal priority of matched $\textit{put}$ operations in $u \cdot v$, (4) all matched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$ and (5) all the other unmatched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$.
-
-Since $h' \sqsubseteq l'$, there exists a sequence $h_{l'}$ of call, return and operations, such that (1) the projection of $h$ into call and return action is $h$, (2) the projection of $h_{l'}$ into operations is $l'$, and (3) for each matched pair of call action $c$ and return action $r$ of $h$, there exists an operation $o$, such that $a$ before $o$, $o$ before $r$, and $o$ is the operation of $c$ and $r$.
-Intuitively, $h_{l'}$ is obtained from $h$ by adding linearization points.
-
-According to $R_{\textit{pr4}}$, there exist sequences $u$ and $v$, such that (1) $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v$, (3) $\textit{pri}_x$ equals the maximal priority of matched $\textit{put}$ operations in $u \cdot v$, (4) all matched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$ and (5) all the other unmatched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$.
-
-Let $l'_1$ be the shortest prefix of $l'$, such that every operation in $l'_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $h_{l'}$.
-Let $\textit{OP}_x$ be the set of $\textit{put}$ operations with priority $\textit{pri}_x$ and their matched $\textit{rm}$ operations in $h$, let $\textit{OP}'_x$ be the set of $\textit{put}$ operations in $\textit{OP}_x$ that do not appear in $l'_1$, and let $\textit{OP}''_x$ be the set of $\textit{rm}$ operations in $\textit{OP}_x$ that do not appear in $l'_1$.
-Let $l_x''=l'_1 \cdot l'_2 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_3$, where $l'_2$ is the projection of $l'$ into operations of $\textit{OP}'_x$, and $l'_3$ is the projection of $l'$ into operations of $\textit{OP}''_x$.
-
-Let $h_x$ be the projection of $h$ into $\textit{OP}_x$.
-Let us prove that $h_x \sqsubseteq l''_x$ by contradiction.
-Assume that there exists operations $o_1,o_2$, such that $o_1$ happens before $o_2$ in $h_x$, and $o_2$ is before $o_1$ in $l''_x$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $(o_1 \in l'_3) \wedge (o_2 \in l'_1 \vee o_2 \in l'_2 \vee o'_2 \in l'_3)$, or $(o_1 \in l'_2) \wedge (o_2 \in l'_1 \vee o_2 \in l'_2)$, or $o_1,o_2 \in l'_1$: Then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_1$: Since the linearization point of $o_2$ is before return action of $o_1$, $o_2$ either overlaps with $o_1$, or happens before $o_1$, which contradicts that $o_1$ happens before $o_2$.
-
-\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_2$: Since $l'_2$ only contains $\textit{put}$ operations with priority $\textit{pri}_x$ and are in $\textit{OP}'_x$, it is not hard to see that each operation in $l'_2$ must overlap with $\textit{put}(x,\textit{pri}_x)$, which contradicts that $o_1$ happens before $o_2$.
-
-\item[-] $o_1 \in l'_3 \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: According to the construction of $l''$, we can see that $1_2 \in l'_1$, which contradicts that $1_2 \in l'_3$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h_x \sqsubseteq l''_x$.
-
-Let $\textit{OP}_{\textit{-x}}$ be the set of $\textit{put}$ operations with non-$\textit{pri}_x$ priority, and their matched $\textit{rm}$ operations.
-Let $h_{\textit{-x}}$ be the projection of $h$ into $\textit{OP}_{\textit{-x}}$.
-Let $l''_{\textit{-x}}$ be the projection of $l'$ into$\textit{OP}_{\textit{-x}}$.
-It is obvious that $h_{\textit{-x}} \sqsubseteq l''_{\textit{-x}}$.
-By Lemma \ref{lemma:merge two linearization}, there exists a sequence $l'''$, such that $h \sqsubseteq l'''$, $l''' \vert{\textit{OP}_x} = l_x''$ and $l''' \vert{\textit{OP}_{\textit{-x}}} = l''_{\textit{-x}}$.
-
-It is obvious that $l''_{\textit{-x}} \in \textit{PQueue}$.
-Let $l'_x$ be the projection of $l'$ into $\textit{OP}_x - \{ \textit{put}(x,\textit{pri}_x) \}$.
-It is obvious that $l'_x \in \textit{PQueue}$ and $l'_1 \in \textit{PQueue}$.
-Since (1) $l'_1 \cdot l'_2 \cdot l'_3$ is obtained from $l'_x$ by making $\textit{put}$ operations in $l'_2$ to happen earlier, and (2) $l'_3$ contains only $\textit{rm}$ operations, we can see that $l'_1 \cdot l'_2 \cdot l'_3 \in \textit{PQueue}$.
-Since $l'_3$ contains only $\textit{rm}$ operations, it is easy to see that $l_x'' \in \textit{PQueue}$.
-Since $\textit{pri}_x$ is the maximal priority of $h$ and $h$ does not contain $\textit{rm}(\textit{Empty})$ operations, the operations in $l_x''$ does not influence the correctness of $l''_{\textit{-x}} \in \textit{PQueue}$.
-Therefore, we can see that $l''' \in \textit{PQueue}$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr5isStepByStepLinearizability}
-\label{lemma:Rpr5 is step-by-step linearizability}
-
-$R_{\textit{pr5}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr5}}$ with operation $o=\textit{rmEmpty}$.
-Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Let $h'=h \setminus o$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-According to $R_{\textit{pr5}}$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{rmEmpty} \cdot v$, where the $\textit{put}$ operations and $\textit{rm}$ operations in $u$ are matched.
-
-Let $D_L$ be the set of items in $u$ and $D_R$ be the set of items in $v$.
-Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
-Let sequence $l'' = l'_L \cdot o \cdot L'_R$.
-Since priority queue is closed under projection (Lemma \ref{lemma:PQ is closed under projection}) and $u$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $l'_L \in \textit{PQueue}$ and the the priority queue is empty after executing $l'_L$.
-Then we know that $l'_L \cdot o \in \textit{PQueue}$.
-Since $l'_R$ is obtained from $l'$ by discarding pairs of matched $\textit{put}$ and $\textit{rm}$ operations, it is easy to see that $L'_R \in \textit{PQueue}$, and then we know that $l'' = l'_L \cdot o \cdot L'_R \in \textit{PQueue}$.
-
-It remains to prove that $h \sqsubseteq l''$.
-To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}.
-Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o_1,o_2 \in l'_L$, or $o_1,o_2 \in l'_R$: Then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] If $o_1=o \wedge o_2 \in l'_L$, or $o_1 \in l'_R \wedge o_2 \in l'_L$, or $o_1 \in l'_R \wedge o_2 = 0$, then $l$ contradicts with happen before relation of $h$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$. \qed
-\end {proof}
-
-The following lemma states that $\textit{PQueue}$ is step-by-step linearizability, it is a direct consequence of Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, Lemma \ref{lemma:Rpr2 is step-by-step linearizability}, Lemma \ref{lemma:Rpr3 is step-by-step linearizability}, Lemma \ref{lemma:Rpr4 is step-by-step linearizability} and Lemma \ref{lemma:Rpr5 is step-by-step linearizability}.
-
-\begin{restatable}{lemma}{PQueueisStepByStepLinearizability}
-\label{lemma:PQueue is step-by-step linearizability}
-
-$\textit{PQueue}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\section{Co-Regular of Priority Queues}
-\label{sec:co-regular of priority queues}
-
-\subsection{Co-Regular of $R_{\textit{pr1}}$}
-\label{subsec:co-regular of Rpr1}
-
-Let us introduce the notion of gap for matched $\textit{put}$ and $\textit{rm}$ operations.
-
-\begin{definition}[gap for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:gap for matched put and rm operations}
-
-Let $h$ be a data-differentiated history and $o_1 = \textit{put}(x,\textit{pri}_x)$, $o_2 = \textit{rm}(x)$ be a pair of matched $\textit{put}$ and $\textit{rm}$ operation of $h$.
-We say that $h$ has a gap on operations $o_1$ and $o_2$, if there is a partition of the operations of $h$ into $L$, $M$ and $R$ satisfying:
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $L \cup M$ has matched $\textit{put}$ and $\textit{rm}$ operations.
-
-\item[-] no operation of $R$ happens before an operation of $L$, $o_1$, $M$ or $o_2$.
-
-\item[-] no operation of $o_2$ happens before an operation of $L$, $o_1$ or $M$.
-
-\item[-] no operation of $M$ happens before an operation of $L$ or $o_1$.
-
-\item[-] no operation of $o_1$ happens before an operation of $L$.
-
-\item[-] no operation in $L \cup M \cup R$ has priority $\textit{pri}_x$.
-\end{itemize}
-\end{definition}
-
-Given a history $h$ and set $O$ of some operations in $h$.
-It is easy to see that there exists a sequence $l$ of operations of $O$, such that $l$ does not contradicts the happen before relation of $h$.
-Therefore, from now on, we can use this fact to directly generate such sequence from such $O$.
-
-The following lemma reduce linearizable with respect to $\textit{MR}_{\textit{pr1}}$ into gap of matched $\textit{put}$ and $\textit{rm}$ operations.
-
-\begin{restatable}{lemma}{GapEqualsLinforRpr1}
-\label{lemma:Gap Equals Lin for Rpr1}
-A data-differentiated history $h$ has a projection $h'$, such that $\textit{last}(h') = R_{\textit{pr1}}$ and $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$, if and only if there exists a projection $h''$ of $h$, such that $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is in $h''$, $\textit{pri}_x$ is the maximal priority of $h''$, and one the the following case holds:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] case $1$: $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$,
-\item[-] case $2$: there is no gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-\end{itemize}
-\end{restatable}
-
-\begin {proof}
-
-Let us prove the $\textit{only if}$ direction by contradiction.
-Assume that there exists such $h'$, while case $1$ and case $2$ do not hold for any projection of $h$.
-Since case $1$ does not hold for any projection of $h$, we know that $\forall y \in D$, $\textit{rm}(y)$ does not happens before $\textit{put}(y,\_)$ in $h$.
-Since case $2$ does not hold for any projection of $h$ and $\textit{last}(h') = R_{\textit{pr1}}$, we can see that $h'$ has gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$, while $\textit{pri}_x$ is the maximal priority in $h'$.
-Since $\textit{last}(h') = R_{\textit{pr1}}$, we can see that $\textit{pri}_x$ is larger than priority of all other $\textit{put}$ operations of $h$.
-
-By the properties of the gap, there is partitions $L$, $M$ and $R$.
-Let $u$, $v$ and $w$ contain the operations of $L$, $M$ and $R$, respectively.
-It is easy to see that $h'$ is linearizable with respect to $l''=u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$ and $l'' \in \textit{MR}_{\textit{pr1}}$.
-This contradicts the fact that $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$.
-
-Let us begin to prove the $\textit{if}$ direction.
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If case $1$ holds for $h''$, then $h \vert_{ \{ x \} }$ is the $h'$ as required.
-
-\item[-] If case $2$ holds, assume by contradiction that, for every projection $h'$ of $h$, if $\textit{last}(h') = R_{\textit{pr1}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr1}}$.
-Since $h''$ is also a projection of $h$, we know that there exist sequences $u$, $v$ and $w$, such that $h'' \sqsubseteq u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, $u \cdot v$ contains matched $\textit{put}$ and $\textit{rm}$ operations, and $\textit{pri}_x$ is larger than priority of $u \cdot v \cdot w$.
-Let $L$, $M$ and $R$ be the operations of $u$, $v$ and $w$, respectively.
-Then, the partitions $L$, $M$ and $R$ form a gap on operation $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-This contradicts the fact that there is no gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\begin{definition}[left-right constraint for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:left-right constraint for matched put and rm operations}
-Given a data-differentiated history $h$, and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-The left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is the graph $G$ where:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] the nodes are the items of $h$, to which we add a node,
-
-\item[-] there is an edge from item $d_1$ to $x$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$ and $\textit{pri}_1 \neq \textit{pri}_x$,
-
-\item[-] there is an edge from $x$ to item $d_1$, if $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$, and the priority of $d_1$ is not $\textit{pri}_x$,
-
-\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{rm}(d_2,\textit{pri}_2)$, $\textit{pri}_1 \neq \textit{pri}_x$, and $\textit{pri}_2 \neq \textit{pri}_x$.
-\end{itemize}
-\end{definition}
-
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-Let $\textit{LMSet}_1(h,x) = \{ o \vert$ $o$ is not an operation of $x$, the priority of $o$ does not equal the priority of $x$, and either it happens before $\textit{put}(x)$ or $\textit{rm}(x)$ in $h$, or there is an operation $o'$ with the same item of $o$, such that $o'$ happens before $\textit{put}(x)$ or $\textit{rm}(x)$ in $h$ $\}$.
-For each $i \geq 1$, let $\textit{LMSet}_{\textit{i+1}}(h,x) = \{ o \vert$ $o$ is not an operation of $x$ or items in $\textit{LMSet}_k(h,x)$ for each $k \leq i$, the priority of $o$ does not equal the priority of $x$, and either it happens before some operation $o' \in \textit{LMSet}_i(h,x)$ in $h$, or there is an operation $o''$ with the same item of $o$ and $o''$ happens before some operation $o' \in \textit{LMSet}_i(h,x)$ in $h$ $\}$.
-%We can see that for each $i$, $\textit{LMSet}_i(h,x)$ contains matched $\textit{put}$ and $\textit{rm}$ operations, and
-We can see that $\textit{LMSet}_i(h,x) \cap \textit{LMSet}_j(h,x) = \emptyset$ for any $i \neq j$.
-Let $\textit{LMSet}(h,x) = \textit{LMSet}_1(h,x) \cup \textit{LMSet}_2(h,x) \cup \ldots$.
-%{\color {red} We say that a history $h$ of priority queue is matched before $x$, if (1) $h$ contains no $\textit{rmEmpty}$ operations, (2) $ h \sqsubseteq l_1 \cdot \textit{put}(x,\_) \cdot l_2 \cdot \textit{rm}(x) \cdot l_3$ for some $l_1,l_2,l_3$, and $l_1 \cdot l_2$ contains matched $\textit{put}$ and $\textit{rm}$ operations.}
-
-The following is an example that shows that, given a data-differentiated history $h$, the set $\textit{LMSet}_i(h,x)$ can be non-empty for each $1 \leq i \leq 3$.
-In \figurename~\ref{fig:his nobound of LMSet} , $\textit{LMSet}_1(h,x) = \{ \textit{put}(a,1),\textit{rm}(a) \}$, $\textit{LMSet}_2(h,x) = \{ \textit{put}(b,1),\textit{rm}(b) \}$ and $\textit{LMSet}_3(h,x) = \{ \textit{put}(a,c),\textit{rm}(c) \}$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{PIC_HIS_NOBOUNDOF_LMSET.pdf}
-%\vspace{-10pt}
-  \caption{An example for non-emptiness of $\textit{LMSet}_1(h,x)$, $\textit{LMSet}_2(h,x)$ and $\textit{LMSet}_3(h,x)$}
-  \label{fig:his nobound of LMSet}
-\end{figure}
-
-\begin{restatable}{lemma}{RmxDoesNotHappenBeforeLMSetForRpr1}
-\label{lemma:Rmx does not happen before LMSet for Rpr1}
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-%, such that $\textit{pri}_x$ is the maximal priority of $h$, and $\textit{last}(h) = R_{\textit{pr1}}$.
-Assume that $\textit{rm}(y)$ does not happen before $\textit{put}(y,\_)$ for each item $y$ in $h$.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-Assume that $G$ has no cycle going through $x$.
-Then, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}(h,x)$.
-\end{restatable}
-
-\begin {proof}
-
-We prove this lemma by induction, and prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x)$, in $\textit{LMSet}_2(h,x)$, $\ldots$.
-Note that, according to the definition of $\textit{LMSet}(h,x)$, $\textit{pri}_x$ is different to priority of all items in $\textit{LMSet}(h,x)$.
-
-\noindent (1) Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x)$ by contradiction.
-Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{LMSet}_1(h,x)$ is an operation of item $d$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d$ does not equals $\textit{pri}_x$).
-
-We use a triple $(t_1,t_2,t_3)$ to represent related information.
-$t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$.
-$t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation.
-$t_2$ and $t_3$ is used for the reason of $o \in \textit{LMSet}_1(h,x)$: $o \in \textit{LMSet}_1(h,x)$, since an operation (of kind $t_2$) of $d$ happens before an operation (of kind $t_3$) of $x$.
-Let us consider all the possible cases of $(t_1,t_2,t_3)$:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.
-
-\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{rm}(x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{rm}(x)$.
-
-\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x) )$.
-By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(x,\textit{pri}_x) ) \vee ( \textit{rm}(d) <_{hb} \textit{put}(d,\_) )$, which is impossible.
-
-\item[-] $(\textit{put},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{rm}(x) )$.
-We can see that $\textit{rm}(d) <_{hb} \textit{rm}(x) <_{hb} \textit{put}(d,\_)$, which contradicts that $\textit{rm}(d)$ does not happen before $\textit{put}(d,\_)$.
-
-\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x) )$.
-We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.
-
-\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{rm}(x) )$.
-We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.
-
-\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.
-
-\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{rm}(x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{rm}(x)$.
-\end{itemize}
-
-This completes the proof for $\textit{LMSet}_1(h,x)$.
-
-\noindent (2) Assume we already prove that for some $j \geq 1$, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_{\textit{j+1}}(h,x)$ by contradiction.
-Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$ is an operation of item $d_{\textit{j+1}}$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d_{\textit{j+1}}$ does not equals $\textit{pri}_x$).
-We use a triple $(t_1,t_2,t_3)$ to represent related information.
-$t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$.
-$t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation.
-$t_2$ and $t_3$ is used for the reason of $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$: $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$, since an operation (of kind $t_2$) of $d_{\textit{j+1}}$ happens before an operation (of kind $t_3$) of $d_j$, where $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{LMSet}_j(h,x)$.
-Let us consider all the possible cases of $(t_1,t_2,t_3)$:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_)$.
-We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-
-\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j,\_)$.
-We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j,\_) ) \wedge ( \textit{rm}(d_j) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-
-\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$.
-By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \vee ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_{\textit{j+1}},\_) )$, which is impossible.
-
-\item[-] $(\textit{put},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$.
-By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \vee ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_{\textit{j+1}},\_) )$, which is impossible.
-
-\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) )$.
-Let us consider the reason of $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{LMSet}_j(h,x)$:
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $( j > 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(_{\textit{j-1}},\_), \textit{rm}(_{\textit{j-1}}) \in \textit{LMSet}_{\textit{j-1}}(h,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, we can see that $\textit{put}(d_{\textit{j+1}},\_) <_{hb} o''$, and then operations of $d_{\textit{j+1}}$ is in $\textit{LMSet}_j(h,x)$, contradicts that operations of $d_{\textit{j+1}}$ is in $\textit{LMSet}_{\textit{j+1}}(h,x)$.
-
-    \item[-] If $( j = 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of $x$: Similar to above case.
-
-    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{LMSet}_{\textit{j-1}}(h,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, we can see that $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} o'' ) \vee ( \textit{rm}(d_j) <_{hb} \textit{put}(d_j,\_) )$, which is impossible.
-
-    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of $x$: Similar to above case.
-    \end{itemize}
-
-\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Let $T_{\textit{ind}}$ be the set of sentences $\{ \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}), \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j),\ldots, \textit{put}(d_{\textit{ind+1}},\_) <_{hb} \textit{rm}(d_{\textit{ind}}) \}$.
-Here each $d_i$ is a item of some operation in $\textit{LMSet}_i(h,x)$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d_i$ does not equals $\textit{pri}_x$ for each $i$).
-Let us prove that from $T_j$ we can obtain contradiction by induction:
-
-    {\bf Base case $1$}: From $T_1$ we can obtain contradiction.
-
-    Let us prove base case $1$:
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{put}(d_1,\_)$ happens $o$, and $o$ is an operation of $x$.
-Then there is a cycle $x \rightarrow d_{\textit{j+1}} \rightarrow \ldots \rightarrow d_1 \rightarrow x$ in $G$, contradicts that $G$ has no cycle going through $x$.
-
-    \item[-] If $\textit{rm}(d_1)$ happens before $o$, and $o$ is an operation of $x$.
-Then since $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1) <_{hb} o$, we can see that $\textit{put}(d_2,\_) <_{hb} o$, and then $\textit{put}(d_2,\_) \in \textit{LMSet}_1(h,x)$, contradicts that $\textit{put}(d_2,\_) \in \textit{LMSet}_2(h,x)$.
-    \end{itemize}
-
-    {\bf Base case $2$}: From $T_2$ we can obtain contradiction.
-
-    Let us prove base case $2$: If $\textit{rm}(d_2) <_{hb} o$, and $o$ is an operation of $d_1$, then since $( \textit{put}(d_3,\_) <_{hb} \textit{rm}(d_2) ) \wedge ( \textit{rm}(d_2) <_{hb} o )$, we know that $\textit{put}(d_3,\_) <_{hb} o$.
-This implies that $\textit{put}(d_3,\_) \in \textit{LMSet}_2(h,x)$, contradicts that $\textit{rm}(d_3,\_) \in \textit{LMSet}_3(h,x)$.
-Therefore, it is only possible that $\textit{put}(d_2,\_)$ happens before an operation of $d_1$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{put}(d_1,\_)$ and $\textit{put}(d_1,\_)$ happens before operations of $x$, then we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, which is impossible.
-
-    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{put}(d_1,\_)$ and $\textit{rm}(d_1)$ happens before operations of $x$, then by interval order, we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, or $\textit{rm}(d_1) <_{hb} \textit{put}(d_1,\_)$, which is impossible.
-
-    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{put}(d_1,\_)$ happens before operations of $x$, then $x \rightarrow d_{\textit{j+1}} \rightarrow \ldots \rightarrow d_1 \rightarrow x$ in $G$, contradicts that $G$ has no cycle going through $x$.
-
-    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1)$ happens before operations of $x$, then we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, which is impossible.
-    \end{itemize}
-
-    {\bf induction step}: Given $\textit{ind} \geq 3$, if from $T_{\textit{ind-1}}$ we can obtain contradiction, then from $T_{\textit{ind}}$ we can also contain contradiction.
-
-    Prove of the induction step: Similarly as base case $2$, we can prove that it is only possible that $\textit{put}(d_{\textit{ind}},\_)$ happens before operations of $d_{\textit{ind-1}}$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$ and $\textit{put}(d_{\textit{ind-1}},\_)$ happens before operations of $d_{\textit{ind-2}}$, then we know that $\textit{put}(d_{\textit{ind}})$ happens before operation of $d_{\textit{ind-2}}$, which is impossible.
-
-    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$ and $\textit{rm}(d_{\textit{ind-1}})$ happens before operations of $d_{\textit{ind-2}}$, then by interval order, we know that $\textit{put}(d_{\textit{ind}},\_)$ happens before operation of $d_{\textit{ind-2}}$, or $\textit{rm}(d_{\textit{ind-1}}) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$, which is impossible.
-
-    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{rm}(d_{\textit{ind-1}})$, then we obtain $T_{\textit{ind-1}}$, which already contain contradiction.
-    \end{itemize}
-
-    By base case $1$, base case $2$ and the induction step, it is easy to see that $T_j$ contains contradiction.
-
-\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$.
-We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-
-\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$.
-We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \wedge ( \textit{rm}(d_j) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-\end{itemize}
-
-This completes the proof for $\textit{LMSet}_{\textit{j+1}}(h,x)$.
-Therefore, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}(h,x) = \textit{LMSet}_1(h,x) \cup \textit{LMSet}_2(h,x) \cup \ldots$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{LMSetHasMatchedPutandRmOperations}
-\label{lemma:LMSet has matched put and rm operations}
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-Assume that $\textit{rm}(y)$ does not happen before $\textit{put}(y,\_)$ for each item $y$ in $h$.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-Assume that $G$ has no cycle going through $x$.
-Then, $\textit{LMSet}(h,x)$ contains only matched $\textit{put}$ and $\textit{rm}$ operations.
-\end{restatable}
-\begin {proof}
-
-We prove this lemma by contradiction.
-Assume that there exists a value, such that $\textit{LMSet}(h,x)$ contains only its $\textit{put}$ operation and does not contain its $\textit{rm}$ operation.
-Then we can see that there exists $d_1,\ldots,d_j$.
-Intuitively, $d_1,\ldots,d_j$ are items in $\textit{LMSet}_1(h,x), \ldots, \textit{LMSet}_i(h,x)$, respectively.
-$\textit{LMSet}(h,x)$ contains $\textit{put}(d_j,\_)$ and does not contain $\textit{rm}(d_j)$.
-And each $d_i$ is the reason of $d_{\textit{i+1}} \in \textit{LMSet}_{\textit{i+1}}(h,x)$.
-Formally, we require that
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] For each $1 \leq i \leq j$, operations of $d_i$ belongs to $\textit{LMSet}_i(h,x)$. %For each $1 \leq i \leq j$, the priority of $d_i$ is not $\textit{pri}_x$.
-
-\item[-] For each $i \neq j$, $\textit{put}(d_i,\_),\textit{rm}(d_i) \in \textit{LMSet}_i(h,x)$. $\textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x)$, and $h$ does not contain $\textit{rm}(d_j)$.
-
-\item[-] An operation of $d_1$ happens before an operation of $x$. For each $1 < i \leq j$, an operation of $d_j$ happens an operation of $d_{\textit{i-1}}$.
-
-\item[-] For each $k$ and $\textit{ind}$, if $k > \textit{ind+1}$, then no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.
-\end{itemize}
-
-According to the definition of $\textit{LMSet}(h,x)$, it is easy to see that such $d_1,\ldots,d_j$ exists. Let us prove the following fact:
-
-\noindent {\bf $\textit{fact}_1$}: Given $1 \leq i < j$, it can not be the case that $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap.
-
-Proof of $\textit{fact}_1$: We prove $\textit{fact}_1$ by contradiction.
-Assume that for some $i \neq j$, $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap.
-Since $\textit{put}(d_i,\_), \textit{rm}(d_i) \in \textit{LMSet}_i(h,x)$, we know that an operation $o_i$ of $d_i$ happens before operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$.
-Moreover, since $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap, it is not hard to see that the call action of $\textit{put}(d_i,\_)$ and the call action of $\textit{rm}(d_i)$ is before the call action of $o_{\textit{i-1}}$.
-Since $\textit{put}(d_{\textit{i+1}},\_), \textit{rm}(d_{\textit{i+1}}) \in \textit{LMSet}_{\textit{i+1}}(h,x)$, we know that an operation $o'_{\textit{i+1}}$ of $d_{\textit{i+1}}$ happens before operation $o'_i$ of $d_i$.
-Then, it is not hard to see that $o'_{\textit{i+1}}$ also happens before $o_{\textit{i-1}}$, which contradicts that For each $k > \textit{ind+1}$, no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.
-
-We already know that an operation of $d_1$ happens before an operation of $x$.
-By $\textit{fact}_1$, we can ensure that $\textit{put}(d_1,\_)$ happens before an operation of $x$, and then $d_1 \rightarrow x$ in $G$.
-For each $1 < i < j$, we know that an operation $o_i$ of $d_i$ happens before an operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$.
-By $\textit{fact}_1$, we can ensure that $o_i=\textit{put}(d_i,\_)$ and $o_{\textit{i-1}}=\textit{rm}(d_{\textit{i-1}})$, and then $d_i \rightarrow d_{\textit{i-1}}$ in $G$.
-Since $h$ contains $\textit{put}(d_j)$ and does not contain $\textit{rm}(d_j)$, we know that $x \rightarrow d_j$ in $G$.
-Then $G$ has a cycle going through $x$, contradicts that $G$ has no cycle going through $x$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{GapEqualsConstraintforRpr1}
-\label{lemma:Gap Equals Constraint for Rpr1}
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-%, where $\textit{pri}_x$ has the maximal priority of $h$, and $\textit{last}(h) = R_{\textit{pr1}}$.
-Assume that for each item in $h$, its $\textit{rm}$ operation does not happen before its $\textit{put}$ operation.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x)$ and $\textit{rm}(x)$.
-There is a gap on $\textit{put}(x)$ and $\textit{rm}(x)$, if and only if $G$ has no cycle going through $x$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{only if}$ direction, assume that there is a gap on $\textit{put}(x)$ and $\textit{rm}(x)$, and let it be $L$, $M$ and $R$.
-Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ in $G$.
-It is obvious that the priority of each $d_i$ is not $\textit{pri}_x$.
-Then our proof proceeds as follows:
-
-According to the definition of left-right constraint, there are two possibilities.
-The first possibility is that, $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$.
-It is obvious that $\textit{rm}(d_1) \in R$, and then since $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_1),\textit{rm}(d_1) \in R$.
-Then,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Since $d_1 \rightarrow d_2$, by definition of $G$, we know that $\textit{put}(d_1)$ happens before $\textit{rm}(d_2)$.
-Since $\textit{put}(d_1) \in R$ and $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_2),\textit{rm}(d_2) \in R$.
-Similarly, for each $1 \leq i \leq m$, we know that $\textit{put}(d_i),\textit{rm}(d_i) \in R$.
-
-\item[-] Since $d_m \rightarrow x$,
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] if $\textit{put}(d_m)$ happens before $\textit{put}(x)$, then we can see that $\textit{put}(d_m) \in L$, which contradicts that $\textit{put}(d_m) \in R$.
-
-    \item[-] if $\textit{put}(d_m)$ happens before $\textit{rm}(x)$, then we can see that $\textit{put}(d_m) \in L \cup M$, which contradicts that $\textit{put}(d_m) \in R$.
-    \end{itemize}
-\end{itemize}
-
-The second possibility is that, $h$ contains one $\textit{put}(d_1,\_)$ operation and no $\textit{rm}(d_1)$ operation.
-Note that for each $j > 1$, $h$ contains $\textit{put}(d_j,\_)$ and $\textit{rm}(d_j)$.
-Since $d_m \rightarrow x$, is is obvious that $\textit{put}(d_m) \in L \cup M$.
-Since $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_m),\textit{rm}(d_m) \in R$.
-Then,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Since $d_{\textit{m-1}} \rightarrow d_m$, by definition of $G$, we know that $\textit{put}(d_{\textit{m-1}})$ happens before $\textit{rm}(d_m)$.
-Since $\textit{rm}(d_m) \in L \cup M$ and $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_{\textit{m-1}}),\textit{rm}(d_{\textit{m-1}}) \in L \cup M$.
-Similarly, for each $1 < i \leq m$, we know that $\textit{put}(d_i),\textit{rm}(d_i) \in L \cup M$, and $\textit{put}(d_1)\in L \cup M$.
-
-\item[-] Since $d_1 \rightarrow d_2$ in $G$, $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2)$.
-Since $\textit{put}(d_2),\textit{rm}(d_2) \in L \cup M$, we can see that $\textit{put}(d_1) \in L \cup M$.
-However, there is one $\textit{put}(d_1,\_)$ operation and no $\textit{rm}(d_1)$ operation in $h$, contradicts that $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
-\end{itemize}
-
-This completes the proof of the $\textit{only if}$ direction.
-
-To prove the $\textit{if}$ direction, assume that $G$ has no cycle going through $x$.
-Let $O_L$ be the set of operations that happen before $\textit{put}(x)$ in $h$.
-It is easy to see that $O_L \subseteq \textit{LMSet}(h,x)$.
-Let $O_M = \textit{LMSet}(h,x) \setminus O_L$.
-Let $O_h$ be the set of operations of $h$, and let $O_R = O_h \setminus \textit{LMSet}(h,x)$.
-Let $L$, $M$ and $R$ be $O_L$, $O_M$ and $O_R$, respectively.
-
-By Lemma \ref{lemma:LMSet has matched put and rm operations}, we can see that $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
-It remains to prove that $L$, $M$ and $R$ is a gap of $\textit{put}(x)$ and $\textit{rm}(x)$.
-We prove this by showing that all the following cases are impossible:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Case $1$: Some operation $o_r \in R$ happens before $\textit{rm}(x)$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $2$: Some operation $o_r \in R$ happens before some operation $o_{lm} \in L \cup M$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $3$: Some operation $o_r \in R$ happens before $\textit{put}(x)$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $4$: $\textit{rm}(x)$ happens before some $o_{lm} \in \textit{LMSet}(h,x) = L \cup M$. By Lemma \ref{lemma:Rmx does not happen before LMSet for Rpr1} we know that this is impossible.
-
-\item[-] Case $5$: $\textit{rm}(x)$ happens before $\textit{put}(x)$. This contradicts that for each item in $h$, its $\textit{rm}$ operation does not happen before its $\textit{put}$ operation.
-
-\item[-] Case $6$: Some operation $o_m \in M$ happens before $\textit{put}(x)$. Then we know that $o_m \in L$, which contradicts that $o_m \in M$.
-
-\item[-] Case $7$: Some operation $o_m \in M$ happens before some operation $o_l \in L$. Then we know that $o_m \in L$, which contradicts that $o_m \in M$.
-
-\item[-] Case $8$: $\textit{put}(x)$ happens before some operation $o_l \in L$. This is impossible.
-\end{itemize}
-
-This completes the proof of the $\textit{if}$ direction.
-
-\qed
-\end {proof}
-
-A automaton $\mathcal{A}_{\textit{Rpr1-1}}$ is given in \figurename~\ref{fig:automata for Rpro-1}, where $c=\textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm,a),$ $\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$.
-$\mathcal{A}_{\textit{Rpr1-1}}$ is used to detect the existence of a item whose $\textit{rm}$ happens before its $\textit{put}$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-1.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-1}}$}
-  \label{fig:automata for Rpro-1}
-\end{figure}
-
-Let us begin to represent several automata that is used for capture the case that, in a sub-history $h'$ of a history $h$, $\textit{last}(h')=R_{\textit{pr1}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$, and the reason is that there is no gap on $\textit{put}(y,\textit{pri}_y)$ and $\textit{rm}(y)$ in $h'$, where $y$ is the item that has maximal priority in $h'$.
-Such case is dealt with by four automata, since there are four possibilities of $h' \vert_{y}$.
-
-The first automaton is $\mathcal{A}_{\textit{Rpr1-pprr}}$, which is given in \figurename~\ref{fig:automata for Rpro-pprr}.
-In \figurename~\ref{fig:automata for Rpro-pprr}, $c_1= \textit{cal}(\textit{put},c,$ $\textit{anyPri}),\textit{ret}(\textit{put},c), \textit{cal}(\textit{rm},c), \textit{ret}(\textit{rm},c)$, $c_2 = c_1 + \textit{cal}(\textit{put},a,\textit{les}_p)$, and $c_3 = c_2 + \textit{ret}(\textit{rm},a)$.
-Here $p$ is a variable for priority, and $\textit{les}_p$ is a predicate that accepts all priorities that are less than the value of $p$.
-Here $\textit{pprr}$ represents that, $h' \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$ (we consider $b$ as $y$ and $p$ as $\textit{pri}_y$).
-In \figurename~\ref{fig:automata for Rpro-pprr}, the three paths that (1) from $q_{\textit{init}}$ to $q_2$ to $q_3$, (2) from $q_{\textit{init}}$ to $q_2$ to $q_{11}$, and (3) from $q_{\textit{init}}$ to $q_{10}$ to $q_{11}$ represents three possible positions of $\textit{ret}(\textit{put},a)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-pprr.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-pprr}}$}
-  \label{fig:automata for Rpro-pprr}
-\end{figure}
-
-The second automaton is $\mathcal{A}_{\textit{Rpr1-prpr}}$, which is given in \figurename~\ref{fig:automata for Rpro-prpr}.
-In $\mathcal{A}_{\textit{Rpr1-prpr}}$, $c_1, c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$.
-Here $\textit{pprr}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{put},b,p) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-prpr.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-prpr}}$}
-  \label{fig:automata for Rpro-prpr}
-\end{figure}
-
-The third automaton is $\mathcal{A}_{\textit{Rpr1-rppr}}$, which is given in \figurename~\ref{fig:automata for Rpro-rppr}.
-In $\mathcal{A}_{\textit{Rpr1-rppr}}$, $c_1, c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$.
-Here $\textit{rppr}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-rppr.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-rppr}}$}
-  \label{fig:automata for Rpro-rppr}
-\end{figure}
-
-The forth automaton is $\mathcal{A}_{\textit{Rpr1-rprp}}$, which is given in \figurename~\ref{fig:automata for Rpro-rprp}.
-In $\mathcal{A}_{\textit{Rpr1-rprp}}$, $c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$, and $c_4=c_1 + \textit{ret}(put,b)$.
-Here $\textit{rprp}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b) \cdot \textit{ret}(\textit{put})$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-rprp.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-rprp}}$}
-  \label{fig:automata for Rpro-rprp}
-\end{figure}
-
-Given a data-differentiated history $h$, we say that a item $x$ with priority $\textit{pri}_x$ in $h$ is covered by items $d_1,\ldots,d_m$ in $h$, if the priorities of $d_1,\ldots,d_m$ is not $\textit{pri}_x$, and
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{put}(d_m,\_)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$,
-
-\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,
-
-\item[-] $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$, or $\textit{rm}(d_1)$ does not exists in $h$
-\end{itemize}
-
-According to the definition of left-right constraint, in a data-differentiated history $h$, there is a cycle going through $x$, if and only if there exists items $d_1,\ldots,d_m$, such that $x$ is covered by $d_1,\ldots,d_m$.
-
-\begin{restatable}{lemma}{Rpr1IsCoRegular}
-\label{lemma:Rpr1 is co-regular}
-$R_{\textit{pr1}}$ is co-regular.
-\end{restatable}
-
-\begin {proof}
-
-Let $\textit{AutS-R}_{\textit{pr1}} = \{ \mathcal{A}_{\textit{Rpr1-1}} , \mathcal{A}_{\textit{Rpr1-pprr}}, \mathcal{A}_{\textit{Rpr1-prpr}}, \mathcal{A}_{\textit{Rpr1-rppr}}, \mathcal{A}_{\textit{Rpr1-rprp}} \}$.
-We need to prove that:
-
-\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr1}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr1}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr1}}$.
-
-By Lemma \ref{lemma:Gap Equals Lin for Rpr1} and Lemma \ref{lemma:Gap Equals Constraint for Rpr1}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:
-
-\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr1}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h''$ is a projection of $h$, such that $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is in $h''$, $\textit{pri}_x$ is the maximal priority of $h''$, and one the the following case holds:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] case $1$: $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$,
-\item[-] case $2$: $x$ is covered by some items $d_1,\ldots,d_m$ in $h''$.
-\end{itemize}
-
-\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\textit{AutS-R}_{\textit{pr1}}$.
-By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$.
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-1}}$: Let $d_1$ a item in $h_2$ such that $r(d_1)=b$ and let $h'' = h_2 \vert_{ \{ d_1 \} }$.
-$d_1$ is the item with maximal priority in $h''$.
-It is obvious that $\textit{last}(h'') = R_{\textit{pr1}}$ and $\textit{rm}(d_1)$ happens before $\textit{put}(d_1,\_)$ in $h''$.
-
-\item[-] If $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-pprr}}$, $\mathcal{A}_{\textit{Rpr1-prpr}}$, $\mathcal{A}_{\textit{Rpr1-rppr}}$, or $\mathcal{A}_{\textit{Rpr1-rprp}}$: Let $d_0$ be a item in $h_2$ such that $r(d_0)=b$, let $d_1,\ldots,d_m$ be the items in $h_2$ such that $r(d_i)=a$ for each $1 \leq i \leq m$.
-Let $h'' = h_1 \vert_{ \{ d_0, d_1, \ldots, d_m \} }$.
-It is obvious that the priority of $d_0$ is larger than the priorities of other operations in $h''$, and then $\textit{last}(h'') = R_{\textit{pr1}}$.
-It is easy to see that $d_0$ is covered by $d_1,\ldots,d_m$.
-\end{itemize}
-
-\noindent The $\textit{if}$ direction: Assume that there exists such $h$, $h''$, $x$ and $\textit{pri}_x$.
-Since $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{pri}_x$ is larger than the priorities of other operations in $h''$.
-Then,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If case $1$ holds: Then, $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$ in $h''$.
-Let $h_1$ be obtained from $h$ by renaming $x$ to $b$ and renaming all other items to $a$.
-By data-independence, $h_1 \in \mathcal{I}$, and it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-1}}$.
-\item[-] case $2$: Then, $x$ is covered by some items $d_1,\ldots,d_m$ in $h''$.
-There are four possibilities of $h'' \vert_{ \{ x \} }$: (1) $\textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{rm},x)$, (2) $\textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},x)$, (3) $\textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},x)$ and (4) $\textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{rm},x) \cdot \textit{ret}(\textit{put})$.
-
-    Let $h_1$ be obtained from $h$ by renaming $x$ into $b$, renaming $d_1,\ldots,d_m$ into $a$ and renaming other items into $c$.
-We can prove that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-pprr}}$, $\mathcal{A}_{\textit{Rpr1-prpr}}$, $\mathcal{A}_{\textit{Rpr1-rppr}}$ or $\mathcal{A}_{\textit{Rpr1-rprp}}$.
-
-    For example, if $h'' \vert_{ \{ x \} } = \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{rm},x)$, and $\textit{ret}(\textit{rm},d_m)$ is before $\textit{call}(\textit{put},x,\textit{pri}_x)$ and is after $\textit{ret}(\textit{put},x)$ in $h''$.
-Then in  $\mathcal{A}_{\textit{Rpr1-pprr}}$, $h_1$ is accepted by transitions from $q_{\textit{init}}$ to $q_2$ to $q_{11}$ and finally to $q_7$ (if $\textit{rm}(d_1)$ does not exists in $h''$) or $q_9$ (if $\textit{rm}(d_1)$ exists in $h''$).
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\subsection{Co-Regular of $R_{\textit{pr2}}$}
-\label{subsec:co-regular of Rpr2}
-
-An automaton $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$ is given in \figurename~\ref{fig:automata for FIFO of single priority}.
-In \figurename~\ref{fig:automata for FIFO of single priority}, let $c_1= \textit{cal}(\textit{put},c,\textit{anyPri}),\textit{ret}(\textit{put},c)$, $\textit{cal}(\textit{rm},c),\textit{ret}(\textit{rm},c)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, and $c_3 = c_2 + \textit{cal}(\textit{rm},a) + \textit{ret}(\textit{rm},a)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_QUEUE_FOR_SINGLE_PRI.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$}
-  \label{fig:automata for FIFO of single priority}
-\end{figure}
-
-According to \cite{Bouajjani:2015}, given a data-independent set $S$ of queue executions, whether every sequence $s$ in $S$ is linearizable to queue can be reduced into checking emptiness of intersection of $S$ and several finite automata.
-Let $\textit{transToQueue}(h)$ be a sequence $h'$ that is generated from $h$ by transforming $\textit{put}$ and $\textit{rm}$ into $\textit{enq}$ and $\textit{deq}$, respectively, and then discarding priorities.
-A priority queue history is called single-priority, if all of its items has only one priority.
-The following lemma states that we can also use finite automata to check ``first in first out property'' for single-priority sub-sequence of priority queue history.
-
-\begin{restatable}{lemma}{AutoForPQwithSignlePri}
-\label{lemma:automata for priority queue with single priority}
-
-Given a data-independent implementations $\mathcal{I}$ of priority queue, let $\mathcal{A} = \{ \mathcal{A}_{\textit{unmRm1}}, \mathcal{A}_{\textit{Rpr0-1}}, \mathcal{A}_{\textit{unmRm2}}, \mathcal{A}_{\textit{FIFO-Sig-Pri}} \}$, then,
-
-$\mathcal{I} \cap \mathcal{A} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority, and $\textit{transToQueue}(e')$ does not linearizable to queue.
-\end{restatable}
-
-\begin {proof}
-
-According to \cite{Bouajjani:2015}, checking linearizable with respect to queue can be reduced into checking the following three conditions for data-differentiated histories:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{deq}(a)$ happens before $\textit{enq}(a)$ (or $\textit{enq}(a)$ does not exist) for some item $a$,
-
-\item[-] there are two or more $\textit{deq}(a)$ for some item $a$,
-
-\item[-] $\textit{enq}(a)$ happens before $\textit{enq}(b)$, and $\textit{deq}(b)$ happens before $\textit{deq}(a)$ (or $\textit{deq}(a)$ does not exists) for some items $a,b$.
-\end{itemize}
-
-It is not hard to see that (1) the first condition is dealt with by $\mathcal{A}_{\textit{unmRm1}}$ (if $\textit{put}(a,\_)$ does not exist) and $\mathcal{A}_{\textit{Rpr0-1}}$ (if $\textit{put}(a,\_)$ exists), (2) the second condition is dealt with by $\mathcal{A}_{\textit{unmRm2}}$, and (3) the third condition is dealt with by $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$.
-\qed
-\end {proof}
-
-By Lemma \ref{lemma:automata for priority queue with single priority}, from now on, it is safe for us to assume that, for each data-differentiated priority queue history $h$ and each $h'$ of projection of $h$, if $h'$ is single-priority, then $\textit{transToQueue}(h')$ is linearizable to queue.
-One consequence of this assumption is that, in $h'$, there does not exists an item $a$, such that $\textit{rm}(a)$ happens before $\textit{put}(a)$.
-
-By the proof of Lemma \ref{lemma:Rpr1 is co-regular}, with the help of $\mathcal{A}_{\textit{Rpr1-pprr}}, \mathcal{A}_{\textit{Rpr1-prpr}}, \mathcal{A}_{\textit{Rpr1-rppr}}$ and $\mathcal{A}_{\textit{Rpr1-rprp}}$, from now on, we can safely assume that given a data-differentiated history $h$ and operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ in $h$, there does not exists any items $d_1,\ldots,d_m$, such that $x$ is not covered by items $d_1,\ldots,d_m$ and the priority of $d_1,\ldots,d_m$ is less that $x$.
-
-However, above two assumptions are not enough for imply the linearizability to priority queue.
-One example is shown in \figurename~\ref{fig:history introduct ob order1}.
-The history $h$ of \figurename~\ref{fig:history introduct ob order1} is not linearizable, even if $h \vert_{1}$ and $h \vert_{2}$ are both linearizable, and either $a$ or $b$ is covered by other items.
-One explanation is that, since $\textit{put}(b,1)$ happens before $\textit{put}(a,1)$, in choosing linearization points, we should make $\textit{rm}(b)$ before $\textit{rm}(a)$.
-However, to satisfy the requirements of $R_{\textit{pr2}}$, the linearization points of $\textit{rm}(a)$ can only locates in the interval from $\textit{ret}(\textit{rm},d)$ to $\textit{cal}(\textit{put},e,2)$.
-But, $\textit{rm}(b)$ is after this entire interval, and we can not locate the linearization point of $\textit{rm}(b)$ in this interval.
-When there is no such phenomenon, such as the history in \figurename~\ref{fig:history introduct ob order2}, the history becomes linearizable.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER1.pdf}
-%\vspace{-10pt}
-  \caption{above assumptions do not implies linearizability (1)}
-  \label{fig:history introduct ob order1}
-\end{figure}
-
-%\begin{figure}[htbp]
-%  \centering
-%  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER3.pdf}
-%\vspace{-10pt}
-%  \caption{above assumptions do not implies linearizability (2)}
-%  \label{fig:history introduct ob order1}
-%\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER2.pdf}
-%\vspace{-10pt}
-  \caption{above assumptions do not implies linearizability (2)}
-  \label{fig:history introduct ob order2}
-\end{figure}
-
-In the following part of this subsection, we use $<_{\textit{ob}}$ order for the case when we can infer that an item is inserted earlier than another item, and we use gap-point to represent the intervals when it is possible to locate linearizaton points of $\textit{rm}(a)$.
-
-\begin{definition}[gap-point for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:gap-point for matched put and rm operations}
-
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x),\textit{rm}(x)$ of $h$, we say $o$ is a gap-point of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$, if we can obtain a sequence $\textit{gapPSeq}(h)$ from $h$ by adding $o$ to some locations between actions of $h$, and satisfies that,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] In $\textit{gapPSeq}(h)$, $o$ is after $\textit{call}(\textit{put},x,\textit{pri}_x)$ and $\textit{call}(\textit{rm},x)$, and is before $\textit{ret}(\textit{rm},x)$.
-\item[-] For each item $y$ with non-$\textit{pri}_x$ priority in $\textit{gapPSeq}(h)$.
-If $\textit{put}(y,\_)$ and $\textit{rm}(y)$ is in $h$, then it can not be the case that $\textit{ret}(\textit{put},y,\_)$ is before $o$ and $\textit{call}(\textit{rm},x)$ is after $o$; Else if $\textit{put}(y,\_)$ is in $h$ and $\textit{rm}(y)$ is not in $h$, then it can not be the case that $\textit{ret}(\textit{put},y,\_)$ is before $o$;
-\end{itemize}
-
-We say that $o$ is a gap-point of priority $\textit{pri}_x$, if we can obtain a sequence $\textit{gapPSeq}(h)$ from $h$ by adding $o$ to some locations between actions of $h$, and satisfies only the latter requirement.
-\end{definition}
-
-\begin{restatable}{lemma}{GapEqualsGapPoint}
-\label{lemma:gap eqals the existence of gap point}
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-There is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, if and only if there is a gap-point $o$ of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{only if}$ direction, we need to prove that: if there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, then there is no gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-
-We prove this by contradiction.
-Assume that there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, but there is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-Since there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, in the time interval starts from $\textit{cal}(\textit{put},x,\textit{pri}_x)$ and $\textit{cal}(\textit{rm},x)$ and ends at $\textit{ret}(\textit{rm},x)$, there is always some item with non-$\textit{pri}_x$ priority in priority queue.
-Therefore, it is not hard to prove that there is a cycle going through the left-right constraint of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, and by Lemma \ref{lemma:Gap Equals Constraint for Rpr0}, there is no gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, which contradicts that there is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-
-To prove the $\textit{if}$ direction, assume that there is a gap-point $o$ of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$ and let $h_1$ be the sequence in definition of gap-point, which is generated from $h$ by adding $o$ into $h$.
-Let sequence $h_2$ be a sequence of call, return and operations, such that the projection of $h_2$ into call and return actions is $h$.
-$h_2$ is generated as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] First, from $h_1$, transform the point $o$ into $\textit{rm}(x)$.
-
-\item[-] Then for all the other items $y$, (1) adding an operation $\textit{put}(y,p)$ at some position between $\textit{call}(\textit{put},y,\_)$ and $\textit{ret}(\textit{put},y)$, and (2) adding an operation $\textit{rm}(y)$ at some position between $\textit{call}(\textit{rm},y)$ and $\textit{ret}(\textit{rm},y)$.
-
-\item[-] For all items $z$ with the same priority as $x$, the positions for adding operations of $z$ into $h$ is arbitrary (but need to satisfy above restriction.
-In the following we implicitly require this).
-
-\item[-] For all items $z$ whose priority is different from that of $x$,
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-
-    \item[-] If $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, and $\textit{ret}(\textit{put},z,\_)$ is before $o$ in $h_1$: Then in $h_2$, $\textit{rm}(z,\_)$ is just after $\textit{call}(\textit{rm},z)$.
-
-    \item[-] If $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, and $o$ is before $\textit{ret}(\textit{put},z,\_)$ $h_1$: Then in $h_2$, $\textit{put}(z,\_)$ is just before $\textit{ret}(\textit{put},z)$.
-
-    \item[-] Otherwise, we can see that $\textit{put}(z,\_)$ and $\textit{rm}(z)$ overlaps in $h$.
-Then we add $o_1=\textit{put}(z,\_)$ and $o_2=\textit{rm}(z)$ in the time interval of interleaving of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ overlaps in $h$, make sure $o_2$ next to $o_1$, and make sure either both of them is before $\textit{rm}(x)$ or both of them is after $\textit{rm}(x)$ in $h_2$.
-    \end{itemize}
-\end{itemize}
-
-According to the definition of gap-point, it is not hard to see that we can construct such $h_2$.
-Let $l$ be the projection of $h_2$ into operations, and $l'$ be the projection of $l$ into $x$ and all items that has priorities not equals that of $x$.
-Let $l'= u \cdot \textit{put}(x,\_) \cdot v \cdot \textit{rm}(x) \cdot w$.
-It is easy to see that $h \sqsubseteq l$, and $u \cdot v$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
-Let $L$, $M$ and $R$ be the set of operations in $u$, $v$ and $w$, respectively.
-Then it is obvious that $L$, $M$ and $R$ is a gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-\qed
-\end {proof}
-
-Given a data-differentiated history $h$ and two items $a,b$, we say that $a <_{\textit{ob}} b$, if the priority of $a$ and $b$ is the same, and one of the following cases holds:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Type $A$: $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$ in $h$,
-
-\item[-] Type $B$: $\textit{rm}(a)$ happens before $\textit{rm}(b)$ in $h$,
-
-\item[-] Type $C$: $\textit{rm}(a)$ happens before $\textit{put}(b,\_)$ in $h$,
-\end{itemize}
-
-Let $<_{\textit{ob}}^*$ be the transitive closure of $<_{\textit{ob}}$.
-Intuitively, $a <_{\textit{ob}} b$ means that $a$ should be inserted earlier than $b$, and $a <_{\textit{ob}}^* b$ means that we can infer that $a$ should be inserted earlier than $b$ with the help of other items.
-We can explicitly write $a <_{\textit{ob}}^A b$ to mean that $a <_{\textit{ob}} b$ and the reason is that $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$.
-Similarly, we can explicitly write $a <_{\textit{ob}}^B b$ and $a <_{\textit{ob}}^C b$ to mean that $a <_{\textit{ob}} b$ and the reason is that $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$ and $\textit{rm}(a)$ happens before $\textit{put}(b,\_)$, respectively.
-
-According to the definition of $<_{\textit{ob}}^*$, if $a <_{\textit{ob}}^* b$, then there exists $c_1,\ldots,c_n$, such that $a <_{\textit{ob}} c_1 <_{\textit{ob}} \ldots <_{\textit{ob}} b$.
-The following lemma states that, the number of intermediate items $c_i$ is in fact bounded.
-
-\begin{restatable}{lemma}{OBOrderHasBoundedLength}
-\label{lemma:ob order has bounded length}
-Given a data-differentiated history $h$.
-Assume that $a <_{\textit{ob}} a_1 <_{\textit{ob}} \ldots <_{\textit{ob}} a_m <_{\textit{ob}} b$, then either $a <_{\textit{ob}} b$, or $a <_{\textit{ob}} a_i <_{\textit{ob}} b$ for some $i$, or $a <_{\textit{ob}} a_i <_{\textit{ob}} a_j <_{\textit{ob}} b$ for some $i,j$.
-\end{restatable}
-
-\begin {proof}
-
-Our proof proceed as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^A$,$<_{\textit{ob}}^B \cdot <_{\textit{ob}}^B$ and $<_{\textit{ob}}^C \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^A c_2 <_{\textit{ob}}^A c_1$, then $\textit{put}(c_3,\_)$ happens before $\textit{put}(c_2,\_)$, and $\textit{put}(c_2,\_)$ happens before $\textit{put}(c_1,\_)$.
-Therefore, it is obvious hat $\textit{put}(c_3,\_)$ happens before $\textit{put}(c_1,\_)$ and $c_3 <_{\textit{ob}}^A c_1$.
-
-    Similarly, if $c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^B c_1$, then $c_3 <_{\textit{ob}}^B c_1$.
-
-    If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^C c_1$: Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
-Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
-
-    Therefore, when we meet successive $<_{\textit{ob}}^A$, it is safe to leave only the first and the last elements and ignore intermediate elements.
-Similar cases hold for $<_{\textit{ob}}^B$ and $<_{\textit{ob}}^C$.
-
-\item[-] $<_{\textit{ob}}^A$ and $<_{\textit{ob}}^C$:
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^A c_2 <_{\textit{ob}}^C c_1$.
-Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{ob}}^A c_2$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
-Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^A c_1$.
-
-    \item[-] ($<_{\textit{ob}}^C \cdot <_{\textit{ob}}^A$): If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^A c_1$.
-Since $c_2 <_{\textit{ob}}^A c_1$, $\textit{put}(c_2,\_)$ happens before $\textit{put}(c_1,\_)$.
-Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{rm}(c_3)$ happens before $\textit{put}(c_2,\_)$.
-Therefore, $\textit{rm}(c_3)$ happens before $\textit{put}(c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
-    \end{itemize}
-
-\item[-] $<_{\textit{ob}}^B$ and $<_{\textit{ob}}^C$:
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] ($<_{\textit{ob}}^B \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^C c_1$.
-Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-It is obvious that $\textit{cal}(\textit{rm},c_2)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$.
-Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
-
-    \item[-] ($<_{\textit{ob}}^C \cdot <_{\textit{ob}}^B$): If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^B c_1$.
-Since $c_2 <_{\textit{ob}}^B c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{rm},c_1)$.
-Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
-Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_1)$, and $c_3 <_{\textit{ob}}^B c_1$.
-    \end{itemize}
-
-\item[-]  ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^B \cdot <_{\textit{ob}}^A$): If $c_4 <_{\textit{ob}}^A c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^A c_1$:
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$: Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$.
-Then $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
-This implies that $c_4 <_{\textit{ob}}^A c_3 <_{\textit{ob}}^C c_1$.
-According to the fact for $<_{\textit{ob}}^A \cdot <_{\textit{ob}}^C$, we know that $c_4  <_{\textit{ob}}^A c_1$.
-
-    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is after $\textit{cal}(\textit{put},c_1,\_)$: Since $c_2 <_{\textit{ob}}^A c_1$, $\textit{ret}(\textit{put},c_2,\_)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{rm}(c_3)$ happens before $\textit{rm}(c_2)$, and then we know that $\textit{put}(c_2,\_)$ can not happen before $\textit{put}(c_3,\_)$.
-Since $\textit{put}(c_2,\_)$ does not happen before $\textit{put}(c_3,\_)$, $\textit{cal}(\textit{put},c_3,\_)$ is before $\textit{ret}(\textit{put},c_2,\_)$.
-Since $c_4 <_{\textit{ob}}^A c_3$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_3,\_)$.
-Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_4 <_{\textit{ob}}^A c_1$.
-    \end{itemize}
-
-\end{itemize}
-
-Based on above results, given $a <_{\textit{ob}}^{b_1} a_1 <_{\textit{ob}} \ldots <_{\textit{ob}}^{b_m} a_m <_{\textit{ob}}^{b_{\textit{m+1}}} b$, where each $b_i$ is in $\{ A,B,C \}$, we can merge relations, until we got one of the following facts:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $a <_{\textit{ob}}^A b$, $a <_{\textit{ob}}^B b$ or $a <_{\textit{ob}}^C b$,
-
-\item[-] $a <_{\textit{ob}}^A a_i <_{\textit{ob}}^B b$, or $a <_{\textit{ob}}^B a_i <_{\textit{ob}}^A b$, for some $i$,
-
-\item[-] $a <_{\textit{ob}}^B a_i <_{\textit{ob}}^A a_j <_{\textit{ob}}^B b$, for some $i$ and $j$,
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-Given a differentiated history $h$ and operations $\textit{put}(x,\_)$ and $\textit{rm}(x)$ of $h$, the time interval from the time point when both $\textit{call}(\textit{put},x,\_)$ and $\textit{cal}(\textit{rm},x)$ happen and ends at $\textit{ret}(\textit{rm},x)$ is called the effective interval of $x$ in $h$.
-
-\begin{restatable}{lemma}{Rpr2AsGPandOBOrder}
-\label{lemma:Rpr2 as gap-point and ob order}
-Given a data-differentiated history $h$.
-There exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr2}}$, and $h'$ does not linearizable to $\textit{MR}_{\textit{pr2}}$, if and only if there exists $h'' \in \textit{proj}(h)$, $\textit{last}(h'') = R_{\textit{pr2}}$, $x$ and $y$ has maximal priority in $h''$, $y <_{\textit{ob}} x$ in $h''$, and the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$ in $h''$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{if}$ direction, let $h_1$ be the history that is obtained from $h''$ by erasing all operations of items that has same priority as $x$, except for operations of $x$ and $y$.
-It is obvious that $\textit{last}(h_1) = R_{\textit{pr2}}$.
-According to $y <_{\textit{ob}} x$,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $y <_{\textit{ob}}^A x$, then $\textit{put}(y,\_)$ happens before $\textit{put}(x,\_)$.
-Since in $R_{\textit{pr2}}$, $\textit{put}(\textit{itm},\_)$ is after $\textit{put}$ operations of items with maximal priority (except for $\textit{itm}$), we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
-\item[-] If $y <_{\textit{ob}}^B x$, then $\textit{rm}(y)$ happens before $\textit{rm}(x)$.
-Since in $R_{\textit{pr2}}$, $\textit{rm}(\textit{itm})$ is the last $\textit{rm}$ operation of items with maximal priority, we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
-
-\item[-] If $y <_{\textit{ob}}^C x$, then $\textit{rm}(y)$ happens before $\textit{put}(x,\_)$.
-Since in $R_{\textit{pr2}}$, $\textit{rm}(\textit{itm})$ is the last $\textit{rm}$ operation of items with maximal priority, we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
-\end{itemize}
-
-Similarly as the proof in Lemma \ref{lemma:gap eqals the existence of gap point}, we can see that the only possible position for locating linearizaton point of $\textit{rm}(x)$ is at gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$.
-Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, then there exists unmatched operations before $\textit{rm}(x)$.
-
-Since the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},x)$ in $h''$ (and also in $h_1$), if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, then $\textit{rm}(x)$ will be before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},x)$.
-
-Therefore, for every sequence $l = u \cdot \textit{put}(x,\_) \cdot v \cdot \textit{rm}(x) \cdot w$, if $h_1 \sqsubseteq l$, then either $u \cdot v$ contains some unmatched operations of priority less than priority of $x$, or $w$ contains $\textit{put}(y,\_)$ or $\textit{rm}(y)$.
-In both cases, $l \notin \textit{MR}_{\textit{pr2}}$.
-
-To prove the $\textit{only if}$ direction, we prove its contrapositive.
-Assume we already know that for each $h'' \in \textit{proj}(h)$ with $\textit{last}(h'') = R_{\textit{pr2}}$, for each $x$ and $y$ has maximal priority in $h''$, if $y <_{\textit{ob}} x$ in $h''$, then the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},x)$ in $h''$.
-We need to prove that for each $h' \in \textit{proj}(h)$, if $\textit{last}(h') = R_{\textit{pr2}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr2}}$.
-
-Let $\textit{pri}$ be the maximal priority of $h'$.
-Recall that we already assume that each single-priority history (if we consider $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$) can be ``linearizable to queue'', and item with larger priority is not covered by items with smaller priority.
-Our proof proceed as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Let $h_{\textit{pri}}$ be the projection of $h$ into operations of priority $\textit{pri}$.
-Since each single-priority history can be ``linearizable to queue'', there exists sequence $l_{\textit{pri}}$, such that $h_{\textit{pri}} \sqsubseteq l_{\textit{pri}}$, and when we treat $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$, $l_{\textit{pri}}$ belongs to queue.
-
-\item[-] Step $1$: Let $a_1$ be the last inserted item of $l_{\textit{pri}}$.
-
-$\textit{Cond1}$: If for each item $b$ with priority $\textit{pri}$ in $h'$, the rightmost gap-point $o$ of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$ does not before any $\textit{cal}(\textit{put},b,\textit{pri})$ or $\textit{cal}(\textit{rm},b)$ in $h'$, then we can
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] Locate the linearization point of $\textit{rm}(a_1)$ at $o$,
-    \item[-] If $\textit{put}(a_1,\textit{pri})$ overlaps with $\textit{rm}(a_1)$, then locate the linearization point of $\textit{put}(a_1,\_)$ just before the linearization point of $\textit{rm}(a_1)$.
-Otherwise, $\textit{put}(a_1,\textit{pri}) <_{\textit{hb}} \textit{rm}(a_1)$, and we locate the linearization point of $\textit{put}(a_1,\_)$ just before its return action.
-    \item[-] Locate linearization points of operations of items of priority $\textit{pri}$ (except for $a_1$) just after the call action of the operation.
-    \item[-] For item $z$ with priority smaller than $\textit{pri}$.
-If both $\textit{cal}(\textit{put},z,\_)$ and $\textit{cal}(\textit{rm},z)$ is before $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just after their call actions.
-If both $\textit{ret}(\textit{put},z)$ and $\textit{ret}(\textit{rm},z)$ is after $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just before their return actions.
-
-        Otherwise, either (1) $\textit{rm}(z) <_{\textit{hb}} \textit{put}(z,\_)$, or (2) $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, $\textit{ret}(\textit{put},z,\_)$ before $o$, and $o$ before $\textit{call}(\textit{rm},z)$.
-The first case violates that each single-priority history can be ``linearizable to queue'', and the second case violates that $o$ is a gap-point.
-    \end{itemize}
-
-    In this way we construct a sequence $l$, and it is easy to see that $h' \sqsubseteq l$ and $l$ satisfies the requirement of $R_{\textit{pr2}}$.
-
-\item[-] Otherwise, there exists $a_2$ with priority $\textit{pri}$, such that the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$ in $h'$.
-We can see that the rightmost gap-point of $\textit{put}(a_2,\textit{pri})$ and $\textit{rm}(a_2)$ is after the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$.
-By assumption we know that $a_2$ does not $<_{\textit{ob}}$ to $a_1$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If for each item $b$ with priority $\textit{pri}$ in $h'$, $a_2$ does not $<_{\textit{ob}}$ to $b$.
-Then we go to step $1$ and treat $a_2$ as $a_1$ (When $\textit{Cond1}$ of step $1$ holds, we break the loop).
-    \item[-] Otherwise, there exists $a_3$ with priority $\textit{pri}$ such that $a_2 <_{\textit{ob}} a_3$.
-By assumption,we know that the rightmost gap-point of $\textit{put}(a_3,\textit{pri})$ and $\textit{rm}(a_3)$ is not before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$.
-Then we can see that the rightmost gap-point of $\textit{put}(a_3,\textit{pri})$ and $\textit{rm}(a_3)$ is after the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$.
-Then we go to step $1$ and treat $a_3$ as $a_1$ (When $\textit{Cond1}$ of step $1$ holds, we break the loop).
-    \end{itemize}
-\end{itemize}
-
-It is not hard to see that, if $b_i$ is the $a_1$ in the $\textit{i-th}$ loop of our proof, $b_j$ is the $a_1$ in the $\textit{j-th}$ loop of our proof, and $i<j$, then the rightmost gap-point of $\textit{put}(b_j,\textit{pri})$ and $\textit{rm}(b_j)$ is after the rightmost gap-point of $\textit{put}(b_i,\textit{pri})$ and $\textit{rm}(b_i)$.
-Thus, we finally stop at some $d$.
-$d$ satisfies that, for each item $c$ with priority $\textit{pri}$, $d$ does not $<_{\textit{ob}}$ to $c$.
-$d$ also satisfies that, for each item $c$ with priority $\textit{pri}$ in $h'$, the rightmost gap-point $o'$ of $\textit{put}(d,\textit{pri})$ and $\textit{rm}(d)$ does not before any $\textit{cal}(\textit{put},c,\textit{pri})$ or $\textit{cal}(\textit{rm},c)$.
-Then, we can
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Locate the linearization point of $\textit{rm}(d)$ at the rightmost gap-point $o'$ of $\textit{put}(d,\textit{pri})$ and $\textit{rm}(d)$,
-
-\item[-] Since for each item $c$ with priority $\textit{pri}$ in $h'$, $o'$ is after any $\textit{cal}(\textit{put},c,\textit{pri})$ and $\textit{cal}(\textit{rm},c)$.
-For each $c$, one of the following cases hold: (1) $\textit{cal}(\textit{rm},c)$ is before $o'$ and $o'$ is before $\textit{ret}(\textit{rm},c)$, (2) $\textit{cal}(\textit{rm},c)$ and $\textit{ret}(\textit{rm},c)$ are before $o'$.
-We locate the linearization point of $\textit{rm}(c)$ just after its call action.
-Such linearization point is obviously before $o'$.
-
-\item[-] If $o'$ is before $\textit{ret}(\textit{put},d,\textit{pri})$, then locate the linearization point of $\textit{put}(d,\textit{pri})$ just before $o'$.
-Else, $\textit{ret}(\textit{put},d,\textit{pri})$ is after $o'$, and we locate the linearization point of $\textit{put}(d,\textit{pri})$ just before the return action of it.
-
-\item[-] If $o'$ is before $\textit{ret}(\textit{put},d,\textit{pri})$, we can see that, for each item $c$ with priority $\textit{pri}$ in $h'$, $\textit{cal}(\textit{put},c,\textit{pri})$ is before $o'$.
-Then, and we locate the linearization point of $\textit{put}(c,\textit{pri})$ just before $o'$ and the linearization point of $\textit{put}(d,\textit{pri})$.
-
-    Otherwise, $\textit{ret}(\textit{put},d,\textit{pri})$ is before $o'$.
-Since $d$ does not $<_{\textit{ob}}$ to $c$, $\textit{put}(d,\textit{pri})$ does not happen before $\textit{put}(c,\textit{pri})$, or we can say, $\textit{ret}(\textit{put},d,\textit{pri})$ is after $\textit{cal}(\textit{put},c,\textit{pri})$.
-Then, we locate the linearization point of $\textit{put}(c,\textit{pri})$ just after its call action.
-\end{itemize}
-
-In this way we construct a sequence $l_d$, and it is easy to see that $h' \sqsubseteq l_d$ and $l_d$ satisfies the requirement of $R_{\textit{pr2}}$ with $d$ acts as $\textit{itm}$.
-This completes the proof of $\textit{if}$ direction.
-\qed
-\end {proof}
-
-Let us begin to represent several automata that is used for capture the case that, in a sub-history $h''$ of $h$, $\textit{last}(h'') = R_{\textit{pr2}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr2}}$, and the reason is that there exists items $a$ and $b$ with maximal priority in $h''$, $a <_{\textit{ob}} b$ in $h''$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
-
-According to Lemma \ref{lemma:ob order has bounded length}, if $a <_{\textit{ob}} b$, then one of the following cases hold: (1) $a <_{\textit{ob}}^A b$, (2) $a <_{\textit{ob}}^B b$, (3) $a <_{\textit{ob}}^C b$, (4) $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$ for some $a_1$, (5) $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$ for some $a_1$, or (6) $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$ for some $a_1$ and $a_2$.
-Although there are many possible enumeration of $a$, $b$, $a_1$ and $b_1$, the following lemma states that we need only to consider five enumeration of operations of $a$, $b$, $a_1$ and $b_1$.
-
-\begin{restatable}{lemma}{FiveEnmuerationisEnoughForRpr2}
-\label{lemma:five enumeration is enough for Rpr2}
-Given a data-differentiated history $h$, $\textit{last}(h) = R_{\textit{pr2}}$, and $h$ does not linearizable to $\textit{MR}_{\textit{pr2}}$.
-Assume that there exists items $a$ and $b$ with maximal priority in $h''$, $a <_{\textit{ob}} b$, and the rightmost gap-point $o$ of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
-Let $a_1$ and $a_2$ be the value when possibly $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$, or $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$.
-Then, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists) and $a_2$ (if exists).
-\end{restatable}
-\begin {proof}
-
-Let us prove by consider all the possible reason of $a <_{\textit{ob}}^A b$.
-According to Lemma \ref{lemma:ob order has bounded length}, we need to consider six reasons: %There may exists items with smaller priority of $a$ and covers some time interval, and assume that we rename them into $d$:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Reason $1$, $a <_{\textit{ob}}^A b$:
-
-    Since $a <_{\textit{ob}}^A b$, $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$.
-Since $o$ is after $\textit{cal}(\textit{put},a,\_)$, we can see that $o$ is before $\textit{cal}(\textit{rm},b)$.
-
-    Since single-priority must satisfy the first in firs out property, $\textit{rm}(b)$ does not happen before $\textit{rm}(a)$, and $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
-If $\textit{ret}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$, since we already assume that there exists gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$, this gap-point is also a gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$, and is after $o$, which contradicts that $o$ is the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$.
-Therefore, $\textit{ret}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$.
-
-    According to above discussion, there are two possible enumeration of operations of $a$ and $b$, as shown in \figurename~\ref{fig:history enumeration 1 for Rpr2} and \figurename~\ref{fig:history enumeration 2 for Rpr2}.
-Here we explicitly draw the leftmost gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$ as $o'$.
-Since the position of $\textit{ret}(\textit{put},b)$ does not influence the correctness, we can simply ignore it.
-
-\item[-] Reason $2$, $a <_{\textit{ob}}^B b$:
-
-    Since $a <_{\textit{ob}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$.
-Since $o$ is after $\textit{cal}(\textit{rm},b)$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$.
-Then, we can see that $o$ is before $\textit{cal}(\textit{put},a,\_)$.
-This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\_)$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible.
-Therefore, we can safely ignore this reason.
-
-\item[-] Reason $3$, $a <_{\textit{ob}}^C b$:
-
-    Since $a <_{\textit{ob}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},b)$.
-Since $o$ is after $\textit{cal}(\textit{put},b)$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$.
-Then, we can see that $o$ is before $\textit{cal}(\textit{put},a,\_)$.
-This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\_)$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible.
-Therefore, we can safely ignore this reason.
-
-\item[-] Reason $4$, $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$:
-
-    Since $a_1 <_{\textit{ob}}^B b$, $\textit{rm}(a_1) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_1)$ is before $\textit{cal}(\textit{rm},b)$.
-Since $\textit{rm}(a_1)$ does not happen before $\textit{put}(a_1)$, $\textit{cal}(\textit{put},a_1,\_)$ is before $\textit{ret}(\textit{rm},a_1)$.
-Since $a <_{\textit{ob}}^A a_1$, $\textit{ret}(\textit{put},a,\_)$ is before $\textit{cal}(\textit{put},a_1,\_)$.
-Therefore, $\textit{ret}(\textit{put},a,\_)$ is before $\textit{cal}(\textit{rm},b)$.
-Since $\textit{cal}(\textit{rm},b)$ is before $o$, we can see that $\textit{ret}(\textit{put},a,\_)$ is before $o$.
-Then, we can see that $o$ is before $\textit{cal}(\textit{rm},a)$.
-
-    If $\textit{cal}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$, then $\textit{rm}(b) <_{\textit{hb}} \textit{rm}(a)$.
-Let $h' = h \vert_{ \{ a,a_1,b \} }$.
-It is not hard to see that $h'$ can not linearizable to a sequence that satisfies the first in first our property, and this contradicts to our assumption that each single-priority history satisfies the first in first out property.
-Therefore, $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
-Similarly as the case of reason $1$, we can see that $\textit{ret}(\textit{rm},b)$ is before $\textit{ret}(\textit{rm},a)$.
-
-    According to above discussion, there are three possible enumeration of operations of $a$, $a_1$ and $b$, as shown in \figurename~\ref{fig:history enumeration 3 for Rpr2}, \figurename~\ref{fig:history enumeration 4 for Rpr2} and \figurename~\ref{fig:history enumeration 5 for Rpr2}.
-Here we explicitly draw the leftmost gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$ as $o'$.
-Since the position of $\textit{cal}(\textit{put},b,\_)$, $\textit{ret}(\textit{put},b)$, $\textit{ret}(\textit{put},a_1,\_)$ and $\textit{cal}(\textit{put},a,\_)$ do not influence the correctness, we can simply ignore it.
-
-\item[-] Reason $5$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$:
-
-    Since $a_1 <_{\textit{ob}}^A b$, $\textit{ret}(\textit{put},a_1,\_)$ is before $\textit{call}(\textit{put},b,\_)$.
-Since $\textit{call}(\textit{put},b,\_)$ is before $o$, we can see that $\textit{ret}(\textit{put},a_1,\_)$ is before $o$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $o$ is before $\textit{cal}(\textit{rm},a)$: Then $o$ is obviously before $\textit{ret}(\textit{rm},a)$.
-Since $a <_{\textit{ob}}^B a_1$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},a_1)$.
-Then we can see that, $o$ is before $\textit{cal}(\textit{rm},a_1)$, and remember that $a_1 <_{\textit{ob}}^A b$.
-Then we can goto the case of reason $1$ and treat $a_1$ as $a$.
-Therefore, we can safely ignore this.
-
-    \item[-] If $o$ is before $\textit{cal}(\textit{put},a,\_)$: Since $\textit{rm}(a)$ does not happen before $\textit{put}(a,\_)$, we can see that $\textit{cal}(\textit{put},a,\_)$ is before $\textit{ret}(\textit{rm},a)$, and then $o$ is before $\textit{ret}(\textit{rm},a)$.
-Then similarly as above case, we can see that $o$ is before $\textit{cal}(\textit{rm},a_1)$, and $a_1 <_{\textit{ob}}^A b$.
-Then we can goto the case of reason $1$ and treat $a_1$ as $a$.
-Therefore, we can safely ignore this.
-    \end{itemize}
-
-\item[-] Reason $6$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$:
-
-    Since $a_2 <_{\textit{ob}}^B b$, $\textit{rm}(a_2) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_2)$ is before $\textit{cal}(\textit{rm},b)$.
-%Since $\textit{rm}(a_2)$ does not happen before $\textit{put}(a_2,\_)$, $\textit{cal}(\textit{put},a_2,\_)$ is before $\textit{ret}(\textit{rm},a_2)$.
-%Since $a_1 <_{\textit{ob}}^A a_2$, $\textit{ret}(\textit{put},a_1,\_)$ is before $\textit{cal}(\textit{put},a_2,\_)$.
-    Since $a_1 <_{\textit{ob}}^A a_2$, $\textit{put}(a_1,\_) <_{\textit{hb}} \textit{put}(a_2,\_)$, and $\textit{rm}(a_2)$ does not happen before $\textit{rm}(a_1)$, and $\textit{cal}(\textit{rm},a_1)$ is before $\textit{ret}(\textit{rm},a_2)$.
-Since $a <_{\textit{ob}}^B a_1$, $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(a_1)$, and $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},a_1)$.
-
-    Therefore, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$.
-Since $\textit{cal}(\textit{rm},b)$ is before $o$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$.
-Then it is easy to see that $o$ is before $\textit{cal}(\textit{put},a,\_)$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\_)$, and $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a,\_)$.
-This is impossible, and we can safely ignore this reason.
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-1.pdf}
-%\vspace{-10pt}
-  \caption{The first possible enumeration.}
-  \label{fig:history enumeration 1 for Rpr2}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-2.pdf}
-%\vspace{-10pt}
-  \caption{The second possible enumeration.}
-  \label{fig:history enumeration 2 for Rpr2}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-3.pdf}
-%\vspace{-10pt}
-  \caption{The third possible enumeration.}
-  \label{fig:history enumeration 3 for Rpr2}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-4.pdf}
-%\vspace{-10pt}
-  \caption{The forth possible enumeration.}
-  \label{fig:history enumeration 4 for Rpr2}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-5.pdf}
-%\vspace{-10pt}
-  \caption{The fifth possible enumeration.}
-  \label{fig:history enumeration 5 for Rpr2}
-\end{figure}
-
-Given a data-differentiated history $h$, two actions $\textit{act}_1$, $\textit{act}_2$ of items of priority $\textit{pri}$, and assume that $\textit{act}_1$ is before $\textit{act}_2$ in $h$.
-We say that $\textit{act}_1$, $\textit{act}_2$ is covered by items $d_1,\ldots,d_m$ in $h$, if the priorities of $d_1,\ldots,d_m$ is not $\textit{pri}$, and
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{ret}(\textit{put},d_m,\_)$ is before $\textit{act}_1$,
-
-\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,
-
-\item[-] $\textit{act}_2$ is before $\textit{cal}(\textit{rm},d_1)$.
-\end{itemize}
-
-According to the definition of gap-point, in \figurename~\ref{fig:history enumeration 1 for Rpr2}, \figurename~\ref{fig:history enumeration 2 for Rpr2}, \figurename~\ref{fig:history enumeration 3 for Rpr2}, \figurename~\ref{fig:history enumeration 4 for Rpr2} or \figurename~\ref{fig:history enumeration 5 for Rpr2}, the location of $o$ and $o'$ is equivalent to that, there exists items $d_1,\ldots,d_m$ in $h$ in $h$, such that $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ is covered by $d_1,\ldots,d_m$, $\textit{cal}(\textit{rm},b)$ is before $\textit{ret}(\textit{put},d_m,\_)$ in $h$, and $\textit{cal}(\textit{rm},d_1)$ is before $\textit{ret}(\textit{rm},a)$ in $h$.
-We say that such $d_1,\ldots,d_m$ constitute the rightmost gap of $\textit{put}(b,\_)$ and $\textit{rm}(b)$.
-
-An automaton $\mathcal{A}_{\textit{Rpr2-1}}$ is given in \figurename~\ref{fig:automata for first enumeration of Rpr2}, and it is constructed for the first enumeration in \figurename~\ref{fig:history enumeration 1 for Rpr2}.
-In \figurename~\ref{fig:automata for first enumeration of Rpr2}, we renames the items that covers $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ into $d$, and rename the renaming items into $e$.
-Here let $c = \textit{cal}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{cal}(\textit{rm},e), \textit{ret}(\textit{rm},e)$, $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_1.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr2-1}}$}
-  \label{fig:automata for first enumeration of Rpr2}
-\end{figure}
-
-An automaton $\mathcal{A}_{\textit{Rpr2-2}}$ is given in \figurename~\ref{fig:automata for second enumeration of Rpr2}, and it is constructed for the second enumeration in \figurename~\ref{fig:history enumeration 2 for Rpr2}.
-In \figurename~\ref{fig:automata for second enumeration of Rpr2}, $c_1$, $c_2$, $c_3$ and $c_4$ is same as that in \figurename~\ref{fig:automata for first enumeration of Rpr2}.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_2.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr2-2}}$}
-  \label{fig:automata for second enumeration of Rpr2}
-\end{figure}
-
-For the third enumeration in \figurename~\ref{fig:history enumeration 3 for Rpr2}.
-Since we want to ensure that $a$ and $b$ are putted only once, we need to explicitly record the positions of $\textit{cal}(\textit{put},a,p)$ and $\textit{cal}(\textit{put},b,p)$.
-Since the positions of $\textit{cal}(\textit{put},a,p)$ and $\textit{cal}(\textit{put},b,p)$ are not fixed, there are finite possible cases to consider, as shown below:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{rm},b)$ and before $\textit{cal}(\textit{rm},a)$: There are two possible positions of $\textit{cal}(\textit{put},a,p)$: (1) before $\textit{cal}(\textit{rm},a_1)$, and (2) after $\textit{cal}(\textit{rm},a_1)$, and before $\textit{ret}(\textit{put},a)$.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$ and before $\textit{cal}(\textit{rm},b)$: same as above case.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{put},a_1,\_)$ and before $\textit{ret}(\textit{rm},a_1)$: same as above case.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{put},a)$ and before $\textit{cal}(\textit{put},a_1,p)$: same as above case.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$: There are three possible positions of $\textit{cal}(\textit{put},a,p)$: (1) after $\textit{cal}(\textit{put},b,p)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{cal}(\textit{put}b,p)$, and (3) before $\textit{call}(\textit{rm},a_1)$.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is before $\textit{cal}(\textit{rm},a_1)$: There are three possible positions of $\textit{cal}(\textit{put},a,p)$: (1) after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{cal}(\textit{put},b,p)$ and before $\textit{cal}(\textit{rm},a_1)$, and (3) before $\textit{cal}(\textit{put},b,p)$.
-\end{itemize}
-
-Therefore, there are fourteen possible cases that satisfy the third enumeration in \figurename~\ref{fig:history enumeration 3 for Rpr2}.
-For each case, we construct an finite automaton.
-Let $\textit{Aut}_{\textit{Rpr2-3}}$ be the set of finite automata that is constructed for above fourteen cases.
-For example, for the case $\textit{ca}_1$ when $\textit{cal}(\textit{put},a,p)$ is before $\textit{cal}(\textit{rm},a_1)$, $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$, and $\textit{cal}(\textit{put},b,p)$ is before $\textit{cal}(\textit{rm},b)$, we construct a finite automaton in \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}.
-In \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}, let $c$ and $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$ the same as that in \figurename~\ref{fig:automata for first enumeration of Rpr2}.
-Let $c_2 = c_1 + \textit{ret}(\textit{put},a_1)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$, $c_4 = c_3 + \textit{ret}(\textit{rm},d)$, and $c_5 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{put},a_1) + \textit{ret}(\textit{rm},d)$.
-The other finite automata in $\textit{Aut}_{\textit{Rpr2-3}}$ can be similarly constructed.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_3.pdf}
-%\vspace{-10pt}
-  \caption{Automaton for $\textit{ca}_1$}
-  \label{fig:automata for ca1 of third enumeration of Rpr2}
-\end{figure}
-
-Similarly, we construct sets $\textit{Aut}_{\textit{Rpr2-4}}$ and $\textit{Aut}_{\textit{Rpr2-5}}$ of finite automata for the forth enumeration in \figurename~\ref{fig:history enumeration 4 for Rpr2} and the fifth enumeration in \figurename~\ref{fig:history enumeration 5 for Rpr2}, respectively.
-
-\begin{restatable}{lemma}{Rpr2IsCoRegular}
-\label{lemma:Rpr2 is co-regular}
-$R_{\textit{pr2}}$ is co-regular.
-\end{restatable}
-
-\begin {proof}
-
-Let $\textit{AutS-R}_{\textit{pr2}} = \{ \mathcal{A}_{\textit{Rpr2-1}}, \mathcal{A}_{\textit{Rpr2-2}} \} \cup \textit{Aut}_{\textit{Rpr2-3}} \cup \textit{Aut}_{\textit{Rpr2-4}} \cup \textit{Aut}_{\textit{Rpr2-5}}$.
-We need to prove that:
-
-\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr2}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr2}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr2}}$.
-
-By Lemma \ref{lemma:Rpr2 as gap-point and ob order}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:
-
-\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr0}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, such that $\textit{last}(h') = R_{\textit{pr2}}$, $a$ and $b$ has maximal priority in $h'$, $a <_{\textit{ob}} b$ in $h'$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
-
-\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\textit{AutS-R}_{\textit{pr2}}$.
-By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$.
-Since $h_1$ is accepted by $\textit{AutS-R}_{\textit{pr2}}$, let $x$ and $y$ be the items that are renamed into $b$ and $a$ by $r$, respectively, and let $d_1,\ldots,d_m$ be the items that are renamed into $d$ by $r$.
-
-let $h'' = h_2 \vert_{ \{ x,y,d_1,\ldots,d_m \} }$.
-It is obvious that $\textit{last}(h'') = R_{\textit{pr2}}$.
-According to our construction of automata in $\textit{AutS-R}_{\textit{pr2}}$, it is not hard to see that $x$ and $y$ has maximal priority in $h_2$, $y <_{\textit{ob}} x$ in $h''$, and the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$ in $h''$.
-%By Lemma \ref{lemma:Rpr2 as gap-point and ob order}, we know that $h''$ is not linearizable to $\textit{MR}_{\textit{pr2}}$.
-
-\noindent The $\textit{if}$ direction: Assume that there exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr2}}$, $a'$ and $b'$ has maximal priority in $h'$, $a' <_{\textit{ob}} b'$ in $h'$, and the rightmost gap-point of $\textit{put}(b',\_)$ and $\textit{rm}(b')$ is before $\textit{cal}(\textit{put},a',\_)$ or $\textit{cal}(\textit{rm},a')$ in $h''$.
-By data-independence, we can obtain history $h_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the items $d_1,\ldots,d_m$ that constitute the rightmost gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, we rename them into $d$, and (3) rename the other items into $e$.
-It is easy to see that $\textit{last}(h_1) = R_{\textit{pr2}}$, $a$ and $b$ has maximal priority in $h_1$, $a <_{\textit{ob}} b$ in $h_1$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
-By Lemma \ref{lemma:five enumeration is enough for Rpr2}, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists) and $a_2$ (if exists).
-Then
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $a <_{\textit{ob}} b$ because of the first enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr2-1}}$.
-
-\item[-] If $a <_{\textit{ob}} b$ because of the second enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr2-2}}$.
-
-\item[-] If $a <_{\textit{ob}} b$ because of the third enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-3}}$.
-
-\item[-] If $a <_{\textit{ob}} b$ because of the forth enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-4}}$.
-
-\item[-] If $a <_{\textit{ob}} b$ because of the fifth enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-5}}$.
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\subsection{Co-Regular of $R_{\textit{pr3}}$}
-\label{subsec:co-regular of Rpr3}
-
-\begin{restatable}{lemma}{Rpr3IsAlwaysCoRegular}
-\label{lemma:Rpr3 is always co-regular}
-Given a data-differentiated history $h$, if $\textit{last}(h) = R_{\textit{pr3}}$, then $h \sqsubseteq \textit{MR}_{\textit{pr3}}$.
-\end{restatable}
-
-\begin {proof}
-
-Since $\textit{last}(h) = R_{\textit{pr3}}$, there exists item $x$ with $\textit{pri}_x$ in $h$, such that $h \vert_{ \{ x \}} = \textit{put}(x,\textit{pri}_x)$, and $\textit{pri}_x$ is larger than any priority of $h$.
-
-For each operation, we locate its linearization point just after its call action.
-In such way we obtain a sequence $l$ of operations, and it is easy to see that $h \sqsubseteq l$.
-It is obvious that $l$ contains $\textit{put}(x,\textit{pri})$, and then $l \in \textit{MR}_{\textit{pr3}}$.
-This completes the proof of this lemma.
-\qed
-\end {proof}
-
-\subsection{Co-Regular of $R_{\textit{pr4}}$}
-\label{subsec:co-regular of Rpr4}
-
-\begin{restatable}{lemma}{Rp4AsHappenBefore}
-\label{lemma:Rpr4 as happen before}
-Given a data-differentiated history $h$.
-There exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr4}}$, and $h'$ does not linearizable to $\textit{MR}_{\textit{pr4}}$, if and only if there exists $h'' \in \textit{proj}(h)$, $\textit{last}(h'') = R_{\textit{pr4}}$, $x$ and $y$ has maximal priority in $h''$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,\_) <_{\textit{hb}} \textit{put}(y,\_)$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{if}$ direction, let $h_1 = h \vert_{ \{ x,y \} }$. It is easy to see that $\textit{last}(h') = R_{\textit{pr4}}$ and $h_1$ does not linearizable to $\textit{MR}_{\textit{pr4}}$.
-
-To prove the $\textit{only if}$ direction, we prove its contrapositive.
-Assume that for each $h'' \in \textit{proj}(h)$, such that $\textit{last}(h'') = R_{\textit{pr4}}$, for each $x$ and $y$ has maximal priority in $h''$, if $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, then $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$.
-We need to prove that, for each $h' \in \textit{proj}(h)$, if $\textit{last}(h') = R_{\textit{pr4}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr4}}$.
-
-Given such $h'$, let $x_1,\ldots,x_m$ be the set of items with maximal priority and has unmatched $\textit{put}$ operations, let $y_1,\ldots,y_n$ be the set of items with maximal priority and has matched $\textit{put}$ and $\textit{rm}$ operations.
-It is easy to see that, for each $x_i$ and $y_j$, $\textit{cal}(\textit{put},y_j)$ is before $\textit{ret}(\textit{put},x_i,\_)$.
-Then we locate the linearization points of $h'$ as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] For each $x_i$, locate the linearization point of $\textit{put}(x_i,\_)$ just before its return action.
-
-\item[-] For each $y_j$, locate the lineariztion point of $\textit{put}(y_j,\_)$ jest after its call action.
-
-\item[-] For other operations, locate their linearization points at an arbitrary location after its call action and before its return action.
-\end{itemize}
-
-Let $l$ be the sequence of linearization points.
-It is easy to see that $h \sqsubseteq l$, and $l \in \textit{MR}_{\textit{pr4}}$.
-Here the $\textit{itm}$ in $R_{\textit{pr4}}$ is chosen to be $\textit{ret}(\textit{put},x_{\textit{ind}},\_)$, where the return action of $x_{\textit{ind}}$ is later than return actions of other $x_i$.
-This completes the proof of this lemma.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr4IsAlwaysCoRegular}
-\label{lemma:Rpr4 is always co-regular}
-Given a data-differentiated history $h$, if $\textit{last}(h) = R_{\textit{pr4}}$, then $h \sqsubseteq \textit{MR}_{\textit{pr4}}$.
-\end{restatable}
-
-\begin {proof}
-
-According to Lemma \ref{lemma:Rpr4 as happen before}, if $\textit{last}(h) = R_{\textit{pr4}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr4}}$, then there exists $x$ and $y$ with maximal priority in $h$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,\_) <_{\textit{hb}} \textit{put}(y,\_)$.
-Let $h_1 = h \vert_{ \{ x,y \} }$.
-It is obvious that $\textit{transToQueue}(h_1)$ does not linearizable to queue.
-This contradicts the assumption that every single-priority history can be ``linearizable to queue'', and thus, we can safely ignore this case.
-\qed
-\end {proof}
-
-\subsection{Co-Regular of $R_{\textit{pr5}}$}
-\label{subsec:co-regular of Rpr5}
-
-$R_{\textit{pr5}} \equiv (u \cdot v \in \textit{PQueue}) \wedge (\textit{matched}(u,\textit{put},\textit{rm}) ) \Rightarrow (u \cdot \textit{rmEmpty} \cdot v \in \textit{PQueue})$
-
-Similar as the notion of gap in \cite{Bouajjani:2015}, we propose the notion of gap of $\textit{em}(\textit{empty})$ operations.
-
-\begin{definition}[gap for $\textit{em}(\textit{empty})$ operations]\label{def:gap for rmEmpty operations}
-
-Let $h$ be a data-differentiated history and $o = \textit{rm}(\textit{empty})$ of $h$.
-We say that $h$ has a gap on operation $o$, if there is a partition of the operations of $h$ into $L$ and $R$ satisfying:
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $L$ has matched $\textit{put}$ and $\textit{rm}$ operations.
-
-\item[-] no operation of $R$ happens before an operation of $L$ or $o$.
-
-\item[-] $o$ does not happen before operations of $L$
-\end{itemize}
-\end{definition}
-
-Compared to the gap on $\textit{put}$ and $\textit{rm}$ operations, here we do not consider the priority of operations in $L$ or $R$.
-
-\begin{restatable}{lemma}{GapEqualsLinforRpr5}
-\label{lemma:Gap Equals Lin for Rpr5}
-A data-differentiated history $h$ has a projection $h'$, such that $\textit{last}(h') = R_{\textit{pr5}}$ and $h'$ does not linearizable to $\textit{MR}_{\textit{pr5}}$, if and only if there exists a projection $h''$ of $h$, such that $\textit{last}(h'') = R_{\textit{pr5}}$, there exists $o = \textit{rm}(\textit{empty})$ in $h''$, and there is no gap of $o$ in $h''$.
-\end{restatable}
-
-\begin {proof}
-
-Let us prove the $\textit{only if}$ direction by contradiction.
-Assume that there exists such $h'$, and for each $\textit{rm}(\textit{empty})$ in $h'$, there is a gap of $o$ in $h''$.
-Since $\textit{last}(h') = R_{\textit{pr5}}$, there is exists at least $\textit{rm}(\textit{empty})$ operation in $h$, and let $o$ be arbitrary one of them.
-Since there is a gap of $o$ in $h$, let $L$ and $R$ be the sets of the gap.
-Let $u$ be a sequence of operations of $L$, and $u$ is generated from $h$ by locating linearization points of each operation in $L$ just after its call action.
-Let $v$ be a sequence of operations of $R$, and $v$ is generated from $h$ by locating linearization points of each operation in $R$ just before its return action.
-It is easy to see that $h \sqsubseteq l = u \cdot o \cdot v$, and $l \in \textit{MR}_{\textit{pr5}}$.
-This contradicts that $h'$ does not linearizable to $\textit{MR}_{\textit{pr5}}$.
-
-To prove the $\textit{if}$ direction, let $h_1$ be a history generate from $h''$ by discarding all other $\textit{rm}(\textit{Empty})$ operations.
-By assumption, it is obvious that there is no gap of $o$ in $h_1$.
-Assume by contradiction that there exists $l'$, such that $h_1 \sqsubseteq l' \in \textit{MR}_{\textit{pr5}}$.
-It is easy to see that $l' = u' \cdot o \cdot v'$, such that $u'$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
-Let $L$ be the set of operations of $u'$, and $R$ be the set of operations in $v'$.
-Since $h_1 \sqsubseteq l'$, it is easy to see that $L$ and $R$ is a gap of $o$ in $h_1$, contradicts that there is no gap of $o$ in $h_1$.
-\qed
-\end {proof}
-
-Similar as the notion of gap in \cite{Bouajjani:2015}, we propose the notion left-right constraint of $\textit{rm}(\textit{empty})$ operation.
-
-\begin{definition}[left-right constraint for $\textit{rm}(\textit{empty})$ operation]\label{def:left-right constraint for rmEmpty operation}
-Given a data-differentiated history $h$, and $o = \textit{rm}(\textit{empty})$ of $h$. The left-right constraint of $o$ is the graph $G$ where:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] the nodes are the items of $h$ or $o$, to which we add a node,
-
-\item[-] there is an edge from item $d_1$ to $o$, if $\textit{put}(d_1,\_)$ happens before $o$,
-
-\item[-] there is an edge from $o$ to item $d_1$, if $o$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$,
-
-\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$.
-\end{itemize}
-\end{definition}
-
-Compared to the left-right constraint of $\textit{put}$ and $\textit{rm}$ operations, here we do not consider the priority.
-We also do not consider other $\textit{rm}(\textit{empty})$ operations in $h$.
-
-Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$.
-Let $\textit{LMSet}_1(h,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some item, either $\textit{op}$ happens before $o$ in $h$, or there is an operation $\textit{op}'$ with the same item of $\textit{op}$, such that $\textit{op}'$ happens before $o$ in $h$ $\}$.
-For each $i \geq 1$, let $\textit{LMSet}_{\textit{i+1}}(h,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some item, $\textit{op}$ is not an operation of items in $\textit{LMSet}_k(h,o)$ for each $k \leq i$, either $\textit{op}$ happens before some operation $o' \in \textit{LMSet}_i(h,o)$ in $h$, or there is an operation $\textit{op}''$ with the same item of $o$ and $\textit{op}''$ happens before some operation $o' \in \textit{LMSet}_i(h,o)$ in $h$ $\}$.
-We can see that $\textit{LMSet}_i(h,o) \cap \textit{LMSet}_j(h,o) = \emptyset$ for any $i \neq j$.
-Let $\textit{LMSet}(h,o) = \textit{LMSet}_1(h,o) \cup \textit{LMSet}_2(h,o) \cup \ldots$.
-
-\begin{restatable}{lemma}{RmEmptyDoesNotHappenBeforeLMSetForRpr5}
-\label{lemma:RmEmpty does not happen before LMSet for Rpr5}
-Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$.
-Let $G$ be the graph representing the left-right constraint of $o$.
-Assume that $G$ has no cycle going through $o$.
-Then, $o$ does not happen before any operation in $\textit{LMSet}(h,o)$.
-\end{restatable}
-
-This Lemma can be similarly proved as Lemma \ref{lemma:Rmx does not happen before LMSet for Rpr1}.
-
-\begin{restatable}{lemma}{LMSetHasMatchedPutandRmOperationsForRpr5}
-\label{lemma:LMSet has matched put and rm operations for Rpr5}
-Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$.
-Let $G$ be the graph representing the left-right constraint of $o$.
-Assume that $G$ has no cycle going through $o$.
-Then, $\textit{LMSet}(h,o)$ contains only matched $\textit{put}$ and $\textit{rm}$ operations.
-\end{restatable}
-
-This Lemma can be similarly proved as Lemma \ref{lemma:LMSet has matched put and rm operations}.
-
-\begin{restatable}{lemma}{GapEqualsConstraintforRpr5}
-\label{lemma:Gap Equals Constraint for Rpr5}
-Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$.
-Let $G$ be the graph representing the left-right constraint of $o$.
-There is a gap on $o$, if and only if $G$ has no cycle going through $o$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{only if}$ direction, assume that there is a gap on $o$, and let it be $L$ and $R$.
-Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow o \rightarrow d_1$ in $G$.
-Since $o \rightarrow d_1$, either $o$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$.
-Since $L$ contains matched $\textit{put}$ and $\textit{rm}$ operations, it is easy to see that operations of $d_1$ is in $R$.
-Since $d_1 \rightarrow d_2$, we can see that $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$.
-Since operations of $d_1$ is in $R$ and $L$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we can see that operations of $d_2$ is in $R$.
-Similarly, we can prove that operations of $d_3,\ldots,d_m$ is in $R$.
-Since $d_m \rightarrow o$, $\textit{put}(d_m,\_)$ happens before $o$, and $\textit{put}(d_m,\_) \in L$.
-This contradicts that operations of $d_m$ is in $R$.
-This completes the proof of the $\textit{only if}$ direction.
-
-To prove the $\textit{if}$ direction, assume that $G$ has no cycle going through $o$. Let $O_h$ be the set of operations of $h$. Let $O_L = \textit{LMSet}(h,o)$, $O_R = O_h \setminus O_L$.
-
-By Lemma \ref{lemma:LMSet has matched put and rm operations for Rpr5}, we can see that $O_L = \textit{LMSet}(h,o)$ contains only matched $\textit{put}$ and $\textit{rm}$ operations.
-It remains to prove that $L$ and $R$ is a gap of $o$.
-We prove this by showing that all the following cases are impossible:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Case $1$: Some operation $o_r \in R$ happens before $o$. Then we know that $o_r \in \textit{LMSet}(h,o)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $2$: Some operation $o_r \in R$ happens before some operation $o_l$. Then we know that $o_r \in \textit{LMSet}(h,o)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $4$: $o$ happens before some $o_l \in \textit{LMSet}(h,o) = L$. By Lemma \ref{lemma:RmEmpty does not happen before LMSet for Rpr5}, we know that this is impossible.
-\end{itemize}
-
-This completes the proof of the $\textit{if}$ direction.
-
-\qed
-\end {proof}
-
-Let us begin to represent an automaton that is used for capture the case that, in a sub-history $h'$ of a history $h$, $\textit{last}(h')=R_{\textit{pr5}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr5}}$, and the reason is that there is no gap on a $\textit{rm}(\textit{empty})$ operation in $h'$.
-The automaton is $\mathcal{A}_{\textit{Rpr5}}$, which is given in \figurename~\ref{fig:automata for Rpr5}.
-In \figurename~\ref{fig:automata for Rpr5}, let $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d), \textit{ret}(\textit{rm},d)$, $c_1 = c + \textit{cal}(\textit{put},b,\textit{anPri})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c + \textit{ret}(\textit{rm},b)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr5.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr5}}$}
-  \label{fig:automata for Rpr5}
-\end{figure}
-
-Given a data-differentiated history $h$, we say that $o = \textit{rm}(\textit{empty})$ in $h$ is covered by items $d_1,\ldots,d_m$ in $h$, if
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{put}(d_m,\_)$ happens before $o$,
-
-\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,
-
-\item[-] $o$ happens before $\textit{rm}(d_1)$, or $\textit{rm}(d_1)$ does not exists in $h$
-\end{itemize}
-
-According to the definition of left-right constraint for $o$, in a data-differentiated history $h$, there is a cycle going through $o$, if and only if there exists items $d_1,\ldots,d_m$, such that $o$ is covered by $d_1,\ldots,d_m$.
-
-\begin{restatable}{lemma}{Rpr5IsCoRegular}
-\label{lemma:Rpr1 5s co-regular}
-$R_{\textit{pr5}}$ is co-regular.
-\end{restatable}
-
-\begin {proof}
-
-We need to prove that:
-
-\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\mathcal{A}_{\textit{Rpr5}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr5}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr5}}$.
-
-By Lemma \ref{lemma:Gap Equals Lin for Rpr5} and Lemma \ref{lemma:Gap Equals Constraint for Rpr5}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:
-
-\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\mathcal{A}_{\textit{Rpr5}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h''$ is a projection of $h$, such that $\textit{last}(h'') = R_{\textit{pr5}}$, $o = \textit{rm}(\textit{empty})$ is in $h''$, and $o$ is covered by some items $d_1,\ldots,d_m$ in $h''$.
-
-\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\mathcal{A}_{\textit{Rpr5}}$.
-By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$.
-Let $d_1,\ldots,d_m$ be the items in $h_2$ such that $r(d_i)=d$ for each $1 \leq i \leq m$.
-Let $h'' = h_2 \vert_{ \{ o, d_0, d_1, \ldots, d_m \} }$.
-It is obvious that $\textit{last}(h'') = R_{\textit{pr5}}$.
-It is easy to see that $o$ is covered by $d_1,\ldots,d_m$.
-
-\noindent The $\textit{if}$ direction: Assume that there exists such $h$, $h''$, $o$ and $d_1,\ldots,d_m$.
-Then, let $h_1$ be obtained from $h$ by renaming $d_1,\ldots,d_m$ into $b$ and renaming other items into $d$.
-We can prove that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr5}}$.
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\section{Relate Queue and Stack with Priority Queue}
-\label{sec:relate queue and stack with priority queue}
-
-\subsection{Relate Queue with Priority Queue}
-\label{subsec:relate queue with priority queue}
-
-Given a history $h_q$ of queue operations, let $\textit{UniPriTrans}(h_q)$ be a history of priority queue, which is generated as follows: transform each $\textit{call}(\textit{enq},a)$, $\textit{ret}(\textit{enq},a)$, $\textit{call}(\textit{deq},a)$ and $\textit{ret}(\textit{deq},a)$ of $h_q$ into $\textit{call}(\textit{put},a,1)$, $\textit{ret}(\textit{put},a)$, $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$.
-
-\begin{restatable}{lemma}{RelateQueuewithPQ}
-\label{lemma:relate queue with priority queue}
-Given a history $h_q$ of queue operations, $h_q$ is linearizable to queue, if and only if $\textit{UniPriTrans}(h_q) \sqsubseteq \textit{PQueue}$.
-\end{restatable}
-
-\begin {proof}
-
-This is obvious, since we already ensure that for each single-priority history $h$, $\textit{transToQueue}(h)$ is linearizable to queue. \qed
-\end {proof}
-
-\subsection{Relate Stack with Priority Queue}
-\label{subsec:relate stack with priority queue}
-
-\section{Conclusion and Future Work}
-\label{sec:conclusion and future work}
-
-Conclusion
-
-%\FloatBarrier
-
-%\bibliography{short}
-%\bibliographystyle{plain}
-%\bibliographystyle{splncs}
-%\bibliography{biblio_cat.bib}
-
-\begin{thebibliography}{50}
-
-\bibitem{Bouajjani:2015}
-Bouajjani, A., Emmi, M., Enea, C., Hamza, J.:
-\newblock On reducing linearizability to state reachability.
-\newblock In: Halld{\'{o}}rsson, M.M. et al. (eds.) ICALP 2015, Part II, pp. 95--107. Springer (2015)
-
-\bibitem{Wolper:1986}
-Wolper, P.:
-\newblock Expressing interesting properties of programs in propositional
-  temporal logic.
-\newblock In {\em POPL'86}. pp.184--193. (1986)
-
-\end{thebibliography}
-\newpage
-
-\appendix
-
-\end{document}

### 2017-02-20 preliminaries.tex
preliminary,data-independence,rules

\section{Preliminaries}
\label{sec:preliminaries}

\redt{In this section, we introduce the notion of sequential executions, concurrent executions, histories and linearizability in \cite{Bouajjani:2015,Wolper:1986}. Since $\textit{put}$ method of priority queue has two arguments, we slightly modified related definitions.}

\redt{We fix several (possibly infinite) set $\mathbb{D}_1,\ldots$ of data values}, and a finite set $\mathbb{M}$ of methods. %We consider that methods have exactly one argument, or one return value.
\redt{Return values are transformed into argument values for uniformity \footnote{\redt{Method return values are guessed nondeterministically, and validated at return points. This can be handled using the assume statements of typical formal specification languages, which only admit executions satisfying a given predicate. The argument value for methods without argument or return values, or with fixed argument/return values, is ignored.}}.}
We identify a subset $\mathbb{M}_{\textit{in}} \subseteq \mathbb{M}$ of input methods in order to differentiated methods taking an argument (e.g., the $\textit{put}$ method which inserts a argument value into a priority queue) from the other methods (e.g., the $\textit{rm}$ method which doesn't take an argument, and returns the item with maximal priority of a queue). A method-event is composed of a method $m \in \mathbb{M}$ and \redt{several data value $x_1 \in \mathbb{D}_1,\ldots$, and is denoted $m(x_1,\ldots)$}. We define the concatenation of method-event sequences $u \cdot v$ in the usual way, and $\epsilon$ denotes the empty sequence.

\begin{definition}\label{def:sequential execution}
A sequential execution is a sequence of method events.
\end{definition}

We also fix an arbitrary infinite set $\mathbb{O}$ of operation (identifiers). A call action is composed of a method $m \in \mathbb{M}$, \redt{several data value $x_1 \in \mathbb{D}_1,\ldots$, an operation $o \in \mathbb{O}$, and is denoted $\textit{cal}_o (m,x_1,\ldots)$. Similarly, a return action is denoted $\textit{ret}_o (m,x_1,\ldots)$.} The operation $o$ is used to match return actions to their call actions.

\begin{definition}\label{def:concurrent execution}
A (concurrent) execution $e$ is a sequence of call and return actions which satisfy a well-formedness property: every return has a call action before it in $e$, \redt{using the same tuple $m,o,x_1,\ldots$}, and an operation $o$ can be used only twice in $e$, once in a call action, and once in a return action.
\end{definition}

\begin{example}\label{example:concurrent execution}
\redt{$\textit{cal}_{o_1}(\textit{put},a,7) \cdot \textit{cal}_{o_2}(\textit{put},b,4) \cdot \textit{ret}_{o_1}(\textit{put},a) \cdot \textit{ret}_{o_2}(\textit{put},b)$ is a concurrent execution, while $\textit{cal}_{o_1}(\textit{put},a,7) \cdot \textit{cal}_{o_2}(\textit{put},b,4) \cdot \textit{ret}_{o_1}(\textit{put},a) \cdot \textit{ret}_{o_1}(\textit{put},b)$ and $\textit{cal}_{o_1}(\textit{put},a,7) \cdot \textit{ret}_{o_1}(\textit{put},a) \cdot \textit{ret}_{o_2}(\textit{put},b)$ are not.}
\end{example}

\begin{definition}\label{def:implementation}
An implementation $\mathcal{I}$ is a set of concurrent executions.
\end{definition}

Implementations represent libraries whose methods are called by external programs. In the remainder of this work, we consider only completed executions, where each call action has a corresponding return action. This simplification is sound when implementation methods can always make progress in isolation \cite{Henzinger:2013}: formally, for any execution $e$ with pending operations, there exists an execution $e'$ obtained by extending $e$ only with the return actions of the pending operations of $e$. Intuitively this means that methods can always return without any help from outside threads, avoiding deadlock.

We simplify reasoning on executions by abstracting them into histories.

\begin{definition}\label{def:histories}
\redt{A history is a labeled partial order $(O,<_{\textit{hb}},l)$ with $O \in \mathbb{O}$ and $l:$ maps each $o \in O$ into $\mathbb{M} \times \mathbb{D}_1$, or $\mathbb{M} \times \mathbb{D}_1 \times \mathbb{D}_2$, or $\ldots$.}
\end{definition}

The order $<_{\textit{hb}}$ is called the happens-before relation, and we say that $o_1$ happens before $o_2$ when $o_1 <_{\textit{hb}} o_2$. Since histories arise from executions, their happens-before relations are interval orders \cite{Bouajjani:2015POPL}: for distinct $o_1,o_2,o_3,o_4$, if $o_1 <_{\textit{hb}} o_2$ and $o_3 <_{\textit{hb}} o_4$, then either $o_1 <_{\textit{hb}} o_4$, or $o_3 <_{\textit{hb}} o_2$. Intuitively, this comes from the fact that concurrent threads share a notion of global time.

The history of an execution $e$ is defined as $O,<_{\textit{hb}},l$ where:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $O$ is the set of operations which appear in $e$,

\item[-] $o_1 <_{\textit{hb}} o_2$, if the return action of $o_1$ is before the call action of $o_2$ in $e$,

\item[-] \redt{an operation $o$ occurring in a call action $\textit{call}_o(m,x)$ is labeled by $m(x)$, the case of multiple arguments are similar.}
\end{itemize}

\begin{example}\label{example:concurrent execution}
\redt{The history of the execution $\textit{cal}_{o_1}(\textit{put},a,7) \cdot \textit{cal}_{o_2}(\textit{put},b,4) \cdot \textit{ret}_{o_1}(\textit{put},a) \cdot \textit{ret}_{o_2}(\textit{put},b)$ is $(\{ o_1,o_2 \}, <_{\textit{hb}}, l)$ with $l(o_1) = \textit{put}(a,7)$, $l(o_2) = \textit{put}(b,4)$ and with $<_{\textit{hb}}$ being the empty order relation, since $o_1$ and $o_2$ overlap.}
\end{example}

Let $h = (O,<_{\textit{hb}},l)$ be a history and \redt{$e$ a sequential execution of length $n$}. We say that $h$ is linearizable with respect to $e$, denoted $h \sqsubseteq e$, if there is a bijection $f: O \rightarrow \{ 1,\ldots,n \}$ s.t.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] if $o_1 <_{\textit{hb}} o_2$, then $f(o_1) <_{\textit{hb}} f(o_2)$,

\item[-] the method event at position $f(o)$ in $e$ is $l(o)$.
\end{itemize}

\begin{definition}\label{def:linearizability}
A history $h$ is linearizable with respect to a set $S$ of sequential executions, denoted $h \sqsubseteq S$, if there exists $e \in S$ such that $h \sqsubseteq e$.
\end{definition}

A set of histories $H$ is linearizable with respect to $S$, denoted $H \sqsubseteq S$, if $h \sqsubseteq S$ for all $h \in H$. We extend these definitions to executions according to their histories. In that context, the set $S$ is called a specification.

### 2017-02-20 rules-PQ.tex
preliminary,data-independence,rules

\section{Inductive Rules of Priority Queue}
\label{sec:inductive rules of priority queue}

Similar as \cite{Bouajjani:2015}, we use four inductive rules to define the set of sequential executions of priority queue. Each rule is of the form $l_1 \cdot \ldots \cdot l_k \in \textit{PQueue} \wedge \textit{Guard}(l_1,\ldots,l_k,\textit{itm},\textit{pri}) \Rightarrow \textit{Expr}(l_1,\ldots,l_k,\textit{itm},\textit{pri}) \in \textit{PQueue}$. Here $\textit{PQueue}$ is the set of sequential executions of priority queue. Here $\textit{itm}$ and $\textit{pri}$ are two variables, and represent an arbitrary item and priority, respectively. $\textit{Guard}(l_1,\ldots,l_k,\textit{itm},\textit{pri})$ is a conjunction of conditions with the following notations:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Given sequential execution $l$, $\textit{rm}(\textit{empty}) \notin l$ is satisfied when each method event of $l$ is not $\textit{rm}(\textit{empty})$.

\item[-] Given sequential execution $l$, $\textit{maxPri}(l)$, $\textit{maxUPri}(l)$ and $\textit{maxMPri}(l)$ returns the maximal priority of all $\textit{put}$ in $l$, all $\textit{put}$ without matched $\textit{rm}$ in $l$, and all $\textit{put}$ with matched $\textit{rm}$ in $l$.

\item[-] Given sequential execution $l$, its sub-sequence $l'$ and a priority $p$, $\textit{putInSeq}(l,l',p)$ is satisfied when all the $\textit{put}$ with priority $p$ of $l$ (if exists) is in $l'$.

\item[-] Given sequential execution $l$, $\textit{matched}(l)$ is satisfied, if (1) for each $a \in \mathbb{D}$, if $\textit{put}(a,\_)$ is in $l$, then $\textit{rm}(a)$ is in $l$, and (2) for each $a \in \mathbb{D}$, if $\textit{rm}(a)$ is in $l$, then $\textit{put}(a,\_)$ is in $l$.
\end{itemize}

$\textit{Expr}(l_1,\ldots,l_k,\textit{itm},\textit{pri})$ is a expression $l'_1 \cdot \ldots \cdot l'_m$, where each $l'_i$ is chosen from (1) $l_j$ for some $j$, (2) method event with item $\textit{itm}$ and priority $\textit{pri}$ and (3) method event $\textit{rm}(\textit{empty})$.

Given $l'_1,\ldots,l'_k$ and expression $e=\textit{Expr}(l_1,\ldots,l_k,\textit{itm},\textit{pri})$, we define $\llbracket e \rrbracket$ as the set of sequential executions which can be obtained from $e$ by replacing $l_i$ with $l'_i$ for each $i$, and replacing $\textit{itm}$ and $\textit{pri}$ with some values in $\mathbb{D}$ and $\mathbb{N}$, respectively. Given a rule $R \equiv l_1 \cdot \ldots \cdot l_k \in \textit{PQueue} \wedge \textit{Guard}(l_1,\ldots,l_k,\textit{itm},\textit{pri}) \Rightarrow \textit{Expr}(l_1,\ldots,l_k,\textit{itm},\textit{pri}) \in \textit{PQueue}$ and a sequential execution $w$, if $w=l'_1 \cdot \ldots \cdot l'_k$, and $\textit{Guard}(l'_1,\ldots,l'_k,a,p)$ holds for some $a \in \mathbb{D}$ and $p \in \mathbb{N}$, then we use $w \xrightarrow{R} w'$ to denote that we can obtain $w'$ from $w$ according to rule $R$, where $w' = \textit{Expr}(l'_1,\ldots,l'_k,a,p)$. Let $\llbracket \textit{PQueue} \rrbracket$ be the set of sequential executions $w$ which can be derived from the empty word: $$\epsilon = w_0 \xrightarrow{R_1} w_1 \ldots \xrightarrow{R_k} w$$ where each $R_i$ is one rules of priority queue. When clear from context, we abuse $\llbracket \textit{PQueue} \rrbracket$ by $\textit{PQueue}$.

\begin{definition}\label{def:inductive rules of priority queue}
$\textit{PQueue}$ is defined by the following rules:
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{PQ}_0 \equiv \epsilon \in \textit{PQueue}$.

\item[-] $\textit{PQ}_1 \equiv (u \cdot v \cdot w \in \textit{PQueue}) \wedge (\textit{rm}(\textit{empty}) \notin u \cdot v \cdot w) \wedge (\textit{pri} \geq \textit{maxPri}(u \cdot v \cdot w)) \wedge (\textit{pri} > \textit{maxUPri}(u \cdot v \cdot w)) \wedge (\textit{matched}(u \cdot v) ) \wedge (\textit{putInSeq}(u \cdot v \cdot w,u,\textit{pri})) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w \in \textit{PQueue})$.

\item[-] $\textit{PQ}_2 \equiv (u \cdot v \in \textit{PQueue}) \wedge (\textit{rm}(\textit{empty}) \notin u \cdot v) \wedge (\textit{pri} \geq \textit{maxPri}(u \cdot v)) \wedge (\textit{putInSeq}(u \cdot v,u,\textit{pri})) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{PQueue})$.

\item[-] $\textit{PQ}_3 \equiv (u \cdot v \in \textit{PQueue}) \wedge (\textit{matched}(u) ) \Rightarrow (u \cdot \textit{rm}(\textit{empty}) \cdot v \in \textit{PQueue})$.
\end{itemize}
\end{definition}

\begin{example}\label{example:generate priority queue executions}
One sequential execution of $\textit{PQueue}$ is generated as follows:

$\epsilon$ $\xrightarrow{\textit{PQ}_1}$ $\textit{put}(c,2) \cdot \textit{rm}(c)$

$\xrightarrow{\textit{PQ}_1}$ $\textit{put}(a,4) \cdot \textit{rm}(a) \cdot \textit{put}(c,2) \cdot \textit{rm}(c)$

$\xrightarrow{\textit{PQ}_1}$ $\textit{put}(a,4) \cdot \textit{rm}(a) \cdot \textit{put}(b,4) \cdot \textit{put}(c,2) \cdot \textit{rm}(c) \cdot \textit{rm}(b)$

$\xrightarrow{\textit{PQ}_1}$ $\textit{put}(a,4) \cdot \textit{rm}(a) \cdot \textit{rm}(\textit{empty}) \cdot \textit{put}(b,4) \cdot \textit{put}(c,2) \cdot \textit{rm}(c) \cdot \textit{rm}(b)$
\end{example}

To facilitate our proof of latter sections, for $\textit{PQ}_1$ we need to distinguish two cases: (1) no matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v \cdot w$ has priority $\textit{pri}$, (2) some matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v \cdot w$ has priority $\textit{pri}$.  Therefore, we separate $\textit{PQ}_1$ into two rules:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{PQ}_1^{>} \equiv (u \cdot v \cdot w \in \textit{PQueue}) \wedge (\textit{rm}(\textit{empty}) \notin u \cdot v \cdot w) \wedge (\textit{pri} > \textit{maxPri}(u \cdot v \cdot w)) \wedge (\textit{pri} > \textit{maxUPri}(u \cdot v \cdot w)) \wedge (\textit{matched}(u \cdot v) ) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w \in \textit{PQueue})$.

\item[-] $\textit{PQ}_1^{=} \equiv (u \cdot v \cdot w \in \textit{PQueue}) \wedge (\textit{rm}(\textit{empty}) \notin u \cdot v \cdot w) \wedge (\textit{pri} = \textit{maxPri}(u \cdot v \cdot w)) \wedge (\textit{pri} > \textit{maxUPri}(u \cdot v \cdot w)) \wedge (\textit{matched}(u \cdot v) ) \wedge (\textit{putInSeq}(u \cdot v \cdot w,u,\textit{pri})) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w \in \textit{PQueue})$.
\end{itemize}

For $\textit{PQ}_2$, we also need to distinguish two cases: (1) no matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ has priority $\textit{pri}$, (2) some matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ has priority $\textit{pri}$. Therefore, we separate $\textit{PQ}_2$ into two rules:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{PQ}_2^{>} \equiv (u \cdot v \in \textit{PQueue}) \wedge (\textit{rm}(\textit{empty}) \notin u \cdot v) \wedge (\textit{pri} \geq \textit{maxPri}(u \cdot v)) \wedge (\textit{pri} > \textit{maxMPri}(u \cdot v)) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{PQueue})$.

\item[-] $\textit{PQ}_2^{=} \equiv (u \cdot v \in \textit{PQueue}) \wedge (\textit{rm}(\textit{empty}) \notin u \cdot v) \wedge (\textit{pri} \geq \textit{maxPri}(u \cdot v)) \wedge (\textit{pri} = \textit{maxMPri}(u \cdot v)) \wedge (\textit{putInSeq}(u \cdot v,u,\textit{pri})) \wedge \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{PQueue})$.
\end{itemize}

Given a sequential execution $w \in \textit{PQueue}$ and let $\{p_1,\ldots,p_k\}$ be the set of priorities in $w$, and assume that $\forall i<j$, we have $p_i < p_j$. According to above example, $w$ is generated from $\epsilon$ by work with the following order: (1) Adding pairs of matched $\textit{put}$ and $\textit{rm}$ with priority $p_1$ by using $\textit{PQ}_1^{>}$ and then several times of $\textit{PQ}_1^{=}$, (2) adding unmatched $\textit{put}$ with priority $p_1$ by using $\textit{PQ}_2^{>}$ and then several times of $\textit{PQ}_2^{=}$, (3) Similarly work with $p_2,\ldots,p_k$, and (4) adding $\textit{rm}(\textit{empty})$.

Thus, given $w$, we define $\textit{last}(w)$ as the last possible rule to generate $w$ according to the rules of priority queues:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $l$ contains $\textit{rm}(\textit{empty})$, then $\textit{last}(l) = \textit{PQ}_3$.

\item[-] Else, if the maximal priority of $l$ is unmatched $\textit{put}$ and matched $\textit{put}$ in $l$, then $\textit{last}(l) = \textit{PQ}_2^{=}$.

\item[-] Else, if the maximal priority of $l$ is unmatched $\textit{put}$ in $l$, then $\textit{last}(l) = \textit{PQ}_2^{>}$.

\item[-] Else, if the maximal priority of $l$ is more than one matched $\textit{put}$, then $\textit{last}(l) = \textit{PQ}_1^{=}$.

\item[-] Else, if the maximal priority of $l$ is one matched $\textit{put}$ in $l$, then $\textit{last}(l) = \textit{PQ}_1^{>}$.

\item[-] Else ($l = \epsilon$), $\textit{last}(l) = \textit{PQ}_0$.
\end{itemize}

### 2017-02-20 paper-LIN-PQ.tex
change paper-LIN-PQ to seperate sections into individual files(just a start)DDDDD

@@
 \usepackage{color}
+\usepackage{amsfonts,amssymb}
 %\usepackage{txfonts}
@@
 
-\author {Chao Wang \and Ahmed Bouajjani \and Constantin Enea}
+\author {Ahmed Bouajjani \and Michael Emmi \and Constantin Enea \and Chao Wang}
 \institute{Institut de Recherche en Informatique Fondamentale, \\Univ. Paris Diderot (Paris 7)
@@
 
-\begin{abstract}
-Abstract
-\end{abstract}
+\input{abstract}
 
-\forget{
-\noindent Keywords: weak memory model, $\textit{linearizability}$,
-$\textit{TSO-to-TSO linearizability}$
-}
+\input{introduction}
 
-\section{Introduction}
-\label{sec:introduction}
-Introduction
+\input{preliminaries}
 
-\smallskip
+\input{data-independence-PQ}
 
-\noindent {\bf Related work}
-%\label{sec:related work}
-\noindent Related work
-
-\section{Inductive Rules for Priority Queues}
-\label{sec:inductive rules for priority queues}
-
-given a sequence $l$ of priority queue operations, we say that $l$ is data-differentiated, if for each item $d \in \mathcal{D}$, $\textit{put}(d,\_)$ appears at most once in $l$.
-Note that a data-differentiated sequence $l$ may contains more than one items with a same priority.
-The subset of data-differentiated operation sequences of a set $S$ is denoted by $S_{\neq}$.
-For priority queue, the renaming functions rename only the items and keep the priorities unchanged.
-For priority queue, we modify the notion of data-independence \cite{Wolper:1986} and use data-differentiated instead of differentiated:
-
-%\vspace{-6pt}
-\begin{definition}[data-independence of priority queue]\label{def:priority-value data-independence}
-The set $S$ of sequences of priority queue operations (resp., histories)  is data independent, if:
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] for all $l \in S$, there exists $l' \in S_{\neq}$, and a renaming $r$, such that $l=r(l')$,
-
-\item[-] for all $l \in S$ and for all renaming $r$, $r(l) \in S$.
-\end{itemize}
-\end{definition}
-
-Similar as in \cite{Bouajjani:2015}, we consider only completed executions, where each call action has a corresponding return action.
-This simplification is sound when implementation methods can always make progress in isolation.
-
-For the inductive rules of priority queue, our rules need to concern the cost of operations. Formally, the inductive rules for priority queues is as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $R_{\textit{pr0}} \equiv \epsilon \in \textit{PQueue}$
-
-\item[-] $R_{\textit{pr1}} \equiv (u \cdot v \cdot w \in \textit{PQueue}) \wedge (u \cdot v \cdot w \in (\textit{put},\textit{rm})^*) \wedge (\textit{pri} >$ priorities of $u \cdot v \cdot w ) \wedge (\textit{matched}(u \cdot v, \textit{put}, \textit{rm}) ) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w \in \textit{PQueue})$
-
-\item[-] $R_{\textit{pr2}} \equiv (u \cdot v \cdot w \in \textit{PQueue}) \wedge (u \cdot v \cdot w \in (\textit{put},\textit{rm})^*) \wedge (\textit{pri} =$ the maximal priority of $u \cdot v \cdot w ) \wedge (\textit{pri} >$ priorities of unmatched items in $u \cdot v \cdot w) \wedge (\textit{matched}(u \cdot v, \textit{put}, \textit{rm}) ) \wedge ($all $\textit{put}$ operation with priority $\textit{pri}$ is in $u) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w \in \textit{PQueue})$
-
-\item[-] $R_{\textit{pr3}} \equiv (u \cdot v \in \textit{PQueue}) \wedge ( u \cdot v \in (\textit{put},\textit{rm})^*) \wedge (\textit{pri} =$ the maximal priority of items in $u \cdot v) \wedge (\textit{pri} >$ priorities of matched items in $u \cdot v \cdot w ) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{PQueue})$
-
-\item[-] $R_{\textit{pr4}} \equiv (u \cdot v \in \textit{PQueue}) \wedge ( u \cdot v \in (\textit{put},\textit{rm})^*) \wedge (\textit{pri} =$ the maximal priority of items in $u \cdot v) \wedge (\textit{pri} =$ the maximal priority of matched items in $u \cdot v \cdot w ) \wedge ($all matched $\textit{put}$ with priority $\textit{pri}$ is in $u) \wedge ($all unmatched $\textit{put}$ with priority $\textit{pri}$ is in $u) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{PQueue})$
-
-\item[-] $R_{\textit{pr5}} \equiv (u \cdot v \in \textit{PQueue}) \wedge (\textit{matched}(u,\textit{put},\textit{rm}) ) \Rightarrow (u \cdot \textit{rmEmpty} \cdot v \in \textit{PQueue})$
-\end{itemize}
-
-Given a sequence $l$ of possibly $\textit{put}$, $\textit{rm}$ and $\textit{rmEmpty}$ operations, we say that $l$ contains unmatched $\textit{rm}$ operations, if the number of $\textit{rm}(d)$ operations is larger than the number of $\textit{put}(d,\_)$ operations for some $d$ in $l$.
-By data-independence, we can detect the existence of unmatched $\textit{rm}$ operations by automata.
-A automaton $\mathcal{A}_{\textit{unmRm1}}$ is given in \figurename~\ref{fig:automata for unmatched rm1}, where $c_1=\textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm),\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$, $c_2=c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c_1 + \textit{cal}(\textit{put},b,\textit{anyPri}) + \textit{ret}(\textit{rm},b)$.
-Here $\textit{anyPri}$ is a specific predicates that is satisfied by any priority.
-$\mathcal{A}_{\textit{unmRm1}}$ is used to detect the existence of a item that has been removed but has never been inserted.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_RM1.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{unmRm1}}$}
-  \label{fig:automata for unmatched rm1}
-\end{figure}
-
-An automaton $\mathcal{A}_{\textit{unmRm2}}$ is given in \figurename~\ref{fig:automata for unmatched rm2}, where $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm),$ $\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$, and $c_1 = c_1 + \textit{ret}(\textit{rm},b)$.
-$\mathcal{A}_{\textit{unmRm1}}$ is used to detect the existence of a item that has been removed more than once but has been inserted once.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_RM2.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{unmRm2}}$}
-  \label{fig:automata for unmatched rm2}
-\end{figure}
-
-The following lemma states that we can detect the existence of unmatched $\textit{rm}$ by $\mathcal{A}_{\textit{unmRm1}}$ and $\mathcal{A}_{\textit{unmRm2}}$.
-
-\begin{restatable}{lemma}{DetectUnmatchedRmByAutomata}
-\label{lemma:detect unmatched rm by automata}
-
-Given a data-independent $\mathcal{I}$, there is some unmatched $\textit{rm}$ in some sequence of $\mathcal{I}$, if and only if $( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm1}} \neq \emptyset) \vee ( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm2}} \neq \emptyset )$.
-\end{restatable}
-
-\begin {proof}
-
-For the $\textit{if}$ direction, assume that $( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm1}} \neq \emptyset) \vee ( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm2}} \neq \emptyset )$.
-Then, there exists a sequence $l \in \mathcal{I}$ and $l$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$ or $\mathcal{A}_{\textit{unmRm2}}$.
-If $l$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$, then $l$ contains at least one $\textit{rm}(b)$ operation and no $\textit{put}(b,\_)$ operation; Else, if $l$ is accepted by $\mathcal{A}_{\textit{unmRm2}}$, then $l$ contains at least two $\textit{rm}(b)$ operations and one $\textit{put}(b,\_)$ operation.
-In both cases, $l$ contains unmatched $\textit{rm}$ operations.
-
-To prove the $\textit{only if}$ direction assume that there exists a sequence $l \in \mathcal{I}$ that contains unmatched $\textit{rm}$ operations.
-By data-independence, there exists data-differentiated sequence $l'$ and renaming function $r$, such that $l=r(l')$.
-We need to prove that in $l'$, one of the following cases holds: (1) $l'$ contains $\textit{rm}(d_1)$ for some $d_1$ and does not contain any $\textit{put}(d_1,\_)$, and (2) $l'$ contains one $\textit{put}(d_1,\_)$ operation and more than one $\textit{rm}(d_1)$ operations.
-We prove this by contradiction.
-Assume by contradiction that both cases do not hold.
-Then, for each value $d$ of $l'$, there is either one $\textit{put}(d,\_)$ and one $\textit{rm}(d)$ in $l'$, or one $\textit{put}(d,\_)$ and no $\textit{rm}(d)$ in $l'$.
-It is obvious that such $l'$ does not contain unmatched $\textit{rm}$ operation, and renaming a sequence that does not contain unmatched $\textit{rm}$ we can only obtain another sequence that does not contain unmatched $\textit{rm}$.
-This contradicts that $l=r(l')$ has unmatched $\textit{rm}$.
-
-Let $l''$ be obtained from $ll$ by renaming $d_1$ to $b$ and renaming other value to $a$.
-If the first case holds, $l''$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$; Otherwise, the second case holds, and $l''$ is accepted by $\mathcal{A}_{\textit{unmRm2}}$.
-\qed
-\end {proof}
-
-Therefore, from now on, we can safely assume that there are no unmatched $\textit{rm}$ operations. The notion of $\textit{last}$ is defined as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $l$ contains $\textit{rmEmpty}$ operations, then $\textit{last}(l) = R_{\textit{pr5}}$.
-\item[-] Else, $l$ does not contain any $\textit{rmEmpty}$ operation.
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If the maximal priority of $l$ is an unmatched $\textit{put}$ operation $o$, and the priority of $o$ equals the priority of some matched $\textit{put}$ operations in $l$, then $\textit{last}(l) = R_{\textit{pr4}}$.
-
-    \item[-] If the maximal priority of $l$ is an unmatched $\textit{put}$ operation $o$, and the priority of $o$ is larger than the priority of all matched $\textit{put}$ operations in $l$, then $\textit{last}(l) = R_{\textit{pr3}}$.
-
-    \item[-] If the maximal priority of $l$ is a matched $\textit{put}$ operation $o$, the priority of $o$ equals the priority of some matched $\textit{put}$ operations in $l$, and the priority of $o$ is larger than the priority of all unmatched $\textit{put}$ operations in $l$, then $\textit{last}(l) = R_{\textit{pr2}}$.
-
-    \item[-] If the maximal priority of $l$ is an matched $\textit{put}$ operation $o$, and the priority of $o$ is larger than the priority of all $\textit{put}$ operations in $l$, then $\textit{last}(l) = R_{\textit{pr1}}$.
-
-    \item[-] Otherwise, $l = \epsilon$, and $\textit{last}(l) = R_{\textit{pr0}}$.
-    \end{itemize}
-\end{itemize}
-
-%The following lemma states that the inductive rules for priority queues are sound and complete.
-
-\section{Step-by-Step Linearizability of Priority Queues}
-\label{sec:step-by-step linearizability of priority queues}
-
-The following lemma states that $\textit{PQueue}$ is closed under projection. This lemma holds since the predicates appears in inductive rules of $\textit{PQueue}$ are closed under projection.
-
-\begin{restatable}{lemma}{PQisClosedUnderProjection}
-\label{lemma:PQ is closed under projection}
-
-$\textit{PQueue}$ is closed under projection.
-
-\end{restatable}
-
-%When the context is clear, we shorten priority-value data-independence as data-independence.
-For priority queue, the notion of step-by-step linearizability is modified as follows:
-
-%\vspace{-6pt}
-\begin{definition}[step-by-step linearizability for priority queue]\label{def:step-by-step linearizability of priority queue}
-A data structure $S=R_1, \ldots, R_n$ is said be to step-by-step linearizable, if for any data-differentiated sequence $e$, if $e$ is linearizable w.r.t.
-$\textit{MR}_{\textit{pri}}$ ($1 \leq i \leq 4$) with witness $x$, we have:
-
-$e \setminus x \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket \Rightarrow e \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$
-
-\noindent if $e$ is linearizable w.r.t. $\textit{MR}_{\textit{pr5}}$ with witness $x$, we have:
-
-$e \setminus x \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket \Rightarrow e \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$
-\end{definition}
-
-Let $\textit{proj}(s)$ denote the projection of $s$ obtained by erasing all the operations of several items from $s$.
-Here we assume that each $\textit{rm}(\textit{empty})$ is differentiated from each other, and we can consider each $\textit{rm}(\textit{empty})$ as having a unique item.
-
-\begin{restatable}{lemma}{PQasMultiInMRpriforSequence}
-\label{lemma:PQ as multi in MRpri for sequence}
-$s \in \textit{PQueue}$, if and only if, $\forall s'$, if $s' \in \textit{proj}(s)$, then $s' \in \textit{MR}_{\textit{pri}}$ for some $0 \leq i \leq 5$.
-\end{restatable}
-
-\begin {proof}
-
-The \textit{only if} direction is a direct consequence of Lemma \ref{lemma:PQ is closed under projection}.
-
-To prove the $\textit{if}$ direction.
-The $\textit{if}$ direction holds trivially when $s = \epsilon$.
-When $s \neq \epsilon$.
-since $s \in \textit{proj}(s)$, we generate sequence $s_1$ from $s$ as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $s$ has $\textit{rm}(\textit{empty})$ operations: $s_1$ is generated from $s$ by erasing some $\textit{rm}(\textit{empty})$ operation.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in $S_1$ and unmatched $\textit{put}$ operations of some items in $S_2$: $s_1$ is generated from $s$ by erasing the unmatched $\textit{put}$ operation of items last putted in $S_2$.
-
-\item[-] If the maximal priority in $s$ is unmatched $\textit{put}$ operations of some items in $S_2$ and $\vert S_2 \vert =1$: $s_1$ is generated from $s$ by erasing this unmatched $\textit{put}$ operation.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert \geq 2$: $s_1$ is generated from $s$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item which is last putted in $S$.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert = 1$: $s_1$ is generated from $s$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item in $S$.
-\end{itemize}
-
-Similarly, we obtain $s_2$ from $s_1$, obtain $s_3$ from $s_2$, $\ldots$, until we obtain some $s_m = \epsilon$.
-It is obvious that $s_m \in \textit{PQueue}$, and by the five rules of $\textit{PQueue}$, it is easy to see that $s_{\textit{m-1}},\ldots,s_1,s \in \textit{PQueue}$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{PQasMultiInMRpriforHistory}
-\label{lemma:PQ as multi in MRpri for history}
-Given a data-differentiated history $h$, assume that $\textit{PQueue}$ is step-by-step linearizability.
-$h \sqsubseteq \textit{PQueue}$, if and only if, $\forall h'$, if $h' \in \textit{proj}(h)$, then $h' \sqsubseteq \textit{MR}_{\textit{pri}}$, where $i = \textit{last}(h')$.
-\end{restatable}
-
-\begin {proof}
-
-The \textit{only if} direction is a direct consequence of Lemma \ref{lemma:PQ is closed under projection}.
-
-To prove the $\textit{if}$ direction.
-Since $h \sqsubseteq \textit{PQueue}$, we can see that $h \sqsubseteq l \in \textit{PQueue}$.
-Each projection $h'$ of $h$ can be linearizable to some projection $l'$ of $l$.
-By Lemma \ref{lemma:PQ as multi in MRpri for sequence}, $l' \in \textit{MR}_{\textit{prj}}$ for some $0 \leq j \leq 5$.
-Since $\textit{last}$ of a sequence is uniquely determined, we can see that $l' \in \textit{MR}_{\textit{pri}}$, where $i = \textit{last}(h')$.
-
-The $\textit{if}$ direction holds trivially when $h = \epsilon$. When $h \neq \epsilon$. since $h \in \textit{proj}(h)$, we generate sequence $h_1$ from $h$ as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $h$ has $\textit{rm}(\textit{empty})$ operations: Then $\textit{last}(h) = R_{\textit{pr5}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr5}}$.
-$h_1$ is generated from $h$ by erasing some $\textit{rm}(\textit{empty})$ operation.
-
-\item[-] If the maximal priority in $h$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in $S_1$ and unmatched $\textit{put}$ operations of some items in $S_2$: Then $\textit{last}(h) = R_{\textit{pr4}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr4}}$.
-$h_1$ is generated from $h$ by erasing a unmatched $\textit{put}(x,\_)$, such that $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$ for each $y \in S_2$.
-
-\item[-] If the maximal priority in $s$ is unmatched $\textit{put}$ operations of some items in $S_2$ and $\vert S_2 \vert =1$: Then $\textit{last}(h) = R_{\textit{pr3}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr3}}$.
-$h_1$ is generated from $h$ by erasing this unmatched $\textit{put}$ operation.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert \geq 2$: Then $\textit{last}(h) = R_{\textit{pr2}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr2}}$.
-$h_1$ is generated from $h$ by erasing $\textit{put}(x,\_)$ and $\textit{rm}(x)$, such that for each $y \in S$, $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$, and $\textit{rm}(x)$ does not happen before $\textit{rm}(y)$.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert = 1$: Then $\textit{last}(h) = R_{\textit{pr1}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr1}}$.
-$h_1$ is generated from $h$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item in $S$.
-\end{itemize}
-
-Similarly, we obtain $h_2$ from $h_1$, obtain $h_3$ from $h_2$, $\ldots$, until we obtain some $h_m = \epsilon$.
-Assume that for each $i$, when generating $h_i$ from $h_{\textit{i-1}}$, we erase operations of item $d_i$.
-It is obvious that $h_m = h_{\textit{m-1}} \setminus d_m \sqsubseteq \textit{PQueue}$, and since $h_{\textit{m-1}} \sqsubseteq \textit{MR}_{\textit{prj}_{\textit{m-1}}}$, where $\textit{last}(h_{\textit{m-1}}) = R_{\textit{prj}_{\textit{m-1}}}$, we can see that $h_{\textit{m-1}} \sqsubseteq \textit{PQueue}$.
-Similarly, we can prove that $h_{\textit{m-1}},\ldots,h_1,h \sqsubseteq \textit{PQueue}$.
-This completes the proof of the $\textit{if}$ direction.
-\qed
-\end {proof}
-
-To prove that $\textit{PQueue}$ is step-by-step linearizability, we use five lemmas for five rules of $\textit{PQueue}$.
-
-\begin{restatable}{lemma}{Rpr1isStepByStepLinearizability}
-\label{lemma:Rpr1 is step-by-step linearizability}
-
-$R_{\textit{pr1}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr1}}$ with witness $x$.
-Assume that  Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Then, let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-According to $R_{\textit{pr1}}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where the $\textit{put}$ operations and $\textit{rm}$ operations in $u \cdot v$ are matched, and $\textit{pri}_x$ is larger than any priority in $u \cdot v \cdot w$.
-Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$.
-Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$ operations, we can see that $D_L \cap D_R = \emptyset$.
-Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
-Let $S_1$ be the set of all operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_{L1}$ be the shortest prefix of $l'_L$ that contains all operations of $S_1$, and let $l'_{L2}$ be the sequence such that $l'_L = l'_{L1} \cdot l'_{L2}$.
-It is obvious that if $\textit{put}(x,\textit{pri}_x)$ happens before an operation $o$ in $h$, then $o$ is in $l'_{L2}$.
-Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.
-
-Since priority queue is closed under projection (Lemma \ref{lemma:PQ is closed under projection}) and $u \cdot v$ contain matched $\textit{put}$ and $\textit{rm}$ operations, we know that $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$.
-Since $\textit{pri}_x$ is larger than any priority in $h$, it is easy to see that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$.
-Since the priority queue is empty after executing $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$, we know that $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R \in \textit{PQueue}$.
-
-It remains to prove that $h \sqsubseteq l''$.
-To prove this, we define a graph $G$ whose nodes are the operations of $h$ and there is an edge from operation $o_1$ to $o_2$, if one of the following case holds
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o_1$ happens-before $o_2$ in h,
-
-\item[-] the method event corresponding to $o_1$ in $l''$ is before the one corresponding to $o_2$.
-\end{itemize}
-
-Assume there is a cycle in $G$.
-According the the property of interval order and the fact that the order of $l''$ is total, we know that there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $o_1,o_2 \in l'_{L1}$, or $o_2 \in l'_{L1} \wedge o_1 \in l'_{L2}$, or $o_1,o_2 \in l'_{L2}$, or $o_1,o_2 \in L'_R$, then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] If $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_{L1}$, then since $\textit{put}(x,\textit{pri}_x)$ happens before $o_2$, we know that $o_2 \in l'_{L2}$, which contradicts that $o_2 \in l'_{L1}$.
-
-\item[-] If $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L1}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L1}$, or $o_1 = \textit{rm}(x) \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 \in L'_R \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 = \textit{rm}(x)$, then $l$ contradicts with happen before relation of $h$.
-
-\item[-] If $o_1 \in l'_{L2} \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, then since $o_1$ happens before $\textit{put}(x,\textit{pri}_x)$, we know that $o_1 \in S_1$ and then $o_1 \in l'_{L1}$, which contradicts that $o_1 \in l'_{L2}$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
-
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr2isStepByStepLinearizability}
-\label{lemma:Rpr2 is step-by-step linearizability}
-
-$R_{\textit{pr2}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr2}}$ with witness $x$.
-Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-Since $h' \sqsubseteq l'$, there exists a sequence $h_{l'}$ of call, return and operations, such that (1) the projection of $h_{l'}$ into call and return action is $h$, (2) the projection of $h_{l'}$ into operations is $l'$, and (3) for each matched pair of call action $c$ and return action $r$ of $h$, there exists an operation $o$, such that $a$ before $o$, $o$ before $r$, and $o$ is the operation of $c$ and $r$.
-Intuitively, $h_{l'}$ is obtained from $h$ by adding linearization points.
-
-According to $R_{\textit{pr2}}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where (1) the $\textit{put}$ operations and $\textit{rm}$ operations in $u \cdot v$ are matched, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v \cdot w$, (3) there exists matched $\textit{put}$ and $\textit{rm}$ operations in $u \cdot v \cdot w$ that has priority $\textit{pri}_x$, (4) $\textit{pri}_x$ is larger than priorities of any unmatched put in $u \cdot v \cdot w$, and (5) all $\textit{put}$ operation with priority $\textit{pri}_x$ is in $u$.
-Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$.
-Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$ operations, we can see that $D_L \cap D_R = \emptyset$.
-Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
-
-Let $l'_{\textit{L0}}$ be the shortest prefix of $l'_L$, such that every operation in $l'_{\textit{L0}}$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $h_{l'}$.
-Let $\textit{OP}_x = \{ \textit{put}(d,\textit{pri}_x) \vert \textit{put}(d,\textit{pri}_x)$ is an operation of $h\}$.
-Let $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t$, where $l_t$ is the projection of $l'$ into operations of $\textit{OP}_x$ that are not in $l'_{\textit{L0}}$.
-Let $\textit{OP}'_x$ be the set of operations in $l_t$.
-Let $l'_{\textit{L2}}$ be the projection of $l'_L$ into operations of $l'_L$ that are not in $l'_{\textit{L1}}$.
-Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.
-
-As in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, we know that $l'_L \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$.
-It is obvious that $l'_{\textit{L0}} \in \textit{PQueue}$, and then $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t \in \textit{PQueue}$.
-Since there is no $\textit{rm}(\textit{Empty})$ in $h$ and $\textit{pri}_x$ is the maximal priority of $h$, $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$.
-Since all $\textit{put}$ operations with priority $\textit{pri}$ is in $l'_{L1}$, we know that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$.
-As in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, we know that $l'' \in \textit{PQueue}$.
-
-To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}.
-Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-
-\item[-] $o_1 \in L'_R \wedge (o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2} \vee o_2 = \textit{rm}(x))$, or $o_1 = \textit{rm}(x) \wedge (o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2})$: Then $l$ contradicts with happen before relation of $h$.
-
-\item[-] $o_1, o_2 \in l'_R$ or $o_1, o_2 \in l'_{L2}$: Then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] $o_1, o_2 \in l'_{L1}$: If $(o_1, o_2 \in l'_{L0}) \vee (o_1,o_2 \in \textit{OP}'_x)$, then $l'$ contradicts with happen before relation of $h$.
-
-    Else, it is the case that $o_1 \in \textit{OP}'_x \wedge o_2 \in l'_{L0}$.
-We can see that $o_1$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and the linearization point of $o_1$ is after the return action of $\textit{put}(x,\textit{pri}_x)$.
-Since $o_1 <_{\textit{hb}} o_2$, it is easy to see that $\textit{put}(x,\textit{pri}_x) <_{\textit{hb}} o_2$, and then $o_2 \notin l'_{L0}$, which contradicts that $o_2 \in l'_{L0}$.
-
-\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_{L1}$: Since $o_1 <_{\textit{hb}} o_2$, it is obvious that $o_2 \notin l'_{L1}$, contradicts that $o_2 \in l'_{L1}$.
-
-\item[-] $o_1 \in l'_{L2} \wedge o_2 \in l'_{L1}$: If $o_2 \in l'_{L0}$, then $l'$ contradicts with happen before relation of $h$.
-
-    Else, it is the case that $o_2 \in \textit{OP}'_x$.
-$o_2$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and the linearization point of $o_2$ is after the return action of $\textit{put}(x,\textit{pri}_x)$.
-Since $o_1 <_{\textit{hb}} o_2$, the return action of $o_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$.
-Then it is easy to see that $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.
-
-\item[-] $o_1 \in l'_{L2} \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: Then $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr3isStepByStepLinearizability}
-\label{lemma:Rpr3 is step-by-step linearizability}
-
-$R_{\textit{pr3}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr3}}$ with witness $x$.
-Assume that  Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Then, let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-According to $R_{\textit{pr3}}$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, where $\textit{pri}_x$ is the maximal priority in $u \cdot v$, and $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ operations in $u \cdot v$ (may equal priority of some unmatched $\textit{put}$ operations in $u \cdot v$).
-Let $S_1$ be the set of all operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_1$ be the shortest prefix of $l'$ that contains all operations of $S_1$, and let $l'_2$ be the sequence such that $l' = l'_1 \cdot l'_2$.
-Let sequence $l'' = l'_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_2$.
-
-Since $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ operations in $u \cdot v$, it is easy to see that $l'' \in \textit{PQueue}$.
-It remains to prove that $h \sqsubseteq l''$.
-To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}.
-Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o_1,o_2 \in l'_1$, or $o_1,o_2 \in l'_2$, or $o_1 \in l'_2 \wedge o_2 \in l'_1$: Then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] $o_1 \in l'_2 \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: $o_1 \in S_1$, and then $o_1 \in l'_1$, which contradicts that $o_1 \in l'_2$.
-
-\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_1$: We can see that in $l'$, $o_2$ must after every operation in $S_1$, and then $o_2 \notin l'_1$, contradicts that $o_2 \in l'_1$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
-\end {proof}
-
-The following lemma states that, from linearization of sub-histories, we can merge them and obtain a linearization (regardless of whether it belongs to sequential specification) of the whole history.
-
-\begin{restatable}{lemma}{MergeTwoLinearization}
-\label{lemma:merge two linearization}
-
-Given a history $h$ and operation set $S_1$, $S_2$, let $h_1 = h \vert{S_1}$ and $h_2 = h \vert{S_2}$.
-Assume that $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, and $S_1 \cup S_2$ contains all operations of $h$.
-Then, there exists a sequence $l$, such that $h \sqsubseteq l$, $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$.
-\end{restatable}
-
-\begin {proof}
-
-Given a history $h$ and a operation $o \in S_2$, let $\textit{MB}(o) = \{ o' \vert o' \in S_1$ and $o'$ happens before $o$ in $h \}$, let $\textit{SBI}(o) = \textit{min}\{ i \vert l_1[0,i]$ contains all elements of $\textit{MB}(o) \}$.
-
-Let $l = s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[n] \cdot s_{\textit{n+1}}$ be generated as follows, where $n = \vert l_2 \vert$:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $s_1 = l_1[0, \textit{SBI}(l_2[1])]$,
-
-\item[-] If $s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[i]$ already contains $l_1(\textit{SBI}(l_2[\textit{i+1}]))$, then $s_{\textit{i+1}} = \epsilon$.
-Otherwise, $s_{\textit{i+1}}$ is a subsequence of $l_1$, which starts from the next of last elements of $s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[i]$ in $l_1$ and ends in $l_1(\textit{SBI}(l_2[\textit{i+1}]))$.
-\end{itemize}
-
-It is obvious that $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$, and it remains to prove that $h \sqsubseteq l$.
-We prove this by contradiction.
-Assume that $h$ is not linearizable with respect to $l$.
-Then there must be two operations of $h$, such that $o_1 <_{hb} o_2$ in $h$ but $o_2$ before $0_1$ in $l$.
-Since $l \vert{S_1} = l_1$, $l \vert{S_2} = l_2$, and $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, it is easy to see that it is impossible that $o_1,o_2 \in S_1$ or $o_1,o_2 \in S_2$.
-There are only two possibilities:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o_1 \in S_1 \wedge o_2 \in S_2$.
-Then we can see that $o_2=l_2[i]$ and $o_1 \in s_j$ for some $i < j$.
-Since $o_1 <_{hb} o_2$, we know that $o_1 \in \textit{SBI}(o_2)$.
-By the construction of $l$, we know that $o_1$ must be in $s_k$ for some $k \leq i$, contradicts that $o_1 \in s_j$ with $i < j$.
-
-\item[-] $o_1 \in S_2 \wedge o_2 \in S_1$.
-Then we can see that $o_2 \in s_i$ and $o_1 = l_2[j]$ for some $i \leq j$.
-It is easy to see that this leads to contradiction when $i = j$.
-For the case of $i \neq j$, we need to satisfy the following requirements: (1) $o_1$ ($l_2[j]$) does not happen before $l_2[i]$, (2) $l_2[i]$ does not happen before $o_2$, (3) $o_2$ is either overlap or happens before $o' \in \textit{MB}(l_2[i])$, and (4) $o' <_{hb} l_2[i]$.
-By enumeration we can see that it is impossible that above four conditions be satisfied while $o_1 <_{hb} o_2$.
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr4isStepByStepLinearizability}
-\label{lemma:Rpr4 is step-by-step linearizability}
-
-$R_{\textit{pr4}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr4}}$ with witness $x$.
-Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-According to $R_{\textit{pr4}}$, there exist sequences $u$ and $v$, such that (1) $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v$, (3) $\textit{pri}_x$ equals the maximal priority of matched $\textit{put}$ operations in $u \cdot v$, (4) all matched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$ and (5) all the other unmatched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$.
-
-Since $h' \sqsubseteq l'$, there exists a sequence $h_{l'}$ of call, return and operations, such that (1) the projection of $h$ into call and return action is $h$, (2) the projection of $h_{l'}$ into operations is $l'$, and (3) for each matched pair of call action $c$ and return action $r$ of $h$, there exists an operation $o$, such that $a$ before $o$, $o$ before $r$, and $o$ is the operation of $c$ and $r$.
-Intuitively, $h_{l'}$ is obtained from $h$ by adding linearization points.
-
-According to $R_{\textit{pr4}}$, there exist sequences $u$ and $v$, such that (1) $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v$, (3) $\textit{pri}_x$ equals the maximal priority of matched $\textit{put}$ operations in $u \cdot v$, (4) all matched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$ and (5) all the other unmatched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$.
-
-Let $l'_1$ be the shortest prefix of $l'$, such that every operation in $l'_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $h_{l'}$.
-Let $\textit{OP}_x$ be the set of $\textit{put}$ operations with priority $\textit{pri}_x$ and their matched $\textit{rm}$ operations in $h$, let $\textit{OP}'_x$ be the set of $\textit{put}$ operations in $\textit{OP}_x$ that do not appear in $l'_1$, and let $\textit{OP}''_x$ be the set of $\textit{rm}$ operations in $\textit{OP}_x$ that do not appear in $l'_1$.
-Let $l_x''=l'_1 \cdot l'_2 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_3$, where $l'_2$ is the projection of $l'$ into operations of $\textit{OP}'_x$, and $l'_3$ is the projection of $l'$ into operations of $\textit{OP}''_x$.
-
-Let $h_x$ be the projection of $h$ into $\textit{OP}_x$.
-Let us prove that $h_x \sqsubseteq l''_x$ by contradiction.
-Assume that there exists operations $o_1,o_2$, such that $o_1$ happens before $o_2$ in $h_x$, and $o_2$ is before $o_1$ in $l''_x$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $(o_1 \in l'_3) \wedge (o_2 \in l'_1 \vee o_2 \in l'_2 \vee o'_2 \in l'_3)$, or $(o_1 \in l'_2) \wedge (o_2 \in l'_1 \vee o_2 \in l'_2)$, or $o_1,o_2 \in l'_1$: Then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_1$: Since the linearization point of $o_2$ is before return action of $o_1$, $o_2$ either overlaps with $o_1$, or happens before $o_1$, which contradicts that $o_1$ happens before $o_2$.
-
-\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_2$: Since $l'_2$ only contains $\textit{put}$ operations with priority $\textit{pri}_x$ and are in $\textit{OP}'_x$, it is not hard to see that each operation in $l'_2$ must overlap with $\textit{put}(x,\textit{pri}_x)$, which contradicts that $o_1$ happens before $o_2$.
-
-\item[-] $o_1 \in l'_3 \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: According to the construction of $l''$, we can see that $1_2 \in l'_1$, which contradicts that $1_2 \in l'_3$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h_x \sqsubseteq l''_x$.
-
-Let $\textit{OP}_{\textit{-x}}$ be the set of $\textit{put}$ operations with non-$\textit{pri}_x$ priority, and their matched $\textit{rm}$ operations.
-Let $h_{\textit{-x}}$ be the projection of $h$ into $\textit{OP}_{\textit{-x}}$.
-Let $l''_{\textit{-x}}$ be the projection of $l'$ into$\textit{OP}_{\textit{-x}}$.
-It is obvious that $h_{\textit{-x}} \sqsubseteq l''_{\textit{-x}}$.
-By Lemma \ref{lemma:merge two linearization}, there exists a sequence $l'''$, such that $h \sqsubseteq l'''$, $l''' \vert{\textit{OP}_x} = l_x''$ and $l''' \vert{\textit{OP}_{\textit{-x}}} = l''_{\textit{-x}}$.
-
-It is obvious that $l''_{\textit{-x}} \in \textit{PQueue}$.
-Let $l'_x$ be the projection of $l'$ into $\textit{OP}_x - \{ \textit{put}(x,\textit{pri}_x) \}$.
-It is obvious that $l'_x \in \textit{PQueue}$ and $l'_1 \in \textit{PQueue}$.
-Since (1) $l'_1 \cdot l'_2 \cdot l'_3$ is obtained from $l'_x$ by making $\textit{put}$ operations in $l'_2$ to happen earlier, and (2) $l'_3$ contains only $\textit{rm}$ operations, we can see that $l'_1 \cdot l'_2 \cdot l'_3 \in \textit{PQueue}$.
-Since $l'_3$ contains only $\textit{rm}$ operations, it is easy to see that $l_x'' \in \textit{PQueue}$.
-Since $\textit{pri}_x$ is the maximal priority of $h$ and $h$ does not contain $\textit{rm}(\textit{Empty})$ operations, the operations in $l_x''$ does not influence the correctness of $l''_{\textit{-x}} \in \textit{PQueue}$.
-Therefore, we can see that $l''' \in \textit{PQueue}$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr5isStepByStepLinearizability}
-\label{lemma:Rpr5 is step-by-step linearizability}
-
-$R_{\textit{pr5}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr5}}$ with operation $o=\textit{rmEmpty}$.
-Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Let $h'=h \setminus o$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-According to $R_{\textit{pr5}}$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{rmEmpty} \cdot v$, where the $\textit{put}$ operations and $\textit{rm}$ operations in $u$ are matched.
-
-Let $D_L$ be the set of items in $u$ and $D_R$ be the set of items in $v$.
-Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
-Let sequence $l'' = l'_L \cdot o \cdot L'_R$.
-Since priority queue is closed under projection (Lemma \ref{lemma:PQ is closed under projection}) and $u$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $l'_L \in \textit{PQueue}$ and the the priority queue is empty after executing $l'_L$.
-Then we know that $l'_L \cdot o \in \textit{PQueue}$.
-Since $l'_R$ is obtained from $l'$ by discarding pairs of matched $\textit{put}$ and $\textit{rm}$ operations, it is easy to see that $L'_R \in \textit{PQueue}$, and then we know that $l'' = l'_L \cdot o \cdot L'_R \in \textit{PQueue}$.
-
-It remains to prove that $h \sqsubseteq l''$.
-To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}.
-Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o_1,o_2 \in l'_L$, or $o_1,o_2 \in l'_R$: Then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] If $o_1=o \wedge o_2 \in l'_L$, or $o_1 \in l'_R \wedge o_2 \in l'_L$, or $o_1 \in l'_R \wedge o_2 = 0$, then $l$ contradicts with happen before relation of $h$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$. \qed
-\end {proof}
-
-The following lemma states that $\textit{PQueue}$ is step-by-step linearizability, it is a direct consequence of Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, Lemma \ref{lemma:Rpr2 is step-by-step linearizability}, Lemma \ref{lemma:Rpr3 is step-by-step linearizability}, Lemma \ref{lemma:Rpr4 is step-by-step linearizability} and Lemma \ref{lemma:Rpr5 is step-by-step linearizability}.
-
-\begin{restatable}{lemma}{PQueueisStepByStepLinearizability}
-\label{lemma:PQueue is step-by-step linearizability}
-
-$\textit{PQueue}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\section{Co-Regular of Priority Queues}
-\label{sec:co-regular of priority queues}
-
-\subsection{Co-Regular of $R_{\textit{pr1}}$}
-\label{subsec:co-regular of Rpr1}
-
-Let us introduce the notion of gap for matched $\textit{put}$ and $\textit{rm}$ operations.
-
-\begin{definition}[gap for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:gap for matched put and rm operations}
-
-Let $h$ be a data-differentiated history and $o_1 = \textit{put}(x,\textit{pri}_x)$, $o_2 = \textit{rm}(x)$ be a pair of matched $\textit{put}$ and $\textit{rm}$ operation of $h$.
-We say that $h$ has a gap on operations $o_1$ and $o_2$, if there is a partition of the operations of $h$ into $L$, $M$ and $R$ satisfying:
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $L \cup M$ has matched $\textit{put}$ and $\textit{rm}$ operations.
-
-\item[-] no operation of $R$ happens before an operation of $L$, $o_1$, $M$ or $o_2$.
-
-\item[-] no operation of $o_2$ happens before an operation of $L$, $o_1$ or $M$.
-
-\item[-] no operation of $M$ happens before an operation of $L$ or $o_1$.
-
-\item[-] no operation of $o_1$ happens before an operation of $L$.
-
-\item[-] no operation in $L \cup M \cup R$ has priority $\textit{pri}_x$.
-\end{itemize}
-\end{definition}
-
-Given a history $h$ and set $O$ of some operations in $h$.
-It is easy to see that there exists a sequence $l$ of operations of $O$, such that $l$ does not contradicts the happen before relation of $h$.
-Therefore, from now on, we can use this fact to directly generate such sequence from such $O$.
-
-The following lemma reduce linearizable with respect to $\textit{MR}_{\textit{pr1}}$ into gap of matched $\textit{put}$ and $\textit{rm}$ operations.
-
-\begin{restatable}{lemma}{GapEqualsLinforRpr1}
-\label{lemma:Gap Equals Lin for Rpr1}
-A data-differentiated history $h$ has a projection $h'$, such that $\textit{last}(h') = R_{\textit{pr1}}$ and $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$, if and only if there exists a projection $h''$ of $h$, such that $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is in $h''$, $\textit{pri}_x$ is the maximal priority of $h''$, and one the the following case holds:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] case $1$: $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$,
-\item[-] case $2$: there is no gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-\end{itemize}
-\end{restatable}
-
-\begin {proof}
-
-Let us prove the $\textit{only if}$ direction by contradiction.
-Assume that there exists such $h'$, while case $1$ and case $2$ do not hold for any projection of $h$.
-Since case $1$ does not hold for any projection of $h$, we know that $\forall y \in D$, $\textit{rm}(y)$ does not happens before $\textit{put}(y,\_)$ in $h$.
-Since case $2$ does not hold for any projection of $h$ and $\textit{last}(h') = R_{\textit{pr1}}$, we can see that $h'$ has gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$, while $\textit{pri}_x$ is the maximal priority in $h'$.
-Since $\textit{last}(h') = R_{\textit{pr1}}$, we can see that $\textit{pri}_x$ is larger than priority of all other $\textit{put}$ operations of $h$.
-
-By the properties of the gap, there is partitions $L$, $M$ and $R$.
-Let $u$, $v$ and $w$ contain the operations of $L$, $M$ and $R$, respectively.
-It is easy to see that $h'$ is linearizable with respect to $l''=u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$ and $l'' \in \textit{MR}_{\textit{pr1}}$.
-This contradicts the fact that $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$.
-
-Let us begin to prove the $\textit{if}$ direction.
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If case $1$ holds for $h''$, then $h \vert_{ \{ x \} }$ is the $h'$ as required.
-
-\item[-] If case $2$ holds, assume by contradiction that, for every projection $h'$ of $h$, if $\textit{last}(h') = R_{\textit{pr1}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr1}}$.
-Since $h''$ is also a projection of $h$, we know that there exist sequences $u$, $v$ and $w$, such that $h'' \sqsubseteq u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, $u \cdot v$ contains matched $\textit{put}$ and $\textit{rm}$ operations, and $\textit{pri}_x$ is larger than priority of $u \cdot v \cdot w$.
-Let $L$, $M$ and $R$ be the operations of $u$, $v$ and $w$, respectively.
-Then, the partitions $L$, $M$ and $R$ form a gap on operation $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-This contradicts the fact that there is no gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\begin{definition}[left-right constraint for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:left-right constraint for matched put and rm operations}
-Given a data-differentiated history $h$, and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-The left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is the graph $G$ where:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] the nodes are the items of $h$, to which we add a node,
-
-\item[-] there is an edge from item $d_1$ to $x$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$ and $\textit{pri}_1 \neq \textit{pri}_x$,
-
-\item[-] there is an edge from $x$ to item $d_1$, if $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$, and the priority of $d_1$ is not $\textit{pri}_x$,
-
-\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{rm}(d_2,\textit{pri}_2)$, $\textit{pri}_1 \neq \textit{pri}_x$, and $\textit{pri}_2 \neq \textit{pri}_x$.
-\end{itemize}
-\end{definition}
-
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-Let $\textit{LMSet}_1(h,x) = \{ o \vert$ $o$ is not an operation of $x$, the priority of $o$ does not equal the priority of $x$, and either it happens before $\textit{put}(x)$ or $\textit{rm}(x)$ in $h$, or there is an operation $o'$ with the same item of $o$, such that $o'$ happens before $\textit{put}(x)$ or $\textit{rm}(x)$ in $h$ $\}$.
-For each $i \geq 1$, let $\textit{LMSet}_{\textit{i+1}}(h,x) = \{ o \vert$ $o$ is not an operation of $x$ or items in $\textit{LMSet}_k(h,x)$ for each $k \leq i$, the priority of $o$ does not equal the priority of $x$, and either it happens before some operation $o' \in \textit{LMSet}_i(h,x)$ in $h$, or there is an operation $o''$ with the same item of $o$ and $o''$ happens before some operation $o' \in \textit{LMSet}_i(h,x)$ in $h$ $\}$.
-%We can see that for each $i$, $\textit{LMSet}_i(h,x)$ contains matched $\textit{put}$ and $\textit{rm}$ operations, and
-We can see that $\textit{LMSet}_i(h,x) \cap \textit{LMSet}_j(h,x) = \emptyset$ for any $i \neq j$.
-Let $\textit{LMSet}(h,x) = \textit{LMSet}_1(h,x) \cup \textit{LMSet}_2(h,x) \cup \ldots$.
-%{\color {red} We say that a history $h$ of priority queue is matched before $x$, if (1) $h$ contains no $\textit{rmEmpty}$ operations, (2) $ h \sqsubseteq l_1 \cdot \textit{put}(x,\_) \cdot l_2 \cdot \textit{rm}(x) \cdot l_3$ for some $l_1,l_2,l_3$, and $l_1 \cdot l_2$ contains matched $\textit{put}$ and $\textit{rm}$ operations.}
-
-The following is an example that shows that, given a data-differentiated history $h$, the set $\textit{LMSet}_i(h,x)$ can be non-empty for each $1 \leq i \leq 3$.
-In \figurename~\ref{fig:his nobound of LMSet} , $\textit{LMSet}_1(h,x) = \{ \textit{put}(a,1),\textit{rm}(a) \}$, $\textit{LMSet}_2(h,x) = \{ \textit{put}(b,1),\textit{rm}(b) \}$ and $\textit{LMSet}_3(h,x) = \{ \textit{put}(a,c),\textit{rm}(c) \}$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{PIC_HIS_NOBOUNDOF_LMSET.pdf}
-%\vspace{-10pt}
-  \caption{An example for non-emptiness of $\textit{LMSet}_1(h,x)$, $\textit{LMSet}_2(h,x)$ and $\textit{LMSet}_3(h,x)$}
-  \label{fig:his nobound of LMSet}
-\end{figure}
-
-\begin{restatable}{lemma}{RmxDoesNotHappenBeforeLMSetForRpr1}
-\label{lemma:Rmx does not happen before LMSet for Rpr1}
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-%, such that $\textit{pri}_x$ is the maximal priority of $h$, and $\textit{last}(h) = R_{\textit{pr1}}$.
-Assume that $\textit{rm}(y)$ does not happen before $\textit{put}(y,\_)$ for each item $y$ in $h$.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-Assume that $G$ has no cycle going through $x$.
-Then, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}(h,x)$.
-\end{restatable}
-
-\begin {proof}
-
-We prove this lemma by induction, and prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x)$, in $\textit{LMSet}_2(h,x)$, $\ldots$.
-Note that, according to the definition of $\textit{LMSet}(h,x)$, $\textit{pri}_x$ is different to priority of all items in $\textit{LMSet}(h,x)$.
-
-\noindent (1) Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x)$ by contradiction.
-Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{LMSet}_1(h,x)$ is an operation of item $d$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d$ does not equals $\textit{pri}_x$).
-
-We use a triple $(t_1,t_2,t_3)$ to represent related information.
-$t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$.
-$t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation.
-$t_2$ and $t_3$ is used for the reason of $o \in \textit{LMSet}_1(h,x)$: $o \in \textit{LMSet}_1(h,x)$, since an operation (of kind $t_2$) of $d$ happens before an operation (of kind $t_3$) of $x$.
-Let us consider all the possible cases of $(t_1,t_2,t_3)$:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.
-
-\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{rm}(x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{rm}(x)$.
-
-\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x) )$.
-By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(x,\textit{pri}_x) ) \vee ( \textit{rm}(d) <_{hb} \textit{put}(d,\_) )$, which is impossible.
-
-\item[-] $(\textit{put},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{rm}(x) )$.
-We can see that $\textit{rm}(d) <_{hb} \textit{rm}(x) <_{hb} \textit{put}(d,\_)$, which contradicts that $\textit{rm}(d)$ does not happen before $\textit{put}(d,\_)$.
-
-\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x) )$.
-We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.
-
-\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{rm}(x) )$.
-We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.
-
-\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.
-
-\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{rm}(x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{rm}(x)$.
-\end{itemize}
-
-This completes the proof for $\textit{LMSet}_1(h,x)$.
-
-\noindent (2) Assume we already prove that for some $j \geq 1$, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_{\textit{j+1}}(h,x)$ by contradiction.
-Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$ is an operation of item $d_{\textit{j+1}}$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d_{\textit{j+1}}$ does not equals $\textit{pri}_x$).
-We use a triple $(t_1,t_2,t_3)$ to represent related information.
-$t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$.
-$t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation.
-$t_2$ and $t_3$ is used for the reason of $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$: $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$, since an operation (of kind $t_2$) of $d_{\textit{j+1}}$ happens before an operation (of kind $t_3$) of $d_j$, where $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{LMSet}_j(h,x)$.
-Let us consider all the possible cases of $(t_1,t_2,t_3)$:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_)$.
-We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-
-\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j,\_)$.
-We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j,\_) ) \wedge ( \textit{rm}(d_j) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-
-\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$.
-By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \vee ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_{\textit{j+1}},\_) )$, which is impossible.
-
-\item[-] $(\textit{put},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$.
-By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \vee ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_{\textit{j+1}},\_) )$, which is impossible.
-
-\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) )$.
-Let us consider the reason of $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{LMSet}_j(h,x)$:
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $( j > 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(_{\textit{j-1}},\_), \textit{rm}(_{\textit{j-1}}) \in \textit{LMSet}_{\textit{j-1}}(h,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, we can see that $\textit{put}(d_{\textit{j+1}},\_) <_{hb} o''$, and then operations of $d_{\textit{j+1}}$ is in $\textit{LMSet}_j(h,x)$, contradicts that operations of $d_{\textit{j+1}}$ is in $\textit{LMSet}_{\textit{j+1}}(h,x)$.
-
-    \item[-] If $( j = 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of $x$: Similar to above case.
-
-    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{LMSet}_{\textit{j-1}}(h,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, we can see that $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} o'' ) \vee ( \textit{rm}(d_j) <_{hb} \textit{put}(d_j,\_) )$, which is impossible.
-
-    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of $x$: Similar to above case.
-    \end{itemize}
-
-\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Let $T_{\textit{ind}}$ be the set of sentences $\{ \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}), \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j),\ldots, \textit{put}(d_{\textit{ind+1}},\_) <_{hb} \textit{rm}(d_{\textit{ind}}) \}$.
-Here each $d_i$ is a item of some operation in $\textit{LMSet}_i(h,x)$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d_i$ does not equals $\textit{pri}_x$ for each $i$).
-Let us prove that from $T_j$ we can obtain contradiction by induction:
-
-    {\bf Base case $1$}: From $T_1$ we can obtain contradiction.
-
-    Let us prove base case $1$:
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{put}(d_1,\_)$ happens $o$, and $o$ is an operation of $x$.
-Then there is a cycle $x \rightarrow d_{\textit{j+1}} \rightarrow \ldots \rightarrow d_1 \rightarrow x$ in $G$, contradicts that $G$ has no cycle going through $x$.
-
-    \item[-] If $\textit{rm}(d_1)$ happens before $o$, and $o$ is an operation of $x$.
-Then since $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1) <_{hb} o$, we can see that $\textit{put}(d_2,\_) <_{hb} o$, and then $\textit{put}(d_2,\_) \in \textit{LMSet}_1(h,x)$, contradicts that $\textit{put}(d_2,\_) \in \textit{LMSet}_2(h,x)$.
-    \end{itemize}
-
-    {\bf Base case $2$}: From $T_2$ we can obtain contradiction.
-
-    Let us prove base case $2$: If $\textit{rm}(d_2) <_{hb} o$, and $o$ is an operation of $d_1$, then since $( \textit{put}(d_3,\_) <_{hb} \textit{rm}(d_2) ) \wedge ( \textit{rm}(d_2) <_{hb} o )$, we know that $\textit{put}(d_3,\_) <_{hb} o$.
-This implies that $\textit{put}(d_3,\_) \in \textit{LMSet}_2(h,x)$, contradicts that $\textit{rm}(d_3,\_) \in \textit{LMSet}_3(h,x)$.
-Therefore, it is only possible that $\textit{put}(d_2,\_)$ happens before an operation of $d_1$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{put}(d_1,\_)$ and $\textit{put}(d_1,\_)$ happens before operations of $x$, then we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, which is impossible.
-
-    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{put}(d_1,\_)$ and $\textit{rm}(d_1)$ happens before operations of $x$, then by interval order, we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, or $\textit{rm}(d_1) <_{hb} \textit{put}(d_1,\_)$, which is impossible.
-
-    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{put}(d_1,\_)$ happens before operations of $x$, then $x \rightarrow d_{\textit{j+1}} \rightarrow \ldots \rightarrow d_1 \rightarrow x$ in $G$, contradicts that $G$ has no cycle going through $x$.
-
-    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1)$ happens before operations of $x$, then we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, which is impossible.
-    \end{itemize}
-
-    {\bf induction step}: Given $\textit{ind} \geq 3$, if from $T_{\textit{ind-1}}$ we can obtain contradiction, then from $T_{\textit{ind}}$ we can also contain contradiction.
-
-    Prove of the induction step: Similarly as base case $2$, we can prove that it is only possible that $\textit{put}(d_{\textit{ind}},\_)$ happens before operations of $d_{\textit{ind-1}}$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$ and $\textit{put}(d_{\textit{ind-1}},\_)$ happens before operations of $d_{\textit{ind-2}}$, then we know that $\textit{put}(d_{\textit{ind}})$ happens before operation of $d_{\textit{ind-2}}$, which is impossible.
-
-    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$ and $\textit{rm}(d_{\textit{ind-1}})$ happens before operations of $d_{\textit{ind-2}}$, then by interval order, we know that $\textit{put}(d_{\textit{ind}},\_)$ happens before operation of $d_{\textit{ind-2}}$, or $\textit{rm}(d_{\textit{ind-1}}) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$, which is impossible.
-
-    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{rm}(d_{\textit{ind-1}})$, then we obtain $T_{\textit{ind-1}}$, which already contain contradiction.
-    \end{itemize}
-
-    By base case $1$, base case $2$ and the induction step, it is easy to see that $T_j$ contains contradiction.
-
-\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$.
-We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-
-\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$.
-We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \wedge ( \textit{rm}(d_j) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-\end{itemize}
-
-This completes the proof for $\textit{LMSet}_{\textit{j+1}}(h,x)$.
-Therefore, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}(h,x) = \textit{LMSet}_1(h,x) \cup \textit{LMSet}_2(h,x) \cup \ldots$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{LMSetHasMatchedPutandRmOperations}
-\label{lemma:LMSet has matched put and rm operations}
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-Assume that $\textit{rm}(y)$ does not happen before $\textit{put}(y,\_)$ for each item $y$ in $h$.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-Assume that $G$ has no cycle going through $x$.
-Then, $\textit{LMSet}(h,x)$ contains only matched $\textit{put}$ and $\textit{rm}$ operations.
-\end{restatable}
-\begin {proof}
-
-We prove this lemma by contradiction.
-Assume that there exists a value, such that $\textit{LMSet}(h,x)$ contains only its $\textit{put}$ operation and does not contain its $\textit{rm}$ operation.
-Then we can see that there exists $d_1,\ldots,d_j$.
-Intuitively, $d_1,\ldots,d_j$ are items in $\textit{LMSet}_1(h,x), \ldots, \textit{LMSet}_i(h,x)$, respectively.
-$\textit{LMSet}(h,x)$ contains $\textit{put}(d_j,\_)$ and does not contain $\textit{rm}(d_j)$.
-And each $d_i$ is the reason of $d_{\textit{i+1}} \in \textit{LMSet}_{\textit{i+1}}(h,x)$.
-Formally, we require that
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] For each $1 \leq i \leq j$, operations of $d_i$ belongs to $\textit{LMSet}_i(h,x)$. %For each $1 \leq i \leq j$, the priority of $d_i$ is not $\textit{pri}_x$.
-
-\item[-] For each $i \neq j$, $\textit{put}(d_i,\_),\textit{rm}(d_i) \in \textit{LMSet}_i(h,x)$. $\textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x)$, and $h$ does not contain $\textit{rm}(d_j)$.
-
-\item[-] An operation of $d_1$ happens before an operation of $x$. For each $1 < i \leq j$, an operation of $d_j$ happens an operation of $d_{\textit{i-1}}$.
-
-\item[-] For each $k$ and $\textit{ind}$, if $k > \textit{ind+1}$, then no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.
-\end{itemize}
-
-According to the definition of $\textit{LMSet}(h,x)$, it is easy to see that such $d_1,\ldots,d_j$ exists. Let us prove the following fact:
-
-\noindent {\bf $\textit{fact}_1$}: Given $1 \leq i < j$, it can not be the case that $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap.
-
-Proof of $\textit{fact}_1$: We prove $\textit{fact}_1$ by contradiction.
-Assume that for some $i \neq j$, $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap.
-Since $\textit{put}(d_i,\_), \textit{rm}(d_i) \in \textit{LMSet}_i(h,x)$, we know that an operation $o_i$ of $d_i$ happens before operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$.
-Moreover, since $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap, it is not hard to see that the call action of $\textit{put}(d_i,\_)$ and the call action of $\textit{rm}(d_i)$ is before the call action of $o_{\textit{i-1}}$.
-Since $\textit{put}(d_{\textit{i+1}},\_), \textit{rm}(d_{\textit{i+1}}) \in \textit{LMSet}_{\textit{i+1}}(h,x)$, we know that an operation $o'_{\textit{i+1}}$ of $d_{\textit{i+1}}$ happens before operation $o'_i$ of $d_i$.
-Then, it is not hard to see that $o'_{\textit{i+1}}$ also happens before $o_{\textit{i-1}}$, which contradicts that For each $k > \textit{ind+1}$, no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.
-
-We already know that an operation of $d_1$ happens before an operation of $x$.
-By $\textit{fact}_1$, we can ensure that $\textit{put}(d_1,\_)$ happens before an operation of $x$, and then $d_1 \rightarrow x$ in $G$.
-For each $1 < i < j$, we know that an operation $o_i$ of $d_i$ happens before an operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$.
-By $\textit{fact}_1$, we can ensure that $o_i=\textit{put}(d_i,\_)$ and $o_{\textit{i-1}}=\textit{rm}(d_{\textit{i-1}})$, and then $d_i \rightarrow d_{\textit{i-1}}$ in $G$.
-Since $h$ contains $\textit{put}(d_j)$ and does not contain $\textit{rm}(d_j)$, we know that $x \rightarrow d_j$ in $G$.
-Then $G$ has a cycle going through $x$, contradicts that $G$ has no cycle going through $x$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{GapEqualsConstraintforRpr1}
-\label{lemma:Gap Equals Constraint for Rpr1}
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-%, where $\textit{pri}_x$ has the maximal priority of $h$, and $\textit{last}(h) = R_{\textit{pr1}}$.
-Assume that for each item in $h$, its $\textit{rm}$ operation does not happen before its $\textit{put}$ operation.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x)$ and $\textit{rm}(x)$.
-There is a gap on $\textit{put}(x)$ and $\textit{rm}(x)$, if and only if $G$ has no cycle going through $x$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{only if}$ direction, assume that there is a gap on $\textit{put}(x)$ and $\textit{rm}(x)$, and let it be $L$, $M$ and $R$.
-Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ in $G$.
-It is obvious that the priority of each $d_i$ is not $\textit{pri}_x$.
-Then our proof proceeds as follows:
-
-According to the definition of left-right constraint, there are two possibilities.
-The first possibility is that, $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$.
-It is obvious that $\textit{rm}(d_1) \in R$, and then since $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_1),\textit{rm}(d_1) \in R$.
-Then,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Since $d_1 \rightarrow d_2$, by definition of $G$, we know that $\textit{put}(d_1)$ happens before $\textit{rm}(d_2)$.
-Since $\textit{put}(d_1) \in R$ and $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_2),\textit{rm}(d_2) \in R$.
-Similarly, for each $1 \leq i \leq m$, we know that $\textit{put}(d_i),\textit{rm}(d_i) \in R$.
-
-\item[-] Since $d_m \rightarrow x$,
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] if $\textit{put}(d_m)$ happens before $\textit{put}(x)$, then we can see that $\textit{put}(d_m) \in L$, which contradicts that $\textit{put}(d_m) \in R$.
-
-    \item[-] if $\textit{put}(d_m)$ happens before $\textit{rm}(x)$, then we can see that $\textit{put}(d_m) \in L \cup M$, which contradicts that $\textit{put}(d_m) \in R$.
-    \end{itemize}
-\end{itemize}
-
-The second possibility is that, $h$ contains one $\textit{put}(d_1,\_)$ operation and no $\textit{rm}(d_1)$ operation.
-Note that for each $j > 1$, $h$ contains $\textit{put}(d_j,\_)$ and $\textit{rm}(d_j)$.
-Since $d_m \rightarrow x$, is is obvious that $\textit{put}(d_m) \in L \cup M$.
-Since $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_m),\textit{rm}(d_m) \in R$.
-Then,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Since $d_{\textit{m-1}} \rightarrow d_m$, by definition of $G$, we know that $\textit{put}(d_{\textit{m-1}})$ happens before $\textit{rm}(d_m)$.
-Since $\textit{rm}(d_m) \in L \cup M$ and $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_{\textit{m-1}}),\textit{rm}(d_{\textit{m-1}}) \in L \cup M$.
-Similarly, for each $1 < i \leq m$, we know that $\textit{put}(d_i),\textit{rm}(d_i) \in L \cup M$, and $\textit{put}(d_1)\in L \cup M$.
-
-\item[-] Since $d_1 \rightarrow d_2$ in $G$, $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2)$.
-Since $\textit{put}(d_2),\textit{rm}(d_2) \in L \cup M$, we can see that $\textit{put}(d_1) \in L \cup M$.
-However, there is one $\textit{put}(d_1,\_)$ operation and no $\textit{rm}(d_1)$ operation in $h$, contradicts that $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
-\end{itemize}
-
-This completes the proof of the $\textit{only if}$ direction.
-
-To prove the $\textit{if}$ direction, assume that $G$ has no cycle going through $x$.
-Let $O_L$ be the set of operations that happen before $\textit{put}(x)$ in $h$.
-It is easy to see that $O_L \subseteq \textit{LMSet}(h,x)$.
-Let $O_M = \textit{LMSet}(h,x) \setminus O_L$.
-Let $O_h$ be the set of operations of $h$, and let $O_R = O_h \setminus \textit{LMSet}(h,x)$.
-Let $L$, $M$ and $R$ be $O_L$, $O_M$ and $O_R$, respectively.
-
-By Lemma \ref{lemma:LMSet has matched put and rm operations}, we can see that $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
-It remains to prove that $L$, $M$ and $R$ is a gap of $\textit{put}(x)$ and $\textit{rm}(x)$.
-We prove this by showing that all the following cases are impossible:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Case $1$: Some operation $o_r \in R$ happens before $\textit{rm}(x)$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $2$: Some operation $o_r \in R$ happens before some operation $o_{lm} \in L \cup M$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $3$: Some operation $o_r \in R$ happens before $\textit{put}(x)$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $4$: $\textit{rm}(x)$ happens before some $o_{lm} \in \textit{LMSet}(h,x) = L \cup M$. By Lemma \ref{lemma:Rmx does not happen before LMSet for Rpr1} we know that this is impossible.
-
-\item[-] Case $5$: $\textit{rm}(x)$ happens before $\textit{put}(x)$. This contradicts that for each item in $h$, its $\textit{rm}$ operation does not happen before its $\textit{put}$ operation.
-
-\item[-] Case $6$: Some operation $o_m \in M$ happens before $\textit{put}(x)$. Then we know that $o_m \in L$, which contradicts that $o_m \in M$.
-
-\item[-] Case $7$: Some operation $o_m \in M$ happens before some operation $o_l \in L$. Then we know that $o_m \in L$, which contradicts that $o_m \in M$.
-
-\item[-] Case $8$: $\textit{put}(x)$ happens before some operation $o_l \in L$. This is impossible.
-\end{itemize}
-
-This completes the proof of the $\textit{if}$ direction.
-
-\qed
-\end {proof}
-
-A automaton $\mathcal{A}_{\textit{Rpr1-1}}$ is given in \figurename~\ref{fig:automata for Rpro-1}, where $c=\textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm,a),$ $\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$.
-$\mathcal{A}_{\textit{Rpr1-1}}$ is used to detect the existence of a item whose $\textit{rm}$ happens before its $\textit{put}$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-1.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-1}}$}
-  \label{fig:automata for Rpro-1}
-\end{figure}
-
-Let us begin to represent several automata that is used for capture the case that, in a sub-history $h'$ of a history $h$, $\textit{last}(h')=R_{\textit{pr1}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$, and the reason is that there is no gap on $\textit{put}(y,\textit{pri}_y)$ and $\textit{rm}(y)$ in $h'$, where $y$ is the item that has maximal priority in $h'$.
-Such case is dealt with by four automata, since there are four possibilities of $h' \vert_{y}$.
-
-The first automaton is $\mathcal{A}_{\textit{Rpr1-pprr}}$, which is given in \figurename~\ref{fig:automata for Rpro-pprr}.
-In \figurename~\ref{fig:automata for Rpro-pprr}, $c_1= \textit{cal}(\textit{put},c,$ $\textit{anyPri}),\textit{ret}(\textit{put},c), \textit{cal}(\textit{rm},c), \textit{ret}(\textit{rm},c)$, $c_2 = c_1 + \textit{cal}(\textit{put},a,\textit{les}_p)$, and $c_3 = c_2 + \textit{ret}(\textit{rm},a)$.
-Here $p$ is a variable for priority, and $\textit{les}_p$ is a predicate that accepts all priorities that are less than the value of $p$.
-Here $\textit{pprr}$ represents that, $h' \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$ (we consider $b$ as $y$ and $p$ as $\textit{pri}_y$).
-In \figurename~\ref{fig:automata for Rpro-pprr}, the three paths that (1) from $q_{\textit{init}}$ to $q_2$ to $q_3$, (2) from $q_{\textit{init}}$ to $q_2$ to $q_{11}$, and (3) from $q_{\textit{init}}$ to $q_{10}$ to $q_{11}$ represents three possible positions of $\textit{ret}(\textit{put},a)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-pprr.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-pprr}}$}
-  \label{fig:automata for Rpro-pprr}
-\end{figure}
-
-The second automaton is $\mathcal{A}_{\textit{Rpr1-prpr}}$, which is given in \figurename~\ref{fig:automata for Rpro-prpr}.
-In $\mathcal{A}_{\textit{Rpr1-prpr}}$, $c_1, c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$.
-Here $\textit{pprr}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{put},b,p) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-prpr.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-prpr}}$}
-  \label{fig:automata for Rpro-prpr}
-\end{figure}
-
-The third automaton is $\mathcal{A}_{\textit{Rpr1-rppr}}$, which is given in \figurename~\ref{fig:automata for Rpro-rppr}.
-In $\mathcal{A}_{\textit{Rpr1-rppr}}$, $c_1, c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$.
-Here $\textit{rppr}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-rppr.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-rppr}}$}
-  \label{fig:automata for Rpro-rppr}
-\end{figure}
-
-The forth automaton is $\mathcal{A}_{\textit{Rpr1-rprp}}$, which is given in \figurename~\ref{fig:automata for Rpro-rprp}.
-In $\mathcal{A}_{\textit{Rpr1-rprp}}$, $c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$, and $c_4=c_1 + \textit{ret}(put,b)$.
-Here $\textit{rprp}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b) \cdot \textit{ret}(\textit{put})$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-rprp.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-rprp}}$}
-  \label{fig:automata for Rpro-rprp}
-\end{figure}
-
-Given a data-differentiated history $h$, we say that a item $x$ with priority $\textit{pri}_x$ in $h$ is covered by items $d_1,\ldots,d_m$ in $h$, if the priorities of $d_1,\ldots,d_m$ is not $\textit{pri}_x$, and
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{put}(d_m,\_)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$,
-
-\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,
-
-\item[-] $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$, or $\textit{rm}(d_1)$ does not exists in $h$
-\end{itemize}
-
-According to the definition of left-right constraint, in a data-differentiated history $h$, there is a cycle going through $x$, if and only if there exists items $d_1,\ldots,d_m$, such that $x$ is covered by $d_1,\ldots,d_m$.
-
-\begin{restatable}{lemma}{Rpr1IsCoRegular}
-\label{lemma:Rpr1 is co-regular}
-$R_{\textit{pr1}}$ is co-regular.
-\end{restatable}
-
-\begin {proof}
-
-Let $\textit{AutS-R}_{\textit{pr1}} = \{ \mathcal{A}_{\textit{Rpr1-1}} , \mathcal{A}_{\textit{Rpr1-pprr}}, \mathcal{A}_{\textit{Rpr1-prpr}}, \mathcal{A}_{\textit{Rpr1-rppr}}, \mathcal{A}_{\textit{Rpr1-rprp}} \}$.
-We need to prove that:
-
-\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr1}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr1}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr1}}$.
-
-By Lemma \ref{lemma:Gap Equals Lin for Rpr1} and Lemma \ref{lemma:Gap Equals Constraint for Rpr1}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:
-
-\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr1}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h''$ is a projection of $h$, such that $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is in $h''$, $\textit{pri}_x$ is the maximal priority of $h''$, and one the the following case holds:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] case $1$: $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$,
-\item[-] case $2$: $x$ is covered by some items $d_1,\ldots,d_m$ in $h''$.
-\end{itemize}
-
-\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\textit{AutS-R}_{\textit{pr1}}$.
-By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$.
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-1}}$: Let $d_1$ a item in $h_2$ such that $r(d_1)=b$ and let $h'' = h_2 \vert_{ \{ d_1 \} }$.
-$d_1$ is the item with maximal priority in $h''$.
-It is obvious that $\textit{last}(h'') = R_{\textit{pr1}}$ and $\textit{rm}(d_1)$ happens before $\textit{put}(d_1,\_)$ in $h''$.
-
-\item[-] If $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-pprr}}$, $\mathcal{A}_{\textit{Rpr1-prpr}}$, $\mathcal{A}_{\textit{Rpr1-rppr}}$, or $\mathcal{A}_{\textit{Rpr1-rprp}}$: Let $d_0$ be a item in $h_2$ such that $r(d_0)=b$, let $d_1,\ldots,d_m$ be the items in $h_2$ such that $r(d_i)=a$ for each $1 \leq i \leq m$.
-Let $h'' = h_1 \vert_{ \{ d_0, d_1, \ldots, d_m \} }$.
-It is obvious that the priority of $d_0$ is larger than the priorities of other operations in $h''$, and then $\textit{last}(h'') = R_{\textit{pr1}}$.
-It is easy to see that $d_0$ is covered by $d_1,\ldots,d_m$.
-\end{itemize}
-
-\noindent The $\textit{if}$ direction: Assume that there exists such $h$, $h''$, $x$ and $\textit{pri}_x$.
-Since $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{pri}_x$ is larger than the priorities of other operations in $h''$.
-Then,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If case $1$ holds: Then, $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$ in $h''$.
-Let $h_1$ be obtained from $h$ by renaming $x$ to $b$ and renaming all other items to $a$.
-By data-independence, $h_1 \in \mathcal{I}$, and it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-1}}$.
-\item[-] case $2$: Then, $x$ is covered by some items $d_1,\ldots,d_m$ in $h''$.
-There are four possibilities of $h'' \vert_{ \{ x \} }$: (1) $\textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{rm},x)$, (2) $\textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},x)$, (3) $\textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},x)$ and (4) $\textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{rm},x) \cdot \textit{ret}(\textit{put})$.
-
-    Let $h_1$ be obtained from $h$ by renaming $x$ into $b$, renaming $d_1,\ldots,d_m$ into $a$ and renaming other items into $c$.
-We can prove that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-pprr}}$, $\mathcal{A}_{\textit{Rpr1-prpr}}$, $\mathcal{A}_{\textit{Rpr1-rppr}}$ or $\mathcal{A}_{\textit{Rpr1-rprp}}$.
-
-    For example, if $h'' \vert_{ \{ x \} } = \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{rm},x)$, and $\textit{ret}(\textit{rm},d_m)$ is before $\textit{call}(\textit{put},x,\textit{pri}_x)$ and is after $\textit{ret}(\textit{put},x)$ in $h''$.
-Then in  $\mathcal{A}_{\textit{Rpr1-pprr}}$, $h_1$ is accepted by transitions from $q_{\textit{init}}$ to $q_2$ to $q_{11}$ and finally to $q_7$ (if $\textit{rm}(d_1)$ does not exists in $h''$) or $q_9$ (if $\textit{rm}(d_1)$ exists in $h''$).
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\subsection{Co-Regular of $R_{\textit{pr2}}$}
-\label{subsec:co-regular of Rpr2}
-
-An automaton $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$ is given in \figurename~\ref{fig:automata for FIFO of single priority}.
-In \figurename~\ref{fig:automata for FIFO of single priority}, let $c_1= \textit{cal}(\textit{put},c,\textit{anyPri}),\textit{ret}(\textit{put},c)$, $\textit{cal}(\textit{rm},c),\textit{ret}(\textit{rm},c)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, and $c_3 = c_2 + \textit{cal}(\textit{rm},a) + \textit{ret}(\textit{rm},a)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_QUEUE_FOR_SINGLE_PRI.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$}
-  \label{fig:automata for FIFO of single priority}
-\end{figure}
-
-According to \cite{Bouajjani:2015}, given a data-independent set $S$ of queue executions, whether every sequence $s$ in $S$ is linearizable to queue can be reduced into checking emptiness of intersection of $S$ and several finite automata.
-Let $\textit{transToQueue}(h)$ be a sequence $h'$ that is generated from $h$ by transforming $\textit{put}$ and $\textit{rm}$ into $\textit{enq}$ and $\textit{deq}$, respectively, and then discarding priorities.
-A priority queue history is called single-priority, if all of its items has only one priority.
-The following lemma states that we can also use finite automata to check ``first in first out property'' for single-priority sub-sequence of priority queue history.
-
-\begin{restatable}{lemma}{AutoForPQwithSignlePri}
-\label{lemma:automata for priority queue with single priority}
-
-Given a data-independent implementations $\mathcal{I}$ of priority queue, let $\mathcal{A} = \{ \mathcal{A}_{\textit{unmRm1}}, \mathcal{A}_{\textit{Rpr0-1}}, \mathcal{A}_{\textit{unmRm2}}, \mathcal{A}_{\textit{FIFO-Sig-Pri}} \}$, then,
-
-$\mathcal{I} \cap \mathcal{A} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority, and $\textit{transToQueue}(e')$ does not linearizable to queue.
-\end{restatable}
-
-\begin {proof}
-
-According to \cite{Bouajjani:2015}, checking linearizable with respect to queue can be reduced into checking the following three conditions for data-differentiated histories:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{deq}(a)$ happens before $\textit{enq}(a)$ (or $\textit{enq}(a)$ does not exist) for some item $a$,
-
-\item[-] there are two or more $\textit{deq}(a)$ for some item $a$,
-
-\item[-] $\textit{enq}(a)$ happens before $\textit{enq}(b)$, and $\textit{deq}(b)$ happens before $\textit{deq}(a)$ (or $\textit{deq}(a)$ does not exists) for some items $a,b$.
-\end{itemize}
-
-It is not hard to see that (1) the first condition is dealt with by $\mathcal{A}_{\textit{unmRm1}}$ (if $\textit{put}(a,\_)$ does not exist) and $\mathcal{A}_{\textit{Rpr0-1}}$ (if $\textit{put}(a,\_)$ exists), (2) the second condition is dealt with by $\mathcal{A}_{\textit{unmRm2}}$, and (3) the third condition is dealt with by $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$.
-\qed
-\end {proof}
-
-By Lemma \ref{lemma:automata for priority queue with single priority}, from now on, it is safe for us to assume that, for each data-differentiated priority queue history $h$ and each $h'$ of projection of $h$, if $h'$ is single-priority, then $\textit{transToQueue}(h')$ is linearizable to queue.
-One consequence of this assumption is that, in $h'$, there does not exists an item $a$, such that $\textit{rm}(a)$ happens before $\textit{put}(a)$.
-
-By the proof of Lemma \ref{lemma:Rpr1 is co-regular}, with the help of $\mathcal{A}_{\textit{Rpr1-pprr}}, \mathcal{A}_{\textit{Rpr1-prpr}}, \mathcal{A}_{\textit{Rpr1-rppr}}$ and $\mathcal{A}_{\textit{Rpr1-rprp}}$, from now on, we can safely assume that given a data-differentiated history $h$ and operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ in $h$, there does not exists any items $d_1,\ldots,d_m$, such that $x$ is not covered by items $d_1,\ldots,d_m$ and the priority of $d_1,\ldots,d_m$ is less that $x$.
-
-However, above two assumptions are not enough for imply the linearizability to priority queue.
-One example is shown in \figurename~\ref{fig:history introduct ob order1}.
-The history $h$ of \figurename~\ref{fig:history introduct ob order1} is not linearizable, even if $h \vert_{1}$ and $h \vert_{2}$ are both linearizable, and either $a$ or $b$ is covered by other items.
-One explanation is that, since $\textit{put}(b,1)$ happens before $\textit{put}(a,1)$, in choosing linearization points, we should make $\textit{rm}(b)$ before $\textit{rm}(a)$.
-However, to satisfy the requirements of $R_{\textit{pr2}}$, the linearization points of $\textit{rm}(a)$ can only locates in the interval from $\textit{ret}(\textit{rm},d)$ to $\textit{cal}(\textit{put},e,2)$.
-But, $\textit{rm}(b)$ is after this entire interval, and we can not locate the linearization point of $\textit{rm}(b)$ in this interval.
-When there is no such phenomenon, such as the history in \figurename~\ref{fig:history introduct ob order2}, the history becomes linearizable.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER1.pdf}
-%\vspace{-10pt}
-  \caption{above assumptions do not implies linearizability (1)}
-  \label{fig:history introduct ob order1}
-\end{figure}
-
-%\begin{figure}[htbp]
-%  \centering
-%  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER3.pdf}
-%\vspace{-10pt}
-%  \caption{above assumptions do not implies linearizability (2)}
-%  \label{fig:history introduct ob order1}
-%\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER2.pdf}
-%\vspace{-10pt}
-  \caption{above assumptions do not implies linearizability (2)}
-  \label{fig:history introduct ob order2}
-\end{figure}
-
-In the following part of this subsection, we use $<_{\textit{ob}}$ order for the case when we can infer that an item is inserted earlier than another item, and we use gap-point to represent the intervals when it is possible to locate linearizaton points of $\textit{rm}(a)$.
-
-\begin{definition}[gap-point for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:gap-point for matched put and rm operations}
-
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x),\textit{rm}(x)$ of $h$, we say $o$ is a gap-point of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$, if we can obtain a sequence $\textit{gapPSeq}(h)$ from $h$ by adding $o$ to some locations between actions of $h$, and satisfies that,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] In $\textit{gapPSeq}(h)$, $o$ is after $\textit{call}(\textit{put},x,\textit{pri}_x)$ and $\textit{call}(\textit{rm},x)$, and is before $\textit{ret}(\textit{rm},x)$.
-\item[-] For each item $y$ with non-$\textit{pri}_x$ priority in $\textit{gapPSeq}(h)$.
-If $\textit{put}(y,\_)$ and $\textit{rm}(y)$ is in $h$, then it can not be the case that $\textit{ret}(\textit{put},y,\_)$ is before $o$ and $\textit{call}(\textit{rm},x)$ is after $o$; Else if $\textit{put}(y,\_)$ is in $h$ and $\textit{rm}(y)$ is not in $h$, then it can not be the case that $\textit{ret}(\textit{put},y,\_)$ is before $o$;
-\end{itemize}
-
-We say that $o$ is a gap-point of priority $\textit{pri}_x$, if we can obtain a sequence $\textit{gapPSeq}(h)$ from $h$ by adding $o$ to some locations between actions of $h$, and satisfies only the latter requirement.
-\end{definition}
-
-\begin{restatable}{lemma}{GapEqualsGapPoint}
-\label{lemma:gap eqals the existence of gap point}
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-There is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, if and only if there is a gap-point $o$ of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{only if}$ direction, we need to prove that: if there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, then there is no gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-
-We prove this by contradiction.
-Assume that there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, but there is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-Since there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, in the time interval starts from $\textit{cal}(\textit{put},x,\textit{pri}_x)$ and $\textit{cal}(\textit{rm},x)$ and ends at $\textit{ret}(\textit{rm},x)$, there is always some item with non-$\textit{pri}_x$ priority in priority queue.
-Therefore, it is not hard to prove that there is a cycle going through the left-right constraint of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, and by Lemma \ref{lemma:Gap Equals Constraint for Rpr0}, there is no gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, which contradicts that there is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-
-To prove the $\textit{if}$ direction, assume that there is a gap-point $o$ of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$ and let $h_1$ be the sequence in definition of gap-point, which is generated from $h$ by adding $o$ into $h$.
-Let sequence $h_2$ be a sequence of call, return and operations, such that the projection of $h_2$ into call and return actions is $h$.
-$h_2$ is generated as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] First, from $h_1$, transform the point $o$ into $\textit{rm}(x)$.
-
-\item[-] Then for all the other items $y$, (1) adding an operation $\textit{put}(y,p)$ at some position between $\textit{call}(\textit{put},y,\_)$ and $\textit{ret}(\textit{put},y)$, and (2) adding an operation $\textit{rm}(y)$ at some position between $\textit{call}(\textit{rm},y)$ and $\textit{ret}(\textit{rm},y)$.
-
-\item[-] For all items $z$ with the same priority as $x$, the positions for adding operations of $z$ into $h$ is arbitrary (but need to satisfy above restriction.
-In the following we implicitly require this).
-
-\item[-] For all items $z$ whose priority is different from that of $x$,
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-
-    \item[-] If $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, and $\textit{ret}(\textit{put},z,\_)$ is before $o$ in $h_1$: Then in $h_2$, $\textit{rm}(z,\_)$ is just after $\textit{call}(\textit{rm},z)$.
-
-    \item[-] If $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, and $o$ is before $\textit{ret}(\textit{put},z,\_)$ $h_1$: Then in $h_2$, $\textit{put}(z,\_)$ is just before $\textit{ret}(\textit{put},z)$.
-
-    \item[-] Otherwise, we can see that $\textit{put}(z,\_)$ and $\textit{rm}(z)$ overlaps in $h$.
-Then we add $o_1=\textit{put}(z,\_)$ and $o_2=\textit{rm}(z)$ in the time interval of interleaving of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ overlaps in $h$, make sure $o_2$ next to $o_1$, and make sure either both of them is before $\textit{rm}(x)$ or both of them is after $\textit{rm}(x)$ in $h_2$.
-    \end{itemize}
-\end{itemize}
-
-According to the definition of gap-point, it is not hard to see that we can construct such $h_2$.
-Let $l$ be the projection of $h_2$ into operations, and $l'$ be the projection of $l$ into $x$ and all items that has priorities not equals that of $x$.
-Let $l'= u \cdot \textit{put}(x,\_) \cdot v \cdot \textit{rm}(x) \cdot w$.
-It is easy to see that $h \sqsubseteq l$, and $u \cdot v$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
-Let $L$, $M$ and $R$ be the set of operations in $u$, $v$ and $w$, respectively.
-Then it is obvious that $L$, $M$ and $R$ is a gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-\qed
-\end {proof}
-
-Given a data-differentiated history $h$ and two items $a,b$, we say that $a <_{\textit{ob}} b$, if the priority of $a$ and $b$ is the same, and one of the following cases holds:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Type $A$: $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$ in $h$,
-
-\item[-] Type $B$: $\textit{rm}(a)$ happens before $\textit{rm}(b)$ in $h$,
-
-\item[-] Type $C$: $\textit{rm}(a)$ happens before $\textit{put}(b,\_)$ in $h$,
-\end{itemize}
-
-Let $<_{\textit{ob}}^*$ be the transitive closure of $<_{\textit{ob}}$.
-Intuitively, $a <_{\textit{ob}} b$ means that $a$ should be inserted earlier than $b$, and $a <_{\textit{ob}}^* b$ means that we can infer that $a$ should be inserted earlier than $b$ with the help of other items.
-We can explicitly write $a <_{\textit{ob}}^A b$ to mean that $a <_{\textit{ob}} b$ and the reason is that $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$.
-Similarly, we can explicitly write $a <_{\textit{ob}}^B b$ and $a <_{\textit{ob}}^C b$ to mean that $a <_{\textit{ob}} b$ and the reason is that $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$ and $\textit{rm}(a)$ happens before $\textit{put}(b,\_)$, respectively.
-
-According to the definition of $<_{\textit{ob}}^*$, if $a <_{\textit{ob}}^* b$, then there exists $c_1,\ldots,c_n$, such that $a <_{\textit{ob}} c_1 <_{\textit{ob}} \ldots <_{\textit{ob}} b$.
-The following lemma states that, the number of intermediate items $c_i$ is in fact bounded.
-
-\begin{restatable}{lemma}{OBOrderHasBoundedLength}
-\label{lemma:ob order has bounded length}
-Given a data-differentiated history $h$.
-Assume that $a <_{\textit{ob}} a_1 <_{\textit{ob}} \ldots <_{\textit{ob}} a_m <_{\textit{ob}} b$, then either $a <_{\textit{ob}} b$, or $a <_{\textit{ob}} a_i <_{\textit{ob}} b$ for some $i$, or $a <_{\textit{ob}} a_i <_{\textit{ob}} a_j <_{\textit{ob}} b$ for some $i,j$.
-\end{restatable}
-
-\begin {proof}
-
-Our proof proceed as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^A$,$<_{\textit{ob}}^B \cdot <_{\textit{ob}}^B$ and $<_{\textit{ob}}^C \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^A c_2 <_{\textit{ob}}^A c_1$, then $\textit{put}(c_3,\_)$ happens before $\textit{put}(c_2,\_)$, and $\textit{put}(c_2,\_)$ happens before $\textit{put}(c_1,\_)$.
-Therefore, it is obvious hat $\textit{put}(c_3,\_)$ happens before $\textit{put}(c_1,\_)$ and $c_3 <_{\textit{ob}}^A c_1$.
-
-    Similarly, if $c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^B c_1$, then $c_3 <_{\textit{ob}}^B c_1$.
-
-    If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^C c_1$: Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
-Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
-
-    Therefore, when we meet successive $<_{\textit{ob}}^A$, it is safe to leave only the first and the last elements and ignore intermediate elements.
-Similar cases hold for $<_{\textit{ob}}^B$ and $<_{\textit{ob}}^C$.
-
-\item[-] $<_{\textit{ob}}^A$ and $<_{\textit{ob}}^C$:
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^A c_2 <_{\textit{ob}}^C c_1$.
-Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{ob}}^A c_2$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
-Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^A c_1$.
-
-    \item[-] ($<_{\textit{ob}}^C \cdot <_{\textit{ob}}^A$): If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^A c_1$.
-Since $c_2 <_{\textit{ob}}^A c_1$, $\textit{put}(c_2,\_)$ happens before $\textit{put}(c_1,\_)$.
-Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{rm}(c_3)$ happens before $\textit{put}(c_2,\_)$.
-Therefore, $\textit{rm}(c_3)$ happens before $\textit{put}(c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
-    \end{itemize}
-
-\item[-] $<_{\textit{ob}}^B$ and $<_{\textit{ob}}^C$:
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] ($<_{\textit{ob}}^B \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^C c_1$.
-Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-It is obvious that $\textit{cal}(\textit{rm},c_2)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$.
-Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
-
-    \item[-] ($<_{\textit{ob}}^C \cdot <_{\textit{ob}}^B$): If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^B c_1$.
-Since $c_2 <_{\textit{ob}}^B c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{rm},c_1)$.
-Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
-Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_1)$, and $c_3 <_{\textit{ob}}^B c_1$.
-    \end{itemize}
-
-\item[-]  ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^B \cdot <_{\textit{ob}}^A$): If $c_4 <_{\textit{ob}}^A c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^A c_1$:
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$: Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$.
-Then $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
-This implies that $c_4 <_{\textit{ob}}^A c_3 <_{\textit{ob}}^C c_1$.
-According to the fact for $<_{\textit{ob}}^A \cdot <_{\textit{ob}}^C$, we know that $c_4  <_{\textit{ob}}^A c_1$.
-
-    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is after $\textit{cal}(\textit{put},c_1,\_)$: Since $c_2 <_{\textit{ob}}^A c_1$, $\textit{ret}(\textit{put},c_2,\_)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{rm}(c_3)$ happens before $\textit{rm}(c_2)$, and then we know that $\textit{put}(c_2,\_)$ can not happen before $\textit{put}(c_3,\_)$.
-Since $\textit{put}(c_2,\_)$ does not happen before $\textit{put}(c_3,\_)$, $\textit{cal}(\textit{put},c_3,\_)$ is before $\textit{ret}(\textit{put},c_2,\_)$.
-Since $c_4 <_{\textit{ob}}^A c_3$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_3,\_)$.
-Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_4 <_{\textit{ob}}^A c_1$.
-    \end{itemize}
-
-\end{itemize}
-
-Based on above results, given $a <_{\textit{ob}}^{b_1} a_1 <_{\textit{ob}} \ldots <_{\textit{ob}}^{b_m} a_m <_{\textit{ob}}^{b_{\textit{m+1}}} b$, where each $b_i$ is in $\{ A,B,C \}$, we can merge relations, until we got one of the following facts:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $a <_{\textit{ob}}^A b$, $a <_{\textit{ob}}^B b$ or $a <_{\textit{ob}}^C b$,
-
-\item[-] $a <_{\textit{ob}}^A a_i <_{\textit{ob}}^B b$, or $a <_{\textit{ob}}^B a_i <_{\textit{ob}}^A b$, for some $i$,
-
-\item[-] $a <_{\textit{ob}}^B a_i <_{\textit{ob}}^A a_j <_{\textit{ob}}^B b$, for some $i$ and $j$,
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-Given a differentiated history $h$ and operations $\textit{put}(x,\_)$ and $\textit{rm}(x)$ of $h$, the time interval from the time point when both $\textit{call}(\textit{put},x,\_)$ and $\textit{cal}(\textit{rm},x)$ happen and ends at $\textit{ret}(\textit{rm},x)$ is called the effective interval of $x$ in $h$.
-
-\begin{restatable}{lemma}{Rpr2AsGPandOBOrder}
-\label{lemma:Rpr2 as gap-point and ob order}
-Given a data-differentiated history $h$.
-There exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr2}}$, and $h'$ does not linearizable to $\textit{MR}_{\textit{pr2}}$, if and only if there exists $h'' \in \textit{proj}(h)$, $\textit{last}(h'') = R_{\textit{pr2}}$, $x$ and $y$ has maximal priority in $h''$, $y <_{\textit{ob}} x$ in $h''$, and the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$ in $h''$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{if}$ direction, let $h_1$ be the history that is obtained from $h''$ by erasing all operations of items that has same priority as $x$, except for operations of $x$ and $y$.
-It is obvious that $\textit{last}(h_1) = R_{\textit{pr2}}$.
-According to $y <_{\textit{ob}} x$,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $y <_{\textit{ob}}^A x$, then $\textit{put}(y,\_)$ happens before $\textit{put}(x,\_)$.
-Since in $R_{\textit{pr2}}$, $\textit{put}(\textit{itm},\_)$ is after $\textit{put}$ operations of items with maximal priority (except for $\textit{itm}$), we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
-\item[-] If $y <_{\textit{ob}}^B x$, then $\textit{rm}(y)$ happens before $\textit{rm}(x)$.
-Since in $R_{\textit{pr2}}$, $\textit{rm}(\textit{itm})$ is the last $\textit{rm}$ operation of items with maximal priority, we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
-
-\item[-] If $y <_{\textit{ob}}^C x$, then $\textit{rm}(y)$ happens before $\textit{put}(x,\_)$.
-Since in $R_{\textit{pr2}}$, $\textit{rm}(\textit{itm})$ is the last $\textit{rm}$ operation of items with maximal priority, we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
-\end{itemize}
-
-Similarly as the proof in Lemma \ref{lemma:gap eqals the existence of gap point}, we can see that the only possible position for locating linearizaton point of $\textit{rm}(x)$ is at gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$.
-Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, then there exists unmatched operations before $\textit{rm}(x)$.
-
-Since the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},x)$ in $h''$ (and also in $h_1$), if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, then $\textit{rm}(x)$ will be before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},x)$.
-
-Therefore, for every sequence $l = u \cdot \textit{put}(x,\_) \cdot v \cdot \textit{rm}(x) \cdot w$, if $h_1 \sqsubseteq l$, then either $u \cdot v$ contains some unmatched operations of priority less than priority of $x$, or $w$ contains $\textit{put}(y,\_)$ or $\textit{rm}(y)$.
-In both cases, $l \notin \textit{MR}_{\textit{pr2}}$.
-
-To prove the $\textit{only if}$ direction, we prove its contrapositive.
-Assume we already know that for each $h'' \in \textit{proj}(h)$ with $\textit{last}(h'') = R_{\textit{pr2}}$, for each $x$ and $y$ has maximal priority in $h''$, if $y <_{\textit{ob}} x$ in $h''$, then the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},x)$ in $h''$.
-We need to prove that for each $h' \in \textit{proj}(h)$, if $\textit{last}(h') = R_{\textit{pr2}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr2}}$.
-
-Let $\textit{pri}$ be the maximal priority of $h'$.
-Recall that we already assume that each single-priority history (if we consider $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$) can be ``linearizable to queue'', and item with larger priority is not covered by items with smaller priority.
-Our proof proceed as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Let $h_{\textit{pri}}$ be the projection of $h$ into operations of priority $\textit{pri}$.
-Since each single-priority history can be ``linearizable to queue'', there exists sequence $l_{\textit{pri}}$, such that $h_{\textit{pri}} \sqsubseteq l_{\textit{pri}}$, and when we treat $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$, $l_{\textit{pri}}$ belongs to queue.
-
-\item[-] Step $1$: Let $a_1$ be the last inserted item of $l_{\textit{pri}}$.
-
-$\textit{Cond1}$: If for each item $b$ with priority $\textit{pri}$ in $h'$, the rightmost gap-point $o$ of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$ does not before any $\textit{cal}(\textit{put},b,\textit{pri})$ or $\textit{cal}(\textit{rm},b)$ in $h'$, then we can
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] Locate the linearization point of $\textit{rm}(a_1)$ at $o$,
-    \item[-] If $\textit{put}(a_1,\textit{pri})$ overlaps with $\textit{rm}(a_1)$, then locate the linearization point of $\textit{put}(a_1,\_)$ just before the linearization point of $\textit{rm}(a_1)$.
-Otherwise, $\textit{put}(a_1,\textit{pri}) <_{\textit{hb}} \textit{rm}(a_1)$, and we locate the linearization point of $\textit{put}(a_1,\_)$ just before its return action.
-    \item[-] Locate linearization points of operations of items of priority $\textit{pri}$ (except for $a_1$) just after the call action of the operation.
-    \item[-] For item $z$ with priority smaller than $\textit{pri}$.
-If both $\textit{cal}(\textit{put},z,\_)$ and $\textit{cal}(\textit{rm},z)$ is before $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just after their call actions.
-If both $\textit{ret}(\textit{put},z)$ and $\textit{ret}(\textit{rm},z)$ is after $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just before their return actions.
-
-        Otherwise, either (1) $\textit{rm}(z) <_{\textit{hb}} \textit{put}(z,\_)$, or (2) $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, $\textit{ret}(\textit{put},z,\_)$ before $o$, and $o$ before $\textit{call}(\textit{rm},z)$.
-The first case violates that each single-priority history can be ``linearizable to queue'', and the second case violates that $o$ is a gap-point.
-    \end{itemize}
-
-    In this way we construct a sequence $l$, and it is easy to see that $h' \sqsubseteq l$ and $l$ satisfies the requirement of $R_{\textit{pr2}}$.
-
-\item[-] Otherwise, there exists $a_2$ with priority $\textit{pri}$, such that the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$ in $h'$.
-We can see that the rightmost gap-point of $\textit{put}(a_2,\textit{pri})$ and $\textit{rm}(a_2)$ is after the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$.
-By assumption we know that $a_2$ does not $<_{\textit{ob}}$ to $a_1$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If for each item $b$ with priority $\textit{pri}$ in $h'$, $a_2$ does not $<_{\textit{ob}}$ to $b$.
-Then we go to step $1$ and treat $a_2$ as $a_1$ (When $\textit{Cond1}$ of step $1$ holds, we break the loop).
-    \item[-] Otherwise, there exists $a_3$ with priority $\textit{pri}$ such that $a_2 <_{\textit{ob}} a_3$.
-By assumption,we know that the rightmost gap-point of $\textit{put}(a_3,\textit{pri})$ and $\textit{rm}(a_3)$ is not before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$.
-Then we can see that the rightmost gap-point of $\textit{put}(a_3,\textit{pri})$ and $\textit{rm}(a_3)$ is after the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$.
-Then we go to step $1$ and treat $a_3$ as $a_1$ (When $\textit{Cond1}$ of step $1$ holds, we break the loop).
-    \end{itemize}
-\end{itemize}
-
-It is not hard to see that, if $b_i$ is the $a_1$ in the $\textit{i-th}$ loop of our proof, $b_j$ is the $a_1$ in the $\textit{j-th}$ loop of our proof, and $i<j$, then the rightmost gap-point of $\textit{put}(b_j,\textit{pri})$ and $\textit{rm}(b_j)$ is after the rightmost gap-point of $\textit{put}(b_i,\textit{pri})$ and $\textit{rm}(b_i)$.
-Thus, we finally stop at some $d$.
-$d$ satisfies that, for each item $c$ with priority $\textit{pri}$, $d$ does not $<_{\textit{ob}}$ to $c$.
-$d$ also satisfies that, for each item $c$ with priority $\textit{pri}$ in $h'$, the rightmost gap-point $o'$ of $\textit{put}(d,\textit{pri})$ and $\textit{rm}(d)$ does not before any $\textit{cal}(\textit{put},c,\textit{pri})$ or $\textit{cal}(\textit{rm},c)$.
-Then, we can
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Locate the linearization point of $\textit{rm}(d)$ at the rightmost gap-point $o'$ of $\textit{put}(d,\textit{pri})$ and $\textit{rm}(d)$,
-
-\item[-] Since for each item $c$ with priority $\textit{pri}$ in $h'$, $o'$ is after any $\textit{cal}(\textit{put},c,\textit{pri})$ and $\textit{cal}(\textit{rm},c)$.
-For each $c$, one of the following cases hold: (1) $\textit{cal}(\textit{rm},c)$ is before $o'$ and $o'$ is before $\textit{ret}(\textit{rm},c)$, (2) $\textit{cal}(\textit{rm},c)$ and $\textit{ret}(\textit{rm},c)$ are before $o'$.
-We locate the linearization point of $\textit{rm}(c)$ just after its call action.
-Such linearization point is obviously before $o'$.
-
-\item[-] If $o'$ is before $\textit{ret}(\textit{put},d,\textit{pri})$, then locate the linearization point of $\textit{put}(d,\textit{pri})$ just before $o'$.
-Else, $\textit{ret}(\textit{put},d,\textit{pri})$ is after $o'$, and we locate the linearization point of $\textit{put}(d,\textit{pri})$ just before the return action of it.
-
-\item[-] If $o'$ is before $\textit{ret}(\textit{put},d,\textit{pri})$, we can see that, for each item $c$ with priority $\textit{pri}$ in $h'$, $\textit{cal}(\textit{put},c,\textit{pri})$ is before $o'$.
-Then, and we locate the linearization point of $\textit{put}(c,\textit{pri})$ just before $o'$ and the linearization point of $\textit{put}(d,\textit{pri})$.
-
-    Otherwise, $\textit{ret}(\textit{put},d,\textit{pri})$ is before $o'$.
-Since $d$ does not $<_{\textit{ob}}$ to $c$, $\textit{put}(d,\textit{pri})$ does not happen before $\textit{put}(c,\textit{pri})$, or we can say, $\textit{ret}(\textit{put},d,\textit{pri})$ is after $\textit{cal}(\textit{put},c,\textit{pri})$.
-Then, we locate the linearization point of $\textit{put}(c,\textit{pri})$ just after its call action.
-\end{itemize}
-
-In this way we construct a sequence $l_d$, and it is easy to see that $h' \sqsubseteq l_d$ and $l_d$ satisfies the requirement of $R_{\textit{pr2}}$ with $d$ acts as $\textit{itm}$.
-This completes the proof of $\textit{if}$ direction.
-\qed
-\end {proof}
-
-Let us begin to represent several automata that is used for capture the case that, in a sub-history $h''$ of $h$, $\textit{last}(h'') = R_{\textit{pr2}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr2}}$, and the reason is that there exists items $a$ and $b$ with maximal priority in $h''$, $a <_{\textit{ob}} b$ in $h''$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
-
-According to Lemma \ref{lemma:ob order has bounded length}, if $a <_{\textit{ob}} b$, then one of the following cases hold: (1) $a <_{\textit{ob}}^A b$, (2) $a <_{\textit{ob}}^B b$, (3) $a <_{\textit{ob}}^C b$, (4) $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$ for some $a_1$, (5) $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$ for some $a_1$, or (6) $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$ for some $a_1$ and $a_2$.
-Although there are many possible enumeration of $a$, $b$, $a_1$ and $b_1$, the following lemma states that we need only to consider five enumeration of operations of $a$, $b$, $a_1$ and $b_1$.
-
-\begin{restatable}{lemma}{FiveEnmuerationisEnoughForRpr2}
-\label{lemma:five enumeration is enough for Rpr2}
-Given a data-differentiated history $h$, $\textit{last}(h) = R_{\textit{pr2}}$, and $h$ does not linearizable to $\textit{MR}_{\textit{pr2}}$.
-Assume that there exists items $a$ and $b$ with maximal priority in $h''$, $a <_{\textit{ob}} b$, and the rightmost gap-point $o$ of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
-Let $a_1$ and $a_2$ be the value when possibly $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$, or $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$.
-Then, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists) and $a_2$ (if exists).
-\end{restatable}
-\begin {proof}
-
-Let us prove by consider all the possible reason of $a <_{\textit{ob}}^A b$.
-According to Lemma \ref{lemma:ob order has bounded length}, we need to consider six reasons: %There may exists items with smaller priority of $a$ and covers some time interval, and assume that we rename them into $d$:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Reason $1$, $a <_{\textit{ob}}^A b$:
-
-    Since $a <_{\textit{ob}}^A b$, $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$.
-Since $o$ is after $\textit{cal}(\textit{put},a,\_)$, we can see that $o$ is before $\textit{cal}(\textit{rm},b)$.
-
-    Since single-priority must satisfy the first in firs out property, $\textit{rm}(b)$ does not happen before $\textit{rm}(a)$, and $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
-If $\textit{ret}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$, since we already assume that there exists gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$, this gap-point is also a gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$, and is after $o$, which contradicts that $o$ is the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$.
-Therefore, $\textit{ret}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$.
-
-    According to above discussion, there are two possible enumeration of operations of $a$ and $b$, as shown in \figurename~\ref{fig:history enumeration 1 for Rpr2} and \figurename~\ref{fig:history enumeration 2 for Rpr2}.
-Here we explicitly draw the leftmost gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$ as $o'$.
-Since the position of $\textit{ret}(\textit{put},b)$ does not influence the correctness, we can simply ignore it.
-
-\item[-] Reason $2$, $a <_{\textit{ob}}^B b$:
-
-    Since $a <_{\textit{ob}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$.
-Since $o$ is after $\textit{cal}(\textit{rm},b)$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$.
-Then, we can see that $o$ is before $\textit{cal}(\textit{put},a,\_)$.
-This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\_)$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible.
-Therefore, we can safely ignore this reason.
-
-\item[-] Reason $3$, $a <_{\textit{ob}}^C b$:
-
-    Since $a <_{\textit{ob}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},b)$.
-Since $o$ is after $\textit{cal}(\textit{put},b)$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$.
-Then, we can see that $o$ is before $\textit{cal}(\textit{put},a,\_)$.
-This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\_)$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible.
-Therefore, we can safely ignore this reason.
-
-\item[-] Reason $4$, $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$:
-
-    Since $a_1 <_{\textit{ob}}^B b$, $\textit{rm}(a_1) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_1)$ is before $\textit{cal}(\textit{rm},b)$.
-Since $\textit{rm}(a_1)$ does not happen before $\textit{put}(a_1)$, $\textit{cal}(\textit{put},a_1,\_)$ is before $\textit{ret}(\textit{rm},a_1)$.
-Since $a <_{\textit{ob}}^A a_1$, $\textit{ret}(\textit{put},a,\_)$ is before $\textit{cal}(\textit{put},a_1,\_)$.
-Therefore, $\textit{ret}(\textit{put},a,\_)$ is before $\textit{cal}(\textit{rm},b)$.
-Since $\textit{cal}(\textit{rm},b)$ is before $o$, we can see that $\textit{ret}(\textit{put},a,\_)$ is before $o$.
-Then, we can see that $o$ is before $\textit{cal}(\textit{rm},a)$.
-
-    If $\textit{cal}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$, then $\textit{rm}(b) <_{\textit{hb}} \textit{rm}(a)$.
-Let $h' = h \vert_{ \{ a,a_1,b \} }$.
-It is not hard to see that $h'$ can not linearizable to a sequence that satisfies the first in first our property, and this contradicts to our assumption that each single-priority history satisfies the first in first out property.
-Therefore, $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
-Similarly as the case of reason $1$, we can see that $\textit{ret}(\textit{rm},b)$ is before $\textit{ret}(\textit{rm},a)$.
-
-    According to above discussion, there are three possible enumeration of operations of $a$, $a_1$ and $b$, as shown in \figurename~\ref{fig:history enumeration 3 for Rpr2}, \figurename~\ref{fig:history enumeration 4 for Rpr2} and \figurename~\ref{fig:history enumeration 5 for Rpr2}.
-Here we explicitly draw the leftmost gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$ as $o'$.
-Since the position of $\textit{cal}(\textit{put},b,\_)$, $\textit{ret}(\textit{put},b)$, $\textit{ret}(\textit{put},a_1,\_)$ and $\textit{cal}(\textit{put},a,\_)$ do not influence the correctness, we can simply ignore it.
-
-\item[-] Reason $5$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$:
-
-    Since $a_1 <_{\textit{ob}}^A b$, $\textit{ret}(\textit{put},a_1,\_)$ is before $\textit{call}(\textit{put},b,\_)$.
-Since $\textit{call}(\textit{put},b,\_)$ is before $o$, we can see that $\textit{ret}(\textit{put},a_1,\_)$ is before $o$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $o$ is before $\textit{cal}(\textit{rm},a)$: Then $o$ is obviously before $\textit{ret}(\textit{rm},a)$.
-Since $a <_{\textit{ob}}^B a_1$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},a_1)$.
-Then we can see that, $o$ is before $\textit{cal}(\textit{rm},a_1)$, and remember that $a_1 <_{\textit{ob}}^A b$.
-Then we can goto the case of reason $1$ and treat $a_1$ as $a$.
-Therefore, we can safely ignore this.
-
-    \item[-] If $o$ is before $\textit{cal}(\textit{put},a,\_)$: Since $\textit{rm}(a)$ does not happen before $\textit{put}(a,\_)$, we can see that $\textit{cal}(\textit{put},a,\_)$ is before $\textit{ret}(\textit{rm},a)$, and then $o$ is before $\textit{ret}(\textit{rm},a)$.
-Then similarly as above case, we can see that $o$ is before $\textit{cal}(\textit{rm},a_1)$, and $a_1 <_{\textit{ob}}^A b$.
-Then we can goto the case of reason $1$ and treat $a_1$ as $a$.
-Therefore, we can safely ignore this.
-    \end{itemize}
-
-\item[-] Reason $6$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$:
-
-    Since $a_2 <_{\textit{ob}}^B b$, $\textit{rm}(a_2) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_2)$ is before $\textit{cal}(\textit{rm},b)$.
-%Since $\textit{rm}(a_2)$ does not happen before $\textit{put}(a_2,\_)$, $\textit{cal}(\textit{put},a_2,\_)$ is before $\textit{ret}(\textit{rm},a_2)$.
-%Since $a_1 <_{\textit{ob}}^A a_2$, $\textit{ret}(\textit{put},a_1,\_)$ is before $\textit{cal}(\textit{put},a_2,\_)$.
-    Since $a_1 <_{\textit{ob}}^A a_2$, $\textit{put}(a_1,\_) <_{\textit{hb}} \textit{put}(a_2,\_)$, and $\textit{rm}(a_2)$ does not happen before $\textit{rm}(a_1)$, and $\textit{cal}(\textit{rm},a_1)$ is before $\textit{ret}(\textit{rm},a_2)$.
-Since $a <_{\textit{ob}}^B a_1$, $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(a_1)$, and $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},a_1)$.
-
-    Therefore, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$.
-Since $\textit{cal}(\textit{rm},b)$ is before $o$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$.
-Then it is easy to see that $o$ is before $\textit{cal}(\textit{put},a,\_)$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\_)$, and $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a,\_)$.
-This is impossible, and we can safely ignore this reason.
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-1.pdf}
-%\vspace{-10pt}
-  \caption{The first possible enumeration.}
-  \label{fig:history enumeration 1 for Rpr2}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-2.pdf}
-%\vspace{-10pt}
-  \caption{The second possible enumeration.}
-  \label{fig:history enumeration 2 for Rpr2}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-3.pdf}
-%\vspace{-10pt}
-  \caption{The third possible enumeration.}
-  \label{fig:history enumeration 3 for Rpr2}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-4.pdf}
-%\vspace{-10pt}
-  \caption{The forth possible enumeration.}
-  \label{fig:history enumeration 4 for Rpr2}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-5.pdf}
-%\vspace{-10pt}
-  \caption{The fifth possible enumeration.}
-  \label{fig:history enumeration 5 for Rpr2}
-\end{figure}
-
-Given a data-differentiated history $h$, two actions $\textit{act}_1$, $\textit{act}_2$ of items of priority $\textit{pri}$, and assume that $\textit{act}_1$ is before $\textit{act}_2$ in $h$.
-We say that $\textit{act}_1$, $\textit{act}_2$ is covered by items $d_1,\ldots,d_m$ in $h$, if the priorities of $d_1,\ldots,d_m$ is not $\textit{pri}$, and
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{ret}(\textit{put},d_m,\_)$ is before $\textit{act}_1$,
-
-\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,
-
-\item[-] $\textit{act}_2$ is before $\textit{cal}(\textit{rm},d_1)$.
-\end{itemize}
-
-According to the definition of gap-point, in \figurename~\ref{fig:history enumeration 1 for Rpr2}, \figurename~\ref{fig:history enumeration 2 for Rpr2}, \figurename~\ref{fig:history enumeration 3 for Rpr2}, \figurename~\ref{fig:history enumeration 4 for Rpr2} or \figurename~\ref{fig:history enumeration 5 for Rpr2}, the location of $o$ and $o'$ is equivalent to that, there exists items $d_1,\ldots,d_m$ in $h$ in $h$, such that $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ is covered by $d_1,\ldots,d_m$, $\textit{cal}(\textit{rm},b)$ is before $\textit{ret}(\textit{put},d_m,\_)$ in $h$, and $\textit{cal}(\textit{rm},d_1)$ is before $\textit{ret}(\textit{rm},a)$ in $h$.
-We say that such $d_1,\ldots,d_m$ constitute the rightmost gap of $\textit{put}(b,\_)$ and $\textit{rm}(b)$.
-
-An automaton $\mathcal{A}_{\textit{Rpr2-1}}$ is given in \figurename~\ref{fig:automata for first enumeration of Rpr2}, and it is constructed for the first enumeration in \figurename~\ref{fig:history enumeration 1 for Rpr2}.
-In \figurename~\ref{fig:automata for first enumeration of Rpr2}, we renames the items that covers $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ into $d$, and rename the renaming items into $e$.
-Here let $c = \textit{cal}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{cal}(\textit{rm},e), \textit{ret}(\textit{rm},e)$, $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_1.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr2-1}}$}
-  \label{fig:automata for first enumeration of Rpr2}
-\end{figure}
-
-An automaton $\mathcal{A}_{\textit{Rpr2-2}}$ is given in \figurename~\ref{fig:automata for second enumeration of Rpr2}, and it is constructed for the second enumeration in \figurename~\ref{fig:history enumeration 2 for Rpr2}.
-In \figurename~\ref{fig:automata for second enumeration of Rpr2}, $c_1$, $c_2$, $c_3$ and $c_4$ is same as that in \figurename~\ref{fig:automata for first enumeration of Rpr2}.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_2.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr2-2}}$}
-  \label{fig:automata for second enumeration of Rpr2}
-\end{figure}
-
-For the third enumeration in \figurename~\ref{fig:history enumeration 3 for Rpr2}.
-Since we want to ensure that $a$ and $b$ are putted only once, we need to explicitly record the positions of $\textit{cal}(\textit{put},a,p)$ and $\textit{cal}(\textit{put},b,p)$.
-Since the positions of $\textit{cal}(\textit{put},a,p)$ and $\textit{cal}(\textit{put},b,p)$ are not fixed, there are finite possible cases to consider, as shown below:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{rm},b)$ and before $\textit{cal}(\textit{rm},a)$: There are two possible positions of $\textit{cal}(\textit{put},a,p)$: (1) before $\textit{cal}(\textit{rm},a_1)$, and (2) after $\textit{cal}(\textit{rm},a_1)$, and before $\textit{ret}(\textit{put},a)$.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$ and before $\textit{cal}(\textit{rm},b)$: same as above case.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{put},a_1,\_)$ and before $\textit{ret}(\textit{rm},a_1)$: same as above case.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{put},a)$ and before $\textit{cal}(\textit{put},a_1,p)$: same as above case.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$: There are three possible positions of $\textit{cal}(\textit{put},a,p)$: (1) after $\textit{cal}(\textit{put},b,p)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{cal}(\textit{put}b,p)$, and (3) before $\textit{call}(\textit{rm},a_1)$.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is before $\textit{cal}(\textit{rm},a_1)$: There are three possible positions of $\textit{cal}(\textit{put},a,p)$: (1) after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{cal}(\textit{put},b,p)$ and before $\textit{cal}(\textit{rm},a_1)$, and (3) before $\textit{cal}(\textit{put},b,p)$.
-\end{itemize}
-
-Therefore, there are fourteen possible cases that satisfy the third enumeration in \figurename~\ref{fig:history enumeration 3 for Rpr2}.
-For each case, we construct an finite automaton.
-Let $\textit{Aut}_{\textit{Rpr2-3}}$ be the set of finite automata that is constructed for above fourteen cases.
-For example, for the case $\textit{ca}_1$ when $\textit{cal}(\textit{put},a,p)$ is before $\textit{cal}(\textit{rm},a_1)$, $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$, and $\textit{cal}(\textit{put},b,p)$ is before $\textit{cal}(\textit{rm},b)$, we construct a finite automaton in \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}.
-In \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}, let $c$ and $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$ the same as that in \figurename~\ref{fig:automata for first enumeration of Rpr2}.
-Let $c_2 = c_1 + \textit{ret}(\textit{put},a_1)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$, $c_4 = c_3 + \textit{ret}(\textit{rm},d)$, and $c_5 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{put},a_1) + \textit{ret}(\textit{rm},d)$.
-The other finite automata in $\textit{Aut}_{\textit{Rpr2-3}}$ can be similarly constructed.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_3.pdf}
-%\vspace{-10pt}
-  \caption{Automaton for $\textit{ca}_1$}
-  \label{fig:automata for ca1 of third enumeration of Rpr2}
-\end{figure}
-
-Similarly, we construct sets $\textit{Aut}_{\textit{Rpr2-4}}$ and $\textit{Aut}_{\textit{Rpr2-5}}$ of finite automata for the forth enumeration in \figurename~\ref{fig:history enumeration 4 for Rpr2} and the fifth enumeration in \figurename~\ref{fig:history enumeration 5 for Rpr2}, respectively.
-
-\begin{restatable}{lemma}{Rpr2IsCoRegular}
-\label{lemma:Rpr2 is co-regular}
-$R_{\textit{pr2}}$ is co-regular.
-\end{restatable}
-
-\begin {proof}
-
-Let $\textit{AutS-R}_{\textit{pr2}} = \{ \mathcal{A}_{\textit{Rpr2-1}}, \mathcal{A}_{\textit{Rpr2-2}} \} \cup \textit{Aut}_{\textit{Rpr2-3}} \cup \textit{Aut}_{\textit{Rpr2-4}} \cup \textit{Aut}_{\textit{Rpr2-5}}$.
-We need to prove that:
-
-\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr2}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr2}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr2}}$.
-
-By Lemma \ref{lemma:Rpr2 as gap-point and ob order}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:
-
-\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr0}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, such that $\textit{last}(h') = R_{\textit{pr2}}$, $a$ and $b$ has maximal priority in $h'$, $a <_{\textit{ob}} b$ in $h'$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
-
-\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\textit{AutS-R}_{\textit{pr2}}$.
-By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$.
-Since $h_1$ is accepted by $\textit{AutS-R}_{\textit{pr2}}$, let $x$ and $y$ be the items that are renamed into $b$ and $a$ by $r$, respectively, and let $d_1,\ldots,d_m$ be the items that are renamed into $d$ by $r$.
-
-let $h'' = h_2 \vert_{ \{ x,y,d_1,\ldots,d_m \} }$.
-It is obvious that $\textit{last}(h'') = R_{\textit{pr2}}$.
-According to our construction of automata in $\textit{AutS-R}_{\textit{pr2}}$, it is not hard to see that $x$ and $y$ has maximal priority in $h_2$, $y <_{\textit{ob}} x$ in $h''$, and the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$ in $h''$.
-%By Lemma \ref{lemma:Rpr2 as gap-point and ob order}, we know that $h''$ is not linearizable to $\textit{MR}_{\textit{pr2}}$.
-
-\noindent The $\textit{if}$ direction: Assume that there exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr2}}$, $a'$ and $b'$ has maximal priority in $h'$, $a' <_{\textit{ob}} b'$ in $h'$, and the rightmost gap-point of $\textit{put}(b',\_)$ and $\textit{rm}(b')$ is before $\textit{cal}(\textit{put},a',\_)$ or $\textit{cal}(\textit{rm},a')$ in $h''$.
-By data-independence, we can obtain history $h_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the items $d_1,\ldots,d_m$ that constitute the rightmost gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, we rename them into $d$, and (3) rename the other items into $e$.
-It is easy to see that $\textit{last}(h_1) = R_{\textit{pr2}}$, $a$ and $b$ has maximal priority in $h_1$, $a <_{\textit{ob}} b$ in $h_1$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
-By Lemma \ref{lemma:five enumeration is enough for Rpr2}, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists) and $a_2$ (if exists).
-Then
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $a <_{\textit{ob}} b$ because of the first enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr2-1}}$.
-
-\item[-] If $a <_{\textit{ob}} b$ because of the second enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr2-2}}$.
-
-\item[-] If $a <_{\textit{ob}} b$ because of the third enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-3}}$.
-
-\item[-] If $a <_{\textit{ob}} b$ because of the forth enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-4}}$.
-
-\item[-] If $a <_{\textit{ob}} b$ because of the fifth enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-5}}$.
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\subsection{Co-Regular of $R_{\textit{pr3}}$}
-\label{subsec:co-regular of Rpr3}
-
-\begin{restatable}{lemma}{Rpr3IsAlwaysCoRegular}
-\label{lemma:Rpr3 is always co-regular}
-Given a data-differentiated history $h$, if $\textit{last}(h) = R_{\textit{pr3}}$, then $h \sqsubseteq \textit{MR}_{\textit{pr3}}$.
-\end{restatable}
-
-\begin {proof}
-
-Since $\textit{last}(h) = R_{\textit{pr3}}$, there exists item $x$ with $\textit{pri}_x$ in $h$, such that $h \vert_{ \{ x \}} = \textit{put}(x,\textit{pri}_x)$, and $\textit{pri}_x$ is larger than any priority of $h$.
-
-For each operation, we locate its linearization point just after its call action.
-In such way we obtain a sequence $l$ of operations, and it is easy to see that $h \sqsubseteq l$.
-It is obvious that $l$ contains $\textit{put}(x,\textit{pri})$, and then $l \in \textit{MR}_{\textit{pr3}}$.
-This completes the proof of this lemma.
-\qed
-\end {proof}
-
-\subsection{Co-Regular of $R_{\textit{pr4}}$}
-\label{subsec:co-regular of Rpr4}
-
-\begin{restatable}{lemma}{Rp4AsHappenBefore}
-\label{lemma:Rpr4 as happen before}
-Given a data-differentiated history $h$.
-There exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr4}}$, and $h'$ does not linearizable to $\textit{MR}_{\textit{pr4}}$, if and only if there exists $h'' \in \textit{proj}(h)$, $\textit{last}(h'') = R_{\textit{pr4}}$, $x$ and $y$ has maximal priority in $h''$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,\_) <_{\textit{hb}} \textit{put}(y,\_)$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{if}$ direction, let $h_1 = h \vert_{ \{ x,y \} }$. It is easy to see that $\textit{last}(h') = R_{\textit{pr4}}$ and $h_1$ does not linearizable to $\textit{MR}_{\textit{pr4}}$.
-
-To prove the $\textit{only if}$ direction, we prove its contrapositive.
-Assume that for each $h'' \in \textit{proj}(h)$, such that $\textit{last}(h'') = R_{\textit{pr4}}$, for each $x$ and $y$ has maximal priority in $h''$, if $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, then $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$.
-We need to prove that, for each $h' \in \textit{proj}(h)$, if $\textit{last}(h') = R_{\textit{pr4}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr4}}$.
-
-Given such $h'$, let $x_1,\ldots,x_m$ be the set of items with maximal priority and has unmatched $\textit{put}$ operations, let $y_1,\ldots,y_n$ be the set of items with maximal priority and has matched $\textit{put}$ and $\textit{rm}$ operations.
-It is easy to see that, for each $x_i$ and $y_j$, $\textit{cal}(\textit{put},y_j)$ is before $\textit{ret}(\textit{put},x_i,\_)$.
-Then we locate the linearization points of $h'$ as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] For each $x_i$, locate the linearization point of $\textit{put}(x_i,\_)$ just before its return action.
-
-\item[-] For each $y_j$, locate the lineariztion point of $\textit{put}(y_j,\_)$ jest after its call action.
-
-\item[-] For other operations, locate their linearization points at an arbitrary location after its call action and before its return action.
-\end{itemize}
-
-Let $l$ be the sequence of linearization points.
-It is easy to see that $h \sqsubseteq l$, and $l \in \textit{MR}_{\textit{pr4}}$.
-Here the $\textit{itm}$ in $R_{\textit{pr4}}$ is chosen to be $\textit{ret}(\textit{put},x_{\textit{ind}},\_)$, where the return action of $x_{\textit{ind}}$ is later than return actions of other $x_i$.
-This completes the proof of this lemma.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr4IsAlwaysCoRegular}
-\label{lemma:Rpr4 is always co-regular}
-Given a data-differentiated history $h$, if $\textit{last}(h) = R_{\textit{pr4}}$, then $h \sqsubseteq \textit{MR}_{\textit{pr4}}$.
-\end{restatable}
-
-\begin {proof}
-
-According to Lemma \ref{lemma:Rpr4 as happen before}, if $\textit{last}(h) = R_{\textit{pr4}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr4}}$, then there exists $x$ and $y$ with maximal priority in $h$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,\_) <_{\textit{hb}} \textit{put}(y,\_)$.
-Let $h_1 = h \vert_{ \{ x,y \} }$.
-It is obvious that $\textit{transToQueue}(h_1)$ does not linearizable to queue.
-This contradicts the assumption that every single-priority history can be ``linearizable to queue'', and thus, we can safely ignore this case.
-\qed
-\end {proof}
-
-\subsection{Co-Regular of $R_{\textit{pr5}}$}
-\label{subsec:co-regular of Rpr5}
-
-$R_{\textit{pr5}} \equiv (u \cdot v \in \textit{PQueue}) \wedge (\textit{matched}(u,\textit{put},\textit{rm}) ) \Rightarrow (u \cdot \textit{rmEmpty} \cdot v \in \textit{PQueue})$
-
-Similar as the notion of gap in \cite{Bouajjani:2015}, we propose the notion of gap of $\textit{em}(\textit{empty})$ operations.
-
-\begin{definition}[gap for $\textit{em}(\textit{empty})$ operations]\label{def:gap for rmEmpty operations}
-
-Let $h$ be a data-differentiated history and $o = \textit{rm}(\textit{empty})$ of $h$.
-We say that $h$ has a gap on operation $o$, if there is a partition of the operations of $h$ into $L$ and $R$ satisfying:
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $L$ has matched $\textit{put}$ and $\textit{rm}$ operations.
-
-\item[-] no operation of $R$ happens before an operation of $L$ or $o$.
-
-\item[-] $o$ does not happen before operations of $L$
-\end{itemize}
-\end{definition}
-
-Compared to the gap on $\textit{put}$ and $\textit{rm}$ operations, here we do not consider the priority of operations in $L$ or $R$.
-
-\begin{restatable}{lemma}{GapEqualsLinforRpr5}
-\label{lemma:Gap Equals Lin for Rpr5}
-A data-differentiated history $h$ has a projection $h'$, such that $\textit{last}(h') = R_{\textit{pr5}}$ and $h'$ does not linearizable to $\textit{MR}_{\textit{pr5}}$, if and only if there exists a projection $h''$ of $h$, such that $\textit{last}(h'') = R_{\textit{pr5}}$, there exists $o = \textit{rm}(\textit{empty})$ in $h''$, and there is no gap of $o$ in $h''$.
-\end{restatable}
-
-\begin {proof}
-
-Let us prove the $\textit{only if}$ direction by contradiction.
-Assume that there exists such $h'$, and for each $\textit{rm}(\textit{empty})$ in $h'$, there is a gap of $o$ in $h''$.
-Since $\textit{last}(h') = R_{\textit{pr5}}$, there is exists at least $\textit{rm}(\textit{empty})$ operation in $h$, and let $o$ be arbitrary one of them.
-Since there is a gap of $o$ in $h$, let $L$ and $R$ be the sets of the gap.
-Let $u$ be a sequence of operations of $L$, and $u$ is generated from $h$ by locating linearization points of each operation in $L$ just after its call action.
-Let $v$ be a sequence of operations of $R$, and $v$ is generated from $h$ by locating linearization points of each operation in $R$ just before its return action.
-It is easy to see that $h \sqsubseteq l = u \cdot o \cdot v$, and $l \in \textit{MR}_{\textit{pr5}}$.
-This contradicts that $h'$ does not linearizable to $\textit{MR}_{\textit{pr5}}$.
-
-To prove the $\textit{if}$ direction, let $h_1$ be a history generate from $h''$ by discarding all other $\textit{rm}(\textit{Empty})$ operations.
-By assumption, it is obvious that there is no gap of $o$ in $h_1$.
-Assume by contradiction that there exists $l'$, such that $h_1 \sqsubseteq l' \in \textit{MR}_{\textit{pr5}}$.
-It is easy to see that $l' = u' \cdot o \cdot v'$, such that $u'$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
-Let $L$ be the set of operations of $u'$, and $R$ be the set of operations in $v'$.
-Since $h_1 \sqsubseteq l'$, it is easy to see that $L$ and $R$ is a gap of $o$ in $h_1$, contradicts that there is no gap of $o$ in $h_1$.
-\qed
-\end {proof}
-
-Similar as the notion of gap in \cite{Bouajjani:2015}, we propose the notion left-right constraint of $\textit{rm}(\textit{empty})$ operation.
-
-\begin{definition}[left-right constraint for $\textit{rm}(\textit{empty})$ operation]\label{def:left-right constraint for rmEmpty operation}
-Given a data-differentiated history $h$, and $o = \textit{rm}(\textit{empty})$ of $h$. The left-right constraint of $o$ is the graph $G$ where:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] the nodes are the items of $h$ or $o$, to which we add a node,
-
-\item[-] there is an edge from item $d_1$ to $o$, if $\textit{put}(d_1,\_)$ happens before $o$,
-
-\item[-] there is an edge from $o$ to item $d_1$, if $o$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$,
-
-\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$.
-\end{itemize}
-\end{definition}
-
-Compared to the left-right constraint of $\textit{put}$ and $\textit{rm}$ operations, here we do not consider the priority.
-We also do not consider other $\textit{rm}(\textit{empty})$ operations in $h$.
-
-Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$.
-Let $\textit{LMSet}_1(h,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some item, either $\textit{op}$ happens before $o$ in $h$, or there is an operation $\textit{op}'$ with the same item of $\textit{op}$, such that $\textit{op}'$ happens before $o$ in $h$ $\}$.
-For each $i \geq 1$, let $\textit{LMSet}_{\textit{i+1}}(h,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some item, $\textit{op}$ is not an operation of items in $\textit{LMSet}_k(h,o)$ for each $k \leq i$, either $\textit{op}$ happens before some operation $o' \in \textit{LMSet}_i(h,o)$ in $h$, or there is an operation $\textit{op}''$ with the same item of $o$ and $\textit{op}''$ happens before some operation $o' \in \textit{LMSet}_i(h,o)$ in $h$ $\}$.
-We can see that $\textit{LMSet}_i(h,o) \cap \textit{LMSet}_j(h,o) = \emptyset$ for any $i \neq j$.
-Let $\textit{LMSet}(h,o) = \textit{LMSet}_1(h,o) \cup \textit{LMSet}_2(h,o) \cup \ldots$.
-
-\begin{restatable}{lemma}{RmEmptyDoesNotHappenBeforeLMSetForRpr5}
-\label{lemma:RmEmpty does not happen before LMSet for Rpr5}
-Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$.
-Let $G$ be the graph representing the left-right constraint of $o$.
-Assume that $G$ has no cycle going through $o$.
-Then, $o$ does not happen before any operation in $\textit{LMSet}(h,o)$.
-\end{restatable}
-
-This Lemma can be similarly proved as Lemma \ref{lemma:Rmx does not happen before LMSet for Rpr1}.
-
-\begin{restatable}{lemma}{LMSetHasMatchedPutandRmOperationsForRpr5}
-\label{lemma:LMSet has matched put and rm operations for Rpr5}
-Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$.
-Let $G$ be the graph representing the left-right constraint of $o$.
-Assume that $G$ has no cycle going through $o$.
-Then, $\textit{LMSet}(h,o)$ contains only matched $\textit{put}$ and $\textit{rm}$ operations.
-\end{restatable}
-
-This Lemma can be similarly proved as Lemma \ref{lemma:LMSet has matched put and rm operations}.
-
-\begin{restatable}{lemma}{GapEqualsConstraintforRpr5}
-\label{lemma:Gap Equals Constraint for Rpr5}
-Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$.
-Let $G$ be the graph representing the left-right constraint of $o$.
-There is a gap on $o$, if and only if $G$ has no cycle going through $o$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{only if}$ direction, assume that there is a gap on $o$, and let it be $L$ and $R$.
-Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow o \rightarrow d_1$ in $G$.
-Since $o \rightarrow d_1$, either $o$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$.
-Since $L$ contains matched $\textit{put}$ and $\textit{rm}$ operations, it is easy to see that operations of $d_1$ is in $R$.
-Since $d_1 \rightarrow d_2$, we can see that $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$.
-Since operations of $d_1$ is in $R$ and $L$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we can see that operations of $d_2$ is in $R$.
-Similarly, we can prove that operations of $d_3,\ldots,d_m$ is in $R$.
-Since $d_m \rightarrow o$, $\textit{put}(d_m,\_)$ happens before $o$, and $\textit{put}(d_m,\_) \in L$.
-This contradicts that operations of $d_m$ is in $R$.
-This completes the proof of the $\textit{only if}$ direction.
-
-To prove the $\textit{if}$ direction, assume that $G$ has no cycle going through $o$. Let $O_h$ be the set of operations of $h$. Let $O_L = \textit{LMSet}(h,o)$, $O_R = O_h \setminus O_L$.
-
-By Lemma \ref{lemma:LMSet has matched put and rm operations for Rpr5}, we can see that $O_L = \textit{LMSet}(h,o)$ contains only matched $\textit{put}$ and $\textit{rm}$ operations.
-It remains to prove that $L$ and $R$ is a gap of $o$.
-We prove this by showing that all the following cases are impossible:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Case $1$: Some operation $o_r \in R$ happens before $o$. Then we know that $o_r \in \textit{LMSet}(h,o)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $2$: Some operation $o_r \in R$ happens before some operation $o_l$. Then we know that $o_r \in \textit{LMSet}(h,o)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $4$: $o$ happens before some $o_l \in \textit{LMSet}(h,o) = L$. By Lemma \ref{lemma:RmEmpty does not happen before LMSet for Rpr5}, we know that this is impossible.
-\end{itemize}
-
-This completes the proof of the $\textit{if}$ direction.
-
-\qed
-\end {proof}
-
-Let us begin to represent an automaton that is used for capture the case that, in a sub-history $h'$ of a history $h$, $\textit{last}(h')=R_{\textit{pr5}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr5}}$, and the reason is that there is no gap on a $\textit{rm}(\textit{empty})$ operation in $h'$.
-The automaton is $\mathcal{A}_{\textit{Rpr5}}$, which is given in \figurename~\ref{fig:automata for Rpr5}.
-In \figurename~\ref{fig:automata for Rpr5}, let $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d), \textit{ret}(\textit{rm},d)$, $c_1 = c + \textit{cal}(\textit{put},b,\textit{anPri})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c + \textit{ret}(\textit{rm},b)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr5.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr5}}$}
-  \label{fig:automata for Rpr5}
-\end{figure}
-
-Given a data-differentiated history $h$, we say that $o = \textit{rm}(\textit{empty})$ in $h$ is covered by items $d_1,\ldots,d_m$ in $h$, if
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{put}(d_m,\_)$ happens before $o$,
-
-\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,
-
-\item[-] $o$ happens before $\textit{rm}(d_1)$, or $\textit{rm}(d_1)$ does not exists in $h$
-\end{itemize}
-
-According to the definition of left-right constraint for $o$, in a data-differentiated history $h$, there is a cycle going through $o$, if and only if there exists items $d_1,\ldots,d_m$, such that $o$ is covered by $d_1,\ldots,d_m$.
-
-\begin{restatable}{lemma}{Rpr5IsCoRegular}
-\label{lemma:Rpr1 5s co-regular}
-$R_{\textit{pr5}}$ is co-regular.
-\end{restatable}
-
-\begin {proof}
-
-We need to prove that:
-
-\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\mathcal{A}_{\textit{Rpr5}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr5}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr5}}$.
-
-By Lemma \ref{lemma:Gap Equals Lin for Rpr5} and Lemma \ref{lemma:Gap Equals Constraint for Rpr5}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:
-
-\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\mathcal{A}_{\textit{Rpr5}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h''$ is a projection of $h$, such that $\textit{last}(h'') = R_{\textit{pr5}}$, $o = \textit{rm}(\textit{empty})$ is in $h''$, and $o$ is covered by some items $d_1,\ldots,d_m$ in $h''$.
-
-\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\mathcal{A}_{\textit{Rpr5}}$.
-By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$.
-Let $d_1,\ldots,d_m$ be the items in $h_2$ such that $r(d_i)=d$ for each $1 \leq i \leq m$.
-Let $h'' = h_2 \vert_{ \{ o, d_0, d_1, \ldots, d_m \} }$.
-It is obvious that $\textit{last}(h'') = R_{\textit{pr5}}$.
-It is easy to see that $o$ is covered by $d_1,\ldots,d_m$.
-
-\noindent The $\textit{if}$ direction: Assume that there exists such $h$, $h''$, $o$ and $d_1,\ldots,d_m$.
-Then, let $h_1$ be obtained from $h$ by renaming $d_1,\ldots,d_m$ into $b$ and renaming other items into $d$.
-We can prove that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr5}}$.
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\section{Relate Queue and Stack with Priority Queue}
-\label{sec:relate queue and stack with priority queue}
-
-\subsection{Relate Queue with Priority Queue}
-\label{subsec:relate queue with priority queue}
-
-Given a history $h_q$ of queue operations, let $\textit{UniPriTrans}(h_q)$ be a history of priority queue, which is generated as follows: transform each $\textit{call}(\textit{enq},a)$, $\textit{ret}(\textit{enq},a)$, $\textit{call}(\textit{deq},a)$ and $\textit{ret}(\textit{deq},a)$ of $h_q$ into $\textit{call}(\textit{put},a,1)$, $\textit{ret}(\textit{put},a)$, $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$.
-
-\begin{restatable}{lemma}{RelateQueuewithPQ}
-\label{lemma:relate queue with priority queue}
-Given a history $h_q$ of queue operations, $h_q$ is linearizable to queue, if and only if $\textit{UniPriTrans}(h_q) \sqsubseteq \textit{PQueue}$.
-\end{restatable}
-
-\begin {proof}
-
-This is obvious, since we already ensure that for each single-priority history $h$, $\textit{transToQueue}(h)$ is linearizable to queue. \qed
-\end {proof}
-
-\subsection{Relate Stack with Priority Queue}
-\label{subsec:relate stack with priority queue}
-
-\section{Conclusion and Future Work}
-\label{sec:conclusion and future work}
-
-Conclusion
+\input{rules-PQ}
 
@@
 
+\bibitem{Bouajjani:2015POPL}
+Bouajjani, A., Emmi, M., Enea, C., Hamza, J.:
+\newblock Tractable Refinement Checking for Concurrent Objects.
+\newblock In: {\em POPL'15}. pp.651--662. (2015)
+
 \bibitem{Wolper:1986}
@@
 
+\bibitem{Henzinger:2013}
+Henzinger A. T., Sezgin A., Vafeiadis, V.:
+\newblock Aspect-Oriented Linearizability Proofs.
+\newblock In {\em CONCUR'13}. pp.242--256. (2013)
+
+\bibitem{Abdulla:2013}
+Abdulla, P. A., Haziza, F., Hol¨ªk, L., Jonsson, B., Rezine, A.:
+\newblock An integrated specification and verification technique for highly concurrent data
+structures.
+\newblock In {\em TACAS'13}. pp.324--338. (2013)
+
 \end{thebibliography}
@@
 
+\input{appendix}
+
 \end{document}

### 2017-02-20 appendix.tex
move other part into appendix from Appendix B

@@
 \end {proof}
+
+\section{Inductive Rules for Priority Queues}
+\label{sec:inductive rules for priority queues}
+
+Given a sequence $l$ of possibly $\textit{put}$, $\textit{rm}$ and $\textit{rmEmpty}$ operations, we say that $l$ contains unmatched $\textit{rm}$ operations, if the number of $\textit{rm}(d)$ operations is larger than the number of $\textit{put}(d,\_)$ operations for some $d$ in $l$.
+By data-independence, we can detect the existence of unmatched $\textit{rm}$ operations by automata.
+A automaton $\mathcal{A}_{\textit{unmRm1}}$ is given in \figurename~\ref{fig:automata for unmatched rm1}, where $c_1=\textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm),\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$, $c_2=c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c_1 + \textit{cal}(\textit{put},b,\textit{anyPri}) + \textit{ret}(\textit{rm},b)$.
+Here $\textit{anyPri}$ is a specific predicates that is satisfied by any priority.
+$\mathcal{A}_{\textit{unmRm1}}$ is used to detect the existence of a item that has been removed but has never been inserted.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_RM1.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{unmRm1}}$}
+  \label{fig:automata for unmatched rm1}
+\end{figure}
+
+An automaton $\mathcal{A}_{\textit{unmRm2}}$ is given in \figurename~\ref{fig:automata for unmatched rm2}, where $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm),$ $\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$, and $c_1 = c_1 + \textit{ret}(\textit{rm},b)$.
+$\mathcal{A}_{\textit{unmRm1}}$ is used to detect the existence of a item that has been removed more than once but has been inserted once.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_RM2.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{unmRm2}}$}
+  \label{fig:automata for unmatched rm2}
+\end{figure}
+
+The following lemma states that we can detect the existence of unmatched $\textit{rm}$ by $\mathcal{A}_{\textit{unmRm1}}$ and $\mathcal{A}_{\textit{unmRm2}}$.
+
+\begin{restatable}{lemma}{DetectUnmatchedRmByAutomata}
+\label{lemma:detect unmatched rm by automata}
+
+Given a data-independent $\mathcal{I}$, there is some unmatched $\textit{rm}$ in some sequence of $\mathcal{I}$, if and only if $( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm1}} \neq \emptyset) \vee ( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm2}} \neq \emptyset )$.
+\end{restatable}
+
+\begin {proof}
+
+For the $\textit{if}$ direction, assume that $( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm1}} \neq \emptyset) \vee ( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm2}} \neq \emptyset )$.
+Then, there exists a sequence $l \in \mathcal{I}$ and $l$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$ or $\mathcal{A}_{\textit{unmRm2}}$.
+If $l$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$, then $l$ contains at least one $\textit{rm}(b)$ operation and no $\textit{put}(b,\_)$ operation; Else, if $l$ is accepted by $\mathcal{A}_{\textit{unmRm2}}$, then $l$ contains at least two $\textit{rm}(b)$ operations and one $\textit{put}(b,\_)$ operation.
+In both cases, $l$ contains unmatched $\textit{rm}$ operations.
+
+To prove the $\textit{only if}$ direction assume that there exists a sequence $l \in \mathcal{I}$ that contains unmatched $\textit{rm}$ operations.
+By data-independence, there exists data-differentiated sequence $l'$ and renaming function $r$, such that $l=r(l')$.
+We need to prove that in $l'$, one of the following cases holds: (1) $l'$ contains $\textit{rm}(d_1)$ for some $d_1$ and does not contain any $\textit{put}(d_1,\_)$, and (2) $l'$ contains one $\textit{put}(d_1,\_)$ operation and more than one $\textit{rm}(d_1)$ operations.
+We prove this by contradiction.
+Assume by contradiction that both cases do not hold.
+Then, for each value $d$ of $l'$, there is either one $\textit{put}(d,\_)$ and one $\textit{rm}(d)$ in $l'$, or one $\textit{put}(d,\_)$ and no $\textit{rm}(d)$ in $l'$.
+It is obvious that such $l'$ does not contain unmatched $\textit{rm}$ operation, and renaming a sequence that does not contain unmatched $\textit{rm}$ we can only obtain another sequence that does not contain unmatched $\textit{rm}$.
+This contradicts that $l=r(l')$ has unmatched $\textit{rm}$.
+
+Let $l''$ be obtained from $ll$ by renaming $d_1$ to $b$ and renaming other value to $a$.
+If the first case holds, $l''$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$; Otherwise, the second case holds, and $l''$ is accepted by $\mathcal{A}_{\textit{unmRm2}}$.
+\qed
+\end {proof}
+
+Therefore, from now on, we can safely assume that there are no unmatched $\textit{rm}$ operations.
+
+\section{Step-by-Step Linearizability of Priority Queues}
+\label{sec:step-by-step linearizability of priority queues}
+
+The following lemma states that $\textit{PQueue}$ is closed under projection. This lemma holds since the predicates appears in inductive rules of $\textit{PQueue}$ are closed under projection.
+
+\begin{restatable}{lemma}{PQisClosedUnderProjection}
+\label{lemma:PQ is closed under projection}
+
+$\textit{PQueue}$ is closed under projection.
+
+\end{restatable}
+
+%When the context is clear, we shorten priority-value data-independence as data-independence.
+For priority queue, the notion of step-by-step linearizability is modified as follows:
+
+%\vspace{-6pt}
+\begin{definition}[step-by-step linearizability for priority queue]\label{def:step-by-step linearizability of priority queue}
+A data structure $S=R_1, \ldots, R_n$ is said be to step-by-step linearizable, if for any data-differentiated sequence $e$, if $e$ is linearizable w.r.t.
+$\textit{MR}_{\textit{pri}}$ ($1 \leq i \leq 4$) with witness $x$, we have:
+
+$e \setminus x \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket \Rightarrow e \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$
+
+\noindent if $e$ is linearizable w.r.t. $\textit{MR}_{\textit{pr5}}$ with witness $x$, we have:
+
+$e \setminus x \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket \Rightarrow e \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$
+\end{definition}
+
+Let $\textit{proj}(s)$ denote the projection of $s$ obtained by erasing all the operations of several items from $s$.
+Here we assume that each $\textit{rm}(\textit{empty})$ is differentiated from each other, and we can consider each $\textit{rm}(\textit{empty})$ as having a unique item.
+
+\begin{restatable}{lemma}{PQasMultiInMRpriforSequence}
+\label{lemma:PQ as multi in MRpri for sequence}
+$s \in \textit{PQueue}$, if and only if, $\forall s'$, if $s' \in \textit{proj}(s)$, then $s' \in \textit{MR}_{\textit{pri}}$ for some $0 \leq i \leq 5$.
+\end{restatable}
+
+\begin {proof}
+
+The \textit{only if} direction is a direct consequence of Lemma \ref{lemma:PQ is closed under projection}.
+
+To prove the $\textit{if}$ direction.
+The $\textit{if}$ direction holds trivially when $s = \epsilon$.
+When $s \neq \epsilon$.
+since $s \in \textit{proj}(s)$, we generate sequence $s_1$ from $s$ as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If $s$ has $\textit{rm}(\textit{empty})$ operations: $s_1$ is generated from $s$ by erasing some $\textit{rm}(\textit{empty})$ operation.
+
+\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in $S_1$ and unmatched $\textit{put}$ operations of some items in $S_2$: $s_1$ is generated from $s$ by erasing the unmatched $\textit{put}$ operation of items last putted in $S_2$.
+
+\item[-] If the maximal priority in $s$ is unmatched $\textit{put}$ operations of some items in $S_2$ and $\vert S_2 \vert =1$: $s_1$ is generated from $s$ by erasing this unmatched $\textit{put}$ operation.
+
+\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert \geq 2$: $s_1$ is generated from $s$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item which is last putted in $S$.
+
+\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert = 1$: $s_1$ is generated from $s$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item in $S$.
+\end{itemize}
+
+Similarly, we obtain $s_2$ from $s_1$, obtain $s_3$ from $s_2$, $\ldots$, until we obtain some $s_m = \epsilon$.
+It is obvious that $s_m \in \textit{PQueue}$, and by the five rules of $\textit{PQueue}$, it is easy to see that $s_{\textit{m-1}},\ldots,s_1,s \in \textit{PQueue}$.
+\qed
+\end {proof}
+
+\begin{restatable}{lemma}{PQasMultiInMRpriforHistory}
+\label{lemma:PQ as multi in MRpri for history}
+Given a data-differentiated history $h$, assume that $\textit{PQueue}$ is step-by-step linearizability.
+$h \sqsubseteq \textit{PQueue}$, if and only if, $\forall h'$, if $h' \in \textit{proj}(h)$, then $h' \sqsubseteq \textit{MR}_{\textit{pri}}$, where $i = \textit{last}(h')$.
+\end{restatable}
+
+\begin {proof}
+
+The \textit{only if} direction is a direct consequence of Lemma \ref{lemma:PQ is closed under projection}.
+
+To prove the $\textit{if}$ direction.
+Since $h \sqsubseteq \textit{PQueue}$, we can see that $h \sqsubseteq l \in \textit{PQueue}$.
+Each projection $h'$ of $h$ can be linearizable to some projection $l'$ of $l$.
+By Lemma \ref{lemma:PQ as multi in MRpri for sequence}, $l' \in \textit{MR}_{\textit{prj}}$ for some $0 \leq j \leq 5$.
+Since $\textit{last}$ of a sequence is uniquely determined, we can see that $l' \in \textit{MR}_{\textit{pri}}$, where $i = \textit{last}(h')$.
+
+The $\textit{if}$ direction holds trivially when $h = \epsilon$. When $h \neq \epsilon$. since $h \in \textit{proj}(h)$, we generate sequence $h_1$ from $h$ as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If $h$ has $\textit{rm}(\textit{empty})$ operations: Then $\textit{last}(h) = R_{\textit{pr5}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr5}}$.
+$h_1$ is generated from $h$ by erasing some $\textit{rm}(\textit{empty})$ operation.
+
+\item[-] If the maximal priority in $h$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in $S_1$ and unmatched $\textit{put}$ operations of some items in $S_2$: Then $\textit{last}(h) = R_{\textit{pr4}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr4}}$.
+$h_1$ is generated from $h$ by erasing a unmatched $\textit{put}(x,\_)$, such that $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$ for each $y \in S_2$.
+
+\item[-] If the maximal priority in $s$ is unmatched $\textit{put}$ operations of some items in $S_2$ and $\vert S_2 \vert =1$: Then $\textit{last}(h) = R_{\textit{pr3}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr3}}$.
+$h_1$ is generated from $h$ by erasing this unmatched $\textit{put}$ operation.
+
+\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert \geq 2$: Then $\textit{last}(h) = R_{\textit{pr2}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr2}}$.
+$h_1$ is generated from $h$ by erasing $\textit{put}(x,\_)$ and $\textit{rm}(x)$, such that for each $y \in S$, $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$, and $\textit{rm}(x)$ does not happen before $\textit{rm}(y)$.
+
+\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert = 1$: Then $\textit{last}(h) = R_{\textit{pr1}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr1}}$.
+$h_1$ is generated from $h$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item in $S$.
+\end{itemize}
+
+Similarly, we obtain $h_2$ from $h_1$, obtain $h_3$ from $h_2$, $\ldots$, until we obtain some $h_m = \epsilon$.
+Assume that for each $i$, when generating $h_i$ from $h_{\textit{i-1}}$, we erase operations of item $d_i$.
+It is obvious that $h_m = h_{\textit{m-1}} \setminus d_m \sqsubseteq \textit{PQueue}$, and since $h_{\textit{m-1}} \sqsubseteq \textit{MR}_{\textit{prj}_{\textit{m-1}}}$, where $\textit{last}(h_{\textit{m-1}}) = R_{\textit{prj}_{\textit{m-1}}}$, we can see that $h_{\textit{m-1}} \sqsubseteq \textit{PQueue}$.
+Similarly, we can prove that $h_{\textit{m-1}},\ldots,h_1,h \sqsubseteq \textit{PQueue}$.
+This completes the proof of the $\textit{if}$ direction.
+\qed
+\end {proof}
+
+To prove that $\textit{PQueue}$ is step-by-step linearizability, we use five lemmas for five rules of $\textit{PQueue}$.
+
+\begin{restatable}{lemma}{Rpr1isStepByStepLinearizability}
+\label{lemma:Rpr1 is step-by-step linearizability}
+
+$R_{\textit{pr1}}$ is step-by-step linearizability.
+
+\end{restatable}
+
+\begin {proof}
+
+Assume that $l$ matches $R_{\textit{pr1}}$ with witness $x$.
+Assume that  Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
+Then, let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
+Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
+
+According to $R_{\textit{pr1}}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where the $\textit{put}$ operations and $\textit{rm}$ operations in $u \cdot v$ are matched, and $\textit{pri}_x$ is larger than any priority in $u \cdot v \cdot w$.
+Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$.
+Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$ operations, we can see that $D_L \cap D_R = \emptyset$.
+Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
+Let $S_1$ be the set of all operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_{L1}$ be the shortest prefix of $l'_L$ that contains all operations of $S_1$, and let $l'_{L2}$ be the sequence such that $l'_L = l'_{L1} \cdot l'_{L2}$.
+It is obvious that if $\textit{put}(x,\textit{pri}_x)$ happens before an operation $o$ in $h$, then $o$ is in $l'_{L2}$.
+Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.
+
+Since priority queue is closed under projection (Lemma \ref{lemma:PQ is closed under projection}) and $u \cdot v$ contain matched $\textit{put}$ and $\textit{rm}$ operations, we know that $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$.
+Since $\textit{pri}_x$ is larger than any priority in $h$, it is easy to see that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$.
+Since the priority queue is empty after executing $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$, we know that $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R \in \textit{PQueue}$.
+
+It remains to prove that $h \sqsubseteq l''$.
+To prove this, we define a graph $G$ whose nodes are the operations of $h$ and there is an edge from operation $o_1$ to $o_2$, if one of the following case holds
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $o_1$ happens-before $o_2$ in h,
+
+\item[-] the method event corresponding to $o_1$ in $l''$ is before the one corresponding to $o_2$.
+\end{itemize}
+
+Assume there is a cycle in $G$.
+According the the property of interval order and the fact that the order of $l''$ is total, we know that there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
+Then, we can see
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If $o_1,o_2 \in l'_{L1}$, or $o_2 \in l'_{L1} \wedge o_1 \in l'_{L2}$, or $o_1,o_2 \in l'_{L2}$, or $o_1,o_2 \in L'_R$, then $l'$ contradicts with happen before relation of $h$.
+
+\item[-] If $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_{L1}$, then since $\textit{put}(x,\textit{pri}_x)$ happens before $o_2$, we know that $o_2 \in l'_{L2}$, which contradicts that $o_2 \in l'_{L1}$.
+
+\item[-] If $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L1}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L1}$, or $o_1 = \textit{rm}(x) \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 \in L'_R \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 = \textit{rm}(x)$, then $l$ contradicts with happen before relation of $h$.
+
+\item[-] If $o_1 \in l'_{L2} \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, then since $o_1$ happens before $\textit{put}(x,\textit{pri}_x)$, we know that $o_1 \in S_1$ and then $o_1 \in l'_{L1}$, which contradicts that $o_1 \in l'_{L2}$.
+\end{itemize}
+
+Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
+
+\end {proof}
+
+\begin{restatable}{lemma}{Rpr2isStepByStepLinearizability}
+\label{lemma:Rpr2 is step-by-step linearizability}
+
+$R_{\textit{pr2}}$ is step-by-step linearizability.
+
+\end{restatable}
+
+\begin {proof}
+
+Assume that $l$ matches $R_{\textit{pr2}}$ with witness $x$.
+Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
+Let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
+Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
+
+Since $h' \sqsubseteq l'$, there exists a sequence $h_{l'}$ of call, return and operations, such that (1) the projection of $h_{l'}$ into call and return action is $h$, (2) the projection of $h_{l'}$ into operations is $l'$, and (3) for each matched pair of call action $c$ and return action $r$ of $h$, there exists an operation $o$, such that $a$ before $o$, $o$ before $r$, and $o$ is the operation of $c$ and $r$.
+Intuitively, $h_{l'}$ is obtained from $h$ by adding linearization points.
+
+According to $R_{\textit{pr2}}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where (1) the $\textit{put}$ operations and $\textit{rm}$ operations in $u \cdot v$ are matched, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v \cdot w$, (3) there exists matched $\textit{put}$ and $\textit{rm}$ operations in $u \cdot v \cdot w$ that has priority $\textit{pri}_x$, (4) $\textit{pri}_x$ is larger than priorities of any unmatched put in $u \cdot v \cdot w$, and (5) all $\textit{put}$ operation with priority $\textit{pri}_x$ is in $u$.
+Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$.
+Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$ operations, we can see that $D_L \cap D_R = \emptyset$.
+Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
+
+Let $l'_{\textit{L0}}$ be the shortest prefix of $l'_L$, such that every operation in $l'_{\textit{L0}}$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $h_{l'}$.
+Let $\textit{OP}_x = \{ \textit{put}(d,\textit{pri}_x) \vert \textit{put}(d,\textit{pri}_x)$ is an operation of $h\}$.
+Let $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t$, where $l_t$ is the projection of $l'$ into operations of $\textit{OP}_x$ that are not in $l'_{\textit{L0}}$.
+Let $\textit{OP}'_x$ be the set of operations in $l_t$.
+Let $l'_{\textit{L2}}$ be the projection of $l'_L$ into operations of $l'_L$ that are not in $l'_{\textit{L1}}$.
+Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.
+
+As in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, we know that $l'_L \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$.
+It is obvious that $l'_{\textit{L0}} \in \textit{PQueue}$, and then $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t \in \textit{PQueue}$.
+Since there is no $\textit{rm}(\textit{Empty})$ in $h$ and $\textit{pri}_x$ is the maximal priority of $h$, $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$.
+Since all $\textit{put}$ operations with priority $\textit{pri}$ is in $l'_{L1}$, we know that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$.
+As in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, we know that $l'' \in \textit{PQueue}$.
+
+To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}.
+Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
+Then, we can see
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+
+\item[-] $o_1 \in L'_R \wedge (o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2} \vee o_2 = \textit{rm}(x))$, or $o_1 = \textit{rm}(x) \wedge (o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2})$: Then $l$ contradicts with happen before relation of $h$.
+
+\item[-] $o_1, o_2 \in l'_R$ or $o_1, o_2 \in l'_{L2}$: Then $l'$ contradicts with happen before relation of $h$.
+
+\item[-] $o_1, o_2 \in l'_{L1}$: If $(o_1, o_2 \in l'_{L0}) \vee (o_1,o_2 \in \textit{OP}'_x)$, then $l'$ contradicts with happen before relation of $h$.
+
+    Else, it is the case that $o_1 \in \textit{OP}'_x \wedge o_2 \in l'_{L0}$.
+We can see that $o_1$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and the linearization point of $o_1$ is after the return action of $\textit{put}(x,\textit{pri}_x)$.
+Since $o_1 <_{\textit{hb}} o_2$, it is easy to see that $\textit{put}(x,\textit{pri}_x) <_{\textit{hb}} o_2$, and then $o_2 \notin l'_{L0}$, which contradicts that $o_2 \in l'_{L0}$.
+
+\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_{L1}$: Since $o_1 <_{\textit{hb}} o_2$, it is obvious that $o_2 \notin l'_{L1}$, contradicts that $o_2 \in l'_{L1}$.
+
+\item[-] $o_1 \in l'_{L2} \wedge o_2 \in l'_{L1}$: If $o_2 \in l'_{L0}$, then $l'$ contradicts with happen before relation of $h$.
+
+    Else, it is the case that $o_2 \in \textit{OP}'_x$.
+$o_2$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and the linearization point of $o_2$ is after the return action of $\textit{put}(x,\textit{pri}_x)$.
+Since $o_1 <_{\textit{hb}} o_2$, the return action of $o_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$.
+Then it is easy to see that $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.
+
+\item[-] $o_1 \in l'_{L2} \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: Then $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.
+\end{itemize}
+
+Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
+\end {proof}
+
+\begin{restatable}{lemma}{Rpr3isStepByStepLinearizability}
+\label{lemma:Rpr3 is step-by-step linearizability}
+
+$R_{\textit{pr3}}$ is step-by-step linearizability.
+
+\end{restatable}
+
+\begin {proof}
+
+Assume that $l$ matches $R_{\textit{pr3}}$ with witness $x$.
+Assume that  Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
+Then, let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
+Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
+
+According to $R_{\textit{pr3}}$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, where $\textit{pri}_x$ is the maximal priority in $u \cdot v$, and $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ operations in $u \cdot v$ (may equal priority of some unmatched $\textit{put}$ operations in $u \cdot v$).
+Let $S_1$ be the set of all operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_1$ be the shortest prefix of $l'$ that contains all operations of $S_1$, and let $l'_2$ be the sequence such that $l' = l'_1 \cdot l'_2$.
+Let sequence $l'' = l'_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_2$.
+
+Since $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ operations in $u \cdot v$, it is easy to see that $l'' \in \textit{PQueue}$.
+It remains to prove that $h \sqsubseteq l''$.
+To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}.
+Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
+Then, we can see
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $o_1,o_2 \in l'_1$, or $o_1,o_2 \in l'_2$, or $o_1 \in l'_2 \wedge o_2 \in l'_1$: Then $l'$ contradicts with happen before relation of $h$.
+
+\item[-] $o_1 \in l'_2 \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: $o_1 \in S_1$, and then $o_1 \in l'_1$, which contradicts that $o_1 \in l'_2$.
+
+\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_1$: We can see that in $l'$, $o_2$ must after every operation in $S_1$, and then $o_2 \notin l'_1$, contradicts that $o_2 \in l'_1$.
+\end{itemize}
+
+Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
+\end {proof}
+
+The following lemma states that, from linearization of sub-histories, we can merge them and obtain a linearization (regardless of whether it belongs to sequential specification) of the whole history.
+
+\begin{restatable}{lemma}{MergeTwoLinearization}
+\label{lemma:merge two linearization}
+
+Given a history $h$ and operation set $S_1$, $S_2$, let $h_1 = h \vert{S_1}$ and $h_2 = h \vert{S_2}$.
+Assume that $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, and $S_1 \cup S_2$ contains all operations of $h$.
+Then, there exists a sequence $l$, such that $h \sqsubseteq l$, $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$.
+\end{restatable}
+
+\begin {proof}
+
+Given a history $h$ and a operation $o \in S_2$, let $\textit{MB}(o) = \{ o' \vert o' \in S_1$ and $o'$ happens before $o$ in $h \}$, let $\textit{SBI}(o) = \textit{min}\{ i \vert l_1[0,i]$ contains all elements of $\textit{MB}(o) \}$.
+
+Let $l = s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[n] \cdot s_{\textit{n+1}}$ be generated as follows, where $n = \vert l_2 \vert$:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $s_1 = l_1[0, \textit{SBI}(l_2[1])]$,
+
+\item[-] If $s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[i]$ already contains $l_1(\textit{SBI}(l_2[\textit{i+1}]))$, then $s_{\textit{i+1}} = \epsilon$.
+Otherwise, $s_{\textit{i+1}}$ is a subsequence of $l_1$, which starts from the next of last elements of $s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[i]$ in $l_1$ and ends in $l_1(\textit{SBI}(l_2[\textit{i+1}]))$.
+\end{itemize}
+
+It is obvious that $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$, and it remains to prove that $h \sqsubseteq l$.
+We prove this by contradiction.
+Assume that $h$ is not linearizable with respect to $l$.
+Then there must be two operations of $h$, such that $o_1 <_{hb} o_2$ in $h$ but $o_2$ before $0_1$ in $l$.
+Since $l \vert{S_1} = l_1$, $l \vert{S_2} = l_2$, and $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, it is easy to see that it is impossible that $o_1,o_2 \in S_1$ or $o_1,o_2 \in S_2$.
+There are only two possibilities:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $o_1 \in S_1 \wedge o_2 \in S_2$.
+Then we can see that $o_2=l_2[i]$ and $o_1 \in s_j$ for some $i < j$.
+Since $o_1 <_{hb} o_2$, we know that $o_1 \in \textit{SBI}(o_2)$.
+By the construction of $l$, we know that $o_1$ must be in $s_k$ for some $k \leq i$, contradicts that $o_1 \in s_j$ with $i < j$.
+
+\item[-] $o_1 \in S_2 \wedge o_2 \in S_1$.
+Then we can see that $o_2 \in s_i$ and $o_1 = l_2[j]$ for some $i \leq j$.
+It is easy to see that this leads to contradiction when $i = j$.
+For the case of $i \neq j$, we need to satisfy the following requirements: (1) $o_1$ ($l_2[j]$) does not happen before $l_2[i]$, (2) $l_2[i]$ does not happen before $o_2$, (3) $o_2$ is either overlap or happens before $o' \in \textit{MB}(l_2[i])$, and (4) $o' <_{hb} l_2[i]$.
+By enumeration we can see that it is impossible that above four conditions be satisfied while $o_1 <_{hb} o_2$.
+\end{itemize}
+
+This completes the proof of this lemma. \qed
+\end {proof}
+
+\begin{restatable}{lemma}{Rpr4isStepByStepLinearizability}
+\label{lemma:Rpr4 is step-by-step linearizability}
+
+$R_{\textit{pr4}}$ is step-by-step linearizability.
+
+\end{restatable}
+
+\begin {proof}
+
+Assume that $l$ matches $R_{\textit{pr4}}$ with witness $x$.
+Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
+Let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
+Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
+
+According to $R_{\textit{pr4}}$, there exist sequences $u$ and $v$, such that (1) $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v$, (3) $\textit{pri}_x$ equals the maximal priority of matched $\textit{put}$ operations in $u \cdot v$, (4) all matched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$ and (5) all the other unmatched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$.
+
+Since $h' \sqsubseteq l'$, there exists a sequence $h_{l'}$ of call, return and operations, such that (1) the projection of $h$ into call and return action is $h$, (2) the projection of $h_{l'}$ into operations is $l'$, and (3) for each matched pair of call action $c$ and return action $r$ of $h$, there exists an operation $o$, such that $a$ before $o$, $o$ before $r$, and $o$ is the operation of $c$ and $r$.
+Intuitively, $h_{l'}$ is obtained from $h$ by adding linearization points.
+
+According to $R_{\textit{pr4}}$, there exist sequences $u$ and $v$, such that (1) $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v$, (3) $\textit{pri}_x$ equals the maximal priority of matched $\textit{put}$ operations in $u \cdot v$, (4) all matched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$ and (5) all the other unmatched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$.
+
+Let $l'_1$ be the shortest prefix of $l'$, such that every operation in $l'_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $h_{l'}$.
+Let $\textit{OP}_x$ be the set of $\textit{put}$ operations with priority $\textit{pri}_x$ and their matched $\textit{rm}$ operations in $h$, let $\textit{OP}'_x$ be the set of $\textit{put}$ operations in $\textit{OP}_x$ that do not appear in $l'_1$, and let $\textit{OP}''_x$ be the set of $\textit{rm}$ operations in $\textit{OP}_x$ that do not appear in $l'_1$.
+Let $l_x''=l'_1 \cdot l'_2 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_3$, where $l'_2$ is the projection of $l'$ into operations of $\textit{OP}'_x$, and $l'_3$ is the projection of $l'$ into operations of $\textit{OP}''_x$.
+
+Let $h_x$ be the projection of $h$ into $\textit{OP}_x$.
+Let us prove that $h_x \sqsubseteq l''_x$ by contradiction.
+Assume that there exists operations $o_1,o_2$, such that $o_1$ happens before $o_2$ in $h_x$, and $o_2$ is before $o_1$ in $l''_x$.
+Then, we can see
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $(o_1 \in l'_3) \wedge (o_2 \in l'_1 \vee o_2 \in l'_2 \vee o'_2 \in l'_3)$, or $(o_1 \in l'_2) \wedge (o_2 \in l'_1 \vee o_2 \in l'_2)$, or $o_1,o_2 \in l'_1$: Then $l'$ contradicts with happen before relation of $h$.
+
+\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_1$: Since the linearization point of $o_2$ is before return action of $o_1$, $o_2$ either overlaps with $o_1$, or happens before $o_1$, which contradicts that $o_1$ happens before $o_2$.
+
+\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_2$: Since $l'_2$ only contains $\textit{put}$ operations with priority $\textit{pri}_x$ and are in $\textit{OP}'_x$, it is not hard to see that each operation in $l'_2$ must overlap with $\textit{put}(x,\textit{pri}_x)$, which contradicts that $o_1$ happens before $o_2$.
+
+\item[-] $o_1 \in l'_3 \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: According to the construction of $l''$, we can see that $1_2 \in l'_1$, which contradicts that $1_2 \in l'_3$.
+\end{itemize}
+
+Therefore, we know that $G$ is acyclic, and then we know that $h_x \sqsubseteq l''_x$.
+
+Let $\textit{OP}_{\textit{-x}}$ be the set of $\textit{put}$ operations with non-$\textit{pri}_x$ priority, and their matched $\textit{rm}$ operations.
+Let $h_{\textit{-x}}$ be the projection of $h$ into $\textit{OP}_{\textit{-x}}$.
+Let $l''_{\textit{-x}}$ be the projection of $l'$ into$\textit{OP}_{\textit{-x}}$.
+It is obvious that $h_{\textit{-x}} \sqsubseteq l''_{\textit{-x}}$.
+By Lemma \ref{lemma:merge two linearization}, there exists a sequence $l'''$, such that $h \sqsubseteq l'''$, $l''' \vert{\textit{OP}_x} = l_x''$ and $l''' \vert{\textit{OP}_{\textit{-x}}} = l''_{\textit{-x}}$.
+
+It is obvious that $l''_{\textit{-x}} \in \textit{PQueue}$.
+Let $l'_x$ be the projection of $l'$ into $\textit{OP}_x - \{ \textit{put}(x,\textit{pri}_x) \}$.
+It is obvious that $l'_x \in \textit{PQueue}$ and $l'_1 \in \textit{PQueue}$.
+Since (1) $l'_1 \cdot l'_2 \cdot l'_3$ is obtained from $l'_x$ by making $\textit{put}$ operations in $l'_2$ to happen earlier, and (2) $l'_3$ contains only $\textit{rm}$ operations, we can see that $l'_1 \cdot l'_2 \cdot l'_3 \in \textit{PQueue}$.
+Since $l'_3$ contains only $\textit{rm}$ operations, it is easy to see that $l_x'' \in \textit{PQueue}$.
+Since $\textit{pri}_x$ is the maximal priority of $h$ and $h$ does not contain $\textit{rm}(\textit{Empty})$ operations, the operations in $l_x''$ does not influence the correctness of $l''_{\textit{-x}} \in \textit{PQueue}$.
+Therefore, we can see that $l''' \in \textit{PQueue}$.
+\qed
+\end {proof}
+
+\begin{restatable}{lemma}{Rpr5isStepByStepLinearizability}
+\label{lemma:Rpr5 is step-by-step linearizability}
+
+$R_{\textit{pr5}}$ is step-by-step linearizability.
+
+\end{restatable}
+
+\begin {proof}
+
+Assume that $l$ matches $R_{\textit{pr5}}$ with operation $o=\textit{rmEmpty}$.
+Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
+Let $h'=h \setminus o$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$.
+Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$ such that $h' \sqsubseteq l'$.
+
+According to $R_{\textit{pr5}}$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{rmEmpty} \cdot v$, where the $\textit{put}$ operations and $\textit{rm}$ operations in $u$ are matched.
+
+Let $D_L$ be the set of items in $u$ and $D_R$ be the set of items in $v$.
+Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
+Let sequence $l'' = l'_L \cdot o \cdot L'_R$.
+Since priority queue is closed under projection (Lemma \ref{lemma:PQ is closed under projection}) and $u$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $l'_L \in \textit{PQueue}$ and the the priority queue is empty after executing $l'_L$.
+Then we know that $l'_L \cdot o \in \textit{PQueue}$.
+Since $l'_R$ is obtained from $l'$ by discarding pairs of matched $\textit{put}$ and $\textit{rm}$ operations, it is easy to see that $L'_R \in \textit{PQueue}$, and then we know that $l'' = l'_L \cdot o \cdot L'_R \in \textit{PQueue}$.
+
+It remains to prove that $h \sqsubseteq l''$.
+To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}.
+Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
+Then, we can see
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $o_1,o_2 \in l'_L$, or $o_1,o_2 \in l'_R$: Then $l'$ contradicts with happen before relation of $h$.
+
+\item[-] If $o_1=o \wedge o_2 \in l'_L$, or $o_1 \in l'_R \wedge o_2 \in l'_L$, or $o_1 \in l'_R \wedge o_2 = 0$, then $l$ contradicts with happen before relation of $h$.
+\end{itemize}
+
+Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$. \qed
+\end {proof}
+
+The following lemma states that $\textit{PQueue}$ is step-by-step linearizability, it is a direct consequence of Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, Lemma \ref{lemma:Rpr2 is step-by-step linearizability}, Lemma \ref{lemma:Rpr3 is step-by-step linearizability}, Lemma \ref{lemma:Rpr4 is step-by-step linearizability} and Lemma \ref{lemma:Rpr5 is step-by-step linearizability}.
+
+\begin{restatable}{lemma}{PQueueisStepByStepLinearizability}
+\label{lemma:PQueue is step-by-step linearizability}
+
+$\textit{PQueue}$ is step-by-step linearizability.
+
+\end{restatable}
+
+\section{Co-Regular of Priority Queues}
+\label{sec:co-regular of priority queues}
+
+\subsection{Co-Regular of $R_{\textit{pr1}}$}
+\label{subsec:co-regular of Rpr1}
+
+Let us introduce the notion of gap for matched $\textit{put}$ and $\textit{rm}$ operations.
+
+\begin{definition}[gap for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:gap for matched put and rm operations}
+
+Let $h$ be a data-differentiated history and $o_1 = \textit{put}(x,\textit{pri}_x)$, $o_2 = \textit{rm}(x)$ be a pair of matched $\textit{put}$ and $\textit{rm}$ operation of $h$.
+We say that $h$ has a gap on operations $o_1$ and $o_2$, if there is a partition of the operations of $h$ into $L$, $M$ and $R$ satisfying:
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $L \cup M$ has matched $\textit{put}$ and $\textit{rm}$ operations.
+
+\item[-] no operation of $R$ happens before an operation of $L$, $o_1$, $M$ or $o_2$.
+
+\item[-] no operation of $o_2$ happens before an operation of $L$, $o_1$ or $M$.
+
+\item[-] no operation of $M$ happens before an operation of $L$ or $o_1$.
+
+\item[-] no operation of $o_1$ happens before an operation of $L$.
+
+\item[-] no operation in $L \cup M \cup R$ has priority $\textit{pri}_x$.
+\end{itemize}
+\end{definition}
+
+Given a history $h$ and set $O$ of some operations in $h$.
+It is easy to see that there exists a sequence $l$ of operations of $O$, such that $l$ does not contradicts the happen before relation of $h$.
+Therefore, from now on, we can use this fact to directly generate such sequence from such $O$.
+
+The following lemma reduce linearizable with respect to $\textit{MR}_{\textit{pr1}}$ into gap of matched $\textit{put}$ and $\textit{rm}$ operations.
+
+\begin{restatable}{lemma}{GapEqualsLinforRpr1}
+\label{lemma:Gap Equals Lin for Rpr1}
+A data-differentiated history $h$ has a projection $h'$, such that $\textit{last}(h') = R_{\textit{pr1}}$ and $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$, if and only if there exists a projection $h''$ of $h$, such that $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is in $h''$, $\textit{pri}_x$ is the maximal priority of $h''$, and one the the following case holds:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] case $1$: $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$,
+\item[-] case $2$: there is no gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
+\end{itemize}
+\end{restatable}
+
+\begin {proof}
+
+Let us prove the $\textit{only if}$ direction by contradiction.
+Assume that there exists such $h'$, while case $1$ and case $2$ do not hold for any projection of $h$.
+Since case $1$ does not hold for any projection of $h$, we know that $\forall y \in D$, $\textit{rm}(y)$ does not happens before $\textit{put}(y,\_)$ in $h$.
+Since case $2$ does not hold for any projection of $h$ and $\textit{last}(h') = R_{\textit{pr1}}$, we can see that $h'$ has gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$, while $\textit{pri}_x$ is the maximal priority in $h'$.
+Since $\textit{last}(h') = R_{\textit{pr1}}$, we can see that $\textit{pri}_x$ is larger than priority of all other $\textit{put}$ operations of $h$.
+
+By the properties of the gap, there is partitions $L$, $M$ and $R$.
+Let $u$, $v$ and $w$ contain the operations of $L$, $M$ and $R$, respectively.
+It is easy to see that $h'$ is linearizable with respect to $l''=u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$ and $l'' \in \textit{MR}_{\textit{pr1}}$.
+This contradicts the fact that $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$.
+
+Let us begin to prove the $\textit{if}$ direction.
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If case $1$ holds for $h''$, then $h \vert_{ \{ x \} }$ is the $h'$ as required.
+
+\item[-] If case $2$ holds, assume by contradiction that, for every projection $h'$ of $h$, if $\textit{last}(h') = R_{\textit{pr1}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr1}}$.
+Since $h''$ is also a projection of $h$, we know that there exist sequences $u$, $v$ and $w$, such that $h'' \sqsubseteq u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, $u \cdot v$ contains matched $\textit{put}$ and $\textit{rm}$ operations, and $\textit{pri}_x$ is larger than priority of $u \cdot v \cdot w$.
+Let $L$, $M$ and $R$ be the operations of $u$, $v$ and $w$, respectively.
+Then, the partitions $L$, $M$ and $R$ form a gap on operation $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
+This contradicts the fact that there is no gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
+\end{itemize}
+
+This completes the proof of this lemma. \qed
+\end {proof}
+
+\begin{definition}[left-right constraint for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:left-right constraint for matched put and rm operations}
+Given a data-differentiated history $h$, and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
+The left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is the graph $G$ where:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] the nodes are the items of $h$, to which we add a node,
+
+\item[-] there is an edge from item $d_1$ to $x$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$ and $\textit{pri}_1 \neq \textit{pri}_x$,
+
+\item[-] there is an edge from $x$ to item $d_1$, if $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$, and the priority of $d_1$ is not $\textit{pri}_x$,
+
+\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{rm}(d_2,\textit{pri}_2)$, $\textit{pri}_1 \neq \textit{pri}_x$, and $\textit{pri}_2 \neq \textit{pri}_x$.
+\end{itemize}
+\end{definition}
+
+Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
+Let $\textit{LMSet}_1(h,x) = \{ o \vert$ $o$ is not an operation of $x$, the priority of $o$ does not equal the priority of $x$, and either it happens before $\textit{put}(x)$ or $\textit{rm}(x)$ in $h$, or there is an operation $o'$ with the same item of $o$, such that $o'$ happens before $\textit{put}(x)$ or $\textit{rm}(x)$ in $h$ $\}$.
+For each $i \geq 1$, let $\textit{LMSet}_{\textit{i+1}}(h,x) = \{ o \vert$ $o$ is not an operation of $x$ or items in $\textit{LMSet}_k(h,x)$ for each $k \leq i$, the priority of $o$ does not equal the priority of $x$, and either it happens before some operation $o' \in \textit{LMSet}_i(h,x)$ in $h$, or there is an operation $o''$ with the same item of $o$ and $o''$ happens before some operation $o' \in \textit{LMSet}_i(h,x)$ in $h$ $\}$.
+%We can see that for each $i$, $\textit{LMSet}_i(h,x)$ contains matched $\textit{put}$ and $\textit{rm}$ operations, and
+We can see that $\textit{LMSet}_i(h,x) \cap \textit{LMSet}_j(h,x) = \emptyset$ for any $i \neq j$.
+Let $\textit{LMSet}(h,x) = \textit{LMSet}_1(h,x) \cup \textit{LMSet}_2(h,x) \cup \ldots$.
+%{\color {red} We say that a history $h$ of priority queue is matched before $x$, if (1) $h$ contains no $\textit{rmEmpty}$ operations, (2) $ h \sqsubseteq l_1 \cdot \textit{put}(x,\_) \cdot l_2 \cdot \textit{rm}(x) \cdot l_3$ for some $l_1,l_2,l_3$, and $l_1 \cdot l_2$ contains matched $\textit{put}$ and $\textit{rm}$ operations.}
+
+The following is an example that shows that, given a data-differentiated history $h$, the set $\textit{LMSet}_i(h,x)$ can be non-empty for each $1 \leq i \leq 3$.
+In \figurename~\ref{fig:his nobound of LMSet} , $\textit{LMSet}_1(h,x) = \{ \textit{put}(a,1),\textit{rm}(a) \}$, $\textit{LMSet}_2(h,x) = \{ \textit{put}(b,1),\textit{rm}(b) \}$ and $\textit{LMSet}_3(h,x) = \{ \textit{put}(a,c),\textit{rm}(c) \}$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.5 \textwidth]{PIC_HIS_NOBOUNDOF_LMSET.pdf}
+%\vspace{-10pt}
+  \caption{An example for non-emptiness of $\textit{LMSet}_1(h,x)$, $\textit{LMSet}_2(h,x)$ and $\textit{LMSet}_3(h,x)$}
+  \label{fig:his nobound of LMSet}
+\end{figure}
+
+\begin{restatable}{lemma}{RmxDoesNotHappenBeforeLMSetForRpr1}
+\label{lemma:Rmx does not happen before LMSet for Rpr1}
+Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
+%, such that $\textit{pri}_x$ is the maximal priority of $h$, and $\textit{last}(h) = R_{\textit{pr1}}$.
+Assume that $\textit{rm}(y)$ does not happen before $\textit{put}(y,\_)$ for each item $y$ in $h$.
+Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
+Assume that $G$ has no cycle going through $x$.
+Then, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}(h,x)$.
+\end{restatable}
+
+\begin {proof}
+
+We prove this lemma by induction, and prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x)$, in $\textit{LMSet}_2(h,x)$, $\ldots$.
+Note that, according to the definition of $\textit{LMSet}(h,x)$, $\textit{pri}_x$ is different to priority of all items in $\textit{LMSet}(h,x)$.
+
+\noindent (1) Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x)$ by contradiction.
+Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{LMSet}_1(h,x)$ is an operation of item $d$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d$ does not equals $\textit{pri}_x$).
+
+We use a triple $(t_1,t_2,t_3)$ to represent related information.
+$t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$.
+$t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation.
+$t_2$ and $t_3$ is used for the reason of $o \in \textit{LMSet}_1(h,x)$: $o \in \textit{LMSet}_1(h,x)$, since an operation (of kind $t_2$) of $d$ happens before an operation (of kind $t_3$) of $x$.
+Let us consider all the possible cases of $(t_1,t_2,t_3)$:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.
+
+\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{rm}(x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{rm}(x)$.
+
+\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x) )$.
+By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(x,\textit{pri}_x) ) \vee ( \textit{rm}(d) <_{hb} \textit{put}(d,\_) )$, which is impossible.
+
+\item[-] $(\textit{put},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{rm}(x) )$.
+We can see that $\textit{rm}(d) <_{hb} \textit{rm}(x) <_{hb} \textit{put}(d,\_)$, which contradicts that $\textit{rm}(d)$ does not happen before $\textit{put}(d,\_)$.
+
+\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x) )$.
+We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.
+
+\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{rm}(x) )$.
+We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.
+
+\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.
+
+\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{rm}(x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{rm}(x)$.
+\end{itemize}
+
+This completes the proof for $\textit{LMSet}_1(h,x)$.
+
+\noindent (2) Assume we already prove that for some $j \geq 1$, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
+Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_{\textit{j+1}}(h,x)$ by contradiction.
+Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$ is an operation of item $d_{\textit{j+1}}$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d_{\textit{j+1}}$ does not equals $\textit{pri}_x$).
+We use a triple $(t_1,t_2,t_3)$ to represent related information.
+$t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$.
+$t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation.
+$t_2$ and $t_3$ is used for the reason of $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$: $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$, since an operation (of kind $t_2$) of $d_{\textit{j+1}}$ happens before an operation (of kind $t_3$) of $d_j$, where $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{LMSet}_j(h,x)$.
+Let us consider all the possible cases of $(t_1,t_2,t_3)$:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_)$.
+We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
+
+\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j,\_)$.
+We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j,\_) ) \wedge ( \textit{rm}(d_j) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
+
+\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$.
+By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \vee ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_{\textit{j+1}},\_) )$, which is impossible.
+
+\item[-] $(\textit{put},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$.
+By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \vee ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_{\textit{j+1}},\_) )$, which is impossible.
+
+\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) )$.
+Let us consider the reason of $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{LMSet}_j(h,x)$:
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $( j > 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(_{\textit{j-1}},\_), \textit{rm}(_{\textit{j-1}}) \in \textit{LMSet}_{\textit{j-1}}(h,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, we can see that $\textit{put}(d_{\textit{j+1}},\_) <_{hb} o''$, and then operations of $d_{\textit{j+1}}$ is in $\textit{LMSet}_j(h,x)$, contradicts that operations of $d_{\textit{j+1}}$ is in $\textit{LMSet}_{\textit{j+1}}(h,x)$.
+
+    \item[-] If $( j = 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of $x$: Similar to above case.
+
+    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{LMSet}_{\textit{j-1}}(h,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, we can see that $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} o'' ) \vee ( \textit{rm}(d_j) <_{hb} \textit{put}(d_j,\_) )$, which is impossible.
+
+    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of $x$: Similar to above case.
+    \end{itemize}
+
+\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Let $T_{\textit{ind}}$ be the set of sentences $\{ \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}), \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j),\ldots, \textit{put}(d_{\textit{ind+1}},\_) <_{hb} \textit{rm}(d_{\textit{ind}}) \}$.
+Here each $d_i$ is a item of some operation in $\textit{LMSet}_i(h,x)$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d_i$ does not equals $\textit{pri}_x$ for each $i$).
+Let us prove that from $T_j$ we can obtain contradiction by induction:
+
+    {\bf Base case $1$}: From $T_1$ we can obtain contradiction.
+
+    Let us prove base case $1$:
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $\textit{put}(d_1,\_)$ happens $o$, and $o$ is an operation of $x$.
+Then there is a cycle $x \rightarrow d_{\textit{j+1}} \rightarrow \ldots \rightarrow d_1 \rightarrow x$ in $G$, contradicts that $G$ has no cycle going through $x$.
+
+    \item[-] If $\textit{rm}(d_1)$ happens before $o$, and $o$ is an operation of $x$.
+Then since $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1) <_{hb} o$, we can see that $\textit{put}(d_2,\_) <_{hb} o$, and then $\textit{put}(d_2,\_) \in \textit{LMSet}_1(h,x)$, contradicts that $\textit{put}(d_2,\_) \in \textit{LMSet}_2(h,x)$.
+    \end{itemize}
+
+    {\bf Base case $2$}: From $T_2$ we can obtain contradiction.
+
+    Let us prove base case $2$: If $\textit{rm}(d_2) <_{hb} o$, and $o$ is an operation of $d_1$, then since $( \textit{put}(d_3,\_) <_{hb} \textit{rm}(d_2) ) \wedge ( \textit{rm}(d_2) <_{hb} o )$, we know that $\textit{put}(d_3,\_) <_{hb} o$.
+This implies that $\textit{put}(d_3,\_) \in \textit{LMSet}_2(h,x)$, contradicts that $\textit{rm}(d_3,\_) \in \textit{LMSet}_3(h,x)$.
+Therefore, it is only possible that $\textit{put}(d_2,\_)$ happens before an operation of $d_1$.
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{put}(d_1,\_)$ and $\textit{put}(d_1,\_)$ happens before operations of $x$, then we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, which is impossible.
+
+    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{put}(d_1,\_)$ and $\textit{rm}(d_1)$ happens before operations of $x$, then by interval order, we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, or $\textit{rm}(d_1) <_{hb} \textit{put}(d_1,\_)$, which is impossible.
+
+    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{put}(d_1,\_)$ happens before operations of $x$, then $x \rightarrow d_{\textit{j+1}} \rightarrow \ldots \rightarrow d_1 \rightarrow x$ in $G$, contradicts that $G$ has no cycle going through $x$.
+
+    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1)$ happens before operations of $x$, then we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, which is impossible.
+    \end{itemize}
+
+    {\bf induction step}: Given $\textit{ind} \geq 3$, if from $T_{\textit{ind-1}}$ we can obtain contradiction, then from $T_{\textit{ind}}$ we can also contain contradiction.
+
+    Prove of the induction step: Similarly as base case $2$, we can prove that it is only possible that $\textit{put}(d_{\textit{ind}},\_)$ happens before operations of $d_{\textit{ind-1}}$.
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$ and $\textit{put}(d_{\textit{ind-1}},\_)$ happens before operations of $d_{\textit{ind-2}}$, then we know that $\textit{put}(d_{\textit{ind}})$ happens before operation of $d_{\textit{ind-2}}$, which is impossible.
+
+    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$ and $\textit{rm}(d_{\textit{ind-1}})$ happens before operations of $d_{\textit{ind-2}}$, then by interval order, we know that $\textit{put}(d_{\textit{ind}},\_)$ happens before operation of $d_{\textit{ind-2}}$, or $\textit{rm}(d_{\textit{ind-1}}) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$, which is impossible.
+
+    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{rm}(d_{\textit{ind-1}})$, then we obtain $T_{\textit{ind-1}}$, which already contain contradiction.
+    \end{itemize}
+
+    By base case $1$, base case $2$ and the induction step, it is easy to see that $T_j$ contains contradiction.
+
+\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$.
+We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
+
+\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$.
+We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \wedge ( \textit{rm}(d_j) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
+\end{itemize}
+
+This completes the proof for $\textit{LMSet}_{\textit{j+1}}(h,x)$.
+Therefore, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}(h,x) = \textit{LMSet}_1(h,x) \cup \textit{LMSet}_2(h,x) \cup \ldots$.
+\qed
+\end {proof}
+
+\begin{restatable}{lemma}{LMSetHasMatchedPutandRmOperations}
+\label{lemma:LMSet has matched put and rm operations}
+Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
+Assume that $\textit{rm}(y)$ does not happen before $\textit{put}(y,\_)$ for each item $y$ in $h$.
+Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
+Assume that $G$ has no cycle going through $x$.
+Then, $\textit{LMSet}(h,x)$ contains only matched $\textit{put}$ and $\textit{rm}$ operations.
+\end{restatable}
+\begin {proof}
+
+We prove this lemma by contradiction.
+Assume that there exists a value, such that $\textit{LMSet}(h,x)$ contains only its $\textit{put}$ operation and does not contain its $\textit{rm}$ operation.
+Then we can see that there exists $d_1,\ldots,d_j$.
+Intuitively, $d_1,\ldots,d_j$ are items in $\textit{LMSet}_1(h,x), \ldots, \textit{LMSet}_i(h,x)$, respectively.
+$\textit{LMSet}(h,x)$ contains $\textit{put}(d_j,\_)$ and does not contain $\textit{rm}(d_j)$.
+And each $d_i$ is the reason of $d_{\textit{i+1}} \in \textit{LMSet}_{\textit{i+1}}(h,x)$.
+Formally, we require that
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] For each $1 \leq i \leq j$, operations of $d_i$ belongs to $\textit{LMSet}_i(h,x)$. %For each $1 \leq i \leq j$, the priority of $d_i$ is not $\textit{pri}_x$.
+
+\item[-] For each $i \neq j$, $\textit{put}(d_i,\_),\textit{rm}(d_i) \in \textit{LMSet}_i(h,x)$. $\textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x)$, and $h$ does not contain $\textit{rm}(d_j)$.
+
+\item[-] An operation of $d_1$ happens before an operation of $x$. For each $1 < i \leq j$, an operation of $d_j$ happens an operation of $d_{\textit{i-1}}$.
+
+\item[-] For each $k$ and $\textit{ind}$, if $k > \textit{ind+1}$, then no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.
+\end{itemize}
+
+According to the definition of $\textit{LMSet}(h,x)$, it is easy to see that such $d_1,\ldots,d_j$ exists. Let us prove the following fact:
+
+\noindent {\bf $\textit{fact}_1$}: Given $1 \leq i < j$, it can not be the case that $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap.
+
+Proof of $\textit{fact}_1$: We prove $\textit{fact}_1$ by contradiction.
+Assume that for some $i \neq j$, $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap.
+Since $\textit{put}(d_i,\_), \textit{rm}(d_i) \in \textit{LMSet}_i(h,x)$, we know that an operation $o_i$ of $d_i$ happens before operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$.
+Moreover, since $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap, it is not hard to see that the call action of $\textit{put}(d_i,\_)$ and the call action of $\textit{rm}(d_i)$ is before the call action of $o_{\textit{i-1}}$.
+Since $\textit{put}(d_{\textit{i+1}},\_), \textit{rm}(d_{\textit{i+1}}) \in \textit{LMSet}_{\textit{i+1}}(h,x)$, we know that an operation $o'_{\textit{i+1}}$ of $d_{\textit{i+1}}$ happens before operation $o'_i$ of $d_i$.
+Then, it is not hard to see that $o'_{\textit{i+1}}$ also happens before $o_{\textit{i-1}}$, which contradicts that For each $k > \textit{ind+1}$, no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.
+
+We already know that an operation of $d_1$ happens before an operation of $x$.
+By $\textit{fact}_1$, we can ensure that $\textit{put}(d_1,\_)$ happens before an operation of $x$, and then $d_1 \rightarrow x$ in $G$.
+For each $1 < i < j$, we know that an operation $o_i$ of $d_i$ happens before an operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$.
+By $\textit{fact}_1$, we can ensure that $o_i=\textit{put}(d_i,\_)$ and $o_{\textit{i-1}}=\textit{rm}(d_{\textit{i-1}})$, and then $d_i \rightarrow d_{\textit{i-1}}$ in $G$.
+Since $h$ contains $\textit{put}(d_j)$ and does not contain $\textit{rm}(d_j)$, we know that $x \rightarrow d_j$ in $G$.
+Then $G$ has a cycle going through $x$, contradicts that $G$ has no cycle going through $x$.
+\qed
+\end {proof}
+
+\begin{restatable}{lemma}{GapEqualsConstraintforRpr1}
+\label{lemma:Gap Equals Constraint for Rpr1}
+Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
+%, where $\textit{pri}_x$ has the maximal priority of $h$, and $\textit{last}(h) = R_{\textit{pr1}}$.
+Assume that for each item in $h$, its $\textit{rm}$ operation does not happen before its $\textit{put}$ operation.
+Let $G$ be the graph representing the left-right constraint of $\textit{put}(x)$ and $\textit{rm}(x)$.
+There is a gap on $\textit{put}(x)$ and $\textit{rm}(x)$, if and only if $G$ has no cycle going through $x$.
+\end{restatable}
+
+\begin {proof}
+
+To prove the $\textit{only if}$ direction, assume that there is a gap on $\textit{put}(x)$ and $\textit{rm}(x)$, and let it be $L$, $M$ and $R$.
+Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ in $G$.
+It is obvious that the priority of each $d_i$ is not $\textit{pri}_x$.
+Then our proof proceeds as follows:
+
+According to the definition of left-right constraint, there are two possibilities.
+The first possibility is that, $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$.
+It is obvious that $\textit{rm}(d_1) \in R$, and then since $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_1),\textit{rm}(d_1) \in R$.
+Then,
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Since $d_1 \rightarrow d_2$, by definition of $G$, we know that $\textit{put}(d_1)$ happens before $\textit{rm}(d_2)$.
+Since $\textit{put}(d_1) \in R$ and $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_2),\textit{rm}(d_2) \in R$.
+Similarly, for each $1 \leq i \leq m$, we know that $\textit{put}(d_i),\textit{rm}(d_i) \in R$.
+
+\item[-] Since $d_m \rightarrow x$,
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] if $\textit{put}(d_m)$ happens before $\textit{put}(x)$, then we can see that $\textit{put}(d_m) \in L$, which contradicts that $\textit{put}(d_m) \in R$.
+
+    \item[-] if $\textit{put}(d_m)$ happens before $\textit{rm}(x)$, then we can see that $\textit{put}(d_m) \in L \cup M$, which contradicts that $\textit{put}(d_m) \in R$.
+    \end{itemize}
+\end{itemize}
+
+The second possibility is that, $h$ contains one $\textit{put}(d_1,\_)$ operation and no $\textit{rm}(d_1)$ operation.
+Note that for each $j > 1$, $h$ contains $\textit{put}(d_j,\_)$ and $\textit{rm}(d_j)$.
+Since $d_m \rightarrow x$, is is obvious that $\textit{put}(d_m) \in L \cup M$.
+Since $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_m),\textit{rm}(d_m) \in R$.
+Then,
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Since $d_{\textit{m-1}} \rightarrow d_m$, by definition of $G$, we know that $\textit{put}(d_{\textit{m-1}})$ happens before $\textit{rm}(d_m)$.
+Since $\textit{rm}(d_m) \in L \cup M$ and $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_{\textit{m-1}}),\textit{rm}(d_{\textit{m-1}}) \in L \cup M$.
+Similarly, for each $1 < i \leq m$, we know that $\textit{put}(d_i),\textit{rm}(d_i) \in L \cup M$, and $\textit{put}(d_1)\in L \cup M$.
+
+\item[-] Since $d_1 \rightarrow d_2$ in $G$, $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2)$.
+Since $\textit{put}(d_2),\textit{rm}(d_2) \in L \cup M$, we can see that $\textit{put}(d_1) \in L \cup M$.
+However, there is one $\textit{put}(d_1,\_)$ operation and no $\textit{rm}(d_1)$ operation in $h$, contradicts that $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
+\end{itemize}
+
+This completes the proof of the $\textit{only if}$ direction.
+
+To prove the $\textit{if}$ direction, assume that $G$ has no cycle going through $x$.
+Let $O_L$ be the set of operations that happen before $\textit{put}(x)$ in $h$.
+It is easy to see that $O_L \subseteq \textit{LMSet}(h,x)$.
+Let $O_M = \textit{LMSet}(h,x) \setminus O_L$.
+Let $O_h$ be the set of operations of $h$, and let $O_R = O_h \setminus \textit{LMSet}(h,x)$.
+Let $L$, $M$ and $R$ be $O_L$, $O_M$ and $O_R$, respectively.
+
+By Lemma \ref{lemma:LMSet has matched put and rm operations}, we can see that $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
+It remains to prove that $L$, $M$ and $R$ is a gap of $\textit{put}(x)$ and $\textit{rm}(x)$.
+We prove this by showing that all the following cases are impossible:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Case $1$: Some operation $o_r \in R$ happens before $\textit{rm}(x)$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.
+
+\item[-] Case $2$: Some operation $o_r \in R$ happens before some operation $o_{lm} \in L \cup M$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.
+
+\item[-] Case $3$: Some operation $o_r \in R$ happens before $\textit{put}(x)$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.
+
+\item[-] Case $4$: $\textit{rm}(x)$ happens before some $o_{lm} \in \textit{LMSet}(h,x) = L \cup M$. By Lemma \ref{lemma:Rmx does not happen before LMSet for Rpr1} we know that this is impossible.
+
+\item[-] Case $5$: $\textit{rm}(x)$ happens before $\textit{put}(x)$. This contradicts that for each item in $h$, its $\textit{rm}$ operation does not happen before its $\textit{put}$ operation.
+
+\item[-] Case $6$: Some operation $o_m \in M$ happens before $\textit{put}(x)$. Then we know that $o_m \in L$, which contradicts that $o_m \in M$.
+
+\item[-] Case $7$: Some operation $o_m \in M$ happens before some operation $o_l \in L$. Then we know that $o_m \in L$, which contradicts that $o_m \in M$.
+
+\item[-] Case $8$: $\textit{put}(x)$ happens before some operation $o_l \in L$. This is impossible.
+\end{itemize}
+
+This completes the proof of the $\textit{if}$ direction.
+
+\qed
+\end {proof}
+
+A automaton $\mathcal{A}_{\textit{Rpr1-1}}$ is given in \figurename~\ref{fig:automata for Rpro-1}, where $c=\textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm,a),$ $\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$.
+$\mathcal{A}_{\textit{Rpr1-1}}$ is used to detect the existence of a item whose $\textit{rm}$ happens before its $\textit{put}$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-1.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-1}}$}
+  \label{fig:automata for Rpro-1}
+\end{figure}
+
+Let us begin to represent several automata that is used for capture the case that, in a sub-history $h'$ of a history $h$, $\textit{last}(h')=R_{\textit{pr1}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$, and the reason is that there is no gap on $\textit{put}(y,\textit{pri}_y)$ and $\textit{rm}(y)$ in $h'$, where $y$ is the item that has maximal priority in $h'$.
+Such case is dealt with by four automata, since there are four possibilities of $h' \vert_{y}$.
+
+The first automaton is $\mathcal{A}_{\textit{Rpr1-pprr}}$, which is given in \figurename~\ref{fig:automata for Rpro-pprr}.
+In \figurename~\ref{fig:automata for Rpro-pprr}, $c_1= \textit{cal}(\textit{put},c,$ $\textit{anyPri}),\textit{ret}(\textit{put},c), \textit{cal}(\textit{rm},c), \textit{ret}(\textit{rm},c)$, $c_2 = c_1 + \textit{cal}(\textit{put},a,\textit{les}_p)$, and $c_3 = c_2 + \textit{ret}(\textit{rm},a)$.
+Here $p$ is a variable for priority, and $\textit{les}_p$ is a predicate that accepts all priorities that are less than the value of $p$.
+Here $\textit{pprr}$ represents that, $h' \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$ (we consider $b$ as $y$ and $p$ as $\textit{pri}_y$).
+In \figurename~\ref{fig:automata for Rpro-pprr}, the three paths that (1) from $q_{\textit{init}}$ to $q_2$ to $q_3$, (2) from $q_{\textit{init}}$ to $q_2$ to $q_{11}$, and (3) from $q_{\textit{init}}$ to $q_{10}$ to $q_{11}$ represents three possible positions of $\textit{ret}(\textit{put},a)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-pprr.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-pprr}}$}
+  \label{fig:automata for Rpro-pprr}
+\end{figure}
+
+The second automaton is $\mathcal{A}_{\textit{Rpr1-prpr}}$, which is given in \figurename~\ref{fig:automata for Rpro-prpr}.
+In $\mathcal{A}_{\textit{Rpr1-prpr}}$, $c_1, c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$.
+Here $\textit{pprr}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{put},b,p) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-prpr.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-prpr}}$}
+  \label{fig:automata for Rpro-prpr}
+\end{figure}
+
+The third automaton is $\mathcal{A}_{\textit{Rpr1-rppr}}$, which is given in \figurename~\ref{fig:automata for Rpro-rppr}.
+In $\mathcal{A}_{\textit{Rpr1-rppr}}$, $c_1, c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$.
+Here $\textit{rppr}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-rppr.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-rppr}}$}
+  \label{fig:automata for Rpro-rppr}
+\end{figure}
+
+The forth automaton is $\mathcal{A}_{\textit{Rpr1-rprp}}$, which is given in \figurename~\ref{fig:automata for Rpro-rprp}.
+In $\mathcal{A}_{\textit{Rpr1-rprp}}$, $c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$, and $c_4=c_1 + \textit{ret}(put,b)$.
+Here $\textit{rprp}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b) \cdot \textit{ret}(\textit{put})$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-rprp.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-rprp}}$}
+  \label{fig:automata for Rpro-rprp}
+\end{figure}
+
+Given a data-differentiated history $h$, we say that a item $x$ with priority $\textit{pri}_x$ in $h$ is covered by items $d_1,\ldots,d_m$ in $h$, if the priorities of $d_1,\ldots,d_m$ is not $\textit{pri}_x$, and
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $\textit{put}(d_m,\_)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$,
+
+\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,
+
+\item[-] $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$, or $\textit{rm}(d_1)$ does not exists in $h$
+\end{itemize}
+
+According to the definition of left-right constraint, in a data-differentiated history $h$, there is a cycle going through $x$, if and only if there exists items $d_1,\ldots,d_m$, such that $x$ is covered by $d_1,\ldots,d_m$.
+
+\begin{restatable}{lemma}{Rpr1IsCoRegular}
+\label{lemma:Rpr1 is co-regular}
+$R_{\textit{pr1}}$ is co-regular.
+\end{restatable}
+
+\begin {proof}
+
+Let $\textit{AutS-R}_{\textit{pr1}} = \{ \mathcal{A}_{\textit{Rpr1-1}} , \mathcal{A}_{\textit{Rpr1-pprr}}, \mathcal{A}_{\textit{Rpr1-prpr}}, \mathcal{A}_{\textit{Rpr1-rppr}}, \mathcal{A}_{\textit{Rpr1-rprp}} \}$.
+We need to prove that:
+
+\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr1}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr1}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr1}}$.
+
+By Lemma \ref{lemma:Gap Equals Lin for Rpr1} and Lemma \ref{lemma:Gap Equals Constraint for Rpr1}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:
+
+\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr1}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h''$ is a projection of $h$, such that $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is in $h''$, $\textit{pri}_x$ is the maximal priority of $h''$, and one the the following case holds:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] case $1$: $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$,
+\item[-] case $2$: $x$ is covered by some items $d_1,\ldots,d_m$ in $h''$.
+\end{itemize}
+
+\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\textit{AutS-R}_{\textit{pr1}}$.
+By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$.
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-1}}$: Let $d_1$ a item in $h_2$ such that $r(d_1)=b$ and let $h'' = h_2 \vert_{ \{ d_1 \} }$.
+$d_1$ is the item with maximal priority in $h''$.
+It is obvious that $\textit{last}(h'') = R_{\textit{pr1}}$ and $\textit{rm}(d_1)$ happens before $\textit{put}(d_1,\_)$ in $h''$.
+
+\item[-] If $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-pprr}}$, $\mathcal{A}_{\textit{Rpr1-prpr}}$, $\mathcal{A}_{\textit{Rpr1-rppr}}$, or $\mathcal{A}_{\textit{Rpr1-rprp}}$: Let $d_0$ be a item in $h_2$ such that $r(d_0)=b$, let $d_1,\ldots,d_m$ be the items in $h_2$ such that $r(d_i)=a$ for each $1 \leq i \leq m$.
+Let $h'' = h_1 \vert_{ \{ d_0, d_1, \ldots, d_m \} }$.
+It is obvious that the priority of $d_0$ is larger than the priorities of other operations in $h''$, and then $\textit{last}(h'') = R_{\textit{pr1}}$.
+It is easy to see that $d_0$ is covered by $d_1,\ldots,d_m$.
+\end{itemize}
+
+\noindent The $\textit{if}$ direction: Assume that there exists such $h$, $h''$, $x$ and $\textit{pri}_x$.
+Since $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{pri}_x$ is larger than the priorities of other operations in $h''$.
+Then,
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If case $1$ holds: Then, $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$ in $h''$.
+Let $h_1$ be obtained from $h$ by renaming $x$ to $b$ and renaming all other items to $a$.
+By data-independence, $h_1 \in \mathcal{I}$, and it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-1}}$.
+\item[-] case $2$: Then, $x$ is covered by some items $d_1,\ldots,d_m$ in $h''$.
+There are four possibilities of $h'' \vert_{ \{ x \} }$: (1) $\textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{rm},x)$, (2) $\textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},x)$, (3) $\textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},x)$ and (4) $\textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{rm},x) \cdot \textit{ret}(\textit{put})$.
+
+    Let $h_1$ be obtained from $h$ by renaming $x$ into $b$, renaming $d_1,\ldots,d_m$ into $a$ and renaming other items into $c$.
+We can prove that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-pprr}}$, $\mathcal{A}_{\textit{Rpr1-prpr}}$, $\mathcal{A}_{\textit{Rpr1-rppr}}$ or $\mathcal{A}_{\textit{Rpr1-rprp}}$.
+
+    For example, if $h'' \vert_{ \{ x \} } = \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{rm},x)$, and $\textit{ret}(\textit{rm},d_m)$ is before $\textit{call}(\textit{put},x,\textit{pri}_x)$ and is after $\textit{ret}(\textit{put},x)$ in $h''$.
+Then in  $\mathcal{A}_{\textit{Rpr1-pprr}}$, $h_1$ is accepted by transitions from $q_{\textit{init}}$ to $q_2$ to $q_{11}$ and finally to $q_7$ (if $\textit{rm}(d_1)$ does not exists in $h''$) or $q_9$ (if $\textit{rm}(d_1)$ exists in $h''$).
+\end{itemize}
+
+This completes the proof of this lemma. \qed
+\end {proof}
+
+\subsection{Co-Regular of $R_{\textit{pr2}}$}
+\label{subsec:co-regular of Rpr2}
+
+An automaton $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$ is given in \figurename~\ref{fig:automata for FIFO of single priority}.
+In \figurename~\ref{fig:automata for FIFO of single priority}, let $c_1= \textit{cal}(\textit{put},c,\textit{anyPri}),\textit{ret}(\textit{put},c)$, $\textit{cal}(\textit{rm},c),\textit{ret}(\textit{rm},c)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, and $c_3 = c_2 + \textit{cal}(\textit{rm},a) + \textit{ret}(\textit{rm},a)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_QUEUE_FOR_SINGLE_PRI.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$}
+  \label{fig:automata for FIFO of single priority}
+\end{figure}
+
+According to \cite{Bouajjani:2015}, given a data-independent set $S$ of queue executions, whether every sequence $s$ in $S$ is linearizable to queue can be reduced into checking emptiness of intersection of $S$ and several finite automata.
+Let $\textit{transToQueue}(h)$ be a sequence $h'$ that is generated from $h$ by transforming $\textit{put}$ and $\textit{rm}$ into $\textit{enq}$ and $\textit{deq}$, respectively, and then discarding priorities.
+A priority queue history is called single-priority, if all of its items has only one priority.
+The following lemma states that we can also use finite automata to check ``first in first out property'' for single-priority sub-sequence of priority queue history.
+
+\begin{restatable}{lemma}{AutoForPQwithSignlePri}
+\label{lemma:automata for priority queue with single priority}
+
+Given a data-independent implementations $\mathcal{I}$ of priority queue, let $\mathcal{A} = \{ \mathcal{A}_{\textit{unmRm1}}, \mathcal{A}_{\textit{Rpr0-1}}, \mathcal{A}_{\textit{unmRm2}}, \mathcal{A}_{\textit{FIFO-Sig-Pri}} \}$, then,
+
+$\mathcal{I} \cap \mathcal{A} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority, and $\textit{transToQueue}(e')$ does not linearizable to queue.
+\end{restatable}
+
+\begin {proof}
+
+According to \cite{Bouajjani:2015}, checking linearizable with respect to queue can be reduced into checking the following three conditions for data-differentiated histories:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $\textit{deq}(a)$ happens before $\textit{enq}(a)$ (or $\textit{enq}(a)$ does not exist) for some item $a$,
+
+\item[-] there are two or more $\textit{deq}(a)$ for some item $a$,
+
+\item[-] $\textit{enq}(a)$ happens before $\textit{enq}(b)$, and $\textit{deq}(b)$ happens before $\textit{deq}(a)$ (or $\textit{deq}(a)$ does not exists) for some items $a,b$.
+\end{itemize}
+
+It is not hard to see that (1) the first condition is dealt with by $\mathcal{A}_{\textit{unmRm1}}$ (if $\textit{put}(a,\_)$ does not exist) and $\mathcal{A}_{\textit{Rpr0-1}}$ (if $\textit{put}(a,\_)$ exists), (2) the second condition is dealt with by $\mathcal{A}_{\textit{unmRm2}}$, and (3) the third condition is dealt with by $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$.
+\qed
+\end {proof}
+
+By Lemma \ref{lemma:automata for priority queue with single priority}, from now on, it is safe for us to assume that, for each data-differentiated priority queue history $h$ and each $h'$ of projection of $h$, if $h'$ is single-priority, then $\textit{transToQueue}(h')$ is linearizable to queue.
+One consequence of this assumption is that, in $h'$, there does not exists an item $a$, such that $\textit{rm}(a)$ happens before $\textit{put}(a)$.
+
+By the proof of Lemma \ref{lemma:Rpr1 is co-regular}, with the help of $\mathcal{A}_{\textit{Rpr1-pprr}}, \mathcal{A}_{\textit{Rpr1-prpr}}, \mathcal{A}_{\textit{Rpr1-rppr}}$ and $\mathcal{A}_{\textit{Rpr1-rprp}}$, from now on, we can safely assume that given a data-differentiated history $h$ and operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ in $h$, there does not exists any items $d_1,\ldots,d_m$, such that $x$ is not covered by items $d_1,\ldots,d_m$ and the priority of $d_1,\ldots,d_m$ is less that $x$.
+
+However, above two assumptions are not enough for imply the linearizability to priority queue.
+One example is shown in \figurename~\ref{fig:history introduct ob order1}.
+The history $h$ of \figurename~\ref{fig:history introduct ob order1} is not linearizable, even if $h \vert_{1}$ and $h \vert_{2}$ are both linearizable, and either $a$ or $b$ is covered by other items.
+One explanation is that, since $\textit{put}(b,1)$ happens before $\textit{put}(a,1)$, in choosing linearization points, we should make $\textit{rm}(b)$ before $\textit{rm}(a)$.
+However, to satisfy the requirements of $R_{\textit{pr2}}$, the linearization points of $\textit{rm}(a)$ can only locates in the interval from $\textit{ret}(\textit{rm},d)$ to $\textit{cal}(\textit{put},e,2)$.
+But, $\textit{rm}(b)$ is after this entire interval, and we can not locate the linearization point of $\textit{rm}(b)$ in this interval.
+When there is no such phenomenon, such as the history in \figurename~\ref{fig:history introduct ob order2}, the history becomes linearizable.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER1.pdf}
+%\vspace{-10pt}
+  \caption{above assumptions do not implies linearizability (1)}
+  \label{fig:history introduct ob order1}
+\end{figure}
+
+%\begin{figure}[htbp]
+%  \centering
+%  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER3.pdf}
+%\vspace{-10pt}
+%  \caption{above assumptions do not implies linearizability (2)}
+%  \label{fig:history introduct ob order1}
+%\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER2.pdf}
+%\vspace{-10pt}
+  \caption{above assumptions do not implies linearizability (2)}
+  \label{fig:history introduct ob order2}
+\end{figure}
+
+In the following part of this subsection, we use $<_{\textit{ob}}$ order for the case when we can infer that an item is inserted earlier than another item, and we use gap-point to represent the intervals when it is possible to locate linearizaton points of $\textit{rm}(a)$.
+
+\begin{definition}[gap-point for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:gap-point for matched put and rm operations}
+
+Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x),\textit{rm}(x)$ of $h$, we say $o$ is a gap-point of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$, if we can obtain a sequence $\textit{gapPSeq}(h)$ from $h$ by adding $o$ to some locations between actions of $h$, and satisfies that,
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] In $\textit{gapPSeq}(h)$, $o$ is after $\textit{call}(\textit{put},x,\textit{pri}_x)$ and $\textit{call}(\textit{rm},x)$, and is before $\textit{ret}(\textit{rm},x)$.
+\item[-] For each item $y$ with non-$\textit{pri}_x$ priority in $\textit{gapPSeq}(h)$.
+If $\textit{put}(y,\_)$ and $\textit{rm}(y)$ is in $h$, then it can not be the case that $\textit{ret}(\textit{put},y,\_)$ is before $o$ and $\textit{call}(\textit{rm},x)$ is after $o$; Else if $\textit{put}(y,\_)$ is in $h$ and $\textit{rm}(y)$ is not in $h$, then it can not be the case that $\textit{ret}(\textit{put},y,\_)$ is before $o$;
+\end{itemize}
+
+We say that $o$ is a gap-point of priority $\textit{pri}_x$, if we can obtain a sequence $\textit{gapPSeq}(h)$ from $h$ by adding $o$ to some locations between actions of $h$, and satisfies only the latter requirement.
+\end{definition}
+
+\begin{restatable}{lemma}{GapEqualsGapPoint}
+\label{lemma:gap eqals the existence of gap point}
+Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
+There is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, if and only if there is a gap-point $o$ of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
+\end{restatable}
+
+\begin {proof}
+
+To prove the $\textit{only if}$ direction, we need to prove that: if there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, then there is no gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
+
+We prove this by contradiction.
+Assume that there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, but there is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
+Since there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, in the time interval starts from $\textit{cal}(\textit{put},x,\textit{pri}_x)$ and $\textit{cal}(\textit{rm},x)$ and ends at $\textit{ret}(\textit{rm},x)$, there is always some item with non-$\textit{pri}_x$ priority in priority queue.
+Therefore, it is not hard to prove that there is a cycle going through the left-right constraint of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, and by Lemma \ref{lemma:Gap Equals Constraint for Rpr0}, there is no gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, which contradicts that there is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
+
+To prove the $\textit{if}$ direction, assume that there is a gap-point $o$ of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$ and let $h_1$ be the sequence in definition of gap-point, which is generated from $h$ by adding $o$ into $h$.
+Let sequence $h_2$ be a sequence of call, return and operations, such that the projection of $h_2$ into call and return actions is $h$.
+$h_2$ is generated as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] First, from $h_1$, transform the point $o$ into $\textit{rm}(x)$.
+
+\item[-] Then for all the other items $y$, (1) adding an operation $\textit{put}(y,p)$ at some position between $\textit{call}(\textit{put},y,\_)$ and $\textit{ret}(\textit{put},y)$, and (2) adding an operation $\textit{rm}(y)$ at some position between $\textit{call}(\textit{rm},y)$ and $\textit{ret}(\textit{rm},y)$.
+
+\item[-] For all items $z$ with the same priority as $x$, the positions for adding operations of $z$ into $h$ is arbitrary (but need to satisfy above restriction.
+In the following we implicitly require this).
+
+\item[-] For all items $z$ whose priority is different from that of $x$,
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+
+    \item[-] If $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, and $\textit{ret}(\textit{put},z,\_)$ is before $o$ in $h_1$: Then in $h_2$, $\textit{rm}(z,\_)$ is just after $\textit{call}(\textit{rm},z)$.
+
+    \item[-] If $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, and $o$ is before $\textit{ret}(\textit{put},z,\_)$ $h_1$: Then in $h_2$, $\textit{put}(z,\_)$ is just before $\textit{ret}(\textit{put},z)$.
+
+    \item[-] Otherwise, we can see that $\textit{put}(z,\_)$ and $\textit{rm}(z)$ overlaps in $h$.
+Then we add $o_1=\textit{put}(z,\_)$ and $o_2=\textit{rm}(z)$ in the time interval of interleaving of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ overlaps in $h$, make sure $o_2$ next to $o_1$, and make sure either both of them is before $\textit{rm}(x)$ or both of them is after $\textit{rm}(x)$ in $h_2$.
+    \end{itemize}
+\end{itemize}
+
+According to the definition of gap-point, it is not hard to see that we can construct such $h_2$.
+Let $l$ be the projection of $h_2$ into operations, and $l'$ be the projection of $l$ into $x$ and all items that has priorities not equals that of $x$.
+Let $l'= u \cdot \textit{put}(x,\_) \cdot v \cdot \textit{rm}(x) \cdot w$.
+It is easy to see that $h \sqsubseteq l$, and $u \cdot v$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
+Let $L$, $M$ and $R$ be the set of operations in $u$, $v$ and $w$, respectively.
+Then it is obvious that $L$, $M$ and $R$ is a gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
+\qed
+\end {proof}
+
+Given a data-differentiated history $h$ and two items $a,b$, we say that $a <_{\textit{ob}} b$, if the priority of $a$ and $b$ is the same, and one of the following cases holds:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Type $A$: $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$ in $h$,
+
+\item[-] Type $B$: $\textit{rm}(a)$ happens before $\textit{rm}(b)$ in $h$,
+
+\item[-] Type $C$: $\textit{rm}(a)$ happens before $\textit{put}(b,\_)$ in $h$,
+\end{itemize}
+
+Let $<_{\textit{ob}}^*$ be the transitive closure of $<_{\textit{ob}}$.
+Intuitively, $a <_{\textit{ob}} b$ means that $a$ should be inserted earlier than $b$, and $a <_{\textit{ob}}^* b$ means that we can infer that $a$ should be inserted earlier than $b$ with the help of other items.
+We can explicitly write $a <_{\textit{ob}}^A b$ to mean that $a <_{\textit{ob}} b$ and the reason is that $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$.
+Similarly, we can explicitly write $a <_{\textit{ob}}^B b$ and $a <_{\textit{ob}}^C b$ to mean that $a <_{\textit{ob}} b$ and the reason is that $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$ and $\textit{rm}(a)$ happens before $\textit{put}(b,\_)$, respectively.
+
+According to the definition of $<_{\textit{ob}}^*$, if $a <_{\textit{ob}}^* b$, then there exists $c_1,\ldots,c_n$, such that $a <_{\textit{ob}} c_1 <_{\textit{ob}} \ldots <_{\textit{ob}} b$.
+The following lemma states that, the number of intermediate items $c_i$ is in fact bounded.
+
+\begin{restatable}{lemma}{OBOrderHasBoundedLength}
+\label{lemma:ob order has bounded length}
+Given a data-differentiated history $h$.
+Assume that $a <_{\textit{ob}} a_1 <_{\textit{ob}} \ldots <_{\textit{ob}} a_m <_{\textit{ob}} b$, then either $a <_{\textit{ob}} b$, or $a <_{\textit{ob}} a_i <_{\textit{ob}} b$ for some $i$, or $a <_{\textit{ob}} a_i <_{\textit{ob}} a_j <_{\textit{ob}} b$ for some $i,j$.
+\end{restatable}
+
+\begin {proof}
+
+Our proof proceed as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^A$,$<_{\textit{ob}}^B \cdot <_{\textit{ob}}^B$ and $<_{\textit{ob}}^C \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^A c_2 <_{\textit{ob}}^A c_1$, then $\textit{put}(c_3,\_)$ happens before $\textit{put}(c_2,\_)$, and $\textit{put}(c_2,\_)$ happens before $\textit{put}(c_1,\_)$.
+Therefore, it is obvious hat $\textit{put}(c_3,\_)$ happens before $\textit{put}(c_1,\_)$ and $c_3 <_{\textit{ob}}^A c_1$.
+
+    Similarly, if $c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^B c_1$, then $c_3 <_{\textit{ob}}^B c_1$.
+
+    If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^C c_1$: Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
+Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
+Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
+Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
+
+    Therefore, when we meet successive $<_{\textit{ob}}^A$, it is safe to leave only the first and the last elements and ignore intermediate elements.
+Similar cases hold for $<_{\textit{ob}}^B$ and $<_{\textit{ob}}^C$.
+
+\item[-] $<_{\textit{ob}}^A$ and $<_{\textit{ob}}^C$:
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^A c_2 <_{\textit{ob}}^C c_1$.
+Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
+Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
+Since $c_3 <_{\textit{ob}}^A c_2$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
+Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^A c_1$.
+
+    \item[-] ($<_{\textit{ob}}^C \cdot <_{\textit{ob}}^A$): If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^A c_1$.
+Since $c_2 <_{\textit{ob}}^A c_1$, $\textit{put}(c_2,\_)$ happens before $\textit{put}(c_1,\_)$.
+Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{rm}(c_3)$ happens before $\textit{put}(c_2,\_)$.
+Therefore, $\textit{rm}(c_3)$ happens before $\textit{put}(c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
+    \end{itemize}
+
+\item[-] $<_{\textit{ob}}^B$ and $<_{\textit{ob}}^C$:
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] ($<_{\textit{ob}}^B \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^C c_1$.
+Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
+It is obvious that $\textit{cal}(\textit{rm},c_2)$ is before $\textit{ret}(\textit{rm},c_2)$.
+Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$.
+Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
+
+    \item[-] ($<_{\textit{ob}}^C \cdot <_{\textit{ob}}^B$): If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^B c_1$.
+Since $c_2 <_{\textit{ob}}^B c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{rm},c_1)$.
+Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
+Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
+Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_1)$, and $c_3 <_{\textit{ob}}^B c_1$.
+    \end{itemize}
+
+\item[-]  ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^B \cdot <_{\textit{ob}}^A$): If $c_4 <_{\textit{ob}}^A c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^A c_1$:
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$: Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$.
+Then $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
+This implies that $c_4 <_{\textit{ob}}^A c_3 <_{\textit{ob}}^C c_1$.
+According to the fact for $<_{\textit{ob}}^A \cdot <_{\textit{ob}}^C$, we know that $c_4  <_{\textit{ob}}^A c_1$.
+
+    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is after $\textit{cal}(\textit{put},c_1,\_)$: Since $c_2 <_{\textit{ob}}^A c_1$, $\textit{ret}(\textit{put},c_2,\_)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
+Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{rm}(c_3)$ happens before $\textit{rm}(c_2)$, and then we know that $\textit{put}(c_2,\_)$ can not happen before $\textit{put}(c_3,\_)$.
+Since $\textit{put}(c_2,\_)$ does not happen before $\textit{put}(c_3,\_)$, $\textit{cal}(\textit{put},c_3,\_)$ is before $\textit{ret}(\textit{put},c_2,\_)$.
+Since $c_4 <_{\textit{ob}}^A c_3$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_3,\_)$.
+Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_4 <_{\textit{ob}}^A c_1$.
+    \end{itemize}
+
+\end{itemize}
+
+Based on above results, given $a <_{\textit{ob}}^{b_1} a_1 <_{\textit{ob}} \ldots <_{\textit{ob}}^{b_m} a_m <_{\textit{ob}}^{b_{\textit{m+1}}} b$, where each $b_i$ is in $\{ A,B,C \}$, we can merge relations, until we got one of the following facts:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $a <_{\textit{ob}}^A b$, $a <_{\textit{ob}}^B b$ or $a <_{\textit{ob}}^C b$,
+
+\item[-] $a <_{\textit{ob}}^A a_i <_{\textit{ob}}^B b$, or $a <_{\textit{ob}}^B a_i <_{\textit{ob}}^A b$, for some $i$,
+
+\item[-] $a <_{\textit{ob}}^B a_i <_{\textit{ob}}^A a_j <_{\textit{ob}}^B b$, for some $i$ and $j$,
+\end{itemize}
+
+This completes the proof of this lemma. \qed
+\end {proof}
+
+Given a differentiated history $h$ and operations $\textit{put}(x,\_)$ and $\textit{rm}(x)$ of $h$, the time interval from the time point when both $\textit{call}(\textit{put},x,\_)$ and $\textit{cal}(\textit{rm},x)$ happen and ends at $\textit{ret}(\textit{rm},x)$ is called the effective interval of $x$ in $h$.
+
+\begin{restatable}{lemma}{Rpr2AsGPandOBOrder}
+\label{lemma:Rpr2 as gap-point and ob order}
+Given a data-differentiated history $h$.
+There exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr2}}$, and $h'$ does not linearizable to $\textit{MR}_{\textit{pr2}}$, if and only if there exists $h'' \in \textit{proj}(h)$, $\textit{last}(h'') = R_{\textit{pr2}}$, $x$ and $y$ has maximal priority in $h''$, $y <_{\textit{ob}} x$ in $h''$, and the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$ in $h''$.
+\end{restatable}
+
+\begin {proof}
+
+To prove the $\textit{if}$ direction, let $h_1$ be the history that is obtained from $h''$ by erasing all operations of items that has same priority as $x$, except for operations of $x$ and $y$.
+It is obvious that $\textit{last}(h_1) = R_{\textit{pr2}}$.
+According to $y <_{\textit{ob}} x$,
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If $y <_{\textit{ob}}^A x$, then $\textit{put}(y,\_)$ happens before $\textit{put}(x,\_)$.
+Since in $R_{\textit{pr2}}$, $\textit{put}(\textit{itm},\_)$ is after $\textit{put}$ operations of items with maximal priority (except for $\textit{itm}$), we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
+\item[-] If $y <_{\textit{ob}}^B x$, then $\textit{rm}(y)$ happens before $\textit{rm}(x)$.
+Since in $R_{\textit{pr2}}$, $\textit{rm}(\textit{itm})$ is the last $\textit{rm}$ operation of items with maximal priority, we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
+
+\item[-] If $y <_{\textit{ob}}^C x$, then $\textit{rm}(y)$ happens before $\textit{put}(x,\_)$.
+Since in $R_{\textit{pr2}}$, $\textit{rm}(\textit{itm})$ is the last $\textit{rm}$ operation of items with maximal priority, we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
+\end{itemize}
+
+Similarly as the proof in Lemma \ref{lemma:gap eqals the existence of gap point}, we can see that the only possible position for locating linearizaton point of $\textit{rm}(x)$ is at gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$.
+Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, then there exists unmatched operations before $\textit{rm}(x)$.
+
+Since the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},x)$ in $h''$ (and also in $h_1$), if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, then $\textit{rm}(x)$ will be before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},x)$.
+
+Therefore, for every sequence $l = u \cdot \textit{put}(x,\_) \cdot v \cdot \textit{rm}(x) \cdot w$, if $h_1 \sqsubseteq l$, then either $u \cdot v$ contains some unmatched operations of priority less than priority of $x$, or $w$ contains $\textit{put}(y,\_)$ or $\textit{rm}(y)$.
+In both cases, $l \notin \textit{MR}_{\textit{pr2}}$.
+
+To prove the $\textit{only if}$ direction, we prove its contrapositive.
+Assume we already know that for each $h'' \in \textit{proj}(h)$ with $\textit{last}(h'') = R_{\textit{pr2}}$, for each $x$ and $y$ has maximal priority in $h''$, if $y <_{\textit{ob}} x$ in $h''$, then the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},x)$ in $h''$.
+We need to prove that for each $h' \in \textit{proj}(h)$, if $\textit{last}(h') = R_{\textit{pr2}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr2}}$.
+
+Let $\textit{pri}$ be the maximal priority of $h'$.
+Recall that we already assume that each single-priority history (if we consider $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$) can be ``linearizable to queue'', and item with larger priority is not covered by items with smaller priority.
+Our proof proceed as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Let $h_{\textit{pri}}$ be the projection of $h$ into operations of priority $\textit{pri}$.
+Since each single-priority history can be ``linearizable to queue'', there exists sequence $l_{\textit{pri}}$, such that $h_{\textit{pri}} \sqsubseteq l_{\textit{pri}}$, and when we treat $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$, $l_{\textit{pri}}$ belongs to queue.
+
+\item[-] Step $1$: Let $a_1$ be the last inserted item of $l_{\textit{pri}}$.
+
+$\textit{Cond1}$: If for each item $b$ with priority $\textit{pri}$ in $h'$, the rightmost gap-point $o$ of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$ does not before any $\textit{cal}(\textit{put},b,\textit{pri})$ or $\textit{cal}(\textit{rm},b)$ in $h'$, then we can
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] Locate the linearization point of $\textit{rm}(a_1)$ at $o$,
+    \item[-] If $\textit{put}(a_1,\textit{pri})$ overlaps with $\textit{rm}(a_1)$, then locate the linearization point of $\textit{put}(a_1,\_)$ just before the linearization point of $\textit{rm}(a_1)$.
+Otherwise, $\textit{put}(a_1,\textit{pri}) <_{\textit{hb}} \textit{rm}(a_1)$, and we locate the linearization point of $\textit{put}(a_1,\_)$ just before its return action.
+    \item[-] Locate linearization points of operations of items of priority $\textit{pri}$ (except for $a_1$) just after the call action of the operation.
+    \item[-] For item $z$ with priority smaller than $\textit{pri}$.
+If both $\textit{cal}(\textit{put},z,\_)$ and $\textit{cal}(\textit{rm},z)$ is before $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just after their call actions.
+If both $\textit{ret}(\textit{put},z)$ and $\textit{ret}(\textit{rm},z)$ is after $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just before their return actions.
+
+        Otherwise, either (1) $\textit{rm}(z) <_{\textit{hb}} \textit{put}(z,\_)$, or (2) $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, $\textit{ret}(\textit{put},z,\_)$ before $o$, and $o$ before $\textit{call}(\textit{rm},z)$.
+The first case violates that each single-priority history can be ``linearizable to queue'', and the second case violates that $o$ is a gap-point.
+    \end{itemize}
+
+    In this way we construct a sequence $l$, and it is easy to see that $h' \sqsubseteq l$ and $l$ satisfies the requirement of $R_{\textit{pr2}}$.
+
+\item[-] Otherwise, there exists $a_2$ with priority $\textit{pri}$, such that the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$ in $h'$.
+We can see that the rightmost gap-point of $\textit{put}(a_2,\textit{pri})$ and $\textit{rm}(a_2)$ is after the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$.
+By assumption we know that $a_2$ does not $<_{\textit{ob}}$ to $a_1$.
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If for each item $b$ with priority $\textit{pri}$ in $h'$, $a_2$ does not $<_{\textit{ob}}$ to $b$.
+Then we go to step $1$ and treat $a_2$ as $a_1$ (When $\textit{Cond1}$ of step $1$ holds, we break the loop).
+    \item[-] Otherwise, there exists $a_3$ with priority $\textit{pri}$ such that $a_2 <_{\textit{ob}} a_3$.
+By assumption,we know that the rightmost gap-point of $\textit{put}(a_3,\textit{pri})$ and $\textit{rm}(a_3)$ is not before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$.
+Then we can see that the rightmost gap-point of $\textit{put}(a_3,\textit{pri})$ and $\textit{rm}(a_3)$ is after the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$.
+Then we go to step $1$ and treat $a_3$ as $a_1$ (When $\textit{Cond1}$ of step $1$ holds, we break the loop).
+    \end{itemize}
+\end{itemize}
+
+It is not hard to see that, if $b_i$ is the $a_1$ in the $\textit{i-th}$ loop of our proof, $b_j$ is the $a_1$ in the $\textit{j-th}$ loop of our proof, and $i<j$, then the rightmost gap-point of $\textit{put}(b_j,\textit{pri})$ and $\textit{rm}(b_j)$ is after the rightmost gap-point of $\textit{put}(b_i,\textit{pri})$ and $\textit{rm}(b_i)$.
+Thus, we finally stop at some $d$.
+$d$ satisfies that, for each item $c$ with priority $\textit{pri}$, $d$ does not $<_{\textit{ob}}$ to $c$.
+$d$ also satisfies that, for each item $c$ with priority $\textit{pri}$ in $h'$, the rightmost gap-point $o'$ of $\textit{put}(d,\textit{pri})$ and $\textit{rm}(d)$ does not before any $\textit{cal}(\textit{put},c,\textit{pri})$ or $\textit{cal}(\textit{rm},c)$.
+Then, we can
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Locate the linearization point of $\textit{rm}(d)$ at the rightmost gap-point $o'$ of $\textit{put}(d,\textit{pri})$ and $\textit{rm}(d)$,
+
+\item[-] Since for each item $c$ with priority $\textit{pri}$ in $h'$, $o'$ is after any $\textit{cal}(\textit{put},c,\textit{pri})$ and $\textit{cal}(\textit{rm},c)$.
+For each $c$, one of the following cases hold: (1) $\textit{cal}(\textit{rm},c)$ is before $o'$ and $o'$ is before $\textit{ret}(\textit{rm},c)$, (2) $\textit{cal}(\textit{rm},c)$ and $\textit{ret}(\textit{rm},c)$ are before $o'$.
+We locate the linearization point of $\textit{rm}(c)$ just after its call action.
+Such linearization point is obviously before $o'$.
+
+\item[-] If $o'$ is before $\textit{ret}(\textit{put},d,\textit{pri})$, then locate the linearization point of $\textit{put}(d,\textit{pri})$ just before $o'$.
+Else, $\textit{ret}(\textit{put},d,\textit{pri})$ is after $o'$, and we locate the linearization point of $\textit{put}(d,\textit{pri})$ just before the return action of it.
+
+\item[-] If $o'$ is before $\textit{ret}(\textit{put},d,\textit{pri})$, we can see that, for each item $c$ with priority $\textit{pri}$ in $h'$, $\textit{cal}(\textit{put},c,\textit{pri})$ is before $o'$.
+Then, and we locate the linearization point of $\textit{put}(c,\textit{pri})$ just before $o'$ and the linearization point of $\textit{put}(d,\textit{pri})$.
+
+    Otherwise, $\textit{ret}(\textit{put},d,\textit{pri})$ is before $o'$.
+Since $d$ does not $<_{\textit{ob}}$ to $c$, $\textit{put}(d,\textit{pri})$ does not happen before $\textit{put}(c,\textit{pri})$, or we can say, $\textit{ret}(\textit{put},d,\textit{pri})$ is after $\textit{cal}(\textit{put},c,\textit{pri})$.
+Then, we locate the linearization point of $\textit{put}(c,\textit{pri})$ just after its call action.
+\end{itemize}
+
+In this way we construct a sequence $l_d$, and it is easy to see that $h' \sqsubseteq l_d$ and $l_d$ satisfies the requirement of $R_{\textit{pr2}}$ with $d$ acts as $\textit{itm}$.
+This completes the proof of $\textit{if}$ direction.
+\qed
+\end {proof}
+
+Let us begin to represent several automata that is used for capture the case that, in a sub-history $h''$ of $h$, $\textit{last}(h'') = R_{\textit{pr2}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr2}}$, and the reason is that there exists items $a$ and $b$ with maximal priority in $h''$, $a <_{\textit{ob}} b$ in $h''$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
+
+According to Lemma \ref{lemma:ob order has bounded length}, if $a <_{\textit{ob}} b$, then one of the following cases hold: (1) $a <_{\textit{ob}}^A b$, (2) $a <_{\textit{ob}}^B b$, (3) $a <_{\textit{ob}}^C b$, (4) $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$ for some $a_1$, (5) $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$ for some $a_1$, or (6) $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$ for some $a_1$ and $a_2$.
+Although there are many possible enumeration of $a$, $b$, $a_1$ and $b_1$, the following lemma states that we need only to consider five enumeration of operations of $a$, $b$, $a_1$ and $b_1$.
+
+\begin{restatable}{lemma}{FiveEnmuerationisEnoughForRpr2}
+\label{lemma:five enumeration is enough for Rpr2}
+Given a data-differentiated history $h$, $\textit{last}(h) = R_{\textit{pr2}}$, and $h$ does not linearizable to $\textit{MR}_{\textit{pr2}}$.
+Assume that there exists items $a$ and $b$ with maximal priority in $h''$, $a <_{\textit{ob}} b$, and the rightmost gap-point $o$ of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
+Let $a_1$ and $a_2$ be the value when possibly $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$, or $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$.
+Then, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists) and $a_2$ (if exists).
+\end{restatable}
+\begin {proof}
+
+Let us prove by consider all the possible reason of $a <_{\textit{ob}}^A b$.
+According to Lemma \ref{lemma:ob order has bounded length}, we need to consider six reasons: %There may exists items with smaller priority of $a$ and covers some time interval, and assume that we rename them into $d$:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Reason $1$, $a <_{\textit{ob}}^A b$:
+
+    Since $a <_{\textit{ob}}^A b$, $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$.
+Since $o$ is after $\textit{cal}(\textit{put},a,\_)$, we can see that $o$ is before $\textit{cal}(\textit{rm},b)$.
+
+    Since single-priority must satisfy the first in firs out property, $\textit{rm}(b)$ does not happen before $\textit{rm}(a)$, and $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
+If $\textit{ret}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$, since we already assume that there exists gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$, this gap-point is also a gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$, and is after $o$, which contradicts that $o$ is the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$.
+Therefore, $\textit{ret}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$.
+
+    According to above discussion, there are two possible enumeration of operations of $a$ and $b$, as shown in \figurename~\ref{fig:history enumeration 1 for Rpr2} and \figurename~\ref{fig:history enumeration 2 for Rpr2}.
+Here we explicitly draw the leftmost gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$ as $o'$.
+Since the position of $\textit{ret}(\textit{put},b)$ does not influence the correctness, we can simply ignore it.
+
+\item[-] Reason $2$, $a <_{\textit{ob}}^B b$:
+
+    Since $a <_{\textit{ob}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$.
+Since $o$ is after $\textit{cal}(\textit{rm},b)$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$.
+Then, we can see that $o$ is before $\textit{cal}(\textit{put},a,\_)$.
+This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\_)$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible.
+Therefore, we can safely ignore this reason.
+
+\item[-] Reason $3$, $a <_{\textit{ob}}^C b$:
+
+    Since $a <_{\textit{ob}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},b)$.
+Since $o$ is after $\textit{cal}(\textit{put},b)$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$.
+Then, we can see that $o$ is before $\textit{cal}(\textit{put},a,\_)$.
+This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\_)$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible.
+Therefore, we can safely ignore this reason.
+
+\item[-] Reason $4$, $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$:
+
+    Since $a_1 <_{\textit{ob}}^B b$, $\textit{rm}(a_1) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_1)$ is before $\textit{cal}(\textit{rm},b)$.
+Since $\textit{rm}(a_1)$ does not happen before $\textit{put}(a_1)$, $\textit{cal}(\textit{put},a_1,\_)$ is before $\textit{ret}(\textit{rm},a_1)$.
+Since $a <_{\textit{ob}}^A a_1$, $\textit{ret}(\textit{put},a,\_)$ is before $\textit{cal}(\textit{put},a_1,\_)$.
+Therefore, $\textit{ret}(\textit{put},a,\_)$ is before $\textit{cal}(\textit{rm},b)$.
+Since $\textit{cal}(\textit{rm},b)$ is before $o$, we can see that $\textit{ret}(\textit{put},a,\_)$ is before $o$.
+Then, we can see that $o$ is before $\textit{cal}(\textit{rm},a)$.
+
+    If $\textit{cal}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$, then $\textit{rm}(b) <_{\textit{hb}} \textit{rm}(a)$.
+Let $h' = h \vert_{ \{ a,a_1,b \} }$.
+It is not hard to see that $h'$ can not linearizable to a sequence that satisfies the first in first our property, and this contradicts to our assumption that each single-priority history satisfies the first in first out property.
+Therefore, $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
+Similarly as the case of reason $1$, we can see that $\textit{ret}(\textit{rm},b)$ is before $\textit{ret}(\textit{rm},a)$.
+
+    According to above discussion, there are three possible enumeration of operations of $a$, $a_1$ and $b$, as shown in \figurename~\ref{fig:history enumeration 3 for Rpr2}, \figurename~\ref{fig:history enumeration 4 for Rpr2} and \figurename~\ref{fig:history enumeration 5 for Rpr2}.
+Here we explicitly draw the leftmost gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$ as $o'$.
+Since the position of $\textit{cal}(\textit{put},b,\_)$, $\textit{ret}(\textit{put},b)$, $\textit{ret}(\textit{put},a_1,\_)$ and $\textit{cal}(\textit{put},a,\_)$ do not influence the correctness, we can simply ignore it.
+
+\item[-] Reason $5$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$:
+
+    Since $a_1 <_{\textit{ob}}^A b$, $\textit{ret}(\textit{put},a_1,\_)$ is before $\textit{call}(\textit{put},b,\_)$.
+Since $\textit{call}(\textit{put},b,\_)$ is before $o$, we can see that $\textit{ret}(\textit{put},a_1,\_)$ is before $o$.
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $o$ is before $\textit{cal}(\textit{rm},a)$: Then $o$ is obviously before $\textit{ret}(\textit{rm},a)$.
+Since $a <_{\textit{ob}}^B a_1$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},a_1)$.
+Then we can see that, $o$ is before $\textit{cal}(\textit{rm},a_1)$, and remember that $a_1 <_{\textit{ob}}^A b$.
+Then we can goto the case of reason $1$ and treat $a_1$ as $a$.
+Therefore, we can safely ignore this.
+
+    \item[-] If $o$ is before $\textit{cal}(\textit{put},a,\_)$: Since $\textit{rm}(a)$ does not happen before $\textit{put}(a,\_)$, we can see that $\textit{cal}(\textit{put},a,\_)$ is before $\textit{ret}(\textit{rm},a)$, and then $o$ is before $\textit{ret}(\textit{rm},a)$.
+Then similarly as above case, we can see that $o$ is before $\textit{cal}(\textit{rm},a_1)$, and $a_1 <_{\textit{ob}}^A b$.
+Then we can goto the case of reason $1$ and treat $a_1$ as $a$.
+Therefore, we can safely ignore this.
+    \end{itemize}
+
+\item[-] Reason $6$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$:
+
+    Since $a_2 <_{\textit{ob}}^B b$, $\textit{rm}(a_2) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_2)$ is before $\textit{cal}(\textit{rm},b)$.
+%Since $\textit{rm}(a_2)$ does not happen before $\textit{put}(a_2,\_)$, $\textit{cal}(\textit{put},a_2,\_)$ is before $\textit{ret}(\textit{rm},a_2)$.
+%Since $a_1 <_{\textit{ob}}^A a_2$, $\textit{ret}(\textit{put},a_1,\_)$ is before $\textit{cal}(\textit{put},a_2,\_)$.
+    Since $a_1 <_{\textit{ob}}^A a_2$, $\textit{put}(a_1,\_) <_{\textit{hb}} \textit{put}(a_2,\_)$, and $\textit{rm}(a_2)$ does not happen before $\textit{rm}(a_1)$, and $\textit{cal}(\textit{rm},a_1)$ is before $\textit{ret}(\textit{rm},a_2)$.
+Since $a <_{\textit{ob}}^B a_1$, $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(a_1)$, and $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},a_1)$.
+
+    Therefore, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$.
+Since $\textit{cal}(\textit{rm},b)$ is before $o$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$.
+Then it is easy to see that $o$ is before $\textit{cal}(\textit{put},a,\_)$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\_)$, and $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a,\_)$.
+This is impossible, and we can safely ignore this reason.
+\end{itemize}
+
+This completes the proof of this lemma. \qed
+\end {proof}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-1.pdf}
+%\vspace{-10pt}
+  \caption{The first possible enumeration.}
+  \label{fig:history enumeration 1 for Rpr2}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-2.pdf}
+%\vspace{-10pt}
+  \caption{The second possible enumeration.}
+  \label{fig:history enumeration 2 for Rpr2}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-3.pdf}
+%\vspace{-10pt}
+  \caption{The third possible enumeration.}
+  \label{fig:history enumeration 3 for Rpr2}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-4.pdf}
+%\vspace{-10pt}
+  \caption{The forth possible enumeration.}
+  \label{fig:history enumeration 4 for Rpr2}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-5.pdf}
+%\vspace{-10pt}
+  \caption{The fifth possible enumeration.}
+  \label{fig:history enumeration 5 for Rpr2}
+\end{figure}
+
+Given a data-differentiated history $h$, two actions $\textit{act}_1$, $\textit{act}_2$ of items of priority $\textit{pri}$, and assume that $\textit{act}_1$ is before $\textit{act}_2$ in $h$.
+We say that $\textit{act}_1$, $\textit{act}_2$ is covered by items $d_1,\ldots,d_m$ in $h$, if the priorities of $d_1,\ldots,d_m$ is not $\textit{pri}$, and
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $\textit{ret}(\textit{put},d_m,\_)$ is before $\textit{act}_1$,
+
+\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,
+
+\item[-] $\textit{act}_2$ is before $\textit{cal}(\textit{rm},d_1)$.
+\end{itemize}
+
+According to the definition of gap-point, in \figurename~\ref{fig:history enumeration 1 for Rpr2}, \figurename~\ref{fig:history enumeration 2 for Rpr2}, \figurename~\ref{fig:history enumeration 3 for Rpr2}, \figurename~\ref{fig:history enumeration 4 for Rpr2} or \figurename~\ref{fig:history enumeration 5 for Rpr2}, the location of $o$ and $o'$ is equivalent to that, there exists items $d_1,\ldots,d_m$ in $h$ in $h$, such that $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ is covered by $d_1,\ldots,d_m$, $\textit{cal}(\textit{rm},b)$ is before $\textit{ret}(\textit{put},d_m,\_)$ in $h$, and $\textit{cal}(\textit{rm},d_1)$ is before $\textit{ret}(\textit{rm},a)$ in $h$.
+We say that such $d_1,\ldots,d_m$ constitute the rightmost gap of $\textit{put}(b,\_)$ and $\textit{rm}(b)$.
+
+An automaton $\mathcal{A}_{\textit{Rpr2-1}}$ is given in \figurename~\ref{fig:automata for first enumeration of Rpr2}, and it is constructed for the first enumeration in \figurename~\ref{fig:history enumeration 1 for Rpr2}.
+In \figurename~\ref{fig:automata for first enumeration of Rpr2}, we renames the items that covers $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ into $d$, and rename the renaming items into $e$.
+Here let $c = \textit{cal}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{cal}(\textit{rm},e), \textit{ret}(\textit{rm},e)$, $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_1.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{Rpr2-1}}$}
+  \label{fig:automata for first enumeration of Rpr2}
+\end{figure}
+
+An automaton $\mathcal{A}_{\textit{Rpr2-2}}$ is given in \figurename~\ref{fig:automata for second enumeration of Rpr2}, and it is constructed for the second enumeration in \figurename~\ref{fig:history enumeration 2 for Rpr2}.
+In \figurename~\ref{fig:automata for second enumeration of Rpr2}, $c_1$, $c_2$, $c_3$ and $c_4$ is same as that in \figurename~\ref{fig:automata for first enumeration of Rpr2}.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_2.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{Rpr2-2}}$}
+  \label{fig:automata for second enumeration of Rpr2}
+\end{figure}
+
+For the third enumeration in \figurename~\ref{fig:history enumeration 3 for Rpr2}.
+Since we want to ensure that $a$ and $b$ are putted only once, we need to explicitly record the positions of $\textit{cal}(\textit{put},a,p)$ and $\textit{cal}(\textit{put},b,p)$.
+Since the positions of $\textit{cal}(\textit{put},a,p)$ and $\textit{cal}(\textit{put},b,p)$ are not fixed, there are finite possible cases to consider, as shown below:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{rm},b)$ and before $\textit{cal}(\textit{rm},a)$: There are two possible positions of $\textit{cal}(\textit{put},a,p)$: (1) before $\textit{cal}(\textit{rm},a_1)$, and (2) after $\textit{cal}(\textit{rm},a_1)$, and before $\textit{ret}(\textit{put},a)$.
+
+\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$ and before $\textit{cal}(\textit{rm},b)$: same as above case.
+
+\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{put},a_1,\_)$ and before $\textit{ret}(\textit{rm},a_1)$: same as above case.
+
+\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{put},a)$ and before $\textit{cal}(\textit{put},a_1,p)$: same as above case.
+
+\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$: There are three possible positions of $\textit{cal}(\textit{put},a,p)$: (1) after $\textit{cal}(\textit{put},b,p)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{cal}(\textit{put}b,p)$, and (3) before $\textit{call}(\textit{rm},a_1)$.
+
+\item[-] If $\textit{cal}(\textit{put},b,p)$ is before $\textit{cal}(\textit{rm},a_1)$: There are three possible positions of $\textit{cal}(\textit{put},a,p)$: (1) after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{cal}(\textit{put},b,p)$ and before $\textit{cal}(\textit{rm},a_1)$, and (3) before $\textit{cal}(\textit{put},b,p)$.
+\end{itemize}
+
+Therefore, there are fourteen possible cases that satisfy the third enumeration in \figurename~\ref{fig:history enumeration 3 for Rpr2}.
+For each case, we construct an finite automaton.
+Let $\textit{Aut}_{\textit{Rpr2-3}}$ be the set of finite automata that is constructed for above fourteen cases.
+For example, for the case $\textit{ca}_1$ when $\textit{cal}(\textit{put},a,p)$ is before $\textit{cal}(\textit{rm},a_1)$, $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$, and $\textit{cal}(\textit{put},b,p)$ is before $\textit{cal}(\textit{rm},b)$, we construct a finite automaton in \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}.
+In \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}, let $c$ and $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$ the same as that in \figurename~\ref{fig:automata for first enumeration of Rpr2}.
+Let $c_2 = c_1 + \textit{ret}(\textit{put},a_1)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$, $c_4 = c_3 + \textit{ret}(\textit{rm},d)$, and $c_5 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{put},a_1) + \textit{ret}(\textit{rm},d)$.
+The other finite automata in $\textit{Aut}_{\textit{Rpr2-3}}$ can be similarly constructed.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_3.pdf}
+%\vspace{-10pt}
+  \caption{Automaton for $\textit{ca}_1$}
+  \label{fig:automata for ca1 of third enumeration of Rpr2}
+\end{figure}
+
+Similarly, we construct sets $\textit{Aut}_{\textit{Rpr2-4}}$ and $\textit{Aut}_{\textit{Rpr2-5}}$ of finite automata for the forth enumeration in \figurename~\ref{fig:history enumeration 4 for Rpr2} and the fifth enumeration in \figurename~\ref{fig:history enumeration 5 for Rpr2}, respectively.
+
+\begin{restatable}{lemma}{Rpr2IsCoRegular}
+\label{lemma:Rpr2 is co-regular}
+$R_{\textit{pr2}}$ is co-regular.
+\end{restatable}
+
+\begin {proof}
+
+Let $\textit{AutS-R}_{\textit{pr2}} = \{ \mathcal{A}_{\textit{Rpr2-1}}, \mathcal{A}_{\textit{Rpr2-2}} \} \cup \textit{Aut}_{\textit{Rpr2-3}} \cup \textit{Aut}_{\textit{Rpr2-4}} \cup \textit{Aut}_{\textit{Rpr2-5}}$.
+We need to prove that:
+
+\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr2}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr2}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr2}}$.
+
+By Lemma \ref{lemma:Rpr2 as gap-point and ob order}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:
+
+\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr0}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, such that $\textit{last}(h') = R_{\textit{pr2}}$, $a$ and $b$ has maximal priority in $h'$, $a <_{\textit{ob}} b$ in $h'$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
+
+\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\textit{AutS-R}_{\textit{pr2}}$.
+By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$.
+Since $h_1$ is accepted by $\textit{AutS-R}_{\textit{pr2}}$, let $x$ and $y$ be the items that are renamed into $b$ and $a$ by $r$, respectively, and let $d_1,\ldots,d_m$ be the items that are renamed into $d$ by $r$.
+
+let $h'' = h_2 \vert_{ \{ x,y,d_1,\ldots,d_m \} }$.
+It is obvious that $\textit{last}(h'') = R_{\textit{pr2}}$.
+According to our construction of automata in $\textit{AutS-R}_{\textit{pr2}}$, it is not hard to see that $x$ and $y$ has maximal priority in $h_2$, $y <_{\textit{ob}} x$ in $h''$, and the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$ in $h''$.
+%By Lemma \ref{lemma:Rpr2 as gap-point and ob order}, we know that $h''$ is not linearizable to $\textit{MR}_{\textit{pr2}}$.
+
+\noindent The $\textit{if}$ direction: Assume that there exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr2}}$, $a'$ and $b'$ has maximal priority in $h'$, $a' <_{\textit{ob}} b'$ in $h'$, and the rightmost gap-point of $\textit{put}(b',\_)$ and $\textit{rm}(b')$ is before $\textit{cal}(\textit{put},a',\_)$ or $\textit{cal}(\textit{rm},a')$ in $h''$.
+By data-independence, we can obtain history $h_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the items $d_1,\ldots,d_m$ that constitute the rightmost gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, we rename them into $d$, and (3) rename the other items into $e$.
+It is easy to see that $\textit{last}(h_1) = R_{\textit{pr2}}$, $a$ and $b$ has maximal priority in $h_1$, $a <_{\textit{ob}} b$ in $h_1$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
+By Lemma \ref{lemma:five enumeration is enough for Rpr2}, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists) and $a_2$ (if exists).
+Then
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If $a <_{\textit{ob}} b$ because of the first enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr2-1}}$.
+
+\item[-] If $a <_{\textit{ob}} b$ because of the second enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr2-2}}$.
+
+\item[-] If $a <_{\textit{ob}} b$ because of the third enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-3}}$.
+
+\item[-] If $a <_{\textit{ob}} b$ because of the forth enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-4}}$.
+
+\item[-] If $a <_{\textit{ob}} b$ because of the fifth enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-5}}$.
+\end{itemize}
+
+This completes the proof of this lemma. \qed
+\end {proof}
+
+\subsection{Co-Regular of $R_{\textit{pr3}}$}
+\label{subsec:co-regular of Rpr3}
+
+\begin{restatable}{lemma}{Rpr3IsAlwaysCoRegular}
+\label{lemma:Rpr3 is always co-regular}
+Given a data-differentiated history $h$, if $\textit{last}(h) = R_{\textit{pr3}}$, then $h \sqsubseteq \textit{MR}_{\textit{pr3}}$.
+\end{restatable}
+
+\begin {proof}
+
+Since $\textit{last}(h) = R_{\textit{pr3}}$, there exists item $x$ with $\textit{pri}_x$ in $h$, such that $h \vert_{ \{ x \}} = \textit{put}(x,\textit{pri}_x)$, and $\textit{pri}_x$ is larger than any priority of $h$.
+
+For each operation, we locate its linearization point just after its call action.
+In such way we obtain a sequence $l$ of operations, and it is easy to see that $h \sqsubseteq l$.
+It is obvious that $l$ contains $\textit{put}(x,\textit{pri})$, and then $l \in \textit{MR}_{\textit{pr3}}$.
+This completes the proof of this lemma.
+\qed
+\end {proof}
+
+\subsection{Co-Regular of $R_{\textit{pr4}}$}
+\label{subsec:co-regular of Rpr4}
+
+\begin{restatable}{lemma}{Rp4AsHappenBefore}
+\label{lemma:Rpr4 as happen before}
+Given a data-differentiated history $h$.
+There exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr4}}$, and $h'$ does not linearizable to $\textit{MR}_{\textit{pr4}}$, if and only if there exists $h'' \in \textit{proj}(h)$, $\textit{last}(h'') = R_{\textit{pr4}}$, $x$ and $y$ has maximal priority in $h''$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,\_) <_{\textit{hb}} \textit{put}(y,\_)$.
+\end{restatable}
+
+\begin {proof}
+
+To prove the $\textit{if}$ direction, let $h_1 = h \vert_{ \{ x,y \} }$. It is easy to see that $\textit{last}(h') = R_{\textit{pr4}}$ and $h_1$ does not linearizable to $\textit{MR}_{\textit{pr4}}$.
+
+To prove the $\textit{only if}$ direction, we prove its contrapositive.
+Assume that for each $h'' \in \textit{proj}(h)$, such that $\textit{last}(h'') = R_{\textit{pr4}}$, for each $x$ and $y$ has maximal priority in $h''$, if $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, then $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$.
+We need to prove that, for each $h' \in \textit{proj}(h)$, if $\textit{last}(h') = R_{\textit{pr4}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr4}}$.
+
+Given such $h'$, let $x_1,\ldots,x_m$ be the set of items with maximal priority and has unmatched $\textit{put}$ operations, let $y_1,\ldots,y_n$ be the set of items with maximal priority and has matched $\textit{put}$ and $\textit{rm}$ operations.
+It is easy to see that, for each $x_i$ and $y_j$, $\textit{cal}(\textit{put},y_j)$ is before $\textit{ret}(\textit{put},x_i,\_)$.
+Then we locate the linearization points of $h'$ as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] For each $x_i$, locate the linearization point of $\textit{put}(x_i,\_)$ just before its return action.
+
+\item[-] For each $y_j$, locate the lineariztion point of $\textit{put}(y_j,\_)$ jest after its call action.
+
+\item[-] For other operations, locate their linearization points at an arbitrary location after its call action and before its return action.
+\end{itemize}
+
+Let $l$ be the sequence of linearization points.
+It is easy to see that $h \sqsubseteq l$, and $l \in \textit{MR}_{\textit{pr4}}$.
+Here the $\textit{itm}$ in $R_{\textit{pr4}}$ is chosen to be $\textit{ret}(\textit{put},x_{\textit{ind}},\_)$, where the return action of $x_{\textit{ind}}$ is later than return actions of other $x_i$.
+This completes the proof of this lemma.
+\qed
+\end {proof}
+
+\begin{restatable}{lemma}{Rpr4IsAlwaysCoRegular}
+\label{lemma:Rpr4 is always co-regular}
+Given a data-differentiated history $h$, if $\textit{last}(h) = R_{\textit{pr4}}$, then $h \sqsubseteq \textit{MR}_{\textit{pr4}}$.
+\end{restatable}
+
+\begin {proof}
+
+According to Lemma \ref{lemma:Rpr4 as happen before}, if $\textit{last}(h) = R_{\textit{pr4}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr4}}$, then there exists $x$ and $y$ with maximal priority in $h$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,\_) <_{\textit{hb}} \textit{put}(y,\_)$.
+Let $h_1 = h \vert_{ \{ x,y \} }$.
+It is obvious that $\textit{transToQueue}(h_1)$ does not linearizable to queue.
+This contradicts the assumption that every single-priority history can be ``linearizable to queue'', and thus, we can safely ignore this case.
+\qed
+\end {proof}
+
+\subsection{Co-Regular of $R_{\textit{pr5}}$}
+\label{subsec:co-regular of Rpr5}
+
+$R_{\textit{pr5}} \equiv (u \cdot v \in \textit{PQueue}) \wedge (\textit{matched}(u,\textit{put},\textit{rm}) ) \Rightarrow (u \cdot \textit{rmEmpty} \cdot v \in \textit{PQueue})$
+
+Similar as the notion of gap in \cite{Bouajjani:2015}, we propose the notion of gap of $\textit{em}(\textit{empty})$ operations.
+
+\begin{definition}[gap for $\textit{em}(\textit{empty})$ operations]\label{def:gap for rmEmpty operations}
+
+Let $h$ be a data-differentiated history and $o = \textit{rm}(\textit{empty})$ of $h$.
+We say that $h$ has a gap on operation $o$, if there is a partition of the operations of $h$ into $L$ and $R$ satisfying:
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $L$ has matched $\textit{put}$ and $\textit{rm}$ operations.
+
+\item[-] no operation of $R$ happens before an operation of $L$ or $o$.
+
+\item[-] $o$ does not happen before operations of $L$
+\end{itemize}
+\end{definition}
+
+Compared to the gap on $\textit{put}$ and $\textit{rm}$ operations, here we do not consider the priority of operations in $L$ or $R$.
+
+\begin{restatable}{lemma}{GapEqualsLinforRpr5}
+\label{lemma:Gap Equals Lin for Rpr5}
+A data-differentiated history $h$ has a projection $h'$, such that $\textit{last}(h') = R_{\textit{pr5}}$ and $h'$ does not linearizable to $\textit{MR}_{\textit{pr5}}$, if and only if there exists a projection $h''$ of $h$, such that $\textit{last}(h'') = R_{\textit{pr5}}$, there exists $o = \textit{rm}(\textit{empty})$ in $h''$, and there is no gap of $o$ in $h''$.
+\end{restatable}
+
+\begin {proof}
+
+Let us prove the $\textit{only if}$ direction by contradiction.
+Assume that there exists such $h'$, and for each $\textit{rm}(\textit{empty})$ in $h'$, there is a gap of $o$ in $h''$.
+Since $\textit{last}(h') = R_{\textit{pr5}}$, there is exists at least $\textit{rm}(\textit{empty})$ operation in $h$, and let $o$ be arbitrary one of them.
+Since there is a gap of $o$ in $h$, let $L$ and $R$ be the sets of the gap.
+Let $u$ be a sequence of operations of $L$, and $u$ is generated from $h$ by locating linearization points of each operation in $L$ just after its call action.
+Let $v$ be a sequence of operations of $R$, and $v$ is generated from $h$ by locating linearization points of each operation in $R$ just before its return action.
+It is easy to see that $h \sqsubseteq l = u \cdot o \cdot v$, and $l \in \textit{MR}_{\textit{pr5}}$.
+This contradicts that $h'$ does not linearizable to $\textit{MR}_{\textit{pr5}}$.
+
+To prove the $\textit{if}$ direction, let $h_1$ be a history generate from $h''$ by discarding all other $\textit{rm}(\textit{Empty})$ operations.
+By assumption, it is obvious that there is no gap of $o$ in $h_1$.
+Assume by contradiction that there exists $l'$, such that $h_1 \sqsubseteq l' \in \textit{MR}_{\textit{pr5}}$.
+It is easy to see that $l' = u' \cdot o \cdot v'$, such that $u'$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
+Let $L$ be the set of operations of $u'$, and $R$ be the set of operations in $v'$.
+Since $h_1 \sqsubseteq l'$, it is easy to see that $L$ and $R$ is a gap of $o$ in $h_1$, contradicts that there is no gap of $o$ in $h_1$.
+\qed
+\end {proof}
+
+Similar as the notion of gap in \cite{Bouajjani:2015}, we propose the notion left-right constraint of $\textit{rm}(\textit{empty})$ operation.
+
+\begin{definition}[left-right constraint for $\textit{rm}(\textit{empty})$ operation]\label{def:left-right constraint for rmEmpty operation}
+Given a data-differentiated history $h$, and $o = \textit{rm}(\textit{empty})$ of $h$. The left-right constraint of $o$ is the graph $G$ where:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] the nodes are the items of $h$ or $o$, to which we add a node,
+
+\item[-] there is an edge from item $d_1$ to $o$, if $\textit{put}(d_1,\_)$ happens before $o$,
+
+\item[-] there is an edge from $o$ to item $d_1$, if $o$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$,
+
+\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$.
+\end{itemize}
+\end{definition}
+
+Compared to the left-right constraint of $\textit{put}$ and $\textit{rm}$ operations, here we do not consider the priority.
+We also do not consider other $\textit{rm}(\textit{empty})$ operations in $h$.
+
+Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$.
+Let $\textit{LMSet}_1(h,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some item, either $\textit{op}$ happens before $o$ in $h$, or there is an operation $\textit{op}'$ with the same item of $\textit{op}$, such that $\textit{op}'$ happens before $o$ in $h$ $\}$.
+For each $i \geq 1$, let $\textit{LMSet}_{\textit{i+1}}(h,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some item, $\textit{op}$ is not an operation of items in $\textit{LMSet}_k(h,o)$ for each $k \leq i$, either $\textit{op}$ happens before some operation $o' \in \textit{LMSet}_i(h,o)$ in $h$, or there is an operation $\textit{op}''$ with the same item of $o$ and $\textit{op}''$ happens before some operation $o' \in \textit{LMSet}_i(h,o)$ in $h$ $\}$.
+We can see that $\textit{LMSet}_i(h,o) \cap \textit{LMSet}_j(h,o) = \emptyset$ for any $i \neq j$.
+Let $\textit{LMSet}(h,o) = \textit{LMSet}_1(h,o) \cup \textit{LMSet}_2(h,o) \cup \ldots$.
+
+\begin{restatable}{lemma}{RmEmptyDoesNotHappenBeforeLMSetForRpr5}
+\label{lemma:RmEmpty does not happen before LMSet for Rpr5}
+Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$.
+Let $G$ be the graph representing the left-right constraint of $o$.
+Assume that $G$ has no cycle going through $o$.
+Then, $o$ does not happen before any operation in $\textit{LMSet}(h,o)$.
+\end{restatable}
+
+This Lemma can be similarly proved as Lemma \ref{lemma:Rmx does not happen before LMSet for Rpr1}.
+
+\begin{restatable}{lemma}{LMSetHasMatchedPutandRmOperationsForRpr5}
+\label{lemma:LMSet has matched put and rm operations for Rpr5}
+Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$.
+Let $G$ be the graph representing the left-right constraint of $o$.
+Assume that $G$ has no cycle going through $o$.
+Then, $\textit{LMSet}(h,o)$ contains only matched $\textit{put}$ and $\textit{rm}$ operations.
+\end{restatable}
+
+This Lemma can be similarly proved as Lemma \ref{lemma:LMSet has matched put and rm operations}.
+
+\begin{restatable}{lemma}{GapEqualsConstraintforRpr5}
+\label{lemma:Gap Equals Constraint for Rpr5}
+Given a data-differentiated history $h$ and $o = \textit{rm}(\textit{empty})$ of $h$.
+Let $G$ be the graph representing the left-right constraint of $o$.
+There is a gap on $o$, if and only if $G$ has no cycle going through $o$.
+\end{restatable}
+
+\begin {proof}
+
+To prove the $\textit{only if}$ direction, assume that there is a gap on $o$, and let it be $L$ and $R$.
+Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow o \rightarrow d_1$ in $G$.
+Since $o \rightarrow d_1$, either $o$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$.
+Since $L$ contains matched $\textit{put}$ and $\textit{rm}$ operations, it is easy to see that operations of $d_1$ is in $R$.
+Since $d_1 \rightarrow d_2$, we can see that $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$.
+Since operations of $d_1$ is in $R$ and $L$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we can see that operations of $d_2$ is in $R$.
+Similarly, we can prove that operations of $d_3,\ldots,d_m$ is in $R$.
+Since $d_m \rightarrow o$, $\textit{put}(d_m,\_)$ happens before $o$, and $\textit{put}(d_m,\_) \in L$.
+This contradicts that operations of $d_m$ is in $R$.
+This completes the proof of the $\textit{only if}$ direction.
+
+To prove the $\textit{if}$ direction, assume that $G$ has no cycle going through $o$. Let $O_h$ be the set of operations of $h$. Let $O_L = \textit{LMSet}(h,o)$, $O_R = O_h \setminus O_L$.
+
+By Lemma \ref{lemma:LMSet has matched put and rm operations for Rpr5}, we can see that $O_L = \textit{LMSet}(h,o)$ contains only matched $\textit{put}$ and $\textit{rm}$ operations.
+It remains to prove that $L$ and $R$ is a gap of $o$.
+We prove this by showing that all the following cases are impossible:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Case $1$: Some operation $o_r \in R$ happens before $o$. Then we know that $o_r \in \textit{LMSet}(h,o)$, which contradicts that $o_r \in R$.
+
+\item[-] Case $2$: Some operation $o_r \in R$ happens before some operation $o_l$. Then we know that $o_r \in \textit{LMSet}(h,o)$, which contradicts that $o_r \in R$.
+
+\item[-] Case $4$: $o$ happens before some $o_l \in \textit{LMSet}(h,o) = L$. By Lemma \ref{lemma:RmEmpty does not happen before LMSet for Rpr5}, we know that this is impossible.
+\end{itemize}
+
+This completes the proof of the $\textit{if}$ direction.
+
+\qed
+\end {proof}
+
+Let us begin to represent an automaton that is used for capture the case that, in a sub-history $h'$ of a history $h$, $\textit{last}(h')=R_{\textit{pr5}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr5}}$, and the reason is that there is no gap on a $\textit{rm}(\textit{empty})$ operation in $h'$.
+The automaton is $\mathcal{A}_{\textit{Rpr5}}$, which is given in \figurename~\ref{fig:automata for Rpr5}.
+In \figurename~\ref{fig:automata for Rpr5}, let $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d), \textit{ret}(\textit{rm},d)$, $c_1 = c + \textit{cal}(\textit{put},b,\textit{anPri})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c + \textit{ret}(\textit{rm},b)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr5.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{Rpr5}}$}
+  \label{fig:automata for Rpr5}
+\end{figure}
+
+Given a data-differentiated history $h$, we say that $o = \textit{rm}(\textit{empty})$ in $h$ is covered by items $d_1,\ldots,d_m$ in $h$, if
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $\textit{put}(d_m,\_)$ happens before $o$,
+
+\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,
+
+\item[-] $o$ happens before $\textit{rm}(d_1)$, or $\textit{rm}(d_1)$ does not exists in $h$
+\end{itemize}
+
+According to the definition of left-right constraint for $o$, in a data-differentiated history $h$, there is a cycle going through $o$, if and only if there exists items $d_1,\ldots,d_m$, such that $o$ is covered by $d_1,\ldots,d_m$.
+
+\begin{restatable}{lemma}{Rpr5IsCoRegular}
+\label{lemma:Rpr1 5s co-regular}
+$R_{\textit{pr5}}$ is co-regular.
+\end{restatable}
+
+\begin {proof}
+
+We need to prove that:
+
+\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\mathcal{A}_{\textit{Rpr5}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr5}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr5}}$.
+
+By Lemma \ref{lemma:Gap Equals Lin for Rpr5} and Lemma \ref{lemma:Gap Equals Constraint for Rpr5}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:
+
+\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\mathcal{A}_{\textit{Rpr5}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h''$ is a projection of $h$, such that $\textit{last}(h'') = R_{\textit{pr5}}$, $o = \textit{rm}(\textit{empty})$ is in $h''$, and $o$ is covered by some items $d_1,\ldots,d_m$ in $h''$.
+
+\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\mathcal{A}_{\textit{Rpr5}}$.
+By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$.
+Let $d_1,\ldots,d_m$ be the items in $h_2$ such that $r(d_i)=d$ for each $1 \leq i \leq m$.
+Let $h'' = h_2 \vert_{ \{ o, d_0, d_1, \ldots, d_m \} }$.
+It is obvious that $\textit{last}(h'') = R_{\textit{pr5}}$.
+It is easy to see that $o$ is covered by $d_1,\ldots,d_m$.
+
+\noindent The $\textit{if}$ direction: Assume that there exists such $h$, $h''$, $o$ and $d_1,\ldots,d_m$.
+Then, let $h_1$ be obtained from $h$ by renaming $d_1,\ldots,d_m$ into $b$ and renaming other items into $d$.
+We can prove that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr5}}$.
+
+This completes the proof of this lemma. \qed
+\end {proof}
+
+\section{Relate Queue and Stack with Priority Queue}
+\label{sec:relate queue and stack with priority queue}
+
+\subsection{Relate Queue with Priority Queue}
+\label{subsec:relate queue with priority queue}
+
+Given a history $h_q$ of queue operations, let $\textit{UniPriTrans}(h_q)$ be a history of priority queue, which is generated as follows: transform each $\textit{call}(\textit{enq},a)$, $\textit{ret}(\textit{enq},a)$, $\textit{call}(\textit{deq},a)$ and $\textit{ret}(\textit{deq},a)$ of $h_q$ into $\textit{call}(\textit{put},a,1)$, $\textit{ret}(\textit{put},a)$, $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$.
+
+\begin{restatable}{lemma}{RelateQueuewithPQ}
+\label{lemma:relate queue with priority queue}
+Given a history $h_q$ of queue operations, $h_q$ is linearizable to queue, if and only if $\textit{UniPriTrans}(h_q) \sqsubseteq \textit{PQueue}$.
+\end{restatable}
+
+\begin {proof}
+
+This is obvious, since we already ensure that for each single-priority history $h$, $\textit{transToQueue}(h)$ is linearizable to queue. \qed
+\end {proof}

### 2017-02-20 data-independence-PQ.tex
move other part into appendix from Appendix B

@@
 
-A priority queue contains two method: $\textit{put}$ and $\textit{rm}$.
-A $\textit{put}$ method has two arguments, while the first argument is an item and the second argument is priority.
-A $\textit{put}$ method is used to put an item into the priority queue with certain priority.
-Here we assume that the item is chosen from a specific (possibly infinite) data domain $\mathbb{D}$ and priority is chosen from the set $\mathbb{N}$ of natural number.
-A $\textit{rm}$ method intends to remove the item with maximal priority in priority queue and then returns it.
-If the priority queue is empty, then $\textit{rm}$ returns $\textit{empty}$.
-If there are multiple items in priority queue with a same maximal priority, then $\textit{rm}$ will return the item that is putted earliest among such items.
-We say that $\textit{put}(a,p)$ matches $\textit{rm}(b)$, if $a = b$.
-
-Let us slightly modify the notion of differentiated and data-independence in \cite{Wolper:1986} and propose data-differentiated sequence and data-independence for priority queues.
+Data-independence \cite{Wolper:1986} can be used to effectively handle unbounded data domain.
+In this section, we slightly modify the notion of data-independence in \cite{Wolper:1986} and propose data-differentiated sequences and data-independence for priority queues.
 

### 2017-02-20 rules-PQ.tex
move other part into appendix from Appendix B

@@
 \label{sec:inductive rules of priority queue}
+
+A priority queue contains two method: $\textit{put}$ and $\textit{rm}$.
+A $\textit{put}$ method has two arguments, while the first argument is an item and the second argument is priority.
+A $\textit{put}$ method is used to put an item into the priority queue with certain priority.
+Here we assume that the item is chosen from a specific (possibly infinite) data domain $\mathbb{D}$ and priority is chosen from the set $\mathbb{N}$ of natural number.
+A $\textit{rm}$ method intends to remove the item with maximal priority in priority queue and then returns it.
+If the priority queue is empty, then $\textit{rm}$ returns $\textit{empty}$.
+If there are multiple items in priority queue with a same maximal priority, then $\textit{rm}$ will return the item that is putted earliest among such items.
+We say that $\textit{put}(a,p)$ matches $\textit{rm}(b)$, if $a = b$.
 

### 2017-02-20 appendix.tex
add step-b-step linearizability, not complete this section

@@
 
-\DataDifferentiatedisEnoughforPQ
+\DataDifferentiatedisEnoughforPQ*
 
@@
 \end {proof}
+
+\section{Proofs in Section \ref{sec:step-by-step linearizability of priority queues}}
+\label{sec:appendix in section step-by-step linearizability of priority queues}
+
+\PQisClosedUnderProjection*
+
+\begin {proof}
+
+This is obvious, since for each conditions in the $\textit{Guard}$ part of the rules of priority queue, if a sequence of sequential executions satisfy it, then its sub-sequence also satisfy it.
+For example, if $\textit{rm}(\textit{empty}) \notin l = u \cdot v \cdot w$ and let $D_l$ be the set of items of $l$, then for each subset $D' \subseteq D$, it is obvious that $\textit{rm}(\textit{empty}) \notin l \vert_{ D' }$.
+Similar cases hold for $\textit{pri} \geq \textit{maxPri}(l)$, $\textit{pri} > \textit{maxUPri}(l)$, $\textit{matched}(l)$, $\textit{putInSeq}(l,l_1,\textit{pri})$ (let $l_1$ be sub-sequence of $l$).
+This completes the proof of this lemma.
+\qed
+\end {proof}
+
+\PQasMultiInMRpriforSequence*
+
+\begin {proof}
+
+The \textit{only if} direction is a direct consequence of Lemma \ref{lemma:PQ is closed under projection}, since it is easy to see that each sequence in $\textit{PQueue}$ also belongs to $\textit{MS}(R)$.
+
+To prove the $\textit{if}$ direction. The $\textit{if}$ direction holds trivially if $e = \epsilon$. When $e \neq \epsilon$, we generate sequence $e_1$ from $e$ as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If $\textit{last}(e) = \textit{PQ}_3$: $e_1$ is generated from $e$ by erasing one $\textit{rm}(\textit{empty})$.
+
+\item[-] Else, if $\textit{last}(e) = \textit{PQ}_2^{=}$, and the maximal priority of $e$ is unmatched $\textit{put}$ with items in $S$ and matched $\textit{put}$: $e_1$ is generated from $e$ by erasing one unmatched $\textit{put}$ which use the item last putted in $S$.
+
+\item[-] Else, if $\textit{last}(e) = \textit{PQ}_2^{>}$, and the maximal priority in $s$ is unmatched $\textit{put}$ with items in $S$: $e_1$ is generated from $e$ by erasing one unmatched $\textit{put}$ which use the item last putted in $S$.
+
+\item[-] Else, if $\textit{last}(e) = \textit{PQ}_1^{=}$, and the maximal priority in $e$ is of more than one pair of matched $\textit{put}$ with items in $S$: $e_1$ is generated from $e$ by erasing matched $\textit{put}$ and $\textit{rm}$ of the item which is last putted in $S$.
+
+\item[-] Else, if $\textit{last}(e) = \textit{PQ}_1^{>}$: $e_1$ is generated from $e$ by erasing a pair of matched $\textit{put}$ and $\textit{rm}$ with maximal priority in $e$.
+\end{itemize}
+
+Similarly, we obtain $e_2$ from $e_1$, obtain $e_3$ from $e_2$, $\ldots$, until we obtain some $e_m = \epsilon$.
+Let $e_0=e$.
+It is obvious that $e_m \in \textit{PQueue}$.
+For $e_{\textit{m-1}}$, since
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $e_m \in \textit{PQueue}$,
+
+\item[-] By assumption, we know that $e_{\textit{m-1}} \in \textit{MS}(R_{\textit{m-1}})$, where $R_{\textit{m-1}} = \textit{last}(R_{\textit{m-1}})$.
+This implies that the guard of $R_{\textit{m-1}}$ is satisfied.
+\end{itemize}
+
+Therefore, we know that $e_{\textit{m-1}} \in \textit{PQueue}$. Similarly, we can prove that $e_{\textit{m-2}},\ldots,e_0 = e \in \textit{PQueue}$. \qed
+\end {proof}
+
+The prove that $\textit{PQueue}$ is step-by-step linearizability, we investigate each rules individually.
+
+Given a data-differentiated execution and its history, we can abuse notation and mix labels and method events with operations themselves, since items are unique in a data-differentiated execution.
+For instance, we will reference an operation labeled by $\textit{put}(p,a)$ as $\textit{put}(p,a)$.
+
+Given an operation $o$ with call action $\textit{cal}_o (\textit{put},a,p)$ and return actions $\textit{ret}_o (\textit{put})$, its method event is $\textit{put}(a,p)$.
+Given an operation $o$ with call action $\textit{cal}_o (\textit{rm})$ and return actions $\textit{ret}_o (\textit{rm},a)$, its method event is $\textit{rm}(a)$.
+
+\begin{restatable}{lemma}{PQ1LarisStepByStepLinearizability}
+\label{lemma:PQ1Lar is step-by-step linearizability}
+If concurrent execution $e$ is linearizable w.r.t. $\textit{MS}(\textit{PQ}_1^{>})$ with witness $x$, then $e \setminus x \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
+\end{restatable}
+
+\begin {proof}
+
+Let $h$ be the data-differentiated history of $e$, and $l$ be an sequential execution such that $h \sqsubseteq l$ and $l$ matches $\textit{PQ}_1^{>}$ with witness $x$.
+Let $h'=h \setminus x$ and assume that $h' \sqsubseteq l' \in \textit{PQueue}$.
+
+According to $\textit{MS}(\textit{PQ}_1^{>})$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where the $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ are matched, and $\textit{pri}_x$ is larger than priority of any method event in $u \cdot v \cdot w$.
+Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$.
+Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$, we can see that $D_L \cap D_R = \emptyset$.
+Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
+Let $S_1$ be the set of method events of operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_{L1}$ be the shortest prefix of $l'_L$ that contains all method events in $S_1$, and let $l'_{L2}$ be the sequence such that $l'_L = l'_{L1} \cdot l'_{L2}$.
+It is obvious that if $\textit{put}(x,\textit{pri}_x)$ happens before an operation $o$ in $h$, then $o$ is in $l'_{L2}$.
+Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.
+
+Since priority queue is closed under projection (Lemma \ref{lemma:PQ is closed under projection}) and $u \cdot v$ contain matched $\textit{put}$ and $\textit{rm}$, we know that $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$.
+Since $\textit{pri}_x$ is larger than any priority in $h$, it is easy to see that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$.
+Since the priority queue is empty after executing $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$, we know that $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R \in \textit{PQueue}$.
+
+It remains to prove that $h \sqsubseteq l''$.
+To prove this, we define a graph $G$ whose nodes are the operations of $h$ and there is an edge from operation $o_1$ to $o_2$, if one of the following case holds
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $o_1$ happens-before $o_2$ in h,
+
+\item[-] the method event corresponding to $o_1$ in $l''$ is before the one corresponding to $o_2$.
+\end{itemize}
+
+Assume there is a cycle in $G$.
+According the the property of interval order and the fact that the order of $l''$ is total, we know that there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
+Then, we can see
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If $o_1,o_2 \in l'_{L1}$, or $o_2 \in l'_{L1} \wedge o_1 \in l'_{L2}$, or $o_1,o_2 \in l'_{L2}$, or $o_1,o_2 \in L'_R$, then $l'$ contradicts with happen before relation of $h$.
+
+\item[-] If $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_{L1}$, then since $\textit{put}(x,\textit{pri}_x)$ happens before $o_2$, we know that $o_2 \in l'_{L2}$, which contradicts that $o_2 \in l'_{L1}$.
+
+\item[-] If $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L1}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L1}$, or $o_1 = \textit{rm}(x) \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 \in L'_R \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 = \textit{rm}(x)$, then $l$ contradicts with happen before relation of $h$.
+
+\item[-] If $o_1 \in l'_{L2} \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, then since $o_1$ happens before $\textit{put}(x,\textit{pri}_x)$, we know that $o_1 \in S_1$ and then $o_1 \in l'_{L1}$, which contradicts that $o_1 \in l'_{L2}$.
+\end{itemize}
+
+Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
+
+\end {proof}
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] if $e$ is linearizable w.r.t.
+$\textit{MS}_{\textit{PQ}_1^{>}}$ ($R \in \{ \textit{PQ}_1^{>}, \textit{PQ}_1^{=}, \textit{PQ}_2^{>},\textit{PQ}_2^{=} \}$) with witness $x$, we have: $e \setminus x \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
+
+\item[-] if $e$ is linearizable w.r.t. $\textit{MS}(\textit{PQ}_3)$ and $o$ is a $\textit{rm}(\textit{empty})$ event, we have:
+
+$e \setminus o \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
+\end{itemize}
+
+\PQueueisStepByStepLinearizability*
 
@@
 
-The following lemma states that $\textit{PQueue}$ is closed under projection. This lemma holds since the predicates appears in inductive rules of $\textit{PQueue}$ are closed under projection.
-
-\begin{restatable}{lemma}{PQisClosedUnderProjection}
-\label{lemma:PQ is closed under projection}
-
-$\textit{PQueue}$ is closed under projection.
-
-\end{restatable}
-
-%When the context is clear, we shorten priority-value data-independence as data-independence.
-For priority queue, the notion of step-by-step linearizability is modified as follows:
-
-%\vspace{-6pt}
-\begin{definition}[step-by-step linearizability for priority queue]\label{def:step-by-step linearizability of priority queue}
-A data structure $S=R_1, \ldots, R_n$ is said be to step-by-step linearizable, if for any data-differentiated sequence $e$, if $e$ is linearizable w.r.t.
-$\textit{MR}_{\textit{pri}}$ ($1 \leq i \leq 4$) with witness $x$, we have:
-
-$e \setminus x \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket \Rightarrow e \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$
-
-\noindent if $e$ is linearizable w.r.t. $\textit{MR}_{\textit{pr5}}$ with witness $x$, we have:
-
-$e \setminus x \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket \Rightarrow e \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$
-\end{definition}
-
-Let $\textit{proj}(s)$ denote the projection of $s$ obtained by erasing all the operations of several items from $s$.
-Here we assume that each $\textit{rm}(\textit{empty})$ is differentiated from each other, and we can consider each $\textit{rm}(\textit{empty})$ as having a unique item.
-
-\begin{restatable}{lemma}{PQasMultiInMRpriforSequence}
-\label{lemma:PQ as multi in MRpri for sequence}
-$s \in \textit{PQueue}$, if and only if, $\forall s'$, if $s' \in \textit{proj}(s)$, then $s' \in \textit{MR}_{\textit{pri}}$ for some $0 \leq i \leq 5$.
-\end{restatable}
-
-\begin {proof}
-
-The \textit{only if} direction is a direct consequence of Lemma \ref{lemma:PQ is closed under projection}.
-
-To prove the $\textit{if}$ direction.
-The $\textit{if}$ direction holds trivially when $s = \epsilon$.
-When $s \neq \epsilon$.
-since $s \in \textit{proj}(s)$, we generate sequence $s_1$ from $s$ as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $s$ has $\textit{rm}(\textit{empty})$ operations: $s_1$ is generated from $s$ by erasing some $\textit{rm}(\textit{empty})$ operation.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in $S_1$ and unmatched $\textit{put}$ operations of some items in $S_2$: $s_1$ is generated from $s$ by erasing the unmatched $\textit{put}$ operation of items last putted in $S_2$.
-
-\item[-] If the maximal priority in $s$ is unmatched $\textit{put}$ operations of some items in $S_2$ and $\vert S_2 \vert =1$: $s_1$ is generated from $s$ by erasing this unmatched $\textit{put}$ operation.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert \geq 2$: $s_1$ is generated from $s$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item which is last putted in $S$.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert = 1$: $s_1$ is generated from $s$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item in $S$.
-\end{itemize}
-
-Similarly, we obtain $s_2$ from $s_1$, obtain $s_3$ from $s_2$, $\ldots$, until we obtain some $s_m = \epsilon$.
-It is obvious that $s_m \in \textit{PQueue}$, and by the five rules of $\textit{PQueue}$, it is easy to see that $s_{\textit{m-1}},\ldots,s_1,s \in \textit{PQueue}$.
-\qed
-\end {proof}
-
 \begin{restatable}{lemma}{PQasMultiInMRpriforHistory}
@@
 To prove that $\textit{PQueue}$ is step-by-step linearizability, we use five lemmas for five rules of $\textit{PQueue}$.
-
-\begin{restatable}{lemma}{Rpr1isStepByStepLinearizability}
-\label{lemma:Rpr1 is step-by-step linearizability}
-
-$R_{\textit{pr1}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr1}}$ with witness $x$.
-Assume that  Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Then, let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-According to $R_{\textit{pr1}}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where the $\textit{put}$ operations and $\textit{rm}$ operations in $u \cdot v$ are matched, and $\textit{pri}_x$ is larger than any priority in $u \cdot v \cdot w$.
-Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$.
-Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$ operations, we can see that $D_L \cap D_R = \emptyset$.
-Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
-Let $S_1$ be the set of all operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_{L1}$ be the shortest prefix of $l'_L$ that contains all operations of $S_1$, and let $l'_{L2}$ be the sequence such that $l'_L = l'_{L1} \cdot l'_{L2}$.
-It is obvious that if $\textit{put}(x,\textit{pri}_x)$ happens before an operation $o$ in $h$, then $o$ is in $l'_{L2}$.
-Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.
-
-Since priority queue is closed under projection (Lemma \ref{lemma:PQ is closed under projection}) and $u \cdot v$ contain matched $\textit{put}$ and $\textit{rm}$ operations, we know that $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$.
-Since $\textit{pri}_x$ is larger than any priority in $h$, it is easy to see that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$.
-Since the priority queue is empty after executing $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$, we know that $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R \in \textit{PQueue}$.
-
-It remains to prove that $h \sqsubseteq l''$.
-To prove this, we define a graph $G$ whose nodes are the operations of $h$ and there is an edge from operation $o_1$ to $o_2$, if one of the following case holds
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o_1$ happens-before $o_2$ in h,
-
-\item[-] the method event corresponding to $o_1$ in $l''$ is before the one corresponding to $o_2$.
-\end{itemize}
-
-Assume there is a cycle in $G$.
-According the the property of interval order and the fact that the order of $l''$ is total, we know that there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $o_1,o_2 \in l'_{L1}$, or $o_2 \in l'_{L1} \wedge o_1 \in l'_{L2}$, or $o_1,o_2 \in l'_{L2}$, or $o_1,o_2 \in L'_R$, then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] If $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_{L1}$, then since $\textit{put}(x,\textit{pri}_x)$ happens before $o_2$, we know that $o_2 \in l'_{L2}$, which contradicts that $o_2 \in l'_{L1}$.
-
-\item[-] If $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L1}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L1}$, or $o_1 = \textit{rm}(x) \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 \in L'_R \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 = \textit{rm}(x)$, then $l$ contradicts with happen before relation of $h$.
-
-\item[-] If $o_1 \in l'_{L2} \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, then since $o_1$ happens before $\textit{put}(x,\textit{pri}_x)$, we know that $o_1 \in S_1$ and then $o_1 \in l'_{L1}$, which contradicts that $o_1 \in l'_{L2}$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
-
-\end {proof}
 
@@
 The following lemma states that $\textit{PQueue}$ is step-by-step linearizability, it is a direct consequence of Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, Lemma \ref{lemma:Rpr2 is step-by-step linearizability}, Lemma \ref{lemma:Rpr3 is step-by-step linearizability}, Lemma \ref{lemma:Rpr4 is step-by-step linearizability} and Lemma \ref{lemma:Rpr5 is step-by-step linearizability}.
-
-\begin{restatable}{lemma}{PQueueisStepByStepLinearizability}
-\label{lemma:PQueue is step-by-step linearizability}
-
-$\textit{PQueue}$ is step-by-step linearizability.
-
-\end{restatable}
 

### 2017-02-20 ss-lin-PQ.tex
add step-b-step linearizability, not complete this section

\section{Step-by-Step Linearizability of Priority Queues}
\label{sec:step-by-step linearizability of priority queues}

In this section we shows that, with the help of a property called step-by-step linearizability, we can partition the concurrent executions which are not linearizable with respect to $\textit{PQueue}$ into a finite number of classes. Intuitively, each such class represents a set of sequential execution that violate one rule of priority queue. Here step-by-step linearizability enables us to build a linearization for an execution $e$ incrementally, using linearizations of projections of $e$. The proof of lemmas in this section can be found in Appendix \ref{sec:appendix in section step-by-step linearizability of priority queues}.

The projection $e \vert{\mathcal{D}}$ of a sequential execution $e$ into a subset $\mathcal{D} \subseteq \mathbb{D}$ of items is obtained from $e$ by erasing all method events with a data value not in $\mathcal{D}$. The set of projections of $e$ is denoted $\textit{proj}(e)$. We write $e \setminus x$ for the projection $e \vert_{ \mathbb{D} \setminus \{ x \} }$. This extends naturally to histories. A set $S$ of sequential executions is closed under projection, if for all $\mathcal{D} \subseteq \mathbb{D}$ and $e \in S$, we have $e \vert_{ \mathcal{D} } \in S$. The following lemma states that $\textit{PQueue}$ is closed under projection, since the predicates used in rules of priority queue are ``closed under projection''.

\begin{restatable}{lemma}{PQisClosedUnderProjection}
\label{lemma:PQ is closed under projection}
$\textit{PQueue}$ is closed under projection.
\end{restatable}

A sequential execution $e$ matches a rule $R$ of priority queue, if $e=\textit{Expr}(l_1, \ldots, l_k,a,b)$, and $\textit{Guard}(l_1, \ldots, l_k,a,b)$ holds. Here $\textit{Guard}$ and $\textit{Expr}$ are of rule $R$, and we call $a$ (if exists) the witness of $e$. We denote by $\textit{MS}(R)$ the set of sequential executions which match $R$. Note that sequences in $\textit{MS}(R)$ only respect rule $R$ and may be not in $\textit{PQueue}$. $e$ is linearizable with respect to $\textit{MS}(R)$ with witness $x$, if $e$ is linearizable with respect to $u \in \textit{MS}(R)$ and $x$ is witness of $u$.

\begin{example}\label{example:match set of R}
$e = \textit{rm}(b) \cdot \textit{put}(b,2) \cdot \textit{put}(a,4) \cdot \textit{rm}(a)$ is in $\textit{MS}(\textit{PQ}_1^{>} )$, but it is obvious that $e \notin \textit{PQueue}$.
\end{example}

The following lemma states that checking inclusion into $\textit{PQueue}$ of a data-differentiated sequence is equivalent to checking inclusion into $\textit{MS}(R)$ for everyone of its projections.

\begin{restatable}{lemma}{PQasMultiInMRpriforSequence}
\label{lemma:PQ as multi in MRpri for sequence}
Given a data-differentiated sequential execution $e$, $e \in \textit{PQueue}$, if and only if, $\forall e' \in \textit{proj}(e)$, $e' \in \textit{MS}(R)$, where $R = \textit{last}(e)$.
\end{restatable}

Lemma \ref{lemma:PQ as multi in MRpri for sequence} simplifies checking inclusion into $\textit{PQueue}$, since checking $\textit{MS}(R)$ only concerns information of one rule, while checking $\textit{PQueue}$ need to consider every method events. We want a similar lemma for checking linearizability with respect to $\textit{PQueue}$. To enable such equivalent characterization, we introduce the notion of step-by-step linearizability for priority queue. The projection $e \vert{O}$ of a concurrent execution $e$ into a set $O$ of operations is obtained from $e$ by erasing all call and return actions of operations not in $O$.  We write $e \setminus o$ for the projection $e \vert_{ O_h \setminus \{ o \} }$, where $O_h$ is the set of operations of $e$. This extends naturally to histories.

\begin{definition}\label{def:step-by-step linearizability of priority queue}
A set $S$ of sequential executions of priority queue is step-by-step linearizable, if for any data-differentiated execution $e$,
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] if $e$ is linearizable w.r.t. $\textit{MS}(R)$ ($R \in \{ \textit{PQ}_1^{>}, \textit{PQ}_1^{=}, \textit{PQ}_2^{>},\textit{PQ}_2^{=} \}$) with witness $x$, we have: $e \setminus x \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.

\item[-] if $e$ is linearizable w.r.t. $\textit{MS}(\textit{PQ}_3)$ and $o$ is a $\textit{rm}(\textit{empty})$ event, we have:

$e \setminus o \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
\end{itemize}
\end{definition}

The following lemma states that $\textit{PQueue}$ is step-by-step linearizability.

\begin{restatable}{lemma}{PQueueisStepByStepLinearizability}
\label{lemma:PQueue is step-by-step linearizability}

$\textit{PQueue}$ is step-by-step linearizability.

\end{restatable}

### 2017-02-21 appendix.tex
I add section of step-by-step linearizability. Now Appendix B is the proof of step-by-step linearizability. In this process, proofs of two lemmas are changed.

@@
 
+\subsection{Proof of Lemma \ref{lemma:PQ is closed under projection}}
+
 \PQisClosedUnderProjection*
@@
 
+\subsection{Proof of Lemma \ref{lemma:PQ as multi in MRpri for sequence}}
+
 \PQasMultiInMRpriforSequence*
@@
 
-\item[-] Else, if $\textit{last}(e) = \textit{PQ}_2^{>}$, and the maximal priority in $s$ is unmatched $\textit{put}$ with items in $S$: $e_1$ is generated from $e$ by erasing one unmatched $\textit{put}$ which use the item last putted in $S$.
+\item[-] Else, if $\textit{last}(e) = \textit{PQ}_2^{>}$, and the maximal priority in $e$ is unmatched $\textit{put}$ with items in $S$: $e_1$ is generated from $e$ by erasing one unmatched $\textit{put}$ which use the item last putted in $S$.
 
@@
 
+\subsection{Proof of Lemma \ref{lemma:PQueue is step-by-step linearizability}}
+
 The prove that $\textit{PQueue}$ is step-by-step linearizability, we investigate each rules individually.
@@
 Given an operation $o$ with call action $\textit{cal}_o (\textit{rm})$ and return actions $\textit{ret}_o (\textit{rm},a)$, its method event is $\textit{rm}(a)$.
+
+Given a data-differentiated execution $e$ and its history $h$, we can obtain a sequence $h'$ from $h$ by adding $\textit{put}(a,p)$ (resp., $\textit{rm}(a)$) between each pair of $\textit{cal}(\textit{put},a)$ and $\textit{ret}(\textit{put},a)$ (resp., $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$).
+The projection of $h'$ into method events is called linearization of $e$ and $h$, and each method event we add in $h'$ can be considered as a linearization point of the corresponding method event.
 
@@
 Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
-Let $S_1$ be the set of method events of operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_{L1}$ be the shortest prefix of $l'_L$ that contains all method events in $S_1$, and let $l'_{L2}$ be the sequence such that $l'_L = l'_{L1} \cdot l'_{L2}$.
+Let $S_1$ be the set of operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_{L1}$ be the shortest prefix of $l'_L$ that contains all method events of operations in $S_1$, and let $l'_{L2}$ be the sequence such that $l'_L = l'_{L1} \cdot l'_{L2}$.
 It is obvious that if $\textit{put}(x,\textit{pri}_x)$ happens before an operation $o$ in $h$, then $o$ is in $l'_{L2}$.
@@
 
-\item[-] If $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L1}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L1}$, or $o_1 = \textit{rm}(x) \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 \in L'_R \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 = \textit{rm}(x)$, then $l$ contradicts with happen before relation of $h$.
+\item[-] If $o_1 = \textit{rm}(x) \wedge (o_2 \in  l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2})$, or $o_1 \in L'_R \wedge ( o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2} \vee o_2 = \textit{rm}(x) )$, then $l$ contradicts with happen before relation of $h$.
 
@@
 
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
-
-\end {proof}
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] if $e$ is linearizable w.r.t.
-$\textit{MS}_{\textit{PQ}_1^{>}}$ ($R \in \{ \textit{PQ}_1^{>}, \textit{PQ}_1^{=}, \textit{PQ}_2^{>},\textit{PQ}_2^{=} \}$) with witness $x$, we have: $e \setminus x \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
-
-\item[-] if $e$ is linearizable w.r.t. $\textit{MS}(\textit{PQ}_3)$ and $o$ is a $\textit{rm}(\textit{empty})$ event, we have:
-
-$e \setminus o \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
-\end{itemize}
+Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \textit{PQueue}$. \qed
+\end {proof}
+
+\begin{restatable}{lemma}{PQ1EqualisStepByStepLinearizability}
+\label{lemma:PQ1Equal is step-by-step linearizability}
+If concurrent execution $e$ is linearizable w.r.t. $\textit{MS}(\textit{PQ}_1^{=})$ with witness $x$, then $e \setminus x \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
+\end{restatable}
+
+\begin {proof}
+
+Let $h$ be the data-differentiated history of $e$, and $l$ be an sequential execution such that $h \sqsubseteq l$ and $l$ matches $\textit{PQ}_1^{=}$ with witness $x$.
+Let $h'=h \setminus x$ and assume that $h' \sqsubseteq l' \in \textit{PQueue}$.
+
+Since $h' \sqsubseteq l'$, there exists a sequence $e_{l'}$ generated from $e$ by adding method events of $e \setminus x$, and $e_{l'}$ satisfies that (1) the projection of $e_{l'}$ into call and return action is $e$, (2) the projection of $e_{l'}$ into method events is $l'$, and (3) for each matched pair of call action $c$ and return action $r$ of $e$ (if their item is not $x$), there exists a method event $o$, such that $c$ before $o$ in $h'_{l'}$, $o$ before $r$ in $h'_{l'}$, and $o$ is the method event of $c$ and $r$.
+Intuitively, $e_{l'}$ is obtained from $h$ by adding linearization points according to $l'$.
+
+According to $\textit{PQ}_1^{=}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where (1) the $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ are matched, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v \cdot w$, (3) there exists matched $\textit{put}$ and $\textit{rm}$ in $u \cdot v \cdot w$ that has priority $\textit{pri}_x$, (4) $\textit{pri}_x$ is larger than priorities of any unmatched put in $u \cdot v \cdot w$, and (5) all $\textit{put}$ with priority $\textit{pri}_x$ is in $u$.
+Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$.
+Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$, we can see that $D_L \cap D_R = \emptyset$.
+Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
+
+Let $l'_{\textit{L0}}$ be the shortest prefix of $l'$, such that the linearization point of every method event in  $l'_{\textit{L0}}$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $e_{l'}$.
+Let $E_{px}$ be the set of $\textit{put}$ with priority $\textit{pri}_x$ in $e \setminus x$.
+Let $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t$, where $l_t$ is the projection of $l'$ into method events of $E_{px}$ that are not in $l'_{\textit{L0}}$.
+Let $E_{l_t}$ be the set of method events in $l_t$.
+Let $l'_{\textit{L2}}$ be the projection of $l'_L$ into method events of $l'_L$ that are not in $l'_{\textit{L1}}$.
+Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.
+
+As in Lemma \ref{lemma:PQ1Lar is step-by-step linearizability}, we know that $l'_L \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$.
+It is obvious that $l'_{\textit{L0}} \in \textit{PQueue}$, and then $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t \in \textit{PQueue}$.
+Since there is no $\textit{rm}(\textit{Empty})$ in $h$ and $\textit{pri}_x$ is the maximal priority of $h$, $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$.
+Since all $\textit{put}$ operations with priority $\textit{pri}_x$ is in $l'_{L1}$, we know that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$.
+Since the priority queue is empty after executing $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$, we know that $l'' \in \textit{PQueue}$.
+
+To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:PQ1Lar is step-by-step linearizability}.
+Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
+Then, we can see
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+
+\item[-] $o_1 \in L'_R \wedge (o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2} \vee o_2 = \textit{rm}(x))$, or $o_1 = \textit{rm}(x) \wedge (o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2})$: Then $l$ contradicts with happen before relation of $h$.
+
+\item[-] $o_1, o_2 \in l'_R$ or $o_1, o_2 \in l'_{L2}$: Then $l'$ contradicts with happen before relation of $h$.
+
+\item[-] $o_1, o_2 \in l'_{L1}$: If $(o_1, o_2 \in l'_{L0}) \vee (o_1,o_2 \in l_t)$, then $l'$ contradicts with happen before relation of $h$.
+
+    Else, it is the case that $o_1 \in l_t \wedge o_2 \in l'_{L0}$.
+We can see that $o_1$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and the linearization point of $o_1$ is after the return action of $\textit{put}(x,\textit{pri}_x)$ in $e_{l'}$.
+Since $o_1 <_{\textit{hb}} o_2$, it is easy to see that $\textit{put}(x,\textit{pri}_x) <_{\textit{hb}} o_2$, then $o_2 \notin l'_{L0}$, which contradicts that $o_2 \in l'_{L0}$.
+
+\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_{L1}$: Since $o_1 <_{\textit{hb}} o_2$, it is obvious that $o_2 \notin l'_{L1}$, contradicts that $o_2 \in l'_{L1}$.
+
+\item[-] $o_1 \in l'_{L2} \wedge o_2 \in l'_{L1}$: If $o_2 \in l'_{L0}$, then $l'$ contradicts with happen before relation of $h$.
+
+    Else, it is the case that $o_2 \in l_t$.
+Since $o_2$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and $o_1 <_{\textit{hb}} o_2$, the return action of $o_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $e$.
+Then it is easy to see that $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.
+
+\item[-] $o_1 \in l'_{L2} \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: Then $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.
+\end{itemize}
+
+Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \textit{PQueue}$. \qed
+\end {proof}
+
+\begin{restatable}{lemma}{PQ2LarisStepByStepLinearizability}
+\label{lemma:PQ2Lar is step-by-step linearizability}
+If concurrent execution $e$ is linearizable w.r.t. $\textit{MS}(\textit{PQ}_2^{>})$ with witness $x$, then $e \setminus x \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
+\end{restatable}
+
+\begin {proof}
+
+Let $h$ be the data-differentiated history of $e$, and $l$ be an sequential execution such that $h \sqsubseteq l$ and $l$ matches $\textit{PQ}_1^{>}$ with witness $x$.
+Let $h'=h \setminus x$ and assume that $h' \sqsubseteq l' \in \textit{PQueue}$.
+
+According to $\textit{PQ}_2^{>}$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, where $\textit{pri}_x$ is the maximal priority in $u \cdot v$, $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ in $u \cdot v$ (may equal priority of some unmatched $\textit{put}$ operations in $u \cdot v$), and all other $\textit{put}$ with priority $\textit{pri}_x$ is in $u$.
+Let $S_1$ be the set of operations of $e$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_1$ be the shortest prefix of $l'$ that contains all method events of operations of $S_1$, and let $l'_2$ be the sequence such that $l' = l'_1 \cdot l'_2$.
+Let sequence $l'' = l'_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_2$.
+
+Since $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ operations in $u \cdot v$, it is easy to see that $l'' \in \textit{PQueue}$.
+It remains to prove that $h \sqsubseteq l''$.
+To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:PQ1Lar is step-by-step linearizability}.
+Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
+Then, we can see
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $o_1,o_2 \in l'_1$, or $o_1,o_2 \in l'_2$, or $o_1 \in l'_2 \wedge o_2 \in l'_1$: Then $l'$ contradicts with happen before relation of $h$.
+
+\item[-] $o_1 \in l'_2 \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: $o_1 \in S_1$, and then $o_1 \in l'_1$, which contradicts that $o_1 \in l'_2$.
+
+\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_1$: We can see that in $l'$, $o_2$ must after every operation in $S_1$, and then $o_2 \notin l'_1$, contradicts that $o_2 \in l'_1$.
+\end{itemize}
+
+Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \textit{PQueue}$. \qed
+\end {proof}
+
+\begin{restatable}{lemma}{PQ2EqualisStepByStepLinearizability}
+\label{lemma:PQ2Equal is step-by-step linearizability}
+If concurrent execution $e$ is linearizable w.r.t. $\textit{MS}(\textit{PQ}_2^{=})$ with witness $x$, then $e \setminus x \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
+\end{restatable}
+
+\begin {proof}
+Let $h$ be the data-differentiated history of $e$, and $l$ be an sequential execution such that $h \sqsubseteq l$ and $l$ matches $\textit{PQ}_1^{>}$ with witness $x$.
+Let $h'=h \setminus x$ and assume that $h' \sqsubseteq l' \in \textit{PQueue}$.
+
+According to $\textit{PQ}_2^{=}$, there exist sequences $u$ and $v$, such that (1) $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v$, (3) $\textit{pri}_x$ equals the maximal priority of matched $\textit{put}$ in $u \cdot v$, (4) all matched $\textit{put}$ with priority $\textit{pri}_x$ is in $u$ and (5) all the other unmatched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$.
+
+Similarly as in Lemma \ref{lemma:PQ1Equal is step-by-step linearizability}, we can obtain a sequence $e_{l'}$ from $e$ by adding linearization points according to $l'$.
+Let $l'_1$ be the shortest prefix of $l'$, such that the linearization point of every method event in $l'_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $e_{l'}$.
+Let $E_{px}$ be the set of $\textit{put}$ with priority $\textit{pri}_x$ in $e \setminus x$.
+Let $l'_2$ be the projection of $l'$ into method events of $E_{px}$ that are not in $l'_1$.
+Let $l_x' = l'_1 \cdot l'_2 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_3$, where $l'_3$ is the projection of $l'$ into method events that are not in $l'_1 \cdot l'_2$.
+
+Let $l_x$ be generated from $l'_x$ by erasing $\textit{put}(x,\textit{pri}_x)$.
+Since (1) $l_x$ can be obtained from $l'$ by making some $\textit{put}$ happen earlier, (2) for each $\textit{rm}(a)$ in $l'$, its corresponding $\textit{put}(a,\_)$ must before it in $l'$, and (3) there are no $\textit{rm}(\textit{empty})$ in $l'$, we can see that $l_x \in \textit{PQueue}$.
+Since $\textit{pri}_x$ is the maximal priority in $l$ and there are no $\textit{rm}(\textit{empty})$ in $l$, we can see that $l'_x \in \textit{PQueue}$.
+
+Let us prove that $h \sqsubseteq l'_x$ by contradiction.
+Assume that there exists operations $o_1,o_2$, such that $o_1$ happens before $o_2$ in $h$, and $o_2$ is before $o_1$ in $l'_x$.
+Then, we can see
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $o_1, o_2 \in l'_1$, or $o_1, o_2 \in l'_2$, or $o_1, o_2 \in l'_3$: Then $l'$ contradicts with happen before relation of $h$.
+
+\item[-] $o_2 \in l'_1 \wedge o_1 \in l'_2$: Then by construction of $l'_1$, it is easy to see that $o_1 \in l'_1$, which contradicts that $o_1 \in l'_2$.
+
+\item[-] $o_2 \in l'_1 \wedge o_1 = \textit{put}(x,\textit{pri}_x)$: Then by construction of $l'_1$, it is easy to see that $o_2 \notin l'_1$, which contradicts that $o_2 \in l'_2$.
+
+\item[-] $o_2 \in l'_1 \wedge o_1 \in l'_3$: Then by construction of $l'_1$, it is easy to see that $o_1 \in l'_1$, which contradicts that $o_1 \in l'_3$.
+
+\item[-] $o_2 \in l'_2 \wedge o_1 = \textit{pt}(x,\textit{pri}_x)$: It is easy to see that each method event in $l'_2$ is a $\textit{put}$ with priority $\textit{pri}$.
+By $l'$ we know that $\textit{pt}(x,\textit{pri}_x)$ does not happen before any of them.
+
+\item[-] $o_2 \in l'_2 \wedge o_1 \in l'_3$: It is easy to see that each method event in $l'_2$ is a $\textit{put}$ $o$ with priority $\textit{pri}$ and $o$ interleaves with $\textit{put}(x,\textit{pri}_x)$.
+Since $o_1$ happens before $o_2$, it is obviou that the return action of $o_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $e$, and then $o_1 \in l'_1$, which contradicts that $o_1 \in l'_3$.
+
+\item[-] $o_2 = \textit{put}(x,\textit{pri}_x) \wedge o_1 \in l'_3$: It is not hard to see that $o_1 \in l'_1$, which contradicts that $o_1 \in l'_3$.
+\end{itemize}
+
+Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq l'_x$. \qed
+\end {proof}
+
+\begin{restatable}{lemma}{PQ3isStepByStepLinearizability}
+\label{lemma:PQ3 is step-by-step linearizability}
+If concurrent execution $e$ is linearizable w.r.t.
+$\textit{MS}(\textit{PQ}_3)$ and $o$ is a $\textit{rm}(\textit{empty})$ event, then $e \setminus o \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
+\end{restatable}
+
+\begin {proof}
+Let $h$ be the data-differentiated history of $e$, $l$ be an sequential execution such that $h \sqsubseteq l$, $l$ matches $\textit{PQ}_3$ and $o$ is a $\textit{rm}$ method event in $h$.
+Let $h'=h \setminus o$ and assume that $h' \sqsubseteq l' \in \textit{PQueue}$.
+
+According to $\textit{PQ}_3$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{rm}(\textit{empty}) \cdot v$, where the $\textit{put}$ operations and $\textit{rm}$ in $u$ are matched.
+
+Let $E_L$ be the set of method events in $u$ and $E_R$ be the set of method events in $v$.
+Let $l'_L = l' \vert_{E_L}$ and $l'_R = l' \vert_{E_R}$.
+Let sequence $l'' = l'_L \cdot o \cdot L'_R$.
+Since priority queue is closed under projection (Lemma \ref{lemma:PQ is closed under projection}) and $u$ contains matched $\textit{put}$ and $\textit{rm}$, we know that $l'_L \in \textit{PQueue}$ and the the priority queue is empty after executing $l'_L$.
+Then we know that $l'_L \cdot o \in \textit{PQueue}$.
+Since $l'_R$ is obtained from $l'$ by discarding pairs of matched $\textit{put}$ and $\textit{rm}$ operations, it is easy to see that $L'_R \in \textit{PQueue}$, and then we know that $l'' = l'_L \cdot o \cdot L'_R \in \textit{PQueue}$.
+
+It remains to prove that $h \sqsubseteq l''$.
+To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:PQ1Lar is step-by-step linearizability}.
+Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
+Then, we can see
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $o_1,o_2 \in l'_L$, or $o_1,o_2 \in l'_R$: Then $l'$ contradicts with happen before relation of $h$.
+
+\item[-] If $o_1=o \wedge o_2 \in l'_L$, or $o_1 \in l'_R \wedge o_2 \in l'_L$, or $o_1 \in l'_R \wedge o_2 = 0$, then $l$ contradicts with happen before relation of $h$.
+\end{itemize}
+
+Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \textit{PQueue}$. \qed
+\end {proof}
+
+The following lemma states that $\textit{PQueue}$ is step-by-step linearizability, it is a direct consequence of Lemma \ref{lemma:PQ1Lar is step-by-step linearizability}, Lemma \ref{lemma:PQ1Equal is step-by-step linearizability}, Lemma \ref{lemma:PQ2Lar is step-by-step linearizability}, Lemma \ref{lemma:PQ2Equal is step-by-step linearizability} and Lemma \ref{lemma:PQ3 is step-by-step linearizability}.
 
 \PQueueisStepByStepLinearizability*
+
+\begin {proof}
+This is a direct consequence of Lemma \ref{lemma:PQ1Lar is step-by-step linearizability}, Lemma \ref{lemma:PQ1Equal is step-by-step linearizability}, Lemma \ref{lemma:PQ2Lar is step-by-step linearizability}, Lemma \ref{lemma:PQ2Equal is step-by-step linearizability} and Lemma \ref{lemma:PQ3 is step-by-step linearizability}.
+\qed
+\end {proof}
+
+\subsection{Proof of Lemma \ref{lemma:PQ as multi in MRpri for history}}
+
+\PQasMultiInMRpriforHistory*
+
+\begin {proof}
+
+To prove the $\textit{only if}$ direction, assume that $e \sqsubseteq l \in \textit{PQueue}$.
+Given $e' = e \vert_{D}$, by Lemma \ref{lemma:PQ is closed under projection}, it is obvious that $e' \sqsubseteq l' = l \vert_{D} \in \textit{PQueue}$.
+Then by Lemma \ref{lemma:PQ as multi in MRpri for sequence} we know that $l' \in \textit{MS}(R)$ where $R = \textit{last}(e')$.
+
+To prove the $\textit{if}$ direction. The $\textit{if}$ direction holds trivially if $e = \epsilon$. When $e \neq \epsilon$, we generate concurrent execution $e_1$ from $e$ as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If $\textit{last}(e) = \textit{PQ}_3$: $e_1$ is generated from $e$ by erasing call and return of one $\textit{rm}(\textit{empty})$.
+
+\item[-] Else, if $\textit{last}(e) = \textit{PQ}_2^{=}$, and the maximal priority of $e$ is unmatched $\textit{put}$ with items in $S$ and matched $\textit{put}$: $e_1$ is generated from $e$ by erasing call and return of one unmatched $\textit{put}$, which is maximal respect to $<_{\textit{hb}}$ among $\textit{put}$ with items in $S$.
+
+\item[-] Else, if $\textit{last}(e) = \textit{PQ}_2^{>}$, and the maximal priority in $s$ is unmatched $\textit{put}$ with items in $S$: $e_1$ is generated from $e$ by erasing call and return of one unmatched $\textit{put}$, which is maximal respect to $<_{\textit{hb}}$ among $\textit{put}$ with items in $S$.
+
+\item[-] Else, if $\textit{last}(e) = \textit{PQ}_1^{=}$, and the maximal priority in $e$ is of more than one pair of matched $\textit{put}$ and $\textit{rm}$ with items in $S$: $e_1$ is generated from $e$ by erasing call and return of one pair of matched $\textit{put}$ and $\textit{rm}$, where such $\textit{put}$ is maximal respect to $<_{\textit{hb}}$ among $\textit{put}$ with items in $S$, and such $\textit{rm}$ is maximal respect to $<_{\textit{hb}}$ among $\textit{rm}$ with items in $S$.
+
+\item[-] Else, if $\textit{last}(e) = \textit{PQ}_1^{>}$: $e_1$ is generated from $e$ by erasing call and return of one pair of matched $\textit{put}$ and $\textit{rm}$ with maximal priority in $e$.
+\end{itemize}
+
+Similarly, we obtain $e_2$ from $e_1$, obtain $e_3$ from $e_2$, $\ldots$, until we obtain some $e_m = \epsilon$.
+Let $e_0=e$.
+It is obvious that $e_m \sqsubseteq \textit{PQueue}$.
+For $e_{\textit{m-1}}$, since
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $e_m \sqsubseteq \textit{PQueue}$,
+
+\item[-] If $\textit{last}(e_{\textit{m-1}}) = R_{\textit{m-1}} \in \{ \textit{PQ}_1^{>}, \textit{PQ}_1^{=}, \textit{PQ}_2^{>}, \textit{PQ}_2^{=} \}$ and $e$ matches $R_{\textit{m-1}}$ with witness $x$: We already know that $e_{\textit{m-1}} \sqsubseteq \textit{MS}(R_{\textit{m-1}})$, $e_{\textit{m-1}} \setminus x \sqsubseteq \textit{PQueue}$, and by step-by-step linearizability of $\textit{PQueue}$, we can see that $e_{\textit{m-1}} \sqsubseteq \textit{PQueue}$.
+
+\item[-] If $\textit{last}(e_{\textit{m-1}}) = R_{\textit{m-1}} = \textit{PQ}_3$ and $o$ is a $\textit{rm}(\textit{empty})$ in $e_{\textit{m-1}}$: We already know that $e_{\textit{m-1}} \sqsubseteq \textit{MS}(R_{\textit{m-1}})$, $e_{\textit{m-1}} \setminus o \sqsubseteq \textit{PQueue}$, and by step-by-step linearizability of $\textit{PQueue}$, we can see that $e_{\textit{m-1}} \sqsubseteq \textit{PQueue}$.
+\end{itemize}
+
+Therefore, we know that $e_{\textit{m-1}} \in \textit{PQueue}$. Similarly, we can prove that $e_{\textit{m-2}},\ldots,e_0 = e \in \textit{PQueue}$. \qed
+\end {proof}
 
@@
 Therefore, from now on, we can safely assume that there are no unmatched $\textit{rm}$ operations.
-
-\section{Step-by-Step Linearizability of Priority Queues}
-\label{sec:step-by-step linearizability of priority queues}
-
-\begin{restatable}{lemma}{PQasMultiInMRpriforHistory}
-\label{lemma:PQ as multi in MRpri for history}
-Given a data-differentiated history $h$, assume that $\textit{PQueue}$ is step-by-step linearizability.
-$h \sqsubseteq \textit{PQueue}$, if and only if, $\forall h'$, if $h' \in \textit{proj}(h)$, then $h' \sqsubseteq \textit{MR}_{\textit{pri}}$, where $i = \textit{last}(h')$.
-\end{restatable}
-
-\begin {proof}
-
-The \textit{only if} direction is a direct consequence of Lemma \ref{lemma:PQ is closed under projection}.
-
-To prove the $\textit{if}$ direction.
-Since $h \sqsubseteq \textit{PQueue}$, we can see that $h \sqsubseteq l \in \textit{PQueue}$.
-Each projection $h'$ of $h$ can be linearizable to some projection $l'$ of $l$.
-By Lemma \ref{lemma:PQ as multi in MRpri for sequence}, $l' \in \textit{MR}_{\textit{prj}}$ for some $0 \leq j \leq 5$.
-Since $\textit{last}$ of a sequence is uniquely determined, we can see that $l' \in \textit{MR}_{\textit{pri}}$, where $i = \textit{last}(h')$.
-
-The $\textit{if}$ direction holds trivially when $h = \epsilon$. When $h \neq \epsilon$. since $h \in \textit{proj}(h)$, we generate sequence $h_1$ from $h$ as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $h$ has $\textit{rm}(\textit{empty})$ operations: Then $\textit{last}(h) = R_{\textit{pr5}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr5}}$.
-$h_1$ is generated from $h$ by erasing some $\textit{rm}(\textit{empty})$ operation.
-
-\item[-] If the maximal priority in $h$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in $S_1$ and unmatched $\textit{put}$ operations of some items in $S_2$: Then $\textit{last}(h) = R_{\textit{pr4}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr4}}$.
-$h_1$ is generated from $h$ by erasing a unmatched $\textit{put}(x,\_)$, such that $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$ for each $y \in S_2$.
-
-\item[-] If the maximal priority in $s$ is unmatched $\textit{put}$ operations of some items in $S_2$ and $\vert S_2 \vert =1$: Then $\textit{last}(h) = R_{\textit{pr3}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr3}}$.
-$h_1$ is generated from $h$ by erasing this unmatched $\textit{put}$ operation.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert \geq 2$: Then $\textit{last}(h) = R_{\textit{pr2}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr2}}$.
-$h_1$ is generated from $h$ by erasing $\textit{put}(x,\_)$ and $\textit{rm}(x)$, such that for each $y \in S$, $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$, and $\textit{rm}(x)$ does not happen before $\textit{rm}(y)$.
-
-\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert = 1$: Then $\textit{last}(h) = R_{\textit{pr1}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr1}}$.
-$h_1$ is generated from $h$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item in $S$.
-\end{itemize}
-
-Similarly, we obtain $h_2$ from $h_1$, obtain $h_3$ from $h_2$, $\ldots$, until we obtain some $h_m = \epsilon$.
-Assume that for each $i$, when generating $h_i$ from $h_{\textit{i-1}}$, we erase operations of item $d_i$.
-It is obvious that $h_m = h_{\textit{m-1}} \setminus d_m \sqsubseteq \textit{PQueue}$, and since $h_{\textit{m-1}} \sqsubseteq \textit{MR}_{\textit{prj}_{\textit{m-1}}}$, where $\textit{last}(h_{\textit{m-1}}) = R_{\textit{prj}_{\textit{m-1}}}$, we can see that $h_{\textit{m-1}} \sqsubseteq \textit{PQueue}$.
-Similarly, we can prove that $h_{\textit{m-1}},\ldots,h_1,h \sqsubseteq \textit{PQueue}$.
-This completes the proof of the $\textit{if}$ direction.
-\qed
-\end {proof}
-
-To prove that $\textit{PQueue}$ is step-by-step linearizability, we use five lemmas for five rules of $\textit{PQueue}$.
-
-\begin{restatable}{lemma}{Rpr2isStepByStepLinearizability}
-\label{lemma:Rpr2 is step-by-step linearizability}
-
-$R_{\textit{pr2}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr2}}$ with witness $x$.
-Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-Since $h' \sqsubseteq l'$, there exists a sequence $h_{l'}$ of call, return and operations, such that (1) the projection of $h_{l'}$ into call and return action is $h$, (2) the projection of $h_{l'}$ into operations is $l'$, and (3) for each matched pair of call action $c$ and return action $r$ of $h$, there exists an operation $o$, such that $a$ before $o$, $o$ before $r$, and $o$ is the operation of $c$ and $r$.
-Intuitively, $h_{l'}$ is obtained from $h$ by adding linearization points.
-
-According to $R_{\textit{pr2}}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where (1) the $\textit{put}$ operations and $\textit{rm}$ operations in $u \cdot v$ are matched, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v \cdot w$, (3) there exists matched $\textit{put}$ and $\textit{rm}$ operations in $u \cdot v \cdot w$ that has priority $\textit{pri}_x$, (4) $\textit{pri}_x$ is larger than priorities of any unmatched put in $u \cdot v \cdot w$, and (5) all $\textit{put}$ operation with priority $\textit{pri}_x$ is in $u$.
-Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$.
-Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$ operations, we can see that $D_L \cap D_R = \emptyset$.
-Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
-
-Let $l'_{\textit{L0}}$ be the shortest prefix of $l'_L$, such that every operation in $l'_{\textit{L0}}$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $h_{l'}$.
-Let $\textit{OP}_x = \{ \textit{put}(d,\textit{pri}_x) \vert \textit{put}(d,\textit{pri}_x)$ is an operation of $h\}$.
-Let $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t$, where $l_t$ is the projection of $l'$ into operations of $\textit{OP}_x$ that are not in $l'_{\textit{L0}}$.
-Let $\textit{OP}'_x$ be the set of operations in $l_t$.
-Let $l'_{\textit{L2}}$ be the projection of $l'_L$ into operations of $l'_L$ that are not in $l'_{\textit{L1}}$.
-Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.
-
-As in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, we know that $l'_L \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$.
-It is obvious that $l'_{\textit{L0}} \in \textit{PQueue}$, and then $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t \in \textit{PQueue}$.
-Since there is no $\textit{rm}(\textit{Empty})$ in $h$ and $\textit{pri}_x$ is the maximal priority of $h$, $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$.
-Since all $\textit{put}$ operations with priority $\textit{pri}$ is in $l'_{L1}$, we know that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$.
-As in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, we know that $l'' \in \textit{PQueue}$.
-
-To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}.
-Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-
-\item[-] $o_1 \in L'_R \wedge (o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2} \vee o_2 = \textit{rm}(x))$, or $o_1 = \textit{rm}(x) \wedge (o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2})$: Then $l$ contradicts with happen before relation of $h$.
-
-\item[-] $o_1, o_2 \in l'_R$ or $o_1, o_2 \in l'_{L2}$: Then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] $o_1, o_2 \in l'_{L1}$: If $(o_1, o_2 \in l'_{L0}) \vee (o_1,o_2 \in \textit{OP}'_x)$, then $l'$ contradicts with happen before relation of $h$.
-
-    Else, it is the case that $o_1 \in \textit{OP}'_x \wedge o_2 \in l'_{L0}$.
-We can see that $o_1$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and the linearization point of $o_1$ is after the return action of $\textit{put}(x,\textit{pri}_x)$.
-Since $o_1 <_{\textit{hb}} o_2$, it is easy to see that $\textit{put}(x,\textit{pri}_x) <_{\textit{hb}} o_2$, and then $o_2 \notin l'_{L0}$, which contradicts that $o_2 \in l'_{L0}$.
-
-\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_{L1}$: Since $o_1 <_{\textit{hb}} o_2$, it is obvious that $o_2 \notin l'_{L1}$, contradicts that $o_2 \in l'_{L1}$.
-
-\item[-] $o_1 \in l'_{L2} \wedge o_2 \in l'_{L1}$: If $o_2 \in l'_{L0}$, then $l'$ contradicts with happen before relation of $h$.
-
-    Else, it is the case that $o_2 \in \textit{OP}'_x$.
-$o_2$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and the linearization point of $o_2$ is after the return action of $\textit{put}(x,\textit{pri}_x)$.
-Since $o_1 <_{\textit{hb}} o_2$, the return action of $o_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$.
-Then it is easy to see that $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.
-
-\item[-] $o_1 \in l'_{L2} \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: Then $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr3isStepByStepLinearizability}
-\label{lemma:Rpr3 is step-by-step linearizability}
-
-$R_{\textit{pr3}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr3}}$ with witness $x$.
-Assume that  Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Then, let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-According to $R_{\textit{pr3}}$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, where $\textit{pri}_x$ is the maximal priority in $u \cdot v$, and $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ operations in $u \cdot v$ (may equal priority of some unmatched $\textit{put}$ operations in $u \cdot v$).
-Let $S_1$ be the set of all operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_1$ be the shortest prefix of $l'$ that contains all operations of $S_1$, and let $l'_2$ be the sequence such that $l' = l'_1 \cdot l'_2$.
-Let sequence $l'' = l'_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_2$.
-
-Since $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ operations in $u \cdot v$, it is easy to see that $l'' \in \textit{PQueue}$.
-It remains to prove that $h \sqsubseteq l''$.
-To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}.
-Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o_1,o_2 \in l'_1$, or $o_1,o_2 \in l'_2$, or $o_1 \in l'_2 \wedge o_2 \in l'_1$: Then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] $o_1 \in l'_2 \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: $o_1 \in S_1$, and then $o_1 \in l'_1$, which contradicts that $o_1 \in l'_2$.
-
-\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_1$: We can see that in $l'$, $o_2$ must after every operation in $S_1$, and then $o_2 \notin l'_1$, contradicts that $o_2 \in l'_1$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
-\end {proof}
-
-The following lemma states that, from linearization of sub-histories, we can merge them and obtain a linearization (regardless of whether it belongs to sequential specification) of the whole history.
-
-\begin{restatable}{lemma}{MergeTwoLinearization}
-\label{lemma:merge two linearization}
-
-Given a history $h$ and operation set $S_1$, $S_2$, let $h_1 = h \vert{S_1}$ and $h_2 = h \vert{S_2}$.
-Assume that $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, and $S_1 \cup S_2$ contains all operations of $h$.
-Then, there exists a sequence $l$, such that $h \sqsubseteq l$, $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$.
-\end{restatable}
-
-\begin {proof}
-
-Given a history $h$ and a operation $o \in S_2$, let $\textit{MB}(o) = \{ o' \vert o' \in S_1$ and $o'$ happens before $o$ in $h \}$, let $\textit{SBI}(o) = \textit{min}\{ i \vert l_1[0,i]$ contains all elements of $\textit{MB}(o) \}$.
-
-Let $l = s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[n] \cdot s_{\textit{n+1}}$ be generated as follows, where $n = \vert l_2 \vert$:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $s_1 = l_1[0, \textit{SBI}(l_2[1])]$,
-
-\item[-] If $s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[i]$ already contains $l_1(\textit{SBI}(l_2[\textit{i+1}]))$, then $s_{\textit{i+1}} = \epsilon$.
-Otherwise, $s_{\textit{i+1}}$ is a subsequence of $l_1$, which starts from the next of last elements of $s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[i]$ in $l_1$ and ends in $l_1(\textit{SBI}(l_2[\textit{i+1}]))$.
-\end{itemize}
-
-It is obvious that $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$, and it remains to prove that $h \sqsubseteq l$.
-We prove this by contradiction.
-Assume that $h$ is not linearizable with respect to $l$.
-Then there must be two operations of $h$, such that $o_1 <_{hb} o_2$ in $h$ but $o_2$ before $0_1$ in $l$.
-Since $l \vert{S_1} = l_1$, $l \vert{S_2} = l_2$, and $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, it is easy to see that it is impossible that $o_1,o_2 \in S_1$ or $o_1,o_2 \in S_2$.
-There are only two possibilities:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o_1 \in S_1 \wedge o_2 \in S_2$.
-Then we can see that $o_2=l_2[i]$ and $o_1 \in s_j$ for some $i < j$.
-Since $o_1 <_{hb} o_2$, we know that $o_1 \in \textit{SBI}(o_2)$.
-By the construction of $l$, we know that $o_1$ must be in $s_k$ for some $k \leq i$, contradicts that $o_1 \in s_j$ with $i < j$.
-
-\item[-] $o_1 \in S_2 \wedge o_2 \in S_1$.
-Then we can see that $o_2 \in s_i$ and $o_1 = l_2[j]$ for some $i \leq j$.
-It is easy to see that this leads to contradiction when $i = j$.
-For the case of $i \neq j$, we need to satisfy the following requirements: (1) $o_1$ ($l_2[j]$) does not happen before $l_2[i]$, (2) $l_2[i]$ does not happen before $o_2$, (3) $o_2$ is either overlap or happens before $o' \in \textit{MB}(l_2[i])$, and (4) $o' <_{hb} l_2[i]$.
-By enumeration we can see that it is impossible that above four conditions be satisfied while $o_1 <_{hb} o_2$.
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr4isStepByStepLinearizability}
-\label{lemma:Rpr4 is step-by-step linearizability}
-
-$R_{\textit{pr4}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr4}}$ with witness $x$.
-Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-According to $R_{\textit{pr4}}$, there exist sequences $u$ and $v$, such that (1) $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v$, (3) $\textit{pri}_x$ equals the maximal priority of matched $\textit{put}$ operations in $u \cdot v$, (4) all matched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$ and (5) all the other unmatched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$.
-
-Since $h' \sqsubseteq l'$, there exists a sequence $h_{l'}$ of call, return and operations, such that (1) the projection of $h$ into call and return action is $h$, (2) the projection of $h_{l'}$ into operations is $l'$, and (3) for each matched pair of call action $c$ and return action $r$ of $h$, there exists an operation $o$, such that $a$ before $o$, $o$ before $r$, and $o$ is the operation of $c$ and $r$.
-Intuitively, $h_{l'}$ is obtained from $h$ by adding linearization points.
-
-According to $R_{\textit{pr4}}$, there exist sequences $u$ and $v$, such that (1) $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v$, (3) $\textit{pri}_x$ equals the maximal priority of matched $\textit{put}$ operations in $u \cdot v$, (4) all matched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$ and (5) all the other unmatched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$.
-
-Let $l'_1$ be the shortest prefix of $l'$, such that every operation in $l'_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $h_{l'}$.
-Let $\textit{OP}_x$ be the set of $\textit{put}$ operations with priority $\textit{pri}_x$ and their matched $\textit{rm}$ operations in $h$, let $\textit{OP}'_x$ be the set of $\textit{put}$ operations in $\textit{OP}_x$ that do not appear in $l'_1$, and let $\textit{OP}''_x$ be the set of $\textit{rm}$ operations in $\textit{OP}_x$ that do not appear in $l'_1$.
-Let $l_x''=l'_1 \cdot l'_2 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_3$, where $l'_2$ is the projection of $l'$ into operations of $\textit{OP}'_x$, and $l'_3$ is the projection of $l'$ into operations of $\textit{OP}''_x$.
-
-Let $h_x$ be the projection of $h$ into $\textit{OP}_x$.
-Let us prove that $h_x \sqsubseteq l''_x$ by contradiction.
-Assume that there exists operations $o_1,o_2$, such that $o_1$ happens before $o_2$ in $h_x$, and $o_2$ is before $o_1$ in $l''_x$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $(o_1 \in l'_3) \wedge (o_2 \in l'_1 \vee o_2 \in l'_2 \vee o'_2 \in l'_3)$, or $(o_1 \in l'_2) \wedge (o_2 \in l'_1 \vee o_2 \in l'_2)$, or $o_1,o_2 \in l'_1$: Then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_1$: Since the linearization point of $o_2$ is before return action of $o_1$, $o_2$ either overlaps with $o_1$, or happens before $o_1$, which contradicts that $o_1$ happens before $o_2$.
-
-\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_2$: Since $l'_2$ only contains $\textit{put}$ operations with priority $\textit{pri}_x$ and are in $\textit{OP}'_x$, it is not hard to see that each operation in $l'_2$ must overlap with $\textit{put}(x,\textit{pri}_x)$, which contradicts that $o_1$ happens before $o_2$.
-
-\item[-] $o_1 \in l'_3 \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: According to the construction of $l''$, we can see that $1_2 \in l'_1$, which contradicts that $1_2 \in l'_3$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h_x \sqsubseteq l''_x$.
-
-Let $\textit{OP}_{\textit{-x}}$ be the set of $\textit{put}$ operations with non-$\textit{pri}_x$ priority, and their matched $\textit{rm}$ operations.
-Let $h_{\textit{-x}}$ be the projection of $h$ into $\textit{OP}_{\textit{-x}}$.
-Let $l''_{\textit{-x}}$ be the projection of $l'$ into$\textit{OP}_{\textit{-x}}$.
-It is obvious that $h_{\textit{-x}} \sqsubseteq l''_{\textit{-x}}$.
-By Lemma \ref{lemma:merge two linearization}, there exists a sequence $l'''$, such that $h \sqsubseteq l'''$, $l''' \vert{\textit{OP}_x} = l_x''$ and $l''' \vert{\textit{OP}_{\textit{-x}}} = l''_{\textit{-x}}$.
-
-It is obvious that $l''_{\textit{-x}} \in \textit{PQueue}$.
-Let $l'_x$ be the projection of $l'$ into $\textit{OP}_x - \{ \textit{put}(x,\textit{pri}_x) \}$.
-It is obvious that $l'_x \in \textit{PQueue}$ and $l'_1 \in \textit{PQueue}$.
-Since (1) $l'_1 \cdot l'_2 \cdot l'_3$ is obtained from $l'_x$ by making $\textit{put}$ operations in $l'_2$ to happen earlier, and (2) $l'_3$ contains only $\textit{rm}$ operations, we can see that $l'_1 \cdot l'_2 \cdot l'_3 \in \textit{PQueue}$.
-Since $l'_3$ contains only $\textit{rm}$ operations, it is easy to see that $l_x'' \in \textit{PQueue}$.
-Since $\textit{pri}_x$ is the maximal priority of $h$ and $h$ does not contain $\textit{rm}(\textit{Empty})$ operations, the operations in $l_x''$ does not influence the correctness of $l''_{\textit{-x}} \in \textit{PQueue}$.
-Therefore, we can see that $l''' \in \textit{PQueue}$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{Rpr5isStepByStepLinearizability}
-\label{lemma:Rpr5 is step-by-step linearizability}
-
-$R_{\textit{pr5}}$ is step-by-step linearizability.
-
-\end{restatable}
-
-\begin {proof}
-
-Assume that $l$ matches $R_{\textit{pr5}}$ with operation $o=\textit{rmEmpty}$.
-Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$.
-Let $h'=h \setminus o$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$.
-Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$ such that $h' \sqsubseteq l'$.
-
-According to $R_{\textit{pr5}}$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{rmEmpty} \cdot v$, where the $\textit{put}$ operations and $\textit{rm}$ operations in $u$ are matched.
-
-Let $D_L$ be the set of items in $u$ and $D_R$ be the set of items in $v$.
-Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.
-Let sequence $l'' = l'_L \cdot o \cdot L'_R$.
-Since priority queue is closed under projection (Lemma \ref{lemma:PQ is closed under projection}) and $u$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $l'_L \in \textit{PQueue}$ and the the priority queue is empty after executing $l'_L$.
-Then we know that $l'_L \cdot o \in \textit{PQueue}$.
-Since $l'_R$ is obtained from $l'$ by discarding pairs of matched $\textit{put}$ and $\textit{rm}$ operations, it is easy to see that $L'_R \in \textit{PQueue}$, and then we know that $l'' = l'_L \cdot o \cdot L'_R \in \textit{PQueue}$.
-
-It remains to prove that $h \sqsubseteq l''$.
-To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}.
-Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
-Then, we can see
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o_1,o_2 \in l'_L$, or $o_1,o_2 \in l'_R$: Then $l'$ contradicts with happen before relation of $h$.
-
-\item[-] If $o_1=o \wedge o_2 \in l'_L$, or $o_1 \in l'_R \wedge o_2 \in l'_L$, or $o_1 \in l'_R \wedge o_2 = 0$, then $l$ contradicts with happen before relation of $h$.
-\end{itemize}
-
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$. \qed
-\end {proof}
-
-The following lemma states that $\textit{PQueue}$ is step-by-step linearizability, it is a direct consequence of Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, Lemma \ref{lemma:Rpr2 is step-by-step linearizability}, Lemma \ref{lemma:Rpr3 is step-by-step linearizability}, Lemma \ref{lemma:Rpr4 is step-by-step linearizability} and Lemma \ref{lemma:Rpr5 is step-by-step linearizability}.
 

### 2017-02-21 ss-lin-PQ.tex
I add section of step-by-step linearizability. Now Appendix B is the proof of step-by-step linearizability. In this process, proofs of two lemmas are changed.

@@
 Here step-by-step linearizability enables us to build a linearization for an execution $e$ incrementally, using linearizations of projections of $e$.
+Our step-by-step linearizability is inspired by the step-by-step linearizabilility of queue and stacks in \cite{Bouajjani:2015}.
 The proof of lemmas in this section can be found in Appendix \ref{sec:appendix in section step-by-step linearizability of priority queues}.
@@
 The set of projections of $e$ is denoted $\textit{proj}(e)$.
+When refer to $\textit{proj}(e)$, we implicitly assume that each $\textit{rm}(\textit{empty})$ in $e$ has a ghost argument that is unique.
 We write $e \setminus x$ for the projection $e \vert_{ \mathbb{D} \setminus \{ x \} }$.
-This extends naturally to histories.
-A set $S$ of sequential executions is closed under projection, if for all $\mathcal{D} \subseteq \mathbb{D}$ and $e \in S$, we have $e \vert_{ \mathcal{D} } \in S$.
+This extends naturally to histories and concurrent executions.
+
+A set $S$ of sequential executions is closed under projection, if for all $\mathcal{D} \subseteq \mathbb{D}$ and $e \in S$, we have $e \vert_{ \mathcal{D} } \in S$..
 The following lemma states that $\textit{PQueue}$ is closed under projection, since the predicates used in rules of priority queue are ``closed under projection''.
@@
 
-The following lemma states that checking inclusion into $\textit{PQueue}$ of a data-differentiated sequence is equivalent to checking inclusion into $\textit{MS}(R)$ for everyone of its projections.
+The following lemma states that for data-differentiated sequential execution, checking inclusion into $\textit{PQueue}$ is equivalent to checking inclusion into $\textit{MS}(R)$ for everyone of its projections.
 
@@
 \label{lemma:PQ as multi in MRpri for sequence}
-Given a data-differentiated sequential execution $e$, $e \in \textit{PQueue}$, if and only if, $\forall e' \in \textit{proj}(e)$, $e' \in \textit{MS}(R)$, where $R = \textit{last}(e)$.
+Given a data-differentiated sequential execution $e$, $e \in \textit{PQueue}$, if and only if, $\forall e' \in \textit{proj}(e)$, $e' \in \textit{MS}(R)$, where $R = \textit{last}(e')$.
 \end{restatable}
@@
 This extends naturally to histories.
+Similarly we can define projection into method events.
 
@@
 
+Given a data-differentiated execution and its history, we can abuse notation and mix labels and method events with operations themselves, since items are unique in a data-differentiated execution.
+For instance, we will reference an operation labeled by $\textit{put}(p,a)$ as $\textit{put}(p,a)$.
 The following lemma states that $\textit{PQueue}$ is step-by-step linearizability.
@@
 \label{lemma:PQueue is step-by-step linearizability}
+$\textit{PQueue}$ is step-by-step linearizability.
+\end{restatable}
 
-$\textit{PQueue}$ is step-by-step linearizability.
+The following lemma states that for data-differentiated execution, checking linearizability with respect to $\textit{PQueue}$ is equivalent to checking linearizability with respect to $\textit{MS}(R)$ for everyone of its projections.
+Roughly speaking, step-by-step linearizability of priority queue paly a important rule for proof of the $\textit{if}$ direction of Lemma \ref{lemma:PQ as multi in MRpri for history}: It guild us how to build a linearization of a whole execution by increasingly construct linearization of sub-execution from a $\epsilon$ execution.
 
+\begin{restatable}{lemma}{PQasMultiInMRpriforHistory}
+\label{lemma:PQ as multi in MRpri for history}
+Given a data-differentiated execution $e$, $e \sqsubseteq \textit{PQueue}$, if and only if, $\forall e' \in \textit{proj}(e)$, $e' \sqsubseteq \textit{MS}(R)$, where $R = \textit{last}(e')$.
 \end{restatable}

### 2017-02-23 appendix.tex
I add section 6.1, which introduce definition of witness automata, co-regular. I give witness automata, which is enough to ensure that each single-priority execution is FIFO. I add Appendix C.1 to contain proof and definition of section 6.1

@@
 
-\section{Inductive Rules for Priority Queues}
-\label{sec:inductive rules for priority queues}
-
-Given a sequence $l$ of possibly $\textit{put}$, $\textit{rm}$ and $\textit{rmEmpty}$ operations, we say that $l$ contains unmatched $\textit{rm}$ operations, if the number of $\textit{rm}(d)$ operations is larger than the number of $\textit{put}(d,\_)$ operations for some $d$ in $l$.
-By data-independence, we can detect the existence of unmatched $\textit{rm}$ operations by automata.
-A automaton $\mathcal{A}_{\textit{unmRm1}}$ is given in \figurename~\ref{fig:automata for unmatched rm1}, where $c_1=\textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm),\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$, $c_2=c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c_1 + \textit{cal}(\textit{put},b,\textit{anyPri}) + \textit{ret}(\textit{rm},b)$.
-Here $\textit{anyPri}$ is a specific predicates that is satisfied by any priority.
-$\mathcal{A}_{\textit{unmRm1}}$ is used to detect the existence of a item that has been removed but has never been inserted.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_RM1.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{unmRm1}}$}
-  \label{fig:automata for unmatched rm1}
-\end{figure}
-
-An automaton $\mathcal{A}_{\textit{unmRm2}}$ is given in \figurename~\ref{fig:automata for unmatched rm2}, where $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm),$ $\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$, and $c_1 = c_1 + \textit{ret}(\textit{rm},b)$.
-$\mathcal{A}_{\textit{unmRm1}}$ is used to detect the existence of a item that has been removed more than once but has been inserted once.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_RM2.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{unmRm2}}$}
-  \label{fig:automata for unmatched rm2}
-\end{figure}
-
-The following lemma states that we can detect the existence of unmatched $\textit{rm}$ by $\mathcal{A}_{\textit{unmRm1}}$ and $\mathcal{A}_{\textit{unmRm2}}$.
-
-\begin{restatable}{lemma}{DetectUnmatchedRmByAutomata}
-\label{lemma:detect unmatched rm by automata}
-
-Given a data-independent $\mathcal{I}$, there is some unmatched $\textit{rm}$ in some sequence of $\mathcal{I}$, if and only if $( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm1}} \neq \emptyset) \vee ( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm2}} \neq \emptyset )$.
-\end{restatable}
-
-\begin {proof}
-
-For the $\textit{if}$ direction, assume that $( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm1}} \neq \emptyset) \vee ( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm2}} \neq \emptyset )$.
-Then, there exists a sequence $l \in \mathcal{I}$ and $l$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$ or $\mathcal{A}_{\textit{unmRm2}}$.
-If $l$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$, then $l$ contains at least one $\textit{rm}(b)$ operation and no $\textit{put}(b,\_)$ operation; Else, if $l$ is accepted by $\mathcal{A}_{\textit{unmRm2}}$, then $l$ contains at least two $\textit{rm}(b)$ operations and one $\textit{put}(b,\_)$ operation.
-In both cases, $l$ contains unmatched $\textit{rm}$ operations.
-
-To prove the $\textit{only if}$ direction assume that there exists a sequence $l \in \mathcal{I}$ that contains unmatched $\textit{rm}$ operations.
-By data-independence, there exists data-differentiated sequence $l'$ and renaming function $r$, such that $l=r(l')$.
-We need to prove that in $l'$, one of the following cases holds: (1) $l'$ contains $\textit{rm}(d_1)$ for some $d_1$ and does not contain any $\textit{put}(d_1,\_)$, and (2) $l'$ contains one $\textit{put}(d_1,\_)$ operation and more than one $\textit{rm}(d_1)$ operations.
-We prove this by contradiction.
-Assume by contradiction that both cases do not hold.
-Then, for each value $d$ of $l'$, there is either one $\textit{put}(d,\_)$ and one $\textit{rm}(d)$ in $l'$, or one $\textit{put}(d,\_)$ and no $\textit{rm}(d)$ in $l'$.
-It is obvious that such $l'$ does not contain unmatched $\textit{rm}$ operation, and renaming a sequence that does not contain unmatched $\textit{rm}$ we can only obtain another sequence that does not contain unmatched $\textit{rm}$.
-This contradicts that $l=r(l')$ has unmatched $\textit{rm}$.
-
-Let $l''$ be obtained from $ll$ by renaming $d_1$ to $b$ and renaming other value to $a$.
-If the first case holds, $l''$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$; Otherwise, the second case holds, and $l''$ is accepted by $\mathcal{A}_{\textit{unmRm2}}$.
-\qed
-\end {proof}
-
-Therefore, from now on, we can safely assume that there are no unmatched $\textit{rm}$ operations.
+\section{Proofs and Definitions in Section \ref{sec:co-regular of priority queues}}
+\label{sec:appendix proof and definition in section co-regular of priority queues}
+
+\subsection{Proofs and Definitions in Subsection \ref{subsec:definition of co-regular}}
+\label{sec:appendix proof and definition in section definition of co-regular}
+
+\cite{Bouajjani:2015} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $a,b$: (1) $\textit{deq}(b) <_{hb} \textit{enq}(b)$, (2) there are are no $\textit{enq}(b)$ and least one $\textit{deq}(b)$, (3) there are are one $\textit{enq}(b)$ and more than one $\textit{deq}(b)$, and (4) $\textit{enq}(a) <_{\textit{hb}} \textit{enq}(b)$, and $\textit{deq}(b) <_{\textit{hb}} \textit{deq}(a)$, or $\textit{deq}(a)$ does not exists.
+
+For each such case, we construct a witness automata.
+Recall that in Section \ref{subsec:definition of co-regular}, we generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}.
+Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri})$, $\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},b),\textit{rm}(\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.6 \textwidth]{figures/PIC_AUTO_FIFO_1.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{SinPri}}^1$}
+  \label{fig:automata for FIFO-1 in appendix}
+\end{figure}
+
+We generate witness automata $\mathcal{A}_{\textit{SinPri}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata for FIFO-2}.
+Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{rm}(\textit{empty})$, $c_2 = c_1 + \textit{cal}(\textit{rm},b) + \textit{ret}(\textit{rm},b)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.3 \textwidth]{figures/PIC_AUTO_FIFO_2.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{SinPri}}^2$}
+  \label{fig:automata for FIFO-2}
+\end{figure}
+
+We generate witness automata $\mathcal{A}_{\textit{SinPri}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata for FIFO-3}.
+Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{rm}(\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},b)$, $c_4 = c_3 + \textit{cal}(\textit{rm},b)$, $c_5 = c_1 + \textit{ret}(\textit{rm},b)$, $c_6 = c_5 + \textit{cal}(\textit{rm},b)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.7 \textwidth]{figures/PIC_AUTO_FIFO_3.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{SinPri}}^3$}
+  \label{fig:automata for FIFO-3}
+\end{figure}
+
+We generate witness automata $\mathcal{A}_{\textit{SinPri}}^4$ for the forth case, and it is shown in \figurename~\ref{fig:automata for FIFO-4}.
+Here $c_1 = c + \textit{cal}(\textit{rm},b)$, and $c_2 = c + \textit{ret}(\textit{put},b) + \textit{cal}(\textit{rm},a) + \textit{ret}(\textit{rm},a)$, where $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d),\textit{ret}(\textit{rm},d),\textit{rm}(\textit{empty})$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.9 \textwidth]{figures/PIC_AUTO_FIFO_4.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{SinPri}}^4$}
+  \label{fig:automata for FIFO-4}
+\end{figure}
+
+Let $\textit{Auts}_{\textit{sinPri}} = \{ \mathcal{A}_{\textit{SinPri}}^1, \mathcal{A}_{\textit{SinPri}}^2, \mathcal{A}_{\textit{SinPri}}^3, \mathcal{A}_{\textit{SinPri}}^4 \}$.
+Let us prove Lemma \ref{lemma:automata for priority queue with single priority}.
+
+\AutoForPQwithSignlePri*
+
+\begin {proof}
+
+\cite{Bouajjani:2015} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $v_a,v_b$: (1) $\textit{deq}(v_b) <_{hb} \textit{enq}(v_b)$, (2) there are are no $\textit{enq}(v_b)$ and least one $\textit{deq}(v_b)$, (3) there are are one $\textit{enq}(v_b)$ and more than one $\textit{deq}(v_b)$, and (4) $\textit{enq}(v_a) <_{\textit{hb}} \textit{enq}(v_b)$, and $\textit{deq}(v_b) <_{\textit{hb}} \textit{deq}(v_a)$, or $\textit{deq}(v_a)$ does not exists.
+
+Let us prove the $\textit{only if}$ direction.
+Assume that there exists execution $e_0 \in \mathcal{I}$ and $e_0$ is accepted by an automaton in $\textit{Auts}_{\textit{sinPri}}$.
+By data-independence, we can see that there exists a data-differentiated $e \in \mathcal{I}$ and renaming function, such that $e_0=r(e)$.
+Let $e'$ be obtained from $e$ by erasing $\textit{rm}(\textit{empty})$ and it is obvious that $e' \in \textit{proj}(e)$.
+It is easy to see that $e'$ satisfies one of above conditions, and then $e'$ is not linearizable w.r.t queue.
+
+Let us prove the $\textit{if}$ direction.
+Assume that exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{transToQueue}(e')$ does not linearizable to queue.
+Then we construct a renaming function $r$ as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If this is because case $1$, case $2$ or case $3$: $r$ maps $v_b$ into $b$ and maps all other values into $a$.
+
+\item[-] If this is because case $4$: $r$ maps $v_a$ and $v_b$ into $a$ and $b$, respectively, and maps all other values into $d$.
+\end{itemize}
+
+Then it is not hard to see that $r(e) \in \mathcal{I}$ and it is accepted by some automaton in $\textit{Auts}_{\textit{sinPri}}$. This completes the proof of this lemma. \qed
+\end {proof}
 
@@
 \label{subsec:co-regular of Rpr2}
-
-An automaton $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$ is given in \figurename~\ref{fig:automata for FIFO of single priority}.
-In \figurename~\ref{fig:automata for FIFO of single priority}, let $c_1= \textit{cal}(\textit{put},c,\textit{anyPri}),\textit{ret}(\textit{put},c)$, $\textit{cal}(\textit{rm},c),\textit{ret}(\textit{rm},c)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, and $c_3 = c_2 + \textit{cal}(\textit{rm},a) + \textit{ret}(\textit{rm},a)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_QUEUE_FOR_SINGLE_PRI.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$}
-  \label{fig:automata for FIFO of single priority}
-\end{figure}
-
-According to \cite{Bouajjani:2015}, given a data-independent set $S$ of queue executions, whether every sequence $s$ in $S$ is linearizable to queue can be reduced into checking emptiness of intersection of $S$ and several finite automata.
-Let $\textit{transToQueue}(h)$ be a sequence $h'$ that is generated from $h$ by transforming $\textit{put}$ and $\textit{rm}$ into $\textit{enq}$ and $\textit{deq}$, respectively, and then discarding priorities.
-A priority queue history is called single-priority, if all of its items has only one priority.
-The following lemma states that we can also use finite automata to check ``first in first out property'' for single-priority sub-sequence of priority queue history.
-
-\begin{restatable}{lemma}{AutoForPQwithSignlePri}
-\label{lemma:automata for priority queue with single priority}
-
-Given a data-independent implementations $\mathcal{I}$ of priority queue, let $\mathcal{A} = \{ \mathcal{A}_{\textit{unmRm1}}, \mathcal{A}_{\textit{Rpr0-1}}, \mathcal{A}_{\textit{unmRm2}}, \mathcal{A}_{\textit{FIFO-Sig-Pri}} \}$, then,
-
-$\mathcal{I} \cap \mathcal{A} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority, and $\textit{transToQueue}(e')$ does not linearizable to queue.
-\end{restatable}
-
-\begin {proof}
-
-According to \cite{Bouajjani:2015}, checking linearizable with respect to queue can be reduced into checking the following three conditions for data-differentiated histories:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{deq}(a)$ happens before $\textit{enq}(a)$ (or $\textit{enq}(a)$ does not exist) for some item $a$,
-
-\item[-] there are two or more $\textit{deq}(a)$ for some item $a$,
-
-\item[-] $\textit{enq}(a)$ happens before $\textit{enq}(b)$, and $\textit{deq}(b)$ happens before $\textit{deq}(a)$ (or $\textit{deq}(a)$ does not exists) for some items $a,b$.
-\end{itemize}
-
-It is not hard to see that (1) the first condition is dealt with by $\mathcal{A}_{\textit{unmRm1}}$ (if $\textit{put}(a,\_)$ does not exist) and $\mathcal{A}_{\textit{Rpr0-1}}$ (if $\textit{put}(a,\_)$ exists), (2) the second condition is dealt with by $\mathcal{A}_{\textit{unmRm2}}$, and (3) the third condition is dealt with by $\mathcal{A}_{\textit{FIFO-Sig-Pri}}$.
-\qed
-\end {proof}
 

### 2017-02-23 co-regular-PQ.tex
I add section 6.1, which introduce definition of witness automata, co-regular. I give witness automata, which is enough to ensure that each single-priority execution is FIFO. I add Appendix C.1 to contain proof and definition of section 6.1

\section{Co-Regular of Priority Queues}
\label{sec:co-regular of priority queues}

A rule $R$ of priority queue is co-regular, if checking linearizability with respect to $\textit{MS}(R)$ of a data-independent implementation $\mathcal{I}$ can be reduced to checking the emptiness of intersection between $\mathcal{I}$ and a set of witness automata. In this section, we propose the definition of witness automata and co-regular. Then we prove that all five rules of priority queues are co-regular, and roughly introduce the proof idea. With the help of step-by-step linearizability and co-regular, we finally reduce the linearizability problem of $\textit{PQueue}$ into emptiness problem of intersection with automata.

\subsection{Definition of Co-Regular}
\label{subsec:definition of co-regular}

A witness automaton is a finite automaton with alphabet $\{ \textit{cal}(\textit{put},d,\textit{pre}), \textit{ret}(\textit{put}d),$ $\textit{cal}(\textit{rm},d),\textit{ret}(\textit{rm},d) \vert d \in \mathbb{D},\textit{red} \in \textit{preWA} \}$. Here $\textit{preWA}$ is the set of predicate of priorities, and it contains

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] A predicate variable $p$, which accepts some specific priority,

\item[-] A predicate $\textit{les}_p$, which accepts all the priorities that are less than the value of $p$,

\item[-] A predicate $\textit{anyPri}$, which accepts any priority.
\end{itemize}

Given a execution $e = \alpha_1 \cdots \alpha_k$ of priority queue and a witness automaton $\mathcal{A}$, we say that $e$ is accepted by $\mathcal{A}$, if

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] There exist transitions $q_0 \xrightarrow{\beta_1} q_1 \ldots \xrightarrow{\beta_k} q_k$ of $\mathcal{A}$, such that $q_0$ is one of initial state of $\mathcal{A}$, and $q_k$ is one of accept state of $\mathcal{A}$. Let $d_p \in \mathbb{D}$.

\item[-] For each $i$, if $\alpha_i = \textit{cal}(\textit{put},a,q)$, then either (1) $\beta_i = \textit{cal}(\textit{put},a,p)$ and $q = d_p$, or (2) $\beta_i = \textit{cal}(\textit{put},a,\textit{les}_p)$ and $q < d_p$, or (3) $beta_i = \textit{cal}(\textit{put},a,\textit{anyPri})$.

\item[-] For each $i$, if $\alpha_i = \textit{ret}(\textit{put},a)$, or $\alpha_i = \textit{cal}(\textit{rm},a)$, or $\alpha_i = \textit{ret}(\textit{rm},a)$, then $\beta_i = \alpha_i$.
\end{itemize}

Note that witness automata does not read operations, since the domains of operations is infinite.

Let us introduce the notion of co-regular:

%\vspace{-6pt}
\begin{definition}\label{def:co-regular of rules of priority queues}
A rule $R$ of priority queue is co-regular, if there are a finite set $\textit{Auts}_{R}$ of witness automata such that, for each data-independence implementation $\mathcal{I}$, we have that

$$ \textit{Auts}_{R} \cap \mathcal{I} = \emptyset \Leftrightarrow \exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), last(e')=R \wedge e \ does \ not \ linearizable \ w.r.t. \ \textit{MS}(R)$$

We say that $\textit{PQueue}$ is regular, if each of its rule is co-regular.
\end{definition}

Before we go to investigate co-regular of each rules, we use the results in \cite{Bouajjani:2015} to simplify our work. \cite{Bouajjani:2015} states that checking linearizability w.r.t queue can be reduced into checking emptiness of intersection between $\mathcal{I}$ and a set of automata. Given a data-differentiated execution $e$, let $e \vert_{i}$ be an execution generated from $e$ by erasing call and return actions of items that does not use priority $i$. We call a priority queue execution with only one priority a single-priority execution. Let $\textit{transToQueue}(e)$ be an execution generated from $e$ by transforming $\textit{put}$ and $\textit{rm}$ into $\textit{enq}$ and $\textit{deq}$, respectively, and then discarding priorities. We can see that for each $e \in \textit{PQueue}$ and each priority $i$, $\textit{transToQueue}(e \vert_{i})$ satisfy FIFO (first in first out) property.

Given an execution of queue, we say that it is differentiated \cite{Wolper:1986}, if each item is enqueued at most once. \cite{Bouajjani:2015} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $a,b$: (1) $\textit{deq}(b) <_{hb} \textit{enq}(b)$, (2) there are are no $\textit{enq}(b)$ and least one $\textit{deq}(b)$, (3) there are are one $\textit{enq}(b)$ and more than one $\textit{deq}(b)$, and (4) $\textit{enq}(a) <_{\textit{hb}} \textit{enq}(b)$, and $\textit{deq}(b) <_{\textit{hb}} \textit{deq}(a)$, or $\textit{deq}(a)$ does not exists. For each such case, we can construct a witness automata for priority queue. For example, for the first case, we generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ in \figurename~\ref{fig:automata for FIFO-1}, here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}), \textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a)$, $\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},b)$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6 \textwidth]{figures/PIC_AUTO_FIFO_1.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{SinPri}}^1$}
  \label{fig:automata for FIFO-1}
\end{figure}

In Appendix \ref{sec:appendix proof and definition in section definition of co-regular}, we construct a set $\textit{Auts}_{\textit{sinPri}}$ of witness automata ($\mathcal{A}_{\textit{SinPri}}^1$ is in it), and shows that they are enough to ensure that for each of its data-differentiated executions, each of its single-priority projection that has no $\textit{rm}(\textit{empty})$ to have ``FIFO'' property, as shown by the following lemma.

\begin{restatable}{lemma}{AutoForPQwithSignlePri}
\label{lemma:automata for priority queue with single priority}

Given a data-independent implementations $\mathcal{I}$ of priority queue, $\mathcal{I} \cap \textit{Auts}_{\textit{sinPri}} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{transToQueue}(e')$ does not linearizable to queue.
\end{restatable}

According to Lemma \ref{lemma:automata for priority queue with single priority}, from now on, it is safe to assume that, for each data-differentiated execution without $\textit{rm}(\textit{empty})$, any of its single-priority projection has ``FIFO'' property. For example, $\textit{rm}(a)$ never happens before $\textit{put}(a,\_)$ for each $a$.

### 2017-02-23 appendix.tex
I add subsection of co-regular of PQ1>. I also add explaination for clarity. Now Appendix C.2 is for this subsection.

@@
 For each such case, we construct a witness automata.
-Recall that in Section \ref{subsec:definition of co-regular}, we generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}.
+We generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}.
 Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri})$, $\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},b),\textit{rm}(\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$.
@@
 
+\subsection{Proofs and Definitions in Subsection \ref{subsec:co-regular of PQ1Lar}}
+\label{sec:appendix proof and definition in section co-regular of PQ1Lar}
+
+We can see that $\textit{UVSet}_i(e,x) \cap \textit{UVSet}_j(e,x) = \emptyset$ for any $i \neq j$.
+
+The following lemma states that $\textit{UVSet}(e,x)$ contains only matched $\textit{put}$ and $\textit{rm}$.
+
+\begin{restatable}{lemma}{UVSetHasMatchedPutandRm}
+\label{lemma:UVSet has matched put and rm}
+Given a data-differentiated execution $e$ with $\textit{last}(e) = \textit{PQ}_1^{>}$.
+Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority.
+Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
+Assume that $G$ has no cycle going through $x$.
+Then, $\textit{UVSet}(e,x)$ contains only matched $\textit{put}$ and $\textit{rm}$.
+\end{restatable}
+\begin {proof}
+
+We prove this lemma by contradiction.
+Assume that there exists a value, such that $\textit{UVSet}(e,x)$ contains only its $\textit{put}$ and does not contain its $\textit{rm}$.
+Then we can see that there exists $d_1,\ldots,d_j$.
+Intuitively, $d_1,\ldots,d_j$ are elements in $\textit{UVSet}_1(e,x), \ldots, \textit{UVSet}_i(e,x)$, respectively.
+$\textit{UVSet}(e,x)$ contains $\textit{put}(d_j,\_)$ and does not contain $\textit{rm}(d_j)$.
+And each $d_i$ is the reason of $d_{\textit{i+1}} \in \textit{UVSet}_{\textit{i+1}}(e,x)$.
+Formally, we require that
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] For each $1 \leq i \leq j$, method events of $d_i$ belongs to $\textit{UVSet}_i(e,x)$.
+
+\item[-] For each $i \neq j$, $\textit{put}(d_i,\_),\textit{rm}(d_i) \in \textit{UVSet}_i(e,x)$. $\textit{put}(d_j,\_) \in \textit{UVSet}_j(e,x)$, and $e$ does not contain $\textit{rm}(d_j)$.
+
+\item[-] An operation of $d_1$ happens before an operations of $x$. For each $1 < i \leq j$, an operation of $d_i$ happens an operation of $d_{\textit{i-1}}$.
+
+\item[-] For each $k$ and $\textit{ind}$, if $k > \textit{ind+1}$, then no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.
+\end{itemize}
+
+According to the definition of $\textit{UVSet}(e,x)$, it is easy to see that such $d_1,\ldots,d_j$ exists. Let us prove the following fact:
+
+\noindent {\bf $\textit{fact}_1$}: Given $1 \leq i < j$, it can not be the case that $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap.
+
+Proof of $\textit{fact}_1$: We prove $\textit{fact}_1$ by contradiction.
+Assume that for some $i \neq j$, $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap.
+Since $\textit{put}(d_i,\_), \textit{rm}(d_i) \in \textit{UVSet}_i(h,x)$, we know that an operation $o_i$ of $d_i$ happens before operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$.
+Moreover, since $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap, it is not hard to see that the call action of $\textit{put}(d_i,\_)$ and the call action of $\textit{rm}(d_i)$ is before the call action of $o_{\textit{i-1}}$.
+Since method events of $d_{\textit{i+1}}$ is in $\textit{UVSet}_{\textit{i+1}}(e,x)$, we know that an operation $o'_{\textit{i+1}}$ of $d_{\textit{i+1}}$ happens before operation $o'_i$ of $d_i$.
+Then, it is not hard to see that $o'_{\textit{i+1}}$ also happens before $o_{\textit{i-1}}$, which contradicts that for each $k > \textit{ind+1}$, no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.
+
+We already know that an operation of $d_1$ happens before an operation of $x$.
+By $\textit{fact}_1$, we can ensure that $\textit{put}(d_1,\_)$ happens before an operation of $x$, and then $d_1 \rightarrow x$ in $G$.
+For each $1 < i \leq j$, we know that an operation $o_i$ of $d_i$ happens before an operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$.
+By $\textit{fact}_1$, we can ensure that $o_i=\textit{put}(d_i,\_)$ and $o_{\textit{i-1}}=\textit{rm}(d_{\textit{i-1}})$, and then $d_i \rightarrow d_{\textit{i-1}}$ in $G$.
+Since $h$ contains $\textit{put}(d_j,\_)$ and does not contain $\textit{rm}(d_j)$, we know that $x \rightarrow d_j$ in $G$.
+Then $G$ has a cycle going through $x$, contradicts that $G$ has no cycle going through $x$.
+\qed
+\end {proof}
+
+The following lemma states that $\textit{UVSet}(e,x)$ does not happen before $\textit{rm}(x)$ when the left-right constraint has no cycle going through $x$.
+
+\begin{restatable}{lemma}{RmxDoesNotHappenBeforeUVSetForPQ1Lar}
+\label{lemma:Rmx does not happen before UVSet for PQ1Lar}
+
+Given a data-differentiated execution $e$ with $\textit{last}(e) = \textit{PQ}_1^{>}$.
+Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority.
+Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
+Assume that $G$ has no cycle going through $x$.
+Then, $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}(e,x)$.
+\end{restatable}
+
+\begin {proof}
+
+We prove this lemma by induction, and prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x)$, in $\textit{UVSet}_2(e,x)$, $\ldots$.
+Note that, by Lemma \ref{lemma:UVSet has matched put and rm}, $\textit{UVSet}(e,x)$ contains only matched $\textit{put}$ and $\textit{rm}$, and it is easy to see that for each $i$, $\textit{UVSet}_i(e,x)$ contains only matched $\textit{put}$ and $\textit{rm}$.
+
+\noindent (1) Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x)$ by contradiction.
+Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{UVSet}_1(e,x)$ is an operation of item $d$.
+%(according to the definition of $\textit{UVSet}_1(h,x)$, the priority of $d$ does not equals $\textit{pri}_x$).
+
+We use a triple $(t_1,t_2,t_3)$ to represent related information.
+$t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$.
+$t_1$ represents whether $o$ is a $\textit{put}$ method event or a $\textit{rm}$ method event.
+$t_2$ and $t_3$ is used for the reason of $o \in \textit{UVSet}_1(e,x)$: $o \in \textit{UVSet}_1(e,x)$, since an operation (of kind $t_2$) of $d$ happens before an operation (of kind $t_3$) of $x$.
+Let us consider all the possible cases of $(t_1,t_2,t_3)$:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.
+
+\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{rm}(x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{rm}(x)$.
+
+\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x) )$.
+By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(x,\textit{pri}_x) ) \vee ( \textit{rm}(d) <_{hb} \textit{put}(d,\_) )$, which is impossible.
+
+\item[-] $(\textit{put},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{rm}(x) )$.
+We can see that $\textit{rm}(d) <_{hb} \textit{rm}(x) <_{hb} \textit{put}(d,\_)$, which contradicts that $\textit{rm}(d)$ does not happen before $\textit{put}(d,\_)$.
+
+\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x) )$.
+We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.
+
+\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{rm}(x) )$.
+We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.
+
+\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.
+
+\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{rm}(x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{rm}(x)$.
+\end{itemize}
+
+This completes the proof for $\textit{UVSet}_1(e,x)$.
+
+\noindent (2) Assume we already prove that for some $j \geq 1$, $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$.
+Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_{\textit{j+1}}(e,x)$ by contradiction.
+Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$ is an operation of item $d_{\textit{j+1}}$.
+We use a triple $(t_1,t_2,t_3)$ to represent related information.
+$t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$.
+$t_1$ represents whether $o$ is a $\textit{put}$ method event or a $\textit{rm}$ method event.
+$t_2$ and $t_3$ is used for the reason of $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$: $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$, since an operation (of kind $t_2$) of $d_{\textit{j+1}}$ happens before an operation (of kind $t_3$) of $d_j$, where $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{UVSet}_j(e,x)$.
+Let us consider all the possible cases of $(t_1,t_2,t_3)$:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_)$.
+We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{UVSet}_j(e,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$.
+
+\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j,\_)$.
+We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j,\_) ) \wedge ( \textit{rm}(d_j) \in \textit{UVSet}_j(e,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$.
+
+\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$.
+By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \vee ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_{\textit{j+1}},\_) )$, which is impossible.
+
+\item[-] $(\textit{put},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$.
+By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \vee ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_{\textit{j+1}},\_) )$, which is impossible.
+
+\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) )$.
+Let us consider the reason of $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{UVSet}_j(e,x)$:
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $( j > 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{UVSet}_{\textit{j-1}}(e,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, we can see that $\textit{put}(d_{\textit{j+1}},\_) <_{hb} o''$, and then operations of $d_{\textit{j+1}}$ is in $\textit{UVSet}_j(e,x)$, contradicts that operations of $d_{\textit{j+1}}$ is in $\textit{UVSet}_{\textit{j+1}}(e,x)$.
+
+    \item[-] If $( j = 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of $x$: Similar to above case.
+
+    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{UVSet}_{\textit{j-1}}(e,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, we can see that $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} o'' ) \vee ( \textit{rm}(d_j) <_{hb} \textit{put}(d_j,\_) )$, which is impossible.
+
+    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of $x$: Similar to above case.
+    \end{itemize}
+
+\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Let $T_{\textit{ind}}$ be the set of sentences $\{ \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}), \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j),\ldots, \textit{put}(d_{\textit{ind+1}},\_) <_{hb} \textit{rm}(d_{\textit{ind}}) \}$.
+Here each $d_i$ is a item of some operation in $\textit{UVSet}_i(e,x)$.
+Let us prove that from $T_j$ we can obtain contradiction by induction:
+
+    {\bf Base case $1$}: From $T_1$ we can obtain contradiction.
+
+    Let us prove base case $1$:
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $\textit{put}(d_1,\_)$ happens $o$, and $o$ is an operation of $x$.
+Then there is a cycle $x \rightarrow d_{\textit{j+1}} \rightarrow \ldots \rightarrow d_1 \rightarrow x$ in $G$, contradicts that $G$ has no cycle going through $x$.
+
+    \item[-] If $\textit{rm}(d_1)$ happens before $o$, and $o$ is an operation of $x$.
+Then since $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1) <_{hb} o$, we can see that $\textit{put}(d_2,\_) <_{hb} o$, and then $\textit{put}(d_2,\_) \in \textit{UVSet}_1(e,x)$, contradicts that $\textit{put}(d_2,\_) \in \textit{UVSet}_2(e,x)$.
+    \end{itemize}
+
+    {\bf Base case $2$}: From $T_2$ we can obtain contradiction.
+
+    Let us prove base case $2$: If $\textit{rm}(d_2) <_{hb} o$, and $o$ is an operation of $d_1$, then since $( \textit{put}(d_3,\_) <_{hb} \textit{rm}(d_2) ) \wedge ( \textit{rm}(d_2) <_{hb} o )$, we know that $\textit{put}(d_3,\_) <_{hb} o$.
+This implies that $\textit{put}(d_3,\_) \in \textit{UVSet}_2(e,x)$, contradicts that $\textit{rm}(d_3,\_) \in \textit{UVSet}_3(e,x)$.
+Therefore, it is only possible that $\textit{put}(d_2,\_)$ happens before an operation of $d_1$.
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{put}(d_1,\_)$ and $\textit{put}(d_1,\_)$ happens before operations of $x$, then we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, which is impossible.
+
+    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{put}(d_1,\_)$ and $\textit{rm}(d_1)$ happens before operations of $x$, then by interval order, we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, or $\textit{rm}(d_1) <_{hb} \textit{put}(d_1,\_)$, which is impossible.
+
+    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{put}(d_1,\_)$ happens before operations of $x$, then $x \rightarrow d_{\textit{j+1}} \rightarrow \ldots \rightarrow d_1 \rightarrow x$ in $G$, contradicts that $G$ has no cycle going through $x$.
+
+    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1)$ happens before operations of $x$, then we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, which is impossible.
+    \end{itemize}
+
+    {\bf induction step}: Given $\textit{ind} \geq 3$, if from $T_{\textit{ind-1}}$ we can obtain contradiction, then from $T_{\textit{ind}}$ we can also contain contradiction.
+
+    Prove of the induction step: Similarly as base case $2$, we can prove that it is only possible that $\textit{put}(d_{\textit{ind}},\_)$ happens before operations of $d_{\textit{ind-1}}$.
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$ and $\textit{put}(d_{\textit{ind-1}},\_)$ happens before operations of $d_{\textit{ind-2}}$, then we know that $\textit{put}(d_{\textit{ind}})$ happens before operation of $d_{\textit{ind-2}}$, which is impossible.
+
+    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$ and $\textit{rm}(d_{\textit{ind-1}})$ happens before operations of $d_{\textit{ind-2}}$, then by interval order, we know that $\textit{put}(d_{\textit{ind}},\_)$ happens before operation of $d_{\textit{ind-2}}$, or $\textit{rm}(d_{\textit{ind-1}}) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$, which is impossible.
+
+    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{rm}(d_{\textit{ind-1}})$, then we obtain $T_{\textit{ind-1}}$, which already contain contradiction.
+    \end{itemize}
+
+    By base case $1$, base case $2$ and the induction step, it is easy to see that for each $i$, $T_i$ contains contradiction.
+Therefore, $T_j$, the case of $(\textit{rm},\textit{put},\textit{rm})$, contains contradiction.
+
+\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$.
+We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{UVSet}_j(e,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$.
+
+\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$.
+We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \wedge ( \textit{rm}(d_j) \in \textit{UVSet}_j(e,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$.
+\end{itemize}
+
+This completes the proof for $\textit{UVSet}_{\textit{j+1}}(e,x)$.
+Therefore, $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}(e,x) = \textit{UVSet}_1(e,x) \cup \textit{UVSet}_2(e,x) \cup \ldots$.
+\qed
+\end {proof}
+
+With Lemma \ref{lemma:UVSet has matched put and rm} and Lemma \ref{lemma:Rmx does not happen before UVSet for PQ1Lar}, we can now prove Lemma \ref{lemma:Lin Equals Constraint for PQ1Lar}.
+
+\LinEqualsConstraintforPQOneLar*
+
+\begin {proof}
+
+To prove the $\textit{only if}$ direction, assume that $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{>})$.
+Let $u$, $v$ and $w$ be the sequences of method events in $\textit{PQ}_1^{>}$, and let $U$, $V$ and $W$ be the set of method events of $u$, $v$ and $w$, respectively.
+Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ in $G$.
+It is obvious that the priority of each $d_i$ is less than $\textit{pri}_x$.
+Then our proof proceeds as follows:
+
+According to the definition of left-right constraint, there are two possibilities.
+The first possibility is that, $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$.
+It is obvious that $\textit{rm}(d_1) \in W$, and then since $U \cup V$ contains matched $\textit{put}$ and $\textit{rm}$, we can see that $\textit{put}(d_1),\textit{rm}(d_1) \in W$.
+Then,
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Since $d_1 \rightarrow d_2$, by definition of $G$, we know that $\textit{put}(d_1)$ happens before $\textit{rm}(d_2)$.
+Since $\textit{put}(d_1) \in W$ and $U \cup V$ contains matched $\textit{put}$ and $\textit{rm}$, we know that $\textit{put}(d_2),\textit{rm}(d_2) \in W$.
+Similarly, for each $1 \leq i \leq m$, we know that $\textit{put}(d_i),\textit{rm}(d_i) \in W$.
+
+\item[-] Since $d_m \rightarrow x$,
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] if $\textit{put}(d_m)$ happens before $\textit{put}(x)$, then we can see that $\textit{put}(d_m) \in U$, which contradicts that $\textit{put}(d_m) \in W$.
+
+    \item[-] if $\textit{put}(d_m)$ happens before $\textit{rm}(x)$, then we can see that $\textit{put}(d_m) \in U \cup V$, which contradicts that $\textit{put}(d_m) \in W$.
+    \end{itemize}
+\end{itemize}
+
+The second possibility is that, $e$ contains one $\textit{put}(d_1,\_)$ and no $\textit{rm}(d_1)$.
+Note that for each $j > 1$, $e$ contains $\textit{put}(d_j,\_)$ and $\textit{rm}(d_j)$.
+Since $d_m \rightarrow x$, is is obvious that $\textit{put}(d_m) \in U \cup V$.
+Since $U \cup V$ contains matched $\textit{put}$ and $\textit{rm}$, we know that $\textit{put}(d_m),\textit{rm}(d_m) \in U \cup V$.
+Then, since $d_{\textit{m-1}} \rightarrow d_m$, by definition of $G$, we know that $\textit{put}(d_{\textit{m-1}})$ happens before $\textit{rm}(d_m)$.
+Since $\textit{rm}(d_m) \in U \cup V$ and $U \cup V$ contains matched $\textit{put}$ and $\textit{rm}$, we know that $\textit{put}(d_{\textit{m-1}}),\textit{rm}(d_{\textit{m-1}}) \in U \cup V$.
+Similarly, for each $1 < i \leq m$, we know that $\textit{put}(d_i),\textit{rm}(d_i) \in U \cup V$, and also $\textit{put}(d_1)\in U \cup V$.
+However, there is one $\textit{put}(d_1,\_)$ and no $\textit{rm}(d_1)$ in $e$, contradicts that $U \cup V$ contains matched $\textit{put}$ and $\textit{rm}$.
+
+This completes the proof of the $\textit{only if}$ direction.
+
+To prove the $\textit{if}$ direction, assume that $G$ has no cycle going through $x$.
+Let $E_u$ be the set of operations that happen before $\textit{put}(x)$ in $e$.
+It is easy to see that $E_u \subseteq \textit{UVSet}(e,x)$.
+Let $E_v = \textit{UVSet}(e,x) \setminus E_u$.
+Let $E_e$ be the set of operations of $e$, and let $E_w = E_e \setminus \textit{UVSet}(e,x)$.
+
+By Lemma \ref{lemma:UVSet has matched put and rm}, we can see that $E_u \cup E_v$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
+It remains to prove that for $E_u$, $\{ \textit{put}(x,\textit{pri}_x) \}$, $E_v$, $\{ \textit{rm}(x) \}$, $E_w$, no elements of the latter set happens before elements of the former set.
+We prove this by showing that all the following cases are impossible:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Case $1$: Some operation $o_w \in E_w$ happens before $\textit{rm}(x)$. Then we know that $o_w \in \textit{UVSet}(e,x) = E_u \cup E_v$, which contradicts that $o_w \in E_w$.
+
+\item[-] Case $2$: Some operation $o_w \in E_w$ happens before some operation $o_{\textit{uv}} \in E_u \cup E_v$.
+Then we know that $o_w \in \textit{UVSet}(e,x) = E_u \cup E_v$, which contradicts that $o_w \in E_w$.
+
+\item[-] Case $3$: Some operation $o_w \in E_w$ happens before $\textit{put}(x)$. Then we know that $o_w \in \textit{UVSet}(e,x) = E_u \cup E_v$, which contradicts that $o_w \in E_w$.
+
+\item[-] Case $4$: $\textit{rm}(x)$ happens before some $o_{\textit{uv}} \in \textit{UVSet}(e,x) = E_u \cup E_v$.
+By Lemma \ref{lemma:Rmx does not happen before UVSet for PQ1Lar} we know that this is impossible.
+
+\item[-] Case $5$: $\textit{rm}(x)$ happens before $\textit{put}(x)$. This contradicts that each single-priority projection satisfy the FIFO property.
+
+\item[-] Case $6$: Some operation $o_v \in E_v$ happens before $\textit{put}(x)$. Then we know that $o_v \in E_u$, which contradicts that $o_v \in E_v$.
+
+\item[-] Case $7$: Some operation $o_v \in E_v$ happens before some operation $o_u \in E_u$. Then we know that $o_v \in E_u$, which contradicts that $o_v \in E_v$.
+
+\item[-] Case $8$: $\textit{put}(x)$ happens before some operation $o_u \in E_u$. This is impossible.
+\end{itemize}
+
+This completes the proof of the $\textit{if}$ direction.
+
+\qed
+\end {proof}
+
+Let us begin to represent witness automata that is used for capture the existence of a data-differentiated execution $e$, it has a projection $e'$, $\textit{last}(e') = \textit{PQ}_1^{>}$, and there exists a cycle going through the item with maximal priority in $e'$.
+By data-independence, we can obtain $e_r$ from $e$ by renaming function, which maps such item to be $b$, maps items that cover it to be $a$, and maps other items into $d$.
+There are four possible enumeration of call and return actions of $\textit{put}(b)$ and $\textit{rm}(b)$.
+For each of them, we generate a witness automaton.
+
+For the case when $e_r \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^1$, as shown in \figurename~\ref{fig:automata APQ1Lar-1}.
+Here $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{cal}(\textit{put},a,\textit{les}_p)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d), \textit{ret}(\textit{rm},d)$.
+The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^1$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=1 \textwidth]{figures/PIC_AUTO_PQ1Lar-pprr.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{l-lar}}^1$}
+  \label{fig:automata APQ1Lar-1}
+\end{figure}
+
+$\mathcal{A}_{\textit{l-lar}}^1$ is used to recognize conditions in \figurename~\ref{fig:his for APQ1Lar-1}.
+Here for simplicity, we only draw operation of $b$, and the return action of the first call of $\textit{ret}(\textit{put},a)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=1 \textwidth]{figures/PIC_HIS_PQ1Lar-pprr.pdf}
+%\vspace{-10pt}
+  \caption{Conditions recognized by $\mathcal{A}_{\textit{l-lar}}^1$}
+  \label{fig:his for APQ1Lar-1}
+\end{figure}
+
+For the case when $e_r \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^2$, as shown in \figurename~\ref{fig:automata APQ1Lar-2}.
+Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$.
+The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^2$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=1 \textwidth]{figures/PIC_AUTO_PQ1Lar-prpr.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{l-lar}}^2$}
+  \label{fig:automata APQ1Lar-2}
+\end{figure}
+
+$\mathcal{A}_{\textit{l-lar}}^2$ is used to recognize conditions in \figurename~\ref{fig:his for APQ1Lar-2}.
+Here for simplicity, we only draw operation of $b$, and the return action of the first call of $\textit{ret}(\textit{put},a)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.7 \textwidth]{figures/PIC_HIS_PQ1Lar-prpr.pdf}
+%\vspace{-10pt}
+  \caption{Conditions recognized by $\mathcal{A}_{\textit{l-lar}}^2$}
+  \label{fig:his for APQ1Lar-2}
+\end{figure}
+
+For the case when $e_r \vert_{b} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^3$, as shown in \figurename~\ref{fig:automata APQ1Lar-3}.
+Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$.
+The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^3$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=1 \textwidth]{figures/PIC_AUTO_PQ1Lar-rppr.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{l-lar}}^3$}
+  \label{fig:automata APQ1Lar-3}
+\end{figure}
+
+$\mathcal{A}_{\textit{l-lar}}^3$ is used to recognize conditions in \figurename~\ref{fig:his for APQ1Lar-3}.
+Here for simplicity, we only draw operation of $b$, and the return action of the first call of $\textit{ret}(\textit{put},a)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.7 \textwidth]{figures/PIC_HIS_PQ1Lar-rppr.pdf}
+%\vspace{-10pt}
+  \caption{Conditions recognized by $\mathcal{A}_{\textit{l-lar}}^3$}
+  \label{fig:his for APQ1Lar-3}
+\end{figure}
+
+For the case when $e_r \vert_{b} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b) \cdot \textit{ret}(\textit{put})$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^4$, as shown in \figurename~\ref{fig:automata APQ1Lar-4}.
+Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$.
+The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^4$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.9 \textwidth]{figures/PIC_AUTO_PQ1Lar-rprp.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{l-lar}}^4$}
+  \label{fig:automata APQ1Lar-4}
+\end{figure}
+
+$\mathcal{A}_{\textit{l-lar}}^4$ is used to recognize conditions in \figurename~\ref{fig:his for APQ1Lar-4}.
+Here for simplicity, we only draw operation of $b$, and the return action of the first call of $\textit{ret}(\textit{put},a)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.7 \textwidth]{figures/PIC_HIS_PQ1Lar-rprp.pdf}
+%\vspace{-10pt}
+  \caption{Conditions recognized by $\mathcal{A}_{\textit{l-lar}}^4$}
+  \label{fig:his for APQ1Lar-4}
+\end{figure}
+
+Let $\textit{Auts}_{\textit{1-lar}} = \{ \mathcal{A}_{\textit{l-lar}}^1, \mathcal{A}_{\textit{l-lar}}^2, \mathcal{A}_{\textit{l-lar}}^3, \mathcal{A}_{\textit{l-lar}}^4 \}$.
+The following lemma states that $\textit{PQ}_1^{>}$ is co-regular.
+
+\PQOneLarisCoRegular*
+
+\begin {proof}
+
+We need to prove that, given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-lar}} \cap \mathcal{I} = \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), last(e')=\textit{PQ}_1^{>} \wedge e$ does not linearizable w.r.t.
+$\textit{MS}(\textit{PQ}_1^{>})$.
+
+By Lemma \ref{lemma:Lin Equals Constraint for PQ1Lar}, we need to prove the following fact:
+
+\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-lar}} \cap \mathcal{I} = \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), last(e')=\textit{PQ}_1^{>} \wedge$ there is a cycle going through $x$ in $G$, where $G$ is the left-right constraint of $e'$, and $x$ is the item with maximal priority in $e'$.
+
+\noindent The $\textit{only if}$ direction: Let us consider the case of $\mathcal{A}_{\textit{l-lar}}^1$.
+Assume that $e_1 \in \mathcal{I}$ is accepted by $\mathcal{A}_{\textit{l-lar}}^1$.
+By data-independence, there exists data-differentiated execution $e \in \mathcal{I}$ and renaming function $r_1$, such that $e_1 = r_1(e)$.
+Assume that $r_1$ maps $d$ into $b$ and maps $f_1,\ldots,f_m$ into $a$.
+Let $e'$ be obtained from $e$ by projection into $\{ d, f_1,\ldots,f_m \}$.
+It is easy to see that $\textit{last}(\textit{PQ}_1^{>})$, and there is a cycle going through $d$ in $G$, where $G$ is the left-right constraint of $e'$.
+The case of $\mathcal{A}_{\textit{l-lar}}^2$, $\mathcal{A}_{\textit{l-lar}}^3$ and $\mathcal{A}_{\textit{l-lar}}^4$ is similar.
+
+\noindent The $\textit{if}$ direction: Given such $e$, $e'$ and $x$.
+Let renaming function $r$ maps $x$ into $b$, maps items cover $x$ into $a$, and maps other items into $d$.
+By data-independence, $r(e) \in \mathcal{I}$.
+Then depending on the case of $r(e) \vert_{b}$, we can see that $r(e)$ is accepted by $\mathcal{A}_{\textit{l-lar}}^1$, $\mathcal{A}_{\textit{l-lar}}^2$, $\mathcal{A}_{\textit{l-lar}}^3$ or $\mathcal{A}_{\textit{l-lar}}^4$.
+\qed
+\end {proof}
+
 \section{Co-Regular of Priority Queues}
 \label{sec:co-regular of priority queues}
-
-\subsection{Co-Regular of $R_{\textit{pr1}}$}
-\label{subsec:co-regular of Rpr1}
-
-Let us introduce the notion of gap for matched $\textit{put}$ and $\textit{rm}$ operations.
-
-\begin{definition}[gap for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:gap for matched put and rm operations}
-
-Let $h$ be a data-differentiated history and $o_1 = \textit{put}(x,\textit{pri}_x)$, $o_2 = \textit{rm}(x)$ be a pair of matched $\textit{put}$ and $\textit{rm}$ operation of $h$.
-We say that $h$ has a gap on operations $o_1$ and $o_2$, if there is a partition of the operations of $h$ into $L$, $M$ and $R$ satisfying:
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $L \cup M$ has matched $\textit{put}$ and $\textit{rm}$ operations.
-
-\item[-] no operation of $R$ happens before an operation of $L$, $o_1$, $M$ or $o_2$.
-
-\item[-] no operation of $o_2$ happens before an operation of $L$, $o_1$ or $M$.
-
-\item[-] no operation of $M$ happens before an operation of $L$ or $o_1$.
-
-\item[-] no operation of $o_1$ happens before an operation of $L$.
-
-\item[-] no operation in $L \cup M \cup R$ has priority $\textit{pri}_x$.
-\end{itemize}
-\end{definition}
-
-Given a history $h$ and set $O$ of some operations in $h$.
-It is easy to see that there exists a sequence $l$ of operations of $O$, such that $l$ does not contradicts the happen before relation of $h$.
-Therefore, from now on, we can use this fact to directly generate such sequence from such $O$.
-
-The following lemma reduce linearizable with respect to $\textit{MR}_{\textit{pr1}}$ into gap of matched $\textit{put}$ and $\textit{rm}$ operations.
-
-\begin{restatable}{lemma}{GapEqualsLinforRpr1}
-\label{lemma:Gap Equals Lin for Rpr1}
-A data-differentiated history $h$ has a projection $h'$, such that $\textit{last}(h') = R_{\textit{pr1}}$ and $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$, if and only if there exists a projection $h''$ of $h$, such that $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is in $h''$, $\textit{pri}_x$ is the maximal priority of $h''$, and one the the following case holds:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] case $1$: $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$,
-\item[-] case $2$: there is no gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-\end{itemize}
-\end{restatable}
-
-\begin {proof}
-
-Let us prove the $\textit{only if}$ direction by contradiction.
-Assume that there exists such $h'$, while case $1$ and case $2$ do not hold for any projection of $h$.
-Since case $1$ does not hold for any projection of $h$, we know that $\forall y \in D$, $\textit{rm}(y)$ does not happens before $\textit{put}(y,\_)$ in $h$.
-Since case $2$ does not hold for any projection of $h$ and $\textit{last}(h') = R_{\textit{pr1}}$, we can see that $h'$ has gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$, while $\textit{pri}_x$ is the maximal priority in $h'$.
-Since $\textit{last}(h') = R_{\textit{pr1}}$, we can see that $\textit{pri}_x$ is larger than priority of all other $\textit{put}$ operations of $h$.
-
-By the properties of the gap, there is partitions $L$, $M$ and $R$.
-Let $u$, $v$ and $w$ contain the operations of $L$, $M$ and $R$, respectively.
-It is easy to see that $h'$ is linearizable with respect to $l''=u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$ and $l'' \in \textit{MR}_{\textit{pr1}}$.
-This contradicts the fact that $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$.
-
-Let us begin to prove the $\textit{if}$ direction.
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If case $1$ holds for $h''$, then $h \vert_{ \{ x \} }$ is the $h'$ as required.
-
-\item[-] If case $2$ holds, assume by contradiction that, for every projection $h'$ of $h$, if $\textit{last}(h') = R_{\textit{pr1}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr1}}$.
-Since $h''$ is also a projection of $h$, we know that there exist sequences $u$, $v$ and $w$, such that $h'' \sqsubseteq u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, $u \cdot v$ contains matched $\textit{put}$ and $\textit{rm}$ operations, and $\textit{pri}_x$ is larger than priority of $u \cdot v \cdot w$.
-Let $L$, $M$ and $R$ be the operations of $u$, $v$ and $w$, respectively.
-Then, the partitions $L$, $M$ and $R$ form a gap on operation $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-This contradicts the fact that there is no gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-\begin{definition}[left-right constraint for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:left-right constraint for matched put and rm operations}
-Given a data-differentiated history $h$, and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-The left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is the graph $G$ where:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] the nodes are the items of $h$, to which we add a node,
-
-\item[-] there is an edge from item $d_1$ to $x$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$ and $\textit{pri}_1 \neq \textit{pri}_x$,
-
-\item[-] there is an edge from $x$ to item $d_1$, if $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$, and the priority of $d_1$ is not $\textit{pri}_x$,
-
-\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{rm}(d_2,\textit{pri}_2)$, $\textit{pri}_1 \neq \textit{pri}_x$, and $\textit{pri}_2 \neq \textit{pri}_x$.
-\end{itemize}
-\end{definition}
-
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-Let $\textit{LMSet}_1(h,x) = \{ o \vert$ $o$ is not an operation of $x$, the priority of $o$ does not equal the priority of $x$, and either it happens before $\textit{put}(x)$ or $\textit{rm}(x)$ in $h$, or there is an operation $o'$ with the same item of $o$, such that $o'$ happens before $\textit{put}(x)$ or $\textit{rm}(x)$ in $h$ $\}$.
-For each $i \geq 1$, let $\textit{LMSet}_{\textit{i+1}}(h,x) = \{ o \vert$ $o$ is not an operation of $x$ or items in $\textit{LMSet}_k(h,x)$ for each $k \leq i$, the priority of $o$ does not equal the priority of $x$, and either it happens before some operation $o' \in \textit{LMSet}_i(h,x)$ in $h$, or there is an operation $o''$ with the same item of $o$ and $o''$ happens before some operation $o' \in \textit{LMSet}_i(h,x)$ in $h$ $\}$.
-%We can see that for each $i$, $\textit{LMSet}_i(h,x)$ contains matched $\textit{put}$ and $\textit{rm}$ operations, and
-We can see that $\textit{LMSet}_i(h,x) \cap \textit{LMSet}_j(h,x) = \emptyset$ for any $i \neq j$.
-Let $\textit{LMSet}(h,x) = \textit{LMSet}_1(h,x) \cup \textit{LMSet}_2(h,x) \cup \ldots$.
-%{\color {red} We say that a history $h$ of priority queue is matched before $x$, if (1) $h$ contains no $\textit{rmEmpty}$ operations, (2) $ h \sqsubseteq l_1 \cdot \textit{put}(x,\_) \cdot l_2 \cdot \textit{rm}(x) \cdot l_3$ for some $l_1,l_2,l_3$, and $l_1 \cdot l_2$ contains matched $\textit{put}$ and $\textit{rm}$ operations.}
-
-The following is an example that shows that, given a data-differentiated history $h$, the set $\textit{LMSet}_i(h,x)$ can be non-empty for each $1 \leq i \leq 3$.
-In \figurename~\ref{fig:his nobound of LMSet} , $\textit{LMSet}_1(h,x) = \{ \textit{put}(a,1),\textit{rm}(a) \}$, $\textit{LMSet}_2(h,x) = \{ \textit{put}(b,1),\textit{rm}(b) \}$ and $\textit{LMSet}_3(h,x) = \{ \textit{put}(a,c),\textit{rm}(c) \}$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{PIC_HIS_NOBOUNDOF_LMSET.pdf}
-%\vspace{-10pt}
-  \caption{An example for non-emptiness of $\textit{LMSet}_1(h,x)$, $\textit{LMSet}_2(h,x)$ and $\textit{LMSet}_3(h,x)$}
-  \label{fig:his nobound of LMSet}
-\end{figure}
-
-\begin{restatable}{lemma}{RmxDoesNotHappenBeforeLMSetForRpr1}
-\label{lemma:Rmx does not happen before LMSet for Rpr1}
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-%, such that $\textit{pri}_x$ is the maximal priority of $h$, and $\textit{last}(h) = R_{\textit{pr1}}$.
-Assume that $\textit{rm}(y)$ does not happen before $\textit{put}(y,\_)$ for each item $y$ in $h$.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-Assume that $G$ has no cycle going through $x$.
-Then, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}(h,x)$.
-\end{restatable}
-
-\begin {proof}
-
-We prove this lemma by induction, and prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x)$, in $\textit{LMSet}_2(h,x)$, $\ldots$.
-Note that, according to the definition of $\textit{LMSet}(h,x)$, $\textit{pri}_x$ is different to priority of all items in $\textit{LMSet}(h,x)$.
-
-\noindent (1) Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x)$ by contradiction.
-Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{LMSet}_1(h,x)$ is an operation of item $d$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d$ does not equals $\textit{pri}_x$).
-
-We use a triple $(t_1,t_2,t_3)$ to represent related information.
-$t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$.
-$t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation.
-$t_2$ and $t_3$ is used for the reason of $o \in \textit{LMSet}_1(h,x)$: $o \in \textit{LMSet}_1(h,x)$, since an operation (of kind $t_2$) of $d$ happens before an operation (of kind $t_3$) of $x$.
-Let us consider all the possible cases of $(t_1,t_2,t_3)$:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.
-
-\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{rm}(x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{rm}(x)$.
-
-\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x) )$.
-By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(x,\textit{pri}_x) ) \vee ( \textit{rm}(d) <_{hb} \textit{put}(d,\_) )$, which is impossible.
-
-\item[-] $(\textit{put},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{rm}(x) )$.
-We can see that $\textit{rm}(d) <_{hb} \textit{rm}(x) <_{hb} \textit{put}(d,\_)$, which contradicts that $\textit{rm}(d)$ does not happen before $\textit{put}(d,\_)$.
-
-\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x) )$.
-We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.
-
-\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{rm}(x) )$.
-We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.
-
-\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.
-
-\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{rm}(x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{rm}(x)$.
-\end{itemize}
-
-This completes the proof for $\textit{LMSet}_1(h,x)$.
-
-\noindent (2) Assume we already prove that for some $j \geq 1$, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_{\textit{j+1}}(h,x)$ by contradiction.
-Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$ is an operation of item $d_{\textit{j+1}}$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d_{\textit{j+1}}$ does not equals $\textit{pri}_x$).
-We use a triple $(t_1,t_2,t_3)$ to represent related information.
-$t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$.
-$t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation.
-$t_2$ and $t_3$ is used for the reason of $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$: $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$, since an operation (of kind $t_2$) of $d_{\textit{j+1}}$ happens before an operation (of kind $t_3$) of $d_j$, where $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{LMSet}_j(h,x)$.
-Let us consider all the possible cases of $(t_1,t_2,t_3)$:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_)$.
-We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-
-\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j,\_)$.
-We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j,\_) ) \wedge ( \textit{rm}(d_j) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-
-\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$.
-By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \vee ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_{\textit{j+1}},\_) )$, which is impossible.
-
-\item[-] $(\textit{put},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$.
-By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \vee ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_{\textit{j+1}},\_) )$, which is impossible.
-
-\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) )$.
-Let us consider the reason of $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{LMSet}_j(h,x)$:
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $( j > 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(_{\textit{j-1}},\_), \textit{rm}(_{\textit{j-1}}) \in \textit{LMSet}_{\textit{j-1}}(h,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, we can see that $\textit{put}(d_{\textit{j+1}},\_) <_{hb} o''$, and then operations of $d_{\textit{j+1}}$ is in $\textit{LMSet}_j(h,x)$, contradicts that operations of $d_{\textit{j+1}}$ is in $\textit{LMSet}_{\textit{j+1}}(h,x)$.
-
-    \item[-] If $( j = 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of $x$: Similar to above case.
-
-    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{LMSet}_{\textit{j-1}}(h,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, we can see that $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} o'' ) \vee ( \textit{rm}(d_j) <_{hb} \textit{put}(d_j,\_) )$, which is impossible.
-
-    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of $x$: Similar to above case.
-    \end{itemize}
-
-\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Let $T_{\textit{ind}}$ be the set of sentences $\{ \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}), \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j),\ldots, \textit{put}(d_{\textit{ind+1}},\_) <_{hb} \textit{rm}(d_{\textit{ind}}) \}$.
-Here each $d_i$ is a item of some operation in $\textit{LMSet}_i(h,x)$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d_i$ does not equals $\textit{pri}_x$ for each $i$).
-Let us prove that from $T_j$ we can obtain contradiction by induction:
-
-    {\bf Base case $1$}: From $T_1$ we can obtain contradiction.
-
-    Let us prove base case $1$:
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{put}(d_1,\_)$ happens $o$, and $o$ is an operation of $x$.
-Then there is a cycle $x \rightarrow d_{\textit{j+1}} \rightarrow \ldots \rightarrow d_1 \rightarrow x$ in $G$, contradicts that $G$ has no cycle going through $x$.
-
-    \item[-] If $\textit{rm}(d_1)$ happens before $o$, and $o$ is an operation of $x$.
-Then since $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1) <_{hb} o$, we can see that $\textit{put}(d_2,\_) <_{hb} o$, and then $\textit{put}(d_2,\_) \in \textit{LMSet}_1(h,x)$, contradicts that $\textit{put}(d_2,\_) \in \textit{LMSet}_2(h,x)$.
-    \end{itemize}
-
-    {\bf Base case $2$}: From $T_2$ we can obtain contradiction.
-
-    Let us prove base case $2$: If $\textit{rm}(d_2) <_{hb} o$, and $o$ is an operation of $d_1$, then since $( \textit{put}(d_3,\_) <_{hb} \textit{rm}(d_2) ) \wedge ( \textit{rm}(d_2) <_{hb} o )$, we know that $\textit{put}(d_3,\_) <_{hb} o$.
-This implies that $\textit{put}(d_3,\_) \in \textit{LMSet}_2(h,x)$, contradicts that $\textit{rm}(d_3,\_) \in \textit{LMSet}_3(h,x)$.
-Therefore, it is only possible that $\textit{put}(d_2,\_)$ happens before an operation of $d_1$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{put}(d_1,\_)$ and $\textit{put}(d_1,\_)$ happens before operations of $x$, then we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, which is impossible.
-
-    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{put}(d_1,\_)$ and $\textit{rm}(d_1)$ happens before operations of $x$, then by interval order, we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, or $\textit{rm}(d_1) <_{hb} \textit{put}(d_1,\_)$, which is impossible.
-
-    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{put}(d_1,\_)$ happens before operations of $x$, then $x \rightarrow d_{\textit{j+1}} \rightarrow \ldots \rightarrow d_1 \rightarrow x$ in $G$, contradicts that $G$ has no cycle going through $x$.
-
-    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1)$ happens before operations of $x$, then we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, which is impossible.
-    \end{itemize}
-
-    {\bf induction step}: Given $\textit{ind} \geq 3$, if from $T_{\textit{ind-1}}$ we can obtain contradiction, then from $T_{\textit{ind}}$ we can also contain contradiction.
-
-    Prove of the induction step: Similarly as base case $2$, we can prove that it is only possible that $\textit{put}(d_{\textit{ind}},\_)$ happens before operations of $d_{\textit{ind-1}}$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$ and $\textit{put}(d_{\textit{ind-1}},\_)$ happens before operations of $d_{\textit{ind-2}}$, then we know that $\textit{put}(d_{\textit{ind}})$ happens before operation of $d_{\textit{ind-2}}$, which is impossible.
-
-    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$ and $\textit{rm}(d_{\textit{ind-1}})$ happens before operations of $d_{\textit{ind-2}}$, then by interval order, we know that $\textit{put}(d_{\textit{ind}},\_)$ happens before operation of $d_{\textit{ind-2}}$, or $\textit{rm}(d_{\textit{ind-1}}) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$, which is impossible.
-
-    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{rm}(d_{\textit{ind-1}})$, then we obtain $T_{\textit{ind-1}}$, which already contain contradiction.
-    \end{itemize}
-
-    By base case $1$, base case $2$ and the induction step, it is easy to see that $T_j$ contains contradiction.
-
-\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$.
-We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-
-\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$.
-We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \wedge ( \textit{rm}(d_j) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.
-\end{itemize}
-
-This completes the proof for $\textit{LMSet}_{\textit{j+1}}(h,x)$.
-Therefore, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}(h,x) = \textit{LMSet}_1(h,x) \cup \textit{LMSet}_2(h,x) \cup \ldots$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{LMSetHasMatchedPutandRmOperations}
-\label{lemma:LMSet has matched put and rm operations}
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-Assume that $\textit{rm}(y)$ does not happen before $\textit{put}(y,\_)$ for each item $y$ in $h$.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
-Assume that $G$ has no cycle going through $x$.
-Then, $\textit{LMSet}(h,x)$ contains only matched $\textit{put}$ and $\textit{rm}$ operations.
-\end{restatable}
-\begin {proof}
-
-We prove this lemma by contradiction.
-Assume that there exists a value, such that $\textit{LMSet}(h,x)$ contains only its $\textit{put}$ operation and does not contain its $\textit{rm}$ operation.
-Then we can see that there exists $d_1,\ldots,d_j$.
-Intuitively, $d_1,\ldots,d_j$ are items in $\textit{LMSet}_1(h,x), \ldots, \textit{LMSet}_i(h,x)$, respectively.
-$\textit{LMSet}(h,x)$ contains $\textit{put}(d_j,\_)$ and does not contain $\textit{rm}(d_j)$.
-And each $d_i$ is the reason of $d_{\textit{i+1}} \in \textit{LMSet}_{\textit{i+1}}(h,x)$.
-Formally, we require that
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] For each $1 \leq i \leq j$, operations of $d_i$ belongs to $\textit{LMSet}_i(h,x)$. %For each $1 \leq i \leq j$, the priority of $d_i$ is not $\textit{pri}_x$.
-
-\item[-] For each $i \neq j$, $\textit{put}(d_i,\_),\textit{rm}(d_i) \in \textit{LMSet}_i(h,x)$. $\textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x)$, and $h$ does not contain $\textit{rm}(d_j)$.
-
-\item[-] An operation of $d_1$ happens before an operation of $x$. For each $1 < i \leq j$, an operation of $d_j$ happens an operation of $d_{\textit{i-1}}$.
-
-\item[-] For each $k$ and $\textit{ind}$, if $k > \textit{ind+1}$, then no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.
-\end{itemize}
-
-According to the definition of $\textit{LMSet}(h,x)$, it is easy to see that such $d_1,\ldots,d_j$ exists. Let us prove the following fact:
-
-\noindent {\bf $\textit{fact}_1$}: Given $1 \leq i < j$, it can not be the case that $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap.
-
-Proof of $\textit{fact}_1$: We prove $\textit{fact}_1$ by contradiction.
-Assume that for some $i \neq j$, $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap.
-Since $\textit{put}(d_i,\_), \textit{rm}(d_i) \in \textit{LMSet}_i(h,x)$, we know that an operation $o_i$ of $d_i$ happens before operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$.
-Moreover, since $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap, it is not hard to see that the call action of $\textit{put}(d_i,\_)$ and the call action of $\textit{rm}(d_i)$ is before the call action of $o_{\textit{i-1}}$.
-Since $\textit{put}(d_{\textit{i+1}},\_), \textit{rm}(d_{\textit{i+1}}) \in \textit{LMSet}_{\textit{i+1}}(h,x)$, we know that an operation $o'_{\textit{i+1}}$ of $d_{\textit{i+1}}$ happens before operation $o'_i$ of $d_i$.
-Then, it is not hard to see that $o'_{\textit{i+1}}$ also happens before $o_{\textit{i-1}}$, which contradicts that For each $k > \textit{ind+1}$, no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.
-
-We already know that an operation of $d_1$ happens before an operation of $x$.
-By $\textit{fact}_1$, we can ensure that $\textit{put}(d_1,\_)$ happens before an operation of $x$, and then $d_1 \rightarrow x$ in $G$.
-For each $1 < i < j$, we know that an operation $o_i$ of $d_i$ happens before an operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$.
-By $\textit{fact}_1$, we can ensure that $o_i=\textit{put}(d_i,\_)$ and $o_{\textit{i-1}}=\textit{rm}(d_{\textit{i-1}})$, and then $d_i \rightarrow d_{\textit{i-1}}$ in $G$.
-Since $h$ contains $\textit{put}(d_j)$ and does not contain $\textit{rm}(d_j)$, we know that $x \rightarrow d_j$ in $G$.
-Then $G$ has a cycle going through $x$, contradicts that $G$ has no cycle going through $x$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{GapEqualsConstraintforRpr1}
-\label{lemma:Gap Equals Constraint for Rpr1}
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-%, where $\textit{pri}_x$ has the maximal priority of $h$, and $\textit{last}(h) = R_{\textit{pr1}}$.
-Assume that for each item in $h$, its $\textit{rm}$ operation does not happen before its $\textit{put}$ operation.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x)$ and $\textit{rm}(x)$.
-There is a gap on $\textit{put}(x)$ and $\textit{rm}(x)$, if and only if $G$ has no cycle going through $x$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{only if}$ direction, assume that there is a gap on $\textit{put}(x)$ and $\textit{rm}(x)$, and let it be $L$, $M$ and $R$.
-Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ in $G$.
-It is obvious that the priority of each $d_i$ is not $\textit{pri}_x$.
-Then our proof proceeds as follows:
-
-According to the definition of left-right constraint, there are two possibilities.
-The first possibility is that, $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$.
-It is obvious that $\textit{rm}(d_1) \in R$, and then since $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_1),\textit{rm}(d_1) \in R$.
-Then,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Since $d_1 \rightarrow d_2$, by definition of $G$, we know that $\textit{put}(d_1)$ happens before $\textit{rm}(d_2)$.
-Since $\textit{put}(d_1) \in R$ and $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_2),\textit{rm}(d_2) \in R$.
-Similarly, for each $1 \leq i \leq m$, we know that $\textit{put}(d_i),\textit{rm}(d_i) \in R$.
-
-\item[-] Since $d_m \rightarrow x$,
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] if $\textit{put}(d_m)$ happens before $\textit{put}(x)$, then we can see that $\textit{put}(d_m) \in L$, which contradicts that $\textit{put}(d_m) \in R$.
-
-    \item[-] if $\textit{put}(d_m)$ happens before $\textit{rm}(x)$, then we can see that $\textit{put}(d_m) \in L \cup M$, which contradicts that $\textit{put}(d_m) \in R$.
-    \end{itemize}
-\end{itemize}
-
-The second possibility is that, $h$ contains one $\textit{put}(d_1,\_)$ operation and no $\textit{rm}(d_1)$ operation.
-Note that for each $j > 1$, $h$ contains $\textit{put}(d_j,\_)$ and $\textit{rm}(d_j)$.
-Since $d_m \rightarrow x$, is is obvious that $\textit{put}(d_m) \in L \cup M$.
-Since $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_m),\textit{rm}(d_m) \in R$.
-Then,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Since $d_{\textit{m-1}} \rightarrow d_m$, by definition of $G$, we know that $\textit{put}(d_{\textit{m-1}})$ happens before $\textit{rm}(d_m)$.
-Since $\textit{rm}(d_m) \in L \cup M$ and $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations, we know that $\textit{put}(d_{\textit{m-1}}),\textit{rm}(d_{\textit{m-1}}) \in L \cup M$.
-Similarly, for each $1 < i \leq m$, we know that $\textit{put}(d_i),\textit{rm}(d_i) \in L \cup M$, and $\textit{put}(d_1)\in L \cup M$.
-
-\item[-] Since $d_1 \rightarrow d_2$ in $G$, $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2)$.
-Since $\textit{put}(d_2),\textit{rm}(d_2) \in L \cup M$, we can see that $\textit{put}(d_1) \in L \cup M$.
-However, there is one $\textit{put}(d_1,\_)$ operation and no $\textit{rm}(d_1)$ operation in $h$, contradicts that $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
-\end{itemize}
-
-This completes the proof of the $\textit{only if}$ direction.
-
-To prove the $\textit{if}$ direction, assume that $G$ has no cycle going through $x$.
-Let $O_L$ be the set of operations that happen before $\textit{put}(x)$ in $h$.
-It is easy to see that $O_L \subseteq \textit{LMSet}(h,x)$.
-Let $O_M = \textit{LMSet}(h,x) \setminus O_L$.
-Let $O_h$ be the set of operations of $h$, and let $O_R = O_h \setminus \textit{LMSet}(h,x)$.
-Let $L$, $M$ and $R$ be $O_L$, $O_M$ and $O_R$, respectively.
-
-By Lemma \ref{lemma:LMSet has matched put and rm operations}, we can see that $L \cup M$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
-It remains to prove that $L$, $M$ and $R$ is a gap of $\textit{put}(x)$ and $\textit{rm}(x)$.
-We prove this by showing that all the following cases are impossible:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Case $1$: Some operation $o_r \in R$ happens before $\textit{rm}(x)$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $2$: Some operation $o_r \in R$ happens before some operation $o_{lm} \in L \cup M$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $3$: Some operation $o_r \in R$ happens before $\textit{put}(x)$. Then we know that $o_r \in \textit{LMSet}(h,x)$, which contradicts that $o_r \in R$.
-
-\item[-] Case $4$: $\textit{rm}(x)$ happens before some $o_{lm} \in \textit{LMSet}(h,x) = L \cup M$. By Lemma \ref{lemma:Rmx does not happen before LMSet for Rpr1} we know that this is impossible.
-
-\item[-] Case $5$: $\textit{rm}(x)$ happens before $\textit{put}(x)$. This contradicts that for each item in $h$, its $\textit{rm}$ operation does not happen before its $\textit{put}$ operation.
-
-\item[-] Case $6$: Some operation $o_m \in M$ happens before $\textit{put}(x)$. Then we know that $o_m \in L$, which contradicts that $o_m \in M$.
-
-\item[-] Case $7$: Some operation $o_m \in M$ happens before some operation $o_l \in L$. Then we know that $o_m \in L$, which contradicts that $o_m \in M$.
-
-\item[-] Case $8$: $\textit{put}(x)$ happens before some operation $o_l \in L$. This is impossible.
-\end{itemize}
-
-This completes the proof of the $\textit{if}$ direction.
-
-\qed
-\end {proof}
-
-A automaton $\mathcal{A}_{\textit{Rpr1-1}}$ is given in \figurename~\ref{fig:automata for Rpro-1}, where $c=\textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm,a),$ $\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$.
-$\mathcal{A}_{\textit{Rpr1-1}}$ is used to detect the existence of a item whose $\textit{rm}$ happens before its $\textit{put}$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-1.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-1}}$}
-  \label{fig:automata for Rpro-1}
-\end{figure}
-
-Let us begin to represent several automata that is used for capture the case that, in a sub-history $h'$ of a history $h$, $\textit{last}(h')=R_{\textit{pr1}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$, and the reason is that there is no gap on $\textit{put}(y,\textit{pri}_y)$ and $\textit{rm}(y)$ in $h'$, where $y$ is the item that has maximal priority in $h'$.
-Such case is dealt with by four automata, since there are four possibilities of $h' \vert_{y}$.
-
-The first automaton is $\mathcal{A}_{\textit{Rpr1-pprr}}$, which is given in \figurename~\ref{fig:automata for Rpro-pprr}.
-In \figurename~\ref{fig:automata for Rpro-pprr}, $c_1= \textit{cal}(\textit{put},c,$ $\textit{anyPri}),\textit{ret}(\textit{put},c), \textit{cal}(\textit{rm},c), \textit{ret}(\textit{rm},c)$, $c_2 = c_1 + \textit{cal}(\textit{put},a,\textit{les}_p)$, and $c_3 = c_2 + \textit{ret}(\textit{rm},a)$.
-Here $p$ is a variable for priority, and $\textit{les}_p$ is a predicate that accepts all priorities that are less than the value of $p$.
-Here $\textit{pprr}$ represents that, $h' \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$ (we consider $b$ as $y$ and $p$ as $\textit{pri}_y$).
-In \figurename~\ref{fig:automata for Rpro-pprr}, the three paths that (1) from $q_{\textit{init}}$ to $q_2$ to $q_3$, (2) from $q_{\textit{init}}$ to $q_2$ to $q_{11}$, and (3) from $q_{\textit{init}}$ to $q_{10}$ to $q_{11}$ represents three possible positions of $\textit{ret}(\textit{put},a)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-pprr.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-pprr}}$}
-  \label{fig:automata for Rpro-pprr}
-\end{figure}
-
-The second automaton is $\mathcal{A}_{\textit{Rpr1-prpr}}$, which is given in \figurename~\ref{fig:automata for Rpro-prpr}.
-In $\mathcal{A}_{\textit{Rpr1-prpr}}$, $c_1, c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$.
-Here $\textit{pprr}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{put},b,p) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-prpr.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-prpr}}$}
-  \label{fig:automata for Rpro-prpr}
-\end{figure}
-
-The third automaton is $\mathcal{A}_{\textit{Rpr1-rppr}}$, which is given in \figurename~\ref{fig:automata for Rpro-rppr}.
-In $\mathcal{A}_{\textit{Rpr1-rppr}}$, $c_1, c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$.
-Here $\textit{rppr}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-rppr.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-rppr}}$}
-  \label{fig:automata for Rpro-rppr}
-\end{figure}
-
-The forth automaton is $\mathcal{A}_{\textit{Rpr1-rprp}}$, which is given in \figurename~\ref{fig:automata for Rpro-rprp}.
-In $\mathcal{A}_{\textit{Rpr1-rprp}}$, $c_2, c_3$ is same as that in $\mathcal{A}_{\textit{Rpr1-pprr}}$, and $c_4=c_1 + \textit{ret}(put,b)$.
-Here $\textit{rprp}$ represents that, $h' \vert_{2} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b) \cdot \textit{ret}(\textit{put})$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_UNMATCHED_Rpr1-rprp.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr1-rprp}}$}
-  \label{fig:automata for Rpro-rprp}
-\end{figure}
-
-Given a data-differentiated history $h$, we say that a item $x$ with priority $\textit{pri}_x$ in $h$ is covered by items $d_1,\ldots,d_m$ in $h$, if the priorities of $d_1,\ldots,d_m$ is not $\textit{pri}_x$, and
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{put}(d_m,\_)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$,
-
-\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,
-
-\item[-] $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$, or $\textit{rm}(d_1)$ does not exists in $h$
-\end{itemize}
-
-According to the definition of left-right constraint, in a data-differentiated history $h$, there is a cycle going through $x$, if and only if there exists items $d_1,\ldots,d_m$, such that $x$ is covered by $d_1,\ldots,d_m$.
-
-\begin{restatable}{lemma}{Rpr1IsCoRegular}
-\label{lemma:Rpr1 is co-regular}
-$R_{\textit{pr1}}$ is co-regular.
-\end{restatable}
-
-\begin {proof}
-
-Let $\textit{AutS-R}_{\textit{pr1}} = \{ \mathcal{A}_{\textit{Rpr1-1}} , \mathcal{A}_{\textit{Rpr1-pprr}}, \mathcal{A}_{\textit{Rpr1-prpr}}, \mathcal{A}_{\textit{Rpr1-rppr}}, \mathcal{A}_{\textit{Rpr1-rprp}} \}$.
-We need to prove that:
-
-\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr1}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr1}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr1}}$.
-
-By Lemma \ref{lemma:Gap Equals Lin for Rpr1} and Lemma \ref{lemma:Gap Equals Constraint for Rpr1}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:
-
-\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr1}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h''$ is a projection of $h$, such that $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is in $h''$, $\textit{pri}_x$ is the maximal priority of $h''$, and one the the following case holds:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] case $1$: $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$,
-\item[-] case $2$: $x$ is covered by some items $d_1,\ldots,d_m$ in $h''$.
-\end{itemize}
-
-\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\textit{AutS-R}_{\textit{pr1}}$.
-By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$.
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-1}}$: Let $d_1$ a item in $h_2$ such that $r(d_1)=b$ and let $h'' = h_2 \vert_{ \{ d_1 \} }$.
-$d_1$ is the item with maximal priority in $h''$.
-It is obvious that $\textit{last}(h'') = R_{\textit{pr1}}$ and $\textit{rm}(d_1)$ happens before $\textit{put}(d_1,\_)$ in $h''$.
-
-\item[-] If $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-pprr}}$, $\mathcal{A}_{\textit{Rpr1-prpr}}$, $\mathcal{A}_{\textit{Rpr1-rppr}}$, or $\mathcal{A}_{\textit{Rpr1-rprp}}$: Let $d_0$ be a item in $h_2$ such that $r(d_0)=b$, let $d_1,\ldots,d_m$ be the items in $h_2$ such that $r(d_i)=a$ for each $1 \leq i \leq m$.
-Let $h'' = h_1 \vert_{ \{ d_0, d_1, \ldots, d_m \} }$.
-It is obvious that the priority of $d_0$ is larger than the priorities of other operations in $h''$, and then $\textit{last}(h'') = R_{\textit{pr1}}$.
-It is easy to see that $d_0$ is covered by $d_1,\ldots,d_m$.
-\end{itemize}
-
-\noindent The $\textit{if}$ direction: Assume that there exists such $h$, $h''$, $x$ and $\textit{pri}_x$.
-Since $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{pri}_x$ is larger than the priorities of other operations in $h''$.
-Then,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If case $1$ holds: Then, $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$ in $h''$.
-Let $h_1$ be obtained from $h$ by renaming $x$ to $b$ and renaming all other items to $a$.
-By data-independence, $h_1 \in \mathcal{I}$, and it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-1}}$.
-\item[-] case $2$: Then, $x$ is covered by some items $d_1,\ldots,d_m$ in $h''$.
-There are four possibilities of $h'' \vert_{ \{ x \} }$: (1) $\textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{rm},x)$, (2) $\textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},x)$, (3) $\textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},x)$ and (4) $\textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{rm},x) \cdot \textit{ret}(\textit{put})$.
-
-    Let $h_1$ be obtained from $h$ by renaming $x$ into $b$, renaming $d_1,\ldots,d_m$ into $a$ and renaming other items into $c$.
-We can prove that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr1-pprr}}$, $\mathcal{A}_{\textit{Rpr1-prpr}}$, $\mathcal{A}_{\textit{Rpr1-rppr}}$ or $\mathcal{A}_{\textit{Rpr1-rprp}}$.
-
-    For example, if $h'' \vert_{ \{ x \} } = \textit{call}(\textit{put},x,\textit{pri}_x) \cdot \textit{ret}(\textit{put}) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{rm},x)$, and $\textit{ret}(\textit{rm},d_m)$ is before $\textit{call}(\textit{put},x,\textit{pri}_x)$ and is after $\textit{ret}(\textit{put},x)$ in $h''$.
-Then in  $\mathcal{A}_{\textit{Rpr1-pprr}}$, $h_1$ is accepted by transitions from $q_{\textit{init}}$ to $q_2$ to $q_{11}$ and finally to $q_7$ (if $\textit{rm}(d_1)$ does not exists in $h''$) or $q_9$ (if $\textit{rm}(d_1)$ exists in $h''$).
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
 

### 2017-02-23 co-regular-PQ.tex
I add subsection of co-regular of PQ1>. I also add explaination for clarity. Now Appendix C.2 is for this subsection.

@@
 For example, $\textit{rm}(a)$ never happens before $\textit{put}(a,\_)$ for each $a$.
+
+\subsection{Co-Regular of $\textit{PQ}_1^{>}$}
+\label{subsec:co-regular of PQ1Lar}
+
+In this subsection, we introduce the idea for proving co-regular of $\textit{PQ}_1^{>}$.
+The proof of this subsection can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of PQ1Lar}.
+The notion of left-right constraint used for priority queue is inspired by left-right constraint of queue \cite{Bouajjani:2015}.
+
+Given a data-differentiated execution $e$ such that $\textit{last}(e) = \textit{PQ}_1^{>}$, although Lemma \ref{lemma:automata for priority queue with single priority} ensures that each single-priority projection of $e$ satisfy the FIFO property, this is still not enough for ensuring that $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{>})$.
+Since it is possible that $e$ does not linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{>})$ because of interaction between actions of multiple priorities.
+
+We give an example of such execution $e$ in \figurename~\ref{fig:introduce gap for PQ1Lar}.
+We call the time interval from $\textit{ret}(\textit{put},x)$ to $\textit{cal}(rm,x)$, or from $\textit{ret}(\textit{put},x)$ when $\textit{cal}(rm,x)$ does not exist, the interval of item $x$.
+In \figurename~\ref{fig:introduce gap for PQ1Lar}, we draw the interval of each item by dashed line.
+The reason of why $e$ does not linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{>})$ is that, each time point from $\textit{cal}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is in interval of some item with less priority than priority of $a$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.6 \textwidth]{figures/PIC-HIS-INTRO-GAP-PQ1L.pdf}
+%\vspace{-10pt}
+  \caption{An execution that does not linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{>})$}
+  \label{fig:introduce gap for PQ1Lar}
+\end{figure}
+
+In this subsection, intuitively we will prove that, as long as we can get rid of the case in \figurename~\ref{fig:introduce gap for PQ1Lar}, we can ensure $\textit{last}(e) = \textit{PQ}_1^{>} \Rightarrow e \sqsubseteq \textit{MS}(\textit{PQ}_1^{>})$.
+
+Let us introduce the notion of left-right constraint, which is a graph, and the existence of cycle though item with maximal priority in it correspond to the existence of the case in \figurename~\ref{fig:introduce gap for PQ1Lar}.
+
+\begin{definition}\label{def:left-right constraint for matched put and rm operations}
+Given a data-differentiated execution $e$ without $\textit{rm}(\textit{empty})$.
+Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$, where $\textit{pri}_x$ is maximal in $e$ and only used by these two method events.
+The left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is the graph $G$ where:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] the nodes are the items of $e$, to which we add a node,
+
+\item[-] there is an edge from item $d_1$ to $x$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$,
+
+\item[-] there is an edge from $x$ to item $d_1$, if $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$,
+
+\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{rm}(d_2,\textit{pri}_2)$.
+\end{itemize}
+\end{definition}
+
+When there is a cycle $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ through item with maximal priority (for example, $x$) in $G$, we say that $x$ is covered by $d_1,\ldots,d_m$.
+To state the effectiveness of left-right constraint, take the execution in \figurename~\ref{fig:introduce gap for PQ1Lar} as an example, we can see that $b$ is covered by $f,e,d$.
+We need to prove that getting rid of cycle though item with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{>})$, as stated by the following lemma:
+
+\begin{restatable}{lemma}{LinEqualsConstraintforPQOneLar}
+\label{lemma:Lin Equals Constraint for PQ1Lar}
+Given a data-differentiated execution $e$ with $\textit{last}(e) = \textit{PQ}_1^{>}$.
+Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority.
+Let $G$ be the graph representing the left-right constraint of $\textit{put}(x)$ and $\textit{rm}(x)$.
+$e \sqsubseteq \textit{MS}(\textit{PQ}_1^{>})$, if and only if $G$ has no cycle going through $x$.
+\end{restatable}
+
+\begin {proof} (Sketch)
+
+The $\textit{only if}$ direction can be easily proved by contradiction and is omitted here.
+To prove the $\textit{if}$ direction, we need to explicitly construct the linearization of $e$, or we can say, we need construct the $u$, $v$ and $w$ in $\textit{PQ}_1^{>}$.
+Let $u$ to be the sequence of all operations that happens before $\textit{put}(x)$.
+The difficulties is how to generate proper $v$.
+
+To generate $v$, we introduce $\textit{UVSet}(e,x)$, which intuitively contains all pairs of method events that should be putted before $\textit{rm}(x)$.
+Let $\textit{UVSet}_1(e,x)= \{ o \vert$ either $o <_{\textit{hb}} \textit{put}(x)$ or $\textit{rm}(x)$, or $\exists o'$ with the same item of $o$, such that $o' <_{\textit{hb}} \textit{put}(x)$ or $\textit{rm}(x)\}$.
+For each $i \geq 1$, let $\textit{LMSet}_{\textit{i+1}}(e,x) = \{ o \vert$ $o \notin \textit{UVSet}_k(e,x)$ for each $k \leq i$, and either $o$ happens before some operation $o' \in \textit{UVSet}_i(e,x)$, or $\exists o''$ with the same item of $o$ and $o''$ happens before some operation $o' \in \textit{UVSet}_i(e,x)\}$.
+Let $\textit{UVSet}(e,x) = \textit{UVSet}_1(e,x) \cup \textit{UVSet}_2(e,x) \cup \ldots$.
+For example, in \figurename~\ref{fig:his nobound of LMSet}, $\textit{UVSet}_1(e,x) = \{ \textit{put}(a,1),\textit{rm}(a) \}$, $\textit{USSet}_2(e,x) = \{ \textit{put}(b,1),\textit{rm}(b) \}$ and $\textit{UVSet}_3(e,x) = \{ \textit{put}(a,c),\textit{rm}(c) \}$.
+Similarly, we can generate execution $e'$ such that, for each $i$, $\textit{UVSet}(e',x) \neq \emptyset$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.5 \textwidth]{figures/PIC_HIS_NOBOUNDOF_LMSET.pdf}
+%\vspace{-10pt}
+  \caption{An example for $\textit{UVSet}(e,x)$}
+  \label{fig:his nobound of LMSet}
+\end{figure}
+
+We let the $v$ to be the sequence of method events that are in $\textit{UVSet}(e,x)$ and not in $u$.
+We let $w$ to be the sequence of remaining method events.
+Then we can prove that such $u$, $v$ and $w$ holds as required.
+\qed
+\end {proof}
+
+According to Lemma \ref{lemma:Lin Equals Constraint for PQ1Lar}, the existence of ``cover item with maximal priority'' is enough for checking violation of linearizability w.r.t $\textit{MS}(\textit{PQ}_1^{>})$.
+Assume $x$ is covered by $d_1,\ldots,d_m$.
+Then we can safely rename $x$ into $b$, rename $d_1,\ldots,d_m$ into $a$ and rename all other item into $c$ by data-independence.
+Such execution can be recognized by witness automata, since between the first $\textit{ret}(\textit{put},a,\_)$ and the last $\textit{cal}(\textit{rm},a)$ (if exists), $\textit{ret}(\textit{put},a,\_)$ and $\textit{cal}(\textit{rm},a)$ occurs in pair and there is no need for count.
+
+There are four possible enumeration of call and return actions of $\textit{put}(b)$ and $\textit{rm}(b)$.
+For each of them, we generate a witness automaton.
+For example, for the case when $e \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^1$, as shown in \figurename~\ref{fig:automata APQ1Lar-1}.
+Here $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{cal}(\textit{put},a,\textit{les}_p)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d), \textit{ret}(\textit{rm},d)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=1 \textwidth]{figures/PIC_AUTO_PQ1Lar-pprr.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{l-lar}}^1$}
+  \label{fig:automata APQ1Lar-1}
+\end{figure}
+
+Given the execution $e$ in \figurename~\ref{fig:introduce gap for PQ1Lar}, let $e'$ be generated from $e$ by renaming $d,e,f$ into $a$.
+Then we can see that $e'$ is accepted by $\mathcal{A}_{\textit{l-lar}}^1$ with a path via $q_{\textit{init}}, q_2, q_3, q_4,q_5,q_7,q_8$.
+
+In Appendix \ref{sec:appendix proof and definition in section co-regular of PQ1Lar}, we construct a set $\textit{Auts}_{\textit{1-lar}}$ of witness automata ($\mathcal{A}_{\textit{l-lar}}^1$ is in it), and use $\textit{Auts}_{\textit{1-lar}}$ to prove that $\textit{PQ}_1^{>}$ is co-regular, as stated by the following lemma.
+
+\begin{restatable}{lemma}{PQOneLarisCoRegular}
+\label{lemma:PQ1Lar is co-regular}
+$\textit{PQ}_1^{>}$ is co-regular.
+\end{restatable}

### 2017-03-04 appendix.tex
I prove the second rule is co-regular. The proof can be found in Appendix C.3

@@
 This is obvious, since for each conditions in the $\textit{Guard}$ part of the rules of priority queue, if a sequence of sequential executions satisfy it, then its sub-sequence also satisfy it.
-For example, if $\textit{rm}(\textit{empty}) \notin l = u \cdot v \cdot w$ and let $D_l$ be the set of items of $l$, then for each subset $D' \subseteq D$, it is obvious that $\textit{rm}(\textit{empty}) \notin l \vert_{ D' }$.
+For example, if $\textit{rm}(\textit{empty}) \notin l = u \cdot v \cdot w$ and let $D_l$ be the set of items of $l$, then for each subset $D' \subseteq D_l$, it is obvious that $\textit{rm}(\textit{empty}) \notin l \vert_{ D' }$.
 Similar cases hold for $\textit{pri} \geq \textit{maxPri}(l)$, $\textit{pri} > \textit{maxUPri}(l)$, $\textit{matched}(l)$, $\textit{putInSeq}(l,l_1,\textit{pri})$ (let $l_1$ be sub-sequence of $l$).
@@
 
-Given a data-differentiated execution $e$ and its history $h$, we can obtain a sequence $h'$ from $h$ by adding $\textit{put}(a,p)$ (resp., $\textit{rm}(a)$) between each pair of $\textit{cal}(\textit{put},a)$ and $\textit{ret}(\textit{put},a)$ (resp., $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$).
+Given a data-differentiated execution $e$ and its history $h$, we can obtain a sequence $h'$ from $h$ by adding $\textit{put}(a,p)$ (resp., $\textit{rm}(a)$, $\textit{rm}(\textit{empty})$) between each pair of $\textit{cal}(\textit{put},a,p)$ and $\textit{ret}(\textit{put},a)$ (resp., $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$, $\textit{cal}(\textit{rm},\textit{empty})$ and $\textit{ret}(\textit{rm},\textit{empty})$).
 The projection of $h'$ into method events is called linearization of $e$ and $h$, and each method event we add in $h'$ can be considered as a linearization point of the corresponding method event.
@@
 Let $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t$, where $l_t$ is the projection of $l'$ into method events of $E_{px}$ that are not in $l'_{\textit{L0}}$.
-Let $E_{l_t}$ be the set of method events in $l_t$.
 Let $l'_{\textit{L2}}$ be the projection of $l'_L$ into method events of $l'_L$ that are not in $l'_{\textit{L1}}$.
@@
 
-To prove the $\textit{if}$ direction. The $\textit{if}$ direction holds trivially if $e = \epsilon$. When $e \neq \epsilon$, we generate concurrent execution $e_1$ from $e$ as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $\textit{last}(e) = \textit{PQ}_3$: $e_1$ is generated from $e$ by erasing call and return of one $\textit{rm}(\textit{empty})$.
-
-\item[-] Else, if $\textit{last}(e) = \textit{PQ}_2^{=}$, and the maximal priority of $e$ is unmatched $\textit{put}$ with items in $S$ and matched $\textit{put}$: $e_1$ is generated from $e$ by erasing call and return of one unmatched $\textit{put}$, which is maximal respect to $<_{\textit{hb}}$ among $\textit{put}$ with items in $S$.
-
-\item[-] Else, if $\textit{last}(e) = \textit{PQ}_2^{>}$, and the maximal priority in $s$ is unmatched $\textit{put}$ with items in $S$: $e_1$ is generated from $e$ by erasing call and return of one unmatched $\textit{put}$, which is maximal respect to $<_{\textit{hb}}$ among $\textit{put}$ with items in $S$.
-
-\item[-] Else, if $\textit{last}(e) = \textit{PQ}_1^{=}$, and the maximal priority in $e$ is of more than one pair of matched $\textit{put}$ and $\textit{rm}$ with items in $S$: $e_1$ is generated from $e$ by erasing call and return of one pair of matched $\textit{put}$ and $\textit{rm}$, where such $\textit{put}$ is maximal respect to $<_{\textit{hb}}$ among $\textit{put}$ with items in $S$, and such $\textit{rm}$ is maximal respect to $<_{\textit{hb}}$ among $\textit{rm}$ with items in $S$.
-
-\item[-] Else, if $\textit{last}(e) = \textit{PQ}_1^{>}$: $e_1$ is generated from $e$ by erasing call and return of one pair of matched $\textit{put}$ and $\textit{rm}$ with maximal priority in $e$.
+To prove the $\textit{if}$ direction.
+The $\textit{if}$ direction holds trivially if $e = \epsilon$.
+When $e \neq \epsilon$, we generate concurrent execution $e_1$ from $e$ as follows: Since $e \in \textit{proj}(e)$ we know that $e \sqsubseteq \textit{MS}(R_e)$, where $R_e = \textit{last}(e)$,
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If $R_e = \textit{PQ}_3$: $e_1$ is generated from $e$ by erasing call and return of one $\textit{rm}(\textit{empty})$.
+
+\item[-] Else, if $R_e = \textit{PQ}_2^{=}, \textit{PQ}_2^{>}, \textit{PQ}_1^{=}, \textit{PQ}_1^{>}$, and the witness is $x$: $e_1$ is generated from $e$ by erasing call and return of method event of $x$.
 \end{itemize}
@@
 We generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}.
-Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri})$, $\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},b),\textit{rm}(\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$.
+Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri})$, $\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},b),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$.
 
@@
 We generate witness automata $\mathcal{A}_{\textit{SinPri}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata for FIFO-2}.
-Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{rm}(\textit{empty})$, $c_2 = c_1 + \textit{cal}(\textit{rm},b) + \textit{ret}(\textit{rm},b)$.
+Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_2 = c_1 + \textit{cal}(\textit{rm},b) + \textit{ret}(\textit{rm},b)$.
 
@@
 We generate witness automata $\mathcal{A}_{\textit{SinPri}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata for FIFO-3}.
-Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{rm}(\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},b)$, $c_4 = c_3 + \textit{cal}(\textit{rm},b)$, $c_5 = c_1 + \textit{ret}(\textit{rm},b)$, $c_6 = c_5 + \textit{cal}(\textit{rm},b)$.
+Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},b)$, $c_4 = c_3 + \textit{cal}(\textit{rm},b)$, $c_5 = c_1 + \textit{ret}(\textit{rm},b)$, $c_6 = c_5 + \textit{cal}(\textit{rm},b)$.
 
@@
 We generate witness automata $\mathcal{A}_{\textit{SinPri}}^4$ for the forth case, and it is shown in \figurename~\ref{fig:automata for FIFO-4}.
-Here $c_1 = c + \textit{cal}(\textit{rm},b)$, and $c_2 = c + \textit{ret}(\textit{put},b) + \textit{cal}(\textit{rm},a) + \textit{ret}(\textit{rm},a)$, where $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d),\textit{ret}(\textit{rm},d),\textit{rm}(\textit{empty})$.
+Here $c_1 = c + \textit{cal}(\textit{rm},b)$, and $c_2 = c + \textit{ret}(\textit{put},b) + \textit{cal}(\textit{rm},a) + \textit{ret}(\textit{rm},a)$, where $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d),\textit{ret}(\textit{rm},d),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.
 
@@
 
-Let us begin to represent witness automata that is used for capture the existence of a data-differentiated execution $e$, it has a projection $e'$, $\textit{last}(e') = \textit{PQ}_1^{>}$, and there exists a cycle going through the item with maximal priority in $e'$.
+Let us begin to represent witness automata that is used for capture the existence of a data-differentiated execution $e$, $e$ has a projection $e'$, $\textit{last}(e') = \textit{PQ}_1^{>}$, and there exists a cycle going through the item with maximal priority in $e'$.
 By data-independence, we can obtain $e_r$ from $e$ by renaming function, which maps such item to be $b$, maps items that cover it to be $a$, and maps other items into $d$.
@@
 
-\section{Co-Regular of Priority Queues}
-\label{sec:co-regular of priority queues}
-
-\subsection{Co-Regular of $R_{\textit{pr2}}$}
-\label{subsec:co-regular of Rpr2}
-
-By Lemma \ref{lemma:automata for priority queue with single priority}, from now on, it is safe for us to assume that, for each data-differentiated priority queue history $h$ and each $h'$ of projection of $h$, if $h'$ is single-priority, then $\textit{transToQueue}(h')$ is linearizable to queue.
-One consequence of this assumption is that, in $h'$, there does not exists an item $a$, such that $\textit{rm}(a)$ happens before $\textit{put}(a)$.
-
-By the proof of Lemma \ref{lemma:Rpr1 is co-regular}, with the help of $\mathcal{A}_{\textit{Rpr1-pprr}}, \mathcal{A}_{\textit{Rpr1-prpr}}, \mathcal{A}_{\textit{Rpr1-rppr}}$ and $\mathcal{A}_{\textit{Rpr1-rprp}}$, from now on, we can safely assume that given a data-differentiated history $h$ and operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ in $h$, there does not exists any items $d_1,\ldots,d_m$, such that $x$ is not covered by items $d_1,\ldots,d_m$ and the priority of $d_1,\ldots,d_m$ is less that $x$.
-
-However, above two assumptions are not enough for imply the linearizability to priority queue.
-One example is shown in \figurename~\ref{fig:history introduct ob order1}.
-The history $h$ of \figurename~\ref{fig:history introduct ob order1} is not linearizable, even if $h \vert_{1}$ and $h \vert_{2}$ are both linearizable, and either $a$ or $b$ is covered by other items.
-One explanation is that, since $\textit{put}(b,1)$ happens before $\textit{put}(a,1)$, in choosing linearization points, we should make $\textit{rm}(b)$ before $\textit{rm}(a)$.
-However, to satisfy the requirements of $R_{\textit{pr2}}$, the linearization points of $\textit{rm}(a)$ can only locates in the interval from $\textit{ret}(\textit{rm},d)$ to $\textit{cal}(\textit{put},e,2)$.
-But, $\textit{rm}(b)$ is after this entire interval, and we can not locate the linearization point of $\textit{rm}(b)$ in this interval.
-When there is no such phenomenon, such as the history in \figurename~\ref{fig:history introduct ob order2}, the history becomes linearizable.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER1.pdf}
-%\vspace{-10pt}
-  \caption{above assumptions do not implies linearizability (1)}
-  \label{fig:history introduct ob order1}
-\end{figure}
-
-%\begin{figure}[htbp]
-%  \centering
-%  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER3.pdf}
-%\vspace{-10pt}
-%  \caption{above assumptions do not implies linearizability (2)}
-%  \label{fig:history introduct ob order1}
-%\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER2.pdf}
-%\vspace{-10pt}
-  \caption{above assumptions do not implies linearizability (2)}
-  \label{fig:history introduct ob order2}
-\end{figure}
-
-In the following part of this subsection, we use $<_{\textit{ob}}$ order for the case when we can infer that an item is inserted earlier than another item, and we use gap-point to represent the intervals when it is possible to locate linearizaton points of $\textit{rm}(a)$.
-
-\begin{definition}[gap-point for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:gap-point for matched put and rm operations}
-
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x),\textit{rm}(x)$ of $h$, we say $o$ is a gap-point of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$, if we can obtain a sequence $\textit{gapPSeq}(h)$ from $h$ by adding $o$ to some locations between actions of $h$, and satisfies that,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] In $\textit{gapPSeq}(h)$, $o$ is after $\textit{call}(\textit{put},x,\textit{pri}_x)$ and $\textit{call}(\textit{rm},x)$, and is before $\textit{ret}(\textit{rm},x)$.
-\item[-] For each item $y$ with non-$\textit{pri}_x$ priority in $\textit{gapPSeq}(h)$.
-If $\textit{put}(y,\_)$ and $\textit{rm}(y)$ is in $h$, then it can not be the case that $\textit{ret}(\textit{put},y,\_)$ is before $o$ and $\textit{call}(\textit{rm},x)$ is after $o$; Else if $\textit{put}(y,\_)$ is in $h$ and $\textit{rm}(y)$ is not in $h$, then it can not be the case that $\textit{ret}(\textit{put},y,\_)$ is before $o$;
-\end{itemize}
-
-We say that $o$ is a gap-point of priority $\textit{pri}_x$, if we can obtain a sequence $\textit{gapPSeq}(h)$ from $h$ by adding $o$ to some locations between actions of $h$, and satisfies only the latter requirement.
-\end{definition}
-
-\begin{restatable}{lemma}{GapEqualsGapPoint}
-\label{lemma:gap eqals the existence of gap point}
-Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$.
-There is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, if and only if there is a gap-point $o$ of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{only if}$ direction, we need to prove that: if there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, then there is no gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-
-We prove this by contradiction.
-Assume that there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, but there is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-Since there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, in the time interval starts from $\textit{cal}(\textit{put},x,\textit{pri}_x)$ and $\textit{cal}(\textit{rm},x)$ and ends at $\textit{ret}(\textit{rm},x)$, there is always some item with non-$\textit{pri}_x$ priority in priority queue.
-Therefore, it is not hard to prove that there is a cycle going through the left-right constraint of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, and by Lemma \ref{lemma:Gap Equals Constraint for Rpr0}, there is no gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, which contradicts that there is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-
-To prove the $\textit{if}$ direction, assume that there is a gap-point $o$ of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$ and let $h_1$ be the sequence in definition of gap-point, which is generated from $h$ by adding $o$ into $h$.
-Let sequence $h_2$ be a sequence of call, return and operations, such that the projection of $h_2$ into call and return actions is $h$.
-$h_2$ is generated as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] First, from $h_1$, transform the point $o$ into $\textit{rm}(x)$.
-
-\item[-] Then for all the other items $y$, (1) adding an operation $\textit{put}(y,p)$ at some position between $\textit{call}(\textit{put},y,\_)$ and $\textit{ret}(\textit{put},y)$, and (2) adding an operation $\textit{rm}(y)$ at some position between $\textit{call}(\textit{rm},y)$ and $\textit{ret}(\textit{rm},y)$.
-
-\item[-] For all items $z$ with the same priority as $x$, the positions for adding operations of $z$ into $h$ is arbitrary (but need to satisfy above restriction.
-In the following we implicitly require this).
-
-\item[-] For all items $z$ whose priority is different from that of $x$,
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-
-    \item[-] If $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, and $\textit{ret}(\textit{put},z,\_)$ is before $o$ in $h_1$: Then in $h_2$, $\textit{rm}(z,\_)$ is just after $\textit{call}(\textit{rm},z)$.
-
-    \item[-] If $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, and $o$ is before $\textit{ret}(\textit{put},z,\_)$ $h_1$: Then in $h_2$, $\textit{put}(z,\_)$ is just before $\textit{ret}(\textit{put},z)$.
-
-    \item[-] Otherwise, we can see that $\textit{put}(z,\_)$ and $\textit{rm}(z)$ overlaps in $h$.
-Then we add $o_1=\textit{put}(z,\_)$ and $o_2=\textit{rm}(z)$ in the time interval of interleaving of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ overlaps in $h$, make sure $o_2$ next to $o_1$, and make sure either both of them is before $\textit{rm}(x)$ or both of them is after $\textit{rm}(x)$ in $h_2$.
-    \end{itemize}
-\end{itemize}
-
-According to the definition of gap-point, it is not hard to see that we can construct such $h_2$.
-Let $l$ be the projection of $h_2$ into operations, and $l'$ be the projection of $l$ into $x$ and all items that has priorities not equals that of $x$.
-Let $l'= u \cdot \textit{put}(x,\_) \cdot v \cdot \textit{rm}(x) \cdot w$.
-It is easy to see that $h \sqsubseteq l$, and $u \cdot v$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
-Let $L$, $M$ and $R$ be the set of operations in $u$, $v$ and $w$, respectively.
-Then it is obvious that $L$, $M$ and $R$ is a gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
-\qed
-\end {proof}
-
-Given a data-differentiated history $h$ and two items $a,b$, we say that $a <_{\textit{ob}} b$, if the priority of $a$ and $b$ is the same, and one of the following cases holds:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Type $A$: $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$ in $h$,
-
-\item[-] Type $B$: $\textit{rm}(a)$ happens before $\textit{rm}(b)$ in $h$,
-
-\item[-] Type $C$: $\textit{rm}(a)$ happens before $\textit{put}(b,\_)$ in $h$,
-\end{itemize}
-
-Let $<_{\textit{ob}}^*$ be the transitive closure of $<_{\textit{ob}}$.
-Intuitively, $a <_{\textit{ob}} b$ means that $a$ should be inserted earlier than $b$, and $a <_{\textit{ob}}^* b$ means that we can infer that $a$ should be inserted earlier than $b$ with the help of other items.
-We can explicitly write $a <_{\textit{ob}}^A b$ to mean that $a <_{\textit{ob}} b$ and the reason is that $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$.
-Similarly, we can explicitly write $a <_{\textit{ob}}^B b$ and $a <_{\textit{ob}}^C b$ to mean that $a <_{\textit{ob}} b$ and the reason is that $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$ and $\textit{rm}(a)$ happens before $\textit{put}(b,\_)$, respectively.
-
-According to the definition of $<_{\textit{ob}}^*$, if $a <_{\textit{ob}}^* b$, then there exists $c_1,\ldots,c_n$, such that $a <_{\textit{ob}} c_1 <_{\textit{ob}} \ldots <_{\textit{ob}} b$.
-The following lemma states that, the number of intermediate items $c_i$ is in fact bounded.
-
-\begin{restatable}{lemma}{OBOrderHasBoundedLength}
-\label{lemma:ob order has bounded length}
-Given a data-differentiated history $h$.
-Assume that $a <_{\textit{ob}} a_1 <_{\textit{ob}} \ldots <_{\textit{ob}} a_m <_{\textit{ob}} b$, then either $a <_{\textit{ob}} b$, or $a <_{\textit{ob}} a_i <_{\textit{ob}} b$ for some $i$, or $a <_{\textit{ob}} a_i <_{\textit{ob}} a_j <_{\textit{ob}} b$ for some $i,j$.
-\end{restatable}
-
-\begin {proof}
-
-Our proof proceed as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^A$,$<_{\textit{ob}}^B \cdot <_{\textit{ob}}^B$ and $<_{\textit{ob}}^C \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^A c_2 <_{\textit{ob}}^A c_1$, then $\textit{put}(c_3,\_)$ happens before $\textit{put}(c_2,\_)$, and $\textit{put}(c_2,\_)$ happens before $\textit{put}(c_1,\_)$.
-Therefore, it is obvious hat $\textit{put}(c_3,\_)$ happens before $\textit{put}(c_1,\_)$ and $c_3 <_{\textit{ob}}^A c_1$.
-
-    Similarly, if $c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^B c_1$, then $c_3 <_{\textit{ob}}^B c_1$.
-
-    If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^C c_1$: Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
-Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
-
-    Therefore, when we meet successive $<_{\textit{ob}}^A$, it is safe to leave only the first and the last elements and ignore intermediate elements.
-Similar cases hold for $<_{\textit{ob}}^B$ and $<_{\textit{ob}}^C$.
-
-\item[-] $<_{\textit{ob}}^A$ and $<_{\textit{ob}}^C$:
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^A c_2 <_{\textit{ob}}^C c_1$.
-Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{ob}}^A c_2$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
-Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^A c_1$.
-
-    \item[-] ($<_{\textit{ob}}^C \cdot <_{\textit{ob}}^A$): If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^A c_1$.
-Since $c_2 <_{\textit{ob}}^A c_1$, $\textit{put}(c_2,\_)$ happens before $\textit{put}(c_1,\_)$.
-Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{rm}(c_3)$ happens before $\textit{put}(c_2,\_)$.
-Therefore, $\textit{rm}(c_3)$ happens before $\textit{put}(c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
-    \end{itemize}
-
-\item[-] $<_{\textit{ob}}^B$ and $<_{\textit{ob}}^C$:
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] ($<_{\textit{ob}}^B \cdot <_{\textit{ob}}^C$): If $c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^C c_1$.
-Since $c_2 <_{\textit{ob}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-It is obvious that $\textit{cal}(\textit{rm},c_2)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$.
-Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
-
-    \item[-] ($<_{\textit{ob}}^C \cdot <_{\textit{ob}}^B$): If $c_3 <_{\textit{ob}}^C c_2 <_{\textit{ob}}^B c_1$.
-Since $c_2 <_{\textit{ob}}^B c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{rm},c_1)$.
-Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{ob}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
-Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_1)$, and $c_3 <_{\textit{ob}}^B c_1$.
-    \end{itemize}
-
-\item[-]  ($<_{\textit{ob}}^A \cdot <_{\textit{ob}}^B \cdot <_{\textit{ob}}^A$): If $c_4 <_{\textit{ob}}^A c_3 <_{\textit{ob}}^B c_2 <_{\textit{ob}}^A c_1$:
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$: Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$.
-Then $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{ob}}^C c_1$.
-This implies that $c_4 <_{\textit{ob}}^A c_3 <_{\textit{ob}}^C c_1$.
-According to the fact for $<_{\textit{ob}}^A \cdot <_{\textit{ob}}^C$, we know that $c_4  <_{\textit{ob}}^A c_1$.
-
-    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is after $\textit{cal}(\textit{put},c_1,\_)$: Since $c_2 <_{\textit{ob}}^A c_1$, $\textit{ret}(\textit{put},c_2,\_)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-Since $c_3 <_{\textit{ob}}^B c_2$, $\textit{rm}(c_3)$ happens before $\textit{rm}(c_2)$, and then we know that $\textit{put}(c_2,\_)$ can not happen before $\textit{put}(c_3,\_)$.
-Since $\textit{put}(c_2,\_)$ does not happen before $\textit{put}(c_3,\_)$, $\textit{cal}(\textit{put},c_3,\_)$ is before $\textit{ret}(\textit{put},c_2,\_)$.
-Since $c_4 <_{\textit{ob}}^A c_3$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_3,\_)$.
-Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_4 <_{\textit{ob}}^A c_1$.
-    \end{itemize}
-
-\end{itemize}
-
-Based on above results, given $a <_{\textit{ob}}^{b_1} a_1 <_{\textit{ob}} \ldots <_{\textit{ob}}^{b_m} a_m <_{\textit{ob}}^{b_{\textit{m+1}}} b$, where each $b_i$ is in $\{ A,B,C \}$, we can merge relations, until we got one of the following facts:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $a <_{\textit{ob}}^A b$, $a <_{\textit{ob}}^B b$ or $a <_{\textit{ob}}^C b$,
-
-\item[-] $a <_{\textit{ob}}^A a_i <_{\textit{ob}}^B b$, or $a <_{\textit{ob}}^B a_i <_{\textit{ob}}^A b$, for some $i$,
-
-\item[-] $a <_{\textit{ob}}^B a_i <_{\textit{ob}}^A a_j <_{\textit{ob}}^B b$, for some $i$ and $j$,
-\end{itemize}
-
-This completes the proof of this lemma. \qed
-\end {proof}
-
-Given a differentiated history $h$ and operations $\textit{put}(x,\_)$ and $\textit{rm}(x)$ of $h$, the time interval from the time point when both $\textit{call}(\textit{put},x,\_)$ and $\textit{cal}(\textit{rm},x)$ happen and ends at $\textit{ret}(\textit{rm},x)$ is called the effective interval of $x$ in $h$.
-
-\begin{restatable}{lemma}{Rpr2AsGPandOBOrder}
-\label{lemma:Rpr2 as gap-point and ob order}
-Given a data-differentiated history $h$.
-There exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr2}}$, and $h'$ does not linearizable to $\textit{MR}_{\textit{pr2}}$, if and only if there exists $h'' \in \textit{proj}(h)$, $\textit{last}(h'') = R_{\textit{pr2}}$, $x$ and $y$ has maximal priority in $h''$, $y <_{\textit{ob}} x$ in $h''$, and the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$ in $h''$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{if}$ direction, let $h_1$ be the history that is obtained from $h''$ by erasing all operations of items that has same priority as $x$, except for operations of $x$ and $y$.
-It is obvious that $\textit{last}(h_1) = R_{\textit{pr2}}$.
-According to $y <_{\textit{ob}} x$,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $y <_{\textit{ob}}^A x$, then $\textit{put}(y,\_)$ happens before $\textit{put}(x,\_)$.
-Since in $R_{\textit{pr2}}$, $\textit{put}(\textit{itm},\_)$ is after $\textit{put}$ operations of items with maximal priority (except for $\textit{itm}$), we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
-\item[-] If $y <_{\textit{ob}}^B x$, then $\textit{rm}(y)$ happens before $\textit{rm}(x)$.
-Since in $R_{\textit{pr2}}$, $\textit{rm}(\textit{itm})$ is the last $\textit{rm}$ operation of items with maximal priority, we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
-
-\item[-] If $y <_{\textit{ob}}^C x$, then $\textit{rm}(y)$ happens before $\textit{put}(x,\_)$.
-Since in $R_{\textit{pr2}}$, $\textit{rm}(\textit{itm})$ is the last $\textit{rm}$ operation of items with maximal priority, we can see that $x$ should be chosen as $\textit{itm}$ in $R_{\textit{pr2}}$.
-\end{itemize}
-
-Similarly as the proof in Lemma \ref{lemma:gap eqals the existence of gap point}, we can see that the only possible position for locating linearizaton point of $\textit{rm}(x)$ is at gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$.
-Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, then there exists unmatched operations before $\textit{rm}(x)$.
-
-Since the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},x)$ in $h''$ (and also in $h_1$), if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, then $\textit{rm}(x)$ will be before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},x)$.
-
-Therefore, for every sequence $l = u \cdot \textit{put}(x,\_) \cdot v \cdot \textit{rm}(x) \cdot w$, if $h_1 \sqsubseteq l$, then either $u \cdot v$ contains some unmatched operations of priority less than priority of $x$, or $w$ contains $\textit{put}(y,\_)$ or $\textit{rm}(y)$.
-In both cases, $l \notin \textit{MR}_{\textit{pr2}}$.
-
-To prove the $\textit{only if}$ direction, we prove its contrapositive.
-Assume we already know that for each $h'' \in \textit{proj}(h)$ with $\textit{last}(h'') = R_{\textit{pr2}}$, for each $x$ and $y$ has maximal priority in $h''$, if $y <_{\textit{ob}} x$ in $h''$, then the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},x)$ in $h''$.
-We need to prove that for each $h' \in \textit{proj}(h)$, if $\textit{last}(h') = R_{\textit{pr2}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr2}}$.
-
-Let $\textit{pri}$ be the maximal priority of $h'$.
-Recall that we already assume that each single-priority history (if we consider $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$) can be ``linearizable to queue'', and item with larger priority is not covered by items with smaller priority.
-Our proof proceed as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Let $h_{\textit{pri}}$ be the projection of $h$ into operations of priority $\textit{pri}$.
-Since each single-priority history can be ``linearizable to queue'', there exists sequence $l_{\textit{pri}}$, such that $h_{\textit{pri}} \sqsubseteq l_{\textit{pri}}$, and when we treat $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$, $l_{\textit{pri}}$ belongs to queue.
-
-\item[-] Step $1$: Let $a_1$ be the last inserted item of $l_{\textit{pri}}$.
-
-$\textit{Cond1}$: If for each item $b$ with priority $\textit{pri}$ in $h'$, the rightmost gap-point $o$ of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$ does not before any $\textit{cal}(\textit{put},b,\textit{pri})$ or $\textit{cal}(\textit{rm},b)$ in $h'$, then we can
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] Locate the linearization point of $\textit{rm}(a_1)$ at $o$,
-    \item[-] If $\textit{put}(a_1,\textit{pri})$ overlaps with $\textit{rm}(a_1)$, then locate the linearization point of $\textit{put}(a_1,\_)$ just before the linearization point of $\textit{rm}(a_1)$.
-Otherwise, $\textit{put}(a_1,\textit{pri}) <_{\textit{hb}} \textit{rm}(a_1)$, and we locate the linearization point of $\textit{put}(a_1,\_)$ just before its return action.
-    \item[-] Locate linearization points of operations of items of priority $\textit{pri}$ (except for $a_1$) just after the call action of the operation.
+\subsection{Proofs and Definitions in Subsection \ref{subsec:co-regular of PQ1Equal}}
+\label{sec:appendix proof and definition in section co-regular of PQ1Equal}
+
+Let $\textit{Items}(e,p)$ be the set of items with priority $p$ in execution $e$. The following lemma states a method to choose $\textit{itm}$ of $\textit{PQ}_1^{=}$.
+
+\begin{restatable}{lemma}{MaximalInPBadGPMakePQ1Equal}
+\label{lemma:maximal in pb and gap-point make a candidate of PQ1Equal}
+Given a data-differentiated execution $e$ with $\textit{last}(e) = \textit{PQ}_1^{=}$.
+Let $\textit{pri}$ be the maximal priority in $e$.
+If there exists an item $x$ with priority $\textit{pri}$, such that for each $y \in \textit{Items}(e,\textit{pri})$, (1) $x$ does not $<_{\textit{pb}}$ to $y$, and (2) the right-most gap-point of $x$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},y)$.
+Then $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
+\end{restatable}
+
+\begin {proof}
+
+Let $o$ be the right-most gap-point of $x$. We locate linearization points of each method event as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Locate the linearization point of $\textit{rm}(x)$ at $o$,
+
+\item[-] If $\textit{put}(x,\textit{pri})$ overlaps with $\textit{rm}(x)$, then locate the linearization point of $\textit{put}(x,\textit{pri})$ just before the linearization point of $\textit{rm}(x)$.
+Otherwise, $\textit{put}(x,\textit{pri}) <_{\textit{hb}} \textit{rm}(x)$, and we locate the linearization point of $\textit{put}(x,\textit{pri})$ just before its return action.
+
+\item[-] Locate linearization points of method event of each $y \in \textit{Items}(e,\textit{pri})$ (except for $x$) just after the call action of the method event.
+
 \item[-] For item $z$ with priority smaller than $\textit{pri}$.
 If both $\textit{cal}(\textit{put},z,\_)$ and $\textit{cal}(\textit{rm},z)$ is before $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just after their call actions.
-If both $\textit{ret}(\textit{put},z)$ and $\textit{ret}(\textit{rm},z)$ is after $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just before their return actions.
-
-        Otherwise, either (1) $\textit{rm}(z) <_{\textit{hb}} \textit{put}(z,\_)$, or (2) $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, $\textit{ret}(\textit{put},z,\_)$ before $o$, and $o$ before $\textit{call}(\textit{rm},z)$.
-The first case violates that each single-priority history can be ``linearizable to queue'', and the second case violates that $o$ is a gap-point.
-    \end{itemize}
-
-    In this way we construct a sequence $l$, and it is easy to see that $h' \sqsubseteq l$ and $l$ satisfies the requirement of $R_{\textit{pr2}}$.
-
-\item[-] Otherwise, there exists $a_2$ with priority $\textit{pri}$, such that the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$ in $h'$.
-We can see that the rightmost gap-point of $\textit{put}(a_2,\textit{pri})$ and $\textit{rm}(a_2)$ is after the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$.
-By assumption we know that $a_2$ does not $<_{\textit{ob}}$ to $a_1$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If for each item $b$ with priority $\textit{pri}$ in $h'$, $a_2$ does not $<_{\textit{ob}}$ to $b$.
-Then we go to step $1$ and treat $a_2$ as $a_1$ (When $\textit{Cond1}$ of step $1$ holds, we break the loop).
-    \item[-] Otherwise, there exists $a_3$ with priority $\textit{pri}$ such that $a_2 <_{\textit{ob}} a_3$.
-By assumption,we know that the rightmost gap-point of $\textit{put}(a_3,\textit{pri})$ and $\textit{rm}(a_3)$ is not before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$.
-Then we can see that the rightmost gap-point of $\textit{put}(a_3,\textit{pri})$ and $\textit{rm}(a_3)$ is after the rightmost gap-point of $\textit{put}(a_1,\textit{pri})$ and $\textit{rm}(a_1)$.
-Then we go to step $1$ and treat $a_3$ as $a_1$ (When $\textit{Cond1}$ of step $1$ holds, we break the loop).
-    \end{itemize}
-\end{itemize}
-
-It is not hard to see that, if $b_i$ is the $a_1$ in the $\textit{i-th}$ loop of our proof, $b_j$ is the $a_1$ in the $\textit{j-th}$ loop of our proof, and $i<j$, then the rightmost gap-point of $\textit{put}(b_j,\textit{pri})$ and $\textit{rm}(b_j)$ is after the rightmost gap-point of $\textit{put}(b_i,\textit{pri})$ and $\textit{rm}(b_i)$.
-Thus, we finally stop at some $d$.
-$d$ satisfies that, for each item $c$ with priority $\textit{pri}$, $d$ does not $<_{\textit{ob}}$ to $c$.
-$d$ also satisfies that, for each item $c$ with priority $\textit{pri}$ in $h'$, the rightmost gap-point $o'$ of $\textit{put}(d,\textit{pri})$ and $\textit{rm}(d)$ does not before any $\textit{cal}(\textit{put},c,\textit{pri})$ or $\textit{cal}(\textit{rm},c)$.
-Then, we can
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Locate the linearization point of $\textit{rm}(d)$ at the rightmost gap-point $o'$ of $\textit{put}(d,\textit{pri})$ and $\textit{rm}(d)$,
-
-\item[-] Since for each item $c$ with priority $\textit{pri}$ in $h'$, $o'$ is after any $\textit{cal}(\textit{put},c,\textit{pri})$ and $\textit{cal}(\textit{rm},c)$.
-For each $c$, one of the following cases hold: (1) $\textit{cal}(\textit{rm},c)$ is before $o'$ and $o'$ is before $\textit{ret}(\textit{rm},c)$, (2) $\textit{cal}(\textit{rm},c)$ and $\textit{ret}(\textit{rm},c)$ are before $o'$.
-We locate the linearization point of $\textit{rm}(c)$ just after its call action.
-Such linearization point is obviously before $o'$.
-
-\item[-] If $o'$ is before $\textit{ret}(\textit{put},d,\textit{pri})$, then locate the linearization point of $\textit{put}(d,\textit{pri})$ just before $o'$.
-Else, $\textit{ret}(\textit{put},d,\textit{pri})$ is after $o'$, and we locate the linearization point of $\textit{put}(d,\textit{pri})$ just before the return action of it.
-
-\item[-] If $o'$ is before $\textit{ret}(\textit{put},d,\textit{pri})$, we can see that, for each item $c$ with priority $\textit{pri}$ in $h'$, $\textit{cal}(\textit{put},c,\textit{pri})$ is before $o'$.
-Then, and we locate the linearization point of $\textit{put}(c,\textit{pri})$ just before $o'$ and the linearization point of $\textit{put}(d,\textit{pri})$.
-
-    Otherwise, $\textit{ret}(\textit{put},d,\textit{pri})$ is before $o'$.
-Since $d$ does not $<_{\textit{ob}}$ to $c$, $\textit{put}(d,\textit{pri})$ does not happen before $\textit{put}(c,\textit{pri})$, or we can say, $\textit{ret}(\textit{put},d,\textit{pri})$ is after $\textit{cal}(\textit{put},c,\textit{pri})$.
-Then, we locate the linearization point of $\textit{put}(c,\textit{pri})$ just after its call action.
-\end{itemize}
-
-In this way we construct a sequence $l_d$, and it is easy to see that $h' \sqsubseteq l_d$ and $l_d$ satisfies the requirement of $R_{\textit{pr2}}$ with $d$ acts as $\textit{itm}$.
+If both $\textit{ret}(\textit{put},z)$ and $\textit{ret}(\textit{rm},z)$ (if exists) is after $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just before their return actions.
+Otherwise, $x$ is in interval of $z$, which contradicts the definition of gap-point.
+\end{itemize}
+
+Let $l$ be the sequence of linearization points constructed above.
+It is obvious that $e \sqsubseteq l$.
+Since for each $y \in \textit{Items}(e,\textit{pri})$, $o$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},x)$, we can see that $\textit{rm}(x)$ is after $\textit{put}(y,\textit{pri})$ and $\textit{rm}(y)$ in $l$.
+It is obvious that $\textit{put}(x,\textit{pri})$ is before $\textit{rm}(x)$ in $l$.
+Since $x$ does not $<_{\textit{pb}}$ to $y$, we can see that no $\textit{put}(y,\textit{pri})$ happens before $\textit{put}(x,\textit{pri})$.
+Then it is easy to see that $\textit{put}(x,\textit{pri})$ is after $\textit{put}(y,\textit{pri})$ in $l$.
+Since $\textit{last}(e) = \textit{PQ}_1^{=}$, all other items in $\textit{Items}(e,\textit{pri})$ has matched $\textit{put}$ and $\textit{rm}$, and it is easy to see that they are all before $\textit{rm}(x)$ in $l$.
+
+For item $z$ with priority smaller than $\textit{pri}$, we can see that there are only two possibilities: (1) $\textit{put}(z,\_)$ and $\textit{rm}(z)$ are both before $\textit{rm}(x)$ in $l$, and (2) $\textit{put}(z,\_)$ and $\textit{rm}(z)$ (if exists) are after before $\textit{rm}(x)$ in $l$.
+Therefore, before $\textit{rm}(x)$ in $l$, the $\textit{put}$ and $\textit{rm}$ of $z$ are matched.
+
+Therefore, it is easy to see that $l \in \textit{MS}(\textit{PQ}_1^{=})$. \qed
+\end {proof}
+
+With Lemma \ref{lemma:maximal in pb and gap-point make a candidate of PQ1Equal}, we can prove the following lemma, which states that getting rid of case in \figurename~\ref{fig:introduce pb order} is enough for ensure $\textit{last}(e) = \textit{PQ}_1^{=} \Rightarrow e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
+
+\PQOneEqualAsPBandGP*
+
+\begin {proof}
+
+To prove the $\textit{if}$ direction, let $e_{x,y}$ be the execution that is obtained from $e$ by erasing all actions of items that has same priority as $x$, except for actions of $x$ and $y$.
+It is obvious that $\textit{last}(e_{x,y}) = \textit{PQ}_1^{=}$.
+Since $y <_{\textit{pb}}^* x$, according to $\textit{PQ}_1^{=}$, we can see that $x$ should be chosen as $\textit{itm}$ in $\textit{PQ}_1^{=}$.
+
+According to Lemma \ref{lemma:Lin Equals Constraint for PQ1Lar} (Here we temporarily forget the existence of $y$), the only possible position for locating linearizaton point of $\textit{rm}(x)$ is at gap-point of $x$.
+Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $x$, then there exists unmatched method event before $\textit{rm}(x)$ with smaller priority.
+Since the rightmost gap-point of $x$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$, if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $x$, then $\textit{rm}(x)$ will be before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},x)$.
+
+Therefore, for every sequence $l = u \cdot \textit{put}(x,\_) \cdot v \cdot \textit{rm}(x) \cdot w$, if $e_{x,y} \sqsubseteq l$, then either $u \cdot v$ contains some unmatched method events of priority less than priority of $x$, or $w$ contains $\textit{put}(y,\_)$ or $\textit{rm}(y)$.
+In both cases, $l \notin \textit{MS}(\textit{PQ}_1^{=})$.
+
+To prove the $\textit{only if}$ direction, we prove its contrapositive.
+Assume we already know that for each $x$ and $y$ has maximal priority in $e$, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},x)$.
+We need to prove that $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
+Let $\textit{pri}$ be the maximal priority of $e$.
+Recall that we already assume that each single-priority execution has FIFO property, and item with larger priority is not covered by items with smaller priority.
+
+Our proof proceed as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Let $e_{\textit{pri}}$ be the projection of $e$ into operations of priority $\textit{pri}$.
+Since each single-priority execution has FIFO property, there exists sequence $l_{\textit{pri}}$, such that $e_{\textit{pri}} \sqsubseteq l_{\textit{pri}}$, and when we treat $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$, $l_{\textit{pri}}$ belongs to queue.
+
+\item[-] Let $a_1$ be the last inserted item of $l_{\textit{pri}}$.
+
+    Step $1$: Check whether for each $b \in \textit{Items}(e,\textit{pri})$, (1) $a_1$ does not $<_{\textit{pb}}$ to $b$, and (2) the right-most gap-point of $a$ is after $\textit{cal}(\textit{put},b,\textit{pri})$ and $\textit{cal}(\textit{rm},b)$.
+
+    It is easy to see that $a_1$ is of priority $\textit{pri}$, and $a_1$ does not $<_{\textit{pb}}$ to any $b \in \textit{Items}(e,\textit{pri})$.
+If for each $b \in \textit{Items}(e,\textit{pri})$, the rightmost gap-point of $a_1$ is after $\textit{cal}(\textit{put},b,\textit{pri})$ and $\textit{cal}(\textit{rm},b)$.
+Then by Lemma \ref{lemma:maximal in pb and gap-point make a candidate of PQ1Equal}, we can obtain that $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
+
+\item[-] Otherwise, there exists $a_2 \in \textit{Items}(e,\textit{pri})$, such that the rightmost gap-point of $a_1$ is before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$ in $e$.
+We can see that each gap-point of $a_2$ is after the rightmost gap-point of $a_1$.%, and thus, the right-most gap-point of $a_2$ is after the rightmost gap-point of $a_1$.
+    By assumption, we know that $a_2$ does not $<_{\textit{pb}}$ to $a_1$.
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If for each item $b \in \textit{Items}(e,\textit{pri})$, $a_2$ does not $<_{\textit{pb}}$ to $b$. Then we go to step $1$ and treat $a_2$ similarly as $a_1$.
+    \item[-] Otherwise, there exists $a_3$ with priority $\textit{pri}$ such that $a_2 <_{\textit{pb}}^* a_3$.
+
+    Since $l_{\textit{pri}}$ has FIFO property, it is easy to see that there is no cycle in $<_{\textit{pb}}$ order.
+It is safe to assume that $a_3$ is maximal in the sense of $<_{\textit{pb}}^*$.
+Or we can say, there does not exists $a_4$, such that $a_3 <_{\textit{pb}}^* a_4$.
+
+    By assumption,we know that the rightmost gap-point of $a_3$ is after $\textit{cal}(\textit{put},a_2,\textit{pri})$ and $\textit{cal}(\textit{rm},a_2)$.
+Therefore, we can see that the rightmost gap-point of $a_3$ is after the rightmost gap-point of $a_3$.
+Then we go to step $1$ and treat $a_3$ similarly as $a_1$.
+    \end{itemize}
+\end{itemize}
+
+Let $a^i$ be the $a_1$ in the $\textit{i-th}$ loop of our proof.
+It is not hard to see that, given $i<j$, the rightmost gap-point of $b_j$ is after the rightmost gap-point of $b_i$.
+Therefore, the loop finally stop at some $a^f$.
+$a^f$ satisfies the check of Step $1$.
+By Lemma \ref{lemma:maximal in pb and gap-point make a candidate of PQ1Equal}, this implies that $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
 This completes the proof of $\textit{if}$ direction.
@@
 
-Let us begin to represent several automata that is used for capture the case that, in a sub-history $h''$ of $h$, $\textit{last}(h'') = R_{\textit{pr2}}$, $h'$ does not linearizable to $\textit{MR}_{\textit{pr2}}$, and the reason is that there exists items $a$ and $b$ with maximal priority in $h''$, $a <_{\textit{ob}} b$ in $h''$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
-
-According to Lemma \ref{lemma:ob order has bounded length}, if $a <_{\textit{ob}} b$, then one of the following cases hold: (1) $a <_{\textit{ob}}^A b$, (2) $a <_{\textit{ob}}^B b$, (3) $a <_{\textit{ob}}^C b$, (4) $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$ for some $a_1$, (5) $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$ for some $a_1$, or (6) $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$ for some $a_1$ and $a_2$.
-Although there are many possible enumeration of $a$, $b$, $a_1$ and $b_1$, the following lemma states that we need only to consider five enumeration of operations of $a$, $b$, $a_1$ and $b_1$.
-
-\begin{restatable}{lemma}{FiveEnmuerationisEnoughForRpr2}
-\label{lemma:five enumeration is enough for Rpr2}
-Given a data-differentiated history $h$, $\textit{last}(h) = R_{\textit{pr2}}$, and $h$ does not linearizable to $\textit{MR}_{\textit{pr2}}$.
-Assume that there exists items $a$ and $b$ with maximal priority in $h''$, $a <_{\textit{ob}} b$, and the rightmost gap-point $o$ of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
-Let $a_1$ and $a_2$ be the value when possibly $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$, or $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$.
-Then, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists) and $a_2$ (if exists).
-\end{restatable}
-\begin {proof}
-
-Let us prove by consider all the possible reason of $a <_{\textit{ob}}^A b$.
-According to Lemma \ref{lemma:ob order has bounded length}, we need to consider six reasons: %There may exists items with smaller priority of $a$ and covers some time interval, and assume that we rename them into $d$:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Reason $1$, $a <_{\textit{ob}}^A b$:
-
-    Since $a <_{\textit{ob}}^A b$, $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$.
-Since $o$ is after $\textit{cal}(\textit{put},a,\_)$, we can see that $o$ is before $\textit{cal}(\textit{rm},b)$.
-
-    Since single-priority must satisfy the first in firs out property, $\textit{rm}(b)$ does not happen before $\textit{rm}(a)$, and $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
-If $\textit{ret}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$, since we already assume that there exists gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$, this gap-point is also a gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$, and is after $o$, which contradicts that $o$ is the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$.
+According to the definition of $<_{\textit{ob}}^*$, if $a <_{\textit{pb}}^* b$, then there exists $a_1,\ldots,a_m$, such that $a <_{\textit{pb}} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}} a_m <_{\textit{pb}} b$.
+The following lemma states that, the number of intermediate items $a_i$ is in fact bounded.
+
+\OBOrderHasBoundedLength*
+
+\begin {proof}
+
+Our proof proceed as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] ($<_{\textit{pb}}^A \cdot <_{\textit{pb}}^A$,$<_{\textit{pb}}^B \cdot <_{\textit{pb}}^B$ and $<_{\textit{pb}}^C \cdot <_{\textit{pb}}^C$): If $c_3 <_{\textit{pb}}^A c_2 <_{\textit{pb}}^A c_1$, then $\textit{put}(c_3,\_)$ happens before $\textit{put}(c_2,\_)$, and $\textit{put}(c_2,\_)$ happens before $\textit{put}(c_1,\_)$.
+Therefore, it is obvious that $\textit{put}(c_3,\_)$ happens before $\textit{put}(c_1,\_)$ and $c_3 <_{\textit{pb}}^A c_1$.
+
+    Similarly, if $c_3 <_{\textit{pb}}^B c_2 <_{\textit{pb}}^B c_1$, then $c_3 <_{\textit{pb}}^B c_1$.
+
+    If $c_3 <_{\textit{pb}}^C c_2 <_{\textit{pb}}^C c_1$: Since $c_2 <_{\textit{pb}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
+Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
+Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
+Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.
+
+    Therefore, when we meet successive $<_{\textit{pb}}^A$, it is safe to leave only the first and the last elements and ignore intermediate elements.
+Similar cases hold for $<_{\textit{pb}}^B$ and $<_{\textit{pb}}^C$.
+
+\item[-] $<_{\textit{pb}}^A$ and $<_{\textit{pb}}^C$:
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] ($<_{\textit{pb}}^A \cdot <_{\textit{pb}}^C$): If $c_3 <_{\textit{pb}}^A c_2 <_{\textit{pb}}^C c_1$.
+Since $c_2 <_{\textit{pb}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
+Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
+Since $c_3 <_{\textit{pb}}^A c_2$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
+Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^A c_1$.
+
+    \item[-] ($<_{\textit{pb}}^C \cdot <_{\textit{pb}}^A$): If $c_3 <_{\textit{pb}}^C c_2 <_{\textit{pb}}^A c_1$.
+Since $c_2 <_{\textit{pb}}^A c_1$, $\textit{ret}(\textit{put},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
+It is obvious that $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{put},c_2)$.
+Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
+Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.
+
+    %Since $c_2 <_{\textit{pb}}^A c_1$, $\textit{put}(c_2,\_)$ happens before $\textit{put}(c_1,\_)$.
+Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{rm}(c_3)$ happens before $\textit{put}(c_2,\_)$.
+Therefore, $\textit{rm}(c_3)$ happens before $\textit{put}(c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.
+    \end{itemize}
+
+\item[-] $<_{\textit{pb}}^B$ and $<_{\textit{pb}}^C$:
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] ($<_{\textit{pb}}^B \cdot <_{\textit{pb}}^C$): If $c_3 <_{\textit{pb}}^B c_2 <_{\textit{pb}}^C c_1$.
+Since $c_2 <_{\textit{pb}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
+It is obvious that $\textit{cal}(\textit{rm},c_2)$ is before $\textit{ret}(\textit{rm},c_2)$.
+Since $c_3 <_{\textit{pb}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$.
+Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.
+
+    \item[-] ($<_{\textit{pb}}^C \cdot <_{\textit{pb}}^B$): If $c_3 <_{\textit{pb}}^C c_2 <_{\textit{pb}}^B c_1$.
+Since $c_2 <_{\textit{pb}}^B c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{rm},c_1)$.
+Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
+Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
+Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_1)$, and $c_3 <_{\textit{pb}}^B c_1$.
+    \end{itemize}
+
+\item[-]  ($<_{\textit{pb}}^A \cdot <_{\textit{pb}}^B \cdot <_{\textit{pb}}^A$): If $c_4 <_{\textit{pb}}^A c_3 <_{\textit{pb}}^B c_2 <_{\textit{pb}}^A c_1$:
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$: Since $c_3 <_{\textit{pb}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$.
+Then $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.
+This implies that $c_4 <_{\textit{pb}}^A c_3 <_{\textit{pb}}^C c_1$.
+According to the fact for $<_{\textit{pb}}^A \cdot <_{\textit{pb}}^C$, we know that $c_4  <_{\textit{pb}}^A c_1$.
+
+    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is after $\textit{cal}(\textit{put},c_1,\_)$: Since $c_2 <_{\textit{pb}}^A c_1$, $\textit{ret}(\textit{put},c_2,\_)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
+Since $c_3 <_{\textit{pb}}^B c_2$, $\textit{rm}(c_3)$ happens before $\textit{rm}(c_2)$, and then we know that $\textit{put}(c_2,\_)$ can not happen before $\textit{put}(c_3,\_)$.
+Since $\textit{put}(c_2,\_)$ does not happen before $\textit{put}(c_3,\_)$, $\textit{cal}(\textit{put},c_3,\_)$ is before $\textit{ret}(\textit{put},c_2,\_)$.
+Since $c_4 <_{\textit{pb}}^A c_3$, $\textit{ret}(\textit{put},c_4)$ is before $\textit{cal}(\textit{put},c_3,\_)$.
+Therefore, $\textit{ret}(\textit{put},c_4)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_4 <_{\textit{pb}}^A c_1$.
+    \end{itemize}
+
+\item[-]  ($<_{\textit{pb}}^B \cdot <_{\textit{pb}}^A \cdot <_{\textit{pb}}^B$): If $c_4 <_{\textit{pb}}^B c_3 <_{\textit{pb}}^A c_2 <_{\textit{pb}}^B c_1$: Since $c_2 <_{\textit{pb}}^B c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{rm},c_1)$.
+Since $c_3 <_{\textit{pb}}^A c_2$, we can see that $\textit{put}(c_3,\_) <_{\textit{hb}} \textit{put}(c_2,\_)$.
+Since each single-priority execution has FIFO property, we know that $\textit{rm}(c_2)$ does not happen before $\textit{rm}(c_3)$, and thus, $\textit{cal}(\textit{rm},c_3)$ is before $\textit{ret}(\textit{rm},c_2)$.
+Since $c_4 <_{\textit{pb}}^B c_3$, $\textit{ret}(\textit{rm},c_4)$ is before $\textit{cal}(\textit{rm},c_3)$.
+Therefore, $\textit{ret}(\textit{rm},c_4)$ is before $\textit{cal}(\textit{rm},c_1)$, and $c_4 <_{\textit{pb}}^B c_1$.
+
+\end{itemize}
+
+Based on above results, given $a <_{\textit{pb}}^{b_1} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}}^{b_m} a_m <_{\textit{pb}}^{b_{\textit{m+1}}} b$, where each $b_i$ is in $\{ A,B,C \}$, we can merge relations, until we got one of the following facts:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$ or $a <_{\textit{pb}}^C b$,
+
+\item[-] $a <_{\textit{pb}}^A a_i <_{\textit{pb}}^B b$, or $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A b$, for some $i$,
+
+\item[-] $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A a_j <_{\textit{pb}}^B b$, for some $i$ and $j$,
+\end{itemize}
+
+This completes the proof of this lemma. \qed
+\end {proof}
+
+There are many enumerations of method events of $a$, $b$ and $a_1$ that may makes $a <_{\textit{pb}}^* b$.
+The following lemma states that with the help of gap-points, the number of potential enumerations can be further reduced into only five.
+
+\FiveEnmuerationisEnoughForPQOneEqual*
+
+\begin {proof}
+
+Let us prove by consider all the possible reason of $a <_{\textit{pb}}^* b$.
+According to Lemma \ref{lemma:ob order has bounded length}, we need to consider five reasons: Let $o$ be the right-most gap-point of $b$.
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Reason $1$, $a <_{\textit{pb}}^A b$:
+
+    Since $a <_{\textit{pb}}^A b$, $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$.
+Since $o$ is after $\textit{cal}(\textit{put},b,\_)$, and thus, after $\textit{cal}(\textit{put},a,\_)$, we can see that $o$ is before $\textit{cal}(\textit{rm},b)$.
+
+    Since single-priority execution must satisfy the FIFO property, $\textit{rm}(b)$ does not happen before $\textit{rm}(a)$, and thus, $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
+If $\textit{ret}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$, since we already assume that there exists gap-point of $a$, this gap-point is also a gap-point of $b$, and is after $o$, which contradicts that $o$ is the rightmost gap-point of $b$.
 Therefore, $\textit{ret}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$.
 
-    According to above discussion, there are two possible enumeration of operations of $a$ and $b$, as shown in \figurename~\ref{fig:history enumeration 1 for Rpr2} and \figurename~\ref{fig:history enumeration 2 for Rpr2}.
-Here we explicitly draw the leftmost gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$ as $o'$.
+    According to above discussion, there are two possible enumeration of operations of $a$ and $b$, as shown in \figurename~\ref{fig:history enumeration 1 for PQ1Equal} and \figurename~\ref{fig:history enumeration 2 for PQ1Equal}.
+Here we explicitly draw the leftmost gap-point of $a$ as $o'$.
 Since the position of $\textit{ret}(\textit{put},b)$ does not influence the correctness, we can simply ignore it.
 
-\item[-] Reason $2$, $a <_{\textit{ob}}^B b$:
-
-    Since $a <_{\textit{ob}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$.
+\item[-] Reason $2$, $a <_{\textit{pb}}^B b$:
+
+    Since $a <_{\textit{pb}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$.
 Since $o$ is after $\textit{cal}(\textit{rm},b)$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$.
@@
 
-\item[-] Reason $3$, $a <_{\textit{ob}}^C b$:
-
-    Since $a <_{\textit{ob}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},b)$.
+\item[-] Reason $3$, $a <_{\textit{pb}}^C b$:
+
+    Since $a <_{\textit{pb}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},b,\_)$.
 Since $o$ is after $\textit{cal}(\textit{put},b)$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$.
@@
 
-\item[-] Reason $4$, $a <_{\textit{ob}}^A a_1 <_{\textit{ob}}^B b$:
-
-    Since $a_1 <_{\textit{ob}}^B b$, $\textit{rm}(a_1) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_1)$ is before $\textit{cal}(\textit{rm},b)$.
+\item[-] Reason $4$, $a <_{\textit{pb}}^A a_1 <_{\textit{pb}}^B b$:
+
+    Since $a_1 <_{\textit{pb}}^B b$, $\textit{rm}(a_1) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_1)$ is before $\textit{cal}(\textit{rm},b)$.
 Since $\textit{rm}(a_1)$ does not happen before $\textit{put}(a_1)$, $\textit{cal}(\textit{put},a_1,\_)$ is before $\textit{ret}(\textit{rm},a_1)$.
-Since $a <_{\textit{ob}}^A a_1$, $\textit{ret}(\textit{put},a,\_)$ is before $\textit{cal}(\textit{put},a_1,\_)$.
+Since $a <_{\textit{pb}}^A a_1$, $\textit{ret}(\textit{put},a,\_)$ is before $\textit{cal}(\textit{put},a_1,\_)$.
 Therefore, $\textit{ret}(\textit{put},a,\_)$ is before $\textit{cal}(\textit{rm},b)$.
@@
 
-    If $\textit{cal}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$, then $\textit{rm}(b) <_{\textit{hb}} \textit{rm}(a)$.
-Let $h' = h \vert_{ \{ a,a_1,b \} }$.
-It is not hard to see that $h'$ can not linearizable to a sequence that satisfies the first in first our property, and this contradicts to our assumption that each single-priority history satisfies the first in first out property.
+    If $\textit{cal}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$, then $e \vert_{ \{ a,a_1,b \} }$ violates the FIFO property.
 Therefore, $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
@@
 
-    According to above discussion, there are three possible enumeration of operations of $a$, $a_1$ and $b$, as shown in \figurename~\ref{fig:history enumeration 3 for Rpr2}, \figurename~\ref{fig:history enumeration 4 for Rpr2} and \figurename~\ref{fig:history enumeration 5 for Rpr2}.
-Here we explicitly draw the leftmost gap-point of $\textit{put}(a,\_)$ and $\textit{rm}(a)$ as $o'$.
-Since the position of $\textit{cal}(\textit{put},b,\_)$, $\textit{ret}(\textit{put},b)$, $\textit{ret}(\textit{put},a_1,\_)$ and $\textit{cal}(\textit{put},a,\_)$ do not influence the correctness, we can simply ignore it.
-
-\item[-] Reason $5$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A b$:
-
-    Since $a_1 <_{\textit{ob}}^A b$, $\textit{ret}(\textit{put},a_1,\_)$ is before $\textit{call}(\textit{put},b,\_)$.
+    According to above discussion, there are three possible enumeration of operations of $a$, $a_1$ and $b$, as shown in \figurename~\ref{fig:history enumeration 3 for PQ1Equal}, \figurename~\ref{fig:history enumeration 4 for PQ1Equal} and \figurename~\ref{fig:history enumeration 5 for PQ1Equal}.
+Here we explicitly draw the leftmost gap-point of $a$ as $o'$.
+Since the position of $\textit{ret}(\textit{put},a_1,\_)$ and $\textit{cal}(\textit{put},a,\_)$ do not influence the correctness, we can simply ignore it.
+We also ignore $\textit{cal}(\textit{put},b,\_)$ and $\textit{ret}(\textit{put},b)$, since the only requirements of them are (1) $\textit{rm}(b)$ does not happen before $\textit{put}(b)$ and (2) $\textit{cal}(\textit{put},b,\_)$ is before $o$.
+
+\item[-] Reason $5$, $a <_{\textit{pb}}^B a_1 <_{\textit{pb}}^A b$:
+
+    Since $a_1 <_{\textit{pb}}^A b$, $\textit{ret}(\textit{put},a_1)$ is before $\textit{call}(\textit{put},b,\_)$.
 Since $\textit{call}(\textit{put},b,\_)$ is before $o$, we can see that $\textit{ret}(\textit{put},a_1,\_)$ is before $o$.
@@
     \item[-] If $o$ is before $\textit{cal}(\textit{rm},a)$: Then $o$ is obviously before $\textit{ret}(\textit{rm},a)$.
-Since $a <_{\textit{ob}}^B a_1$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},a_1)$.
-Then we can see that, $o$ is before $\textit{cal}(\textit{rm},a_1)$, and remember that $a_1 <_{\textit{ob}}^A b$.
+Since $a <_{\textit{pb}}^B a_1$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},a_1)$.
+Then we can see that, $o$ is before $\textit{cal}(\textit{rm},a_1)$, and remember that $a_1 <_{\textit{pb}}^A b$.
 Then we can goto the case of reason $1$ and treat $a_1$ as $a$.
@@
     \item[-] If $o$ is before $\textit{cal}(\textit{put},a,\_)$: Since $\textit{rm}(a)$ does not happen before $\textit{put}(a,\_)$, we can see that $\textit{cal}(\textit{put},a,\_)$ is before $\textit{ret}(\textit{rm},a)$, and then $o$ is before $\textit{ret}(\textit{rm},a)$.
-Then similarly as above case, we can see that $o$ is before $\textit{cal}(\textit{rm},a_1)$, and $a_1 <_{\textit{ob}}^A b$.
+Then similarly as above case, we can see that $o$ is before $\textit{cal}(\textit{rm},a_1)$, and $a_1 <_{\textit{pb}}^A b$.
 Then we can goto the case of reason $1$ and treat $a_1$ as $a$.
@@
     \end{itemize}
-
-\item[-] Reason $6$, $a <_{\textit{ob}}^B a_1 <_{\textit{ob}}^A a_2 <_{\textit{ob}}^B b$:
-
-    Since $a_2 <_{\textit{ob}}^B b$, $\textit{rm}(a_2) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_2)$ is before $\textit{cal}(\textit{rm},b)$.
-%Since $\textit{rm}(a_2)$ does not happen before $\textit{put}(a_2,\_)$, $\textit{cal}(\textit{put},a_2,\_)$ is before $\textit{ret}(\textit{rm},a_2)$.
-%Since $a_1 <_{\textit{ob}}^A a_2$, $\textit{ret}(\textit{put},a_1,\_)$ is before $\textit{cal}(\textit{put},a_2,\_)$.
-    Since $a_1 <_{\textit{ob}}^A a_2$, $\textit{put}(a_1,\_) <_{\textit{hb}} \textit{put}(a_2,\_)$, and $\textit{rm}(a_2)$ does not happen before $\textit{rm}(a_1)$, and $\textit{cal}(\textit{rm},a_1)$ is before $\textit{ret}(\textit{rm},a_2)$.
-Since $a <_{\textit{ob}}^B a_1$, $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(a_1)$, and $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},a_1)$.
-
-    Therefore, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$.
-Since $\textit{cal}(\textit{rm},b)$ is before $o$, we can see that $\textit{ret}(\textit{rm},a)$ is before $o$.
-Then it is easy to see that $o$ is before $\textit{cal}(\textit{put},a,\_)$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\_)$, and $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a,\_)$.
-This is impossible, and we can safely ignore this reason.
 \end{itemize}
@@
   \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-1.pdf}
+  \includegraphics[width=0.4 \textwidth]{figures/PIC-HIS-PQ1Equal-1.pdf}
 %\vspace{-10pt}
   \caption{The first possible enumeration.}
-  \label{fig:history enumeration 1 for Rpr2}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-2.pdf}
+  \label{fig:history enumeration 1 for PQ1Equal}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.4 \textwidth]{figures/PIC-HIS-PQ1Equal-2.pdf}
 %\vspace{-10pt}
   \caption{The second possible enumeration.}
-  \label{fig:history enumeration 2 for Rpr2}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-3.pdf}
+  \label{fig:history enumeration 2 for PQ1Equal}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.4 \textwidth]{figures/PIC-HIS-PQ1Equal-3.pdf}
 %\vspace{-10pt}
   \caption{The third possible enumeration.}
-  \label{fig:history enumeration 3 for Rpr2}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-4.pdf}
+  \label{fig:history enumeration 3 for PQ1Equal}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.4 \textwidth]{figures/PIC-HIS-PQ1Equal-4.pdf}
 %\vspace{-10pt}
   \caption{The forth possible enumeration.}
-  \label{fig:history enumeration 4 for Rpr2}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{PIC-HIS-Apr2-A-5.pdf}
+  \label{fig:history enumeration 4 for PQ1Equal}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.4 \textwidth]{figures/PIC-HIS-PQ1Equal-5.pdf}
 %\vspace{-10pt}
   \caption{The fifth possible enumeration.}
-  \label{fig:history enumeration 5 for Rpr2}
-\end{figure}
-
-Given a data-differentiated history $h$, two actions $\textit{act}_1$, $\textit{act}_2$ of items of priority $\textit{pri}$, and assume that $\textit{act}_1$ is before $\textit{act}_2$ in $h$.
-We say that $\textit{act}_1$, $\textit{act}_2$ is covered by items $d_1,\ldots,d_m$ in $h$, if the priorities of $d_1,\ldots,d_m$ is not $\textit{pri}$, and
+  \label{fig:history enumeration 5 for PQ1Equal}
+\end{figure}
+
+Let us begin to represent several witness automata that is used to capture the existence of a data-differentiated execution $e$, $e$ has a projection $e'$, $\textit{last}(e') = \textit{PQ}_1^{=}$, there exists items $a$ and $b$ with maximal priority in $e'$, $a <_{\textit{pb}} b$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$.
+
+Given a data-differentiated execution $e$, two actions $\textit{act}_1$, $\textit{act}_2$ of maximal priority in $e$, and assume that $\textit{act}_1$ is before $\textit{act}_2$ in $e$.
+we say that $\textit{act}_1$, $\textit{act}_2$ is covered by items $d_1,\ldots,d_m$ in $e$, if the priorities of $d_1,\ldots,d_m$ is smaller than that of $\textit{act}_1$ and $\textit{act}_2$, and
 
@@
 
-According to the definition of gap-point, in \figurename~\ref{fig:history enumeration 1 for Rpr2}, \figurename~\ref{fig:history enumeration 2 for Rpr2}, \figurename~\ref{fig:history enumeration 3 for Rpr2}, \figurename~\ref{fig:history enumeration 4 for Rpr2} or \figurename~\ref{fig:history enumeration 5 for Rpr2}, the location of $o$ and $o'$ is equivalent to that, there exists items $d_1,\ldots,d_m$ in $h$ in $h$, such that $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ is covered by $d_1,\ldots,d_m$, $\textit{cal}(\textit{rm},b)$ is before $\textit{ret}(\textit{put},d_m,\_)$ in $h$, and $\textit{cal}(\textit{rm},d_1)$ is before $\textit{ret}(\textit{rm},a)$ in $h$.
-We say that such $d_1,\ldots,d_m$ constitute the rightmost gap of $\textit{put}(b,\_)$ and $\textit{rm}(b)$.
-
-An automaton $\mathcal{A}_{\textit{Rpr2-1}}$ is given in \figurename~\ref{fig:automata for first enumeration of Rpr2}, and it is constructed for the first enumeration in \figurename~\ref{fig:history enumeration 1 for Rpr2}.
-In \figurename~\ref{fig:automata for first enumeration of Rpr2}, we renames the items that covers $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ into $d$, and rename the renaming items into $e$.
-Here let $c = \textit{cal}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{cal}(\textit{rm},e), \textit{ret}(\textit{rm},e)$, $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_1.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr2-1}}$}
-  \label{fig:automata for first enumeration of Rpr2}
-\end{figure}
-
-An automaton $\mathcal{A}_{\textit{Rpr2-2}}$ is given in \figurename~\ref{fig:automata for second enumeration of Rpr2}, and it is constructed for the second enumeration in \figurename~\ref{fig:history enumeration 2 for Rpr2}.
-In \figurename~\ref{fig:automata for second enumeration of Rpr2}, $c_1$, $c_2$, $c_3$ and $c_4$ is same as that in \figurename~\ref{fig:automata for first enumeration of Rpr2}.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_2.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{Rpr2-2}}$}
-  \label{fig:automata for second enumeration of Rpr2}
-\end{figure}
-
-For the third enumeration in \figurename~\ref{fig:history enumeration 3 for Rpr2}.
+According to Lemma \ref{lemma:PQ1Equal as pb order and gap-point}, Lemma \ref{lemma:ob order has bounded length} and Lemma \ref{lemma:five enumeration is enough for PQ1Equal}, it is not hard to prove that, given a data-differentiated execution $e$ with $\textit{last}(e) = \textit{PQ}_1^{=}$, $e$ does not linearizable with respect to $\textit{MS}(\textit{PQ}_1^{=})$, if and only if, one of enumerations holds in $e$ (permit renaming), while $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ is covered by some $d_1,\ldots,d_m$, $\textit{cal}(\textit{rm},b)$ is before $\textit{ret}(\textit{put},d_m,\_)$, and $\textit{cal}(\textit{rm},d_1)$ is before $\textit{ret}(\textit{rm},a)$.
+We say that such $d_1,\ldots,d_m$ constitute the rightmost gap of $b$.
+
+An automaton $\mathcal{A}_{\textit{l-eq}}^1$ is given in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}, and it is constructed for the first enumeration in \figurename~\ref{fig:history enumeration 1 for PQ1Equal}.
+Here we rename the items that covers $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ into $d$, and rename the remanning items into $e$.
+In this figure, $c = \textit{cal}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{cal}(\textit{rm},e), \textit{ret}(\textit{rm},e)$, $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.8 \textwidth]{figures/PIC_AUTO_PQ1Equ-1.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{l-eq}}^1$}
+  \label{fig:automata for first enumeration of PQ1Equal}
+\end{figure}
+
+An automaton $\mathcal{A}_{\textit{l-eq}}^2$ is given in \figurename~\ref{fig:automata for second enumeration of PQ1Equal}, and it is constructed for the second enumeration in \figurename~\ref{fig:history enumeration 2 for PQ1Equal}.
+In \figurename~\ref{fig:automata for second enumeration of PQ1Equal}, $c_1$, $c_2$, $c_3$ and $c_4$ is same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.8 \textwidth]{figures/PIC_AUTO_PQ1Equ-2.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{l-eq}}^2$}
+  \label{fig:automata for second enumeration of PQ1Equal}
+\end{figure}
+
+For the third enumeration in \figurename~\ref{fig:history enumeration 3 for PQ1Equal}.
 Since we want to ensure that $a$ and $b$ are putted only once, we need to explicitly record the positions of $\textit{cal}(\textit{put},a,p)$ and $\textit{cal}(\textit{put},b,p)$.
@@
 
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$: There are three possible positions of $\textit{cal}(\textit{put},a,p)$: (1) after $\textit{cal}(\textit{put},b,p)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{cal}(\textit{put}b,p)$, and (3) before $\textit{call}(\textit{rm},a_1)$.
+\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$: There are three possible positions of $\textit{cal}(\textit{put},a,p)$: (1) after $\textit{cal}(\textit{put},b,p)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{cal}(\textit{put},b,p)$, and (3) before $\textit{call}(\textit{rm},a_1)$.
 
@@
 
-Therefore, there are fourteen possible cases that satisfy the third enumeration in \figurename~\ref{fig:history enumeration 3 for Rpr2}.
+Therefore, there are fourteen possible cases that satisfy the third enumeration in \figurename~\ref{fig:history enumeration 3 for PQ1Equal}.
 For each case, we construct an finite automaton.
-Let $\textit{Aut}_{\textit{Rpr2-3}}$ be the set of finite automata that is constructed for above fourteen cases.
-For example, for the case $\textit{ca}_1$ when $\textit{cal}(\textit{put},a,p)$ is before $\textit{cal}(\textit{rm},a_1)$, $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$, and $\textit{cal}(\textit{put},b,p)$ is before $\textit{cal}(\textit{rm},b)$, we construct a finite automaton in \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}.
-In \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}, let $c$ and $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$ the same as that in \figurename~\ref{fig:automata for first enumeration of Rpr2}.
+Let $\textit{Auts}_{\textit{1-eq}}^{3}$ be the set of finite automata that is constructed for above fourteen cases.
+For example, for the case $\textit{ca}_1$ when $\textit{cal}(\textit{put},a,p)$ is before $\textit{cal}(\textit{rm},a_1)$, $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$, and $\textit{cal}(\textit{put},b,p)$ is before $\textit{cal}(\textit{rm},b)$, we construct a finite automaton $\mathcal{A}_{\textit{l-eq}}^{\textit{3-1}}$ in \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}.
+In \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}, let $c$ and $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$ the same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}.
 Let $c_2 = c_1 + \textit{ret}(\textit{put},a_1)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$, $c_4 = c_3 + \textit{ret}(\textit{rm},d)$, and $c_5 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{put},a_1) + \textit{ret}(\textit{rm},d)$.
-The other finite automata in $\textit{Aut}_{\textit{Rpr2-3}}$ can be similarly constructed.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC_AUTO_Rpr2_3.pdf}
-%\vspace{-10pt}
-  \caption{Automaton for $\textit{ca}_1$}
+The other witness automata in $\textit{Auts}_{\textit{1-eq}}^{3}$ can be similarly constructed.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.8 \textwidth]{figures/PIC_AUTO_PQ1Equ-3-1.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{l-eq}}^{\textit{3-1}}$}
   \label{fig:automata for ca1 of third enumeration of Rpr2}
@@
 
-Similarly, we construct sets $\textit{Aut}_{\textit{Rpr2-4}}$ and $\textit{Aut}_{\textit{Rpr2-5}}$ of finite automata for the forth enumeration in \figurename~\ref{fig:history enumeration 4 for Rpr2} and the fifth enumeration in \figurename~\ref{fig:history enumeration 5 for Rpr2}, respectively.
-
-\begin{restatable}{lemma}{Rpr2IsCoRegular}
-\label{lemma:Rpr2 is co-regular}
-$R_{\textit{pr2}}$ is co-regular.
-\end{restatable}
-
-\begin {proof}
-
-Let $\textit{AutS-R}_{\textit{pr2}} = \{ \mathcal{A}_{\textit{Rpr2-1}}, \mathcal{A}_{\textit{Rpr2-2}} \} \cup \textit{Aut}_{\textit{Rpr2-3}} \cup \textit{Aut}_{\textit{Rpr2-4}} \cup \textit{Aut}_{\textit{Rpr2-5}}$.
-We need to prove that:
-
-\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr2}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, $\textit{last}(h') = R_{\textit{pr2}}$ and $h$ does not linearizable to $\textit{MR}_{\textit{pr2}}$.
-
-By Lemma \ref{lemma:Rpr2 as gap-point and ob order}, $\textit{fact}_1$ is equivalent to $\textit{fact}_2$:
-
-\noindent {\bf $\textit{fact}_2$}: Given a data-independent implementation $\mathcal{I}$, $\textit{AutS-R}_{\textit{pr0}} \cap \mathcal{I} \neq \emptyset$, if and only if there exists data-differentiated history $h \in \mathcal{I}$, $h'$ is a projection of $h$, such that $\textit{last}(h') = R_{\textit{pr2}}$, $a$ and $b$ has maximal priority in $h'$, $a <_{\textit{ob}} b$ in $h'$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
-
-\noindent The $\textit{only if}$ direction: Assume that $h_1 \in \mathcal{I}$ is accepted by $\textit{AutS-R}_{\textit{pr2}}$.
-By data-independence, there exists data-differentiated sequence $h_2 \in \mathcal{I}$ and a renaming function $r$, such that $h_1=r(h_2)$.
-Since $h_1$ is accepted by $\textit{AutS-R}_{\textit{pr2}}$, let $x$ and $y$ be the items that are renamed into $b$ and $a$ by $r$, respectively, and let $d_1,\ldots,d_m$ be the items that are renamed into $d$ by $r$.
-
-let $h'' = h_2 \vert_{ \{ x,y,d_1,\ldots,d_m \} }$.
-It is obvious that $\textit{last}(h'') = R_{\textit{pr2}}$.
-According to our construction of automata in $\textit{AutS-R}_{\textit{pr2}}$, it is not hard to see that $x$ and $y$ has maximal priority in $h_2$, $y <_{\textit{ob}} x$ in $h''$, and the rightmost gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$ in $h''$.
-%By Lemma \ref{lemma:Rpr2 as gap-point and ob order}, we know that $h''$ is not linearizable to $\textit{MR}_{\textit{pr2}}$.
-
-\noindent The $\textit{if}$ direction: Assume that there exists $h' \in \textit{proj}(h)$, such that $\textit{last}(h') = R_{\textit{pr2}}$, $a'$ and $b'$ has maximal priority in $h'$, $a' <_{\textit{ob}} b'$ in $h'$, and the rightmost gap-point of $\textit{put}(b',\_)$ and $\textit{rm}(b')$ is before $\textit{cal}(\textit{put},a',\_)$ or $\textit{cal}(\textit{rm},a')$ in $h''$.
-By data-independence, we can obtain history $h_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the items $d_1,\ldots,d_m$ that constitute the rightmost gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, we rename them into $d$, and (3) rename the other items into $e$.
-It is easy to see that $\textit{last}(h_1) = R_{\textit{pr2}}$, $a$ and $b$ has maximal priority in $h_1$, $a <_{\textit{ob}} b$ in $h_1$, and the rightmost gap-point of $\textit{put}(b,\_)$ and $\textit{rm}(b)$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $h''$.
-By Lemma \ref{lemma:five enumeration is enough for Rpr2}, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists) and $a_2$ (if exists).
+Similarly, we construct sets $\textit{Auts}_{\textit{1-eq}}^{4}$ and $\textit{Auts}_{\textit{1-eq}}^{5}$ of witness automata for the forth enumeration in \figurename~\ref{fig:history enumeration 4 for PQ1Equal} and the fifth enumeration in \figurename~\ref{fig:history enumeration 5 for PQ1Equal}, respectively.
+
+Let $\textit{Auts}_{\textit{1-eq}} = \{ \mathcal{A}_{\textit{l-eq}}^1, \mathcal{A}_{\textit{l-eq}}^2 \} \cup \textit{Auts}_{\textit{1-eq}}^{3} \cup \textit{Auts}_{\textit{1-eq}}^{4} \cup \textit{Auts}_{\textit{1-eq}}^{5}$.
+The following lemma states that $\textit{PQ}_1^{=}$ is co-regular.
+
+\PQOneEqualIsCoRegular*
+
+\begin {proof}
+
+We need to prove that, given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-eq}} \cap \mathcal{I} = \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), last(e')=\textit{PQ}_1^{=} \wedge e$ does not linearizable w.r.t.
+$\textit{MS}(\textit{PQ}_1^{=})$.
+
+By Lemma \ref{lemma:PQ1Equal as pb order and gap-point}, we need to prove the following fact:
+
+\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-eq}} \cap \mathcal{I} = \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), last(e')=\textit{PQ}_1^{=}$, $a$ and $b$ are two items with maximal priority in $e'$, $a <_{\textit{pb}}^* b$ in $e'$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $e'$.
+
+\noindent The $\textit{only if}$ direction: Assume that $e_1 \in \mathcal{I}$ is accepted by some witness automata in $\textit{Auts}_{\textit{1-eq}}$.
+By data-independence, there exists data-differentiated execution $e_2 \in \mathcal{I}$ and a renaming function $r$, such that $e_1=r(e_2)$.
+Since $e_1$ is accepted by some witness automata in  $\textit{Auts}_{\textit{1-eq}}$, let $x$, $y$ and $z$ (if exists) be the items that are renamed into $b$, $a$ and $a_1$ (if exists) by $r$, respectively, and let $d_1,\ldots,d_m$ be the items that are renamed into $d$ by $r$.
+
+let $e'' = e_2 \vert_{ \{ x,y,z,d_1,\ldots,d_m \} }$.
+It is obvious that $\textit{last}(e'') = \textit{PQ}_1^{=}$.
+According to our construction of automata in $\textit{Auts}_{\textit{1-eq}}$, it is not hard to see that $x$ and $y$ has maximal priority in $h_2$, $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$ in $e''$.
+
+\noindent The $\textit{if}$ direction: Assume that there exists $e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $last(e')=\textit{PQ}_1^{=}$, $a'$ and $b'$ are two items with maximal priority in $e'$, $a' <_{\textit{pb}}^* b'$ in $e'$, and the rightmost gap-point of $b'$ is before $\textit{cal}(\textit{put},a',\_)$ or $\textit{cal}(\textit{rm},a')$ in $e'$.
+By data-independence, we can obtain execution $e_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the items $d_1,\ldots,d_m$ that constitute the rightmost gap of $b'$, we rename them into $d$, (3) if $a' <_{\textit{pb}}^A a'_1 <_{\textit{pb}}^B b$, we rename $a'_1$ into $a_1$, and (4) rename the other items into $e$.
+It is easy to see that $\textit{last}(e_1) = \textit{PQ}_1^{=}$, $a$ and $b$ has maximal priority in $e_1$, $a <_{\textit{pb}}^* b$ in $e_1$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$ in $e_1$.
+By Lemma \ref{lemma:five enumeration is enough for PQ1Equal}, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists).
 Then
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] If $a <_{\textit{ob}} b$ because of the first enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr2-1}}$.
-
-\item[-] If $a <_{\textit{ob}} b$ because of the second enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{Rpr2-2}}$.
-
-\item[-] If $a <_{\textit{ob}} b$ because of the third enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-3}}$.
-
-\item[-] If $a <_{\textit{ob}} b$ because of the forth enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-4}}$.
-
-\item[-] If $a <_{\textit{ob}} b$ because of the fifth enumeration, it is easy to see that $h_1$ is accepted by some automaton in $\textit{Aut}_{\textit{Rpr2-5}}$.
+\item[-] If $a <_{\textit{pb}}^* b$ because of the first enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{l-eq}}^1$.
+
+\item[-] If $a <_{\textit{pb}}^* b$ because of the second enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{l-eq}}^2$.
+
+\item[-] If $a <_{\textit{pb}}^* b$ because of the third enumeration, it is easy to see that $h_1$ is accepted by some witness automaton in $\textit{Auts}_{\textit{1-eq}}^{3}$.
+
+\item[-] If $a <_{\textit{pb}}^* b$ because of the forth enumeration, it is easy to see that $h_1$ is accepted by some witness automaton in $\textit{Auts}_{\textit{1-eq}}^{4}$.
+
+\item[-] If $a <_{\textit{pb}}^* b$ because of the fifth enumeration, it is easy to see that $h_1$ is accepted by some witness automaton in $\textit{Auts}_{\textit{1-eq}}^{5}$.
 \end{itemize}

### 2017-03-04 co-regular-PQ.tex
I prove the second rule is co-regular. The proof can be found in Appendix C.3

@@
-\section{Co-Regular of Priority Queues}
+\section{Reducing Linearizability of Priority Queues into State Reachability}
 \label{sec:co-regular of priority queues}
@@
 
-A witness automaton is a finite automaton with alphabet $\{ \textit{cal}(\textit{put},d,\textit{pre}), \textit{ret}(\textit{put}d),$ $\textit{cal}(\textit{rm},d),\textit{ret}(\textit{rm},d) \vert d \in \mathbb{D},\textit{red} \in \textit{preWA} \}$.
-Here $\textit{preWA}$ is the set of predicate of priorities, and it contains
+A witness automaton is a finite automaton with alphabet $\{ \textit{cal}(\textit{put},d,\textit{pred}), \textit{ret}(\textit{put},d),$ $\textit{cal}(\textit{rm},d),\textit{ret}(\textit{rm},d) \vert d \in \mathbb{D},\textit{pred} \in \textit{predWA} \}$.
+Here $\textit{predWA}$ is the set of predicate of priorities, and it contains
 
@@
 
-\item[-] For each $i$, if $\alpha_i = \textit{ret}(\textit{put},a)$, or $\alpha_i = \textit{cal}(\textit{rm},a)$, or $\alpha_i = \textit{ret}(\textit{rm},a)$, then $\beta_i = \alpha_i$.
+\item[-] For each $i$, if $\alpha_i = \textit{ret}(\textit{put},a)$, $\textit{cal}(\textit{rm},a)$, $\textit{ret}(\textit{rm},a)$, $\textit{cal}(\textit{rm},\textit{empty})$ or $\textit{ret}(\textit{rm},\textit{empty})$, then $\beta_i = \alpha_i$.
 \end{itemize}
@@
 
-$$ \textit{Auts}_{R} \cap \mathcal{I} = \emptyset \Leftrightarrow \exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), last(e')=R \wedge e \ does \ not \ linearizable \ w.r.t. \ \textit{MS}(R)$$
+$$ \textit{Auts}_{R} \cap \mathcal{I} \neq \emptyset \Leftrightarrow \exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), last(e')=R \wedge e' \ does \ not \ linearizable \ w.r.t.
+\ \textit{MS}(R)$$
 
@@
 Given an execution of queue, we say that it is differentiated \cite{Wolper:1986}, if each item is enqueued at most once.
-\cite{Bouajjani:2015} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $a,b$: (1) $\textit{deq}(b) <_{hb} \textit{enq}(b)$, (2) there are are no $\textit{enq}(b)$ and least one $\textit{deq}(b)$, (3) there are are one $\textit{enq}(b)$ and more than one $\textit{deq}(b)$, and (4) $\textit{enq}(a) <_{\textit{hb}} \textit{enq}(b)$, and $\textit{deq}(b) <_{\textit{hb}} \textit{deq}(a)$, or $\textit{deq}(a)$ does not exists.
+\cite{Bouajjani:2015} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $a,b$: (1) $\textit{deq}(b) <_{hb} \textit{enq}(b)$, (2) there are no $\textit{enq}(b)$ and at least one $\textit{deq}(b)$, (3) there are one $\textit{enq}(b)$ and more than one $\textit{deq}(b)$, and (4) $\textit{enq}(a) <_{\textit{hb}} \textit{enq}(b)$, and $\textit{deq}(b) <_{\textit{hb}} \textit{deq}(a)$, or $\textit{deq}(a)$ does not exists.
 For each such case, we can construct a witness automata for priority queue.
-For example, for the first case, we generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ in \figurename~\ref{fig:automata for FIFO-1}, here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}), \textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a)$, $\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},b)$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$.
+For example, for the first case, we generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ in \figurename~\ref{fig:automata for FIFO-1}, here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}), \textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a)$, $\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},b),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$.
 
@@
 In \figurename~\ref{fig:introduce gap for PQ1Lar}, we draw the interval of each item by dashed line.
-The reason of why $e$ does not linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{>})$ is that, each time point from $\textit{cal}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is in interval of some item with less priority than priority of $a$.
+The reason of why $e$ does not linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{>})$ is that, each time point from $\textit{cal}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is in interval of some item with smaller priority than priority of $a$.
 
@@
 Let $\textit{UVSet}(e,x) = \textit{UVSet}_1(e,x) \cup \textit{UVSet}_2(e,x) \cup \ldots$.
-For example, in \figurename~\ref{fig:his nobound of LMSet}, $\textit{UVSet}_1(e,x) = \{ \textit{put}(a,1),\textit{rm}(a) \}$, $\textit{USSet}_2(e,x) = \{ \textit{put}(b,1),\textit{rm}(b) \}$ and $\textit{UVSet}_3(e,x) = \{ \textit{put}(a,c),\textit{rm}(c) \}$.
+For example, in \figurename~\ref{fig:his nobound of LMSet}, $\textit{UVSet}_1(e,x) = \{ \textit{put}(a,1),\textit{rm}(a) \}$, $\textit{USSet}_2(e,x) = \{ \textit{put}(b,1),\textit{rm}(b) \}$ and $\textit{UVSet}_3(e,x) = \{ \textit{put}(c,1),\textit{rm}(c) \}$.
 Similarly, we can generate execution $e'$ such that, for each $i$, $\textit{UVSet}(e',x) \neq \emptyset$.
@@
 \end{restatable}
+
+\subsection{Co-Regular of $\textit{PQ}_1^{=}$}
+\label{subsec:co-regular of PQ1Equal}
+
+In this subsection, we introduce the idea for proving co-regular of $\textit{PQ}_1^{=}$.
+The proof of this subsection can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of PQ1Equal}.
+
+Given a data-differentiated execution $e$ such that $\textit{last}(e) = \textit{PQ}_1^{=}$.
+Lemma \ref{lemma:automata for priority queue with single priority} ensures that each single-priority projection of $e$ satisfy the FIFO property, and with the help of Lemma \ref{lemma:PQ1Lar is co-regular} and $\textit{Auts}_{\textit{1-lar}}$, we can ensure that each item in $e$ can not be covered by items with smaller priority.
+However, these are sill not enough for ensure that $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
+This is because that there are multiple matched $\textit{put}$ (assume they use items in $S$), and all the possible linearization point of $\textit{rm}(b)$ may be disabled by $\textit{rm}(b)$ for some $a,b \in S$.
+
+We give an example of such execution $e$ in \figurename~\ref{fig:introduce pb order}.
+The execution $e$ of \figurename~\ref{fig:introduce pb order} is not linearizable, even if $h \vert_{1}$ and $h \vert_{4}$ are both linearizable, and either $a$ or $b$ is covered by other items.
+To explain this, me need to observe the following three points:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Since $\textit{put}(a,1)$ happens before $\textit{put}(b,1)$, we know that $a$ is ``putted earlier'' than $b$, and therefore, the linearization points of $\textit{rm}(a)$ should before the linearization point of $\textit{rm}(b)$.
+
+\item[-] The time intervals satisfy the following conditions (we identify them with dotted lines in \figurename~\ref{fig:introduce pb order}) are possible position to locate the linearization point of $\textit{rm}(b)$ according to Lemma \ref{lemma:Lin Equals Constraint for PQ1Lar} (Here we temporarily forget the existence of $a$): (1) between from $\textit{cal}(\textit{rm},b)$ and $\textit{ret}(\textit{rm},b)$, (2) does not in interval of any item with smaller priority.
+
+\item[-] To satisfy requirement of the second condition, the linearization points of $\textit{rm}(b)$ should be in the time interval of dotted line.
+However, all time point in dotted line is before $\textit{cal}(\textit{rm},a)$ and is thus disabled by $\textit{cal}(\textit{rm},a)$.
+\end{itemize}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.6 \textwidth]{figures/PIC-HIS-INTRO-PB-ORDER.pdf}
+%\vspace{-10pt}
+  \caption{An execution that does not linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{=})$}
+  \label{fig:introduce pb order}
+\end{figure}
+
+In this subsection, intuitively we will prove that, as long as we can get rid of the case in \figurename~\ref{fig:introduce pb order}, we can ensure $\textit{last}(e) = \textit{PQ}_1^{=} \Rightarrow e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
+
+Let us introduce $<_{\textit{pb}}$ (represents put-before) order to formally state that ``an item is putted before another''.
+Given a data-differentiated execution $e$ and two items $a,b$ with maximal priority in $e$,  we say that $a <_{\textit{pb}} b$, if one of the following cases holds:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$ in $e$,
+
+\item[-] $\textit{rm}(a)$ happens before $\textit{rm}(b)$ in $e$,
+
+\item[-] $\textit{rm}(a)$ happens before $\textit{put}(b,\_)$ in $e$,
+\end{itemize}
+
+Sometimes we use $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$ and $a <_{\textit{pb}}^C b$ to explicitly distinguish above three cases.
+Let $<_{\textit{pb}}^*$ be the transitive closure of $<_{\textit{pb}}$.
+Intuitively, $a <_{\textit{pb}}^* b$ means that we can infer that $a$ should be inserted earlier than $b$ with the help of possibly other items.
+
+Let us introduce gap-point to formally define the time point in dotted line of \figurename~\ref{fig:introduce pb order}.
+
+\begin{definition}\label{def:gap-point for matched put and rm operations}
+Given a data-differentiated execution $e$ and two method events $\textit{put}(x,\textit{pri}_x),\textit{rm}(x)$ of $e$, where $\textit{pri}_x$ is the maximal priority in $e$.
+We say that a time-point $o$ is a gap-point of $x$, if
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $o$ is after $\textit{call}(\textit{put},x,\textit{pri}_x)$ and $\textit{call}(\textit{rm},x)$, and is before $\textit{ret}(\textit{rm},x)$.
+
+\item[-] $o$ is not in interval of any items with smaller priority.
+\end{itemize}
+\end{definition}
+
+Going back to \figurename~\ref{fig:introduce pb order}, we can effectively characterize it using $<_{\textit{pb}}$ order and gap-point as follows: $a <_{\textit{pb}} b$, while the right-most gap-point of $b$ is before $\textit{cal}(\textit{rm},a)$.
+We need to prove that getting rid of such case is enough fo ensure linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{=})$, as stated by the following lemma.
+Let $\textit{Items}(e,p)$ be the set of items with priority $p$ in execution $e$.
+
+\begin{restatable}{lemma}{PQOneEqualAsPBandGP}
+\label{lemma:PQ1Equal as pb order and gap-point}
+Given a data-differentiated execution $e$  with $\textit{last}(e) = \textit{PQ}_1^{=}$.
+$e$ does not linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{=})$, if and only if there exists $x$ and $y$ with maximal priority in $e$, such that $y <_{\textit{pb}}^* x$ in $e$, and the rightmost gap-point of $x$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$ in $e$.
+\end{restatable}
+
+\begin {proof} (Sketch)
+
+We have already intuitively explain the proof of the $\textit{if}$ direction.
+
+To prove the $\textit{only if}$ direction, we prove its contrapositive.
+Assume we already know that for each $x$ and $y$ has maximal priority in $e$, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},x)$.
+We need to prove that $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
+Or we can say, we need to explicitly construct linearization of $e$.
+Let $\textit{pri}$ be the maximal priority of $e$.
+
+We introduce another lemma (Lemma \ref{lemma:maximal in pb and gap-point make a candidate of PQ1Equal}), which states that: If $\exists g_1 \in \textit{Items}(e,\textit{pri})$, such that $\forall g_2 \in \textit{Items}(e,\textit{pri})$, (1) $g_1$ does not $<_{\textit{pb}}$ to $g_2$, and (2) the right-most gap-point of $g_1$ is after $\textit{cal}(\textit{put},g_2,\_)$ and $\textit{cal}(\textit{rm},g_2)$.
+Then $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
+The proof of this lemma also tell us how to construct linearization in such case.
+
+Since every single-priority projection has FIFO property, let $l_{\textit{pri}}$ be the linearization of projection of $e$ into actions of priority $\textit{pri}$. Then our proof proceeds as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] We start with $a_1$, the last inserted item of $l_{\textit{pri}}$.
+
+\item[-] Step $1$: check if $a_1$ satisfy the conditions of \ref{lemma:maximal in pb and gap-point make a candidate of PQ1Equal}.
+
+    It is obvious that $a_1$ satisfy the first condition.
+If the second condition is also satisfied, by Lemma \ref{lemma:maximal in pb and gap-point make a candidate of PQ1Equal}, we can obtain that $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
+
+\item[-] Otherwise, $\exists a_2 \in \textit{Items}(e,\textit{pri})$, such that the rightmost gap-point of $a_1$ is before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$ in $e$.
+We can see that each gap-point of $a_2$ is after the rightmost gap-point of $a_1$.
+By assumption, we know that $a_2$ does not $<_{\textit{pb}}$ to $a_1$.
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $\forall b \in \textit{Items}(e,\textit{pri})$, $a_2$ does not $<_{\textit{pb}}$ to $b$. Then we go to step $1$ and treat $a_2$ similarly as $a_1$.
+    \item[-] Otherwise, there exists $a_3$ with priority $\textit{pri}$ such that $a_2 <_{\textit{pb}}^* a_3$.
+
+    Since $l_{\textit{pri}}$ has FIFO property, it is easy to see that there is no cycle in $<_{\textit{pb}}$ order.
+It is safe to assume that $a_3$ is maximal in the sense of $<_{\textit{pb}}^*$.
+Or we can say, there does not exists $a_4$, such that $a_3 <_{\textit{pb}}^* a_4$.
+
+    By assumption,we know that the rightmost gap-point of $a_3$ is after $\textit{cal}(\textit{put},a_2,\textit{pri})$ and $\textit{cal}(\textit{rm},a_2)$.
+Therefore, we can see that the rightmost gap-point of $a_3$ is after the rightmost gap-point of $a_1$.
+Then we go to step $1$ and treat $a_3$ similarly as $a_1$.
+    \end{itemize}
+\end{itemize}
+
+Let $a^i$ be the $a_1$ in the $\textit{i-th}$ loop of our proof.
+It is not hard to see that, given $i<j$, the rightmost gap-point of $a^j$ is after the rightmost gap-point of $a^i$.
+Therefore, the loop finally stop at some $a^f$.
+$a^f$ satisfies the check of Step $1$.
+By Lemma \ref{lemma:maximal in pb and gap-point make a candidate of PQ1Equal}, this implies that $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
+This completes the proof of $\textit{if}$ direction.
+\qed
+\end {proof}
+
+The result of Lemma \ref{lemma:PQ1Equal as pb order and gap-point} is not quite suitable for verificaion with automata.
+The reason is that we need to ensure $y <_{\textit{pb}}^* x$, while according to definition of $<_{\textit{pb}}^*$, there may be arbitrary intermediate items $a_1,\ldots,a_m$, such that $a <_{\textit{pb}} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}} a_m <_{\textit{pb}} b$.
+It is hard to store unbounded $a_i$ by automata.
+Fortunately, we find that the number of intermediate items $a_i$ is in fact bounded, as stated by the following lemma.
+
+\begin{restatable}{lemma}{OBOrderHasBoundedLength}
+\label{lemma:ob order has bounded length}
+Given a data-differentiated execution $h$. Assume that $a <_{\textit{pb}} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}} a_m <_{\textit{pb}} b$, then one of the following cases holds:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$ or $a <_{\textit{pb}}^C b$,
+
+\item[-] $a <_{\textit{pb}}^A a_i <_{\textit{pb}}^B b$, or $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A b$, for some $i$.
+%\item[-] $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A a_j <_{\textit{pb}}^B b$, for some $i$ and $j$,
+\end{itemize}
+\end{restatable}
+
+With Lemma \ref{lemma:ob order has bounded length}, now we can use witness automata to detect $a <_{\textit{pb}}^* b$ by enumerating all possible enumerations of $a$, $b$ and an intermediate item $a_1$, and then check if it satisfy one cases of \ref{lemma:ob order has bounded length}.
+
+The number of potential enumerations can be further reduced.
+The reason is that Lemma \ref{lemma:PQ1Equal as pb order and gap-point} requires both $<_{\textit{pb}}^*$ and gap-point, while some combination of $<_{\textit{pb}}^*$ and gap-points are conflict.
+For example, if $a <_{\textit{pb}}^B b$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$, then since (1) $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(b)$, (2) each gap-point of $b$ is after $\textit{cal}(\textit{rm},b)$ (also $\textit{ret}(\textit{rm},a)$), we can see that the right-most gap-point of $b$ is before $\textit{cal}(\textit{put},a,\_)$ and after $\textit{ret}(\textit{rm},a)$, which implies $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a,\_)$ and is impossible.
+Therefore, we finally reduce the number of potential enumeration into only five, as shown by the following lemma:
+
+\begin{restatable}{lemma}{FiveEnmuerationisEnoughForPQOneEqual}
+\label{lemma:five enumeration is enough for PQ1Equal}
+Given a data-differentiated execution $e$ with $\textit{last}(e) = \textit{PQ}_1^{=}$.
+Let $a$ and $b$ be items with maximal priority in $e$.
+Assume that $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$.
+Then, there are five possible enumeration of method events of $a$, $b$, $a_1$ (if exists), where $a_1$ is the possible intermediate items for obtain $a <_{\textit{pb}}^* b$.
+\end{restatable}
+
+The five enumerations is shown in \figurename~\ref{fig:five enumerations}, where we use $o$ to explicitly denote the right-most gap-point of $b$.
+In this figure, $e_1$ and $e_2$ comes from $a <_{\textit{pb}}^A b$, $e_3,e_4,e_5$ comes from $a <_{\textit{pb}}^A a_1 <_{\textit{pb}}^B b$, while other reasons of $a <_{\textit{pb}}^* b$ turns out to be either redundant or impossible.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=1 \textwidth]{figures/PIC-HIS-FiveEnumerations.pdf}
+%\vspace{-10pt}
+  \caption{Five possible enumerations}
+  \label{fig:five enumerations}
+\end{figure}
+
+For each enumeration in \figurename~\ref{fig:five enumerations}, we can generate several automata to capture it.
+For example, to capture the first enumeration, we generate witness automaton $\mathcal{A}_{\textit{l-eq}}^1$, as shown in \figurename~\ref{fig:automata for first enumeration of PQ1Equal not in appendix}.
+Here we rename the items that ``covers the time interval from $\textit{cal}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$'' into $d$ (see Appendix \ref{sec:appendix proof and definition in section co-regular of PQ1Equal}), and rename the renaming items into $e$.
+In this figure, $c = \textit{cal}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{cal}(\textit{rm},e), \textit{ret}(\textit{rm},e)$, $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.8 \textwidth]{figures/PIC_AUTO_PQ1Equ-1.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{l-eq}}^1$}
+  \label{fig:automata for first enumeration of PQ1Equal not in appendix}
+\end{figure}
+
+In Appendix \ref{sec:appendix proof and definition in section co-regular of PQ1Equal}, we construct a set $\textit{Auts}_{\textit{1-eq}}$ of witness automata ($\mathcal{A}_{\textit{l-eq}}^1$ is in it), and use $\textit{Auts}_{\textit{1-eq}}$ to prove that $\textit{PQ}_1^{=}$ is co-regular, as stated by the following lemma.
+
+\begin{restatable}{lemma}{PQOneEqualIsCoRegular}
+\label{lemma:Rpr2 is co-regular}
+$\textit{PQ}_1^{=}$ is co-regular.
+\end{restatable}

### 2017-03-04 data-independence-PQ.tex
I prove the second rule is co-regular. The proof can be found in Appendix C.3

@@
 Let $\_$ denote an element, of which the value is irrelevant.
-A sequential execution $e$ of priority is said to be data-differentiated if, for all $d \in \mathbb{D}$, there is at most one method event $\textit{put}(d,\_)$ in $e$.
+A sequential execution $e$ of priority queue is said to be data-differentiated if, for all $d \in \mathbb{D}$, there is at most one method event $\textit{put}(d,\_)$ in $e$.
 Note that a data-differentiated sequential execution $e$ may contains more than one items with a same priority.

### 2017-03-04 rules-PQ.tex
slightly modify a rule
I prove the second rule is co-regular. The proof can be found in Appendix C.3

@@
 \setlength{\itemsep}{0.5pt}
-\item[-] $\textit{PQ}_2^{>} \equiv (u \cdot v \in \textit{PQueue}) \wedge (\textit{rm}(\textit{empty}) \notin u \cdot v) \wedge (\textit{pri} \geq \textit{maxPri}(u \cdot v)) \wedge (\textit{pri} > \textit{maxMPri}(u \cdot v)) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{PQueue})$.
+\item[-] $\textit{PQ}_2^{>} \equiv (u \cdot v \in \textit{PQueue}) \wedge (\textit{rm}(\textit{empty}) \notin u \cdot v) \wedge (\textit{pri} \geq \textit{maxPri}(u \cdot v)) \wedge (\textit{pri} > \textit{maxMPri}(u \cdot v)) \wedge (\textit{putInSeq}(u \cdot v,u,\textit{pri})) \Rightarrow (u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{PQueue})$.
 
@@
 Given a sequential execution $w \in \textit{PQueue}$ and let $\{p_1,\ldots,p_k\}$ be the set of priorities in $w$, and assume that $\forall i<j$, we have $p_i < p_j$.
-According to above example, $w$ is generated from $\epsilon$ by work with the following order: (1) Adding pairs of matched $\textit{put}$ and $\textit{rm}$ with priority $p_1$ by using $\textit{PQ}_1^{>}$ and then several times of $\textit{PQ}_1^{=}$, (2) adding unmatched $\textit{put}$ with priority $p_1$ by using $\textit{PQ}_2^{>}$ and then several times of $\textit{PQ}_2^{=}$, (3) Similarly work with $p_2,\ldots,p_k$, and (4) adding $\textit{rm}(\textit{empty})$.
+According to above example, $w$ is generated from $\epsilon$ by work with the following order: {\color {red}(1) If there are matched $\textit{put}$ and $\textit{rm}$ with priority $p_1$ in $w$: Adding pairs of matched $\textit{put}$ and $\textit{rm}$ with priority $p_1$ by using one time of $\textit{PQ}_1^{>}$ and then (possibly) several times of $\textit{PQ}_1^{=}$, and then (possibly) adding unmatched $\textit{put}$ with priority $p_1$ by using $\textit{PQ}_2^{=}$, (2) If there are no matched $\textit{put}$ and $\textit{rm}$ with priority $p_1$ in $w$: Adding unmatched $\textit{put}$ with priority $p_1$ by using $\textit{PQ}_2^{>}$ for several times, (3) Similarly work with $p_2,\ldots,p_k$, and (4) adding $\textit{rm}(\textit{empty})$.}
 

### 2017-03-04 ss-lin-PQ.tex
I prove the second rule is co-regular. The proof can be found in Appendix C.3

@@
 
-A set $S$ of sequential executions is closed under projection, if for all $\mathcal{D} \subseteq \mathbb{D}$ and $e \in S$, we have $e \vert_{ \mathcal{D} } \in S$..
+A set $S$ of sequential executions is closed under projection, if for all $\mathcal{D} \subseteq \mathbb{D}$ and $e \in S$, we have $e \vert_{ \mathcal{D} } \in S$.
 The following lemma states that $\textit{PQueue}$ is closed under projection, since the predicates used in rules of priority queue are ``closed under projection''.
@@
 The following lemma states that for data-differentiated execution, checking linearizability with respect to $\textit{PQueue}$ is equivalent to checking linearizability with respect to $\textit{MS}(R)$ for everyone of its projections.
-Roughly speaking, step-by-step linearizability of priority queue paly a important rule for proof of the $\textit{if}$ direction of Lemma \ref{lemma:PQ as multi in MRpri for history}: It guild us how to build a linearization of a whole execution by increasingly construct linearization of sub-execution from a $\epsilon$ execution.
+Roughly speaking, step-by-step linearizability of priority queue play a important rule for proof of the $\textit{if}$ direction of Lemma \ref{lemma:PQ as multi in MRpri for history}: It guide us how to build a linearization of a whole execution by increasingly construct linearization of sub-execution from a $\epsilon$ execution.
 

### 2017-03-07 co-regular-PQ.tex
I prove the co-regular of five rules

@@
 
-We say that $\textit{PQueue}$ is regular, if each of its rule is co-regular.
+We say that $\textit{PQueue}$ is co-regular, if each of its rule is co-regular.
 \end{definition}
@@
 \end{restatable}
+
+\subsection{Combine Step-by-Step linearizability and Co-Regular}
+\label{subsec:combine step-by-step linearizability and co-regular}
+
+We also prove that $\textit{PQ}_2^{>}$, $\textit{PQ}_2^{=}$ and $\textit{PQ}_3$ are co-regular.
+The case of $\textit{PQ}_2^{>}$ is quite simple, since whenever $\textit{last}(e) = \textit{PQ}_2^{>}$, $e \sqsubseteq \textit{MS}(\textit{PQ}_2^{>})$.
+For the case of $\textit{PQ}_2^{=}$, we find a necessary and sufficient condition for violating $\textit{MS}(\textit{PQ}_2^{>})$, and shows that such condition violates the assumptions that every single-priority execution is FIFO and can be safely ignored.
+For the case of $\textit{PQ}_3$, it can be similarly proved as the rule for $\textit{deq}(\textit{empty})$ in \cite{Bouajjani:2015}.
+We leave the detailed proof of co-regular of these three rules in Appendix \ref{subsec:appendix co-regular of PQ2Lar}, Appendix \ref{subsec:appendix co-regular of PQ2Equal} and Appendix \ref{subsec:co-regular of PQ3}, respectively.
+
+Now we can see that $\textit{PQueue}$ is co-regular, which is a direct consequence of co-regular of each of its rules, and is stated below.
+
+\begin{restatable}{lemma}{PQueueIsCoRegular}
+\label{lemma:PQueue is co-regular}
+$\textit{PQueue}$ is co-regular.
+\end{restatable}
+
+Since we have already prove that $\textit{PQueue}$ is step-by-step linearizability and co-regular, we can now reduce the verification of linearizability with respect to $\textit{PQueue}$ into a reachability problem, as illustrated by the following theorem.
+Here $\textit{Auts}$ is the set of witness automata of this section.
+
+\begin{restatable}{theorem}{ReducePQueueIntoStateReachability}
+\label{lemma:reduce PQueue into state reachability}
+Given a data-independence implementation $\mathcal{I}$. $\mathcal{I} \sqsubseteq \textit{PQueue}$, if and only if, $\mathcal{I} \cap \textit{Auts} = \emptyset$.
+\end{restatable}

### 2017-03-07 relate-PQ-with-Queue-Stack.tex
I prove the co-regular of five rules

\section{Reduce Queue and Stack Executions into Priority Queue Executions}
\label{sec:reduce queue and stack executions into priority queue executions}

In this section, we shows how to reduce queue and stack executions into priority queue executions in polynomial time.

\subsection{Reduce Queue Executions into Priority Queue Executions}
\label{subsec:reduce queue executions into priority queue executions}

Since we have already guarantee that each single-priority execution of priority queue has FIFO property, to reduce queue executions into priority queue executions, we simply give every item a same priority. Given a execution $e_q$ of queue, let $\textit{TransQ}(e_q)$ be an execution of priority queue, which is generated as follows: for each $a$, transform $\textit{call}(\textit{enq},a)$, $\textit{ret}(\textit{enq},a)$, $\textit{call}(\textit{deq},a)$ and $\textit{ret}(\textit{deq},a)$ of $e_q$ into $\textit{call}(\textit{put},a,1)$, $\textit{ret}(\textit{put},a)$, $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$. The following lemma is obvious.

\begin{restatable}{lemma}{RelateQueuewithPQ}
\label{lemma:relate queue with priority queue}
Given an execution $e_q$ of queue, $e_q$ is linearizable w.r.t queue, if and only if $\textit{TransQ}(e_q) \sqsubseteq \textit{PQueue}$.
\end{restatable}

### 2017-04-04 appendix-EPQ.tex
add tex

\section{Proof in Section \ref{sec:inductive rules of extended priority queue}}
\label{sec:appendix in section inductive rules of extended priority queue}

A $\textit{labelled transition system}$ ($LTS$) is a tuple $\mathcal{A}=(Q,\Sigma,\rightarrow,q_0)$, where $Q$ is a set of states, $\Sigma$ is an alphabet of transition labels, $\rightarrow\subseteq Q\times\Sigma\times Q$ is a transition relation and $q_0$ is the initial state.

Let us model extended priority queue as an LTS $\textit{LTS}_e = (Q,\Sigma,\rightarrow,q_0)$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Each state of $Q$ is a function from $\mathbb{P}$ into a finite sequence over $\mathbb{D}$.

\item[-] The initial state $q_0$ is a function that maps each element in $\mathbb{P}$ into $\epsilon$.

\item[-] $\Sigma = \{ \textit{put}(a,p),\textit{rm}(a),\textit{rm}(\textit{empty}) \vert a \in \mathbb{D}, p \in \mathbb{P} \}$.

\item[-] The transition relation $\rightarrow$ is defined as follows:

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] $q_1 \xrightarrow{\textit{put}(a,p)} q_2$, if $q_1$ maps $p$ into some finite sequence $l$, and $q_2$ is the same as $q_1$, except for $p$, where it maps $p$ into $a \cdot l$.

    \item[-] $q_1 \xrightarrow{\textit{rm}(a)} q_2$, if $q_1$ maps $p$ into $l \cdot a$ for some finite sequence $l$, and $q_2$ is the same as $q_1$, except for $p$, where it maps $p$ into $l$. We also require that for each priority $p'$ such that $p' <_{\mathbb{P}} p$, $q_1$ and $q_2$ map $p'$ into $\epsilon$.

    \item[-] $q_1 \xrightarrow{\textit{rm}(\textit{empty})} q_2$, if $q_1 = q_2$, and they maps each element in $\mathbb{P}$ into $\epsilon$.
    \end{itemize}
\end{itemize}

A path of an LTS is a finite transition sequence $q_0\xrightarrow{\beta_1}q_1\overset{\beta_2}{\longrightarrow}\ldots\overset{\beta_k}{\longrightarrow}q_k$ for $k\geq 0$, where $q_0$ is the initial state of the LTS. A trace of an LTS is a finite sequence $\beta_1 \cdot \beta_2 \cdot \ldots \cdot \beta_k$, where $k \geq 0$ if there exists a path $q_0\overset{\beta_1}{\longrightarrow}q_1\overset{\beta_2}{\longrightarrow}\ldots\overset{\beta_k}{\longrightarrow}q_k$ of the LTS. Let $\textit{EPQ}_s$ be the set of traces of $\textit{LTS}_e$. The following lemma states that the language generated by our rules equals the set of traces of $\textit{EPQ}_s$.

\EPQRulesAndSemantics*

\begin {proof}

To prove that $\textit{EPQ} \subseteq \textit{EPQ}_s$, we prove that each sequence in $\textit{EPQ}$ is also in $\textit{EPQ}_s$ by induction:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] It is obvious that $\epsilon \in \textit{EPQ}_s$.

\item[-] If $l_1 \in \textit{EPQ}_s$ and $l_1 \xrightarrow{\textit{EPQ}_1} l_2$. Then we need to prove that $l_2 \in \textit{EPQ}_s$. Let $l_1 = u \cdot v \cdot w$, such that $\textit{Guard}(u,v,w,\textit{item},\textit{pri})$ of $\textit{EPQ}_1$ holds, and $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm})$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$, $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_j$ and $w = \alpha_{\textit{j+1}} \cdot \ldots \cdot \alpha_m$. Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_j} q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{EPQ}_s$. For each $i \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $\textit{pri}$ into $\textit{item} \cdot l_k$ and $q_k$ maps $\textit{pri}$ into $l_k$ for some finite sequence $l_k$.

    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(\textit{itm},\textit{pri})} q'_i$. Since (1) all $\textit{put}$ with priority $\textit{pri}$ is in $u$, and (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $\textit{pri}$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq j$) with a newest $\textit{itm}$ with priority $\textit{pri}$. Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j$ are transitions of $\textit{LTS}_e$. Since $\textit{matched-C}(u \cdot v)$ holds, we can see that $q_j$ maps each priority that is smaller than $\textit{pri}$ into $\epsilon$ and maps $\textit{pri}$ into $\epsilon$, and $q'_j$ maps each priority that is smaller than $\textit{pri}$ into $\epsilon$ and maps $\textit{pri}$ into $\textit{itm}$. Then, we can see that $q'_j \xrightarrow{\textit{rm}(\textit{itm})} q_j$. We already know that that $q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$. Therefore, we can see that $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \in \textit{EPQ}_s$.

\item[-] If $l_1 \in \textit{EPQ}_s$ and $l_1 \xrightarrow{\textit{EPQ}_2} l_2$. Then we need to prove that $l_2 \in \textit{EPQ}_s$. Let $l_1 = u \cdot v$, such that $\textit{Guard}(u,v,\textit{item},\textit{pri})$ of $\textit{EPQ}_2$ holds, and $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$. Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{EPQ}_s$. For each $i \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $\textit{pri}$ into $\textit{item} \cdot l_k$ and $q_k$ maps $\textit{pri}$ into $l_k$ for some finite sequence $l_k$.

    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that$q_i \xrightarrow{\textit{put}(\textit{itm},\textit{pri})} q'_i$. Since (1) all $\textit{put}$ with priority $\textit{pri}$ is in $u$, (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $\textit{pri}$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq m$) with a newest $\textit{itm}$ with priority $\textit{pri}$. Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q'_m$ are transitions of $\textit{LTS}_e$. Therefore, we can see that $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{EPQ}_s$.

\item[-] If $l_1 \in \textit{EPQ}_s$ and $l_1 \xrightarrow{\textit{EPQ}_3} l_2$. Then we need to prove that $l_2 \in \textit{EPQ}_s$. Let $l_1 = u \cdot v$, such that $\textit{Guard}(u,v)$ of $\textit{EPQ}_3$ holds, and $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$. Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{EPQ}_s$.

    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$. Since $\textit{matched-All}(u)$ holds, we can see that $q_i$ maps each element in $\mathbb{P}$ into $\epsilon$, and then $q_i \xrightarrow{\textit{rm}(\textit{empty})} q_i$. We already know that $q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q_m$. Therefore, we can see that $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v \in \textit{EPQ}_s$.
\end{itemize}

To prove that $\textit{EPQ}_s \subseteq \textit{EPQ}$, we show that given $l_2 \in \textit{EPQ}_s$ and $l_2 \neq \epsilon$, how to construct a sequence $l_1$, such that $l_1 \xrightarrow{R} l_2$ for some rule $R$, and $l_1 \in \textit{EPQ}_s$. Based on this, we can decompose a sequence of $\textit{EPQ}_s$ into $\epsilon$, while the reverse process is the reason of why this sequence is in $\textit{EPQ}$. Note that from a $l_2$ we may construct more than one $l_1$, and this does not influence the correctness of our proof.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $l_2$ contains $\textit{rm}(\textit{empty})$: Assume that $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v$. It is easy to see that $\textit{matched-All}(u)$ holds. Let $l_1 = u \cdot v$. It is easy to see that $l_1 \xrightarrow{\textit{EPQ}_3} l_2$, and $l_1$ satisfy the guard of $\textit{EPQ}_3$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$. Since We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{rm}(\textit{empty})} q'_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is transitions of $\textit{LTS}_e$. Since $\textit{matched-All}(u)$ holds, we can see that $q_i = q'_i$, and they map each element in $\mathbb{P}$ into $\epsilon$. Then we can see that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is transitions of $\textit{LTS}_e$, and $l_1 \in \textit{EPQ}_s$.

\item[-] If $l_2$ does not contain $\textit{rm}(\textit{empty})$, $\textit{pri}$ is one of maximal priority of $l_2$, and items in $l_2$ with priority $\textit{pri}$ are unmatched $\textit{put}$ and (possibly) matched $\textit{put}$:

    Assume that $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v$, such that all $\textit{put}$ with priority $\textit{pri}$ of $u \cdot v$ is in $u$. Let $l_1 = u \cdot v$. According to construction of $\textit{LTS}_e$, we can see that $l_1$ satisfies the guard of $\textit{EPQ}_2$, and $l_1 \xrightarrow{\textit{EPQ}_2} l_2$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$. We already know that $\textit{pa} = q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{put}(\textit{itm},\textit{pri})} q_{\textit{i+}} \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q_m$ are transitions of $\textit{LTS}_e$. For each $\textit{i+1} \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $\textit{pri}$ into some $\textit{itm} \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $\textit{pri}$ into $l_k$. Since (1) all $\textit{put}$ with priority $\textit{pri}$ of $u \cdot v$ is in $u$ and (2) $\textit{pri}$ is one of maximal priority of $l_2$, it is safe to remove $\textit{itm}$ without influence other transitions of $\textit{pa}$. Or we can say, $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q'_m$ are transitions of $\textit{LTS}_e$. Therefore, $l_1 \in \textit{EPQ}_s$.

\item[-] If $l_2$ does not contain $\textit{rm}(\textit{empty})$, $\textit{pri}$ is one of maximal priority of $l_2$, and items in $l_2$ with priority $\textit{pri}$ are matched $\textit{put}$:

    Assume that $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w$, such that all $\textit{put}$ with priority $\textit{pri}$ of $u \cdot v \cdot w$ is in $u$. Let $l_1 = u \cdot v \cdot w$. According to construction of $\textit{LTS}_e$, we can see that $l_1$ satisfies the guard of $\textit{EPQ}_1$. We can also see that $l_1 \xrightarrow{\textit{EPQ}_1} l_2$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$, $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_j$ and $w = \alpha_{\textit{j+1}} \cdot \ldots \cdot \alpha_m$. We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{put}(\textit{itm},\textit{pri})} q_{\textit{i+}} \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q_j \xrightarrow{\textit{rm}(\textit{itm})} q_{\textit{j+}} \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$. For each $\textit{i+1} \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $\textit{pri}$ into $\textit{itm} \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $\textit{pri}$ into $l_k$. Since (1) $\textit{pri}$ is one of maximal priority in $l_2$, (2) $\textit{itm}$ is the newest item with priority $\textit{pri}$ in $l_2$, and (3) $\textit{itm}$ is not removed until $\textit{rm}(\textit{itm})$, we know that whether we keep $\textit{itm}$ or remove it will not influence transitions from $q_{\textit{i+1}}$ to $q_j$. Then we can see that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ are transitions of $\textit{LTS}_e$. Therefore, $l_1 \in \textit{EPQ}_s$.
\end{itemize}

This completes the proof of this lemma. \qed
\end {proof}

\section{Proof in Section \ref{sec:data-independence of extended priority queue}}
\label{sec:appendix in section data-independence of EPQ}

\DataDifferentiatedisEnoughforPQ*

\begin {proof}

To prove the $\textit{only if}$ direction, given a data-differentiated execution $e \in \mathcal{I}_{\neq}$. By assumption, it is linearizable with respect to a sequential execution $l \in S$, and the bijection between the operations of $e$ and the method events of $l$ ensures that $l$ is differentiated and belongs to $S_{\neq}$.

To prove the $\textit{if}$ direction, given an execution $e \in \mathcal{I}$. By data independence of $\mathcal{I}$, we know that there exists $e' \in \mathcal{I}_{\neq}$ and a renaming function $r$, such that $r(e') = e$. By assumption, $e'$ is linearizable with respect to a sequential execution $l' \in S_{\neq}$. Let $l=r(l')$. By data independence of $S$ it is easy to see that $l \in S$, and it is easy to see that $e \sqsubseteq l$  using the same bijection used for $e' \sqsubseteq l'$. \qed
\end {proof}

\section{Proofs in Section \ref{sec:step-by-step linearizability of extended priority queues}}
\label{sec:appendix in section step-by-step linearizability of extended priority queues}

\subsection{Proof of Lemma \ref{lemma:EPQ is closed under projection}}

\EPQisClosedUnderProjection*

\begin {proof}

This is obvious, since for each conditions in the $\textit{Guard}$ part of the rules of priority queue, if a sequence of sequential executions satisfy it, then its sub-sequence also satisfy it. For example, if $\textit{noRE}(l)$ holds for some $l = u \cdot v \cdot w$ and let $D_l$ be the set of items of $l$, then for each subset $D' \subseteq D_l$, it is obvious that $\textit{rm}(\textit{empty}) \notin l \vert_{ D' }$ and $\textit{noRE}(l')$ holds. Similar cases hold for other predicates of the four rules of $\textit{EPQ}$, such as $\textit{LEI}$, $\textit{LI-U}$, $\textit{matched-C}$, $\textit{putInSeq}(l,l_1,\textit{pri})$ and $\textit{matched-All}$. This completes the proof of this lemma. \qed
\end {proof}

\subsection{Proof of Lemma \ref{lemma:EPQ as multi in MRpri for sequence}}

\EPQasMultiInMRpriforSequence*

\begin {proof}

The \textit{only if} direction is obvious and can be similarly proved as the $\textit{EPQ}_s \subseteq \textit{EPQ}$ direction of Lemma \ref{lemma:EPQ rules and semantics}.

To prove the $\textit{if}$ direction, we proceed as follows: From $e1=e$, we generate a sequence $e_2$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\textit{PQ}_3 \in \textit{last}(e_1)$: Then we can see that $e_1$ contains at least one $\textit{rm}(\textit{empty})$. $e_2$ is generated from $e_1$ by erasing one $\textit{rm}(\textit{empty})$.

\item[-] Else, if $\textit{PQ}_2^{=} \in \textit{last}(e_1)$: Then we can see that one of the maximal priority of $e_1$ is unmatched $\textit{put}$ and (possibly) matched $\textit{put}$. Assume the set of the items of these unmatched $\textit{put}$ is $S$. $e_2$ is generated from $e_1$ by erasing one unmatched $\textit{put}$ which use the item last putted in $S$.

\item[-] Else, if $\textit{PQ}_2^{>} \in \textit{last}(e_1)$: Then we can see that one of the maximal priority in $e_1$ is unmatched $\textit{put}$. Assume the set of the items of these unmatched $\textit{put}$ is $S$. $e_2$ is generated from $e_1$ by erasing one unmatched $\textit{put}$ which use the item last putted in $S$.

\item[-] Else, if $\textit{PQ}_1^{=} \in \textit{last}(e_1)$: Then we can see that one of the maximal priority in $e_1$ is of more than one pair of matched $\textit{put}$. Assume the set of the items of these matched $\textit{put}$ is $S$. $e_2$ is generated from $e_1$ by erasing matched $\textit{put}$ and $\textit{rm}$ of the item which is last putted in $S$.

\item[-] Else, if $\textit{PQ}_1^{>} \in \textit{last}(e_1)$: Then we can see that one of the maximal priority in $e_1$ is of one pair of matched $\textit{put}$. $e_2$ is generated from $e_1$ by erasing this pair of matched $\textit{put}$ and $\textit{rm}$.
\end{itemize}

Similarly, for each $i > 1$, we obtain $e_{\textit{i+1}}$ from $e_i$, until we obtain $e_m = \epsilon$ for some $m$. It is obvious that $e_m \in \textit{EPQ}$. For $e_{\textit{m-1}}$, since

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $e_m \in \textit{EPQ}$,

\item[-] By assumption, we know that $e_{\textit{m-1}} \in \textit{MS}(R_{\textit{m-1}})$, where $R_{\textit{m-1}} \in \textit{last}(e_{\textit{m-1}})$. This implies that the guard of $R_{\textit{m-1}}$ is satisfied.
\end{itemize}

Therefore, we know that $e_{\textit{m-1}} \in \textit{EPQ}$. Similarly, we can prove that $e_{\textit{m-2}},\ldots,e_1 = e \in \textit{EPQ}$. \qed
\end {proof}

\subsection{Proof of Lemma \ref{lemma:EPQ is step-by-step linearizability}}

The prove that $\textit{EPQ}$ is step-by-step linearizability, we investigate each rules individually.

Given a data-differentiated execution and its history, we can abuse notation and mix labels and method events with operations themselves, since items are unique in a data-differentiated execution. For instance, we will reference an operation labeled by $\textit{put}(p,a)$ as $\textit{put}(p,a)$.

Given an operation $o$ with call action $\textit{cal}_o (\textit{put},a,p)$ and return actions $\textit{ret}_o (\textit{put})$, its method event is $\textit{put}(a,p)$. Given an operation $o$ with call action $\textit{cal}_o (\textit{rm})$ and return actions $\textit{ret}_o (\textit{rm},a)$, its method event is $\textit{rm}(a)$.

Given a data-differentiated execution $e$ and its history $h$, we can obtain a sequence $h'$ from $h$ by adding $\textit{put}(a,p)$ (resp., $\textit{rm}(a)$, $\textit{rm}(\textit{empty})$) between each pair of $\textit{cal}(\textit{put},a,p)$ and $\textit{ret}(\textit{put},a)$ (resp., $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$, $\textit{cal}(\textit{rm},\textit{empty})$ and $\textit{ret}(\textit{rm},\textit{empty})$). The projection of $h'$ into method events is called linearization of $e$ and $h$, and each method event we add in $h'$ can be considered as a linearization point of the corresponding method event. We call such $h'$ an execution with linearization points of $e$.

\begin{restatable}{lemma}{EPQ1isStepByStepLinearizability}
\label{lemma:EPQ1 is step-by-step linearizability}
If a data-differentiated concurrent execution $e$ is linearizable w.r.t. $\textit{MS}(\textit{EPQ}_1)$ with witness $x$, then $e \setminus x \sqsubseteq \textit{EPQ} \Rightarrow e \sqsubseteq \textit{EPQ}$.
\end{restatable}

\begin {proof}
Let $h$ be the data-differentiated history of $e$, and $l$ be an sequential execution such that $h \sqsubseteq l$ and $l$ matches $\textit{EPQ}_1$ with witness $x$. Let the priority of $x$ be $\textit{pri}_x$, and let $h'=h \setminus x$ and assume that $h' \sqsubseteq l' \in \textit{EPQ}$. Let $e_{\textit{lp}}$ be an execution with linearization points of $e$ and the linearization points is added according to $l'$. Or we can say, $e_{\textit{lp}}$ is generated from $e$ by instrumenting linearization points, and the projection of $e_{\textit{lp}}$ into method event is $l'$.

According to $\textit{MS}(\textit{EPQ}_1)$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$ and $u$, $v$, $w$, $x$ and $\textit{pri}_x$ satisfy the guard of $\textit{EPQ}_1$. Let $l'_v$ be the shortest prefix of $l'$ that contains all method event of $u \cdot v$.

Let $U$, $V$ and $W$ be the set of operations of $u$, $v$ and $w$, respectively. Let us change $U$, $V$ and $W$ by loops as follows: In the first loop, we start from the first $W$-element of $l'_v$, and let it be $o_h$,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Case $1$: If in $e_{\textit{lp}}$, the linearization point of $o_h$ is before $\textit{ret}(\textit{rm},x)$, and no element in $W$ happens before $o_h$. Then, we erase $o_h$ from $W$ and put it into $U \cdot V$.

\item[-] Case $2$: Else, if in $e_{\textit{lp}}$, the linearization point of $o_h$ is before $\textit{ret}(\textit{rm},x)$, and there exists $o_w \in W$, such that $o_w <_{\textit{hb}} o_h$. Then, in $l'_v$, we erase all such operations from $U \cdot V$ and put them into $W$, and then stop the process of changing $U$, $V$ and $W$. These operations should satisfy (1) their linearization point is after $o_h$ in $e_{\textit{lp}}$, and (2) the priority of their item is incomparable with $\textit{pri}_x$.

\item[-] Case $3$: Else, if in $e_{\textit{lp}}$, the linearization point of $o_h$ is after $\textit{ret}(\textit{rm},x)$. Then, in $l'_v$, we erase all such operations from $U \cdot V$ and put them into $W$, and then stop the process of changing $U$, $V$ and $W$. These operations should satisfy (1) their linearization point is after $o_h$ in $e_{\textit{lp}}$, and (2) the priority of their item is incomparable with $\textit{pri}_x$.
\end{itemize}

Our process proceed, until either all element in $l'_v$ are in new $U$ or new $V$, or case $2$ or case $3$ holds and this process terminates. Let $U'$, $V'$ and $W'$ be the new version of $U$, $V$ and $W$ after the process terminates, respectively. Let $O_+$ be the set of operations that are moved into $U' \cup V'$ in the process, and let $O_-$ be the set of operations that are moved into $W'$ in the process.

Let $l'_{\textit{u'v'}}$ be the projection of $l'$ into $U' \cup V'$, let $O_x$ be the set of $\textit{put}(\_,\textit{pri}_x)$ while the item is not $x$ in $h$. Let $l''_a$ be the longest prefix of $l'_{\textit{u'v'}}$, where linearization of each operation of $l''_a$ is before $\textit{ret}(\textit{put},b)$ in $e_{\textit{lp}}$. Let $l''_d$ be the projection of $l'$ into operations of $O_x$ that are not in $l''_a$. Let $l''_1 = l''_a \cdot l''_d$. Let $l''_2$ be the projection of $l'$ into operations of $l'_{\textit{u'v'}}$ that are not in $l''_1$. Let $l''_3$ be the projection of $l'$ into operations of $l'_{\textit{u'v'}}$ that are not in $l''_1$ and $l''_2$. Let $l'' = l''_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3$.

To prove $h \sqsubseteq l''$, we define a graph $G$ whose nodes are the operations of $h$ and there is an edge from operation $o_1$ to $o_2$, if one of the following case holds

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $o_1$ happens-before $o_2$ in h,

\item[-] the method event corresponding to $o_1$ in $l''$ is before the one corresponding to $o_2$.
\end{itemize}

Assume there is a cycle in $G$. According the the property of interval order and the fact that the order of $l''$ is total, we know that there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$. Then, we consider all possible case of $o_1$ and $o_2$ as follows: Let $O_a$ and $O_d$ be the set of operations in $l''_a$ and $l''_d$, respectively.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $o_2 \in l''_1 \wedge o_1 \in l''_1$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_1,o_2 \in O_a$ or $o_1,o_2 \in O_d$: Then $l'$ contradicts with happen before relation of $h$.

    \item[-] If $o_2 \in O_a \wedge o_1 \in O_d$: Then the order of linearization points of $e_{\textit{lp}}$ contradicts with happen before relation of $h$.
    \end{itemize}

\item[-] If $o_2 \in l''_1 \wedge o_1 = \textit{put}(x,\textit{pri}_x)$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_1 \in O_a$: This is impossible, since the linearization point of operations in $O_a$ is before $\textit{ret}(\textit{put},b)$ in $e_{\textit{lp}}$.

    \item[-] If $o_1 \in O_d$: Then $l$ contradicts with happen before relation of $h$.
    \end{itemize}

\item[-] If $o_2 \in l''_1 \wedge o_1 \in l''_2$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_2 \in O_a$: This violates the order of linearization point in $e_{\textit{lp}}$.

    \item[-] If $o_2 \in O_d$: According to $l$, we can see that $\textit{put}(x,\textit{pri}_x)$ does not happen before any operation in $O_x$. Then we can see that the linearization point of $o_1$ is before $\textit{ret}(\textit{put},b)$ and $o_1 \in O_a$. This violates that $o_1 \in l''_2$.
    \end{itemize}

\item[-] If $o_2 \in l''_1 \wedge o_1 = \textit{rm}(x)$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_2 \in U \cup V$: Then $l$ contradicts with happen before relation of $h$.

    \item[-] If $o_2 \in O_+$: This is impossible, since the linearization point of operations in $O_+$ is before $\textit{ret}(\textit{rm},b)$ in $e_{\textit{lp}}$.
    \end{itemize}

\item[-] If $o_2 \in l''_1 \wedge o_1 \in l''_3$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_1 \in W \wedge o_2 \in U \cup V$: Then $l$ contradicts with happen before relation of $h$.

    \item[-] If $o_1 \in W \wedge o_2 \in O_+$: This is impossible, since according to construction of $O_+$, we can see that $o_2 \in O_-$.

    \item[-] If $o_1 \in O_- \wedge o_2 \in U \cup V$:
        \begin{itemize}
        \setlength{\itemsep}{0.5pt}
        \item[-] If the reason of $o_1 \in O_-$ is case $2$: Let $o_h$ be as in case $2$. Then there exists $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{cal}(o_2)$. Therefore, we can see that $o_w <_{\textit{hb}} o_2$, and then $l$ contradicts with happen before relation of $h$.

        \item[-] If the reason of $o_1 \in O_-$ is case $3$: Let $o_h$ be as in case $3$. Then in $e_{\textit{lp}}$, $\textit{ret}(\textit{rm},x)$ is before the linearization point of $o_h$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{cal}(o_2)$. Therefore, we can see that $\textit{rm}(x) <_{\textit{hb}} o_2$, and then $l$ contradicts with happen before relation of $h$.
        \end{itemize}
    \item[-] If $o_1 \in O_- \wedge O_2 \in O_+$: This is impossible, since in $e_{\textit{lp}}$, the linearization points of operations in $O_+$ is before the linearization points of operations in $O_-$.
    \end{itemize}

\item[-] If $o_2 = \textit{put}(x,\textit{pri}_x) \wedge o_1 \in l''_2$: This is impossible, since in $e_{\textit{lp}}$, the linearization points of operations in $l''_2$ is after $\textit{ret}(\textit{put},x)$.

\item[-] If $o_2 = \textit{put}(x,\textit{pri}_x) \wedge o_1 = \textit{rm}(x)$: Then $l$ contradicts with happen before relation of $h$.

\item[-] If $o_2 = \textit{put}(x,\textit{pri}_x) \wedge o_1 \in l''_3$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_1 \in W$: Then $l$ contradicts with happen before relation of $h$.

    \item[-] If $o_1 \in O_-$:
         \begin{itemize}
         \setlength{\itemsep}{0.5pt}
         \item[-] If the reason of $o_1 \in O_-$ is case $2$: Let $o_h$ be as in case $2$. Then there exists $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{cal}(\textit{put},x,\textit{pri}_x)$. Therefore, we can see that $o_w <_{\textit{hb}} \textit{put}(x,\textit{pri}_x)$, and then $l$ contradicts with happen before relation of $h$.

         \item[-] If the reason of $o_1 \in O_-$ is case $3$: Let $o_h$ be as in case $3$. Then in $e_{\textit{lp}}$, $\textit{ret}(\textit{rm},x)$ is before the linearization point of $o_h$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{cal}(\textit{put},x,\textit{pri}_x)$. Therefore, we can see that $\textit{rm}(x) <_{\textit{hb}} \textit{put}(x,\textit{pri}_x)$, and then $l$ contradicts with happen before relation of $h$.
         \end{itemize}
    \end{itemize}

\item[-] If $o_2 \in l''_2 \wedge o_1 \in l''_2$: Then $l'$ contradicts with happen before relation of $h$.

\item[-] If $o_2 \in l''_2 \wedge o_1 = \textit{rm}(x)$: We can prove this similarly as the case of $o_2 \in l''_1 \wedge o_1 = \textit{rm}(x)$.

\item[-] If $o_2 \in l''_2 \wedge o_1 \in l''_3$: We can prove this similarly as the case of $o_2 \in l''_1 \wedge o_1 \in l''_3$.

\item[-] If $o_2 = \textit{rm}(x) \wedge o_1 \in l''_3$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_1 \in W$: Then $l$ contradicts with happen before relation of $h$.

    \item[-] If $o_1 \in O_-$:
         \begin{itemize}
         \setlength{\itemsep}{0.5pt}
         \item[-] If the reason of $o_1 \in O_-$ is case $2$: Let $o_h$ be as in case $2$. Then there exists $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$.

             Since $l$ is consistent with the happen before order of $h$, we can see that $\textit{cal}(\textit{rm},x)$ is before $\textit{ret}(o_w)$. Therefore, we can see that the linearization point of $o_1$ is after $\textit{cal}(\textit{rm},x)$, and then it is impossible that $o_1 <_{\textit{hb}} \textit{rm}(x)$.

         \item[-] If the reason of $o_1 \in O_-$ is case $3$: Let $o_h$ be as in case $3$. Then in $e_{\textit{lp}}$, $\textit{ret}(\textit{rm},x)$ is before the linearization point of $o_h$, and the linearization point of $o_h$ is before the linearization point of $o_1$. Therefore, we can see that the linearization point of $o_1$ is after $\textit{ret}(\textit{rm},x)$, and then it is impossible that $o_1 <_{\textit{hb}} \textit{rm}(x)$.
         \end{itemize}
    \end{itemize}

\item[-] If $o_2 \in l''_3 \wedge o_1 \in l''_3$: Then $l'$ contradicts with happen before relation of $h$.
\end{itemize}

Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq l''$.

It remains to prove that $l'' \in \textit{EPQ}$. Let $O_c$ be the set of operations in $h$, whose items have priority that are comparable with $\textit{pri}_x$, and let $O_i$ be the set of operations in $h$, whose items have priority that are incomparable with $\textit{pri}_x$. Then we obtain that $l'' \in \textit{EPQ}$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Since $l' \in \textit{EPQ}$ and $l'_v$ is a prefix of $l'$, it is obvious that $l'_v \in \textit{EPQ}$.

\item[-] $l'_{\textit{u'v'}}$ can be obtained from $l'_v$ as follows: From some time point $t$, discard all the $O_i$ operations after $t$ in $l'_v$. Since $\textit{pri}_x$ is one of maximal priority in $h$, we can see that $\forall p_i \in O_i$ and $\forall p_c \in O_c$, either $p_c <_{\mathbb{P}} p_i$, or $p_c$ is incomparable with $p_i$. Therefore, this process does not influence correctness of $l'_{\textit{u'v'}}$, and then $l'_{\textit{u'v'}} \in \textit{EPQ}$.

\item[-] $l''_1 \cdot l''_2$ can be obtained from $l'_{\textit{u'v'}}$ as follows: Execute until reaching some time point $t$, then first execute all $O_x$ operations after $t$, and then execute remanning operations. Since $O_x$ only contains $\textit{put}(\_,\textit{pri}_x)$ and $\textit{pri}_x$ is one of maximal priority in $h$, we can see that $l''_1 \cdot l''_2 \in \textit{EPQ}$.

\item[-] Let $l''_f$ be the projection of $l'$ into $O_-$, and let $l''_g$ be the projection of $l'$ into operations not in $l''_1$, $l''_2$ or $l''_f$. It is easy to see that $l''_3 = l''_f \cdot l''_g$.

\item[-] $l'_{\textit{u'v'}} \cdot l''_f$ can be obtained from $l'_v$ as follows: (1) Execute until reaching some time point $t$, (2) from $t$, execute only $O_c$ operations until the end of $l'_v$, and (3) execute $O_i$ operations after $t$ in $l'_v$. Since $\textit{pri}_x$ is one of maximal priority in $h$, we can see that $l'_{\textit{u'v'}} \cdot l''_f \in \textit{EPQ}$.

\item[-] It is easy to see that $l' = l'_v \cdot l''_g \in \textit{EPQ}$. We already know that $l'_{\textit{u'v'}} \cdot l''_f \in \textit{EPQ}$, $l'_v \in \textit{EPQ}$. It is not hard to see that the content of extended priority queue after executing $l'_{\textit{u'v'}} \cdot l''_f$ is the same as that after executing $l'_v$. Therefore, it is easy to prove that $l'_{\textit{u'v'}} \cdot l''_f \cdot l''_g \in \textit{EPQ}$.

\item[-] We already know that $l''_1 \cdot l''_2\in \textit{EPQ}$ and $l'_{\textit{u'v'}} \in \textit{EPQ}$, and it is not hard to see that the the content of extended priority queue after executing $l''_1 \cdot l''_2$ is the same as that after executing $l'_{\textit{u'v'}}$. Since $l'_{\textit{u'v'}} \cdot l''_f \cdot l''_g \in \textit{EPQ}$, it is easy to see that $l''_1 \cdot l''_2 \cdot l''_f \cdot l''_g \in \textit{EPQ}$. Or we can say, $l''_1 \cdot \l''_2 \cdot l''_3 \in \textit{EPQ}$. Since $u$, $v$, $w$, $x$ and $\textit{pri}_x$ satisfy the guard of $\textit{EPQ}_1$, it is easy to see that $l'' \in \textit{EPQ}$.
\end{itemize}

Therefore, we prove that $h \sqsubseteq l'' \in \textit{EPQ}$. This completes the proof of this lemma. \qed
\end {proof}

\begin{restatable}{lemma}{EPQ2isStepByStepLinearizability}
\label{lemma:EPQ2 is step-by-step linearizability}
If a data-differentiated concurrent execution $e$ is linearizable w.r.t. $\textit{MS}(\textit{EPQ}_2)$ with witness $x$, then $e \setminus x \sqsubseteq \textit{EPQ} \Rightarrow e \sqsubseteq \textit{EPQ}$.
\end{restatable}

\begin {proof}

Let $h$ be the data-differentiated history of $e$, and $l$ be an sequential execution such that $h \sqsubseteq l$ and $l$ matches $\textit{EPQ}_2$ with witness $x$. Let the priority of $x$ be $\textit{pri}_x$, and let $h'=h \setminus x$ and assume that $h' \sqsubseteq l' \in \textit{EPQ}$. Let $e_{\textit{lp}}$ be an execution with linearization points of $e$ and the linearization points is added according to $l'$. Or we can say, $e_{\textit{lp}}$ is generated from $e$ by instrumenting linearization points, and the projection of $e_{\textit{lp}}$ into method event is $l'$.

According to $\textit{MS}(\textit{EPQ}_2)$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$ and $u$, $v$, $x$ and $\textit{pri}_x$ satisfy the guard of $\textit{EPQ}_2$.

Let $l''_a$ be the longest prefix of $l'$ such that linearization point of each operation of $l''_a$ is before $\textit{ret}(\textit{put},x)$ in $e_{\textit{lp}}$. Let $O_x$ be the set of $\textit{put}(\_,\textit{pri}_x)$ while the item is not $x$ in $h$. Let $l''_s$ be the projection of $l'$ into operations of $O_x$ that are not in $l''_a$. Let $l''_1 = l''_a \cdot l''_s$. Let $l''_2$ be the projection of $l'$ into operations of $l'$ that are not in $l''_1$. Let $l'' = l''_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l''_2$.

To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:EPQ1 is step-by-step linearizability}. Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$. Then, we consider all possible case of $o_1$ and $o_2$ as follows: Let $O_a$ and $O_s$ be the set of operations in $l''_a$ and $l''_s$, respectively.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $o_2 \in l''_1 \wedge o_1 \in l''_1$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_1,o_2 \in O_a$ or $o_1,o_2 \in O_s$: Then $l'$ contradicts with happen before relation of $h$.

    \item[-] If $o_2 \in O_a \wedge o_1 \in O_s$: It is not hard to see that $\textit{put}(x,\textit{pri}_x) <_{\textit{hb}} o_2$. But then it is impossible that in $e_{\textit{lp}}$, the linearization point of $o_2$ be located before $\textit{ret}(\textit{put},x)$.
    \end{itemize}

\item[-] If $o_2 \in l''_1 \wedge o_1 = \textit{put}(x,\textit{pri}_x)$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_2 \in O_a$: This is impossible, since in $e_{\textit{lp}}$, the linearization point of $o_1$ is before $\textit{ret}(\textit{put},x)$.

    \item[-] If $o_2 \in O_s$: This is impossible, since such $o_1 = \textit{put}(\_,\textit{pri}_x)$, and $l$ is consistent with happen before relation of $h$.
    \end{itemize}

\item[-] If $o_2 \in l''_1 \wedge o_1 \in l''_2$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_2 \in O_a$: This is impossible, since in $e_{\textit{lp}}$, the linearization point of operation in $l''_a$ is before the linearization point of operations in $l''_2$.

    \item[-] If $o_2 \in O_s$: Since no $\textit{put}(\_,\textit{pri}_x)$ happens before $\textit{put}(x,\textit{pri}_x)$ in $h$, $\textit{cal}(o_2)$ is before $\textit{ret}(\textit{put},x)$. Since $o_1 <_{\textit{hb}} o_2$, we can see that $\textit{ret}(o_1)$ is before $\textit{cal}(o_2)$, and then $\textit{ret}(o_1)$ is before $\textit{ret}(\textit{put},x)$. Then the linearization point of $o_1$ can only be before $\textit{ret}(\textit{put},x)$, and $o_1 \in l''_a$, which contradicts that $o_1 \in l''_2$.
    \end{itemize}

\item[-] If $o_2 = \textit{put}(x,\textit{pri}_x) \wedge o_1 \in l''_2$: Then since the linearization point of $o_1$ can only be before $\textit{ret}(\textit{put},x)$, we can see that $o_1 \in l''_a$, which contradicts that $o_1 \in l''_2$.

\item[-] If $o_2 \in l''_2 \wedge o_1 \in l''_2$: Then $l'$ contradicts with happen before relation of $h$.
\end{itemize}

Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq l''$.

It remains to prove that $l'' \in \textit{EPQ}$. $l''_a \cdot l''_s \cdot l''_2$ can be obtained from $l'$ as follows: Execute until reaching some time point $t$, then first execute all $O_x$ operations after $t$, and then execute remanning operations. Since $O_x$ only contains $\textit{put}(x,\textit{pri}_x)$ and $\textit{pri}_x$ is one of maximal priority in $h$, we can see that $l''_1 \cdot l''_2 = l''_a \cdot l''_s \cdot l''_2 \in \textit{EPQ}$. Since $u$, $v$, $x$ and $\textit{pri}_x$ satisfy the guard of $\textit{EPQ}_2$, it is easy to see that $l'' \in \textit{EPQ}$.

Therefore, we prove that $h \sqsubseteq l'' \in \textit{EPQ}$. This completes the proof of this lemma.\qed
\end {proof}

\begin{restatable}{lemma}{EPQ3isStepByStepLinearizability}
\label{lemma:EPQ3 is step-by-step linearizability}
If a data-differentiated concurrent execution $e$ is linearizable w.r.t. $\textit{MS}(\textit{EPQ}_3)$ and $o$ is a $\textit{rm}(\textit{empty})$ event, then $e \setminus o \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
\end{restatable}

\begin {proof}
Let $h$ be the data-differentiated history of $e$, $l$ be an sequential execution such that $h \sqsubseteq l$, $l$ matches $\textit{EPQ}_3$ and $o$ is a $\textit{rm}$ method event in $h$. Let $h'=h \setminus o$ and assume that $h' \sqsubseteq l' \in \textit{EPQ}$.

According to $\textit{EPQ}_3$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{rm}(\textit{empty}) \cdot v$, where all the $\textit{put}$ operations and $\textit{rm}$ in $u$ are matched.

Let $E_L$ be the set of method events in $u$ and $E_R$ be the set of method events in $v$. Let $l'_L = l' \vert_{E_L}$ and $l'_R = l' \vert_{E_R}$. Let sequence $l'' = l'_L \cdot o \cdot L'_R$. Since priority queue is closed under projection (Lemma \ref{lemma:EPQ is closed under projection}) and all the $\textit{put}$ operations and $\textit{rm}$ in $u$ are matched, we know that $l'_L \in \textit{EPQ}$ and the the priority queue is empty after executing $l'_L$. Then we know that $l'_L \cdot \textit{rm}(\textit{empty}) \in \textit{EPQ}$. Since $l'_R$ is obtained from $l'$ by discarding pairs of matched $\textit{put}$ and $\textit{rm}$ operations, it is easy to see that $L'_R \in \textit{EPQ}$, and then we know that $l'' = l'_L \cdot o \cdot L'_R \in \textit{EPQ}$.

It remains to prove that $h \sqsubseteq l''$. To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:EPQ1 is step-by-step linearizability}. Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$. Then, we consider all possible case of $o_1$ and $o_2$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $o_1,o_2 \in l'_L$, or $o_1,o_2 \in l'_R$: Then $l'$ contradicts with happen before relation of $h$.

\item[-] If $o_1=o \wedge o_2 \in l'_L$, or $o_1 \in l'_R \wedge o_2 \in l'_L$, or $o_1 \in l'_R \wedge o_2 = o$, then $l$ contradicts with happen before relation of $h$.
\end{itemize}

Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \textit{EPQ}$. \qed
\end {proof}

The following lemma states that $\textit{EPQ}$ is step-by-step linearizability, it is a direct consequence of Lemma \ref{lemma:EPQ1 is step-by-step linearizability}, Lemma \ref{lemma:EPQ2 is step-by-step linearizability} and Lemma \ref{lemma:EPQ3 is step-by-step linearizability}.

\EPQueueisStepByStepLinearizability*

\begin {proof}
This is a direct consequence of Lemma \ref{lemma:EPQ1 is step-by-step linearizability}, Lemma \ref{lemma:EPQ2 is step-by-step linearizability} and Lemma \ref{lemma:EPQ3 is step-by-step linearizability}. \qed
\end {proof}

\subsection{Proof of Lemma \ref{lemma:EPQ as multi in MRpri for history}}

\EPQasMultiInMRpriforHistory*

\begin {proof}

To prove the $\textit{only if}$ direction, assume that $e \sqsubseteq l \in \textit{EPQ}$. Given $e' = e \vert_{D}$ and $l' = l \vert_{D}$, it is easy to see that $e' \sqsubseteq l'$, and by Lemma \ref{lemma:EPQ is closed under projection}, we can see that $l' \in \textit{EPQ}$. Then by Lemma \ref{lemma:EPQ as multi in MRpri for sequence} we know that for each $R \in \textit{last}(l')$, we have $l' \in \textit{MS}(R)$.

To prove the $\textit{if}$ direction, given $e_1 = e$, we generate sequence $e_2$ from $e_1$ as follows: Since $e_1 \in \textit{proj}(e)$, we know that for each $R_1 \in \textit{last}(e_1)$, we have $e_1 \sqsubseteq \textit{MS}(R_1)$. We choose an arbitrary $R_1$ in $\textit{last}(e_1)$,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $R_1 = \textit{EPQ}_3$: $e_2$ is generated from $e_1$ by erasing call and return of one $\textit{rm}(\textit{empty})$ operation.

\item[-] Else, if $R_1 = \textit{EPQ}_2^{=}, \textit{EPQ}_2^{>}, \textit{EPQ}_1^{=}, \textit{EPQ}_1^{>}$, and the witness is $x$: $e_2$ is generated from $e_1$ by erasing call and return of method event of item $x$.
\end{itemize}

Similarly, for each $i > 1$, we obtain $e_{\textit{i+1}}$ from $e_i$, until we obtain $e_m = \epsilon$ for some $m$. It is obvious that $e_m \sqsubseteq \textit{EPQ}$. For $e_{\textit{m-1}}$, since

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $e_m \sqsubseteq \textit{EPQ}$,

\item[-] If $\textit{last}(e_{\textit{m-1}}) = R_{\textit{m-1}} \in \{ \textit{EPQ}_1^{>}, \textit{EPQ}_1^{=}, \textit{EPQ}_2^{>}, \textit{EPQ}_2^{=} \}$ and $e_{\textit{m-1}}$ matches $R_{\textit{m-1}}$ with witness $x$: We already know that $e_{\textit{m-1}} \sqsubseteq \textit{MS}(R_{\textit{m-1}})$, $e_m = e_{\textit{m-1}} \setminus x \sqsubseteq \textit{EPQ}$, and by step-by-step linearizability of $\textit{EPQ}$ (Lemma \ref{lemma:EPQ is step-by-step linearizability}), we can see that $e_{\textit{m-1}} \sqsubseteq \textit{EPQ}$.

\item[-] If $\textit{last}(e_{\textit{m-1}}) = R_{\textit{m-1}} = \textit{EPQ}_3$ and $o$ is a $\textit{rm}(\textit{empty})$ in $e_{\textit{m-1}}$ that is removed in the process of constructing $e_m$: We already know that $e_{\textit{m-1}} \sqsubseteq \textit{MS}(R_{\textit{m-1}})$, $e_m = e_{\textit{m-1}} \setminus o \sqsubseteq \textit{EPQ}$, and by step-by-step linearizability of $\textit{EPQ}$ (Lemma \ref{lemma:EPQ is step-by-step linearizability}), we can see that $e_{\textit{m-1}} \sqsubseteq \textit{EPQ}$.
\end{itemize}

Therefore, we know that $e_{\textit{m-1}} \in \textit{EPQ}$. Similarly, we can prove that $e_{\textit{m-2}},\ldots,e_1 = e \in \textit{EPQ}$. \qed
\end {proof}

\section{Proofs and Definitions in Section \ref{sec:co-regular of extended priority queues}}
\label{sec:appendix proof and definition in section co-regular of extended priority queues}

\subsection{Proofs and Definitions in Subsection \ref{subsec:definition of co-regular}}
\label{sec:appendix proof and definition in section definition of co-regular}

\cite{Bouajjani:2015} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $a,b$: (1) $\textit{deq}(b) <_{hb} \textit{enq}(b)$, (2) there are are no $\textit{enq}(b)$ and at least one $\textit{deq}(b)$, (3) there are are one $\textit{enq}(b)$ and more than one $\textit{deq}(b)$, and (4) $\textit{enq}(a) <_{\textit{hb}} \textit{enq}(b)$, and $\textit{deq}(b) <_{\textit{hb}} \textit{deq}(a)$, or $\textit{deq}(a)$ does not exists.

For each such case, we construct a witness automata. We generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}. Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri})$, $\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},b),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6 \textwidth]{figures/PIC_AUTO_FIFO_1.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{SinPri}}^1$}
  \label{fig:automata for FIFO-1 in appendix}
\end{figure}

We generate witness automata $\mathcal{A}_{\textit{SinPri}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata for FIFO-2}. Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_2 = c_1 + \textit{cal}(\textit{rm},b) + \textit{ret}(\textit{rm},b)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.3 \textwidth]{figures/PIC_AUTO_FIFO_2.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{SinPri}}^2$}
  \label{fig:automata for FIFO-2}
\end{figure}

We generate witness automata $\mathcal{A}_{\textit{SinPri}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata for FIFO-3}. Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},b)$, $c_4 = c_3 + \textit{cal}(\textit{rm},b)$, $c_5 = c_1 + \textit{ret}(\textit{rm},b)$, $c_6 = c_5 + \textit{cal}(\textit{rm},b)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.7 \textwidth]{figures/PIC_AUTO_FIFO_3.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{SinPri}}^3$}
  \label{fig:automata for FIFO-3}
\end{figure}

We generate witness automata $\mathcal{A}_{\textit{SinPri}}^4$ for the forth case, and it is shown in \figurename~\ref{fig:automata for FIFO-4}. Here $c_1 = c + \textit{cal}(\textit{rm},b)$, and $c_2 = c + \textit{ret}(\textit{put},b) + \textit{cal}(\textit{rm},a) + \textit{ret}(\textit{rm},a)$, where $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d),\textit{ret}(\textit{rm},d),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.9 \textwidth]{figures/PIC_AUTO_FIFO_4.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{SinPri}}^4$}
  \label{fig:automata for FIFO-4}
\end{figure}

Let $\textit{Auts}_{\textit{sinPri}} = \{ \mathcal{A}_{\textit{SinPri}}^1, \mathcal{A}_{\textit{SinPri}}^2, \mathcal{A}_{\textit{SinPri}}^3, \mathcal{A}_{\textit{SinPri}}^4 \}$. Let us prove Lemma \ref{lemma:automata for extended priority queue with single priority}.

\AutoForEPQwithSignlePri*

\begin {proof}

\cite{Bouajjani:2015} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $v_a,v_b$: (1) $\textit{deq}(v_b) <_{hb} \textit{enq}(v_b)$, (2) there are are no $\textit{enq}(v_b)$ and at least one $\textit{deq}(v_b)$, (3) there are are one $\textit{enq}(v_b)$ and more than one $\textit{deq}(v_b)$, and (4) $\textit{enq}(v_a) <_{\textit{hb}} \textit{enq}(v_b)$, and $\textit{deq}(v_b) <_{\textit{hb}} \textit{deq}(v_a)$, or $\textit{deq}(v_a)$ does not exists.

Let us prove the $\textit{only if}$ direction. Assume that there exists execution $e_0 \in \mathcal{I}$ and $e_0$ is accepted by an automaton in $\textit{Auts}_{\textit{sinPri}}$. By data-independence, we can see that there exists a data-differentiated $e \in \mathcal{I}$ and renaming function, such that $e_0=r(e)$. Let $e'$ be obtained from $e$ by first removing $\textit{rm}(\textit{empty})$, and then,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $e_0$ is accepted by $\mathcal{A}_{\textit{SinPri}}^1$, $\mathcal{A}_{\textit{SinPri}}^2$ or $\mathcal{A}_{\textit{SinPri}}^3$: Then remove all items that are not renamed into $b$ by $r$.

\item[-] If $e_0$ is accepted by $\mathcal{A}_{\textit{SinPri}}^4$: Then remove all items that are not renamed into $a$ or $b$ by $r$.
\end{itemize}

It is obvious that $e' \in \textit{proj}(e)$. It is easy to see that $\textit{transToQueue}(e')$ satisfies one of above conditions, and then $\textit{transToQueue}(e')$ is not linearizable w.r.t queue.

Let us prove the $\textit{if}$ direction. Assume that exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{transToQueue}(e')$ does not linearizable to queue. Then we construct a renaming function $r$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If this is because case $1$, case $2$ or case $3$: $r$ maps $v_b$ into $b$ and maps all other values into $a$.

\item[-] If this is because case $4$: $r$ maps $v_a$ and $v_b$ into $a$ and $b$, respectively, and maps all other values into $d$.
\end{itemize}

Then it is not hard to see that $r(e) \in \mathcal{I}$ and it is accepted by some automaton in $\textit{Auts}_{\textit{sinPri}}$. This completes the proof of this lemma. \qed
\end {proof}

\subsection{Proofs and Definitions in Subsection \ref{subsec:co-regular of EPQ1Lar}}
\label{sec:appendix proof and definition in section co-regular of EPQ1Lar}

The following lemma states that, from linearization of sub-histories, we can merge them and obtain a linearization (regardless of whether it belongs to sequential specification) of the whole history.

\begin{restatable}{lemma}{MergeTwoLinearization}
\label{lemma:merge two linearization}

Given a history $h$, operation sets $S_1$, $S_2$ and sequences $l_1$ and $l_2$. Let $h_1 = h \vert{S_1}$ and $h_2 = h \vert{S_2}$. Assume that $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, and $S_1 \cup S_2$ contains all operations of $h$. Then, there exists a sequence $l$, such that $h \sqsubseteq l$, $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$.
\end{restatable}

\begin {proof}

Given a history $h$ and a operation $o \in S_2$, let $\textit{MB}(o) = \{ o' \vert o' \in S_1$ and $o'$ happens before $o$ in $h \}$, let $\textit{SBI}(o) = \textit{min}\{ i \vert l_1[0,i]$ contains all elements of $\textit{MB}(o) \}$.

Let $l = s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[n] \cdot s_{\textit{n+1}}$ be generated as follows, where $n = \vert l_2 \vert$:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $s_1 = l_1[0, \textit{SBI}(l_2[1])]$,

\item[-] If $s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[i]$ already contains $l_1(\textit{SBI}(l_2[\textit{i+1}]))$, then $s_{\textit{i+1}} = \epsilon$. Otherwise, $s_{\textit{i+1}}$ is a subsequence of $l_1$, which starts from the next of last elements of $s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[i]$ in $l_1$ and ends in $l_1(\textit{SBI}(l_2[\textit{i+1}]))$.
\end{itemize}

It is obvious that $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$, and it remains to prove that $h \sqsubseteq l$. We prove this by contradiction. Assume that $h$ is not linearizable with respect to $l$. Then there must be two operations of $h$, such that $o_1 <_{hb} o_2$ in $h$ but $o_2$ before $0_1$ in $l$. Since $l \vert{S_1} = l_1$, $l \vert{S_2} = l_2$, and $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, it is easy to see that it is impossible that $o_1,o_2 \in S_1$ or $o_1,o_2 \in S_2$. There are only two possibilities:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $o_1 \in S_1 \wedge o_2 \in S_2$. Then we can see that $o_2=l_2[i]$ and $o_1 \in s_j$ for some $i < j$. Since $o_1 <_{hb} o_2$, we know that $o_1 \in \textit{SBI}(o_2)$. By the construction of $l$, we know that $o_1$ must be in $s_k$ for some $k \leq i$, contradicts that $o_1 \in s_j$ with $i < j$.

\item[-] $o_1 \in S_2 \wedge o_2 \in S_1$. Then we can see that $o_2 \in s_i$ and $o_1 = l_2[j]$ for some $i \leq j$. It is easy to see that this leads to contradiction when $i = j$. For the case of $i \neq j$, we need to satisfy the following requirements: (1) $o_1$ ($l_2[j]$) does not happen before $l_2[i]$, (2) $l_2[i]$ does not happen before $o_2$, (3) $o_2$ is either overlap or happens before $o' \in \textit{MB}(l_2[i])$, and (4) $o' <_{hb} l_2[i]$. By enumeration we can see that it is impossible that above four conditions be satisfied while $o_1 <_{hb} o_2$.
\end{itemize}

This completes the proof of this lemma. \qed
\end {proof}

With Lemma \ref{lemma:merge two linearization}, we can now prove Lemma \ref{lemma:pri execution is enough}.

\priExecutionIsEnough*
\begin {proof}

We deal with the case of $R = \textit{EPQ}_1^{>}$ , and other cases can be similarly dealt with.

To prove the $\textit{only if}$ direction, given $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$ and such $\textit{pri}$ and $x$. Since $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$ with witness $x$, we know that $e \sqsubseteq u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$, where $u$, $v$, $w$, $x$ and $\textit{pri}$ satisfy the guard of $\textit{EPQ}_1^{>}$. Let $u'$, $v'$ and $w'$ be obtained from $u$, $v$ and $w$ by erasing all items with priority incomparable with $\textit{pri}$, respectively. It is not hard to see that $u'$, $v'$, $w'$, $x$ and $\textit{pri}$ satisfy the guard of $\textit{EPQ}_1^{>}$, and then $e \sqsubseteq l = u' \cdot \textit{put}(x,\textit{pri}) \cdot v' \cdot \textit{rm}(x) \cdot w' \in \textit{MS}(\textit{EPQ}_1^{>})$.

To prove the $\textit{if}$ direction, given $e' = \textit{pri-Exec}(e)$ and such $x$ and $\textit{pri}$. Since $e' \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$ with witness $x$, we know that $e' \sqsubseteq l_1 = u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$, where $u$, $v$, $w$, $x$ and $\textit{pri}$ satisfy the guard of $\textit{EPQ}_1^{>}$. Let $O_c$ be the set of operations in $e$ that have priorities comparable with $\textit{pri}$, and Let $O_i$ be the set of operations in $e$ that have priorities incomparable with $\textit{pri}$. It is obvious that $l_1$ is the linearization of $e \vert_{O_c}$. By Lemma \ref{lemma:merge two linearization}, there exists sequence $l$, such that $e \sqsubseteq l$, and $l \vert_{O_c} = l_1$. Then $l = u' \cdot \textit{put}(x,\textit{pri}) \cdot v' \cdot \textit{rm}(x) \cdot w'$, where $u' \vert_{O_c} = u$, $v' \vert_{O_c} = v$ and $w' \vert_{O_c} = w$. Since $\textit{pri}$ is one of maximal priorities in $e$, and the predicates of guards of $\textit{EPQ}_1^{>}$ does not restrict $O_i$, it is easy to see that $l \in \textit{MS}(\textit{EPQ}_1^{>})$ and then $e \sqsubseteq l$ with witness $x$. \qed
\end {proof}

We can see that $\textit{UVSet}_i(e,x) \cap \textit{UVSet}_j(e,x) = \emptyset$ for any $i \neq j$.

The following lemma states that $\textit{UVSet}(e,x)$ contains only matched $\textit{put}$ and $\textit{rm}$.

\begin{restatable}{lemma}{UVSetHasMatchedPutandRm}
\label{lemma:UVSet has matched put and rm}
Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{>}$. Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority. Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$. Assume that $G$ has no cycle going through $x$. Then, $\textit{UVSet}(e,x)$ contains only matched $\textit{put}$ and $\textit{rm}$.
\end{restatable}
\begin {proof}

We prove this lemma by contradiction. Assume that there exists a value, such that $\textit{UVSet}(e,x)$ contains only its $\textit{put}$ and does not contain its $\textit{rm}$. Then we can see that there exists $d_1,\ldots,d_j$. Intuitively, $d_1,\ldots,d_j$ are elements in $\textit{UVSet}_1(e,x), \ldots, \textit{UVSet}_i(e,x)$, respectively. $\textit{UVSet}(e,x)$ contains $\textit{put}(d_j,\_)$ and does not contain $\textit{rm}(d_j)$. And each $d_i$ is the reason of $d_{\textit{i+1}} \in \textit{UVSet}_{\textit{i+1}}(e,x)$. Formally, we require that

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] For each $1 \leq i \leq j$, method events of $d_i$ belongs to $\textit{UVSet}_i(e,x)$.

\item[-] For each $i \neq j$, $\textit{put}(d_i,\_),\textit{rm}(d_i) \in \textit{UVSet}_i(e,x)$. $\textit{put}(d_j,\_) \in \textit{UVSet}_j(e,x)$, and $e$ does not contain $\textit{rm}(d_j)$.

\item[-] An operation of $d_1$ happens before an operations of $x$. For each $1 < i \leq j$, an operation of $d_i$ happens an operation of $d_{\textit{i-1}}$.

\item[-] For each $k$ and $\textit{ind}$, if $k > \textit{ind+1}$, then no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.
\end{itemize}

According to the definition of $\textit{UVSet}(e,x)$, it is easy to see that such $d_1,\ldots,d_j$ exists. Let us prove the following fact:

\noindent {\bf $\textit{fact}_1$}: Given $1 \leq i < j$, it can not be the case that $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap.

Proof of $\textit{fact}_1$: We prove $\textit{fact}_1$ by contradiction. Assume that for some $i \neq j$, $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap. Since $\textit{put}(d_i,\_), \textit{rm}(d_i) \in \textit{UVSet}_i(h,x)$, we know that an operation $o_i$ of $d_i$ happens before operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$. Moreover, since $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap, it is not hard to see that the call action of $\textit{put}(d_i,\_)$ and the call action of $\textit{rm}(d_i)$ is before the call action of $o_{\textit{i-1}}$. Since method events of $d_{\textit{i+1}}$ is in $\textit{UVSet}_{\textit{i+1}}(e,x)$, we know that an operation $o'_{\textit{i+1}}$ of $d_{\textit{i+1}}$ happens before operation $o'_i$ of $d_i$. Then, it is not hard to see that $o'_{\textit{i+1}}$ also happens before $o_{\textit{i-1}}$, which contradicts that for each $k > \textit{ind+1}$, no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.

We already know that an operation of $d_1$ happens before an operation of $x$. By $\textit{fact}_1$, we can ensure that $\textit{put}(d_1,\_)$ happens before an operation of $x$, and then $d_1 \rightarrow x$ in $G$. For each $1 < i \leq j$, we know that an operation $o_i$ of $d_i$ happens before an operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$. By $\textit{fact}_1$, we can ensure that $o_i=\textit{put}(d_i,\_)$ and $o_{\textit{i-1}}=\textit{rm}(d_{\textit{i-1}})$, and then $d_i \rightarrow d_{\textit{i-1}}$ in $G$. Since $h$ contains $\textit{put}(d_j,\_)$ and does not contain $\textit{rm}(d_j)$, we know that $x \rightarrow d_j$ in $G$. Then $G$ has a cycle going through $x$, contradicts that $G$ has no cycle going through $x$. \qed
\end {proof}

The following lemma states that $\textit{UVSet}(e,x)$ does not happen before $\textit{rm}(x)$ when the left-right constraint has no cycle going through $x$.

\begin{restatable}{lemma}{RmxDoesNotHappenBeforeUVSetForEPQ1Lar}
\label{lemma:Rmx does not happen before UVSet for EPQ1Lar}

Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{>}$. Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority. Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$. Assume that $G$ has no cycle going through $x$. Then, $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}(e,x)$.
\end{restatable}

\begin {proof}

We prove this lemma by induction, and prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x)$, in $\textit{UVSet}_2(e,x)$, $\ldots$. Note that, by Lemma \ref{lemma:UVSet has matched put and rm}, $\textit{UVSet}(e,x)$ contains only matched $\textit{put}$ and $\textit{rm}$, and it is easy to see that for each $i$, $\textit{UVSet}_i(e,x)$ contains only matched $\textit{put}$ and $\textit{rm}$.

\noindent (1) Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x)$ by contradiction. Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{UVSet}_1(e,x)$ is an operation of item $d$. %(according to the definition of $\textit{UVSet}_1(h,x)$, the priority of $d$ does not equals $\textit{pri}_x$).

We use a triple $(t_1,t_2,t_3)$ to represent related information. $t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$. $t_1$ represents whether $o$ is a $\textit{put}$ method event or a $\textit{rm}$ method event. $t_2$ and $t_3$ is used for the reason of $o \in \textit{UVSet}_1(e,x)$: $o \in \textit{UVSet}_1(e,x)$, since an operation (of kind $t_2$) of $d$ happens before an operation (of kind $t_3$) of $x$. Let us consider all the possible cases of $(t_1,t_2,t_3)$:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.

\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{rm}(x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{rm}(x)$.

\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x) )$. By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(x,\textit{pri}_x) ) \vee ( \textit{rm}(d) <_{hb} \textit{put}(d,\_) )$, which is impossible.

\item[-] $(\textit{put},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{rm}(x) )$. We can see that $\textit{rm}(d) <_{hb} \textit{rm}(x) <_{hb} \textit{put}(d,\_)$, which contradicts that $\textit{rm}(d)$ does not happen before $\textit{put}(d,\_)$.

\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x) )$. We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.

\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{rm}(x) )$. We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.

\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.

\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{rm}(x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{rm}(x)$.
\end{itemize}

This completes the proof for $\textit{UVSet}_1(e,x)$.

\noindent (2) Assume we already prove that for some $j \geq 1$, $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$. Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_{\textit{j+1}}(e,x)$ by contradiction. Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$ is an operation of item $d_{\textit{j+1}}$. We use a triple $(t_1,t_2,t_3)$ to represent related information. $t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$. $t_1$ represents whether $o$ is a $\textit{put}$ method event or a $\textit{rm}$ method event. $t_2$ and $t_3$ is used for the reason of $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$: $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$, since an operation (of kind $t_2$) of $d_{\textit{j+1}}$ happens before an operation (of kind $t_3$) of $d_j$, where $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{UVSet}_j(e,x)$. Let us consider all the possible cases of $(t_1,t_2,t_3)$:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_)$. We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{UVSet}_j(e,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$.

\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j,\_)$. We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j,\_) ) \wedge ( \textit{rm}(d_j) \in \textit{UVSet}_j(e,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$.

\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$. By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \vee ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_{\textit{j+1}},\_) )$, which is impossible.

\item[-] $(\textit{put},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$. By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \vee ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_{\textit{j+1}},\_) )$, which is impossible.

\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) )$. Let us consider the reason of $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{UVSet}_j(e,x)$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $( j > 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{UVSet}_{\textit{j-1}}(e,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, we can see that $\textit{put}(d_{\textit{j+1}},\_) <_{hb} o''$, and then operations of $d_{\textit{j+1}}$ is in $\textit{UVSet}_j(e,x)$, contradicts that operations of $d_{\textit{j+1}}$ is in $\textit{UVSet}_{\textit{j+1}}(e,x)$.

    \item[-] If $( j = 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of $x$: Similar to above case.

    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{UVSet}_{\textit{j-1}}(e,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, we can see that $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} o'' ) \vee ( \textit{rm}(d_j) <_{hb} \textit{put}(d_j,\_) )$, which is impossible.

    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of $x$: Similar to above case.
    \end{itemize}

\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Let $T_{\textit{ind}}$ be the set of sentences $\{ \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}), \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j),\ldots, \textit{put}(d_{\textit{ind+1}},\_) <_{hb} \textit{rm}(d_{\textit{ind}}) \}$. Here each $d_i$ is a item of some operation in $\textit{UVSet}_i(e,x)$. Let us prove that from $T_j$ we can obtain contradiction by induction:

    {\bf Base case $1$}: From $T_1$ we can obtain contradiction.

    Let us prove base case $1$:

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $\textit{put}(d_1,\_)$ happens $o$, and $o$ is an operation of $x$. Then there is a cycle $x \rightarrow d_{\textit{j+1}} \rightarrow \ldots \rightarrow d_1 \rightarrow x$ in $G$, contradicts that $G$ has no cycle going through $x$.

    \item[-] If $\textit{rm}(d_1)$ happens before $o$, and $o$ is an operation of $x$. Then since $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1) <_{hb} o$, we can see that $\textit{put}(d_2,\_) <_{hb} o$, and then $\textit{put}(d_2,\_) \in \textit{UVSet}_1(e,x)$, contradicts that $\textit{put}(d_2,\_) \in \textit{UVSet}_2(e,x)$.
    \end{itemize}

    {\bf Base case $2$}: From $T_2$ we can obtain contradiction.

    Let us prove base case $2$: If $\textit{rm}(d_2) <_{hb} o$, and $o$ is an operation of $d_1$, then since $( \textit{put}(d_3,\_) <_{hb} \textit{rm}(d_2) ) \wedge ( \textit{rm}(d_2) <_{hb} o )$, we know that $\textit{put}(d_3,\_) <_{hb} o$. This implies that $\textit{put}(d_3,\_) \in \textit{UVSet}_2(e,x)$, contradicts that $\textit{rm}(d_3,\_) \in \textit{UVSet}_3(e,x)$. Therefore, it is only possible that $\textit{put}(d_2,\_)$ happens before an operation of $d_1$.

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{put}(d_1,\_)$ and $\textit{put}(d_1,\_)$ happens before operations of $x$, then we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, which is impossible.

    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{put}(d_1,\_)$ and $\textit{rm}(d_1)$ happens before operations of $x$, then by interval order, we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, or $\textit{rm}(d_1) <_{hb} \textit{put}(d_1,\_)$, which is impossible.

    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{put}(d_1,\_)$ happens before operations of $x$, then $x \rightarrow d_{\textit{j+1}} \rightarrow \ldots \rightarrow d_1 \rightarrow x$ in $G$, contradicts that $G$ has no cycle going through $x$.

    \item[-] If $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1)$ happens before operations of $x$, then we know that $\textit{put}(d_2,\_)$ happens before operation of $x$, which is impossible.
    \end{itemize}

    {\bf induction step}: Given $\textit{ind} \geq 3$, if from $T_{\textit{ind-1}}$ we can obtain contradiction, then from $T_{\textit{ind}}$ we can also contain contradiction.

    Prove of the induction step: Similarly as base case $2$, we can prove that it is only possible that $\textit{put}(d_{\textit{ind}},\_)$ happens before operations of $d_{\textit{ind-1}}$.

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$ and $\textit{put}(d_{\textit{ind-1}},\_)$ happens before operations of $d_{\textit{ind-2}}$, then we know that $\textit{put}(d_{\textit{ind}})$ happens before operation of $d_{\textit{ind-2}}$, which is impossible.

    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$ and $\textit{rm}(d_{\textit{ind-1}})$ happens before operations of $d_{\textit{ind-2}}$, then by interval order, we know that $\textit{put}(d_{\textit{ind}},\_)$ happens before operation of $d_{\textit{ind-2}}$, or $\textit{rm}(d_{\textit{ind-1}}) <_{hb} \textit{put}(d_{\textit{ind-1}},\_)$, which is impossible.

    \item[-] If $\textit{put}(d_{\textit{ind}},\_) <_{hb} \textit{rm}(d_{\textit{ind-1}})$, then we obtain $T_{\textit{ind-1}}$, which already contain contradiction.
    \end{itemize}

    By base case $1$, base case $2$ and the induction step, it is easy to see that for each $i$, $T_i$ contains contradiction. Therefore, $T_j$, the case of $(\textit{rm},\textit{put},\textit{rm})$, contains contradiction.

\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$. We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{UVSet}_j(e,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$.

\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$. We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \wedge ( \textit{rm}(d_j) \in \textit{UVSet}_j(e,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$.
\end{itemize}

This completes the proof for $\textit{UVSet}_{\textit{j+1}}(e,x)$. Therefore, $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}(e,x) = \textit{UVSet}_1(e,x) \cup \textit{UVSet}_2(e,x) \cup \ldots$. \qed
\end {proof}

With Lemma \ref{lemma:UVSet has matched put and rm} and Lemma \ref{lemma:Rmx does not happen before UVSet for EPQ1Lar}, we can now prove Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar}.

\LinEqualsConstraintforEPQOneLar*

\begin {proof}

To prove the $\textit{only if}$ direction, assume that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$. Let $u$, $v$ and $w$ be the sequences of method events in $\textit{EPQ}_1^{>}$, and let $U$, $V$ and $W$ be the set of method events of $u$, $v$ and $w$, respectively. Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ in $G$. It is obvious that the priority of each $d_i$ is smaller than $\textit{pri}_x$. Then our proof proceeds as follows:

According to the definition of left-right constraint, there are two possibilities. The first possibility is that, $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$. It is obvious that $\textit{rm}(d_1) \in W$, and then since $U \cup V$ contains matched $\textit{put}$ and $\textit{rm}$, we can see that $\textit{put}(d_1),\textit{rm}(d_1) \in W$. Then,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Since $d_1 \rightarrow d_2$, by definition of $G$, we know that $\textit{put}(d_1)$ happens before $\textit{rm}(d_2)$. Since $\textit{put}(d_1) \in W$ and $U \cup V$ contains matched $\textit{put}$ and $\textit{rm}$, we know that $\textit{put}(d_2),\textit{rm}(d_2) \in W$. Similarly, for each $1 \leq i \leq m$, we know that $\textit{put}(d_i),\textit{rm}(d_i) \in W$.

\item[-] Since $d_m \rightarrow x$,
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] if $\textit{put}(d_m)$ happens before $\textit{put}(x)$, then we can see that $\textit{put}(d_m) \in U$, which contradicts that $\textit{put}(d_m) \in W$.

    \item[-] if $\textit{put}(d_m)$ happens before $\textit{rm}(x)$, then we can see that $\textit{put}(d_m) \in U \cup V$, which contradicts that $\textit{put}(d_m) \in W$.
    \end{itemize}
\end{itemize}

The second possibility is that, $e$ contains one $\textit{put}(d_1,\_)$ and no $\textit{rm}(d_1)$. Note that for each $j > 1$, $e$ contains $\textit{put}(d_j,\_)$ and $\textit{rm}(d_j)$. Since $d_m \rightarrow x$, is is obvious that $\textit{put}(d_m) \in U \cup V$. Since $U \cup V$ contains matched $\textit{put}$ and $\textit{rm}$, we know that $\textit{put}(d_m),\textit{rm}(d_m) \in U \cup V$. Then, since $d_{\textit{m-1}} \rightarrow d_m$, by definition of $G$, we know that $\textit{put}(d_{\textit{m-1}})$ happens before $\textit{rm}(d_m)$. Since $\textit{rm}(d_m) \in U \cup V$ and $U \cup V$ contains matched $\textit{put}$ and $\textit{rm}$, we know that $\textit{put}(d_{\textit{m-1}}),\textit{rm}(d_{\textit{m-1}}) \in U \cup V$. Similarly, for each $1 < i \leq m$, we know that $\textit{put}(d_i),\textit{rm}(d_i) \in U \cup V$, and also $\textit{put}(d_1)\in U \cup V$. However, there is one $\textit{put}(d_1,\_)$ and no $\textit{rm}(d_1)$ in $e$, contradicts that $U \cup V$ contains matched $\textit{put}$ and $\textit{rm}$.

This completes the proof of the $\textit{only if}$ direction.

To prove the $\textit{if}$ direction, assume that $G$ has no cycle going through $x$. Let $E_u$ be the set of operations that happen before $\textit{put}(x)$ in $e$. It is easy to see that $E_u \subseteq \textit{UVSet}(e,x)$. Let $E_v = \textit{UVSet}(e,x) \setminus E_u$. Let $E_e$ be the set of operations of $e$, and let $E_w = E_e \setminus \textit{UVSet}(e,x)$.

By Lemma \ref{lemma:UVSet has matched put and rm}, we can see that $E_u \cup E_v$ contains matched $\textit{put}$ and $\textit{rm}$ operations. It remains to prove that for $E_u$, $\{ \textit{put}(x,\textit{pri}_x) \}$, $E_v$, $\{ \textit{rm}(x) \}$, $E_w$, no elements of the latter set happens before elements of the former set. We prove this by showing that all the following cases are impossible:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Case $1$: Some operation $o_w \in E_w$ happens before $\textit{rm}(x)$. Then we know that $o_w \in \textit{UVSet}(e,x) = E_u \cup E_v$, which contradicts that $o_w \in E_w$.

\item[-] Case $2$: Some operation $o_w \in E_w$ happens before some operation $o_{\textit{uv}} \in E_u \cup E_v$. Then we know that $o_w \in \textit{UVSet}(e,x) = E_u \cup E_v$, which contradicts that $o_w \in E_w$.

\item[-] Case $3$: Some operation $o_w \in E_w$ happens before $\textit{put}(x)$. Then we know that $o_w \in \textit{UVSet}(e,x) = E_u \cup E_v$, which contradicts that $o_w \in E_w$.

\item[-] Case $4$: $\textit{rm}(x)$ happens before some $o_{\textit{uv}} \in \textit{UVSet}(e,x) = E_u \cup E_v$. By Lemma \ref{lemma:Rmx does not happen before UVSet for EPQ1Lar} we know that this is impossible.

\item[-] Case $5$: $\textit{rm}(x)$ happens before $\textit{put}(x)$. This contradicts that each single-priority projection satisfy the FIFO property.

\item[-] Case $6$: Some operation $o_v \in E_v$ happens before $\textit{put}(x)$. Then we know that $o_v \in E_u$, which contradicts that $o_v \in E_v$.

\item[-] Case $7$: Some operation $o_v \in E_v$ happens before some operation $o_u \in E_u$. Then we know that $o_v \in E_u$, which contradicts that $o_v \in E_v$.

\item[-] Case $8$: $\textit{put}(x)$ happens before some operation $o_u \in E_u$. This is impossible.
\end{itemize}

This completes the proof of the $\textit{if}$ direction.

\qed
\end {proof}

Let us begin to represent witness automata that is used for capture the existence of a data-differentiated execution $e$, $e$ has a $\_$-projection $e'$, $\textit{last}(e') = \textit{PQ}_1^{>}$, and there exists a cycle going through the item with maximal priority in $e'$. By data-independence, we can obtain $e_r$ from $e$ by renaming function, which maps such item to be $b$, maps items that cover it to be $a$, and maps other items into $d$. There are four possible enumeration of call and return actions of $\textit{put}(b)$ and $\textit{rm}(b)$. For each of them, we generate a witness automaton.

For the case when $e_r \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^1$, as shown in \figurename~\ref{fig:automata APQ1Lar-1}. Here $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{cal}(\textit{put},a,\textit{les}_p)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$. The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^1$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1 \textwidth]{figures/PIC_AUTO_PQ1Lar-pprr.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{l-lar}}^1$}
  \label{fig:automata APQ1Lar-1}
\end{figure}

$\mathcal{A}_{\textit{l-lar}}^1$ is used to recognize conditions in \figurename~\ref{fig:his for APQ1Lar-1}. Here for simplicity, we only draw operation of $b$, and the first $\textit{ret}(\textit{put},a)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1 \textwidth]{figures/PIC_HIS_PQ1Lar-pprr.pdf}
%\vspace{-10pt}
  \caption{Conditions recognized by $\mathcal{A}_{\textit{l-lar}}^1$}
  \label{fig:his for APQ1Lar-1}
\end{figure}

For the case when $e_r \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^2$, as shown in \figurename~\ref{fig:automata APQ1Lar-2}. Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$. The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^2$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1 \textwidth]{figures/PIC_AUTO_PQ1Lar-prpr.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{l-lar}}^2$}
  \label{fig:automata APQ1Lar-2}
\end{figure}

$\mathcal{A}_{\textit{l-lar}}^2$ is used to recognize conditions in \figurename~\ref{fig:his for APQ1Lar-2}. Here for simplicity, we only draw operation of $b$, and the first $\textit{ret}(\textit{put},a)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.7 \textwidth]{figures/PIC_HIS_PQ1Lar-prpr.pdf}
%\vspace{-10pt}
  \caption{Conditions recognized by $\mathcal{A}_{\textit{l-lar}}^2$}
  \label{fig:his for APQ1Lar-2}
\end{figure}

For the case when $e_r \vert_{b} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^3$, as shown in \figurename~\ref{fig:automata APQ1Lar-3}. Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$. The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^3$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1 \textwidth]{figures/PIC_AUTO_PQ1Lar-rppr.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{l-lar}}^3$}
  \label{fig:automata APQ1Lar-3}
\end{figure}

$\mathcal{A}_{\textit{l-lar}}^3$ is used to recognize conditions in \figurename~\ref{fig:his for APQ1Lar-3}. Here for simplicity, we only draw operation of $b$, and the first $\textit{ret}(\textit{put},a)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.7 \textwidth]{figures/PIC_HIS_PQ1Lar-rppr.pdf}
%\vspace{-10pt}
  \caption{Conditions recognized by $\mathcal{A}_{\textit{l-lar}}^3$}
  \label{fig:his for APQ1Lar-3}
\end{figure}

For the case when $e_r \vert_{b} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b) \cdot \textit{ret}(\textit{put})$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^4$, as shown in \figurename~\ref{fig:automata APQ1Lar-4}. Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$, and $c_4 = c_1 + \textit{ret}(\textit{put},b)$. The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^4$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.9 \textwidth]{figures/PIC_AUTO_PQ1Lar-rprp.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{l-lar}}^4$}
  \label{fig:automata APQ1Lar-4}
\end{figure}

$\mathcal{A}_{\textit{l-lar}}^4$ is used to recognize conditions in \figurename~\ref{fig:his for APQ1Lar-4}. Here for simplicity, we only draw operation of $b$, and the first $\textit{ret}(\textit{put},a)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.7 \textwidth]{figures/PIC_HIS_PQ1Lar-rprp.pdf}
%\vspace{-10pt}
  \caption{Conditions recognized by $\mathcal{A}_{\textit{l-lar}}^4$}
  \label{fig:his for APQ1Lar-4}
\end{figure}

Let $\textit{Auts}_{\textit{1-lar}} = \{ \mathcal{A}_{\textit{l-lar}}^1, \mathcal{A}_{\textit{l-lar}}^2, \mathcal{A}_{\textit{l-lar}}^3, \mathcal{A}_{\textit{l-lar}}^4 \}$. The following lemma states that $\textit{EPQ}_1^{>}$ is co-regular.

\EPQOneLarisCoRegular*

\begin {proof}

We need to prove that, given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-lar}} \cap \mathcal{I} \neq \emptyset$, if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e),$ $\textit{EPQ}_1^{>} \in last(e') \wedge e'$ does not linearizable w.r.t. $\textit{MS}(\textit{EPQ}_1^{>})$.

By Lemma \ref{lemma:pri execution is enough} and Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar}, we need to prove the following fact:

\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-lar}} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e)$, $last(e')=\textit{EPQ}_1^{>}$, $x$ is the item with maximal priority $\textit{pri}$ in $e'$, $e'$ is a $\textit{pri}$-execution. And there is a cycle going through $x$ in $G$, where $G$ is the left-right constraint of $e'$.

\noindent The $\textit{only if}$ direction: Let us consider the case of $\mathcal{A}_{\textit{l-lar}}^1$. Assume that $e_1 \in \mathcal{I}$ is accepted by $\mathcal{A}_{\textit{l-lar}}^1$. By data-independence, there exists data-differentiated execution $e \in \mathcal{I}$ and renaming function $r_1$, such that $e_1 = r_1(e)$. Assume that $r_1$ maps $d$ into $b$ and maps $f_1,\ldots,f_m$ into $a$. Let $e'$ be obtained from $e$ by projection into $\{ d, f_1,\ldots,f_m \}$. Assume that the priority of $b$ is $p$. It is easy to see that $e'$ is a $p$-execution, $\textit{last}(e') = \textit{EPQ}_1^{>}$, and there is a cycle going through $d$ in $G$, where $G$ is the left-right constraint of $e'$. The case of $\mathcal{A}_{\textit{l-lar}}^2$, $\mathcal{A}_{\textit{l-lar}}^3$ and $\mathcal{A}_{\textit{l-lar}}^4$ can be similarly proved.

\noindent The $\textit{if}$ direction: Given such $e$, $e'$ and $x$. Let renaming function $r$ maps $x$ into $b$, maps items cover $x$ into $a$, and maps other items into $d$. By data-independence, $r(e) \in \mathcal{I}$. Then depending on the cases of $r(e) \vert_{b}$, we can see that $r(e)$ is accepted by $\mathcal{A}_{\textit{l-lar}}^1$, $\mathcal{A}_{\textit{l-lar}}^2$, $\mathcal{A}_{\textit{l-lar}}^3$ or $\mathcal{A}_{\textit{l-lar}}^4$. \qed
\end {proof}

\subsection{Proofs and Definitions in Subsection \ref{subsec:co-regular of EPQ1Equal}}
\label{sec:appendix proof and definition in section co-regular of EPQ1Equal}

Let $\textit{Items}(e,p)$ be the set of items with priority $p$ in execution $e$. The following lemma states a method to choose $\textit{itm}$ of $\textit{EPQ}_1^{=}$.

\begin{restatable}{lemma}{MaximalInPBadGPMakePQ1Equal}
\label{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}
Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{=}$. If there exists an item $x$ with priority $\textit{pri}$, such that for each $y \in \textit{Items}(e,\textit{pri})$, (1) $x$ does not $<_{\textit{pb}}$ to $y$, and (2) the right-most gap-point of $x$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},y)$. Then $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
\end{restatable}

\begin {proof}

Let $o$ be the right-most gap-point of $x$. We locate linearization points of each method event as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Locate the linearization point of $\textit{rm}(x)$ at $o$,

\item[-] If $\textit{put}(x,\textit{pri})$ overlaps with $\textit{rm}(x)$, then locate the linearization point of $\textit{put}(x,\textit{pri})$ just before the linearization point of $\textit{rm}(x)$. Otherwise, $\textit{put}(x,\textit{pri}) <_{\textit{hb}} \textit{rm}(x)$, and we locate the linearization point of $\textit{put}(x,\textit{pri})$ just before its return action.

\item[-] Locate linearization points of method event of each $y \in \textit{Items}(e,\textit{pri})$ (except for $x$) just after the call action of the method event.

\item[-] For item $z$ with priority smaller than $\textit{pri}$. If both $\textit{cal}(\textit{put},z,\_)$ and $\textit{cal}(\textit{rm},z)$ is before $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just after their call actions. If both $\textit{ret}(\textit{put},z)$ and $\textit{ret}(\textit{rm},z)$ (if exists) is after $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just before their return actions. Otherwise, $x$ is in interval of $z$, which contradicts the definition of gap-point, and is impossible.
\end{itemize}

Let $l$ be the sequence of linearization points constructed above. It is obvious that $e \sqsubseteq l$. Since for each $y \in \textit{Items}(e,\textit{pri})$, $o$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},x)$, we can see that $\textit{rm}(x)$ is after $\textit{put}(y,\textit{pri})$ and $\textit{rm}(y)$ in $l$. It is obvious that $\textit{put}(x,\textit{pri})$ is before $\textit{rm}(x)$ in $l$. Since $x$ does not $<_{\textit{pb}}$ to $y$, we can see that no $\textit{put}(y,\textit{pri})$ happens before $\textit{put}(x,\textit{pri})$. Then it is easy to see that $\textit{put}(x,\textit{pri})$ is after $\textit{put}(y,\textit{pri})$ in $l$. Since $\textit{last}(e) = \textit{EPQ}_1^{=}$, all other items in $\textit{Items}(e,\textit{pri})$ has matched $\textit{put}$ and $\textit{rm}$, and it is easy to see that their $\textit{put}$ and $\textit{rm}$ (except for that of $x$) are all before $\textit{rm}(x)$ in $l$.

For item $z$ with priority smaller than $\textit{pri}$, we can see that there are only two possibilities: (1) $\textit{put}(z,\_)$ and $\textit{rm}(z)$ are both before $\textit{rm}(x)$ in $l$, and (2) $\textit{put}(z,\_)$ and $\textit{rm}(z)$ (if exists) are after before $\textit{rm}(x)$ in $l$. Therefore, before $\textit{rm}(x)$ in $l$, the $\textit{put}$ and $\textit{rm}$ of $z$ are matched.

Therefore, it is easy to see that $l \in \textit{MS}(\textit{EPQ}_1^{=})$. \qed
\end {proof}

With Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, we can prove the following lemma, which states that getting rid of case in \figurename~\ref{fig:introduce pb order} is enough for ensure $\textit{last}(e) = \textit{EPQ}_1^{=} \Rightarrow e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.

\EPQOneEqualAsPBandGP*

\begin {proof}

To prove the $\textit{if}$ direction, let $e_{x,y}$ be the execution that is obtained from $e$ by erasing all actions of items that has same priority as $x$, except for actions of $x$ and $y$. It is obvious that $\textit{last}(e_{x,y}) = \textit{EPQ}_1^{=}$. Since $y <_{\textit{pb}}^* x$, according to $\textit{EPQ}_1^{=}$, we can see that $x$ should be chosen as $\textit{itm}$ in $\textit{EPQ}_1^{=}$.

According to Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar} (Here we temporarily forget the existence of $y$), the only possible position for locating linearizaton point of $\textit{rm}(x)$ is at gap-point of $x$. Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $x$, then there exists unmatched method event before $\textit{rm}(x)$ with smaller priority. Since the rightmost gap-point of $x$ is before $\textit{cal}(\textit{put},y,\textit{pri})$ or $\textit{cal}(\textit{rm},y)$, if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $x$, then $\textit{rm}(x)$ will be before $\textit{cal}(\textit{put},y,\textit{pri})$ or $\textit{cal}(\textit{rm},x)$.

Therefore, for every sequence $l = u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$, if $e_{x,y} \sqsubseteq l$, then either $u \cdot v$ contains some unmatched method events of priority smaller than $\textit{pri}$, or $w$ contains $\textit{put}(y,\textit{pri})$ or $\textit{rm}(y)$. In both cases, $l \notin \textit{MS}(\textit{EPQ}_1^{=})$.

To prove the $\textit{only if}$ direction, we prove its contrapositive. Assume we already know that for each $x$ and $y$ has maximal priority in $e$, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{cal}(\textit{put},y,\textit{pri})$ and $\textit{cal}(\textit{rm},x)$. We need to prove that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$. Recall that we already assume that each single-priority execution has FIFO property, and item with larger priority is not covered by items with smaller priority.

Our proof proceed as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Let $e_{\textit{pri}}$ be the projection of $e$ into operations of priority $\textit{pri}$. Since each single-priority execution has FIFO property, there exists sequence $l_{\textit{pri}}$, such that $e_{\textit{pri}} \sqsubseteq l_{\textit{pri}}$, and when we treat $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$, $l_{\textit{pri}}$ belongs to queue.

\item[-] Let $a_1$ be the last inserted item of $l_{\textit{pri}}$.

    Step $1$: Check whether for each $b \in \textit{Items}(e,\textit{pri})$, (1) $a_1$ does not $<_{\textit{pb}}$ to $b$, and (2) the right-most gap-point of $a$ is after $\textit{cal}(\textit{put},b,\textit{pri})$ and $\textit{cal}(\textit{rm},b)$.

    It is easy to see that $a_1$ is of priority $\textit{pri}$, and $a_1$ does not $<_{\textit{pb}}$ to any $b \in \textit{Items}(e,\textit{pri})$. If for each $b \in \textit{Items}(e,\textit{pri})$, the rightmost gap-point of $a_1$ is after $\textit{cal}(\textit{put},b,\textit{pri})$ and $\textit{cal}(\textit{rm},b)$. Then by Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, we can obtain that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.

\item[-] Otherwise, there exists $a_2 \in \textit{Items}(e,\textit{pri})$, such that the rightmost gap-point of $a_1$ is before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$ in $e$. We can see that each gap-point of $a_2$ is after the rightmost gap-point of $a_1$.%, and thus, the right-most gap-point of $a_2$ is after the rightmost gap-point of $a_1$.
    By assumption, we know that $a_2$ does not $<_{\textit{pb}}$ to $a_1$.

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If for each item $b \in \textit{Items}(e,\textit{pri})$, $a_2$ does not $<_{\textit{pb}}$ to $b$. Then we go to step $1$ and treat $a_2$ similarly as $a_1$.
    \item[-] Otherwise, there exists $a_3$ with priority $\textit{pri}$ such that $a_2 <_{\textit{pb}}^* a_3$.

    Since $l_{\textit{pri}}$ has FIFO property, it is easy to see that there is no cycle in $<_{\textit{pb}}$ order. It is safe to assume that $a_3$ is maximal in the sense of $<_{\textit{pb}}^*$. Or we can say, there does not exists $a_4$, such that $a_3 <_{\textit{pb}}^* a_4$.

    By assumption,we know that the rightmost gap-point of $a_3$ is after $\textit{cal}(\textit{put},a_2,\textit{pri})$ and $\textit{cal}(\textit{rm},a_2)$. Therefore, we can see that the rightmost gap-point of $a_3$ is after the rightmost gap-point of $a_1$. Then we go to step $1$ and treat $a_3$ similarly as $a_1$.
    \end{itemize}
\end{itemize}

Let $a^i$ be the $a_1$ in the $\textit{i-th}$ loop of our proof. It is not hard to see that, given $i<j$, the rightmost gap-point of $a^j$ is after the rightmost gap-point of $a^i$. Therefore, the loop finally stop at some $a^f$. $a^f$ satisfies the check of Step $1$. By Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, this implies that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$. This completes the proof of $\textit{if}$ direction. \qed
\end {proof}

According to the definition of $<_{\textit{ob}}^*$, if $a <_{\textit{pb}}^* b$, then there exists $a_1,\ldots,a_m$, such that $a <_{\textit{pb}} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}} a_m <_{\textit{pb}} b$. The following lemma states that, the number of intermediate items $a_i$ is in fact bounded.

\OBOrderHasBoundedLength*

\begin {proof}

Our proof proceed as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] ($<_{\textit{pb}}^A \cdot <_{\textit{pb}}^A$,$<_{\textit{pb}}^B \cdot <_{\textit{pb}}^B$ and $<_{\textit{pb}}^C \cdot <_{\textit{pb}}^C$): If $c_3 <_{\textit{pb}}^A c_2 <_{\textit{pb}}^A c_1$, then $\textit{put}(c_3,\_)$ happens before $\textit{put}(c_2,\_)$, and $\textit{put}(c_2,\_)$ happens before $\textit{put}(c_1,\_)$. Therefore, it is obvious that $\textit{put}(c_3,\_)$ happens before $\textit{put}(c_1,\_)$ and $c_3 <_{\textit{pb}}^A c_1$.

    Similarly, if $c_3 <_{\textit{pb}}^B c_2 <_{\textit{pb}}^B c_1$, then $c_3 <_{\textit{pb}}^B c_1$.

    If $c_3 <_{\textit{pb}}^C c_2 <_{\textit{pb}}^C c_1$: Since $c_2 <_{\textit{pb}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$. Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$. Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$. Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.

    Therefore, when we meet successive $<_{\textit{pb}}^A$, it is safe to leave only the first and the last elements and ignore intermediate elements. Similar cases hold for $<_{\textit{pb}}^B$ and $<_{\textit{pb}}^C$.

\item[-] $<_{\textit{pb}}^A$ and $<_{\textit{pb}}^C$:

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] ($<_{\textit{pb}}^A \cdot <_{\textit{pb}}^C$): If $c_3 <_{\textit{pb}}^A c_2 <_{\textit{pb}}^C c_1$. Since $c_2 <_{\textit{pb}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$. Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$. Since $c_3 <_{\textit{pb}}^A c_2$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$. Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^A c_1$.

    \item[-] ($<_{\textit{pb}}^C \cdot <_{\textit{pb}}^A$): If $c_3 <_{\textit{pb}}^C c_2 <_{\textit{pb}}^A c_1$. Since $c_2 <_{\textit{pb}}^A c_1$, $\textit{ret}(\textit{put},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$. It is obvious that $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{put},c_2)$. Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$. Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.

    %Since $c_2 <_{\textit{pb}}^A c_1$, $\textit{put}(c_2,\_)$ happens before $\textit{put}(c_1,\_)$. Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{rm}(c_3)$ happens before $\textit{put}(c_2,\_)$. Therefore, $\textit{rm}(c_3)$ happens before $\textit{put}(c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.
    \end{itemize}

\item[-] $<_{\textit{pb}}^B$ and $<_{\textit{pb}}^C$:

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] ($<_{\textit{pb}}^B \cdot <_{\textit{pb}}^C$): If $c_3 <_{\textit{pb}}^B c_2 <_{\textit{pb}}^C c_1$. Since $c_2 <_{\textit{pb}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$. It is obvious that $\textit{cal}(\textit{rm},c_2)$ is before $\textit{ret}(\textit{rm},c_2)$. Since $c_3 <_{\textit{pb}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$. Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.

    \item[-] ($<_{\textit{pb}}^C \cdot <_{\textit{pb}}^B$): If $c_3 <_{\textit{pb}}^C c_2 <_{\textit{pb}}^B c_1$. Since $c_2 <_{\textit{pb}}^B c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{rm},c_1)$. Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$. Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$. Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_1)$, and $c_3 <_{\textit{pb}}^B c_1$.
    \end{itemize}

\item[-]  ($<_{\textit{pb}}^A \cdot <_{\textit{pb}}^B \cdot <_{\textit{pb}}^A$): If $c_4 <_{\textit{pb}}^A c_3 <_{\textit{pb}}^B c_2 <_{\textit{pb}}^A c_1$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$: Since $c_3 <_{\textit{pb}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$. Then $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$. This implies that $c_4 <_{\textit{pb}}^A c_3 <_{\textit{pb}}^C c_1$. According to the fact for $<_{\textit{pb}}^A \cdot <_{\textit{pb}}^C$, we know that $c_4  <_{\textit{pb}}^A c_1$.

    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is after $\textit{cal}(\textit{put},c_1,\_)$: Since $c_2 <_{\textit{pb}}^A c_1$, $\textit{ret}(\textit{put},c_2,\_)$ is before $\textit{cal}(\textit{put},c_1,\_)$. Since $c_3 <_{\textit{pb}}^B c_2$, $\textit{rm}(c_3)$ happens before $\textit{rm}(c_2)$, and then we know that $\textit{put}(c_2,\_)$ can not happen before $\textit{put}(c_3,\_)$. Since $\textit{put}(c_2,\_)$ does not happen before $\textit{put}(c_3,\_)$, $\textit{cal}(\textit{put},c_3,\_)$ is before $\textit{ret}(\textit{put},c_2,\_)$. Since $c_4 <_{\textit{pb}}^A c_3$, $\textit{ret}(\textit{put},c_4)$ is before $\textit{cal}(\textit{put},c_3,\_)$. Therefore, $\textit{ret}(\textit{put},c_4)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_4 <_{\textit{pb}}^A c_1$.
    \end{itemize}

\item[-]  ($<_{\textit{pb}}^B \cdot <_{\textit{pb}}^A \cdot <_{\textit{pb}}^B$): If $c_4 <_{\textit{pb}}^B c_3 <_{\textit{pb}}^A c_2 <_{\textit{pb}}^B c_1$: Since $c_2 <_{\textit{pb}}^B c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{rm},c_1)$. Since $c_3 <_{\textit{pb}}^A c_2$, we can see that $\textit{put}(c_3,\_) <_{\textit{hb}} \textit{put}(c_2,\_)$. Since each single-priority execution has FIFO property, we know that $\textit{rm}(c_2)$ does not happen before $\textit{rm}(c_3)$, and thus, $\textit{cal}(\textit{rm},c_3)$ is before $\textit{ret}(\textit{rm},c_2)$. Since $c_4 <_{\textit{pb}}^B c_3$, $\textit{ret}(\textit{rm},c_4)$ is before $\textit{cal}(\textit{rm},c_3)$. Therefore, $\textit{ret}(\textit{rm},c_4)$ is before $\textit{cal}(\textit{rm},c_1)$, and $c_4 <_{\textit{pb}}^B c_1$.

\end{itemize}

Based on above results, given $a <_{\textit{pb}}^{b_1} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}}^{b_m} a_m <_{\textit{pb}}^{b_{\textit{m+1}}} b$, where each $b_i$ is in $\{ A,B,C \}$, we can merge relations, until we got one of the following facts:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$ or $a <_{\textit{pb}}^C b$,

\item[-] $a <_{\textit{pb}}^A a_i <_{\textit{pb}}^B b$, or $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A b$, for some $i$,
\end{itemize}

This completes the proof of this lemma. \qed
\end {proof}

There are many enumerations of method events of $a$, $b$ and $a_1$ that may makes $a <_{\textit{pb}}^* b$. The following lemma states that with the help of gap-points, the number of potential enumerations can be further reduced into only five.

\FiveEnmuerationisEnoughForEPQOneEqual*

\begin {proof}

Let us prove by consider all the possible reason of $a <_{\textit{pb}}^* b$. According to Lemma \ref{lemma:ob order has bounded length}, we need to consider five reasons: Let $o$ be the right-most gap-point of $b$.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Reason $1$, $a <_{\textit{pb}}^A b$:

    Since $a <_{\textit{pb}}^A b$, $\textit{put}(a,\textit{pri}) <_{\textit{hb}} \textit{put}(b,\textit{pri})$. Since $o$ is after $\textit{cal}(\textit{put},b,\textit{pri})$, and thus, after $\textit{cal}(\textit{put},a,\textit{pri})$, we can see that $o$ is before $\textit{cal}(\textit{rm},b)$.

    Since single-priority execution must satisfy the FIFO property, $\textit{rm}(b)$ does not happen before $\textit{rm}(a)$, and thus, $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$. If $\textit{cal}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$, then $o$ is also a gap-point of $a$ and contradicts our assumption. So we know that $\textit{cal}(\textit{rm},a)$ is after $\textit{cal}(\textit{rm},b)$. If $\textit{ret}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$, since we already assume that there exists gap-point of $a$, this gap-point is also a gap-point of $b$, and is after $o$, which contradicts that $o$ is the rightmost gap-point of $b$. Therefore, $\textit{ret}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$.

    According to above discussion, there are two possible enumeration of operations of $a$ and $b$, as shown in \figurename~\ref{fig:history enumeration 1 for PQ1Equal} and \figurename~\ref{fig:history enumeration 2 for PQ1Equal}. Here we explicitly draw the leftmost gap-point of $a$ as $o'$. Since the position of $\textit{ret}(\textit{put},b)$ does not influence the correctness, we can simply ignore it.

\item[-] Reason $2$, $a <_{\textit{pb}}^B b$:

    Since $a <_{\textit{pb}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$. Since $o$ is after $\textit{cal}(\textit{rm},b)$, we can see that $o$ is before $\textit{cal}(\textit{put},a,\textit{pri})$. This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\textit{pri})$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible. Therefore, we can safely ignore this reason.

\item[-] Reason $3$, $a <_{\textit{pb}}^C b$:

    Since $a <_{\textit{pb}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},b,\textit{pri})$. Since $o$ is after $\textit{cal}(\textit{put},b)$, we can see that $o$ is before $\textit{cal}(\textit{put},a,\textit{pri})$. This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\textit{pri})$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible. Therefore, we can safely ignore this reason.

\item[-] Reason $4$, $a <_{\textit{pb}}^A a_1 <_{\textit{pb}}^B b$:

    Since $a_1 <_{\textit{pb}}^B b$, $\textit{rm}(a_1) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_1)$ is before $\textit{cal}(\textit{rm},b)$. Since $\textit{rm}(a_1)$ does not happen before $\textit{put}(a_1)$, $\textit{cal}(\textit{put},a_1,\textit{pri})$ is before $\textit{ret}(\textit{rm},a_1)$. Since $a <_{\textit{pb}}^A a_1$, $\textit{ret}(\textit{put},a,\textit{pri})$ is before $\textit{cal}(\textit{put},a_1,\textit{pri})$. Therefore, $\textit{ret}(\textit{put},a,\textit{pri})$ is before $\textit{cal}(\textit{rm},b)$. Since $\textit{cal}(\textit{rm},b)$ is before $o$, we can see that $o$ is before $\textit{cal}(\textit{rm},a)$.

    If $\textit{cal}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$, then $e \vert_{ \{ a,a_1,b \} }$ violates the FIFO property. Therefore, $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$. Similarly as the case of reason $1$, we can see that $\textit{ret}(\textit{rm},b)$ is before $\textit{ret}(\textit{rm},a)$.

    According to above discussion, there are three possible enumeration of operations of $a$, $a_1$ and $b$, as shown in \figurename~\ref{fig:history enumeration 3 for PQ1Equal}, \figurename~\ref{fig:history enumeration 4 for PQ1Equal} and \figurename~\ref{fig:history enumeration 5 for PQ1Equal}. Here we explicitly draw the leftmost gap-point of $a$ as $o'$. Since the position of $\textit{ret}(\textit{put},a_1,\textit{pri})$ and $\textit{cal}(\textit{put},a,\textit{pri})$ do not influence the correctness, we can simply ignore it. We also ignore $\textit{cal}(\textit{put},b,\textit{pri})$ and $\textit{ret}(\textit{put},b)$, since the only requirements of them are (1) $\textit{rm}(b)$ does not happen before $\textit{put}(b)$ and (2) $\textit{cal}(\textit{put},b,\textit{pri})$ is before $o$.

\item[-] Reason $5$, $a <_{\textit{pb}}^B a_1 <_{\textit{pb}}^A b$:

    Since $a_1 <_{\textit{pb}}^A b$, $\textit{ret}(\textit{put},a_1)$ is before $\textit{call}(\textit{put},b,\textit{pri})$. Since $\textit{call}(\textit{put},b,\textit{pri})$ is before $o$, we can see that $\textit{ret}(\textit{put},a_1)$ is before $o$.

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o$ is before $\textit{cal}(\textit{rm},a)$: Then $o$ is obviously before $\textit{ret}(\textit{rm},a)$. Since $a <_{\textit{pb}}^B a_1$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},a_1)$. Then we can see that, $o$ is before $\textit{cal}(\textit{rm},a_1)$, and remember that $a_1 <_{\textit{pb}}^A b$. Then we can goto the case of reason $1$ and treat $a_1$ as $a$. Therefore, we can safely ignore this.

    \item[-] If $o$ is before $\textit{cal}(\textit{put},a,\textit{pri})$: Since $\textit{rm}(a)$ does not happen before $\textit{put}(a,\textit{pri})$, we can see that $\textit{cal}(\textit{put},a,\textit{pri})$ is before $\textit{ret}(\textit{rm},a)$, and then $o$ is before $\textit{ret}(\textit{rm},a)$. Then similarly as above case, we can see that $o$ is before $\textit{cal}(\textit{rm},a_1)$, and $a_1 <_{\textit{pb}}^A b$. Then we can goto the case of reason $1$ and treat $a_1$ as $a$. Therefore, we can safely ignore this.
    \end{itemize}
\end{itemize}

This completes the proof of this lemma. \qed
\end {proof}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.4 \textwidth]{figures/PIC-HIS-PQ1Equal-1.pdf}
%\vspace{-10pt}
  \caption{The first possible enumeration.}
  \label{fig:history enumeration 1 for PQ1Equal}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.4 \textwidth]{figures/PIC-HIS-PQ1Equal-2.pdf}
%\vspace{-10pt}
  \caption{The second possible enumeration.}
  \label{fig:history enumeration 2 for PQ1Equal}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.4 \textwidth]{figures/PIC-HIS-PQ1Equal-3.pdf}
%\vspace{-10pt}
  \caption{The third possible enumeration.}
  \label{fig:history enumeration 3 for PQ1Equal}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.4 \textwidth]{figures/PIC-HIS-PQ1Equal-4.pdf}
%\vspace{-10pt}
  \caption{The forth possible enumeration.}
  \label{fig:history enumeration 4 for PQ1Equal}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.4 \textwidth]{figures/PIC-HIS-PQ1Equal-5.pdf}
%\vspace{-10pt}
  \caption{The fifth possible enumeration.}
  \label{fig:history enumeration 5 for PQ1Equal}
\end{figure}

Let us begin to represent several witness automata that is used to capture the existence of a data-differentiated $\_$-execution $e$, $e$ has a projection $e'$, $\textit{last}(e') = \textit{EPQ}_1^{=}$, there exists items $a$ and $b$ with maximal priority in $e'$, $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$.

Given a data-differentiated $\_$-execution $e$, two actions $\textit{act}_1$, $\textit{act}_2$ of maximal priority in $e$, and assume that $\textit{act}_1$ is before $\textit{act}_2$ in $e$.
we say that $\textit{act}_1$, $\textit{act}_2$ is covered by items $d_1,\ldots,d_m$ in $e$, if the priorities of $d_1,\ldots,d_m$ is smaller than that of $\textit{act}_1$ and $\textit{act}_2$, and

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{ret}(\textit{put},d_m,\_)$ is before $\textit{act}_1$,

\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,

\item[-] $\textit{act}_2$ is before $\textit{cal}(\textit{rm},d_1)$.
\end{itemize}

According to Lemma \ref{lemma:EPQ1Equal as pb order and gap-point}, Lemma \ref{lemma:ob order has bounded length} and Lemma \ref{lemma:five enumeration is enough for EPQ1Equal}, it is not hard to prove that, given a data-differentiated $\_$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{=}$, $e$ does not linearizable with respect to $\textit{EMS}(\textit{PQ}_1^{=})$, if and only if, one of enumerations holds in $e$ (permit renaming), while $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ is covered by some $d_1,\ldots,d_m$, $\textit{cal}(\textit{rm},b)$ is before $\textit{ret}(\textit{put},d_m,\_)$, and $\textit{cal}(\textit{rm},d_1)$ is before $\textit{ret}(\textit{rm},a)$. We say that such $d_1,\ldots,d_m$ constitute the rightmost gap of $b$.

An automaton $\mathcal{A}_{\textit{l-eq}}^1$ is given in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}, and it is constructed for the first enumeration in \figurename~\ref{fig:history enumeration 1 for PQ1Equal}. Here we rename the items that covers $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ into $d$, and rename the remanning items into $e$. In this figure, $c = \textit{cal}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{cal}(\textit{rm},e), \textit{ret}(\textit{rm},e),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{cal}(\textit{put},d),\textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8 \textwidth]{figures/PIC_AUTO_PQ1Equ-1.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{l-eq}}^1$}
  \label{fig:automata for first enumeration of PQ1Equal}
\end{figure}

An automaton $\mathcal{A}_{\textit{l-eq}}^2$ is given in \figurename~\ref{fig:automata for second enumeration of PQ1Equal}, and it is constructed for the second enumeration in \figurename~\ref{fig:history enumeration 2 for PQ1Equal}. In \figurename~\ref{fig:automata for second enumeration of PQ1Equal}, $c_1$, $c_2$, $c_3$ and $c_4$ is same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8 \textwidth]{figures/PIC_AUTO_PQ1Equ-2.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{l-eq}}^2$}
  \label{fig:automata for second enumeration of PQ1Equal}
\end{figure}

For the third enumeration in \figurename~\ref{fig:history enumeration 3 for PQ1Equal}. Since we want to ensure that $a$ and $b$ are putted only once, we need to explicitly record the positions of $\textit{cal}(\textit{put},a,p)$ and $\textit{cal}(\textit{put},b,p)$. Since the positions of $\textit{cal}(\textit{put},a,p)$ and $\textit{cal}(\textit{put},b,p)$ are not fixed, there are finite possible cases to consider, as shown below:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{rm},b)$ and before $\textit{cal}(\textit{rm},a)$: There are two possible positions of $\textit{cal}(\textit{put},a,p)$: (1) before $\textit{cal}(\textit{rm},a_1)$, and (2) after $\textit{cal}(\textit{rm},a_1)$, and before $\textit{ret}(\textit{put},a)$.

\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$ and before $\textit{cal}(\textit{rm},b)$: same as above case.

\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{put},a_1,\_)$ and before $\textit{ret}(\textit{rm},a_1)$: same as above case.

\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{put},a)$ and before $\textit{cal}(\textit{put},a_1,p)$: same as above case.

\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$: There are three possible positions of $\textit{cal}(\textit{put},a,p)$: (1) after $\textit{cal}(\textit{put},b,p)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{cal}(\textit{put},b,p)$, and (3) before $\textit{call}(\textit{rm},a_1)$.

\item[-] If $\textit{cal}(\textit{put},b,p)$ is before $\textit{cal}(\textit{rm},a_1)$: There are three possible positions of $\textit{cal}(\textit{put},a,p)$: (1) after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{cal}(\textit{put},b,p)$ and before $\textit{cal}(\textit{rm},a_1)$, and (3) before $\textit{cal}(\textit{put},b,p)$.
\end{itemize}

Therefore, there are fourteen possible cases that satisfy the third enumeration in \figurename~\ref{fig:history enumeration 3 for PQ1Equal}. For each case, we construct an finite automaton. Let $\textit{Auts}_{\textit{1-eq}}^{3}$ be the set of finite automata that is constructed for above fourteen cases. For example, for the case $\textit{ca}_1$ when $\textit{cal}(\textit{put},a,p)$ is before $\textit{cal}(\textit{rm},a_1)$, $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$, and $\textit{cal}(\textit{put},b,p)$ is before $\textit{cal}(\textit{rm},b)$, we construct a finite automaton $\mathcal{A}_{\textit{l-eq}}^{\textit{3-1}}$ in \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}. In \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}, let $c$ and $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$ the same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}. Let $c_2 = c_1 + \textit{ret}(\textit{put},a_1)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$, $c_4 = c_3 + \textit{cal}(\textit{put},d) + \textit{ret}(\textit{rm},d)$, and $c_5 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{put},a_1) + \textit{ret}(\textit{rm},d)$. The other witness automata in $\textit{Auts}_{\textit{1-eq}}^{3}$ can be similarly constructed.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8 \textwidth]{figures/PIC_AUTO_PQ1Equ-3-1.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{l-eq}}^{\textit{3-1}}$}
  \label{fig:automata for ca1 of third enumeration of Rpr2}
\end{figure}

Similarly, we construct sets $\textit{Auts}_{\textit{1-eq}}^{4}$ and $\textit{Auts}_{\textit{1-eq}}^{5}$ of witness automata for the forth enumeration in \figurename~\ref{fig:history enumeration 4 for PQ1Equal} and the fifth enumeration in \figurename~\ref{fig:history enumeration 5 for PQ1Equal}, respectively.

Let $\textit{Auts}_{\textit{1-eq}} = \{ \mathcal{A}_{\textit{l-eq}}^1, \mathcal{A}_{\textit{l-eq}}^2 \} \cup \textit{Auts}_{\textit{1-eq}}^{3} \cup \textit{Auts}_{\textit{1-eq}}^{4} \cup \textit{Auts}_{\textit{1-eq}}^{5}$. The following lemma states that $\textit{PQ}_1^{=}$ is co-regular.

\EPQOneEqualIsCoRegular*

\begin {proof}

We need to prove that, given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$, if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e),$ $\textit{EPQ}_1^{=} \in \textit{last}(e') \wedge e'$ does not linearizable w.r.t. $\textit{MS}(\textit{EPQ}_1^{=})$.

By Lemma \ref{lemma:pri execution is enough} and Lemma \ref{lemma:EPQ1Equal as pb order and gap-point}, we need to prove the following fact:

\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e)$, $\textit{last}(e')=\textit{EPQ}_1^{=}$, $a$ and $b$ are two items with maximal priority $\textit{pri}$ in $e'$, $e'$ is a $\textit{pri}$-execution, $a <_{\textit{pb}}^* b$ in $e'$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\textit{pri})$ or $\textit{cal}(\textit{rm},a)$ in $e'$.

\noindent The $\textit{only if}$ direction: Assume that $e_1 \in \mathcal{I}$ is accepted by some witness automata in $\textit{Auts}_{\textit{1-eq}}$. By data-independence, there exists data-differentiated execution $e_2 \in \mathcal{I}$ and a renaming function $r$, such that $e_1=r(e_2)$. Since $e_1$ is accepted by some witness automata in  $\textit{Auts}_{\textit{1-eq}}$, let $x$, $y$ and $z$ (if exists) be the items that are renamed into $b$, $a$ and $a_1$ (if exists) by $r$, respectively, and let $d_1,\ldots,d_m$ be the items that are renamed into $d$ by $r$.

let $e'' = e_2 \vert_{ \{ x,y,z,d_1,\ldots,d_m \} }$. It is obvious that $e'' \in \textit{proj}(e_2)$ is a $\textit{pri}$-execution, $\textit{last}(e'') = \textit{EPQ}_1^{=}$. According to our construction of automata in $\textit{Auts}_{\textit{1-eq}}$, it is not hard to see that $x$ and $y$ has maximal priority in $h_2$, $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is before $\textit{cal}(\textit{put},y,\textit{pri})$ or $\textit{cal}(\textit{rm},y)$ in $e''$.

\noindent The $\textit{if}$ direction: Assume that there exists $e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $last(e')=\textit{PQ}_1^{=}$, $a'$ and $b'$ are two items with maximal priority $\textit{pri}$ in $e'$, $e'$ is a $\textit{pri}$-execution, $a' <_{\textit{pb}}^* b'$ in $e'$, and the rightmost gap-point of $b'$ is before $\textit{cal}(\textit{put},a',\textit{pri})$ or $\textit{cal}(\textit{rm},a')$ in $e'$. By data-independence, we can obtain execution $e_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the items $d_1,\ldots,d_m$ that constitute the rightmost gap of $b'$, we rename them into $d$, (3) if $a' <_{\textit{pb}}^A a'_1 <_{\textit{pb}}^B b$, we rename $a'_1$ into $a_1$, and (4) rename the other items into $e$. It is easy to see that $\textit{last}(e_1) = \textit{EPQ}_1^{=}$, $a$ and $b$ has maximal priority in $e_1$, $a <_{\textit{pb}}^* b$ in $e_1$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\textit{pri})$ or $\textit{cal}(\textit{rm},a)$ in $e_1$. By Lemma \ref{lemma:five enumeration is enough for EPQ1Equal}, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists). Then

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $a <_{\textit{pb}}^* b$ because of the first enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{l-eq}}^1$.

\item[-] If $a <_{\textit{pb}}^* b$ because of the second enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{l-eq}}^2$.

\item[-] If $a <_{\textit{pb}}^* b$ because of the third enumeration, it is easy to see that $h_1$ is accepted by some witness automaton in $\textit{Auts}_{\textit{1-eq}}^{3}$.

\item[-] If $a <_{\textit{pb}}^* b$ because of the forth enumeration, it is easy to see that $h_1$ is accepted by some witness automaton in $\textit{Auts}_{\textit{1-eq}}^{4}$.

\item[-] If $a <_{\textit{pb}}^* b$ because of the fifth enumeration, it is easy to see that $h_1$ is accepted by some witness automaton in $\textit{Auts}_{\textit{1-eq}}^{5}$.
\end{itemize}

This completes the proof of this lemma. \qed
\end {proof}

\subsection{Co-Regular of $\textit{EPQ}_2^{>}$}
\label{subsec:appendix co-regular of EPQ2Lar}

\begin{restatable}{lemma}{EPQ2LarIsAlwaysCoRegular}
\label{lemma:EPQ2Lar is always co-regular}
Given a data-differentiated $\_$-execution $e$, if $\textit{last}(e) = \textit{EPQ}_2^{>}$, then $e \sqsubseteq \textit{MS}(\textit{EPQ}_2^{>})$.
\end{restatable}

\begin {proof}
Since $\textit{last}(e) = \textit{EPQ}_2^{>}$, the actions with maximal priority in $e$ is some unmatched $\textit{put}$. Therefore, no matter how we locate linearization points, we can always obtain a sequence $l$ of method events that contains unmatched $\textit{put}$ with maximal priority, and this satisfy the guard of $\textit{MS}(\textit{EPQ}_2^{>})$. This completes the proof of this lemma. \qed
\end {proof}

\subsection{Co-Regular of $\textit{EPQ}_2^{=}$}
\label{subsec:appendix co-regular of EPQ2Equal}

\begin{restatable}{lemma}{EPQ2EqualAsHappenBefore}
\label{lemma:EPQ2Equal as happen before}
Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_2^{=}$. $e$ does not linearizable to $\textit{MS}(\textit{EPQ}_2^{=})$, if and only if there exists $x$ and $y$ with priority $\textit{pri}$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,\textit{pri}) <_{\textit{hb}} \textit{put}(y,\textit{pri})$.
\end{restatable}

\begin {proof}

The $\textit{if}$ direction is obvious.

To prove the $\textit{only if}$ direction, we prove its contrapositive. Assume that for each pair of $x$ and $y$ with maximal priority in $e$, if $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, then $\textit{put}(x,\textit{pri})$ does not happen before $\textit{put}(y,\textit{pri})$. We need to prove that $e \sqsubseteq \textit{MS}(\textit{EPQ}_2^{=})$.

Let $x_1,\ldots,x_m$ be the set of items with priority $\textit{pri}$ and has unmatched $\textit{put}$ in $e$, let $y_1,\ldots,y_n$ be the set of items with priority $\textit{pri}$ and has matched $\textit{put}$ and $\textit{rm}$ in $e$. By assumption, we know that $\textit{cal}(\textit{put},y_i,\textit{pri})$ is before $\textit{ret}(\textit{put},x_j)$ for each $i,j$. Then we explicitly construction the linearization of $e$ by locating the linearization points of $e$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] For each $x_i$, locate the linearization point of $\textit{put}(x_i,\textit{pri})$ just before its return action.

\item[-] For each $y_j$, locate the lineariztion point of $\textit{put}(y_j,\textit{pri})$ jest after its call action.

\item[-] For other method events, locate their linearization points at an arbitrary location after its call action and before its return action.
\end{itemize}

Let $l$ be the sequence of linearization points. It is easy to see that $e \sqsubseteq l$. Since linearization points of $\textit{put}(x_i,\textit{pri})$ is after the linearization point of $\textit{put}(y_j,\textit{pri})$ for each $i,j$, it is easy to see that $l \in \textit{MS}(\textit{EPQ}_2^{=})$. This completes the proof of this lemma. \qed
\end {proof}

Lemma \ref{lemma:EPQ2Equal as happen before} shows how to check violation to $\textit{MS}(\textit{EPQ}_2^{=})$. However, the case in Lemma \ref{lemma:EPQ2Equal as happen before} violates our assumption that each single-priority execution is FIFO. Therefore, we know that $\textit{EPQ}_2^{=}$ is always co-regular, as states by the following lemma.

\begin{restatable}{lemma}{EPQ2EqualIsAlwaysCoRegular}
\label{lemma:EPQ2Equal is always co-regular}
Given a data-differentiated $\textit{pri}$-execution $e$, if $\textit{last}(e) = \textit{EPQ}_2^{=}$, then $e \sqsubseteq \textit{MS}(\textit{EPQ}_2^{=})$.
\end{restatable}

\begin {proof}

According to Lemma \ref{lemma:EPQ2Equal as happen before}, if $\textit{last}(e) = \textit{EPQ}_2^{=}$ and $e$ does not linearizable to $\textit{MS}(\textit{EPQ}_2^{=})$, then there exists $x$ and $y$ with priority $\textit{pri}$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$, and $\textit{put}(x,\textit{pri}) <_{\textit{hb}} \textit{put}(y,\textit{pri})$. Let $e_1 = e \vert_{ \{ x,y \} }$. It is obvious that $e_1$ does not satisfy FIFO property. This contradicts the assumption that every single-priority execution has FIFO property, and thus, we can safely ignore this case. \qed
\end {proof}

\subsection{Co-Regular of $\textit{EPQ}_3$}
\label{subsec:co-regular of EPQ3}

In this subsection we prove that $\textit{EPQ}_3$ is co-regular. The notion of left-right constraint of $\textit{rm}(\textit{empty})$ is inspired by left-right constraint of queue \cite{Bouajjani:2015}.

\begin{definition}\label{def:left-right constraint for rmEmpty operation}
Given a data-differentiated execution $e$, and $o = \textit{rm}(\textit{empty})$ of $e$. The left-right constraint of $o$ is the graph $G$ where:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] the nodes are the items of $e$ or $o$, to which we add a node,

\item[-] there is an edge from item $d_1$ to $o$, if $\textit{put}(d_1,\_)$ happens before $o$,

\item[-] there is an edge from $o$ to item $d_1$, if $o$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$,

\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$.
\end{itemize}
\end{definition}

Given a data-differentiated execution $e$ and $o = \textit{rm}(\textit{empty})$ of $e$, it is obvious that $\textit{last}(e) = \textit{EPQ}_3$. Let $\textit{USet}_1(e,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some item, and either $\textit{op} <_{\textit{hb}} o$, or there is $\textit{op}'$ with the same item of $\textit{op}$, such that $\textit{op}' <_{\textit{hb}} o \}$. For each $i \geq 1$, let $\textit{USet}_{\textit{i+1}}(e,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some item, $\textit{op}$ is not in $\textit{USet}_k(e,o)$ for each $k \leq i$, and either $\textit{op}$ happens before some $o' \in \textit{USet}_i(e,o)$, or there is $\textit{op}''$ with the same item of $o$ and $\textit{op}''$ happens before some $o' \in \textit{USet}_i(e,o) \}$. We can see that $\textit{USet}_i(e,o) \cap \textit{USet}_j(e,o) = \emptyset$ for any $i \neq j$. Let $\textit{USet}(e,o) = \textit{USet}_1(e,o) \cup \textit{USet}_2(e,o) \cup \ldots$.

Similarly as $\textit{UVSet}$, we can prove the following two lemmas for $\textit{USet}$.

\begin{restatable}{lemma}{USetHasMatchedPutandRm}
\label{lemma:USet has matched put and rm}
Given a data-differentiated execution $e$ with $\textit{last}(e) = \textit{EPQ}_3$. Let $o$ be a $\textit{rm}(\textit{empty})$ of $e$. Let $G$ be the graph representing the left-right constraint of $o$. Assume that $G$ has no cycle going through $o$. Then, $\textit{USet}(e,o)$ contains only matched $\textit{put}$ and $\textit{rm}$.
\end{restatable}

This Lemma can be similarly proved as Lemma \ref{lemma:UVSet has matched put and rm}.

\begin{restatable}{lemma}{RmxDoesNotHappenBeforeUSetForEPQ3}
\label{lemma:Rmx does not happen before USet for EPQ3}
Given a data-differentiated $\_$execution $e$ with $\textit{last}(e) = \textit{EPQ}_3$. Let $o$ be a $\textit{rm}(\textit{empty})$ of $e$. Let $G$ be the graph representing the left-right constraint of $o$. Assume that $G$ has no cycle going through $o$. Then, $o$ does not happen before any operation in $\textit{USet}(e,o)$.
\end{restatable}

This Lemma can be similarly proved as Lemma \ref{lemma:Rmx does not happen before UVSet for EPQ1Lar}.

Then we can prove that getting rid of cycle though $o$ in left-right constraint is enough for ensure linearizable w.r.t $\textit{MS}(\textit{EPQ}_3)$, as stated by the following lemma.

\begin{restatable}{lemma}{LINEqualsConstraintforEPQ3}
\label{lemma:Lin Equals Constraint for EPQ3}
Given a data-differentiated execution $e$ with $\textit{last}(e) = \textit{EPQ}_3$. $e$ does not linearizable w.r.t $\textit{MS}(\textit{PQ}_3)$, if and only if there exists $o = \textit{rm}(\textit{empty})$ in $e$, $G$ has a cycle going through $o$, where $G$ is the graph representing the left-right constraint of $o$.
\end{restatable}

\begin {proof}

To prove the $\textit{if}$ direction, assume that there is such a cycle. Assume by contradiction that $e \sqsubseteq \textit{MS}(\textit{EPQ}_3)$, and let $U$ and $V$ be the set of operations in $u$ and $v$. Let the cycle be $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow o \rightarrow d_1$ in $G$. Since $d_m \rightarrow o$, $\textit{put}(d_m,\_)$ happens before $o$, and it is easy to see that $\textit{put}(d_m,\_)$ is in $U$. Since $U$ contains matched $\textit{put}$ and $\textit{rm}$, we can see that operations of $d_m$ is in $U$. Similarly, we can see that method events of $d_{\textit{m-1}},\ldots,d_1$ is in $U$. If $\textit{rm}(d_1)$ does not exists, then this contradicts that $U$ contains matched $\textit{put}$ and $\textit{rm}$. Else, if $\textit{rm}(d_1)$ exists, since $o$ happens before $\textit{rm}(d_1)$, we can see that $\textit{rm}(d_1) \in V$, which contradicts that $\textit{rm}(d_1) \in U$. This completes the proof of the $\textit{if}$ direction.

To prove the $\textit{only if}$ direction, we prove its contrapositive. Assume that for each such $o$ and $G$, $G$ has no cycle going through $o$. Let $O$ be the set of operations of $e$, except for $\textit{rm}(\textit{empty})$. Let $O_L = \textit{USet}(e,o)$, $O_R = O \setminus O_L$.

By Lemma \ref{lemma:USet has matched put and rm}, we can see that $O_L = \textit{USet}(e,o)$ contains only matched $\textit{put}$ and $\textit{rm}$. Let $O'_L$ be the union of $O_L$ and all the $\textit{rm}(\textit{empty})$ that happens before some operations in $O_L \cup \{ o \}$. Let $O'_R$ be the union of $O_R$ and the remanning $\textit{rm}(\textit{empty})$. It remains to prove that for $O'_L$, $\{ o \}$, $O'_R$, no elements of the latter set happens before elements of the former set. We prove this by showing that all the following cases are impossible:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Case $1$: If some operation $o_r \in O'_R$ happens before $o$. Then we can see that $o_r \in \textit{USet}(e,o)$ or is a $\textit{rm}(\textit{empty})$ that happens before $o$, and then $o_r \in O'_L$, which contradicts that $o_r \in O'_R$.

\item[-] Case $2$: If some operation $o_r \in O'_R$ happens before some operation $o_l \in O'_L$. Then we know that $o_r \in \textit{USet}(e,o)$ or is a $\textit{rm}(\textit{empty})$ that happens before some operations in $O_L \cup \{ o \}$, and then $o_r \in O'_L$, which contradicts that $o_r \in O'_R$.

\item[-] Case $3$: If $o$ happens before some $o_l \in O'_L$. If $o_l \in \textit{USet}(e,o)$, then by Lemma \ref{lemma:Rmx does not happen before USet for EPQ3} we know that this is impossible. Else, $o_l$ is a $\textit{rm}(\textit{empty})$ that happens before some operations in $O_L \cup \{ o \}$, and $o$ happens before some operations in $O_L \cup \{ o \}$, which is impossible by Lemma \ref{lemma:Rmx does not happen before USet for EPQ3}.
\end{itemize}

This completes the proof of the $\textit{only if}$ direction.

\qed
\end {proof}

Let us begin to represent an automaton that is used for capture the case that, in a sub-execution $e'$ of an execution $e$, $\textit{last}(e')=\textit{EPQ}_3$, $e'$ does not linearizable to $\textit{MS}(\textit{EPQ}_3)$, and the reason is that there is a cycle going through some $\textit{rm}(\textit{empty})$ $o$ in the left-right constraint of $o$. The automaton is $\mathcal{A}_{\textit{EPQ}}^3$, which is given in \figurename~\ref{fig:automata for PQ3}. In \figurename~\ref{fig:automata for PQ3}, let $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d), \textit{ret}(\textit{rm},d),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{cal}(\textit{put},b,\textit{anyPri})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c + \textit{ret}(\textit{rm},b)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8 \textwidth]{figures/PIC_AUTO_PQ3.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{EPQ}}^3$}
  \label{fig:automata for PQ3}
\end{figure}

Given a data-differentiated execution $e$, we say that $o = \textit{rm}(\textit{empty})$ in $e$ is covered by items $d_1,\ldots,d_m$ in $h$, if

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{put}(d_m,\_)$ happens before $o$,

\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,

\item[-] $o$ happens before $\textit{rm}(d_1)$, or $\textit{rm}(d_1)$ does not exists in $e$
\end{itemize}

According to the definition of left-right constraint for $o$, in a data-differentiated execution $e$, there is a cycle going through $o$, if and only if there exists items $d_1,\ldots,d_m$, such that $o$ is covered by $d_1,\ldots,d_m$.

\begin{restatable}{lemma}{EPQ3IsCoRegular}
\label{lemma:EPQ3 is co-regular}
$\textit{EPQ}_3$ is co-regular.
\end{restatable}

\begin {proof}

We need to prove that, given a data-independence implementation $\mathcal{I}$, $\mathcal{A}_{\textit{EPQ}}^3 \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), last(e')=\textit{EPQ}_3 \wedge e$ does not linearizable w.r.t. $\textit{MS}(\textit{EPQ}_3)$.

By Lemma \ref{lemma:Lin Equals Constraint for EPQ3}, we need to prove the following fact:

\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\mathcal{A}_{\textit{EPQ}}^3 \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), last(e')=\textit{EPQ}_3$, $o = \textit{rm}(\textit{empty})$ is in $e'$, and $o$ is covered by some items $d_1,\ldots,d_m$ in $e'$.

\noindent The $\textit{only if}$ direction: Assume that $e_1 \in \mathcal{I}$ is accepted by $\mathcal{A}_{\textit{EPQ}}^3$. By data-independence, there exists data-differentiated execution $e_2 \in \mathcal{I}$ and a renaming function $r$, such that $e_1=r(e_2)$. Let $d_1,\ldots,d_m$ be the items in $e_2$ such that $r(d_i)=b$ for each $1 \leq i \leq m$. Let $e_3 = e_2 \vert_{ \{ o, d_1, \ldots, d_m \} }$. It is obvious that $e_3 \in \textit{proj}(e_2)$ and $\textit{last}(e_3) = \textit{EPQ}_3$. It is easy to see that $o$ is covered by $d_1,\ldots,d_m$.

\noindent The $\textit{if}$ direction: Assume that there exists such $e$, $e'$, $o$ and $d_1,\ldots,d_m$. Then, let $e_1$ be obtained from $e$ by renaming $d_1,\ldots,d_m$ into $b$ and renaming other items into $d$. By data-independence, $e_1 \in \mathcal{I}$. It is easy to see that $e_1$ is accepted by $\mathcal{A}_{\textit{EPQ}}^3$.

This completes the proof of this lemma. \qed
\end {proof}

\section{Proofs and Definitions in Section \ref{sec:relate other data structures with extended priority queue}}
\label{sec:appendix proof and definition in section relate other data structures with extended priority queue}

\subsection{Proofs and Definitions in Subsection \ref{subsec:relate multiSet with extended priority queue}}
\label{subsec:appendix proof and definition in section relate multiset with extended priority queue}

In this section, we use the notions of inductive rules, $\textit{last}$ of a sequential execution, step-by-step and co-regular in \cite{Bouajjani:2015}. Since they are quite similar to the corresponding notions of extended priority queues in Section \ref{sec:inductive rules of extended priority queue}, Section \ref{sec:step-by-step linearizability of extended priority queues} and Section \ref{sec:co-regular of extended priority queues}, we do not introduce their definitions here.

Let us first define three predicates:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Given a sequential execution $l$ of multi-set, $\textit{noDE}(l)$ is satisfied when each method event of $l$ is not $\textit{delete}(\textit{empty})$.

\item[-] Given a sequential execution $l$ of multi-set, $\textit{matched-MS}(l)$ is satisfied, if (1) for each item $a \in \mathbb{D}$, if $\textit{insert}(a)$ is in $l$, then $\textit{delete}(a)$ is in $l$, and (2) for each item $a \in \mathbb{D}$, if $\textit{delete}(a)$ is in $l$, then $\textit{insert}(a)$ is in $l$.

\item[-] Given a sequential execution $l$ of multi-set, $l \in \textit{Insert}^*$ is satisfied when each method event of $l$ is a $\textit{insert}$ event.
\end{itemize}

Let $\textit{MSet}$ be the set of sequential executions $w$ which can be derived from the empty word by inductive rules of multi-set. $\textit{MSet}$ is defined by the following inductive rules:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{MSet}_0 \equiv \epsilon \in \textit{MSet}$.

\item[-] $\textit{MSet}_1 \equiv (u \in \textit{MSet}) \wedge
(u \in \textit{Insert}^*)
\Rightarrow
(u \cdot \textit{insert}(\textit{itm}) \in \textit{MSet})$.

\item[-] $\textit{MSet}_2 \equiv
(u \cdot v \cdot w \in \textit{MSet}) \wedge
(\textit{noDE}(u \cdot v \cdot w))
\Rightarrow
(u \cdot \textit{insert}(\textit{itm}) \cdot v \cdot \textit{delete}(\textit{itm}) \cdot w \in \textit{MSet})$.

\item[-] $\textit{MSet}_3 \equiv
(u \cdot v \in \textit{MSet}) \wedge
(\textit{matched-MS}(u) )
\Rightarrow
(u \cdot \textit{delete}(\textit{empty}) \cdot v \in \textit{MSet})$.
\end{itemize}

Thus, given a sequential execution $e$, we define $\textit{last}(e)$ as the last possible rule to generate $e$ according to the rules of multi-set:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $e$ contains $\textit{delete}(\textit{empty})$, then $\textit{last}(e) = \textit{MS}_3$.

\item[-] Else, if $e$ contains $\textit{delete}$, then $\textit{last}(e) = \textit{MSet}_2$.

\item[-] Else, if $e$ contains only $\textit{insert}$, then $\textit{last}(e) = \textit{MSet}_1$.

\item[-] Else ($e = \epsilon$), $\textit{last}(e) = \textit{MSet}_0$.
\end{itemize}

The following three lemmas state that the rules for multi-set are step-by-step linearizability.

\begin{restatable}{lemma}{MS1isStepByStepLinearizability}
\label{lemma:MS1 is step-by-step linearizability}
If a differentiated concurrent execution $e$ is linearizable w.r.t. $\textit{MS}(\textit{MSet}_1)$ with witness $x$, then $e \setminus x \sqsubseteq \textit{MSet} \Rightarrow e \sqsubseteq \textit{MSet}$.
\end{restatable}

\begin {proof}
Let $h$ be the data-differentiated history of $e$, and $l$ be an sequential execution such that $h \sqsubseteq l$ and $l$ matches $\textit{MSet}_1$ with witness $x$. Let $h'=h \setminus x$ and assume that $h' \sqsubseteq l' \in \textit{MSet}$. Let $e_{\textit{lp}}$ be an execution with linearization points of $e$ and the linearization points is added according to $l'$. Or we can say, $e_{\textit{lp}}$ is generated from $e$ by instrumenting linearization points, and the projection of $e_{\textit{lp}}$ into method event is $l'$.

Let sequence $e'_{\textit{lp}}$ be generated from $e_{\textit{lp}}$ by adding $\textit{insert}(x)$ at an arbitrary time point between $\textit{cal}(\textit{insert},x)$ and $\textit{ret}(\textit{insert},x)$. Let $l''$ be the projection of $e'_{\textit{lp}}$ into method events.

It is easy to see that $h \sqsubseteq l''$. Since $l''$ is obtained from $l'$ by adding one $\textit{insert}(x)$, we can see that $l''$ contains only $\textit{insert}$, and then $l'' \in \textit{MSet}$. \qed
\end {proof}

\begin{restatable}{lemma}{MS2isStepByStepLinearizability}
\label{lemma:MS2 is step-by-step linearizability}
If a differentiated concurrent execution $e$ is linearizable w.r.t. $\textit{MS}(\textit{MSet}_2)$ with witness $x$, then $e \setminus x \sqsubseteq \textit{MSet} \Rightarrow e \sqsubseteq \textit{MSet}$.
\end{restatable}

\begin {proof}
Let $h$ be the data-differentiated history of $e$, and $l$ be an sequential execution such that $h \sqsubseteq l$ and $l$ matches $\textit{MSet}_1$ with witness $x$. Let $h'=h \setminus x$ and assume that $h' \sqsubseteq l' \in \textit{MSet}$. Let $e_{\textit{lp}}$ be an execution with linearization points of $e$ and the linearization points is added according to $l'$. Or we can say, $e_{\textit{lp}}$ is generated from $e$ by instrumenting linearization points, and the projection of $e_{\textit{lp}}$ into method event is $l'$.

It is easy to see that $\textit{delete}(x)$ does not happen before $\textit{insert}(x)$, and then $\textit{cal}(\textit{insert},x)$ is before $\textit{ret}(\textit{delete},x)$ in $e$. Let sequence $e'_{\textit{lp}}$ be generated from $e_{\textit{lp}}$ by adding $\textit{insert}(x)$ just after $\textit{cal}(\textit{insert},x)$ and adding $\textit{delete}(x)$ just before $\textit{ret}(\textit{delete},x)$. Let $l''$ be the projection of $e'_{\textit{lp}}$ into method events.

It is easy to see that $h \sqsubseteq l''$. Since (1) $l' \in \textit{MSet}$, (2) $l''$ is obtained from $l'$ by adding one $\textit{insert}(x)$ and one $\textit{delete}(x)$, while $\textit{insert}(x)$ is before $\textit{delete}(x)$ (3)and $l'$ does not contain $\textit{delete}(\textit{empty})$, we can see that $l'' \in \textit{MSet}$. \qed
\end {proof}

\begin{restatable}{lemma}{MS3isStepByStepLinearizability}
\label{lemma:MS3 is step-by-step linearizability}
If a differentiated concurrent execution $e$ is linearizable w.r.t. $\textit{MS}(\textit{MSet}_3)$ and $o$ is a $\textit{delete}(\textit{empty})$ event, then $e \setminus o \sqsubseteq \textit{MSet} \Rightarrow e \sqsubseteq \textit{MSet}$.
\end{restatable}

\begin {proof}
This Lemma can be similarly proved as Lemma \ref{lemma:EPQ3 is step-by-step linearizability}. \qed
\end {proof}

The following lemma states that $\textit{MSet}_1$ is always co-regular.

\begin{restatable}{lemma}{MS1IsAlwaysCoRegular}
\label{lemma:MS1 is always co-regular}
Given a differentiated execution $e$, if $\textit{last}(e) = \textit{MSet}_1$, then $e \sqsubseteq \textit{MS}(\textit{MSet}_1)$.
\end{restatable}

\begin {proof}
Since $\textit{last}(e) = \textit{MSet}_1$, there are only $\textit{insert}$ in $e$. Then no matter how we locate linearization points of operations of $e$, we can always obtain a sequence in $\textit{MS}(\textit{MSet}_1)$. This completes the proof of this lemma. \qed
\end {proof}

The following lemma shows how to detect violation to $\textit{MS}(\textit{MSet}_2)$.

\begin{restatable}{lemma}{ReduceMS2intoOneValue}
\label{lemma:reduce MS2 into one value}
Given a differentiated execution $e$, there exists some $e' \in \textit{proj}(e)$, such that $\textit{last}(e') = \textit{MSet}_2$ and $e'$ doe not linearizable w.r.t $\textit{MS}(\textit{MSet}_2)$, if and only if one of the following case holds for some $x \in \mathbb{D}$.
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{delete}(x)$ is in $e$ while $\textit{insert}(x)$ is not in $e$.

\item[-] there is more than one $\textit{delete}(x)$ in $e$ and one $\textit{insert}(x)$ in $e$.

\item[-] $\textit{delete}(x) <_{\textit{hb}} \textit{insert}(x)$ in $e$.
\end{itemize}
\end{restatable}

\begin {proof}

The $\textit{only if}$ is obvious and omitted here.

To prove the $\textit{if}$ direction, we prove its contrapositive. Assume that for each item $x$ of $e$, the following conditions are satisfied:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\textit{delete}(x)$ is in $e$, then $\textit{insert}(x)$ is also in $e$.

\item[-] There is at most $\textit{delete}(x)$ in $e$.

\item[-] $\textit{delete}(x)$ does not happen before $\textit{insert}(x)$.
\end{itemize}

Assume by contradiction that there exists some $e' \in \textit{proj}(e)$, such that $\textit{last}(e') = \textit{MSet}_2$ and $e'$ doe not linearizable w.r.t $\textit{MS}(\textit{MSet}_2)$. Since $\textit{last}(e') = \textit{MSet}_2$, there exists $\textit{delete}$ in $e$. Let this $\textit{delete}$ operation be $\textit{delete}(a)$. By assumption, we know that in $e$ there exists one $\textit{insert}(a)$ and one $\textit{delete}(a)$, and $\textit{delete}(a)$ does not happen before $\textit{insert}(a)$. Then we know that $\textit{cal}(\textit{insert},x)$ is before $\textit{ret}(\textit{delete},x)$.

Let $e_{\textit{lp}}$ be generated from $e$ by (1) put the linearization of $\textit{insert}(x)$ just after $\textit{cal}(\textit{insert},x)$, (2) put the linearization of $\textit{delete}(x)$ just before $\textit{ret}(\textit{insert},x)$, and (3) for other operations, put its linearization point at an arbitrary between its call and return actions. Let $l'$ be the projection of $e_{\textit{lp}}$ into method events. It is obvious that $h \sqsubseteq l'$. According to our construction of $e_{\textit{lp}}$, we can see that in $l'$, $\textit{insert}(x)$ is before $\textit{delete}(x)$. Therefore, we can see that $l' \in \textit{MS}(\textit{MSet}_2)$, which contradicts our assumption. This completes the proof of the $\textit{id}$ direction. \qed
\end {proof}

According to Lemma \ref{lemma:reduce MS2 into one value}, to check violation to $\textit{MS}(\textit{MSet}_2)$, we need to consider three cases for some $b \in \mathbb{D}$: (1) there is $\textit{delete}(b)$ but there is not $\textit{insert}(x)$, (2) there is more than one $\textit{delete}(b)$ and one $\textit{insert}(b)$ in $e$, and (3) $\textit{delete}(b) <_{\textit{hb}} \textit{insert}(b)$.

For each such case, we construct a witness automata. We generate witness automata $\mathcal{A}_{\textit{MS}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata 1 for MS-2 in appendix}. Here $c_1 = \textit{cal}(\textit{insert},a),\textit{ret}(\textit{insert},a)$, $\textit{cal}(\textit{delete},a),\textit{ret}(\textit{delete},a),
\textit{cal}(\textit{delete},\textit{empty}),\textit{ret}(\textit{delete},\textit{empty})$, $c_2 = c_1 + \textit{cal}(\textit{delete},b) + \textit{ret}(\textit{delete},b)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.3 \textwidth]{figures/PIC_AUTO_MS_1.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{MS}}^1$}
  \label{fig:automata 1 for MS-2 in appendix}
\end{figure}

We generate witness automata $\mathcal{A}_{\textit{MS}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata 2 for MS-2 in appendix}. Here $c_1 = \textit{cal}(\textit{insert},a),\textit{ret}(\textit{insert},a), \textit{cal}(\textit{delete},a),\textit{ret}(\textit{delete},a),\textit{cal}(\textit{delete},\textit{empty})$, $\textit{ret}(\textit{delete},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{insert},b)$, $c_3 = c_2 + \textit{ret}(\textit{delete},b)$, $c_4 = c_3 + \textit{cal}(\textit{delete},b)$, $c_5 = c_1 + \textit{ret}(\textit{delete},b)$, $c_6 = c_5 + \textit{cal}(\textit{delete},b)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.7 \textwidth]{figures/PIC_AUTO_MS_2.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{MS}}^2$}
  \label{fig:automata 2 for MS-2 in appendix}
\end{figure}

We generate witness automata $\mathcal{A}_{\textit{MS}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata 3 for MS-2 in appendix}. Here $c_1 = \textit{cal}(\textit{insert},a)$, $\textit{ret}(\textit{insert},a)$, $\textit{cal}(\textit{delete},a)$, $\textit{ret}(\textit{delete},a),\textit{cal}(\textit{delete},b)$, $\textit{cal}($ $\textit{delete},\textit{empty}),\textit{ret}(\textit{delete},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{delete},b)$, $c_3 = c_2 + \textit{ret}(\textit{insert},b)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6 \textwidth]{figures/PIC_AUTO_MS_3.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{MS}}^3$}
  \label{fig:automata 3 for MS-2 in appendix}
\end{figure}

Let $\textit{Auts}_{\textit{2-ms}} = \{ \mathcal{A}_{\textit{MS}}^1, \mathcal{A}_{\textit{MS}}^2, \mathcal{A}_{\textit{MS}}^3 \}$. The following lemma states that $\textit{MS}_2$ is co-regular.

\begin{restatable}{lemma}{MS2IsCoRegular}
\label{lemma:MS2 is co-regular}
$\textit{MSet}_2$ is co-regular.
\end{restatable}

\begin {proof}

According to \cite{Bouajjani:2015}, we need to prove that, given a independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{2-ms}} \cap \mathcal{I} \neq \emptyset$, if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e),$ $\textit{last}(e') = \textit{MSet}_2 \wedge e'$ does not linearizable w.r.t. $\textit{MS}(\textit{MSet}_2)$.

By Lemma \ref{lemma:reduce MS2 into one value}, we need to prove the following fact:

\noindent {\bf $\textit{fact}_1$}: given a independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{2-ms}} \cap \mathcal{I} \neq \emptyset$, if and only if $\exists e \in \mathcal{I}_{\neq}$, and one of the following case holds for some $x \in \mathbb{D}$.
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{delete}(x)$ is in $e$ while $\textit{insert}(x)$ is not in $e$.

\item[-] there is more than one $\textit{delete}(x)$ in $e$ and one $\textit{insert}(x)$ in $e$.

\item[-] $\textit{delete}(x) <_{\textit{hb}} \textit{insert}(x)$ in $e$.
\end{itemize}

\noindent The $\textit{only if}$ direction: Assume that $e_1 \in \mathcal{I}$ is accepted by some witness automata in $\textit{Auts}_{\textit{1-eq}}$. By data-independence, there exists data-differentiated execution $e_2 \in \mathcal{I}$ and a renaming function $r$, such that $e_1=r(e_2)$. Since $e_1$ is accepted by some witness automata in  $\textit{Auts}_{\textit{1-eq}}$, let $y$ be the item that are renamed into $b$ by $r$. Then it is not hard to see that $y$ satisfies one of three conditions in $e_2$.

\noindent The $\textit{if}$ direction: Assume that there exists such $e \in \mathcal{I}_{\neq}$ and $x$. Let renaming function $r$ maps $x$ into $b$ and all other items into $a$. By data-independence, we can see that $r(e) \in \mathcal{I}$. Then it is easy to see that $r(e)$ is accepted by some automaton in $\textit{Auts}_{\textit{2-ms}}$. \qed
\end {proof}

Similar to Lemma \ref{lemma:EPQ3 is co-regular}, we can prove that $\textit{MSet}_3$ is co-regular, as stated by the following lemma.

\begin{restatable}{lemma}{MSet3IsCoRegular}
\label{lemma:MSet3 is co-regular}
$\textit{MSet}_3$ is co-regular.
\end{restatable}

Similar as witness automata for $\textit{EPQ}_3$, we have witness automata $\mathcal{A}_{\textit{MS}}^4$ for $\textit{MSet}_3$, which is shown in \figurename~\ref{fig:automata 4 for MS-3 in appendix}. In \figurename~\ref{fig:automata 4 for MS-3 in appendix}, let $c = \textit{cal}(\textit{insert},d),\textit{ret}(\textit{insert},d), \textit{cal}(\textit{delete},d)$, $\textit{ret}(\textit{delete},d),\textit{cal}(\textit{delete},\textit{empty}),\textit{ret}(\textit{delete},\textit{empty})$, $c_1 = c + \textit{cal}(\textit{insert},b)$, $c_2 = c_1 + \textit{ret}(\textit{delete},b)$, and $c_3 = c + \textit{ret}(\textit{delete},b)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8 \textwidth]{figures/PIC_AUTO_MS_4.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{MS}}^4$}
  \label{fig:automata 4 for MS-3 in appendix}
\end{figure}

The following lemma shows that our transformation from multi-set into extended priority queue is correct.

\RelateMultiSetwithEPQ*

\begin {proof}

For the $\textit{only if}$ direction, given an execution $e_m \in \mathcal{I}_m$ of multi-set, such that $e_m$ is accepted by some automaton in $\textit{Auts}_{\textit{MS}}$. If $e_m$ is accepted by $\mathcal{A}_{\textit{MS}}^1$, $\mathcal{A}_{\textit{MS}}^2$, $\mathcal{A}_{\textit{MS}}^3$ or $\mathcal{A}_{\textit{MS}}^4$, then it is easy to see that $\textit{MStoEPQ}(e_m)$ is accepted by $\mathcal{A}_{\textit{SinPri}}^2$, $\mathcal{A}_{\textit{SinPri}}^3$, $\mathcal{A}_{\textit{SinPri}}^1$ or $\mathcal{A}_{\textit{EPQ}}^3$, respectively.

For the $\textit{if}$ direction, given an execution $e_{\textit{epq}}$ of extended priority queue, such that $e_{\textit{epq}} = \textit{MStoEPQ}(e_m)$ for some $e_m$ of multi-set, and $e_{\textit{epq}}$ is accepted by some automaton in $\textit{Auts}_{\textit{EPQ}}$.

According to definition of $\textit{MStoEPQ}$, we can see that (1) in $e_{\textit{epq}}$ there does not exist two items with comparable priorities, and (2) in $e_{\textit{epq}}$, there does not exists two items with a same priority. Therefore, the witness automata for $\textit{EPQ}_1$ and the witness automaton $\mathcal{A}_{\textit{SinPri}}^4$ can be safely ignored. Then, if $e_{\textit{epq}}$ is accepted by $\mathcal{A}_{\textit{SinPri}}^1$, $\mathcal{A}_{\textit{SinPri}}^2$, $\mathcal{A}_{\textit{SinPri}}^3$ or $\mathcal{A}_{\textit{EPQ}}^3$, then it is easy to see that $e_m$ is accepted by $\mathcal{A}_{\textit{MS}}^3$, $\mathcal{A}_{\textit{MS}}^1$, $\mathcal{A}_{\textit{MS}}^2$ and $\mathcal{A}_{\textit{MS}}^4$, respectively. \qed
\end {proof}

\subsection{Proofs and Definitions in Subsection \ref{subsec:relate stack with extended priority queue}}
\label{subsec:appendix proof and definition in section relate stack with extended priority queue}

### 2017-04-04 co-regular-EPQ.tex
add tex

--- 2017-03-04 co-regular-PQ.tex
@@
-\section{Reducing Linearizability of Priority Queues into State Reachability}
-\label{sec:co-regular of priority queues}
+\section{Reducing Linearizability of Extended Priority Queues into State Reachability}
+\label{sec:co-regular of extended priority queues}
 
-A rule $R$ of priority queue is co-regular, if checking linearizability with respect to $\textit{MS}(R)$ of a data-independent implementation $\mathcal{I}$ can be reduced to checking the emptiness of intersection between $\mathcal{I}$ and a set of witness automata.
+A rule $R$ of extended priority queue is co-regular, if checking linearizability with respect to $\textit{MS}(R)$ of a data-independent implementation $\mathcal{I}$ can be reduced to checking the emptiness of intersection between $\mathcal{I}$ and a set of witness automata.
 In this section, we propose the definition of witness automata and co-regular.
-Then we prove that all five rules of priority queues are co-regular, and roughly introduce the proof idea.
-With the help of step-by-step linearizability and co-regular, we finally reduce the linearizability problem of $\textit{PQueue}$ into emptiness problem of intersection with automata.
+Then we prove that all five rules of extended priority queues are co-regular, and roughly introduce the proof idea.
+With the help of step-by-step linearizability and co-regular, we finally reduce the linearizability problem of $\textit{EPQ}$ into emptiness problem of intersection with automata.
 
@@
 
-A witness automaton is a finite automaton with alphabet $\{ \textit{cal}(\textit{put},d,\textit{pred}), \textit{ret}(\textit{put},d),$ $\textit{cal}(\textit{rm},d),\textit{ret}(\textit{rm},d) \vert d \in \mathbb{D},\textit{pred} \in \textit{predWA} \}$.
+A witness automaton is a finite automaton with alphabet $\{ \textit{cal}(\textit{put},d,\textit{pred}), \textit{ret}(\textit{put},d),$ $\textit{cal}(\textit{rm},d),\textit{ret}(\textit{rm},d) \vert d \in \mathbb{D} \cup \{ \textit{empty} \},\textit{pred} \in \textit{predWA} \}$.
 Here $\textit{predWA}$ is the set of predicate of priorities, and it contains
@@
 
-\item[-] A predicate $\textit{les}_p$, which accepts all the priorities that are less than the value of $p$,
+\item[-] A predicate $\textit{les}_p$, which accepts all the priorities that are smaller than the value of $p$ according to $<_{\mathbb{P}}$,
 
@@
 
-Given a execution $e = \alpha_1 \cdots \alpha_k$ of priority queue and a witness automaton $\mathcal{A}$, we say that $e$ is accepted by $\mathcal{A}$, if
+Given a execution $e = \alpha_1 \cdots \alpha_k$ of extended priority queue and a witness automaton $\mathcal{A}$, we say that $e$ is accepted by $\mathcal{A}$, if
 
@@
 \item[-] There exist transitions $q_0 \xrightarrow{\beta_1} q_1 \ldots \xrightarrow{\beta_k} q_k$ of $\mathcal{A}$, such that $q_0$ is one of initial state of $\mathcal{A}$, and $q_k$ is one of accept state of $\mathcal{A}$.
-Let $d_p \in \mathbb{D}$.
+We choose value of $p$ as some value $d_p \in \mathbb{D}$.
 
-\item[-] For each $i$, if $\alpha_i = \textit{cal}(\textit{put},a,q)$, then either (1) $\beta_i = \textit{cal}(\textit{put},a,p)$ and $q = d_p$, or (2) $\beta_i = \textit{cal}(\textit{put},a,\textit{les}_p)$ and $q < d_p$, or (3) $beta_i = \textit{cal}(\textit{put},a,\textit{anyPri})$.
+\item[-] For each $i$, if $\alpha_i = \textit{cal}(\textit{put},a,q)$, then either (1) $\beta_i = \textit{cal}(\textit{put},a,p)$ and $q = d_p$, or (2) $\beta_i = \textit{cal}(\textit{put},a,\textit{les}_p)$ and $q <_{\mathbb{P}} d_p$, or (3) $beta_i = \textit{cal}(\textit{put},a,\textit{anyPri})$.
 
@@
 %\vspace{-6pt}
-\begin{definition}\label{def:co-regular of rules of priority queues}
-A rule $R$ of priority queue is co-regular, if there are a finite set $\textit{Auts}_{R}$ of witness automata such that, for each data-independence implementation $\mathcal{I}$, we have that
+\begin{definition}\label{def:co-regular of rules of extended priority queues}
+A rule $R$ of extended priority queue is co-regular, if there are a finite set $\textit{Auts}_{R}$ of witness automata such that, for each data-independence implementation $\mathcal{I}$, we have that
 
-$$ \textit{Auts}_{R} \cap \mathcal{I} \neq \emptyset \Leftrightarrow \exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), last(e')=R \wedge e' \ does \ not \ linearizable \ w.r.t.
+$$\textit{Auts}_{R} \cap \mathcal{I} \neq \emptyset \Leftrightarrow \exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), R \in last(e') \wedge e' \ does \ not \ linearizable \ w.r.t.
 \ \textit{MS}(R)$$
 
-We say that $\textit{PQueue}$ is regular, if each of its rule is co-regular.
+We say that $\textit{EPQ}$ is co-regular, if each of its rule is co-regular.
 \end{definition}
@@
 \cite{Bouajjani:2015} states that checking linearizability w.r.t queue can be reduced into checking emptiness of intersection between $\mathcal{I}$ and a set of automata.
-Given a data-differentiated execution $e$, let $e \vert_{i}$ be an execution generated from $e$ by erasing call and return actions of items that does not use priority $i$.
-We call a priority queue execution with only one priority a single-priority execution.
+Given a data-differentiated execution $e$, let $e \vert_{i}$ be an execution generated from $e$ by erasing call and return actions of items that does not use priority $i$ (does not influence $\textit{rm}(\textit{empty})$).
+We call a extended priority queue execution with only one priority a single-priority execution.
 Let $\textit{transToQueue}(e)$ be an execution generated from $e$ by transforming $\textit{put}$ and $\textit{rm}$ into $\textit{enq}$ and $\textit{deq}$, respectively, and then discarding priorities.
-We can see that for each $e \in \textit{PQueue}$ and each priority $i$, $\textit{transToQueue}(e \vert_{i})$ satisfy FIFO (first in first out) property.
+We can see that for each $e \in \textit{EPQ}$ and each priority $i$, $\textit{transToQueue}(e \vert_{i})$ satisfy FIFO (first in first out) property.
 
@@
 \cite{Bouajjani:2015} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $a,b$: (1) $\textit{deq}(b) <_{hb} \textit{enq}(b)$, (2) there are no $\textit{enq}(b)$ and at least one $\textit{deq}(b)$, (3) there are one $\textit{enq}(b)$ and more than one $\textit{deq}(b)$, and (4) $\textit{enq}(a) <_{\textit{hb}} \textit{enq}(b)$, and $\textit{deq}(b) <_{\textit{hb}} \textit{deq}(a)$, or $\textit{deq}(a)$ does not exists.
-For each such case, we can construct a witness automata for priority queue.
+For each such case, we can construct a witness automata for extended priority queue.
 For example, for the first case, we generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ in \figurename~\ref{fig:automata for FIFO-1}, here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}), \textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a)$, $\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},b),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$.
@@
 
-In Appendix \ref{sec:appendix proof and definition in section definition of co-regular}, we construct a set $\textit{Auts}_{\textit{sinPri}}$ of witness automata ($\mathcal{A}_{\textit{SinPri}}^1$ is in it), and shows that they are enough to ensure that for each of its data-differentiated executions, each of its single-priority projection that has no $\textit{rm}(\textit{empty})$ to have ``FIFO'' property, as shown by the following lemma.
+In Appendix \ref{sec:appendix proof and definition in section definition of co-regular}, we construct a set $\textit{Auts}_{\textit{sinPri}}$ of witness automata ($\mathcal{A}_{\textit{SinPri}}^1$ is in it), and shows that they are enough to ensure that for each data-differentiated executions, each of its single-priority projection without $\textit{rm}(\textit{empty})$ to have ``FIFO'' property, as shown by the following lemma.
 
-\begin{restatable}{lemma}{AutoForPQwithSignlePri}
-\label{lemma:automata for priority queue with single priority}
+\begin{restatable}{lemma}{AutoForEPQwithSignlePri}
+\label{lemma:automata for extended priority queue with single priority}
 
-Given a data-independent implementations $\mathcal{I}$ of priority queue, $\mathcal{I} \cap \textit{Auts}_{\textit{sinPri}} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{transToQueue}(e')$ does not linearizable to queue.
+Given a data-independent implementations $\mathcal{I}$ of extended priority queue, $\mathcal{I} \cap \textit{Auts}_{\textit{sinPri}} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{transToQueue}(e')$ does not linearizable to queue.
 \end{restatable}
 
-According to Lemma \ref{lemma:automata for priority queue with single priority}, from now on, it is safe to assume that, for each data-differentiated execution without $\textit{rm}(\textit{empty})$, any of its single-priority projection has ``FIFO'' property.
+According to Lemma \ref{lemma:automata for extended priority queue with single priority}, from now on, it is safe to assume that, for each data-differentiated execution without $\textit{rm}(\textit{empty})$, any of its single-priority projection has ``FIFO'' property.
 For example, $\textit{rm}(a)$ never happens before $\textit{put}(a,\_)$ for each $a$.
 
-\subsection{Co-Regular of $\textit{PQ}_1^{>}$}
-\label{subsec:co-regular of PQ1Lar}
+\subsection{Co-Regular of $\textit{EPQ}_1^{>}$}
+\label{subsec:co-regular of EPQ1Lar}
 
 In this subsection, we introduce the idea for proving co-regular of $\textit{PQ}_1^{>}$.
-The proof of this subsection can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of PQ1Lar}.
-The notion of left-right constraint used for priority queue is inspired by left-right constraint of queue \cite{Bouajjani:2015}.
+The proof of this subsection can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
+The notion of left-right constraint used for extended priority queue is inspired by left-right constraint of queue \cite{Bouajjani:2015}.
 
-Given a data-differentiated execution $e$ such that $\textit{last}(e) = \textit{PQ}_1^{>}$, although Lemma \ref{lemma:automata for priority queue with single priority} ensures that each single-priority projection of $e$ satisfy the FIFO property, this is still not enough for ensuring that $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{>})$.
-Since it is possible that $e$ does not linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{>})$ because of interaction between actions of multiple priorities.
+Given a data-differentiated execution $e$, we say that $e$ is a $\textit{pri}$-execution, if $\textit{pri}$ is the maximal priority of $e$, and $\textit{pri}$ is larger than or equal to the priority of all other items of $e$.
+Given a data-differentiated execution $e$ and assume that $\textit{pri}$ is one of the maximal priority of $e$, let $\textit{pri-Exec}(e)$ be an execution obtained from $e$ by erasing all operations of items whose priority are incomparable with $\textit{pri}$.
+The following lemma states that when checking co-regular of $\textit{EPQ}_1^{>}$, $\textit{EPQ}_1^{=}$, $\textit{EPQ}_2^{>}$ and $\textit{EPQ}_2^{=}$, we need only consider $\textit{pri}$-executions.
 
-We give an example of such execution $e$ in \figurename~\ref{fig:introduce gap for PQ1Lar}.
+\begin{restatable}{lemma}{priExecutionIsEnough}
+\label{lemma:pri execution is enough}
+Given a data-differentiated execution $e$ and $R = \textit{EPQ}_1^{>}$ (resp., $R = \textit{EPQ}_1^{=}$, $R=\textit{EPQ}_2^{>}$, or $R = \textit{EPQ}_2^{=}$).
+$e \sqsubseteq \textit{MS}(R)$ with witness $x$, if and only if $\textit{pri-Exec}(e) \sqsubseteq \textit{MS}(R)$ with witness $x$, where $\textit{pri}$ is the priority of $x$.
+%for each $\textit{pri}$ that is (1) maximal in $e$ and (2) items of $\textit{pri}$ are only one pair of matched $\textit{put}$ in $e$ (resp., more than one pair of matched $\textit{put}$, unmatched $\textit{put}$, or unmatched $\textit{put}$ and matched $\textit{put}$).
+\end{restatable}
+
+Lemma \ref{lemma:pri execution is enough} simplifies our proof by make us safely ignore all the items that has priorities incomparable with $\textit{pri}$.
+When construct witness automata, priorities which are incomparable with $\textit{pri}$ can be safely represented by $\textit{anyPri}$.
+Given a $\textit{pri}$-execution $e'$, we can see that $\textit{last}(e')$ contains only one rules.
+
+Given a data-differentiated $\_$-execution $e$ such that $\textit{last}(e) = \textit{EPQ}_1^{>}$, although Lemma \ref{lemma:automata for extended priority queue with single priority} ensures that each single-priority projection of $e$ satisfy the FIFO property, this is still not enough for ensuring that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$.
+Since it is possible that $e$ does not linearizable w.r.t $\textit{MS}(\textit{EPQ}_1^{>})$ because of interaction between actions of multiple priorities.
+
+We give an example of such execution $e$ in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
 We call the time interval from $\textit{ret}(\textit{put},x)$ to $\textit{cal}(rm,x)$, or from $\textit{ret}(\textit{put},x)$ when $\textit{cal}(rm,x)$ does not exist, the interval of item $x$.
-In \figurename~\ref{fig:introduce gap for PQ1Lar}, we draw the interval of each item by dashed line.
-The reason of why $e$ does not linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{>})$ is that, each time point from $\textit{cal}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is in interval of some item with smaller priority than priority of $a$.
+In \figurename~\ref{fig:introduce gap for EPQ1Lar}, we draw the interval of each item by dashed line.
+Here we assume that $p_1 <_{\mathbb{P}} p_4$, $p_2 <_{\mathbb{P}} p_4$ and $p_3 <_{\mathbb{P}} p_4$.
+The reason of why $e$ does not linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{>})$ is that, each time point from $\textit{cal}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is in interval of some item with smaller priority.
 
@@
   \centering
-  \includegraphics[width=0.6 \textwidth]{figures/PIC-HIS-INTRO-GAP-PQ1L.pdf}
+  \includegraphics[width=0.6 \textwidth]{figures/PIC-HIS-INTRO-GAP-EPQ1L.pdf}
 %\vspace{-10pt}
-  \caption{An execution that does not linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{>})$}
-  \label{fig:introduce gap for PQ1Lar}
+  \caption{An execution that does not linearizable w.r.t $\textit{MS}(\textit{EPQ}_1^{>})$}
+  \label{fig:introduce gap for EPQ1Lar}
 \end{figure}
 
-In this subsection, intuitively we will prove that, as long as we can get rid of the case in \figurename~\ref{fig:introduce gap for PQ1Lar}, we can ensure $\textit{last}(e) = \textit{PQ}_1^{>} \Rightarrow e \sqsubseteq \textit{MS}(\textit{PQ}_1^{>})$.
+In this subsection, intuitively we will prove that, as long as we can get rid of the case in \figurename~\ref{fig:introduce gap for EPQ1Lar}, we can ensure $\textit{last}(e) = \textit{EPQ}_1^{>} \Rightarrow e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$.
 
-Let us introduce the notion of left-right constraint, which is a graph, and the existence of cycle though item with maximal priority in it correspond to the existence of the case in \figurename~\ref{fig:introduce gap for PQ1Lar}.
+Let us introduce the notion of left-right constraint, which is a graph, and the existence of cycle though item with maximal priority in it correspond to the existence of the case in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
 
 \begin{definition}\label{def:left-right constraint for matched put and rm operations}
-Given a data-differentiated execution $e$ without $\textit{rm}(\textit{empty})$.
-Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$, where $\textit{pri}_x$ is maximal in $e$ and only used by these two method events.
+Given a data-differentiated $\textit{pri}_x$-execution $e$ without $\textit{rm}(\textit{empty})$.
+Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be the only  method events of $e$ with priority $\textit{pri}_x$.
 The left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is the graph $G$ where:
@@
 
-\item[-] there is an edge from item $d_1$ to $x$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$,
+\item[-] there is an edge from item $d_1$ to $x$, if $\textit{put}(d_1,\_)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$,
 
@@
 
-\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{rm}(d_2,\textit{pri}_2)$.
+\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$.
 \end{itemize}
@@
 When there is a cycle $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ through item with maximal priority (for example, $x$) in $G$, we say that $x$ is covered by $d_1,\ldots,d_m$.
-To state the effectiveness of left-right constraint, take the execution in \figurename~\ref{fig:introduce gap for PQ1Lar} as an example, we can see that $b$ is covered by $f,e,d$.
-We need to prove that getting rid of cycle though item with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{>})$, as stated by the following lemma:
+To state the effectiveness of left-right constraint, take the execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} as an example, we can see that $b$ is covered by $f,e,d$.
+We need to prove that getting rid of cycle though item with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\textit{MS}(\textit{EPQ}_1^{>})$, as stated by the following lemma:
 
-\begin{restatable}{lemma}{LinEqualsConstraintforPQOneLar}
-\label{lemma:Lin Equals Constraint for PQ1Lar}
-Given a data-differentiated execution $e$ with $\textit{last}(e) = \textit{PQ}_1^{>}$.
+\begin{restatable}{lemma}{LinEqualsConstraintforEPQOneLar}
+\label{lemma:Lin Equals Constraint for EPQ1Lar}
+Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{>}$.
 Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority.
 Let $G$ be the graph representing the left-right constraint of $\textit{put}(x)$ and $\textit{rm}(x)$.
-$e \sqsubseteq \textit{MS}(\textit{PQ}_1^{>})$, if and only if $G$ has no cycle going through $x$.
+$e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$, if and only if $G$ has no cycle going through $x$.
 \end{restatable}
@@
 The $\textit{only if}$ direction can be easily proved by contradiction and is omitted here.
-To prove the $\textit{if}$ direction, we need to explicitly construct the linearization of $e$, or we can say, we need construct the $u$, $v$ and $w$ in $\textit{PQ}_1^{>}$.
+To prove the $\textit{if}$ direction, we need to explicitly construct the linearization of $e$, or we can say, we need construct the $u$, $v$ and $w$ in $\textit{EPQ}_1^{>}$.
 Let $u$ to be the sequence of all operations that happens before $\textit{put}(x)$.
@@
 Let $\textit{UVSet}(e,x) = \textit{UVSet}_1(e,x) \cup \textit{UVSet}_2(e,x) \cup \ldots$.
-For example, in \figurename~\ref{fig:his nobound of LMSet}, $\textit{UVSet}_1(e,x) = \{ \textit{put}(a,1),\textit{rm}(a) \}$, $\textit{USSet}_2(e,x) = \{ \textit{put}(b,1),\textit{rm}(b) \}$ and $\textit{UVSet}_3(e,x) = \{ \textit{put}(c,1),\textit{rm}(c) \}$.
+For example, assume that $p_1 <_{\mathbb{P}} p_2$, then in \figurename~\ref{fig:his nobound of UVSet}, $\textit{UVSet}_1(e,x) = \{ \textit{put}(a,p_1),\textit{rm}(a) \}$, $\textit{USSet}_2(e,x) = \{ \textit{put}(b,p_1),\textit{rm}(b) \}$ and $\textit{UVSet}_3(e,x) = \{ \textit{put}(c,p_1),\textit{rm}(c) \}$.
 Similarly, we can generate execution $e'$ such that, for each $i$, $\textit{UVSet}(e',x) \neq \emptyset$.
@@
   \centering
-  \includegraphics[width=0.5 \textwidth]{figures/PIC_HIS_NOBOUNDOF_LMSET.pdf}
+  \includegraphics[width=0.5 \textwidth]{figures/PIC_HIS_NOBOUNDOF_UVSET.pdf}
 %\vspace{-10pt}
   \caption{An example for $\textit{UVSet}(e,x)$}
-  \label{fig:his nobound of LMSet}
+  \label{fig:his nobound of UVSet}
 \end{figure}
@@
 
-According to Lemma \ref{lemma:Lin Equals Constraint for PQ1Lar}, the existence of ``cover item with maximal priority'' is enough for checking violation of linearizability w.r.t $\textit{MS}(\textit{PQ}_1^{>})$.
+According to Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar}, the existence of ``An item covered by items of smaller priority'' is enough for checking violation of linearizability w.r.t $\textit{MS}(\textit{EPQ}_1^{>})$.
 Assume $x$ is covered by $d_1,\ldots,d_m$.
@@
 For example, for the case when $e \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^1$, as shown in \figurename~\ref{fig:automata APQ1Lar-1}.
-Here $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{cal}(\textit{put},a,\textit{les}_p)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d), \textit{ret}(\textit{rm},d)$.
+Here $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{cal}(\textit{put},a,\textit{les}_p)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d), \textit{ret}(\textit{rm},d),\textit{cal}(\textit{rm},\textit{empty}), \textit{ret}(\textit{rm},\textit{empty})$.
 
@@
 
-Given the execution $e$ in \figurename~\ref{fig:introduce gap for PQ1Lar}, let $e'$ be generated from $e$ by renaming $d,e,f$ into $a$.
+Given the execution $e$ in \figurename~\ref{fig:introduce gap for EPQ1Lar}, let $e'$ be generated from $e$ by renaming $d,e,f$ into $a$.
 Then we can see that $e'$ is accepted by $\mathcal{A}_{\textit{l-lar}}^1$ with a path via $q_{\textit{init}}, q_2, q_3, q_4,q_5,q_7,q_8$.
 
-In Appendix \ref{sec:appendix proof and definition in section co-regular of PQ1Lar}, we construct a set $\textit{Auts}_{\textit{1-lar}}$ of witness automata ($\mathcal{A}_{\textit{l-lar}}^1$ is in it), and use $\textit{Auts}_{\textit{1-lar}}$ to prove that $\textit{PQ}_1^{>}$ is co-regular, as stated by the following lemma.
+In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we construct a set $\textit{Auts}_{\textit{1-lar}}$ of witness automata ($\mathcal{A}_{\textit{l-lar}}^1$ is in it), and use $\textit{Auts}_{\textit{1-lar}}$ to prove that $\textit{EPQ}_1^{>}$ is co-regular, as stated by the following lemma.
 
-\begin{restatable}{lemma}{PQOneLarisCoRegular}
-\label{lemma:PQ1Lar is co-regular}
-$\textit{PQ}_1^{>}$ is co-regular.
+\begin{restatable}{lemma}{EPQOneLarisCoRegular}
+\label{lemma:EPQ1Lar is co-regular}
+$\textit{EPQ}_1^{>}$ is co-regular.
 \end{restatable}
 
-\subsection{Co-Regular of $\textit{PQ}_1^{=}$}
-\label{subsec:co-regular of PQ1Equal}
+\subsection{Co-Regular of $\textit{EPQ}_1^{=}$}
+\label{subsec:co-regular of EPQ1Equal}
 
-In this subsection, we introduce the idea for proving co-regular of $\textit{PQ}_1^{=}$.
-The proof of this subsection can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of PQ1Equal}.
+In this subsection, we introduce the idea for proving co-regular of $\textit{EPQ}_1^{=}$.
+The proof of this subsection can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}.
 
-Given a data-differentiated execution $e$ such that $\textit{last}(e) = \textit{PQ}_1^{=}$.
-Lemma \ref{lemma:automata for priority queue with single priority} ensures that each single-priority projection of $e$ satisfy the FIFO property, and with the help of Lemma \ref{lemma:PQ1Lar is co-regular} and $\textit{Auts}_{\textit{1-lar}}$, we can ensure that each item in $e$ can not be covered by items with smaller priority.
-However, these are sill not enough for ensure that $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
-This is because that there are multiple matched $\textit{put}$ (assume they use items in $S$), and all the possible linearization point of $\textit{rm}(b)$ may be disabled by $\textit{rm}(b)$ for some $a,b \in S$.
+Given a data-differentiated $\_$-execution $e$ such that $\textit{last}(e) = \textit{EPQ}_1^{=}$.
+Lemma \ref{lemma:automata for extended priority queue with single priority} ensures that each single-priority projection of $e$ satisfy the FIFO property, and with the help of Lemma \ref{lemma:EPQ1Lar is co-regular} and $\textit{Auts}_{\textit{1-lar}}$, we can ensure that each item in $e$ can not be covered by items with smaller priority.
+However, these are sill not enough for ensure that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
+This is because that given items $a$ and $b$ with maximal priority, it is possible that all the possible linearization point of $\textit{rm}(b)$ may be disabled by $\textit{rm}(b)$.
 
 We give an example of such execution $e$ in \figurename~\ref{fig:introduce pb order}.
-The execution $e$ of \figurename~\ref{fig:introduce pb order} is not linearizable, even if $h \vert_{1}$ and $h \vert_{4}$ are both linearizable, and either $a$ or $b$ is covered by other items.
+The execution $e$ of \figurename~\ref{fig:introduce pb order} is not linearizable, even if $h \vert_{p_1}$ and $h \vert_{p_4}$ are both linearizable, and either $a$ or $b$ is covered by items with smaller priority.
 To explain this, me need to observe the following three points:
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] Since $\textit{put}(a,1)$ happens before $\textit{put}(b,1)$, we know that $a$ is ``putted earlier'' than $b$, and therefore, the linearization points of $\textit{rm}(a)$ should before the linearization point of $\textit{rm}(b)$.
+\item[-] Since $\textit{put}(a,p_4)$ happens before $\textit{put}(b,p_4)$, we know that $a$ is ``putted earlier'' than $b$, and therefore, the linearization points of $\textit{rm}(a)$ should before the linearization point of $\textit{rm}(b)$.
 
-\item[-] The time intervals satisfy the following conditions (we identify them with dotted lines in \figurename~\ref{fig:introduce pb order}) are possible position to locate the linearization point of $\textit{rm}(b)$ according to Lemma \ref{lemma:Lin Equals Constraint for PQ1Lar} (Here we temporarily forget the existence of $a$): (1) between from $\textit{cal}(\textit{rm},b)$ and $\textit{ret}(\textit{rm},b)$, (2) does not in interval of any item with smaller priority.
+\item[-] The time intervals satisfy the following conditions (we identify them with dotted lines in \figurename~\ref{fig:introduce pb order}) are possible position to locate the linearization point of $\textit{rm}(b)$ according to Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar} (Here we temporarily forget the existence of $a$): (1) between $\textit{cal}(\textit{rm},b)$ and $\textit{ret}(\textit{rm},b)$, (2) does not in interval of any item with smaller priority.
 
@@
   \centering
-  \includegraphics[width=0.6 \textwidth]{figures/PIC-HIS-INTRO-PB-ORDER.pdf}
+  \includegraphics[width=0.6 \textwidth]{figures/PIC-HIS-INTRO-PB-ORDER-EPQ.pdf}
 %\vspace{-10pt}
-  \caption{An execution that does not linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{=})$}
+  \caption{An execution that does not linearizable w.r.t $\textit{MS}(\textit{EPQ}_1^{=})$}
   \label{fig:introduce pb order}
@@
 
-In this subsection, intuitively we will prove that, as long as we can get rid of the case in \figurename~\ref{fig:introduce pb order}, we can ensure $\textit{last}(e) = \textit{PQ}_1^{=} \Rightarrow e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
+In this subsection, intuitively we will prove that, as long as we can get rid of the case in \figurename~\ref{fig:introduce pb order}, we can ensure $\textit{last}(e) = \textit{EPQ}_1^{=} \Rightarrow e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
 
-Let us introduce $<_{\textit{pb}}$ (represents put-before) order to formally state that ``an item is putted before another''.
-Given a data-differentiated execution $e$ and two items $a,b$ with maximal priority in $e$,  we say that $a <_{\textit{pb}} b$, if one of the following cases holds:
+Let us introduce $<_{\textit{pb}}$ (represents put-before) order to formally state that ``an item is putted before another item''.
+Given a data-differentiated $\_$-execution $e$ and two items $a,b$ with maximal priority in $e$, we say that $a <_{\textit{pb}} b$, if one of the following cases holds:
 
@@
 \begin{definition}\label{def:gap-point for matched put and rm operations}
-Given a data-differentiated execution $e$ and two method events $\textit{put}(x,\textit{pri}_x),\textit{rm}(x)$ of $e$, where $\textit{pri}_x$ is the maximal priority in $e$.
-We say that a time-point $o$ is a gap-point of $x$, if
+Given a data-differentiated $\textit{pri}_x$-execution $e$ and two method events $\textit{put}(x,\textit{pri}_x),\textit{rm}(x)$ of $e$. We say that a time-point $o$ is a gap-point of $x$, if
 
@@
 
-\begin{restatable}{lemma}{PQOneEqualAsPBandGP}
-\label{lemma:PQ1Equal as pb order and gap-point}
-Given a data-differentiated execution $e$  with $\textit{last}(e) = \textit{PQ}_1^{=}$.
-$e$ does not linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{=})$, if and only if there exists $x$ and $y$ with maximal priority in $e$, such that $y <_{\textit{pb}}^* x$ in $e$, and the rightmost gap-point of $x$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$ in $e$.
+\begin{restatable}{lemma}{EPQOneEqualAsPBandGP}
+\label{lemma:EPQ1Equal as pb order and gap-point}
+Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{=}$.
+$e$ does not linearizable w.r.t $\textit{MS}(\textit{EPQ}_1^{=})$, if and only if there exists $x$ and $y$ with maximal priority $\textit{pri}$ in $e$, such that $y <_{\textit{pb}}^* x$ in $e$, and the rightmost gap-point of $x$ is before $\textit{cal}(\textit{put},y,\textit{pri})$ or $\textit{cal}(\textit{rm},y)$ in $e$.
 \end{restatable}
@@
 To prove the $\textit{only if}$ direction, we prove its contrapositive.
-Assume we already know that for each $x$ and $y$ has maximal priority in $e$, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},x)$.
-We need to prove that $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
+Assume we already know that for each $x$ and $y$ has maximal priority in $e$, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{cal}(\textit{put},y,\textit{pri})$ and $\textit{cal}(\textit{rm},x)$.
+We need to prove that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
 Or we can say, we need to explicitly construct linearization of $e$.
-Let $\textit{pri}$ be the maximal priority of $e$.
 
-We introduce another lemma (Lemma \ref{lemma:maximal in pb and gap-point make a candidate of PQ1Equal}), which states that: If $\exists g_1 \in \textit{Items}(e,\textit{pri})$, such that $\forall g_2 \in \textit{Items}(e,\textit{pri})$, (1) $g_1$ does not $<_{\textit{pb}}$ to $g_2$, and (2) the right-most gap-point of $g_1$ is after $\textit{cal}(\textit{put},g_2,\_)$ and $\textit{cal}(\textit{rm},g_2)$.
-Then $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
+We introduce another lemma (Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}), which states that: If $\exists g_1 \in \textit{Items}(e,\textit{pri})$, such that $\forall g_2 \in \textit{Items}(e,\textit{pri})$, (1) $g_1$ does not $<_{\textit{pb}}$ to $g_2$, and (2) the right-most gap-point of $g_1$ is after $\textit{cal}(\textit{put},g_2,\textit{pri})$ and $\textit{cal}(\textit{rm},g_2)$.
+Then $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
 The proof of this lemma also tell us how to construct linearization in such case.
@@
 
-\item[-] Step $1$: check if $a_1$ satisfy the conditions of \ref{lemma:maximal in pb and gap-point make a candidate of PQ1Equal}.
+\item[-] Step $1$: check if $a_1$ satisfy the conditions of \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}.
 
     It is obvious that $a_1$ satisfy the first condition.
-If the second condition is also satisfied, by Lemma \ref{lemma:maximal in pb and gap-point make a candidate of PQ1Equal}, we can obtain that $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
+If the second condition is also satisfied, by Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, we can obtain that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
 
@@
 $a^f$ satisfies the check of Step $1$.
-By Lemma \ref{lemma:maximal in pb and gap-point make a candidate of PQ1Equal}, this implies that $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
+By Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, this implies that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
 This completes the proof of $\textit{if}$ direction.
@@
 
-The result of Lemma \ref{lemma:PQ1Equal as pb order and gap-point} is not quite suitable for verificaion with automata.
-The reason is that we need to ensure $y <_{\textit{pb}}^* x$, while according to definition of $<_{\textit{pb}}^*$, there may be arbitrary intermediate items $a_1,\ldots,a_m$, such that $a <_{\textit{pb}} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}} a_m <_{\textit{pb}} b$.
+The result of Lemma \ref{lemma:EPQ1Equal as pb order and gap-point} is not quite suitable for verificaion with automata.
+The reason is that we need to ensure $y <_{\textit{pb}}^* x$, while according to definition of $<_{\textit{pb}}^*$, there may be arbitrary intermediate items $a_1,\ldots,a_m$, such that $x <_{\textit{pb}} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}} a_m <_{\textit{pb}} y$.
 It is hard to store unbounded $a_i$ by automata.
@@
 The number of potential enumerations can be further reduced.
-The reason is that Lemma \ref{lemma:PQ1Equal as pb order and gap-point} requires both $<_{\textit{pb}}^*$ and gap-point, while some combination of $<_{\textit{pb}}^*$ and gap-points are conflict.
+The reason is that Lemma \ref{lemma:EPQ1Equal as pb order and gap-point} requires both $<_{\textit{pb}}^*$ and gap-point, while some combination of $<_{\textit{pb}}^*$ and gap-points are conflict.
 For example, if $a <_{\textit{pb}}^B b$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$, then since (1) $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(b)$, (2) each gap-point of $b$ is after $\textit{cal}(\textit{rm},b)$ (also $\textit{ret}(\textit{rm},a)$), we can see that the right-most gap-point of $b$ is before $\textit{cal}(\textit{put},a,\_)$ and after $\textit{ret}(\textit{rm},a)$, which implies $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a,\_)$ and is impossible.
@@
 
-\begin{restatable}{lemma}{FiveEnmuerationisEnoughForPQOneEqual}
-\label{lemma:five enumeration is enough for PQ1Equal}
-Given a data-differentiated execution $e$ with $\textit{last}(e) = \textit{PQ}_1^{=}$.
-Let $a$ and $b$ be items with maximal priority in $e$.
-Assume that $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$.
+\begin{restatable}{lemma}{FiveEnmuerationisEnoughForEPQOneEqual}
+\label{lemma:five enumeration is enough for EPQ1Equal}
+Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{=}$.
+Let $a$ and $b$ be items with maximal priority $\textit{pri}$.
+Assume that $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\textit{pri})$ or $\textit{cal}(\textit{rm},a)$.
 Then, there are five possible enumeration of method events of $a$, $b$, $a_1$ (if exists), where $a_1$ is the possible intermediate items for obtain $a <_{\textit{pb}}^* b$.
@@
 For example, to capture the first enumeration, we generate witness automaton $\mathcal{A}_{\textit{l-eq}}^1$, as shown in \figurename~\ref{fig:automata for first enumeration of PQ1Equal not in appendix}.
-Here we rename the items that ``covers the time interval from $\textit{cal}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$'' into $d$ (see Appendix \ref{sec:appendix proof and definition in section co-regular of PQ1Equal}), and rename the renaming items into $e$.
-In this figure, $c = \textit{cal}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{cal}(\textit{rm},e), \textit{ret}(\textit{rm},e)$, $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.
+Here we rename the items that ``covers the time interval from $\textit{cal}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$'' (see Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}) into $d$, and rename the renaming items into $e$.
+In this figure, $c = \textit{cal}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{cal}(\textit{rm},e), \textit{ret}(\textit{rm},e),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.
 
@@
 
-In Appendix \ref{sec:appendix proof and definition in section co-regular of PQ1Equal}, we construct a set $\textit{Auts}_{\textit{1-eq}}$ of witness automata ($\mathcal{A}_{\textit{l-eq}}^1$ is in it), and use $\textit{Auts}_{\textit{1-eq}}$ to prove that $\textit{PQ}_1^{=}$ is co-regular, as stated by the following lemma.
+In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}, we construct a set $\textit{Auts}_{\textit{1-eq}}$ of witness automata ($\mathcal{A}_{\textit{l-eq}}^1$ is in it), and use $\textit{Auts}_{\textit{1-eq}}$ to prove that $\textit{EPQ}_1^{=}$ is co-regular, as stated by the following lemma.
 
-\begin{restatable}{lemma}{PQOneEqualIsCoRegular}
-\label{lemma:Rpr2 is co-regular}
-$\textit{PQ}_1^{=}$ is co-regular.
+\begin{restatable}{lemma}{EPQOneEqualIsCoRegular}
+\label{lemma:EPQ1Equal is co-regular}
+$\textit{EPQ}_1^{=}$ is co-regular.
 \end{restatable}
+
+\subsection{Combine Step-by-Step linearizability and Co-Regular}
+\label{subsec:combine step-by-step linearizability and co-regular}
+
+We also prove that $\textit{EPQ}_2^{>}$, $\textit{EPQ}_2^{=}$ and $\textit{EPQ}_3$ are co-regular.
+The case of $\textit{EPQ}_2^{>}$ is quite simple, since whenever $\textit{last}(e) = \textit{EPQ}_2^{>}$, $e \sqsubseteq \textit{MS}(\textit{EPQ}_2^{>})$.
+For the case of $\textit{EPQ}_2^{=}$, we find a necessary and sufficient condition for violating $\textit{MS}(\textit{EPQ}_2^{>})$, and shows that such condition violates the assumptions that every single-priority execution is FIFO and can be safely ignored.
+For the case of $\textit{EPQ}_3$, it can be similarly proved as the rule for $\textit{deq}(\textit{empty})$ in \cite{Bouajjani:2015}.
+We leave the detailed proof of co-regular of these three rules in Appendix \ref{subsec:appendix co-regular of EPQ2Lar}, Appendix \ref{subsec:appendix co-regular of EPQ2Equal} and Appendix \ref{subsec:co-regular of EPQ3}, respectively.
+
+Now we can see that $\textit{EPQ}$ is co-regular, which is a direct consequence of co-regular of each of its rules, and is stated below.
+
+\begin{restatable}{lemma}{EPQIsCoRegular}
+\label{lemma:EPQ is co-regular}
+$\textit{EPQ}$ is co-regular.
+\end{restatable}
+
+Since we have already prove that $\textit{EPQ}$ is step-by-step linearizability and co-regular, we can now reduce the verification of linearizability with respect to $\textit{EPQ}$ into a reachability problem, as illustrated by the following theorem.
+Here $\textit{Auts}_{\textit{EPQ}}$ is the set of witness automata of this section.
+
+\begin{restatable}{theorem}{ReduceEPQIntoStateReachability}
+\label{lemma:reduce EPQ into state reachability}
+Given a data-independence implementation $\mathcal{I}$. $\mathcal{I} \sqsubseteq \textit{EPQ}$, if and only if, $\mathcal{I} \cap \textit{Auts}_{\textit{EPQ}} = \emptyset$.
+\end{restatable}

### 2017-04-04 data-independence-EPQ.tex
add tex

--- 2017-02-20 data-independence-PQ.tex
@@
-\section{Data-Independence of Priority Queue}
-\label{sec:data-independence of priority queue}
+\section{Data-Independence of Extended Priority Queue}
+\label{sec:data-independence of extended priority queue}
 
 Data-independence \cite{Wolper:1986} can be used to effectively handle unbounded data domain.
-In this section, we slightly modify the notion of data-independence in \cite{Wolper:1986} and propose data-differentiated sequences and data-independence for priority queues.
+In this section, we slightly modify the notion of data-independence in \cite{Wolper:1986} and propose data-differentiated sequences and data-independence for extended priority queues.
 
 Let $\_$ denote an element, of which the value is irrelevant.
-A sequential execution $e$ of priority is said to be data-differentiated if, for all $d \in \mathbb{D}$, there is at most one method event $\textit{put}(d,\_)$ in $e$.
+A sequential execution $e$ of extended priority queue is said to be data-differentiated if, for all $d \in \mathbb{D}$, there is at most one method event $\textit{put}(d,\_)$ in $e$.
 Note that a data-differentiated sequential execution $e$ may contains more than one items with a same priority.
@@
 
-A renaming function $r$ for priority queue is a function from $\mathbb{D}$ to $\mathbb{D}$.
+A renaming function $r$ for extended priority queue is a function from $\mathbb{D}$ to $\mathbb{D}$.
 Given a sequential execution (resp., execution or history) $e$, we denote by $r(e)$ the sequential execution (resp., execution or history) obtained from $e$ by replacing every item $x$ by $r(x)$.
 Note that here the renaming functions rename only the items and keep the priorities unchanged.
+The intuitive explanation is that renaming items will not influence the executions of program, while renaming priorities may influence the executions of program.
 
@@
 \begin{definition}\label{def:priority-value data-independence}
-The set of sequential executions (resp., executions or histories) $S$ is data-independent, if:
+A set of sequential executions (resp., executions or histories) $S$ is data-independent, if:
 \begin{itemize}
@@
 Thus, in the remainder of the paper, we focus on characterizing linearizability for data-differentiated executions, rather than arbitrary ones.
-The proof of this lemma can be found in Appendix \ref{sec:appendix in section data-independence of PQ}.
+The proof of this lemma can be found in Appendix \ref{sec:appendix in section data-independence of EPQ}.
 

### 2017-04-04 preliminaries-EPQ.tex
add tex

--- 2017-02-20 preliminaries.tex
@@
 \redt{We fix several (possibly infinite) set $\mathbb{D}_1,\ldots$ of data values}, and a finite set $\mathbb{M}$ of methods. %We consider that methods have exactly one argument, or one return value.
-\redt{Return values are transformed into argument values for uniformity \footnote{\redt{Method return values are guessed nondeterministically, and validated at return points.
+%\redt{Return values are transformed into argument values for uniformity \footnote{\redt{Method return values are guessed nondeterministically, and validated at return points.
 This can be handled using the assume statements of typical formal specification languages, which only admit executions satisfying a given predicate.

### 2017-04-04 relate-PQ-with-Queue-Stack-MultiSet-EPQ.tex
add tex

\section{Relate Other Data Structures with Extended Priority Queue}
\label{sec:relate other data structures with extended priority queue}

In this section, we shows how to reduce queue and stack executions into priority queue executions in polynomial time.

\subsection{Relate Queue with Extended Priority Queue}
\label{subsec:relate queue with extended priority queue}

Given a queue execution $e_q$, we can obtain an execution $\textit{QtoEPQ}(e_q)$ of extended priority queue by (1) transforming $\textit{enq}(a)$ and $\textit{deq}(a)$ into $\textit{put}(a,p)$ and $\textit{rm}(a)$ for each $a \in \mathbb{D}$, and (2) transforming $\textit{deq}(\textit{empty})$ into $\textit{rm}(\textit{empty})$. Note that in $\textit{QtoEPQ}(e_q)$, we assign all items with a same priority $p \in \mathbb{P}$. Since we have already guarantee that each single-priority execution of priority queue has FIFO property, it is easy to see that $e_q$ is linearizability w.r.t queue, if and only if $\textit{QtoEPQ}(e_q)$ is linearizable w.r.t $\textit{EPQ}$, as stated by the following lemma:

\begin{restatable}{lemma}{RelateQueuewithEPQ}
\label{lemma:relate queue with extended priority queue}
Given an execution $e_q$ of queue, $e_q$ is linearizable w.r.t queue, if and only if $\textit{QtoEPQ}(e_q) \sqsubseteq \textit{EPQ}$.
\end{restatable}

\subsection{Relate Multi-Set with Extended Priority Queue}
\label{subsec:relate multiSet with extended priority queue}

A multi-set contains two method: $\textit{insert}$ and $\textit{delete}$. A $\textit{insert}$ method has one argument, and is used to insert an item into the multi-set. We also assume that the item is chosen from a specific (possibly infinite) data domain $\mathbb{D}$. A $\textit{delete}$ method intends to remove an arbitrary item in multi-set and then returns it. If the multi-set is empty, $\textit{delete}$ returns $\textit{empty}$. Note that when $\textit{delete}$ returns an item, there is no restriction (such as FIFO, LIFO or priorities) for how to choose this item in multi-set.

Let $\textit{MSet}$ be the set of sequential executions of multi-set. In Appendix \ref{subsec:appendix proof and definition in section relate multiset with extended priority queue}, we give inductive rules of $\textit{MSet}$, and prove that these rules are step-by-step linearizability and co-regular. We also give witness automata for $\textit{MSet}$. Here we use the notion of inductive rules, step-by-step linearizability and co-regular in \cite{Bouajjani:2015}.

Since multi-set does not add any constraint for selecting item when deleting items, we simulate such behavior by giving each item an unique priority where such priorities are pair-wise incomparable. To make it simple, we just ignore items that are putted more than one times. Such ignorance is correct due to data-independence. Formally, given a multi-set execution $e_m$, we can obtain an execution $\textit{MStoEPQ}(e_m)$ of extended priority queue by (1) erasing operations of items which are putted more than once in $e_m$, (2) transforming $\textit{insert}(a)$ and $\textit{delete}(a)$ into $\textit{put}(a,p_a)$ and $\textit{rm}(a)$ for each $a \in \mathbb{D}$, and (3) transforming $\textit{delete}(\textit{empty})$ into $\textit{rm}(\textit{empty})$. Here we need to ensure that for each $a \neq b$, $p_a$ and $p_b$ must be incomparable. Given a set $S$ of executions of multi-set, let $\textit{MStoEPQ}(S)$ be the set generated by applying $\textit{MStoEPQ}$ to each sequence in $S$. Let $\textit{Auts}_{\textit{EPQ}}$ and $\textit{Auts}_{\textit{MS}}$ be the set of witness automata for extended priority queue and for multi-set, respectively. The following lemma stats that our witness automata for extended priority queue are enough for checking violations w.r.t $\textit{MSet}$.

\begin{restatable}{lemma}{RelateMultiSetwithEPQ}
\label{lemma:relate multi set with extended priority queue}
Given an data-independent implementation $\mathcal{I}_m$ of multi-set, $\mathcal{I}_m \cap \textit{Auts}_{\textit{MS}} \neq \emptyset$, if and only if $\textit{MStoEPQ}(\mathcal{I}_m) \cap \textit{Auts}_{\textit{EPQ}} \neq \emptyset$.
\end{restatable}

\subsection{Relate Stack with Extended Priority Queue}
\label{subsec:relate stack with extended priority queue}

A stack contains two method: $\textit{push}$ and $\textit{pop}$. A $\textit{push}$ method has one argument, and is used to insert an item into the stack. A $\textit{pop}$ method intends to remove an item that is newest in stack and then returns it. If the stack is empty, $\textit{pop}$ returns $\textit{empty}$.

Given an execution $e_s$ of stack, we can obtain an execution $\textit{StoEPQ}(e_s)$ of extended priority queue by transforming $\textit{push}(a)$ and $\textit{pop}(a)$ into $\textit{put}(a,p_a)$ and $\textit{rm}(a)$, and transforming $\textit{pop}(\textit{empty})$ into $\textit{rm}(\textit{empty})$. Here let us show the process of assigning priorities for items. Given $e_s$, the process of assign priorities is as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Let $e'$ be the projection of $e_s$ into $\textit{push}$. For each $i$, associate the $\textit{i-th}$ element of $e'$ with $e$.

\item[-] For each $\textit{push}_o(a)$ in $e$, we assign priority $(i,j)$ to it. Here $i$ and $j$ are the number associated with the call and return action of $\textit{push}_o(a)$. 
\end{itemize}

Then the order $<_{\mathbb{P}}$ is defined as follows: $(i_1,i_2) <_{\mathbb{P}} (j_1,j_2)$, if $j_1 > i_2$, or we can say, if the $\textit{put}$ of the former happens before the $\textit{put}$ of the latter. In \figurename~\ref{fig:execution with intervals}, we associate numbers with the call and return actions of $\textit{put}$, and then assign priorities to items. We can see that the priority of $c$ is smaller than the priority of $a$ according to $<_{\mathbb{P}}$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6 \textwidth]{figures/PIC-HIS-INTRO-Interval.pdf}
%\vspace{-10pt}
  \caption{Execution with intervals}
  \label{fig:execution with intervals}
\end{figure}

### 2017-04-04 rules-EPQ.tex
add tex

\section{Inductive Rules of Extended Priority Queue}
\label{sec:inductive rules of extended priority queue}

A priority queue contains two method: $\textit{put}$ and $\textit{rm}$. A $\textit{put}$ method has two arguments, while the first argument is an item and the second argument is its priority. A $\textit{put}$ method is used to put an item into the priority queue with certain priority. Here we assume that the item is chosen from a specific (possibly infinite) data domain $\mathbb{D}$ and priority is chosen from a (possibly infinite) set $\mathbb{P}$. {\color {red}Moreover, we assume that there is a strict partial-order $<_\mathbb{P}$ among elements in $\mathbb{P}$.} A $\textit{rm}$ method intends to remove the item with minimal priority (with respect to $<_\mathbb{P}$) in priority queue and then returns it. It works as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If the priority queue is empty, then $\textit{rm}$ returns $\textit{empty}$.

\item[-] {\color {red}Else, $\textit{rm}$ returns a oldest element of one of minimal priorities. Formally, there are a set $S$ of items in priority queue. $S$ can be divided into several group, such that (1) items in each group have same priority, (2) the priorities of each two groups is incomparable and (3) no priority of items in $S$ can be larger than items not in $S$. Each group is the set of items of some minimal priority. Then, $\textit{rm}$ returns an item of some group, and this item must be putted earliest in this group. However, the chosen of group is arbitrary.}
\end{itemize}

We say that $\textit{put}(a,p)$ matches $\textit{rm}(b)$, if $a = b$. Our priority queue is an extension of common priority queue, where the priority is chosen from the set $\mathbb{N}$ of natural numbers, or from a set with total order. To distinguish our priority queue with common priority queue, we explicitly call our priority queue the extended priority queue.

Similar as \cite{Bouajjani:2015}, we use inductive rules to define the set of sequential executions of extended priority queue. Each rule is of the form $l_1 \cdot \ldots \cdot l_k \in \textit{EPQ} \wedge \textit{Guard}(l_1,\ldots,l_k,\textit{itm},\textit{pri}) \Rightarrow \textit{Expr}(l_1,\ldots,l_k,\textit{itm},\textit{pri}) \in \textit{EPQ}$. Here $\textit{EPQ}$ is the set of sequential executions of extended priority queue. Here $\textit{itm}$ and $\textit{pri}$ are two variables, and represent an arbitrary item and priority, respectively. $\textit{Guard}(l_1,\ldots,l_k,\textit{itm},\textit{pri})$ is a conjunction of conditions with the following notations:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Given sequential execution $l$, $\textit{noRE}(l)$ is satisfied when each method event of $l$ is not $\textit{rm}(\textit{empty})$.

\item[-] Given sequential execution $l$, $\textit{LEI}(\textit{pri},l)$ is satisfied, if for priority of every item of $l$, $\textit{pri}$ is either larger, or equal, or incomparable with it. Here $L$ represents larger, $E$ represents equal, and $I$ represents incomparable. Similarly, we can define $\textit{LI}(\textit{pri},l)$. We use $U$ to represent items of unmatched $\textit{put}$, and $\textit{LI-U}(\textit{pri},l)$ is satisfied, if for priority of every item of unmatched $\textit{put}$ in $l$, $\textit{pri}$ is either larger, or incomparable with it. We use $M$ to represent items of matched $\textit{put}$, and define $\textit{LI-M}$ and $\textit{LEI-M}$ similarly.

    {\color {red} We use $E'$ to emphasize that equal must holds somewhere. For example, $\textit{LE'I}(\textit{pri},l)$ holds, if (1) for priority of every item of $l$, $\textit{pri}$ is either larger, or equal, or incomparable with it, and (2) $\textit{pri}$ indeed equals priority of some items of $l$. $\textit{LE'I-M}$ is similarly defined.}

\item[-] Given sequential execution $l$, its sub-sequence $l'$ and a priority $p$, $\textit{putInSeq}(l,l',p)$ is satisfied when all the $\textit{put}$ with priority $p$ of $l$ (if exists) is in $l'$.

\item[-] {\color {red}Given sequential execution $l$ and priority $p$, $\textit{matched-C}(l,p)$ is satisfied, if (1) for each item $a$ whose priority is comparable with $p$, if $\textit{put}(a,\_)$ is in $l$, then $\textit{rm}(a)$ is in $l$, and (2) for each item $a$ whose priority is comparable with $p$, if $\textit{rm}(a)$ is in $l$, then $\textit{put}(a,\_)$ is in $l$. Similarly, we can define $\textit{matched-All}(l)$, where all items in $l$, instead of items with priority comparable with some priority in $l$, is considered and matched.}
\end{itemize}

$\textit{Expr}(l_1,\ldots,l_k,\textit{itm},\textit{pri})$ is a expression $l'_1 \cdot \ldots \cdot l'_m$, where each $l'_i$ is chosen from (1) $l_j$ for some $j$, (2) method event with item $\textit{itm}$ and priority $\textit{pri}$ and (3) method event $\textit{rm}(\textit{empty})$.
%Given $l'_1,\ldots,l'_k$ and expression $e=\textit{Expr}(l_1,\ldots,l_k,\textit{itm},\textit{pri})$, we define $\llbracket e \rrbracket$ as the set of sequential executions which can be obtained from $e$ by replacing $l_i$ with $l'_i$ for each $i$, and replacing $\textit{itm}$ and $\textit{pri}$ with some values in $\mathbb{D}$ and $\mathbb{P}$, respectively.
Given a rule $R \equiv l_1 \cdot \ldots \cdot l_k \in \textit{EPQ} \wedge \textit{Guard}(l_1,\ldots,l_k,\textit{itm},\textit{pri}) \Rightarrow \textit{Expr}(l_1,\ldots,l_k,\textit{itm},\textit{pri}) \in \textit{EPQ}$ and a sequential execution $w$, if $w=l'_1 \cdot \ldots \cdot l'_k$, and $\textit{Guard}(l'_1,\ldots,l'_k,a,p)$ holds for some $a \in \mathbb{D}$ and $p \in \mathbb{P}$, then we use $w \xrightarrow{R} w'$ to denote that we can obtain $w'$ from $w$ according to rule $R$, where $w' = \textit{Expr}(l'_1,\ldots,l'_k,a,p)$. Let $\llbracket \textit{EPQ} \rrbracket$ be the set of sequential executions $w$ which can be derived from the empty word: $$\epsilon = w_0 \xrightarrow{R_1} w_1 \ldots \xrightarrow{R_k} w$$ where each $R_i$ is one rules of the extended priority queue. When clear from context, we abuse $\llbracket \textit{EPQ} \rrbracket$ by $\textit{EPQ}$.

\begin{definition}\label{def:inductive rules of priority queue}
$\textit{EPQ}$ is defined by the following rules:
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{EPQ}_0 \equiv \epsilon \in \textit{EPQ}$.

\item[-] $\textit{EPQ}_1 \equiv (u \cdot v \cdot w \in \textit{EPQ}) \wedge
(\textit{noRE}(u \cdot v \cdot w)) \wedge
(\textit{LEI}(\textit{pri}, u \cdot v \cdot w)) \wedge
(\textit{LI-U}(\textit{pri},u \cdot v \cdot w)) \wedge
(\textit{matched-C}(u \cdot v,\textit{pri}) ) \wedge
%(\textit{matched}(w) ) \wedge
(\textit{putInSeq}(u \cdot v \cdot w,u,\textit{pri}))
\Rightarrow
(u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w \in \textit{EPQ})$.

\item[-] $\textit{EPQ}_2 \equiv
(u \cdot v \in \textit{EPQ}) \wedge
(\textit{noRE}(u \cdot v)) \wedge
(\textit{LEI}(\textit{pri},u \cdot v)) \wedge
(\textit{putInSeq}(u \cdot v,u,\textit{pri}))
\Rightarrow
(u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{EPQ})$.

\item[-] $\textit{EPQ}_3 \equiv
(u \cdot v \in \textit{EPQ}) \wedge
(\textit{matched-All}(u) )
\Rightarrow
(u \cdot \textit{rm}(\textit{empty}) \cdot v \in \textit{EPQ})$.
\end{itemize}
\end{definition}

\begin{example}\label{example:generate extended priority queue executions}
Given priorities $p_1,p_2,p_3$ with orders $p_1 <_{\mathbb{P}} p_2$ and $p_1 <_{\mathbb{P}} p_3$, one sequential execution of $\textit{EPQ}$ is generated as follows:

$\epsilon$ $\xrightarrow{\textit{EPQ}_1}$ $\textit{put}(a,p_1) \cdot \textit{rm}(a)$

$\xrightarrow{\textit{EPQ}_2}$ $\textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{put}(b,p_1)$

$\xrightarrow{\textit{EPQ}_1}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(b,p_1)$

$\xrightarrow{\textit{EPQ}_2}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(b,p_1)$

$\xrightarrow{\textit{EPQ}_1}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)$

$\xrightarrow{\textit{EPQ}_3}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{rm}(\textit{empty}) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)$
\end{example}

To facilitate our proof of latter sections, we need to separate $\textit{EPQ}_1$ into two cases: (1) no matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v \cdot w$ has priority $\textit{pri}$, (2) some matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v \cdot w$ has priority $\textit{pri}$.  Therefore, we separate $\textit{EPQ}_1$ into two rules:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{EPQ}_1^{>} \equiv (u \cdot v \cdot w \in \textit{EPQ}) \wedge
(\textit{noRE}(u \cdot v \cdot w)) \wedge
(\textit{LI}(\textit{pri}, u \cdot v \cdot w)) \wedge
(\textit{LI-U}(\textit{pri},u \cdot v \cdot w)) \wedge
(\textit{matched-C}(u \cdot v,\textit{pri}) ) %\wedge
%(\textit{matched}(w) )
\Rightarrow
(u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w \in \textit{EPQ})$.

\item[-] $\textit{EPQ}_1^{=} \equiv (u \cdot v \cdot w \in \textit{EPQ}) \wedge
(\textit{noRE}(u \cdot v \cdot w)) \wedge
(\textit{LE'I}(\textit{pri}, u \cdot v \cdot w)) \wedge
(\textit{LI-U}(\textit{pri},u \cdot v \cdot w)) \wedge
(\textit{matched-C}(u \cdot v,\textit{pri}) ) %\wedge
%(\textit{matched}(w) )
\wedge
(\textit{putInSeq}(u \cdot v \cdot w,u,\textit{pri}))
\Rightarrow
(u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w \in \textit{EPQ})$.
\end{itemize}

For $\textit{EPQ}_2$, we also need to distinguish two cases: (1) no matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ has priority $\textit{pri}$, (2) some matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ has priority $\textit{pri}$. Therefore, we separate $\textit{EPQ}_2$ into two rules:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{EPQ}_2^{>} \equiv
(u \cdot v \in \textit{EPQ}) \wedge
(\textit{noRE}(u \cdot v)) \wedge
(\textit{LEI}(\textit{pri},u \cdot v)) \wedge
(\textit{LI-M}(\textit{pri},u \cdot v)) \wedge
(\textit{putInSeq}(u \cdot v,u,\textit{pri}))
\Rightarrow
(u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{EPQ})$.

\item[-] $\textit{EPQ}_2^{=} \equiv
(u \cdot v \in \textit{EPQ}) \wedge
(\textit{noRE}(u \cdot v)) \wedge
(\textit{LEI}(\textit{pri},u \cdot v)) \wedge
(\textit{LE'I-M}(\textit{pri},u \cdot v)) \wedge
(\textit{putInSeq}(u \cdot v,u,\textit{pri}))
\Rightarrow
(u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{EPQ})$.
\end{itemize}

To persuade readers that our rules is indeed the rules of extended priority queue, in Appendix \ref{sec:appendix in section inductive rules of extended priority queue}, we give a semantical version definition $\textit{EPQ}_s$ of extended priority queue, and shows that the language generated by our rules equals the set of traces of $\textit{EPQ}_s$. To model the possible behaviors of extended priority queue, we model it as an labelled transition system (shortened as LTS) $\textit{LTS}_e$. Each state of $\textit{LTS}_e$ is a function from $\mathbb{P}$ into sequences over $\mathbb{D}$, and represents a snapshot of contents of extended priority queue. $\textit{EPQ}_s$ is the set of traces of $\textit{LTS}_e$. The definition of $\textit{EPQ}_s$ and the proof of the following lemma can be found in Appendix \ref{sec:appendix in section inductive rules of extended priority queue}.

\begin{restatable}{lemma}{EPQRulesAndSemantics}
\label{lemma:EPQ rules and semantics}
$\textit{EPQ} = \textit{EPQ}_s$.
\end{restatable}

Given a sequential execution $w \in \textit{EPQ}$ and let $P$ be the set of priorities in $w$. According to above example, $w$ is generated from $\epsilon$ as follows: Let $P$ be the set of priorities of $w$ and $p$ be one of minimal priority in $P$. Then we start to loop. In each round of the loop there are two possibilities: (1) If there are matched $\textit{put}$ and $\textit{rm}$ with priority $p$ in $w$: Add pairs of matched $\textit{put}$ and $\textit{rm}$ with priority $p$ by using one time of $\textit{EPQ}_1^{>}$ and then (possibly) several times of $\textit{EPQ}_1^{=}$, and then (possibly) add unmatched $\textit{put}$ with priority $p$ by using $\textit{EPQ}_2^{=}$, (2) If there are no matched $\textit{put}$ and $\textit{rm}$ with priority $p$ in $w$: Add unmatched $\textit{put}$ with priority $p$ by using $\textit{EPQ}_2^{>}$ for several times. Then, let $P = P \setminus \{ p \}$, and restart this loop, until $P = \emptyset$. Finally, add $\textit{rm}(\textit{empty})$ by using $\textit{EPQ}_3$ for several times. We can see that, the order for generating $w$ may be not fixed, since some priorities are incomparable.

Thus, given $w$, we define $\textit{last}(w)$ as the set of last possible rule to generate $w$ according to the rules of extended priority queues:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $w$ contains $\textit{rm}(\textit{empty})$, then $\textit{last}(w) = \{ \textit{EPQ}_3 \}$.

\item[-] Else, if items of several unmatched $\textit{put}$ and matched $\textit{put}$ have a maximal priority of $w$, then $\textit{last}(w)$ contains $\textit{EPQ}_2^{=}$.

\item[-] Else, if items of only several unmatched $\textit{put}$ have a maximal priority of $w$, then $\textit{last}(w)$ contains $\textit{EPQ}_2^{>}$.

\item[-] Else, if items of only more than one matched $\textit{put}$ have a maximal priority of $w$, then $\textit{last}(w)$ contains $\textit{EPQ}_1^{=}$.

\item[-] Else, if items of only one matched $\textit{put}$ has a maximal priority of $w$, then $\textit{last}(w)$ contains $\textit{EPQ}_1^{>}$.

\item[-] Else ($w = \epsilon$), $\textit{last}(w) = \{ \textit{EPQ}_0 \}$.
\end{itemize}

Note that $\textit{last}(w)$ may contains more than one rules. For example, given $w = \textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)$ and orders $p_1 <_{\mathbb{P}} p_2$ and $p_1 <_{\mathbb{P}} p_3$, then $\textit{last}(w) = \{  \textit{EPQ}_2^{=}, \textit{EPQ}_1^{>} \}$. The notion of $\textit{last}$ can be extended into execution and histories: given a history $h$, $\textit{last}(h) = \textit{last}(u)$, where $u$ is any sequential execution such that $h \sqsubseteq u$. When $\textit{last}(e)$ contains only one rule $R$, we write $\textit{last}(e)=R$ for simplicity.

### 2017-04-04 ss-lin-EPQ.tex
add two tex

\section{Step-by-Step Linearizability of Extended Priority Queues}
\label{sec:step-by-step linearizability of extended priority queues}

In this section we shows that, with the help of a property called step-by-step linearizability, we can partition the concurrent executions which are not linearizable with respect to $\textit{EPQ}$ into a finite number of classes. Intuitively, each such class represents a set of sequential execution that violate one rule of extended priority queue. Here step-by-step linearizability enables us to build a linearization for an execution $e$ incrementally, using linearizations of projections of $e$. Our step-by-step linearizability is inspired by the step-by-step linearizabilility of queue and stacks in \cite{Bouajjani:2015}. The proof of lemmas in this section can be found in Appendix \ref{sec:appendix in section step-by-step linearizability of extended priority queues}.

The projection $e \vert{\mathcal{D}}$ of a sequential execution $e$ into a subset $\mathcal{D} \subseteq \mathbb{D}$ of items is obtained from $e$ by erasing all method events with a data value not in $\mathcal{D}$. The set of projections of $e$ is denoted $\textit{proj}(e)$. When refer to $\textit{proj}(e)$, we implicitly assume that each $\textit{rm}(\textit{empty})$ in $e$ has a ghost argument that is unique. We write $e \setminus x$ for the projection $e \vert_{ \mathbb{D} \setminus \{ x \} }$. This extends naturally to histories and concurrent executions.

A set $S$ of sequential executions is closed under projection, if for all $\mathcal{D} \subseteq \mathbb{D}$ and $e \in S$, we have $e \vert_{ \mathcal{D} } \in S$. The following lemma states that $\textit{EPQ}$ is closed under projection, since the predicates used in rules of extended priority queue are ``closed under projection''.

\begin{restatable}{lemma}{EPQisClosedUnderProjection}
\label{lemma:EPQ is closed under projection}
$\textit{EPQ}$ is closed under projection.
\end{restatable}

A sequential execution $e$ matches a rule $R$ of extended priority queue, if $e=\textit{Expr}(l_1, \ldots, l_k,a,b)$, and $\textit{Guard}(l_1, \ldots, l_k,a,b)$ holds. Here $\textit{Guard}$ and $\textit{Expr}$ are of rule $R$, and we call $a$ (if exists) the witness of $e$. We denote by $\textit{MS}(R)$ the set of sequential executions which match $R$. Note that sequences in $\textit{MS}(R)$ only respect rule $R$ and may be not in $\textit{EPQ}$. $e$ is linearizable with respect to $\textit{MS}(R)$ with witness $x$, if $e$ is linearizable with respect to $u \in \textit{MS}(R)$ and $x$ is the witness of $u$.

\begin{example}\label{example:match set of R}
Assume that $p_1 <_{\mathbb{P}} p_2$, we can see that $e = \textit{rm}(b) \cdot \textit{put}(b,p_1) \cdot \textit{put}(a,p_2) \cdot \textit{rm}(a)$ is in $\textit{MS}(\textit{EPQ}_1^{>} )$, but it is obvious that $e \notin \textit{EPQ}$.
\end{example}

The following lemma states that for data-differentiated sequential execution, checking inclusion into $\textit{EPQ}$ is equivalent to checking inclusion into $\textit{MS}(R)$ for everyone of its projections.

\begin{restatable}{lemma}{EPQasMultiInMRpriforSequence}
\label{lemma:EPQ as multi in MRpri for sequence}
Given a data-differentiated sequential execution $e$, $e \in \textit{EPQ}$, if and only if, $\forall e' \in \textit{proj}(e)$ and $\forall R \in \textit{last}(e')$, we have $e' \in \textit{MS}(R)$.
\end{restatable}

Lemma \ref{lemma:EPQ as multi in MRpri for sequence} simplifies checking inclusion into $\textit{EPQ}$, since checking $\textit{MS}(R)$ only concerns information of one rule, while checking $\textit{EPQ}$ need to consider every method events. We want a similar lemma for checking linearizability with respect to $\textit{EPQ}$. To enable such equivalent characterization, we introduce the notion of step-by-step linearizability for extended priority queue. The projection $e \vert{O}$ of a concurrent execution $e$ into a set $O$ of operations is obtained from $e$ by erasing all call and return actions of non-$O$ operations. We write $e \setminus o$ for the projection $e \vert_{ O_h \setminus \{ o \} }$, where $O_h$ is the set of operations of $e$. This extends naturally to histories. Similarly we can define projection into method events.

\begin{definition}\label{def:step-by-step linearizability of extended priority queue}
A set $S$ of sequential executions of extended priority queue is step-by-step linearizable, if for any data-differentiated execution $e$,
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] if $e$ is linearizable w.r.t. $\textit{MS}(R)$ ($R \in \{ \textit{EPQ}_1^{>}, \textit{EPQ}_1^{=}, \textit{EPQ}_2^{>},\textit{EPQ}_2^{=} \}$) with witness $x$, we have: $e \setminus x \sqsubseteq \textit{EPQ} \Rightarrow e \sqsubseteq \textit{EPQ}$.

\item[-] if $e$ is linearizable w.r.t. $\textit{MS}(\textit{EPQ}_3)$ and $o$ is a $\textit{rm}(\textit{empty})$ event, we have:

$e \setminus o \sqsubseteq \textit{EPQ} \Rightarrow e \sqsubseteq \textit{EPQ}$.
\end{itemize}
\end{definition}

Given a data-differentiated execution and its history, we can abuse notation and mix labels and method events with operations themselves, since items are unique in a data-differentiated execution. For instance, we will reference an operation labeled by $\textit{put}(p,a)$ as $\textit{put}(p,a)$. The following lemma states that $\textit{EPQ}$ is step-by-step linearizability.

\begin{restatable}{lemma}{EPQueueisStepByStepLinearizability}
\label{lemma:EPQ is step-by-step linearizability}
$\textit{EPQ}$ is step-by-step linearizability.
\end{restatable}

The following lemma states that for data-differentiated execution, checking linearizability with respect to $\textit{EPQ}$ is equivalent to checking linearizability with respect to $\textit{MS}(R)$ for everyone of its projections. Roughly speaking, step-by-step linearizability of extended priority queue play a important rule for proof of the $\textit{if}$ direction of Lemma \ref{lemma:EPQ as multi in MRpri for history}: It guide us how to build a linearization of a whole execution by increasingly construct linearization of sub-execution from $\epsilon$ execution.

\begin{restatable}{lemma}{EPQasMultiInMRpriforHistory}
\label{lemma:EPQ as multi in MRpri for history}
Given a data-differentiated execution $e$, $e \sqsubseteq \textit{EPQ}$, if and only if, $\forall e' \in \textit{proj}(e)$ and $\forall R \in \textit{last}(e')$, we have $e' \sqsubseteq \textit{MS}(R)$.
\end{restatable}

### 2017-04-14 appendix-EPQ.tex
I change proof of EPQ1 and add explanation to it

@@
 
+Before we prove the step-by-step linearizability of $\textit{MS}(\textit{EPQ}_1)$ and $\textit{MS}(\textit{EPQ}_2)$, we introduce several lemmas, which are used to ensure some sub-sequences of $\textit{EPQ}$ still belongs to $\textit{EPQ}$.
+Given a data-differentiated sequence $l$ and one of its maximal priority $\textit{pri}$, let $O_c(l,\textit{pri})$ and $O_i(l,\textit{pri})$ be the set of operations with priorities comparable with $\textit{pri}$ and incomparable with $\textit{pri}$ in $l$, respectively.
+Similarly we can define $D_c(l,\textit{pri})$ and $D_i(l,\textit{pri})$ for items instead of operations.
+%Let $O_x(l)$ be the set of $\textit{put}(y,\textit{pri})$ with $y \neq x$ in $l$, where $\textit{pri}$ is the priority of $x$ in $l$.
+We can see that each priority of items in $O_i(l,\textit{pri})$ is either larger or incomparable with priorities of items in  $O_c(l,\textit{pri})$.
+
+The following lemma shows that if a new sequence is generated by erasing some operations in $O_c(l,\textit{pri})$ while keeping the remaining $O_c(l,\textit{pri})$ sub-sequences in $\textit{EPQ}$, then this new sequence is still in $\textit{EPQ}$.
+Note that this is different from projection on items.
+
+\begin{restatable}{lemma}{EraseOcStillinEPQ}
+\label{lemma:erase Oc still in EPQ}
+Given a data-differentiated sequential execution $l \in \textit{EPQ}$ and a maximal priority $\textit{pri}$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$.
+Let $l'$ be generated from $l$ by discarding some operations in $O_c(l,\textit{pri})$, and $l' \vert_{ O_c(l,\textit{pri}) } \in \textit{EPQ}$.
+Then, $l' \in \textit{EPQ}$.
+\end{restatable}
+
+\begin {proof}
+Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{LTS}_e$.
+Assume that $l'$ is generated from $l$ by discarding $o_{\textit{ind1}},\ldots,o_{\textit{indn}}$.
+Let $D$ be the set such that $D$ contains $a$, if $\textit{put}(a,\_)$ is in $o_{\textit{ind1}},\ldots,o_{\textit{indn}}$.
+For each $i$, let $q'_i$ be generated from $q_i$ by erasing items in $D$.
+
+For each $q'_j$ with $j \neq \textit{ind}\_\textit{-1}$, if $o_{\textit{j+1}}$ is $\textit{put}$, then it is obvious that $q'_j \xrightarrow{o_{\textit{j+1}}} q'_{\textit{j+1}}$.
+Else, assume $o_{\textit{j+1}} = \textit{rm}(a)$,
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If $\textit{rm}(a) \in O_c(l,\textit{pri})$: By assumption, $l' \vert_{ O_c(l,\textit{pri}) } \in \textit{EPQ}$.
+Therefore, $a$ is in $q'_j$ and is the should-be-removed item in $O_c(l,\textit{pri})$.
+Since each priority of items in $O_i(l,\textit{pri})$ is either larger or incomparable with priorities of items in $O_c(l,\textit{pri})$, we can removed $a_c$ from $q'_j$, and then $q'_j \xrightarrow{o_{\textit{j+1}}} q'_{\textit{j+1}}$.
+
+\item[-] If $\textit{rm}(a) \in O_i(l,\textit{pri})$: By assumption we know that $q_j \xrightarrow{\textit{rm}(a)} q_{\textit{j+1}}$.
+Since $q'_j$ contains the same $D_i(l,\textit{pri})$ items as $q_j$ and $q'_j$ contains less $D_c(l,\textit{pri})$ items than $q_j$, we can see that $q'_j \xrightarrow{o_{\textit{j+1}}} q'_{\textit{j+1}}$.
+\end{itemize}
+
+For each $q'_{\textit{indj-1}}$, it is easy to see that $q'_{\textit{indj-1}} = q'_{\textit{indj}}$.
+Therefore, we can see that $l' = o_1 \cdot \ldots o_{\textit{ind1-1}} \cdot o_{\textit{ind1+1}} \cdot \ldots \in \textit{EPQ}$.
+\qed
+\end {proof}
+
+The following lemma shows that if a new sequence is generated by from some time point, erasing operations in $O_i(l,\textit{pri})$, then this new sequence is still in $\textit{EPQ}$.
+
+\begin{restatable}{lemma}{EraseOiFromSomeTimePointStillinEPQ}
+\label{lemma:erase Oi from some time point still in EPQ}
+Given a data-differentiated sequential execution $l \in \textit{EPQ}$ and a maximal priority $\textit{pri}$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$.
+Let $l'$ be generated from $l$ by discarding operations in $O_i(l,\textit{pri})$ from some time point, then, $l' \in \textit{EPQ}$.
+\end{restatable}
+
+\begin {proof}
+Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{LTS}_e$.
+Assume that $l'$ is generated from $l$ by discarding all operations $o_i$ if (1) $o_i \in O_i(l,\textit{pri})$ and (2) $i \geq k$ for a specific index $k$.
+Let $D$ be a set such that $a \in D$, if $\textit{put}(a,\_)$ is in $l$ and not in $l'$.
+For each $0 \leq i \leq m$, let $q'_i$ be generated from $q_i$ by erasing items in $D$.
+
+Let $l'=o'_1 \cdot \ldots \cdot o'_n$, and let $f$ be a function, such that $f(i)=j$, if $o'_i = o_j$.
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] We can see that $f$ maps each $0 \leq i \leq \textit{k-1}$ into $i$, and $q'_0 \xrightarrow{o_1} q'_1 \ldots \xrightarrow{o_{\textit{k-1}}} q'_{\textit{k-1}}$.
+
+\item[-] It is easy to see that $q'_{\textit{k-1}} = q'_{f(k)-1}$, and for each $i>k$, $q'_{f(k)} = q'_{f(k+1)-1}$.
+
+\item[-] If $o_{f(k)}$ is a $\textit{put}$ operation, then it is obvious that $q'_{f(k)-1} \xrightarrow{o_k} q'_{f(k)}$.
+Else, if $o_{f(k)} = \textit{rm}(a)$, we can see $a$ is in $q'_{f(k)-1}$, and since (1) $q'_{f(k)-1}$ contains the same $D_c(l,\textit{pri})$ items as $q_{f(k)-1}$ and $q'_{f(k)-1}$ contains less $D_i(l,\textit{pri})$ items than $q_{f(k)-1}$, and (2) each priority of items in $O_i(l,\textit{pri})$ is either larger or incomparable with priorities of items in  $O_c(l,\textit{pri})$, we can see that $q'_{f(k)-1} \xrightarrow{o_k} q'_{f(k)}$.
+Similarly we can prove the case of $o_{f(j)}$ with $j > k$.
+\end{itemize}
+
+This completes the proof of this lemma. \qed
+\end {proof}
+
+The following lemma shows that we can make $\textit{put}$ with maximal priority to happen earlier.
+
+\begin{restatable}{lemma}{MakePutWithMaxPriorityHappenEarlier}
+\label{lemma:make put with maximal priority happen earlier}
+Given a data-differentiated sequential execution $l \in \textit{EPQ}$ and a maximal priority $\textit{pri}$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$.
+Let $l=l_1 \cdot l_2$.
+Let $l_3$ be the projection of $l_2$ into $\{ \textit{put}(\_,\textit{pri}) \}$, and $l_4$ be the projection of $l_2$ into other operations.
+Then, $l'=l_1 \cdot l_3 \cdot l_4 \in \textit{EPQ}$.
+\end{restatable}
+
+\begin {proof}
+Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{LTS}_e$.
+Let $l_1 = o_1 \cdot \ldots \cdot o_n$, let $l_2 = o_{\textit{n+1}} \cdot \ldots \cdot o_m$, let $l_3 = o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$, let $l_4 = o'_{\textit{k+1}} \cdot \ldots \cdot o'_m$.
+Let $f$ be a function, such that $f(i)=j$, if $o'_i = o_j$.
+
+Let $q'_i$ be constructed as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] For $0 \leq i \leq n$, let $q'_i = q_i$.
+
+\item[-] For $\textit{n+1} \leq i \leq k$, let $q'_i$ be obtained from $q_n$ by adding items in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ with priority $\textit{pri}$ and in the same order.
+
+\item[-] For $\textit{k+1} \leq i \leq m$, let $q'_i$ be obtained from $q_{f(i)}$ by adding items which are (1) with priority $\textit{pri}$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(i)}$.
+The order of adding them is the same as $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$.
+\end{itemize}
+
+Then, our proof proceeds as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] It is obvious that $q'_0 \xrightarrow{o_1} q'_1 \ldots \xrightarrow{o_n} q'_n$ and $q'_n \xrightarrow{o_{\textit{n+1}}} q'_{\textit{n+1}} \ldots \xrightarrow{o_k} q'_k$.
+
+\item[-] For $q'_{\textit{k+1}}$: We already know that $q_{f(k+1)-1} \xrightarrow{o_{f(k+1)}} q_{f(k+1)}$, and it is easy to see that $q_{f(k+1)-1}$ is obtained from $q_n$ by adding items in $o_{\textit{n+1}} \cdot \ldots \cdot o_{f(k+1)-1}$.
+
+    We can see that $q'_k$ is obtained from $q_n$ by adding items in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$, and $q'_{k+1}$ is obtained from $q_{f(k+1)}$ by adding items which are (1) with priority $\textit{pri}$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(k+1)}$.
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $o_{f(k+1)}$ is an operation of non-$\textit{pri}$ items, then $q'_{k+1}$ is obtained from $q_{f(k+1)}$ by adding items in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$.
+Since non-$\textit{pri}$ priority is either smaller or incomparable with $\textit{pri}$, we can see that $q'_k \xrightarrow{o_{f(k+1)}} q'_{\textit{k+1}}$.
+
+    \item[-] Otherwise, it is only possible that $o_{f(k+1)} = \textit{rm}(a)$ for some item $a$ with priority $\textit{pri}$.
+We can see that $q'_{k+1}$ is obtained from $q_{f(k+1)}$ by adding items in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$ and then remove $a$.
+Since $q_{f(k+1)-1} \xrightarrow{o_{f(k+1)}} q_{f(k+1)}$, in $q_{f(k+1)-1}$( and also in $q'_k$), there is no item with priority less than $\textit{pri}$, and $a$ is the first-input item of priority $\textit{pri}$.
+Therefore, we can see that $q'_k \xrightarrow{o_{f(k+1)}} q'_{\textit{k+1}}$.
+    \end{itemize}
+
+\item[-] For $q'_{\textit{k+i}}$ with $i>1$: We already know that $q_{f(k+i)-1} \xrightarrow{o_{f(k+i)}} q_{f(k+i)}$, and $q_{f(k+i)-1}$ is obtained from $q_{f(k+i-1)}$ by adding items in $o_{f(k+i-1)+1} \cdot \ldots \cdot o_{f(k+i)-1}$.
+
+    We can see that $q'_{\textit{k+i}-1}$ (resp., $q'_{\textit{k+i}}$) is obtained from $q_{f(k+i-1)}$ (resp., $q_{f(k+i)}$) by adding items which are (1) with priority $\textit{pri}$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(k+i-1)}$ (resp., $o_1 \cdot \ldots \cdot o_{f(k+i)}$).
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $o_{f(k+i)}$ is an operation of non-$\textit{pri}$ items, then $q'_{k+i}$ is obtained from $q_{f(k+i)}$ by adding items which are (1) with priority $\textit{pri}$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(k+i-1)}$.
+Since non-$\textit{pri}$ priority is either smaller or incomparable with $\textit{pri}$, we can see that $q'_{\textit{k+i-1}} \xrightarrow{o_{f(k+i)}} q'_{\textit{k+i}}$.
+
+    \item[-] Otherwise, it is only possible that $o_{f(k+i)} = \textit{rm}(a)$ for some item $a$ with priority $\textit{pri}$.
+We can see that $q'_{k+i}$ is obtained from $q_{f(k+i)}$ by adding items in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$ and then remove $a$.
+Since $q_{f(k+i)-1} \xrightarrow{o_{f(k+i)}} q_{f(k+i)}$, in $q_{f(k+i)-1}$( and also in $q'_{\textit{k+i-1}}$), there is no item with priority less than $\textit{pri}$, and $a$ is the first-input item of priority $\textit{pri}$.
+Therefore, we can see that $q'_{\textit{k+i-1}} \xrightarrow{o_{f(k+i)}} q'_{\textit{k+i}}$.
+    \end{itemize}
+\end{itemize}
+
+This completes the proof of this lemma. \qed
+\end {proof}
+
+The following lemma shows that if a new sequence is generated by make some $O_i(l,\textit{pri})$ behaviors to happen earlier, then this new sequence is still in $\textit{EPQ}$.
+
+\begin{restatable}{lemma}{MakeOiHappenEarlierStillinEPQ}
+\label{lemma:make Oi happen earlier still in EPQ}
+Given a data-differentiated sequential execution $l \in \textit{EPQ}$ and a maximal priority $\textit{pri}$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$.
+Let $l \vert_{ O_i(l,\textit{pri}) } = l_1 \cdot l_2$, let $l' = l_1 \cdot l_3$, where $l_3$ is the projection of $l$ into non-$l_1$ operations.
+Then, $l' \in \textit{EPQ}$.
+\end{restatable}
+
+\begin {proof}
+Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{LTS}_e$.
+Let $l_1 = o'_1 \cdot \ldots \cdot o'_n$, let $l_3 = o'_{\textit{n+1}} \cdot \ldots \cdot o'_m$.
+Let $f$ be a function, such that $f(i)=j$, if $o'_i = o_j$.
+Let $D$ be the se of items which are added and not removed in $o'_1 \cdot \ldots \cdot o'_n$.
+
+Let $q'_i$ be constructed as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] It is easy to see that $l_1 \in \textit{EPQ}$, and let $q'_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n$ be the path of $l_1$ in $\textit{LTS}_e$.
+
+\item[-] For $\textit{n+1} \leq i \leq m$, let $q'_i$ be obtained from $q_{f(i)}$ by adding items in $D$. The order of adding them is the same as $o'_1 \cdot \ldots \cdot o'_n$.
+\end{itemize}
+
+Then, our proof proceeds as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] We already know that $q'_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n$.
+
+\item[-] For $q'_{\textit{n+i}}$: We already know that $q_{f(n+i)-1} \xrightarrow{o_{f(n+i)}} q_{f(n+i)}$, and it is easy to see that $q_{f(n+i)-1}$ is obtained from $q_{f(n+i-1)}$ by adding $D$-items in $o_{f(n+i-1)+1} \cdot \ldots \cdot q_{f(n+i)-1}$.
+
+     We can see that $q'_{\textit{n+i}-1}$ (resp., $q'_{\textit{n+i}}$) is obtained from $q_{f(n+i-1)}$ (resp., $q_{f(n+i)}$) by adding remanning items in $D$.
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $o_{f(n+1)}$ is an operation of $O_c(l,\textit{pri})$ items, since priority in $O_i(l,\textit{pri})$ is either larger or incomparable with priority in $O_c(l,\textit{pri})$, we can see that $q'_n \xrightarrow{o_{f(n+1)}} q'_{\textit{n+1}}$.
+
+    \item[-] Otherwise, it is only possible that $o_{f(k+1)} = \textit{rm}(a)$ for some item $a$ in $O_c(l,\textit{pri})$.
+Since $q_{f(n+i)-1} \xrightarrow{o_{f(n+i)}} q_{f(n+i)}$, in $q_{f(n+i)-1}$( and also in $q'_{\textit{n+i-1}}$), there is no item with priority less than $\textit{pri}$, and $a$ is the first-input item of priority $\textit{pri}$.
+Therefore, we can see that $q'_{\textit{n+i-1}} \xrightarrow{o_{f(n+i)}} q'_{\textit{n+i}}$.
+    \end{itemize}
+\end{itemize}
+
+This completes the proof of this lemma. \qed
+\end {proof}
+
+The following lemma shows that if a new sequence is generated by replacing a prefix with another one which make the priority queue has same content, then this new sequence is still in $\textit{EPQ}$.
+
+\begin{restatable}{lemma}{ReplaceEquivalentPrefixStillinEPQ}
+\label{lemma:replace equivalent prefix still in EPQ}
+Given a data-differentiated sequential execution $l \in \textit{EPQ}$.
+Let $l=l_1 \cdot l_2$.
+Given $l_3 \in \textit{EPQ}$.
+Assume that the priority queue has same content after executing $l_1$ and $l_3$.
+Let $l' = l_3 \cdot l_2$.
+Then, $l' \in \textit{EPQ}$.
+\end{restatable}
+
+\begin {proof}
+Let $l=o_1 \cdot \ldots \cdot o_m$ and let $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{LTS}_e$.
+Let $l_1 = o_1 \cdot \ldots \cdot l_k$, $l_3 = o'_1 \cdot \ldots \cdot o'_n$ and let $q_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n$ be the path of $l_3$ in $\textit{LTS}_e$.
+
+By assumption we know that $q_k = q'_n$.
+Then it is not hard to see that $q_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n \xrightarrow{o_{\textit{k+1}}} q_{\textit{k+1}} \ldots \xrightarrow{o_m} q_m$ is a path in $\textit{LTS}_e$, and then $l' \in \textit{EPQ}_s$.
+By Lemma \ref{lemma:EPQ rules and semantics}, we know that $l' \in \textit{EPQ}$.
+\qed
+\end {proof}
+
+With the help of Lemma \ref{lemma:erase Oc still in EPQ}, Lemma \ref{lemma:erase Oi from some time point still in EPQ}, Lemma \ref{lemma:make put with maximal priority happen earlier}, Lemma \ref{lemma:make Oi happen earlier still in EPQ} and Lemma \ref{lemma:replace equivalent prefix still in EPQ}, we can now prove that $\textit{EPQ}_1$ is step-by-step linearizability.
+
 \begin{restatable}{lemma}{EPQ1isStepByStepLinearizability}
@@
 Let $U$, $V$ and $W$ be the set of operations of $u$, $v$ and $w$, respectively.
-Let us change $U$, $V$ and $W$ by loops as follows: In the first loop, we start from the first $W$-element of $l'_v$, and let it be $o_h$,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Case $1$: If in $e_{\textit{lp}}$, the linearization point of $o_h$ is before $\textit{ret}(\textit{rm},x)$, and no element in $W$ happens before $o_h$.
-Then, we erase $o_h$ from $W$ and put it into $U \cdot V$.
-
-\item[-] Case $2$: Else, if in $e_{\textit{lp}}$, the linearization point of $o_h$ is before $\textit{ret}(\textit{rm},x)$, and there exists $o_w \in W$, such that $o_w <_{\textit{hb}} o_h$.
-Then, in $l'_v$, we erase all such operations from $U \cdot V$ and put them into $W$, and then stop the process of changing $U$, $V$ and $W$.
-These operations should satisfy (1) their linearization point is after $o_h$ in $e_{\textit{lp}}$, and (2) the priority of their item is incomparable with $\textit{pri}_x$.
+Let us change $O_i(l,\textit{pri}_x)$ elements in $U$, $V$ and $W$, while keep $O_c(l,\textit{pri}_x)$ elements unchanged.
+We proceed by several loops: In the first loop, we start from the first $O_i(l,\textit{pri}_x)$-element of $l'_v$, and let it be $o_h$,
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Case $1$: If in $e_{\textit{lp}}$, the linearization point of $o_h$ is before $\textit{ret}(\textit{rm},x)$, and no $O_c(l,\textit{pri}_x)$-element in $W$ happens before $o_h$.
+Then, $o_h$ is in the new version of $U \cup V$.
+
+\item[-] Case $2$: Else, if in $e_{\textit{lp}}$, the linearization point of $o_h$ is before $\textit{ret}(\textit{rm},x)$, and there exists $O_c(l,\textit{pri}_x)$-element $o_w$ in $W$, such that $o_w <_{\textit{hb}} o_h$.
+Then, in $l'_v$, we put $o_h$ and all $O_i(l,\textit{pri}_x)$-element whose linearization points is after the linearization point of $o_h$ into new version of $W$, and then stop the process of changing $U$, $V$ and $W$.
 
 \item[-] Case $3$: Else, if in $e_{\textit{lp}}$, the linearization point of $o_h$ is after $\textit{ret}(\textit{rm},x)$.
-Then, in $l'_v$, we erase all such operations from $U \cdot V$ and put them into $W$, and then stop the process of changing $U$, $V$ and $W$.
-These operations should satisfy (1) their linearization point is after $o_h$ in $e_{\textit{lp}}$, and (2) the priority of their item is incomparable with $\textit{pri}_x$.
-\end{itemize}
-
-Our process proceed, until either all element in $l'_v$ are in new $U$ or new $V$, or case $2$ or case $3$ holds and this process terminates.
-Let $U'$, $V'$ and $W'$ be the new version of $U$, $V$ and $W$ after the process terminates, respectively.
+Then, in $l'_v$, we put $o_h$ and all $O_i(l,\textit{pri}_x)$-element whose linearization points is after the linearization point of $o_h$ into new version of $W$, and then stop the process of changing $U$, $V$ and $W$.
+\end{itemize}
+
+In the next loop, we consider the second $O_i(l,\textit{pri}_x)$-element of $l'_v$, and so on.
+Our process proceed, until either all element in $l'_v$ are in new version of $U \cup V$, or case $2$ or case $3$ happens and this process terminates.
+Let $U' \cup V'$ and $W'$ be the new version of $U \cup V$ and $W$ after the process terminates, respectively.
 Let $O_+$ be the set of operations that are moved into $U' \cup V'$ in the process, and let $O_-$ be the set of operations that are moved into $W'$ in the process.
@@
 Let $l''_2$ be the projection of $l'$ into operations of $l'_{\textit{u'v'}}$ that are not in $l''_1$.
-Let $l''_3$ be the projection of $l'$ into operations of $l'_{\textit{u'v'}}$ that are not in $l''_1$ and $l''_2$.
+Let $l''_3$ be the projection of $l'$ into operations which are not in $l'_{\textit{u'v'}}$.
 Let $l'' = l''_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3$.
@@
     \setlength{\itemsep}{0.5pt}
-    \item[-] If $o_1 \in O_a$: This is impossible, since the linearization point of operations in $O_a$ is before $\textit{ret}(\textit{put},b)$ in $e_{\textit{lp}}$.
-
-    \item[-] If $o_1 \in O_d$: Then $l$ contradicts with happen before relation of $h$.
+    \item[-] If $o_2 \in O_a$: This is impossible, since the linearization point of operations in $O_a$ is before $\textit{ret}(\textit{put},b)$ in $e_{\textit{lp}}$.
+
+    \item[-] If $o_2 \in O_d$: Then $l$ contradicts with happen before relation of $h$.
     \end{itemize}
@@
     \item[-] If $o_2 \in O_d$: According to $l$, we can see that $\textit{put}(x,\textit{pri}_x)$ does not happen before any operation in $O_x$.
-Then we can see that the linearization point of $o_1$ is before $\textit{ret}(\textit{put},b)$ and $o_1 \in O_a$.
+Then we can see that the linearization point of $o_1$ is before $\textit{ret}(\textit{put},x)$ and $o_1 \in O_a$.
 This violates that $o_1 \in l''_2$.
@@
 
-    \item[-] If $o_2 \in O_+$: This is impossible, since the linearization point of operations in $O_+$ is before $\textit{ret}(\textit{rm},b)$ in $e_{\textit{lp}}$.
+    \item[-] If $o_2 \in O_+$: This is impossible, since the linearization point of operations in $O_+$ is before $\textit{ret}(\textit{rm},x)$ in $e_{\textit{lp}}$.
     \end{itemize}
@@
 
-    \item[-] If $o_1 \in W \wedge o_2 \in O_+$: This is impossible, since according to construction of $O_+$, we can see that $o_2 \in O_-$.
+    \item[-] If $o_1 \in W \wedge o_2 \in O_+$:
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $o_1 \in O_i(l,\textit{pri}_x)$: Then according to the construction process of $U' \cup V'$ and $W'$, we can see that $o_1 \in O_+$ and then $o_1 \in U' \cup V'$, which contradicts that $o_1 \in l''_3$.
+
+    \item[-] If $o_1 \in O_c(l,\textit{pri}_x)$: Then according to the construction process of $U' \cup V'$ and $W'$, we can see that $o_2 \in O_-$, which contradicts that $o_2 \in O_+$.
+    \end{itemize}
 
@@
         \item[-] If the reason of $o_1 \in O_-$ is case $2$: Let $o_h$ be as in case $2$.
-Then there exists $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{cal}(o_2)$.
+Then there exists $O_c(l,\textit{pri}_x)$-element $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{cal}(o_2)$.
 Therefore, we can see that $o_w <_{\textit{hb}} o_2$, and then $l$ contradicts with happen before relation of $h$.
@@
          \item[-] If the reason of $o_1 \in O_-$ is case $2$: Let $o_h$ be as in case $2$.
-Then there exists $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{cal}(\textit{put},x,\textit{pri}_x)$.
+Then there exists $O_c(l,\textit{pri}_x)$-element $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{cal}(\textit{put},x,\textit{pri}_x)$.
 Therefore, we can see that $o_w <_{\textit{hb}} \textit{put}(x,\textit{pri}_x)$, and then $l$ contradicts with happen before relation of $h$.
@@
          \item[-] If the reason of $o_1 \in O_-$ is case $2$: Let $o_h$ be as in case $2$.
-Then there exists $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$.
+Then there exists $O_c(l,\textit{pri}_x)$-element $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$.
 
@@
 
-It remains to prove that $l'' \in \textit{EPQ}$.
-Let $O_c$ be the set of operations in $h$, whose items have priority that are comparable with $\textit{pri}_x$, and let $O_i$ be the set of operations in $h$, whose items have priority that are incomparable with $\textit{pri}_x$.
-Then we obtain that $l'' \in \textit{EPQ}$ as follows:
+It remains to prove that $l'' \in \textit{EPQ}$. The process for proving $l'' \in \textit{EPQ}$ is as follows:
 
@@
 
-\item[-] $l'_{\textit{u'v'}}$ can be obtained from $l'_v$ as follows: From some time point $t$, discard all the $O_i$ operations after $t$ in $l'_v$.
-Since $\textit{pri}_x$ is one of maximal priority in $h$, we can see that $\forall p_i \in O_i$ and $\forall p_c \in O_c$, either $p_c <_{\mathbb{P}} p_i$, or $p_c$ is incomparable with $p_i$.
-Therefore, this process does not influence correctness of $l'_{\textit{u'v'}}$, and then $l'_{\textit{u'v'}} \in \textit{EPQ}$.
+\item[-] $l'_{\textit{u'v'}}$ can be obtained from $l'_v$ as follows:
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] Discard $O_c(l,\textit{pri}_x)$-element that are in $W$ and keep $O_c(l,\textit{pri}_x)$-element in $U \cup V$ unchanged.
+
+    \item[-] From some time point, discard all the $O_i(l,\textit{pri}_x)$ operations after this time point.
+    \end{itemize}
+
+From $\textit{matched-C}(u \cdot v,\textit{pri}_x)$, we can see that $O_c(l,\textit{pri}_x)$-element in $U \cdot V$ is matched, and then it is easy to see that the projection of $l'_v$ in to $O_c(l,\textit{pri}_x)$-element in $U \cdot V$ is still in $\textit{EPQ}$.
+By Lemma \ref{lemma:erase Oc still in EPQ} and Lemma \ref{lemma:erase Oi from some time point still in EPQ}, we can see that $l'_{\textit{u'v'}} \in \textit{EPQ}$.
 
 \item[-] $l''_1 \cdot l''_2$ can be obtained from $l'_{\textit{u'v'}}$ as follows: Execute until reaching some time point $t$, then first execute all $O_x$ operations after $t$, and then execute remanning operations.
-Since $O_x$ only contains $\textit{put}(\_,\textit{pri}_x)$ and $\textit{pri}_x$ is one of maximal priority in $h$, we can see that $l''_1 \cdot l''_2 \in \textit{EPQ}$.
-
-\item[-] Let $l''_f$ be the projection of $l'$ into $O_-$, and let $l''_g$ be the projection of $l'$ into operations not in $l''_1$, $l''_2$ or $l''_f$.
-It is easy to see that $l''_3 = l''_f \cdot l''_g$.
-
-\item[-] $l'_{\textit{u'v'}} \cdot l''_f$ can be obtained from $l'_v$ as follows: (1) Execute until reaching some time point $t$, (2) from $t$, execute only $O_c$ operations until the end of $l'_v$, and (3) execute $O_i$ operations after $t$ in $l'_v$.
-Since $\textit{pri}_x$ is one of maximal priority in $h$, we can see that $l'_{\textit{u'v'}} \cdot l''_f \in \textit{EPQ}$.
-
-\item[-] It is easy to see that $l' = l'_v \cdot l''_g \in \textit{EPQ}$.
-We already know that $l'_{\textit{u'v'}} \cdot l''_f \in \textit{EPQ}$, $l'_v \in \textit{EPQ}$.
-It is not hard to see that the content of extended priority queue after executing $l'_{\textit{u'v'}} \cdot l''_f$ is the same as that after executing $l'_v$.
-Therefore, it is easy to prove that $l'_{\textit{u'v'}} \cdot l''_f \cdot l''_g \in \textit{EPQ}$.
-
-\item[-] We already know that $l''_1 \cdot l''_2\in \textit{EPQ}$ and $l'_{\textit{u'v'}} \in \textit{EPQ}$, and it is not hard to see that the the content of extended priority queue after executing $l''_1 \cdot l''_2$ is the same as that after executing $l'_{\textit{u'v'}}$.
-Since $l'_{\textit{u'v'}} \cdot l''_f \cdot l''_g \in \textit{EPQ}$, it is easy to see that $l''_1 \cdot l''_2 \cdot l''_f \cdot l''_g \in \textit{EPQ}$.
-Or we can say, $l''_1 \cdot \l''_2 \cdot l''_3 \in \textit{EPQ}$.
-Since $u$, $v$, $w$, $x$ and $\textit{pri}_x$ satisfy the guard of $\textit{EPQ}_1$, it is easy to see that $l'' \in \textit{EPQ}$.
+By Lemma \ref{lemma:make put with maximal priority happen earlier}, we can see that $l''_1 \cdot l''_2 \in \textit{EPQ}$.
+
+\item[-] Let $l'_e$ be obtained from $l'$ by discarding $O_c(l,\textit{pri}_x)$-element in $U \cdot V$. By Lemma \ref{lemma:EPQ is closed under projection}, we can see that $l'_e \in \textit{EPQ}$.
+
+\item[-] Let $l'_e \vert_{O_i(l,\textit{pri}_x)}  = l'_f \cdot l'_g$, where $l'_f$ is the projection of $O_i(l,\textit{pri}_x)$-element in $U \cdot V$.
+Let $l'_h$ be obtained from $l'_e$ by discarding operations in $l'_f$.
+By Lemma \ref{lemma:make Oi happen earlier still in EPQ}, we can see that $l'_f \cdot l'_h \in \textit{EPQ}$.
+
+\item[-] By Lemma \ref{lemma:EPQ is closed under projection}, it is obvious that $l'_f \in \textit{EPQ}$.
+From $\textit{matched-C}(u \cdot v,\textit{pri}_x)$, we can see that the content of priority queue after executing $l''_1 \cdot l''_2$ is the same as after executing $l'_f$.
+By Lemma \ref{lemma:replace equivalent prefix still in EPQ}, we can see that $l''_1 \cdot l''_2 \cdot l'_h \in \textit{EPQ}$.
+It is easy to see that $l'_h = l''_3$, and then $l''_1 \cdot l''_2 \cdot l''_3 \in \textit{EPQ}$.
+
+\item[-] Since $u$, $v$, $w$, $x$ and $\textit{pri}_x$ satisfy the guard of $\textit{EPQ}_1$, it is easy to see that $l'' = l''_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3 \in \textit{EPQ}$.
 \end{itemize}
@@
     \item[-] If $o_2 \in O_a \wedge o_1 \in O_s$: It is not hard to see that $\textit{put}(x,\textit{pri}_x) <_{\textit{hb}} o_2$.
-But then it is impossible that in $e_{\textit{lp}}$, the linearization point of $o_2$ be located before $\textit{ret}(\textit{put},x)$.
+Then, it is impossible to locate the linearization point of $o_2$ before $\textit{ret}(\textit{put},x)$ in $e_{\textit{lp}}$.
     \end{itemize}
@@
 
-    \item[-] If $o_2 \in O_s$: This is impossible, since such $o_1 = \textit{put}(\_,\textit{pri}_x)$, and $l$ is consistent with happen before relation of $h$.
+    \item[-] If $o_2 \in O_s$: This is impossible, since $o_2 = \textit{put}(\_,\textit{pri}_x)$, and $l$ is consistent with happen before relation of $h$.
     \end{itemize}
@@
 $l''_a \cdot l''_s \cdot l''_2$ can be obtained from $l'$ as follows: Execute until reaching some time point $t$, then first execute all $O_x$ operations after $t$, and then execute remanning operations.
-Since $O_x$ only contains $\textit{put}(x,\textit{pri}_x)$ and $\textit{pri}_x$ is one of maximal priority in $h$, we can see that $l''_1 \cdot l''_2 = l''_a \cdot l''_s \cdot l''_2 \in \textit{EPQ}$.
-Since $u$, $v$, $x$ and $\textit{pri}_x$ satisfy the guard of $\textit{EPQ}_2$, it is easy to see that $l'' \in \textit{EPQ}$.
+By Lemma \ref{lemma:make put with maximal priority happen earlier}, we can see that $l''_1 \cdot l''_2 = l''_a \cdot l''_s \cdot l''_2 \in \textit{EPQ}$.
+Since $u$, $v$, $x$ and $\textit{pri}_x$ satisfy the guard of $\textit{EPQ}_2$, it is easy to see that $l'' = l''_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l''_2 \in \textit{EPQ}$.
 
@@
 
+\forget{
 \section{Proofs and Definitions in Section \ref{sec:relate other data structures with extended priority queue}}
@@
 \label{subsec:appendix proof and definition in section relate stack with extended priority queue}
+}

### 2017-04-14 ss-lin-EPQ.tex
I change proof of EPQ1 and add explanation to it

@@
 
+Let us briefly explain the idea of proving step-by-step linearizability of $\textit{EPQ}_1$ with an example, while the other two rules is much simpler to deal with.
+Given a data-differentiated concurrent execution $e \sqsubseteq l \in \textit{MS}(\textit{EPQ}_1)$ with witness $x$ and assume that $e \setminus x \sqsubseteq l' \in \textit{EPQ}$, we explicitly construct a sequence $l''$ and prove that $e \sqsubseteq l'' \in \textit{EPQ}$.
+The construction of $l''$ is as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] In \figurename~\ref{fig:concurrent execution for EPQ1}, we give an example of $e$, and we explicitly draw the linearization points according to $l'$.
+Assume that $p_1 <_{\mathbb{P}} p_2$ and $p_1 <_{\mathbb{P}} p_3$.
+According to $\textit{EPQ}_1$, there exists $u$, $v$ and $w$, such that $l = u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$.
+We use $\textit{put}(z_2,p_3)-w$ to denote that $\textit{put}(z_2,p_3)$ is in $w$.
+
+\item[-] Let $O_c$ and $O_i$ be the set of operations with priorities comparable with $\textit{pri}_x$ and incomparable with $\textit{pri}_x$ in $l$, respectively.
+It seems that we can make $l''= l''_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3$, where $l''_1$, $l''_2$ and $l''_3$ are sequences of operations chosen from $u$, $v$ and $w$, respectively.
+However, this is incorrect.
+The reason is that there is no restriction to $O_i$ operations in $u \cdot v$, and thus, there is no guarantee that the projection of $l'$ into $O_i$ operations in $u \cdot v$ being correct.
+In our example, we can see that such projection is $\textit{put}(z_1,p_3) \cdot \textit{rm}(z_2)$ and is incorrect.
+
+\item[-] Let $U$, $V$ and $W$ be the set of operations in $u$, $v$ and $w$, respectively. We use a two-steps method to construct $U'$, $V'$ and $W'$, which are the new version of $U$, $V$ and $W$.
+
+\item[-] The first step is to obtain $U' \cup V'$ and $W'$.
+We leave $O_c$ operations unchanged for $U \cdot V$ and $W$.
+We try to put as many as possible $O_i$ operations into $U' \cup V'$ (regardless whether they belong to $U \cdot V$ or $W$), until we meet a $O_i$ operation $o$ that happens before a $O_c$ operation in $W$.
+We move $O_i$ operations that before $o$ in $l'$ into $U' \cdot V'$, and move $o$ and $O_i$ operations that after $o$ in $l'$ into $W'$.
+In this example, $o$ is $\textit{put}(x_3,p_2)$ (emphasized by adding vertical dashed line), we move $\textit{put}(z_1,p_3)$ and $\textit{put}(z_2,p_3)$ into $U' \cdot V'$, and move $\textit{rm}(z_1)$ and $\textit{rm}(z_2)$ into $W'$ (emphasized by adding boxes).
+
+\item[-] The second step is to obtain $U'$ and $V'$.
+$U'$ contains the following two kinds of operations: (1) Operations whose linearization points are before $\textit{ret}(\textit{put},x)$, and (2) other $\textit{put}$ operations with priority $\textit{pri}_x$.
+$V'$ contains the remanning operations.
+In this example, $U'$ contains $\textit{put}(z_1,p_3)$ and $\textit{put}(x_2,p_2)$.
+
+\item[-] $l''_1$ is the concatenation of projection of $l'$ into the first and second kinds of operations in $U'$.
+$l''_2$ and $l''_3$ are projection of $l'$ into $V'$ and $W'$, respectively.
+In this example, $l''_1 = \textit{put}(z_1,p_3) \cdot \textit{put}(x_2,p_2)$, $l''_2 = \textit{put}(z_2,p_3) \cdot \textit{rm}(x_2) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(y_1)$, and $l''_3 = \textit{put}(x_3,p_2) \cdot \textit{rm}(z_1) \cdot \textit{rm}(z_2)$.
+In \figurename~\ref{fig:concurrent execution with new linearization points for EPQ1}, we add linarization points according to $l''$, and we can see that $l''$ holds as required.
+\end{itemize}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.5 \textwidth]{figures/PIC-HIS-EPQ1.pdf}
+%\vspace{-10pt}
+  \caption{Concurrent execution $e$}
+  \label{fig:concurrent execution for EPQ1}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.5 \textwidth]{figures/PIC-HIS-EPQ1-NEWLP.pdf}
+%\vspace{-10pt}
+  \caption{Concurrent execution $e$ with linearization points according to $l''$}
+  \label{fig:concurrent execution with new linearization points for EPQ1}
+\end{figure}
+
 The following lemma states that for data-differentiated execution, checking linearizability with respect to $\textit{EPQ}$ is equivalent to checking linearizability with respect to $\textit{MS}(R)$ for everyone of its projections.

### 2017-04-14 ss-lin-EPQ.tex
change explanation

--- 2017-04-04 ss-lin-EPQ.tex
@@
 
+Let us briefly explain the idea of proving step-by-step linearizability of $\textit{EPQ}_1$ with an example, while the other two rules is much simpler to deal with.
+Given a data-differentiated concurrent execution $e \sqsubseteq l = u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w \in \textit{MS}(\textit{EPQ}_1)$ with witness $x$ and assume that $e \setminus x \sqsubseteq l' \in \textit{EPQ}$, we explicitly construct a sequence $l''= l''_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3$ and prove that $e \sqsubseteq l'' \in \textit{EPQ}$.
+$e$ is shown in \figurename~\ref{fig:concurrent execution for EPQ1} and we explicitly draw the linearization points according to $l'$.
+Here $\textit{pri}_x=p_2$, and assume that $p_1 <_{\mathbb{P}} p_2$ and $p_1 <_{\mathbb{P}} p_3$.
+Here $u = \epsilon$, $v = \textit{put}(z_1,p_3) \cdot \textit{put}(x_2,p_2) \cdot \textit{rm}(y_1) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(z_2) \cdot \textit{rm}(x_2)$, and $w=\textit{rm}(z_2) \cdot \textit{put}(x_3,p_2) \cdot \textit{rm}(z_1)$.
+We use $o-w$ to emphasize that $o$ is in $w$.
+
+The construction of $l''$ is as follows. Our construction does not rely how to choose linearization points of $l'$ in $e$.
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] At first glance, we can construct $l''_1$, $l''_2$ and $l''_3$ as the projection of $l'$ into operations of $u$, $v$ and $w$, respectively.
+However, this is incorrect.
+To explain this, let $\textit{pri}_x$-comparable operations (resp., $\textit{pri}_x$-incomparable operations) be the operations with items whose priority is comparable with $\textit{pri}_x$ (resp., incomparable with $\textit{pri}_x$).
+According to $\textit{EPQ}_1$, there is no restriction to $\textit{pri}_x$-incomparable operations operations in $u \cdot v$, and thus, there is no guarantee that the projection of $l'$ into $\textit{pri}_x$-incomparable operations operations in $u \cdot v$ being correct.
+In this example, we can see that such projection is $\textit{put}(z_1,p_3) \cdot \textit{rm}(z_2)$ and is incorrect.
+
+\item[-] Let us construct set $U'$, $V'$ and $W'$, such that $l''_1$, $l''_2$ and $l''_3$ are projection of $l'$ into $U'$, $V'$ and $W'$, respectively. The construction contains two steps:
+
+\item[-] The first step is to define $W'$.
+The $\textit{pri}_x$-comparable operations in $W'$ is same as that in $W$.
+To obtain $\textit{pri}_x$-incomparable operations in $W'$, we try to find an operation $o$ which either happens before some $\textit{pri}_x$-comparable operations in $W$, or with linearization points after $\textit{rm}(x)$.
+In this example, $o$ is $\textit{put}(x_3,p_2)$ (emphasized by adding vertical dashed line).
+Then, we put $o$ and all $\textit{pri}_x$-incomparable operations in $l'$ whose linearization points are after $o$ into $W'$.
+In this example, $W'$ contains $\textit{put}(x_3,p_2)$, $\textit{rm}(z_1)$ and $\textit{rm}(z_2)$.
+We use boxes to emphasize they are put into $W'$.
+
+\item[-] The second step is to define $U'$ and $V'$.
+$U'$ contains the following two kinds of operations: (1) Operations whose linearization points are before $\textit{ret}(\textit{put},x)$, and (2) other $\textit{put}$ operations with priority $\textit{pri}_x$.
+$V'$ contains the remanning operations.
+In this example, $U'$ contains $\textit{put}(z_1,p_3)$ and $\textit{put}(x_2,p_2)$.
+
+\item[-] In this example, $l''_1 = \textit{put}(z_1,p_3) \cdot \textit{put}(x_2,p_2)$, $l''_2 = \textit{put}(z_2,p_3) \cdot \textit{rm}(x_2) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(y_1)$, and $l''_3 = \textit{put}(x_3,p_2) \cdot \textit{rm}(z_1) \cdot \textit{rm}(z_2)$.
+In \figurename~\ref{fig:concurrent execution with new linearization points for EPQ1}, we add linarization points according to $l''$, and we can see that $l''$ holds as required.
+\end{itemize}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.5 \textwidth]{figures/PIC-HIS-EPQ1.pdf}
+%\vspace{-10pt}
+  \caption{Concurrent execution $e$}
+  \label{fig:concurrent execution for EPQ1}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.5 \textwidth]{figures/PIC-HIS-EPQ1-NEWLP.pdf}
+%\vspace{-10pt}
+  \caption{Concurrent execution $e$ with linearization points according to $l''$}
+  \label{fig:concurrent execution with new linearization points for EPQ1}
+\end{figure}
+
 The following lemma states that for data-differentiated execution, checking linearizability with respect to $\textit{EPQ}$ is equivalent to checking linearizability with respect to $\textit{MS}(R)$ for everyone of its projections.

### 2017-04-18 Seq-EPQ-dataInd.tex
add files

\section{Priority Queue and Data-Independence}
\label{sec:priority queue and data-independence}

A priority queue with partially-ordered priorities (priority queue for short) contains two method: $\textit{put}$ and $\textit{rm}$. A $\textit{put}$ method has two arguments, while the first argument is an item and the second argument is its priority. A $\textit{put}$ method is used to put an item into the priority queue with certain priority. Here the item is chosen from a specific (possibly infinite) data domain $\mathbb{D}$ and priority is chosen from a (possibly infinite) set $\mathbb{P}$. Moreover, there is a strict partial-order $\prec$ among elements in $\mathbb{P}$. A $\textit{rm}$ method intends to remove the item with minimal priority (w.r.t $\prec$) in priority queue and then returns it. It works as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If the priority queue is empty, then $\textit{rm}$ returns $\textit{empty}$.

\item[-] Else, $\textit{rm}$ choose one of minimal priority of items in priority queue, and returns the earliest putted item of this priority. Note that if there are more than one incomparable and minimal candidate priorities, then the chosen of priority is arbitrary.
\end{itemize}

For example, assume that $p_1 \prec p_2$, then $\textit{put}(a,p_2) \cdot \textit{put}(b,p_1) \cdot \textit{rm}(b)$ is a correct behavior of priority queue, since $b$ has smaller priority.

Let us introduce the notion of sequential executions, which modelled the specification of concurrent libraries. We fix a (possibly infinite) set $\mathbb{D}$ of data values, a finite set $\mathbb{M}$ of methods, and an infinite set $\mathbb{O}$ of operation (identifiers). Given $m \in \mathbb{M}$, $x \in \mathbb{D}$, and $o \in \mathbb{O}$, a call action $\textit{call}_o (m,x)$ represents an invocation to method $m$ with argument $x$, while a return action $\textit{ret}_o (m,x)$ represents a response from method $m$ with return value $x$. \footnote{Call actions with more than one arguments are similarly defined.} Here $o$ is used to match return actions to their call actions: $\textit{cal}_o (m,x)$ matches $\textit{ret}_{o'} (m',x')$, if $o=o' \wedge m=m'$. A sequential execution is a sequence of call and return actions, while each call action is immediately followed by its matching return action. Let $\cdot$ be the concatenation of sequences. To ease the reading, given a sequential execution $e=\textit{call}_{o_1}(m_1,a_1) \cdot \textit{rm}_{o_1}(m_1,b_1) \cdot \ldots \cdot \textit{call}_{o_n}(m_n,a_n) \cdot \textit{rm}_{o_n}(m_n,b_n)$, when the context is clear, we can write it as $e=m_1(a_1,b_1) \cdot \ldots \cdot m_n(a_n,b_n)$. We write $m(a,b)$ as $m(a)$ or $m(b)$, when there is no return value or there is no argument, respectively.

Data-independence \cite{Wolper:1986} can be used to effectively handle unbounded data domain. In this paper, we slightly modify the notion of data-independence in \cite{Wolper:1986} and propose data-independence for priority queues. Let $\_$ denote an element, of which the value is irrelevant. A sequential execution $e$ of priority queue is data-differentiated if, for all $d \in \mathbb{D}$, there is at most one $\textit{put}(d,\_)$ in $e$. Note that a data-differentiated $e$ may contains more than one items with a same priority. The subset of data-differentiated sequential executions of a set $S$ is denoted by $S_{\neq}$.

A renaming function $r$ for priority queue is a function from $\mathbb{D}$ to $\mathbb{D}$. Given a sequential execution $e$, we denote by $r(e)$ the sequential execution obtained from $e$ by replacing every item $x$ by $r(x)$. Note that here the renaming functions rename only the items and keep the priorities unchanged. The explanation is that renaming items will not influence the executions of program, while renaming priorities may do.

%\vspace{-6pt}
\begin{definition}\label{def:priority-value data-independence}
A set of sequential executions $S$ is data-independent, if:
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] for all $e \in S$, there exists $e' \in S'$, and a renaming function $r$, such that $e=r(e')$,

\item[-] for all $e \in S$ and for all renaming $r$, $r(e) \in S$.
\end{itemize}
\end{definition}

In latter section, we shows that with the help of data-independence, we can concentrate on only data-differentiated executions when checking correctness.

### 2017-04-18 appendix-EPQ.tex
add files

@@
 
-\EPQisClosedUnderProjection*
+\begin{restatable}{lemma}{EPQisClosedUnderProjection}
+\label{lemma:EPQ is closed under projection}
+$\textit{EPQ}$ is closed under projection.
+\end{restatable}
 
@@
 
-\EPQasMultiInMRpriforSequence*
+%\EPQasMultiInMRpriforSequence*
+
+\begin{restatable}{lemma}{EPQasMultiInMRpriforSequence}
+\label{lemma:EPQ as multi in MRpri for sequence}
+Given a data-differentiated sequential execution $e$, $e \in \textit{EPQ}$, if and only if, $\forall e' \in \textit{proj}(e)$ and $\forall R \in \textit{last}(e')$, we have $e' \in \textit{MS}(R)$.
+\end{restatable}
 
@@
 
-\EPQueueisStepByStepLinearizability*
+%\EPQueueisStepByStepLinearizability*
+
+\begin{restatable}{lemma}{EPQueueisStepByStepLinearizability}
+\label{lemma:EPQ is step-by-step linearizability}
+$\textit{PQ}$ is step-by-step linearizability.
+\end{restatable}
 
@@
 
+To facilitate our proof of latter sections, we need to separate $\textit{EPQ}_1$ into two cases: (1) no matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v \cdot w$ has priority $\textit{pri}$, (2) some matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v \cdot w$ has priority $\textit{pri}$.
+Therefore, we separate $\textit{EPQ}_1$ into two cases:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $\textit{EPQ}_1^{>}(u,v,w,\textit{pri}) \equiv
+(\textit{noRE}(u \cdot v \cdot w)) \wedge
+(\textit{LI}(\textit{pri}, u \cdot v \cdot w)) \wedge
+(\textit{LI-U}(\textit{pri},u \cdot v \cdot w)) \wedge
+(\textit{matched-C}(u \cdot v,\textit{pri}) )$
+
+\item[-] $\textit{EPQ}_1^{=}(u,v,w,\textit{pri}) \equiv
+(\textit{noRE}(u \cdot v \cdot w)) \wedge
+(\textit{LE'I}(\textit{pri}, u \cdot v \cdot w)) \wedge
+(\textit{LI-U}(\textit{pri},u \cdot v \cdot w)) \wedge
+(\textit{matched-C}(u \cdot v,\textit{pri}) ) \wedge
+(\textit{putInSeq}(u \cdot v \cdot w,u,\textit{pri}))$
+\end{itemize}
+
+For $\textit{EPQ}_2$, we also need to distinguish two cases: (1) no matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ has priority $\textit{pri}$, (2) some matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ has priority $\textit{pri}$.
+Therefore, we separate $\textit{EPQ}_2$ into two rules:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $\textit{EPQ}_2^{>}(u,v) \equiv
+(\textit{noRE}(u \cdot v)) \wedge
+(\textit{LEI}(\textit{pri},u \cdot v)) \wedge
+(\textit{LI-M}(\textit{pri},u \cdot v)) \wedge
+(\textit{putInSeq}(u \cdot v,u,\textit{pri}))$.
+
+\item[-] $\textit{EPQ}_2^{=}(u,v) \equiv
+(\textit{noRE}(u \cdot v)) \wedge
+(\textit{LEI}(\textit{pri},u \cdot v)) \wedge
+(\textit{LE'I-M}(\textit{pri},u \cdot v)) \wedge
+(\textit{putInSeq}(u \cdot v,u,\textit{pri}))$.
+\end{itemize}
+
 \subsection{Proofs and Definitions in Subsection \ref{subsec:definition of co-regular}}
 \label{sec:appendix proof and definition in section definition of co-regular}
+
+Before we go to investigate co-regular of each rules, we use the results in \cite{Bouajjani:2015} to simplify our work.
+\cite{Bouajjani:2015} states that checking linearizability w.r.t queue can be reduced into checking emptiness of intersection between $\mathcal{I}$ and a set of automata.
+Given a data-differentiated execution $e$, let $e \vert_{i}$ be an execution generated from $e$ by erasing call and return actions of items that does not use priority $i$ (does not influence $\textit{rm}(\textit{empty})$).
+We call a extended priority queue execution with only one priority a single-priority execution.
+Let $\textit{transToQueue}(e)$ be an execution generated from $e$ by transforming $\textit{put}$ and $\textit{rm}$ into $\textit{enq}$ and $\textit{deq}$, respectively, and then discarding priorities.
+We can see that for each $e \in \textit{EPQ}$ and each priority $i$, $\textit{transToQueue}(e \vert_{i})$ satisfy FIFO (first in first out) property.
+
+Given an execution of queue, we say that it is differentiated \cite{Wolper:1986}, if each item is enqueued at most once.
+\cite{Bouajjani:2015} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $a,b$: (1) $\textit{deq}(b) <_{hb} \textit{enq}(b)$, (2) there are no $\textit{enq}(b)$ and at least one $\textit{deq}(b)$, (3) there are one $\textit{enq}(b)$ and more than one $\textit{deq}(b)$, and (4) $\textit{enq}(a) <_{\textit{hb}} \textit{enq}(b)$, and $\textit{deq}(b) <_{\textit{hb}} \textit{deq}(a)$, or $\textit{deq}(a)$ does not exists.
+For each such case, we can construct a witness automata for extended priority queue.
 
@@
 \label{sec:appendix proof and definition in section co-regular of EPQ1Lar}
+
+The notion of left-right constraint used for extended priority queue is inspired by left-right constraint of queue \cite{Bouajjani:2015}.
+Let us introduce the notion of left-right constraint, which is a graph, and the existence of cycle though item with maximal priority in it correspond to the existence of the case in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
+
+\begin{definition}\label{def:left-right constraint for matched put and rm operations}
+Given a data-differentiated $\textit{pri}_x$-execution $e$ without $\textit{rm}(\textit{empty})$.
+Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be the only  method events of $e$ with priority $\textit{pri}_x$.
+The left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is the graph $G$ where:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] the nodes are the items of $e$, to which we add a node,
+
+\item[-] there is an edge from item $d_1$ to $x$, if $\textit{put}(d_1,\_)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$,
+
+\item[-] there is an edge from $x$ to item $d_1$, if $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$,
+
+\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$.
+\end{itemize}
+\end{definition}
+
+When there is a cycle $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ through item with maximal priority (for example, $x$) in $G$, we say that $x$ is covered by $d_1,\ldots,d_m$.
+To state the effectiveness of left-right constraint, take the execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} as an example, we can see that $b$ is covered by $f,e,d$.
+We need to prove that getting rid of cycle though item with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\textit{MS}(\textit{EPQ}_1^{>})$, as stated by the following lemma:
+
+\begin{restatable}{lemma}{LinEqualsConstraintforEPQOneLar}
+\label{lemma:Lin Equals Constraint for EPQ1Lar}
+Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{>}$.
+Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority.
+Let $G$ be the graph representing the left-right constraint of $\textit{put}(x)$ and $\textit{rm}(x)$.
+$e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$, if and only if $G$ has no cycle going through $x$.
+\end{restatable}
 

### 2017-04-18 recur-process-concurrent.tex
add files

\section{Checking Linearizability By Recursive Procedure}
\label{sec:checking linearizablity by recursive procedure}

In this section, we extend $\textit{check-PQ}$ with linearizability, and use it to simplify checking linearizability w.r.t $\textit{PQ}$. This inspired us to partition the concurrent executions which are not linearizable w.r.t $\textit{PQ}$ into violation of $\textit{MS}(R)$. To ensure the correctness of $\textit{check-PQ}$, we step-by-step linearizability to incrementally build linearization. Our step-by-step linearizability is inspired by the step-by-step linearizabilility of queue and stacks in \cite{Bouajjani:2015}. The related definitions and proofs of lemmas in this section can be found in Appendix \ref{sec:appendix in section step-by-step linearizability of extended priority queues}.

\subsection{Linearizability}
\label{subsec:linearizability}

A (concurrent) execution $e$ is a sequence of call and return actions which satisfy a well-formedness property: every return has a matching call action before it in $e$, and an operation $o$ can be used only twice in $e$, once in a call action, and once in a return action. The definition of data-differentiated, renaming function, data-independence and projection extends to (concurrent) executions. For instance, an execution is data-differentiated if, for all $d \in \mathbb{D}$, there is at most one $\textit{cal}_{\_}(\textit{put},d,\_)$.

An implementation $\mathcal{I}$ is a set of concurrent executions. Implementations represent libraries whose methods are called by external programs. In the remainder of this work, we consider only completed executions, where each call action has a corresponding return action. This simplification is sound when implementation methods can always make progress in isolation \cite{Henzinger:2013}: formally, for any execution $e$ with pending operations, there exists an execution $e'$ obtained by extending $e$ only with the return actions of the pending operations of $e$. Intuitively this means that methods can always return without any help from outside threads, avoiding deadlock.

Given execution $e$, the happen-before relation $<_{\textit{hb}}$ is defined as follows: $o_1 <_{\textit{hb}} o_2$, if the return action of $o_1$ is before the call action of $o_2$ in $e$. The happens-before relation are interval orders \cite{Bouajjani:2015POPL}: for distinct $o_1,o_2,o_3,o_4$, if $o_1 <_{\textit{hb}} o_2$ and $o_3 <_{\textit{hb}} o_4$, then either $o_1 <_{\textit{hb}} o_4$, or $o_3 <_{\textit{hb}} o_2$. Intuitively, this comes from the fact that concurrent threads share a notion of global time.

Linearizability \cite{Herlihy:1990} is the \emph{de facto} standard correctness condition for concurrent data structures. Given concurrent execution $e$ and sequential execution $s$ of length $n$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if they have same set of operations, and there is a bijection $f$ between operations s.t.

\begin{itemize}
\item[-] if $o_1 <_{\textit{hb}} o_2$, then $f(o_1) <_{\textit{hb}} f(o_2)$,

\item[-] $o$ and $f(o)$ have same call and return actions.
\end{itemize}

A concurrent execution $e$ is linearizable w.r.t a set $S$ of sequential executions, denoted $e \sqsubseteq S$, if there exists $s \in S$ such that $e \sqsubseteq s$. A set of concurrent executions $E$ is linearizable w.r.t $S$, denoted $E \sqsubseteq S$, if $e \sqsubseteq S$ for all $e \in H$.

The following lemma states that with the help of data-independence, it is enough to consider only data-differentiated executions when checking linearizability. This is similar to that in \cite{Abdulla:2013}, where they use the notion of data-independence in \cite{Wolper:1986}. Thus, in the remainder of the paper, we focus on characterizing linearizability for data-differentiated executions. The proof of this lemma can be found in Appendix \ref{sec:appendix in section data-independence of EPQ}.

\begin{restatable}{lemma}{DataDifferentiatedisEnoughforPQ}
\label{lemma:data differentiated is enough for PQ}
A data-independent implementation $\mathcal{I}$ is linearizable w.r.t data-independent set $S$ of sequential executions, if and only if $\mathcal{I}_{\neq}$ is linearizable with respect to $S_{\neq}$.
\end{restatable}

\subsection{Modified $\textit{check-PQ}$}
\label{subsec:recursive procedure con-check-EPQ}

Let us extend $\textit{check-PQ}$ with linearizability to deal with linearizability w.r.t $\textit{PQ}$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] The input $e$ is a concurrent execution.

\item[-] $\textit{last}(e)$ is the union of $\textit{last}(u)$ for any sequential execution $u$ such that $e \sqsubseteq u$. $e \sqsubseteq \textit{MS}(R)$ with witness $x$ (resp., $o$), if $e \sqsubseteq u \in \textit{MS}(R)$ and $x$ (resp., $o$) is the witness of $u$.

\item[-] $P_i(e,o)$ and $P_i(e,x)$ holds, if $e \sqsubseteq \textit{MS}(\textit{PQ}_i)$ withe witness $o$ or $x$, respectively. 
\end{itemize}

The following lemma states that $\textit{check-PQ}$ is correct.

\begin{restatable}{lemma}{ConCheckEPQIsCorrect}
\label{lemma:con-check-EPQ is correct}
Given a data-differentiated execution $e$, $e \sqsubseteq \textit{EPQ}$, if and only if, $\textit{check-PQ} = \textit{true}$.
\end{restatable}

\begin {proof} (Sketch)
It is not hard to see that $\textit{check-PQ}(e) = \textit{false} \Rightarrow e \not\sqsubseteq \textit{PQ}$, since $\textit{PQ}_{\neq} \subset \textit{MS}(R)_{\neq}$. To ensure that $\textit{check-PQ}(e) = \textit{true} \Rightarrow e \sqsubseteq \textit{PQ}$, we require $\textit{PQ}$ to satisfy the following property.

for any data-differentiated execution $e$,
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] if $e \sqsubseteq \textit{MS}(R)$ ($R \in \{ \textit{PQ}_1, \textit{PQ}_2 \}$) with witness $x$, we have: $e \setminus x \sqsubseteq \textit{PQ} \Rightarrow e \sqsubseteq \textit{PQ}$.

\item[-] if $e \sqsubseteq \textit{MS}(\textit{PQ}_3)$ and $o$ is a $\textit{rm}(\textit{empty})$, we have: $e \setminus o \sqsubseteq \textit{PQ} \Rightarrow e \sqsubseteq \textit{PQ}$.
\end{itemize}

With this property, we can build a linearization of a whole execution by increasingly construct linearization of sub-execution from $\epsilon$, and this convince us that $\textit{check-PQ}(e) = \textit{true} \Rightarrow e \sqsubseteq \textit{PQ}$.

Let us briefly explain the idea of proving this property for $\textit{PQ}_1$ with an example. Given a data-differentiated concurrent execution $e \sqsubseteq u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \textit{MS}(\textit{PQ}_1)$ with witness $x$ and assume that $e \setminus x \sqsubseteq l' \in \textit{PQ}$, as shown in \figurename~\ref{fig:concurrent execution for EPQ1} (a), we explicitly construct a sequence $l''= l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3$ and prove that $e \sqsubseteq l'' \in \textit{PQ}$. In this example, $p_1 \prec p$, $p_1 \prec p_2$, $u=\epsilon$, $w$ contains $\textit{rm}(z_2), \textit{put}(x_3,p_2), \textit{rm}(z_1)$ (emphasize by $\textit{rm}(z_2)-w$), and the remanning operations are in $v$. We also explicitly draw the linearization points according to $l'$. We construct $l''$ as follows, and this method does not rely on the locations of linearization points of $l'$ and operations of $x$. Here we implicitly mix operation $o$ (when $\textit{call}_o(m,a)$, $\textit{ret}_o(m,b)$ in $e$) with $m(a,b)$.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Let $p$-comparable operations (resp., $p$-incomparable operations) be the set of operations with items whose priority is comparable with $p$ (resp., incomparable with $p$). It seems that we can construct $l''_1$, $l''_2$ and $l''_3$ as the projection of $l'$ into operations of $u$, $v$ and $w$, respectively. However, this is incorrect, since $\textit{PQ}_1$ has no restriction to $p$-incomparable operations operations in $u \cdot v$, and thus, there is no guarantee that the projection of $l'$ into $p$-incomparable operations operations in $u \cdot v$ being correct. In this example, such projection is $\textit{put}(z_1,p_3) \cdot \textit{rm}(z_2)$ and is incorrect.

\item[-] Let us construct sets $U'$, $V'$ and $W'$, such then make $l''_1$, $l''_2$ and $l''_3$ the projections of $l'$ into $U'$, $V'$ and $W'$, respectively. This contains two steps:

\item[-] The first step is to define $W'$. The $p$-comparable operations in $W'$ is same as that in $w$. To obtain $p$-incomparable operations in $W'$, we try to find a $p$-incomparable operation $o$ which either happens before some $p$-comparable operations in $w$, or with linearization points after $\textit{rm}(x)$. Then, we put $o$ and all $p$-incomparable operations after $o$ in $l'$ into $W'$ (emphasized by boxes in example). In this example, $o$ is $\textit{rm}(z_1)$, and $W'$ contains $\textit{put}(x_3,p)$, $\textit{rm}(z_1)$ and $\textit{rm}(z_2)$. We use boxes to emphasize they are put into $W'$.

\item[-] The second step is to define $U'$ and $V'$. $U'$ contains the following two kinds of operations: (1) Operations whose linearization points are before $\textit{ret}(\textit{put},x)$, and (2) other $\textit{put}$ operations with priority $p$. $V'$ contains the remanning operations. In this example, $U'$ contains $\textit{put}(z_1,p_3)$ and $\textit{put}(x_2,p_2)$.

\item[-] In this example, $l''_1 = \textit{put}(z_1,p_2) \cdot \textit{put}(x_2,p)$, $l''_2 = \textit{put}(z_2,p_2) \cdot \textit{rm}(x_2) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(y_1)$, and $l''_3 = \textit{put}(x_3,p) \cdot \textit{rm}(z_1) \cdot \textit{rm}(z_2)$. In \figurename~\ref{fig:concurrent execution for EPQ1} (b), we add linarization points according to $l''$, and we can see that $l''$ holds as required.
\end{itemize}

This completes the proof of this lemma.

\qed
\end {proof}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1 \textwidth]{figures/PIC-HIS-EPQ1-TwoHis.pdf}
%\vspace{-10pt}
  \caption{The process of obtaining linearization of $e$}
  \label{fig:concurrent execution for EPQ1}
\end{figure}

### 2017-04-18 recur-process.tex
add files

\section{Checking Inclusion By Recursive Procedure}
\label{sec:checking inclusion by recursive procedure}

Given a data-differentiated sequential execution $e$, it is costly to check whether $e$ belongs to priority queue, since for each element of $e$, we should keep the current content of priority queue and try to modify the priority queue according to this element. Such process is too complex. In this section, we propose another method for checking inclusion of priority queue by using a recursive procedure $\textit{check-PQ}$. For checking a sequence, every time we only check a much simpler property, and then recursively check the remanning sequences obtained by erasing one or two elements. Each step of our method is easier to deal with.

$\textit{put}(a,p)$ matches $\textit{rm}(b)$, if $a = b$. To introduce $\textit{check-PQ}$, let us introduce several predicates. Given sequential execution $l$ and priority $p$:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{rm}(\textit{empty}) \notin l$ is satisfied when $l$ does not contain $\textit{rm}(\textit{empty})$.

\item[-] Let $\textit{priorities}(l)$ and $\textit{unmatched-priorities}(l)$ be the set of priorities of putted items and items of unmatched put in $l$, respectively. $p \prec \textit{priorities}(l)$ (resp., $p \preceq \textit{priorities}(l)$), if $p \prec p'$ (resp., $p \preceq p'$)for some $p' \in \textit{priorities}(l)$. Similarly we can define $p \preceq \textit{unmatched-priorities}(l)$.

\item[-] $\textit{matched}_{\prec}(l,p)$ is satisfied, if for each item of $l$ with priorities smaller than $p$, it has matched $\textit{put}$ and $\textit{rm}$. Similarly, we can define $\textit{matched}(l)$, where we consider all items in $l$ instead of items with certain priorities.
\end{itemize}

Let us introduce some notations: given sequential executions $u,v,w$ and priority $p$,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $\textit{PQ}_1(u,v,w,p) \equiv
(\textit{rm}(\textit{empty}) \notin u \cdot v \cdot w) \wedge
(p \not\prec \textit{priorities}(u \cdot v \cdot w)) \wedge
(p \not\preceq \textit{unmatched-priorities}(u \cdot v \cdot w)) \wedge
(\textit{matched}_{\prec}(u \cdot v,p) ) \wedge
(p \notin \textit{priorities}(v \cdot w))$

\item[-] $\textit{PQ}_2(u,v,p) \equiv
(\textit{rm}(\textit{empty}) \notin u \cdot v) \wedge
(p \not\prec \textit{priorities}(u \cdot v \cdot w)) \wedge
(p \notin \textit{priorities}(v \cdot w))$

\item[-] $\textit{PQ}_3(u,v) \equiv
\textit{matched}(u \cdot v)$
\end{itemize}

Let us defined set $\textit{MS}(R)$ ($R \in \textit{PQ}_1,\textit{PQ}_2,\textit{PQ}_3$), which is the set of data-differentiated sequential executions that respect $R$.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $e \in \textit{MS}(\textit{PQ}_1)$, if $e = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$ and $\textit{PQ}_1(u,v,w,p)$ holds for some item $x$ and a maximal priority $p$,

\item[-] $e \in \textit{MS}(\textit{PQ}_2)$, if $e = u \cdot \textit{put}(x,p) \cdot v$ and $\textit{PQ}_2(u,v,p)$ holds for some item $x$ and a maximal priority $p$,

\item[-] $e \in \textit{MS}(\textit{PQ}_3)$, if $e = u \cdot \textit{rm}(\textit{empty}) \cdot v$ and $\textit{PQ}_3(u,v)$ holds.
\end{itemize}

We call such $x$ and $\textit{rm}(\textit{empty})$ (if exists) the witness of $e$. Note that sequences in $\textit{MS}(R)$ does not guaranteed to be a priority queue execution. Let us introduce the notion $\textit{last}(e)$, which is a set and used as the branch condition of $\textit{check-PQ}$. Here $e$ is a data-differentiated sequential execution.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $e$ contains $\textit{rm}(\textit{empty})$, then $\textit{last}(e) = \{ \textit{PQ}_3 \}$.

\item[-] Else, if for a maximal priority $p$ of $e$, $\textit{unmatched-priorities}(e) \neq \emptyset$, then $\textit{PQ}_2 \in \textit{last}(e)$.

\item[-] Else, if for a maximal priority $p$ of $e$, $\textit{unmatched-priorities}(e) = \emptyset$, then $\textit{PQ}_1 \in \textit{last}(e)$.
\end{itemize}

When $\textit{last}(e)$ contains only one element $R$, we write $\textit{last}(w)=R$ for simplicity.

The projection $e \vert{\mathcal{D}}$ of $e$ into $\mathcal{D} \subseteq \mathbb{D}$ is obtained from $e$ by erasing all elements with a data value not in $\mathcal{D}$. We write $e \setminus x$ for the projection $e \vert_{ \mathbb{D} \setminus \{ x \} }$. The projection $e \vert{O}$ of $e$ into $O$ is obtained from $e$ by erasing all call and return actions of non-$O$ operations. We write $e \setminus o$ for the projection $e \vert_{ O_e \setminus \{ o \} }$, where $O_e$ is the set of operations of $e$.

The recursive procedure $\textit{con-check-EPQ}$ is given as follows. Here $P_i(e,o)$ and $P_i(e,x)$ is the predicates of $e \in \textit{MS}(\textit{PQ}_i)$ withe witness $o$ or $x$, respectively.

%\begin{minipage}{.5\textwidth}
\begin{algorithm}[H]
\KwIn {a data-differentiated sequential execution $e$}

\If {$e = \epsilon$}
{\KwRet $\textit{true}$;}

\If {$\textit{last}(e) = \textit{PQ}_3$}
{
    \If {$\exists o=\textit{rm}(\textit{empty})$, $P_3(e,o)$ holds}
    {
        \KwRet $\textit{check-PQ}(e \setminus o)$;
    }
    \KwRet $\textit{false}$;
}

\If {$\textit{PQ}_i \in \textit{last}(e)$ for $i \in \{1,2\}$}
{
    \If {$\exists l$, $P_i(e,x)$ holds}
    {
        \KwRet $\textit{check-PQ}(e \setminus x)$;
    }
    \KwRet $\textit{false}$;
}
\caption{$\textit{check-PQ}$}
\label{Method-check-PQ}
\end{algorithm}
%\end{minipage}
%\hfill

Let us use an example to show how $\textit{check-PQ}$ works. Here we use $w_1 \xrightarrow{R} w_2$ to denote that in $\textit{check-PQ}(w_1)$, we choose the branch of $R$ and recursively call $\textit{check-PQ}(w_2)$.

\begin{example}\label{example:generate extended priority queue executions}
Given priorities $p_1,p_2,p_3$ with orders $p_1 \prec p_2$ and $p_1 \prec p_3$,

\noindent $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{rm}(\textit{empty}) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)$

$\xrightarrow{\textit{EPQ}_3}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)$

$\xrightarrow{\textit{EPQ}_1}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(b,p_1)$

$\xrightarrow{\textit{EPQ}_2}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(b,p_1)$

$\xrightarrow{\textit{EPQ}_1}$ $\textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{put}(b,p_1)$

$\xrightarrow{\textit{EPQ}_2}$ $\textit{put}(a,p_1) \cdot \textit{rm}(a)$

$\xrightarrow{\textit{EPQ}_1}$ $\epsilon$
\end{example}

Let $\textit{PQ}$ be set of sequences obtained by renaming some data-differentiated sequential execution that is accepted by $\textit{check-PQ}$. To persuade readers that $\textit{PQ}$ is indeed the set of behaviors of priority queue, in Appendix \ref{sec:appendix in section inductive rules of extended priority queue}, we give a semantical version definition $\textit{PQ}_s$ of priority queue, and shows that $\textit{PQ} = \textit{PQ}_s$. To model the possible behaviors of priority queue, we model it as an labelled transition system (shortened as LTS) $\textit{LTS}_e$. Each state of $\textit{LTS}_e$ is a function from $\mathbb{P}$ into sequences over $\mathbb{D}$, and represents a snapshot of contents of priority queue. $\textit{PQ}_s$ is the set of traces of $\textit{LTS}_e$. The definition of $\textit{PQ}_s$ and the proof of the following lemma can be found in Appendix \ref{sec:appendix in section inductive rules of extended priority queue}.

\begin{restatable}{lemma}{EPQRulesAndSemantics}
\label{lemma:EPQ rules and semantics}
$\textit{PQ} = \textit{PQ}_s$.
\end{restatable}

By Example \ref{example:generate extended priority queue executions} we can see how to generate a sequence $e \in \textit{EP}$ from $\epsilon$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] First we add non-$\textit{rm}(\textit{empty})$ elements with a loop: Let $P$ be the set of priorities of items in $e$. In the first round of the loop, we choose a minimal priorities $p \in P$, we add items by using $\textit{PQ}_1$ and $\textit{PQ}_2$, let $P = P \setminus \{p\}$, and begins the next round of loop.

\item[-] Then, we add $\textit{rm}(\textit{empty})$ by using $\textit{PQ}_3$.
\end{itemize}

### 2017-04-18 co-regular-EPQ.tex
add files

@@
 
-A rule $R$ of extended priority queue is co-regular, if checking linearizability with respect to $\textit{MS}(R)$ of a data-independent implementation $\mathcal{I}$ can be reduced to checking the emptiness of intersection between $\mathcal{I}$ and a set of witness automata.
-In this section, we propose the definition of witness automata and co-regular.
-Then we prove that all five rules of extended priority queues are co-regular, and roughly introduce the proof idea.
-With the help of step-by-step linearizability and co-regular, we finally reduce the linearizability problem of $\textit{EPQ}$ into emptiness problem of intersection with automata.
+In this section, we use the notation co-regular to reduce checking linearizability w.r.t $\textit{MS}(R)$ into checking the emptiness of intersection between implementation and a set of witness automata.
+With the help of step-by-step linearizability and co-regular, we finally reduce the linearizability problem of $\textit{EPQ}$ into state reachability problem.
 
-\subsection{Definition of Co-Regular}
+\subsection{Witness Automata and Co-Regular}
 \label{subsec:definition of co-regular}
 
-A witness automaton is a finite automaton with alphabet $\{ \textit{cal}(\textit{put},d,\textit{pred}), \textit{ret}(\textit{put},d),$ $\textit{cal}(\textit{rm},d),\textit{ret}(\textit{rm},d) \vert d \in \mathbb{D} \cup \{ \textit{empty} \},\textit{pred} \in \textit{predWA} \}$.
-Here $\textit{predWA}$ is the set of predicate of priorities, and it contains
+$\textit{con-check-EPQ}$ inspires us that, for data-differentiated executions, checking linearizability with respect to $\textit{EPQ}$ is equivalent to checking linearizability with respect to $\textit{MS}(R)$ for everyone of its projections, as stated by the following lemma.
 
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] A predicate variable $p$, which accepts some specific priority,
+The set of projections of $e$ is denoted $\textit{proj}(e)$. When refer to $\textit{proj}(e)$, we implicitly assume that each $\textit{rm}(\textit{empty})$ in $e$ has a ghost argument that is unique.
 
-\item[-] A predicate $\textit{les}_p$, which accepts all the priorities that are smaller than the value of $p$ according to $<_{\mathbb{P}}$,
+\begin{restatable}{lemma}{EPQasMultiInMRpriforHistory}
+\label{lemma:EPQ as multi in MRpri for history}
+Given a data-differentiated execution $e$, $e \sqsubseteq \textit{EPQ}$, if and only if, $\forall e' \in \textit{proj}(e)$ and $\forall R \in \textit{last}(e')$, we have $e' \sqsubseteq \textit{MS}(R)$.
+\end{restatable}
 
-\item[-] A predicate $\textit{anyPri}$, which accepts any priority.
-\end{itemize}
-
-Given a execution $e = \alpha_1 \cdots \alpha_k$ of extended priority queue and a witness automaton $\mathcal{A}$, we say that $e$ is accepted by $\mathcal{A}$, if
+A witness automaton is a finite automaton with priority variable.
+It uses alphabet $\{ \textit{cal}(\textit{put},d,\textit{pred}), \textit{ret}(\textit{put},d),$ $\textit{cal}(\textit{rm},d),\textit{ret}(\textit{rm},d) \vert d \in \mathbb{D} \cup \{ \textit{empty} \},\textit{pred} \in \textit{predWA} \}$.
+Here $\textit{predWA} = \{p, \textit{les}_p, \textit{anyPri}\}$, where $p$ is a predicate variable, $\textit{les}_p$ represents priorities that are smaller than the value of $p$, and $\textit{anyPri}$ represents any priority.
+Given an execution $e = \alpha_1 \cdots \alpha_k$ of extended priority queue and a witness automaton $\mathcal{A}$, we say that $e$ is accepted by $\mathcal{A}$, if
 
@@
 
-\item[-] For each $i$, if $\alpha_i = \textit{cal}(\textit{put},a,q)$, then either (1) $\beta_i = \textit{cal}(\textit{put},a,p)$ and $q = d_p$, or (2) $\beta_i = \textit{cal}(\textit{put},a,\textit{les}_p)$ and $q <_{\mathbb{P}} d_p$, or (3) $beta_i = \textit{cal}(\textit{put},a,\textit{anyPri})$.
+\item[-] For each $i$, if $\alpha_i = \textit{cal}(\textit{put},a,q)$, then either (1) $q = d_p$ and $\beta_i = \textit{cal}(\textit{put},a,p)$, or (2) $q <_{\mathbb{P}} d_p$ and $\beta_i = \textit{cal}(\textit{put},a,\textit{les}_p)$, or (3) $\beta_i = \textit{cal}(\textit{put},a,\textit{anyPri})$.
 
@@
 
-Note that witness automata does not read operations, since the domains of operations is infinite.
-
-Let us introduce the notion of co-regular:
+Note that due to data-independence, witness automata are used to detect some renamed violations, instead of store and checking directly.
+Let us introduce the notion of co-regular, which reduce checking linearizable w.r.t $\textit{MS}(R)$ into checking set of witness automata.
 
@@
 \begin{definition}\label{def:co-regular of rules of extended priority queues}
-A rule $R$ of extended priority queue is co-regular, if there are a finite set $\textit{Auts}_{R}$ of witness automata such that, for each data-independence implementation $\mathcal{I}$, we have that
+Given $R \in \{ \textit{EPQ}_1^{>}, \textit{EPQ}_1^{=}, \textit{EPQ}_2^{>}, \textit{EPQ}_2^{=}, \textit{EPQ}_3 \}$.
+$R$ is co-regular, if there are a finite set $\textit{Auts}_{R}$ of witness automata such that, for each data-independence implementation $\mathcal{I}$, we have that
 
-$$\textit{Auts}_{R} \cap \mathcal{I} \neq \emptyset \Leftrightarrow \exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), R \in last(e') \wedge e' \ does \ not \ linearizable \ w.r.t.
-\ \textit{MS}(R)$$
+$$\textit{Auts}_{R} \cap \mathcal{I} \neq \emptyset \Leftrightarrow \exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), R \in last(e') \wedge e' \not\sqsubseteq \textit{MS}(R)$$
 
-We say that $\textit{EPQ}$ is co-regular, if each of its rule is co-regular.
+We say that $\textit{EPQ}$ is co-regular, if all five candidates of $R$ are co-regular.
 \end{definition}
 
-Before we go to investigate co-regular of each rules, we use the results in \cite{Bouajjani:2015} to simplify our work.
-\cite{Bouajjani:2015} states that checking linearizability w.r.t queue can be reduced into checking emptiness of intersection between $\mathcal{I}$ and a set of automata.
-Given a data-differentiated execution $e$, let $e \vert_{i}$ be an execution generated from $e$ by erasing call and return actions of items that does not use priority $i$ (does not influence $\textit{rm}(\textit{empty})$).
-We call a extended priority queue execution with only one priority a single-priority execution.
+We simply the proof of co-regular of $\textit{EPQ}$ in two side.
+The first side uses the results in \cite{Bouajjani:2015} to ensure FIFO (first in first out) property of each single-priority projections.
+A execution where only items of one priority is putted is called a single-priority execution.
+It is obvious that each single-priority projection of sequences in $\textit{EPQ}$ satisfy FIFO property.
 Let $\textit{transToQueue}(e)$ be an execution generated from $e$ by transforming $\textit{put}$ and $\textit{rm}$ into $\textit{enq}$ and $\textit{deq}$, respectively, and then discarding priorities.
-We can see that for each $e \in \textit{EPQ}$ and each priority $i$, $\textit{transToQueue}(e \vert_{i})$ satisfy FIFO (first in first out) property.
-
-Given an execution of queue, we say that it is differentiated \cite{Wolper:1986}, if each item is enqueued at most once.
-\cite{Bouajjani:2015} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $a,b$: (1) $\textit{deq}(b) <_{hb} \textit{enq}(b)$, (2) there are no $\textit{enq}(b)$ and at least one $\textit{deq}(b)$, (3) there are one $\textit{enq}(b)$ and more than one $\textit{deq}(b)$, and (4) $\textit{enq}(a) <_{\textit{hb}} \textit{enq}(b)$, and $\textit{deq}(b) <_{\textit{hb}} \textit{deq}(a)$, or $\textit{deq}(a)$ does not exists.
-For each such case, we can construct a witness automata for extended priority queue.
-For example, for the first case, we generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ in \figurename~\ref{fig:automata for FIFO-1}, here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}), \textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a)$, $\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},b),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.6 \textwidth]{figures/PIC_AUTO_FIFO_1.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{SinPri}}^1$}
-  \label{fig:automata for FIFO-1}
-\end{figure}
-
-In Appendix \ref{sec:appendix proof and definition in section definition of co-regular}, we construct a set $\textit{Auts}_{\textit{sinPri}}$ of witness automata ($\mathcal{A}_{\textit{SinPri}}^1$ is in it), and shows that they are enough to ensure that for each data-differentiated executions, each of its single-priority projection without $\textit{rm}(\textit{empty})$ to have ``FIFO'' property, as shown by the following lemma.
+In Appendix \ref{sec:appendix proof and definition in section definition of co-regular}, according to the result of queue in \cite{Bouajjani:2015}, we construct a set $\textit{Auts}_{\textit{sinPri}}$ of witness automata, and shows that they are enough to ensure that for each data-differentiated executions, each of its single-priority projection without $\textit{rm}(\textit{empty})$ to have ``FIFO'' property, as shown by the following lemma.
 
@@
 
-\subsection{Co-Regular of $\textit{EPQ}_1^{>}$}
-\label{subsec:co-regular of EPQ1Lar}
-
-In this subsection, we introduce the idea for proving co-regular of $\textit{PQ}_1^{>}$.
-The proof of this subsection can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
-The notion of left-right constraint used for extended priority queue is inspired by left-right constraint of queue \cite{Bouajjani:2015}.
-
+The second side is to ignore the incomparable priorities.
 Given a data-differentiated execution $e$, we say that $e$ is a $\textit{pri}$-execution, if $\textit{pri}$ is the maximal priority of $e$, and $\textit{pri}$ is larger than or equal to the priority of all other items of $e$.
 Given a data-differentiated execution $e$ and assume that $\textit{pri}$ is one of the maximal priority of $e$, let $\textit{pri-Exec}(e)$ be an execution obtained from $e$ by erasing all operations of items whose priority are incomparable with $\textit{pri}$.
-The following lemma states that when checking co-regular of $\textit{EPQ}_1^{>}$, $\textit{EPQ}_1^{=}$, $\textit{EPQ}_2^{>}$ and $\textit{EPQ}_2^{=}$, we need only consider $\textit{pri}$-executions.
+The following lemma states that when checking co-regular of $\textit{EPQ}$, it is safe to consider $\textit{pri}$-executions.
+We can see that $\textit{last}(e')$ contains at most one element for $\textit{pri}$-execution $e'$.
 
@@
 
-Lemma \ref{lemma:pri execution is enough} simplifies our proof by make us safely ignore all the items that has priorities incomparable with $\textit{pri}$.
-When construct witness automata, priorities which are incomparable with $\textit{pri}$ can be safely represented by $\textit{anyPri}$.
-Given a $\textit{pri}$-execution $e'$, we can see that $\textit{last}(e')$ contains only one rules.
+\subsection{Co-Regular of $\textit{EPQ}_1^{>}$}
+\label{subsec:co-regular of EPQ1Lar}
 
-Given a data-differentiated $\_$-execution $e$ such that $\textit{last}(e) = \textit{EPQ}_1^{>}$, although Lemma \ref{lemma:automata for extended priority queue with single priority} ensures that each single-priority projection of $e$ satisfy the FIFO property, this is still not enough for ensuring that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$.
-Since it is possible that $e$ does not linearizable w.r.t $\textit{MS}(\textit{EPQ}_1^{>})$ because of interaction between actions of multiple priorities.
+In this subsection, we briefly introduce the idea for proving co-regular of $\textit{PQ}_1^{>}$.
+The proof of this subsection can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
+The notion of left-right constraint used for extended priority queue is inspired by left-right constraint of queue \cite{Bouajjani:2015}.
 
-We give an example of such execution $e$ in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
+Given a data-differentiated $\_$-execution $e$ such that $\textit{last}(e) = \textit{EPQ}_1^{>}$, Lemma \ref{lemma:automata for extended priority queue with single priority} is not enough for ensuring that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$.
+This is because that it is possible that $e \not\sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$ because of interaction between actions of multiple priorities.
+One such example is shown in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
 We call the time interval from $\textit{ret}(\textit{put},x)$ to $\textit{cal}(rm,x)$, or from $\textit{ret}(\textit{put},x)$ when $\textit{cal}(rm,x)$ does not exist, the interval of item $x$.
@@
 Here we assume that $p_1 <_{\mathbb{P}} p_4$, $p_2 <_{\mathbb{P}} p_4$ and $p_3 <_{\mathbb{P}} p_4$.
-The reason of why $e$ does not linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{>})$ is that, each time point from $\textit{cal}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is in interval of some item with smaller priority.
+The reason of $e \not\sqsubseteq \textit{MS}(\textit{PQ}_1^{>})$ is that, each time point from $\textit{cal}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is in interval of some item with smaller priority.
 
@@
   \centering
-  \includegraphics[width=0.6 \textwidth]{figures/PIC-HIS-INTRO-GAP-EPQ1L.pdf}
+  \includegraphics[width=0.4 \textwidth]{figures/PIC-HIS-INTRO-GAP-EPQ1L.pdf}
 %\vspace{-10pt}
@@
 
-In this subsection, intuitively we will prove that, as long as we can get rid of the case in \figurename~\ref{fig:introduce gap for EPQ1Lar}, we can ensure $\textit{last}(e) = \textit{EPQ}_1^{>} \Rightarrow e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$.
-
-Let us introduce the notion of left-right constraint, which is a graph, and the existence of cycle though item with maximal priority in it correspond to the existence of the case in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
-
-\begin{definition}\label{def:left-right constraint for matched put and rm operations}
-Given a data-differentiated $\textit{pri}_x$-execution $e$ without $\textit{rm}(\textit{empty})$.
-Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be the only  method events of $e$ with priority $\textit{pri}_x$.
-The left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is the graph $G$ where:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] the nodes are the items of $e$, to which we add a node,
-
-\item[-] there is an edge from item $d_1$ to $x$, if $\textit{put}(d_1,\_)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$,
-
-\item[-] there is an edge from $x$ to item $d_1$, if $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$,
-
-\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$.
-\end{itemize}
-\end{definition}
-
-When there is a cycle $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ through item with maximal priority (for example, $x$) in $G$, we say that $x$ is covered by $d_1,\ldots,d_m$.
-To state the effectiveness of left-right constraint, take the execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} as an example, we can see that $b$ is covered by $f,e,d$.
-We need to prove that getting rid of cycle though item with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\textit{MS}(\textit{EPQ}_1^{>})$, as stated by the following lemma:
+In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we introduce the notion of left-right constraint, which is a graph, and the existence of cycle though item with maximal priority in it formally modelled the case in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
+When there is a cycle $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ through item $x$ with maximal priority in left-right constraint, we say that $x$ is covered by $d_1,\ldots,d_m$ For example, in \figurename~\ref{fig:introduce gap for EPQ1Lar}, $b$ is covered by $f$, $e$ and $d$.
+We then prove that getting rid of cycle though item with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\textit{MS}(\textit{EPQ}_1^{>})$, as stated by the following lemma:
 
@@
 
-\begin {proof} (Sketch)
-
-The $\textit{only if}$ direction can be easily proved by contradiction and is omitted here.
-To prove the $\textit{if}$ direction, we need to explicitly construct the linearization of $e$, or we can say, we need construct the $u$, $v$ and $w$ in $\textit{EPQ}_1^{>}$.
-Let $u$ to be the sequence of all operations that happens before $\textit{put}(x)$.
-The difficulties is how to generate proper $v$.
-
-To generate $v$, we introduce $\textit{UVSet}(e,x)$, which intuitively contains all pairs of method events that should be putted before $\textit{rm}(x)$.
-Let $\textit{UVSet}_1(e,x)= \{ o \vert$ either $o <_{\textit{hb}} \textit{put}(x)$ or $\textit{rm}(x)$, or $\exists o'$ with the same item of $o$, such that $o' <_{\textit{hb}} \textit{put}(x)$ or $\textit{rm}(x)\}$.
-For each $i \geq 1$, let $\textit{LMSet}_{\textit{i+1}}(e,x) = \{ o \vert$ $o \notin \textit{UVSet}_k(e,x)$ for each $k \leq i$, and either $o$ happens before some operation $o' \in \textit{UVSet}_i(e,x)$, or $\exists o''$ with the same item of $o$ and $o''$ happens before some operation $o' \in \textit{UVSet}_i(e,x)\}$.
-Let $\textit{UVSet}(e,x) = \textit{UVSet}_1(e,x) \cup \textit{UVSet}_2(e,x) \cup \ldots$.
-For example, assume that $p_1 <_{\mathbb{P}} p_2$, then in \figurename~\ref{fig:his nobound of UVSet}, $\textit{UVSet}_1(e,x) = \{ \textit{put}(a,p_1),\textit{rm}(a) \}$, $\textit{USSet}_2(e,x) = \{ \textit{put}(b,p_1),\textit{rm}(b) \}$ and $\textit{UVSet}_3(e,x) = \{ \textit{put}(c,p_1),\textit{rm}(c) \}$.
-Similarly, we can generate execution $e'$ such that, for each $i$, $\textit{UVSet}(e',x) \neq \emptyset$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5 \textwidth]{figures/PIC_HIS_NOBOUNDOF_UVSET.pdf}
-%\vspace{-10pt}
-  \caption{An example for $\textit{UVSet}(e,x)$}
-  \label{fig:his nobound of UVSet}
-\end{figure}
-
-We let the $v$ to be the sequence of method events that are in $\textit{UVSet}(e,x)$ and not in $u$.
-We let $w$ to be the sequence of remaining method events.
-Then we can prove that such $u$, $v$ and $w$ holds as required.
-\qed
-\end {proof}
-
-According to Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar}, the existence of ``An item covered by items of smaller priority'' is enough for checking violation of linearizability w.r.t $\textit{MS}(\textit{EPQ}_1^{>})$.
-Assume $x$ is covered by $d_1,\ldots,d_m$.
-Then we can safely rename $x$ into $b$, rename $d_1,\ldots,d_m$ into $a$ and rename all other item into $c$ by data-independence.
+Assume that $x$ is covered by $d_1,\ldots,d_m$, we can safely rename $x$ into $b$, rename $d_1,\ldots,d_m$ into $a$ and rename all other item into $c$ by data-independence.
 Such execution can be recognized by witness automata, since between the first $\textit{ret}(\textit{put},a,\_)$ and the last $\textit{cal}(\textit{rm},a)$ (if exists), $\textit{ret}(\textit{put},a,\_)$ and $\textit{cal}(\textit{rm},a)$ occurs in pair and there is no need for count.
-
-There are four possible enumeration of call and return actions of $\textit{put}(b)$ and $\textit{rm}(b)$.
-For each of them, we generate a witness automaton.
-For example, for the case when $e \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^1$, as shown in \figurename~\ref{fig:automata APQ1Lar-1}.
-Here $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{cal}(\textit{put},a,\textit{les}_p)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d), \textit{ret}(\textit{rm},d),\textit{cal}(\textit{rm},\textit{empty}), \textit{ret}(\textit{rm},\textit{empty})$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{figures/PIC_AUTO_PQ1Lar-pprr.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{l-lar}}^1$}
-  \label{fig:automata APQ1Lar-1}
-\end{figure}
-
-Given the execution $e$ in \figurename~\ref{fig:introduce gap for EPQ1Lar}, let $e'$ be generated from $e$ by renaming $d,e,f$ into $a$.
-Then we can see that $e'$ is accepted by $\mathcal{A}_{\textit{l-lar}}^1$ with a path via $q_{\textit{init}}, q_2, q_3, q_4,q_5,q_7,q_8$.
-
-In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we construct a set $\textit{Auts}_{\textit{1-lar}}$ of witness automata ($\mathcal{A}_{\textit{l-lar}}^1$ is in it), and use $\textit{Auts}_{\textit{1-lar}}$ to prove that $\textit{EPQ}_1^{>}$ is co-regular, as stated by the following lemma.
+In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we construct a set $\textit{Auts}_{\textit{1-lar}}$ of witness automata, and use $\textit{Auts}_{\textit{1-lar}}$ to prove that $\textit{EPQ}_1^{>}$ is co-regular, as stated by the following lemma.
 
@@
 
-In this subsection, we introduce the idea for proving co-regular of $\textit{EPQ}_1^{=}$.
+In this subsection, we briefly introduce the idea for proving co-regular of $\textit{EPQ}_1^{=}$.
 The proof of this subsection can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}.
@@
 Given a data-differentiated $\_$-execution $e$ such that $\textit{last}(e) = \textit{EPQ}_1^{=}$.
-Lemma \ref{lemma:automata for extended priority queue with single priority} ensures that each single-priority projection of $e$ satisfy the FIFO property, and with the help of Lemma \ref{lemma:EPQ1Lar is co-regular} and $\textit{Auts}_{\textit{1-lar}}$, we can ensure that each item in $e$ can not be covered by items with smaller priority.
-However, these are sill not enough for ensure that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
-This is because that given items $a$ and $b$ with maximal priority, it is possible that all the possible linearization point of $\textit{rm}(b)$ may be disabled by $\textit{rm}(b)$.
-
+Lemma \ref{lemma:automata for extended priority queue with single priority} and $\textit{Auts}_{\textit{1-lar}}$ are sill not enough for ensure that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
+This is because that given items $a$ and $b$ with maximal priority, it is possible that all the possible linearization point of $\textit{rm}(b)$ are disabled by $\textit{rm}(b)$.
 We give an example of such execution $e$ in \figurename~\ref{fig:introduce pb order}.
-The execution $e$ of \figurename~\ref{fig:introduce pb order} is not linearizable, even if $h \vert_{p_1}$ and $h \vert_{p_4}$ are both linearizable, and either $a$ or $b$ is covered by items with smaller priority.
-To explain this, me need to observe the following three points:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Since $\textit{put}(a,p_4)$ happens before $\textit{put}(b,p_4)$, we know that $a$ is ``putted earlier'' than $b$, and therefore, the linearization points of $\textit{rm}(a)$ should before the linearization point of $\textit{rm}(b)$.
-
-\item[-] The time intervals satisfy the following conditions (we identify them with dotted lines in \figurename~\ref{fig:introduce pb order}) are possible position to locate the linearization point of $\textit{rm}(b)$ according to Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar} (Here we temporarily forget the existence of $a$): (1) between $\textit{cal}(\textit{rm},b)$ and $\textit{ret}(\textit{rm},b)$, (2) does not in interval of any item with smaller priority.
-
-\item[-] To satisfy requirement of the second condition, the linearization points of $\textit{rm}(b)$ should be in the time interval of dotted line.
-However, all time point in dotted line is before $\textit{cal}(\textit{rm},a)$ and is thus disabled by $\textit{cal}(\textit{rm},a)$.
-\end{itemize}
+The execution $e$ of \figurename~\ref{fig:introduce pb order} is not linearizable w.r.t $\textit{EPQ}$, even if $h \vert_{p_1}$ and $h \vert_{p_4}$ both does, and either $a$ or $b$ is covered by items with smaller priority.
+The explanation of $e \not\sqsubseteq \textit{EPQ}$ is as follows: Since $\textit{put}(a,p_4) <_{\textit{hb}} \textit{put}(b,p_4)$, the linearization points of $\textit{rm}(a)$ should before the linearization point of $\textit{rm}(b)$.
+According to Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar}, the time intervals identified with dotted lines in \figurename~\ref{fig:introduce pb order} are are possible position to locate the linearization point of $\textit{rm}(b)$.
+However, each of them is before $\textit{cal}(\textit{rm},a)$.
 
@@
 
-In this subsection, intuitively we will prove that, as long as we can get rid of the case in \figurename~\ref{fig:introduce pb order}, we can ensure $\textit{last}(e) = \textit{EPQ}_1^{=} \Rightarrow e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
-
-Let us introduce $<_{\textit{pb}}$ (represents put-before) order to formally state that ``an item is putted before another item''.
-Given a data-differentiated $\_$-execution $e$ and two items $a,b$ with maximal priority in $e$, we say that $a <_{\textit{pb}} b$, if one of the following cases holds:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$ in $e$,
-
-\item[-] $\textit{rm}(a)$ happens before $\textit{rm}(b)$ in $e$,
-
-\item[-] $\textit{rm}(a)$ happens before $\textit{put}(b,\_)$ in $e$,
-\end{itemize}
-
+Let us introduce put-before order $<_{\textit{pb}}$ to formally state that ``an item is putted before another item''.
+Given a data-differentiated $\_$-execution $e$ and two items $a,b$ with maximal priority, $a <_{\textit{pb}} b$, if one of the following cases holds: (1) $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$, (2) $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(b)$, or (3) $\textit{rm}(a) <_{\textit{hb}} \textit{put}(b,\_)$.
 Sometimes we use $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$ and $a <_{\textit{pb}}^C b$ to explicitly distinguish above three cases.
 Let $<_{\textit{pb}}^*$ be the transitive closure of $<_{\textit{pb}}$.
-Intuitively, $a <_{\textit{pb}}^* b$ means that we can infer that $a$ should be inserted earlier than $b$ with the help of possibly other items.
+Intuitively, $a <_{\textit{pb}}^* b$ means that we can infer that $a$ should be putted earlier than $b$ with the help of several other items.
 
@@
 \begin{definition}\label{def:gap-point for matched put and rm operations}
-Given a data-differentiated $\textit{pri}_x$-execution $e$ and two method events $\textit{put}(x,\textit{pri}_x),\textit{rm}(x)$ of $e$. We say that a time-point $o$ is a gap-point of $x$, if
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $o$ is after $\textit{call}(\textit{put},x,\textit{pri}_x)$ and $\textit{call}(\textit{rm},x)$, and is before $\textit{ret}(\textit{rm},x)$.
-
-\item[-] $o$ is not in interval of any items with smaller priority.
-\end{itemize}
+Given a data-differentiated $\textit{pri}_x$-execution $e$ and two method events $\textit{put}(x,\textit{pri}_x),\textit{rm}(x)$ of $e$.
+We say that a time-point $o$ is a gap-point of $x$, if $o$ is after $\textit{call}(\textit{put},x,\textit{pri}_x)$ and $\textit{call}(\textit{rm},x)$, is before $\textit{ret}(\textit{rm},x)$, and is not in interval of any items with smaller priority.
 \end{definition}
 
-Going back to \figurename~\ref{fig:introduce pb order}, we can effectively characterize it using $<_{\textit{pb}}$ order and gap-point as follows: $a <_{\textit{pb}} b$, while the right-most gap-point of $b$ is before $\textit{cal}(\textit{rm},a)$.
-We need to prove that getting rid of such case is enough fo ensure linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{=})$, as stated by the following lemma.
-Let $\textit{Items}(e,p)$ be the set of items with priority $p$ in execution $e$.
+The case of \figurename~\ref{fig:introduce pb order} can be formally described as follows: $a <_{\textit{pb}}^* b$, while the right-most gap-point of $b$ is before $\textit{cal}(\textit{rm},a)$ or $\textit{cal}(\textit{put},a,p_4)$.
+The following lemma states that getting rid of such case is enough for ensure linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{=})$.
 
@@
 Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{=}$.
-$e$ does not linearizable w.r.t $\textit{MS}(\textit{EPQ}_1^{=})$, if and only if there exists $x$ and $y$ with maximal priority $\textit{pri}$ in $e$, such that $y <_{\textit{pb}}^* x$ in $e$, and the rightmost gap-point of $x$ is before $\textit{cal}(\textit{put},y,\textit{pri})$ or $\textit{cal}(\textit{rm},y)$ in $e$.
+$e \not\sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$, if and only if there exists $x$ and $y$ with maximal priority $\textit{pri}$ in $e$, such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is before $\textit{cal}(\textit{put},y,\textit{pri})$ or $\textit{cal}(\textit{rm},y)$ in $e$.
 \end{restatable}
@@
 We have already intuitively explain the proof of the $\textit{if}$ direction.
+To prove the $\textit{only if}$ direction, we prove its contrapositive.
+%, where we already know that $\forall x, y$ with maximal priority, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{cal}(\textit{put},y,\textit{pri})$ and $\textit{cal}(\textit{rm},x)$.
+We need to prove that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
+%Or we can say, we need to explicitly construct linearization of $e$.
 
-To prove the $\textit{only if}$ direction, we prove its contrapositive.
-Assume we already know that for each $x$ and $y$ has maximal priority in $e$, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{cal}(\textit{put},y,\textit{pri})$ and $\textit{cal}(\textit{rm},x)$.
-We need to prove that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
-Or we can say, we need to explicitly construct linearization of $e$.
-
-We introduce another lemma (Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}), which states that: If $\exists g_1 \in \textit{Items}(e,\textit{pri})$, such that $\forall g_2 \in \textit{Items}(e,\textit{pri})$, (1) $g_1$ does not $<_{\textit{pb}}$ to $g_2$, and (2) the right-most gap-point of $g_1$ is after $\textit{cal}(\textit{put},g_2,\textit{pri})$ and $\textit{cal}(\textit{rm},g_2)$.
+Let $\textit{Items}(e,p)$ be the set of items with priority $p$ in $e$.
+We introduce another lemma (Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal} in Appendix), which states that: If $\exists g_1 \in \textit{Items}(e,\textit{pri})$, such that $\forall g_2 \in \textit{Items}(e,\textit{pri})$, (1) $g_1 \nless_{\textit{pb}} g_2$, and (2) the right-most gap-point of $g_1$ is after $\textit{cal}(\textit{put},g_2,\textit{pri})$ and $\textit{cal}(\textit{rm},g_2)$.
 Then $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
-The proof of this lemma also tell us how to construct linearization in such case.
-
-Since every single-priority projection has FIFO property, let $l_{\textit{pri}}$ be the linearization of projection of $e$ into actions of priority $\textit{pri}$. Then our proof proceeds as follows:
+%The proof of this lemma also tell us how to construct linearization in such case.
+The remaining of our proof is a loop for searching $g_1$ in $e$.
+Initially we set $a_1$ to be the last inserted item in the $l_{\textit{pri}}$, linearization of the projection of $e$ into $\textit{pri}$.
 
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] We start with $a_1$, the last inserted item of $l_{\textit{pri}}$.
-
-\item[-] Step $1$: check if $a_1$ satisfy the conditions of \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}.
-
-    It is obvious that $a_1$ satisfy the first condition.
-If the second condition is also satisfied, by Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, we can obtain that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
+%\item[-] We start with $a_1$, the last inserted item in the linearization of the projection of $e$ into $\textit{pri}$.
+\item[-] check if $a_1$ satisfy the conditions of Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}. If it does, then this loop terminates.
 
 \item[-] Otherwise, $\exists a_2 \in \textit{Items}(e,\textit{pri})$, such that the rightmost gap-point of $a_1$ is before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$ in $e$.
-We can see that each gap-point of $a_2$ is after the rightmost gap-point of $a_1$.
-By assumption, we know that $a_2$ does not $<_{\textit{pb}}$ to $a_1$.
+By assumption, we know that $a_2 \nless_{\textit{pb}} a_1$, and each gap-point of $a_2$ is after the rightmost gap-point of $a_1$.
 
@@
     \setlength{\itemsep}{0.5pt}
-    \item[-] If $\forall b \in \textit{Items}(e,\textit{pri})$, $a_2$ does not $<_{\textit{pb}}$ to $b$. Then we go to step $1$ and treat $a_2$ similarly as $a_1$.
-    \item[-] Otherwise, there exists $a_3$ with priority $\textit{pri}$ such that $a_2 <_{\textit{pb}}^* a_3$.
-
-    Since $l_{\textit{pri}}$ has FIFO property, it is easy to see that there is no cycle in $<_{\textit{pb}}$ order.
-It is safe to assume that $a_3$ is maximal in the sense of $<_{\textit{pb}}^*$.
-Or we can say, there does not exists $a_4$, such that $a_3 <_{\textit{pb}}^* a_4$.
+    \item[-] If $\forall b \in \textit{Items}(e,\textit{pri})$, $a_2$ does not $<_{\textit{pb}}$ to $b$. Then we go to next round of the loop and make $a_2$ to be the ``new-round $a_1$''.
+    \item[-] Otherwise, there exists $a_3 \in \textit{Items}(e,\textit{pri})$ such that $a_2 <_{\textit{pb}}^* a_3$.
+Since $l_{\textit{pri}}$ has FIFO property, there is no cycle in $<_{\textit{pb}}$, and it is safe to assume that $a_3$ is maximal w.r.t $<_{\textit{pb}}^*$.
 
@@
 Therefore, we can see that the rightmost gap-point of $a_3$ is after the rightmost gap-point of $a_1$.
-Then we go to step $1$ and treat $a_3$ similarly as $a_1$.
+Then we go to next round of the loop and make $a_3$ to be the ``new-round $a_1$''.
     \end{itemize}
@@
 
-Let $a^i$ be the $a_1$ in the $\textit{i-th}$ loop of our proof.
-It is not hard to see that, given $i<j$, the rightmost gap-point of $a^j$ is after the rightmost gap-point of $a^i$.
-Therefore, the loop finally stop at some $a^f$.
-$a^f$ satisfies the check of Step $1$.
-By Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, this implies that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
-This completes the proof of $\textit{if}$ direction.
+Let $a^i$ be the $a_1$ in the $\textit{i-th}$ round of loop.
+We can see that the rightmost gap-point of $a^j$ is after the rightmost gap-point of $a^i$ for each $i<j$.
+Therefore, the loop finally terminates at some $a^f$ that satisfies the condition of Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}.
+This implies that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
 \qed
@@
 
-The result of Lemma \ref{lemma:EPQ1Equal as pb order and gap-point} is not quite suitable for verificaion with automata.
-The reason is that we need to ensure $y <_{\textit{pb}}^* x$, while according to definition of $<_{\textit{pb}}^*$, there may be arbitrary intermediate items $a_1,\ldots,a_m$, such that $x <_{\textit{pb}} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}} a_m <_{\textit{pb}} y$.
-It is hard to store unbounded $a_i$ by automata.
-Fortunately, we find that the number of intermediate items $a_i$ is in fact bounded, as stated by the following lemma.
+The result of Lemma \ref{lemma:EPQ1Equal as pb order and gap-point} is not quite suitable for verificaion with automata, because obtaining $y <_{\textit{pb}}^* x$ may requires arbitrary many intermediate items, and it is hard to store unbounded items by automata.
+Fortunately, we prove that the number of intermediate items $a_i$ is essentially bounded, as stated by the following lemma.
 
@@
 
-With Lemma \ref{lemma:ob order has bounded length}, now we can use witness automata to detect $a <_{\textit{pb}}^* b$ by enumerating all possible enumerations of $a$, $b$ and an intermediate item $a_1$, and then check if it satisfy one cases of \ref{lemma:ob order has bounded length}.
-
-The number of potential enumerations can be further reduced.
-The reason is that Lemma \ref{lemma:EPQ1Equal as pb order and gap-point} requires both $<_{\textit{pb}}^*$ and gap-point, while some combination of $<_{\textit{pb}}^*$ and gap-points are conflict.
-For example, if $a <_{\textit{pb}}^B b$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$, then since (1) $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(b)$, (2) each gap-point of $b$ is after $\textit{cal}(\textit{rm},b)$ (also $\textit{ret}(\textit{rm},a)$), we can see that the right-most gap-point of $b$ is before $\textit{cal}(\textit{put},a,\_)$ and after $\textit{ret}(\textit{rm},a)$, which implies $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a,\_)$ and is impossible.
-Therefore, we finally reduce the number of potential enumeration into only five, as shown by the following lemma:
+By Lemma \ref{lemma:ob order has bounded length}, we can use witness automata to detect $a <_{\textit{pb}}^* b$ by enumerating all possible enumerations of operations of $a$, $b$ and possibly $a_1$.
+Since some combinations of $<_{\textit{pb}}^*$ and gap-points are conflict, we finally reduce the number of potential enumerations of operations of $a$, $b$ and possibly $a_1$ into only five, as shown by the following lemma:
 
@@
 
-The five enumerations is shown in \figurename~\ref{fig:five enumerations}, where we use $o$ to explicitly denote the right-most gap-point of $b$.
-In this figure, $e_1$ and $e_2$ comes from $a <_{\textit{pb}}^A b$, $e_3,e_4,e_5$ comes from $a <_{\textit{pb}}^A a_1 <_{\textit{pb}}^B b$, while other reasons of $a <_{\textit{pb}}^* b$ turns out to be either redundant or impossible.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1 \textwidth]{figures/PIC-HIS-FiveEnumerations.pdf}
-%\vspace{-10pt}
-  \caption{Five possible enumerations}
-  \label{fig:five enumerations}
-\end{figure}
-
-For each enumeration in \figurename~\ref{fig:five enumerations}, we can generate several automata to capture it.
-For example, to capture the first enumeration, we generate witness automaton $\mathcal{A}_{\textit{l-eq}}^1$, as shown in \figurename~\ref{fig:automata for first enumeration of PQ1Equal not in appendix}.
-Here we rename the items that ``covers the time interval from $\textit{cal}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$'' (see Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}) into $d$, and rename the renaming items into $e$.
-In this figure, $c = \textit{cal}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{cal}(\textit{rm},e), \textit{ret}(\textit{rm},e),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{figures/PIC_AUTO_PQ1Equ-1.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{l-eq}}^1$}
-  \label{fig:automata for first enumeration of PQ1Equal not in appendix}
-\end{figure}
-
-In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}, we construct a set $\textit{Auts}_{\textit{1-eq}}$ of witness automata ($\mathcal{A}_{\textit{l-eq}}^1$ is in it), and use $\textit{Auts}_{\textit{1-eq}}$ to prove that $\textit{EPQ}_1^{=}$ is co-regular, as stated by the following lemma.
+In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}, we construct a set $\textit{Auts}_{\textit{1-eq}}$ of witness automata, and prove that $\textit{EPQ}_1^{=}$ is co-regular, as stated by the following lemma.
 
@@
 
+A enumeration of Lemma \ref{lemma:five enumeration is enough for EPQ1Equal} and its witness automata is shown in \figurename~\ref{fig:an enumeration and its witness automaton}.
+$o$ is the rightmost gap-point of $b$, we rename the items that ``covers the time interval from $\textit{cal}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$'' (see Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}) into $d$, and rename the renaming items into $e$.
+In this figure, $c = \textit{cal}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{cal}(\textit{rm},e), \textit{ret}(\textit{rm},e),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.8 \textwidth]{figures/PIC-Enumeration-WitnessAutomata.pdf}
+%\vspace{-10pt}
+  \caption{An enumerations and its witness automaton}
+  \label{fig:an enumeration and its witness automaton}
+\end{figure}
+
 \subsection{Combine Step-by-Step linearizability and Co-Regular}
@@
 
-We also prove that $\textit{EPQ}_2^{>}$, $\textit{EPQ}_2^{=}$ and $\textit{EPQ}_3$ are co-regular.
-The case of $\textit{EPQ}_2^{>}$ is quite simple, since whenever $\textit{last}(e) = \textit{EPQ}_2^{>}$, $e \sqsubseteq \textit{MS}(\textit{EPQ}_2^{>})$.
-For the case of $\textit{EPQ}_2^{=}$, we find a necessary and sufficient condition for violating $\textit{MS}(\textit{EPQ}_2^{>})$, and shows that such condition violates the assumptions that every single-priority execution is FIFO and can be safely ignored.
-For the case of $\textit{EPQ}_3$, it can be similarly proved as the rule for $\textit{deq}(\textit{empty})$ in \cite{Bouajjani:2015}.
-We leave the detailed proof of co-regular of these three rules in Appendix \ref{subsec:appendix co-regular of EPQ2Lar}, Appendix \ref{subsec:appendix co-regular of EPQ2Equal} and Appendix \ref{subsec:co-regular of EPQ3}, respectively.
-
-Now we can see that $\textit{EPQ}$ is co-regular, which is a direct consequence of co-regular of each of its rules, and is stated below.
+We also prove that $\textit{EPQ}_2^{>}$, $\textit{EPQ}_2^{=}$ and $\textit{EPQ}_3$ are co-regular in Appendix \ref{subsec:appendix co-regular of EPQ2Lar}, Appendix \ref{subsec:appendix co-regular of EPQ2Equal} and Appendix \ref{subsec:co-regular of EPQ3}, respectively.
+Then we can see that $\textit{EPQ}$ is co-regular.
 
@@
 
-Since we have already prove that $\textit{EPQ}$ is step-by-step linearizability and co-regular, we can now reduce the verification of linearizability with respect to $\textit{EPQ}$ into a reachability problem, as illustrated by the following theorem.
-Here $\textit{Auts}_{\textit{EPQ}}$ is the set of witness automata of this section.
+By Lemma \ref{lemma:EPQ as multi in MRpri for history} and Lemma \ref{lemma:EPQ is co-regular}, we can finally reduce the verification of linearizability w.r.t $\textit{EPQ}$ into a reachability problem, as illustrated by the following theorem, where $\textit{Auts}_{\textit{EPQ}}$ is the set of witness automata of this section.
 

### 2017-04-18 appendix-EPQ.tex
change appendix

@@
 
-Let us model extended priority queue as an LTS $\textit{LTS}_e = (Q,\Sigma,\rightarrow,q_0)$ as follows:
+Let us model priority queue as an LTS $\textit{LTS}_e = (Q,\Sigma,\rightarrow,q_0)$ as follows:
 
@@
     \item[-] $q_1 \xrightarrow{\textit{rm}(a)} q_2$, if $q_1$ maps $p$ into $l \cdot a$ for some finite sequence $l$, and $q_2$ is the same as $q_1$, except for $p$, where it maps $p$ into $l$.
-We also require that for each priority $p'$ such that $p' <_{\mathbb{P}} p$, $q_1$ and $q_2$ map $p'$ into $\epsilon$.
+We also require that for each priority $p'$ such that $p' \prec p$, $q_1$ and $q_2$ map $p'$ into $\epsilon$.
 
@@
 A trace of an LTS is a finite sequence $\beta_1 \cdot \beta_2 \cdot \ldots \cdot \beta_k$, where $k \geq 0$ if there exists a path $q_0\overset{\beta_1}{\longrightarrow}q_1\overset{\beta_2}{\longrightarrow}\ldots\overset{\beta_k}{\longrightarrow}q_k$ of the LTS.
-Let $\textit{EPQ}_s$ be the set of traces of $\textit{LTS}_e$.
-The following lemma states that the language generated by our rules equals the set of traces of $\textit{EPQ}_s$.
+Let $\textit{PQ}_s$ be the set of traces of $\textit{LTS}_e$.
+The following lemma states that $\textit{PQ} = \textit{PQ}_s$.
 
@@
 
-To prove that $\textit{EPQ} \subseteq \textit{EPQ}_s$, we prove that each sequence in $\textit{EPQ}$ is also in $\textit{EPQ}_s$ by induction:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] It is obvious that $\epsilon \in \textit{EPQ}_s$.
-
-\item[-] If $l_1 \in \textit{EPQ}_s$ and $l_1 \xrightarrow{\textit{EPQ}_1} l_2$.
-Then we need to prove that $l_2 \in \textit{EPQ}_s$.
-Let $l_1 = u \cdot v \cdot w$, such that $\textit{Guard}(u,v,w,\textit{item},\textit{pri})$ of $\textit{EPQ}_1$ holds, and $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm})$.
+To prove that $\textit{PQ} \subseteq \textit{PQ}_s$, we first prove that $\textit{PQ}_{\neq} \subseteq \textit{PQ}_{s\neq}$ by induction.
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] It is obvious that $\epsilon \in \textit{PQ}_s$.
+
+\item[-] If $l_1 \in \textit{PQ}_{s\neq}$ and $l_2 \xrightarrow{\textit{PQ}_1} l_1$.
+Then we need to prove that $l_2 \in \textit{PQ}_s$.
+We know that $l_1 = u \cdot v \cdot w$, such that $\textit{PQ}_1(u,v,w,p)$ holds, and $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v \cdot \textit{rm}(\textit{itm})$.
 
     Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$, $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_j$ and $w = \alpha_{\textit{j+1}} \cdot \ldots \cdot \alpha_m$.
-Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_j} q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{EPQ}_s$.
-For each $i \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $\textit{pri}$ into $\textit{item} \cdot l_k$ and $q_k$ maps $\textit{pri}$ into $l_k$ for some finite sequence $l_k$.
-
-    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(\textit{itm},\textit{pri})} q'_i$.
-Since (1) all $\textit{put}$ with priority $\textit{pri}$ is in $u$, and (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $\textit{pri}$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq j$) with a newest $\textit{itm}$ with priority $\textit{pri}$.
+Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_j} q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}_s$.
+For each $i \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $p$ into $\textit{item} \cdot l_k$ and $q_k$ maps $\textit{pri}$ into $l_k$ for some finite sequence $l_k$.
+
+    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(\textit{itm},p)} q'_i$.
+Since (1) all $\textit{put}$ with priority $p$ is in $u$, and (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $p$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq j$) with a newest $\textit{itm}$ with priority $p$.
 Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j$ are transitions of $\textit{LTS}_e$.
-Since $\textit{matched-C}(u \cdot v)$ holds, we can see that $q_j$ maps each priority that is smaller than $\textit{pri}$ into $\epsilon$ and maps $\textit{pri}$ into $\epsilon$, and $q'_j$ maps each priority that is smaller than $\textit{pri}$ into $\epsilon$ and maps $\textit{pri}$ into $\textit{itm}$.
+Since $\textit{matched}_{\prec}(u \cdot v,p)$ holds, we can see that $q_j$ maps each priority that is smaller than $p$ into $\epsilon$ and maps $p$ into $\epsilon$, and $q'_j$ maps each priority that is smaller than $p$ into $\epsilon$ and maps $p$ into $\textit{itm}$.
 Then, we can see that $q'_j \xrightarrow{\textit{rm}(\textit{itm})} q_j$.
 We already know that that $q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$.
-Therefore, we can see that $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \in \textit{EPQ}_s$.
-
-\item[-] If $l_1 \in \textit{EPQ}_s$ and $l_1 \xrightarrow{\textit{EPQ}_2} l_2$.
-Then we need to prove that $l_2 \in \textit{EPQ}_s$.
-Let $l_1 = u \cdot v$, such that $\textit{Guard}(u,v,\textit{item},\textit{pri})$ of $\textit{EPQ}_2$ holds, and $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v$.
+Therefore, we can see that $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v \cdot \textit{rm}(\textit{itm}) \in \textit{PQ}_s$.
+
+\item[-] If $l_1 \in \textit{PQ}_{s\neq}$ and $l_2 \xrightarrow{\textit{PQ}_2} l_1$.
+Then we need to prove that $l_2 \in \textit{PQ}_s$.
+We know that $l_1 = u \cdot v$, such that $\textit{PQ}_2(u,v,p)$ holds, and $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v$.
 
     Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$.
-Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{EPQ}_s$.
-For each $i \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $\textit{pri}$ into $\textit{item} \cdot l_k$ and $q_k$ maps $\textit{pri}$ into $l_k$ for some finite sequence $l_k$.
-
-    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that$q_i \xrightarrow{\textit{put}(\textit{itm},\textit{pri})} q'_i$.
-Since (1) all $\textit{put}$ with priority $\textit{pri}$ is in $u$, (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $\textit{pri}$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq m$) with a newest $\textit{itm}$ with priority $\textit{pri}$.
+Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}_s$.
+For each $i \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $p$ into $\textit{item} \cdot l_k$ and $q_k$ maps $p$ into $l_k$ for some finite sequence $l_k$.
+
+    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(\textit{itm},p)} q'_i$.
+Since (1) all $\textit{put}$ with priority $p$ is in $u$, (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $p$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq m$) with a newest $\textit{itm}$ with priority $p$.
 Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q'_m$ are transitions of $\textit{LTS}_e$.
-Therefore, we can see that $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{EPQ}_s$.
-
-\item[-] If $l_1 \in \textit{EPQ}_s$ and $l_1 \xrightarrow{\textit{EPQ}_3} l_2$.
-Then we need to prove that $l_2 \in \textit{EPQ}_s$.
-Let $l_1 = u \cdot v$, such that $\textit{Guard}(u,v)$ of $\textit{EPQ}_3$ holds, and $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v$.
+Therefore, we can see that $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v \in \textit{PQ}_s$.
+
+\item[-] If $l_1 \in \textit{PQ}_{s\neq}$ and $l_2 \xrightarrow{\textit{PQ}_3} l_1$.
+Then we need to prove that $l_2 \in \textit{PQ}_s$.
+We know that $l_1 = u \cdot v$, such that $\textit{PQ}_3(u,v)$ holds, and $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v$.
 
     Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$.
-Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{EPQ}_s$.
+Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}_s$.
 
     We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$.
-Since $\textit{matched-All}(u)$ holds, we can see that $q_i$ maps each element in $\mathbb{P}$ into $\epsilon$, and then $q_i \xrightarrow{\textit{rm}(\textit{empty})} q_i$.
+Since $\textit{matched}(u)$ holds, we can see that $q_i$ maps each element in $\mathbb{P}$ into $\epsilon$, and then $q_i \xrightarrow{\textit{rm}(\textit{empty})} q_i$.
 We already know that $q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q_m$.
-Therefore, we can see that $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v \in \textit{EPQ}_s$.
-\end{itemize}
-
-To prove that $\textit{EPQ}_s \subseteq \textit{EPQ}$, we show that given $l_2 \in \textit{EPQ}_s$ and $l_2 \neq \epsilon$, how to construct a sequence $l_1$, such that $l_1 \xrightarrow{R} l_2$ for some rule $R$, and $l_1 \in \textit{EPQ}_s$.
-Based on this, we can decompose a sequence of $\textit{EPQ}_s$ into $\epsilon$, while the reverse process is the reason of why this sequence is in $\textit{EPQ}$.
+Therefore, we can see that $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v \in \textit{PQ}_s$.
+\end{itemize}
+
+To prove that $\textit{PQ} \subseteq \textit{PQ}_s$.
+It is easy to see that $\textit{PQ}$ and $\textit{PQ}_s$ are both data-independent.
+Given $e_1 \in \textit{PQ}$, by data-independence we know that there is a data-differentiated sequential execution $e' \in \textit{PQ}$ and renaming function $f$, such that $e=f(e')$.
+By $\textit{PQ}_{\neq} \subseteq \textit{PQ}_{s\neq}$, we know that $e' \in \textit{PQ}_{s\neq}$.
+By data-independence, we know that $e \in \textit{PQ}_s$.
+
+To prove that $\textit{PQ}_s \subseteq \textit{PQ}$, let us first prove $\textit{PQ}_{s\neq} \subseteq \textit{PQ}_{\neq}$.
+We show that given $l_2 \in \textit{PQ}_{s\neq}$, how to construct a sequence $l_1$, such that $l_2 \xrightarrow{R} l_1$ for some $R$, and $l_1 \in \textit{PQ}_s$.
+Based on this, we can decompose a sequence of $\textit{PQ}_s$ into $\epsilon$, and this process ensures that this sequence is in $\textit{PQ}$.
 Note that from a $l_2$ we may construct more than one $l_1$, and this does not influence the correctness of our proof.
@@
 \item[-] If $l_2$ contains $\textit{rm}(\textit{empty})$: Assume that $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v$.
-It is easy to see that $\textit{matched-All}(u)$ holds.
+It is easy to see that $\textit{matched}(u)$ holds.
 Let $l_1 = u \cdot v$.
-It is easy to see that $l_1 \xrightarrow{\textit{EPQ}_3} l_2$, and $l_1$ satisfy the guard of $\textit{EPQ}_3$.
+It is easy to see that $l_2 \xrightarrow{\textit{PQ}_3} l_1$, and $\textit{PQ}_3(u,v)$ holds.
 
@@
 Since We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{rm}(\textit{empty})} q'_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is transitions of $\textit{LTS}_e$.
-Since $\textit{matched-All}(u)$ holds, we can see that $q_i = q'_i$, and they map each element in $\mathbb{P}$ into $\epsilon$.
-Then we can see that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is transitions of $\textit{LTS}_e$, and $l_1 \in \textit{EPQ}_s$.
-
-\item[-] If $l_2$ does not contain $\textit{rm}(\textit{empty})$, $\textit{pri}$ is one of maximal priority of $l_2$, and items in $l_2$ with priority $\textit{pri}$ are unmatched $\textit{put}$ and (possibly) matched $\textit{put}$:
-
-    Assume that $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v$, such that all $\textit{put}$ with priority $\textit{pri}$ of $u \cdot v$ is in $u$.
+It is easy to see that $q_i = q'_i$, and they map each element in $\mathbb{P}$ into $\epsilon$.
+Then we can see that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is transitions of $\textit{LTS}_e$, and $l_1 \in \textit{PQ}_s$.
+
+\item[-] If $l_2$ does not contain $\textit{rm}(\textit{empty})$, $p$ is one of maximal priority of $l_2$, and items in $l_2$ with priority $p$ are unmatched $\textit{put}$ and (possibly) matched $\textit{put}$:
+
+    Assume that $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v$, such that all $\textit{put}$ with priority $p$ of $u \cdot v$ is in $u$.
 Let $l_1 = u \cdot v$.
-According to construction of $\textit{LTS}_e$, we can see that $l_1$ satisfies the guard of $\textit{EPQ}_2$, and $l_1 \xrightarrow{\textit{EPQ}_2} l_2$.
+According to construction of $\textit{LTS}_e$, we can see that $\textit{PQ}_2(u,v,p)$ holds, and $l_2 \xrightarrow{\textit{PQ}_2} l_1$.
 
     Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$.
-We already know that $\textit{pa} = q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{put}(\textit{itm},\textit{pri})} q_{\textit{i+}} \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q_m$ are transitions of $\textit{LTS}_e$.
-For each $\textit{i+1} \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $\textit{pri}$ into some $\textit{itm} \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $\textit{pri}$ into $l_k$.
-Since (1) all $\textit{put}$ with priority $\textit{pri}$ of $u \cdot v$ is in $u$ and (2) $\textit{pri}$ is one of maximal priority of $l_2$, it is safe to remove $\textit{itm}$ without influence other transitions of $\textit{pa}$.
+We already know that $\textit{pa} = q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{put}(\textit{itm},p)} q_{\textit{i+}} \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q_m$ are transitions of $\textit{LTS}_e$.
+For each $\textit{i+1} \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $p$ into some $\textit{itm} \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $p$ into $l_k$.
+Since (1) all $\textit{put}$ with priority $p$ of $u \cdot v$ is in $u$ and (2) $p$ is one of maximal priority of $l_2$, it is safe to remove $\textit{itm}$ without influence other transitions of $\textit{pa}$.
 Or we can say, $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q'_m$ are transitions of $\textit{LTS}_e$.
-Therefore, $l_1 \in \textit{EPQ}_s$.
-
-\item[-] If $l_2$ does not contain $\textit{rm}(\textit{empty})$, $\textit{pri}$ is one of maximal priority of $l_2$, and items in $l_2$ with priority $\textit{pri}$ are matched $\textit{put}$:
-
-    Assume that $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w$, such that all $\textit{put}$ with priority $\textit{pri}$ of $u \cdot v \cdot w$ is in $u$.
+Therefore, $l_1 \in \textit{PQ}_s$.
+
+\item[-] If $l_2$ does not contain $\textit{rm}(\textit{empty})$, $p$ is one of maximal priority of $l_2$, and items in $l_2$ with priority $p$ are matched $\textit{put}$:
+
+    Assume that $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w$, such that all $\textit{put}$ with priority $p$ of $u \cdot v \cdot w$ is in $u$.
 Let $l_1 = u \cdot v \cdot w$.
-According to construction of $\textit{LTS}_e$, we can see that $l_1$ satisfies the guard of $\textit{EPQ}_1$.
-We can also see that $l_1 \xrightarrow{\textit{EPQ}_1} l_2$.
+According to construction of $\textit{LTS}_e$, we can see that $\textit{PQ}_1(u,v,w,p)$ holds, and $l_2 \xrightarrow{\textit{PQ}_1} l_1$.
 
@@
 We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{put}(\textit{itm},\textit{pri})} q_{\textit{i+}} \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q_j \xrightarrow{\textit{rm}(\textit{itm})} q_{\textit{j+}} \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$.
-For each $\textit{i+1} \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $\textit{pri}$ into $\textit{itm} \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $\textit{pri}$ into $l_k$.
-Since (1) $\textit{pri}$ is one of maximal priority in $l_2$, (2) $\textit{itm}$ is the newest item with priority $\textit{pri}$ in $l_2$, and (3) $\textit{itm}$ is not removed until $\textit{rm}(\textit{itm})$, we know that whether we keep $\textit{itm}$ or remove it will not influence transitions from $q_{\textit{i+1}}$ to $q_j$.
+For each $\textit{i+1} \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $p$ into $\textit{itm} \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $p$ into $l_k$.
+Since (1) $p$ is one of maximal priority in $l_2$, (2) $\textit{itm}$ is the newest item with priority $p$ in $l_2$, and (3) $\textit{itm}$ is not removed until $\textit{rm}(\textit{itm})$, we know that whether we keep $\textit{itm}$ or remove it will not influence transitions from $q_{\textit{i+1}}$ to $q_j$.
 Then we can see that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ are transitions of $\textit{LTS}_e$.
-Therefore, $l_1 \in \textit{EPQ}_s$.
-\end{itemize}
-
-This completes the proof of this lemma. \qed
+Therefore, $l_1 \in \textit{PQ}_s$.
+\end{itemize}
+
+To prove that $\textit{PQ}_s \subseteq \textit{PQ}$.
+Given $e_1 \in \textit{PQ}_s$, by data-independence we know that there is a data-differentiated sequential execution $e' \in \textit{PQ}_s$ and renaming function $f$, such that $e=f(e')$.
+By $\textit{PQ}_{s\neq} \subseteq \textit{PQ}_{\neq}$, we know that $e' \in \textit{PQ}_{\neq}$.
+By data-independence, we know that $e \in \textit{PQ}$.
+\qed
 \end {proof}

### 2017-04-18 Seq-EPQ-dataInd.tex
preliminaries

@@
-\section{Priority Queue and Data-Independence}
+%!TEX root = draft.tex
+\newcommand{\seqPQ}{\mathsf{SeqPQ}}
+
+\section{The Priority Queue ADT}
 \label{sec:priority queue and data-independence}
 
-A priority queue with partially-ordered priorities (priority queue for short) contains two method: $\textit{put}$ and $\textit{rm}$.
+We consider priority queues whose interface contains two methods $\textit{put}$ and $\textit{rm}$ for adding and respectively, removing a value.
+Each value is assigned with a priority when being added to the data structure (using a call to  $\textit{put}$) and the remove method $\textit{rm}$ removes a value with a minimal priority.
+For generality, we assume that the set of priorities is partially-ordered and that incomparable priorities can be removed in any order.
+In case multiple values are assigned with the same priority, $\textit{rm}$ returns the least recent value (according to a FIFO semantics).
+Also, when the set of values stored in the priority queue is empty, $\textit{rm}$ returns the distinguished value $\textit{empty}$.
+Concurrent priority queues allow the methods $\textit{put}$ and $\textit{rm}$ to be called concurrently from different threads.
+However, every method invocation should give the illusion that it takes place instantaneously at some point between its invocation and its return.
+We formalize (concurrent) executions and implementations in Section~\ref{ssec:exec}, Section~\ref{ssec:semantic_prop} introduces a set of properties satisfied by all the  implementations we are aware of, and Section~\ref{ssec:lin} defines the standard correctness criterion for concurrent ADTs such as the priority queue known as \emph{linearizability}\cite{journals/toplas/HerlihyW90}.
+
+\subsection{Executions}\label{ssec:exec}
+
+We fix a (possibly infinite) set $\mathbb{D}$ of data values, a (possibly infinite) set $\mathbb{P}$ of priorities, a partial order $\prec$ among elements in $\mathbb{P}$, and an infinite set $\mathbb{O}$ of operation identifiers.
+The latter are used to match call and return actions of the same invocation.
+Call actions $\textit{call}_o(\textit{put},a,p)$ and $\textit{call}_o(\textit{rm},a')$ with $a\in \mathbb{D}$, $a'\in \mathbb{D}\cup\{\textit{empty}\}$, $p \in \mathbb{P}$, and $o \in \mathbb{O}$, combine a method name and a set of parameters with an operation identifier.
+The return value of a remove is transformed to an argument value for uniformity~\footnote{Method return values are guessed nondeterministically, and validated at return points.
+This can be handled using the {\tt assume} statements of typical formal specification languages, which only admit executions satisfying a given predicate.}.
+The return actions are denoted in a similar way as $\textit{ret}_o(\textit{put},a,p)$ and respectively, $\textit{ret}_o(\textit{rm},a')$.
+
+An \emph{execution} $e$ is a sequence of call and return actions which satisfy the following well-formedness properties: each return is preceded by a matching call (i.e., having the same operation identifier), and each operation identifier is used in at most one call/return.
+We assume every set of executions is closed under isomorphic renaming of operation identifiers.
+An execution is called \emph{sequential} when no two operations overlap, i.e., each call action is immediately followed by its matching return action, and \emph{concurrent} otherwise.
+To ease the reading, we write a sequential execution as a sequence of $\textit{put}(a,p)$ and $\textit{rm}(a)$ symbols representing a pair of actions $\textit{call}_o(\textit{put},a,p)\cdot \textit{ret}_o(\textit{put},a,p)$ and $\textit{call}_o(\textit{rm},a)\cdot \textit{ret}_o(\textit{rm},a)$, respectively (where $o\in\mathbb{O}$).
+For example, given two priorities $p_1 \prec p_2$, $\textit{put}(a,p_2) \cdot \textit{put}(b,p_1) \cdot \textit{rm}(b)$ is a sequential execution of the priority queue ($\textit{rm}$ returns $b$ because it has smaller priority).
+The set of all the possible sequential executions (obtained by adding and removing arbitrarily values with arbitrary priorities) is denoted by $\seqPQ$.
+
+An implementation $\mathcal{I}$ is a set of executions.
+Implementations represent libraries whose methods are called by external programs.
+In the remainder of this work, we consider only completed executions, where each call action has a corresponding return action.
+This simplification is sound when implementation methods can always make progress in isolation: formally, for any execution $e$ with pending operations, there exists an execution $e'$ obtained by extending $e$ only with the return actions of the pending operations of $e$.
+Intuitively this means that methods can always return without any help from outside threads, avoiding deadlock.
+
+%a finite set $\mathbb{M}$ of methods, and an infinite set $\mathbb{O}$ of operation (identifiers).
+Given $m \in \mathbb{M}$, $x \in \mathbb{D}$, and $o \in \mathbb{O}$, a call action $\textit{call}_o (m,x)$ represents an invocation to method $m$ with argument $x$, while a return action $\textit{ret}_o (m,x)$ represents a response from method $m$ with return value $x$.
+\footnote{Call actions with more than one arguments are similarly defined.} Here $o$ is used to match return actions to their call actions: $\textit{cal}_o (m,x)$ matches $\textit{ret}_{o'} (m',x')$, if $o=o' \wedge m=m'$.
+A sequential execution is a sequence of call and return actions, while each call action is immediately followed by its matching return action.
+Let $\cdot$ be the concatenation of sequences.
+To ease the reading, given a sequential execution $e=\textit{call}_{o_1}(m_1,a_1) \cdot \textit{rm}_{o_1}(m_1,b_1) \cdot \ldots \cdot \textit{call}_{o_n}(m_n,a_n) \cdot \textit{rm}_{o_n}(m_n,b_n)$, when the context is clear, we can write it as $e=m_1(a_1,b_1) \cdot \ldots \cdot m_n(a_n,b_n)$.
+We write $m(a,b)$ as $m(a)$ or $m(b)$, when there is no return value or there is no argument, respectively.
+
+%A priority queue with partially-ordered priorities (priority queue for short) contains two method: $\textit{put}$ and $\textit{rm}$.
 A $\textit{put}$ method has two arguments, while the first argument is an item and the second argument is its priority.
@@
 
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If the priority queue is empty, then $\textit{rm}$ returns $\textit{empty}$.
+%\begin{itemize}
+%\setlength{\itemsep}{0.5pt}
+%\item[-] If the priority queue is empty, then $\textit{rm}$ returns $\textit{empty}$.
+%
+%\item[-] Else, $\textit{rm}$ choose one of minimal priority of items in priority queue, and returns the earliest putted item of this priority.
+Note that if there are more than one incomparable and minimal candidate priorities, then the chosen of priority is arbitrary.
+%\end{itemize}
 
-\item[-] Else, $\textit{rm}$ choose one of minimal priority of items in priority queue, and returns the earliest putted item of this priority.
-Note that if there are more than one incomparable and minimal candidate priorities, then the chosen of priority is arbitrary.
-\end{itemize}
-
-For example, assume that $p_1 \prec p_2$, then $\textit{put}(a,p_2) \cdot \textit{put}(b,p_1) \cdot \textit{rm}(b)$ is a correct behavior of priority queue, since $b$ has smaller priority.
-
-Let us introduce the notion of sequential executions, which modelled the specification of concurrent libraries.
+%Let us introduce the notion of sequential executions, which modelled the specification of concurrent libraries.
 We fix a (possibly infinite) set $\mathbb{D}$ of data values, a finite set $\mathbb{M}$ of methods, and an infinite set $\mathbb{O}$ of operation (identifiers).
@@
 
-Data-independence \cite{Wolper:1986} can be used to effectively handle unbounded data domain.
+\subsection{Semantic Properties of Priority Queues}\label{ssec:semantic_prop}
+
+We define two properties which are satisfied by priority queue implementations and which are important for the results that follow: (1) \emph{data independence}~\cite{conf/popl/Wolper86,conf/tacas/AbdullaHHJR13} states that priority queue behaviors do not depend on the actual values which are added to the queue, and (2) \emph{closure under projection}~\cite{DBLP:conf/icalp/BouajjaniEEH15} states that intuitively, remove operations can return the same values no matter how many other different values are in the queue, assuming they don't have more important priorities.
+
+%Data-independence \cite{Wolper:1986} can be used to effectively handle unbounded data domain.
 In this paper, we slightly modify the notion of data-independence in \cite{Wolper:1986} and propose data-independence for priority queues.
 Let $\_$ denote an element, of which the value is irrelevant.
-A sequential execution $e$ of priority queue is data-differentiated if, for all $d \in \mathbb{D}$, there is at most one $\textit{put}(d,\_)$ in $e$.
-Note that a data-differentiated $e$ may contains more than one items with a same priority.
-The subset of data-differentiated sequential executions of a set $S$ is denoted by $S_{\neq}$.
+An execution $e$ is \emph{data-differentiated} if every value is added at most once, i.e., for each $d \in \mathbb{D}$, $e$ contains at most one action $\textit{call}_o(\textit{put},d,p)$ with $o\in\mathbb{O}$ and $p\in \mathbb{P}$.
+Note that this property concerns only values, a data-differentiated execution $e$ may contain more than one value with the same priority.
+The subset of data-differentiated executions of a set of executions $E$ is denoted by $E_{\neq}$.
 
-A renaming function $r$ for priority queue is a function from $\mathbb{D}$ to $\mathbb{D}$.
-Given a sequential execution $e$, we denote by $r(e)$ the sequential execution obtained from $e$ by replacing every item $x$ by $r(x)$.
-Note that here the renaming functions rename only the items and keep the priorities unchanged.
-The explanation is that renaming items will not influence the executions of program, while renaming priorities may do.
+A renaming function $r$ is a function from $\mathbb{D}$ to $\mathbb{D}$.
+Given an execution $e$, we denote by $r(e)$ the execution obtained from $e$ by replacing every data value $x$ by $r(x)$.
+Note that $r$ renames only the values and keep the priorities unchanged.
+Intuitively, renaming values has no influence on the behavior of the priority queue, contrary to renaming priorities.
 
@@
 \begin{definition}\label{def:priority-value data-independence}
-A set of sequential executions $S$ is data-independent, if:
+A set of executions $E$ is \emph{data independent} iff
 \begin{itemize}
 \setlength{\itemsep}{0.5pt}
-\item[-] for all $e \in S$, there exists $e' \in S'$, and a renaming function $r$, such that $e=r(e')$,
+\item[-] for all $e \in E$, there exists $e' \in E_{\neq}$ and a renaming function $r$, such that $e=r(e')$,
 
-\item[-] for all $e \in S$ and for all renaming $r$, $r(e) \in S$.
+\item[-] for all $e \in E$ and for all renamings $r$, $r(e) \in E$.
 \end{itemize}
@@
 
-In latter section, we shows that with the help of data-independence, we can concentrate on only data-differentiated executions when checking correctness.
+The following lemma is a direct consequence of definitions.
+
+\begin{lemma}
+$\seqPQ$ is data independent.
+\end{lemma}
+
+Beyond the sequential executions, every (concurrent) implementation of the priority queue that we are aware of is data-independent.
+Therefore, from now on, we consider only data-independent implementations.
+This assumption enables a reduction from checking the correctness of an implementation $\mathcal{I}$ to checking the correctness of only its data-differentiated executions in $\mathcal{I}_{\neq}$.
+
+TODO MOTIVATE THE FOLLOWING
+
+The projection $e \vert D$ of an execution $e$ to a set of values $D \subseteq \mathbb{D}$ is obtained from $e$ by erasing all actions with a data value not in $D$.
+We write $e \setminus x$ for the projection $e \vert_{ \mathbb{D} \setminus \{ x \} }$.
+
+\begin{lemma}
+$\seqPQ$ is closed under projection, i.e., for every execution $e\in \seqPQ$ and $D \subseteq \mathbb{D}$, $e \vert D\in \seqPQ$.
+\end{lemma}
+
+\subsection{Linearizability}\label{ssec:lin}
+
+%A (concurrent) execution $e$ is a sequence of call and return actions which satisfy a well-formedness property: every return has a matching call action before it in $e$, and an operation $o$ can be used only twice in $e$, once in a call action, and once in a return action.
+The definition of data-differentiated, renaming function, data-independence and projection extends to (concurrent) executions.
+For instance, an execution is data-differentiated if, for all $d \in \mathbb{D}$, there is at most one $\textit{cal}_{\_}(\textit{put},d,\_)$.
+%
+%An implementation $\mathcal{I}$ is a set of concurrent executions.
+Implementations represent libraries whose methods are called by external programs.
+In the remainder of this work, we consider only completed executions, where each call action has a corresponding return action.
+This simplification is sound when implementation methods can always make progress in isolation \cite{Henzinger:2013}: formally, for any execution $e$ with pending operations, there exists an execution $e'$ obtained by extending $e$ only with the return actions of the pending operations of $e$.
+Intuitively this means that methods can always return without any help from outside threads, avoiding deadlock.
+
+TODO WORK THE FOLLOWING
+
+Given execution $e$, the happen-before relation $<_{\textit{hb}}$ is defined as follows: $o_1 <_{\textit{hb}} o_2$, if the return action of $o_1$ is before the call action of $o_2$ in $e$.
+The happens-before relation are interval orders \cite{DBLP:conf/popl/BouajjaniEEH15}: for distinct $o_1,o_2,o_3,o_4$, if $o_1 <_{\textit{hb}} o_2$ and $o_3 <_{\textit{hb}} o_4$, then either $o_1 <_{\textit{hb}} o_4$, or $o_3 <_{\textit{hb}} o_2$.
+Intuitively, this comes from the fact that concurrent threads share a notion of global time.
+
+Linearizability \cite{journals/toplas/HerlihyW90} is the \emph{de facto} standard correctness condition for concurrent data structures.
+Given concurrent execution $e$ and sequential execution $s$ of length $n$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if they have same set of operations, and there is a bijection $f$ between operations s.t.
+
+\begin{itemize}
+\item[-] if $o_1 <_{\textit{hb}} o_2$, then $f(o_1) <_{\textit{hb}} f(o_2)$,
+
+\item[-] $o$ and $f(o)$ have same call and return actions.
+\end{itemize}
+
+A concurrent execution $e$ is linearizable w.r.t a set $S$ of sequential executions, denoted $e \sqsubseteq S$, if there exists $s \in S$ such that $e \sqsubseteq s$.
+A set of concurrent executions $E$ is linearizable w.r.t $S$, denoted $E \sqsubseteq S$, if $e \sqsubseteq S$ for all $e \in H$.
+
+The following lemma states that with the help of data-independence, it is enough to consider only data-differentiated executions when checking linearizability.
+This is similar to that in \cite{Abdulla:2013}, where they use the notion of data-independence in \cite{Wolper:1986}.
+Thus, in the remainder of the paper, we focus on characterizing linearizability for data-differentiated executions.
+The proof of this lemma can be found in Appendix \ref{sec:appendix in section data-independence of EPQ}.
+
+\begin{restatable}{lemma}{DataDifferentiatedisEnoughforPQ}
+\label{lemma:data differentiated is enough for PQ}
+A data-independent implementation $\mathcal{I}$ is linearizable w.r.t data-independent set $S$ of sequential executions, if and only if $\mathcal{I}_{\neq}$ is linearizable with respect to $S_{\neq}$.
+\end{restatable}

### 2017-04-18 recur-process-concurrent.tex
preliminaries

@@
 For instance, an execution is data-differentiated if, for all $d \in \mathbb{D}$, there is at most one $\textit{cal}_{\_}(\textit{put},d,\_)$.
-
-An implementation $\mathcal{I}$ is a set of concurrent executions.
-Implementations represent libraries whose methods are called by external programs.
-In the remainder of this work, we consider only completed executions, where each call action has a corresponding return action.
-This simplification is sound when implementation methods can always make progress in isolation \cite{Henzinger:2013}: formally, for any execution $e$ with pending operations, there exists an execution $e'$ obtained by extending $e$ only with the return actions of the pending operations of $e$.
-Intuitively this means that methods can always return without any help from outside threads, avoiding deadlock.
 

### 2017-04-18 checking-executions.tex
sequential executions

%!TEX root = draft.tex
\section{Checking Linearizability of Priority Queue Executions}
\label{sec:checking inclusion by recursive procedure}

While checking linearizability of a given execution is known to be NP-complete in general~\cite{journals/siamcomp/GibbonsK97} we show in this section that it is polynomial-time for data-differentiated executions of the priority queue. For pedagogical reasons, Section~\ref{ssec:seq_exec} introduces a polynomial-time recursive procedure for checking whether a data-differentiated \emph{sequential} execution is admitted by the priority queue which is then extended to the concurrent case in Section~\ref{ssec:conc_exec}.

\subsection{Checking Data-Differentiated Sequential Executions}\label{ssec:seq_exec}

Checking whether a data-differentiated sequential execution belongs to $\seqPQ$ could be implemented by checking membership into the language accepted by $LTS??$. The recursive procedure $\textit{Check-PQ}$ outlined in Algorithm~\ref{alg:seq_check} is an alternative to this membership test. Roughly, it selects one or two operations in the input execution, checks whether their return values are correct by ignoring the order between the other operations other than how they are ordered w.r.t. the selected ones, and calls itself recursively on the execution without the selected operations. Since the $\textit{rm}(\textit{empty})$ operations are read-only (they don't affect the state of the queue), they are selected first. Ensuring that an operation $o=\textit{rm}(\textit{empty})$ is correct boils down to checking that every $\textit{put}(a,p)$ operation before $o$ is matched to a $\textit{rm}(a)$ operation which also occurs before $o$. Once there are no more $\textit{rm}(\textit{empty})$ operations, the procedure selects either a $\textit{put}$ operation adding a value with maximal priority which is not removed, or a pair of $\textit{put}$ and $\textit{rm}$ operations adding and removing the same maximal priority value. Their correctness can be checked using a similar argument as for $\textit{rm}(\textit{empty})$ operations.

\begin{algorithm}[t]
\KwIn {A data-differentiated sequential execution $e$}
\KwOut{$\mathsf{true}$ iff $e\in \seqPQ$}

\If {$e = \epsilon$}
{\Return $\mathsf{true}$\;}

\If {$\mathsf{HasEmptyRemoves}(e)$}
{
    \If {$\exists\ o=\textit{rm}(\textit{empty})\in e$ such that $\mathsf{EmptyRemoveSeq}(e,o)$ holds}
    {
        \KwRet $\textit{Check-PQ}(e \setminus o)$\;
    }
    \Else {\KwRet $\mathsf{false}$\;}
}
\ElseIf{$\mathsf{HasUnmatchedMaxPriority}(e)$}
{
    \If {$\exists\ x \in \textit{values}(e)$ such that $\mathsf{UnmatchedMaxPrioritySeq}(e,x)$ holds}
    {
        \KwRet $\textit{Check-PQ}(e \setminus x)$\;
    }
    \Else {\KwRet $\mathsf{false}$\;}
}

\Else
{
    \If {$\exists\ x \in \textit{values}(e)$ such that $\mathsf{MatchedMaxPrioritySeq}(e,x)$ holds}
    {
        \KwRet $\textit{Check-PQ}(e \setminus x)$\;
    }
    \Else {\KwRet $\mathsf{false}$\;}
}
\caption{$\textit{Check-PQ}$}
\label{alg:seq_check}
\end{algorithm}

In formal terms, the operations which are selected depend on the following set of predicates on executions:
{\small
\begin{align*}
\mathsf{HasEmptyRemoves}(e)=\mathsf{true} & \mbox{ iff  $e$ contains a $\textit{rm}(\textit{empty})$ operation} \\
\mathsf{HasUnmatchedMaxPriority}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{unmatched-priorities}(e)$ for a maximal priority} \\
&\hspace{4mm}\mbox{$p\in priorities(e)$}
\end{align*}}
where $\textit{priorities}(l)$, resp., $\textit{unmatched-priorities}(l)$, is the set of priorities occurring in $e$, resp., in $\textit{put}$ operations of $e$ for which there is no $\textit{rm}$ operation removing the same value. We call the latter \emph{unmatched} put operations. A put operation which is not unmatched is called \emph{matched}. For readability, we consider the following syntactic sugar $\mathsf{HasMatchedMaxPriority}(e)=\neg \mathsf{HasEmptyRemoves}(e)\land \neg \mathsf{HasUnmatchedMaxPriority}(e)$.

The predicates defining the correctness of the selected operations are defined as follows:
{\small
\begin{align*}
\mathsf{EmptyRemoveSeq}(e,o)=\mathsf{true} & \mbox{ iff  $e= u\cdot o\cdot v$ and $\textit{matched}(u)$} \\
\mathsf{UnmatchedMaxPrioritySeq}(e,x)=\mathsf{true} & \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v$, $p\not\prec \textit{priorities}(u\cdot v)$, and $p\not\in \textit{priorities}(v)$} \\
\mathsf{MatchedMaxPrioritySeq}(e,x)=\mathsf{true} & \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v\cdot \textit{rm}(x)\cdot w$, $p\not\prec \textit{priorities}(u\cdot v\cdot w)$,} \\
&\hspace{4mm}\mbox{$p\not\preceq \textit{unmatched-priorities}(u\cdot v\cdot w)$, $\textit{matched}_\prec(u\cdot v,p)$,} \\
&\hspace{4mm}\mbox{and $p\not\in \textit{priorities}(v\cdot w)$} 
\end{align*}
}

\vspace{-3mm}
\noindent
where $p\prec \textit{priorities}(e)$ when $p\prec p'$ for some $p'\in \textit{priorities}(e)$ (and similarly for $p\prec \textit{unmatched-priorities}(e)$ or $p\preceq \textit{unmatched-priorities}(e)$), 
$\textit{matched}_\prec(e,p)$ holds when each value with priority strictly smaller than $p$ is removed in $e$, and $\textit{matched}(e)$ holds when $\textit{matched}_\prec(e,p)$ holds for each $p\in\mathbb{P}$.

When $o$ is a $\textit{rm}(\textit{empty})$ operation, we write $e\setminus o$ for the maximal subsequence of $e$ which doesn't contain $o$.

% and items of unmatched put in $l$, respectively. $p \prec \textit{priorities}(l)$ (resp., $p \preceq \textit{priorities}(l)$), if $p \prec p'$ (resp., $p \preceq p'$)for some $p' \in \textit{priorities}(l)$. Similarly we can define $p \preceq \textit{unmatched-priorities}(l)$.
%
%
%Essentially, this procedure chooses either a $\textit{rm}(\textit{empty})$ operation or a set of operations adding/removing a maximal priority in the input execution and checks whether they satisfy certain properties
%
%
% proceeds by 
%
% The projection $e \vert{O}$ of $e$ into $O$ is obtained from $e$ by erasing all call and return actions of non-$O$ operations. We write $e \setminus o$ for the projection $e \vert_{ O_e \setminus \{ o \} }$, where $O_e$ is the set of operations of $e$.
%
%Given a data-differentiated sequential execution $e$, it is costly to check whether $e$ belongs to priority queue, since for each element of $e$, we should keep the current content of priority queue and try to modify the priority queue according to this element. Such process is too complex. In this section, we propose another method for checking inclusion of priority queue by using a recursive procedure $\textit{check-PQ}$. For checking a sequence, every time we only check a much simpler property, and then recursively check the remanning sequences obtained by erasing one or two elements. Each step of our method is easier to deal with.
%
%$\textit{put}(a,p)$ matches $\textit{rm}(b)$, if $a = b$. To introduce $\textit{check-PQ}$, let us introduce several predicates. Given sequential execution $l$ and priority $p$:
%
%\begin{itemize}
%\setlength{\itemsep}{0.5pt}
%\item[-] $\textit{rm}(\textit{empty}) \notin l$ is satisfied when $l$ does not contain $\textit{rm}(\textit{empty})$.
%
%\item[-] Let $\textit{priorities}(l)$ and $\textit{unmatched-priorities}(l)$ be the set of priorities of putted items and items of unmatched put in $l$, respectively. $p \prec \textit{priorities}(l)$ (resp., $p \preceq \textit{priorities}(l)$), if $p \prec p'$ (resp., $p \preceq p'$)for some $p' \in \textit{priorities}(l)$. Similarly we can define $p \preceq \textit{unmatched-priorities}(l)$.
%
%\item[-] $\textit{matched}_{\prec}(l,p)$ is satisfied, if for each item of $l$ with priorities smaller than $p$, it has matched $\textit{put}$ and $\textit{rm}$. Similarly, we can define $\textit{matched}(l)$, where we consider all items in $l$ instead of items with certain priorities.
%\end{itemize}
%
%Let us introduce some notations: given sequential executions $u,v,w$ and priority $p$,
%
%\begin{itemize}
%\setlength{\itemsep}{0.5pt}
%\item[-] $\textit{PQ}_1(u,v,w,p) \equiv
%(\textit{rm}(\textit{empty}) \notin u \cdot v \cdot w) \wedge
%(p \not\prec \textit{priorities}(u \cdot v \cdot w)) \wedge
%(p \not\preceq \textit{unmatched-priorities}(u \cdot v \cdot w)) \wedge
%(\textit{matched}_{\prec}(u \cdot v,p) ) \wedge
%(p \notin \textit{priorities}(v \cdot w))$
%
%\item[-] $\textit{PQ}_2(u,v,p) \equiv
%(\textit{rm}(\textit{empty}) \notin u \cdot v) \wedge
%(p \not\prec \textit{priorities}(u \cdot v \cdot w)) \wedge
%(p \notin \textit{priorities}(v \cdot w))$
%
%\item[-] $\textit{PQ}_3(u,v) \equiv
%\textit{matched}(u \cdot v)$
%\end{itemize}
%
%
%Let us defined set $\textit{MS}(R)$ ($R \in \textit{PQ}_1,\textit{PQ}_2,\textit{PQ}_3$), which is the set of data-differentiated sequential executions that respect $R$.
%
%\begin{itemize}
%\setlength{\itemsep}{0.5pt}
%\item[-] $e \in \textit{MS}(\textit{PQ}_1)$, if $e = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$ and $\textit{PQ}_1(u,v,w,p)$ holds for some item $x$ and a maximal priority $p$,
%
%\item[-] $e \in \textit{MS}(\textit{PQ}_2)$, if $e = u \cdot \textit{put}(x,p) \cdot v$ and $\textit{PQ}_2(u,v,p)$ holds for some item $x$ and a maximal priority $p$,
%
%\item[-] $e \in \textit{MS}(\textit{PQ}_3)$, if $e = u \cdot \textit{rm}(\textit{empty}) \cdot v$ and $\textit{PQ}_3(u,v)$ holds.
%\end{itemize}
%
%We call such $x$ and $\textit{rm}(\textit{empty})$ (if exists) the witness of $e$. Note that sequences in $\textit{MS}(R)$ does not guaranteed to be a priority queue execution. Let us introduce the notion $\textit{last}(e)$, which is a set and used as the branch condition of $\textit{check-PQ}$. Here $e$ is a data-differentiated sequential execution.
%
%\begin{itemize}
%\setlength{\itemsep}{0.5pt}
%\item[-] If $e$ contains $\textit{rm}(\textit{empty})$, then $\textit{last}(e) = \{ \textit{PQ}_3 \}$.
%
%\item[-] Else, if for a maximal priority $p$ of $e$, $\textit{unmatched-priorities}(e) \neq \emptyset$, then $\textit{PQ}_2 \in \textit{last}(e)$.
%
%\item[-] Else, if for a maximal priority $p$ of $e$, $\textit{unmatched-priorities}(e) = \emptyset$, then $\textit{PQ}_1 \in \textit{last}(e)$.
%\end{itemize}
%
%When $\textit{last}(e)$ contains only one element $R$, we write $\textit{last}(w)=R$ for simplicity.
%
%The projection $e \vert{\mathcal{D}}$ of $e$ into $\mathcal{D} \subseteq \mathbb{D}$ is obtained from $e$ by erasing all elements with a data value not in $\mathcal{D}$. We write $e \setminus x$ for the projection $e \vert_{ \mathbb{D} \setminus \{ x \} }$. The projection $e \vert{O}$ of $e$ into $O$ is obtained from $e$ by erasing all call and return actions of non-$O$ operations. We write $e \setminus o$ for the projection $e \vert_{ O_e \setminus \{ o \} }$, where $O_e$ is the set of operations of $e$.
%
%The recursive procedure $\textit{con-check-EPQ}$ is given as follows. Here $P_i(e,o)$ and $P_i(e,x)$ is the predicates of $e \in \textit{MS}(\textit{PQ}_i)$ withe witness $o$ or $x$, respectively.
%
%\begin{minipage}{.5\textwidth}
%\begin{algorithm}[H]
%\KwIn {a data-differentiated sequential execution $e$}
%
%\If {$e = \epsilon$}
%{\KwRet $\textit{true}$;}
%
%\If {$\textit{last}(e) = \textit{PQ}_3$}
%{
%    \If {$\exists o=\textit{rm}(\textit{empty})$, $P_3(e,o)$ holds}
%    {
%        \KwRet $\textit{check-PQ}(e \setminus o)$;
%    }
%    \KwRet $\textit{false}$;
%}
%
%\If {$\textit{PQ}_i \in \textit{last}(e)$ for $i \in \{1,2\}$}
%{
%    \If {$\exists l$, $P_i(e,x)$ holds}
%    {
%        \KwRet $\textit{check-PQ}(e \setminus x)$;
%    }
%    \KwRet $\textit{false}$;
%}
%\caption{$\textit{check-PQ}$}
%\label{Method-check-PQ}
%\end{algorithm}
%\end{minipage}
%\hfill

TODO MODIFY THE FOLLOWING EXAMPLE

Let us use an example to show how $\textit{check-PQ}$ works. Here we use $w_1 \xrightarrow{R} w_2$ to denote that in $\textit{check-PQ}(w_1)$, we choose the branch of $R$ and recursively call $\textit{check-PQ}(w_2)$.

\begin{example}\label{example:generate extended priority queue executions}
Given priorities $p_1,p_2,p_3$ with orders $p_1 \prec p_2$ and $p_1 \prec p_3$,

\noindent $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{rm}(\textit{empty}) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)$

$\xrightarrow{\textit{EPQ}_3}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)$

$\xrightarrow{\textit{EPQ}_1}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(b,p_1)$

$\xrightarrow{\textit{EPQ}_2}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(b,p_1)$

$\xrightarrow{\textit{EPQ}_1}$ $\textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{put}(b,p_1)$

$\xrightarrow{\textit{EPQ}_2}$ $\textit{put}(a,p_1) \cdot \textit{rm}(a)$

$\xrightarrow{\textit{EPQ}_1}$ $\epsilon$
\end{example}

The following lemma states the correctness of $\textit{Check-PQ}$ (the proof can be found in Appendix~\ref{}).

\begin{lemma}
Algorithm~\ref{alg:seq_check} returns $\mathsf{true}$ on an execution $e$ iff $e\in \seqPQ$.
\end{lemma}

%Let $\textit{PQ}$ be set of sequences obtained by renaming some data-differentiated sequential execution that is accepted by $\textit{check-PQ}$. To persuade readers that $\textit{PQ}$ is indeed the set of behaviors of priority queue, in Appendix \ref{sec:appendix in section inductive rules of extended priority queue}, we give a semantical version definition $\textit{PQ}_s$ of priority queue, and shows that $\textit{PQ} = \textit{PQ}_s$. To model the possible behaviors of priority queue, we model it as an labelled transition system (shortened as LTS) $\textit{LTS}_e$. Each state of $\textit{LTS}_e$ is a function from $\mathbb{P}$ into sequences over $\mathbb{D}$, and represents a snapshot of contents of priority queue. $\textit{PQ}_s$ is the set of traces of $\textit{LTS}_e$. The definition of $\textit{PQ}_s$ and the proof of the following lemma can be found in Appendix \ref{sec:appendix in section inductive rules of extended priority queue}.
%
%\begin{restatable}{lemma}{EPQRulesAndSemantics}
%\label{lemma:EPQ rules and semantics}
%$\textit{PQ} = \textit{PQ}_s$.
%\end{restatable}
%
%By Example \ref{example:generate extended priority queue executions} we can see how to generate a sequence $e \in \textit{EP}$ from $\epsilon$ as follows:
%
%\begin{itemize}
%\setlength{\itemsep}{0.5pt}
%\item[-] First we add non-$\textit{rm}(\textit{empty})$ elements with a loop: Let $P$ be the set of priorities of items in $e$. In the first round of the loop, we choose a minimal priorities $p \in P$, we add items by using $\textit{PQ}_1$ and $\textit{PQ}_2$, let $P = P \setminus \{p\}$, and begins the next round of loop.
%
%\item[-] Then, we add $\textit{rm}(\textit{empty})$ by using $\textit{PQ}_3$.
%\end{itemize}

\subsection{Checking Data-Differentiated Concurrent Executions}\label{ssec:conc_exec}

Let us extend $\textit{check-PQ}$ with linearizability to deal with linearizability w.r.t $\textit{PQ}$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] The input $e$ is a concurrent execution.

\item[-] $\textit{last}(e)$ is the union of $\textit{last}(u)$ for any sequential execution $u$ such that $e \sqsubseteq u$. $e \sqsubseteq \textit{MS}(R)$ with witness $x$ (resp., $o$), if $e \sqsubseteq u \in \textit{MS}(R)$ and $x$ (resp., $o$) is the witness of $u$.

\item[-] $P_i(e,o)$ and $P_i(e,x)$ holds, if $e \sqsubseteq \textit{MS}(\textit{PQ}_i)$ withe witness $o$ or $x$, respectively. 
\end{itemize}

The following lemma states that $\textit{check-PQ}$ is correct.

\begin{restatable}{lemma}{ConCheckEPQIsCorrect}
\label{lemma:con-check-EPQ is correct}
Given a data-differentiated execution $e$, $e \sqsubseteq \textit{EPQ}$, if and only if, $\textit{check-PQ} = \textit{true}$.
\end{restatable}

\begin {proof} (Sketch)
It is not hard to see that $\textit{check-PQ}(e) = \textit{false} \Rightarrow e \not\sqsubseteq \textit{PQ}$, since $\textit{PQ}_{\neq} \subset \textit{MS}(R)_{\neq}$. To ensure that $\textit{check-PQ}(e) = \textit{true} \Rightarrow e \sqsubseteq \textit{PQ}$, we require $\textit{PQ}$ to satisfy the following property.

for any data-differentiated execution $e$,
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] if $e \sqsubseteq \textit{MS}(R)$ ($R \in \{ \textit{PQ}_1, \textit{PQ}_2 \}$) with witness $x$, we have: $e \setminus x \sqsubseteq \textit{PQ} \Rightarrow e \sqsubseteq \textit{PQ}$.

\item[-] if $e \sqsubseteq \textit{MS}(\textit{PQ}_3)$ and $o$ is a $\textit{rm}(\textit{empty})$, we have: $e \setminus o \sqsubseteq \textit{PQ} \Rightarrow e \sqsubseteq \textit{PQ}$.
\end{itemize}

With this property, we can build a linearization of a whole execution by increasingly construct linearization of sub-execution from $\epsilon$, and this convince us that $\textit{check-PQ}(e) = \textit{true} \Rightarrow e \sqsubseteq \textit{PQ}$.

Let us briefly explain the idea of proving this property for $\textit{PQ}_1$ with an example. Given a data-differentiated concurrent execution $e \sqsubseteq u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \textit{MS}(\textit{PQ}_1)$ with witness $x$ and assume that $e \setminus x \sqsubseteq l' \in \textit{PQ}$, as shown in \figurename~\ref{fig:concurrent execution for EPQ1} (a), we explicitly construct a sequence $l''= l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3$ and prove that $e \sqsubseteq l'' \in \textit{PQ}$. In this example, $p_1 \prec p$, $p_1 \prec p_2$, $u=\epsilon$, $w$ contains $\textit{rm}(z_2), \textit{put}(x_3,p_2), \textit{rm}(z_1)$ (emphasize by $\textit{rm}(z_2)-w$), and the remanning operations are in $v$. We also explicitly draw the linearization points according to $l'$. We construct $l''$ as follows, and this method does not rely on the locations of linearization points of $l'$ and operations of $x$. Here we implicitly mix operation $o$ (when $\textit{call}_o(m,a)$, $\textit{ret}_o(m,b)$ in $e$) with $m(a,b)$.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Let $p$-comparable operations (resp., $p$-incomparable operations) be the set of operations with items whose priority is comparable with $p$ (resp., incomparable with $p$). It seems that we can construct $l''_1$, $l''_2$ and $l''_3$ as the projection of $l'$ into operations of $u$, $v$ and $w$, respectively. However, this is incorrect, since $\textit{PQ}_1$ has no restriction to $p$-incomparable operations operations in $u \cdot v$, and thus, there is no guarantee that the projection of $l'$ into $p$-incomparable operations operations in $u \cdot v$ being correct. In this example, such projection is $\textit{put}(z_1,p_3) \cdot \textit{rm}(z_2)$ and is incorrect.

\item[-] Let us construct sets $U'$, $V'$ and $W'$, such then make $l''_1$, $l''_2$ and $l''_3$ the projections of $l'$ into $U'$, $V'$ and $W'$, respectively. This contains two steps:

\item[-] The first step is to define $W'$. The $p$-comparable operations in $W'$ is same as that in $w$. To obtain $p$-incomparable operations in $W'$, we try to find a $p$-incomparable operation $o$ which either happens before some $p$-comparable operations in $w$, or with linearization points after $\textit{rm}(x)$. Then, we put $o$ and all $p$-incomparable operations after $o$ in $l'$ into $W'$ (emphasized by boxes in example). In this example, $o$ is $\textit{rm}(z_1)$, and $W'$ contains $\textit{put}(x_3,p)$, $\textit{rm}(z_1)$ and $\textit{rm}(z_2)$. We use boxes to emphasize they are put into $W'$.

\item[-] The second step is to define $U'$ and $V'$. $U'$ contains the following two kinds of operations: (1) Operations whose linearization points are before $\textit{ret}(\textit{put},x)$, and (2) other $\textit{put}$ operations with priority $p$. $V'$ contains the remanning operations. In this example, $U'$ contains $\textit{put}(z_1,p_3)$ and $\textit{put}(x_2,p_2)$.

\item[-] In this example, $l''_1 = \textit{put}(z_1,p_2) \cdot \textit{put}(x_2,p)$, $l''_2 = \textit{put}(z_2,p_2) \cdot \textit{rm}(x_2) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(y_1)$, and $l''_3 = \textit{put}(x_3,p) \cdot \textit{rm}(z_1) \cdot \textit{rm}(z_2)$. In \figurename~\ref{fig:concurrent execution for EPQ1} (b), we add linarization points according to $l''$, and we can see that $l''$ holds as required.
\end{itemize}

This completes the proof of this lemma.

\qed
\end {proof}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1 \textwidth]{figures/PIC-HIS-EPQ1-TwoHis.pdf}
%\vspace{-10pt}
  \caption{The process of obtaining linearization of $e$}
  \label{fig:concurrent execution for EPQ1}
\end{figure}

### 2017-04-18 priority-queues.tex
sequential executions

--- 2017-04-18 Seq-EPQ-dataInd.tex
@@
 For example, given two priorities $p_1 \prec p_2$, $\textit{put}(a,p_2) \cdot \textit{put}(b,p_1) \cdot \textit{rm}(b)$ is a sequential execution of the priority queue ($\textit{rm}$ returns $b$ because it has smaller priority).
-The set of all the possible sequential executions (obtained by adding and removing arbitrarily values with arbitrary priorities) is denoted by $\seqPQ$.
+
+We define $\seqPQ$, the set of sequential priority queue executions semantically via labelled transition system ($LTS$).
+An LTS is a tuple $(Q,\Sigma,\rightarrow,q_0)$, where $Q$ is a set of states, $\Sigma$ is an alphabet of transition labels, $\rightarrow\subseteq Q\times\Sigma\times Q$ is a transition relation and $q_0$ is the initial state.
+We modelled priority queue as an LTS $\textit{LTS}_e$ as follows: Each state of $\textit{LTS}_e$ is a function from $\mathbb{P}$ into a finite sequence over $\mathbb{D}$ and represents a snapshot of priority queue content; The transition label are $\textit{put}$ and $\textit{rm}$ operations; Each transition modified state according to operations.
+For example, $q_1 \xrightarrow{\textit{rm}(\textit{empty})} q_2$, if $q_1 = q_2$, and they maps each element in $\mathbb{P}$ into $\epsilon$.
+Let $\seqPQ$ be the set of traces of $\textit{LTS}_e$.
 
@@
 Linearizability \cite{journals/toplas/HerlihyW90} is the \emph{de facto} standard correctness condition for concurrent data structures.
-Given concurrent execution $e$ and sequential execution $s$ of length $n$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if they have same set of operations, and there is a bijection $f$ between operations s.t.
+Given concurrent execution $e$ and sequential execution $s$ of length $n$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if there is a bijection $f: O_1 \rightarrow O_2$, where $O_1$ and $O_2$ are set of operations of $e$ and $s$, respectively, and $f$ s.t.
 

### 2017-04-18 checking-executions.tex
sequential executions
example for the sequential procedure

@@
 the selected ones, and calls itself recursively on the execution without the selected operations.
+
+We explain how the procedure works on the following execution:
+\begin{align}
+\hspace{-5mm}\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{rm}(\textit{empty}) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)\label{eq:ex_rec1}
+\end{align}
+where $p_1$, $p_2$, $p_3$ are priorities such that $p_1 \prec p_2$ and $p_1 \prec p_3$, and $p_2$ and $p_3$ are incomparable.
 Since the $\textit{rm}(\textit{empty})$ operations are read-only (they don't affect the state of the queue), they are selected first.
-Ensuring that an operation $o=\textit{rm}(\textit{empty})$ is correct boils down to checking that every $\textit{put}(a,p)$ operation before $o$ is matched to a $\textit{rm}(a)$ operation which also occurs before $o$.
-Once there are no more $\textit{rm}(\textit{empty})$ operations, the procedure selects either a $\textit{put}$ operation adding a value with maximal priority which is not removed, or a pair of $\textit{put}$ and $\textit{rm}$ operations adding and removing the same maximal priority value.
-Their correctness can be checked using a similar argument as for $\textit{rm}(\textit{empty})$ operations.
+Ensuring that an operation $o=\textit{rm}(\textit{empty})$ is correct boils down to checking that every $\textit{put}(x,p)$ operation before $o$ is matched to a $\textit{rm}(x)$ operation which also occurs before $o$.
+This is true in this case for $x\in \{a,c\}$.
+Therefore, the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
+\begin{align*}
+\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)
+\end{align*}
+When there are no more $\textit{rm}(\textit{empty})$ operations, the procedure selects either a $\textit{put}$ operation adding a value with maximal priority which is not removed, or a pair of $\textit{put}$ and $\textit{rm}$ operations adding and removing the same maximal priority value.
+For instance, since $p_2$ is a maximal priority, it may select the operation $\textit{put}(d,p_2)$.
+This operation is trivially correct
+and the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of 
+\begin{align*}
+\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)
+\end{align*}
+Now, all the maximal priority values are removed and the procedure selects a pair of $\textit{put}$/$\textit{rm}$ operations, for instance, $\textit{put}(c,p_2)$ and $\textit{rm}(c)$.
+In this case, the value returned by $\textit{rm}$ is correct if all the values of priority smaller than $p_2$ added before $\textit{rm}(c)$ are also removed before $\textit{rm}(c)$.
+In this case, the only value of priority smaller than $p_2$ is $a$ and satisfies this property.
+Applying a similar reasoning for all the remaining values, it can be proved that this execution is correct.
 
@@
 $\textit{matched}_\prec(e,p)$ holds when each value with priority strictly smaller than $p$ is removed in $e$, and $\textit{matched}(e)$ holds when $\textit{matched}_\prec(e,p)$ holds for each $p\in\mathbb{P}$.
+Compared to the example presented at the beginning of the section, these predicates take into consideration that multiple values with the same priority are removed in FIFO order: the predicate $\mathsf{MatchedMaxPrioritySeq}(e,x)$ holds when $x$ is the last value with priority $p$ added in $e$.
 
@@
 
-TODO MODIFY THE FOLLOWING EXAMPLE
-
-Let us use an example to show how $\textit{check-PQ}$ works.
+%TODO MODIFY THE FOLLOWING EXAMPLE
+%
+%Let us use an example to show how $\textit{check-PQ}$ works.
 Here we use $w_1 \xrightarrow{R} w_2$ to denote that in $\textit{check-PQ}(w_1)$, we choose the branch of $R$ and recursively call $\textit{check-PQ}(w_2)$.
-
-\begin{example}\label{example:generate extended priority queue executions}
-Given priorities $p_1,p_2,p_3$ with orders $p_1 \prec p_2$ and $p_1 \prec p_3$,
-
-\noindent $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{rm}(\textit{empty}) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)$
-
-$\xrightarrow{\textit{EPQ}_3}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)$
-
-$\xrightarrow{\textit{EPQ}_1}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(b,p_1)$
-
-$\xrightarrow{\textit{EPQ}_2}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(b,p_1)$
-
-$\xrightarrow{\textit{EPQ}_1}$ $\textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{put}(b,p_1)$
-
-$\xrightarrow{\textit{EPQ}_2}$ $\textit{put}(a,p_1) \cdot \textit{rm}(a)$
-
-$\xrightarrow{\textit{EPQ}_1}$ $\epsilon$
-\end{example}
+%
+%\begin{example}\label{example:generate extended priority queue executions}
+%Given priorities $p_1,p_2,p_3$ with orders $p_1 \prec p_2$ and $p_1 \prec p_3$,
+%
+%\noindent $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{rm}(\textit{empty}) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)$
+%
+%$\xrightarrow{\textit{EPQ}_3}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)$
+%
+%$\xrightarrow{\textit{EPQ}_1}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(b,p_1)$
+%
+%$\xrightarrow{\textit{EPQ}_2}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(b,p_1)$
+%
+%$\xrightarrow{\textit{EPQ}_1}$ $\textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{put}(b,p_1)$
+%
+%$\xrightarrow{\textit{EPQ}_2}$ $\textit{put}(a,p_1) \cdot \textit{rm}(a)$
+%
+%$\xrightarrow{\textit{EPQ}_1}$ $\epsilon$
+%\end{example}
 
@@
 
-\begin{lemma}
+\begin{restatable}{lemma}{EPQRulesAndSemantics}
+\label{lemma:EPQ rules and semantics}
 Algorithm~\ref{alg:seq_check} returns $\mathsf{true}$ on an execution $e$ iff $e\in \seqPQ$.
-\end{lemma}
+\end{restatable}
+
+%
+%\begin{lemma}
+%Algorithm~\ref{alg:seq_check} returns $\mathsf{true}$ on an execution $e$ iff $e\in \seqPQ$.
+%\end{lemma}
 

### 2017-04-18 checking-executions.tex
example for the sequential procedure

@@
 {
-    \If {$\exists\ o=\textit{rm}(\textit{empty})\in e$ such that $\mathsf{EmptyRemoveSeq}(e,o)$ holds}
+    \If {$\exists\ o=\textit{rm}(\textit{empty})\in e$ such that $\mathsf{EmptyRemove\text{-}Seq}(e,o)$ holds}
     {
@@
     }
-    \Else {\KwRet $\mathsf{false}$\;}
+    %\Else {\KwRet $\mathsf{false}$\;}
 }
@@
 {
-    \If {$\exists\ x \in \textit{values}(e)$ such that $\mathsf{UnmatchedMaxPrioritySeq}(e,x)$ holds}
+    \If {$\exists\ x \in \textit{values}(e)$ such that $\mathsf{UnmatchedMaxPriority\text{-}Seq}(e,x)$ holds}
     {
@@
 {
-    \If {$\exists\ x \in \textit{values}(e)$ such that $\mathsf{MatchedMaxPrioritySeq}(e,x)$ holds}
+    \If {$\exists\ x \in \textit{values}(e)$ such that $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ holds}
     {
@@
 \end{align*}}
-where $\textit{priorities}(l)$, resp., $\textit{unmatched-priorities}(l)$, is the set of priorities occurring in $e$, resp., in $\textit{put}$ operations of $e$ for which there is no $\textit{rm}$ operation removing the same value.
+where $\textit{priorities}(l)$, resp., $\textit{unmatched-priorities}(l)$, is the set of priorities occurring in $\textit{put}$ operations of $e$, resp., in $\textit{put}$ operations of $e$ for which there is no $\textit{rm}$ operation removing the same value.
 We call the latter \emph{unmatched} put operations.
@@
 \begin{align*}
-\mathsf{EmptyRemoveSeq}(e,o)=\mathsf{true} & \mbox{ iff  $e= u\cdot o\cdot v$ and $\textit{matched}(u)$} \\
-\mathsf{UnmatchedMaxPrioritySeq}(e,x)=\mathsf{true} & \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v$, $p\not\prec \textit{priorities}(u\cdot v)$, and $p\not\in \textit{priorities}(v)$} \\
-\mathsf{MatchedMaxPrioritySeq}(e,x)=\mathsf{true} & \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v\cdot \textit{rm}(x)\cdot w$, $p\not\prec \textit{priorities}(u\cdot v\cdot w)$,} \\
+\mathsf{EmptyRemove\text{-}Seq}(e,o)=\mathsf{true} & \mbox{ iff  $e= u\cdot o\cdot v$ and $\textit{matched}(u)$} \\
+\mathsf{UnmatchedMaxPriority\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v$, $p\not\prec \textit{priorities}(u\cdot v)$, and $p\not\in \textit{priorities}(v)$} \\
+\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v\cdot \textit{rm}(x)\cdot w$, $p\not\prec \textit{priorities}(u\cdot v\cdot w)$,} \\
 &\hspace{4mm}\mbox{$p\not\preceq \textit{unmatched-priorities}(u\cdot v\cdot w)$, $\textit{matched}_\prec(u\cdot v,p)$,} \\
@@
 When $o$ is a $\textit{rm}(\textit{empty})$ operation, we write $e\setminus o$ for the maximal subsequence of $e$ which doesn't contain $o$.
+For an execution $e$, $\textit{values}(e)$ is the set of values occurring in call/return actions of $e$.
 

### 2017-04-18 checking-executions.tex
concurrent executions

@@
 Checking whether a data-differentiated sequential execution belongs to $\seqPQ$ could be implemented by checking membership into the language accepted by $LTS??$.
-The recursive procedure $\textit{Check-PQ}$ outlined in Algorithm~\ref{alg:seq_check} is an alternative to this membership test.
+The recursive procedure $\textit{Check-PQ-Seq}$ outlined in Algorithm~\ref{alg:seq_check} is an alternative to this membership test.
 Roughly, it selects one or two operations in the input execution, checks whether their return values are correct by ignoring the order between the other operations other than how they are ordered w.r.t.
@@
     {
-        \KwRet $\textit{Check-PQ}(e \setminus o)$\;
+        \KwRet $\textit{Check-PQ-Seq}(e \setminus o)$\;
     }
@@
     {
-        \KwRet $\textit{Check-PQ}(e \setminus x)$\;
+        \KwRet $\textit{Check-PQ-Seq}(e \setminus x)$\;
     }
@@
 }
-\caption{$\textit{Check-PQ}$}
+\caption{$\textit{Check-PQ-Seq}$}
 \label{alg:seq_check}
@@
 
-The following lemma states the correctness of $\textit{Check-PQ}$ (the proof can be found in Appendix~\ref{}).
+The following lemma states the correctness of $\textit{Check-PQ-Seq}$ (the proof can be found in Appendix~\ref{}).
 
@@
 \label{lemma:EPQ rules and semantics}
-Algorithm~\ref{alg:seq_check} returns $\mathsf{true}$ on an execution $e$ iff $e\in \seqPQ$.
+$\textit{Check-PQ-Seq}(e)=\mathsf{true}$ iff $e\in \seqPQ$, for every data-differentiated sequential execution $e$.
 \end{restatable}
@@
 
-Let us extend $\textit{check-PQ}$ with linearizability to deal with linearizability w.r.t $\textit{PQ}$ as follows:
+The extension of $\textit{Check-PQ-Seq}$ to concurrent executions, checking whether they are linearizable w.r.t. $\seqPQ$, is obtained by replacing every predicate $\Gamma\mathsf{\text{-}Seq}$ with
+\begin{align*}
+\Gamma\mathsf{\text{-}Conc}(e,\alpha) = \mathsf{true}\mbox{ iff there exists a sequential execution $s$ such that $e\sqsubseteq s$ and $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$}
+\end{align*}
+for each $\Gamma\in \{\mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$.
+Let $\textit{Check-PQ-Conc}$ denote the thus obtained procedure (we assume recursive calls are modified accordingly).
 
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] The input $e$ is a concurrent execution.
+% with linearizability to deal with linearizability w.r.t $\textit{PQ}$ as follows: 
+%
+%\begin{itemize}
+%\setlength{\itemsep}{0.5pt}
+%\item[-] The input $e$ is a concurrent execution. 
+%
+%\item[-] $\textit{last}(e)$ is the union of $\textit{last}(u)$ for any sequential execution $u$ such that $e \sqsubseteq u$.
+$e \sqsubseteq \textit{MS}(R)$ with witness $x$ (resp., $o$), if $e \sqsubseteq u \in \textit{MS}(R)$ and $x$ (resp., $o$) is the witness of $u$.
+%    
+%\item[-] $P_i(e,o)$ and $P_i(e,x)$ holds, if $e \sqsubseteq \textit{MS}(\textit{PQ}_i)$ withe witness $o$ or $x$, respectively. 
+%\end{itemize}
 
-\item[-] $\textit{last}(e)$ is the union of $\textit{last}(u)$ for any sequential execution $u$ such that $e \sqsubseteq u$.
-$e \sqsubseteq \textit{MS}(R)$ with witness $x$ (resp., $o$), if $e \sqsubseteq u \in \textit{MS}(R)$ and $x$ (resp., $o$) is the witness of $u$.
-
-\item[-] $P_i(e,o)$ and $P_i(e,x)$ holds, if $e \sqsubseteq \textit{MS}(\textit{PQ}_i)$ withe witness $o$ or $x$, respectively. 
-\end{itemize}
-
-The following lemma states that $\textit{check-PQ}$ is correct.
+The following lemma states the correctness of $\textit{Check-PQ-Conc}$.
+Completeness follows easily from the properties of $\seqPQ$.
+Thus, if $\textit{Check-PQ-Conc}(e) = \textit{false}$, then there exists a set of values $D$ such that one of the following doesn't hold: $\mathsf{EmptyRemove\text{-}Conc}(e \vert D)$, $\mathsf{UnmatchedMaxPriority\text{-}Conc}(e \vert D,x)$ for all maximal priority values $x$, or $\mathsf{MatchedMaxPriority\text{-}Conc}(e \vert D,x)$ for all maximal priority values $x$.
+It can be easily seen that we get $e \vert D\not\sqsubseteq \seqPQ$ in all cases, which by the closure under projection of $\seqPQ$ implies, $e \not\sqsubseteq \seqPQ$ (since every linearization of $e$ includes as a subsequence a linearization of $e \vert D$).
 
@@
 \label{lemma:con-check-EPQ is correct}
-Given a data-differentiated execution $e$, $e \sqsubseteq \textit{EPQ}$, if and only if, $\textit{check-PQ} = \textit{true}$.
+$\textit{Check-PQ-Conc}(e)=\mathsf{true}$ iff $e \sqsubseteq \seqPQ$, for every data-differentiated execution $e$.
 \end{restatable}
 
-\begin {proof} (Sketch)
-It is not hard to see that $\textit{check-PQ}(e) = \textit{false} \Rightarrow e \not\sqsubseteq \textit{PQ}$, since $\textit{PQ}_{\neq} \subset \textit{MS}(R)_{\neq}$.
-To ensure that $\textit{check-PQ}(e) = \textit{true} \Rightarrow e \sqsubseteq \textit{PQ}$, we require $\textit{PQ}$ to satisfy the following property.
+Proving soundness is highly non-trivial and one of the main technical contributions of this paper (see Appendix~\ref{} for a complete proof).
+The main technical difficulty is showing that for any execution $e$, any linearization of $e\setminus x$ for some maximal priority value $x$ can be extended to a linearization of $e$ provided that $\mathsf{UnmatchedMaxPriority}$ or $\mathsf{MatchedMaxPriority}$ holds (depending on whether there are values with the same priority as $x$ in $e$ which are not removed).
+%
+%
+%To ensure that $\textit{check-PQ}(e) = \textit{true} \Rightarrow e \sqsubseteq \textit{PQ}$, we require $\textit{PQ}$ to satisfy the following property.  
+%
+%for any data-differentiated execution $e$,
+%\begin{itemize}
+%\setlength{\itemsep}{0.5pt}
+%\item[-] if $e \sqsubseteq \textit{MS}(R)$ ($R \in \{ \textit{PQ}_1, \textit{PQ}_2 \}$) with witness $x$, we have: $e \setminus x \sqsubseteq \textit{PQ} \Rightarrow e \sqsubseteq \textit{PQ}$.
+%
+%\item[-] if $e \sqsubseteq \textit{MS}(\textit{PQ}_3)$ and $o$ is a $\textit{rm}(\textit{empty})$, we have: $e \setminus o \sqsubseteq \textit{PQ} \Rightarrow e \sqsubseteq \textit{PQ}$.
+%\end{itemize}
+%
+%With this property, we can build a linearization of a whole execution by increasingly construct linearization of sub-execution from $\epsilon$, and this convince us that $\textit{check-PQ}(e) = \textit{true} \Rightarrow e \sqsubseteq \textit{PQ}$.
+%
 
-for any data-differentiated execution $e$,
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] if $e \sqsubseteq \textit{MS}(R)$ ($R \in \{ \textit{PQ}_1, \textit{PQ}_2 \}$) with witness $x$, we have: $e \setminus x \sqsubseteq \textit{PQ} \Rightarrow e \sqsubseteq \textit{PQ}$.
-
-\item[-] if $e \sqsubseteq \textit{MS}(\textit{PQ}_3)$ and $o$ is a $\textit{rm}(\textit{empty})$, we have: $e \setminus o \sqsubseteq \textit{PQ} \Rightarrow e \sqsubseteq \textit{PQ}$.
-\end{itemize}
-
-With this property, we can build a linearization of a whole execution by increasingly construct linearization of sub-execution from $\epsilon$, and this convince us that $\textit{check-PQ}(e) = \textit{true} \Rightarrow e \sqsubseteq \textit{PQ}$.
-
-Let us briefly explain the idea of proving this property for $\textit{PQ}_1$ with an example.
+TODO UPDATE THE NOTATIONS IN THIS EXAMPLE
+Let us briefly explain the idea of proving this property when $\mathsf{MatchedMaxPriority}$ holds with an example.
 Given a data-differentiated concurrent execution $e \sqsubseteq u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \textit{MS}(\textit{PQ}_1)$ with witness $x$ and assume that $e \setminus x \sqsubseteq l' \in \textit{PQ}$, as shown in \figurename~\ref{fig:concurrent execution for EPQ1} (a), we explicitly construct a sequence $l''= l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3$ and prove that $e \sqsubseteq l'' \in \textit{PQ}$.
@@
 
-This completes the proof of this lemma.
+%\begin {proof} (Sketch)
+%It is not hard to see that $\textit{check-PQ}(e) = \textit{false} \Rightarrow e \not\sqsubseteq \textit{PQ}$, since $\textit{PQ}_{\neq} \subset \textit{MS}(R)_{\neq}$.
 
-\qed
-\end {proof}
+%This completes the proof of this lemma. 
+%
+%
+%\qed
+%\end {proof}
 

### 2017-04-19 checking-executions.tex
modify

@@
 TODO UPDATE THE NOTATIONS IN THIS EXAMPLE
+
 Let us briefly explain the idea of proving this property when $\mathsf{MatchedMaxPriority}$ holds with an example.
-Given a data-differentiated concurrent execution $e \sqsubseteq u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \textit{MS}(\textit{PQ}_1)$ with witness $x$ and assume that $e \setminus x \sqsubseteq l' \in \textit{PQ}$, as shown in \figurename~\ref{fig:concurrent execution for EPQ1} (a), we explicitly construct a sequence $l''= l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3$ and prove that $e \sqsubseteq l'' \in \textit{PQ}$.
+Assume $e \sqsubseteq l=u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, $\mathsf{MatchedMaxPriority}(l,x)$ holds, and $e \setminus x \sqsubseteq l' \in \textit{PQ}$, as shown in \figurename~\ref{fig:concurrent execution for EPQ1} (a), we explicitly construct a sequence $l''= l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3$ and prove that $e \sqsubseteq l'' \in \textit{PQ}$.
 In this example, $p_1 \prec p$, $p_1 \prec p_2$, $u=\epsilon$, $w$ contains $\textit{rm}(z_2), \textit{put}(x_3,p_2), \textit{rm}(z_1)$ (emphasize by $\textit{rm}(z_2)-w$), and the remanning operations are in $v$.
@@
 It seems that we can construct $l''_1$, $l''_2$ and $l''_3$ as the projection of $l'$ into operations of $u$, $v$ and $w$, respectively.
-However, this is incorrect, since $\textit{PQ}_1$ has no restriction to $p$-incomparable operations operations in $u \cdot v$, and thus, there is no guarantee that the projection of $l'$ into $p$-incomparable operations operations in $u \cdot v$ being correct.
+However, this is incorrect, since $\mathsf{MatchedMaxPriority}(l,x)$ has no restriction to $p$-incomparable operations operations in $u \cdot v$, and thus, there is no guarantee that the projection of $l'$ into $p$-incomparable operations operations in $u \cdot v$ being correct.
 In this example, such projection is $\textit{put}(z_1,p_3) \cdot \textit{rm}(z_2)$ and is incorrect.
@@
 Then, we put $o$ and all $p$-incomparable operations after $o$ in $l'$ into $W'$ (emphasized by boxes in example).
-In this example, $o$ is $\textit{rm}(z_1)$, and $W'$ contains $\textit{put}(x_3,p)$, $\textit{rm}(z_1)$ and $\textit{rm}(z_2)$.
-We use boxes to emphasize they are put into $W'$.
+In this example, $o$ is $\textit{rm}(z_1)$, and $p$-incomparable operations after $o$ in $l'$ is $\textit{rm}(z_2)$, which are emphasized by adding boxes.
+In this process, whether a $p$-incomparable operation is in $W'$ or not only relies whether it is before or after $o$ in $l'$.
 

### 2017-04-19 checking-executions.tex
modify

@@
 the selected ones, and calls itself recursively on the execution without the selected operations.
+{\color {blue} Here we implicitly mix operation $o$ (when $\textit{call}_o(m,a)$, $\textit{ret}_o(m,b)$ in $e$) with $m(a,b)$.}
 
@@
 \end{align*}
-When there are no more $\textit{rm}(\textit{empty})$ operations, the procedure selects either a $\textit{put}$ operation adding a value with maximal priority which is not removed, or a pair of $\textit{put}$ and $\textit{rm}$ operations adding and removing the same maximal priority value.
-For instance, since $p_2$ is a maximal priority, it may select the operation $\textit{put}(d,p_2)$.
-This operation is trivially correct
-and the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
+{\color{blue}When there are no more $\textit{rm}(\textit{empty})$ operations, the procedure try to select a $\textit{put}$ operation adding a value with maximal priority which is not removed, and then a pair of $\textit{put}$ and $\textit{rm}$ operations adding and removing the same maximal priority value.} For instance, since $p_2$ is a maximal priority, it may select the operation $\textit{put}(d,p_2)$.
+{\color {blue}This operation is correct if $d$ is the last value with priority $p_2$,} and the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
 \begin{align*}
@@
 \end{align*}
-Now, all the maximal priority values are removed and the procedure selects a pair of $\textit{put}$/$\textit{rm}$ operations, for instance, $\textit{put}(c,p_2)$ and $\textit{rm}(c)$.
+{\color {blue}Now, all the putted value that is not removed of priority $p_2$ do not exists, and the procedure selects a pair of $\textit{put}$/$\textit{rm}$ operations of priority $p_2$,} for instance, $\textit{put}(c,p_2)$ and $\textit{rm}(c)$.
 In this case, the value returned by $\textit{rm}$ is correct if all the values of priority smaller than $p_2$ added before $\textit{rm}(c)$ are also removed before $\textit{rm}(c)$.
@@
 The following lemma states the correctness of $\textit{Check-PQ-Seq}$ (the proof can be found in Appendix~\ref{}).
+{\color {blue}Therefore, in latter section we can abuse $\seqPQ$ by the set obtained by renaming sequences that are accepted by $Check-PQ-Seq$.}
 
@@
 Completeness follows easily from the properties of $\seqPQ$.
-Thus, if $\textit{Check-PQ-Conc}(e) = \textit{false}$, then there exists a set of values $D$ such that one of the following doesn't hold: $\mathsf{EmptyRemove\text{-}Conc}(e \vert D)$, $\mathsf{UnmatchedMaxPriority\text{-}Conc}(e \vert D,x)$ for all maximal priority values $x$, or $\mathsf{MatchedMaxPriority\text{-}Conc}(e \vert D,x)$ for all maximal priority values $x$.
-It can be easily seen that we get $e \vert D\not\sqsubseteq \seqPQ$ in all cases, which by the closure under projection of $\seqPQ$ implies, $e \not\sqsubseteq \seqPQ$ (since every linearization of $e$ includes as a subsequence a linearization of $e \vert D$).
+Thus, if $\textit{Check-PQ-Conc}(e) = \textit{false}$, then there exists a set of values $D$ such that one of the following doesn't hold: $\mathsf{EmptyRemove\text{-}Conc}(e \vert D)$, {\color {blue}$\mathsf{UnmatchedMaxPriority\text{-}Conc}(e \vert D,x)$ for some maximal priority values $x$, or $\mathsf{MatchedMaxPriority\text{-}Conc}(e \vert D,x)$ for some maximal priority values $x$.} It can be easily seen that we get $e \vert D\not\sqsubseteq \seqPQ$ in all cases, which by the closure under projection of $\seqPQ$ implies, $e \not\sqsubseteq \seqPQ$ (since every linearization of $e$ includes as a subsequence a linearization of $e \vert D$).
 
@@
 We construct $l''$ as follows, and this method does not rely on the locations of linearization points of $l'$ and operations of $x$.
-Here we implicitly mix operation $o$ (when $\textit{call}_o(m,a)$, $\textit{ret}_o(m,b)$ in $e$) with $m(a,b)$.
 

### 2017-04-19 checking-executions.tex
modify

@@
 The following lemma states the correctness of $\textit{Check-PQ-Seq}$ (the proof can be found in Appendix~\ref{}).
-{\color {blue}Therefore, in latter section we can abuse $\seqPQ$ by the set obtained by renaming sequences that are accepted by $Check-PQ-Seq$.}
 
@@
 Completeness follows easily from the properties of $\seqPQ$.
-Thus, if $\textit{Check-PQ-Conc}(e) = \textit{false}$, then there exists a set of values $D$ such that one of the following doesn't hold: $\mathsf{EmptyRemove\text{-}Conc}(e \vert D)$, {\color {blue}$\mathsf{UnmatchedMaxPriority\text{-}Conc}(e \vert D,x)$ for some maximal priority values $x$, or $\mathsf{MatchedMaxPriority\text{-}Conc}(e \vert D,x)$ for some maximal priority values $x$.} It can be easily seen that we get $e \vert D\not\sqsubseteq \seqPQ$ in all cases, which by the closure under projection of $\seqPQ$ implies, $e \not\sqsubseteq \seqPQ$ (since every linearization of $e$ includes as a subsequence a linearization of $e \vert D$).
+Thus, if $\textit{Check-PQ-Conc}(e) = \textit{false}$, then there exists a set of values $D$ such that one of the following doesn't hold: $\mathsf{EmptyRemove\text{-}Conc}(e \vert D)$, {\color {blue}$\mathsf{UnmatchedMaxPriority\text{-}Conc}(e \vert D,x)$ for all values that are unmatched and with maximal priority $x$, or $\mathsf{MatchedMaxPriority\text{-}Conc}(e \vert D,x)$ for all maximal priority values $x$ where values of $x$ are all matched.} It can be easily seen that we get $e \vert D\not\sqsubseteq \seqPQ$ in all cases, which by the closure under projection of $\seqPQ$ implies, $e \not\sqsubseteq \seqPQ$ (since every linearization of $e$ includes as a subsequence a linearization of $e \vert D$).
 

### 2017-04-19 introduction.tex
intro sketch

@@
+%!TEX root = draft.tex
 \section{Introduction}
 \label{sec:introduction}
-I am introduction
+
+Efficient implementations of concurrent objects are hard to get right.
+Verifying them is difficult, checking a single execution is NP-complete~\cite{journals/siamcomp/GibbonsK97} and checking a finite-state implementation is undecidable~\cite{conf/esop/BouajjaniEEH13}.
+
+The set of objects considered in previous work consists of stacks, queues, registers.
+
+Here, we consider another important object, the priority queues, which are used to TODO, and for which numerous implementations exist CITE (research articles and languages/libraries providing priority queue implementations).
+This object is much less studied compared to the other ones in the verification literature.
+
+Priority queues are collections and in that sense, they could be found similar to stacks and queues.
+The main difference which complicates the process of inventing verification techniques is that the order in which elements are removed is not fixed by the happens-before like in stacks and queues but by parameters, the priorities, which come from an unbounded domain.
+TODO an example.
+Say that for generality, we consider partially-ordered priorities and that for single priority we take FIFO.
+In java, same priority values are removed in an arbitrary order, however, this can be modeled using our partial orders.
+TODO SAY HOW
+
+Building on previous work~\cite{DBLP:conf/icalp/BouajjaniEEH15}, we define decision procedures for priority queues. These procedures are designed in several steps:
+1) defining recursive procedures recognizing valid sequences, and extending them to concurrent executions.
+These procedures take values one by one and check some property local to that value, ignoring how operations on other values are ordered between them.
+This works only for data-differentiated, but this is complete by data-independence.
+2) checking whether a fixed value violates that property can be done using a specific class of register automata, TODO say how simple they are
+3) we distinguish between that value being removed or not, and we have to also consider the case of remove(empty). This results in 3-4 automata describing all the possible violations.
+
+Although we follow the same schema as in~\cite{DBLP:conf/icalp/BouajjaniEEH15}, these results require establishing key results which are specific to priority queues and which are not implied by those for stacks and queues (say that more details are given in the related work)
+
+This gives a reduction to reachability that works for arbitrary implementations, and a decision procedure for finite-state implementations
 

### 2017-04-19 co-regular-EPQ.tex
modify

@@
-\section{Reducing Linearizability of Extended Priority Queues into State Reachability}
+\section{Reducing Linearizability of Priority Queues into State Reachability}
 \label{sec:co-regular of extended priority queues}
 
-In this section, we use the notation co-regular to reduce checking linearizability w.r.t $\textit{MS}(R)$ into checking the emptiness of intersection between implementation and a set of witness automata.
-With the help of step-by-step linearizability and co-regular, we finally reduce the linearizability problem of $\textit{EPQ}$ into state reachability problem.
+In this section, we divide checking linearizability w.r.t $\seqPQ$ into checking linearizablity w.r.t $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$, each of which can be done in polynomial-time.
+Due to data-independence, violations to each $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$ can be captured by register automata.
+This reduces checking linearizability w.r.t $\seqPQ$ into state reachability problem.
 
-\subsection{Witness Automata and Co-Regular}
+\subsection{Co-Regular}
 \label{subsec:definition of co-regular}
 
-$\textit{con-check-EPQ}$ inspires us that, for data-differentiated executions, checking linearizability with respect to $\textit{EPQ}$ is equivalent to checking linearizability with respect to $\textit{MS}(R)$ for everyone of its projections, as stated by the following lemma.
-
-The set of projections of $e$ is denoted $\textit{proj}(e)$. When refer to $\textit{proj}(e)$, we implicitly assume that each $\textit{rm}(\textit{empty})$ in $e$ has a ghost argument that is unique.
+Let $\Gamma\mathsf{\text{-}Condition}(e)$ be $\mathsf{HasEmptyRemoves}(e)$, $\neg \mathsf{HasEmptyRemoves}(e) \wedge \mathsf{HasUnmatchedMaxPriority}(e)$, and $\mathsf{HasMatchedMaxPriority}(e)$, for $\Gamma = \mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}$, respectively.
+$\textit{Check-PQ-Conc}$ inspires us that, whenever $\Gamma\mathsf{\text{-}Condition}(e)$ holds, there should exist some $\alpha$, such that  $\Gamma\mathsf{\text{-}Conc}(e,\alpha)$ holds.
+The following lemma states this formally.
+Let $\textit{proj}(e)$ be the set of projections of $e$ into set of values.
+When refer to $\textit{proj}(e)$, we implicitly assume that each $\textit{rm}(\textit{empty})$ in $e$ has a ghost argument that is unique.
 
@@
 \label{lemma:EPQ as multi in MRpri for history}
-Given a data-differentiated execution $e$, $e \sqsubseteq \textit{EPQ}$, if and only if, $\forall e' \in \textit{proj}(e)$ and $\forall R \in \textit{last}(e')$, we have $e' \sqsubseteq \textit{MS}(R)$.
+Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$, if and only if, $\forall e' \in \textit{proj}(e)$, $\Gamma\mathsf{\text{-}Condition}(e) \Rightarrow \Gamma\mathsf{\text{-}Conc}(e,\alpha)$ holds for some $\alpha$
 \end{restatable}
 
-A witness automaton is a finite automaton with priority variable.
-It uses alphabet $\{ \textit{cal}(\textit{put},d,\textit{pred}), \textit{ret}(\textit{put},d),$ $\textit{cal}(\textit{rm},d),\textit{ret}(\textit{rm},d) \vert d \in \mathbb{D} \cup \{ \textit{empty} \},\textit{pred} \in \textit{predWA} \}$.
-Here $\textit{predWA} = \{p, \textit{les}_p, \textit{anyPri}\}$, where $p$ is a predicate variable, $\textit{les}_p$ represents priorities that are smaller than the value of $p$, and $\textit{anyPri}$ represents any priority.
-Given an execution $e = \alpha_1 \cdots \alpha_k$ of extended priority queue and a witness automaton $\mathcal{A}$, we say that $e$ is accepted by $\mathcal{A}$, if
+Register automata is a kind of finite automata with registers used to deal with infinite data domain, where register can be assigned values and do simple operations.
+In this paper we use a restricted kind of register automata, which contains only one register, whose value is obtained once by guessing and never changed.
+Our register machine contains a register $\textit{reg}$.
+Its transition labels are $\textit{call}(\textit{put},d,p,g), \textit{ret}(\textit{put},d,p),$ $\textit{call}(\textit{rm},d),\textit{ret}(\textit{rm},d)$, where $d \in \mathbb{D} \cup \{ \textit{empty} \}$, and $g$ and $\top$ are guards.
+A guard is a predicate chosen from $\{\textit{equalReg},\textit{lesReg},\top \}$, which represents equals the value of $\textit{reg}$, less than the value of $\textit{reg}$, and always holds, respectively.
+Given an execution $e = \alpha_1 \cdots \alpha_k$ of priority queue, register automata $\mathcal{A}$ accepts $e$, if there exists value $p \in \in \mathbb{D}$ and transitions $q_0 \xrightarrow{\beta_1} q_1 \ldots \xrightarrow{\beta_k} q_k$ of $\mathcal{A}$ from a initial state to a final state, such that for each $i$
 
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] There exist transitions $q_0 \xrightarrow{\beta_1} q_1 \ldots \xrightarrow{\beta_k} q_k$ of $\mathcal{A}$, such that $q_0$ is one of initial state of $\mathcal{A}$, and $q_k$ is one of accept state of $\mathcal{A}$.
-We choose value of $p$ as some value $d_p \in \mathbb{D}$.
+\item[-] If $\alpha_i = \textit{call}(\textit{put},a,p)$, then $\beta_i= \textit{call}(\textit{put},a,p,g)$ and $g(q)$ holds.
 
-\item[-] For each $i$, if $\alpha_i = \textit{cal}(\textit{put},a,q)$, then either (1) $q = d_p$ and $\beta_i = \textit{cal}(\textit{put},a,p)$, or (2) $q <_{\mathbb{P}} d_p$ and $\beta_i = \textit{cal}(\textit{put},a,\textit{les}_p)$, or (3) $\beta_i = \textit{cal}(\textit{put},a,\textit{anyPri})$.
-
-\item[-] For each $i$, if $\alpha_i = \textit{ret}(\textit{put},a)$, $\textit{cal}(\textit{rm},a)$, $\textit{ret}(\textit{rm},a)$, $\textit{cal}(\textit{rm},\textit{empty})$ or $\textit{ret}(\textit{rm},\textit{empty})$, then $\beta_i = \alpha_i$.
+\item[-] Else, $\alpha_i = \beta_i$. 
 \end{itemize}
 
-Note that due to data-independence, witness automata are used to detect some renamed violations, instead of store and checking directly.
-Let us introduce the notion of co-regular, which reduce checking linearizable w.r.t $\textit{MS}(R)$ into checking set of witness automata.
+Note that due to data-independence, register automata are used to detect some renamed violations, instead of store and checking directly.
+
+Let us introduce the notion of co-regular, which reduce checking linearizable w.r.t $\Gamma\mathsf{\text{-}Conc}(e,\alpha)$ into checking set of witness automata.
 

### 2017-04-19 co-regular-EPQ.tex
modify

@@
 Our register machine contains a register $\textit{reg}$.
-Its transition labels are $\textit{call}(\textit{put},d,p,g), \textit{ret}(\textit{put},d,p),$ $\textit{call}(\textit{rm},d),\textit{ret}(\textit{rm},d)$, where $d \in \mathbb{D} \cup \{ \textit{empty} \}$, and $g$ and $\top$ are guards.
+Its transition labels are $\textit{call}(\textit{put},d,g), \textit{ret}(\textit{put},d,p),$ $\textit{call}(\textit{rm},d),\textit{ret}(\textit{rm},d)$, where $d \in \mathbb{D} \cup \{ \textit{empty} \}$, and $g$ and $\top$ are guards.
 A guard is a predicate chosen from $\{\textit{equalReg},\textit{lesReg},\top \}$, which represents equals the value of $\textit{reg}$, less than the value of $\textit{reg}$, and always holds, respectively.
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] If $\alpha_i = \textit{call}(\textit{put},a,p)$, then $\beta_i= \textit{call}(\textit{put},a,p,g)$ and $g(q)$ holds.
+\item[-] If $\alpha_i = \textit{call}(\textit{put},a,p)$, then $\beta_i= \textit{call}(\textit{put},a,g)$ and $g(q)$ holds.
 

### 2017-04-19 co-regular-EPQ.tex
modify

@@
 \label{lemma:EPQ as multi in MRpri for history}
-Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$, if and only if, $\forall e' \in \textit{proj}(e)$, $\Gamma\mathsf{\text{-}Condition}(e) \Rightarrow \Gamma\mathsf{\text{-}Conc}(e,\alpha)$ holds for some $\alpha$
+Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$, if and only if, $\forall e' \in \textit{proj}(e)$, $\Gamma\mathsf{\text{-}Condition}(e) \Rightarrow \Gamma\mathsf{\text{-}Conc}(e,\alpha)$ holds for some $\alpha$ for each $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$.
 \end{restatable}
@@
 Note that due to data-independence, register automata are used to detect some renamed violations, instead of store and checking directly.
-
 Let us introduce the notion of co-regular, which reduce checking linearizable w.r.t $\Gamma\mathsf{\text{-}Conc}(e,\alpha)$ into checking set of witness automata.
@@
 \begin{definition}\label{def:co-regular of rules of extended priority queues}
-Given $R \in \{ \textit{EPQ}_1^{>}, \textit{EPQ}_1^{=}, \textit{EPQ}_2^{>}, \textit{EPQ}_2^{=}, \textit{EPQ}_3 \}$.
-$R$ is co-regular, if there are a finite set $\textit{Auts}_{R}$ of witness automata such that, for each data-independence implementation $\mathcal{I}$, we have that
+$\seqPQ$ is co-regular, if for each $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$, there is a finite set $\textit{Auts}_{\Gamma}$ of register automata, such that for each data-independence implementation $\mathcal{I}$, we have that:
 
-$$\textit{Auts}_{R} \cap \mathcal{I} \neq \emptyset \Leftrightarrow \exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), R \in last(e') \wedge e' \not\sqsubseteq \textit{MS}(R)$$
-
-We say that $\textit{EPQ}$ is co-regular, if all five candidates of $R$ are co-regular.
+$\textit{Auts}_{\Gamma} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $\Gamma\mathsf{\text{-}Condition}(e')$ holds but $\Gamma\mathsf{\text{-}Conc}(e',\alpha)$ does not hold for some $\alpha$
 \end{definition}
 
-We simply the proof of co-regular of $\textit{EPQ}$ in two side.
-The first side uses the results in \cite{Bouajjani:2015} to ensure FIFO (first in first out) property of each single-priority projections.
-A execution where only items of one priority is putted is called a single-priority execution.
-It is obvious that each single-priority projection of sequences in $\textit{EPQ}$ satisfy FIFO property.
-Let $\textit{transToQueue}(e)$ be an execution generated from $e$ by transforming $\textit{put}$ and $\textit{rm}$ into $\textit{enq}$ and $\textit{deq}$, respectively, and then discarding priorities.
+We use two point to simply our proof co-regular.
+The first point is to ues the results in \cite{Bouajjani:2015} to ensure FIFO (first in first out) property of each single-priority projections.
+A execution where only values of one priority is putted is called a single-priority execution.
+It is obvious that each single-priority projection of sequences in $\seqPQ$ satisfy FIFO property.
+Let $\textit{toQueue}(e)$ be an execution generated from $e$ by transforming $\textit{put}$ and $\textit{rm}$ into $\textit{enq}$ and $\textit{deq}$, respectively, and then discarding priorities.
 In Appendix \ref{sec:appendix proof and definition in section definition of co-regular}, according to the result of queue in \cite{Bouajjani:2015}, we construct a set $\textit{Auts}_{\textit{sinPri}}$ of witness automata, and shows that they are enough to ensure that for each data-differentiated executions, each of its single-priority projection without $\textit{rm}(\textit{empty})$ to have ``FIFO'' property, as shown by the following lemma.
@@
 
-According to Lemma \ref{lemma:automata for extended priority queue with single priority}, from now on, it is safe to assume that, for each data-differentiated execution without $\textit{rm}(\textit{empty})$, any of its single-priority projection has ``FIFO'' property.
-For example, $\textit{rm}(a)$ never happens before $\textit{put}(a,\_)$ for each $a$.
+According to Lemma \ref{lemma:automata for extended priority queue with single priority}, from now on, it is safe to assume that, for each data-differentiated execution, any of its single-priority projection without $\textit{rm}(\textit{empty})$ has ``FIFO'' property.
+This effectively exclude some obvious violations, such as $\textit{rm}(a)$ happens before $\textit{put}(a,\_)$.
 
-The second side is to ignore the incomparable priorities.
-Given a data-differentiated execution $e$, we say that $e$ is a $\textit{pri}$-execution, if $\textit{pri}$ is the maximal priority of $e$, and $\textit{pri}$ is larger than or equal to the priority of all other items of $e$.
-Given a data-differentiated execution $e$ and assume that $\textit{pri}$ is one of the maximal priority of $e$, let $\textit{pri-Exec}(e)$ be an execution obtained from $e$ by erasing all operations of items whose priority are incomparable with $\textit{pri}$.
-The following lemma states that when checking co-regular of $\textit{EPQ}$, it is safe to consider $\textit{pri}$-executions.
-We can see that $\textit{last}(e')$ contains at most one element for $\textit{pri}$-execution $e'$.
+The second point is to ignore the incomparable priorities.
+Given a data-differentiated execution $e$, we say that $e$ is a $p$-execution, if $p$ is larger than or equal to the priority of values of $e$.
+Given a data-differentiated execution $e$ and assume that $p$ is one of the maximal priority of $e$, let $\textit{p-Exec}(e)$ be an execution obtained from $e$ by erasing all operations of values whose priority are incomparable with $\textit{pri}$.
+The following lemma states that when checking co-regular of $\seqPQ$, it is safe to consider only $p$-executions.
 
@@
 \label{lemma:pri execution is enough}
-Given a data-differentiated execution $e$ and $R = \textit{EPQ}_1^{>}$ (resp., $R = \textit{EPQ}_1^{=}$, $R=\textit{EPQ}_2^{>}$, or $R = \textit{EPQ}_2^{=}$).
-$e \sqsubseteq \textit{MS}(R)$ with witness $x$, if and only if $\textit{pri-Exec}(e) \sqsubseteq \textit{MS}(R)$ with witness $x$, where $\textit{pri}$ is the priority of $x$.
-%for each $\textit{pri}$ that is (1) maximal in $e$ and (2) items of $\textit{pri}$ are only one pair of matched $\textit{put}$ in $e$ (resp., more than one pair of matched $\textit{put}$, unmatched $\textit{put}$, or unmatched $\textit{put}$ and matched $\textit{put}$).
+Given a data-differentiated execution $e$ and $\Gamma\in \{\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$.
+$\Gamma\mathsf{\text{-}Conc}(e,x)$ holds, if and only if $\Gamma\mathsf{\text{-}Conc}(\textit{p-Exec}(e),x)$ holds, where $p$ is the priority of $x$.
 \end{restatable}
+\begin {proof} (Sketch)
+
+To prove the $\textit{only if}$ direction, given $e \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \Gamma\mathsf{\text{-}Seq}(s,\alpha)$.
+Since there is no restriction to $p$-incomparable operations in $u$, $v$ and $w$, we can safely erase all $p$-incomparable values.
+
+To prove the $\textit{if}$ direction, we use a lemma in Appendix, which said that if the projection an execution into a set $O_1$ of operations has linearization $l_1$ and the projection of the execution into $O_2$, the remanning operation, has linearization $l_2$, then the execution has a linearization whose projection into $O_1$ is $l_1$ and projection into $O_2$ is $l_2$.
+
+Given $\textit{p-Exec}(e) \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \Gamma\mathsf{\text{-}Seq}(s,\alpha)$.
+By above lemma we know that $e$ has a linearization $l'$, such that the projection of $l'$ into $p$-incomparable values.
+It is not hard to see that the priorities comparable with $p$ are either smaller or incomparable with the priorities incomparable with $p$.
+Thus, $l' \in \seqPQ$.
+\qed
+\end {proof}
+
+Lemma \ref{lemma:pri execution is enough} make our co-regular proof more clear. One consequence of it is that for $e=\textit{p-Exec}(e)$, at most one $\Gamma\mathsf{\text{-}Condition}(e')$ holds.
 

### 2017-04-19 co-regular-EPQ.tex
modify

@@
 Let $\Gamma\mathsf{\text{-}Condition}(e)$ be $\mathsf{HasEmptyRemoves}(e)$, $\neg \mathsf{HasEmptyRemoves}(e) \wedge \mathsf{HasUnmatchedMaxPriority}(e)$, and $\mathsf{HasMatchedMaxPriority}(e)$, for $\Gamma = \mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}$, respectively.
-$\textit{Check-PQ-Conc}$ inspires us that, whenever $\Gamma\mathsf{\text{-}Condition}(e)$ holds, there should exist some $\alpha$, such that  $\Gamma\mathsf{\text{-}Conc}(e,\alpha)$ holds.
+$\textit{Check-PQ-Conc}$ inspires us that, when $\Gamma\mathsf{\text{-}Condition}(e)$ holds, there should exist some $\alpha$, such that  $\Gamma\mathsf{\text{-}Conc}(e,\alpha)$ holds.
 The following lemma states this formally.
@@
 In this paper we use a restricted kind of register automata, which contains only one register, whose value is obtained once by guessing and never changed.
-Our register machine contains a register $\textit{reg}$.
-Its transition labels are $\textit{call}(\textit{put},d,g), \textit{ret}(\textit{put},d,p),$ $\textit{call}(\textit{rm},d),\textit{ret}(\textit{rm},d)$, where $d \in \mathbb{D} \cup \{ \textit{empty} \}$, and $g$ and $\top$ are guards.
-A guard is a predicate chosen from $\{\textit{equalReg},\textit{lesReg},\top \}$, which represents equals the value of $\textit{reg}$, less than the value of $\textit{reg}$, and always holds, respectively.
-Given an execution $e = \alpha_1 \cdots \alpha_k$ of priority queue, register automata $\mathcal{A}$ accepts $e$, if there exists value $p \in \in \mathbb{D}$ and transitions $q_0 \xrightarrow{\beta_1} q_1 \ldots \xrightarrow{\beta_k} q_k$ of $\mathcal{A}$ from a initial state to a final state, such that for each $i$
+Our register machine contains a register $r$.
+Its transition labels are $r=*,\textit{call}(\textit{put},d,g), \textit{ret}(\textit{put},d,g),$ $\textit{call}(\textit{rm},d),\textit{ret}(\textit{rm},d)$, where $d \in \mathbb{D} \cup \{ \textit{empty} \}$, and $g$ and $\top$ are guards.
+The first transition from initial state is a $r=*$ transition, and is used to assign an arbitrary value to $r$.
+A guard is a predicate chosen from $\{=r,\leq r,\top \}$, which represents equals the value of $\textit{reg}$, less than the value of $\textit{reg}$, and always holds, respectively.
+Given an execution $e = \alpha_1 \cdots \alpha_k$ of priority queue, register automata $\mathcal{A}$ accepts $e$, if there exists value $d_r\in \mathbb{D}$ and transitions $q_0 \xrightarrow{\beta_1} q_1 \ldots \xrightarrow{\beta_k} q_k$ of $\mathcal{A}$ from a initial state to a final state, such that for each $i$
 
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] If $\alpha_i = \textit{call}(\textit{put},a,p)$, then $\beta_i= \textit{call}(\textit{put},a,g)$ and $g(q)$ holds.
+\item[-] If $\alpha_i = \textit{call}(\textit{put},a,p)$, then $\beta_i= \textit{call}(\textit{put},a,g)$ and $g$ holds with $p$ and $d_r$. Similarly for $\textit{ret}(\textit{put},a,p)$.
 
@@
 \begin{definition}\label{def:co-regular of rules of extended priority queues}
-$\seqPQ$ is co-regular, if for each $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$, there is a finite set $\textit{Auts}_{\Gamma}$ of register automata, such that for each data-independence implementation $\mathcal{I}$, we have that:
+Given $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$.
+$\Gamma$ is co-regular, if there is a finite set $\textit{Auts}_{\Gamma}$ of register automata, such that for each data-independence implementation $\mathcal{I}$, we have that:
 
-$\textit{Auts}_{\Gamma} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $\Gamma\mathsf{\text{-}Condition}(e')$ holds but $\Gamma\mathsf{\text{-}Conc}(e',\alpha)$ does not hold for some $\alpha$
+$\textit{Auts}_{\Gamma} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $\Gamma\mathsf{\text{-}Condition}(e')$ holds but $\Gamma\mathsf{\text{-}Conc}(e',\alpha)$ does not hold for some $\alpha$.
+
+$\seqPQ$ is co-regular, if each $\Gamma$ is co-regular. 
 \end{definition}
@@
 
-Given a data-independent implementations $\mathcal{I}$ of extended priority queue, $\mathcal{I} \cap \textit{Auts}_{\textit{sinPri}} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{transToQueue}(e')$ does not linearizable to queue.
+Given a data-independent implementations $\mathcal{I}$ of extended priority queue, $\mathcal{I} \cap \textit{Auts}_{\textit{sinPri}} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{toQueue}(e')$ does not linearizable to queue.
 \end{restatable}
@@
 The second point is to ignore the incomparable priorities.
-Given a data-differentiated execution $e$, we say that $e$ is a $p$-execution, if $p$ is larger than or equal to the priority of values of $e$.
-Given a data-differentiated execution $e$ and assume that $p$ is one of the maximal priority of $e$, let $\textit{p-Exec}(e)$ be an execution obtained from $e$ by erasing all operations of values whose priority are incomparable with $\textit{pri}$.
-The following lemma states that when checking co-regular of $\seqPQ$, it is safe to consider only $p$-executions.
+Given a data-differentiated execution $e$ and one of its maximal priority $p$, $e\vert_{\preceq p}$ is an execution obtained from $e$ by erasing operations of items with priority incomparable with $p$.
+The following lemma states that when checking co-regular of $\seqPQ$, it is safe to ignore all incomparable priorities.
 
@@
 \label{lemma:pri execution is enough}
-Given a data-differentiated execution $e$ and $\Gamma\in \{\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$.
-$\Gamma\mathsf{\text{-}Conc}(e,x)$ holds, if and only if $\Gamma\mathsf{\text{-}Conc}(\textit{p-Exec}(e),x)$ holds, where $p$ is the priority of $x$.
+Given a data-differentiated execution $e$ and $\Gamma\in \{\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$.
+$\Gamma\mathsf{\text{-}Conc}(e,x)$ holds, if and only if $\Gamma\mathsf{\text{-}Conc}(e\vert_{\preceq p},x)$ holds, where $p$ is the priority of $x$.
 \end{restatable}
@@
 
-Given $\textit{p-Exec}(e) \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \Gamma\mathsf{\text{-}Seq}(s,\alpha)$.
+Given $e\vert_{\preceq p} \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \Gamma\mathsf{\text{-}Seq}(s,\alpha)$.
 By above lemma we know that $e$ has a linearization $l'$, such that the projection of $l'$ into $p$-incomparable values.
@@
 
-Lemma \ref{lemma:pri execution is enough} make our co-regular proof more clear. One consequence of it is that for $e=\textit{p-Exec}(e)$, at most one $\Gamma\mathsf{\text{-}Condition}(e')$ holds.
+Lemma \ref{lemma:pri execution is enough} makes our co-regular proof more clear. One consequence of it is that for $e=\textit{p-Exec}(e)$, at most one $\Gamma\mathsf{\text{-}Condition}(e')$ holds.
 
-\subsection{Co-Regular of $\textit{EPQ}_1^{>}$}
+\subsection{Co-Regular of $\mathsf{MatchedMaxPriority}^{>}$}
 \label{subsec:co-regular of EPQ1Lar}
 
-In this subsection, we briefly introduce the idea for proving co-regular of $\textit{PQ}_1^{>}$.
+To facilitate the proof of co-Regular of $\mathsf{MatchedMaxPriority}$, we separate $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ into the following two sub-predicates, where the first one represents that $x$ is the only value with priority $p$ in $e$ , and the second one represents that there are more than one value with priority $p$ in $e$.
+Here $p\not\prec'\textit{priorities}(l)$ means that $p$ is either larger, equal, or incomparable with priorities in $l$, and $p$ must equals with some priority in $l$.
+
+{\small
+\begin{align*}
+\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v\cdot \textit{rm}(x)\cdot w$, $p\not\preceq \textit{priorities}(u\cdot v\cdot w)$,} \\
+&\hspace{4mm}\mbox{$p\not\preceq \textit{unmatched-priorities}(u\cdot v\cdot w)$, $\textit{matched}_\prec(u\cdot v,p)$,} \\
+&\hspace{4mm}\mbox{and $p\not\in \textit{priorities}(v\cdot w)$} \\
+\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v\cdot \textit{rm}(x)\cdot w$, $p\not\prec' \textit{priorities}(u\cdot v\cdot w)$,} \\
+&\hspace{4mm}\mbox{$p\not\preceq \textit{unmatched-priorities}(u\cdot v\cdot w)$, $\textit{matched}_\prec(u\cdot v,p)$,} \\
+&\hspace{4mm}\mbox{and $p\not\in \textit{priorities}(v\cdot w)$} 
+\end{align*}
+}
+
+The corresponding conditions are defined as follows: $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Condition}(e)$
+
+{\small
+\begin{align*}
+\mathsf{HasEmptyRemoves}(e)=\mathsf{true} & \mbox{ iff  $e$ contains a $\textit{rm}(\textit{empty})$ operation} \\
+\mathsf{HasUnmatchedMaxPriority}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{unmatched-priorities}(e)$ for a maximal priority} \\
+&\hspace{4mm}\mbox{$p\in priorities(e)$}
+\end{align*}}
+
+$\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ is obtained from $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Conc}(e,x)$ by using $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(e,x)$ instead of $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$.
+And similarly for $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Conc}(e,x)$.
+
+In this subsection, we briefly introduce the idea for proving co-regular of $\mathsf{MatchedMaxPriority}^{>}$.
 The proof of this subsection can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
-The notion of left-right constraint used for extended priority queue is inspired by left-right constraint of queue \cite{Bouajjani:2015}.
+The notion of left-right constraint used for priority queue is inspired by left-right constraint of queue \cite{Bouajjani:2015}.
 
-Given a data-differentiated $\_$-execution $e$ such that $\textit{last}(e) = \textit{EPQ}_1^{>}$, Lemma \ref{lemma:automata for extended priority queue with single priority} is not enough for ensuring that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$.
+$\Gamma\mathsf{\text{-}Condition}(e')$ holds but $\Gamma\mathsf{\text{-}Conc}(e',\alpha)$ does not hold
+
+Given a data-differentiated $\_$-execution $e$ such that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Condition}(e)$ holds, Lemma \ref{lemma:automata for extended priority queue with single priority} is not enough for ensuring that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ for some $x$.
 This is because that it is possible that $e \not\sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$ because of interaction between actions of multiple priorities.

### 2017-04-19 introduction.tex
intro sketch
added references to the intro

@@
 
-Efficient implementations of concurrent objects are hard to get right.
+Efficient implementations of concurrent objects are essential and hard to get right.
 Verifying them is difficult, checking a single execution is NP-complete~\cite{journals/siamcomp/GibbonsK97} and checking a finite-state implementation is undecidable~\cite{conf/esop/BouajjaniEEH13}.
@@
 
-Here, we consider another important object, the priority queues, which are used to TODO, and for which numerous implementations exist CITE (research articles and languages/libraries providing priority queue implementations).
+Here, we consider another important object, the priority queues, which are essential for applications such as task scheduling and discrete event simulation, and for which numerous implementations have been proposed, e.g.,~\cite{DBLP:conf/ppopp/AlistarhKLS15,DBLP:conf/wdag/CalciuMH14,DBLP:conf/opodis/LindenJ13,DBLP:conf/podc/ShavitZ99,DBLP:conf/ipps/ShavitL00}.
 This object is much less studied compared to the other ones in the verification literature.

### 2017-04-19 co-regular-EPQ.tex
modify

@@
 The first transition from initial state is a $r=*$ transition, and is used to assign an arbitrary value to $r$.
-A guard is a predicate chosen from $\{=r,\leq r,\top \}$, which represents equals the value of $\textit{reg}$, less than the value of $\textit{reg}$, and always holds, respectively.
+A guard is a predicate chosen from $\{=r,\leq r,\top \}$, which represents equals the value of $r$, less than the value of $r$, and always holds, respectively.
 Given an execution $e = \alpha_1 \cdots \alpha_k$ of priority queue, register automata $\mathcal{A}$ accepts $e$, if there exists value $d_r\in \mathbb{D}$ and transitions $q_0 \xrightarrow{\beta_1} q_1 \ldots \xrightarrow{\beta_k} q_k$ of $\mathcal{A}$ from a initial state to a final state, such that for each $i$
@@
 Given $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$.
-$\Gamma$ is co-regular, if there is a finite set $\textit{Auts}_{\Gamma}$ of register automata, such that for each data-independence implementation $\mathcal{I}$, we have that:
+$\Gamma$ is co-regular, if there is a finite set $\textit{Auts}_{\Gamma}$ of register automata, such that for each data-independent implementation $\mathcal{I}$, we have that:
 
@@
 
-We use two point to simply our proof co-regular.
-The first point is to ues the results in \cite{Bouajjani:2015} to ensure FIFO (first in first out) property of each single-priority projections.
+We use two point to simply our proof of co-regular.
+The first point is to use the results in \cite{Bouajjani:2015} to ensure FIFO (first in first out) property of each single-priority projections.
 A execution where only values of one priority is putted is called a single-priority execution.
@@
 The second point is to ignore the incomparable priorities.
-Given a data-differentiated execution $e$ and one of its maximal priority $p$, $e\vert_{\preceq p}$ is an execution obtained from $e$ by erasing operations of items with priority incomparable with $p$.
+Given a data-differentiated execution $e$ and one of its maximal priority $p$, $e\vert_{\preceq p}$ is an execution obtained from $e$ by erasing operations of values with priority incomparable with $p$.
 The following lemma states that when checking co-regular of $\seqPQ$, it is safe to ignore all incomparable priorities.
@@
 
-To prove the $\textit{only if}$ direction, given $e \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \Gamma\mathsf{\text{-}Seq}(s,\alpha)$.
+To prove the $\textit{only if}$ direction, given $e \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
 Since there is no restriction to $p$-incomparable operations in $u$, $v$ and $w$, we can safely erase all $p$-incomparable values.
+Similar case holds for $\Gamma=\mathsf{UnmatchedMaxPriority}$.
 
-To prove the $\textit{if}$ direction, we use a lemma in Appendix, which said that if the projection an execution into a set $O_1$ of operations has linearization $l_1$ and the projection of the execution into $O_2$, the remanning operation, has linearization $l_2$, then the execution has a linearization whose projection into $O_1$ is $l_1$ and projection into $O_2$ is $l_2$.
+To prove the $\textit{if}$ direction, we use a lemma in Appendix, which said that if the projection of an execution into a set $O_1$ of operations has linearization $l_1$ and the projection of the execution into $O_2$, the remanning operations, has linearization $l_2$, then the execution has a linearization whose projection into $O_1$ is $l_1$ and projection into $O_2$ is $l_2$.
 
-Given $e\vert_{\preceq p} \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \Gamma\mathsf{\text{-}Seq}(s,\alpha)$.
-By above lemma we know that $e$ has a linearization $l'$, such that the projection of $l'$ into $p$-incomparable values.
-It is not hard to see that the priorities comparable with $p$ are either smaller or incomparable with the priorities incomparable with $p$.
-Thus, $l' \in \seqPQ$.
+Given $e\vert_{\preceq p} \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
+By above lemma we know that $e$ has a linearization $l' = u' \cdot \textit{put}(x,p) \cdot v' \cdot \textit{rm}(x) \cdot w'$, such that the projection of $l'$ into $p$-incomparable values is l.
+We can see that $u$, $v$ and $w$ are sub-sequence of $u'$, $v'$ an $w'$, respectively.
+Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$ does not have restriction to $p$-incomparable values, we can see that $l' \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
 \qed
@@
 
-Lemma \ref{lemma:pri execution is enough} makes our co-regular proof more clear. One consequence of it is that for $e=\textit{p-Exec}(e)$, at most one $\Gamma\mathsf{\text{-}Condition}(e')$ holds.
+Lemma \ref{lemma:pri execution is enough} makes our co-regular proof more clear.
+One consequence of it is that for given $e\vert_{\preceq p}$, $\Gamma\mathsf{\text{-}Condition}(e')$ holds only for one $\Gamma$.
 
@@
 To facilitate the proof of co-Regular of $\mathsf{MatchedMaxPriority}$, we separate $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ into the following two sub-predicates, where the first one represents that $x$ is the only value with priority $p$ in $e$ , and the second one represents that there are more than one value with priority $p$ in $e$.
-Here $p\not\prec'\textit{priorities}(l)$ means that $p$ is either larger, equal, or incomparable with priorities in $l$, and $p$ must equals with some priority in $l$.
+Here $\textit{occurOnce}(p,l)$ holds if there is only one value with priority $p$ in $l$.
 
@@
 \begin{align*}
-\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v\cdot \textit{rm}(x)\cdot w$, $p\not\preceq \textit{priorities}(u\cdot v\cdot w)$,} \\
-&\hspace{4mm}\mbox{$p\not\preceq \textit{unmatched-priorities}(u\cdot v\cdot w)$, $\textit{matched}_\prec(u\cdot v,p)$,} \\
-&\hspace{4mm}\mbox{and $p\not\in \textit{priorities}(v\cdot w)$} \\
-\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v\cdot \textit{rm}(x)\cdot w$, $p\not\prec' \textit{priorities}(u\cdot v\cdot w)$,} \\
-&\hspace{4mm}\mbox{$p\not\preceq \textit{unmatched-priorities}(u\cdot v\cdot w)$, $\textit{matched}_\prec(u\cdot v,p)$,} \\
-&\hspace{4mm}\mbox{and $p\not\in \textit{priorities}(v\cdot w)$} 
+\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(e,x)=\mathsf{true}$,} \\
+&\hspace{4mm}\mbox{$\textit{occurOnce}(p,e)$ where $p$ is the priority of $x$} \\
+\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(e,x)=\mathsf{true}$,} \\
+&\hspace{4mm}\mbox{$\neg\textit{occurOnce}(p,e)$ where $p$ is the priority of $x$} \\
 \end{align*}
@@
 
-The corresponding conditions are defined as follows: $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Condition}(e)$
+$\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ is obtained from $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Conc}(e,x)$ by using $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(e,x)$ instead of $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$.
+And similarly for $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Conc}(e,x)$.
+
+We can also define two conditions, where $\textit{unmatched-priorities}(l)$ is the set of priorities occurring in $\textit{put}$ operations of $e$ for which there is $\textit{rm}$ operation removing the same value:
 
@@
 \begin{align*}
-\mathsf{HasEmptyRemoves}(e)=\mathsf{true} & \mbox{ iff  $e$ contains a $\textit{rm}(\textit{empty})$ operation} \\
-\mathsf{HasUnmatchedMaxPriority}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{unmatched-priorities}(e)$ for a maximal priority} \\
-&\hspace{4mm}\mbox{$p\in priorities(e)$}
+\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Condition}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{matched-priorities}(e)$ for a maximal priority} \\
+&\hspace{4mm}\mbox{$p\in priorities(e)$, $\textit{occurOnce}(p,e)$} \\
+\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{matched-priorities}(e)$ for a maximal priority} \\
+&\hspace{4mm}\mbox{$p\in priorities(e)$, $\neg\textit{occurOnce}(p,e)$} 
 \end{align*}}
-
-$\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ is obtained from $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Conc}(e,x)$ by using $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(e,x)$ instead of $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$.
-And similarly for $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Conc}(e,x)$.
 
@@
 
-$\Gamma\mathsf{\text{-}Condition}(e')$ holds but $\Gamma\mathsf{\text{-}Conc}(e',\alpha)$ does not hold
-
 Given a data-differentiated $\_$-execution $e$ such that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Condition}(e)$ holds, Lemma \ref{lemma:automata for extended priority queue with single priority} is not enough for ensuring that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ for some $x$.
-This is because that it is possible that $e \not\sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$ because of interaction between actions of multiple priorities.
+This is because that it is possible that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ does not holds because of interaction between actions of multiple priorities.
 One such example is shown in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
-We call the time interval from $\textit{ret}(\textit{put},x)$ to $\textit{cal}(rm,x)$, or from $\textit{ret}(\textit{put},x)$ when $\textit{cal}(rm,x)$ does not exist, the interval of item $x$.
-In \figurename~\ref{fig:introduce gap for EPQ1Lar}, we draw the interval of each item by dashed line.
-Here we assume that $p_1 <_{\mathbb{P}} p_4$, $p_2 <_{\mathbb{P}} p_4$ and $p_3 <_{\mathbb{P}} p_4$.
-The reason of $e \not\sqsubseteq \textit{MS}(\textit{PQ}_1^{>})$ is that, each time point from $\textit{cal}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is in interval of some item with smaller priority.
+We call the time interval from $\textit{ret}(\textit{put},x)$ to $\textit{cal}(rm,x)$, or from $\textit{ret}(\textit{put},x)$ when $\textit{cal}(rm,x)$ does not exist, the interval of value $x$.
+In \figurename~\ref{fig:introduce gap for EPQ1Lar}, we draw the interval of each value by dashed line.
+Here we assume that $p_1 \prec p_4$, $p_2 \prec p_4$ and $p_3 \prec p_4$.
+We can not find a sequence $s$ where $e \sqsubseteq s$ and $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(s,b)$ holds, since each time point from $\textit{cal}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is in interval of some item with smaller priority, and $\textit{rm}(b)$ can not happen in interval of item with smaller priority.
 
@@
 %\vspace{-10pt}
-  \caption{An execution that does not linearizable w.r.t $\textit{MS}(\textit{EPQ}_1^{>})$}
+  \caption{An execution that does not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$}
   \label{fig:introduce gap for EPQ1Lar}
@@
 In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we introduce the notion of left-right constraint, which is a graph, and the existence of cycle though item with maximal priority in it formally modelled the case in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
-When there is a cycle $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ through item $x$ with maximal priority in left-right constraint, we say that $x$ is covered by $d_1,\ldots,d_m$ For example, in \figurename~\ref{fig:introduce gap for EPQ1Lar}, $b$ is covered by $f$, $e$ and $d$.
-We then prove that getting rid of cycle though item with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\textit{MS}(\textit{EPQ}_1^{>})$, as stated by the following lemma:
+For example, in \figurename~\ref{fig:introduce gap for EPQ1Lar}, $b$ is covered by $f$, $e$ and $d$.
+We then prove that getting rid of cycle though item with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$, or we can say, ensure $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ for some $x$.
 
-\begin{restatable}{lemma}{LinEqualsConstraintforEPQOneLar}
-\label{lemma:Lin Equals Constraint for EPQ1Lar}
-Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{>}$.
-Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x)$ and $\textit{rm}(x)$.
-$e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$, if and only if $G$ has no cycle going through $x$.
-\end{restatable}
-
+When there is a cycle $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ through item $x$ with maximal priority in left-right constraint, we say that $x$ is covered by $d_1,\ldots,d_m$.
 Assume that $x$ is covered by $d_1,\ldots,d_m$, we can safely rename $x$ into $b$, rename $d_1,\ldots,d_m$ into $a$ and rename all other item into $c$ by data-independence.
-Such execution can be recognized by witness automata, since between the first $\textit{ret}(\textit{put},a,\_)$ and the last $\textit{cal}(\textit{rm},a)$ (if exists), $\textit{ret}(\textit{put},a,\_)$ and $\textit{cal}(\textit{rm},a)$ occurs in pair and there is no need for count.
-In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we construct a set $\textit{Auts}_{\textit{1-lar}}$ of witness automata, and use $\textit{Auts}_{\textit{1-lar}}$ to prove that $\textit{EPQ}_1^{>}$ is co-regular, as stated by the following lemma.
+Such execution can be recognized by witness automata, since between the first $\textit{ret}(\textit{put},a,\_)$ and the last $\textit{call}(\textit{rm},a)$ (if exists), $\textit{ret}(\textit{put},a,\_)$ and $\textit{call}(\textit{rm},a)$ occurs in pair and there is no need for count.
+In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we construct a set $\textit{Auts}_{\textit{1-lar}}$ of witness automata, and use $\textit{Auts}_{\textit{1-lar}}$ to prove that $\mathsf{MatchedMaxPriority}^{>}$ is co-regular, as stated by the following lemma.
 
@@
 \label{lemma:EPQ1Lar is co-regular}
-$\textit{EPQ}_1^{>}$ is co-regular.
+$\mathsf{MatchedMaxPriority}^{>}$ is co-regular.
 \end{restatable}

### 2017-04-19 appendix-EPQ.tex
modify

@@
 
-\FiveEnmuerationisEnoughForEPQOneEqual*
-
+\begin{restatable}{lemma}{FiveEnmuerationisEnoughForEPQOneEqual}
+\label{lemma:five enumeration is enough for EPQ1Equal}
+Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{=}$.
+Let $a$ and $b$ be items with maximal priority $\textit{pri}$.
+Assume that $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\textit{pri})$ or $\textit{cal}(\textit{rm},a)$.
+Then, there are five possible enumeration of method events of $a$, $b$, $a_1$ (if exists), where $a_1$ is the possible intermediate items for obtain $a <_{\textit{pb}}^* b$.
+\end{restatable}
 \begin {proof}

### 2017-04-19 co-regular-EPQ.tex
modify

@@
 When refer to $\textit{proj}(e)$, we implicitly assume that each $\textit{rm}(\textit{empty})$ in $e$ has a ghost argument that is unique.
+We say that execution $l$ is linearizable w.r.t $\Gamma$, if we can find a sequence $s$ where $e \sqsubseteq s$ and $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$ holds for some $\alpha$.
 
@@
 \label{lemma:EPQ as multi in MRpri for history}
-Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$, if and only if, $\forall e' \in \textit{proj}(e)$, $\Gamma\mathsf{\text{-}Condition}(e) \Rightarrow \Gamma\mathsf{\text{-}Conc}(e,\alpha)$ holds for some $\alpha$ for each $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$.
+Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$, if and only if, $\forall e' \in \textit{proj}(e)$, $\forall \Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$,  $\Gamma\mathsf{\text{-}Condition}(e)$ $\Rightarrow$ $e$ is linearizable w.r.t $\Gamma$.
 \end{restatable}
@@
 The first transition from initial state is a $r=*$ transition, and is used to assign an arbitrary value to $r$.
-A guard is a predicate chosen from $\{=r,\leq r,\top \}$, which represents equals the value of $r$, less than the value of $r$, and always holds, respectively.
+A guard is a predicate chosen from $\{=r,< r,\top \}$, which represents equals the value of $r$, less than the value of $r$, and always holds, respectively.
 Given an execution $e = \alpha_1 \cdots \alpha_k$ of priority queue, register automata $\mathcal{A}$ accepts $e$, if there exists value $d_r\in \mathbb{D}$ and transitions $q_0 \xrightarrow{\beta_1} q_1 \ldots \xrightarrow{\beta_k} q_k$ of $\mathcal{A}$ from a initial state to a final state, such that for each $i$
@@
 
-$\textit{Auts}_{\Gamma} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $\Gamma\mathsf{\text{-}Condition}(e')$ holds but $\Gamma\mathsf{\text{-}Conc}(e',\alpha)$ does not hold for some $\alpha$.
+$\textit{Auts}_{\Gamma} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $\Gamma\mathsf{\text{-}Condition}(e')$ holds but $e$ does not linearizable w.r.t $\Gamma$.
 
@@
 
-Given a data-differentiated $\_$-execution $e$ such that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Condition}(e)$ holds, Lemma \ref{lemma:automata for extended priority queue with single priority} is not enough for ensuring that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ for some $x$.
+Given a data-differentiated $\_$-execution $e$ such that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Condition}(e)$ holds, Lemma \ref{lemma:automata for extended priority queue with single priority} is not enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$.
 This is because that it is possible that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ does not holds because of interaction between actions of multiple priorities.
@@
 Here we assume that $p_1 \prec p_4$, $p_2 \prec p_4$ and $p_3 \prec p_4$.
-We can not find a sequence $s$ where $e \sqsubseteq s$ and $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(s,b)$ holds, since each time point from $\textit{cal}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is in interval of some item with smaller priority, and $\textit{rm}(b)$ can not happen in interval of item with smaller priority.
+We can not find a sequence $s$ where $e \sqsubseteq s$ and $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(s,b)$ holds, since each time point from $\textit{cal}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is in interval of some value with smaller priority, and $\textit{rm}(b)$ can not happen in interval of value with smaller priority.
 
@@
 
-In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we introduce the notion of left-right constraint, which is a graph, and the existence of cycle though item with maximal priority in it formally modelled the case in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
+In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we introduce the notion of left-right constraint, which is a graph, and the existence of cycle though value with maximal priority in it formally modelled the case in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
 For example, in \figurename~\ref{fig:introduce gap for EPQ1Lar}, $b$ is covered by $f$, $e$ and $d$.
-We then prove that getting rid of cycle though item with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$, or we can say, ensure $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ for some $x$.
+We then prove that getting rid of cycle though value with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$.
 
-When there is a cycle $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ through item $x$ with maximal priority in left-right constraint, we say that $x$ is covered by $d_1,\ldots,d_m$.
-Assume that $x$ is covered by $d_1,\ldots,d_m$, we can safely rename $x$ into $b$, rename $d_1,\ldots,d_m$ into $a$ and rename all other item into $c$ by data-independence.
+When there is a cycle $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ through value $x$ with maximal priority in left-right constraint, we say that $x$ is covered by $d_1,\ldots,d_m$.
+Assume that $x$ is covered by $d_1,\ldots,d_m$, we can safely rename $x$ into $b$, rename $d_1,\ldots,d_m$ into $a$ and rename all other value into $c$ by data-independence.
 Such execution can be recognized by witness automata, since between the first $\textit{ret}(\textit{put},a,\_)$ and the last $\textit{call}(\textit{rm},a)$ (if exists), $\textit{ret}(\textit{put},a,\_)$ and $\textit{call}(\textit{rm},a)$ occurs in pair and there is no need for count.
-In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we construct a set $\textit{Auts}_{\textit{1-lar}}$ of witness automata, and use $\textit{Auts}_{\textit{1-lar}}$ to prove that $\mathsf{MatchedMaxPriority}^{>}$ is co-regular, as stated by the following lemma.
+In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we construct a set $\textit{Auts}_{\textit{1-lar}}$ of register automata, and use $\textit{Auts}_{\textit{1-lar}}$ to prove that $\mathsf{MatchedMaxPriority}^{>}$ is co-regular, as stated by the following lemma.
 
@@
 
-\subsection{Co-Regular of $\textit{EPQ}_1^{=}$}
+\subsection{Co-Regular of $\mathsf{MatchedMaxPriority}^{=}$}
 \label{subsec:co-regular of EPQ1Equal}
 
-In this subsection, we briefly introduce the idea for proving co-regular of $\textit{EPQ}_1^{=}$.
+In this subsection, we briefly introduce the idea for proving co-regular of $\mathsf{MatchedMaxPriority}^{=}$.
 The proof of this subsection can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}.
 
-Given a data-differentiated $\_$-execution $e$ such that $\textit{last}(e) = \textit{EPQ}_1^{=}$.
-Lemma \ref{lemma:automata for extended priority queue with single priority} and $\textit{Auts}_{\textit{1-lar}}$ are sill not enough for ensure that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
-This is because that given items $a$ and $b$ with maximal priority, it is possible that all the possible linearization point of $\textit{rm}(b)$ are disabled by $\textit{rm}(b)$.
+Given a data-differentiated $\_$-execution $e$ such that $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)$ holds, Lemma \ref{lemma:automata for extended priority queue with single priority} and $\textit{Auts}_{\textit{1-lar}}$ are sill not enough for ensure that $e$ linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$.
+This is because that given values $a$ and $b$ with maximal priority, it is possible that all the possible linearization points of $\textit{rm}(b)$ are disabled by $\textit{rm}(b)$.
 We give an example of such execution $e$ in \figurename~\ref{fig:introduce pb order}.
-The execution $e$ of \figurename~\ref{fig:introduce pb order} is not linearizable w.r.t $\textit{EPQ}$, even if $h \vert_{p_1}$ and $h \vert_{p_4}$ both does, and either $a$ or $b$ is covered by items with smaller priority.
-The explanation of $e \not\sqsubseteq \textit{EPQ}$ is as follows: Since $\textit{put}(a,p_4) <_{\textit{hb}} \textit{put}(b,p_4)$, the linearization points of $\textit{rm}(a)$ should before the linearization point of $\textit{rm}(b)$.
-According to Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar}, the time intervals identified with dotted lines in \figurename~\ref{fig:introduce pb order} are are possible position to locate the linearization point of $\textit{rm}(b)$.
-However, each of them is before $\textit{cal}(\textit{rm},a)$.
+The execution $e$ of \figurename~\ref{fig:introduce pb order} is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$, even if $h \vert_{p_1}$ and $h \vert_{p_4}$ both linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$, and either $a$ or $b$ is covered by values with smaller priority.
+The explanation of $e$ not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ is as follows: Since $\textit{put}(a,p_4) <_{\textit{hb}} \textit{put}(b,p_4)$, the linearization points of $\textit{rm}(a)$ should before the linearization point of $\textit{rm}(b)$.
+According to the previous section, which says that linearizaton point of larger priority can be happen in the interval of values with smaller priority.
+Thus, the time intervals identified with dotted lines in \figurename~\ref{fig:introduce pb order} are are possible position to locate the linearization point of $\textit{rm}(b)$.
+However, each of them is before $\textit{call}(\textit{rm},a)$.
 
@@
 
-Let us introduce put-before order $<_{\textit{pb}}$ to formally state that ``an item is putted before another item''.
-Given a data-differentiated $\_$-execution $e$ and two items $a,b$ with maximal priority, $a <_{\textit{pb}} b$, if one of the following cases holds: (1) $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$, (2) $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(b)$, or (3) $\textit{rm}(a) <_{\textit{hb}} \textit{put}(b,\_)$.
+One point need to be modeled in \figurename~\ref{fig:introduce pb order} is that ``a is putted before b''.
+Let us introduce put-before order $<_{\textit{pb}}$ to formally state that ``an value is putted before another value''.
+Given a data-differentiated $\_$-execution $e$ and two values $a,b$ with maximal priority, $a <_{\textit{pb}} b$, if one of the following cases holds: (1) $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$, (2) $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(b)$, or (3) $\textit{rm}(a) <_{\textit{hb}} \textit{put}(b,\_)$.
 Sometimes we use $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$ and $a <_{\textit{pb}}^C b$ to explicitly distinguish above three cases.
 Let $<_{\textit{pb}}^*$ be the transitive closure of $<_{\textit{pb}}$.
-Intuitively, $a <_{\textit{pb}}^* b$ means that we can infer that $a$ should be putted earlier than $b$ with the help of several other items.
+Intuitively, $a <_{\textit{pb}}^* b$ means that we can infer that $a$ is putted before $b$ with the help of several intermediate values.
 
-Let us introduce gap-point to formally define the time point in dotted line of \figurename~\ref{fig:introduce pb order}.
+Another point need to be modeled in \figurename~\ref{fig:introduce pb order} is the possible time points of $\textit{rm}(b)$, which is formally modeled by gap-point.
 
 \begin{definition}\label{def:gap-point for matched put and rm operations}
-Given a data-differentiated $\textit{pri}_x$-execution $e$ and two method events $\textit{put}(x,\textit{pri}_x),\textit{rm}(x)$ of $e$.
-We say that a time-point $o$ is a gap-point of $x$, if $o$ is after $\textit{call}(\textit{put},x,\textit{pri}_x)$ and $\textit{call}(\textit{rm},x)$, is before $\textit{ret}(\textit{rm},x)$, and is not in interval of any items with smaller priority.
+Given a data-differentiated $p$-execution $e$ and two method events $\textit{put}(x,p),\textit{rm}(x)$ of $e$.
+We say that a time-point $o$ is a gap-point of $x$, if $o$ is after $\textit{call}(\textit{put},x,p)$ and $\textit{call}(\textit{rm},x)$, is before $\textit{ret}(\textit{rm},x)$, and is not in interval of any values with smaller priority.
 \end{definition}
 
-The case of \figurename~\ref{fig:introduce pb order} can be formally described as follows: $a <_{\textit{pb}}^* b$, while the right-most gap-point of $b$ is before $\textit{cal}(\textit{rm},a)$ or $\textit{cal}(\textit{put},a,p_4)$.
-The following lemma states that getting rid of such case is enough for ensure linearizable w.r.t $\textit{MS}(\textit{PQ}_1^{=})$.
+The case of \figurename~\ref{fig:introduce pb order} can be formally described as follows: $a <_{\textit{pb}}^* b$, while the right-most gap-point of $b$ is before $\textit{call}(\textit{rm},a)$ or $\textit{call}(\textit{put},a,p_4)$.
+The following lemma states that getting rid of such case is enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$.
 
@@
 \label{lemma:EPQ1Equal as pb order and gap-point}
-Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{=}$.
-$e \not\sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$, if and only if there exists $x$ and $y$ with maximal priority $\textit{pri}$ in $e$, such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is before $\textit{cal}(\textit{put},y,\textit{pri})$ or $\textit{cal}(\textit{rm},y)$ in $e$.
+Given a data-differentiated $p$-execution $e$ where $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)$ holds.
+$e$ is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$, if and only if there exists $x$ and $y$ with maximal priority $p$, such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is before $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$ in $e$.
 \end{restatable}
@@
 
-Let $\textit{Items}(e,p)$ be the set of items with priority $p$ in $e$.
-We introduce another lemma (Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal} in Appendix), which states that: If $\exists g_1 \in \textit{Items}(e,\textit{pri})$, such that $\forall g_2 \in \textit{Items}(e,\textit{pri})$, (1) $g_1 \nless_{\textit{pb}} g_2$, and (2) the right-most gap-point of $g_1$ is after $\textit{cal}(\textit{put},g_2,\textit{pri})$ and $\textit{cal}(\textit{rm},g_2)$.
-Then $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
-%The proof of this lemma also tell us how to construct linearization in such case.
-The remaining of our proof is a loop for searching $g_1$ in $e$.
-Initially we set $a_1$ to be the last inserted item in the $l_{\textit{pri}}$, linearization of the projection of $e$ into $\textit{pri}$.
+Let $\textit{Items}(e,p)$ be the set of values with priority $p$ in $e$.
+We introduce another lemma (Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal} in Appendix), which states that if we can find a value $\exists g_1 \in \textit{Items}(e,p)$, such that $\forall g_2 \in \textit{Items}(e,p)$, (1) $g_1 \nless_{\textit{pb}} g_2$, and (2) the right-most gap-point of $g_1$ is after $\textit{call}(\textit{put},g_2,p)$ and $\textit{call}(\textit{rm},g_2)$.
+Then $e$ is linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$.
+The remaining of our proof is a loop for searching such $g_1$ in $e$.
+Let $e_p$ be the projection of $e$ into values of priority $p$, and let $l_p$ be the linearization of $e$ which satisfy the requirements of queue.
+Let the initial value of $a_1$ be the last inserted value in $l_p$.
 
@@
 \setlength{\itemsep}{0.5pt}
-%\item[-] We start with $a_1$, the last inserted item in the linearization of the projection of $e$ into $\textit{pri}$.
 \item[-] check if $a_1$ satisfy the conditions of Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}. If it does, then this loop terminates.
 
-\item[-] Otherwise, $\exists a_2 \in \textit{Items}(e,\textit{pri})$, such that the rightmost gap-point of $a_1$ is before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$ in $e$.
+\item[-] Otherwise, $\exists a_2 \in \textit{Items}(e,p)$, such that the rightmost gap-point of $a_1$ is before $\textit{cal}(\textit{put},a_2,p)$ or $\textit{cal}(\textit{rm},a_2)$ in $e$.
 By assumption, we know that $a_2 \nless_{\textit{pb}} a_1$, and each gap-point of $a_2$ is after the rightmost gap-point of $a_1$.
@@
     \setlength{\itemsep}{0.5pt}
-    \item[-] If $\forall b \in \textit{Items}(e,\textit{pri})$, $a_2$ does not $<_{\textit{pb}}$ to $b$. Then we go to next round of the loop and make $a_2$ to be the ``new-round $a_1$''.
-    \item[-] Otherwise, there exists $a_3 \in \textit{Items}(e,\textit{pri})$ such that $a_2 <_{\textit{pb}}^* a_3$.
-Since $l_{\textit{pri}}$ has FIFO property, there is no cycle in $<_{\textit{pb}}$, and it is safe to assume that $a_3$ is maximal w.r.t $<_{\textit{pb}}^*$.
+    \item[-] If $\forall b \in \textit{Items}(e,p)$, $a_2\not<_{\textit{pb}}b$. Then we go to next round of the loop and make $a_2$ to be the ``new-round $a_1$''.
+    \item[-] Otherwise, there exists $a_3 \in \textit{Items}(e,p)$ such that $a_2 <_{\textit{pb}}^* a_3$.
+It is not hard to see that there is no cycle in $<_{\textit{pb}}$, and it is safe to assume that $a_3$ is maximal w.r.t $<_{\textit{pb}}^*$.
 
-        By assumption,we know that the rightmost gap-point of $a_3$ is after $\textit{cal}(\textit{put},a_2,\textit{pri})$ and $\textit{cal}(\textit{rm},a_2)$.
+        By assumption,we know that the rightmost gap-point of $a_3$ is after $\textit{call}(\textit{put},a_2,p)$ and $\textit{call}(\textit{rm},a_2)$.
 Therefore, we can see that the rightmost gap-point of $a_3$ is after the rightmost gap-point of $a_1$.
@@
 Therefore, the loop finally terminates at some $a^f$ that satisfies the condition of Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}.
-This implies that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
+This implies that $e$ is linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ and completes the proof of the $\textit{only if}$ direction.
 \qed
@@
 
-The result of Lemma \ref{lemma:EPQ1Equal as pb order and gap-point} is not quite suitable for verificaion with automata, because obtaining $y <_{\textit{pb}}^* x$ may requires arbitrary many intermediate items, and it is hard to store unbounded items by automata.
-Fortunately, we prove that the number of intermediate items $a_i$ is essentially bounded, as stated by the following lemma.
+The result of Lemma \ref{lemma:EPQ1Equal as pb order and gap-point} still needs to be impove, since obtaining $y <_{\textit{pb}}^* x$ may requires arbitrary many intermediate values, and it is hard to store unbounded values by automata.
+Fortunately, we prove that the number of intermediate values is essentially bounded, as stated by the following lemma.
 
@@
 \label{lemma:ob order has bounded length}
-Given a data-differentiated execution $h$. Assume that $a <_{\textit{pb}} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}} a_m <_{\textit{pb}} b$, then one of the following cases holds:
+Given a data-differentiated execution $e$. Assume that $a <_{\textit{pb}} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}} a_m <_{\textit{pb}} b$, then one of the following cases holds:
 
@@
 
-By Lemma \ref{lemma:ob order has bounded length}, we can use witness automata to detect $a <_{\textit{pb}}^* b$ by enumerating all possible enumerations of operations of $a$, $b$ and possibly $a_1$.
-Since some combinations of $<_{\textit{pb}}^*$ and gap-points are conflict, we finally reduce the number of potential enumerations of operations of $a$, $b$ and possibly $a_1$ into only five, as shown by the following lemma:
+By Lemma \ref{lemma:ob order has bounded length}, we can use register automata to detect $a <_{\textit{pb}}^* b$ by enumerating all possible enumerations of operations of $a$, $b$ and possibly $a_1$.
+Since some combinations of $<_{\textit{pb}}^*$ and gap-points are conflict, we finally reduce the number of potential enumerations of operations of $a$, $b$ and possibly $a_1$ into only five in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}.
 
-\begin{restatable}{lemma}{FiveEnmuerationisEnoughForEPQOneEqual}
-\label{lemma:five enumeration is enough for EPQ1Equal}
-Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{=}$.
-Let $a$ and $b$ be items with maximal priority $\textit{pri}$.
-Assume that $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\textit{pri})$ or $\textit{cal}(\textit{rm},a)$.
-Then, there are five possible enumeration of method events of $a$, $b$, $a_1$ (if exists), where $a_1$ is the possible intermediate items for obtain $a <_{\textit{pb}}^* b$.
-\end{restatable}
-
-In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}, we construct a set $\textit{Auts}_{\textit{1-eq}}$ of witness automata, and prove that $\textit{EPQ}_1^{=}$ is co-regular, as stated by the following lemma.
+In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}, we construct a set $\textit{Auts}_{\textit{1-eq}}$ of register automata, and prove that $\mathsf{MatchedMaxPriority}^{=}$ is co-regular, as stated by the following lemma.
 
@@
 \label{lemma:EPQ1Equal is co-regular}
-$\textit{EPQ}_1^{=}$ is co-regular.
+$\mathsf{MatchedMaxPriority}^{=}$ is co-regular.
 \end{restatable}
 
-A enumeration of Lemma \ref{lemma:five enumeration is enough for EPQ1Equal} and its witness automata is shown in \figurename~\ref{fig:an enumeration and its witness automaton}.
-$o$ is the rightmost gap-point of $b$, we rename the items that ``covers the time interval from $\textit{cal}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$'' (see Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}) into $d$, and rename the renaming items into $e$.
-In this figure, $c = \textit{cal}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{cal}(\textit{rm},e), \textit{ret}(\textit{rm},e),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.
+One such enumeration and its register automata is shown in \figurename~\ref{fig:an enumeration and its witness automaton}.
+$o$ is the rightmost gap-point of $b$, we rename the values that ``covers the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$'' (see Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}) into $d$, and rename the renaming values into $e$.
+In this figure, $c = \textit{call}(\textit{put},e,\top),\textit{ret}(\textit{put},e,\top)$, $\textit{call}(\textit{rm},e), \textit{ret}(\textit{rm},e),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{call}(\textit{put},d,<r)$, $c_2 = c_1 + \textit{ret}(\textit{put},b,r)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b,r) + \textit{ret}(\textit{rm},d)$.
 
@@
 %\vspace{-10pt}
-  \caption{An enumerations and its witness automaton}
+  \caption{An enumerations and its register automaton}
   \label{fig:an enumeration and its witness automaton}
@@
 
-\subsection{Combine Step-by-Step linearizability and Co-Regular}
+\subsection{Reduce into State Reachability}
 \label{subsec:combine step-by-step linearizability and co-regular}
 
-We also prove that $\textit{EPQ}_2^{>}$, $\textit{EPQ}_2^{=}$ and $\textit{EPQ}_3$ are co-regular in Appendix \ref{subsec:appendix co-regular of EPQ2Lar}, Appendix \ref{subsec:appendix co-regular of EPQ2Equal} and Appendix \ref{subsec:co-regular of EPQ3}, respectively.
-Then we can see that $\textit{EPQ}$ is co-regular.
+We also prove that $\mathsf{UnmatchedMaxPriority}$ are co-regular in Appendix \ref{subsec:appendix co-regular of EPQ2Lar}, Appendix \ref{subsec:appendix co-regular of EPQ2Equal}, and prove that $\mathsf{EmptyRemove}$ are co-regular in Appendix \ref{subsec:co-regular of EPQ3}.
+Then we can see that $\seqPQ$ is co-regular.
 
@@
 \label{lemma:EPQ is co-regular}
-$\textit{EPQ}$ is co-regular.
+$\seqPQ$ is co-regular.
 \end{restatable}
 
-By Lemma \ref{lemma:EPQ as multi in MRpri for history} and Lemma \ref{lemma:EPQ is co-regular}, we can finally reduce the verification of linearizability w.r.t $\textit{EPQ}$ into a reachability problem, as illustrated by the following theorem, where $\textit{Auts}_{\textit{EPQ}}$ is the set of witness automata of this section.
+By Lemma \ref{lemma:EPQ as multi in MRpri for history} and Lemma \ref{lemma:EPQ is co-regular}, we can finally reduce the verification of linearizability w.r.t $\seqPQ$ into a reachability problem, as illustrated by the following theorem, where $\textit{Auts}_{\textit{EPQ}}$ is the set of witness automata of this section.
 
@@
 \label{lemma:reduce EPQ into state reachability}
-Given a data-independence implementation $\mathcal{I}$. $\mathcal{I} \sqsubseteq \textit{EPQ}$, if and only if, $\mathcal{I} \cap \textit{Auts}_{\textit{EPQ}} = \emptyset$.
+Given a data-independence implementation $\mathcal{I}$. $\mathcal{I} \sqsubseteq \seqPQ$, if and only if, $\mathcal{I} \cap \textit{Auts}_{\textit{EPQ}} = \emptyset$.
 \end{restatable}
+
+When the data domain and priority domain are both fixed to be finite set, the register automata can be simulated with finite automata, and by Lemma \ref{lemma:reduce EPQ into state reachability}, checking linearizable w.r.t $\seqPQ$ is decidable.

### 2017-04-20 priority-queues.tex
modify
Section 2

--- 2017-04-18 Seq-EPQ-dataInd.tex
@@
 Also, when the set of values stored in the priority queue is empty, $\textit{rm}$ returns the distinguished value $\textit{empty}$.
-Concurrent priority queues allow the methods $\textit{put}$ and $\textit{rm}$ to be called concurrently from different threads.
+Concurrent implementations of the priority queue allow the methods $\textit{put}$ and $\textit{rm}$ to be called concurrently from different threads.
 However, every method invocation should give the illusion that it takes place instantaneously at some point between its invocation and its return.
-We formalize (concurrent) executions and implementations in Section~\ref{ssec:exec}, Section~\ref{ssec:semantic_prop} introduces a set of properties satisfied by all the  implementations we are aware of, and Section~\ref{ssec:lin} defines the standard correctness criterion for concurrent ADTs such as the priority queue known as \emph{linearizability}\cite{journals/toplas/HerlihyW90}.
+We formalize (concurrent) executions and implementations in Section~\ref{ssec:exec}, Section~\ref{ssec:semantic_prop} introduces a set of properties satisfied by all the  implementations we are aware of, and Section~\ref{ssec:lin} defines the standard correctness criterion for concurrent implementations of ADTs known as \emph{linearizability}~\cite{journals/toplas/HerlihyW90}.
 
@@
 The latter are used to match call and return actions of the same invocation.
-Call actions $\textit{call}_o(\textit{put},a,p)$ and $\textit{call}_o(\textit{rm},a')$ with $a\in \mathbb{D}$, $a'\in \mathbb{D}\cup\{\textit{empty}\}$, $p \in \mathbb{P}$, and $o \in \mathbb{O}$, combine a method name and a set of parameters with an operation identifier.
+Call actions $\textit{call}_o(\textit{put},a,p)$ and $\textit{call}_o(\textit{rm},a')$ with $a\in \mathbb{D}$, $a'\in \mathbb{D}\cup\{\textit{empty}\}$, $p \in \mathbb{P}$, and $o \in \mathbb{O}$, combine a method name and a set of arguments with an operation identifier.
 The return value of a remove is transformed to an argument value for uniformity~\footnote{Method return values are guessed nondeterministically, and validated at return points.
@@
 For example, given two priorities $p_1 \prec p_2$, $\textit{put}(a,p_2) \cdot \textit{put}(b,p_1) \cdot \textit{rm}(b)$ is a sequential execution of the priority queue ($\textit{rm}$ returns $b$ because it has smaller priority).
-The set of all the possible sequential executions (obtained by adding and removing arbitrarily values with arbitrary priorities) is denoted by $\seqPQ$.
+
+We define $\seqPQ$, the set of sequential priority queue executions, semantically via labelled transition system ($LTS$).
+An LTS is a tuple $A=(Q,\Sigma,\rightarrow,q_0)$, where $Q$ is a set of states, $\Sigma$ is an alphabet of transition labels, $\rightarrow\subseteq Q\times\Sigma\times Q$ is a transition relation and $q_0$ is the initial state.
+We model priority queue as an LTS $\textit{PQ}$ defined as follows: each state of $\textit{PQ}$ is a mapping associating priorities in $\mathbb{P}$ with sequences of values in $\mathbb{D}$ representing a snapshot of the priority queue (for each priority, the values are ordered as they were inserted); the transition labels are $\textit{put}(a,p)$ and $\textit{rm}(a)$ operations; Each transition modifies the state as expected.
+For example, $q_1 \xrightarrow{\textit{rm}(\textit{empty})} q_2$ if $q_1 = q_2$, and $q_1$ and $q_2$ map each priority to the empty sequence $\epsilon$.
+Then, $\seqPQ$ is the set of words accepted by $\textit{PQ}$.
+The detailed definition of $\textit{PQ}$ can be found in Appendix \ref{A}.
 
@@
 
-TODO MOTIVATE THE FOLLOWING
-
-The projection $e \vert D$ of an execution $e$ to a set of values $D \subseteq \mathbb{D}$ is obtained from $e$ by erasing all actions with a data value not in $D$.
+Besides data independence, the sequential behaviors of the priority queue satisfy the following closure property: a behavior remains valid when removing all the operations with an argument in some set of values $D \subseteq \mathbb{D}$ and any $\textit{rm}(\textit{empty})$ operation (since they are read-only and they don't affect the queue's state).
+In order to distinguish between different $\textit{rm}(\textit{empty})$ operations while simplifying the technical exposition, we assume that they receive as argument a value, i.e., call actions are of the form $\textit{call}_o(\textit{rm},\textit{empty},a)$ for some $a\in \mathbb{D}$.
+We will make explicit this argument only when needed in our technical development.
+The projection $e \vert D$ of an execution $e$ to a set of values $D \subseteq \mathbb{D}$ is obtained from $e$ by erasing all call/return actions with an argument in $D$.
 We write $e \setminus x$ for the projection $e \vert_{ \mathbb{D} \setminus \{ x \} }$.
@@
 Intuitively this means that methods can always return without any help from outside threads, avoiding deadlock.
-
-TODO WORK THE FOLLOWING
-
-Given execution $e$, the happen-before relation $<_{\textit{hb}}$ is defined as follows: $o_1 <_{\textit{hb}} o_2$, if the return action of $o_1$ is before the call action of $o_2$ in $e$.
-The happens-before relation are interval orders \cite{DBLP:conf/popl/BouajjaniEEH15}: for distinct $o_1,o_2,o_3,o_4$, if $o_1 <_{\textit{hb}} o_2$ and $o_3 <_{\textit{hb}} o_4$, then either $o_1 <_{\textit{hb}} o_4$, or $o_3 <_{\textit{hb}} o_2$.
+We recall the notion of \emph{linearizability}~\cite{journals/toplas/HerlihyW90} which is the \emph{de facto} standard correctness condition for concurrent data structures.
+Given an execution $e$, the happen-before relation $<_{\textit{hb}}$ between operations~\footnote{In general, we refer to operations using their identifiers.} is defined as follows: $o_1 <_{\textit{hb}} o_2$, if the return action of $o_1$ occurs before the call action of $o_2$ in $e$.
+The happens-before relation is an interval order \cite{DBLP:conf/popl/BouajjaniEEH15}: for distinct $o_1,o_2,o_3,o_4$, if $o_1 <_{\textit{hb}} o_2$ and $o_3 <_{\textit{hb}} o_4$, then either $o_1 <_{\textit{hb}} o_4$, or $o_3 <_{\textit{hb}} o_2$.
 Intuitively, this comes from the fact that concurrent threads share a notion of global time.
 
-Linearizability \cite{journals/toplas/HerlihyW90} is the \emph{de facto} standard correctness condition for concurrent data structures.
-Given concurrent execution $e$ and sequential execution $s$ of length $n$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if they have same set of operations, and there is a bijection $f$ between operations s.t.
+Given a (concurrent) execution $e$ and a sequential execution $s$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if there is a bijection $f: O_1 \rightarrow O_2$, where $O_1$ and $O_2$ are the set of operations of $e$ and $s$, respectively, such that (1) $o$ and $f(o)$ have the same call and return actions, and (2) if $o_1 <_{\textit{hb}} o_2$, then $f(o_1) <_{\textit{hb}} f(o_2)$.
+A (concurrent) execution $e$ is linearizable w.r.t a set $S$ of sequential executions, denoted $e \sqsubseteq S$, if there exists $s \in S$ such that $e \sqsubseteq s$.
+A set of concurrent executions $E$ is linearizable w.r.t $S$, denoted $E \sqsubseteq S$, if $e \sqsubseteq S$ for all $e \in E$.
 
-\begin{itemize}
-\item[-] if $o_1 <_{\textit{hb}} o_2$, then $f(o_1) <_{\textit{hb}} f(o_2)$,
-
-\item[-] $o$ and $f(o)$ have same call and return actions.
-\end{itemize}
-
-A concurrent execution $e$ is linearizable w.r.t a set $S$ of sequential executions, denoted $e \sqsubseteq S$, if there exists $s \in S$ such that $e \sqsubseteq s$.
-A set of concurrent executions $E$ is linearizable w.r.t $S$, denoted $E \sqsubseteq S$, if $e \sqsubseteq S$ for all $e \in H$.
-
-The following lemma states that with the help of data-independence, it is enough to consider only data-differentiated executions when checking linearizability.
-This is similar to that in \cite{Abdulla:2013}, where they use the notion of data-independence in \cite{Wolper:1986}.
+The following lemma states that by data-independence, it is enough to consider only data-differentiated executions when checking linearizability.
+This is similar to that in \cite{conf/tacas/AbdullaHHJR13,DBLP:conf/icalp/BouajjaniEEH15}, where they use the notion of data-independence in \cite{conf/popl/Wolper86}.
 Thus, in the remainder of the paper, we focus on characterizing linearizability for data-differentiated executions.
-The proof of this lemma can be found in Appendix \ref{sec:appendix in section data-independence of EPQ}.
+%The proof of this lemma can be found in Appendix \ref{sec:appendix in section data-independence of EPQ}.
 
@@
 \label{lemma:data differentiated is enough for PQ}
-A data-independent implementation $\mathcal{I}$ is linearizable w.r.t data-independent set $S$ of sequential executions, if and only if $\mathcal{I}_{\neq}$ is linearizable with respect to $S_{\neq}$.
+A data-independent implementation $\mathcal{I}$ is linearizable w.r.t a data-independent set $S$ of sequential executions, if and only if $\mathcal{I}_{\neq}$ is linearizable w.r.t. $S_{\neq}$.
 \end{restatable}

### 2017-04-20 checking-executions.tex
small stuff

@@
 
-\subsection{Checking Data-Differentiated Sequential Executions}\label{ssec:seq_exec}
+\subsection{Characterizing Data-Differentiated Sequential Executions}\label{ssec:seq_exec}
 
@@
 
-\subsection{Checking Data-Differentiated Concurrent Executions}\label{ssec:conc_exec}
+\subsection{Checking Linearizability of Data-Differentiated Concurrent Executions}\label{ssec:conc_exec}
 

### 2017-04-20 introduction.tex
nodify

@@
 \noindent {\bf Related work}
-%\label{sec:related work}
-\noindent Related work
+\noindent The most related work of our paper is ~\cite{DBLP:conf/icalp/BouajjaniEEH15}.
+Compared to ~\cite{DBLP:conf/icalp/BouajjaniEEH15}, our work are specific to priority queue and thus have points not occur in queues and stacks.
+Due to the partial-order priorities, to prove the correctness of recursive procedure, we need to alter the positions of linearization points of incomparable items while keeps the linearization points of comparable items unchanged.
+For checking violations of one local property, we just need to find a value, such that at every possible time point to locate its remove operation, there is already values with smaller priorities in priority queue.
+For checking violations of another local property, we just need to find two values with maximal priority, such that every possible time point to locate remove operation of one value is disabled by remove operation of another value.
+Such phenomenons are unique for priority queue.

### 2017-04-20 appendix-EPQ.tex
co-regularity

@@
 
-\AutoForEPQwithSignlePri*
+\begin{lemma}\label{lemma:automata for extended priority queue with single priority}
+Given a data-independent implementations $\mathcal{I}$ of extended priority queue, $\mathcal{I} \cap \textit{Auts}_{\textit{sinPri}} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{toQueue}(e')$ does not linearizable to queue.
+\end{lemma}
+
+%\begin{restatable}{lemma}{AutoForEPQwithSignlePri}
+%\label{lemma:automata for extended priority queue with single priority}
+%
+%Given a data-independent implementations $\mathcal{I}$ of extended priority queue, $\mathcal{I} \cap \textit{Auts}_{\textit{sinPri}} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{toQueue}(e')$ does not linearizable to queue.
+%\end{restatable}
+
+%\AutoForEPQwithSignlePri*
 

### 2017-04-20 checking-executions.tex
co-regularity

@@
 
+TODO REPHRASE THIS
 While checking linearizability of a given execution is known to be NP-complete in general~\cite{journals/siamcomp/GibbonsK97} we show in this section that it is polynomial-time for data-differentiated executions of the priority queue.
@@
 
-\If {$\mathsf{HasEmptyRemoves}(e)$}
+\If {$\mathsf{Has\text{-}EmptyRemoves}(e)$}
 {
@@
 }
-\ElseIf{$\mathsf{HasUnmatchedMaxPriority}(e)$}
+\ElseIf{$\mathsf{Has\text{-}UnmatchedMaxPriority}(e)$}
 {
@@
 \begin{align*}
-\mathsf{HasEmptyRemoves}(e)=\mathsf{true} & \mbox{ iff  $e$ contains a $\textit{rm}(\textit{empty})$ operation} \\
-\mathsf{HasUnmatchedMaxPriority}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{unmatched-priorities}(e)$ for a maximal priority} \\
+\mathsf{Has\text{-}EmptyRemoves}(e)=\mathsf{true} & \mbox{ iff  $e$ contains a $\textit{rm}(\textit{empty})$ operation} \\
+\mathsf{Has\text{-}UnmatchedMaxPriority}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{unmatched-priorities}(e)$ for a maximal priority} \\
 &\hspace{4mm}\mbox{$p\in priorities(e)$}
@@
 A put operation which is not unmatched is called \emph{matched}.
-For readability, we consider the following syntactic sugar $\mathsf{HasMatchedMaxPriority}(e)=\neg \mathsf{HasEmptyRemoves}(e)\land \neg \mathsf{HasUnmatchedMaxPriority}(e)$.
+For simplicity, we consider the following syntactic sugar $\mathsf{Has\text{-}MatchedMaxPriority}(e)=\neg \mathsf{Has\text{-}EmptyRemoves}(e)\land \neg \mathsf{Has\text{-}UnmatchedMaxPriority}(e)$.
+By an abuse of notation, we also assume that  $\mathsf{Has\text{-}UnmatchedMaxPriority}(e) \Rightarrow \neg \mathsf{Has\text{-}EmptyRemoves}(e)$ (this is sound by the order of the conditionals in $\textit{Check-PQ-Seq}$.
 
@@
 
-\begin{figure}[htbp]
+\begin{figure}[t]
   \centering
@@
 \end{figure}
+
+Section~\ref{sec:co-regular of extended priority queues} introduces a characterization of concurrent priority queue violations using a set of \emph{non-recursive} automata (i.e., whose states consist of a fixed number of registers) whose product is equivalent to $\textit{Check-PQ-Conc}$ (modulo renaming of values which is possible by data-independence).
+Since $\seqPQ$ is closed under projection (Lemma~\ref{lem:closure_proj}), the recursion in $\textit{Check-PQ-Conc}$ can be eliminated by checking that each projection of a given execution $e$ passes a non-recursive version of $\textit{Check-PQ-Conc}$ where every recursive call is replaced by $\mathsf{true}$.
+More precisely, every occurrence of $\text{{\bf return}}\ \textit{Check-PQ-Conc}$ is replaced by  $\text{{\bf return}}\ \mathsf{true}$.
+Let $\textit{Check-PQ-Conc-NonRec}$ be the thus obtained procedure.
+
+\begin{restatable}{lemma}{EPQasMultiInMRpriforHistory}
+\label{lemma:EPQ as multi in MRpri for history}
+Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$ if and only if for each $e' \in \textit{proj}(e)$, $\textit{Check-PQ-Conc-NonRec}(e)$ returns $\mathsf{true}$.
+\end{restatable}
+
+%The following section shows that the linearizability check in the above lemma can be implemented using a set of register automata.
+
+%In this section, we divide checking linearizability w.r.t $\seqPQ$ into checking linearizablity w.r.t $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$, each of which can be done in polynomial-time.
+Due to data-independence, violations to each $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$ can be captured by register automata.
+This reduces checking linearizability w.r.t $\seqPQ$ into state reachability problem.
+
+%\begin{restatable}{lemma}{EPQasMultiInMRpriforHistory}
+%\label{lemma:EPQ as multi in MRpri for history}
+%Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$ if and only if for each $e' \in \textit{proj}(e)$, $\mathsf{Has\text{-}}\Gamma(e')$ implies $\Gamma\mathsf{\text{-}Conc}(e')$.
+%
+%$\forall \Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$,  $\Gamma\mathsf{\text{-}Condition}(e)$ $\Rightarrow$ $e$ is linearizable w.r.t $\Gamma$. 
+%\end{restatable}
+%
+%
+%
+%\subsection{Co-Regular}
+%\label{subsec:definition of co-regular}
+%
+%Let $\Gamma\mathsf{\text{-}Condition}(e)$ be $\mathsf{Has\text{-}EmptyRemoves}(e)$, $\neg \mathsf{Has\text{-}EmptyRemoves}(e) \wedge \mathsf{HasUnmatchedMaxPriority}(e)$, and $\mathsf{HasMatchedMaxPriority}(e)$, for $\Gamma = \mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}$, respectively.
+$\textit{Check-PQ-Conc}$ inspires us that, when $\Gamma\mathsf{\text{-}Condition}(e)$ holds, there should exist some $\alpha$, such that  $\Gamma\mathsf{\text{-}Conc}(e,\alpha)$ holds.
+The following lemma states this formally.
+Let $\textit{proj}(e)$ be the set of projections of $e$ into set of values.
+When refer to $\textit{proj}(e)$, we implicitly assume that each $\textit{rm}(\textit{empty})$ in $e$ has a ghost argument that is unique.
+We say that execution $l$ is linearizable w.r.t $\Gamma$, if we can find a sequence $s$ where $e \sqsubseteq s$ and $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$ holds for some $\alpha$.
+%
+%\begin{restatable}{lemma}{EPQasMultiInMRpriforHistory}
+%\label{lemma:EPQ as multi in MRpri for history}
+%Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$, if and only if, $\forall e' \in \textit{proj}(e)$, $\forall \Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$,  $\Gamma\mathsf{\text{-}Condition}(e)$ $\Rightarrow$ $e$ is linearizable w.r.t $\Gamma$.
+%\end{restatable}

### 2017-04-20 co-regular-EPQ.tex
co-regularity

@@
-\section{Reducing Linearizability of Priority Queues into State Reachability}
+%!TEX root = draft.tex
+\section{Reducing Linearizability of Priority Queues to Reachability}
 \label{sec:co-regular of extended priority queues}
 
-In this section, we divide checking linearizability w.r.t $\seqPQ$ into checking linearizablity w.r.t $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$, each of which can be done in polynomial-time.
+We show that the set of executions for which some projection fails the test $\textit{Check-PQ-Conc-NonRec}$ can be characterized using a set of register automata, modulo a value renaming.
+The possibility of renaming values (which is complete for checking data independent implementations) allows to simplify the reasoning about projections.
+Thus, we assume that all the operations which are not in the projection failing this test use the same distinguished value $\top$, different from those used in the projection.
+Then, it is enough to find an automata characterization for the set of executions $e$ for which $\textit{Check-PQ-Conc-NonRec}$ fails, or equivalently, for which one
+%a projection $e'$ fails $\textit{Check-PQ-Conc-NonRec}$ when one 
+of the following three formulas is false:
+\begin{align*}
+\hspace{-5mm}
+\Gamma(e) := \mathsf{Has\text{-}}\Gamma(e) \Rightarrow \exists \alpha.\ \Gamma\mathsf{\text{-}Conc}(e,\alpha)\mbox{ with $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$}
+\end{align*}
+Extending the automaton corresponding to some $\Gamma$ with self-loops that allow any operation with parameter $\top$ results in an automaton satisfying the following property called \emph{$\Gamma$-completeness}.
+
+\begin{definition}
+For $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$, an automaton $A$ is called \emph{$\Gamma$-complete} when for each data-independent implementation $\mathcal{I}$:
+
+$A \cap \mathcal{I} \neq \emptyset$ if and only if there exists $ e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$ such that $\Gamma(e')$ doesn't hold.
+\end{definition}
+
+The following result shows that deriving $\Gamma$-complete automata for each $\Gamma$ enables an effective reduction of checking linearizability of concurrent priority queue implementations to state reachability.
+It is a direct consequence of the above definitions.
+
+\begin{restatable}{theorem}{ReduceEPQIntoStateReachability}
+\label{lemma:reduce EPQ into state reachability}
+Let $\mathcal{I}$ be a data-independent implementation, and $A(\Gamma)$ a $\Gamma$-complete automaton for each $\Gamma$. Then,
+$\mathcal{I} \sqsubseteq \seqPQ$ if and only if $\mathcal{I} \cap A(\Gamma) = \emptyset$ for all $\Gamma$.
+\end{restatable}
+
+Section~\ref{ssec:aut} describes a $\mathsf{MatchedMaxPriority}$-complete automaton (TODO the other automata are discussed in Appendix~\ref{} or Icalp??) while Section~\ref{} discusses decidability results implied by the theorem above.
+
+W.l.o.g., we assume that every implementation $\mathcal{I}$ behaves correctly, i.e., as a FIFO queue, when only values with the same priority are observed.
+More precisely, we assume that for every execution $e\in\mathcal{I}$ and every priority $p\in\mathbb{P}$, the projection of $e$ to values with priority $p$ is linearizable.
+This property can be checked separately using the techniques introduced for regular FIFO queues in~\cite{DBLP:conf/icalp/BouajjaniEEH15} (see Appendix~\ref{} for more details).
+Note that this assumption excludes some obvious violations, such as a $\textit{rm}(a)$ operation happens before a $\textit{put}(a,p)$ operation, for some $p$.
+
+Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions where any two occurring priorities are comparable (w.r.t.
+$\prec$).
+This is possible because any data-differentiated execution for which $\Gamma(e)$ is false has such a projection.
+Formally, given a data-differentiated execution $e$ and $p$ a maximal priority in $e$, $e\vert_{\preceq p}$ is the projection of $e$ to the set of values with priorities comparable to $p$.
+Then,
+
+\begin{restatable}{lemma}{priExecutionIsEnough}
+\label{lemma:pri execution is enough}
+Let $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$ and $e$ be a data-differentiated execution.
+Then, $\Gamma(e)$ is false if and only if $\Gamma(e\vert_{\preceq p})$ is false for some maximal priority $p$ in $e$.
+\end{restatable}
+\begin {proof} (Sketch)
+TODO UPDATE THE FOLLOWING
+To prove the $\textit{only if}$ direction, given $e \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
+Since there is no restriction to $p$-incomparable operations in $u$, $v$ and $w$, we can safely erase all $p$-incomparable values.
+Similar case holds for $\Gamma=\mathsf{UnmatchedMaxPriority}$.
+
+To prove the $\textit{if}$ direction, we use a lemma in Appendix, which said that if the projection of an execution into a set $O_1$ of operations has linearization $l_1$ and the projection of the execution into $O_2$, the remanning operations, has linearization $l_2$, then the execution has a linearization whose projection into $O_1$ is $l_1$ and projection into $O_2$ is $l_2$.
+
+Given $e\vert_{\preceq p} \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
+By above lemma we know that $e$ has a linearization $l' = u' \cdot \textit{put}(x,p) \cdot v' \cdot \textit{rm}(x) \cdot w'$, such that the projection of $l'$ into $p$-incomparable values is l.
+We can see that $u$, $v$ and $w$ are sub-sequence of $u'$, $v'$ an $w'$, respectively.
+Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$ does not have restriction to $p$-incomparable values, we can see that $l' \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
+\qed
+\end {proof}
+
+%When the data domain and priority domain are both fixed to be finite set, the register automata can be simulated with finite automata, and by Lemma \ref{lemma:reduce EPQ into state reachability}, checking linearizable w.r.t $\seqPQ$ is decidable.
+
+%
+%
+% and 
+%
+%Let $A(\Gamma)$ be a $\Gamma$-complete automaton for each $\Gamma$. The following theorem 
+%
+%
+%We will describe this automaton only in the case of $\Gamma=\mathsf{MatchedMaxPriority}$. The rest can be found in Appendix~\ref{}. These automata satisfy the following property 
+%
+%the main technical challenge is to show that each of the following sets of executions can be characterized in terms of automata (modulo renaming of values):
+%
+%
+% define a set of register automata whose product with a given implementation $\mathcal{I}$ accepts a linearizability violation whenever $\mathcal{I}$ is not linearizable.
+%
+%are precise enough to recognize every non-linearizable implementation.
+%
+%In this section, we divide checking linearizability w.r.t $\seqPQ$ into checking linearizablity w.r.t $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$, each of which can be done in polynomial-time.
 Due to data-independence, violations to each $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$ can be captured by register automata.
 This reduces checking linearizability w.r.t $\seqPQ$ into state reachability problem.
+%
+%
+%
+%\subsection{Co-Regular}
+%\label{subsec:definition of co-regular}
+%
 
-\subsection{Co-Regular}
-\label{subsec:definition of co-regular}
-
-Let $\Gamma\mathsf{\text{-}Condition}(e)$ be $\mathsf{HasEmptyRemoves}(e)$, $\neg \mathsf{HasEmptyRemoves}(e) \wedge \mathsf{HasUnmatchedMaxPriority}(e)$, and $\mathsf{HasMatchedMaxPriority}(e)$, for $\Gamma = \mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}$, respectively.
+%Let $\Gamma\mathsf{\text{-}Condition}(e)$ be $\mathsf{HasEmptyRemoves}(e)$, $\neg \mathsf{HasEmptyRemoves}(e) \wedge \mathsf{HasUnmatchedMaxPriority}(e)$, and $\mathsf{HasMatchedMaxPriority}(e)$, for $\Gamma = \mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}$, respectively.
 $\textit{Check-PQ-Conc}$ inspires us that, when $\Gamma\mathsf{\text{-}Condition}(e)$ holds, there should exist some $\alpha$, such that  $\Gamma\mathsf{\text{-}Conc}(e,\alpha)$ holds.
@@
 
-\begin{restatable}{lemma}{EPQasMultiInMRpriforHistory}
-\label{lemma:EPQ as multi in MRpri for history}
-Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$, if and only if, $\forall e' \in \textit{proj}(e)$, $\forall \Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$,  $\Gamma\mathsf{\text{-}Condition}(e)$ $\Rightarrow$ $e$ is linearizable w.r.t $\Gamma$.
-\end{restatable}
+%\begin{restatable}{lemma}{EPQasMultiInMRpriforHistory}
+%\label{lemma:EPQ as multi in MRpri for history}
+%Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$, if and only if, $\forall e' \in \textit{proj}(e)$, $\forall \Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$,  $\Gamma\mathsf{\text{-}Condition}(e)$ $\Rightarrow$ $e$ is linearizable w.r.t $\Gamma$.
+%\end{restatable}
+
+%Note that due to data-independence, register automata are used to detect some renamed violations, instead of store and checking directly.
+Let us introduce the notion of co-regular, which reduce checking linearizable w.r.t $\Gamma\mathsf{\text{-}Conc}(e,\alpha)$ into checking set of witness automata.
+%
+%%\vspace{-6pt}
+%\begin{definition}\label{def:co-regular of rules of extended priority queues}
+%Given $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$.
+$\Gamma$ is co-regular, if there is a finite set $\textit{Auts}_{\Gamma}$ of register automata, such that for each data-independent implementation $\mathcal{I}$, we have that:
+%
+%$\textit{Auts}_{\Gamma} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $\Gamma\mathsf{\text{-}Condition}(e')$ holds but $e$ does not linearizable w.r.t $\Gamma$.
+%
+%$\seqPQ$ is co-regular, if each $\Gamma$ is co-regular.
+%\end{definition}
+
+\subsection{A $\mathsf{MatchedMaxPriority}$-complete automaton}\label{ssec:aut}
+
+The formula $\mathsf{MatchedMaxPriority}$ is false for some execution $e$ when all the values of maximal priorities in $e$ are removed and there exists a value $x$ of maximal priority $p$ s.t.
+$\mathsf{MatchedMaxPriority\text{-}Conc}(e,x)$ is false.
+We consider two cases depending on whether $x$ is the only value with priority $p$.
+We denote by $\mathsf{MatchedMaxPriority}^{>}$ the strengthening of $\mathsf{MatchedMaxPriority}$ with this condition (i.e., all the values other than $x$ have a priority strictly smaller than $p$) and $\mathsf{MatchedMaxPriority}^{=}$ for the strengthening of the same formula with the negation of this condition (i.e., there exists another value $y$ with priority $p$).
 
@@
 
-Note that due to data-independence, register automata are used to detect some renamed violations, instead of store and checking directly.
-Let us introduce the notion of co-regular, which reduce checking linearizable w.r.t $\Gamma\mathsf{\text{-}Conc}(e,\alpha)$ into checking set of witness automata.
+\subsubsection{A $\mathsf{MatchedMaxPriority}^>$-complete automaton}
+\label{subsec:co-regular of EPQ1Lar}
 
-%\vspace{-6pt}
-\begin{definition}\label{def:co-regular of rules of extended priority queues}
-Given $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$.
-$\Gamma$ is co-regular, if there is a finite set $\textit{Auts}_{\Gamma}$ of register automata, such that for each data-independent implementation $\mathcal{I}$, we have that:
-
-$\textit{Auts}_{\Gamma} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $\Gamma\mathsf{\text{-}Condition}(e')$ holds but $e$ does not linearizable w.r.t $\Gamma$.
-
-$\seqPQ$ is co-regular, if each $\Gamma$ is co-regular.
-\end{definition}
-
-We use two point to simply our proof of co-regular.
+%We use two point to simply our proof of co-regular.
 The first point is to use the results in \cite{Bouajjani:2015} to ensure FIFO (first in first out) property of each single-priority projections.
@@
 In Appendix \ref{sec:appendix proof and definition in section definition of co-regular}, according to the result of queue in \cite{Bouajjani:2015}, we construct a set $\textit{Auts}_{\textit{sinPri}}$ of witness automata, and shows that they are enough to ensure that for each data-differentiated executions, each of its single-priority projection without $\textit{rm}(\textit{empty})$ to have ``FIFO'' property, as shown by the following lemma.
-
-\begin{restatable}{lemma}{AutoForEPQwithSignlePri}
-\label{lemma:automata for extended priority queue with single priority}
-
-Given a data-independent implementations $\mathcal{I}$ of extended priority queue, $\mathcal{I} \cap \textit{Auts}_{\textit{sinPri}} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{toQueue}(e')$ does not linearizable to queue.
-\end{restatable}
-
-According to Lemma \ref{lemma:automata for extended priority queue with single priority}, from now on, it is safe to assume that, for each data-differentiated execution, any of its single-priority projection without $\textit{rm}(\textit{empty})$ has ``FIFO'' property.
+%
+%\begin{restatable}{lemma}{AutoForEPQwithSignlePri}
+%\label{lemma:automata for extended priority queue with single priority}
+%
+%Given a data-independent implementations $\mathcal{I}$ of extended priority queue, $\mathcal{I} \cap \textit{Auts}_{\textit{sinPri}} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{toQueue}(e')$ does not linearizable to queue.
+%\end{restatable}
+%
+%According to Lemma \ref{lemma:automata for extended priority queue with single priority}, from now on, it is safe to assume that, for each data-differentiated execution, any of its single-priority projection without $\textit{rm}(\textit{empty})$ has ``FIFO'' property.
 This effectively exclude some obvious violations, such as $\textit{rm}(a)$ happens before $\textit{put}(a,\_)$.
 
-The second point is to ignore the incomparable priorities.
-Given a data-differentiated execution $e$ and one of its maximal priority $p$, $e\vert_{\preceq p}$ is an execution obtained from $e$ by erasing operations of values with priority incomparable with $p$.
-The following lemma states that when checking co-regular of $\seqPQ$, it is safe to ignore all incomparable priorities.
-
-\begin{restatable}{lemma}{priExecutionIsEnough}
-\label{lemma:pri execution is enough}
-Given a data-differentiated execution $e$ and $\Gamma\in \{\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$.
-$\Gamma\mathsf{\text{-}Conc}(e,x)$ holds, if and only if $\Gamma\mathsf{\text{-}Conc}(e\vert_{\preceq p},x)$ holds, where $p$ is the priority of $x$.
-\end{restatable}
-\begin {proof} (Sketch)
-
-To prove the $\textit{only if}$ direction, given $e \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
-Since there is no restriction to $p$-incomparable operations in $u$, $v$ and $w$, we can safely erase all $p$-incomparable values.
-Similar case holds for $\Gamma=\mathsf{UnmatchedMaxPriority}$.
-
-To prove the $\textit{if}$ direction, we use a lemma in Appendix, which said that if the projection of an execution into a set $O_1$ of operations has linearization $l_1$ and the projection of the execution into $O_2$, the remanning operations, has linearization $l_2$, then the execution has a linearization whose projection into $O_1$ is $l_1$ and projection into $O_2$ is $l_2$.
-
-Given $e\vert_{\preceq p} \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
-By above lemma we know that $e$ has a linearization $l' = u' \cdot \textit{put}(x,p) \cdot v' \cdot \textit{rm}(x) \cdot w'$, such that the projection of $l'$ into $p$-incomparable values is l.
-We can see that $u$, $v$ and $w$ are sub-sequence of $u'$, $v'$ an $w'$, respectively.
-Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$ does not have restriction to $p$-incomparable values, we can see that $l' \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
-\qed
-\end {proof}
-
-Lemma \ref{lemma:pri execution is enough} makes our co-regular proof more clear.
+%Lemma \ref{lemma:pri execution is enough} makes our co-regular proof more clear.
 One consequence of it is that for given $e\vert_{\preceq p}$, $\Gamma\mathsf{\text{-}Condition}(e')$ holds only for one $\Gamma$.
-
-\subsection{Co-Regular of $\mathsf{MatchedMaxPriority}^{>}$}
-\label{subsec:co-regular of EPQ1Lar}
+%
+%
+%
+%
+%
+%
+%\subsection{Co-Regular of $\mathsf{MatchedMaxPriority}^{>}$}
+%\label{subsec:co-regular of EPQ1Lar}
 
@@
 
-\subsection{Co-Regular of $\mathsf{MatchedMaxPriority}^{=}$}
+\subsection{A $\mathsf{MatchedMaxPriority}^=$-complete automaton}
 \label{subsec:co-regular of EPQ1Equal}

### 2017-04-20 priority-queues.tex
small stuff
co-regularity

@@
 We write $e \setminus x$ for the projection $e \vert_{ \mathbb{D} \setminus \{ x \} }$.
+Let $\textit{proj}(e)$ be the set of all projections of $e$ to a set of values.
 
-\begin{lemma}
-$\seqPQ$ is closed under projection, i.e., for every execution $e\in \seqPQ$ and $D \subseteq \mathbb{D}$, $e \vert D\in \seqPQ$.
+\begin{lemma}\label{lem:closure_proj}
+$\seqPQ$ is closed under projection, i.e., $\textit{proj}(e)\subseteq \seqPQ$ for each $e\in \seqPQ$.
 \end{lemma}
@@
 This is similar to that in \cite{conf/tacas/AbdullaHHJR13,DBLP:conf/icalp/BouajjaniEEH15}, where they use the notion of data-independence in \cite{conf/popl/Wolper86}.
-Thus, in the remainder of the paper, we focus on characterizing linearizability for data-differentiated executions.
+Section~\ref{sec:checking inclusion by recursive procedure} will focus on characterizing linearizability for data-differentiated executions.
 %The proof of this lemma can be found in Appendix \ref{sec:appendix in section data-independence of EPQ}.

### 2017-04-20 co-regular-EPQ.tex
first part of co-regular

@@
 \end{align*}
-Extending the automaton corresponding to some $\Gamma$ with self-loops that allow any operation with parameter $\top$ results in an automaton satisfying the following property called \emph{$\Gamma$-completeness}.
+Intuitively, $\Gamma(e)$ states that $e$ is linearizable w.r.t.
+the set of sequential executions described by $\Gamma\mathsf{\text{-}Seq}$ (provided that $\mathsf{Has\text{-}}\Gamma(e)$ holds).
+Therefore, by an abuse of terminology, an execution $e$ satisfying $\Gamma(e)$ is called \emph{$\Gamma$-linearizable}.
+Extending the automaton characterizing executions which are not $\Gamma$-linearizable, with self-loops that allow any operation with parameter $\top$ results in an automaton satisfying the following property called \emph{$\Gamma$-completeness}.
 
@@
 
-$A \cap \mathcal{I} \neq \emptyset$ if and only if there exists $ e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$ such that $\Gamma(e')$ doesn't hold.
+$A \cap \mathcal{I} \neq \emptyset$ if and only if there exists $ e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$ such that $e'$ is not $\Gamma$-linearizable.
 \end{definition}
@@
 W.l.o.g., we assume that every implementation $\mathcal{I}$ behaves correctly, i.e., as a FIFO queue, when only values with the same priority are observed.
-More precisely, we assume that for every execution $e\in\mathcal{I}$ and every priority $p\in\mathbb{P}$, the projection of $e$ to values with priority $p$ is linearizable.
+More precisely, we assume that for every execution $e\in\mathcal{I}$ and every priority $p\in\mathbb{P}$, the projection of $e$ to values with priority $p$ is linearizable (w.r.t.
+$\seqPQ$).
 This property can be checked separately using the techniques introduced for regular FIFO queues in~\cite{DBLP:conf/icalp/BouajjaniEEH15} (see Appendix~\ref{} for more details).
@@
 $\prec$).
-This is possible because any data-differentiated execution for which $\Gamma(e)$ is false has such a projection.
+This is possible because any data-differentiated execution which is not $\Gamma$-linearizable has such a projection.
 Formally, given a data-differentiated execution $e$ and $p$ a maximal priority in $e$, $e\vert_{\preceq p}$ is the projection of $e$ to the set of values with priorities comparable to $p$.
@@
 Let $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$ and $e$ be a data-differentiated execution.
-Then, $\Gamma(e)$ is false if and only if $\Gamma(e\vert_{\preceq p})$ is false for some maximal priority $p$ in $e$.
+Then, $e$ is not $\Gamma$-linearizable if and only if $e\vert_{\preceq p}$ is not $\Gamma$-linearizable, for some maximal priority $p$ in $e$.
 \end{restatable}
@@
 
-The formula $\mathsf{MatchedMaxPriority}$ is false for some execution $e$ when all the values of maximal priorities in $e$ are removed and there exists a value $x$ of maximal priority $p$ s.t.
-$\mathsf{MatchedMaxPriority\text{-}Conc}(e,x)$ is false.
-We consider two cases depending on whether $x$ is the only value with priority $p$.
-We denote by $\mathsf{MatchedMaxPriority}^{>}$ the strengthening of $\mathsf{MatchedMaxPriority}$ with this condition (i.e., all the values other than $x$ have a priority strictly smaller than $p$) and $\mathsf{MatchedMaxPriority}^{=}$ for the strengthening of the same formula with the negation of this condition (i.e., there exists another value $y$ with priority $p$).
+A differentiated execution $e$ is not $\mathsf{MatchedMaxPriority}$-linearizable when all the $\textit{put}$ operations in $e$ using the maximal priority $p$ are matched, and $e$ is not linearizable w.r.t.
+the set of sequential executions satisfying $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ for some value $x$ of priority $p$.
+We consider two cases depending on whether $e$ contains exactly one value with priority $p$ or at least two values.
+We denote by $\mathsf{MatchedMaxPriority}^{>}$ the strengthening of $\mathsf{MatchedMaxPriority}$ with the condition that all the values other than $x$ have a priority strictly smaller than $p$ (corresponding to the first case), and by $\mathsf{MatchedMaxPriority}^{=}$ the strengthening of the same formula with the negation of this condition (corresponding to the second case).
+We use particular instances of register automata~\cite{finite-memory, the paper I found} whose states include only two registers, one for storing a priority guessed at the initial state, and one for storing the priority of the current action in the execution.
+The transitions can check equality or the order relation $\prec$ between the values stored in the two registers.
+Instead of formalizing the full class of register automata, we consider a simpler class which suffices our needs.
+More precisely, we consider a class of labeled transition systems whose states consist of a finite control part and a register $r$ interpreted to elements of $\mathbb{P}$.
+The transition labels can be one of the following:
+\begin{itemize}
+	\item $r=*$ for storing an arbitrary value to $r$, 
+	\item $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$ for reading call/return actions of a remove, 
+	\item $\textit{call}(\textit{put},d,g)$ where $g\in\{=r,\prec r,true\}$ is a guard, for reading a call action $\textit{call}(\textit{put},d,p)$ of a put and checking whether $p$ is either equal to or smaller than the value stored in $r$, or arbitrary,
+	\item $\textit{ret}(\textit{put},d,true)$ for reading a return action $\textit{ret}(\textit{put},d,p)$ for some $p$.
+\end{itemize}
+The set of words accepted by such a transition system can be defined as usual.
 
-Register automata is a kind of finite automata with registers used to deal with infinite data domain, where register can be assigned values and do simple operations.
+%$\textit{call}(\textit{put},d,g)$ $\textit{ret}(\textit{put},d,g),$ $\textit{call}(\textit{rm},d),\textit{ret}(\textit{rm},d)$, where $d \in \mathbb{D} \cup \{ \textit{empty} \}$, and $g$ and $\top$ are guards
+%
+% which are complete for each case. Since we don't require the full generality of register automata, 
+%
+%
+%We define $\Gamma$-complete automata for both cases and their union will give a $\mathsf{MatchedMaxPriority}$-complete automaton
+%
+%
+%
+%Register automata is a kind of finite automata with registers used to deal with infinite data domain, where register can be assigned values and do simple operations.
 In this paper we use a restricted kind of register automata, which contains only one register, whose value is obtained once by guessing and never changed.
@@
 Given an execution $e = \alpha_1 \cdots \alpha_k$ of priority queue, register automata $\mathcal{A}$ accepts $e$, if there exists value $d_r\in \mathbb{D}$ and transitions $q_0 \xrightarrow{\beta_1} q_1 \ldots \xrightarrow{\beta_k} q_k$ of $\mathcal{A}$ from a initial state to a final state, such that for each $i$
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $\alpha_i = \textit{call}(\textit{put},a,p)$, then $\beta_i= \textit{call}(\textit{put},a,g)$ and $g$ holds with $p$ and $d_r$. Similarly for $\textit{ret}(\textit{put},a,p)$.
-
-\item[-] Else, $\alpha_i = \beta_i$.
-\end{itemize}
+%
+%\begin{itemize}
+%\setlength{\itemsep}{0.5pt}
+%\item[-] If $\alpha_i = \textit{call}(\textit{put},a,p)$, then $\beta_i= \textit{call}(\textit{put},a,g)$ and $g$ holds with $p$ and $d_r$. Similarly for $\textit{ret}(\textit{put},a,p)$.
+%
+%\item[-] Else, $\alpha_i = \beta_i$.
+%\end{itemize}
 
@@
 
-To facilitate the proof of co-Regular of $\mathsf{MatchedMaxPriority}$, we separate $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ into the following two sub-predicates, where the first one represents that $x$ is the only value with priority $p$ in $e$ , and the second one represents that there are more than one value with priority $p$ in $e$.
+%To facilitate the proof of co-Regular of $\mathsf{MatchedMaxPriority}$, we separate $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ into the following two sub-predicates, where the first one represents that $x$ is the only value with priority $p$ in $e$ , and the second one represents that there are more than one value with priority $p$ in $e$.
 Here $\textit{occurOnce}(p,l)$ holds if there is only one value with priority $p$ in $l$.
-
-{\small
-\begin{align*}
-\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(e,x)=\mathsf{true}$,} \\
-&\hspace{4mm}\mbox{$\textit{occurOnce}(p,e)$ where $p$ is the priority of $x$} \\
-\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(e,x)=\mathsf{true}$,} \\
-&\hspace{4mm}\mbox{$\neg\textit{occurOnce}(p,e)$ where $p$ is the priority of $x$} \\
-\end{align*}
-}
-
-$\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ is obtained from $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Conc}(e,x)$ by using $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(e,x)$ instead of $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$.
+%
+%{\small
+%\begin{align*}
+%\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(e,x)=\mathsf{true}$,} \\
+%&\hspace{4mm}\mbox{$\textit{occurOnce}(p,e)$ where $p$ is the priority of $x$} \\
+%\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(e,x)=\mathsf{true}$,} \\
+%&\hspace{4mm}\mbox{$\neg\textit{occurOnce}(p,e)$ where $p$ is the priority of $x$} \\
+%\end{align*}
+%}
+%
+%$\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ is obtained from $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Conc}(e,x)$ by using $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(e,x)$ instead of $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$.
 And similarly for $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Conc}(e,x)$.
-
-We can also define two conditions, where $\textit{unmatched-priorities}(l)$ is the set of priorities occurring in $\textit{put}$ operations of $e$ for which there is $\textit{rm}$ operation removing the same value:
-
-{\small
-\begin{align*}
-\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Condition}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{matched-priorities}(e)$ for a maximal priority} \\
-&\hspace{4mm}\mbox{$p\in priorities(e)$, $\textit{occurOnce}(p,e)$} \\
-\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{matched-priorities}(e)$ for a maximal priority} \\
-&\hspace{4mm}\mbox{$p\in priorities(e)$, $\neg\textit{occurOnce}(p,e)$} 
-\end{align*}}
-
-In this subsection, we briefly introduce the idea for proving co-regular of $\mathsf{MatchedMaxPriority}^{>}$.
-The proof of this subsection can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
-The notion of left-right constraint used for priority queue is inspired by left-right constraint of queue \cite{Bouajjani:2015}.
-
-Given a data-differentiated $\_$-execution $e$ such that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Condition}(e)$ holds, Lemma \ref{lemma:automata for extended priority queue with single priority} is not enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$.
+%
+%We can also define two conditions, where $\textit{unmatched-priorities}(l)$ is the set of priorities occurring in $\textit{put}$ operations of $e$ for which there is $\textit{rm}$ operation removing the same value:
+%
+%{\small
+%\begin{align*}
+%\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Condition}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{matched-priorities}(e)$ for a maximal priority} \\
+%&\hspace{4mm}\mbox{$p\in priorities(e)$, $\textit{occurOnce}(p,e)$} \\
+%\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{matched-priorities}(e)$ for a maximal priority} \\
+%&\hspace{4mm}\mbox{$p\in priorities(e)$, $\neg\textit{occurOnce}(p,e)$} 
+%\end{align*}} 
+We give a typical example of an execution $e$ which is not $\mathsf{MatchedMaxPriority}^>$-linearizable in 
+%Given a data-differentiated execution $e$ such that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Condition}(e)$ holds, Lemma \ref{lemma:automata for extended priority queue with single priority} is not enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$.
 This is because that it is possible that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ does not holds because of interaction between actions of multiple priorities.
-One such example is shown in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
-We call the time interval from $\textit{ret}(\textit{put},x)$ to $\textit{cal}(rm,x)$, or from $\textit{ret}(\textit{put},x)$ when $\textit{cal}(rm,x)$ does not exist, the interval of value $x$.
+One such example is shown in
+\figurename~\ref{fig:introduce gap for EPQ1Lar}.
+Intuitively, this is a violation because during the whole execution of $\textit{rm}(b)$, the priority queue stores a smaller priority value (which should be removed before $b$).
+To be more precise, we define \emph{the interval of a value $x$} as the time interval from the return of a put $\textit{ret}(\textit{put},x,p)$ to the call of the matching remove $\textit{call}(rm,x)$, or to the end of the execution if such a call action doesn't exist.
+Intuitively, it represents the time interval in which a value is guaranteed to be stored into the concurrent priority queue.
 In \figurename~\ref{fig:introduce gap for EPQ1Lar}, we draw the interval of each value by dashed line.
 Here we assume that $p_1 \prec p_4$, $p_2 \prec p_4$ and $p_3 \prec p_4$.
-We can not find a sequence $s$ where $e \sqsubseteq s$ and $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(s,b)$ holds, since each time point from $\textit{cal}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is in interval of some value with smaller priority, and $\textit{rm}(b)$ can not happen in interval of value with smaller priority.
+We can not find a sequence $s$ where $e \sqsubseteq s$ and $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,b)$ holds, since each time point from $\textit{call}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is included in the interval of some smaller priority value, and $\textit{rm}(b)$ can not happen in interval of a smaller priority value.
 
 \begin{figure}[htbp]
+TODO UPDATE THE ARGUMENTS OF ACTIONS
   \centering
@@
 \end{figure}
+
+TODO INTRODUCE THE LEFT-RIGHT CONSTRAINT
 
@@
 \end{restatable}
+
+TODO GIVE THE AUTOMATON
+
+More details can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
+TODO MOVE THIS TO THE RELATED WORK The notion of left-right constraint used for priority queue is inspired by left-right constraint of queue \cite{Bouajjani:2015}.
 

### 2017-04-20 co-regular-EPQ.tex
first part of co-regular
modify

@@
 
-Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions where any two occurring priorities are comparable (w.r.t.
+{\color {blue}Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions which contains only one maximal priority.
+This is possible because any data-differentiated execution for which $\Gamma(e)$ is false has such a projection.}
+%Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions where any two occurring priorities are comparable (w.r.t.
 $\prec$).
 This is possible because any data-differentiated execution which is not $\Gamma$-linearizable has such a projection.
-Formally, given a data-differentiated execution $e$ and $p$ a maximal priority in $e$, $e\vert_{\preceq p}$ is the projection of $e$ to the set of values with priorities comparable to $p$.
-Then,
+Formally, given a data-differentiated execution $e$ and $p$ a maximal priority in $e$, $e\vert_{\preceq p}$ is the projection of $e$ to the set of values with priorities comparable to $p$. Then,
 
@@
 
-A differentiated execution $e$ is not $\mathsf{MatchedMaxPriority}$-linearizable when all the $\textit{put}$ operations in $e$ using the maximal priority $p$ are matched, and $e$ is not linearizable w.r.t.
+{\color {blue}When $\mathsf{MatchedMaxPriority}$ is false for some execution $e$, then there are two possibilities:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] There exists value $x$, let $e_x$ be obtained from $e$ when all the values of maximal priorities in $e$ but $x$ are removed. Then $\mathsf{MatchedMaxPriority\text{-}Conc}(e_x,x)$ is false.
+
+\item[-] There exists set $X$ of values, let $e'$ be obtained from $e$ when all the values of maximal priorities in $e$ but values in $X$ are removed.
+Then, although $\mathsf{MatchedMaxPriority\text{-}Conc}(e'_x,x)$ is true for each $x \in X$, there is no value $y$ to make $\mathsf{MatchedMaxPriority\text{-}Conc}(e',y)$ to be true.
+\end{itemize}
+
+It is surprised to see that these two possibilities can be captured by two sub-cases of $\mathsf{MatchedMaxPriority}$, depending on whether is only one value $x$ with maximal priority $p$.
+We denote by $\mathsf{MatchedMaxPriority}^{>}$ the strengthening of $\mathsf{MatchedMaxPriority}$ with this condition (i.e., all the values other than $x$ have a priority strictly smaller than $p$) and $\mathsf{MatchedMaxPriority}^{=}$ for the strengthening of the same formula with the negation of this condition (i.e., there exists another value $y$ with priority $p$).
+}
+
+%A differentiated execution $e$ is not $\mathsf{MatchedMaxPriority}$-linearizable when all the $\textit{put}$ operations in $e$ using the maximal priority $p$ are matched, and $e$ is not linearizable w.r.t.
 the set of sequential executions satisfying $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ for some value $x$ of priority $p$.
@@
 
-TODO INTRODUCE THE LEFT-RIGHT CONSTRAINT
+TODO GIVE THE FORMAL DEFINITION OF THE LEFT-RIGHT CONSTRAINT
 

### 2017-04-20 co-regular-EPQ.tex
first part of co-regular

--- 2017-04-20 co-regular-EPQ.tex
@@
 \begin {proof} (Sketch)
-TODO UPDATE THE FOLLOWING
-To prove the $\textit{only if}$ direction, given $e \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
-Since there is no restriction to $p$-incomparable operations in $u$, $v$ and $w$, we can safely erase all $p$-incomparable values.
-Similar case holds for $\Gamma=\mathsf{UnmatchedMaxPriority}$.
+To prove the $\textit{only if}$ direction, let $e$ be an execution linearizable w.r.t.
+$l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
+Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$ imposes no restriction on the operations in $u$, $v$ and $w$ with priorities incomparable to $p$, erasing all these operations results in a sequential execution which still satisfies this property.
+Similarly, for $\Gamma=\mathsf{UnmatchedMaxPriority}$.
 
-To prove the $\textit{if}$ direction, we use a lemma in Appendix, which said that if the projection of an execution into a set $O_1$ of operations has linearization $l_1$ and the projection of the execution into $O_2$, the remanning operations, has linearization $l_2$, then the execution has a linearization whose projection into $O_1$ is $l_1$ and projection into $O_2$ is $l_2$.
+The $\textit{if}$ direction follows from the fact that if the projection of an execution to a set of operations $O_1$ has a linearization $l_1$ and the projection of the same execution to the remaining set of operations has a linearization $l_2$, then the execution has a linearization which is defined as an interleaving of $l_1$ and $l_2$ (see Appendix~\ref{} for more details).
 

### 2017-04-20 co-regular-EPQ.tex

@@
 
-Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions where any two occurring priorities are comparable (w.r.t.
+Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions which contains only one maximal priority.
+This is possible because any data-differentiated execution for which $\Gamma(e)$ is false has such a projection.
+%Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions where any two occurring priorities are comparable (w.r.t.
 $\prec$).
 This is possible because any data-differentiated execution which is not $\Gamma$-linearizable has such a projection.
-Formally, given a data-differentiated execution $e$ and $p$ a maximal priority in $e$, $e\vert_{\preceq p}$ is the projection of $e$ to the set of values with priorities comparable to $p$.
-Then,
+Formally, given a data-differentiated execution $e$ and $p$ a maximal priority in $e$, $e\vert_{\preceq p}$ is the projection of $e$ to the set of values with priorities comparable to $p$. Then,
 
@@
 \begin {proof} (Sketch)
-To prove the $\textit{only if}$ direction, let $e$ be an execution linearizable w.r.t.
+To prove the $\textit{only if}$ direction, let $e$ be a data-differentiated execution linearizable w.r.t.
 $l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
@@
 
-Given $e\vert_{\preceq p} \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
-By above lemma we know that $e$ has a linearization $l' = u' \cdot \textit{put}(x,p) \cdot v' \cdot \textit{rm}(x) \cdot w'$, such that the projection of $l'$ into $p$-incomparable values is l.
-We can see that $u$, $v$ and $w$ are sub-sequence of $u'$, $v'$ an $w'$, respectively.
-Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$ does not have restriction to $p$-incomparable values, we can see that $l' \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
-\qed
+Thus, let $e$ be an execution such that $e\vert_{\preceq p}$ is linearizable w.r.t.
+$l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
+By the property above, we know that $e$ has a linearization $l' = u' \cdot \textit{put}(x,p) \cdot v' \cdot \textit{rm}(x) \cdot w'$, such that the projection of $l'$ to values of priority comparable to $p$ is $l$.
+%We can see that $u$, $v$ and $w$ are subsequences of $u'$, $v'$ an $w'$, respectively. 
+Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$ does not have a condition on values of priority incomparable to $p$, we obtain that $l' \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
 \end {proof}
@@
 \subsection{A $\mathsf{MatchedMaxPriority}$-complete automaton}\label{ssec:aut}
+
+%{\color {blue}When $\mathsf{MatchedMaxPriority}$ is false for some execution $e$, then there are two possibilities:
+%
+%\begin{itemize}
+%\setlength{\itemsep}{0.5pt}
+%\item[-] There exists value $x$, let $e_x$ be obtained from $e$ when all the values of maximal priorities in $e$ but $x$ are removed. Then $\mathsf{MatchedMaxPriority\text{-}Conc}(e_x,x)$ is false.
+%
+%\item[-] There exists set $X$ of values, let $e'$ be obtained from $e$ when all the values of maximal priorities in $e$ but values in $X$ are removed.
+Then, although $\mathsf{MatchedMaxPriority\text{-}Conc}(e'_x,x)$ is true for each $x \in X$, there is no value $y$ to make $\mathsf{MatchedMaxPriority\text{-}Conc}(e',y)$ to be true.
+%\end{itemize}
+%
+%It is surprised to see that these two possibilities can be captured by two sub-cases of $\mathsf{MatchedMaxPriority}$, depending on whether is only one value $x$ with maximal priority $p$.
+We denote by $\mathsf{MatchedMaxPriority}^{>}$ the strengthening of $\mathsf{MatchedMaxPriority}$ with this condition (i.e., all the values other than $x$ have a priority strictly smaller than $p$) and $\mathsf{MatchedMaxPriority}^{=}$ for the strengthening of the same formula with the negation of this condition (i.e., there exists another value $y$ with priority $p$).
+%}
 

### 2017-04-20 co-regular-EPQ.tex
modify

@@
 Let $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$ and $e$ be a data-differentiated execution.
-Then, $e$ is not $\Gamma$-linearizable if and only if $e\vert_{\preceq p}$ is not $\Gamma$-linearizable, for some maximal priority $p$ in $e$.
+Then, $e$ is not $\Gamma$-linearizable if and only if $e\vert_{\preceq p}$ is not $\Gamma$-linearizable, for {\color{blue}any} maximal priority $p$ in $e$.
 \end{restatable}
 \begin {proof} (Sketch) 
+{\color {blue}We prove an equivalent property: $e$ is $\Gamma$-linearizable if and only if $e\vert_{\preceq p}$ is $\Gamma$-linearizable for some maximal priority $p$ in $e$.} %We consider the case of $\mathsf{MatchedMaxPriority}$ and the case of $\mathsf{UnmatchedMaxPriority}$ is similar.
+
 To prove the $\textit{only if}$ direction, let $e$ be a data-differentiated execution linearizable w.r.t.
-$l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
-Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$ imposes no restriction on the operations in $u$, $v$ and $w$ with priorities incomparable to $p$, erasing all these operations results in a sequential execution which still satisfies this property.
+$l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,x)$.
+Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,x)$ imposes no restriction on the operations in $u$, $v$ and $w$ with priorities incomparable to $p$, erasing all these operations results in a sequential execution which still satisfies this property.
 Similarly, for $\Gamma=\mathsf{UnmatchedMaxPriority}$.
@@
 Thus, let $e$ be an execution such that $e\vert_{\preceq p}$ is linearizable w.r.t.
-$l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
+$l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,x)$.
 By the property above, we know that $e$ has a linearization $l' = u' \cdot \textit{put}(x,p) \cdot v' \cdot \textit{rm}(x) \cdot w'$, such that the projection of $l'$ to values of priority comparable to $p$ is $l$.
 %We can see that $u$, $v$ and $w$ are subsequences of $u'$, $v'$ an $w'$, respectively.
-Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$ does not have a condition on values of priority incomparable to $p$, we obtain that $l' \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
+Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,x)$ does not have a condition on values of priority incomparable to $p$, we obtain that $l' \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
 \end {proof}
@@
 A differentiated execution $e$ is not $\mathsf{MatchedMaxPriority}$-linearizable when all the $\textit{put}$ operations in $e$ using the maximal priority $p$ are matched, and $e$ is not linearizable w.r.t.
-the set of sequential executions satisfying $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ for some value $x$ of priority $p$.
+the set of sequential executions satisfying $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ for {\color {blue}any value $x$ of priority $p$}.
 We consider two cases depending on whether $e$ contains exactly one value with priority $p$ or at least two values.
@@
 %\vspace{-10pt}
-  \caption{An execution that does not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$}
+  \caption{An execution that is not $\mathsf{MatchedMaxPriority}^{>}$-linearizable}
   \label{fig:introduce gap for EPQ1Lar}
@@
 TODO GIVE THE FORMAL DEFINITION OF THE LEFT-RIGHT CONSTRAINT
+{\color {blue}Let us introduce left-right constraint to detect the phenomenon in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
 
-In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we introduce the notion of left-right constraint, which is a graph, and the existence of cycle though value with maximal priority in it formally modelled the case in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
-For example, in \figurename~\ref{fig:introduce gap for EPQ1Lar}, $b$ is covered by $f$, $e$ and $d$.
+\begin{definition}\label{def:left-right constraint for matched put and rm operations}
+Given a data-differentiated $e$ without $\textit{rm}(\textit{empty})$, and let $\textit{put}(x,p)$ and $\textit{rm}(x)$ be the only operations with priority $p$.
+The left-right constraint of $\textit{put}(x,p)$ and $\textit{rm}(x)$ is the graph $G$ where:
+
+\begin{itemize}
+\item the nodes are the values of $e$, to which we add a node,
+\item there is an edge from $d_1$ to $x$, if $\textit{put}(d_1,\_) <_{\textit{hb}} \textit{put}(x,p)$ or $\textit{rm}(x)$,
+\item there is an edge from $x$ to $d_1$, if $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists,
+\item there is an edge from $d_1$ to $d_2$, if $\textit{put}(d_1,\_) <_{\textit{hb}} \textit{rm}(d_2,\_)$.
+\end{itemize} 
+\end{definition}
+
+The existence of cycle though value with maximal priority in left-right constraint formally modelled the phenomenon in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
+For example, in \figurename~\ref{fig:introduce gap for EPQ1Lar}, $f \rightarrow e \rightarrow d \rightarrow b \rightarrow f$ is a cycle in left-right constraint.
 We then prove that getting rid of cycle though value with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$.
+}
 
@@
 Assume that $x$ is covered by $d_1,\ldots,d_m$, we can safely rename $x$ into $b$, rename $d_1,\ldots,d_m$ into $a$ and rename all other value into $c$ by data-independence.
-Such execution can be recognized by witness automata, since between the first $\textit{ret}(\textit{put},a,\_)$ and the last $\textit{call}(\textit{rm},a)$ (if exists), $\textit{ret}(\textit{put},a,\_)$ and $\textit{call}(\textit{rm},a)$ occurs in pair and there is no need for count.
-In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we construct a set $\textit{Auts}_{\textit{1-lar}}$ of register automata, and use $\textit{Auts}_{\textit{1-lar}}$ to prove that $\mathsf{MatchedMaxPriority}^{>}$ is co-regular, as stated by the following lemma.
+Such execution can be recognized by register automata, since between the first $\textit{ret}(\textit{put},a,\_)$ and the last $\textit{call}(\textit{rm},a)$ (if exists), $\textit{ret}(\textit{put},a,\_)$ and $\textit{call}(\textit{rm},a)$ occurs in pair and there is no need for count.
+In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we construct a set $\textit{Auts}_{\textit{1-lar}}$ of register automata, {\color {blue}and prove that $\textit{Auts}_{\textit{1-lar}}$ is $\mathsf{MatchedMaxPriority}^{>}$-complete, as stated by the following lemma.
 
@@
 \label{lemma:EPQ1Lar is co-regular}
-$\mathsf{MatchedMaxPriority}^{>}$ is co-regular.
+$\textit{Auts}_{\textit{1-lar}}$ is $\mathsf{MatchedMaxPriority}^{>}$-complete.
 \end{restatable}
+}
 
+{\color {blue}
+For example, to capture non-$\mathsf{MatchedMaxPriority}^{>}$-linearizable execution $e$ where $e \vert_{b} = \textit{call}(\textit{put},b,\_) \cdot \textit{ret}(\textit{put},b,\_) \cdot \textit{call}(\textit{rm},b) \cdot \textit{ret}(\textit{rm},b)$, we use the register automaton in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
+Here $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{call}(\textit{put},a,=r)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{call}(\textit{put},d,\top),\textit{ret}(\textit{put},d,\top), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.
+There are three paths in $\mathcal{A}_{\textit{l-lar}}^1$, since there are three possible position of  the first $\textit{ret}(\textit{put},a)$ in $e$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=1 \textwidth]{figures/PIC_AUTO_PQ1Lar-pprr.pdf}
+%\vspace{-10pt}
+  \caption{Automaton $\mathcal{A}_{\textit{l-lar}}^1$}
+  \label{fig:automata APQ1Lar-1 in paper}
+\end{figure}
+}
 TODO GIVE THE AUTOMATON
 
-More details can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
-TODO MOVE THIS TO THE RELATED WORK The notion of left-right constraint used for priority queue is inspired by left-right constraint of queue \cite{Bouajjani:2015}.
+More details can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}. %TODO MOVE THIS TO THE RELATED WORK
 

### 2017-04-20 introduction.tex
modify

@@
 Due to the partial-order priorities, to prove the correctness of recursive procedure, we need to alter the positions of linearization points of incomparable items while keeps the linearization points of comparable items unchanged.
-For checking violations of one local property, we just need to find a value, such that at every possible time point to locate its remove operation, there is already values with smaller priorities in priority queue.
+{\color {blue}In the proof process we use the notion of left-right constraint, which is inspired by left-right constraint of queue \cite{Bouajjani:2015}.} For checking violations of one local property, we just need to find a value, such that at every possible time point to locate its remove operation, there is already values with smaller priorities in priority queue.
 For checking violations of another local property, we just need to find two values with maximal priority, such that every possible time point to locate remove operation of one value is disabled by remove operation of another value.

### 2017-04-21 co-regular-EPQ.tex
first co-regular

@@
 \label{lemma:pri execution is enough}
-Let $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$ and $e$ be a data-differentiated execution.
-Then, $e$ is not $\Gamma$-linearizable if and only if $e\vert_{\preceq p}$ is not $\Gamma$-linearizable, for {\color{blue}any} maximal priority $p$ in $e$.
+Let $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$ and $e$ a data-differentiated execution.
+Then, $e$ is $\Gamma$-linearizable iff $e\vert_{\preceq p}$ is $\Gamma$-linearizable for some maximal priority $p$ in $e$.
 \end{restatable}
 \begin {proof} (Sketch) 
-{\color {blue}We prove an equivalent property: $e$ is $\Gamma$-linearizable if and only if $e\vert_{\preceq p}$ is $\Gamma$-linearizable for some maximal priority $p$ in $e$.} %We consider the case of $\mathsf{MatchedMaxPriority}$ and the case of $\mathsf{UnmatchedMaxPriority}$ is similar.
-
+%{\color {blue}We prove an equivalent property: $e$ is $\Gamma$-linearizable if and only if $e\vert_{\preceq p}$ is $\Gamma$-linearizable for some maximal priority $p$ in $e$.} %We consider the case of $\mathsf{MatchedMaxPriority}$ and the case of $\mathsf{UnmatchedMaxPriority}$ is similar.
 To prove the $\textit{only if}$ direction, let $e$ be a data-differentiated execution linearizable w.r.t.
@@
 A differentiated execution $e$ is not $\mathsf{MatchedMaxPriority}$-linearizable when all the $\textit{put}$ operations in $e$ using the maximal priority $p$ are matched, and $e$ is not linearizable w.r.t.
-the set of sequential executions satisfying $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ for {\color {blue}any value $x$ of priority $p$}.
+the set of sequential executions satisfying $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ for each value $x$ of priority $p$.
 We consider two cases depending on whether $e$ contains exactly one value with priority $p$ or at least two values.
@@
 In \figurename~\ref{fig:introduce gap for EPQ1Lar}, we draw the interval of each value by dashed line.
-Here we assume that $p_1 \prec p_4$, $p_2 \prec p_4$ and $p_3 \prec p_4$.
-We can not find a sequence $s$ where $e \sqsubseteq s$ and $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,b)$ holds, since each time point from $\textit{call}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is included in the interval of some smaller priority value, and $\textit{rm}(b)$ can not happen in interval of a smaller priority value.
+Here we assume that $p_1 \prec p_4$, $p_2 \prec p_4$, and $p_3 \prec p_4$.
+We can not find a sequence $s$ where $e \sqsubseteq s$ and $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,b)$ holds, since each time point from $\textit{call}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is included in the interval of some smaller priority value, and $\textit{rm}(b)$ can't take effect in the interval of a smaller priority value.
 
@@
 
-TODO GIVE THE FORMAL DEFINITION OF THE LEFT-RIGHT CONSTRAINT
-{\color {blue}Let us introduce left-right constraint to detect the phenomenon in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
+To formalize the scenario in \figurename~\ref{fig:introduce gap for EPQ1Lar} we use the notion of \emph{left-right constraint} defined below.
 
 \begin{definition}\label{def:left-right constraint for matched put and rm operations}
-Given a data-differentiated $e$ without $\textit{rm}(\textit{empty})$, and let $\textit{put}(x,p)$ and $\textit{rm}(x)$ be the only operations with priority $p$.
-The left-right constraint of $\textit{put}(x,p)$ and $\textit{rm}(x)$ is the graph $G$ where:
-
+Let $e$ be a data-differentiated execution which contains only one maximal priority $p$, and only one value $x$ of priority $p$ (and no $\textit{rm}(\textit{empty})$ operations). 
+% and let $\textit{put}(x,p)$ and $\textit{rm}(x)$ be the only operations with priority $p$. 
+The \emph{left-right constraint of $x$} is the graph $G$ where: %$\textit{put}(x,p)$ and $\textit{rm}(x)$ is the graph $G$ where:
 \begin{itemize}
-\item the nodes are the values of $e$, to which we add a node,
-\item there is an edge from $d_1$ to $x$, if $\textit{put}(d_1,\_) <_{\textit{hb}} \textit{put}(x,p)$ or $\textit{rm}(x)$,
-\item there is an edge from $x$ to $d_1$, if $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists,
+\item the nodes are the values occurring in $e$, %, to which we add a node,
+\item there is an edge from $d_1$ to $x$, if $\textit{put}(d_1,\_) <_{\textit{hb}} \textit{put}(x,p)$ or $\textit{put}(d_1,\_) <_{\textit{hb}} \textit{rm}(x)$,
+\item there is an edge from $x$ to $d_1$, if $\textit{rm}(x)<_{\textit{hb}}\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists,
 \item there is an edge from $d_1$ to $d_2$, if $\textit{put}(d_1,\_) <_{\textit{hb}} \textit{rm}(d_2,\_)$.
@@
 
-The existence of cycle though value with maximal priority in left-right constraint formally modelled the phenomenon in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
+The execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} is not $\mathsf{MatchedMaxPriority}^>$-linearizable because the left-right constraint of the maximal priority value $b$ contains a cycle: $f \rightarrow e \rightarrow d \rightarrow b \rightarrow f$.
+We prove that the presence of such a cycle is equivalent to the execution not being $\mathsf{MatchedMaxPriority}^>$-linearizable (see Appendix~\ref{}).
+%The existence of cycle though value with maximal priority in left-right constraint formally modelled the phenomenon in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
 For example, in \figurename~\ref{fig:introduce gap for EPQ1Lar}, $f \rightarrow e \rightarrow d \rightarrow b \rightarrow f$ is a cycle in left-right constraint.
 We then prove that getting rid of cycle though value with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$.
-}
+When the left-right constraint of the maximal priority value $x$ contains a cycle of the form $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ for some $d_1$,$\ldots$,$d_n\in \mathbb{D}$, we say that $x$ is \emph{covered} by $d_1,\ldots,d_m$.
+The shape of such a cycle can be detected using our class of automata, the only complication being the unbounded number of values $d_1$,$\ldots$,$d_n$.
+However, by data independence, whenever an implementation contains such an execution it also contains an execution where all the values $d_1$,$\ldots$,$d_n$ are renamed to the same value $a$, and $x$ is renamed to $b$.
+Therefore, our automata can be defined over a fixed set of values $a$, $b$, and $\top$ (for the operations outside of the non-linearizable projection).
 
-When there is a cycle $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ through value $x$ with maximal priority in left-right constraint, we say that $x$ is covered by $d_1,\ldots,d_m$.
+TODO WHY IT IS A SET OF AUTOMATA ?
+SAY WHAT ARE THE CASES THAT NEED TO BE COVERED.
+SAY WHAT IS THE CASE WE GIVE IN THE FIGURE.
+USE $\top$ AS MENTIONED AT THE BEGINNING OF THE SECTION INSTEAD OF 	$c$.
+USE BIG CAPS FOR SETS OF TRANSITION LABELS.
+
 Assume that $x$ is covered by $d_1,\ldots,d_m$, we can safely rename $x$ into $b$, rename $d_1,\ldots,d_m$ into $a$ and rename all other value into $c$ by data-independence.

### 2017-04-21 introduction.tex
introduction

--- 2017-04-19 introduction.tex
@@
 
-Efficient implementations of concurrent objects are essential and hard to get right.
+Modern computer software is increasingly concurrent. Interactive applications
+and services necessitate reactive asynchronous operations to handle requests
+immediately as they happen, rather than waiting for long-running operations to
+complete. Furthermore, as processor manufacturers approach clock-speed limits,
+performance improvements are more-often achieved by parallelizing operations
+across multiple processor cores.
+
+%Efficient implementations of concurrent objects are essential and hard to get right.
 Verifying them is difficult, checking a single execution is NP-complete~\cite{journals/siamcomp/GibbonsK97} and checking a finite-state implementation is undecidable~\cite{conf/esop/BouajjaniEEH13}.
 
-The set of objects considered in previous work consists of stacks, queues, registers.
+%The set of objects considered in previous work consists of stacks, queues, registers.
 
-Here, we consider another important object, the priority queues, which are essential for applications such as task scheduling and discrete event simulation, and for which numerous implementations have been proposed, e.g.,~\cite{DBLP:conf/ppopp/AlistarhKLS15,DBLP:conf/wdag/CalciuMH14,DBLP:conf/opodis/LindenJ13,DBLP:conf/podc/ShavitZ99,DBLP:conf/ipps/ShavitL00}.
-This object is much less studied compared to the other ones in the verification literature.
+ Multithreaded software is typically built with specialized “concurrent
+  objects” like atomic integers, queues, maps, priority queues. These objects’ methods are
+  designed to confom to better established sequential specifications, a property known as \emph{linearizability}~\cite{journals/toplas/HerlihyW90},
+  despite being optimized to avoid blocking and exploit parallelism, e.g.,~by
+  using atomic machine instructions like compare-and-swap.
+Intuitively, linearizability asks that every individual operation appears to take place instantaneously at some point between its invocation and its return.
+Verifying linearizability is intrinsically hard, and undecidable in general~\cite{conf/esop/BouajjaniEEH13}.
+% (even for finite-state implementations).
+However, recent work~\cite{DBLP:conf/icalp/BouajjaniEEH15} has shown that for particular classes of objects, 
+i.e., registers, mutexes, queues, and stacks, the problem of verifying linearizability becomes decidable (for finite-state implementations).
 
-Priority queues are collections and in that sense, they could be found similar to stacks and queues.
-The main difference which complicates the process of inventing verification techniques is that the order in which elements are removed is not fixed by the happens-before like in stacks and queues but by parameters, the priorities, which come from an unbounded domain.
-TODO an example.
-Say that for generality, we consider partially-ordered priorities and that for single priority we take FIFO.
-In java, same priority values are removed in an arbitrary order, however, this can be modeled using our partial orders.
-TODO SAY HOW
+In this paper, we consider another important object, namely the priority queue, which is essential for applications such as task scheduling and discrete event simulation.
+Numerous implementations have been proposed in the research literature, e.g.,~\cite{DBLP:conf/ppopp/AlistarhKLS15,DBLP:conf/wdag/CalciuMH14,DBLP:conf/opodis/LindenJ13,DBLP:conf/podc/ShavitZ99,DBLP:conf/ipps/ShavitL00}, and concrete implementations exist in many modern languages like C++ or Java.
+%
+%This object is much less studied compared to the other ones in the verification literature.
+%
+Priority queues are collections providing $\textit{put}$ and $\textit{rm}$ methods for adding and removing values.
+Every added value is associated to a priority and a remove operation returns a minimal priority value.
+For generality, we consider a partially-ordered set of priorities.
+Values with incomparable priorities can be removed in any order, and values having the same priority are removed in the FIFO order.
+Implementations like the PriorityBlockingQueue in Java where same priority values are removed in an arbitrary order can be modeled in our framework by renaming equal priorities to incomparable priorities (while preserving the order constraints).
 
-Building on previous work~\cite{DBLP:conf/icalp/BouajjaniEEH15}, we define decision procedures for priority queues. These procedures are designed in several steps:
-1) defining recursive procedures recognizing valid sequences, and extending them to concurrent executions.
+Compared to previously studied collections like stacks and queues, the main challenge in dealing with priority queues is that the order in which values are removed is not fixed by the happens-before between add/remove operations, but by parameters of the $\textit{put}$ operations (the priorities) which come from an unbounded domain.
+For instance, the sequential behavior $\textit{put}(a,p_1)\cdot \textit{put}(b,p_3)\cdot \textit{put}(c,p_2)\cdot \textit{rm}(a,p_1)\cdot \textit{rm}(c,p_2)$ where the priority $p_1$ is less than $p_2$ which is less than $p_3$ is not admitted neither by the regular queue nor the stack.
+
+%While it was proved that the violations of concurrent stacks and queues can be captured using finite-state automata~\cite{DBLP:conf/icalp/BouajjaniEEH15}, we show that characterizing the violations of the priority queue needs \emph{register automata} where registers are used to store and compare priorities.
+
+%This gives a reduction to reachability that works for arbitrary implementations, and a decision procedure for finite-state implementations
+
+Following the approach in~\cite{DBLP:conf/icalp/BouajjaniEEH15}, we give a characterization of concurrent priority queue behaviors violating linearizability in terms of automata.
+This characterization enables a reduction of checking linearizability for arbitrary implementations to reachability or invariant checking, and implies decidability for checking linearizability of finite-state implementations.
+However, differently from the case of stacks and queues where finite-state automata are sufficient, we show that the case of priority queues needs \emph{register automata} where registers are used to store and compare priorities.
+
+This characterization is obtained in several steps.
+We first define a recursive procedure which recognizes valid sequential executions which is then extended to recognize linearizable concurrent executions.
+Intuitively, for an input execution $e$, this procedure handles values occurring in $e$ one by one, starting with values of maximal priority (which are to be removed the latest).
+For each value $x$, it checks whether $e$ satisfies some property ``local'' to that value, which is agnostic to how the operations adding or removing other values are ordered between them (w.r.t.
+the happens-before), other than how they are ordered w.r.t.
+the operations on $x$.
+When this property holds, the same is done for the rest of the execution, without the operations on $x$.
+This procedure works only for executions where a value is added at most once, but this is not a limitation for \emph{data-independent} implementations whose behavior doesn't depend on the values that are added or removed.
+In fact, all the implementations that we are aware of are data-independent.
+
+Checking whether an execution violates this ``local'' property for a value $x$ can be done using a class of register automata~\cite{DBLP:journals/tcs/KaminskiF94,DBLP:conf/icalp/Cerans94,DBLP:conf/stacs/SegoufinT11} (transition systems where the states consist of a fixed set of registers that can receive values and be compared).
+Actually, only two registers are needed: one register $r_1$ for storing a priority guessed at the initial state, and one register $r_2$ for reading priorities as they occur in the execution and comparing them with the one stored in $r_1$.
+We show that registers storing values added to or removed from the priority queue are not needed, since any data-independent implementation admits a violation to linearizability whenever it admits a violation where the number of values is constant, and at most 4 (the number of priorities can still be unbounded).
+
+%Building on previous work~\cite{DBLP:conf/icalp/BouajjaniEEH15}, we define decision procedures for priority queues.
+%These procedures are designed in several steps:
+%1) defining recursive procedures recognizing valid sequences, and extending them to concurrent executions.
 These procedures take values one by one and check some property local to that value, ignoring how operations on other values are ordered between them.
 This works only for data-differentiated, but this is complete by data-independence.
-2) checking whether a fixed value violates that property can be done using a specific class of register automata, TODO say how simple they are
-3) we distinguish between that value being removed or not, and we have to also consider the case of remove(empty). This results in 3-4 automata describing all the possible violations.
-
-Although we follow the same schema as in~\cite{DBLP:conf/icalp/BouajjaniEEH15}, these results require establishing key results which are specific to priority queues and which are not implied by those for stacks and queues (say that more details are given in the related work)
-
-This gives a reduction to reachability that works for arbitrary implementations, and a decision procedure for finite-state implementations
-
-\smallskip
-
-\noindent {\bf Related work}
-%\label{sec:related work}
-\noindent Related work
+%2) checking whether a fixed value violates that property can be done using a specific class of register automata, TODO say how simple they are
+%3) we distinguish between that value being removed or not, and we have to also consider the case of remove(empty). This results in 3-4 automata describing all the possible violations.

### 2017-04-21 related.tex
introduction

\section{Related work}

Although we follow the same schema as in~\cite{DBLP:conf/icalp/BouajjaniEEH15}, these results require establishing key results which are specific to priority queues and which are not implied by those for stacks and queues (say that more details are given in the related work)

\noindent The most related work of our paper is ~\cite{DBLP:conf/icalp/BouajjaniEEH15}. Compared to ~\cite{DBLP:conf/icalp/BouajjaniEEH15}, our work are specific to priority queue and thus have points not occur in queues and stacks. Due to the partial-order priorities, to prove the correctness of recursive procedure, we need to alter the positions of linearization points of incomparable items while keeps the linearization points of comparable items unchanged. {\color {blue}In the proof process we use the notion of left-right constraint, which is inspired by left-right constraint of queue \cite{Bouajjani:2015}.} For checking violations of one local property, we just need to find a value, such that at every possible time point to locate its remove operation, there is already values with smaller priorities in priority queue. For checking violations of another local property, we just need to find two values with maximal priority, such that every possible time point to locate remove operation of one value is disabled by remove operation of another value. Such phenomenons are unique for priority queue.

### 2017-04-21 co-regular-EPQ.tex
introduction
modify

@@
 
-Assume that $x$ is covered by $d_1,\ldots,d_m$, we can safely rename $x$ into $b$, rename $d_1,\ldots,d_m$ into $a$ and rename all other value into $c$ by data-independence.
-Such execution can be recognized by register automata, since between the first $\textit{ret}(\textit{put},a,\_)$ and the last $\textit{call}(\textit{rm},a)$ (if exists), $\textit{ret}(\textit{put},a,\_)$ and $\textit{call}(\textit{rm},a)$ occurs in pair and there is no need for count.
-In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we construct a set $\textit{Auts}_{\textit{1-lar}}$ of register automata, {\color {blue}and prove that $\textit{Auts}_{\textit{1-lar}}$ is $\mathsf{MatchedMaxPriority}^{>}$-complete, as stated by the following lemma.
-
-\begin{restatable}{lemma}{EPQOneLarisCoRegular}
-\label{lemma:EPQ1Lar is co-regular}
-$\textit{Auts}_{\textit{1-lar}}$ is $\mathsf{MatchedMaxPriority}^{>}$-complete.
-\end{restatable}
+{\color {blue} Assume that $x$ is covered by $d_1,\ldots,d_m$, we can safely rename $x$ into $b$, rename $d_1,\ldots,d_m$ into $a$ and rename all other value into $c$ by data-independence.
+Since there are four cases of projection of execution to $x$ when their is a cycle through $x$.
+To ease the reading, for each of such case we generate a register automata, and then let the $\mathsf{MatchedMaxPriority}^{>}$-complete automaton be the union of the four register automata.
+For example, for the case when the projection of $e$ to $x$ is as in \figurename~\ref{fig:introduce gap for EPQ1Lar} (rename $b$ into $x$), we construct a register automaton $\mathcal{A}_{\textit{l-lar}}^1$ in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
+Here transition labels $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{call}(\textit{put},a,=r)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{call}(\textit{put},d,\top),\textit{ret}(\textit{put},d,\top), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.
+There are three paths in \figurename~\ref{fig:introduce gap for EPQ1Lar} from $q_{\textit{init}}$ to $q_7$, and they comes from that there are three possible positions for the first $\textit{call}(\textit{put},a,\_)$ in the captured execution, which is shown in \figurename~\ref{fig:executions APQ1Lar-1 in paper}.
+In \figurename~\ref{fig:automata APQ1Lar-1 in paper}, the paths $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_{10}$ and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$ corresponds to the first, second and third cases in \figurename~\ref{fig:executions APQ1Lar-1 in paper}, respectively.
 }
-
-{\color {blue}
-For example, to capture non-$\mathsf{MatchedMaxPriority}^{>}$-linearizable execution $e$ where $e \vert_{b} = \textit{call}(\textit{put},b,\_) \cdot \textit{ret}(\textit{put},b,\_) \cdot \textit{call}(\textit{rm},b) \cdot \textit{ret}(\textit{rm},b)$, we use the register automaton in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
-Here $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{call}(\textit{put},a,=r)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{call}(\textit{put},d,\top),\textit{ret}(\textit{put},d,\top), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.
-There are three paths in $\mathcal{A}_{\textit{l-lar}}^1$, since there are three possible position of  the first $\textit{ret}(\textit{put},a)$ in $e$.
 
@@
 \end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=1 \textwidth]{figures/PIC_HIS_PQ1Lar-pprr-in-paper.pdf}
+%\vspace{-10pt}
+  \caption{Three possible positions of the first $\textit{call}(\textit{put},a,\_)$ in captured execution}
+  \label{fig:executions APQ1Lar-1 in paper}
+\end{figure}
+
+{\color {blue}
+Let $\mathcal{A}_{\textit{1-lar}}$ be the union of the four register automata mentioned above, then we prove that $\mathcal{A}_{\textit{1-lar}}$ is $\mathsf{MatchedMaxPriority}^{>}$-complete.
+The detailed construction and proof can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
 }
-TODO GIVE THE AUTOMATON
+
+\begin{restatable}{lemma}{EPQOneLarisCoRegular}
+\label{lemma:EPQ1Lar is co-regular}
+$\textit{Auts}_{\textit{1-lar}}$ is $\mathsf{MatchedMaxPriority}^{>}$-complete.
+\end{restatable}
 
@@
 
-\subsection{A $\mathsf{MatchedMaxPriority}^=$-complete automaton}
+\subsubsection{A $\mathsf{MatchedMaxPriority}^=$-complete automaton}
 \label{subsec:co-regular of EPQ1Equal}

### 2017-04-21 introduction.tex
introduction

@@
 Priority queues are collections providing $\textit{put}$ and $\textit{rm}$ methods for adding and removing values.
-Every added value is associated to a priority and a remove operation returns a minimal priority value.
+Every value added is associated to a priority and a remove operation returns a minimal priority value.
 For generality, we consider a partially-ordered set of priorities.
@@
 
-Compared to previously studied collections like stacks and queues, the main challenge in dealing with priority queues is that the order in which values are removed is not fixed by the happens-before between add/remove operations, but by parameters of the $\textit{put}$ operations (the priorities) which come from an unbounded domain.
-For instance, the sequential behavior $\textit{put}(a,p_1)\cdot \textit{put}(b,p_3)\cdot \textit{put}(c,p_2)\cdot \textit{rm}(a,p_1)\cdot \textit{rm}(c,p_2)$ where the priority $p_1$ is less than $p_2$ which is less than $p_3$ is not admitted neither by the regular queue nor the stack.
+Compared to previously studied collections like stacks and queues, the main challenge in dealing with priority queues is that the order in which values are removed is not fixed by the happens-before between add/remove operations (e.g., in the case of a queue, the values are removed in the order in which they were inserted), but by parameters of the $\textit{put}$ operations (the priorities) which come from an unbounded domain.
+For instance, the sequential behavior $\textit{put}(a,p_1)\cdot \textit{put}(b,p_3)\cdot \textit{put}(c,p_2)\cdot \textit{rm}(a,p_1)\cdot \textit{rm}(c,p_2)$ where the priority $p_1$ is less than $p_2$ which is less than $p_3$, is not admitted neither by the regular queue nor the stack.
 
@@
 Intuitively, for an input execution $e$, this procedure handles values occurring in $e$ one by one, starting with values of maximal priority (which are to be removed the latest).
-For each value $x$, it checks whether $e$ satisfies some property ``local'' to that value, which is agnostic to how the operations adding or removing other values are ordered between them (w.r.t.
+For each value $x$, it checks whether $e$ satisfies some property ``local'' to that value, i.e., which is agnostic to how the operations adding or removing other values are ordered between them (w.r.t.
 the happens-before), other than how they are ordered w.r.t.
@@
 
-Checking whether an execution violates this ``local'' property for a value $x$ can be done using a class of register automata~\cite{DBLP:journals/tcs/KaminskiF94,DBLP:conf/icalp/Cerans94,DBLP:conf/stacs/SegoufinT11} (transition systems where the states consist of a fixed set of registers that can receive values and be compared).
+Next, we show that checking whether an execution violates this ``local'' property for a value $x$ can be done using a class of register automata~\cite{DBLP:journals/tcs/KaminskiF94,DBLP:conf/icalp/Cerans94,DBLP:conf/stacs/SegoufinT11} (transition systems where the states consist of a fixed set of registers that can receive values and be compared).
 Actually, only two registers are needed: one register $r_1$ for storing a priority guessed at the initial state, and one register $r_2$ for reading priorities as they occur in the execution and comparing them with the one stored in $r_1$.
 We show that registers storing values added to or removed from the priority queue are not needed, since any data-independent implementation admits a violation to linearizability whenever it admits a violation where the number of values is constant, and at most 4 (the number of priorities can still be unbounded).
+
+The remainder of this article is organized as follows.
+Section~\ref{sec:priority queue and data-independence} describes the priority queue ADT, lists several semantic properties like data-independence that are satisfied by implementations of this ADT, and formalizes the notion of linearizability.
+Sections~\ref{sec:checking inclusion by recursive procedure} defines a recursive procedure for checking linearizability of concurrent priority queue behaviors.
+Section~\ref{sec:co-regular of extended priority queues} gives an automata characterization of the violations to linearizability, and
+Section~\ref{sec:related} discusses related work.
+%, and
+%Section~\ref{sec:conclusion} gives concluding remarks.
 

### 2017-04-21 priority-queues.tex
modify

@@
 We will make explicit this argument only when needed in our technical development.
-The projection $e \vert D$ of an execution $e$ to a set of values $D \subseteq \mathbb{D}$ is obtained from $e$ by erasing all call/return actions with an argument in $D$.
+The projection $e \vert D$ of an execution $e$ to a set of values $D \subseteq \mathbb{D}$ is obtained from $e$ by erasing all call/return actions with an argument not in $D$.
 We write $e \setminus x$ for the projection $e \vert_{ \mathbb{D} \setminus \{ x \} }$.

### 2017-04-21 abstract.tex
several things

@@
+%!TEX root = draft.tex
 \begin{abstract}
-I am abstract
+Efficient implementations of concurrent objects such as atomic collections are essential to modern computing.
+Programming such objects is error prone: in minimizing the synchronization overhead between concurrent object invocations, one risks the conformance to sequential specifications -- or in formal terms, one risks violating linearizability.
+Unfortunately, verifying linearizability is undecidable in general, even on classes of implementations where the usual control-state reachability is decidable.
+In this work we consider concurrent priority queues which are fundamental to many multi-threaded applications such as task scheduling or discrete event simulation, and show that verifying linearizability of such implementations can be reduced to control-state reachability.
+This reduction entails the first decidability results for verifying concurrent priority queues in the context of an unbounded number of threads, and it enables the application of existing safety-verification tools for establishing their correctness.
 \end{abstract}

### 2017-04-21 related.tex
several things

@@
-\section{Related work}
+%!TEX root = draft.tex
+\section{Related work}\label{sec:related}
 
-Although we follow the same schema as in~\cite{DBLP:conf/icalp/BouajjaniEEH15}, these results require establishing key results which are specific to priority queues and which are not implied by those for stacks and queues (say that more details are given in the related work)
+The theoretical limits of checking linearizability have been investigated in previous works. 
+Checking linearizability of a single execution w.r.t. an arbitrary ADT is NP-complete~\cite{journals/siamcomp/GibbonsK97} while checking linearizability of all the executions  
+of a finite-state implementation w.r.t. an arbitrary ADT 
+specification (given as a regular language) is EXPSPACE-complete when the number of program 
+threads is bounded~\cite{journals/iandc/AlurMP00,netys-lin}, and
+undecidable otherwise~\cite{conf/esop/BouajjaniEEH13}.
 
-\noindent The most related work of our paper is ~\cite{DBLP:conf/icalp/BouajjaniEEH15}.
+Existing automated methods for proving linearizability of a concurrent object
+implementation are also based on reductions to safety
+verification, e.g.,~\cite{conf/tacas/AbdullaHHJR13, conf/concur/HenzingerSV13,
+conf/cav/Vafeiadis10}. The approach in~\cite{conf/cav/Vafeiadis10} considers
+implementations where 
+operations' \emph{linearization points}
+are 
+% fixed to particular source-code locations.
+manually specified.
+Essentially, this approach instruments the
+implementation with ghost variables simulating the ADT specification at
+linearization points. This approach is incomplete since not all implementations
+have fixed linearization points. Aspect-oriented
+proofs~\cite{conf/concur/HenzingerSV13} reduce linearizability to the
+verification of four simpler safety properties. However, this approach has only
+been applied to queues, and has not produced a fully automated
+and complete proof technique. The work in~\cite{Dodds:2015:SCT:2676726.2676963} proves 
+linearizability of stack implementations with an automated proof assistant. 
+Their approach does not lead to full automation however, e.g.,~by reduction to 
+safety verification.
+
+Our previous work~\cite{DBLP:conf/icalp/BouajjaniEEH15}
+shows that checking linearizability of finite-state implementations of concurrent queues and stacks is decidable.
+Roughly, we follow the same schema: the recursive procedure in Section~\ref{ssec:seq_exec} is similar to the inductive rules in~\cite{DBLP:conf/icalp/BouajjaniEEH15}, and its extension to concurrent executions in Section~\ref{ssec:conc_exec} corresponds to the notion of step-by-step linearizability in~\cite{DBLP:conf/icalp/BouajjaniEEH15}.
+Although similar in nature, defining these procedures and establishing their correctness require proof techniques which are specific to the priority queue semantics.
+The order in which values are removed from a priority queue is encoded in their priorities which come from an unbounded domain, and not in the happens-before order as in the case of stacks and queues.
+Therefore, the results we introduce in this paper cannot be inferred from those in~\cite{DBLP:conf/icalp/BouajjaniEEH15}.
+At a technical level, characterizing the priority queue violations requires a more expressive class of automata (with registers) than the finite-state automata in~\cite{DBLP:conf/icalp/BouajjaniEEH15}.
+
+%
+%these results require establishing key results which are specific to priority queues and which are not implied by those for stacks and queues (say that more details are given in the related work)
+%
+%\noindent The most related work of our paper is ~\cite{DBLP:conf/icalp/BouajjaniEEH15}.
 Compared to ~\cite{DBLP:conf/icalp/BouajjaniEEH15}, our work are specific to priority queue and thus have points not occur in queues and stacks.

### 2017-04-21 checking-executions.tex

@@
 \label{lemma:EPQ as multi in MRpri for history}
-Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$ if and only if for each $e' \in \textit{proj}(e)$, $\textit{Check-PQ-Conc-NonRec}(e)$ returns $\mathsf{true}$.
+Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$ if and only if for each $e' \in \textit{proj}(e)$, $\textit{Check-PQ-Conc-NonRec}(e')$ returns $\mathsf{true}$.
 \end{restatable}

### 2017-04-21 appendix-EPQ.tex
modify

@@
-\section{Proof in Section \ref{sec:inductive rules of extended priority queue}}
-\label{sec:appendix in section inductive rules of extended priority queue}
+\section{Definition of $\seqPQ$ and Proof of Lemma \ref{lemma:EPQ rules and semantics}}
+\label{sec:appendix definition of seqPQ and proof of Lemma EQP rules and semantics}
 
@@
 
-Let us model priority queue as an LTS $\textit{LTS}_e = (Q,\Sigma,\rightarrow,q_0)$ as follows:
+Let us model priority queue as an LTS $\textit{PQ} = (Q,\Sigma,\rightarrow,q_0)$ as follows:
 
@@
 A trace of an LTS is a finite sequence $\beta_1 \cdot \beta_2 \cdot \ldots \cdot \beta_k$, where $k \geq 0$ if there exists a path $q_0\overset{\beta_1}{\longrightarrow}q_1\overset{\beta_2}{\longrightarrow}\ldots\overset{\beta_k}{\longrightarrow}q_k$ of the LTS.
-Let $\textit{PQ}_s$ be the set of traces of $\textit{LTS}_e$.
-The following lemma states that $\textit{PQ} = \textit{PQ}_s$.
+Let $\seqPQ$ be the set of traces of $\textit{PQ}$.
+The following lemma states that $\seqPQ$ is indeed the set of sequences obtained by renaming sequences accepted by $\textit{Check-PQ-Seq}$.
+
+$\textit{Check-PQ-Seq}(e)=\mathsf{true}$ iff $e\in \seqPQ$, for every data-differentiated sequential execution $e$..
 
@@
 
-To prove that $\textit{PQ} \subseteq \textit{PQ}_s$, we first prove that $\textit{PQ}_{\neq} \subseteq \textit{PQ}_{s\neq}$ by induction.
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] It is obvious that $\epsilon \in \textit{PQ}_s$.
-
-\item[-] If $l_1 \in \textit{PQ}_{s\neq}$ and $l_2 \xrightarrow{\textit{PQ}_1} l_1$.
-Then we need to prove that $l_2 \in \textit{PQ}_s$.
-We know that $l_1 = u \cdot v \cdot w$, such that $\textit{PQ}_1(u,v,w,p)$ holds, and $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v \cdot \textit{rm}(\textit{itm})$.
+Let $\textit{SeqPQ}_f$ be the set of data-differentiated sequences, such that $e \in \textit{SeqPQ}_f$, if $\textit{Check-PQ-Seq}(e)=\mathsf{true}$.
+We need to prove that $\textit{SeqPQ}_f = \seqPQ_{\neq}$.
+Given $\Gamma\in \{\mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, let us use $l_2 \xrightarrow{\Gamma} l_1$ to mean that when we use $\textit{Check-PQ-Seq}$ to check $l_2$, we choose the branch of $\mathsf{Has\text{-}\Gamma}$ and finally recursively call $\textit{Check-PQ-Seq}$ to check $l_1$.
+
+We prove $\textit{SeqPQ}_f \subseteq \seqPQ_{\neq}$ by induction.
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] It is obvious that $\epsilon \in \seqPQ$.
+
+\item[-] If $l_1 \in \seqPQ_{\neq}$ and $l_2 \xrightarrow{\mathsf{MatchedMaxPriority}} l_1$.
+Then we need to prove that $l_2 \in \seqPQ$.
+We know that $l_1 = u \cdot v \cdot w$, such that $\mathsf{MatchedMaxPriority\text{-}Seq}(l_2,x)$ holds and $l_2 = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$.
 
     Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$, $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_j$ and $w = \alpha_{\textit{j+1}} \cdot \ldots \cdot \alpha_m$.
-Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_j} q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}_s$.
+Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_j} q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}$.
 For each $i \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $p$ into $\textit{item} \cdot l_k$ and $q_k$ maps $\textit{pri}$ into $l_k$ for some finite sequence $l_k$.
 
-    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(\textit{itm},p)} q'_i$.
-Since (1) all $\textit{put}$ with priority $p$ is in $u$, and (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $p$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq j$) with a newest $\textit{itm}$ with priority $p$.
-Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j$ are transitions of $\textit{LTS}_e$.
-Since $\textit{matched}_{\prec}(u \cdot v,p)$ holds, we can see that $q_j$ maps each priority that is smaller than $p$ into $\epsilon$ and maps $p$ into $\epsilon$, and $q'_j$ maps each priority that is smaller than $p$ into $\epsilon$ and maps $p$ into $\textit{itm}$.
-Then, we can see that $q'_j \xrightarrow{\textit{rm}(\textit{itm})} q_j$.
+    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(x,p)} q'_i$.
+Since (1) all $\textit{put}$ with priority $p$ is in $u$, and (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $p$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq j$) with a newest $x$ with priority $p$.
+Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j$ are transitions of $\textit{PQ}$.
+Since $\textit{matched}_{\prec}(u \cdot v,p)$ holds, we can see that $q_j$ maps each priority that is smaller than $p$ into $\epsilon$ and maps $p$ into $\epsilon$, and $q'_j$ maps each priority that is smaller than $p$ into $\epsilon$ and maps $p$ into $x$.
+Then, we can see that $q'_j \xrightarrow{\textit{rm}(x)} q_j$.
 We already know that that $q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$.
-Therefore, we can see that $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v \cdot \textit{rm}(\textit{itm}) \in \textit{PQ}_s$.
-
-\item[-] If $l_1 \in \textit{PQ}_{s\neq}$ and $l_2 \xrightarrow{\textit{PQ}_2} l_1$.
-Then we need to prove that $l_2 \in \textit{PQ}_s$.
-We know that $l_1 = u \cdot v$, such that $\textit{PQ}_2(u,v,p)$ holds, and $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v$.
+Therefore, we can see that $l_2 = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \in \seqPQ$.
+
+\item[-] If $l_1 \in \seqPQ_{\neq}$ and $l_2 \xrightarrow{\mathsf{UnmatchedMaxPriority}} l_1$.
+Then we need to prove that $l_2 \in \seqPQ$.
+We know that $l_1 = u \cdot v$, such that $\mathsf{UnmatchedMaxPriority\text{-}Seq}(l_2,x)$ holds and $l_2 = u \cdot \textit{put}(x,p) \cdot v$.
 
     Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$.
-Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}_s$.
+Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}$.
 For each $i \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $p$ into $\textit{item} \cdot l_k$ and $q_k$ maps $p$ into $l_k$ for some finite sequence $l_k$.
 
-    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(\textit{itm},p)} q'_i$.
-Since (1) all $\textit{put}$ with priority $p$ is in $u$, (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $p$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq m$) with a newest $\textit{itm}$ with priority $p$.
-Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q'_m$ are transitions of $\textit{LTS}_e$.
-Therefore, we can see that $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v \in \textit{PQ}_s$.
-
-\item[-] If $l_1 \in \textit{PQ}_{s\neq}$ and $l_2 \xrightarrow{\textit{PQ}_3} l_1$.
-Then we need to prove that $l_2 \in \textit{PQ}_s$.
-We know that $l_1 = u \cdot v$, such that $\textit{PQ}_3(u,v)$ holds, and $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v$.
+    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(x,p)} q'_i$.
+Since (1) all $\textit{put}$ with priority $p$ is in $u$, (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $p$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq m$) with a newest $x$ with priority $p$.
+Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q'_m$ are transitions of $\textit{PQ}$.
+Therefore, we can see that $l_2 = u \cdot \textit{put}(x,p) \cdot v \in \seqPQ$.
+
+\item[-] If $l_1 \in \seqPQ_{\neq}$ and $l_2 \xrightarrow{\mathsf{EmptyRemove}} l_1$.
+Then we need to prove that $l_2 \in \seqPQ$.
+We know that $l_1 = u \cdot v$, such that $\mathsf{EmptyRemove\text{-}Seq}(l_2,o)$ holds, and $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v$ with $o=\textit{rm}(\textit{empty})$.
 
     Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$.
-Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}_s$.
+Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}$.
 
@@
 We already know that $q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q_m$.
-Therefore, we can see that $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v \in \textit{PQ}_s$.
-\end{itemize}
-
-To prove that $\textit{PQ} \subseteq \textit{PQ}_s$.
-It is easy to see that $\textit{PQ}$ and $\textit{PQ}_s$ are both data-independent.
-Given $e_1 \in \textit{PQ}$, by data-independence we know that there is a data-differentiated sequential execution $e' \in \textit{PQ}$ and renaming function $f$, such that $e=f(e')$.
-By $\textit{PQ}_{\neq} \subseteq \textit{PQ}_{s\neq}$, we know that $e' \in \textit{PQ}_{s\neq}$.
-By data-independence, we know that $e \in \textit{PQ}_s$.
-
-To prove that $\textit{PQ}_s \subseteq \textit{PQ}$, let us first prove $\textit{PQ}_{s\neq} \subseteq \textit{PQ}_{\neq}$.
-We show that given $l_2 \in \textit{PQ}_{s\neq}$, how to construct a sequence $l_1$, such that $l_2 \xrightarrow{R} l_1$ for some $R$, and $l_1 \in \textit{PQ}_s$.
-Based on this, we can decompose a sequence of $\textit{PQ}_s$ into $\epsilon$, and this process ensures that this sequence is in $\textit{PQ}$.
+Therefore, we can see that $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v \in \seqPQ$.
+\end{itemize}
+
+To prove that $\seqPQ_{\neq} \subseteq \textit{SeqPQ}_f$, we show that given $l_2 \in \seqPQ_{\neq}$, how to construct a sequence $l_1$, such that $l_2 \xrightarrow{\Gamma} l_1$ for some $\Gamma$, and $l_1 \in \seqPQ$.
+Based on this, we can decompose a sequence of $\seqPQ$ into $\epsilon$, and this process ensures that this sequence is in $\textit{SeqPQ}_f$.
 Note that from a $l_2$ we may construct more than one $l_1$, and this does not influence the correctness of our proof.
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] If $l_2$ contains $\textit{rm}(\textit{empty})$: Assume that $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v$.
+\item[-] If $\mathsf{Has\text{-}EmptyRemoves}(l_2)$: Assume that $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v$.
 It is easy to see that $\textit{matched}(u)$ holds.
 Let $l_1 = u \cdot v$.
-It is easy to see that $l_2 \xrightarrow{\textit{PQ}_3} l_1$, and $\textit{PQ}_3(u,v)$ holds.
+It is easy to see that $l_2 \xrightarrow{\mathsf{EmptyRemove}} l_1$, and $\mathsf{UnmatchedMaxPriority\text{-}Seq}(l_2,o)$ holds for some $o=\textit{rm}(\textit{empty})$.
 
     Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$.
-Since We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{rm}(\textit{empty})} q'_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is transitions of $\textit{LTS}_e$.
+Since We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{rm}(\textit{empty})} q'_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is transitions of $\textit{PQ}$.
 It is easy to see that $q_i = q'_i$, and they map each element in $\mathbb{P}$ into $\epsilon$.
-Then we can see that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is transitions of $\textit{LTS}_e$, and $l_1 \in \textit{PQ}_s$.
-
-\item[-] If $l_2$ does not contain $\textit{rm}(\textit{empty})$, $p$ is one of maximal priority of $l_2$, and items in $l_2$ with priority $p$ are unmatched $\textit{put}$ and (possibly) matched $\textit{put}$:
-
-    Assume that $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v$, such that all $\textit{put}$ with priority $p$ of $u \cdot v$ is in $u$.
+Then we can see that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is transitions of $\textit{PQ}$, and $l_1 \in \seqPQ$.
+
+\item[-] If $\mathsf{Has\text{-}UnmatchedMaxPriority}(l_2)$: Assume that $l_2 = u \cdot \textit{put}(x,p) \cdot v$, such that all $\textit{put}$ with priority $p$ of $u \cdot v$ is in $u$.
 Let $l_1 = u \cdot v$.
-According to construction of $\textit{LTS}_e$, we can see that $\textit{PQ}_2(u,v,p)$ holds, and $l_2 \xrightarrow{\textit{PQ}_2} l_1$.
+According to construction of $\textit{PQ}$, we can see that $\mathsf{UnmatchedMaxPriority\text{-}Seq}(l_2,x)$ holds, and $l_2 \xrightarrow{\mathsf{UnmatchedMaxPriority}} l_1$.
 
     Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$.
-We already know that $\textit{pa} = q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{put}(\textit{itm},p)} q_{\textit{i+}} \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q_m$ are transitions of $\textit{LTS}_e$.
-For each $\textit{i+1} \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $p$ into some $\textit{itm} \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $p$ into $l_k$.
-Since (1) all $\textit{put}$ with priority $p$ of $u \cdot v$ is in $u$ and (2) $p$ is one of maximal priority of $l_2$, it is safe to remove $\textit{itm}$ without influence other transitions of $\textit{pa}$.
-Or we can say, $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q'_m$ are transitions of $\textit{LTS}_e$.
-Therefore, $l_1 \in \textit{PQ}_s$.
-
-\item[-] If $l_2$ does not contain $\textit{rm}(\textit{empty})$, $p$ is one of maximal priority of $l_2$, and items in $l_2$ with priority $p$ are matched $\textit{put}$:
-
-    Assume that $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w$, such that all $\textit{put}$ with priority $p$ of $u \cdot v \cdot w$ is in $u$.
+We already know that $\textit{pa} = q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{put}(x,p)} q_{\textit{i+}} \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q_m$ are transitions of $\textit{PQ}$.
+For each $\textit{i+1} \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $p$ into some $x \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $p$ into $l_k$.
+Since (1) all $\textit{put}$ with priority $p$ of $u \cdot v$ is in $u$ and (2) $p$ is one of maximal priority of $l_2$, it is safe to remove $x$ without influence other transitions of $\textit{pa}$.
+Or we can say, $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q'_m$ are transitions of $\textit{PQ}$.
+Therefore, $l_1 \in \seqPQ$.
+
+\item[-] If $\mathsf{Has\text{-}MatchedMaxPriority}(l_2)$: Assume that $l_2 = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, such that all $\textit{put}$ with priority $p$ of $u \cdot v \cdot w$ is in $u$.
 Let $l_1 = u \cdot v \cdot w$.
-According to construction of $\textit{LTS}_e$, we can see that $\textit{PQ}_1(u,v,w,p)$ holds, and $l_2 \xrightarrow{\textit{PQ}_1} l_1$.
+According to construction of $\textit{PQ}$, we can see that $\mathsf{MatchedMaxPriority\text{-}Seq}(l_2,x)$ holds, and $l_2 \xrightarrow{\mathsf{MatchedMaxPriority}} l_1$.
 
     Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$, $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_j$ and $w = \alpha_{\textit{j+1}} \cdot \ldots \cdot \alpha_m$.
-We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{put}(\textit{itm},\textit{pri})} q_{\textit{i+}} \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q_j \xrightarrow{\textit{rm}(\textit{itm})} q_{\textit{j+}} \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$.
-For each $\textit{i+1} \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $p$ into $\textit{itm} \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $p$ into $l_k$.
-Since (1) $p$ is one of maximal priority in $l_2$, (2) $\textit{itm}$ is the newest item with priority $p$ in $l_2$, and (3) $\textit{itm}$ is not removed until $\textit{rm}(\textit{itm})$, we know that whether we keep $\textit{itm}$ or remove it will not influence transitions from $q_{\textit{i+1}}$ to $q_j$.
-Then we can see that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ are transitions of $\textit{LTS}_e$.
-Therefore, $l_1 \in \textit{PQ}_s$.
-\end{itemize}
-
-To prove that $\textit{PQ}_s \subseteq \textit{PQ}$.
-Given $e_1 \in \textit{PQ}_s$, by data-independence we know that there is a data-differentiated sequential execution $e' \in \textit{PQ}_s$ and renaming function $f$, such that $e=f(e')$.
-By $\textit{PQ}_{s\neq} \subseteq \textit{PQ}_{\neq}$, we know that $e' \in \textit{PQ}_{\neq}$.
-By data-independence, we know that $e \in \textit{PQ}$.
-\qed
+We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{put}(x,p)} q_{\textit{i+}} \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q_j \xrightarrow{\textit{rm}(x)} q_{\textit{j+}} \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$.
+For each $\textit{i+1} \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $p$ into $x \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $p$ into $l_k$.
+Since (1) $p$ is one of maximal priority in $l_2$, (2) $x$ is the newest item with priority $p$ in $l_2$, and (3) $x$ is not removed until $\textit{rm}(x)$, we know that whether we keep $x$ or remove it will not influence transitions from $q_{\textit{i+1}}$ to $q_j$.
+Then we can see that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ are transitions of $\textit{PQ}$.
+Therefore, $l_1 \in \seqPQ$.
+\end{itemize}
+
+This completes the proof of this lemma. \qed
 \end {proof}
@@
 \begin {proof}
-Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{LTS}_e$.
+Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{PQ}$.
 Assume that $l'$ is generated from $l$ by discarding $o_{\textit{ind1}},\ldots,o_{\textit{indn}}$.
@@
 \begin {proof}
-Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{LTS}_e$.
+Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{PQ}$.
 Assume that $l'$ is generated from $l$ by discarding all operations $o_i$ if (1) $o_i \in O_i(l,\textit{pri})$ and (2) $i \geq k$ for a specific index $k$.
@@
 \begin {proof}
-Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{LTS}_e$.
+Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{PQ}$.
 Let $l_1 = o_1 \cdot \ldots \cdot o_n$, let $l_2 = o_{\textit{n+1}} \cdot \ldots \cdot o_m$, let $l_3 = o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$, let $l_4 = o'_{\textit{k+1}} \cdot \ldots \cdot o'_m$.
@@
 \begin {proof}
-Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{LTS}_e$.
+Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{PQ}$.
 Let $l_1 = o'_1 \cdot \ldots \cdot o'_n$, let $l_3 = o'_{\textit{n+1}} \cdot \ldots \cdot o'_m$.
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] It is easy to see that $l_1 \in \textit{EPQ}$, and let $q'_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n$ be the path of $l_1$ in $\textit{LTS}_e$.
+\item[-] It is easy to see that $l_1 \in \textit{EPQ}$, and let $q'_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n$ be the path of $l_1$ in $\textit{PQ}$.
 
@@
 \begin {proof}
-Let $l=o_1 \cdot \ldots \cdot o_m$ and let $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{LTS}_e$.
-Let $l_1 = o_1 \cdot \ldots \cdot l_k$, $l_3 = o'_1 \cdot \ldots \cdot o'_n$ and let $q_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n$ be the path of $l_3$ in $\textit{LTS}_e$.
+Let $l=o_1 \cdot \ldots \cdot o_m$ and let $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{PQ}$.
+Let $l_1 = o_1 \cdot \ldots \cdot l_k$, $l_3 = o'_1 \cdot \ldots \cdot o'_n$ and let $q_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n$ be the path of $l_3$ in $\textit{PQ}$.
 
 By assumption we know that $q_k = q'_n$.
-Then it is not hard to see that $q_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n \xrightarrow{o_{\textit{k+1}}} q_{\textit{k+1}} \ldots \xrightarrow{o_m} q_m$ is a path in $\textit{LTS}_e$, and then $l' \in \textit{EPQ}_s$.
+Then it is not hard to see that $q_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n \xrightarrow{o_{\textit{k+1}}} q_{\textit{k+1}} \ldots \xrightarrow{o_m} q_m$ is a path in $\textit{PQ}$, and then $l' \in \textit{EPQ}_s$.
 By Lemma \ref{lemma:EPQ rules and semantics}, we know that $l' \in \textit{EPQ}$.

### 2017-04-21 checking-executions.tex
modify

@@
 
-The following lemma states the correctness of $\textit{Check-PQ-Seq}$ (the proof can be found in Appendix~\ref{}).
+The following lemma states the correctness of $\textit{Check-PQ-Seq}$ (the proof can be found in Appendix~\ref{sec:appendix definition of seqPQ and proof of Lemma EQP rules and semantics}).
 

### 2017-04-21 appendix-EPQ.tex
modify

@@
 
-\section{Proof in Section \ref{sec:data-independence of extended priority queue}}
-\label{sec:appendix in section data-independence of EPQ}
+\section{Proofs in Section \ref{sec:priority queue and data-independence}}
+\label{sec:appendix proofs in section priority queue and data-independence}
+
+\EPQClosedUnderProjection*
+
+\begin {proof}
+
+By Lemma \ref{lemma:EPQ rules and semantics}, $\seqPQ$ is equivalent to the set of sequences obtained by renaming sequences accepted by $\textit{Check-PQ-Seq}$.
+It is easy to see that for the predicates of $\textit{Check-PQ-Seq}$, if a sequential execution satisfy it, then its sub-sequence also satisfy it.
+For example, if $\textit{matched}(u)$ holds, then $\textit{matched}(u \vert D)$ holds for each set $D$ of values.
+This completes the proof of this lemma.
+\qed
+\end {proof}
 
@@
 \subsection{Proof of Lemma \ref{lemma:EPQ is closed under projection}}
-
-\begin{restatable}{lemma}{EPQisClosedUnderProjection}
-\label{lemma:EPQ is closed under projection}
-$\textit{EPQ}$ is closed under projection.
-\end{restatable}
-
-\begin {proof}
-
-This is obvious, since for each conditions in the $\textit{Guard}$ part of the rules of priority queue, if a sequence of sequential executions satisfy it, then its sub-sequence also satisfy it.
-For example, if $\textit{noRE}(l)$ holds for some $l = u \cdot v \cdot w$ and let $D_l$ be the set of items of $l$, then for each subset $D' \subseteq D_l$, it is obvious that $\textit{rm}(\textit{empty}) \notin l \vert_{ D' }$ and $\textit{noRE}(l')$ holds.
-Similar cases hold for other predicates of the four rules of $\textit{EPQ}$, such as $\textit{LEI}$, $\textit{LI-U}$, $\textit{matched-C}$, $\textit{putInSeq}(l,l_1,\textit{pri})$ and $\textit{matched-All}$.
-This completes the proof of this lemma.
-\qed
-\end {proof}
 

### 2017-04-21 priority-queues.tex
modify

@@
 Then, $\seqPQ$ is the set of words accepted by $\textit{PQ}$.
-The detailed definition of $\textit{PQ}$ can be found in Appendix \ref{A}.
+The detailed definition of $\textit{PQ}$ can be found in Appendix \ref{sec:appendix definition of seqPQ and proof of Lemma EQP rules and semantics}.
 
@@
 Let $\textit{proj}(e)$ be the set of all projections of $e$ to a set of values.
+{\color {blue} The proof of the following lemma can be found in Appendix \ref{sec:appendix proofs in section priority queue and data-independence}.}
 
-\begin{lemma}\label{lem:closure_proj}
+%\begin{lemma}\label{lem:closure_proj}
+%$\seqPQ$ is closed under projection, i.e., $\textit{proj}(e)\subseteq \seqPQ$ for each $e\in \seqPQ$.
+%\end{lemma}
+\begin{restatable}{lemma}{EPQClosedUnderProjection}
+\label{lem:closure_proj}
 $\seqPQ$ is closed under projection, i.e., $\textit{proj}(e)\subseteq \seqPQ$ for each $e\in \seqPQ$.
-\end{lemma}
+\end{restatable}
 
@@
 The following lemma states that by data-independence, it is enough to consider only data-differentiated executions when checking linearizability.
-This is similar to that in \cite{conf/tacas/AbdullaHHJR13,DBLP:conf/icalp/BouajjaniEEH15}, where they use the notion of data-independence in \cite{conf/popl/Wolper86}.
+{\color {blue}It is proved in Appendix \ref{sec:appendix proofs in section priority queue and data-independence}.} This is similar to that in \cite{conf/tacas/AbdullaHHJR13,DBLP:conf/icalp/BouajjaniEEH15}, where they use the notion of data-independence in \cite{conf/popl/Wolper86}.
 Section~\ref{sec:checking inclusion by recursive procedure} will focus on characterizing linearizability for data-differentiated executions.

### 2017-04-21 co-regular-EPQ.tex
introduction
modify
small things

@@
 
-Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions which contains only one maximal priority.
+Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions which contain only one maximal priority.
 This is possible because any data-differentiated execution for which $\Gamma(e)$ is false has such a projection.
@@
 \begin{figure}[htbp]
-TODO UPDATE THE ARGUMENTS OF ACTIONS
   \centering
@@
 %\vspace{-10pt}
-  \caption{An execution that is not $\mathsf{MatchedMaxPriority}^{>}$-linearizable}
+  \caption{An execution that is not $\mathsf{MatchedMaxPriority}^{>}$-linearizable.
+We represent each operation as a time interval whose left, resp., right, bound corresponds to the call, resp., return action.
+Operations adding and removing the same value are aligned vertically.}
   \label{fig:introduce gap for EPQ1Lar}
@@
 When the left-right constraint of the maximal priority value $x$ contains a cycle of the form $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ for some $d_1$,$\ldots$,$d_n\in \mathbb{D}$, we say that $x$ is \emph{covered} by $d_1,\ldots,d_m$.
-The shape of such a cycle can be detected using our class of automata, the only complication being the unbounded number of values $d_1$,$\ldots$,$d_n$.
+The shape of such a cycle (i.e., the alternation between call/return actions of $\textit{put}$/$\textit{rm}$ operations) can be detected using our class of automata, the only complication being the unbounded number of values $d_1$,$\ldots$,$d_n$.
 However, by data independence, whenever an implementation contains such an execution it also contains an execution where all the values $d_1$,$\ldots$,$d_n$ are renamed to the same value $a$, and $x$ is renamed to $b$.
@@
 For example, for the case when the projection of $e$ to $x$ is as in \figurename~\ref{fig:introduce gap for EPQ1Lar} (rename $b$ into $x$), we construct a register automaton $\mathcal{A}_{\textit{l-lar}}^1$ in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
-Here transition labels $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{call}(\textit{put},a,=r)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{call}(\textit{put},d,\top),\textit{ret}(\textit{put},d,\top), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.
+Here transition labels $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{call}(\textit{put},a,=r)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.
 There are three paths in \figurename~\ref{fig:introduce gap for EPQ1Lar} from $q_{\textit{init}}$ to $q_7$, and they comes from that there are three possible positions for the first $\textit{call}(\textit{put},a,\_)$ in the captured execution, which is shown in \figurename~\ref{fig:executions APQ1Lar-1 in paper}.

### 2017-04-21 co-regular-EPQ.tex
started the = case

@@
 the set of sequential executions described by $\Gamma\mathsf{\text{-}Seq}$ (provided that $\mathsf{Has\text{-}}\Gamma(e)$ holds).
-Therefore, by an abuse of terminology, an execution $e$ satisfying $\Gamma(e)$ is called \emph{$\Gamma$-linearizable}.
+Therefore, by an abuse of terminology, an execution $e$ satisfying $\Gamma(e)$ is called \emph{linearizable w.r.t.
+$\Gamma$}, or \emph{$\Gamma$-linearizable}.
 Extending the automaton characterizing executions which are not $\Gamma$-linearizable, with self-loops that allow any operation with parameter $\top$ results in an automaton satisfying the following property called \emph{$\Gamma$-completeness}.
@@
 Intuitively, it represents the time interval in which a value is guaranteed to be stored into the concurrent priority queue.
+Concretely, for a standard indexing of actions in an execution, a time interval is a closed interval between the indexes of two actions in the execution.
 In \figurename~\ref{fig:introduce gap for EPQ1Lar}, we draw the interval of each value by dashed line.
@@
 
-In this subsection, we briefly introduce the idea for proving co-regular of $\mathsf{MatchedMaxPriority}^{=}$.
+%In this subsection, we briefly introduce the idea for proving co-regular of $\mathsf{MatchedMaxPriority}^{=}$.
 The proof of this subsection can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}.
-
-Given a data-differentiated $\_$-execution $e$ such that $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)$ holds, Lemma \ref{lemma:automata for extended priority queue with single priority} and $\textit{Auts}_{\textit{1-lar}}$ are sill not enough for ensure that $e$ linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$.
-This is because that given values $a$ and $b$ with maximal priority, it is possible that all the possible linearization points of $\textit{rm}(b)$ are disabled by $\textit{rm}(b)$.
-We give an example of such execution $e$ in \figurename~\ref{fig:introduce pb order}.
-The execution $e$ of \figurename~\ref{fig:introduce pb order} is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$, even if $h \vert_{p_1}$ and $h \vert_{p_4}$ both linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$, and either $a$ or $b$ is covered by values with smaller priority.
-The explanation of $e$ not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ is as follows: Since $\textit{put}(a,p_4) <_{\textit{hb}} \textit{put}(b,p_4)$, the linearization points of $\textit{rm}(a)$ should before the linearization point of $\textit{rm}(b)$.
-According to the previous section, which says that linearizaton point of larger priority can be happen in the interval of values with smaller priority.
-Thus, the time intervals identified with dotted lines in \figurename~\ref{fig:introduce pb order} are are possible position to locate the linearization point of $\textit{rm}(b)$.
-However, each of them is before $\textit{call}(\textit{rm},a)$.
+%Given a data-differentiated $\_$-execution $e$ such that $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)$ holds, Lemma \ref{lemma:automata for extended priority queue with single priority} and $\textit{Auts}_{\textit{1-lar}}$ are sill not enough for ensure that $e$ linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$.
+When an execution contains at least two values of maximal priority, the acyclicity of the left-right constraints (for all the maximal priority values) is not enough to conclude that the execution is $\mathsf{MatchedMaxPriority}$-linearizable.
+Intuitively, there may exist a value $a$ which is added before another value $b$ such that all the possible linearization points of $\textit{rm}(b)$ are disabled by the position of $\textit{rm}(a)$ in the happens-before.
+We give an example of such an execution $e$ in \figurename~\ref{fig:introduce pb order}.
+This execution  is not linearizable w.r.t.
+$\mathsf{MatchedMaxPriority}$ (or $\mathsf{MatchedMaxPriority}^{=}$) even if
+%$h \vert_{\prec p_1}$ and $h \vert_{p_4}$ both linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$, and 
+neither $a$ nor $b$ are covered by values with smaller priority. %The explanation of $e$ not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ is as follows: 
+Since $\textit{put}(a,p_4) <_{\textit{hb}} \textit{put}(b,p_4)$ and values of the same priority are removed in FIFO order, $\textit{rm}(a)$ should be linearized before $\textit{rm}(b)$ (i.e., this execution should be linearizable w.r.t.
+a sequence where $\textit{rm}(a)$ occurs before $\textit{rm}(b)$).
+%According to the previous section, which says that linearizaton point of larger priority can be happen in the interval of values with smaller priority. 
+Since $\textit{rm}(b)$ cannot take effect during the interval of a smaller priority value, it could be linearized only in one of the two time intervals pictured with dotted lines in \figurename~\ref{fig:introduce pb order}.
+However, each of these time intervals ends before $\textit{call}(\textit{rm},a)$, and thus $\textit{rm}(a)$ cannot be linearized before $\textit{rm}(b)$.
 
@@
 %\vspace{-10pt}
-  \caption{An execution that does not linearizable w.r.t $\textit{MS}(\textit{EPQ}_1^{=})$}
+  \caption{An execution that is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^=$.}
   \label{fig:introduce pb order}
@@
 
-One point need to be modeled in \figurename~\ref{fig:introduce pb order} is that ``a is putted before b''.
-Let us introduce put-before order $<_{\textit{pb}}$ to formally state that ``an value is putted before another value''.
-Given a data-differentiated $\_$-execution $e$ and two values $a,b$ with maximal priority, $a <_{\textit{pb}} b$, if one of the following cases holds: (1) $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$, (2) $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(b)$, or (3) $\textit{rm}(a) <_{\textit{hb}} \textit{put}(b,\_)$.
-Sometimes we use $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$ and $a <_{\textit{pb}}^C b$ to explicitly distinguish above three cases.
+To recognize the scenarios in \figurename~\ref{fig:introduce pb order}, we introduce an order $<_{\textit{pb}}$ between values which intuitively, can be thought of as ``a value $a$ is putted before another value $b$''.
+%Let us introduce put-before order $<_{\textit{pb}}$ to formally state that ``an value is putted before another value''.
+Thus, given a data-differentiated execution $e$ and two values $a$ and $b$ of maximal priority, $a <_{\textit{pb}} b$ if one of the following holds: (1) $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$, (2) $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(b)$, or (3) $\textit{rm}(a) <_{\textit{hb}} \textit{put}(b,\_)$.
+Sometimes we use $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$ and $a <_{\textit{pb}}^C b$ to explicitly distinguish between these three cases.
 Let $<_{\textit{pb}}^*$ be the transitive closure of $<_{\textit{pb}}$.
-Intuitively, $a <_{\textit{pb}}^* b$ means that we can infer that $a$ is putted before $b$ with the help of several intermediate values.
+%Intuitively, $a <_{\textit{pb}}^* b$ means that we can infer that $a$ is putted before $b$ with the help of several intermediate values.
 
-Another point need to be modeled in \figurename~\ref{fig:introduce pb order} is the possible time points of $\textit{rm}(b)$, which is formally modeled by gap-point.
+Then, to model the time intervals in which a remove operation like $\textit{rm}(b)$ in \figurename~\ref{fig:introduce pb order}, can be linearized (outside of intervals of smaller priority values) we introduce the notion of gap-point.
+This notion relies on a standard indexing of actions in an execution, starting with 0.
 
 \begin{definition}\label{def:gap-point for matched put and rm operations}
-Given a data-differentiated $p$-execution $e$ and two method events $\textit{put}(x,p),\textit{rm}(x)$ of $e$.
-We say that a time-point $o$ is a gap-point of $x$, if $o$ is after $\textit{call}(\textit{put},x,p)$ and $\textit{call}(\textit{rm},x)$, is before $\textit{ret}(\textit{rm},x)$, and is not in interval of any values with smaller priority.
+Let $e$ be a data-differentiated execution which contains only one maximal priority $p$, and $\textit{put}(x,p)$ and $\textit{rm}(x)$ two operations in $e$.
+An index $i\in [0,|e|-1]$ is a \emph{gap-point of $x$} if $i$ is greater than or equal to the index of both $\textit{call}(\textit{put},x,p)$ and $\textit{call}(\textit{rm},x)$, smaller than the index of $\textit{ret}(\textit{rm},x)$, and it is not included in the interval of some value with priority smaller than $p$.
 \end{definition}
 
-The case of \figurename~\ref{fig:introduce pb order} can be formally described as follows: $a <_{\textit{pb}}^* b$, while the right-most gap-point of $b$ is before $\textit{call}(\textit{rm},a)$ or $\textit{call}(\textit{put},a,p_4)$.
-The following lemma states that getting rid of such case is enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$.
+The case of \figurename~\ref{fig:introduce pb order} can be formally described as follows: $a <_{\textit{pb}}^* b$ while the right-most gap-point of $b$ is before $\textit{call}(\textit{rm},a)$ or $\textit{call}(\textit{put},a,p_4)$.
+The following lemma states that these conditions are enough to characterize non-linearizability w.r.t $\mathsf{MatchedMaxPriority}^{=}$.
 
@@
 \label{lemma:EPQ1Equal as pb order and gap-point}
-Given a data-differentiated $p$-execution $e$ where $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)$ holds.
-$e$ is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$, if and only if there exists $x$ and $y$ with maximal priority $p$, such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is before $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$ in $e$.
+Let $e$ be a data-differentiated execution which contains only one maximal priority $p$ such that $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds.
+%Given a data-differentiated $p$-execution $e$ where $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)$ holds. 
+Then, $e$ is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ iff $e$ contains two values $x$ and $y$ of maximal priority $p$ such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is strictly smaller than the index of both $\textit{call}(\textit{put},y,p)$ and $\textit{call}(\textit{rm},y)$.
 \end{restatable}
@@
 \begin {proof} (Sketch)
-
 We have already intuitively explain the proof of the $\textit{if}$ direction.
@@
 This implies that $e$ is linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ and completes the proof of the $\textit{only if}$ direction.
-\qed
 \end {proof}

### 2017-04-21 appendix-EPQ.tex
modify

@@
 Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_j} q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}$.
-For each $i \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $p$ into $\textit{item} \cdot l_k$ and $q_k$ maps $\textit{pri}$ into $l_k$ for some finite sequence $l_k$.
+For each $i \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $p$ into $x \cdot l_k$ and $q_k$ maps $p$ into $l_k$ for some finite sequence $l_k$.
 
     We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(x,p)} q'_i$.
-Since (1) all $\textit{put}$ with priority $p$ is in $u$, and (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $p$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq j$) with a newest $x$ with priority $p$.
+Since (1) all $\textit{put}$ with priority $p$ is in $u$, and (2) in $u \cdot v$, only values with priority either incomparable, or less, or equal than $p$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq j$) with a newest $x$ with priority $p$.
 Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j$ are transitions of $\textit{PQ}$.
@@
 Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}$.
-For each $i \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $p$ into $\textit{item} \cdot l_k$ and $q_k$ maps $p$ into $l_k$ for some finite sequence $l_k$.
+For each $i \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $p$ into $x \cdot l_k$ and $q_k$ maps $p$ into $l_k$ for some finite sequence $l_k$.
 
     We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(x,p)} q'_i$.
-Since (1) all $\textit{put}$ with priority $p$ is in $u$, (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $p$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq m$) with a newest $x$ with priority $p$.
+Since (1) all $\textit{put}$ with priority $p$ is in $u$, (2) in $u \cdot v$, only values with priority either incomparable, or less, or equal than $p$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq m$) with a newest $x$ with priority $p$.
 Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q'_m$ are transitions of $\textit{PQ}$.
@@
 For each $\textit{i+1} \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $p$ into $x \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $p$ into $l_k$.
-Since (1) $p$ is one of maximal priority in $l_2$, (2) $x$ is the newest item with priority $p$ in $l_2$, and (3) $x$ is not removed until $\textit{rm}(x)$, we know that whether we keep $x$ or remove it will not influence transitions from $q_{\textit{i+1}}$ to $q_j$.
+Since (1) $p$ is one of maximal priority in $l_2$, (2) $x$ is the newest value with priority $p$ in $l_2$, and (3) $x$ is not removed until $\textit{rm}(x)$, we know that whether we keep $x$ or remove it will not influence transitions from $q_{\textit{i+1}}$ to $q_j$.
 Then we can see that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ are transitions of $\textit{PQ}$.
@@
 To prove the $\textit{only if}$ direction, given a data-differentiated execution $e \in \mathcal{I}_{\neq}$.
-By assumption, it is linearizable with respect to a sequential execution $l \in S$, and the bijection between the operations of $e$ and the method events of $l$ ensures that $l$ is differentiated and belongs to $S_{\neq}$.
+By assumption, it is linearizable with respect to a sequential execution $l \in S$, and the bijection between the operations of $e$ and the operations of $l$ ensures that $l$ is differentiated and belongs to $S_{\neq}$.
 
@@
 
-\section{Proofs in Section \ref{sec:step-by-step linearizability of extended priority queues}}
-\label{sec:appendix in section step-by-step linearizability of extended priority queues}
-
-\subsection{Proof of Lemma \ref{lemma:EPQ is closed under projection}}
-
-\subsection{Proof of Lemma \ref{lemma:EPQ as multi in MRpri for sequence}}
-
-%\EPQasMultiInMRpriforSequence*
-
-\begin{restatable}{lemma}{EPQasMultiInMRpriforSequence}
-\label{lemma:EPQ as multi in MRpri for sequence}
-Given a data-differentiated sequential execution $e$, $e \in \textit{EPQ}$, if and only if, $\forall e' \in \textit{proj}(e)$ and $\forall R \in \textit{last}(e')$, we have $e' \in \textit{MS}(R)$.
-\end{restatable}
-
-\begin {proof}
-
-The \textit{only if} direction is obvious and can be similarly proved as the $\textit{EPQ}_s \subseteq \textit{EPQ}$ direction of Lemma \ref{lemma:EPQ rules and semantics}.
-
-To prove the $\textit{if}$ direction, we proceed as follows: From $e1=e$, we generate a sequence $e_2$ as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $\textit{PQ}_3 \in \textit{last}(e_1)$: Then we can see that $e_1$ contains at least one $\textit{rm}(\textit{empty})$.
-$e_2$ is generated from $e_1$ by erasing one $\textit{rm}(\textit{empty})$.
-
-\item[-] Else, if $\textit{PQ}_2^{=} \in \textit{last}(e_1)$: Then we can see that one of the maximal priority of $e_1$ is unmatched $\textit{put}$ and (possibly) matched $\textit{put}$.
-Assume the set of the items of these unmatched $\textit{put}$ is $S$.
-$e_2$ is generated from $e_1$ by erasing one unmatched $\textit{put}$ which use the item last putted in $S$.
-
-\item[-] Else, if $\textit{PQ}_2^{>} \in \textit{last}(e_1)$: Then we can see that one of the maximal priority in $e_1$ is unmatched $\textit{put}$.
-Assume the set of the items of these unmatched $\textit{put}$ is $S$.
-$e_2$ is generated from $e_1$ by erasing one unmatched $\textit{put}$ which use the item last putted in $S$.
-
-\item[-] Else, if $\textit{PQ}_1^{=} \in \textit{last}(e_1)$: Then we can see that one of the maximal priority in $e_1$ is of more than one pair of matched $\textit{put}$.
-Assume the set of the items of these matched $\textit{put}$ is $S$.
-$e_2$ is generated from $e_1$ by erasing matched $\textit{put}$ and $\textit{rm}$ of the item which is last putted in $S$.
-
-\item[-] Else, if $\textit{PQ}_1^{>} \in \textit{last}(e_1)$: Then we can see that one of the maximal priority in $e_1$ is of one pair of matched $\textit{put}$.
-$e_2$ is generated from $e_1$ by erasing this pair of matched $\textit{put}$ and $\textit{rm}$.
-\end{itemize}
-
-Similarly, for each $i > 1$, we obtain $e_{\textit{i+1}}$ from $e_i$, until we obtain $e_m = \epsilon$ for some $m$. It is obvious that $e_m \in \textit{EPQ}$. For $e_{\textit{m-1}}$, since
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $e_m \in \textit{EPQ}$,
-
-\item[-] By assumption, we know that $e_{\textit{m-1}} \in \textit{MS}(R_{\textit{m-1}})$, where $R_{\textit{m-1}} \in \textit{last}(e_{\textit{m-1}})$.
-This implies that the guard of $R_{\textit{m-1}}$ is satisfied.
-\end{itemize}
-
-Therefore, we know that $e_{\textit{m-1}} \in \textit{EPQ}$. Similarly, we can prove that $e_{\textit{m-2}},\ldots,e_1 = e \in \textit{EPQ}$. \qed
-\end {proof}
-
-\subsection{Proof of Lemma \ref{lemma:EPQ is step-by-step linearizability}}
-
-The prove that $\textit{EPQ}$ is step-by-step linearizability, we investigate each rules individually.
-
-Given a data-differentiated execution and its history, we can abuse notation and mix labels and method events with operations themselves, since items are unique in a data-differentiated execution.
-For instance, we will reference an operation labeled by $\textit{put}(p,a)$ as $\textit{put}(p,a)$.
-
-Given an operation $o$ with call action $\textit{cal}_o (\textit{put},a,p)$ and return actions $\textit{ret}_o (\textit{put})$, its method event is $\textit{put}(a,p)$.
-Given an operation $o$ with call action $\textit{cal}_o (\textit{rm})$ and return actions $\textit{ret}_o (\textit{rm},a)$, its method event is $\textit{rm}(a)$.
-
-Given a data-differentiated execution $e$ and its history $h$, we can obtain a sequence $h'$ from $h$ by adding $\textit{put}(a,p)$ (resp., $\textit{rm}(a)$, $\textit{rm}(\textit{empty})$) between each pair of $\textit{cal}(\textit{put},a,p)$ and $\textit{ret}(\textit{put},a)$ (resp., $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$, $\textit{cal}(\textit{rm},\textit{empty})$ and $\textit{ret}(\textit{rm},\textit{empty})$).
-The projection of $h'$ into method events is called linearization of $e$ and $h$, and each method event we add in $h'$ can be considered as a linearization point of the corresponding method event.
-We call such $h'$ an execution with linearization points of $e$.
-
-Before we prove the step-by-step linearizability of $\textit{MS}(\textit{EPQ}_1)$ and $\textit{MS}(\textit{EPQ}_2)$, we introduce several lemmas, which are used to ensure some sub-sequences of $\textit{EPQ}$ still belongs to $\textit{EPQ}$.
-Given a data-differentiated sequence $l$ and one of its maximal priority $\textit{pri}$, let $O_c(l,\textit{pri})$ and $O_i(l,\textit{pri})$ be the set of operations with priorities comparable with $\textit{pri}$ and incomparable with $\textit{pri}$ in $l$, respectively.
-Similarly we can define $D_c(l,\textit{pri})$ and $D_i(l,\textit{pri})$ for items instead of operations.
-%Let $O_x(l)$ be the set of $\textit{put}(y,\textit{pri})$ with $y \neq x$ in $l$, where $\textit{pri}$ is the priority of $x$ in $l$.
-We can see that each priority of items in $O_i(l,\textit{pri})$ is either larger or incomparable with priorities of items in  $O_c(l,\textit{pri})$.
-
-The following lemma shows that if a new sequence is generated by erasing some operations in $O_c(l,\textit{pri})$ while keeping the remaining $O_c(l,\textit{pri})$ sub-sequences in $\textit{EPQ}$, then this new sequence is still in $\textit{EPQ}$.
-Note that this is different from projection on items.
+\section{Proofs in Section \ref{sec:checking inclusion by recursive procedure}}
+\label{sec:appendix in section checking lnclusion by recursive procedure}
+
+\subsection{Definition of Step-by-Step linearizablity}
+\label{sec:appendix definition of step-by-step linearizability}
+
+We introduce the notion of step-by-step linearizability, which means that from a linearization of $e \setminus x$ that satisfy the requirements of priority queue, we can obtain a linearization of $e$ that satisfy the requirements of priority queue.
+Its formal definition is as follows:
+
+%\vspace{-6pt}
+\begin{definition}\label{def:step-by-step linearizability}
+Given $\Gamma\in \{\mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, $\Gamma$ is step-by-step linearizability, if for each data-differentiated execution $e$ where $\Gamma\mathsf{\text{-}Conc}(e,\alpha)$ for some $\alpha$, then $e \setminus \alpha \sqsubseteq \seqPQ \Rightarrow e \sqsubseteq \seqPQ$.
+
+$\seqPQ$ is step-by-step linearizability, if each $\Gamma$ is step-by-step linearizability. 
+\end{definition}
+
+Our notion of step-by-step linearizability is inspired by the step-by-step linearizability in ~\cite{DBLP:conf/icalp/BouajjaniEEH15}.
+
+Given a data-differentiated execution $e$, we can obtain a sequence $e'$ from $e$ by adding $\textit{put}(a,p)$ (resp., $\textit{rm}(a)$, $\textit{rm}(\textit{empty})$) between each pair of $\textit{call}(\textit{put},a,p)$ and $\textit{ret}(\textit{put},a,p)$ (resp., $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$, $\textit{call}(\textit{rm},\textit{empty})$ and $\textit{ret}(\textit{rm},\textit{empty})$).
+Such $e'$ is called an execution with linearization points, and we call the projection of $e'$ into $m(a,b)$ the linearization of $e$.
+
+\subsection{Obtaining New Sequences while Ensuring that they are in $\seqPQ$}
+\label{sec:obtaining new sequences while ensureing that they are in SeqPQ}
+
+Before we prove the step-by-step linearizability of $\seqPQ$, we introduce several lemmas, which are used to ensure some sub-sequences of $\seqPQ$ still belongs to $\seqPQ$.
+Given a data-differentiated sequence $l$ and one of its maximal priority $p$, let $O_c(l,p)$ and $O_i(l,p)$ be the set of operations with priorities comparable with $p$ and incomparable with $p$ in $l$, respectively.
+Similarly we can define $D_c(l,p)$ and $D_i(l,p)$ for se of values instead of set of operations.
+We can see that each priority of values in $O_i(l,p)$ is either larger or incomparable with priorities of values in  $O_c(l,p)$.
+
+The following lemma shows that if a new sequence is generated by erasing some operations in $O_c(l,p)$ while keeping the remaining $O_c(l,p)$ sub-sequences in $\seqPQ$, then this new sequence is still in $\seqPQ$.
+Note that this is different from projection on value.
 
@@
 \label{lemma:erase Oc still in EPQ}
-Given a data-differentiated sequential execution $l \in \textit{EPQ}$ and a maximal priority $\textit{pri}$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$.
-Let $l'$ be generated from $l$ by discarding some operations in $O_c(l,\textit{pri})$, and $l' \vert_{ O_c(l,\textit{pri}) } \in \textit{EPQ}$.
-Then, $l' \in \textit{EPQ}$.
+Given a data-differentiated sequential execution $l \in \seqPQ$ and a maximal priority $p$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$.
+Let $l'$ be generated from $l$ by discarding some operations in $O_c(l,p)$, and $l' \vert_{ O_c(l,p) } \in \seqPQ$.
+Then, $l' \in \seqPQ$.
 \end{restatable}
@@
 Let $D$ be the set such that $D$ contains $a$, if $\textit{put}(a,\_)$ is in $o_{\textit{ind1}},\ldots,o_{\textit{indn}}$.
-For each $i$, let $q'_i$ be generated from $q_i$ by erasing items in $D$.
+For each $i$, let $q'_i$ be generated from $q_i$ by erasing values in $D$.
 
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] If $\textit{rm}(a) \in O_c(l,\textit{pri})$: By assumption, $l' \vert_{ O_c(l,\textit{pri}) } \in \textit{EPQ}$.
-Therefore, $a$ is in $q'_j$ and is the should-be-removed item in $O_c(l,\textit{pri})$.
-Since each priority of items in $O_i(l,\textit{pri})$ is either larger or incomparable with priorities of items in $O_c(l,\textit{pri})$, we can removed $a_c$ from $q'_j$, and then $q'_j \xrightarrow{o_{\textit{j+1}}} q'_{\textit{j+1}}$.
-
-\item[-] If $\textit{rm}(a) \in O_i(l,\textit{pri})$: By assumption we know that $q_j \xrightarrow{\textit{rm}(a)} q_{\textit{j+1}}$.
-Since $q'_j$ contains the same $D_i(l,\textit{pri})$ items as $q_j$ and $q'_j$ contains less $D_c(l,\textit{pri})$ items than $q_j$, we can see that $q'_j \xrightarrow{o_{\textit{j+1}}} q'_{\textit{j+1}}$.
+\item[-] If $\textit{rm}(a) \in O_c(l,p)$: By assumption, $l' \vert_{ O_c(l,p) } \in \seqPQ$.
+Therefore, $a$ is in $q'_j$ and is the should-be-removed value in $O_c(l,p)$.
+Since each priority of values in $O_i(l,p)$ is either larger or incomparable with priorities of values in $O_c(l,p)$, we can removed $a_c$ from $q'_j$, and then $q'_j \xrightarrow{o_{\textit{j+1}}} q'_{\textit{j+1}}$.
+
+\item[-] If $\textit{rm}(a) \in O_i(l,p)$: By assumption we know that $q_j \xrightarrow{\textit{rm}(a)} q_{\textit{j+1}}$.
+Since $q'_j$ contains the same $D_i(l,p)$ values as $q_j$ and $q'_j$ contains less $D_c(l,p)$ values than $q_j$, we can see that $q'_j \xrightarrow{o_{\textit{j+1}}} q'_{\textit{j+1}}$.
 \end{itemize}
@@
 For each $q'_{\textit{indj-1}}$, it is easy to see that $q'_{\textit{indj-1}} = q'_{\textit{indj}}$.
-Therefore, we can see that $l' = o_1 \cdot \ldots o_{\textit{ind1-1}} \cdot o_{\textit{ind1+1}} \cdot \ldots \in \textit{EPQ}$.
+Therefore, we can see that $l' = o_1 \cdot \ldots o_{\textit{ind1-1}} \cdot o_{\textit{ind1+1}} \cdot \ldots \in \seqPQ$.
 \qed
@@
 
-The following lemma shows that if a new sequence is generated by from some time point, erasing operations in $O_i(l,\textit{pri})$, then this new sequence is still in $\textit{EPQ}$.
+The following lemma shows that if a new sequence is generated by from some time point, erasing operations in $O_i(l,p)$, then this new sequence is still in $\seqPQ$.
 
@@
 \label{lemma:erase Oi from some time point still in EPQ}
-Given a data-differentiated sequential execution $l \in \textit{EPQ}$ and a maximal priority $\textit{pri}$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$.
-Let $l'$ be generated from $l$ by discarding operations in $O_i(l,\textit{pri})$ from some time point, then, $l' \in \textit{EPQ}$.
+Given a data-differentiated sequential execution $l \in \seqPQ$ and a maximal priority $p$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$.
+Let $l'$ be generated from $l$ by discarding operations in $O_i(l,p)$ from some time point, then, $l' \in \seqPQ$.
 \end{restatable}
@@
 Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{PQ}$.
-Assume that $l'$ is generated from $l$ by discarding all operations $o_i$ if (1) $o_i \in O_i(l,\textit{pri})$ and (2) $i \geq k$ for a specific index $k$.
+Assume that $l'$ is generated from $l$ by discarding all operations $o_i$ if (1) $o_i \in O_i(l,p)$ and (2) $i \geq k$ for a specific index $k$.
 Let $D$ be a set such that $a \in D$, if $\textit{put}(a,\_)$ is in $l$ and not in $l'$.
-For each $0 \leq i \leq m$, let $q'_i$ be generated from $q_i$ by erasing items in $D$.
+For each $0 \leq i \leq m$, let $q'_i$ be generated from $q_i$ by erasing values in $D$.
 
@@
 \item[-] If $o_{f(k)}$ is a $\textit{put}$ operation, then it is obvious that $q'_{f(k)-1} \xrightarrow{o_k} q'_{f(k)}$.
-Else, if $o_{f(k)} = \textit{rm}(a)$, we can see $a$ is in $q'_{f(k)-1}$, and since (1) $q'_{f(k)-1}$ contains the same $D_c(l,\textit{pri})$ items as $q_{f(k)-1}$ and $q'_{f(k)-1}$ contains less $D_i(l,\textit{pri})$ items than $q_{f(k)-1}$, and (2) each priority of items in $O_i(l,\textit{pri})$ is either larger or incomparable with priorities of items in  $O_c(l,\textit{pri})$, we can see that $q'_{f(k)-1} \xrightarrow{o_k} q'_{f(k)}$.
+Else, if $o_{f(k)} = \textit{rm}(a)$, we can see $a$ is in $q'_{f(k)-1}$, and since (1) $q'_{f(k)-1}$ contains the same $D_c(l,p)$ values as $q_{f(k)-1}$ and $q'_{f(k)-1}$ contains less $D_i(l,p)$ values than $q_{f(k)-1}$, and (2) each priority of values in $O_i(l,p)$ is either larger or incomparable with priorities of values in  $O_c(l,p)$, we can see that $q'_{f(k)-1} \xrightarrow{o_k} q'_{f(k)}$.
 Similarly we can prove the case of $o_{f(j)}$ with $j > k$.
@@
 \label{lemma:make put with maximal priority happen earlier}
-Given a data-differentiated sequential execution $l \in \textit{EPQ}$ and a maximal priority $\textit{pri}$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$.
+Given a data-differentiated sequential execution $l \in \seqPQ$ and a maximal priority $p$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$.
 Let $l=l_1 \cdot l_2$.
-Let $l_3$ be the projection of $l_2$ into $\{ \textit{put}(\_,\textit{pri}) \}$, and $l_4$ be the projection of $l_2$ into other operations.
-Then, $l'=l_1 \cdot l_3 \cdot l_4 \in \textit{EPQ}$.
+Let $l_3$ be the projection of $l_2$ into $\{ \textit{put}(\_,p) \}$, and $l_4$ be the projection of $l_2$ into other operations.
+Then, $l'=l_1 \cdot l_3 \cdot l_4 \in \seqPQ$.
 \end{restatable}
@@
 
-\item[-] For $\textit{n+1} \leq i \leq k$, let $q'_i$ be obtained from $q_n$ by adding items in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ with priority $\textit{pri}$ and in the same order.
-
-\item[-] For $\textit{k+1} \leq i \leq m$, let $q'_i$ be obtained from $q_{f(i)}$ by adding items which are (1) with priority $\textit{pri}$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(i)}$.
+\item[-] For $\textit{n+1} \leq i \leq k$, let $q'_i$ be obtained from $q_n$ by adding values in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ with priority $p$ and in the same order.
+
+\item[-] For $\textit{k+1} \leq i \leq m$, let $q'_i$ be obtained from $q_{f(i)}$ by adding values which are (1) with priority $p$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(i)}$.
 The order of adding them is the same as $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$.
@@
 
-\item[-] For $q'_{\textit{k+1}}$: We already know that $q_{f(k+1)-1} \xrightarrow{o_{f(k+1)}} q_{f(k+1)}$, and it is easy to see that $q_{f(k+1)-1}$ is obtained from $q_n$ by adding items in $o_{\textit{n+1}} \cdot \ldots \cdot o_{f(k+1)-1}$.
-
-    We can see that $q'_k$ is obtained from $q_n$ by adding items in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$, and $q'_{k+1}$ is obtained from $q_{f(k+1)}$ by adding items which are (1) with priority $\textit{pri}$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(k+1)}$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $o_{f(k+1)}$ is an operation of non-$\textit{pri}$ items, then $q'_{k+1}$ is obtained from $q_{f(k+1)}$ by adding items in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$.
-Since non-$\textit{pri}$ priority is either smaller or incomparable with $\textit{pri}$, we can see that $q'_k \xrightarrow{o_{f(k+1)}} q'_{\textit{k+1}}$.
-
-    \item[-] Otherwise, it is only possible that $o_{f(k+1)} = \textit{rm}(a)$ for some item $a$ with priority $\textit{pri}$.
-We can see that $q'_{k+1}$ is obtained from $q_{f(k+1)}$ by adding items in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$ and then remove $a$.
-Since $q_{f(k+1)-1} \xrightarrow{o_{f(k+1)}} q_{f(k+1)}$, in $q_{f(k+1)-1}$( and also in $q'_k$), there is no item with priority less than $\textit{pri}$, and $a$ is the first-input item of priority $\textit{pri}$.
+\item[-] For $q'_{\textit{k+1}}$: We already know that $q_{f(k+1)-1} \xrightarrow{o_{f(k+1)}} q_{f(k+1)}$, and it is easy to see that $q_{f(k+1)-1}$ is obtained from $q_n$ by adding values in $o_{\textit{n+1}} \cdot \ldots \cdot o_{f(k+1)-1}$.
+
+    We can see that $q'_k$ is obtained from $q_n$ by adding values in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$, and $q'_{k+1}$ is obtained from $q_{f(k+1)}$ by adding values which are (1) with priority $p$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(k+1)}$.
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $o_{f(k+1)}$ is an operation of non-$p$ values, then $q'_{k+1}$ is obtained from $q_{f(k+1)}$ by adding values in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$.
+Since non-$p$ priority is either smaller or incomparable with $p$, we can see that $q'_k \xrightarrow{o_{f(k+1)}} q'_{\textit{k+1}}$.
+
+    \item[-] Otherwise, it is only possible that $o_{f(k+1)} = \textit{rm}(a)$ for some value $a$ with priority $p$.
+We can see that $q'_{k+1}$ is obtained from $q_{f(k+1)}$ by adding values in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$ and then remove $a$.
+Since $q_{f(k+1)-1} \xrightarrow{o_{f(k+1)}} q_{f(k+1)}$, in $q_{f(k+1)-1}$( and also in $q'_k$), there is no value with priority less than $p$, and $a$ is the first-input value of priority $p$.
 Therefore, we can see that $q'_k \xrightarrow{o_{f(k+1)}} q'_{\textit{k+1}}$.
@@
 
-\item[-] For $q'_{\textit{k+i}}$ with $i>1$: We already know that $q_{f(k+i)-1} \xrightarrow{o_{f(k+i)}} q_{f(k+i)}$, and $q_{f(k+i)-1}$ is obtained from $q_{f(k+i-1)}$ by adding items in $o_{f(k+i-1)+1} \cdot \ldots \cdot o_{f(k+i)-1}$.
-
-    We can see that $q'_{\textit{k+i}-1}$ (resp., $q'_{\textit{k+i}}$) is obtained from $q_{f(k+i-1)}$ (resp., $q_{f(k+i)}$) by adding items which are (1) with priority $\textit{pri}$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(k+i-1)}$ (resp., $o_1 \cdot \ldots \cdot o_{f(k+i)}$).
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $o_{f(k+i)}$ is an operation of non-$\textit{pri}$ items, then $q'_{k+i}$ is obtained from $q_{f(k+i)}$ by adding items which are (1) with priority $\textit{pri}$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(k+i-1)}$.
-Since non-$\textit{pri}$ priority is either smaller or incomparable with $\textit{pri}$, we can see that $q'_{\textit{k+i-1}} \xrightarrow{o_{f(k+i)}} q'_{\textit{k+i}}$.
-
-    \item[-] Otherwise, it is only possible that $o_{f(k+i)} = \textit{rm}(a)$ for some item $a$ with priority $\textit{pri}$.
-We can see that $q'_{k+i}$ is obtained from $q_{f(k+i)}$ by adding items in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$ and then remove $a$.
-Since $q_{f(k+i)-1} \xrightarrow{o_{f(k+i)}} q_{f(k+i)}$, in $q_{f(k+i)-1}$( and also in $q'_{\textit{k+i-1}}$), there is no item with priority less than $\textit{pri}$, and $a$ is the first-input item of priority $\textit{pri}$.
+\item[-] For $q'_{\textit{k+i}}$ with $i>1$: We already know that $q_{f(k+i)-1} \xrightarrow{o_{f(k+i)}} q_{f(k+i)}$, and $q_{f(k+i)-1}$ is obtained from $q_{f(k+i-1)}$ by adding values in $o_{f(k+i-1)+1} \cdot \ldots \cdot o_{f(k+i)-1}$.
+
+    We can see that $q'_{\textit{k+i}-1}$ (resp., $q'_{\textit{k+i}}$) is obtained from $q_{f(k+i-1)}$ (resp., $q_{f(k+i)}$) by adding values which are (1) with priority $p$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(k+i-1)}$ (resp., $o_1 \cdot \ldots \cdot o_{f(k+i)}$).
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $o_{f(k+i)}$ is an operation of non-$p$ values, then $q'_{k+i}$ is obtained from $q_{f(k+i)}$ by adding values which are (1) with priority $p$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(k+i-1)}$.
+Since non-$p$ priority is either smaller or incomparable with $p$, we can see that $q'_{\textit{k+i-1}} \xrightarrow{o_{f(k+i)}} q'_{\textit{k+i}}$.
+
+    \item[-] Otherwise, it is only possible that $o_{f(k+i)} = \textit{rm}(a)$ for some value $a$ with priority $p$.
+We can see that $q'_{k+i}$ is obtained from $q_{f(k+i)}$ by adding values in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$ and then remove $a$.
+Since $q_{f(k+i)-1} \xrightarrow{o_{f(k+i)}} q_{f(k+i)}$, in $q_{f(k+i)-1}$( and also in $q'_{\textit{k+i-1}}$), there is no value with priority less than $p$, and $a$ is the first-input value of priority $p$.
 Therefore, we can see that $q'_{\textit{k+i-1}} \xrightarrow{o_{f(k+i)}} q'_{\textit{k+i}}$.
@@
 
-The following lemma shows that if a new sequence is generated by make some $O_i(l,\textit{pri})$ behaviors to happen earlier, then this new sequence is still in $\textit{EPQ}$.
+The following lemma shows that if a new sequence is generated by make some $O_i(l,p)$ behaviors to happen earlier, then this new sequence is still in $\seqPQ$.
 
@@
 \label{lemma:make Oi happen earlier still in EPQ}
-Given a data-differentiated sequential execution $l \in \textit{EPQ}$ and a maximal priority $\textit{pri}$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$.
-Let $l \vert_{ O_i(l,\textit{pri}) } = l_1 \cdot l_2$, let $l' = l_1 \cdot l_3$, where $l_3$ is the projection of $l$ into non-$l_1$ operations.
-Then, $l' \in \textit{EPQ}$.
+Given a data-differentiated sequential execution $l \in \seqPQ$ and a maximal priority $p$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$.
+Let $l \vert_{ O_i(l,p) } = l_1 \cdot l_2$, let $l' = l_1 \cdot l_3$, where $l_3$ is the projection of $l$ into non-$l_1$ operations.
+Then, $l' \in \seqPQ$.
 \end{restatable}
@@
 Let $f$ be a function, such that $f(i)=j$, if $o'_i = o_j$.
-Let $D$ be the se of items which are added and not removed in $o'_1 \cdot \ldots \cdot o'_n$.
+Let $D$ be the set of values which are added and not removed in $o'_1 \cdot \ldots \cdot o'_n$.
 
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] It is easy to see that $l_1 \in \textit{EPQ}$, and let $q'_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n$ be the path of $l_1$ in $\textit{PQ}$.
-
-\item[-] For $\textit{n+1} \leq i \leq m$, let $q'_i$ be obtained from $q_{f(i)}$ by adding items in $D$. The order of adding them is the same as $o'_1 \cdot \ldots \cdot o'_n$.
+\item[-] It is easy to see that $l_1 \in \seqPQ$, and let $q'_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n$ be the path of $l_1$ in $\textit{PQ}$.
+
+\item[-] For $\textit{n+1} \leq i \leq m$, let $q'_i$ be obtained from $q_{f(i)}$ by adding values in $D$. The order of adding them is the same as $o'_1 \cdot \ldots \cdot o'_n$.
 \end{itemize}
@@
 
-\item[-] For $q'_{\textit{n+i}}$: We already know that $q_{f(n+i)-1} \xrightarrow{o_{f(n+i)}} q_{f(n+i)}$, and it is easy to see that $q_{f(n+i)-1}$ is obtained from $q_{f(n+i-1)}$ by adding $D$-items in $o_{f(n+i-1)+1} \cdot \ldots \cdot q_{f(n+i)-1}$.
-
-     We can see that $q'_{\textit{n+i}-1}$ (resp., $q'_{\textit{n+i}}$) is obtained from $q_{f(n+i-1)}$ (resp., $q_{f(n+i)}$) by adding remanning items in $D$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $o_{f(n+1)}$ is an operation of $O_c(l,\textit{pri})$ items, since priority in $O_i(l,\textit{pri})$ is either larger or incomparable with priority in $O_c(l,\textit{pri})$, we can see that $q'_n \xrightarrow{o_{f(n+1)}} q'_{\textit{n+1}}$.
-
-    \item[-] Otherwise, it is only possible that $o_{f(k+1)} = \textit{rm}(a)$ for some item $a$ in $O_c(l,\textit{pri})$.
-Since $q_{f(n+i)-1} \xrightarrow{o_{f(n+i)}} q_{f(n+i)}$, in $q_{f(n+i)-1}$( and also in $q'_{\textit{n+i-1}}$), there is no item with priority less than $\textit{pri}$, and $a$ is the first-input item of priority $\textit{pri}$.
+\item[-] For $q'_{\textit{n+i}}$: We already know that $q_{f(n+i)-1} \xrightarrow{o_{f(n+i)}} q_{f(n+i)}$, and it is easy to see that $q_{f(n+i)-1}$ is obtained from $q_{f(n+i-1)}$ by adding $D$-values in $o_{f(n+i-1)+1} \cdot \ldots \cdot q_{f(n+i)-1}$.
+
+     We can see that $q'_{\textit{n+i}-1}$ (resp., $q'_{\textit{n+i}}$) is obtained from $q_{f(n+i-1)}$ (resp., $q_{f(n+i)}$) by adding remanning values in $D$.
+
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $o_{f(n+1)}$ is an operation of $O_c(l,p)$ values, since priority in $O_i(l,p)$ is either larger or incomparable with priority in $O_c(l,p)$, we can see that $q'_n \xrightarrow{o_{f(n+1)}} q'_{\textit{n+1}}$.
+
+    \item[-] Otherwise, it is only possible that $o_{f(k+1)} = \textit{rm}(a)$ for some value $a$ in $O_c(l,p)$.
+Since $q_{f(n+i)-1} \xrightarrow{o_{f(n+i)}} q_{f(n+i)}$, in $q_{f(n+i)-1}$( and also in $q'_{\textit{n+i-1}}$), there is no value with priority less than $p$, and $a$ is the first-input value of priority $p$.
 Therefore, we can see that $q'_{\textit{n+i-1}} \xrightarrow{o_{f(n+i)}} q'_{\textit{n+i}}$.
@@
 
-The following lemma shows that if a new sequence is generated by replacing a prefix with another one which make the priority queue has same content, then this new sequence is still in $\textit{EPQ}$.
+The following lemma shows that if a new sequence is generated by replacing a prefix with another one which make the priority queue has same content, then this new sequence is still in $\seqPQ$.
 
@@
 \label{lemma:replace equivalent prefix still in EPQ}
-Given a data-differentiated sequential execution $l \in \textit{EPQ}$.
+Given a data-differentiated sequential execution $l \in \seqPQ$.
 Let $l=l_1 \cdot l_2$.
-Given $l_3 \in \textit{EPQ}$.
+Given $l_3 \in \seqPQ$.
 Assume that the priority queue has same content after executing $l_1$ and $l_3$.
 Let $l' = l_3 \cdot l_2$.
-Then, $l' \in \textit{EPQ}$.
+Then, $l' \in \seqPQ$.
 \end{restatable}
@@
 By assumption we know that $q_k = q'_n$.
-Then it is not hard to see that $q_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n \xrightarrow{o_{\textit{k+1}}} q_{\textit{k+1}} \ldots \xrightarrow{o_m} q_m$ is a path in $\textit{PQ}$, and then $l' \in \textit{EPQ}_s$.
-By Lemma \ref{lemma:EPQ rules and semantics}, we know that $l' \in \textit{EPQ}$.
+Then it is not hard to see that $q_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n \xrightarrow{o_{\textit{k+1}}} q_{\textit{k+1}} \ldots \xrightarrow{o_m} q_m$ is a path in $\textit{PQ}$, and then $l' \in \seqPQ_s$.
+By Lemma \ref{lemma:EPQ rules and semantics}, we know that $l' \in \seqPQ$.
 \qed
@@
 
-With the help of Lemma \ref{lemma:erase Oc still in EPQ}, Lemma \ref{lemma:erase Oi from some time point still in EPQ}, Lemma \ref{lemma:make put with maximal priority happen earlier}, Lemma \ref{lemma:make Oi happen earlier still in EPQ} and Lemma \ref{lemma:replace equivalent prefix still in EPQ}, we can now prove that $\textit{EPQ}_1$ is step-by-step linearizability.
+\subsection{Proving Step-by-Step Linearizability of $\seqPQ$}
+\label{sec:obtaining new sequences while ensureing that they are in SeqPQ}
+
+With the help of Lemma \ref{lemma:erase Oc still in EPQ}, Lemma \ref{lemma:erase Oi from some time point still in EPQ}, Lemma \ref{lemma:make put with maximal priority happen earlier}, Lemma \ref{lemma:make Oi happen earlier still in EPQ} and Lemma \ref{lemma:replace equivalent prefix still in EPQ}, we can now prove that $\mathsf{MatchedMaxPriority}$ is step-by-step linearizability.
+Here we use $\textit{call}(o)$ and $\textit{ret}(o)$ as the call and return action of operation $o$, respectively.
 
@@
 \label{lemma:EPQ1 is step-by-step linearizability}
-If a data-differentiated concurrent execution $e$ is linearizable w.r.t.
-$\textit{MS}(\textit{EPQ}_1)$ with witness $x$, then $e \setminus x \sqsubseteq \textit{EPQ} \Rightarrow e \sqsubseteq \textit{EPQ}$.
-\end{restatable}
-
-\begin {proof}
-Let $h$ be the data-differentiated history of $e$, and $l$ be an sequential execution such that $h \sqsubseteq l$ and $l$ matches $\textit{EPQ}_1$ with witness $x$.
-Let the priority of $x$ be $\textit{pri}_x$, and let $h'=h \setminus x$ and assume that $h' \sqsubseteq l' \in \textit{EPQ}$.
+For each data-differentiated execution $e$ where $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Conc}(e,x)$ for some $x$, then $e \setminus x \sqsubseteq \seqPQ \Rightarrow e \sqsubseteq \seqPQ$.
+\end{restatable}
+
+\begin {proof}
+
+By assumption we know that there exists $l$, such that $e \sqsubseteq l$ and $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds.
+Let $p$ be the priority of $x$, then $l=u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$ for some $u$, $v$ and $w$.
+Let $e' = e \setminus x$.
+By assumption there exists sequence $l'$, such that $e' \sqsubseteq l' \in \seqPQ$.
 Let $e_{\textit{lp}}$ be an execution with linearization points of $e$ and the linearization points is added according to $l'$.
-Or we can say, $e_{\textit{lp}}$ is generated from $e$ by instrumenting linearization points, and the projection of $e_{\textit{lp}}$ into method event is $l'$.
-
-According to $\textit{MS}(\textit{EPQ}_1)$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$ and $u$, $v$, $w$, $x$ and $\textit{pri}_x$ satisfy the guard of $\textit{EPQ}_1$.
-Let $l'_v$ be the shortest prefix of $l'$ that contains all method event of $u \cdot v$.
+Or we can say, $e_{\textit{lp}}$ is generated from $e$ by instrumenting linearization points, and the projection of $e_{\textit{lp}}$ into operations is $l'$.
+Let $l'_v$ be the shortest prefix of $l'$ that contains all operation of $u \cdot v$.
 
 Let $U$, $V$ and $W$ be the set of operations of $u$, $v$ and $w$, respectively.
-Let us change $O_i(l,\textit{pri}_x)$ elements in $U$, $V$ and $W$, while keep $O_c(l,\textit{pri}_x)$ elements unchanged.
-We proceed by several loops: In the first loop, we start from the first $O_i(l,\textit{pri}_x)$-element of $l'_v$, and let it be $o_h$,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Case $1$: If in $e_{\textit{lp}}$, the linearization point of $o_h$ is before $\textit{ret}(\textit{rm},x)$, and no $O_c(l,\textit{pri}_x)$-element in $W$ happens before $o_h$.
+Let us change $O_i(l,p)$ elements in $U$, $V$ and $W$, while keep $O_c(l,p)$ elements unchanged.
+We proceed by a loop: In the first round of the loop, we start from the first $O_i(l,p)$-value of $l'_v$, and let it be $o_h$,
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Case $1$: If in $e_{\textit{lp}}$, the linearization point of $o_h$ is before $\textit{ret}(\textit{rm},x)$, and no $O_c(l,p)$-value in $W$ happens before $o_h$.
 Then, $o_h$ is in the new version of $U \cup V$.
 
-\item[-] Case $2$: Else, if in $e_{\textit{lp}}$, the linearization point of $o_h$ is before $\textit{ret}(\textit{rm},x)$, and there exists $O_c(l,\textit{pri}_x)$-element $o_w$ in $W$, such that $o_w <_{\textit{hb}} o_h$.
-Then, in $l'_v$, we put $o_h$ and all $O_i(l,\textit{pri}_x)$-element whose linearization points is after the linearization point of $o_h$ into new version of $W$, and then stop the process of changing $U$, $V$ and $W$.
+\item[-] Case $2$: Else, if in $e_{\textit{lp}}$, the linearization point of $o_h$ is before $\textit{ret}(\textit{rm},x)$, and there exists $O_c(l,p)$-value $o_w$ in $W$, such that $o_w <_{\textit{hb}} o_h$.
+Then, in $l'_v$, we put $o_h$ and all $O_i(l,p)$-value whose linearization points is after the linearization point of $o_h$ into new version of $W$, and then stop the process of changing $U$, $V$ and $W$.
 
 \item[-] Case $3$: Else, if in $e_{\textit{lp}}$, the linearization point of $o_h$ is after $\textit{ret}(\textit{rm},x)$.
-Then, in $l'_v$, we put $o_h$ and all $O_i(l,\textit{pri}_x)$-element whose linearization points is after the linearization point of $o_h$ into new version of $W$, and then stop the process of changing $U$, $V$ and $W$.
-\end{itemize}
-
-In the next loop, we consider the second $O_i(l,\textit{pri}_x)$-element of $l'_v$, and so on.
+Then, in $l'_v$, we put $o_h$ and all $O_i(l,p)$-value whose linearization points is after the linearization point of $o_h$ into new version of $W$, and then stop the process of changing $U$, $V$ and $W$.
+\end{itemize}
+
+In the next round of the loop, we consider the second $O_i(l,p)$-value of $l'_v$, and so on.
 Our process proceed, until either all element in $l'_v$ are in new version of $U \cup V$, or case $2$ or case $3$ happens and this process terminates.
@@
 
-Let $l'_{\textit{u'v'}}$ be the projection of $l'$ into $U' \cup V'$, let $O_x$ be the set of $\textit{put}(\_,\textit{pri}_x)$ while the item is not $x$ in $h$.
+Let $l'_{\textit{u'v'}}$ be the projection of $l'$ into $U' \cup V'$, let $O_x$ be the set of $\textit{put}(\_,p)$ while the item is not $x$ in $h$.
 Let $l''_a$ be the longest prefix of $l'_{\textit{u'v'}}$, where linearization of each operation of $l''_a$ is before $\textit{ret}(\textit{put},b)$ in $e_{\textit{lp}}$.
@@
 Let $l''_3$ be the projection of $l'$ into operations which are not in $l'_{\textit{u'v'}}$.
-Let $l'' = l''_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3$.
-
-To prove $h \sqsubseteq l''$, we define a graph $G$ whose nodes are the operations of $h$ and there is an edge from operation $o_1$ to $o_2$, if one of the following case holds
+Let $l'' = l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3$.
+
+To prove $e \sqsubseteq l''$, we define a graph $G$ whose nodes are the operations of $h$ and there is an edge from operation $o_1$ to $o_2$, if one of the following case holds
 
@@
 
-\item[-] the method event corresponding to $o_1$ in $l''$ is before the one corresponding to $o_2$.
+\item[-] the operation corresponding to $o_1$ in $l''$ is before the one corresponding to $o_2$.
 \end{itemize}
@@
 Assume there is a cycle in $G$.
-According the the property of interval order and the fact that the order of $l''$ is total, we know that there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
+According the the property of interval order and the fact that the order of $l''$ is total, we know that there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding operations are in the opposite order in $l''$.
 Then, we consider all possible case of $o_1$ and $o_2$ as follows: Let $O_a$ and $O_d$ be the set of operations in $l''_a$ and $l''_d$, respectively.
@@
 
-\item[-] If $o_2 \in l''_1 \wedge o_1 = \textit{put}(x,\textit{pri}_x)$:
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $o_2 \in O_a$: This is impossible, since the linearization point of operations in $O_a$ is before $\textit{ret}(\textit{put},b)$ in $e_{\textit{lp}}$.
+\item[-] If $o_2 \in l''_1 \wedge o_1 = \textit{put}(x,p)$:
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $o_2 \in O_a$: This is impossible, since the linearization point of operations in $O_a$ is before $\textit{ret}(\textit{put},x,p)$ in $e_{\textit{lp}}$.
 
@@
 
-    \item[-] If $o_2 \in O_d$: According to $l$, we can see that $\textit{put}(x,\textit{pri}_x)$ does not happen before any operation in $O_x$.
-Then we can see that the linearization point of $o_1$ is before $\textit{ret}(\textit{put},x)$ and $o_1 \in O_a$.
+    \item[-] If $o_2 \in O_d$: According to $l$, we can see that $\textit{put}(x,p)$ does not happen before any operation in $O_x$.
+Then we can see that the linearization point of $o_1$ is before $\textit{ret}(\textit{put},x,p)$ and $o_1 \in O_a$.
 This violates that $o_1 \in l''_2$.
@@
     \setlength{\itemsep}{0.5pt}
-    \item[-] If $o_1 \in O_i(l,\textit{pri}_x)$: Then according to the construction process of $U' \cup V'$ and $W'$, we can see that $o_1 \in O_+$ and then $o_1 \in U' \cup V'$, which contradicts that $o_1 \in l''_3$.
-
-    \item[-] If $o_1 \in O_c(l,\textit{pri}_x)$: Then according to the construction process of $U' \cup V'$ and $W'$, we can see that $o_2 \in O_-$, which contradicts that $o_2 \in O_+$.
+    \item[-] If $o_1 \in O_i(l,p)$: Then according to the construction process of $U' \cup V'$ and $W'$, we can see that $o_1 \in O_+$ and then $o_1 \in U' \cup V'$, which contradicts that $o_1 \in l''_3$.
+
+    \item[-] If $o_1 \in O_c(l,p)$: Then according to the construction process of $U' \cup V'$ and $W'$, we can see that $o_2 \in O_-$, which contradicts that $o_2 \in O_+$.
     \end{itemize}
@@
         \item[-] If the reason of $o_1 \in O_-$ is case $2$: Let $o_h$ be as in case $2$.
-Then there exists $O_c(l,\textit{pri}_x)$-element $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{cal}(o_2)$.
+Then there exists $O_c(l,p)$-value $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{call}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{call}(o_2)$.
 Therefore, we can see that $o_w <_{\textit{hb}} o_2$, and then $l$ contradicts with happen before relation of $h$.
@@
         \item[-] If the reason of $o_1 \in O_-$ is case $3$: Let $o_h$ be as in case $3$.
-Then in $e_{\textit{lp}}$, $\textit{ret}(\textit{rm},x)$ is before the linearization point of $o_h$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{cal}(o_2)$.
+Then in $e_{\textit{lp}}$, $\textit{ret}(\textit{rm},x)$ is before the linearization point of $o_h$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{call}(o_2)$.
 Therefore, we can see that $\textit{rm}(x) <_{\textit{hb}} o_2$, and then $l$ contradicts with happen before relation of $h$.
@@
 
-\item[-] If $o_2 = \textit{put}(x,\textit{pri}_x) \wedge o_1 \in l''_2$: This is impossible, since in $e_{\textit{lp}}$, the linearization points of operations in $l''_2$ is after $\textit{ret}(\textit{put},x)$.
-
-\item[-] If $o_2 = \textit{put}(x,\textit{pri}_x) \wedge o_1 = \textit{rm}(x)$: Then $l$ contradicts with happen before relation of $h$.
-
-\item[-] If $o_2 = \textit{put}(x,\textit{pri}_x) \wedge o_1 \in l''_3$:
+\item[-] If $o_2 = \textit{put}(x,p) \wedge o_1 \in l''_2$: This is impossible, since in $e_{\textit{lp}}$, the linearization points of operations in $l''_2$ is after $\textit{ret}(\textit{put},x,p)$.
+
+\item[-] If $o_2 = \textit{put}(x,p) \wedge o_1 = \textit{rm}(x)$: Then $l$ contradicts with happen before relation of $h$.
+
+\item[-] If $o_2 = \textit{put}(x,p) \wedge o_1 \in l''_3$:
     \begin{itemize}
@@
          \item[-] If the reason of $o_1 \in O_-$ is case $2$: Let $o_h$ be as in case $2$.
-Then there exists $O_c(l,\textit{pri}_x)$-element $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{cal}(\textit{put},x,\textit{pri}_x)$.
-Therefore, we can see that $o_w <_{\textit{hb}} \textit{put}(x,\textit{pri}_x)$, and then $l$ contradicts with happen before relation of $h$.
+Then there exists $O_c(l,p)$-value $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{call}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{call}(\textit{put},x,p)$.
+Therefore, we can see that $o_w <_{\textit{hb}} \textit{put}(x,p)$, and then $l$ contradicts with happen before relation of $h$.
 
          \item[-] If the reason of $o_1 \in O_-$ is case $3$: Let $o_h$ be as in case $3$.
-Then in $e_{\textit{lp}}$, $\textit{ret}(\textit{rm},x)$ is before the linearization point of $o_h$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{cal}(\textit{put},x,\textit{pri}_x)$.
-Therefore, we can see that $\textit{rm}(x) <_{\textit{hb}} \textit{put}(x,\textit{pri}_x)$, and then $l$ contradicts with happen before relation of $h$.
+Then in $e_{\textit{lp}}$, $\textit{ret}(\textit{rm},x)$ is before the linearization point of $o_h$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{call}(\textit{put},x,p)$.
+Therefore, we can see that $\textit{rm}(x) <_{\textit{hb}} \textit{put}(x,p)$, and then $l$ contradicts with happen before relation of $h$.
          \end{itemize}
@@
          \item[-] If the reason of $o_1 \in O_-$ is case $2$: Let $o_h$ be as in case $2$.
-Then there exists $O_c(l,\textit{pri}_x)$-element $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$.
-
-             Since $l$ is consistent with the happen before order of $h$, we can see that $\textit{cal}(\textit{rm},x)$ is before $\textit{ret}(o_w)$.
-Therefore, we can see that the linearization point of $o_1$ is after $\textit{cal}(\textit{rm},x)$, and then it is impossible that $o_1 <_{\textit{hb}} \textit{rm}(x)$.
+Then there exists $O_c(l,p)$-value $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{call}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$.
+
+             Since $l$ is consistent with the happen before order of $h$, we can see that $\textit{call}(\textit{rm},x)$ is before $\textit{ret}(o_w)$.
+Therefore, we can see that the linearization point of $o_1$ is after $\textit{call}(\textit{rm},x)$, and then it is impossible that $o_1 <_{\textit{hb}} \textit{rm}(x)$.
 
@@
 
-It remains to prove that $l'' \in \textit{EPQ}$. The process for proving $l'' \in \textit{EPQ}$ is as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Since $l' \in \textit{EPQ}$ and $l'_v$ is a prefix of $l'$, it is obvious that $l'_v \in \textit{EPQ}$.
+It remains to prove that $l'' \in \seqPQ$. The process for proving $l'' \in \seqPQ$ is as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Since $l' \in \seqPQ$ and $l'_v$ is a prefix of $l'$, it is obvious that $l'_v \in \seqPQ$.
 
@@
     \setlength{\itemsep}{0.5pt}
-    \item[-] Discard $O_c(l,\textit{pri}_x)$-element that are in $W$ and keep $O_c(l,\textit{pri}_x)$-element in $U \cup V$ unchanged.
-
-    \item[-] From some time point, discard all the $O_i(l,\textit{pri}_x)$ operations after this time point.
-    \end{itemize}
-
-From $\textit{matched-C}(u \cdot v,\textit{pri}_x)$, we can see that $O_c(l,\textit{pri}_x)$-element in $U \cdot V$ is matched, and then it is easy to see that the projection of $l'_v$ in to $O_c(l,\textit{pri}_x)$-element in $U \cdot V$ is still in $\textit{EPQ}$.
-By Lemma \ref{lemma:erase Oc still in EPQ} and Lemma \ref{lemma:erase Oi from some time point still in EPQ}, we can see that $l'_{\textit{u'v'}} \in \textit{EPQ}$.
+    \item[-] Discard $O_c(l,p)$-value that are in $W$ and keep $O_c(l,p)$-value in $U \cup V$ unchanged.
+
+    \item[-] From some time point, discard all the $O_i(l,p)$ operations after this time point.
+    \end{itemize}
+
+From $\textit{matched}_{\prec}(u \cdot v,p)$, we can see that $O_c(l,p)$-value in $U \cdot V$ is matched, and then it is easy to see that the projection of $l'_v$ in to $O_c(l,p)$-value in $U \cdot V$ is still in $\seqPQ$.
+By Lemma \ref{lemma:erase Oc still in EPQ} and Lemma \ref{lemma:erase Oi from some time point still in EPQ}, we can see that $l'_{\textit{u'v'}} \in \seqPQ$.
 
 \item[-] $l''_1 \cdot l''_2$ can be obtained from $l'_{\textit{u'v'}}$ as follows: Execute until reaching some time point $t$, then first execute all $O_x$ operations after $t$, and then execute remanning operations.
-By Lemma \ref{lemma:make put with maximal priority happen earlier}, we can see that $l''_1 \cdot l''_2 \in \textit{EPQ}$.
-
-\item[-] Let $l'_e$ be obtained from $l'$ by discarding $O_c(l,\textit{pri}_x)$-element in $U \cdot V$. By Lemma \ref{lemma:EPQ is closed under projection}, we can see that $l'_e \in \textit{EPQ}$.
-
-\item[-] Let $l'_e \vert_{O_i(l,\textit{pri}_x)}  = l'_f \cdot l'_g$, where $l'_f$ is the projection of $O_i(l,\textit{pri}_x)$-element in $U \cdot V$.
+By Lemma \ref{lemma:make put with maximal priority happen earlier}, we can see that $l''_1 \cdot l''_2 \in \seqPQ$.
+
+\item[-] Let $l'_e$ be obtained from $l'$ by discarding $O_c(l,p)$-value in $U \cdot V$. By Lemma \ref{lem:closure_proj}, we can see that $l'_e \in \seqPQ$.
+
+\item[-] Let $l'_e \vert_{O_i(l,p)}  = l'_f \cdot l'_g$, where $l'_f$ is the projection of $O_i(l,p)$-value in $U \cdot V$.
 Let $l'_h$ be obtained from $l'_e$ by discarding operations in $l'_f$.
-By Lemma \ref{lemma:make Oi happen earlier still in EPQ}, we can see that $l'_f \cdot l'_h \in \textit{EPQ}$.
-
-\item[-] By Lemma \ref{lemma:EPQ is closed under projection}, it is obvious that $l'_f \in \textit{EPQ}$.
-From $\textit{matched-C}(u \cdot v,\textit{pri}_x)$, we can see that the content of priority queue after executing $l''_1 \cdot l''_2$ is the same as after executing $l'_f$.
-By Lemma \ref{lemma:replace equivalent prefix still in EPQ}, we can see that $l''_1 \cdot l''_2 \cdot l'_h \in \textit{EPQ}$.
-It is easy to see that $l'_h = l''_3$, and then $l''_1 \cdot l''_2 \cdot l''_3 \in \textit{EPQ}$.
-
-\item[-] Since $u$, $v$, $w$, $x$ and $\textit{pri}_x$ satisfy the guard of $\textit{EPQ}_1$, it is easy to see that $l'' = l''_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3 \in \textit{EPQ}$.
-\end{itemize}
-
-Therefore, we prove that $h \sqsubseteq l'' \in \textit{EPQ}$. This completes the proof of this lemma. \qed
+By Lemma \ref{lemma:make Oi happen earlier still in EPQ}, we can see that $l'_f \cdot l'_h \in \seqPQ$.
+
+\item[-] By Lemma \ref{lem:closure_proj}, it is obvious that $l'_f \in \seqPQ$.
+From $\textit{matched}_{\prec}(u \cdot v,p)$, we can see that the content of priority queue after executing $l''_1 \cdot l''_2$ is the same as after executing $l'_f$.
+By Lemma \ref{lemma:replace equivalent prefix still in EPQ}, we can see that $l''_1 \cdot l''_2 \cdot l'_h \in \seqPQ$.
+It is easy to see that $l'_h = l''_3$, and then $l''_1 \cdot l''_2 \cdot l''_3 \in \seqPQ$.
+
+\item[-] Since $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds, it is easy to see that $l'' = l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3 \in \seqPQ$.
+\end{itemize}
+
+Therefore, we prove that $e \sqsubseteq l'' \in \seqPQ$. This completes the proof of this lemma. \qed
 \end {proof}
@@
 \label{lemma:EPQ2 is step-by-step linearizability}
-If a data-differentiated concurrent execution $e$ is linearizable w.r.t.
-$\textit{MS}(\textit{EPQ}_2)$ with witness $x$, then $e \setminus x \sqsubseteq \textit{EPQ} \Rightarrow e \sqsubseteq \textit{EPQ}$.
-\end{restatable}
-
-\begin {proof}
-
-Let $h$ be the data-differentiated history of $e$, and $l$ be an sequential execution such that $h \sqsubseteq l$ and $l$ matches $\textit{EPQ}_2$ with witness $x$.
-Let the priority of $x$ be $\textit{pri}_x$, and let $h'=h \setminus x$ and assume that $h' \sqsubseteq l' \in \textit{EPQ}$.
+For each data-differentiated execution $e$ where $\mathsf{UnmatchedMaxPriority}\mathsf{\text{-}Conc}(e,x)$ for some $x$, then $e \setminus x \sqsubseteq \seqPQ \Rightarrow e \sqsubseteq \seqPQ$.
+\end{restatable}
+
+\begin {proof}
+
+By assumption we know that there exists $l$, such that $e \sqsubseteq l$ and $\mathsf{UnmatchedMaxPriority\text{-}Seq}(l,x)$ holds.
+Let $p$ be the priority of $x$, then $l=u \cdot \textit{put}(x,p) \cdot v$ for some $u$ and $v$.
+Let $e' = e \setminus x$.
+By assumption there exists sequence $l'$, such that $e' \sqsubseteq l' \in \seqPQ$.
 Let $e_{\textit{lp}}$ be an execution with linearization points of $e$ and the linearization points is added according to $l'$.
-Or we can say, $e_{\textit{lp}}$ is generated from $e$ by instrumenting linearization points, and the projection of $e_{\textit{lp}}$ into method event is $l'$.
-
-According to $\textit{MS}(\textit{EPQ}_2)$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$ and $u$, $v$, $x$ and $\textit{pri}_x$ satisfy the guard of $\textit{EPQ}_2$.
-
-Let $l''_a$ be the longest prefix of $l'$ such that linearization point of each operation of $l''_a$ is before $\textit{ret}(\textit{put},x)$ in $e_{\textit{lp}}$.
-Let $O_x$ be the set of $\textit{put}(\_,\textit{pri}_x)$ while the item is not $x$ in $h$.
+Or we can say, $e_{\textit{lp}}$ is generated from $e$ by instrumenting linearization points, and the projection of $e_{\textit{lp}}$ into operations is $l'$.
+Let $l'_v$ be the shortest prefix of $l'$ that contains all operations of $u \cdot v$.
+
+Let $l''_a$ be the longest prefix of $l'$ such that linearization point of each operation of $l''_a$ is before $\textit{ret}(\textit{put},x,p)$ in $e_{\textit{lp}}$.
+Let $O_x$ be the set of $\textit{put}(\_,p)$ while the value is not $x$ in $h$.
 Let $l''_s$ be the projection of $l'$ into operations of $O_x$ that are not in $l''_a$.
@@
 Let $l''_2$ be the projection of $l'$ into operations of $l'$ that are not in $l''_1$.
-Let $l'' = l''_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l''_2$.
+Let $l'' = l''_1 \cdot \textit{put}(x,p) \cdot l''_2$.
 
 To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:EPQ1 is step-by-step linearizability}.
-Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
+Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding operations are in the opposite order in $l''$.
 Then, we consider all possible case of $o_1$ and $o_2$ as follows: Let $O_a$ and $O_s$ be the set of operations in $l''_a$ and $l''_s$, respectively.
@@
 
-    \item[-] If $o_2 \in O_a \wedge o_1 \in O_s$: It is not hard to see that $\textit{put}(x,\textit{pri}_x) <_{\textit{hb}} o_2$.
-Then, it is impossible to locate the linearization point of $o_2$ before $\textit{ret}(\textit{put},x)$ in $e_{\textit{lp}}$.
-    \end{itemize}
-
-\item[-] If $o_2 \in l''_1 \wedge o_1 = \textit{put}(x,\textit{pri}_x)$:
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $o_2 \in O_a$: This is impossible, since in $e_{\textit{lp}}$, the linearization point of $o_1$ is before $\textit{ret}(\textit{put},x)$.
-
-    \item[-] If $o_2 \in O_s$: This is impossible, since $o_2 = \textit{put}(\_,\textit{pri}_x)$, and $l$ is consistent with happen before relation of $h$.
+    \item[-] If $o_2 \in O_a \wedge o_1 \in O_s$: It is not hard to see that $\textit{put}(x,p) <_{\textit{hb}} o_2$.
+Then, it is impossible to locate the linearization point of $o_2$ before $\textit{ret}(\textit{put},x,p)$ in $e_{\textit{lp}}$.
+    \end{itemize}
+
+\item[-] If $o_2 \in l''_1 \wedge o_1 = \textit{put}(x,p)$:
+    \begin{itemize}
+    \setlength{\itemsep}{0.5pt}
+    \item[-] If $o_2 \in O_a$: This is impossible, since in $e_{\textit{lp}}$, the linearization point of $o_1$ is before $\textit{ret}(\textit{put},x,p)$.
+
+    \item[-] If $o_2 \in O_s$: This is impossible, since $o_2 = \textit{put}(\_,p)$, and $l$ is consistent with happen before relation of $h$.
     \end{itemize}
@@
 
-    \item[-] If $o_2 \in O_s$: Since no $\textit{put}(\_,\textit{pri}_x)$ happens before $\textit{put}(x,\textit{pri}_x)$ in $h$, $\textit{cal}(o_2)$ is before $\textit{ret}(\textit{put},x)$.
-Since $o_1 <_{\textit{hb}} o_2$, we can see that $\textit{ret}(o_1)$ is before $\textit{cal}(o_2)$, and then $\textit{ret}(o_1)$ is before $\textit{ret}(\textit{put},x)$.
-Then the linearization point of $o_1$ can only be before $\textit{ret}(\textit{put},x)$, and $o_1 \in l''_a$, which contradicts that $o_1 \in l''_2$.
-    \end{itemize}
-
-\item[-] If $o_2 = \textit{put}(x,\textit{pri}_x) \wedge o_1 \in l''_2$: Then since the linearization point of $o_1$ can only be before $\textit{ret}(\textit{put},x)$, we can see that $o_1 \in l''_a$, which contradicts that $o_1 \in l''_2$.
+    \item[-] If $o_2 \in O_s$: Since no $\textit{put}(\_,p)$ happens before $\textit{put}(x,p)$ in $h$, $\textit{call}(o_2)$ is before $\textit{ret}(\textit{put},x,p)$.
+Since $o_1 <_{\textit{hb}} o_2$, we can see that $\textit{ret}(o_1)$ is before $\textit{call}(o_2)$, and then $\textit{ret}(o_1)$ is before $\textit{ret}(\textit{put},x,p)$.
+Then the linearization point of $o_1$ can only be before $\textit{ret}(\textit{put},x,p)$, and $o_1 \in l''_a$, which contradicts that $o_1 \in l''_2$.
+    \end{itemize}
+
+\item[-] If $o_2 = \textit{put}(x,p) \wedge o_1 \in l''_2$: Then since the linearization point of $o_1$ can only be before $\textit{ret}(\textit{put},x,p)$, we can see that $o_1 \in l''_a$, which contradicts that $o_1 \in l''_2$.
 
@@
 
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq l''$.
-
-It remains to prove that $l'' \in \textit{EPQ}$.
+Therefore, we know that $G$ is acyclic, and then we know that $e \sqsubseteq l''$.
+
+It remains to prove that $l'' \in \seqPQ$.
 $l''_a \cdot l''_s \cdot l''_2$ can be obtained from $l'$ as follows: Execute until reaching some time point $t$, then first execute all $O_x$ operations after $t$, and then execute remanning operations.
-By Lemma \ref{lemma:make put with maximal priority happen earlier}, we can see that $l''_1 \cdot l''_2 = l''_a \cdot l''_s \cdot l''_2 \in \textit{EPQ}$.
-Since $u$, $v$, $x$ and $\textit{pri}_x$ satisfy the guard of $\textit{EPQ}_2$, it is easy to see that $l'' = l''_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l''_2 \in \textit{EPQ}$.
-
-Therefore, we prove that $h \sqsubseteq l'' \in \textit{EPQ}$. This completes the proof of this lemma.\qed
+By Lemma \ref{lemma:make put with maximal priority happen earlier}, we can see that $l''_1 \cdot l''_2 = l''_a \cdot l''_s \cdot l''_2 \in \seqPQ$.
+Since $\mathsf{UnmatchedMaxPriority\text{-}Seq}(l,x)$ holds, it is easy to see that $l'' = l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \in \seqPQ$.
+
+Therefore, we prove that $e \sqsubseteq l'' \in \seqPQ$. This completes the proof of this lemma.\qed
 \end {proof}
@@
 \label{lemma:EPQ3 is step-by-step linearizability}
-If a data-differentiated concurrent execution $e$ is linearizable w.r.t.
-$\textit{MS}(\textit{EPQ}_3)$ and $o$ is a $\textit{rm}(\textit{empty})$ event, then $e \setminus o \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
-\end{restatable}
-
-\begin {proof}
-Let $h$ be the data-differentiated history of $e$, $l$ be an sequential execution such that $h \sqsubseteq l$, $l$ matches $\textit{EPQ}_3$ and $o$ is a $\textit{rm}$ method event in $h$.
-Let $h'=h \setminus o$ and assume that $h' \sqsubseteq l' \in \textit{EPQ}$.
-
-According to $\textit{EPQ}_3$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{rm}(\textit{empty}) \cdot v$, where all the $\textit{put}$ operations and $\textit{rm}$ in $u$ are matched.
-
-Let $E_L$ be the set of method events in $u$ and $E_R$ be the set of method events in $v$.
+For each data-differentiated execution $e$ where $\mathsf{EmptyRemove}\mathsf{\text{-}Conc}(e,x)$ for some $o=\textit{rm}(\textit{empty})$, then $e \setminus o \sqsubseteq \seqPQ \Rightarrow e \sqsubseteq \seqPQ$.
+\end{restatable}
+
+\begin {proof}
+
+By assumption we know that there exists $l$, such that $e \sqsubseteq l$ and $\mathsf{EmptyRemove\text{-}Seq}(l,o)$ holds.
+Then $l=u \cdot o \cdot v$ for some $u$ and $v$.
+Let $e' = e \setminus o$.
+By assumption there exists sequence $l'$, such that $e' \sqsubseteq l' \in \seqPQ$.
+
+Let $E_L$ be the set of operations in $u$ and $E_R$ be the set of operations in $v$.
 Let $l'_L = l' \vert_{E_L}$ and $l'_R = l' \vert_{E_R}$.
 Let sequence $l'' = l'_L \cdot o \cdot L'_R$.
-Since priority queue is closed under projection (Lemma \ref{lemma:EPQ is closed under projection}) and all the $\textit{put}$ operations and $\textit{rm}$ in $u$ are matched, we know that $l'_L \in \textit{EPQ}$ and the the priority queue is empty after executing $l'_L$.
-Then we know that $l'_L \cdot \textit{rm}(\textit{empty}) \in \textit{EPQ}$.
-Since $l'_R$ is obtained from $l'$ by discarding pairs of matched $\textit{put}$ and $\textit{rm}$ operations, it is easy to see that $L'_R \in \textit{EPQ}$, and then we know that $l'' = l'_L \cdot o \cdot L'_R \in \textit{EPQ}$.
+Since priority queue is closed under projection (Lemma \ref{lem:closure_proj}) and all the $\textit{put}$ operations and $\textit{rm}$ in $u$ are matched, we know that $l'_L \in \seqPQ$ and the the priority queue is empty after executing $l'_L$.
+Then we know that $l'_L \cdot \textit{rm}(\textit{empty}) \in \seqPQ$.
+Since $l'_R$ is obtained from $l'$ by discarding pairs of matched $\textit{put}$ and $\textit{rm}$ operations, it is easy to see that $L'_R \in \seqPQ$, and then we know that $l'' = l'_L \cdot o \cdot L'_R \in \seqPQ$.
 
@@
 To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:EPQ1 is step-by-step linearizability}.
-Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$.
+Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding operations are in the opposite order in $l''$.
 Then, we consider all possible case of $o_1$ and $o_2$ as follows:
@@
 
-Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \textit{EPQ}$. \qed
-\end {proof}
-
-The following lemma states that $\textit{EPQ}$ is step-by-step linearizability, it is a direct consequence of Lemma \ref{lemma:EPQ1 is step-by-step linearizability}, Lemma \ref{lemma:EPQ2 is step-by-step linearizability} and Lemma \ref{lemma:EPQ3 is step-by-step linearizability}.
+Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \seqPQ$. \qed
+\end {proof}
+
+The following lemma states that $\seqPQ$ is step-by-step linearizability, which is a direct consequence of Lemma \ref{lemma:EPQ1 is step-by-step linearizability}, Lemma \ref{lemma:EPQ2 is step-by-step linearizability} and Lemma \ref{lemma:EPQ3 is step-by-step linearizability}.
 
@@
 \label{lemma:EPQ is step-by-step linearizability}
-$\textit{PQ}$ is step-by-step linearizability.
+$\seqPQ$ is step-by-step linearizability.
 \end{restatable}
@@
 
+\subsection{Proof of Lemma \ref{lemma:con-check-EPQ is correct}}
+\label{sec:appendix subsection proof of lemma con-check-EPQ is correct}
+
+\ConCheckEPQIsCorrect*
+
+\begin {proof}
+
+$\textit{Check-PQ-Conc}(e)=\mathsf{true}$ iff $e \sqsubseteq \seqPQ$, for every data-differentiated execution $e$.
+
+To prove the $\textit{if}$ direction, given a data-differentiated $e \sqsubseteq l \in \seqPQ$. Then we being a loop as follows,
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If $\mathsf{Has\text{-}EmptyRemoves}(e)$, then $l = u \cdot o \cdot v$ for some $o=\textit{rm}(\textit{empty})$.
+It is not hard to see that $\mathsf{EmptyRemove\text{-}Seq}(l,o)$.
+Let $e'$ be the projection of $e$ into operations of $u \cdot v$.
+It is easy to see that $e' \sqsubseteq u \cdot v$ and then $e' \sqsubseteq \seqPQ$.
+Then we start the next round and choose $e'$ to be the ``$e$ in the next round''.
+
+\item[-] \item[-] If $\mathsf{Has\text{-}UnmatchedMaxPriority}(e)$, then there exists a maximal priority $p$ in $e$, such that $p$ has unmatched $\textit{put}$.
+Let $l = u \cdot \textit{put}(x,p) \cdot v$, where in $v$ there is no $\textit{put}$ with priority $p$.
+It is not hard to see that $\mathsf{UnmatchedMaxPriority\text{-}Seq}(l,x)$.
+Let $e'$ be the projection of $e$ into operations of $u \cdot v$.
+It is easy to see that $e' \sqsubseteq u \cdot v$ and then $e' \sqsubseteq \seqPQ$.
+Then we start the next round and choose $e'$ to be the ``$e$ in the next round''.
+
+\end{itemize}
+
+where $\mathsf{Has\text{-}\Gamma}(e)$ holds and .
+
+The proof of the $\textit{if}$ direction is quite similar to the proof of $\seqPQ_{\neq} \subseteq \textit{SeqPQ}_f$ direction of Lemma \ref{lemma:EPQ rules and semantics} and is omitted here.
+
+To prove the $\textit{only if}$ direction, given a data-differentiated execution $e$ and assume that $\textit{Check-PQ-Conc}(e)=\mathsf{true}$.
+
+$e \sqsubseteq \seqPQ$, according to the definition of $\textit{PQ}$, we can always find
+
+\qed
+\end {proof}
+
 \subsection{Proof of Lemma \ref{lemma:EPQ as multi in MRpri for history}}
@@
 
-To prove the $\textit{only if}$ direction, assume that $e \sqsubseteq l \in \textit{EPQ}$.
-Given $e' = e \vert_{D}$ and $l' = l \vert_{D}$, it is easy to see that $e' \sqsubseteq l'$, and by Lemma \ref{lemma:EPQ is closed under projection}, we can see that $l' \in \textit{EPQ}$.
+To prove the $\textit{only if}$ direction, assume that $e \sqsubseteq l \in \seqPQ$.
+Given $e' = e \vert_{D}$ and $l' = l \vert_{D}$, it is easy to see that $e' \sqsubseteq l'$, and by Lemma \ref{lemma:EPQ is closed under projection}, we can see that $l' \in \seqPQ$.
 Then by Lemma \ref{lemma:EPQ as multi in MRpri for sequence} we know that for each $R \in \textit{last}(l')$, we have $l' \in \textit{MS}(R)$.
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] If $R_1 = \textit{EPQ}_3$: $e_2$ is generated from $e_1$ by erasing call and return of one $\textit{rm}(\textit{empty})$ operation.
-
-\item[-] Else, if $R_1 = \textit{EPQ}_2^{=}, \textit{EPQ}_2^{>}, \textit{EPQ}_1^{=}, \textit{EPQ}_1^{>}$, and the witness is $x$: $e_2$ is generated from $e_1$ by erasing call and return of method event of item $x$.
-\end{itemize}
-
-Similarly, for each $i > 1$, we obtain $e_{\textit{i+1}}$ from $e_i$, until we obtain $e_m = \epsilon$ for some $m$. It is obvious that $e_m \sqsubseteq \textit{EPQ}$. For $e_{\textit{m-1}}$, since
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $e_m \sqsubseteq \textit{EPQ}$,
-
-\item[-] If $\textit{last}(e_{\textit{m-1}}) = R_{\textit{m-1}} \in \{ \textit{EPQ}_1^{>}, \textit{EPQ}_1^{=}, \textit{EPQ}_2^{>}, \textit{EPQ}_2^{=} \}$ and $e_{\textit{m-1}}$ matches $R_{\textit{m-1}}$ with witness $x$: We already know that $e_{\textit{m-1}} \sqsubseteq \textit{MS}(R_{\textit{m-1}})$, $e_m = e_{\textit{m-1}} \setminus x \sqsubseteq \textit{EPQ}$, and by step-by-step linearizability of $\textit{EPQ}$ (Lemma \ref{lemma:EPQ is step-by-step linearizability}), we can see that $e_{\textit{m-1}} \sqsubseteq \textit{EPQ}$.
-
-\item[-] If $\textit{last}(e_{\textit{m-1}}) = R_{\textit{m-1}} = \textit{EPQ}_3$ and $o$ is a $\textit{rm}(\textit{empty})$ in $e_{\textit{m-1}}$ that is removed in the process of constructing $e_m$: We already know that $e_{\textit{m-1}} \sqsubseteq \textit{MS}(R_{\textit{m-1}})$, $e_m = e_{\textit{m-1}} \setminus o \sqsubseteq \textit{EPQ}$, and by step-by-step linearizability of $\textit{EPQ}$ (Lemma \ref{lemma:EPQ is step-by-step linearizability}), we can see that $e_{\textit{m-1}} \sqsubseteq \textit{EPQ}$.
-\end{itemize}
-
-Therefore, we know that $e_{\textit{m-1}} \in \textit{EPQ}$. Similarly, we can prove that $e_{\textit{m-2}},\ldots,e_1 = e \in \textit{EPQ}$. \qed
+\item[-] If $R_1 = \seqPQ_3$: $e_2$ is generated from $e_1$ by erasing call and return of one $\textit{rm}(\textit{empty})$ operation.
+
+\item[-] Else, if $R_1 = \seqPQ_2^{=}, \seqPQ_2^{>}, \seqPQ_1^{=}, \seqPQ_1^{>}$, and the witness is $x$: $e_2$ is generated from $e_1$ by erasing call and return of method event of item $x$.
+\end{itemize}
+
+Similarly, for each $i > 1$, we obtain $e_{\textit{i+1}}$ from $e_i$, until we obtain $e_m = \epsilon$ for some $m$. It is obvious that $e_m \sqsubseteq \seqPQ$. For $e_{\textit{m-1}}$, since
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $e_m \sqsubseteq \seqPQ$,
+
+\item[-] If $\textit{last}(e_{\textit{m-1}}) = R_{\textit{m-1}} \in \{ \seqPQ_1^{>}, \seqPQ_1^{=}, \seqPQ_2^{>}, \seqPQ_2^{=} \}$ and $e_{\textit{m-1}}$ matches $R_{\textit{m-1}}$ with witness $x$: We already know that $e_{\textit{m-1}} \sqsubseteq \textit{MS}(R_{\textit{m-1}})$, $e_m = e_{\textit{m-1}} \setminus x \sqsubseteq \seqPQ$, and by step-by-step linearizability of $\seqPQ$ (Lemma \ref{lemma:EPQ is step-by-step linearizability}), we can see that $e_{\textit{m-1}} \sqsubseteq \seqPQ$.
+
+\item[-] If $\textit{last}(e_{\textit{m-1}}) = R_{\textit{m-1}} = \seqPQ_3$ and $o$ is a $\textit{rm}(\textit{empty})$ in $e_{\textit{m-1}}$ that is removed in the process of constructing $e_m$: We already know that $e_{\textit{m-1}} \sqsubseteq \textit{MS}(R_{\textit{m-1}})$, $e_m = e_{\textit{m-1}} \setminus o \sqsubseteq \seqPQ$, and by step-by-step linearizability of $\seqPQ$ (Lemma \ref{lemma:EPQ is step-by-step linearizability}), we can see that $e_{\textit{m-1}} \sqsubseteq \seqPQ$.
+\end{itemize}
+
+Therefore, we know that $e_{\textit{m-1}} \in \seqPQ$. Similarly, we can prove that $e_{\textit{m-2}},\ldots,e_1 = e \in \seqPQ$. \qed
+\end {proof}
+
+\subsection{Proof of Lemma \ref{lemma:EPQ as multi in MRpri for sequence}}
+
+%\EPQasMultiInMRpriforSequence*
+
+\begin{restatable}{lemma}{EPQasMultiInMRpriforSequence}
+\label{lemma:EPQ as multi in MRpri for sequence}
+Given a data-differentiated sequential execution $e$, $e \in \seqPQ$, if and only if, $\forall e' \in \textit{proj}(e)$ and $\forall R \in \textit{last}(e')$, we have $e' \in \textit{MS}(R)$.
+\end{restatable}
+
+\begin {proof}
+
+The \textit{only if} direction is obvious and can be similarly proved as the $\seqPQ_s \subseteq \seqPQ$ direction of Lemma \ref{lemma:EPQ rules and semantics}.
+
+To prove the $\textit{if}$ direction, we proceed as follows: From $e1=e$, we generate a sequence $e_2$ as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If $\textit{PQ}_3 \in \textit{last}(e_1)$: Then we can see that $e_1$ contains at least one $\textit{rm}(\textit{empty})$.
+$e_2$ is generated from $e_1$ by erasing one $\textit{rm}(\textit{empty})$.
+
+\item[-] Else, if $\textit{PQ}_2^{=} \in \textit{last}(e_1)$: Then we can see that one of the maximal priority of $e_1$ is unmatched $\textit{put}$ and (possibly) matched $\textit{put}$.
+Assume the set of the items of these unmatched $\textit{put}$ is $S$.
+$e_2$ is generated from $e_1$ by erasing one unmatched $\textit{put}$ which use the item last putted in $S$.
+
+\item[-] Else, if $\textit{PQ}_2^{>} \in \textit{last}(e_1)$: Then we can see that one of the maximal priority in $e_1$ is unmatched $\textit{put}$.
+Assume the set of the items of these unmatched $\textit{put}$ is $S$.
+$e_2$ is generated from $e_1$ by erasing one unmatched $\textit{put}$ which use the item last putted in $S$.
+
+\item[-] Else, if $\textit{PQ}_1^{=} \in \textit{last}(e_1)$: Then we can see that one of the maximal priority in $e_1$ is of more than one pair of matched $\textit{put}$.
+Assume the set of the items of these matched $\textit{put}$ is $S$.
+$e_2$ is generated from $e_1$ by erasing matched $\textit{put}$ and $\textit{rm}$ of the item which is last putted in $S$.
+
+\item[-] Else, if $\textit{PQ}_1^{>} \in \textit{last}(e_1)$: Then we can see that one of the maximal priority in $e_1$ is of one pair of matched $\textit{put}$.
+$e_2$ is generated from $e_1$ by erasing this pair of matched $\textit{put}$ and $\textit{rm}$.
+\end{itemize}
+
+Similarly, for each $i > 1$, we obtain $e_{\textit{i+1}}$ from $e_i$, until we obtain $e_m = \epsilon$ for some $m$. It is obvious that $e_m \in \seqPQ$. For $e_{\textit{m-1}}$, since
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $e_m \in \seqPQ$,
+
+\item[-] By assumption, we know that $e_{\textit{m-1}} \in \textit{MS}(R_{\textit{m-1}})$, where $R_{\textit{m-1}} \in \textit{last}(e_{\textit{m-1}})$.
+This implies that the guard of $R_{\textit{m-1}}$ is satisfied.
+\end{itemize}
+
+Therefore, we know that $e_{\textit{m-1}} \in \seqPQ$. Similarly, we can prove that $e_{\textit{m-2}},\ldots,e_1 = e \in \seqPQ$. \qed
 \end {proof}
@@
 
-To facilitate our proof of latter sections, we need to separate $\textit{EPQ}_1$ into two cases: (1) no matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v \cdot w$ has priority $\textit{pri}$, (2) some matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v \cdot w$ has priority $\textit{pri}$.
-Therefore, we separate $\textit{EPQ}_1$ into two cases:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{EPQ}_1^{>}(u,v,w,\textit{pri}) \equiv
+To facilitate our proof of latter sections, we need to separate $\seqPQ_1$ into two cases: (1) no matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v \cdot w$ has priority $\textit{pri}$, (2) some matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v \cdot w$ has priority $\textit{pri}$.
+Therefore, we separate $\seqPQ_1$ into two cases:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $\seqPQ_1^{>}(u,v,w,\textit{pri}) \equiv
 (\textit{noRE}(u \cdot v \cdot w)) \wedge
@@
 
-\item[-] $\textit{EPQ}_1^{=}(u,v,w,\textit{pri}) \equiv
+\item[-] $\seqPQ_1^{=}(u,v,w,\textit{pri}) \equiv
 (\textit{noRE}(u \cdot v \cdot w)) \wedge
@@
 
-For $\textit{EPQ}_2$, we also need to distinguish two cases: (1) no matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ has priority $\textit{pri}$, (2) some matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ has priority $\textit{pri}$.
-Therefore, we separate $\textit{EPQ}_2$ into two rules:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{EPQ}_2^{>}(u,v) \equiv
+For $\seqPQ_2$, we also need to distinguish two cases: (1) no matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ has priority $\textit{pri}$, (2) some matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ has priority $\textit{pri}$.
+Therefore, we separate $\seqPQ_2$ into two rules:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $\seqPQ_2^{>}(u,v) \equiv
 (\textit{noRE}(u \cdot v)) \wedge
@@
 
-\item[-] $\textit{EPQ}_2^{=}(u,v) \equiv
+\item[-] $\seqPQ_2^{=}(u,v) \equiv
 (\textit{noRE}(u \cdot v)) \wedge
@@
 Let $\textit{transToQueue}(e)$ be an execution generated from $e$ by transforming $\textit{put}$ and $\textit{rm}$ into $\textit{enq}$ and $\textit{deq}$, respectively, and then discarding priorities.
-We can see that for each $e \in \textit{EPQ}$ and each priority $i$, $\textit{transToQueue}(e \vert_{i})$ satisfy FIFO (first in first out) property.
+We can see that for each $e \in \seqPQ$ and each priority $i$, $\textit{transToQueue}(e \vert_{i})$ satisfy FIFO (first in first out) property.
 
@@
 To state the effectiveness of left-right constraint, take the execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} as an example, we can see that $b$ is covered by $f,e,d$.
-We need to prove that getting rid of cycle though item with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\textit{MS}(\textit{EPQ}_1^{>})$, as stated by the following lemma:
+We need to prove that getting rid of cycle though item with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\textit{MS}(\seqPQ_1^{>})$, as stated by the following lemma:
 
@@
 \label{lemma:Lin Equals Constraint for EPQ1Lar}
-Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{>}$.
+Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{>}$.
 Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority.
 Let $G$ be the graph representing the left-right constraint of $\textit{put}(x)$ and $\textit{rm}(x)$.
-$e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$, if and only if $G$ has no cycle going through $x$.
+$e \sqsubseteq \textit{MS}(\seqPQ_1^{>})$, if and only if $G$ has no cycle going through $x$.
 \end{restatable}
@@
 
-We deal with the case of $R = \textit{EPQ}_1^{>}$ , and other cases can be similarly dealt with.
-
-To prove the $\textit{only if}$ direction, given $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$ and such $\textit{pri}$ and $x$.
-Since $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$ with witness $x$, we know that $e \sqsubseteq u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$, where $u$, $v$, $w$, $x$ and $\textit{pri}$ satisfy the guard of $\textit{EPQ}_1^{>}$.
+We deal with the case of $R = \seqPQ_1^{>}$ , and other cases can be similarly dealt with.
+
+To prove the $\textit{only if}$ direction, given $e \sqsubseteq \textit{MS}(\seqPQ_1^{>})$ and such $\textit{pri}$ and $x$.
+Since $e \sqsubseteq \textit{MS}(\seqPQ_1^{>})$ with witness $x$, we know that $e \sqsubseteq u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$, where $u$, $v$, $w$, $x$ and $\textit{pri}$ satisfy the guard of $\seqPQ_1^{>}$.
 Let $u'$, $v'$ and $w'$ be obtained from $u$, $v$ and $w$ by erasing all items with priority incomparable with $\textit{pri}$, respectively.
-It is not hard to see that $u'$, $v'$, $w'$, $x$ and $\textit{pri}$ satisfy the guard of $\textit{EPQ}_1^{>}$, and then $e \sqsubseteq l = u' \cdot \textit{put}(x,\textit{pri}) \cdot v' \cdot \textit{rm}(x) \cdot w' \in \textit{MS}(\textit{EPQ}_1^{>})$.
+It is not hard to see that $u'$, $v'$, $w'$, $x$ and $\textit{pri}$ satisfy the guard of $\seqPQ_1^{>}$, and then $e \sqsubseteq l = u' \cdot \textit{put}(x,\textit{pri}) \cdot v' \cdot \textit{rm}(x) \cdot w' \in \textit{MS}(\seqPQ_1^{>})$.
 
 To prove the $\textit{if}$ direction, given $e' = \textit{pri-Exec}(e)$ and such $x$ and $\textit{pri}$.
-Since $e' \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$ with witness $x$, we know that $e' \sqsubseteq l_1 = u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$, where $u$, $v$, $w$, $x$ and $\textit{pri}$ satisfy the guard of $\textit{EPQ}_1^{>}$.
+Since $e' \sqsubseteq \textit{MS}(\seqPQ_1^{>})$ with witness $x$, we know that $e' \sqsubseteq l_1 = u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$, where $u$, $v$, $w$, $x$ and $\textit{pri}$ satisfy the guard of $\seqPQ_1^{>}$.
 Let $O_c$ be the set of operations in $e$ that have priorities comparable with $\textit{pri}$, and Let $O_i$ be the set of operations in $e$ that have priorities incomparable with $\textit{pri}$.
@@
 Then $l = u' \cdot \textit{put}(x,\textit{pri}) \cdot v' \cdot \textit{rm}(x) \cdot w'$, where $u' \vert_{O_c} = u$, $v' \vert_{O_c} = v$ and $w' \vert_{O_c} = w$.
-Since $\textit{pri}$ is one of maximal priorities in $e$, and the predicates of guards of $\textit{EPQ}_1^{>}$ does not restrict $O_i$, it is easy to see that $l \in \textit{MS}(\textit{EPQ}_1^{>})$ and then $e \sqsubseteq l$ with witness $x$.
+Since $\textit{pri}$ is one of maximal priorities in $e$, and the predicates of guards of $\seqPQ_1^{>}$ does not restrict $O_i$, it is easy to see that $l \in \textit{MS}(\seqPQ_1^{>})$ and then $e \sqsubseteq l$ with witness $x$.
 \qed
@@
 \label{lemma:UVSet has matched put and rm}
-Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{>}$.
+Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{>}$.
 Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority.
@@
 
-Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{>}$.
+Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{>}$.
 Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority.
@@
 
-To prove the $\textit{only if}$ direction, assume that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{>})$.
-Let $u$, $v$ and $w$ be the sequences of method events in $\textit{EPQ}_1^{>}$, and let $U$, $V$ and $W$ be the set of method events of $u$, $v$ and $w$, respectively.
+To prove the $\textit{only if}$ direction, assume that $e \sqsubseteq \textit{MS}(\seqPQ_1^{>})$.
+Let $u$, $v$ and $w$ be the sequences of method events in $\seqPQ_1^{>}$, and let $U$, $V$ and $W$ be the set of method events of $u$, $v$ and $w$, respectively.
 Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ in $G$.
@@
 Let $\textit{Auts}_{\textit{1-lar}} = \{ \mathcal{A}_{\textit{l-lar}}^1, \mathcal{A}_{\textit{l-lar}}^2, \mathcal{A}_{\textit{l-lar}}^3, \mathcal{A}_{\textit{l-lar}}^4 \}$.
-The following lemma states that $\textit{EPQ}_1^{>}$ is co-regular.
+The following lemma states that $\seqPQ_1^{>}$ is co-regular.
 
@@
 
-We need to prove that, given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-lar}} \cap \mathcal{I} \neq \emptyset$, if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e),$ $\textit{EPQ}_1^{>} \in last(e') \wedge e'$ does not linearizable w.r.t.
-$\textit{MS}(\textit{EPQ}_1^{>})$.
+We need to prove that, given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-lar}} \cap \mathcal{I} \neq \emptyset$, if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e),$ $\seqPQ_1^{>} \in last(e') \wedge e'$ does not linearizable w.r.t.
+$\textit{MS}(\seqPQ_1^{>})$.
 
@@
 
-\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-lar}} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e)$, $last(e')=\textit{EPQ}_1^{>}$, $x$ is the item with maximal priority $\textit{pri}$ in $e'$, $e'$ is a $\textit{pri}$-execution.
+\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-lar}} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e)$, $last(e')=\seqPQ_1^{>}$, $x$ is the item with maximal priority $\textit{pri}$ in $e'$, $e'$ is a $\textit{pri}$-execution.
 And there is a cycle going through $x$ in $G$, where $G$ is the left-right constraint of $e'$.
@@
 Assume that the priority of $b$ is $p$.
-It is easy to see that $e'$ is a $p$-execution, $\textit{last}(e') = \textit{EPQ}_1^{>}$, and there is a cycle going through $d$ in $G$, where $G$ is the left-right constraint of $e'$.
+It is easy to see that $e'$ is a $p$-execution, $\textit{last}(e') = \seqPQ_1^{>}$, and there is a cycle going through $d$ in $G$, where $G$ is the left-right constraint of $e'$.
 The case of $\mathcal{A}_{\textit{l-lar}}^2$, $\mathcal{A}_{\textit{l-lar}}^3$ and $\mathcal{A}_{\textit{l-lar}}^4$ can be similarly proved.
@@
 
-Let $\textit{Items}(e,p)$ be the set of items with priority $p$ in execution $e$. The following lemma states a method to choose $\textit{itm}$ of $\textit{EPQ}_1^{=}$.
+Let $\textit{Items}(e,p)$ be the set of items with priority $p$ in execution $e$. The following lemma states a method to choose $\textit{itm}$ of $\seqPQ_1^{=}$.
 
@@
 \label{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}
-Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{=}$.
+Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{=}$.
 If there exists an item $x$ with priority $\textit{pri}$, such that for each $y \in \textit{Items}(e,\textit{pri})$, (1) $x$ does not $<_{\textit{pb}}$ to $y$, and (2) the right-most gap-point of $x$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},y)$.
-Then $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
+Then $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
 \end{restatable}
@@
 Then it is easy to see that $\textit{put}(x,\textit{pri})$ is after $\textit{put}(y,\textit{pri})$ in $l$.
-Since $\textit{last}(e) = \textit{EPQ}_1^{=}$, all other items in $\textit{Items}(e,\textit{pri})$ has matched $\textit{put}$ and $\textit{rm}$, and it is easy to see that their $\textit{put}$ and $\textit{rm}$ (except for that of $x$) are all before $\textit{rm}(x)$ in $l$.
+Since $\textit{last}(e) = \seqPQ_1^{=}$, all other items in $\textit{Items}(e,\textit{pri})$ has matched $\textit{put}$ and $\textit{rm}$, and it is easy to see that their $\textit{put}$ and $\textit{rm}$ (except for that of $x$) are all before $\textit{rm}(x)$ in $l$.
 
@@
 
-Therefore, it is easy to see that $l \in \textit{MS}(\textit{EPQ}_1^{=})$. \qed
-\end {proof}
-
-With Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, we can prove the following lemma, which states that getting rid of case in \figurename~\ref{fig:introduce pb order} is enough for ensure $\textit{last}(e) = \textit{EPQ}_1^{=} \Rightarrow e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
+Therefore, it is easy to see that $l \in \textit{MS}(\seqPQ_1^{=})$. \qed
+\end {proof}
+
+With Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, we can prove the following lemma, which states that getting rid of case in \figurename~\ref{fig:introduce pb order} is enough for ensure $\textit{last}(e) = \seqPQ_1^{=} \Rightarrow e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
 
@@
 To prove the $\textit{if}$ direction, let $e_{x,y}$ be the execution that is obtained from $e$ by erasing all actions of items that has same priority as $x$, except for actions of $x$ and $y$.
-It is obvious that $\textit{last}(e_{x,y}) = \textit{EPQ}_1^{=}$.
-Since $y <_{\textit{pb}}^* x$, according to $\textit{EPQ}_1^{=}$, we can see that $x$ should be chosen as $\textit{itm}$ in $\textit{EPQ}_1^{=}$.
+It is obvious that $\textit{last}(e_{x,y}) = \seqPQ_1^{=}$.
+Since $y <_{\textit{pb}}^* x$, according to $\seqPQ_1^{=}$, we can see that $x$ should be chosen as $\textit{itm}$ in $\seqPQ_1^{=}$.
 
@@
 Therefore, for every sequence $l = u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$, if $e_{x,y} \sqsubseteq l$, then either $u \cdot v$ contains some unmatched method events of priority smaller than $\textit{pri}$, or $w$ contains $\textit{put}(y,\textit{pri})$ or $\textit{rm}(y)$.
-In both cases, $l \notin \textit{MS}(\textit{EPQ}_1^{=})$.
+In both cases, $l \notin \textit{MS}(\seqPQ_1^{=})$.
 
@@
 Assume we already know that for each $x$ and $y$ has maximal priority in $e$, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{cal}(\textit{put},y,\textit{pri})$ and $\textit{cal}(\textit{rm},x)$.
-We need to prove that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
+We need to prove that $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
 Recall that we already assume that each single-priority execution has FIFO property, and item with larger priority is not covered by items with smaller priority.
@@
 If for each $b \in \textit{Items}(e,\textit{pri})$, the rightmost gap-point of $a_1$ is after $\textit{cal}(\textit{put},b,\textit{pri})$ and $\textit{cal}(\textit{rm},b)$.
-Then by Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, we can obtain that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
+Then by Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, we can obtain that $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
 
@@
 $a^f$ satisfies the check of Step $1$.
-By Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, this implies that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
+By Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, this implies that $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
 This completes the proof of $\textit{if}$ direction.
@@
 \label{lemma:five enumeration is enough for EPQ1Equal}
-Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{=}$.
+Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{=}$.
 Let $a$ and $b$ be items with maximal priority $\textit{pri}$.
@@
 
-Let us begin to represent several witness automata that is used to capture the existence of a data-differentiated $\_$-execution $e$, $e$ has a projection $e'$, $\textit{last}(e') = \textit{EPQ}_1^{=}$, there exists items $a$ and $b$ with maximal priority in $e'$, $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$.
+Let us begin to represent several witness automata that is used to capture the existence of a data-differentiated $\_$-execution $e$, $e$ has a projection $e'$, $\textit{last}(e') = \seqPQ_1^{=}$, there exists items $a$ and $b$ with maximal priority in $e'$, $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$.
 
@@
 
-According to Lemma \ref{lemma:EPQ1Equal as pb order and gap-point}, Lemma \ref{lemma:ob order has bounded length} and Lemma \ref{lemma:five enumeration is enough for EPQ1Equal}, it is not hard to prove that, given a data-differentiated $\_$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_1^{=}$, $e$ does not linearizable with respect to $\textit{EMS}(\textit{PQ}_1^{=})$, if and only if, one of enumerations holds in $e$ (permit renaming), while $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ is covered by some $d_1,\ldots,d_m$, $\textit{cal}(\textit{rm},b)$ is before $\textit{ret}(\textit{put},d_m,\_)$, and $\textit{cal}(\textit{rm},d_1)$ is before $\textit{ret}(\textit{rm},a)$.
+According to Lemma \ref{lemma:EPQ1Equal as pb order and gap-point}, Lemma \ref{lemma:ob order has bounded length} and Lemma \ref{lemma:five enumeration is enough for EPQ1Equal}, it is not hard to prove that, given a data-differentiated $\_$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{=}$, $e$ does not linearizable with respect to $\textit{EMS}(\textit{PQ}_1^{=})$, if and only if, one of enumerations holds in $e$ (permit renaming), while $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ is covered by some $d_1,\ldots,d_m$, $\textit{cal}(\textit{rm},b)$ is before $\textit{ret}(\textit{put},d_m,\_)$, and $\textit{cal}(\textit{rm},d_1)$ is before $\textit{ret}(\textit{rm},a)$.
 We say that such $d_1,\ldots,d_m$ constitute the rightmost gap of $b$.
@@
 
-We need to prove that, given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$, if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e),$ $\textit{EPQ}_1^{=} \in \textit{last}(e') \wedge e'$ does not linearizable w.r.t.
-$\textit{MS}(\textit{EPQ}_1^{=})$.
+We need to prove that, given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$, if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e),$ $\seqPQ_1^{=} \in \textit{last}(e') \wedge e'$ does not linearizable w.r.t.
+$\textit{MS}(\seqPQ_1^{=})$.
 
@@
 
-\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e)$, $\textit{last}(e')=\textit{EPQ}_1^{=}$, $a$ and $b$ are two items with maximal priority $\textit{pri}$ in $e'$, $e'$ is a $\textit{pri}$-execution, $a <_{\textit{pb}}^* b$ in $e'$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\textit{pri})$ or $\textit{cal}(\textit{rm},a)$ in $e'$.
+\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e)$, $\textit{last}(e')=\seqPQ_1^{=}$, $a$ and $b$ are two items with maximal priority $\textit{pri}$ in $e'$, $e'$ is a $\textit{pri}$-execution, $a <_{\textit{pb}}^* b$ in $e'$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\textit{pri})$ or $\textit{cal}(\textit{rm},a)$ in $e'$.
 
@@
 let $e'' = e_2 \vert_{ \{ x,y,z,d_1,\ldots,d_m \} }$.
-It is obvious that $e'' \in \textit{proj}(e_2)$ is a $\textit{pri}$-execution, $\textit{last}(e'') = \textit{EPQ}_1^{=}$.
+It is obvious that $e'' \in \textit{proj}(e_2)$ is a $\textit{pri}$-execution, $\textit{last}(e'') = \seqPQ_1^{=}$.
 According to our construction of automata in $\textit{Auts}_{\textit{1-eq}}$, it is not hard to see that $x$ and $y$ has maximal priority in $h_2$, $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is before $\textit{cal}(\textit{put},y,\textit{pri})$ or $\textit{cal}(\textit{rm},y)$ in $e''$.
@@
 By data-independence, we can obtain execution $e_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the items $d_1,\ldots,d_m$ that constitute the rightmost gap of $b'$, we rename them into $d$, (3) if $a' <_{\textit{pb}}^A a'_1 <_{\textit{pb}}^B b$, we rename $a'_1$ into $a_1$, and (4) rename the other items into $e$.
-It is easy to see that $\textit{last}(e_1) = \textit{EPQ}_1^{=}$, $a$ and $b$ has maximal priority in $e_1$, $a <_{\textit{pb}}^* b$ in $e_1$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\textit{pri})$ or $\textit{cal}(\textit{rm},a)$ in $e_1$.
+It is easy to see that $\textit{last}(e_1) = \seqPQ_1^{=}$, $a$ and $b$ has maximal priority in $e_1$, $a <_{\textit{pb}}^* b$ in $e_1$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\textit{pri})$ or $\textit{cal}(\textit{rm},a)$ in $e_1$.
 By Lemma \ref{lemma:five enumeration is enough for EPQ1Equal}, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists).
@@
 
-\subsection{Co-Regular of $\textit{EPQ}_2^{>}$}
+\subsection{Co-Regular of $\seqPQ_2^{>}$}
 \label{subsec:appendix co-regular of EPQ2Lar}
@@
 \label{lemma:EPQ2Lar is always co-regular}
-Given a data-differentiated $\_$-execution $e$, if $\textit{last}(e) = \textit{EPQ}_2^{>}$, then $e \sqsubseteq \textit{MS}(\textit{EPQ}_2^{>})$.
-\end{restatable}
-
-\begin {proof}
-Since $\textit{last}(e) = \textit{EPQ}_2^{>}$, the actions with maximal priority in $e$ is some unmatched $\textit{put}$.
-Therefore, no matter how we locate linearization points, we can always obtain a sequence $l$ of method events that contains unmatched $\textit{put}$ with maximal priority, and this satisfy the guard of $\textit{MS}(\textit{EPQ}_2^{>})$.
+Given a data-differentiated $\_$-execution $e$, if $\textit{last}(e) = \seqPQ_2^{>}$, then $e \sqsubseteq \textit{MS}(\seqPQ_2^{>})$.
+\end{restatable}
+
+\begin {proof}
+Since $\textit{last}(e) = \seqPQ_2^{>}$, the actions with maximal priority in $e$ is some unmatched $\textit{put}$.
+Therefore, no matter how we locate linearization points, we can always obtain a sequence $l$ of method events that contains unmatched $\textit{put}$ with maximal priority, and this satisfy the guard of $\textit{MS}(\seqPQ_2^{>})$.
 This completes the proof of this lemma.
@@
 
-\subsection{Co-Regular of $\textit{EPQ}_2^{=}$}
+\subsection{Co-Regular of $\seqPQ_2^{=}$}
 \label{subsec:appendix co-regular of EPQ2Equal}
@@
 \label{lemma:EPQ2Equal as happen before}
-Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \textit{EPQ}_2^{=}$.
-$e$ does not linearizable to $\textit{MS}(\textit{EPQ}_2^{=})$, if and only if there exists $x$ and $y$ with priority $\textit{pri}$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,\textit{pri}) <_{\textit{hb}} \textit{put}(y,\textit{pri})$.
+Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \seqPQ_2^{=}$.
+$e$ does not linearizable to $\textit{MS}(\seqPQ_2^{=})$, if and only if there exists $x$ and $y$ with priority $\textit{pri}$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,\textit{pri}) <_{\textit{hb}} \textit{put}(y,\textit{pri})$.
 \end{restatable}
@@
 Assume that for each pair of $x$ and $y$ with maximal priority in $e$, if $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, then $\textit{put}(x,\textit{pri})$ does not happen before $\textit{put}(y,\textit{pri})$.
-We need to prove that $e \sqsubseteq \textit{MS}(\textit{EPQ}_2^{=})$.
+We need to prove that $e \sqsubseteq \textit{MS}(\seqPQ_2^{=})$.
 
@@
 It is easy to see that $e \sqsubseteq l$.
-Since linearization points of $\textit{put}(x_i,\textit{pri})$ is after the linearization point of $\textit{put}(y_j,\textit{pri})$ for each $i,j$, it is easy to see that $l \in \textit{MS}(\textit{EPQ}_2^{=})$.
+Since linearization points of $\textit{put}(x_i,\textit{pri})$ is after the linearization point of $\textit{put}(y_j,\textit{pri})$ for each $i,j$, it is easy to see that $l \in \textit{MS}(\seqPQ_2^{=})$.
 This completes the proof of this lemma.
@@
 
-Lemma \ref{lemma:EPQ2Equal as happen before} shows how to check violation to $\textit{MS}(\textit{EPQ}_2^{=})$.
+Lemma \ref{lemma:EPQ2Equal as happen before} shows how to check violation to $\textit{MS}(\seqPQ_2^{=})$.
 However, the case in Lemma \ref{lemma:EPQ2Equal as happen before} violates our assumption that each single-priority execution is FIFO.
-Therefore, we know that $\textit{EPQ}_2^{=}$ is always co-regular, as states by the following lemma.
+Therefore, we know that $\seqPQ_2^{=}$ is always co-regular, as states by the following lemma.
 
@@
 \label{lemma:EPQ2Equal is always co-regular}
-Given a data-differentiated $\textit{pri}$-execution $e$, if $\textit{last}(e) = \textit{EPQ}_2^{=}$, then $e \sqsubseteq \textit{MS}(\textit{EPQ}_2^{=})$.
-\end{restatable}
-
-\begin {proof}
-
-According to Lemma \ref{lemma:EPQ2Equal as happen before}, if $\textit{last}(e) = \textit{EPQ}_2^{=}$ and $e$ does not linearizable to $\textit{MS}(\textit{EPQ}_2^{=})$, then there exists $x$ and $y$ with priority $\textit{pri}$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$, and $\textit{put}(x,\textit{pri}) <_{\textit{hb}} \textit{put}(y,\textit{pri})$.
+Given a data-differentiated $\textit{pri}$-execution $e$, if $\textit{last}(e) = \seqPQ_2^{=}$, then $e \sqsubseteq \textit{MS}(\seqPQ_2^{=})$.
+\end{restatable}
+
+\begin {proof}
+
+According to Lemma \ref{lemma:EPQ2Equal as happen before}, if $\textit{last}(e) = \seqPQ_2^{=}$ and $e$ does not linearizable to $\textit{MS}(\seqPQ_2^{=})$, then there exists $x$ and $y$ with priority $\textit{pri}$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$, and $\textit{put}(x,\textit{pri}) <_{\textit{hb}} \textit{put}(y,\textit{pri})$.
 Let $e_1 = e \vert_{ \{ x,y \} }$.
@@
 
-\subsection{Co-Regular of $\textit{EPQ}_3$}
+\subsection{Co-Regular of $\seqPQ_3$}
 \label{subsec:co-regular of EPQ3}
 
-In this subsection we prove that $\textit{EPQ}_3$ is co-regular.
-The notion of left-right constraint of $\textit{rm}(\textit{empty})$ is inspired by left-right constraint of queue \cite{Bouajjani:2015}.
+In this subsection we prove that $\seqPQ_3$ is co-regular. The notion of left-right constraint of $\textit{rm}(\textit{empty})$ is inspired by left-right constraint of queue \cite{Bouajjani:2015}.
 
@@
 
-Given a data-differentiated execution $e$ and $o = \textit{rm}(\textit{empty})$ of $e$, it is obvious that $\textit{last}(e) = \textit{EPQ}_3$.
+Given a data-differentiated execution $e$ and $o = \textit{rm}(\textit{empty})$ of $e$, it is obvious that $\textit{last}(e) = \seqPQ_3$.
 Let $\textit{USet}_1(e,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some item, and either $\textit{op} <_{\textit{hb}} o$, or there is $\textit{op}'$ with the same item of $\textit{op}$, such that $\textit{op}' <_{\textit{hb}} o \}$.
@@
 \label{lemma:USet has matched put and rm}
-Given a data-differentiated execution $e$ with $\textit{last}(e) = \textit{EPQ}_3$.
+Given a data-differentiated execution $e$ with $\textit{last}(e) = \seqPQ_3$.
 Let $o$ be a $\textit{rm}(\textit{empty})$ of $e$.
@@
 \label{lemma:Rmx does not happen before USet for EPQ3}
-Given a data-differentiated $\_$execution $e$ with $\textit{last}(e) = \textit{EPQ}_3$.
+Given a data-differentiated $\_$execution $e$ with $\textit{last}(e) = \seqPQ_3$.
 Let $o$ be a $\textit{rm}(\textit{empty})$ of $e$.
@@
 
-Then we can prove that getting rid of cycle though $o$ in left-right constraint is enough for ensure linearizable w.r.t $\textit{MS}(\textit{EPQ}_3)$, as stated by the following lemma.
+Then we can prove that getting rid of cycle though $o$ in left-right constraint is enough for ensure linearizable w.r.t $\textit{MS}(\seqPQ_3)$, as stated by the following lemma.
 
@@
 \label{lemma:Lin Equals Constraint for EPQ3}
-Given a data-differentiated execution $e$ with $\textit{last}(e) = \textit{EPQ}_3$.
+Given a data-differentiated execution $e$ with $\textit{last}(e) = \seqPQ_3$.
 $e$ does not linearizable w.r.t $\textit{MS}(\textit{PQ}_3)$, if and only if there exists $o = \textit{rm}(\textit{empty})$ in $e$, $G$ has a cycle going through $o$, where $G$ is the graph representing the left-right constraint of $o$.
@@
 To prove the $\textit{if}$ direction, assume that there is such a cycle.
-Assume by contradiction that $e \sqsubseteq \textit{MS}(\textit{EPQ}_3)$, and let $U$ and $V$ be the set of operations in $u$ and $v$.
+Assume by contradiction that $e \sqsubseteq \textit{MS}(\seqPQ_3)$, and let $U$ and $V$ be the set of operations in $u$ and $v$.
 Let the cycle be $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow o \rightarrow d_1$ in $G$.
@@
 
-Let us begin to represent an automaton that is used for capture the case that, in a sub-execution $e'$ of an execution $e$, $\textit{last}(e')=\textit{EPQ}_3$, $e'$ does not linearizable to $\textit{MS}(\textit{EPQ}_3)$, and the reason is that there is a cycle going through some $\textit{rm}(\textit{empty})$ $o$ in the left-right constraint of $o$.
-The automaton is $\mathcal{A}_{\textit{EPQ}}^3$, which is given in \figurename~\ref{fig:automata for PQ3}.
+Let us begin to represent an automaton that is used for capture the case that, in a sub-execution $e'$ of an execution $e$, $\textit{last}(e')=\seqPQ_3$, $e'$ does not linearizable to $\textit{MS}(\seqPQ_3)$, and the reason is that there is a cycle going through some $\textit{rm}(\textit{empty})$ $o$ in the left-right constraint of $o$.
+The automaton is $\mathcal{A}_{\seqPQ}^3$, which is given in \figurename~\ref{fig:automata for PQ3}.
 In \figurename~\ref{fig:automata for PQ3}, let $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d), \textit{ret}(\textit{rm},d),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{cal}(\textit{put},b,\textit{anyPri})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c + \textit{ret}(\textit{rm},b)$.
@@
 %\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{EPQ}}^3$}
+  \caption{Automaton $\mathcal{A}_{\seqPQ}^3$}
   \label{fig:automata for PQ3}
@@
 \label{lemma:EPQ3 is co-regular}
-$\textit{EPQ}_3$ is co-regular.
-\end{restatable}
-
-\begin {proof}
-
-We need to prove that, given a data-independence implementation $\mathcal{I}$, $\mathcal{A}_{\textit{EPQ}}^3 \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), last(e')=\textit{EPQ}_3 \wedge e$ does not linearizable w.r.t.
-$\textit{MS}(\textit{EPQ}_3)$.
+$\seqPQ_3$ is co-regular.
+\end{restatable}
+
+\begin {proof}
+
+We need to prove that, given a data-independence implementation $\mathcal{I}$, $\mathcal{A}_{\seqPQ}^3 \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), last(e')=\seqPQ_3 \wedge e$ does not linearizable w.r.t.
+$\textit{MS}(\seqPQ_3)$.
 
@@
 
-\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\mathcal{A}_{\textit{EPQ}}^3 \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), last(e')=\textit{EPQ}_3$, $o = \textit{rm}(\textit{empty})$ is in $e'$, and $o$ is covered by some items $d_1,\ldots,d_m$ in $e'$.
-
-\noindent The $\textit{only if}$ direction: Assume that $e_1 \in \mathcal{I}$ is accepted by $\mathcal{A}_{\textit{EPQ}}^3$.
+\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\mathcal{A}_{\seqPQ}^3 \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), last(e')=\seqPQ_3$, $o = \textit{rm}(\textit{empty})$ is in $e'$, and $o$ is covered by some items $d_1,\ldots,d_m$ in $e'$.
+
+\noindent The $\textit{only if}$ direction: Assume that $e_1 \in \mathcal{I}$ is accepted by $\mathcal{A}_{\seqPQ}^3$.
 By data-independence, there exists data-differentiated execution $e_2 \in \mathcal{I}$ and a renaming function $r$, such that $e_1=r(e_2)$.
@@
 Let $e_3 = e_2 \vert_{ \{ o, d_1, \ldots, d_m \} }$.
-It is obvious that $e_3 \in \textit{proj}(e_2)$ and $\textit{last}(e_3) = \textit{EPQ}_3$.
+It is obvious that $e_3 \in \textit{proj}(e_2)$ and $\textit{last}(e_3) = \seqPQ_3$.
 It is easy to see that $o$ is covered by $d_1,\ldots,d_m$.
@@
 By data-independence, $e_1 \in \mathcal{I}$.
-It is easy to see that $e_1$ is accepted by $\mathcal{A}_{\textit{EPQ}}^3$.
+It is easy to see that $e_1$ is accepted by $\mathcal{A}_{\seqPQ}^3$.
 
@@
 
-Similar as witness automata for $\textit{EPQ}_3$, we have witness automata $\mathcal{A}_{\textit{MS}}^4$ for $\textit{MSet}_3$, which is shown in \figurename~\ref{fig:automata 4 for MS-3 in appendix}.
+Similar as witness automata for $\seqPQ_3$, we have witness automata $\mathcal{A}_{\textit{MS}}^4$ for $\textit{MSet}_3$, which is shown in \figurename~\ref{fig:automata 4 for MS-3 in appendix}.
 In \figurename~\ref{fig:automata 4 for MS-3 in appendix}, let $c = \textit{cal}(\textit{insert},d),\textit{ret}(\textit{insert},d), \textit{cal}(\textit{delete},d)$, $\textit{ret}(\textit{delete},d),\textit{cal}(\textit{delete},\textit{empty}),\textit{ret}(\textit{delete},\textit{empty})$, $c_1 = c + \textit{cal}(\textit{insert},b)$, $c_2 = c_1 + \textit{ret}(\textit{delete},b)$, and $c_3 = c + \textit{ret}(\textit{delete},b)$.
@@
 For the $\textit{only if}$ direction, given an execution $e_m \in \mathcal{I}_m$ of multi-set, such that $e_m$ is accepted by some automaton in $\textit{Auts}_{\textit{MS}}$.
-If $e_m$ is accepted by $\mathcal{A}_{\textit{MS}}^1$, $\mathcal{A}_{\textit{MS}}^2$, $\mathcal{A}_{\textit{MS}}^3$ or $\mathcal{A}_{\textit{MS}}^4$, then it is easy to see that $\textit{MStoEPQ}(e_m)$ is accepted by $\mathcal{A}_{\textit{SinPri}}^2$, $\mathcal{A}_{\textit{SinPri}}^3$, $\mathcal{A}_{\textit{SinPri}}^1$ or $\mathcal{A}_{\textit{EPQ}}^3$, respectively.
-
-For the $\textit{if}$ direction, given an execution $e_{\textit{epq}}$ of extended priority queue, such that $e_{\textit{epq}} = \textit{MStoEPQ}(e_m)$ for some $e_m$ of multi-set, and $e_{\textit{epq}}$ is accepted by some automaton in $\textit{Auts}_{\textit{EPQ}}$.
+If $e_m$ is accepted by $\mathcal{A}_{\textit{MS}}^1$, $\mathcal{A}_{\textit{MS}}^2$, $\mathcal{A}_{\textit{MS}}^3$ or $\mathcal{A}_{\textit{MS}}^4$, then it is easy to see that $\textit{MStoEPQ}(e_m)$ is accepted by $\mathcal{A}_{\textit{SinPri}}^2$, $\mathcal{A}_{\textit{SinPri}}^3$, $\mathcal{A}_{\textit{SinPri}}^1$ or $\mathcal{A}_{\seqPQ}^3$, respectively.
+
+For the $\textit{if}$ direction, given an execution $e_{\textit{epq}}$ of extended priority queue, such that $e_{\textit{epq}} = \textit{MStoEPQ}(e_m)$ for some $e_m$ of multi-set, and $e_{\textit{epq}}$ is accepted by some automaton in $\textit{Auts}_{\seqPQ}$.
 
 According to definition of $\textit{MStoEPQ}$, we can see that (1) in $e_{\textit{epq}}$ there does not exist two items with comparable priorities, and (2) in $e_{\textit{epq}}$, there does not exists two items with a same priority.
-Therefore, the witness automata for $\textit{EPQ}_1$ and the witness automaton $\mathcal{A}_{\textit{SinPri}}^4$ can be safely ignored.
-Then, if $e_{\textit{epq}}$ is accepted by $\mathcal{A}_{\textit{SinPri}}^1$, $\mathcal{A}_{\textit{SinPri}}^2$, $\mathcal{A}_{\textit{SinPri}}^3$ or $\mathcal{A}_{\textit{EPQ}}^3$, then it is easy to see that $e_m$ is accepted by $\mathcal{A}_{\textit{MS}}^3$, $\mathcal{A}_{\textit{MS}}^1$, $\mathcal{A}_{\textit{MS}}^2$ and $\mathcal{A}_{\textit{MS}}^4$, respectively.
+Therefore, the witness automata for $\seqPQ_1$ and the witness automaton $\mathcal{A}_{\textit{SinPri}}^4$ can be safely ignored.
+Then, if $e_{\textit{epq}}$ is accepted by $\mathcal{A}_{\textit{SinPri}}^1$, $\mathcal{A}_{\textit{SinPri}}^2$, $\mathcal{A}_{\textit{SinPri}}^3$ or $\mathcal{A}_{\seqPQ}^3$, then it is easy to see that $e_m$ is accepted by $\mathcal{A}_{\textit{MS}}^3$, $\mathcal{A}_{\textit{MS}}^1$, $\mathcal{A}_{\textit{MS}}^2$ and $\mathcal{A}_{\textit{MS}}^4$, respectively.
 \qed

### 2017-04-21 checking-executions.tex
something

@@
 
-TODO REPHRASE THIS
-While checking linearizability of a given execution is known to be NP-complete in general~\cite{journals/siamcomp/GibbonsK97} we show in this section that it is polynomial-time for data-differentiated executions of the priority queue.
-For pedagogical reasons, Section~\ref{ssec:seq_exec} introduces a polynomial-time recursive procedure for checking whether a data-differentiated \emph{sequential} execution is admitted by the priority queue which is then extended to the concurrent case in Section~\ref{ssec:conc_exec}.
+%TODO REPHRASE THIS
+%While checking linearizability of a given execution is known to be NP-complete in general~\cite{journals/siamcomp/GibbonsK97} we show in this section that it is polynomial-time for data-differentiated executions of the priority queue.
+We define a recursive procedure for checking linearizability of an execution w.r.t. $\seqPQ$.
+To ease the exposition, Section~\ref{ssec:seq_exec} introduces a recursive procedure for checking whether a data-differentiated \emph{sequential} execution is admitted by the priority queue which is then extended to the concurrent case in Section~\ref{ssec:conc_exec}.
 
@@
 
-Checking whether a data-differentiated sequential execution belongs to $\seqPQ$ could be implemented by checking membership into the language accepted by $LTS??$.
+Checking whether a data-differentiated sequential execution belongs to $\seqPQ$ could be implemented by checking membership into the language accepted by the LTS $PQ$.
 The recursive procedure $\textit{Check-PQ-Seq}$ outlined in Algorithm~\ref{alg:seq_check} is an alternative to this membership test.
@@
 the selected ones, and calls itself recursively on the execution without the selected operations.
-{\color {blue} Here we implicitly mix operation $o$ (when $\textit{call}_o(m,a)$, $\textit{ret}_o(m,b)$ in $e$) with $m(a,b)$.}
+%{\color {blue} Here we implicitly mix operation $o$ (when $\textit{call}_o(m,a)$, $\textit{ret}_o(m,b)$ in $e$) with $m(a,b)$.}
 
@@
 where $p_1$, $p_2$, $p_3$ are priorities such that $p_1 \prec p_2$ and $p_1 \prec p_3$, and $p_2$ and $p_3$ are incomparable.
-Since the $\textit{rm}(\textit{empty})$ operations are read-only (they don't affect the state of the queue), they are selected first.
+Since the $\textit{rm}(\textit{empty})$ operations\footnote{An $m(a)$-operation in an execution $e$ is an operation identifier $o$ s.t.
+$e$ contains the actions $\textit{call}_o(m,a)$, $\textit{ret}_o(m,a)$.} are read-only (they don't affect the state of the queue), they are selected first.
 Ensuring that an operation $o=\textit{rm}(\textit{empty})$ is correct boils down to checking that every $\textit{put}(x,p)$ operation before $o$ is matched to a $\textit{rm}(x)$ operation which also occurs before $o$.
@@
 \end{align*}
-{\color{blue}When there are no more $\textit{rm}(\textit{empty})$ operations, the procedure try to select a $\textit{put}$ operation adding a value with maximal priority which is not removed, and then a pair of $\textit{put}$ and $\textit{rm}$ operations adding and removing the same maximal priority value.} For instance, since $p_2$ is a maximal priority, it may select the operation $\textit{put}(d,p_2)$.
-{\color {blue}This operation is correct if $d$ is the last value with priority $p_2$,} and the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
+When there are no more $\textit{rm}(\textit{empty})$ operations, the procedure selects a $\textit{put}$ operation adding a value with maximal priority which is not removed, and then a pair of $\textit{put}$ and $\textit{rm}$ operations adding and removing the same maximal priority value.
+For instance, since $p_2$ is a maximal priority, it may select the operation $\textit{put}(d,p_2)$.
+{This operation is correct if $d$ is the last value with priority $p_2$, and the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
 \begin{align*}
@@
 \end{align*}
-{\color {blue}Now, all the putted value that is not removed of priority $p_2$ do not exists, and the procedure selects a pair of $\textit{put}$/$\textit{rm}$ operations of priority $p_2$,} for instance, $\textit{put}(c,p_2)$ and $\textit{rm}(c)$.
-In this case, the value returned by $\textit{rm}$ is correct if all the values of priority smaller than $p_2$ added before $\textit{rm}(c)$ are also removed before $\textit{rm}(c)$.
-In this case, the only value of priority smaller than $p_2$ is $a$ and satisfies this property.
+Since there is no other value of priority $p_2$ which is not removed, the procedure selects a pair of $\textit{put}$/$\textit{rm}$ operations with an argument of priority $p_2$,} for instance, $\textit{put}(c,p_2)$ and $\textit{rm}(c)$.
+The value returned by $\textit{rm}$ is correct if all the values of priority smaller than $p_2$ added before $\textit{rm}(c)$ are also removed before $\textit{rm}(c)$.
+In this case, $a$ is the only value of priority smaller than $p_2$ and it satisfies this property.
 Applying a similar reasoning for all the remaining values, it can be proved that this execution is correct.
@@
 Completeness follows easily from the properties of $\seqPQ$.
-Thus, if $\textit{Check-PQ-Conc}(e) = \textit{false}$, then there exists a set of values $D$ such that one of the following doesn't hold: $\mathsf{EmptyRemove\text{-}Conc}(e \vert D)$, {\color {blue}$\mathsf{UnmatchedMaxPriority\text{-}Conc}(e \vert D,x)$ for all values that are unmatched and with maximal priority $x$, or $\mathsf{MatchedMaxPriority\text{-}Conc}(e \vert D,x)$ for all maximal priority values $x$ where values of $x$ are all matched.} It can be easily seen that we get $e \vert D\not\sqsubseteq \seqPQ$ in all cases, which by the closure under projection of $\seqPQ$ implies, $e \not\sqsubseteq \seqPQ$ (since every linearization of $e$ includes as a subsequence a linearization of $e \vert D$).
+Thus, if $\textit{Check-PQ-Conc}(e) = \textit{false}$, then there exists a set $D$ of values s.t.
+either  $\mathsf{EmptyRemove\text{-}Conc}(e \vert D)$ is false, or $\mathsf{UnmatchedMaxPriority\text{-}Conc}(e \vert D,x)$ is false for all the values $x$ of maximal priority that are not removed (and there exists at least one such value), or $\mathsf{MatchedMaxPriority\text{-}Conc}(e \vert D,x)$ is false for all the values $x$ of maximal priority (and these values are all removed in $e \vert D$).
+It can be easily seen that we get $e \vert D\not\sqsubseteq \seqPQ$ in all cases, which by the closure under projection of $\seqPQ$ implies, $e \not\sqsubseteq \seqPQ$ (since every linearization of $e$ includes as a subsequence a linearization of $e \vert D$).
 
@@
 %
-
-TODO UPDATE THE NOTATIONS IN THIS EXAMPLE
 

### 2017-04-21 co-regular-EPQ.tex
modify
something

@@
 
-The following result shows that deriving $\Gamma$-complete automata for each $\Gamma$ enables an effective reduction of checking linearizability of concurrent priority queue implementations to state reachability.
-It is a direct consequence of the above definitions.
+Section~\ref{ssec:aut} describes a $\mathsf{MatchedMaxPriority}$-complete automaton, the other automata being defined in Appendix~\ref{} or Icalp??. Therefore, the following holds.
 
-\begin{restatable}{theorem}{ReduceEPQIntoStateReachability}
-\label{lemma:reduce EPQ into state reachability}
-Let $\mathcal{I}$ be a data-independent implementation, and $A(\Gamma)$ a $\Gamma$-complete automaton for each $\Gamma$. Then,
-$\mathcal{I} \sqsubseteq \seqPQ$ if and only if $\mathcal{I} \cap A(\Gamma) = \emptyset$ for all $\Gamma$.
-\end{restatable}
+\begin{lemma}
+There exists a $\Gamma$-complete automaton for each $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$.
+\end{lemma}
 
-Section~\ref{ssec:aut} describes a $\mathsf{MatchedMaxPriority}$-complete automaton (TODO the other automata are discussed in Appendix~\ref{} or Icalp??) while Section~\ref{} discusses decidability results implied by the theorem above.
-
-W.l.o.g., we assume that every implementation $\mathcal{I}$ behaves correctly, i.e., as a FIFO queue, when only values with the same priority are observed.
+When defining $\Gamma$-complete automata, we assume that every implementation $\mathcal{I}$ behaves correctly, i.e., as a FIFO queue, when only values with the same priority are observed.
 More precisely, we assume that for every execution $e\in\mathcal{I}$ and every priority $p\in\mathbb{P}$, the projection of $e$ to values with priority $p$ is linearizable (w.r.t.
@@
 \end {proof}
+
+The following result shows that $\Gamma$-complete automata enable an effective reduction of checking linearizability of concurrent priority queue implementations to state reachability.
+It is a direct consequence of the above definitions.
+
+\begin{restatable}{theorem}{ReduceEPQIntoStateReachability}
+\label{lemma:reduce EPQ into state reachability}
+Let $\mathcal{I}$ be a data-independent implementation, and $A(\Gamma)$ a $\Gamma$-complete automaton for each $\Gamma$. Then,
+$\mathcal{I} \sqsubseteq \seqPQ$ if and only if $\mathcal{I} \cap A(\Gamma) = \emptyset$ for all $\Gamma$.
+\end{restatable}
+
+Section~\ref{subsec:combine step-by-step linearizability and co-regular} discusses decidability results implied by the theorem above.
+
+%Section~\ref{ssec:aut} discusses
 
@@
 
-TODO WHY IT IS A SET OF AUTOMATA ?
+%TODO WHY IT IS A SET OF AUTOMATA ?
 SAY WHAT ARE THE CASES THAT NEED TO BE COVERED.
@@
 USE BIG CAPS FOR SETS OF TRANSITION LABELS.
+TODO COMPLETE THE FOLLOWING
+Defining a $\mathsf{MatchedMaxPriority}^>$-complete automaton needs to consider ??
+cases depending on the order between ??.
+For each case, we define an automaton recognizing the induced set of violations.
+For example, the automaton corresponding to the case ??
+is pictured in in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
+More details can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
 
@@
 
-{\color {blue}
-Let $\mathcal{A}_{\textit{1-lar}}$ be the union of the four register automata mentioned above, then we prove that $\mathcal{A}_{\textit{1-lar}}$ is $\mathsf{MatchedMaxPriority}^{>}$-complete.
+%{\color {blue}
+%Let $\mathcal{A}_{\textit{1-lar}}$ be the union of the four register automata mentioned above, then we prove that $\mathcal{A}_{\textit{1-lar}}$ is $\mathsf{MatchedMaxPriority}^{>}$-complete.
 The detailed construction and proof can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
-}
-
-\begin{restatable}{lemma}{EPQOneLarisCoRegular}
-\label{lemma:EPQ1Lar is co-regular}
-$\textit{Auts}_{\textit{1-lar}}$ is $\mathsf{MatchedMaxPriority}^{>}$-complete.
-\end{restatable}
-
-More details can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}. %TODO MOVE THIS TO THE RELATED WORK
+%}
+%
+%\begin{restatable}{lemma}{EPQOneLarisCoRegular}
+%\label{lemma:EPQ1Lar is co-regular}
+%$\textit{Auts}_{\textit{1-lar}}$ is $\mathsf{MatchedMaxPriority}^{>}$-complete.
+%\end{restatable}
 
@@
 
-The result of Lemma \ref{lemma:EPQ1Equal as pb order and gap-point} still needs to be impove, since obtaining $y <_{\textit{pb}}^* x$ may requires arbitrary many intermediate values, and it is hard to store unbounded values by automata.
+To characterize violations to linearizability w.r.t.
+$\mathsf{MatchedMaxPriority}^{=}$ using an automaton that tracks a bounded number of values, we show that the number of values needed to witness that $y <_{\textit{pb}}^* x$ for some $x$ and $y$ is bounded.
+%
+%While Lemma \ref{lemma:EPQ1Equal as pb order and gap-point} doesn't imply a bound for 
+%the number of values needs to witness that 
+%
+%The result of Lemma \ref{lemma:EPQ1Equal as pb order and gap-point} still needs to be impove, since obtaining $y <_{\textit{pb}}^* x$ may requires arbitrary many intermediate values, and it is hard to store unbounded values by automata.
 Fortunately, we prove that the number of intermediate values is essentially bounded, as stated by the following lemma.
@@
 \label{lemma:ob order has bounded length}
-Given a data-differentiated execution $e$. Assume that $a <_{\textit{pb}} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}} a_m <_{\textit{pb}} b$, then one of the following cases holds:
-
+Let $e$ be a data-differentiated execution such that $a <_{\textit{pb}} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}} a_m <_{\textit{pb}} b$ holds for some set of values $a$, $a_1$,$\ldots$,$a_m$, $b$.
+Then, one of the following holds:
 \begin{itemize}
 \setlength{\itemsep}{0.5pt}
-\item[-] $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$ or $a <_{\textit{pb}}^C b$,
+\item[-] $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$, or $a <_{\textit{pb}}^C b$,
 
-\item[-] $a <_{\textit{pb}}^A a_i <_{\textit{pb}}^B b$, or $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A b$, for some $i$.
+\item[-] $a <_{\textit{pb}}^A a_i <_{\textit{pb}}^B b$ or $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A b$, for some $i$.
 %\item[-] $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A a_j <_{\textit{pb}}^B b$, for some $i$ and $j$,
@@
 
-By Lemma \ref{lemma:ob order has bounded length}, we can use register automata to detect $a <_{\textit{pb}}^* b$ by enumerating all possible enumerations of operations of $a$, $b$ and possibly $a_1$.
-Since some combinations of $<_{\textit{pb}}^*$ and gap-points are conflict, we finally reduce the number of potential enumerations of operations of $a$, $b$ and possibly $a_1$ into only five in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}.
+To characterize violations to linearizability w.r.t.
+$\mathsf{MatchedMaxPriority}^{=}$, one has to consider all the possible orders between call/return actions of the operations on values $a$, $b$, and $a_i$ in Lemma \ref{lemma:ob order has bounded length}, and the right-most gap point of $b$.
+Excluding the inconsistent cases, we are left with 5 possible orders that are listed in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}.
+For each case, we define an automaton recognizing the induced set of violations.
+For instance, one such case and the corresponding register automaton is shown in \figurename~\ref{fig:an enumeration and its witness automaton} ($o$ denotes the rightmost gap-point of $b$).
+In this case, the conditions in Lemma~\ref{lemma:EPQ1Equal as pb order and gap-point} are equivalent to the fact that intuitively, the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$ is covered by lower priority values (and thus, there is no gap-point of $b$ which occurs after $\textit{call}(\textit{rm},a)$).
+Using again the data-independence property, these lower priority values can all be renamed to a fixed value $d$, and the other values to a fixed value $e$.
 
-In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}, we construct a set $\textit{Auts}_{\textit{1-eq}}$ of register automata, and prove that $\mathsf{MatchedMaxPriority}^{=}$ is co-regular, as stated by the following lemma.
-
-\begin{restatable}{lemma}{EPQOneEqualIsCoRegular}
-\label{lemma:EPQ1Equal is co-regular}
-$\mathsf{MatchedMaxPriority}^{=}$ is co-regular.
-\end{restatable}
-
-One such enumeration and its register automata is shown in \figurename~\ref{fig:an enumeration and its witness automaton}.
-$o$ is the rightmost gap-point of $b$, we rename the values that ``covers the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$'' (see Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}) into $d$, and rename the renaming values into $e$.
-In this figure, $c = \textit{call}(\textit{put},e,\top),\textit{ret}(\textit{put},e,\top)$, $\textit{call}(\textit{rm},e), \textit{ret}(\textit{rm},e),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{call}(\textit{put},d,<r)$, $c_2 = c_1 + \textit{ret}(\textit{put},b,r)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b,r) + \textit{ret}(\textit{rm},d)$.
+%We denote by $o$ the rightmost gap-point of $b$, we rename the values that ``cover the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$'' (see Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}) into $d$, and rename the renaming values into $e$.
 
@@
 %\vspace{-10pt}
-  \caption{An enumerations and its register automaton}
+  \caption{A case in deriving a $\mathsf{MatchedMaxPriority}^{=}$-complete automaton.
+We use the following notations: $c = \textit{call}(\textit{put},e,\top),\textit{ret}(\textit{put},e,\top)$, $\textit{call}(\textit{rm},e), \textit{ret}(\textit{rm},e),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{call}(\textit{put},d,<r)$, $c_2 = c_1 + \textit{ret}(\textit{put},b,r)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b,r) + \textit{ret}(\textit{rm},d)$.}
   \label{fig:an enumeration and its witness automaton}
@@
 
-\subsection{Reduce into State Reachability}
+%we can use register automata to detect $a <_{\textit{pb}}^* b$ by enumerating all possible enumerations of operations of $a$, $b$ and possibly $a_1$.
+Since some combinations of $<_{\textit{pb}}^*$ and gap-points are conflict, we finally reduce the number of potential enumerations of operations of $a$, $b$ and possibly $a_1$ into only five in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}.
+
+%In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}, we construct a set $\textit{Auts}_{\textit{1-eq}}$ of register automata, and prove that $\mathsf{MatchedMaxPriority}^{=}$ is co-regular, as stated by the following lemma.
+%
+%\begin{restatable}{lemma}{EPQOneEqualIsCoRegular}
+%\label{lemma:EPQ1Equal is co-regular}
+%$\mathsf{MatchedMaxPriority}^{=}$ is co-regular.
+%\end{restatable}
+
+\subsection{Decidability Result}
 \label{subsec:combine step-by-step linearizability and co-regular}
 
-We also prove that $\mathsf{UnmatchedMaxPriority}$ are co-regular in Appendix \ref{subsec:appendix co-regular of EPQ2Lar}, Appendix \ref{subsec:appendix co-regular of EPQ2Equal}, and prove that $\mathsf{EmptyRemove}$ are co-regular in Appendix \ref{subsec:co-regular of EPQ3}.
+We describe a class $\mathcal{C}$ of data-independent implementations for which linearizability w.r.t.
+$\seqPQ$ is decidable.
+The implementations in $\mathcal{C}$ allow an unbounded number of values but a bounded number of priorities.
+Each method manipulates a finite number of local variables which store Boolean values, or data values from $\mathbb{D}$.
+Methods communicate through a finite number of shared variables that also store Boolean values, or data values from $\mathbb{D}$.
+Data values may be assigned, but never used in program predicates (e.g., in the conditions of if-then-else and while statements) so as to ensure data independence.
+This class captures typical implementations, or finite-state abstractions thereof, e.g., obtained via predicate abstraction.
+Since the $\Gamma$-complete automata $A(\Gamma)$ we define use a fixed set $D=\{a,b,d,e,\top\}$ of values, we have that $\mathcal{I}\cap A(\Gamma)\neq\emptyset$ for some $\Gamma$ iff $\mathcal{I}_D\cap A(\Gamma)\neq\emptyset$ where $\mathcal{I}_D$ is the subset of $\mathcal{I}$ that uses only values in $D$.
+The set of executions $\mathcal{I}_D$ can be represented by a Petri Net since bounding the set of values allows us to represent each thread with a finite-state machine.
+For a fixed set of priorities $P$, the register automata $A(\Gamma)$ can be transformed to finite-state automata since the number of possible valuations of the registers is bounded.
+
+\begin{theorem}
+Verifying whether an implementation in $\mathcal{C}$ is linearizable w.r.t. $\seqPQ$ is PSPACE-complete for a fixed number of threads, and EXPSPACE-complete otherwise.
+\end{theorem}
+
+%We also prove that $\mathsf{UnmatchedMaxPriority}$ are co-regular in Appendix \ref{subsec:appendix co-regular of EPQ2Lar}, Appendix \ref{subsec:appendix co-regular of EPQ2Equal}, and prove that $\mathsf{EmptyRemove}$ are co-regular in Appendix \ref{subsec:co-regular of EPQ3}.
 Then we can see that $\seqPQ$ is co-regular.
-
-\begin{restatable}{lemma}{EPQIsCoRegular}
-\label{lemma:EPQ is co-regular}
-$\seqPQ$ is co-regular.
-\end{restatable}
-
-By Lemma \ref{lemma:EPQ as multi in MRpri for history} and Lemma \ref{lemma:EPQ is co-regular}, we can finally reduce the verification of linearizability w.r.t $\seqPQ$ into a reachability problem, as illustrated by the following theorem, where $\textit{Auts}_{\textit{EPQ}}$ is the set of witness automata of this section.
-
-\begin{restatable}{theorem}{ReduceEPQIntoStateReachability}
-\label{lemma:reduce EPQ into state reachability}
-Given a data-independence implementation $\mathcal{I}$. $\mathcal{I} \sqsubseteq \seqPQ$, if and only if, $\mathcal{I} \cap \textit{Auts}_{\textit{EPQ}} = \emptyset$.
-\end{restatable}
-
-When the data domain and priority domain are both fixed to be finite set, the register automata can be simulated with finite automata, and by Lemma \ref{lemma:reduce EPQ into state reachability}, checking linearizable w.r.t $\seqPQ$ is decidable.
+%
+%\begin{restatable}{lemma}{EPQIsCoRegular}
+%\label{lemma:EPQ is co-regular}
+%$\seqPQ$ is co-regular.
+%\end{restatable}
+%
+%By Lemma \ref{lemma:EPQ as multi in MRpri for history} and Lemma \ref{lemma:EPQ is co-regular}, we can finally reduce the verification of linearizability w.r.t $\seqPQ$ into a reachability problem, as illustrated by the following theorem, where $\textit{Auts}_{\textit{EPQ}}$ is the set of witness automata of this section.
+%
+%\begin{restatable}{theorem}{ReduceEPQIntoStateReachability}
+%\label{lemma:reduce EPQ into state reachability}
+%Given a data-independence implementation $\mathcal{I}$. $\mathcal{I} \sqsubseteq \seqPQ$, if and only if, $\mathcal{I} \cap \textit{Auts}_{\textit{EPQ}} = \emptyset$.
+%\end{restatable}
+%
+%When the data domain and priority domain are both fixed to be finite set, the register automata can be simulated with finite automata, and by Lemma \ref{lemma:reduce EPQ into state reachability}, checking linearizable w.r.t $\seqPQ$ is decidable.

### 2017-04-21 introduction.tex
something

@@
 Actually, only two registers are needed: one register $r_1$ for storing a priority guessed at the initial state, and one register $r_2$ for reading priorities as they occur in the execution and comparing them with the one stored in $r_1$.
-We show that registers storing values added to or removed from the priority queue are not needed, since any data-independent implementation admits a violation to linearizability whenever it admits a violation where the number of values is constant, and at most 4 (the number of priorities can still be unbounded).
+We show that registers storing values added to or removed from the priority queue are not needed, since any data-independent implementation admits a violation to linearizability whenever it admits a violation where the number of values is constant, and at most 5 (the number of priorities can still be unbounded).
 

### 2017-04-21 priority-queues.tex
something

@@
 Let $\textit{proj}(e)$ be the set of all projections of $e$ to a set of values.
-{\color {blue} The proof of the following lemma can be found in Appendix \ref{sec:appendix proofs in section priority queue and data-independence}.}
+The proof of the following lemma can be found in Appendix \ref{sec:appendix proofs in section priority queue and data-independence}.
 
@@
 
-The following lemma states that by data-independence, it is enough to consider only data-differentiated executions when checking linearizability.
-{\color {blue}It is proved in Appendix \ref{sec:appendix proofs in section priority queue and data-independence}.} This is similar to that in \cite{conf/tacas/AbdullaHHJR13,DBLP:conf/icalp/BouajjaniEEH15}, where they use the notion of data-independence in \cite{conf/popl/Wolper86}.
+The following lemma states that by data-independence, it is enough to consider only data-differentiated executions when checking linearizability (see Appendix \ref{sec:appendix proofs in section priority queue and data-independence}).
+This is similar to that in \cite{conf/tacas/AbdullaHHJR13,DBLP:conf/icalp/BouajjaniEEH15}, where they use the notion of data-independence in \cite{conf/popl/Wolper86}.
 Section~\ref{sec:checking inclusion by recursive procedure} will focus on characterizing linearizability for data-differentiated executions.

### 2017-04-21 appendix-EPQ.tex
something
modify

@@
 The following lemma states that $\seqPQ$ is indeed the set of sequences obtained by renaming sequences accepted by $\textit{Check-PQ-Seq}$.
-
-$\textit{Check-PQ-Seq}(e)=\mathsf{true}$ iff $e\in \seqPQ$, for every data-differentiated sequential execution $e$..
+Given $\Gamma\in \{\mathsf{EmptyRemove}, \mathsf{Unmatched-}$ $\mathsf{MaxPriority}, \mathsf{MatchedMaxPriority}\}$, let us use $l_2 \xrightarrow{\Gamma} l_1$ to mean that when we use $\textit{Check-PQ-Seq}$ to check $l_2$, we choose the branch of $\mathsf{Has\text{-}\Gamma}$ and finally recursively call $\textit{Check-PQ-Seq}$ to check $l_1$.
 
@@
 
-Let $\textit{SeqPQ}_f$ be the set of data-differentiated sequences, such that $e \in \textit{SeqPQ}_f$, if $\textit{Check-PQ-Seq}(e)=\mathsf{true}$.
+Let $\textit{SeqPQ}_f$ be the set of data-differentiated sequences, such that $e \in \textit{SeqPQ}_f$, if $\textit{Check-PQ-}$ $\textit{Seq}(e)=\mathsf{true}$.
 We need to prove that $\textit{SeqPQ}_f = \seqPQ_{\neq}$.
-Given $\Gamma\in \{\mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, let us use $l_2 \xrightarrow{\Gamma} l_1$ to mean that when we use $\textit{Check-PQ-Seq}$ to check $l_2$, we choose the branch of $\mathsf{Has\text{-}\Gamma}$ and finally recursively call $\textit{Check-PQ-Seq}$ to check $l_1$.
-
 We prove $\textit{SeqPQ}_f \subseteq \seqPQ_{\neq}$ by induction.
@@
 
-$\textit{Check-PQ-Conc}(e)=\mathsf{true}$ iff $e \sqsubseteq \seqPQ$, for every data-differentiated execution $e$.
-
 To prove the $\textit{if}$ direction, given a data-differentiated $e \sqsubseteq l \in \seqPQ$. Then we being a loop as follows,
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] If $\mathsf{Has\text{-}EmptyRemoves}(e)$, then $l = u \cdot o \cdot v$ for some $o=\textit{rm}(\textit{empty})$.
-It is not hard to see that $\mathsf{EmptyRemove\text{-}Seq}(l,o)$.
+\item[-] If $\mathsf{Has\text{-}EmptyRemoves}(e)$ holds, then $l = u \cdot o \cdot v$ for some $o=\textit{rm}(\textit{empty})$.
+It is not hard to see that $\mathsf{EmptyRemove\text{-}Seq}(l,o)$ holds.
 Let $e'$ be the projection of $e$ into operations of $u \cdot v$.
@@
 
-\item[-] \item[-] If $\mathsf{Has\text{-}UnmatchedMaxPriority}(e)$, then there exists a maximal priority $p$ in $e$, such that $p$ has unmatched $\textit{put}$.
+\item[-] If $\mathsf{Has\text{-}UnmatchedMaxPriority}(e)$ holds, then there exists a maximal priority $p$ in $e$, such that $p$ has unmatched $\textit{put}$.
 Let $l = u \cdot \textit{put}(x,p) \cdot v$, where in $v$ there is no $\textit{put}$ with priority $p$.
-It is not hard to see that $\mathsf{UnmatchedMaxPriority\text{-}Seq}(l,x)$.
+It is not hard to see that $\mathsf{UnmatchedMaxPriority\text{-}Seq}(l,x)$ holds.
 Let $e'$ be the projection of $e$ into operations of $u \cdot v$.
@@
 
-\end{itemize}
-
-where $\mathsf{Has\text{-}\Gamma}(e)$ holds and .
-
-The proof of the $\textit{if}$ direction is quite similar to the proof of $\seqPQ_{\neq} \subseteq \textit{SeqPQ}_f$ direction of Lemma \ref{lemma:EPQ rules and semantics} and is omitted here.
-
-To prove the $\textit{only if}$ direction, given a data-differentiated execution $e$ and assume that $\textit{Check-PQ-Conc}(e)=\mathsf{true}$.
-
-$e \sqsubseteq \seqPQ$, according to the definition of $\textit{PQ}$, we can always find
-
+\item[-] If $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, then there exists a maximal priority $p$ in $e$, such that $p$ has only matched $\textit{put}$.
+Let $l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, where in $v \cdot w$ there is no $\textit{put}$ with priority $p$.
+It is not hard to see that $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds.
+Let $e'$ be the projection of $e$ into operations of $u \cdot v \cdot w$.
+It is easy to see that $e' \sqsubseteq u \cdot v \cdot w$ and then $e' \sqsubseteq \seqPQ$.
+Then we start the next round and choose $e'$ to be the ``$e$ in the next round''.
+\end{itemize}
+
+The loop terminates when at some round $e=\epsilon$. This process ensures that $\textit{Check-PQ-Conc}(e)=\mathsf{true}$.
+
+To prove the $\textit{only if}$ direction, given a data-differentiated execution $e$ and assume that $\textit{Check-PQ-}$ $\textit{Conc}(e)=\mathsf{true}$.
+Then we have $e \xrightarrow{\Gamma_1} e_1 \ldots \xrightarrow{\Gamma_m} e_m$ with $e_0=e$ and $e_m=\epsilon$.
+Assume that for each $i$, we found $\mathsf{\Gamma\text{-}Seq}(e_i,\alpha_i)$ holds for some $\alpha_i$.
+From $\mathsf{\Gamma\text{-}Seq}(e_{\textit{m-1}},\alpha_{\textit{m-1}})$ holds and $e_m = e_{\textit{m-1}} \setminus \alpha_{\textit{m-1}} \sqsubseteq \seqPQ$, by Lemma \ref{lemma:EPQ is step-by-step linearizability}, we can see that $e_{\textit{m-1}} \sqsubseteq \seqPQ$.
+Similarly, we can prove that $e_{\textit{m-2}},\ldots,e_0=e \sqsubseteq \seqPQ$.
 \qed
@@
 To prove the $\textit{only if}$ direction, assume that $e \sqsubseteq l \in \seqPQ$.
-Given $e' = e \vert_{D}$ and $l' = l \vert_{D}$, it is easy to see that $e' \sqsubseteq l'$, and by Lemma \ref{lemma:EPQ is closed under projection}, we can see that $l' \in \seqPQ$.
-Then by Lemma \ref{lemma:EPQ as multi in MRpri for sequence} we know that for each $R \in \textit{last}(l')$, we have $l' \in \textit{MS}(R)$.
-
-To prove the $\textit{if}$ direction, given $e_1 = e$, we generate sequence $e_2$ from $e_1$ as follows: Since $e_1 \in \textit{proj}(e)$, we know that for each $R_1 \in \textit{last}(e_1)$, we have $e_1 \sqsubseteq \textit{MS}(R_1)$.
-We choose an arbitrary $R_1$ in $\textit{last}(e_1)$,
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $R_1 = \seqPQ_3$: $e_2$ is generated from $e_1$ by erasing call and return of one $\textit{rm}(\textit{empty})$ operation.
-
-\item[-] Else, if $R_1 = \seqPQ_2^{=}, \seqPQ_2^{>}, \seqPQ_1^{=}, \seqPQ_1^{>}$, and the witness is $x$: $e_2$ is generated from $e_1$ by erasing call and return of method event of item $x$.
-\end{itemize}
-
-Similarly, for each $i > 1$, we obtain $e_{\textit{i+1}}$ from $e_i$, until we obtain $e_m = \epsilon$ for some $m$. It is obvious that $e_m \sqsubseteq \seqPQ$. For $e_{\textit{m-1}}$, since
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $e_m \sqsubseteq \seqPQ$,
-
-\item[-] If $\textit{last}(e_{\textit{m-1}}) = R_{\textit{m-1}} \in \{ \seqPQ_1^{>}, \seqPQ_1^{=}, \seqPQ_2^{>}, \seqPQ_2^{=} \}$ and $e_{\textit{m-1}}$ matches $R_{\textit{m-1}}$ with witness $x$: We already know that $e_{\textit{m-1}} \sqsubseteq \textit{MS}(R_{\textit{m-1}})$, $e_m = e_{\textit{m-1}} \setminus x \sqsubseteq \seqPQ$, and by step-by-step linearizability of $\seqPQ$ (Lemma \ref{lemma:EPQ is step-by-step linearizability}), we can see that $e_{\textit{m-1}} \sqsubseteq \seqPQ$.
-
-\item[-] If $\textit{last}(e_{\textit{m-1}}) = R_{\textit{m-1}} = \seqPQ_3$ and $o$ is a $\textit{rm}(\textit{empty})$ in $e_{\textit{m-1}}$ that is removed in the process of constructing $e_m$: We already know that $e_{\textit{m-1}} \sqsubseteq \textit{MS}(R_{\textit{m-1}})$, $e_m = e_{\textit{m-1}} \setminus o \sqsubseteq \seqPQ$, and by step-by-step linearizability of $\seqPQ$ (Lemma \ref{lemma:EPQ is step-by-step linearizability}), we can see that $e_{\textit{m-1}} \sqsubseteq \seqPQ$.
-\end{itemize}
-
-Therefore, we know that $e_{\textit{m-1}} \in \seqPQ$. Similarly, we can prove that $e_{\textit{m-2}},\ldots,e_1 = e \in \seqPQ$. \qed
-\end {proof}
-
-\subsection{Proof of Lemma \ref{lemma:EPQ as multi in MRpri for sequence}}
-
-%\EPQasMultiInMRpriforSequence*
-
-\begin{restatable}{lemma}{EPQasMultiInMRpriforSequence}
-\label{lemma:EPQ as multi in MRpri for sequence}
-Given a data-differentiated sequential execution $e$, $e \in \seqPQ$, if and only if, $\forall e' \in \textit{proj}(e)$ and $\forall R \in \textit{last}(e')$, we have $e' \in \textit{MS}(R)$.
-\end{restatable}
-
-\begin {proof}
-
-The \textit{only if} direction is obvious and can be similarly proved as the $\seqPQ_s \subseteq \seqPQ$ direction of Lemma \ref{lemma:EPQ rules and semantics}.
-
-To prove the $\textit{if}$ direction, we proceed as follows: From $e1=e$, we generate a sequence $e_2$ as follows:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $\textit{PQ}_3 \in \textit{last}(e_1)$: Then we can see that $e_1$ contains at least one $\textit{rm}(\textit{empty})$.
-$e_2$ is generated from $e_1$ by erasing one $\textit{rm}(\textit{empty})$.
-
-\item[-] Else, if $\textit{PQ}_2^{=} \in \textit{last}(e_1)$: Then we can see that one of the maximal priority of $e_1$ is unmatched $\textit{put}$ and (possibly) matched $\textit{put}$.
-Assume the set of the items of these unmatched $\textit{put}$ is $S$.
-$e_2$ is generated from $e_1$ by erasing one unmatched $\textit{put}$ which use the item last putted in $S$.
-
-\item[-] Else, if $\textit{PQ}_2^{>} \in \textit{last}(e_1)$: Then we can see that one of the maximal priority in $e_1$ is unmatched $\textit{put}$.
-Assume the set of the items of these unmatched $\textit{put}$ is $S$.
-$e_2$ is generated from $e_1$ by erasing one unmatched $\textit{put}$ which use the item last putted in $S$.
-
-\item[-] Else, if $\textit{PQ}_1^{=} \in \textit{last}(e_1)$: Then we can see that one of the maximal priority in $e_1$ is of more than one pair of matched $\textit{put}$.
-Assume the set of the items of these matched $\textit{put}$ is $S$.
-$e_2$ is generated from $e_1$ by erasing matched $\textit{put}$ and $\textit{rm}$ of the item which is last putted in $S$.
-
-\item[-] Else, if $\textit{PQ}_1^{>} \in \textit{last}(e_1)$: Then we can see that one of the maximal priority in $e_1$ is of one pair of matched $\textit{put}$.
-$e_2$ is generated from $e_1$ by erasing this pair of matched $\textit{put}$ and $\textit{rm}$.
-\end{itemize}
-
-Similarly, for each $i > 1$, we obtain $e_{\textit{i+1}}$ from $e_i$, until we obtain $e_m = \epsilon$ for some $m$. It is obvious that $e_m \in \seqPQ$. For $e_{\textit{m-1}}$, since
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $e_m \in \seqPQ$,
-
-\item[-] By assumption, we know that $e_{\textit{m-1}} \in \textit{MS}(R_{\textit{m-1}})$, where $R_{\textit{m-1}} \in \textit{last}(e_{\textit{m-1}})$.
-This implies that the guard of $R_{\textit{m-1}}$ is satisfied.
-\end{itemize}
-
-Therefore, we know that $e_{\textit{m-1}} \in \seqPQ$. Similarly, we can prove that $e_{\textit{m-2}},\ldots,e_1 = e \in \seqPQ$. \qed
+Given $e' = e \vert_{D}$ and $l' = l \vert_{D}$, it is easy to see that $e' \sqsubseteq l'$, and by Lemma \ref{lem:closure_proj}, we can see that $l' \in \seqPQ$.
+Then it is obvious that $\textit{Check-PQ-Conc-NonRec}(e')$ returns $\mathsf{true}$.
+
+To prove the $\textit{if}$ direction, assume that for each $e' \in \textit{proj}(e)$, $\textit{Check-PQ-Conc-NonRec}(e')$ returns $\mathsf{true}$.
+Then similarly as by Lemma \ref{lemma:con-check-EPQ is correct}, we can implies that $\textit{Check-PQ-Conc}(e)=\mathsf{true}$, and then by Lemma \ref{lemma:con-check-EPQ is correct}, we know that $e \sqsubseteq \seqPQ$.
+\qed
 \end {proof}
@@
 
-\EPQOneLarisCoRegular*
+%\EPQOneLarisCoRegular*
+\begin{restatable}{lemma}{EPQOneLarisCoRegular}
+\label{lemma:EPQ1Lar is co-regular}
+$\textit{Auts}_{\textit{1-lar}}$ is $\mathsf{MatchedMaxPriority}^{>}$-complete.
+\end{restatable}
 
@@
 
-\EPQOneEqualIsCoRegular*
+\begin{restatable}{lemma}{EPQOneEqualIsCoRegular}
+\label{lemma:EPQ1Equal is co-regular}
+$\mathsf{MatchedMaxPriority}^{=}$ is co-regular.
+\end{restatable}
 

### 2017-04-21 appendix-EPQ.tex
modify

@@
 
-To facilitate our proof of latter sections, we need to separate $\seqPQ_1$ into two cases: (1) no matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v \cdot w$ has priority $\textit{pri}$, (2) some matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v \cdot w$ has priority $\textit{pri}$.
-Therefore, we separate $\seqPQ_1$ into two cases:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\seqPQ_1^{>}(u,v,w,\textit{pri}) \equiv
-(\textit{noRE}(u \cdot v \cdot w)) \wedge
-(\textit{LI}(\textit{pri}, u \cdot v \cdot w)) \wedge
-(\textit{LI-U}(\textit{pri},u \cdot v \cdot w)) \wedge
-(\textit{matched-C}(u \cdot v,\textit{pri}) )$
-
-\item[-] $\seqPQ_1^{=}(u,v,w,\textit{pri}) \equiv
-(\textit{noRE}(u \cdot v \cdot w)) \wedge
-(\textit{LE'I}(\textit{pri}, u \cdot v \cdot w)) \wedge
-(\textit{LI-U}(\textit{pri},u \cdot v \cdot w)) \wedge
-(\textit{matched-C}(u \cdot v,\textit{pri}) ) \wedge
-(\textit{putInSeq}(u \cdot v \cdot w,u,\textit{pri}))$
-\end{itemize}
-
-For $\seqPQ_2$, we also need to distinguish two cases: (1) no matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ has priority $\textit{pri}$, (2) some matched pair of $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ has priority $\textit{pri}$.
-Therefore, we separate $\seqPQ_2$ into two rules:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\seqPQ_2^{>}(u,v) \equiv
-(\textit{noRE}(u \cdot v)) \wedge
-(\textit{LEI}(\textit{pri},u \cdot v)) \wedge
-(\textit{LI-M}(\textit{pri},u \cdot v)) \wedge
-(\textit{putInSeq}(u \cdot v,u,\textit{pri}))$.
-
-\item[-] $\seqPQ_2^{=}(u,v) \equiv
-(\textit{noRE}(u \cdot v)) \wedge
-(\textit{LEI}(\textit{pri},u \cdot v)) \wedge
-(\textit{LE'I-M}(\textit{pri},u \cdot v)) \wedge
-(\textit{putInSeq}(u \cdot v,u,\textit{pri}))$.
-\end{itemize}
-
-\subsection{Proofs and Definitions in Subsection \ref{subsec:definition of co-regular}}
-\label{sec:appendix proof and definition in section definition of co-regular}
-
-Before we go to investigate co-regular of each rules, we use the results in \cite{Bouajjani:2015} to simplify our work.
-\cite{Bouajjani:2015} states that checking linearizability w.r.t queue can be reduced into checking emptiness of intersection between $\mathcal{I}$ and a set of automata.
-Given a data-differentiated execution $e$, let $e \vert_{i}$ be an execution generated from $e$ by erasing call and return actions of items that does not use priority $i$ (does not influence $\textit{rm}(\textit{empty})$).
+To facilitate our proof, we consider two cases of $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ depending on whether $e$ contains exactly one value with priority $p$ or at least two values.
+We denote by $\mathsf{MatchedMax-}$ $\mathsf{Priority}^{>}(e,x)$ the strengthening of $\mathsf{MatchedMaxPriority}(e,x)$ with the condition that all the values other than $x$ have a priority strictly smaller than $p$, and by $\mathsf{MatchedMaxPriority}^{=}(e,x)$ the strengthening of the same formula with the negation of this condition.
+
+Similarly, we consider two cases of $\mathsf{UnmatchedMaxPriority\text{-}Seq}(e,x)$ depending on whether $e$ some values of priority $p$ has matched $\textit{put}$.
+We denote by $\mathsf{UnmatchedMaxPriority}^{>}(e,x)$ the strengthening of $\mathsf{UnmatchedMaxPriority}(e,x)$ with the condition that non value of priority $p$ has matched $\textit{put}$, and by $\mathsf{UnmatchedMaxPriority}^{=}(e,x)$ the strengthening of the same formula with the negation of this condition.
+
+\subsection{Lemma and Register Automata For FIFO of Single-Priority Executions}
+\label{sec:appendix lemma and register automata for FIFO of single-priority executions}
+
+Before we go to investigate $\Gamma$-linearizable, we can use the result in \cite{DBLP:conf/icalp/BouajjaniEEH15} to simplify our work.
+\cite{DBLP:conf/icalp/BouajjaniEEH15} states that checking linearizability w.r.t queue can be reduced into checking emptiness of intersection between $\mathcal{I}$ and a set of automata.
+Given a data-differentiated execution $e$, let $e \vert_{i}$ be an execution generated from $e$ by erasing call and return actions of values that does not use priority $i$ (does not influence $\textit{rm}(\textit{empty})$).
 We call a extended priority queue execution with only one priority a single-priority execution.
@@
 
-Given an execution of queue, we say that it is differentiated \cite{Wolper:1986}, if each item is enqueued at most once.
-\cite{Bouajjani:2015} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $a,b$: (1) $\textit{deq}(b) <_{hb} \textit{enq}(b)$, (2) there are no $\textit{enq}(b)$ and at least one $\textit{deq}(b)$, (3) there are one $\textit{enq}(b)$ and more than one $\textit{deq}(b)$, and (4) $\textit{enq}(a) <_{\textit{hb}} \textit{enq}(b)$, and $\textit{deq}(b) <_{\textit{hb}} \textit{deq}(a)$, or $\textit{deq}(a)$ does not exists.
-For each such case, we can construct a witness automata for extended priority queue.
-
-\cite{Bouajjani:2015} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $a,b$: (1) $\textit{deq}(b) <_{hb} \textit{enq}(b)$, (2) there are are no $\textit{enq}(b)$ and at least one $\textit{deq}(b)$, (3) there are are one $\textit{enq}(b)$ and more than one $\textit{deq}(b)$, and (4) $\textit{enq}(a) <_{\textit{hb}} \textit{enq}(b)$, and $\textit{deq}(b) <_{\textit{hb}} \textit{deq}(a)$, or $\textit{deq}(a)$ does not exists.
-
-For each such case, we construct a witness automata.
+Given an execution of queue, we say that it is differentiated, if each item is enqueued at most once.
+\cite{DBLP:conf/icalp/BouajjaniEEH15} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $a,b$: (1) $\textit{deq}(b) <_{hb} \textit{enq}(b)$, (2) there are no $\textit{enq}(b)$ and at least one $\textit{deq}(b)$, (3) there are one $\textit{enq}(b)$ and more than one $\textit{deq}(b)$, and (4) $\textit{enq}(a) <_{\textit{hb}} \textit{enq}(b)$, and $\textit{deq}(b) <_{\textit{hb}} \textit{deq}(a)$, or $\textit{deq}(a)$ does not exists.
+For each such case, we can construct a register automata for extended priority queue.
+
 We generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}.
-Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri})$, $\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},b),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$.
+Here $C_1 = \{ \textit{call}(\textit{put},a,\top)$, $\textit{ret}(\textit{put},a,\top), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},b),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2$ $= C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{put},b,\top) \}$.
 
@@
 We generate witness automata $\mathcal{A}_{\textit{SinPri}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata for FIFO-2}.
-Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_2 = c_1 + \textit{cal}(\textit{rm},b) + \textit{ret}(\textit{rm},b)$.
+Here $C_1 = \{ \textit{call}(\textit{put},a,\top),\textit{ret}(\textit{put},a,\top), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{cal}(\textit{rm},b) + \textit{ret}(\textit{rm},b) \}$.
 

### 2017-04-21 appendix-EPQ.tex
modify

@@
 
-We generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}.
-Here $C_1 = \{ \textit{call}(\textit{put},a,\top)$, $\textit{ret}(\textit{put},a,\top), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},b),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2$ $= C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{put},b,\top) \}$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.6 \textwidth]{figures/PIC_AUTO_FIFO_1.pdf}
+We generate register automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}.
+Here $C_1 = \{ \textit{call}(\textit{put},a,\top)$, $\textit{ret}(\textit{put},a,\top), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},b),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2$ $= C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{put},b,\top) \}$.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.5 \textwidth]{figures/PIC_AUTO_FIFO_1.pdf}
 %\vspace{-10pt}
@@
 
-We generate witness automata $\mathcal{A}_{\textit{SinPri}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata for FIFO-2}.
-Here $C_1 = \{ \textit{call}(\textit{put},a,\top),\textit{ret}(\textit{put},a,\top), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{cal}(\textit{rm},b) + \textit{ret}(\textit{rm},b) \}$.
+We generate register automata $\mathcal{A}_{\textit{SinPri}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata for FIFO-2}.
+Here $C_1 = \{ \textit{call}(\textit{put},a,\top),\textit{ret}(\textit{put},a,\top), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{call}(\textit{rm},b) + \textit{ret}(\textit{rm},b) \}$.
 
@@
 
-We generate witness automata $\mathcal{A}_{\textit{SinPri}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata for FIFO-3}.
-Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},b)$, $c_4 = c_3 + \textit{cal}(\textit{rm},b)$, $c_5 = c_1 + \textit{ret}(\textit{rm},b)$, $c_6 = c_5 + \textit{cal}(\textit{rm},b)$.
+We generate register automata $\mathcal{A}_{\textit{SinPri}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata for FIFO-3}.
+Here $C_1 = \{ \textit{call}(\textit{put},a,\top),\textit{ret}(\textit{put},a,\top), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,\top) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_4 = C_3 \cup \{ \textit{call}(\textit{rm},b) \}$, $C_5 = C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_6 = C_5 \cup \{ \textit{call}(\textit{rm},b) \}$.
 
@@
 
-We generate witness automata $\mathcal{A}_{\textit{SinPri}}^4$ for the forth case, and it is shown in \figurename~\ref{fig:automata for FIFO-4}.
-Here $c_1 = c + \textit{cal}(\textit{rm},b)$, and $c_2 = c + \textit{ret}(\textit{put},b) + \textit{cal}(\textit{rm},a) + \textit{ret}(\textit{rm},a)$, where $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d),\textit{ret}(\textit{rm},d),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.
+We generate register automata $\mathcal{A}_{\textit{SinPri}}^4$ for the forth case, and it is shown in \figurename~\ref{fig:automata for FIFO-4}.
+Here $C_1 = C \cup \{ \textit{call}(\textit{rm},b) \}$, and $C_2 = C \cup \{ \textit{ret}(\textit{put},b,\top), \textit{call}(\textit{rm},a), \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},d,\top),\textit{ret}(\textit{put},d,\top), \textit{call}(\textit{rm},d),\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.
 
@@
 
-Let $\textit{Auts}_{\textit{sinPri}} = \{ \mathcal{A}_{\textit{SinPri}}^1, \mathcal{A}_{\textit{SinPri}}^2, \mathcal{A}_{\textit{SinPri}}^3, \mathcal{A}_{\textit{SinPri}}^4 \}$.
+Let $\mathcal{A}_{\textit{sinPri}}$ be the union of $\mathcal{A}_{\textit{SinPri}}^1, \mathcal{A}_{\textit{SinPri}}^2, \mathcal{A}_{\textit{SinPri}}^3, \mathcal{A}_{\textit{SinPri}}^4$.
 Let us prove Lemma \ref{lemma:automata for extended priority queue with single priority}.
@@
 \begin{lemma}\label{lemma:automata for extended priority queue with single priority}
-Given a data-independent implementations $\mathcal{I}$ of extended priority queue, $\mathcal{I} \cap \textit{Auts}_{\textit{sinPri}} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{toQueue}(e')$ does not linearizable to queue.
+Given a data-independent implementations $\mathcal{I}$ of extended priority queue, $\mathcal{I} \cap \mathcal{A}_{\textit{sinPri}} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{transToQueue}(e')$ does not linearizable to queue.
 \end{lemma}
@@
 
-\cite{Bouajjani:2015} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $v_a,v_b$: (1) $\textit{deq}(v_b) <_{hb} \textit{enq}(v_b)$, (2) there are are no $\textit{enq}(v_b)$ and at least one $\textit{deq}(v_b)$, (3) there are are one $\textit{enq}(v_b)$ and more than one $\textit{deq}(v_b)$, and (4) $\textit{enq}(v_a) <_{\textit{hb}} \textit{enq}(v_b)$, and $\textit{deq}(v_b) <_{\textit{hb}} \textit{deq}(v_a)$, or $\textit{deq}(v_a)$ does not exists.
+\cite{DBLP:conf/icalp/BouajjaniEEH15} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $v_a,v_b$: (1) $\textit{deq}(v_b) <_{hb} \textit{enq}(v_b)$, (2) there are are no $\textit{enq}(v_b)$ and at least one $\textit{deq}(v_b)$, (3) there are are one $\textit{enq}(v_b)$ and more than one $\textit{deq}(v_b)$, and (4) $\textit{enq}(v_a) <_{\textit{hb}} \textit{enq}(v_b)$, and $\textit{deq}(v_b) <_{\textit{hb}} \textit{deq}(v_a)$, or $\textit{deq}(v_a)$ does not exists.
 
 Let us prove the $\textit{only if}$ direction.
-Assume that there exists execution $e_0 \in \mathcal{I}$ and $e_0$ is accepted by an automaton in $\textit{Auts}_{\textit{sinPri}}$.
+Assume that there exists execution $e_0 \in \mathcal{I}$ and $e_0$ is accepted by $\mathcal{A}_{\textit{SinPri}}^1, \mathcal{A}_{\textit{SinPri}}^2, \mathcal{A}_{\textit{SinPri}}^3$ or $\mathcal{A}_{\textit{SinPri}}^4$.
 By data-independence, we can see that there exists a data-differentiated $e \in \mathcal{I}$ and renaming function, such that $e_0=r(e)$.
@@
 
-Then it is not hard to see that $r(e) \in \mathcal{I}$ and it is accepted by some automaton in $\textit{Auts}_{\textit{sinPri}}$. This completes the proof of this lemma. \qed
+Then it is not hard to see that $r(e) \in \mathcal{I}$ and it is accepted by $\mathcal{A}_{\textit{SinPri}}^1, \mathcal{A}_{\textit{SinPri}}^2, \mathcal{A}_{\textit{SinPri}}^3$ or $\mathcal{A}_{\textit{SinPri}}^4$.
+This completes the proof of this lemma.
+\qed
 \end {proof}
@@
 
-Let us begin to represent witness automata that is used for capture the existence of a data-differentiated execution $e$, $e$ has a $\_$-projection $e'$, $\textit{last}(e') = \textit{PQ}_1^{>}$, and there exists a cycle going through the item with maximal priority in $e'$.
+Let us begin to represent register automata that is used for capture the existence of a data-differentiated execution $e$, $e$ has a $\_$-projection $e'$, $\textit{last}(e') = \textit{PQ}_1^{>}$, and there exists a cycle going through the item with maximal priority in $e'$.
 By data-independence, we can obtain $e_r$ from $e$ by renaming function, which maps such item to be $b$, maps items that cover it to be $a$, and maps other items into $d$.
 There are four possible enumeration of call and return actions of $\textit{put}(b)$ and $\textit{rm}(b)$.
-For each of them, we generate a witness automaton.
-
-For the case when $e_r \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^1$, as shown in \figurename~\ref{fig:automata APQ1Lar-1}.
-Here $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{cal}(\textit{put},a,\textit{les}_p)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.
+For each of them, we generate a register automaton.
+
+For the case when $e_r \vert_{b} = \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^1$, as shown in \figurename~\ref{fig:automata APQ1Lar-1}.
+Here $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{call}(\textit{put},a,\textit{les}_p)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{call}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.
 The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^1$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.
@@
 
-For the case when $e_r \vert_{b} = \textit{cal}(\textit{put},b,p) \cdot \textit{cal}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^2$, as shown in \figurename~\ref{fig:automata APQ1Lar-2}.
+For the case when $e_r \vert_{b} = \textit{call}(\textit{put},b,p) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^2$, as shown in \figurename~\ref{fig:automata APQ1Lar-2}.
 Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$.
@@
 
-For the case when $e_r \vert_{b} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^3$, as shown in \figurename~\ref{fig:automata APQ1Lar-3}.
+For the case when $e_r \vert_{b} = \textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^3$, as shown in \figurename~\ref{fig:automata APQ1Lar-3}.
 Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$.
@@
 
-For the case when $e_r \vert_{b} = \textit{cal}(\textit{rm}) \cdot \textit{cal}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b) \cdot \textit{ret}(\textit{put})$, we generate witness automaton $\mathcal{A}_{\textit{l-lar}}^4$, as shown in \figurename~\ref{fig:automata APQ1Lar-4}.
+For the case when $e_r \vert_{b} = \textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b) \cdot \textit{ret}(\textit{put})$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^4$, as shown in \figurename~\ref{fig:automata APQ1Lar-4}.
 Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$, and $c_4 = c_1 + \textit{ret}(\textit{put},b)$.
@@
 Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{=}$.
-If there exists an item $x$ with priority $\textit{pri}$, such that for each $y \in \textit{Items}(e,\textit{pri})$, (1) $x$ does not $<_{\textit{pb}}$ to $y$, and (2) the right-most gap-point of $x$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},y)$.
+If there exists an item $x$ with priority $\textit{pri}$, such that for each $y \in \textit{Items}(e,\textit{pri})$, (1) $x$ does not $<_{\textit{pb}}$ to $y$, and (2) the right-most gap-point of $x$ is after $\textit{call}(\textit{put},y,\_)$ and $\textit{call}(\textit{rm},y)$.
 Then $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
@@
 \item[-] For item $z$ with priority smaller than $\textit{pri}$.
-If both $\textit{cal}(\textit{put},z,\_)$ and $\textit{cal}(\textit{rm},z)$ is before $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just after their call actions.
+If both $\textit{call}(\textit{put},z,\_)$ and $\textit{call}(\textit{rm},z)$ is before $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just after their call actions.
 If both $\textit{ret}(\textit{put},z)$ and $\textit{ret}(\textit{rm},z)$ (if exists) is after $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just before their return actions.
@@
 It is obvious that $e \sqsubseteq l$.
-Since for each $y \in \textit{Items}(e,\textit{pri})$, $o$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},x)$, we can see that $\textit{rm}(x)$ is after $\textit{put}(y,\textit{pri})$ and $\textit{rm}(y)$ in $l$.
+Since for each $y \in \textit{Items}(e,\textit{pri})$, $o$ is after $\textit{call}(\textit{put},y,\_)$ and $\textit{call}(\textit{rm},x)$, we can see that $\textit{rm}(x)$ is after $\textit{put}(y,\textit{pri})$ and $\textit{rm}(y)$ in $l$.
 It is obvious that $\textit{put}(x,\textit{pri})$ is before $\textit{rm}(x)$ in $l$.
@@
 Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $x$, then there exists unmatched method event before $\textit{rm}(x)$ with smaller priority.
-Since the rightmost gap-point of $x$ is before $\textit{cal}(\textit{put},y,\textit{pri})$ or $\textit{cal}(\textit{rm},y)$, if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $x$, then $\textit{rm}(x)$ will be before $\textit{cal}(\textit{put},y,\textit{pri})$ or $\textit{cal}(\textit{rm},x)$.
+Since the rightmost gap-point of $x$ is before $\textit{call}(\textit{put},y,\textit{pri})$ or $\textit{call}(\textit{rm},y)$, if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $x$, then $\textit{rm}(x)$ will be before $\textit{call}(\textit{put},y,\textit{pri})$ or $\textit{call}(\textit{rm},x)$.
 
@@
 To prove the $\textit{only if}$ direction, we prove its contrapositive.
-Assume we already know that for each $x$ and $y$ has maximal priority in $e$, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{cal}(\textit{put},y,\textit{pri})$ and $\textit{cal}(\textit{rm},x)$.
+Assume we already know that for each $x$ and $y$ has maximal priority in $e$, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{call}(\textit{put},y,\textit{pri})$ and $\textit{call}(\textit{rm},x)$.
 We need to prove that $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
@@
 
-    Step $1$: Check whether for each $b \in \textit{Items}(e,\textit{pri})$, (1) $a_1$ does not $<_{\textit{pb}}$ to $b$, and (2) the right-most gap-point of $a$ is after $\textit{cal}(\textit{put},b,\textit{pri})$ and $\textit{cal}(\textit{rm},b)$.
+    Step $1$: Check whether for each $b \in \textit{Items}(e,\textit{pri})$, (1) $a_1$ does not $<_{\textit{pb}}$ to $b$, and (2) the right-most gap-point of $a$ is after $\textit{call}(\textit{put},b,\textit{pri})$ and $\textit{call}(\textit{rm},b)$.
 
     It is easy to see that $a_1$ is of priority $\textit{pri}$, and $a_1$ does not $<_{\textit{pb}}$ to any $b \in \textit{Items}(e,\textit{pri})$.
-If for each $b \in \textit{Items}(e,\textit{pri})$, the rightmost gap-point of $a_1$ is after $\textit{cal}(\textit{put},b,\textit{pri})$ and $\textit{cal}(\textit{rm},b)$.
+If for each $b \in \textit{Items}(e,\textit{pri})$, the rightmost gap-point of $a_1$ is after $\textit{call}(\textit{put},b,\textit{pri})$ and $\textit{call}(\textit{rm},b)$.
 Then by Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, we can obtain that $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
 
-\item[-] Otherwise, there exists $a_2 \in \textit{Items}(e,\textit{pri})$, such that the rightmost gap-point of $a_1$ is before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$ in $e$.
+\item[-] Otherwise, there exists $a_2 \in \textit{Items}(e,\textit{pri})$, such that the rightmost gap-point of $a_1$ is before $\textit{call}(\textit{put},a_2,\textit{pri})$ or $\textit{call}(\textit{rm},a_2)$ in $e$.
 We can see that each gap-point of $a_2$ is after the rightmost gap-point of $a_1$.%, and thus, the right-most gap-point of $a_2$ is after the rightmost gap-point of $a_1$.
@@
 
-    By assumption,we know that the rightmost gap-point of $a_3$ is after $\textit{cal}(\textit{put},a_2,\textit{pri})$ and $\textit{cal}(\textit{rm},a_2)$.
+    By assumption,we know that the rightmost gap-point of $a_3$ is after $\textit{call}(\textit{put},a_2,\textit{pri})$ and $\textit{call}(\textit{rm},a_2)$.
 Therefore, we can see that the rightmost gap-point of $a_3$ is after the rightmost gap-point of $a_1$.
@@
 
-    If $c_3 <_{\textit{pb}}^C c_2 <_{\textit{pb}}^C c_1$: Since $c_2 <_{\textit{pb}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
-Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.
+    If $c_3 <_{\textit{pb}}^C c_2 <_{\textit{pb}}^C c_1$: Since $c_2 <_{\textit{pb}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{call}(\textit{put},c_1,\_)$.
+Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{call}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
+Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{call}(\textit{put},c_2,\_)$.
+Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{call}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.
 
@@
     \item[-] ($<_{\textit{pb}}^A \cdot <_{\textit{pb}}^C$): If $c_3 <_{\textit{pb}}^A c_2 <_{\textit{pb}}^C c_1$.
-Since $c_2 <_{\textit{pb}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{pb}}^A c_2$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
-Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^A c_1$.
+Since $c_2 <_{\textit{pb}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{call}(\textit{put},c_1,\_)$.
+Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{call}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
+Since $c_3 <_{\textit{pb}}^A c_2$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{call}(\textit{put},c_2,\_)$.
+Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{call}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^A c_1$.
 
     \item[-] ($<_{\textit{pb}}^C \cdot <_{\textit{pb}}^A$): If $c_3 <_{\textit{pb}}^C c_2 <_{\textit{pb}}^A c_1$.
-Since $c_2 <_{\textit{pb}}^A c_1$, $\textit{ret}(\textit{put},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-It is obvious that $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{put},c_2)$.
-Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
-Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.
+Since $c_2 <_{\textit{pb}}^A c_1$, $\textit{ret}(\textit{put},c_2)$ is before $\textit{call}(\textit{put},c_1,\_)$.
+It is obvious that $\textit{call}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{put},c_2)$.
+Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{call}(\textit{put},c_2,\_)$.
+Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{call}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.
 
@@
     \item[-] ($<_{\textit{pb}}^B \cdot <_{\textit{pb}}^C$): If $c_3 <_{\textit{pb}}^B c_2 <_{\textit{pb}}^C c_1$.
-Since $c_2 <_{\textit{pb}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
-It is obvious that $\textit{cal}(\textit{rm},c_2)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{pb}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$.
-Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.
+Since $c_2 <_{\textit{pb}}^C c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{call}(\textit{put},c_1,\_)$.
+It is obvious that $\textit{call}(\textit{rm},c_2)$ is before $\textit{ret}(\textit{rm},c_2)$.
+Since $c_3 <_{\textit{pb}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{call}(\textit{rm},c_2)$.
+Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{call}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.
 
     \item[-] ($<_{\textit{pb}}^C \cdot <_{\textit{pb}}^B$): If $c_3 <_{\textit{pb}}^C c_2 <_{\textit{pb}}^B c_1$.
-Since $c_2 <_{\textit{pb}}^B c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{rm},c_1)$.
-Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{cal}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_2,\_)$.
-Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_1)$, and $c_3 <_{\textit{pb}}^B c_1$.
+Since $c_2 <_{\textit{pb}}^B c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{call}(\textit{rm},c_1)$.
+Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{call}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
+Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{call}(\textit{put},c_2,\_)$.
+Therefore, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{call}(\textit{rm},c_1)$, and $c_3 <_{\textit{pb}}^B c_1$.
     \end{itemize}
@@
     \setlength{\itemsep}{0.5pt}
-    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{put},c_1,\_)$: Since $c_3 <_{\textit{pb}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{rm},c_2)$.
-Then $\textit{ret}(\textit{rm},c_3)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.
+    \item[-] If $\textit{call}(\textit{rm},c_2)$ is before $\textit{call}(\textit{put},c_1,\_)$: Since $c_3 <_{\textit{pb}}^B c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{call}(\textit{rm},c_2)$.
+Then $\textit{ret}(\textit{rm},c_3)$ is before $\textit{call}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^C c_1$.
 This implies that $c_4 <_{\textit{pb}}^A c_3 <_{\textit{pb}}^C c_1$.
@@
 
-    \item[-] If $\textit{cal}(\textit{rm},c_2)$ is after $\textit{cal}(\textit{put},c_1,\_)$: Since $c_2 <_{\textit{pb}}^A c_1$, $\textit{ret}(\textit{put},c_2,\_)$ is before $\textit{cal}(\textit{put},c_1,\_)$.
+    \item[-] If $\textit{call}(\textit{rm},c_2)$ is after $\textit{call}(\textit{put},c_1,\_)$: Since $c_2 <_{\textit{pb}}^A c_1$, $\textit{ret}(\textit{put},c_2,\_)$ is before $\textit{call}(\textit{put},c_1,\_)$.
 Since $c_3 <_{\textit{pb}}^B c_2$, $\textit{rm}(c_3)$ happens before $\textit{rm}(c_2)$, and then we know that $\textit{put}(c_2,\_)$ can not happen before $\textit{put}(c_3,\_)$.
-Since $\textit{put}(c_2,\_)$ does not happen before $\textit{put}(c_3,\_)$, $\textit{cal}(\textit{put},c_3,\_)$ is before $\textit{ret}(\textit{put},c_2,\_)$.
-Since $c_4 <_{\textit{pb}}^A c_3$, $\textit{ret}(\textit{put},c_4)$ is before $\textit{cal}(\textit{put},c_3,\_)$.
-Therefore, $\textit{ret}(\textit{put},c_4)$ is before $\textit{cal}(\textit{put},c_1,\_)$, and $c_4 <_{\textit{pb}}^A c_1$.
-    \end{itemize}
-
-\item[-]  ($<_{\textit{pb}}^B \cdot <_{\textit{pb}}^A \cdot <_{\textit{pb}}^B$): If $c_4 <_{\textit{pb}}^B c_3 <_{\textit{pb}}^A c_2 <_{\textit{pb}}^B c_1$: Since $c_2 <_{\textit{pb}}^B c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{cal}(\textit{rm},c_1)$.
+Since $\textit{put}(c_2,\_)$ does not happen before $\textit{put}(c_3,\_)$, $\textit{call}(\textit{put},c_3,\_)$ is before $\textit{ret}(\textit{put},c_2,\_)$.
+Since $c_4 <_{\textit{pb}}^A c_3$, $\textit{ret}(\textit{put},c_4)$ is before $\textit{call}(\textit{put},c_3,\_)$.
+Therefore, $\textit{ret}(\textit{put},c_4)$ is before $\textit{call}(\textit{put},c_1,\_)$, and $c_4 <_{\textit{pb}}^A c_1$.
+    \end{itemize}
+
+\item[-]  ($<_{\textit{pb}}^B \cdot <_{\textit{pb}}^A \cdot <_{\textit{pb}}^B$): If $c_4 <_{\textit{pb}}^B c_3 <_{\textit{pb}}^A c_2 <_{\textit{pb}}^B c_1$: Since $c_2 <_{\textit{pb}}^B c_1$, $\textit{ret}(\textit{rm},c_2)$ is before $\textit{call}(\textit{rm},c_1)$.
 Since $c_3 <_{\textit{pb}}^A c_2$, we can see that $\textit{put}(c_3,\_) <_{\textit{hb}} \textit{put}(c_2,\_)$.
-Since each single-priority execution has FIFO property, we know that $\textit{rm}(c_2)$ does not happen before $\textit{rm}(c_3)$, and thus, $\textit{cal}(\textit{rm},c_3)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_4 <_{\textit{pb}}^B c_3$, $\textit{ret}(\textit{rm},c_4)$ is before $\textit{cal}(\textit{rm},c_3)$.
-Therefore, $\textit{ret}(\textit{rm},c_4)$ is before $\textit{cal}(\textit{rm},c_1)$, and $c_4 <_{\textit{pb}}^B c_1$.
+Since each single-priority execution has FIFO property, we know that $\textit{rm}(c_2)$ does not happen before $\textit{rm}(c_3)$, and thus, $\textit{call}(\textit{rm},c_3)$ is before $\textit{ret}(\textit{rm},c_2)$.
+Since $c_4 <_{\textit{pb}}^B c_3$, $\textit{ret}(\textit{rm},c_4)$ is before $\textit{call}(\textit{rm},c_3)$.
+Therefore, $\textit{ret}(\textit{rm},c_4)$ is before $\textit{call}(\textit{rm},c_1)$, and $c_4 <_{\textit{pb}}^B c_1$.
 
@@
 Let $a$ and $b$ be items with maximal priority $\textit{pri}$.
-Assume that $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\textit{pri})$ or $\textit{cal}(\textit{rm},a)$.
+Assume that $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,\textit{pri})$ or $\textit{call}(\textit{rm},a)$.
 Then, there are five possible enumeration of method events of $a$, $b$, $a_1$ (if exists), where $a_1$ is the possible intermediate items for obtain $a <_{\textit{pb}}^* b$.
@@
     Since $a <_{\textit{pb}}^A b$, $\textit{put}(a,\textit{pri}) <_{\textit{hb}} \textit{put}(b,\textit{pri})$.
-Since $o$ is after $\textit{cal}(\textit{put},b,\textit{pri})$, and thus, after $\textit{cal}(\textit{put},a,\textit{pri})$, we can see that $o$ is before $\textit{cal}(\textit{rm},b)$.
-
-    Since single-priority execution must satisfy the FIFO property, $\textit{rm}(b)$ does not happen before $\textit{rm}(a)$, and thus, $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
-If $\textit{cal}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$, then $o$ is also a gap-point of $a$ and contradicts our assumption.
-So we know that $\textit{cal}(\textit{rm},a)$ is after $\textit{cal}(\textit{rm},b)$.
+Since $o$ is after $\textit{call}(\textit{put},b,\textit{pri})$, and thus, after $\textit{call}(\textit{put},a,\textit{pri})$, we can see that $o$ is before $\textit{call}(\textit{rm},b)$.
+
+    Since single-priority execution must satisfy the FIFO property, $\textit{rm}(b)$ does not happen before $\textit{rm}(a)$, and thus, $\textit{call}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
+If $\textit{call}(\textit{rm},a)$ is before $\textit{call}(\textit{rm},b)$, then $o$ is also a gap-point of $a$ and contradicts our assumption.
+So we know that $\textit{call}(\textit{rm},a)$ is after $\textit{call}(\textit{rm},b)$.
 If $\textit{ret}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$, since we already assume that there exists gap-point of $a$, this gap-point is also a gap-point of $b$, and is after $o$, which contradicts that $o$ is the rightmost gap-point of $b$.
@@
 
-    Since $a <_{\textit{pb}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},b)$.
-Since $o$ is after $\textit{cal}(\textit{rm},b)$, we can see that $o$ is before $\textit{cal}(\textit{put},a,\textit{pri})$.
-This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\textit{pri})$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible.
+    Since $a <_{\textit{pb}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{call}(\textit{rm},b)$.
+Since $o$ is after $\textit{call}(\textit{rm},b)$, we can see that $o$ is before $\textit{call}(\textit{put},a,\textit{pri})$.
+This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{call}(\textit{put},a,\textit{pri})$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible.
 Therefore, we can safely ignore this reason.
@@
 
-    Since $a <_{\textit{pb}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},b,\textit{pri})$.
-Since $o$ is after $\textit{cal}(\textit{put},b)$, we can see that $o$ is before $\textit{cal}(\textit{put},a,\textit{pri})$.
-This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{put},a,\textit{pri})$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible.
+    Since $a <_{\textit{pb}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{call}(\textit{put},b,\textit{pri})$.
+Since $o$ is after $\textit{call}(\textit{put},b)$, we can see that $o$ is before $\textit{call}(\textit{put},a,\textit{pri})$.
+This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{call}(\textit{put},a,\textit{pri})$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible.
 Therefore, we can safely ignore this reason.
@@
 
-    Since $a_1 <_{\textit{pb}}^B b$, $\textit{rm}(a_1) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_1)$ is before $\textit{cal}(\textit{rm},b)$.
-Since $\textit{rm}(a_1)$ does not happen before $\textit{put}(a_1)$, $\textit{cal}(\textit{put},a_1,\textit{pri})$ is before $\textit{ret}(\textit{rm},a_1)$.
-Since $a <_{\textit{pb}}^A a_1$, $\textit{ret}(\textit{put},a,\textit{pri})$ is before $\textit{cal}(\textit{put},a_1,\textit{pri})$.
-Therefore, $\textit{ret}(\textit{put},a,\textit{pri})$ is before $\textit{cal}(\textit{rm},b)$.
-Since $\textit{cal}(\textit{rm},b)$ is before $o$, we can see that $o$ is before $\textit{cal}(\textit{rm},a)$.
-
-    If $\textit{cal}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$, then $e \vert_{ \{ a,a_1,b \} }$ violates the FIFO property.
-Therefore, $\textit{cal}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
+    Since $a_1 <_{\textit{pb}}^B b$, $\textit{rm}(a_1) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_1)$ is before $\textit{call}(\textit{rm},b)$.
+Since $\textit{rm}(a_1)$ does not happen before $\textit{put}(a_1)$, $\textit{call}(\textit{put},a_1,\textit{pri})$ is before $\textit{ret}(\textit{rm},a_1)$.
+Since $a <_{\textit{pb}}^A a_1$, $\textit{ret}(\textit{put},a,\textit{pri})$ is before $\textit{call}(\textit{put},a_1,\textit{pri})$.
+Therefore, $\textit{ret}(\textit{put},a,\textit{pri})$ is before $\textit{call}(\textit{rm},b)$.
+Since $\textit{call}(\textit{rm},b)$ is before $o$, we can see that $o$ is before $\textit{call}(\textit{rm},a)$.
+
+    If $\textit{call}(\textit{rm},a)$ is after $\textit{ret}(\textit{rm},b)$, then $e \vert_{ \{ a,a_1,b \} }$ violates the FIFO property.
+Therefore, $\textit{call}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
 Similarly as the case of reason $1$, we can see that $\textit{ret}(\textit{rm},b)$ is before $\textit{ret}(\textit{rm},a)$.
@@
 Here we explicitly draw the leftmost gap-point of $a$ as $o'$.
-Since the position of $\textit{ret}(\textit{put},a_1,\textit{pri})$ and $\textit{cal}(\textit{put},a,\textit{pri})$ do not influence the correctness, we can simply ignore it.
-We also ignore $\textit{cal}(\textit{put},b,\textit{pri})$ and $\textit{ret}(\textit{put},b)$, since the only requirements of them are (1) $\textit{rm}(b)$ does not happen before $\textit{put}(b)$ and (2) $\textit{cal}(\textit{put},b,\textit{pri})$ is before $o$.
+Since the position of $\textit{ret}(\textit{put},a_1,\textit{pri})$ and $\textit{call}(\textit{put},a,\textit{pri})$ do not influence the correctness, we can simply ignore it.
+We also ignore $\textit{call}(\textit{put},b,\textit{pri})$ and $\textit{ret}(\textit{put},b)$, since the only requirements of them are (1) $\textit{rm}(b)$ does not happen before $\textit{put}(b)$ and (2) $\textit{call}(\textit{put},b,\textit{pri})$ is before $o$.
 
@@
     \setlength{\itemsep}{0.5pt}
-    \item[-] If $o$ is before $\textit{cal}(\textit{rm},a)$: Then $o$ is obviously before $\textit{ret}(\textit{rm},a)$.
-Since $a <_{\textit{pb}}^B a_1$, $\textit{ret}(\textit{rm},a)$ is before $\textit{cal}(\textit{rm},a_1)$.
-Then we can see that, $o$ is before $\textit{cal}(\textit{rm},a_1)$, and remember that $a_1 <_{\textit{pb}}^A b$.
+    \item[-] If $o$ is before $\textit{call}(\textit{rm},a)$: Then $o$ is obviously before $\textit{ret}(\textit{rm},a)$.
+Since $a <_{\textit{pb}}^B a_1$, $\textit{ret}(\textit{rm},a)$ is before $\textit{call}(\textit{rm},a_1)$.
+Then we can see that, $o$ is before $\textit{call}(\textit{rm},a_1)$, and remember that $a_1 <_{\textit{pb}}^A b$.
 Then we can goto the case of reason $1$ and treat $a_1$ as $a$.
@@
 
-    \item[-] If $o$ is before $\textit{cal}(\textit{put},a,\textit{pri})$: Since $\textit{rm}(a)$ does not happen before $\textit{put}(a,\textit{pri})$, we can see that $\textit{cal}(\textit{put},a,\textit{pri})$ is before $\textit{ret}(\textit{rm},a)$, and then $o$ is before $\textit{ret}(\textit{rm},a)$.
-Then similarly as above case, we can see that $o$ is before $\textit{cal}(\textit{rm},a_1)$, and $a_1 <_{\textit{pb}}^A b$.
+    \item[-] If $o$ is before $\textit{call}(\textit{put},a,\textit{pri})$: Since $\textit{rm}(a)$ does not happen before $\textit{put}(a,\textit{pri})$, we can see that $\textit{call}(\textit{put},a,\textit{pri})$ is before $\textit{ret}(\textit{rm},a)$, and then $o$ is before $\textit{ret}(\textit{rm},a)$.
+Then similarly as above case, we can see that $o$ is before $\textit{call}(\textit{rm},a_1)$, and $a_1 <_{\textit{pb}}^A b$.
 Then we can goto the case of reason $1$ and treat $a_1$ as $a$.
@@
 
-Let us begin to represent several witness automata that is used to capture the existence of a data-differentiated $\_$-execution $e$, $e$ has a projection $e'$, $\textit{last}(e') = \seqPQ_1^{=}$, there exists items $a$ and $b$ with maximal priority in $e'$, $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\_)$ or $\textit{cal}(\textit{rm},a)$.
+Let us begin to represent several register automata that is used to capture the existence of a data-differentiated $\_$-execution $e$, $e$ has a projection $e'$, $\textit{last}(e') = \seqPQ_1^{=}$, there exists items $a$ and $b$ with maximal priority in $e'$, $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,\_)$ or $\textit{call}(\textit{rm},a)$.
 
@@
 
-\item[-] $\textit{act}_2$ is before $\textit{cal}(\textit{rm},d_1)$.
-\end{itemize}
-
-According to Lemma \ref{lemma:EPQ1Equal as pb order and gap-point}, Lemma \ref{lemma:ob order has bounded length} and Lemma \ref{lemma:five enumeration is enough for EPQ1Equal}, it is not hard to prove that, given a data-differentiated $\_$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{=}$, $e$ does not linearizable with respect to $\textit{EMS}(\textit{PQ}_1^{=})$, if and only if, one of enumerations holds in $e$ (permit renaming), while $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ is covered by some $d_1,\ldots,d_m$, $\textit{cal}(\textit{rm},b)$ is before $\textit{ret}(\textit{put},d_m,\_)$, and $\textit{cal}(\textit{rm},d_1)$ is before $\textit{ret}(\textit{rm},a)$.
+\item[-] $\textit{act}_2$ is before $\textit{call}(\textit{rm},d_1)$.
+\end{itemize}
+
+According to Lemma \ref{lemma:EPQ1Equal as pb order and gap-point}, Lemma \ref{lemma:ob order has bounded length} and Lemma \ref{lemma:five enumeration is enough for EPQ1Equal}, it is not hard to prove that, given a data-differentiated $\_$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{=}$, $e$ does not linearizable with respect to $\textit{EMS}(\textit{PQ}_1^{=})$, if and only if, one of enumerations holds in $e$ (permit renaming), while $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ is covered by some $d_1,\ldots,d_m$, $\textit{call}(\textit{rm},b)$ is before $\textit{ret}(\textit{put},d_m,\_)$, and $\textit{call}(\textit{rm},d_1)$ is before $\textit{ret}(\textit{rm},a)$.
 We say that such $d_1,\ldots,d_m$ constitute the rightmost gap of $b$.
@@
 An automaton $\mathcal{A}_{\textit{l-eq}}^1$ is given in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}, and it is constructed for the first enumeration in \figurename~\ref{fig:history enumeration 1 for PQ1Equal}.
-Here we rename the items that covers $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ into $d$, and rename the remanning items into $e$.
-In this figure, $c = \textit{cal}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{cal}(\textit{rm},e), \textit{ret}(\textit{rm},e),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{cal}(\textit{put},d),\textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.
+Here we rename the items that covers $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ into $d$, and rename the remanning items into $e$.
+In this figure, $c = \textit{call}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{call}(\textit{rm},e), \textit{ret}(\textit{rm},e),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{call}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{call}(\textit{put},d),\textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.
 
@@
 For the third enumeration in \figurename~\ref{fig:history enumeration 3 for PQ1Equal}.
-Since we want to ensure that $a$ and $b$ are putted only once, we need to explicitly record the positions of $\textit{cal}(\textit{put},a,p)$ and $\textit{cal}(\textit{put},b,p)$.
-Since the positions of $\textit{cal}(\textit{put},a,p)$ and $\textit{cal}(\textit{put},b,p)$ are not fixed, there are finite possible cases to consider, as shown below:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{rm},b)$ and before $\textit{cal}(\textit{rm},a)$: There are two possible positions of $\textit{cal}(\textit{put},a,p)$: (1) before $\textit{cal}(\textit{rm},a_1)$, and (2) after $\textit{cal}(\textit{rm},a_1)$, and before $\textit{ret}(\textit{put},a)$.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$ and before $\textit{cal}(\textit{rm},b)$: same as above case.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{put},a_1,\_)$ and before $\textit{ret}(\textit{rm},a_1)$: same as above case.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{put},a)$ and before $\textit{cal}(\textit{put},a_1,p)$: same as above case.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$: There are three possible positions of $\textit{cal}(\textit{put},a,p)$: (1) after $\textit{cal}(\textit{put},b,p)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{cal}(\textit{put},b,p)$, and (3) before $\textit{call}(\textit{rm},a_1)$.
-
-\item[-] If $\textit{cal}(\textit{put},b,p)$ is before $\textit{cal}(\textit{rm},a_1)$: There are three possible positions of $\textit{cal}(\textit{put},a,p)$: (1) after $\textit{cal}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{cal}(\textit{put},b,p)$ and before $\textit{cal}(\textit{rm},a_1)$, and (3) before $\textit{cal}(\textit{put},b,p)$.
+Since we want to ensure that $a$ and $b$ are putted only once, we need to explicitly record the positions of $\textit{call}(\textit{put},a,p)$ and $\textit{call}(\textit{put},b,p)$.
+Since the positions of $\textit{call}(\textit{put},a,p)$ and $\textit{call}(\textit{put},b,p)$ are not fixed, there are finite possible cases to consider, as shown below:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{call}(\textit{rm},b)$ and before $\textit{call}(\textit{rm},a)$: There are two possible positions of $\textit{call}(\textit{put},a,p)$: (1) before $\textit{call}(\textit{rm},a_1)$, and (2) after $\textit{call}(\textit{rm},a_1)$, and before $\textit{ret}(\textit{put},a)$.
+
+\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$ and before $\textit{call}(\textit{rm},b)$: same as above case.
+
+\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{call}(\textit{put},a_1,\_)$ and before $\textit{ret}(\textit{rm},a_1)$: same as above case.
+
+\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{ret}(\textit{put},a)$ and before $\textit{call}(\textit{put},a_1,p)$: same as above case.
+
+\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{call}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$: There are three possible positions of $\textit{call}(\textit{put},a,p)$: (1) after $\textit{call}(\textit{put},b,p)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{call}(\textit{rm},a_1)$ and before $\textit{call}(\textit{put},b,p)$, and (3) before $\textit{call}(\textit{rm},a_1)$.
+
+\item[-] If $\textit{call}(\textit{put},b,p)$ is before $\textit{call}(\textit{rm},a_1)$: There are three possible positions of $\textit{call}(\textit{put},a,p)$: (1) after $\textit{call}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{call}(\textit{put},b,p)$ and before $\textit{call}(\textit{rm},a_1)$, and (3) before $\textit{call}(\textit{put},b,p)$.
 \end{itemize}
@@
 Let $\textit{Auts}_{\textit{1-eq}}^{3}$ be the set of finite automata that is constructed for above fourteen cases.
-For example, for the case $\textit{ca}_1$ when $\textit{cal}(\textit{put},a,p)$ is before $\textit{cal}(\textit{rm},a_1)$, $\textit{cal}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$, and $\textit{cal}(\textit{put},b,p)$ is before $\textit{cal}(\textit{rm},b)$, we construct a finite automaton $\mathcal{A}_{\textit{l-eq}}^{\textit{3-1}}$ in \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}.
-In \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}, let $c$ and $c_1 = c + \textit{cal}(\textit{put},d,\textit{les}_p)$ the same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}.
-Let $c_2 = c_1 + \textit{ret}(\textit{put},a_1)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$, $c_4 = c_3 + \textit{cal}(\textit{put},d) + \textit{ret}(\textit{rm},d)$, and $c_5 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{put},a_1) + \textit{ret}(\textit{rm},d)$.
-The other witness automata in $\textit{Auts}_{\textit{1-eq}}^{3}$ can be similarly constructed.
+For example, for the case $\textit{ca}_1$ when $\textit{call}(\textit{put},a,p)$ is before $\textit{call}(\textit{rm},a_1)$, $\textit{call}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$, and $\textit{call}(\textit{put},b,p)$ is before $\textit{call}(\textit{rm},b)$, we construct a finite automaton $\mathcal{A}_{\textit{l-eq}}^{\textit{3-1}}$ in \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}.
+In \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}, let $c$ and $c_1 = c + \textit{call}(\textit{put},d,\textit{les}_p)$ the same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}.
+Let $c_2 = c_1 + \textit{ret}(\textit{put},a_1)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$, $c_4 = c_3 + \textit{call}(\textit{put},d) + \textit{ret}(\textit{rm},d)$, and $c_5 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{put},a_1) + \textit{ret}(\textit{rm},d)$.
+The other register automata in $\textit{Auts}_{\textit{1-eq}}^{3}$ can be similarly constructed.
 
@@
 
-Similarly, we construct sets $\textit{Auts}_{\textit{1-eq}}^{4}$ and $\textit{Auts}_{\textit{1-eq}}^{5}$ of witness automata for the forth enumeration in \figurename~\ref{fig:history enumeration 4 for PQ1Equal} and the fifth enumeration in \figurename~\ref{fig:history enumeration 5 for PQ1Equal}, respectively.
+Similarly, we construct sets $\textit{Auts}_{\textit{1-eq}}^{4}$ and $\textit{Auts}_{\textit{1-eq}}^{5}$ of register automata for the forth enumeration in \figurename~\ref{fig:history enumeration 4 for PQ1Equal} and the fifth enumeration in \figurename~\ref{fig:history enumeration 5 for PQ1Equal}, respectively.
 
@@
 
-\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e)$, $\textit{last}(e')=\seqPQ_1^{=}$, $a$ and $b$ are two items with maximal priority $\textit{pri}$ in $e'$, $e'$ is a $\textit{pri}$-execution, $a <_{\textit{pb}}^* b$ in $e'$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\textit{pri})$ or $\textit{cal}(\textit{rm},a)$ in $e'$.
-
-\noindent The $\textit{only if}$ direction: Assume that $e_1 \in \mathcal{I}$ is accepted by some witness automata in $\textit{Auts}_{\textit{1-eq}}$.
+\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e)$, $\textit{last}(e')=\seqPQ_1^{=}$, $a$ and $b$ are two items with maximal priority $\textit{pri}$ in $e'$, $e'$ is a $\textit{pri}$-execution, $a <_{\textit{pb}}^* b$ in $e'$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,\textit{pri})$ or $\textit{call}(\textit{rm},a)$ in $e'$.
+
+\noindent The $\textit{only if}$ direction: Assume that $e_1 \in \mathcal{I}$ is accepted by some register automata in $\textit{Auts}_{\textit{1-eq}}$.
 By data-independence, there exists data-differentiated execution $e_2 \in \mathcal{I}$ and a renaming function $r$, such that $e_1=r(e_2)$.
-Since $e_1$ is accepted by some witness automata in  $\textit{Auts}_{\textit{1-eq}}$, let $x$, $y$ and $z$ (if exists) be the items that are renamed into $b$, $a$ and $a_1$ (if exists) by $r$, respectively, and let $d_1,\ldots,d_m$ be the items that are renamed into $d$ by $r$.
+Since $e_1$ is accepted by some register automata in  $\textit{Auts}_{\textit{1-eq}}$, let $x$, $y$ and $z$ (if exists) be the items that are renamed into $b$, $a$ and $a_1$ (if exists) by $r$, respectively, and let $d_1,\ldots,d_m$ be the items that are renamed into $d$ by $r$.
 
@@
 It is obvious that $e'' \in \textit{proj}(e_2)$ is a $\textit{pri}$-execution, $\textit{last}(e'') = \seqPQ_1^{=}$.
-According to our construction of automata in $\textit{Auts}_{\textit{1-eq}}$, it is not hard to see that $x$ and $y$ has maximal priority in $h_2$, $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is before $\textit{cal}(\textit{put},y,\textit{pri})$ or $\textit{cal}(\textit{rm},y)$ in $e''$.
-
-\noindent The $\textit{if}$ direction: Assume that there exists $e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $last(e')=\textit{PQ}_1^{=}$, $a'$ and $b'$ are two items with maximal priority $\textit{pri}$ in $e'$, $e'$ is a $\textit{pri}$-execution, $a' <_{\textit{pb}}^* b'$ in $e'$, and the rightmost gap-point of $b'$ is before $\textit{cal}(\textit{put},a',\textit{pri})$ or $\textit{cal}(\textit{rm},a')$ in $e'$.
+According to our construction of automata in $\textit{Auts}_{\textit{1-eq}}$, it is not hard to see that $x$ and $y$ has maximal priority in $h_2$, $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is before $\textit{call}(\textit{put},y,\textit{pri})$ or $\textit{call}(\textit{rm},y)$ in $e''$.
+
+\noindent The $\textit{if}$ direction: Assume that there exists $e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $last(e')=\textit{PQ}_1^{=}$, $a'$ and $b'$ are two items with maximal priority $\textit{pri}$ in $e'$, $e'$ is a $\textit{pri}$-execution, $a' <_{\textit{pb}}^* b'$ in $e'$, and the rightmost gap-point of $b'$ is before $\textit{call}(\textit{put},a',\textit{pri})$ or $\textit{call}(\textit{rm},a')$ in $e'$.
 By data-independence, we can obtain execution $e_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the items $d_1,\ldots,d_m$ that constitute the rightmost gap of $b'$, we rename them into $d$, (3) if $a' <_{\textit{pb}}^A a'_1 <_{\textit{pb}}^B b$, we rename $a'_1$ into $a_1$, and (4) rename the other items into $e$.
-It is easy to see that $\textit{last}(e_1) = \seqPQ_1^{=}$, $a$ and $b$ has maximal priority in $e_1$, $a <_{\textit{pb}}^* b$ in $e_1$, and the rightmost gap-point of $b$ is before $\textit{cal}(\textit{put},a,\textit{pri})$ or $\textit{cal}(\textit{rm},a)$ in $e_1$.
+It is easy to see that $\textit{last}(e_1) = \seqPQ_1^{=}$, $a$ and $b$ has maximal priority in $e_1$, $a <_{\textit{pb}}^* b$ in $e_1$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,\textit{pri})$ or $\textit{call}(\textit{rm},a)$ in $e_1$.
 By Lemma \ref{lemma:five enumeration is enough for EPQ1Equal}, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists).
@@
 
-\item[-] If $a <_{\textit{pb}}^* b$ because of the third enumeration, it is easy to see that $h_1$ is accepted by some witness automaton in $\textit{Auts}_{\textit{1-eq}}^{3}$.
-
-\item[-] If $a <_{\textit{pb}}^* b$ because of the forth enumeration, it is easy to see that $h_1$ is accepted by some witness automaton in $\textit{Auts}_{\textit{1-eq}}^{4}$.
-
-\item[-] If $a <_{\textit{pb}}^* b$ because of the fifth enumeration, it is easy to see that $h_1$ is accepted by some witness automaton in $\textit{Auts}_{\textit{1-eq}}^{5}$.
+\item[-] If $a <_{\textit{pb}}^* b$ because of the third enumeration, it is easy to see that $h_1$ is accepted by some register automaton in $\textit{Auts}_{\textit{1-eq}}^{3}$.
+
+\item[-] If $a <_{\textit{pb}}^* b$ because of the forth enumeration, it is easy to see that $h_1$ is accepted by some register automaton in $\textit{Auts}_{\textit{1-eq}}^{4}$.
+
+\item[-] If $a <_{\textit{pb}}^* b$ because of the fifth enumeration, it is easy to see that $h_1$ is accepted by some register automaton in $\textit{Auts}_{\textit{1-eq}}^{5}$.
 \end{itemize}
@@
 Let $x_1,\ldots,x_m$ be the set of items with priority $\textit{pri}$ and has unmatched $\textit{put}$ in $e$, let $y_1,\ldots,y_n$ be the set of items with priority $\textit{pri}$ and has matched $\textit{put}$ and $\textit{rm}$ in $e$.
-By assumption, we know that $\textit{cal}(\textit{put},y_i,\textit{pri})$ is before $\textit{ret}(\textit{put},x_j)$ for each $i,j$.
+By assumption, we know that $\textit{call}(\textit{put},y_i,\textit{pri})$ is before $\textit{ret}(\textit{put},x_j)$ for each $i,j$.
 Then we explicitly construction the linearization of $e$ by locating the linearization points of $e$ as follows:
@@
 The automaton is $\mathcal{A}_{\seqPQ}^3$, which is given in \figurename~\ref{fig:automata for PQ3}.
-In \figurename~\ref{fig:automata for PQ3}, let $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d), \textit{ret}(\textit{rm},d),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{cal}(\textit{put},b,\textit{anyPri})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c + \textit{ret}(\textit{rm},b)$.
+In \figurename~\ref{fig:automata for PQ3}, let $c = \textit{call}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{call}(\textit{rm},d), \textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{call}(\textit{put},b,\textit{anyPri})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c + \textit{ret}(\textit{rm},b)$.
 
@@
 
-Let sequence $e'_{\textit{lp}}$ be generated from $e_{\textit{lp}}$ by adding $\textit{insert}(x)$ at an arbitrary time point between $\textit{cal}(\textit{insert},x)$ and $\textit{ret}(\textit{insert},x)$.
+Let sequence $e'_{\textit{lp}}$ be generated from $e_{\textit{lp}}$ by adding $\textit{insert}(x)$ at an arbitrary time point between $\textit{call}(\textit{insert},x)$ and $\textit{ret}(\textit{insert},x)$.
 Let $l''$ be the projection of $e'_{\textit{lp}}$ into method events.
@@
 
-It is easy to see that $\textit{delete}(x)$ does not happen before $\textit{insert}(x)$, and then $\textit{cal}(\textit{insert},x)$ is before $\textit{ret}(\textit{delete},x)$ in $e$.
-Let sequence $e'_{\textit{lp}}$ be generated from $e_{\textit{lp}}$ by adding $\textit{insert}(x)$ just after $\textit{cal}(\textit{insert},x)$ and adding $\textit{delete}(x)$ just before $\textit{ret}(\textit{delete},x)$.
+It is easy to see that $\textit{delete}(x)$ does not happen before $\textit{insert}(x)$, and then $\textit{call}(\textit{insert},x)$ is before $\textit{ret}(\textit{delete},x)$ in $e$.
+Let sequence $e'_{\textit{lp}}$ be generated from $e_{\textit{lp}}$ by adding $\textit{insert}(x)$ just after $\textit{call}(\textit{insert},x)$ and adding $\textit{delete}(x)$ just before $\textit{ret}(\textit{delete},x)$.
 Let $l''$ be the projection of $e'_{\textit{lp}}$ into method events.
@@
 By assumption, we know that in $e$ there exists one $\textit{insert}(a)$ and one $\textit{delete}(a)$, and $\textit{delete}(a)$ does not happen before $\textit{insert}(a)$.
-Then we know that $\textit{cal}(\textit{insert},x)$ is before $\textit{ret}(\textit{delete},x)$.
-
-Let $e_{\textit{lp}}$ be generated from $e$ by (1) put the linearization of $\textit{insert}(x)$ just after $\textit{cal}(\textit{insert},x)$, (2) put the linearization of $\textit{delete}(x)$ just before $\textit{ret}(\textit{insert},x)$, and (3) for other operations, put its linearization point at an arbitrary between its call and return actions.
+Then we know that $\textit{call}(\textit{insert},x)$ is before $\textit{ret}(\textit{delete},x)$.
+
+Let $e_{\textit{lp}}$ be generated from $e$ by (1) put the linearization of $\textit{insert}(x)$ just after $\textit{call}(\textit{insert},x)$, (2) put the linearization of $\textit{delete}(x)$ just before $\textit{ret}(\textit{insert},x)$, and (3) for other operations, put its linearization point at an arbitrary between its call and return actions.
 Let $l'$ be the projection of $e_{\textit{lp}}$ into method events.
@@
 
-For each such case, we construct a witness automata.
-We generate witness automata $\mathcal{A}_{\textit{MS}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata 1 for MS-2 in appendix}.
-Here $c_1 = \textit{cal}(\textit{insert},a),\textit{ret}(\textit{insert},a)$, $\textit{cal}(\textit{delete},a),\textit{ret}(\textit{delete},a),
-\textit{cal}(\textit{delete},\textit{empty}),\textit{ret}(\textit{delete},\textit{empty})$, $c_2 = c_1 + \textit{cal}(\textit{delete},b) + \textit{ret}(\textit{delete},b)$.
+For each such case, we construct a register automata.
+We generate register automata $\mathcal{A}_{\textit{MS}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata 1 for MS-2 in appendix}.
+Here $c_1 = \textit{call}(\textit{insert},a),\textit{ret}(\textit{insert},a)$, $\textit{call}(\textit{delete},a),\textit{ret}(\textit{delete},a),
+\textit{call}(\textit{delete},\textit{empty}),\textit{ret}(\textit{delete},\textit{empty})$, $c_2 = c_1 + \textit{call}(\textit{delete},b) + \textit{ret}(\textit{delete},b)$.
 
@@
 
-We generate witness automata $\mathcal{A}_{\textit{MS}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata 2 for MS-2 in appendix}.
-Here $c_1 = \textit{cal}(\textit{insert},a),\textit{ret}(\textit{insert},a), \textit{cal}(\textit{delete},a),\textit{ret}(\textit{delete},a),\textit{cal}(\textit{delete},\textit{empty})$, $\textit{ret}(\textit{delete},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{insert},b)$, $c_3 = c_2 + \textit{ret}(\textit{delete},b)$, $c_4 = c_3 + \textit{cal}(\textit{delete},b)$, $c_5 = c_1 + \textit{ret}(\textit{delete},b)$, $c_6 = c_5 + \textit{cal}(\textit{delete},b)$.
+We generate register automata $\mathcal{A}_{\textit{MS}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata 2 for MS-2 in appendix}.
+Here $c_1 = \textit{call}(\textit{insert},a),\textit{ret}(\textit{insert},a), \textit{call}(\textit{delete},a),\textit{ret}(\textit{delete},a),\textit{call}(\textit{delete},\textit{empty})$, $\textit{ret}(\textit{delete},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{insert},b)$, $c_3 = c_2 + \textit{ret}(\textit{delete},b)$, $c_4 = c_3 + \textit{call}(\textit{delete},b)$, $c_5 = c_1 + \textit{ret}(\textit{delete},b)$, $c_6 = c_5 + \textit{call}(\textit{delete},b)$.
 
@@
 
-We generate witness automata $\mathcal{A}_{\textit{MS}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata 3 for MS-2 in appendix}.
-Here $c_1 = \textit{cal}(\textit{insert},a)$, $\textit{ret}(\textit{insert},a)$, $\textit{cal}(\textit{delete},a)$, $\textit{ret}(\textit{delete},a),\textit{cal}(\textit{delete},b)$, $\textit{cal}($ $\textit{delete},\textit{empty}),\textit{ret}(\textit{delete},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{delete},b)$, $c_3 = c_2 + \textit{ret}(\textit{insert},b)$.
+We generate register automata $\mathcal{A}_{\textit{MS}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata 3 for MS-2 in appendix}.
+Here $c_1 = \textit{call}(\textit{insert},a)$, $\textit{ret}(\textit{insert},a)$, $\textit{call}(\textit{delete},a)$, $\textit{ret}(\textit{delete},a),\textit{call}(\textit{delete},b)$, $\textit{call}($ $\textit{delete},\textit{empty}),\textit{ret}(\textit{delete},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{delete},b)$, $c_3 = c_2 + \textit{ret}(\textit{insert},b)$.
 
@@
 
-\noindent The $\textit{only if}$ direction: Assume that $e_1 \in \mathcal{I}$ is accepted by some witness automata in $\textit{Auts}_{\textit{1-eq}}$.
+\noindent The $\textit{only if}$ direction: Assume that $e_1 \in \mathcal{I}$ is accepted by some register automata in $\textit{Auts}_{\textit{1-eq}}$.
 By data-independence, there exists data-differentiated execution $e_2 \in \mathcal{I}$ and a renaming function $r$, such that $e_1=r(e_2)$.
-Since $e_1$ is accepted by some witness automata in  $\textit{Auts}_{\textit{1-eq}}$, let $y$ be the item that are renamed into $b$ by $r$.
+Since $e_1$ is accepted by some register automata in  $\textit{Auts}_{\textit{1-eq}}$, let $y$ be the item that are renamed into $b$ by $r$.
 Then it is not hard to see that $y$ satisfies one of three conditions in $e_2$.
@@
 
-Similar as witness automata for $\seqPQ_3$, we have witness automata $\mathcal{A}_{\textit{MS}}^4$ for $\textit{MSet}_3$, which is shown in \figurename~\ref{fig:automata 4 for MS-3 in appendix}.
-In \figurename~\ref{fig:automata 4 for MS-3 in appendix}, let $c = \textit{cal}(\textit{insert},d),\textit{ret}(\textit{insert},d), \textit{cal}(\textit{delete},d)$, $\textit{ret}(\textit{delete},d),\textit{cal}(\textit{delete},\textit{empty}),\textit{ret}(\textit{delete},\textit{empty})$, $c_1 = c + \textit{cal}(\textit{insert},b)$, $c_2 = c_1 + \textit{ret}(\textit{delete},b)$, and $c_3 = c + \textit{ret}(\textit{delete},b)$.
+Similar as register automata for $\seqPQ_3$, we have register automata $\mathcal{A}_{\textit{MS}}^4$ for $\textit{MSet}_3$, which is shown in \figurename~\ref{fig:automata 4 for MS-3 in appendix}.
+In \figurename~\ref{fig:automata 4 for MS-3 in appendix}, let $c = \textit{call}(\textit{insert},d),\textit{ret}(\textit{insert},d), \textit{call}(\textit{delete},d)$, $\textit{ret}(\textit{delete},d),\textit{call}(\textit{delete},\textit{empty}),\textit{ret}(\textit{delete},\textit{empty})$, $c_1 = c + \textit{call}(\textit{insert},b)$, $c_2 = c_1 + \textit{ret}(\textit{delete},b)$, and $c_3 = c + \textit{ret}(\textit{delete},b)$.
 
@@
 According to definition of $\textit{MStoEPQ}$, we can see that (1) in $e_{\textit{epq}}$ there does not exist two items with comparable priorities, and (2) in $e_{\textit{epq}}$, there does not exists two items with a same priority.
-Therefore, the witness automata for $\seqPQ_1$ and the witness automaton $\mathcal{A}_{\textit{SinPri}}^4$ can be safely ignored.
+Therefore, the register automata for $\seqPQ_1$ and the register automaton $\mathcal{A}_{\textit{SinPri}}^4$ can be safely ignored.
 Then, if $e_{\textit{epq}}$ is accepted by $\mathcal{A}_{\textit{SinPri}}^1$, $\mathcal{A}_{\textit{SinPri}}^2$, $\mathcal{A}_{\textit{SinPri}}^3$ or $\mathcal{A}_{\seqPQ}^3$, then it is easy to see that $e_m$ is accepted by $\mathcal{A}_{\textit{MS}}^3$, $\mathcal{A}_{\textit{MS}}^1$, $\mathcal{A}_{\textit{MS}}^2$ and $\mathcal{A}_{\textit{MS}}^4$, respectively.

### 2017-04-21 appendix-EPQ.tex
modify

@@
 
-\subsection{Proofs and Definitions in Subsection \ref{subsec:co-regular of EPQ1Lar}}
-\label{sec:appendix proof and definition in section co-regular of EPQ1Lar}
-
-The notion of left-right constraint used for extended priority queue is inspired by left-right constraint of queue \cite{Bouajjani:2015}.
-Let us introduce the notion of left-right constraint, which is a graph, and the existence of cycle though item with maximal priority in it correspond to the existence of the case in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
-
-\begin{definition}\label{def:left-right constraint for matched put and rm operations}
-Given a data-differentiated $\textit{pri}_x$-execution $e$ without $\textit{rm}(\textit{empty})$.
-Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be the only  method events of $e$ with priority $\textit{pri}_x$.
-The left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is the graph $G$ where:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] the nodes are the items of $e$, to which we add a node,
-
-\item[-] there is an edge from item $d_1$ to $x$, if $\textit{put}(d_1,\_)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$,
-
-\item[-] there is an edge from $x$ to item $d_1$, if $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$,
-
-\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$.
-\end{itemize}
-\end{definition}
-
-When there is a cycle $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ through item with maximal priority (for example, $x$) in $G$, we say that $x$ is covered by $d_1,\ldots,d_m$.
-To state the effectiveness of left-right constraint, take the execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} as an example, we can see that $b$ is covered by $f,e,d$.
-We need to prove that getting rid of cycle though item with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\textit{MS}(\seqPQ_1^{>})$, as stated by the following lemma:
-
-\begin{restatable}{lemma}{LinEqualsConstraintforEPQOneLar}
-\label{lemma:Lin Equals Constraint for EPQ1Lar}
-Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{>}$.
-Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x)$ and $\textit{rm}(x)$.
-$e \sqsubseteq \textit{MS}(\seqPQ_1^{>})$, if and only if $G$ has no cycle going through $x$.
-\end{restatable}
+\subsection{Proof of Lemma \ref{lemma:pri execution is enough}}
+\label{sec:appendix proof of Lemma pri execution is enough}
 
@@
 
-Given a history $h$, operation sets $S_1$, $S_2$ and sequences $l_1$ and $l_2$.
-Let $h_1 = h \vert{S_1}$ and $h_2 = h \vert{S_2}$.
-Assume that $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, and $S_1 \cup S_2$ contains all operations of $h$.
-Then, there exists a sequence $l$, such that $h \sqsubseteq l$, $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$.
-\end{restatable}
-
-\begin {proof}
-
-Given a history $h$ and a operation $o \in S_2$, let $\textit{MB}(o) = \{ o' \vert o' \in S_1$ and $o'$ happens before $o$ in $h \}$, let $\textit{SBI}(o) = \textit{min}\{ i \vert l_1[0,i]$ contains all elements of $\textit{MB}(o) \}$.
+Given an execution $e$, operation sets $S_1$, $S_2$ and sequences $l_1$ and $l_2$.
+Let $e_1 = e \vert{S_1}$ and $e_2 = e \vert{S_2}$.
+Assume that $e_1 \sqsubseteq l_1$, $e_2 \sqsubseteq l_2$, and $S_1 \cup S_2$ contains all operations of $e$.
+Then, there exists a sequence $l$, such that $e \sqsubseteq l$, $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$.
+\end{restatable}
+
+\begin {proof}
+
+Given an execution $e$ and a operation $o \in S_2$, let $\textit{MB}(o) = \{ o' \vert o' \in S_1$ and $o'$ happens before $o$ in $h \}$, let $\textit{SBI}(o) = \textit{min}\{ i \vert l_1[0,i]$ contains all elements of $\textit{MB}(o) \}$.
 
@@
 
-It is obvious that $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$, and it remains to prove that $h \sqsubseteq l$.
+It is obvious that $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$, and it remains to prove that $e \sqsubseteq l$.
 We prove this by contradiction.
-Assume that $h$ is not linearizable with respect to $l$.
-Then there must be two operations of $h$, such that $o_1 <_{hb} o_2$ in $h$ but $o_2$ before $0_1$ in $l$.
+Assume that $e \not \sqsubseteq l$.
+Then there must be two operations $o_1$, $o_2$ of $e$, such that $o_1 <_{hb} o_2$ in $e$ but $o_2$ before $0_1$ in $l$.
 Since $l \vert{S_1} = l_1$, $l \vert{S_2} = l_2$, and $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, it is easy to see that it is impossible that $o_1,o_2 \in S_1$ or $o_1,o_2 \in S_2$.
@@
 
-We deal with the case of $R = \seqPQ_1^{>}$ , and other cases can be similarly dealt with.
-
-To prove the $\textit{only if}$ direction, given $e \sqsubseteq \textit{MS}(\seqPQ_1^{>})$ and such $\textit{pri}$ and $x$.
-Since $e \sqsubseteq \textit{MS}(\seqPQ_1^{>})$ with witness $x$, we know that $e \sqsubseteq u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$, where $u$, $v$, $w$, $x$ and $\textit{pri}$ satisfy the guard of $\seqPQ_1^{>}$.
-Let $u'$, $v'$ and $w'$ be obtained from $u$, $v$ and $w$ by erasing all items with priority incomparable with $\textit{pri}$, respectively.
-It is not hard to see that $u'$, $v'$, $w'$, $x$ and $\textit{pri}$ satisfy the guard of $\seqPQ_1^{>}$, and then $e \sqsubseteq l = u' \cdot \textit{put}(x,\textit{pri}) \cdot v' \cdot \textit{rm}(x) \cdot w' \in \textit{MS}(\seqPQ_1^{>})$.
-
-To prove the $\textit{if}$ direction, given $e' = \textit{pri-Exec}(e)$ and such $x$ and $\textit{pri}$.
-Since $e' \sqsubseteq \textit{MS}(\seqPQ_1^{>})$ with witness $x$, we know that $e' \sqsubseteq l_1 = u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$, where $u$, $v$, $w$, $x$ and $\textit{pri}$ satisfy the guard of $\seqPQ_1^{>}$.
-Let $O_c$ be the set of operations in $e$ that have priorities comparable with $\textit{pri}$, and Let $O_i$ be the set of operations in $e$ that have priorities incomparable with $\textit{pri}$.
+We deal with the case of $\Gamma = \mathsf{MatchedMaxPriority}^{>}$ , and other cases can be similarly dealt with.
+
+To prove the $\textit{only if}$ direction, assume that $e \sqsubseteq l$, $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds and $p$ is the priority of $x$.
+We can see that $e \sqsubseteq u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$.
+Let $u'$, $v'$ and $w'$ be obtained from $u$, $v$ and $w$ by erasing all items with priority incomparable with $p$, respectively.
+Since the predicates in $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ does not restrict the values of priorities incomparable with $p$, is not hard to see that $e \sqsubseteq l'$ and $\mathsf{MatchedMaxPriority\text{-}Seq}(l',x)$ holds, where $l' = u' \cdot \textit{put}(x,\textit{pri}) \cdot v' \cdot \textit{rm}(x) \cdot w'$.
+
+To prove the $\textit{if}$ direction, given $e' = e \vert_{\prec p}$.
+By assumption, $e' \sqsubseteq l$, $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds and $p$ is the priority of $x$, and then we can see that $e' \sqsubseteq l_1 = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$.
+Let $O_c$ be the set of operations in $e$ that have priorities comparable with $p$, and Let $O_i$ be the set of operations in $e$ that have priorities incomparable with $p$.
 It is obvious that $l_1$ is the linearization of $e \vert_{O_c}$.
 By Lemma \ref{lemma:merge two linearization}, there exists sequence $l$, such that $e \sqsubseteq l$, and $l \vert_{O_c} = l_1$.
-Then $l = u' \cdot \textit{put}(x,\textit{pri}) \cdot v' \cdot \textit{rm}(x) \cdot w'$, where $u' \vert_{O_c} = u$, $v' \vert_{O_c} = v$ and $w' \vert_{O_c} = w$.
-Since $\textit{pri}$ is one of maximal priorities in $e$, and the predicates of guards of $\seqPQ_1^{>}$ does not restrict $O_i$, it is easy to see that $l \in \textit{MS}(\seqPQ_1^{>})$ and then $e \sqsubseteq l$ with witness $x$.
+Then $l = u' \cdot \textit{put}(x,p) \cdot v' \cdot \textit{rm}(x) \cdot w'$, where $u' \vert_{O_c} = u$, $v' \vert_{O_c} = v$ and $w' \vert_{O_c} = w$.
+Since $p$ is one of maximal priorities in $e$, and the predicates in $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ does not restrict the values with priority $O_i$, it is easy to see that $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds.
 \qed
@@
 
-We can see that $\textit{UVSet}_i(e,x) \cap \textit{UVSet}_j(e,x) = \emptyset$ for any $i \neq j$.
+\subsection{Proofs and Definitions in Subsection \ref{subsec:co-regular of EPQ1Lar}}
+\label{sec:appendix proof and definition in section co-regular of EPQ1Lar}
+
+Let us introduce $\textit{UVSet}(e,x)$, which intuitively contains all pairs of operations that should be putted before $\textit{rm}(x)$ when construction linearization of $x$ according to $\mathsf{MatchedMaxPriority}$.
+Let $\textit{UVSet}_1(e$, $x) = o \vert$ either $o <{\textit{hb}} \textit{put}(x,\_)$ or $\textit{rm}(x)$, or there exists $o'$ with the same value of $o$, such that $o' <{\textit{hb}} \textit{put}(x)$ or $\textit{rm}(x) \}$.
+For each $i > 1$, let $\textit{UVSet}_{\textit{i+1}}(e,x) = o \vert$ $o \notin \textit{UVSet}_k$ for each $k \leq i$, and either $o$ happens before some operation $o' \in \textit{UVSet}_i(e,x)$, or there exists $o''$ with the same value of $o$, and $o''$ happens before some operation $o' \in \textit{UVSet}_i(e,x) \}$.
+Let $\textit{UVSet}(e,x) = \textit{UVSet}_1(e,x) \cup \ldots$.
+Note that it is possible that $\textit{UVSet}_i(e,x) \cap \textit{UVSet}_j(e,x) = \emptyset$ for any $i \neq j$.
 
@@
 
-\LinEqualsConstraintforEPQOneLar*
+\begin{restatable}{lemma}{LinEqualsConstraintforEPQOneLar}
+\label{lemma:Lin Equals Constraint for EPQ1Lar}
+Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{>}$.
+Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority.
+Let $G$ be the graph representing the left-right constraint of $\textit{put}(x)$ and $\textit{rm}(x)$.
+$e \sqsubseteq \textit{MS}(\seqPQ_1^{>})$, if and only if $G$ has no cycle going through $x$.
+\end{restatable}
 

### 2017-04-21 appendix-EPQ.tex
modify

@@
 We generate register automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}.
-Here $C_1 = \{ \textit{call}(\textit{put},a,\top)$, $\textit{ret}(\textit{put},a,\top), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},b),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2$ $= C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{put},b,\top) \}$.
+Here $C_1 = \{ \textit{call}(\textit{put},a,\textit{true})$, $\textit{ret}(\textit{put},a,\textit{true}), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},b),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2$ $= C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{put},b,\textit{true}) \}$.
 
@@
 We generate register automata $\mathcal{A}_{\textit{SinPri}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata for FIFO-2}.
-Here $C_1 = \{ \textit{call}(\textit{put},a,\top),\textit{ret}(\textit{put},a,\top), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{call}(\textit{rm},b) + \textit{ret}(\textit{rm},b) \}$.
+Here $C_1 = \{ \textit{call}(\textit{put},a,\textit{true}),\textit{ret}(\textit{put},a,\textit{true}), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{call}(\textit{rm},b) + \textit{ret}(\textit{rm},b) \}$.
 
@@
 We generate register automata $\mathcal{A}_{\textit{SinPri}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata for FIFO-3}.
-Here $C_1 = \{ \textit{call}(\textit{put},a,\top),\textit{ret}(\textit{put},a,\top), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,\top) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_4 = C_3 \cup \{ \textit{call}(\textit{rm},b) \}$, $C_5 = C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_6 = C_5 \cup \{ \textit{call}(\textit{rm},b) \}$.
+Here $C_1 = \{ \textit{call}(\textit{put},a,\textit{true}),\textit{ret}(\textit{put},a,\textit{true}), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,\textit{true}) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_4 = C_3 \cup \{ \textit{call}(\textit{rm},b) \}$, $C_5 = C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_6 = C_5 \cup \{ \textit{call}(\textit{rm},b) \}$.
 
@@
 We generate register automata $\mathcal{A}_{\textit{SinPri}}^4$ for the forth case, and it is shown in \figurename~\ref{fig:automata for FIFO-4}.
-Here $C_1 = C \cup \{ \textit{call}(\textit{rm},b) \}$, and $C_2 = C \cup \{ \textit{ret}(\textit{put},b,\top), \textit{call}(\textit{rm},a), \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},d,\top),\textit{ret}(\textit{put},d,\top), \textit{call}(\textit{rm},d),\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.
+Here $C_1 = C \cup \{ \textit{call}(\textit{rm},b) \}$, and $C_2 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{call}(\textit{rm},a), \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},d,\textit{true}),\textit{ret}(\textit{put},d,\textit{true}), \textit{call}(\textit{rm},d),\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.
 
@@
 
+Therefore, from now on, it is safe to consider only the data-differentiated sequences with only one maximal priority.
+
 \subsection{Proofs and Definitions in Subsection \ref{subsec:co-regular of EPQ1Lar}}
@@
 \label{lemma:UVSet has matched put and rm}
-Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{>}$.
-Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
+Given a data-differentiated execution $e$ where $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, let $p$ be its maximal priority and $\textit{put}(x,p),\textit{rm}(x)$ are only operations of priority $p$ in $e$.
+Let $G$ be the graph representing the left-right constraint of $x$.
 Assume that $G$ has no cycle going through $x$.
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] For each $1 \leq i \leq j$, method events of $d_i$ belongs to $\textit{UVSet}_i(e,x)$.
+\item[-] For each $1 \leq i \leq j$, operations of $d_i$ belongs to $\textit{UVSet}_i(e,x)$.
 
@@
 Moreover, since $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap, it is not hard to see that the call action of $\textit{put}(d_i,\_)$ and the call action of $\textit{rm}(d_i)$ is before the call action of $o_{\textit{i-1}}$.
-Since method events of $d_{\textit{i+1}}$ is in $\textit{UVSet}_{\textit{i+1}}(e,x)$, we know that an operation $o'_{\textit{i+1}}$ of $d_{\textit{i+1}}$ happens before operation $o'_i$ of $d_i$.
+Since operations of $d_{\textit{i+1}}$ is in $\textit{UVSet}_{\textit{i+1}}(e,x)$, we know that an operation $o'_{\textit{i+1}}$ of $d_{\textit{i+1}}$ happens before operation $o'_i$ of $d_i$.
 Then, it is not hard to see that $o'_{\textit{i+1}}$ also happens before $o_{\textit{i-1}}$, which contradicts that for each $k > \textit{ind+1}$, no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.
@@
 
-Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{>}$.
-Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
+Given a data-differentiated execution $e$ where $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, let $p$ be its maximal priority and $\textit{put}(x,p),\textit{rm}(x)$ are only operations of priority $p$ in $e$.
+Let $G$ be the graph representing the left-right constraint of x.
 Assume that $G$ has no cycle going through $x$.
@@
 Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{UVSet}_1(e,x)$ is an operation of item $d$.
-%(according to the definition of $\textit{UVSet}_1(h,x)$, the priority of $d$ does not equals $\textit{pri}_x$).
 
@@
 $t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$.
-$t_1$ represents whether $o$ is a $\textit{put}$ method event or a $\textit{rm}$ method event.
+$t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation.
 $t_2$ and $t_3$ is used for the reason of $o \in \textit{UVSet}_1(e,x)$: $o \in \textit{UVSet}_1(e,x)$, since an operation (of kind $t_2$) of $d$ happens before an operation (of kind $t_3$) of $x$.
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.
+\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d,\_) <_{hb} \textit{put}(x,p)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,p)$.
 
@@
 
-\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x) )$.
-By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(x,\textit{pri}_x) ) \vee ( \textit{rm}(d) <_{hb} \textit{put}(d,\_) )$, which is impossible.
+\item[-] $(\textit{put},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{put}(d,\_) ) \wedge ( \textit{rm}(d) <_{hb} \textit{put}(x,p) )$.
+By interval order, we know that $( \textit{rm}(x) <_{hb} \textit{put}(x,p) ) \vee ( \textit{rm}(d) <_{hb} \textit{put}(d,\_) )$, which is impossible.
 
@@
 
-\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{put}(x,\textit{pri}_x) )$.
+\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d) ) \wedge ( \textit{put}(d,\_) <_{hb} \textit{put}(x,p) )$.
 We can see that $x$ and $d$ has circle in $G$, contradicts that $G$ has no cycle going through $x$.
@@
 
-\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{put}(x,\textit{pri}_x)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,\textit{pri}_x)$.
+\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{rm}(d) <_{hb} \textit{put}(x,p)$, contradicts that $\textit{rm}(x)$ does not happen before $\textit{put}(x,p)$.
 
@@
 $t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$.
-$t_1$ represents whether $o$ is a $\textit{put}$ method event or a $\textit{rm}$ method event.
+$t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation.
 $t_2$ and $t_3$ is used for the reason of $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$: $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$, since an operation (of kind $t_2$) of $d_{\textit{j+1}}$ happens before an operation (of kind $t_3$) of $d_j$, where $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{UVSet}_j(e,x)$.
@@
 
-With Lemma \ref{lemma:UVSet has matched put and rm} and Lemma \ref{lemma:Rmx does not happen before UVSet for EPQ1Lar}, we can now prove Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar}.
+With Lemma \ref{lemma:UVSet has matched put and rm} and Lemma \ref{lemma:Rmx does not happen before UVSet for EPQ1Lar}, we can now prove Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar}, which give the equivalent characterization of non $\mathsf{MatchedMaxPriority}$-linearizable executions.
 
@@
 \label{lemma:Lin Equals Constraint for EPQ1Lar}
-Given a data-differentiated $\textit{pri}_x$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{>}$.
-Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority.
-Let $G$ be the graph representing the left-right constraint of $\textit{put}(x)$ and $\textit{rm}(x)$.
-$e \sqsubseteq \textit{MS}(\seqPQ_1^{>})$, if and only if $G$ has no cycle going through $x$.
-\end{restatable}
-
-\begin {proof}
-
-To prove the $\textit{only if}$ direction, assume that $e \sqsubseteq \textit{MS}(\seqPQ_1^{>})$.
-Let $u$, $v$ and $w$ be the sequences of method events in $\seqPQ_1^{>}$, and let $U$, $V$ and $W$ be the set of method events of $u$, $v$ and $w$, respectively.
+Given a data-differentiated execution $e$ where $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, let $p$ be its maximal priority and $\textit{put}(x,p),\textit{rm}(x)$ are only operations of priority $p$ in $e$.
+Let $G$ be the graph representing the left-right constraint of $x$.
+$e$ is $\mathsf{MatchedMaxPriority}$-linearizable, if and only if $G$ has no cycle going through $x$.
+\end{restatable}
+
+\begin {proof}
+
+To prove the $\textit{only if}$ direction, assume that $e \sqsubseteq l$ and $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds.
+Then $l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, and let $U$, $V$ and $W$ be the set of operations of $u$, $v$ and $w$, respectively.
 Assume by contradiction that, there is a cycle $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ in $G$.
-It is obvious that the priority of each $d_i$ is smaller than $\textit{pri}_x$.
+It is obvious that the priority of each $d_i$ is smaller than $p$.
 Then our proof proceeds as follows:
@@
 By Lemma \ref{lemma:UVSet has matched put and rm}, we can see that $E_u \cup E_v$ contains matched $\textit{put}$ and $\textit{rm}$ operations.
-It remains to prove that for $E_u$, $\{ \textit{put}(x,\textit{pri}_x) \}$, $E_v$, $\{ \textit{rm}(x) \}$, $E_w$, no elements of the latter set happens before elements of the former set.
+It remains to prove that for $E_u$, $\{ \textit{put}(x,p) \}$, $E_v$, $\{ \textit{rm}(x) \}$, $E_w$, no elements of the latter set happens before elements of the former set.
 We prove this by showing that all the following cases are impossible:
@@
 
-Let us begin to represent register automata that is used for capture the existence of a data-differentiated execution $e$, $e$ has a $\_$-projection $e'$, $\textit{last}(e') = \textit{PQ}_1^{>}$, and there exists a cycle going through the item with maximal priority in $e'$.
+Let us begin to represent register automata that is used for capture the existence of a data-differentiated execution $e$, let $p$ be a maximal priority of $e$, and $e'$ be the projection of $e$ into values with priorities comparable with $p$.
+We also require that $\mathsf{Has\text{-}MatchedMaxPriority}(e')$ holds and there exists a cycle going through the value with maximal priority in $e'$.
 By data-independence, we can obtain $e_r$ from $e$ by renaming function, which maps such item to be $b$, maps items that cover it to be $a$, and maps other items into $d$.
-There are four possible enumeration of call and return actions of $\textit{put}(b)$ and $\textit{rm}(b)$.
+There are four possible enumeration of call and return actions of $\textit{put}(b,p)$ and $\textit{rm}(b)$.
 For each of them, we generate a register automaton.
 
-For the case when $e_r \vert_{b} = \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^1$, as shown in \figurename~\ref{fig:automata APQ1Lar-1}.
-Here $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{call}(\textit{put},a,\textit{les}_p)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{call}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.
+For the case when $e_r \vert_{b} = \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{put},b,p) \cdot \textit{call}(\textit{rm},b) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^1$, as shown in \figurename~\ref{fig:automata APQ1Lar-1}.
+Here $C_1 = C \cup \{ \textit{ret}(\textit{rm},a) \}$, $C_2 = C \cup \{ \textit{call}(\textit{put},a,<r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.
 The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^1$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.
@@
 
-Let $o$ be the right-most gap-point of $x$. We locate linearization points of each method event as follows:
+Let $o$ be the right-most gap-point of $x$. We locate linearization points of each operation as follows:
 
@@
 
-\item[-] Locate linearization points of method event of each $y \in \textit{Items}(e,\textit{pri})$ (except for $x$) just after the call action of the method event.
+\item[-] Locate linearization points of operation of each $y \in \textit{Items}(e,\textit{pri})$ (except for $x$) just after the call action of the operation.
 
@@
 According to Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar} (Here we temporarily forget the existence of $y$), the only possible position for locating linearizaton point of $\textit{rm}(x)$ is at gap-point of $x$.
-Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $x$, then there exists unmatched method event before $\textit{rm}(x)$ with smaller priority.
+Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $x$, then there exists unmatched operation before $\textit{rm}(x)$ with smaller priority.
 Since the rightmost gap-point of $x$ is before $\textit{call}(\textit{put},y,\textit{pri})$ or $\textit{call}(\textit{rm},y)$, if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $x$, then $\textit{rm}(x)$ will be before $\textit{call}(\textit{put},y,\textit{pri})$ or $\textit{call}(\textit{rm},x)$.
 
-Therefore, for every sequence $l = u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$, if $e_{x,y} \sqsubseteq l$, then either $u \cdot v$ contains some unmatched method events of priority smaller than $\textit{pri}$, or $w$ contains $\textit{put}(y,\textit{pri})$ or $\textit{rm}(y)$.
+Therefore, for every sequence $l = u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$, if $e_{x,y} \sqsubseteq l$, then either $u \cdot v$ contains some unmatched operations of priority smaller than $\textit{pri}$, or $w$ contains $\textit{put}(y,\textit{pri})$ or $\textit{rm}(y)$.
 In both cases, $l \notin \textit{MS}(\seqPQ_1^{=})$.
@@
 
-There are many enumerations of method events of $a$, $b$ and $a_1$ that may makes $a <_{\textit{pb}}^* b$.
+There are many enumerations of operations of $a$, $b$ and $a_1$ that may makes $a <_{\textit{pb}}^* b$.
 The following lemma states that with the help of gap-points, the number of potential enumerations can be further reduced into only five.
@@
 Assume that $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,\textit{pri})$ or $\textit{call}(\textit{rm},a)$.
-Then, there are five possible enumeration of method events of $a$, $b$, $a_1$ (if exists), where $a_1$ is the possible intermediate items for obtain $a <_{\textit{pb}}^* b$.
+Then, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists), where $a_1$ is the possible intermediate items for obtain $a <_{\textit{pb}}^* b$.
 \end{restatable}
@@
 Since $\textit{last}(e) = \seqPQ_2^{>}$, the actions with maximal priority in $e$ is some unmatched $\textit{put}$.
-Therefore, no matter how we locate linearization points, we can always obtain a sequence $l$ of method events that contains unmatched $\textit{put}$ with maximal priority, and this satisfy the guard of $\textit{MS}(\seqPQ_2^{>})$.
+Therefore, no matter how we locate linearization points, we can always obtain a sequence $l$ of operations that contains unmatched $\textit{put}$ with maximal priority, and this satisfy the guard of $\textit{MS}(\seqPQ_2^{>})$.
 This completes the proof of this lemma.
@@
 
-\item[-] For other method events, locate their linearization points at an arbitrary location after its call action and before its return action.
+\item[-] For other operations, locate their linearization points at an arbitrary location after its call action and before its return action.
 \end{itemize}
@@
 Since $U$ contains matched $\textit{put}$ and $\textit{rm}$, we can see that operations of $d_m$ is in $U$.
-Similarly, we can see that method events of $d_{\textit{m-1}},\ldots,d_1$ is in $U$.
+Similarly, we can see that operations of $d_{\textit{m-1}},\ldots,d_1$ is in $U$.
 If $\textit{rm}(d_1)$ does not exists, then this contradicts that $U$ contains matched $\textit{put}$ and $\textit{rm}$.
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] Given a sequential execution $l$ of multi-set, $\textit{noDE}(l)$ is satisfied when each method event of $l$ is not $\textit{delete}(\textit{empty})$.
+\item[-] Given a sequential execution $l$ of multi-set, $\textit{noDE}(l)$ is satisfied when each operation of $l$ is not $\textit{delete}(\textit{empty})$.
 
@@
 
-\item[-] Given a sequential execution $l$ of multi-set, $l \in \textit{Insert}^*$ is satisfied when each method event of $l$ is a $\textit{insert}$ event.
+\item[-] Given a sequential execution $l$ of multi-set, $l \in \textit{Insert}^*$ is satisfied when each operation of $l$ is a $\textit{insert}$ event.
 \end{itemize}
@@
 Let $e_{\textit{lp}}$ be an execution with linearization points of $e$ and the linearization points is added according to $l'$.
-Or we can say, $e_{\textit{lp}}$ is generated from $e$ by instrumenting linearization points, and the projection of $e_{\textit{lp}}$ into method event is $l'$.
+Or we can say, $e_{\textit{lp}}$ is generated from $e$ by instrumenting linearization points, and the projection of $e_{\textit{lp}}$ into operation is $l'$.
 
 Let sequence $e'_{\textit{lp}}$ be generated from $e_{\textit{lp}}$ by adding $\textit{insert}(x)$ at an arbitrary time point between $\textit{call}(\textit{insert},x)$ and $\textit{ret}(\textit{insert},x)$.
-Let $l''$ be the projection of $e'_{\textit{lp}}$ into method events.
+Let $l''$ be the projection of $e'_{\textit{lp}}$ into operations.
 
@@
 Let $e_{\textit{lp}}$ be an execution with linearization points of $e$ and the linearization points is added according to $l'$.
-Or we can say, $e_{\textit{lp}}$ is generated from $e$ by instrumenting linearization points, and the projection of $e_{\textit{lp}}$ into method event is $l'$.
+Or we can say, $e_{\textit{lp}}$ is generated from $e$ by instrumenting linearization points, and the projection of $e_{\textit{lp}}$ into operation is $l'$.
 
@@
 Let sequence $e'_{\textit{lp}}$ be generated from $e_{\textit{lp}}$ by adding $\textit{insert}(x)$ just after $\textit{call}(\textit{insert},x)$ and adding $\textit{delete}(x)$ just before $\textit{ret}(\textit{delete},x)$.
-Let $l''$ be the projection of $e'_{\textit{lp}}$ into method events.
+Let $l''$ be the projection of $e'_{\textit{lp}}$ into operations.
 
@@
 Let $e_{\textit{lp}}$ be generated from $e$ by (1) put the linearization of $\textit{insert}(x)$ just after $\textit{call}(\textit{insert},x)$, (2) put the linearization of $\textit{delete}(x)$ just before $\textit{ret}(\textit{insert},x)$, and (3) for other operations, put its linearization point at an arbitrary between its call and return actions.
-Let $l'$ be the projection of $e_{\textit{lp}}$ into method events.
+Let $l'$ be the projection of $e_{\textit{lp}}$ into operations.
 It is obvious that $h \sqsubseteq l'$.

### 2017-04-21 appendix-EPQ.tex
modify

@@
 For the case when $e_r \vert_{b} = \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{put},b,p) \cdot \textit{call}(\textit{rm},b) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^1$, as shown in \figurename~\ref{fig:automata APQ1Lar-1}.
-Here $C_1 = C \cup \{ \textit{ret}(\textit{rm},a) \}$, $C_2 = C \cup \{ \textit{call}(\textit{put},a,<r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.
-The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^1$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.
+Here $C_1 = C \cup \{ \textit{ret}(\textit{rm},a) \}$, $C_2 = C \cup \{ \textit{call}(\textit{put},a,<r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},\top)$, $\textit{ret}(\textit{rm},\top),$ $\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.
+The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^1$ comes from the positions of the first $\textit{ret}(\textit{put},a,\_)$.
 
@@
 $\mathcal{A}_{\textit{l-lar}}^1$ is used to recognize conditions in \figurename~\ref{fig:his for APQ1Lar-1}.
-Here for simplicity, we only draw operation of $b$, and the first $\textit{ret}(\textit{put},a)$.
+Here for simplicity, we only draw operation of $b$, and the first $\textit{ret}(\textit{put},a,\_)$.
 
@@
 
-For the case when $e_r \vert_{b} = \textit{call}(\textit{put},b,p) \cdot \textit{call}(\textit{rm}) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^2$, as shown in \figurename~\ref{fig:automata APQ1Lar-2}.
-Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$.
-The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^2$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.
+For the case when $e_r \vert_{b} = \textit{call}(\textit{put},b,p) \cdot \textit{call}(\textit{rm},b) \cdot \textit{ret}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^2$, as shown in \figurename~\ref{fig:automata APQ1Lar-2}.
+Here $C_1,C_2,C_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$.
+The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^2$ comes from the positions of the first $\textit{ret}(\textit{put},a,\_)$.
 
@@
 $\mathcal{A}_{\textit{l-lar}}^2$ is used to recognize conditions in \figurename~\ref{fig:his for APQ1Lar-2}.
-Here for simplicity, we only draw operation of $b$, and the first $\textit{ret}(\textit{put},a)$.
+Here for simplicity, we only draw operation of $b$, and the first $\textit{ret}(\textit{put},a,\_)$.
 
@@
 
-For the case when $e_r \vert_{b} = \textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{put}) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^3$, as shown in \figurename~\ref{fig:automata APQ1Lar-3}.
-Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$.
-The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^3$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.
+For the case when $e_r \vert_{b} = \textit{call}(\textit{rm},b) \cdot \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^3$, as shown in \figurename~\ref{fig:automata APQ1Lar-3}.
+Here $C_1,C_2,C_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$.
+The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^3$ comes from the positions of the first $\textit{ret}(\textit{put},a,\_)$.
 
@@
 $\mathcal{A}_{\textit{l-lar}}^3$ is used to recognize conditions in \figurename~\ref{fig:his for APQ1Lar-3}.
-Here for simplicity, we only draw operation of $b$, and the first $\textit{ret}(\textit{put},a)$.
+Here for simplicity, we only draw operation of $b$, and the first $\textit{ret}(\textit{put},a,\_)$.
 
@@
 
-For the case when $e_r \vert_{b} = \textit{call}(\textit{rm}) \cdot \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b) \cdot \textit{ret}(\textit{put})$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^4$, as shown in \figurename~\ref{fig:automata APQ1Lar-4}.
-Here $c_1,c_2,c_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$, and $c_4 = c_1 + \textit{ret}(\textit{put},b)$.
-The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^4$ comes from the positions of the first $\textit{ret}(\textit{put},a)$.
+For the case when $e_r \vert_{b} = \textit{call}(\textit{rm},b) \cdot \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b) \cdot \textit{ret}(\textit{put},b,p)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^4$, as shown in \figurename~\ref{fig:automata APQ1Lar-4}.
+Here $C_1,C_2,C_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$, and $C_4 = C_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$.
+The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^4$ comes from the positions of the first $\textit{ret}(\textit{put},a,\_)$.
 
@@
 $\mathcal{A}_{\textit{l-lar}}^4$ is used to recognize conditions in \figurename~\ref{fig:his for APQ1Lar-4}.
-Here for simplicity, we only draw operation of $b$, and the first $\textit{ret}(\textit{put},a)$.
+Here for simplicity, we only draw operation of $b$, and the first $\textit{ret}(\textit{put},a,\_)$.
 
@@
 
-Let $\textit{Auts}_{\textit{1-lar}} = \{ \mathcal{A}_{\textit{l-lar}}^1, \mathcal{A}_{\textit{l-lar}}^2, \mathcal{A}_{\textit{l-lar}}^3, \mathcal{A}_{\textit{l-lar}}^4 \}$.
-The following lemma states that $\seqPQ_1^{>}$ is co-regular.
+Let $\mathcal{A}_{\textit{1-lar}}$ be the union of $\mathcal{A}_{\textit{l-lar}}^1, \mathcal{A}_{\textit{l-lar}}^2, \mathcal{A}_{\textit{l-lar}}^3$ and $\mathcal{A}_{\textit{l-lar}}^4$.
+The following lemma states that $\mathcal{A}_{\textit{1-lar}}$ is $\mathsf{MatchedMaxPriority}^{>}$-complete.
 
@@
 \label{lemma:EPQ1Lar is co-regular}
-$\textit{Auts}_{\textit{1-lar}}$ is $\mathsf{MatchedMaxPriority}^{>}$-complete.
-\end{restatable}
-
-\begin {proof}
-
-We need to prove that, given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-lar}} \cap \mathcal{I} \neq \emptyset$, if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e),$ $\seqPQ_1^{>} \in last(e') \wedge e'$ does not linearizable w.r.t.
-$\textit{MS}(\seqPQ_1^{>})$.
+$\mathcal{A}_{\textit{1-lar}}$ is $\mathsf{MatchedMaxPriority}^{>}$-complete.
+\end{restatable}
+
+\begin {proof}
+
+We need to prove that, given a data-independent implementation $\mathcal{I}$.
+$\mathcal{A}_{\textit{1-lar}} \cap \mathcal{I} \neq \emptyset$ if and only if there exists $e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$ such that $e'$ is not $\mathsf{MatchedMaxPriority}^{>}$-linearizable.
 
@@
 
-\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-lar}} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e)$, $last(e')=\seqPQ_1^{>}$, $x$ is the item with maximal priority $\textit{pri}$ in $e'$, $e'$ is a $\textit{pri}$-execution.
-And there is a cycle going through $x$ in $G$, where $G$ is the left-right constraint of $e'$.
+\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$.
+$\mathcal{A}_{\textit{1-lar}} \cap \mathcal{I} \neq \emptyset$ if and only if there exists $e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$, $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, $x$ is the value with maximal priority in $e'$, $e'$ has only one maximal priority, and there is a cycle going through $x$ in $G$, where $G$ is the left-right constraint of $x$ in $e'$.
 
@@
 Assume that the priority of $b$ is $p$.
-It is easy to see that $e'$ is a $p$-execution, $\textit{last}(e') = \seqPQ_1^{>}$, and there is a cycle going through $d$ in $G$, where $G$ is the left-right constraint of $e'$.
+It is easy to see that $e'$ has only one maximal priority, $\mathsf{Has\text{-}MatchedMaxPriority}(e')$ holds, and there is a cycle going through $d$ in $G$, where $G$ is the left-right constraint of $e'$.
 The case of $\mathcal{A}_{\textit{l-lar}}^2$, $\mathcal{A}_{\textit{l-lar}}^3$ and $\mathcal{A}_{\textit{l-lar}}^4$ can be similarly proved.

### 2017-04-21 appendix-EPQ.tex
modify

@@
 
-\subsection{Proofs and Definitions in Subsection \ref{subsec:co-regular of EPQ1Lar}}
+\subsection{Proofs, Definitions and Register Automata in Subsection \ref{subsec:co-regular of EPQ1Lar}}
 \label{sec:appendix proof and definition in section co-regular of EPQ1Lar}
@@
 
-\subsection{Proofs and Definitions in Subsection \ref{subsec:co-regular of EPQ1Equal}}
+\subsection{Proofs, Definitions and Register Automata in Subsection \ref{subsec:co-regular of EPQ1Equal}}
 \label{sec:appendix proof and definition in section co-regular of EPQ1Equal}
 
-Let $\textit{Items}(e,p)$ be the set of items with priority $p$ in execution $e$. The following lemma states a method to choose $\textit{itm}$ of $\seqPQ_1^{=}$.
+Let $\textit{Items}(e,p)$ be the set of items with priority $p$ in execution $e$. The following lemma states a method to build the linearization of $e$ w.r.t $\mathsf{MatchedMaxPriority}^{=}$.
 
@@
 \label{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}
-Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{=}$.
-If there exists an item $x$ with priority $\textit{pri}$, such that for each $y \in \textit{Items}(e,\textit{pri})$, (1) $x$ does not $<_{\textit{pb}}$ to $y$, and (2) the right-most gap-point of $x$ is after $\textit{call}(\textit{put},y,\_)$ and $\textit{call}(\textit{rm},y)$.
-Then $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
+Given a data-differentiated execution $e$ where $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e)$ holds and $p$ is its maximal priority.
+If there exists a value $x$ with priority $p$, such that for each $y \in \textit{Items}(e,p)$, (1) $x\not<_{\textit{pb}}y$, and (2) the right-most gap-point of $x$ is after $\textit{call}(\textit{put},y,p)$ and $\textit{call}(\textit{rm},y)$.
+Then $e$ is $\mathsf{MatchedMaxPriority}^{=}$-linearizable.
 \end{restatable}
@@
 
-\item[-] If $\textit{put}(x,\textit{pri})$ overlaps with $\textit{rm}(x)$, then locate the linearization point of $\textit{put}(x,\textit{pri})$ just before the linearization point of $\textit{rm}(x)$.
-Otherwise, $\textit{put}(x,\textit{pri}) <_{\textit{hb}} \textit{rm}(x)$, and we locate the linearization point of $\textit{put}(x,\textit{pri})$ just before its return action.
-
-\item[-] Locate linearization points of operation of each $y \in \textit{Items}(e,\textit{pri})$ (except for $x$) just after the call action of the operation.
-
-\item[-] For item $z$ with priority smaller than $\textit{pri}$.
+\item[-] If $\textit{put}(x,p)$ overlaps with $\textit{rm}(x)$, then locate the linearization point of $\textit{put}(x,p)$ just before the linearization point of $\textit{rm}(x)$.
+Otherwise, $\textit{put}(x,p) <_{\textit{hb}} \textit{rm}(x)$, and we locate the linearization point of $\textit{put}(x,p)$ just before its return action.
+
+\item[-] Locate linearization points of operation of each $y \in \textit{Items}(e,p)$ (except for $x$) just after the call action of the operation.
+
+\item[-] For item $z$ with priority smaller than $p$.
 If both $\textit{call}(\textit{put},z,\_)$ and $\textit{call}(\textit{rm},z)$ is before $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just after their call actions.
@@
 It is obvious that $e \sqsubseteq l$.
-Since for each $y \in \textit{Items}(e,\textit{pri})$, $o$ is after $\textit{call}(\textit{put},y,\_)$ and $\textit{call}(\textit{rm},x)$, we can see that $\textit{rm}(x)$ is after $\textit{put}(y,\textit{pri})$ and $\textit{rm}(y)$ in $l$.
-It is obvious that $\textit{put}(x,\textit{pri})$ is before $\textit{rm}(x)$ in $l$.
-Since $x$ does not $<_{\textit{pb}}$ to $y$, we can see that no $\textit{put}(y,\textit{pri})$ happens before $\textit{put}(x,\textit{pri})$.
-Then it is easy to see that $\textit{put}(x,\textit{pri})$ is after $\textit{put}(y,\textit{pri})$ in $l$.
-Since $\textit{last}(e) = \seqPQ_1^{=}$, all other items in $\textit{Items}(e,\textit{pri})$ has matched $\textit{put}$ and $\textit{rm}$, and it is easy to see that their $\textit{put}$ and $\textit{rm}$ (except for that of $x$) are all before $\textit{rm}(x)$ in $l$.
-
-For item $z$ with priority smaller than $\textit{pri}$, we can see that there are only two possibilities: (1) $\textit{put}(z,\_)$ and $\textit{rm}(z)$ are both before $\textit{rm}(x)$ in $l$, and (2) $\textit{put}(z,\_)$ and $\textit{rm}(z)$ (if exists) are after before $\textit{rm}(x)$ in $l$.
+Since for each $y \in \textit{Items}(e,p)$, $o$ is after $\textit{call}(\textit{put},y,\_)$ and $\textit{call}(\textit{rm},x)$, we can see that $\textit{rm}(x)$ is after $\textit{put}(y,p)$ and $\textit{rm}(y)$ in $l$.
+It is obvious that $\textit{put}(x,p)$ is before $\textit{rm}(x)$ in $l$.
+Since $x$ does not $<_{\textit{pb}}$ to $y$, we can see that no $\textit{put}(y,p)$ happens before $\textit{put}(x,p)$.
+Then it is easy to see that $\textit{put}(x,p)$ is after $\textit{put}(y,p)$ in $l$.
+Since $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e)$ holds, all other items in $\textit{Items}(e,p)$ has matched $\textit{put}$ and $\textit{rm}$, and it is easy to see that their $\textit{put}$ and $\textit{rm}$ (except for that of $x$) are all before $\textit{rm}(x)$ in $l$.
+
+For item $z$ with priority smaller than $p$, we can see that there are only two possibilities: (1) $\textit{put}(z,\_)$ and $\textit{rm}(z)$ are both before $\textit{rm}(x)$ in $l$, and (2) $\textit{put}(z,\_)$ and $\textit{rm}(z)$ (if exists) are after before $\textit{rm}(x)$ in $l$.
 Therefore, before $\textit{rm}(x)$ in $l$, the $\textit{put}$ and $\textit{rm}$ of $z$ are matched.
 
-Therefore, it is easy to see that $l \in \textit{MS}(\seqPQ_1^{=})$. \qed
-\end {proof}
-
-With Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, we can prove the following lemma, which states that getting rid of case in \figurename~\ref{fig:introduce pb order} is enough for ensure $\textit{last}(e) = \seqPQ_1^{=} \Rightarrow e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
+Therefore, it is easy to see that $\mathsf{MatchedMaxPriority^{=}\text{-}Seq}(l,x)$ holds and $e$ is $\mathsf{MatchedMaxPriority}^{=}$-linearizable. \qed 
+\end {proof}
+
+With Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, we can prove the following lemma, which states that getting rid of case in \figurename~\ref{fig:introduce pb order} is enough for ensure that $\mathsf{MatchedMaxPriority}^{=}(e)$ holds.
 
@@
 Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $x$, then there exists unmatched operation before $\textit{rm}(x)$ with smaller priority.
-Since the rightmost gap-point of $x$ is before $\textit{call}(\textit{put},y,\textit{pri})$ or $\textit{call}(\textit{rm},y)$, if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $x$, then $\textit{rm}(x)$ will be before $\textit{call}(\textit{put},y,\textit{pri})$ or $\textit{call}(\textit{rm},x)$.
-
-Therefore, for every sequence $l = u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$, if $e_{x,y} \sqsubseteq l$, then either $u \cdot v$ contains some unmatched operations of priority smaller than $\textit{pri}$, or $w$ contains $\textit{put}(y,\textit{pri})$ or $\textit{rm}(y)$.
+Since the rightmost gap-point of $x$ is before $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$, if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $x$, then $\textit{rm}(x)$ will be before $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},x)$.
+
+Therefore, for every sequence $l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, if $e_{x,y} \sqsubseteq l$, then either $u \cdot v$ contains some unmatched operations of priority smaller than $p$, or $w$ contains $\textit{put}(y,p)$ or $\textit{rm}(y)$.
 In both cases, $l \notin \textit{MS}(\seqPQ_1^{=})$.
@@
 To prove the $\textit{only if}$ direction, we prove its contrapositive.
-Assume we already know that for each $x$ and $y$ has maximal priority in $e$, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{call}(\textit{put},y,\textit{pri})$ and $\textit{call}(\textit{rm},x)$.
+Assume we already know that for each $x$ and $y$ has maximal priority in $e$, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{call}(\textit{put},y,p)$ and $\textit{call}(\textit{rm},x)$.
 We need to prove that $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] Let $e_{\textit{pri}}$ be the projection of $e$ into operations of priority $\textit{pri}$.
-Since each single-priority execution has FIFO property, there exists sequence $l_{\textit{pri}}$, such that $e_{\textit{pri}} \sqsubseteq l_{\textit{pri}}$, and when we treat $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$, $l_{\textit{pri}}$ belongs to queue.
-
-\item[-] Let $a_1$ be the last inserted item of $l_{\textit{pri}}$.
-
-    Step $1$: Check whether for each $b \in \textit{Items}(e,\textit{pri})$, (1) $a_1$ does not $<_{\textit{pb}}$ to $b$, and (2) the right-most gap-point of $a$ is after $\textit{call}(\textit{put},b,\textit{pri})$ and $\textit{call}(\textit{rm},b)$.
-
-    It is easy to see that $a_1$ is of priority $\textit{pri}$, and $a_1$ does not $<_{\textit{pb}}$ to any $b \in \textit{Items}(e,\textit{pri})$.
-If for each $b \in \textit{Items}(e,\textit{pri})$, the rightmost gap-point of $a_1$ is after $\textit{call}(\textit{put},b,\textit{pri})$ and $\textit{call}(\textit{rm},b)$.
+\item[-] Let $e_{p}$ be the projection of $e$ into operations of priority $p$.
+Since each single-priority execution has FIFO property, there exists sequence $l_{p}$, such that $e_{p} \sqsubseteq l_{p}$, and when we treat $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$, $l_{p}$ belongs to queue.
+
+\item[-] Let $a_1$ be the last inserted item of $l_{p}$.
+
+    Step $1$: Check whether for each $b \in \textit{Items}(e,p)$, (1) $a_1$ does not $<_{\textit{pb}}$ to $b$, and (2) the right-most gap-point of $a$ is after $\textit{call}(\textit{put},b,p)$ and $\textit{call}(\textit{rm},b)$.
+
+    It is easy to see that $a_1$ is of priority $p$, and $a_1$ does not $<_{\textit{pb}}$ to any $b \in \textit{Items}(e,p)$.
+If for each $b \in \textit{Items}(e,p)$, the rightmost gap-point of $a_1$ is after $\textit{call}(\textit{put},b,p)$ and $\textit{call}(\textit{rm},b)$.
 Then by Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, we can obtain that $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
 
-\item[-] Otherwise, there exists $a_2 \in \textit{Items}(e,\textit{pri})$, such that the rightmost gap-point of $a_1$ is before $\textit{call}(\textit{put},a_2,\textit{pri})$ or $\textit{call}(\textit{rm},a_2)$ in $e$.
+\item[-] Otherwise, there exists $a_2 \in \textit{Items}(e,p)$, such that the rightmost gap-point of $a_1$ is before $\textit{call}(\textit{put},a_2,p)$ or $\textit{call}(\textit{rm},a_2)$ in $e$.
 We can see that each gap-point of $a_2$ is after the rightmost gap-point of $a_1$.%, and thus, the right-most gap-point of $a_2$ is after the rightmost gap-point of $a_1$.
@@
     \setlength{\itemsep}{0.5pt}
-    \item[-] If for each item $b \in \textit{Items}(e,\textit{pri})$, $a_2$ does not $<_{\textit{pb}}$ to $b$. Then we go to step $1$ and treat $a_2$ similarly as $a_1$.
-    \item[-] Otherwise, there exists $a_3$ with priority $\textit{pri}$ such that $a_2 <_{\textit{pb}}^* a_3$.
-
-    Since $l_{\textit{pri}}$ has FIFO property, it is easy to see that there is no cycle in $<_{\textit{pb}}$ order.
+    \item[-] If for each item $b \in \textit{Items}(e,p)$, $a_2$ does not $<_{\textit{pb}}$ to $b$. Then we go to step $1$ and treat $a_2$ similarly as $a_1$.
+    \item[-] Otherwise, there exists $a_3$ with priority $p$ such that $a_2 <_{\textit{pb}}^* a_3$.
+
+    Since $l_{p}$ has FIFO property, it is easy to see that there is no cycle in $<_{\textit{pb}}$ order.
 It is safe to assume that $a_3$ is maximal in the sense of $<_{\textit{pb}}^*$.
@@
 
-    By assumption,we know that the rightmost gap-point of $a_3$ is after $\textit{call}(\textit{put},a_2,\textit{pri})$ and $\textit{call}(\textit{rm},a_2)$.
+    By assumption,we know that the rightmost gap-point of $a_3$ is after $\textit{call}(\textit{put},a_2,p)$ and $\textit{call}(\textit{rm},a_2)$.
 Therefore, we can see that the rightmost gap-point of $a_3$ is after the rightmost gap-point of $a_1$.
@@
 \label{lemma:five enumeration is enough for EPQ1Equal}
-Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{=}$.
-Let $a$ and $b$ be items with maximal priority $\textit{pri}$.
-Assume that $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,\textit{pri})$ or $\textit{call}(\textit{rm},a)$.
+Given a data-differentiated $p$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{=}$.
+Let $a$ and $b$ be items with maximal priority $p$.
+Assume that $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$.
 Then, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists), where $a_1$ is the possible intermediate items for obtain $a <_{\textit{pb}}^* b$.
@@
 
-    Since $a <_{\textit{pb}}^A b$, $\textit{put}(a,\textit{pri}) <_{\textit{hb}} \textit{put}(b,\textit{pri})$.
-Since $o$ is after $\textit{call}(\textit{put},b,\textit{pri})$, and thus, after $\textit{call}(\textit{put},a,\textit{pri})$, we can see that $o$ is before $\textit{call}(\textit{rm},b)$.
+    Since $a <_{\textit{pb}}^A b$, $\textit{put}(a,p) <_{\textit{hb}} \textit{put}(b,p)$.
+Since $o$ is after $\textit{call}(\textit{put},b,p)$, and thus, after $\textit{call}(\textit{put},a,p)$, we can see that $o$ is before $\textit{call}(\textit{rm},b)$.
 
@@
     Since $a <_{\textit{pb}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{call}(\textit{rm},b)$.
-Since $o$ is after $\textit{call}(\textit{rm},b)$, we can see that $o$ is before $\textit{call}(\textit{put},a,\textit{pri})$.
-This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{call}(\textit{put},a,\textit{pri})$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible.
+Since $o$ is after $\textit{call}(\textit{rm},b)$, we can see that $o$ is before $\textit{call}(\textit{put},a,p)$.
+This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{call}(\textit{put},a,p)$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible.
 Therefore, we can safely ignore this reason.
@@
 
-    Since $a <_{\textit{pb}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{call}(\textit{put},b,\textit{pri})$.
-Since $o$ is after $\textit{call}(\textit{put},b)$, we can see that $o$ is before $\textit{call}(\textit{put},a,\textit{pri})$.
-This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{call}(\textit{put},a,\textit{pri})$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible.
+    Since $a <_{\textit{pb}}^B b$, $\textit{ret}(\textit{rm},a)$ is before $\textit{call}(\textit{put},b,p)$.
+Since $o$ is after $\textit{call}(\textit{put},b)$, we can see that $o$ is before $\textit{call}(\textit{put},a,p)$.
+This implies that $\textit{ret}(\textit{rm},a)$ is before $\textit{call}(\textit{put},a,p)$, and then $\textit{rm}(a) <_{\textit{hb}} \textit{put}(a)$, which is impossible.
 Therefore, we can safely ignore this reason.
@@
     Since $a_1 <_{\textit{pb}}^B b$, $\textit{rm}(a_1) <_{\textit{hb}} \textit{rm}(b)$, and $\textit{ret}(\textit{rm},a_1)$ is before $\textit{call}(\textit{rm},b)$.
-Since $\textit{rm}(a_1)$ does not happen before $\textit{put}(a_1)$, $\textit{call}(\textit{put},a_1,\textit{pri})$ is before $\textit{ret}(\textit{rm},a_1)$.
-Since $a <_{\textit{pb}}^A a_1$, $\textit{ret}(\textit{put},a,\textit{pri})$ is before $\textit{call}(\textit{put},a_1,\textit{pri})$.
-Therefore, $\textit{ret}(\textit{put},a,\textit{pri})$ is before $\textit{call}(\textit{rm},b)$.
+Since $\textit{rm}(a_1)$ does not happen before $\textit{put}(a_1)$, $\textit{call}(\textit{put},a_1,p)$ is before $\textit{ret}(\textit{rm},a_1)$.
+Since $a <_{\textit{pb}}^A a_1$, $\textit{ret}(\textit{put},a,p)$ is before $\textit{call}(\textit{put},a_1,p)$.
+Therefore, $\textit{ret}(\textit{put},a,p)$ is before $\textit{call}(\textit{rm},b)$.
 Since $\textit{call}(\textit{rm},b)$ is before $o$, we can see that $o$ is before $\textit{call}(\textit{rm},a)$.
@@
 Here we explicitly draw the leftmost gap-point of $a$ as $o'$.
-Since the position of $\textit{ret}(\textit{put},a_1,\textit{pri})$ and $\textit{call}(\textit{put},a,\textit{pri})$ do not influence the correctness, we can simply ignore it.
-We also ignore $\textit{call}(\textit{put},b,\textit{pri})$ and $\textit{ret}(\textit{put},b)$, since the only requirements of them are (1) $\textit{rm}(b)$ does not happen before $\textit{put}(b)$ and (2) $\textit{call}(\textit{put},b,\textit{pri})$ is before $o$.
+Since the position of $\textit{ret}(\textit{put},a_1,p)$ and $\textit{call}(\textit{put},a,p)$ do not influence the correctness, we can simply ignore it.
+We also ignore $\textit{call}(\textit{put},b,p)$ and $\textit{ret}(\textit{put},b)$, since the only requirements of them are (1) $\textit{rm}(b)$ does not happen before $\textit{put}(b)$ and (2) $\textit{call}(\textit{put},b,p)$ is before $o$.
 
@@
 
-    Since $a_1 <_{\textit{pb}}^A b$, $\textit{ret}(\textit{put},a_1)$ is before $\textit{call}(\textit{put},b,\textit{pri})$.
-Since $\textit{call}(\textit{put},b,\textit{pri})$ is before $o$, we can see that $\textit{ret}(\textit{put},a_1)$ is before $o$.
+    Since $a_1 <_{\textit{pb}}^A b$, $\textit{ret}(\textit{put},a_1)$ is before $\textit{call}(\textit{put},b,p)$.
+Since $\textit{call}(\textit{put},b,p)$ is before $o$, we can see that $\textit{ret}(\textit{put},a_1)$ is before $o$.
 
@@
 
-    \item[-] If $o$ is before $\textit{call}(\textit{put},a,\textit{pri})$: Since $\textit{rm}(a)$ does not happen before $\textit{put}(a,\textit{pri})$, we can see that $\textit{call}(\textit{put},a,\textit{pri})$ is before $\textit{ret}(\textit{rm},a)$, and then $o$ is before $\textit{ret}(\textit{rm},a)$.
+    \item[-] If $o$ is before $\textit{call}(\textit{put},a,p)$: Since $\textit{rm}(a)$ does not happen before $\textit{put}(a,p)$, we can see that $\textit{call}(\textit{put},a,p)$ is before $\textit{ret}(\textit{rm},a)$, and then $o$ is before $\textit{ret}(\textit{rm},a)$.
 Then similarly as above case, we can see that $o$ is before $\textit{call}(\textit{rm},a_1)$, and $a_1 <_{\textit{pb}}^A b$.
@@
 
-\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e)$, $\textit{last}(e')=\seqPQ_1^{=}$, $a$ and $b$ are two items with maximal priority $\textit{pri}$ in $e'$, $e'$ is a $\textit{pri}$-execution, $a <_{\textit{pb}}^* b$ in $e'$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,\textit{pri})$ or $\textit{call}(\textit{rm},a)$ in $e'$.
+\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e)$, $\textit{last}(e')=\seqPQ_1^{=}$, $a$ and $b$ are two items with maximal priority $p$ in $e'$, $e'$ is a $p$-execution, $a <_{\textit{pb}}^* b$ in $e'$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$ in $e'$.
 
@@
 let $e'' = e_2 \vert_{ \{ x,y,z,d_1,\ldots,d_m \} }$.
-It is obvious that $e'' \in \textit{proj}(e_2)$ is a $\textit{pri}$-execution, $\textit{last}(e'') = \seqPQ_1^{=}$.
-According to our construction of automata in $\textit{Auts}_{\textit{1-eq}}$, it is not hard to see that $x$ and $y$ has maximal priority in $h_2$, $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is before $\textit{call}(\textit{put},y,\textit{pri})$ or $\textit{call}(\textit{rm},y)$ in $e''$.
-
-\noindent The $\textit{if}$ direction: Assume that there exists $e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $last(e')=\textit{PQ}_1^{=}$, $a'$ and $b'$ are two items with maximal priority $\textit{pri}$ in $e'$, $e'$ is a $\textit{pri}$-execution, $a' <_{\textit{pb}}^* b'$ in $e'$, and the rightmost gap-point of $b'$ is before $\textit{call}(\textit{put},a',\textit{pri})$ or $\textit{call}(\textit{rm},a')$ in $e'$.
+It is obvious that $e'' \in \textit{proj}(e_2)$ is a $p$-execution, $\textit{last}(e'') = \seqPQ_1^{=}$.
+According to our construction of automata in $\textit{Auts}_{\textit{1-eq}}$, it is not hard to see that $x$ and $y$ has maximal priority in $h_2$, $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is before $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$ in $e''$.
+
+\noindent The $\textit{if}$ direction: Assume that there exists $e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $last(e')=\textit{PQ}_1^{=}$, $a'$ and $b'$ are two items with maximal priority $p$ in $e'$, $e'$ is a $p$-execution, $a' <_{\textit{pb}}^* b'$ in $e'$, and the rightmost gap-point of $b'$ is before $\textit{call}(\textit{put},a',p)$ or $\textit{call}(\textit{rm},a')$ in $e'$.
 By data-independence, we can obtain execution $e_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the items $d_1,\ldots,d_m$ that constitute the rightmost gap of $b'$, we rename them into $d$, (3) if $a' <_{\textit{pb}}^A a'_1 <_{\textit{pb}}^B b$, we rename $a'_1$ into $a_1$, and (4) rename the other items into $e$.
-It is easy to see that $\textit{last}(e_1) = \seqPQ_1^{=}$, $a$ and $b$ has maximal priority in $e_1$, $a <_{\textit{pb}}^* b$ in $e_1$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,\textit{pri})$ or $\textit{call}(\textit{rm},a)$ in $e_1$.
+It is easy to see that $\textit{last}(e_1) = \seqPQ_1^{=}$, $a$ and $b$ has maximal priority in $e_1$, $a <_{\textit{pb}}^* b$ in $e_1$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$ in $e_1$.
 By Lemma \ref{lemma:five enumeration is enough for EPQ1Equal}, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists).
@@
 \label{lemma:EPQ2Equal as happen before}
-Given a data-differentiated $\textit{pri}$-execution $e$ with $\textit{last}(e) = \seqPQ_2^{=}$.
-$e$ does not linearizable to $\textit{MS}(\seqPQ_2^{=})$, if and only if there exists $x$ and $y$ with priority $\textit{pri}$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,\textit{pri}) <_{\textit{hb}} \textit{put}(y,\textit{pri})$.
+Given a data-differentiated $p$-execution $e$ with $\textit{last}(e) = \seqPQ_2^{=}$.
+$e$ does not linearizable to $\textit{MS}(\seqPQ_2^{=})$, if and only if there exists $x$ and $y$ with priority $p$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,p) <_{\textit{hb}} \textit{put}(y,p)$.
 \end{restatable}
@@
 To prove the $\textit{only if}$ direction, we prove its contrapositive.
-Assume that for each pair of $x$ and $y$ with maximal priority in $e$, if $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, then $\textit{put}(x,\textit{pri})$ does not happen before $\textit{put}(y,\textit{pri})$.
+Assume that for each pair of $x$ and $y$ with maximal priority in $e$, if $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, then $\textit{put}(x,p)$ does not happen before $\textit{put}(y,p)$.
 We need to prove that $e \sqsubseteq \textit{MS}(\seqPQ_2^{=})$.
 
-Let $x_1,\ldots,x_m$ be the set of items with priority $\textit{pri}$ and has unmatched $\textit{put}$ in $e$, let $y_1,\ldots,y_n$ be the set of items with priority $\textit{pri}$ and has matched $\textit{put}$ and $\textit{rm}$ in $e$.
-By assumption, we know that $\textit{call}(\textit{put},y_i,\textit{pri})$ is before $\textit{ret}(\textit{put},x_j)$ for each $i,j$.
+Let $x_1,\ldots,x_m$ be the set of items with priority $p$ and has unmatched $\textit{put}$ in $e$, let $y_1,\ldots,y_n$ be the set of items with priority $p$ and has matched $\textit{put}$ and $\textit{rm}$ in $e$.
+By assumption, we know that $\textit{call}(\textit{put},y_i,p)$ is before $\textit{ret}(\textit{put},x_j)$ for each $i,j$.
 Then we explicitly construction the linearization of $e$ by locating the linearization points of $e$ as follows:
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] For each $x_i$, locate the linearization point of $\textit{put}(x_i,\textit{pri})$ just before its return action.
-
-\item[-] For each $y_j$, locate the lineariztion point of $\textit{put}(y_j,\textit{pri})$ jest after its call action.
+\item[-] For each $x_i$, locate the linearization point of $\textit{put}(x_i,p)$ just before its return action.
+
+\item[-] For each $y_j$, locate the lineariztion point of $\textit{put}(y_j,p)$ jest after its call action.
 
@@
 It is easy to see that $e \sqsubseteq l$.
-Since linearization points of $\textit{put}(x_i,\textit{pri})$ is after the linearization point of $\textit{put}(y_j,\textit{pri})$ for each $i,j$, it is easy to see that $l \in \textit{MS}(\seqPQ_2^{=})$.
+Since linearization points of $\textit{put}(x_i,p)$ is after the linearization point of $\textit{put}(y_j,p)$ for each $i,j$, it is easy to see that $l \in \textit{MS}(\seqPQ_2^{=})$.
 This completes the proof of this lemma.
@@
 \label{lemma:EPQ2Equal is always co-regular}
-Given a data-differentiated $\textit{pri}$-execution $e$, if $\textit{last}(e) = \seqPQ_2^{=}$, then $e \sqsubseteq \textit{MS}(\seqPQ_2^{=})$.
-\end{restatable}
-
-\begin {proof}
-
-According to Lemma \ref{lemma:EPQ2Equal as happen before}, if $\textit{last}(e) = \seqPQ_2^{=}$ and $e$ does not linearizable to $\textit{MS}(\seqPQ_2^{=})$, then there exists $x$ and $y$ with priority $\textit{pri}$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$, and $\textit{put}(x,\textit{pri}) <_{\textit{hb}} \textit{put}(y,\textit{pri})$.
+Given a data-differentiated $p$-execution $e$, if $\textit{last}(e) = \seqPQ_2^{=}$, then $e \sqsubseteq \textit{MS}(\seqPQ_2^{=})$.
+\end{restatable}
+
+\begin {proof}
+
+According to Lemma \ref{lemma:EPQ2Equal as happen before}, if $\textit{last}(e) = \seqPQ_2^{=}$ and $e$ does not linearizable to $\textit{MS}(\seqPQ_2^{=})$, then there exists $x$ and $y$ with priority $p$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$, and $\textit{put}(x,p) <_{\textit{hb}} \textit{put}(y,p)$.
 Let $e_1 = e \vert_{ \{ x,y \} }$.

### 2017-04-21 checking-executions.tex
modify

@@
 
-Proving soundness is highly non-trivial and one of the main technical contributions of this paper (see Appendix~\ref{} for a complete proof).
+Proving soundness is highly non-trivial and one of the main technical contributions of this paper (see Appendix~\ref{sec:appendix subsection proof of lemma con-check-EPQ is correct} for a complete proof).
 The main technical difficulty is showing that for any execution $e$, any linearization of $e\setminus x$ for some maximal priority value $x$ can be extended to a linearization of $e$ provided that $\mathsf{UnmatchedMaxPriority}$ or $\mathsf{MatchedMaxPriority}$ holds (depending on whether there are values with the same priority as $x$ in $e$ which are not removed).

### 2017-04-21 co-regular-EPQ.tex
modify

@@
 
-Section~\ref{ssec:aut} describes a $\mathsf{MatchedMaxPriority}$-complete automaton, the other automata being defined in Appendix~\ref{} or Icalp??. Therefore, the following holds.
+Section~\ref{ssec:aut} describes a $\mathsf{MatchedMaxPriority}$-complete automaton, the other automata being defined in Appendix~\ref{sec:appendix lemma and register automata for FIFO of single-priority executions}, %Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar},
+%Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal} 
+and Appendix \ref{subsec:co-regular of EPQ3}. Therefore, the following holds.
 
@@
 $\seqPQ$).
-This property can be checked separately using the techniques introduced for regular FIFO queues in~\cite{DBLP:conf/icalp/BouajjaniEEH15} (see Appendix~\ref{} for more details).
+This property can be checked separately using the techniques introduced for regular FIFO queues in~\cite{DBLP:conf/icalp/BouajjaniEEH15} (see Appendix~\ref{sec:appendix lemma and register automata for FIFO of single-priority executions} for more details).
 Note that this assumption excludes some obvious violations, such as a $\textit{rm}(a)$ operation happens before a $\textit{put}(a,p)$ operation, for some $p$.
@@
 
-The $\textit{if}$ direction follows from the fact that if the projection of an execution to a set of operations $O_1$ has a linearization $l_1$ and the projection of the same execution to the remaining set of operations has a linearization $l_2$, then the execution has a linearization which is defined as an interleaving of $l_1$ and $l_2$ (see Appendix~\ref{} for more details).
+The $\textit{if}$ direction follows from the fact that if the projection of an execution to a set of operations $O_1$ has a linearization $l_1$ and the projection of the same execution to the remaining set of operations has a linearization $l_2$, then the execution has a linearization which is defined as an interleaving of $l_1$ and $l_2$ (see Appendix~\ref{sec:appendix proof of Lemma pri execution is enough} for more details).
 
@@
 The execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} is not $\mathsf{MatchedMaxPriority}^>$-linearizable because the left-right constraint of the maximal priority value $b$ contains a cycle: $f \rightarrow e \rightarrow d \rightarrow b \rightarrow f$.
-We prove that the presence of such a cycle is equivalent to the execution not being $\mathsf{MatchedMaxPriority}^>$-linearizable (see Appendix~\ref{}).
+We prove that the presence of such a cycle is equivalent to the execution not being $\mathsf{MatchedMaxPriority}^>$-linearizable (see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}).
 %The existence of cycle though value with maximal priority in left-right constraint formally modelled the phenomenon in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
@@
 
-{\color {blue} Assume that $x$ is covered by $d_1,\ldots,d_m$, we can safely rename $x$ into $b$, rename $d_1,\ldots,d_m$ into $a$ and rename all other value into $c$ by data-independence.
-Since there are four cases of projection of execution to $x$ when their is a cycle through $x$.
-To ease the reading, for each of such case we generate a register automata, and then let the $\mathsf{MatchedMaxPriority}^{>}$-complete automaton be the union of the four register automata.
-For example, for the case when the projection of $e$ to $x$ is as in \figurename~\ref{fig:introduce gap for EPQ1Lar} (rename $b$ into $x$), we construct a register automaton $\mathcal{A}_{\textit{l-lar}}^1$ in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
-Here transition labels $c_1 = c + \textit{ret}(\textit{rm},a)$, $c_2 = c + \textit{call}(\textit{put},a,=r)$, $c_3 = c_2 + \textit{ret}(\textit{rm},a)$, where $c = \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.
-There are three paths in \figurename~\ref{fig:introduce gap for EPQ1Lar} from $q_{\textit{init}}$ to $q_7$, and they comes from that there are three possible positions for the first $\textit{call}(\textit{put},a,\_)$ in the captured execution, which is shown in \figurename~\ref{fig:executions APQ1Lar-1 in paper}.
-In \figurename~\ref{fig:automata APQ1Lar-1 in paper}, the paths $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_{10}$ and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$ corresponds to the first, second and third cases in \figurename~\ref{fig:executions APQ1Lar-1 in paper}, respectively.
+{\color {blue} 
+Defining a $\mathsf{MatchedMaxPriority}^>$-complete automaton needs to capture the executions when a value $x$ is covered by values with smaller priority.
+By data-independence, we rename $x$ into $b$, rename the items covers it into $a$ and rename all other value into $\top$.
+There are four possibility of projection of captured execution into $b$, and for clarity we construct a register automaton for each of these case.
+For example, one such case is as in \figurename~\ref{fig:introduce gap for EPQ1Lar} (consider $b$ as $x$), and the corresponding register automaton is shown in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
+There are three possible place of the first $\textit{call}(\textit{put},a,\_)$ in the captured execution, as shown in \figurename~\ref{fig:executions APQ1Lar-1 in paper}.
+These three possible places correspond to three paths in $\mathcal{A}_{\textit{l-lar}}^1$: $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_{10}$ and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$.
+More details can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
 }
@@
 %\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{l-lar}}^1$}
+  \caption{Register automaton $\mathcal{A}_{\textit{l-lar}}^1$.
+We use the following notations: $C_1 = C \cup \{ \textit{ret}(\textit{rm},a) \}$, $C_2 = C \cup \{ \textit{call}(\textit{put},a,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.}
   \label{fig:automata APQ1Lar-1 in paper}
@@
 In this case, the conditions in Lemma~\ref{lemma:EPQ1Equal as pb order and gap-point} are equivalent to the fact that intuitively, the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$ is covered by lower priority values (and thus, there is no gap-point of $b$ which occurs after $\textit{call}(\textit{rm},a)$).
-Using again the data-independence property, these lower priority values can all be renamed to a fixed value $d$, and the other values to a fixed value $e$.
+Using again the data-independence property, these lower priority values can all be renamed to a fixed value $d$, and the other values to a fixed value $\top$.
 
@@
   \caption{A case in deriving a $\mathsf{MatchedMaxPriority}^{=}$-complete automaton.
-We use the following notations: $c = \textit{call}(\textit{put},e,\top),\textit{ret}(\textit{put},e,\top)$, $\textit{call}(\textit{rm},e), \textit{ret}(\textit{rm},e),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{call}(\textit{put},d,<r)$, $c_2 = c_1 + \textit{ret}(\textit{put},b,r)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b,r) + \textit{ret}(\textit{rm},d)$.}
+We use the following notations: $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true})$, $\textit{call}(\textit{rm},\top), \textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_1 = C \cup \textit{call}(\textit{put},d,<r) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},d) \}$, $C_4 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},d) \}$.}
   \label{fig:an enumeration and its witness automaton}

### 2017-04-22 co-regular-EPQ.tex
sec 4.1.1.

@@
 USE BIG CAPS FOR SETS OF TRANSITION LABELS.
-TODO COMPLETE THE FOLLOWING
-Defining a $\mathsf{MatchedMaxPriority}^>$-complete automaton needs to consider ??
-cases depending on the order between ??.
-For each case, we define an automaton recognizing the induced set of violations.
-For example, the automaton corresponding to the case ??
+To define a $\mathsf{MatchedMaxPriority}^>$-complete automaton we need to consider several cases depending on the order between the call/return actions of the $\textit{put}$/$\textit{rm}$ operations that add and respectively, remove the value $b$.
+For example, the case where the put happens-before the remove (as in \figurename~\ref{fig:introduce gap for EPQ1Lar}) is pictured in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
+This automaton captures the three possible ways of ordering the first action $\textit{ret}(\textit{put},a,\_)$ w.r.t.
+the actions with value $b$, which are pictured in \figurename~\ref{fig:executions APQ1Lar-1 in paper} (this action cannot occur after $\textit{call}(\textit{rm},b,\_)$ since $b$ must be covered by the $a$s).
+The paths corresponding to these three possible orders are: $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_{10}$ and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$.
+Automata for the other cases can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
+
+%For each case, we define an automaton recognizing the induced set of violations. 
+%For example, the automaton corresponding to the case ??
 is pictured in in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
 More details can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
-
-{\color {blue} 
-Defining a $\mathsf{MatchedMaxPriority}^>$-complete automaton needs to capture the executions when a value $x$ is covered by values with smaller priority.
+%
+%
+%{\color {blue} 
+%Defining a $\mathsf{MatchedMaxPriority}^>$-complete automaton needs to capture the executions when a value $x$ is covered by values with smaller priority.
 By data-independence, we rename $x$ into $b$, rename the items covers it into $a$ and rename all other value into $\top$.
@@
 More details can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
-}
+%}
 

### 2017-04-22 checking-executions.tex
example

@@
 When there are no more $\textit{rm}(\textit{empty})$ operations, the procedure selects a $\textit{put}$ operation adding a value with maximal priority which is not removed, and then a pair of $\textit{put}$ and $\textit{rm}$ operations adding and removing the same maximal priority value.
-For instance, since $p_2$ is a maximal priority, it may select the operation $\textit{put}(d,p_2)$.
+For instance, since $p_2$ is a maximal priority, it selects the operation $\textit{put}(d,p_2)$.
 {This operation is correct if $d$ is the last value with priority $p_2$, and the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
@@
 \end{align*}
-Since there is no other value of priority $p_2$ which is not removed, the procedure selects a pair of $\textit{put}$/$\textit{rm}$ operations with an argument of priority $p_2$,} for instance, $\textit{put}(c,p_2)$ and $\textit{rm}(c)$.
+Since there is no other value of maximal priority which is not removed, the procedure selects a pair of $\textit{put}$/$\textit{rm}$ operations with an argument of maximal priority $p_2$,} for instance, $\textit{put}(c,p_2)$ and $\textit{rm}(c)$.
 The value returned by $\textit{rm}$ is correct if all the values of priority smaller than $p_2$ added before $\textit{rm}(c)$ are also removed before $\textit{rm}(c)$.
@@
 
-The following lemma states the correctness of $\textit{Check-PQ-Seq}$ (the proof can be found in Appendix~\ref{sec:appendix definition of seqPQ and proof of Lemma EQP rules and semantics}).
+The following lemma states the correctness of $\textit{Check-PQ-Seq}$ (see Appendix~\ref{sec:appendix definition of seqPQ and proof of Lemma EQP rules and semantics} for the proof).
 
@@
 
-Let us briefly explain the idea of proving this property when $\mathsf{MatchedMaxPriority}$ holds with an example.
-Assume $e \sqsubseteq l=u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, $\mathsf{MatchedMaxPriority}(l,x)$ holds, and $e \setminus x \sqsubseteq l' \in \textit{PQ}$, as shown in \figurename~\ref{fig:concurrent execution for EPQ1} (a), we explicitly construct a sequence $l''= l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3$ and prove that $e \sqsubseteq l'' \in \textit{PQ}$.
-In this example, $p_1 \prec p$, $p_1 \prec p_2$, $u=\epsilon$, $w$ contains $\textit{rm}(z_2), \textit{put}(x_3,p_2), \textit{rm}(z_1)$ (emphasize by $\textit{rm}(z_2)-w$), and the remanning operations are in $v$.
+We explain the proof of this property on the execution $e$ in \figurename~\ref{fig:concurrent execution for EPQ1}(a) where $p_1 \prec p$, $p_1 \prec p_2$, and the predicate $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds.
+Assume that there exist two sequential executions $l$ and $l'$ such that $e \sqsubseteq l=u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds, and $e \setminus x \sqsubseteq l' \in \seqPQ$.
+Let $u=\epsilon$, $w$ be any sequence containing the set of operations $\textit{rm}(z_2)$, $\textit{put}(x_3,p_2)$, and $\textit{rm}(z_1)$ (we distinguish them by adding the suffix $-w$ to their name, e.g., $\textit{rm}(z_2)-w$), and $v$ be any sequence containing the remaining operations.
+In general, the linearization $l'$ can be defined by choosing for each operation, a point in time between its call and return actions, called \emph{linearization point}.
+The order between the linearization points defines the sequence $l'$.
+\figurename~\ref{fig:concurrent execution for EPQ1}(a) draws linearization points for the operations in $e \setminus x$ defining the linearization $l'$~\footnote{In general, there may exist multiple ways of choosing linearization points to define the same linearization.
+Our construction is agnostic to this choice.}.
+%(the order of these linearization points defines the sequence $l'$). 
+We show how to construct a sequence $l''= l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3\in\seqPQ$ such that $e \sqsubseteq l''$. 
+%In this example, .
 We also explicitly draw the linearization points according to $l'$.
 We construct $l''$ as follows, and this method does not rely on the locations of linearization points of $l'$ and operations of $x$.
-
 \begin{itemize}
 \setlength{\itemsep}{0.5pt}
-\item[-] Let $p$-comparable operations (resp., $p$-incomparable operations) be the set of operations with items whose priority is comparable with $p$ (resp., incomparable with $p$).
-It seems that we can construct $l''_1$, $l''_2$ and $l''_3$ as the projection of $l'$ into operations of $u$, $v$ and $w$, respectively.
-However, this is incorrect, since $\mathsf{MatchedMaxPriority}(l,x)$ has no restriction to $p$-incomparable operations operations in $u \cdot v$, and thus, there is no guarantee that the projection of $l'$ into $p$-incomparable operations operations in $u \cdot v$ being correct.
-In this example, such projection is $\textit{put}(z_1,p_3) \cdot \textit{rm}(z_2)$ and is incorrect.
+\item[-] An operation is called $p$-comparable (resp., $p$-incomparable) when it receives as argument a value of priority comparable to $p$ (resp., incomparable to $p$).
+We could try to define $l''_1$, $l''_2$ and $l''_3$ as the projection of $l'$ to the operations in $u$, $v$ and $w$, respectively.
+However, this is incorrect, since $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ imposes no restriction on  $p$-incomparable operations in $u \cdot v$, and thus, there is no guarantee that the projection of $l'$ to $p$-incomparable operations in $u \cdot v$ is correct.
+In this example, this projection is $\textit{put}(z_1,p_3) \cdot \textit{rm}(z_2)$ which is incorrect.
 
-\item[-] Let us construct sets $U'$, $V'$ and $W'$, such then make $l''_1$, $l''_2$ and $l''_3$ the projections of $l'$ into $U'$, $V'$ and $W'$, respectively. This contains two steps:
+\item[-] We define the sets of operations $U'$, $V'$ and $W'$ such that $l''_1$, $l''_2$ and $l''_3$ are the projections of $l'$ to $U'$, $V'$, and $W'$, respectively. This is done in two steps:
 
 \item[-] The first step is to define $W'$.
-The $p$-comparable operations in $W'$ is same as that in $w$.
-To obtain $p$-incomparable operations in $W'$, we try to find a $p$-incomparable operation $o$ which either happens before some $p$-comparable operations in $w$, or with linearization points after $\textit{rm}(x)$.
-Then, we put $o$ and all $p$-incomparable operations after $o$ in $l'$ into $W'$ (emphasized by boxes in example).
-In this example, $o$ is $\textit{rm}(z_1)$, and $p$-incomparable operations after $o$ in $l'$ is $\textit{rm}(z_2)$, which are emphasized by adding boxes.
-In this process, whether a $p$-incomparable operation is in $W'$ or not only relies whether it is before or after $o$ in $l'$.
+The $p$-comparable operations in $W'$ are the same as in $w$.
+To identify the $p$-incomparable operations in $W'$, we search for a $p$-incomparable operation $o$ which either happens before some $p$-comparable operation in $w$, or whose linearization point occurs after $\textit{rm}(x)$.
+We add to $W'$ the operation $o$ and all the $p$-incomparable operations occurring after $o$ in $l'$.
+In this example, $o$ is $\textit{rm}(z_1)$ and the only $p$-incomparable operation occurring after $o$ in $l'$ is $\textit{rm}(z_2)$ (they are surrounded by boxes in the figure).
+In this process, whether a $p$-incomparable operation is in $W'$ or not only relies on whether it is before or after $o$ in $l'$.
 
 \item[-] The second step is to define $U'$ and $V'$.
-$U'$ contains the following two kinds of operations: (1) Operations whose linearization points are before $\textit{ret}(\textit{put},x)$, and (2) other $\textit{put}$ operations with priority $p$.
-$V'$ contains the remanning operations.
+$U'$ contains two kinds of operations: (1) operations whose linearization points are before $\textit{ret}(\textit{put},x)$, and (2) other $\textit{put}$ operations with priority $p$.
+$V'$ contains the remaining operations.
 In this example, $U'$ contains $\textit{put}(z_1,p_3)$ and $\textit{put}(x_2,p_2)$.
 
-\item[-] In this example, $l''_1 = \textit{put}(z_1,p_2) \cdot \textit{put}(x_2,p)$, $l''_2 = \textit{put}(z_2,p_2) \cdot \textit{rm}(x_2) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(y_1)$, and $l''_3 = \textit{put}(x_3,p) \cdot \textit{rm}(z_1) \cdot \textit{rm}(z_2)$.
-In \figurename~\ref{fig:concurrent execution for EPQ1} (b), we add linarization points according to $l''$, and we can see that $l''$ holds as required.
+\item[-] In conclusion, we have that $l''_1 = \textit{put}(z_1,p_2) \cdot \textit{put}(x_2,p)$, $l''_2 = \textit{put}(z_2,p_2) \cdot \textit{rm}(x_2) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(y_1)$, and $l''_3 = \textit{put}(x_3,p) \cdot \textit{rm}(z_1) \cdot \textit{rm}(z_2)$.
+\figurename~\ref{fig:concurrent execution for EPQ1}(b) draws linearization points for each operation in $e$ defining the linearization $l''$.
 \end{itemize}
@@
 %\end {proof}
-
 \begin{figure}[t]

### 2017-04-22 checking-executions.tex
modify

@@
 Assume that there exist two sequential executions $l$ and $l'$ such that $e \sqsubseteq l=u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds, and $e \setminus x \sqsubseteq l' \in \seqPQ$.
-Let $u=\epsilon$, $w$ be any sequence containing the set of operations $\textit{rm}(z_2)$, $\textit{put}(x_3,p_2)$, and $\textit{rm}(z_1)$ (we distinguish them by adding the suffix $-w$ to their name, e.g., $\textit{rm}(z_2)-w$), and $v$ be any sequence containing the remaining operations.
+Let $u=\epsilon$, $w$ be any sequence containing the set of operations $\textit{put}(z_2,p_2)$, $\textit{put}(x_3,p)$ and $\textit{rm}(z_1)$ (we distinguish them by adding the suffix $-w$ to their name, e.g., $\textit{rm}(z_1)-w$), and $v$ be any sequence containing the remaining operations.
 In general, the linearization $l'$ can be defined by choosing for each operation, a point in time between its call and return actions, called \emph{linearization point}.
@@
 However, this is incorrect, since $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ imposes no restriction on  $p$-incomparable operations in $u \cdot v$, and thus, there is no guarantee that the projection of $l'$ to $p$-incomparable operations in $u \cdot v$ is correct.
-In this example, this projection is $\textit{put}(z_1,p_3) \cdot \textit{rm}(z_2)$ which is incorrect.
+In this example, this projection is $\textit{put}(z_1,p_2) \cdot \textit{rm}(z_2)$ which is incorrect.
 
@@
 \item[-] The second step is to define $U'$ and $V'$.
-$U'$ contains two kinds of operations: (1) operations whose linearization points are before $\textit{ret}(\textit{put},x)$, and (2) other $\textit{put}$ operations with priority $p$.
+$U'$ contains two kinds of operations: (1) operations whose linearization points are before $\textit{ret}(\textit{put},x,p)$, and (2) other $\textit{put}$ operations with priority $p$.
 $V'$ contains the remaining operations.
-In this example, $U'$ contains $\textit{put}(z_1,p_3)$ and $\textit{put}(x_2,p_2)$.
+In this example, $U'$ contains $\textit{put}(z_1,p_2)$ and $\textit{put}(x_2,p)$.
 

### 2017-04-22 co-regular-EPQ.tex
some stuff

@@
 
-Section~\ref{ssec:aut} describes a $\mathsf{MatchedMaxPriority}$-complete automaton, the other automata being defined in Appendix~\ref{sec:appendix lemma and register automata for FIFO of single-priority executions}, %Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar},
+Section~\ref{ssec:aut} describes a $\mathsf{MatchedMaxPriority}$-complete automaton, the other automata being defined in Appendix~\ref{sec:appendix lemma and register automata for FIFO of single-priority executions} %Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar},
 %Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal} 
@@
 This is possible because any data-differentiated execution which is not $\Gamma$-linearizable has such a projection.
-Formally, given a data-differentiated execution $e$ and $p$ a maximal priority in $e$, $e\vert_{\preceq p}$ is the projection of $e$ to the set of values with priorities comparable to $p$. Then,
+Formally, given a data-differentiated execution $e$ and $p$ a maximal priority in $e$, $e\vert_{\preceq p}$ is the projection of $e$ to the set of values with priorities smaller than $p$. Then,
 
@@
 
-The following result shows that $\Gamma$-complete automata enable an effective reduction of checking linearizability of concurrent priority queue implementations to state reachability.
+The following shows that $\Gamma$-complete automata enable an effective reduction of checking linearizability of concurrent priority queue implementations to state reachability.
 It is a direct consequence of the above definitions.
+Section~\ref{subsec:combine step-by-step linearizability and co-regular} discusses decidability results implied by this reduction.
 
@@
 \label{lemma:reduce EPQ into state reachability}
-Let $\mathcal{I}$ be a data-independent implementation, and $A(\Gamma)$ a $\Gamma$-complete automaton for each $\Gamma$. Then,
+Let $\mathcal{I}$ be a data-independent implementation, and $A(\Gamma)$ be a $\Gamma$-complete automaton for each $\Gamma$. Then,
 $\mathcal{I} \sqsubseteq \seqPQ$ if and only if $\mathcal{I} \cap A(\Gamma) = \emptyset$ for all $\Gamma$.
 \end{restatable}
-
-Section~\ref{subsec:combine step-by-step linearizability and co-regular} discusses decidability results implied by the theorem above.
 
@@
 We denote by $\mathsf{MatchedMaxPriority}^{>}$ the strengthening of $\mathsf{MatchedMaxPriority}$ with the condition that all the values other than $x$ have a priority strictly smaller than $p$ (corresponding to the first case), and by $\mathsf{MatchedMaxPriority}^{=}$ the strengthening of the same formula with the negation of this condition (corresponding to the second case).
-We use particular instances of register automata~\cite{finite-memory, the paper I found} whose states include only two registers, one for storing a priority guessed at the initial state, and one for storing the priority of the current action in the execution.
+We use particular instances of register automata~\cite{DBLP:journals/tcs/KaminskiF94,DBLP:conf/icalp/Cerans94,DBLP:conf/stacs/SegoufinT11} whose states include only two registers, one for storing a priority guessed at the initial state, and one for storing the priority of the current action in the execution.
 The transitions can check equality or the order relation $\prec$ between the values stored in the two registers.
@@
 	\item $\textit{call}(\textit{put},d,g)$ where $g\in\{=r,\prec r,true\}$ is a guard, for reading a call action $\textit{call}(\textit{put},d,p)$ of a put and checking whether $p$ is either equal to or smaller than the value stored in $r$, or arbitrary,
-	\item $\textit{ret}(\textit{put},d,true)$ for reading a return action $\textit{ret}(\textit{put},d,p)$ for some $p$.
+	\item $\textit{ret}(\textit{put},d,true)$ for reading a return action $\textit{ret}(\textit{put},d,p)$ for any $p$.
 \end{itemize}
@@
 The execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} is not $\mathsf{MatchedMaxPriority}^>$-linearizable because the left-right constraint of the maximal priority value $b$ contains a cycle: $f \rightarrow e \rightarrow d \rightarrow b \rightarrow f$.
-We prove that the presence of such a cycle is equivalent to the execution not being $\mathsf{MatchedMaxPriority}^>$-linearizable (see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}).
+The presence of such a cycle is equivalent to the execution not being $\mathsf{MatchedMaxPriority}^>$-linearizable (see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}).
 %The existence of cycle though value with maximal priority in left-right constraint formally modelled the phenomenon in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
@@
 However, by data independence, whenever an implementation contains such an execution it also contains an execution where all the values $d_1$,$\ldots$,$d_n$ are renamed to the same value $a$, and $x$ is renamed to $b$.
-Therefore, our automata can be defined over a fixed set of values $a$, $b$, and $\top$ (for the operations outside of the non-linearizable projection).
+Therefore, our automata can be defined over a fixed set of values $a$, $b$, and $\top$ (recall that $\top$ is used for the operations outside of the non-linearizable projection).
 
@@
 This automaton captures the three possible ways of ordering the first action $\textit{ret}(\textit{put},a,\_)$ w.r.t.
-the actions with value $b$, which are pictured in \figurename~\ref{fig:executions APQ1Lar-1 in paper} (this action cannot occur after $\textit{call}(\textit{rm},b,\_)$ since $b$ must be covered by the $a$s).
+the actions with value $b$, which are pictured in \figurename~\ref{fig:executions APQ1Lar-1 in paper} (this action cannot occur after $\textit{call}(\textit{rm},b,\_)$ since $b$ must be covered by the $a$-s).
 The paths corresponding to these three possible orders are: $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_{10}$ and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$.
@@
 The case of \figurename~\ref{fig:introduce pb order} can be formally described as follows: $a <_{\textit{pb}}^* b$ while the right-most gap-point of $b$ is before $\textit{call}(\textit{rm},a)$ or $\textit{call}(\textit{put},a,p_4)$.
-The following lemma states that these conditions are enough to characterize non-linearizability w.r.t $\mathsf{MatchedMaxPriority}^{=}$.
+The following lemma states that these conditions are enough to characterize non-linearizability w.r.t $\mathsf{MatchedMaxPriority}^{=}$ (see Appendix\ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}).
 
@@
 
-\begin {proof} (Sketch)
-We have already intuitively explain the proof of the $\textit{if}$ direction.
+%\begin {proof} (Sketch)
+%We have already gave an intuitive explanation for the $\textit{if}$ direction.
 To prove the $\textit{only if}$ direction, we prove its contrapositive.
@@
 %Or we can say, we need to explicitly construct linearization of $e$.
-
-Let $\textit{Items}(e,p)$ be the set of values with priority $p$ in $e$.
-We introduce another lemma (Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal} in Appendix), which states that if we can find a value $\exists g_1 \in \textit{Items}(e,p)$, such that $\forall g_2 \in \textit{Items}(e,p)$, (1) $g_1 \nless_{\textit{pb}} g_2$, and (2) the right-most gap-point of $g_1$ is after $\textit{call}(\textit{put},g_2,p)$ and $\textit{call}(\textit{rm},g_2)$.
+%
+%Let $\textit{values}(e,p)$ be the set of values with priority $p$ in $e$.
+We introduce another lemma (Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal} in Appendix), which states that if we can find a value $\exists g_1 \in \textit{values}(e,p)$, such that $\forall g_2 \in \textit{values}(e,p)$, (1) $g_1 \nless_{\textit{pb}} g_2$, and (2) the right-most gap-point of $g_1$ is after $\textit{call}(\textit{put},g_2,p)$ and $\textit{call}(\textit{rm},g_2)$.
 Then $e$ is linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$.
-The remaining of our proof is a loop for searching such $g_1$ in $e$.
+%The remaining of our proof is a loop for searching such $g_1$ in $e$.
 Let $e_p$ be the projection of $e$ into values of priority $p$, and let $l_p$ be the linearization of $e$ which satisfy the requirements of queue.
 Let the initial value of $a_1$ be the last inserted value in $l_p$.
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] check if $a_1$ satisfy the conditions of Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}. If it does, then this loop terminates.
-
-\item[-] Otherwise, $\exists a_2 \in \textit{Items}(e,p)$, such that the rightmost gap-point of $a_1$ is before $\textit{cal}(\textit{put},a_2,p)$ or $\textit{cal}(\textit{rm},a_2)$ in $e$.
+%
+%\begin{itemize}
+%\setlength{\itemsep}{0.5pt}
+%\item[-] check if $a_1$ satisfy the conditions of Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}. If it does, then this loop terminates.
+%
+%\item[-] Otherwise, $\exists a_2 \in \textit{values}(e,p)$, such that the rightmost gap-point of $a_1$ is before $\textit{cal}(\textit{put},a_2,p)$ or $\textit{cal}(\textit{rm},a_2)$ in $e$.
 By assumption, we know that $a_2 \nless_{\textit{pb}} a_1$, and each gap-point of $a_2$ is after the rightmost gap-point of $a_1$.
-
-    \begin{itemize}
-    \setlength{\itemsep}{0.5pt}
-    \item[-] If $\forall b \in \textit{Items}(e,p)$, $a_2\not<_{\textit{pb}}b$. Then we go to next round of the loop and make $a_2$ to be the ``new-round $a_1$''.
-    \item[-] Otherwise, there exists $a_3 \in \textit{Items}(e,p)$ such that $a_2 <_{\textit{pb}}^* a_3$.
+%
+%    \begin{itemize}
+%    \setlength{\itemsep}{0.5pt}
+%    \item[-] If $\forall b \in \textit{values}(e,p)$, $a_2\not<_{\textit{pb}}b$. Then we go to next round of the loop and make $a_2$ to be the ``new-round $a_1$''.
+%    \item[-] Otherwise, there exists $a_3 \in \textit{values}(e,p)$ such that $a_2 <_{\textit{pb}}^* a_3$.
 It is not hard to see that there is no cycle in $<_{\textit{pb}}$, and it is safe to assume that $a_3$ is maximal w.r.t $<_{\textit{pb}}^*$.
-
-        By assumption,we know that the rightmost gap-point of $a_3$ is after $\textit{call}(\textit{put},a_2,p)$ and $\textit{call}(\textit{rm},a_2)$.
+%
+%        By assumption,we know that the rightmost gap-point of $a_3$ is after $\textit{call}(\textit{put},a_2,p)$ and $\textit{call}(\textit{rm},a_2)$.
 Therefore, we can see that the rightmost gap-point of $a_3$ is after the rightmost gap-point of $a_1$.
 Then we go to next round of the loop and make $a_3$ to be the ``new-round $a_1$''.
-    \end{itemize}
-\end{itemize}
-
-Let $a^i$ be the $a_1$ in the $\textit{i-th}$ round of loop.
+%    \end{itemize}
+%\end{itemize}
+%
+%Let $a^i$ be the $a_1$ in the $\textit{i-th}$ round of loop.
 We can see that the rightmost gap-point of $a^j$ is after the rightmost gap-point of $a^i$ for each $i<j$.
@@
 This implies that $e$ is linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ and completes the proof of the $\textit{only if}$ direction.
-\end {proof}
+%\end {proof}
 

### 2017-04-22 appendix-EPQ.tex

@@
 
-Let $l'_{\textit{u'v'}}$ be the projection of $l'$ into $U' \cup V'$, let $O_x$ be the set of $\textit{put}(\_,p)$ while the item is not $x$ in $h$.
+Let $l'_{\textit{u'v'}}$ be the projection of $l'$ into $U' \cup V'$, let $O_x$ be the set of $\textit{put}(\_,p)$ while the value is not $x$ in $h$.
 Let $l''_a$ be the longest prefix of $l'_{\textit{u'v'}}$, where linearization of each operation of $l''_a$ is before $\textit{ret}(\textit{put},b)$ in $e_{\textit{lp}}$.
@@
 
-Given an execution of queue, we say that it is differentiated, if each item is enqueued at most once.
+Given an execution of queue, we say that it is differentiated, if each value is enqueued at most once.
 \cite{DBLP:conf/icalp/BouajjaniEEH15} states that, given a differentiated queue execution $e$ without $\textit{deq}(\textit{empty})$, $e$ is not linearizable with respect to queue, if one of the following cases holds for some $a,b$: (1) $\textit{deq}(b) <_{hb} \textit{enq}(b)$, (2) there are no $\textit{enq}(b)$ and at least one $\textit{deq}(b)$, (3) there are one $\textit{enq}(b)$ and more than one $\textit{deq}(b)$, and (4) $\textit{enq}(a) <_{\textit{hb}} \textit{enq}(b)$, and $\textit{deq}(b) <_{\textit{hb}} \textit{deq}(a)$, or $\textit{deq}(a)$ does not exists.
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] If $e_0$ is accepted by $\mathcal{A}_{\textit{SinPri}}^1$, $\mathcal{A}_{\textit{SinPri}}^2$ or $\mathcal{A}_{\textit{SinPri}}^3$: Then remove all items that are not renamed into $b$ by $r$.
-
-\item[-] If $e_0$ is accepted by $\mathcal{A}_{\textit{SinPri}}^4$: Then remove all items that are not renamed into $a$ or $b$ by $r$.
+\item[-] If $e_0$ is accepted by $\mathcal{A}_{\textit{SinPri}}^1$, $\mathcal{A}_{\textit{SinPri}}^2$ or $\mathcal{A}_{\textit{SinPri}}^3$: Then remove all values that are not renamed into $b$ by $r$.
+
+\item[-] If $e_0$ is accepted by $\mathcal{A}_{\textit{SinPri}}^4$: Then remove all values that are not renamed into $a$ or $b$ by $r$.
 \end{itemize}
@@
 We can see that $e \sqsubseteq u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$.
-Let $u'$, $v'$ and $w'$ be obtained from $u$, $v$ and $w$ by erasing all items with priority incomparable with $p$, respectively.
+Let $u'$, $v'$ and $w'$ be obtained from $u$, $v$ and $w$ by erasing all values with priority incomparable with $p$, respectively.
 Since the predicates in $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ does not restrict the values of priorities incomparable with $p$, is not hard to see that $e \sqsubseteq l'$ and $\mathsf{MatchedMaxPriority\text{-}Seq}(l',x)$ holds, where $l' = u' \cdot \textit{put}(x,\textit{pri}) \cdot v' \cdot \textit{rm}(x) \cdot w'$.
@@
 \noindent (1) Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x)$ by contradiction.
-Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{UVSet}_1(e,x)$ is an operation of item $d$.
+Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{UVSet}_1(e,x)$ is an operation of value $d$.
 
@@
 Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_{\textit{j+1}}(e,x)$ by contradiction.
-Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$ is an operation of item $d_{\textit{j+1}}$.
+Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$ is an operation of value $d_{\textit{j+1}}$.
 We use a triple $(t_1,t_2,t_3)$ to represent related information.
@@
     \setlength{\itemsep}{0.5pt}
-    \item[-] If $( j > 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{UVSet}_{\textit{j-1}}(e,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, we can see that $\textit{put}(d_{\textit{j+1}},\_) <_{hb} o''$, and then operations of $d_{\textit{j+1}}$ is in $\textit{UVSet}_j(e,x)$, contradicts that operations of $d_{\textit{j+1}}$ is in $\textit{UVSet}_{\textit{j+1}}(e,x)$.
+    \item[-] If $( j > 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of value $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{UVSet}_{\textit{j-1}}(e,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, we can see that $\textit{put}(d_{\textit{j+1}},\_) <_{hb} o''$, and then operations of $d_{\textit{j+1}}$ is in $\textit{UVSet}_j(e,x)$, contradicts that operations of $d_{\textit{j+1}}$ is in $\textit{UVSet}_{\textit{j+1}}(e,x)$.
 
@@
 
-    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{UVSet}_{\textit{j-1}}(e,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, we can see that $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} o'' ) \vee ( \textit{rm}(d_j) <_{hb} \textit{put}(d_j,\_) )$, which is impossible.
+    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of value $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{UVSet}_{\textit{j-1}}(e,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, we can see that $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} o'' ) \vee ( \textit{rm}(d_j) <_{hb} \textit{put}(d_j,\_) )$, which is impossible.
 
@@
 \item[-] $(\textit{rm},\textit{put},\textit{rm})$: Let $T_{\textit{ind}}$ be the set of sentences $\{ \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}), \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j),\ldots, \textit{put}(d_{\textit{ind+1}},\_) <_{hb} \textit{rm}(d_{\textit{ind}}) \}$.
-Here each $d_i$ is a item of some operation in $\textit{UVSet}_i(e,x)$.
+Here each $d_i$ is a value of some operation in $\textit{UVSet}_i(e,x)$.
 Let us prove that from $T_j$ we can obtain contradiction by induction:
@@
 We also require that $\mathsf{Has\text{-}MatchedMaxPriority}(e')$ holds and there exists a cycle going through the value with maximal priority in $e'$.
-By data-independence, we can obtain $e_r$ from $e$ by renaming function, which maps such item to be $b$, maps items that cover it to be $a$, and maps other items into $d$.
+By data-independence, we can obtain $e_r$ from $e$ by renaming function, which maps such value to be $b$, maps values that cover it to be $a$, and maps other values into $d$.
 There are four possible enumeration of call and return actions of $\textit{put}(b,p)$ and $\textit{rm}(b)$.
@@
 \noindent The $\textit{if}$ direction: Given such $e$, $e'$ and $x$.
-Let renaming function $r$ maps $x$ into $b$, maps items cover $x$ into $a$, and maps other items into $d$.
+Let renaming function $r$ maps $x$ into $b$, maps values cover $x$ into $a$, and maps other values into $d$.
 By data-independence, $r(e) \in \mathcal{I}$.
@@
 \begin {proof}
-
-To prove the $\textit{if}$ direction, let $e_{x,y}$ be the execution that is obtained from $e$ by erasing all actions of items that has same priority as $x$, except for actions of $x$ and $y$.
-It is obvious that $\textit{last}(e_{x,y}) = \seqPQ_1^{=}$.
-Since $y <_{\textit{pb}}^* x$, according to $\seqPQ_1^{=}$, we can see that $x$ should be chosen as $\textit{itm}$ in $\seqPQ_1^{=}$.
+To prove the $\textit{if}$ direction, let $e_{x,y}$ be the execution that is obtained from $e$ by erasing all actions of values that has same priority as $x$, except for actions of $x$ and $y$.
+It is obvious that $\mathsf{Has\text{-}MatchedMaxP-}$ $\mathsf{riority}^{=}(e_{x,y})$ holds.
+Since $y <_{\textit{pb}}^* x$, we can see that $x$ should be chosen as $\alpha$ in $\mathsf{MatchedMaxPriority}^{=}$ .
 
@@
 Therefore, for every sequence $l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, if $e_{x,y} \sqsubseteq l$, then either $u \cdot v$ contains some unmatched operations of priority smaller than $p$, or $w$ contains $\textit{put}(y,p)$ or $\textit{rm}(y)$.
-In both cases, $l \notin \textit{MS}(\seqPQ_1^{=})$.
+In both cases, $\mathsf{MatchedMaxPriority^{=}\text{-}Seq}(l,x)$ does not hold.
 
@@
 Assume we already know that for each $x$ and $y$ has maximal priority in $e$, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{call}(\textit{put},y,p)$ and $\textit{call}(\textit{rm},x)$.
-We need to prove that $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
-Recall that we already assume that each single-priority execution has FIFO property, and item with larger priority is not covered by items with smaller priority.
+We need to prove that $e$ is $\mathsf{MatchedMaxPriority}^{=}$-linearizable.
+Recall that we already assume that each single-priority execution has FIFO property, and value with larger priority is not covered by values with smaller priority.
 
@@
 
-\item[-] Let $a_1$ be the last inserted item of $l_{p}$.
+\item[-] Let $a_1$ be the last inserted value of $l_{p}$.
 
@@
 If for each $b \in \textit{Items}(e,p)$, the rightmost gap-point of $a_1$ is after $\textit{call}(\textit{put},b,p)$ and $\textit{call}(\textit{rm},b)$.
-Then by Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, we can obtain that $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
+Then by Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, we can obtain that $e$ is $\mathsf{MatchedMaxPriority}^{=}$-linearizable.
 
@@
     \setlength{\itemsep}{0.5pt}
-    \item[-] If for each item $b \in \textit{Items}(e,p)$, $a_2$ does not $<_{\textit{pb}}$ to $b$. Then we go to step $1$ and treat $a_2$ similarly as $a_1$.
+    \item[-] If for each value $b \in \textit{Items}(e,p)$, $a_2$ does not $<_{\textit{pb}}$ to $b$. Then we go to step $1$ and treat $a_2$ similarly as $a_1$.
     \item[-] Otherwise, there exists $a_3$ with priority $p$ such that $a_2 <_{\textit{pb}}^* a_3$.
@@
 $a^f$ satisfies the check of Step $1$.
-By Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, this implies that $e \sqsubseteq \textit{MS}(\seqPQ_1^{=})$.
+By Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}, this implies that $e$ is $\mathsf{MatchedMaxPriority}^{=}$-linearizable.
 This completes the proof of $\textit{if}$ direction.
@@
 Since $\textit{rm}(c_2)$ does not happen before $\textit{put}(c_2,\_)$, $\textit{call}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{rm},c_2)$.
-Since $c_3 <_{\textit{pb}}^A c_2$, $\textit{ret}(\textit{put},c_3)$ is before $\textit{call}(\textit{put},c_2,\_)$.
+Since $c_3 <_{\textit{pb}}^A c_2$, $\textit{ret}(\textit{put},c_3,\_)$ is before $\textit{call}(\textit{put},c_2,\_)$.
 Therefore, $\textit{ret}(\textit{put},c_3)$ is before $\textit{call}(\textit{put},c_1,\_)$, and $c_3 <_{\textit{pb}}^A c_1$.
@@
     \item[-] ($<_{\textit{pb}}^C \cdot <_{\textit{pb}}^A$): If $c_3 <_{\textit{pb}}^C c_2 <_{\textit{pb}}^A c_1$.
-Since $c_2 <_{\textit{pb}}^A c_1$, $\textit{ret}(\textit{put},c_2)$ is before $\textit{call}(\textit{put},c_1,\_)$.
-It is obvious that $\textit{call}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{put},c_2)$.
+Since $c_2 <_{\textit{pb}}^A c_1$, $\textit{ret}(\textit{put},c_2,\_)$ is before $\textit{call}(\textit{put},c_1,\_)$.
+It is obvious that $\textit{call}(\textit{put},c_2,\_)$ is before $\textit{ret}(\textit{put},c_2,\_)$.
 Since $c_3 <_{\textit{pb}}^C c_2$, $\textit{ret}(\textit{rm},c_3)$ is before $\textit{call}(\textit{put},c_2,\_)$.
@@
 Since $\textit{put}(c_2,\_)$ does not happen before $\textit{put}(c_3,\_)$, $\textit{call}(\textit{put},c_3,\_)$ is before $\textit{ret}(\textit{put},c_2,\_)$.
-Since $c_4 <_{\textit{pb}}^A c_3$, $\textit{ret}(\textit{put},c_4)$ is before $\textit{call}(\textit{put},c_3,\_)$.
-Therefore, $\textit{ret}(\textit{put},c_4)$ is before $\textit{call}(\textit{put},c_1,\_)$, and $c_4 <_{\textit{pb}}^A c_1$.
+Since $c_4 <_{\textit{pb}}^A c_3$, $\textit{ret}(\textit{put},c_4,\_)$ is before $\textit{call}(\textit{put},c_3,\_)$.
+Therefore, $\textit{ret}(\textit{put},c_4,\_)$ is before $\textit{call}(\textit{put},c_1,\_)$, and $c_4 <_{\textit{pb}}^A c_1$.
     \end{itemize}
@@
 
-Based on above results, given $a <_{\textit{pb}}^{b_1} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}}^{b_m} a_m <_{\textit{pb}}^{b_{\textit{m+1}}} b$, where each $b_i$ is in $\{ A,B,C \}$, we can merge relations, until we got one of the following facts:
+Based on above results, given $a <_{\textit{pb}}^{b_1} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}}^{b_m} a_m <_{\textit{pb}}^{b_{\textit{m+1}}} b$, where each $b_i$ is in $\{ A,B,C \}$, we can merge relations, until we get one of the following facts:
 
@@
 There are many enumerations of operations of $a$, $b$ and $a_1$ that may makes $a <_{\textit{pb}}^* b$.
-The following lemma states that with the help of gap-points, the number of potential enumerations can be further reduced into only five.
+The following lemma states that since some of them is not consistent with the requirements of gap-point, the number of potential enumerations can be further reduced into only five.
 
@@
 \label{lemma:five enumeration is enough for EPQ1Equal}
-Given a data-differentiated $p$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{=}$.
+Given a data-differentiated $p$-execution $e$ where $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e)$ holds.
 Let $a$ and $b$ be items with maximal priority $p$.
 Assume that $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$.
-Then, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists), where $a_1$ is the possible intermediate items for obtain $a <_{\textit{pb}}^* b$.
-\end{restatable}
-\begin {proof}
-
-Let us prove by consider all the possible reason of $a <_{\textit{pb}}^* b$.
+Then, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists), where $a_1$ is the possible intermediate value for obtain $a <_{\textit{pb}}^* b$.
+\end{restatable}
+\begin {proof}
+
+Let us prove by consider all the possible reasons of $a <_{\textit{pb}}^* b$.
 According to Lemma \ref{lemma:ob order has bounded length}, we need to consider five reasons: Let $o$ be the right-most gap-point of $b$.
@@
 
-    Since single-priority execution must satisfy the FIFO property, $\textit{rm}(b)$ does not happen before $\textit{rm}(a)$, and thus, $\textit{call}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
+    Since single-priority execution must satisfy the FIFO property, $\textit{rm}(b) \not <_{\textit{hb}} \textit{rm}(a)$, and thus, $\textit{call}(\textit{rm},a)$ is before $\textit{ret}(\textit{rm},b)$.
 If $\textit{call}(\textit{rm},a)$ is before $\textit{call}(\textit{rm},b)$, then $o$ is also a gap-point of $a$ and contradicts our assumption.
@@
 Here we explicitly draw the leftmost gap-point of $a$ as $o'$.
-Since the position of $\textit{ret}(\textit{put},b)$ does not influence the correctness, we can simply ignore it.
+Since the position of $\textit{ret}(\textit{put},b,\_)$ does not influence the correctness, we can simply ignore it.
 
@@
 Since the position of $\textit{ret}(\textit{put},a_1,p)$ and $\textit{call}(\textit{put},a,p)$ do not influence the correctness, we can simply ignore it.
-We also ignore $\textit{call}(\textit{put},b,p)$ and $\textit{ret}(\textit{put},b)$, since the only requirements of them are (1) $\textit{rm}(b)$ does not happen before $\textit{put}(b)$ and (2) $\textit{call}(\textit{put},b,p)$ is before $o$.
+We also ignore $\textit{call}(\textit{put},b,p)$ and $\textit{ret}(\textit{put},b,\_)$, since the only requirements of them are (1) $\textit{rm}(b) \not <_{\textit{hb}} \textit{put}(b)$ and (2) $\textit{call}(\textit{put},b,p)$ is before $o$.
 
@@
     Since $a_1 <_{\textit{pb}}^A b$, $\textit{ret}(\textit{put},a_1)$ is before $\textit{call}(\textit{put},b,p)$.
-Since $\textit{call}(\textit{put},b,p)$ is before $o$, we can see that $\textit{ret}(\textit{put},a_1)$ is before $o$.
+Since $\textit{call}(\textit{put},b,p)$ is before $o$, we can see that $\textit{ret}(\textit{put},a_1,\_)$ is before $o$.
 
@@
 
-Let us begin to represent several register automata that is used to capture the existence of a data-differentiated $\_$-execution $e$, $e$ has a projection $e'$, $\textit{last}(e') = \seqPQ_1^{=}$, there exists items $a$ and $b$ with maximal priority in $e'$, $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,\_)$ or $\textit{call}(\textit{rm},a)$.
-
-Given a data-differentiated $\_$-execution $e$, two actions $\textit{act}_1$, $\textit{act}_2$ of maximal priority in $e$, and assume that $\textit{act}_1$ is before $\textit{act}_2$ in $e$.
+Given a data-differentiated execution $e$ with only one maximal priority, two actions $\textit{act}_1$, $\textit{act}_2$ of maximal priority in $e$, and assume that $\textit{act}_1$ is before $\textit{act}_2$ in $e$.
 we say that $\textit{act}_1$, $\textit{act}_2$ is covered by items $d_1,\ldots,d_m$ in $e$, if the priorities of $d_1,\ldots,d_m$ is smaller than that of $\textit{act}_1$ and $\textit{act}_2$, and
@@
 
-\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_)$ happens before $\textit{rm}(d_i)$,
+\item[-] For each $i < 1 \leq m$,$\textit{put}(d_{\textit{i-1}},\_) <_{\textit{hb}} \textit{rm}(d_i)$,
 
@@
 
-According to Lemma \ref{lemma:EPQ1Equal as pb order and gap-point}, Lemma \ref{lemma:ob order has bounded length} and Lemma \ref{lemma:five enumeration is enough for EPQ1Equal}, it is not hard to prove that, given a data-differentiated $\_$-execution $e$ with $\textit{last}(e) = \seqPQ_1^{=}$, $e$ does not linearizable with respect to $\textit{EMS}(\textit{PQ}_1^{=})$, if and only if, one of enumerations holds in $e$ (permit renaming), while $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ is covered by some $d_1,\ldots,d_m$, $\textit{call}(\textit{rm},b)$ is before $\textit{ret}(\textit{put},d_m,\_)$, and $\textit{call}(\textit{rm},d_1)$ is before $\textit{ret}(\textit{rm},a)$.
+According to Lemma \ref{lemma:EPQ1Equal as pb order and gap-point}, Lemma \ref{lemma:ob order has bounded length} and Lemma \ref{lemma:five enumeration is enough for EPQ1Equal}, it is not hard to prove that, given a data-differentiated execution $e$ with only one maximal priority and $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e)$ holds.
+$e$ is not $\mathsf{MatchedMaxPriority}$-linearizable, if and only if, one of enumerations holds in $e$ (permit renaming), while $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ is covered by some $d_1,\ldots,d_m$, $\textit{call}(\textit{rm},b)$ is before $\textit{ret}(\textit{put},d_m,\_)$, and $\textit{call}(\textit{rm},d_1)$ is before $\textit{ret}(\textit{rm},a)$.
 We say that such $d_1,\ldots,d_m$ constitute the rightmost gap of $b$.
+
+Let us begin to represent several register automata that is used to capture the existence of a data-differentiated execution $e$, $e$ has a projection $e'$, $e'$ has only one maximal priority, $\mathsf{Has\text{-}MatchedMax-}$ $\mathsf{Priority}(e')$ holds, there exists items $a$ and $b$ with maximal priority in $e'$, $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,\_)$ or $\textit{call}(\textit{rm},a)$.
 

### 2017-04-22 co-regular-EPQ.tex
some stuff

@@
 The case of \figurename~\ref{fig:introduce pb order} can be formally described as follows: $a <_{\textit{pb}}^* b$ while the right-most gap-point of $b$ is before $\textit{call}(\textit{rm},a)$ or $\textit{call}(\textit{put},a,p_4)$.
-The following lemma states that these conditions are enough to characterize non-linearizability w.r.t $\mathsf{MatchedMaxPriority}^{=}$ (see Appendix\ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}).
+The following lemma states that these conditions are enough to characterize non-linearizability w.r.t $\mathsf{MatchedMaxPriority}^{=}$ (see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}).
 
@@
 
+\smallskip
 To characterize violations to linearizability w.r.t.

### 2017-04-22 appendix-EPQ.tex
modify

@@
 
-\EPQRulesAndSemantics*
+%\EPQRulesAndSemantics*
+
+$\newline$
+
+{\noindent \bf Lemma \ref{lemma:EPQ rules and semantics}}: $\textit{Check-PQ-Seq}(e)=\mathsf{true}$ iff $e\in \seqPQ$, for every data-differentiated sequential execution $e$.
 
@@
 
-\EPQClosedUnderProjection*
+%\EPQClosedUnderProjection*
+
+{\noindent \bf Lemma \ref{lem:closure_proj}:} $\seqPQ$ is closed under projection, i.e., $\textit{proj}(e)\subseteq \seqPQ$ for each $e\in \seqPQ$.
 
@@
 
-\DataDifferentiatedisEnoughforPQ*
+%\DataDifferentiatedisEnoughforPQ*
+
+{\noindent \bf Lemma \ref{lemma:data differentiated is enough for PQ}:} A data-independent implementation $\mathcal{I}$ is linearizable w.r.t a data-independent set $S$ of sequential executions, if and only if $\mathcal{I}_{\neq}$ is linearizable w.r.t.
+$S_{\neq}$.
 
@@
 
-\ConCheckEPQIsCorrect*
+%\ConCheckEPQIsCorrect*
+
+{\noindent \bf Lemma \ref{lemma:con-check-EPQ is correct}}: $\textit{Check-PQ-Conc}(e)=\mathsf{true}$ iff $e \sqsubseteq \seqPQ$, for every data-differentiated execution $e$.
 
@@
 
-\EPQasMultiInMRpriforHistory*
+{\noindent \bf Lemma \ref{lemma:EPQ as multi in MRpri for history}}: Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$ if and only if for each $e' \in \textit{proj}(e)$, $\textit{Check-PQ-Conc-NonRec}(e')$ returns $\mathsf{true}$.
+
+%\EPQasMultiInMRpriforHistory*
 
@@
 
-\priExecutionIsEnough*
+$\newline$
+
+{\noindent \bf Lemma \ref{lemma:pri execution is enough}}: Let $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$ and $e$ a data-differentiated execution.
+Then, $e$ is $\Gamma$-linearizable iff $e\vert_{\preceq p}$ is $\Gamma$-linearizable for some maximal priority $p$ in $e$.
+
+%\priExecutionIsEnough*
 \begin {proof}
@@
 \noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$.
-$\mathcal{A}_{\textit{1-lar}} \cap \mathcal{I} \neq \emptyset$ if and only if there exists $e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$, $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, $x$ is the value with maximal priority in $e'$, $e'$ has only one maximal priority, and there is a cycle going through $x$ in $G$, where $G$ is the left-right constraint of $x$ in $e'$.
+$\mathcal{A}_{\textit{1-lar}} \cap \mathcal{I} \neq \emptyset$ if and only if there exists $e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$, $\mathsf{Has\text{-}MatchedMaxPriority}(e')$ holds, $x$ is the value with maximal priority in $e'$, $e'$ has only one maximal priority, and there is a cycle going through $x$ in $G$, where $G$ is the left-right constraint of $x$ in $e'$.
 
@@
 
-\EPQOneEqualAsPBandGP*
+%\EPQOneEqualAsPBandGP*
+
+$\newline$
+{\noindent \bf Lemma \ref{lemma:EPQ1Equal as pb order and gap-point}}: Let $e$ be a data-differentiated execution which contains only one maximal priority $p$ such that $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds.
+%Given a data-differentiated $p$-execution $e$ where $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)$ holds.
+Then, $e$ is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ iff $e$ contains two values $x$ and $y$ of maximal priority $p$ such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is strictly smaller than the index of both $\textit{call}(\textit{put},y,p)$ and $\textit{call}(\textit{rm},y)$.
 
@@
 
-\OBOrderHasBoundedLength*
+%\OBOrderHasBoundedLength*
+
+$\newline$
+{\noindent \bf Lemma \ref{lemma:ob order has bounded length}}: Let $e$ be a data-differentiated execution such that $a <_{\textit{pb}} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}} a_m <_{\textit{pb}} b$ holds for some set of values $a$, $a_1$,$\ldots$,$a_m$, $b$.
+Then, one of the following holds:
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$, or $a <_{\textit{pb}}^C b$,
+
+\item[-] $a <_{\textit{pb}}^A a_i <_{\textit{pb}}^B b$ or $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A b$, for some $i$.
+%\item[-] $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A a_j <_{\textit{pb}}^B b$, for some $i$ and $j$,
+\end{itemize}
 
@@
 An automaton $\mathcal{A}_{\textit{l-eq}}^1$ is given in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}, and it is constructed for the first enumeration in \figurename~\ref{fig:history enumeration 1 for PQ1Equal}.
-Here we rename the items that covers $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ into $d$, and rename the remanning items into $e$.
-In this figure, $c = \textit{call}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{call}(\textit{rm},e), \textit{ret}(\textit{rm},e),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{call}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{call}(\textit{put},d),\textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.
+Here we rename the values that cover $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ into $d$, and rename the remanning values into $\top$.
+In this figure, $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true})$, $\textit{call}(\textit{rm},\top), \textit{ret}(\textit{rm},\top),$ $\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_1 = C \cup \{ \textit{call}(\textit{put},d,<r) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $C_3 = C_2 \cup \{ \textit{call}(\textit{put},d,<r),\textit{ret}(\textit{rm},d) \}$, $C_4 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},d) \}$.
 
@@
 An automaton $\mathcal{A}_{\textit{l-eq}}^2$ is given in \figurename~\ref{fig:automata for second enumeration of PQ1Equal}, and it is constructed for the second enumeration in \figurename~\ref{fig:history enumeration 2 for PQ1Equal}.
-In \figurename~\ref{fig:automata for second enumeration of PQ1Equal}, $c_1$, $c_2$, $c_3$ and $c_4$ is same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}.
+In \figurename~\ref{fig:automata for second enumeration of PQ1Equal}, $C_1$, $C_2$, $C_3$ and $C_4$ is same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}.
 
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{call}(\textit{rm},b)$ and before $\textit{call}(\textit{rm},a)$: There are two possible positions of $\textit{call}(\textit{put},a,p)$: (1) before $\textit{call}(\textit{rm},a_1)$, and (2) after $\textit{call}(\textit{rm},a_1)$, and before $\textit{ret}(\textit{put},a)$.
+\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{call}(\textit{rm},b)$ and before $\textit{call}(\textit{rm},a)$: There are two possible positions of $\textit{call}(\textit{put},a,p)$: (1) before $\textit{call}(\textit{rm},a_1)$, and (2) after $\textit{call}(\textit{rm},a_1)$, and before $\textit{ret}(\textit{put},a,p)$.
 
@@
 
-\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{call}(\textit{put},a_1,\_)$ and before $\textit{ret}(\textit{rm},a_1)$: same as above case.
-
-\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{ret}(\textit{put},a)$ and before $\textit{call}(\textit{put},a_1,p)$: same as above case.
-
-\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{call}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$: There are three possible positions of $\textit{call}(\textit{put},a,p)$: (1) after $\textit{call}(\textit{put},b,p)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{call}(\textit{rm},a_1)$ and before $\textit{call}(\textit{put},b,p)$, and (3) before $\textit{call}(\textit{rm},a_1)$.
-
-\item[-] If $\textit{call}(\textit{put},b,p)$ is before $\textit{call}(\textit{rm},a_1)$: There are three possible positions of $\textit{call}(\textit{put},a,p)$: (1) after $\textit{call}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{call}(\textit{put},b,p)$ and before $\textit{call}(\textit{rm},a_1)$, and (3) before $\textit{call}(\textit{put},b,p)$.
+\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{call}(\textit{put},a_1,p)$ and before $\textit{ret}(\textit{rm},a_1)$: same as above case.
+
+\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{ret}(\textit{put},a,p)$ and before $\textit{call}(\textit{put},a_1,p)$: same as above case.
+
+\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{call}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a,p)$: There are three possible positions of $\textit{call}(\textit{put},a,p)$: (1) after $\textit{call}(\textit{put},b,p)$ and before $\textit{ret}(\textit{put},a,p)$, (2) after $\textit{call}(\textit{rm},a_1)$ and before $\textit{call}(\textit{put},b,p)$, and (3) before $\textit{call}(\textit{rm},a_1)$.
+
+\item[-] If $\textit{call}(\textit{put},b,p)$ is before $\textit{call}(\textit{rm},a_1)$: There are three possible positions of $\textit{call}(\textit{put},a,p)$: (1) after $\textit{call}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a,p)$, (2) after $\textit{call}(\textit{put},b,p)$ and before $\textit{call}(\textit{rm},a_1)$, and (3) before $\textit{call}(\textit{put},b,p)$.
 \end{itemize}
@@
 For each case, we construct an finite automaton.
-Let $\textit{Auts}_{\textit{1-eq}}^{3}$ be the set of finite automata that is constructed for above fourteen cases.
+Let $\mathcal{A}_{\textit{1-eq}}^{3}$ be the union of register automata that is constructed for above fourteen cases.
 For example, for the case $\textit{ca}_1$ when $\textit{call}(\textit{put},a,p)$ is before $\textit{call}(\textit{rm},a_1)$, $\textit{call}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$, and $\textit{call}(\textit{put},b,p)$ is before $\textit{call}(\textit{rm},b)$, we construct a finite automaton $\mathcal{A}_{\textit{l-eq}}^{\textit{3-1}}$ in \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}.
-In \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}, let $c$ and $c_1 = c + \textit{call}(\textit{put},d,\textit{les}_p)$ the same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}.
-Let $c_2 = c_1 + \textit{ret}(\textit{put},a_1)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$, $c_4 = c_3 + \textit{call}(\textit{put},d) + \textit{ret}(\textit{rm},d)$, and $c_5 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{put},a_1) + \textit{ret}(\textit{rm},d)$.
-The other register automata in $\textit{Auts}_{\textit{1-eq}}^{3}$ can be similarly constructed.
+In \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}, let $C$ and $C_1$ the same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}.
+Let $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},a_1,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $C_4 = C_3 \cup \{ \textit{call}(\textit{put},d,<r), \textit{ret}(\textit{rm},d) \}$, and $C_5 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{put},a_1,=r), \textit{ret}(\textit{rm},d) \}$.
+Other register automata can be similarly constructed.
 
@@
 
-Similarly, we construct sets $\textit{Auts}_{\textit{1-eq}}^{4}$ and $\textit{Auts}_{\textit{1-eq}}^{5}$ of register automata for the forth enumeration in \figurename~\ref{fig:history enumeration 4 for PQ1Equal} and the fifth enumeration in \figurename~\ref{fig:history enumeration 5 for PQ1Equal}, respectively.
-
-Let $\textit{Auts}_{\textit{1-eq}} = \{ \mathcal{A}_{\textit{l-eq}}^1, \mathcal{A}_{\textit{l-eq}}^2 \} \cup \textit{Auts}_{\textit{1-eq}}^{3} \cup \textit{Auts}_{\textit{1-eq}}^{4} \cup \textit{Auts}_{\textit{1-eq}}^{5}$.
-The following lemma states that $\textit{PQ}_1^{=}$ is co-regular.
+Similarly, we construct sets $\mathcal{A}_{\textit{1-eq}}^{4}$ and $\mathcal{A}_{\textit{1-eq}}^{5}$ of union of register automata for the forth enumeration in \figurename~\ref{fig:history enumeration 4 for PQ1Equal} and the fifth enumeration in \figurename~\ref{fig:history enumeration 5 for PQ1Equal}, respectively.
+
+Let $\mathcal{A}_{\textit{1-eq}}$ be the union of $\mathcal{A}_{\textit{l-eq}}^1, \mathcal{A}_{\textit{l-eq}}^2,\mathcal{A}_{\textit{1-eq}}^{3}, \mathcal{A}_{\textit{1-eq}}^{4}$ and $\mathcal{A}_{\textit{1-eq}}^{5}$.
+The following lemma states that $\mathcal{A}_{\textit{1-eq}}$ is $\mathsf{MatchedMaxPriority}^{=}$-complete.
 
@@
 \label{lemma:EPQ1Equal is co-regular}
-$\mathsf{MatchedMaxPriority}^{=}$ is co-regular.
-\end{restatable}
-
-\begin {proof}
-
-We need to prove that, given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$, if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e),$ $\seqPQ_1^{=} \in \textit{last}(e') \wedge e'$ does not linearizable w.r.t.
-$\textit{MS}(\seqPQ_1^{=})$.
+$\mathcal{A}_{\textit{1-eq}}$ is $\mathsf{MatchedMaxPriority}^{=}$-complete.
+\end{restatable}
+
+\begin {proof}
+
+We need to prove that, given a data-independent implementation $\mathcal{I}$.
+$\mathcal{A}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$ if and only if there exists $e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$ such that $e'$ is not $\mathsf{MatchedMaxPriority}^{=}$-linearizable.
 
@@
 
-\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e)$, $\textit{last}(e')=\seqPQ_1^{=}$, $a$ and $b$ are two items with maximal priority $p$ in $e'$, $e'$ is a $p$-execution, $a <_{\textit{pb}}^* b$ in $e'$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$ in $e'$.
-
-\noindent The $\textit{only if}$ direction: Assume that $e_1 \in \mathcal{I}$ is accepted by some register automata in $\textit{Auts}_{\textit{1-eq}}$.
+\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$.
+$\mathcal{A}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$ if and only if there exists $e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$, $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e')$ holds, $x$ is the value with maximal priority in $e'$, $e'$ has only one maximal priority, $a$ and $b$ are two items with maximal priority $p$ in $e'$, $a <_{\textit{pb}}^* b$ in $e'$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$ in $e'$.
+
+\noindent The $\textit{only if}$ direction: Assume that $e_1 \in \mathcal{I}$ is accepted by some register automata in $\mathcal{A}_{\textit{1-eq}}$.
 By data-independence, there exists data-differentiated execution $e_2 \in \mathcal{I}$ and a renaming function $r$, such that $e_1=r(e_2)$.
-Since $e_1$ is accepted by some register automata in  $\textit{Auts}_{\textit{1-eq}}$, let $x$, $y$ and $z$ (if exists) be the items that are renamed into $b$, $a$ and $a_1$ (if exists) by $r$, respectively, and let $d_1,\ldots,d_m$ be the items that are renamed into $d$ by $r$.
+Since $e_1$ is accepted by some register automata in  $\mathcal{A}_{\textit{1-eq}}$, let $x$, $y$ and $z$ (if exists) be the items that are renamed into $b$, $a$ and $a_1$ (if exists) by $r$, respectively, and let $d_1,\ldots,d_m$ be the items that are renamed into $d$ by $r$.
 
 let $e'' = e_2 \vert_{ \{ x,y,z,d_1,\ldots,d_m \} }$.
-It is obvious that $e'' \in \textit{proj}(e_2)$ is a $p$-execution, $\textit{last}(e'') = \seqPQ_1^{=}$.
+It is obvious that $e'' \in \textit{proj}(e_2)$, $e''$ has only one maximal priority, and $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e'')$ holds.
 According to our construction of automata in $\textit{Auts}_{\textit{1-eq}}$, it is not hard to see that $x$ and $y$ has maximal priority in $h_2$, $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is before $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$ in $e''$.
 
-\noindent The $\textit{if}$ direction: Assume that there exists $e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $last(e')=\textit{PQ}_1^{=}$, $a'$ and $b'$ are two items with maximal priority $p$ in $e'$, $e'$ is a $p$-execution, $a' <_{\textit{pb}}^* b'$ in $e'$, and the rightmost gap-point of $b'$ is before $\textit{call}(\textit{put},a',p)$ or $\textit{call}(\textit{rm},a')$ in $e'$.
-By data-independence, we can obtain execution $e_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the items $d_1,\ldots,d_m$ that constitute the rightmost gap of $b'$, we rename them into $d$, (3) if $a' <_{\textit{pb}}^A a'_1 <_{\textit{pb}}^B b$, we rename $a'_1$ into $a_1$, and (4) rename the other items into $e$.
-It is easy to see that $\textit{last}(e_1) = \seqPQ_1^{=}$, $a$ and $b$ has maximal priority in $e_1$, $a <_{\textit{pb}}^* b$ in $e_1$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$ in $e_1$.
+\noindent The $\textit{if}$ direction: Assume that there exists $e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e')$ holds $e'$ has only one maximal priority, $a'$ and $b'$ are two items with maximal priority $p$ in $e'$, $a' <_{\textit{pb}}^* b'$ in $e'$, and the rightmost gap-point of $b'$ is before $\textit{call}(\textit{put},a',p)$ or $\textit{call}(\textit{rm},a')$ in $e'$.
+By data-independence, we can obtain execution $e_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the items $d_1,\ldots,d_m$ that constitute the rightmost gap of $b'$, we rename them into $d$, (3) if $a' <_{\textit{pb}}^A a'_1 <_{\textit{pb}}^B b$, we rename $a'_1$ into $a_1$, and (4) rename the other items into $\top$.
+It is easy to see that $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e_1)$ holds, $a$ and $b$ has maximal priority in $e_1$, $a <_{\textit{pb}}^* b$ in $e_1$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$ in $e_1$.
 By Lemma \ref{lemma:five enumeration is enough for EPQ1Equal}, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists).
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] If $a <_{\textit{pb}}^* b$ because of the first enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{l-eq}}^1$.
-
-\item[-] If $a <_{\textit{pb}}^* b$ because of the second enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{l-eq}}^2$.
-
-\item[-] If $a <_{\textit{pb}}^* b$ because of the third enumeration, it is easy to see that $h_1$ is accepted by some register automaton in $\textit{Auts}_{\textit{1-eq}}^{3}$.
-
-\item[-] If $a <_{\textit{pb}}^* b$ because of the forth enumeration, it is easy to see that $h_1$ is accepted by some register automaton in $\textit{Auts}_{\textit{1-eq}}^{4}$.
-
-\item[-] If $a <_{\textit{pb}}^* b$ because of the fifth enumeration, it is easy to see that $h_1$ is accepted by some register automaton in $\textit{Auts}_{\textit{1-eq}}^{5}$.
+\item[-] If $a <_{\textit{pb}}^* b$ because of the first enumeration, it is easy to see that $e_1$ is accepted by $\mathcal{A}_{\textit{l-eq}}^1$.
+
+\item[-] If $a <_{\textit{pb}}^* b$ because of the second enumeration, it is easy to see that $e_1$ is accepted by $\mathcal{A}_{\textit{l-eq}}^2$.
+
+\item[-] If $a <_{\textit{pb}}^* b$ because of the third enumeration, it is easy to see that $e_1$ is accepted by some register automaton in $\mathcal{A}_{\textit{1-eq}}^{3}$.
+
+\item[-] If $a <_{\textit{pb}}^* b$ because of the forth enumeration, it is easy to see that $e_1$ is accepted by some register automaton in $\mathcal{A}_{\textit{1-eq}}^{4}$.
+
+\item[-] If $a <_{\textit{pb}}^* b$ because of the fifth enumeration, it is easy to see that $e_1$ is accepted by some register automaton in $\mathcal{A}_{\textit{1-eq}}^{5}$.
 \end{itemize}
@@
 
-\subsection{Co-Regular of $\seqPQ_2^{>}$}
+\subsection{$\mathsf{UnmatchedMaxPriority}^{>}(e)$ Always Holds}
 \label{subsec:appendix co-regular of EPQ2Lar}
@@
 \label{lemma:EPQ2Lar is always co-regular}
-Given a data-differentiated $\_$-execution $e$, if $\textit{last}(e) = \seqPQ_2^{>}$, then $e \sqsubseteq \textit{MS}(\seqPQ_2^{>})$.
-\end{restatable}
-
-\begin {proof}
-Since $\textit{last}(e) = \seqPQ_2^{>}$, the actions with maximal priority in $e$ is some unmatched $\textit{put}$.
-Therefore, no matter how we locate linearization points, we can always obtain a sequence $l$ of operations that contains unmatched $\textit{put}$ with maximal priority, and this satisfy the guard of $\textit{MS}(\seqPQ_2^{>})$.
+
+Given a data-differentiated execution $e$ with only one maximal priority, if $\mathsf{Has\text{-}}\mathsf{Unmat-}$ $\mathsf{chedMaxPriority}^{>}(e)$ holds, then $e \sqsubseteq l$ for some $l$ where $\mathsf{UnmatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(l,x)$ holds.
+\end{restatable}
+
+\begin {proof}
+Since $\mathsf{Has\text{-}}\mathsf{UnmatchedMaxPriority}^{>}(e)$ holds, the actions with maximal priority in $e$ is some unmatched $\textit{put}$.
+Therefore, no matter how we locate linearization points, we can always obtain a sequence $l$ of operations that contains unmatched $\textit{put}$ with maximal priority, and this satisfy the requirements of $\mathsf{UnmatchedMaxPriority}^{>}$.
 This completes the proof of this lemma.
@@
 
-\subsection{Co-Regular of $\seqPQ_2^{=}$}
+\subsection{$\mathsf{UnmatchedMaxPriority}^{=}(e)$ Always Holds}
 \label{subsec:appendix co-regular of EPQ2Equal}
@@
 \label{lemma:EPQ2Equal as happen before}
-Given a data-differentiated $p$-execution $e$ with $\textit{last}(e) = \seqPQ_2^{=}$.
-$e$ does not linearizable to $\textit{MS}(\seqPQ_2^{=})$, if and only if there exists $x$ and $y$ with priority $p$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,p) <_{\textit{hb}} \textit{put}(y,p)$.
+Given a data-differentiated execution $e$ with only one priority and $\mathsf{Has\text{-}}\mathsf{Unmatched-}$ $\mathsf{MaxPriority}^{=}(e)$ holds.
+$e$ is not $\mathsf{UnmatchedMaxPriority}^{=}(e)$-linearizable, if and only if there exists $x$ and $y$ with maximal priority $p$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,p) <_{\textit{hb}} \textit{put}(y,p)$.
 \end{restatable}
@@
 Assume that for each pair of $x$ and $y$ with maximal priority in $e$, if $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, then $\textit{put}(x,p)$ does not happen before $\textit{put}(y,p)$.
-We need to prove that $e \sqsubseteq \textit{MS}(\seqPQ_2^{=})$.
-
-Let $x_1,\ldots,x_m$ be the set of items with priority $p$ and has unmatched $\textit{put}$ in $e$, let $y_1,\ldots,y_n$ be the set of items with priority $p$ and has matched $\textit{put}$ and $\textit{rm}$ in $e$.
-By assumption, we know that $\textit{call}(\textit{put},y_i,p)$ is before $\textit{ret}(\textit{put},x_j)$ for each $i,j$.
+We need to prove that $e$ is $\mathsf{UnmatchedMaxPriority}^{=}(e)$-linearizable.
+
+Let $x_1,\ldots,x_m$ be the set of values with priority $p$ and has unmatched $\textit{put}$ in $e$, let $y_1,\ldots,y_n$ be the set of values with priority $p$ and has matched $\textit{put}$ and $\textit{rm}$ in $e$.
+By assumption, we know that $\textit{call}(\textit{put},y_i,p)$ is before $\textit{ret}(\textit{put},x_j,p)$ for each $i,j$.
 Then we explicitly construction the linearization of $e$ by locating the linearization points of $e$ as follows:
@@
 It is easy to see that $e \sqsubseteq l$.
-Since linearization points of $\textit{put}(x_i,p)$ is after the linearization point of $\textit{put}(y_j,p)$ for each $i,j$, it is easy to see that $l \in \textit{MS}(\seqPQ_2^{=})$.
+Since linearization points of $\textit{put}(x_i,p)$ is after the linearization point of $\textit{put}(y_j,p)$ for each $i,j$, it is easy to see that $\mathsf{UnmatchedMaxPriority}^{=}\mathsf{\text{-}Seq}(l,x')$ holds where $x' = x_{\_}$ and its return action is after other values in $x_1,\ldots,x_m$.
 This completes the proof of this lemma.
@@
 
-Lemma \ref{lemma:EPQ2Equal as happen before} shows how to check violation to $\textit{MS}(\seqPQ_2^{=})$.
+Lemma \ref{lemma:EPQ2Equal as happen before} shows how to check non $\mathsf{UnmatchedMaxPriority}^{=}(e)$-linearizable exections.
 However, the case in Lemma \ref{lemma:EPQ2Equal as happen before} violates our assumption that each single-priority execution is FIFO.
-Therefore, we know that $\seqPQ_2^{=}$ is always co-regular, as states by the following lemma.
+Therefore, we know that $\mathsf{UnmatchedMaxPriority}^{=}(e)$ always holds, as states by the following lemma.
 
@@
 \label{lemma:EPQ2Equal is always co-regular}
-Given a data-differentiated $p$-execution $e$, if $\textit{last}(e) = \seqPQ_2^{=}$, then $e \sqsubseteq \textit{MS}(\seqPQ_2^{=})$.
-\end{restatable}
-
-\begin {proof}
-
-According to Lemma \ref{lemma:EPQ2Equal as happen before}, if $\textit{last}(e) = \seqPQ_2^{=}$ and $e$ does not linearizable to $\textit{MS}(\seqPQ_2^{=})$, then there exists $x$ and $y$ with priority $p$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$, and $\textit{put}(x,p) <_{\textit{hb}} \textit{put}(y,p)$.
+Given a data-differentiated execution $e$ with only one maximal priority, if $\mathsf{Has\text{-}}\mathsf{Unmat-}$ $\mathsf{chedMaxPriority}^{=}(e)$ holds, then $e \sqsubseteq l$ for some $l$ where $\mathsf{UnmatchedMaxPriority}^{=}\mathsf{\text{-}Seq}(l,x)$ holds.
+\end{restatable}
+
+\begin {proof}
+
+According to Lemma \ref{lemma:EPQ2Equal as happen before}, if $\mathsf{Has\text{-}}\mathsf{UnmatchedMaxPriority}^{=}(e)$ holds and $e$ is not $\mathsf{Unmatched-}$ $\mathsf{MaxPriority}^{=}$-linearizable, then there exists $x$ and $y$ with maximal priority $p$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$, and $\textit{put}(x,p) <_{\textit{hb}} \textit{put}(y,p)$.
 Let $e_1 = e \vert_{ \{ x,y \} }$.
@@
 
-\subsection{Co-Regular of $\seqPQ_3$}
+\subsection{Proofs, Definitions and Register Automata for $\mathsf{EmptyRemove}$}
 \label{subsec:co-regular of EPQ3}
 
-In this subsection we prove that $\seqPQ_3$ is co-regular. The notion of left-right constraint of $\textit{rm}(\textit{empty})$ is inspired by left-right constraint of queue \cite{Bouajjani:2015}.
+In this subsection we construct $\mathsf{EmptyRemove}$-complete register automata.
+The notion of left-right constraint of $\textit{rm}(\textit{empty})$ is inspired by left-right constraint of queue \cite{DBLP:conf/icalp/BouajjaniEEH15}.
 
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] the nodes are the items of $e$ or $o$, to which we add a node,
-
-\item[-] there is an edge from item $d_1$ to $o$, if $\textit{put}(d_1,\_)$ happens before $o$,
-
-\item[-] there is an edge from $o$ to item $d_1$, if $o$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$,
-
-\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$.
+\item[-] the nodes are the values of $e$ or $o$, to which we add a node,
+
+\item[-] there is an edge from value $d_1$ to $o$, if $\textit{put}(d_1,\_)$ happens before $o$,
+
+\item[-] there is an edge from $o$ to value $d_1$, if $o$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$,
+
+\item[-] there is an edge from value $d_1$ to value $d_2$, if $\textit{put}(d_1,\_)$ happens before $\textit{rm}(d_2,\_)$.
 \end{itemize}
@@
 
-Given a data-differentiated execution $e$ and $o = \textit{rm}(\textit{empty})$ of $e$, it is obvious that $\textit{last}(e) = \seqPQ_3$.
-Let $\textit{USet}_1(e,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some item, and either $\textit{op} <_{\textit{hb}} o$, or there is $\textit{op}'$ with the same item of $\textit{op}$, such that $\textit{op}' <_{\textit{hb}} o \}$.
-For each $i \geq 1$, let $\textit{USet}_{\textit{i+1}}(e,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some item, $\textit{op}$ is not in $\textit{USet}_k(e,o)$ for each $k \leq i$, and either $\textit{op}$ happens before some $o' \in \textit{USet}_i(e,o)$, or there is $\textit{op}''$ with the same item of $o$ and $\textit{op}''$ happens before some $o' \in \textit{USet}_i(e,o) \}$.
+Given a data-differentiated execution $e$ and $o = \textit{rm}(\textit{empty})$ of $e$, it is obvious that $\mathsf{Has\text{-}}\mathsf{EmptyRe-}$ $\mathsf{move}^{=}(e)$ holds.
+Let $\textit{USet}_1(e,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some value, and either $\textit{op} <_{\textit{hb}} o$, or there is $\textit{op}'$ with the same value of $\textit{op}$, such that $\textit{op}' <_{\textit{hb}} o \}$.
+For each $i \geq 1$, let $\textit{USet}_{\textit{i+1}}(e,o) = \{ \textit{op} \vert$ $\textit{op}$ is an operation of some value, $\textit{op}$ is not in $\textit{USet}_k(e,o)$ for each $k \leq i$, and either $\textit{op}$ happens before some $o' \in \textit{USet}_i(e,o)$, or there is $\textit{op}''$ with the same value of $o$ and $\textit{op}''$ happens before some $o' \in \textit{USet}_i(e,o) \}$.
 We can see that $\textit{USet}_i(e,o) \cap \textit{USet}_j(e,o) = \emptyset$ for any $i \neq j$.
@@
 \label{lemma:USet has matched put and rm}
-Given a data-differentiated execution $e$ with $\textit{last}(e) = \seqPQ_3$.
+
+Given a data-differentiated execution $e$ where $\mathsf{Has\text{-}EmptyRemove}(e)$ holds.
 Let $o$ be a $\textit{rm}(\textit{empty})$ of $e$.
@@
 \label{lemma:Rmx does not happen before USet for EPQ3} 
-Given a data-differentiated $\_$execution $e$ with $\textit{last}(e) = \seqPQ_3$.
+Given a data-differentiated execution $e$ where $\mathsf{Has\text{-}EmptyRemove}(e)$ holds.
 Let $o$ be a $\textit{rm}(\textit{empty})$ of $e$.
@@
 \label{lemma:Lin Equals Constraint for EPQ3}
-Given a data-differentiated execution $e$ with $\textit{last}(e) = \seqPQ_3$.
-$e$ does not linearizable w.r.t $\textit{MS}(\textit{PQ}_3)$, if and only if there exists $o = \textit{rm}(\textit{empty})$ in $e$, $G$ has a cycle going through $o$, where $G$ is the graph representing the left-right constraint of $o$.
+Given a data-differentiated execution $e$ where $\mathsf{Has\text{-}EmptyRemove}(e)$ holds.
+$e$ is not $\mathsf{EmptyRemove}$-linearizable, if and only if there exists $o = \textit{rm}(\textit{empty})$ in $e$, $G$ has a cycle going through $o$, where $G$ is the graph representing the left-right constraint of $o$.
 \end{restatable}
@@
 To prove the $\textit{if}$ direction, assume that there is such a cycle.
-Assume by contradiction that $e \sqsubseteq \textit{MS}(\seqPQ_3)$, and let $U$ and $V$ be the set of operations in $u$ and $v$.
+Assume by contradiction that $e \sqsubseteq l= u \cdot o \cdot v$ and $\mathsf{EmptyRemove\text{-}Seq}(l,o)$ holds.
+Let $U$ and $V$ be the set of operations in $u$ and $v$.
 Let the cycle be $d_1 \rightarrow d_2 \rightarrow \ldots \rightarrow d_m \rightarrow o \rightarrow d_1$ in $G$.
@@
 
-Let us begin to represent an automaton that is used for capture the case that, in a sub-execution $e'$ of an execution $e$, $\textit{last}(e')=\seqPQ_3$, $e'$ does not linearizable to $\textit{MS}(\seqPQ_3)$, and the reason is that there is a cycle going through some $\textit{rm}(\textit{empty})$ $o$ in the left-right constraint of $o$.
+Let us begin to represent a register automaton that is used for capture the case that, in a sub-execution $e'$ of an execution $e$, $\mathsf{Has\text{-}EmptyRemove}(e')$ holds, $e'$ is not $\mathsf{EmptyRemove}$-linearizable, and the reason is that there is a cycle going through some $\textit{rm}(\textit{empty})$ $o$ in the left-right constraint of $o$.
 The automaton is $\mathcal{A}_{\seqPQ}^3$, which is given in \figurename~\ref{fig:automata for PQ3}.
-In \figurename~\ref{fig:automata for PQ3}, let $c = \textit{call}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{call}(\textit{rm},d), \textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{call}(\textit{put},b,\textit{anyPri})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c + \textit{ret}(\textit{rm},b)$.
+In \figurename~\ref{fig:automata for PQ3}, let $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}($ $\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},\top), \textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_1 = C \cup \{ \textit{call}(\textit{put},b,\textit{true}) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, and $C_3 = C \cup \{ \textit{ret}(\textit{rm},b) \}$.
 
@@
 
-Given a data-differentiated execution $e$, we say that $o = \textit{rm}(\textit{empty})$ in $e$ is covered by items $d_1,\ldots,d_m$ in $h$, if
+Given a data-differentiated execution $e$, we say that $o = \textit{rm}(\textit{empty})$ in $e$ is covered by items $d_1,\ldots,d_m$, if
 
@@
 \label{lemma:EPQ3 is co-regular}
-$\seqPQ_3$ is co-regular.
-\end{restatable}
-
-\begin {proof}
-
-We need to prove that, given a data-independence implementation $\mathcal{I}$, $\mathcal{A}_{\seqPQ}^3 \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), last(e')=\seqPQ_3 \wedge e$ does not linearizable w.r.t.
-$\textit{MS}(\seqPQ_3)$.
+$\mathcal{A}_{\seqPQ}^3$ is $\mathsf{EmptyRemove}^{=}$-complete. 
+\end{restatable}
+
+\begin {proof}
+
+We need to prove that, given a data-independent implementation $\mathcal{I}$.
+$\mathcal{A}_{\seqPQ}^3 \cap \mathcal{I} \neq \emptyset$ if and only if there exists $e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$ such that $e'$ is not $\mathsf{EmptyRemove}^{=}$-linearizable.
 
@@
 
-\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\mathcal{A}_{\seqPQ}^3 \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e), last(e')=\seqPQ_3$, $o = \textit{rm}(\textit{empty})$ is in $e'$, and $o$ is covered by some items $d_1,\ldots,d_m$ in $e'$.
+\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$.
+$\mathcal{A}_{\seqPQ}^3 \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, $\mathsf{Has\text{-}EmptyRemove}(e')$ holds, $o = \textit{rm}(\textit{empty})$ is in $e'$, and $o$ is covered by some items $d_1,\ldots,d_m$ in $e'$.
 
@@
 \noindent The $\textit{if}$ direction: Assume that there exists such $e$, $e'$, $o$ and $d_1,\ldots,d_m$.
-Then, let $e_1$ be obtained from $e$ by renaming $d_1,\ldots,d_m$ into $b$ and renaming other items into $d$.
+Then, let $e_1$ be obtained from $e$ by renaming $d_1,\ldots,d_m$ into $b$ and renaming other items into $\top$.
 By data-independence, $e_1 \in \mathcal{I}$.
@@
 \end {proof}
-
-\forget{
-\section{Proofs and Definitions in Section \ref{sec:relate other data structures with extended priority queue}}
-\label{sec:appendix proof and definition in section relate other data structures with extended priority queue}
-
-\subsection{Proofs and Definitions in Subsection \ref{subsec:relate multiSet with extended priority queue}}
-\label{subsec:appendix proof and definition in section relate multiset with extended priority queue}
-
-In this section, we use the notions of inductive rules, $\textit{last}$ of a sequential execution, step-by-step and co-regular in \cite{Bouajjani:2015}.
-Since they are quite similar to the corresponding notions of extended priority queues in Section \ref{sec:inductive rules of extended priority queue}, Section \ref{sec:step-by-step linearizability of extended priority queues} and Section \ref{sec:co-regular of extended priority queues}, we do not introduce their definitions here.
-
-Let us first define three predicates:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Given a sequential execution $l$ of multi-set, $\textit{noDE}(l)$ is satisfied when each operation of $l$ is not $\textit{delete}(\textit{empty})$.
-
-\item[-] Given a sequential execution $l$ of multi-set, $\textit{matched-MS}(l)$ is satisfied, if (1) for each item $a \in \mathbb{D}$, if $\textit{insert}(a)$ is in $l$, then $\textit{delete}(a)$ is in $l$, and (2) for each item $a \in \mathbb{D}$, if $\textit{delete}(a)$ is in $l$, then $\textit{insert}(a)$ is in $l$.
-
-\item[-] Given a sequential execution $l$ of multi-set, $l \in \textit{Insert}^*$ is satisfied when each operation of $l$ is a $\textit{insert}$ event.
-\end{itemize}
-
-Let $\textit{MSet}$ be the set of sequential executions $w$ which can be derived from the empty word by inductive rules of multi-set. $\textit{MSet}$ is defined by the following inductive rules:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{MSet}_0 \equiv \epsilon \in \textit{MSet}$.
-
-\item[-] $\textit{MSet}_1 \equiv (u \in \textit{MSet}) \wedge
-(u \in \textit{Insert}^*)
-\Rightarrow
-(u \cdot \textit{insert}(\textit{itm}) \in \textit{MSet})$.
-
-\item[-] $\textit{MSet}_2 \equiv
-(u \cdot v \cdot w \in \textit{MSet}) \wedge
-(\textit{noDE}(u \cdot v \cdot w))
-\Rightarrow
-(u \cdot \textit{insert}(\textit{itm}) \cdot v \cdot \textit{delete}(\textit{itm}) \cdot w \in \textit{MSet})$.
-
-\item[-] $\textit{MSet}_3 \equiv
-(u \cdot v \in \textit{MSet}) \wedge
-(\textit{matched-MS}(u) )
-\Rightarrow
-(u \cdot \textit{delete}(\textit{empty}) \cdot v \in \textit{MSet})$.
-\end{itemize}
-
-Thus, given a sequential execution $e$, we define $\textit{last}(e)$ as the last possible rule to generate $e$ according to the rules of multi-set:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $e$ contains $\textit{delete}(\textit{empty})$, then $\textit{last}(e) = \textit{MS}_3$.
-
-\item[-] Else, if $e$ contains $\textit{delete}$, then $\textit{last}(e) = \textit{MSet}_2$.
-
-\item[-] Else, if $e$ contains only $\textit{insert}$, then $\textit{last}(e) = \textit{MSet}_1$.
-
-\item[-] Else ($e = \epsilon$), $\textit{last}(e) = \textit{MSet}_0$.
-\end{itemize}
-
-The following three lemmas state that the rules for multi-set are step-by-step linearizability.
-
-\begin{restatable}{lemma}{MS1isStepByStepLinearizability}
-\label{lemma:MS1 is step-by-step linearizability}
-If a differentiated concurrent execution $e$ is linearizable w.r.t.
-$\textit{MS}(\textit{MSet}_1)$ with witness $x$, then $e \setminus x \sqsubseteq \textit{MSet} \Rightarrow e \sqsubseteq \textit{MSet}$.
-\end{restatable}
-
-\begin {proof}
-Let $h$ be the data-differentiated history of $e$, and $l$ be an sequential execution such that $h \sqsubseteq l$ and $l$ matches $\textit{MSet}_1$ with witness $x$.
-Let $h'=h \setminus x$ and assume that $h' \sqsubseteq l' \in \textit{MSet}$.
-Let $e_{\textit{lp}}$ be an execution with linearization points of $e$ and the linearization points is added according to $l'$.
-Or we can say, $e_{\textit{lp}}$ is generated from $e$ by instrumenting linearization points, and the projection of $e_{\textit{lp}}$ into operation is $l'$.
-
-Let sequence $e'_{\textit{lp}}$ be generated from $e_{\textit{lp}}$ by adding $\textit{insert}(x)$ at an arbitrary time point between $\textit{call}(\textit{insert},x)$ and $\textit{ret}(\textit{insert},x)$.
-Let $l''$ be the projection of $e'_{\textit{lp}}$ into operations.
-
-It is easy to see that $h \sqsubseteq l''$.
-Since $l''$ is obtained from $l'$ by adding one $\textit{insert}(x)$, we can see that $l''$ contains only $\textit{insert}$, and then $l'' \in \textit{MSet}$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{MS2isStepByStepLinearizability}
-\label{lemma:MS2 is step-by-step linearizability}
-If a differentiated concurrent execution $e$ is linearizable w.r.t.
-$\textit{MS}(\textit{MSet}_2)$ with witness $x$, then $e \setminus x \sqsubseteq \textit{MSet} \Rightarrow e \sqsubseteq \textit{MSet}$.
-\end{restatable}
-
-\begin {proof}
-Let $h$ be the data-differentiated history of $e$, and $l$ be an sequential execution such that $h \sqsubseteq l$ and $l$ matches $\textit{MSet}_1$ with witness $x$.
-Let $h'=h \setminus x$ and assume that $h' \sqsubseteq l' \in \textit{MSet}$.
-Let $e_{\textit{lp}}$ be an execution with linearization points of $e$ and the linearization points is added according to $l'$.
-Or we can say, $e_{\textit{lp}}$ is generated from $e$ by instrumenting linearization points, and the projection of $e_{\textit{lp}}$ into operation is $l'$.
-
-It is easy to see that $\textit{delete}(x)$ does not happen before $\textit{insert}(x)$, and then $\textit{call}(\textit{insert},x)$ is before $\textit{ret}(\textit{delete},x)$ in $e$.
-Let sequence $e'_{\textit{lp}}$ be generated from $e_{\textit{lp}}$ by adding $\textit{insert}(x)$ just after $\textit{call}(\textit{insert},x)$ and adding $\textit{delete}(x)$ just before $\textit{ret}(\textit{delete},x)$.
-Let $l''$ be the projection of $e'_{\textit{lp}}$ into operations.
-
-It is easy to see that $h \sqsubseteq l''$.
-Since (1) $l' \in \textit{MSet}$, (2) $l''$ is obtained from $l'$ by adding one $\textit{insert}(x)$ and one $\textit{delete}(x)$, while $\textit{insert}(x)$ is before $\textit{delete}(x)$ (3)and $l'$ does not contain $\textit{delete}(\textit{empty})$, we can see that $l'' \in \textit{MSet}$.
-\qed
-\end {proof}
-
-\begin{restatable}{lemma}{MS3isStepByStepLinearizability}
-\label{lemma:MS3 is step-by-step linearizability}
-If a differentiated concurrent execution $e$ is linearizable w.r.t.
-$\textit{MS}(\textit{MSet}_3)$ and $o$ is a $\textit{delete}(\textit{empty})$ event, then $e \setminus o \sqsubseteq \textit{MSet} \Rightarrow e \sqsubseteq \textit{MSet}$.
-\end{restatable}
-
-\begin {proof}
-This Lemma can be similarly proved as Lemma \ref{lemma:EPQ3 is step-by-step linearizability}. \qed
-\end {proof}
-
-The following lemma states that $\textit{MSet}_1$ is always co-regular.
-
-\begin{restatable}{lemma}{MS1IsAlwaysCoRegular}
-\label{lemma:MS1 is always co-regular}
-Given a differentiated execution $e$, if $\textit{last}(e) = \textit{MSet}_1$, then $e \sqsubseteq \textit{MS}(\textit{MSet}_1)$.
-\end{restatable}
-
-\begin {proof}
-Since $\textit{last}(e) = \textit{MSet}_1$, there are only $\textit{insert}$ in $e$.
-Then no matter how we locate linearization points of operations of $e$, we can always obtain a sequence in $\textit{MS}(\textit{MSet}_1)$.
-This completes the proof of this lemma.
-\qed
-\end {proof}
-
-The following lemma shows how to detect violation to $\textit{MS}(\textit{MSet}_2)$.
-
-\begin{restatable}{lemma}{ReduceMS2intoOneValue}
-\label{lemma:reduce MS2 into one value}
-Given a differentiated execution $e$, there exists some $e' \in \textit{proj}(e)$, such that $\textit{last}(e') = \textit{MSet}_2$ and $e'$ doe not linearizable w.r.t $\textit{MS}(\textit{MSet}_2)$, if and only if one of the following case holds for some $x \in \mathbb{D}$.
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{delete}(x)$ is in $e$ while $\textit{insert}(x)$ is not in $e$.
-
-\item[-] there is more than one $\textit{delete}(x)$ in $e$ and one $\textit{insert}(x)$ in $e$.
-
-\item[-] $\textit{delete}(x) <_{\textit{hb}} \textit{insert}(x)$ in $e$.
-\end{itemize}
-\end{restatable}
-
-\begin {proof}
-
-The $\textit{only if}$ is obvious and omitted here.
-
-To prove the $\textit{if}$ direction, we prove its contrapositive. Assume that for each item $x$ of $e$, the following conditions are satisfied:
-
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If $\textit{delete}(x)$ is in $e$, then $\textit{insert}(x)$ is also in $e$.
-
-\item[-] There is at most $\textit{delete}(x)$ in $e$.
-
-\item[-] $\textit{delete}(x)$ does not happen before $\textit{insert}(x)$.
-\end{itemize}
-
-Assume by contradiction that there exists some $e' \in \textit{proj}(e)$, such that $\textit{last}(e') = \textit{MSet}_2$ and $e'$ doe not linearizable w.r.t $\textit{MS}(\textit{MSet}_2)$.
-Since $\textit{last}(e') = \textit{MSet}_2$, there exists $\textit{delete}$ in $e$.
-Let this $\textit{delete}$ operation be $\textit{delete}(a)$.
-By assumption, we know that in $e$ there exists one $\textit{insert}(a)$ and one $\textit{delete}(a)$, and $\textit{delete}(a)$ does not happen before $\textit{insert}(a)$.
-Then we know that $\textit{call}(\textit{insert},x)$ is before $\textit{ret}(\textit{delete},x)$.
-
-Let $e_{\textit{lp}}$ be generated from $e$ by (1) put the linearization of $\textit{insert}(x)$ just after $\textit{call}(\textit{insert},x)$, (2) put the linearization of $\textit{delete}(x)$ just before $\textit{ret}(\textit{insert},x)$, and (3) for other operations, put its linearization point at an arbitrary between its call and return actions.
-Let $l'$ be the projection of $e_{\textit{lp}}$ into operations.
-It is obvious that $h \sqsubseteq l'$.
-According to our construction of $e_{\textit{lp}}$, we can see that in $l'$, $\textit{insert}(x)$ is before $\textit{delete}(x)$.
-Therefore, we can see that $l' \in \textit{MS}(\textit{MSet}_2)$, which contradicts our assumption.
-This completes the proof of the $\textit{id}$ direction.
-\qed
-\end {proof}
-
-According to Lemma \ref{lemma:reduce MS2 into one value}, to check violation to $\textit{MS}(\textit{MSet}_2)$, we need to consider three cases for some $b \in \mathbb{D}$: (1) there is $\textit{delete}(b)$ but there is not $\textit{insert}(x)$, (2) there is more than one $\textit{delete}(b)$ and one $\textit{insert}(b)$ in $e$, and (3) $\textit{delete}(b) <_{\textit{hb}} \textit{insert}(b)$.
-
-For each such case, we construct a register automata.
-We generate register automata $\mathcal{A}_{\textit{MS}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata 1 for MS-2 in appendix}.
-Here $c_1 = \textit{call}(\textit{insert},a),\textit{ret}(\textit{insert},a)$, $\textit{call}(\textit{delete},a),\textit{ret}(\textit{delete},a),
-\textit{call}(\textit{delete},\textit{empty}),\textit{ret}(\textit{delete},\textit{empty})$, $c_2 = c_1 + \textit{call}(\textit{delete},b) + \textit{ret}(\textit{delete},b)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.3 \textwidth]{figures/PIC_AUTO_MS_1.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{MS}}^1$}
-  \label{fig:automata 1 for MS-2 in appendix}
-\end{figure}
-
-We generate register automata $\mathcal{A}_{\textit{MS}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata 2 for MS-2 in appendix}.
-Here $c_1 = \textit{call}(\textit{insert},a),\textit{ret}(\textit{insert},a), \textit{call}(\textit{delete},a),\textit{ret}(\textit{delete},a),\textit{call}(\textit{delete},\textit{empty})$, $\textit{ret}(\textit{delete},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{insert},b)$, $c_3 = c_2 + \textit{ret}(\textit{delete},b)$, $c_4 = c_3 + \textit{call}(\textit{delete},b)$, $c_5 = c_1 + \textit{ret}(\textit{delete},b)$, $c_6 = c_5 + \textit{call}(\textit{delete},b)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.7 \textwidth]{figures/PIC_AUTO_MS_2.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{MS}}^2$}
-  \label{fig:automata 2 for MS-2 in appendix}
-\end{figure}
-
-We generate register automata $\mathcal{A}_{\textit{MS}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata 3 for MS-2 in appendix}.
-Here $c_1 = \textit{call}(\textit{insert},a)$, $\textit{ret}(\textit{insert},a)$, $\textit{call}(\textit{delete},a)$, $\textit{ret}(\textit{delete},a),\textit{call}(\textit{delete},b)$, $\textit{call}($ $\textit{delete},\textit{empty}),\textit{ret}(\textit{delete},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{delete},b)$, $c_3 = c_2 + \textit{ret}(\textit{insert},b)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.6 \textwidth]{figures/PIC_AUTO_MS_3.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{MS}}^3$}
-  \label{fig:automata 3 for MS-2 in appendix}
-\end{figure}
-
-Let $\textit{Auts}_{\textit{2-ms}} = \{ \mathcal{A}_{\textit{MS}}^1, \mathcal{A}_{\textit{MS}}^2, \mathcal{A}_{\textit{MS}}^3 \}$. The following lemma states that $\textit{MS}_2$ is co-regular.
-
-\begin{restatable}{lemma}{MS2IsCoRegular}
-\label{lemma:MS2 is co-regular}
-$\textit{MSet}_2$ is co-regular.
-\end{restatable}
-
-\begin {proof}
-
-According to \cite{Bouajjani:2015}, we need to prove that, given a independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{2-ms}} \cap \mathcal{I} \neq \emptyset$, if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e),$ $\textit{last}(e') = \textit{MSet}_2 \wedge e'$ does not linearizable w.r.t.
-$\textit{MS}(\textit{MSet}_2)$.
-
-By Lemma \ref{lemma:reduce MS2 into one value}, we need to prove the following fact:
-
-\noindent {\bf $\textit{fact}_1$}: given a independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{2-ms}} \cap \mathcal{I} \neq \emptyset$, if and only if $\exists e \in \mathcal{I}_{\neq}$, and one of the following case holds for some $x \in \mathbb{D}$.
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] $\textit{delete}(x)$ is in $e$ while $\textit{insert}(x)$ is not in $e$.
-
-\item[-] there is more than one $\textit{delete}(x)$ in $e$ and one $\textit{insert}(x)$ in $e$.
-
-\item[-] $\textit{delete}(x) <_{\textit{hb}} \textit{insert}(x)$ in $e$.
-\end{itemize}
-
-\noindent The $\textit{only if}$ direction: Assume that $e_1 \in \mathcal{I}$ is accepted by some register automata in $\textit{Auts}_{\textit{1-eq}}$.
-By data-independence, there exists data-differentiated execution $e_2 \in \mathcal{I}$ and a renaming function $r$, such that $e_1=r(e_2)$.
-Since $e_1$ is accepted by some register automata in  $\textit{Auts}_{\textit{1-eq}}$, let $y$ be the item that are renamed into $b$ by $r$.
-Then it is not hard to see that $y$ satisfies one of three conditions in $e_2$.
-
-\noindent The $\textit{if}$ direction: Assume that there exists such $e \in \mathcal{I}_{\neq}$ and $x$.
-Let renaming function $r$ maps $x$ into $b$ and all other items into $a$.
-By data-independence, we can see that $r(e) \in \mathcal{I}$.
-Then it is easy to see that $r(e)$ is accepted by some automaton in $\textit{Auts}_{\textit{2-ms}}$.
-\qed
-\end {proof}
-
-Similar to Lemma \ref{lemma:EPQ3 is co-regular}, we can prove that $\textit{MSet}_3$ is co-regular, as stated by the following lemma.
-
-\begin{restatable}{lemma}{MSet3IsCoRegular}
-\label{lemma:MSet3 is co-regular}
-$\textit{MSet}_3$ is co-regular.
-\end{restatable}
-
-Similar as register automata for $\seqPQ_3$, we have register automata $\mathcal{A}_{\textit{MS}}^4$ for $\textit{MSet}_3$, which is shown in \figurename~\ref{fig:automata 4 for MS-3 in appendix}.
-In \figurename~\ref{fig:automata 4 for MS-3 in appendix}, let $c = \textit{call}(\textit{insert},d),\textit{ret}(\textit{insert},d), \textit{call}(\textit{delete},d)$, $\textit{ret}(\textit{delete},d),\textit{call}(\textit{delete},\textit{empty}),\textit{ret}(\textit{delete},\textit{empty})$, $c_1 = c + \textit{call}(\textit{insert},b)$, $c_2 = c_1 + \textit{ret}(\textit{delete},b)$, and $c_3 = c + \textit{ret}(\textit{delete},b)$.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{figures/PIC_AUTO_MS_4.pdf}
-%\vspace{-10pt}
-  \caption{Automaton $\mathcal{A}_{\textit{MS}}^4$}
-  \label{fig:automata 4 for MS-3 in appendix}
-\end{figure}
-
-The following lemma shows that our transformation from multi-set into extended priority queue is correct.
-
-\RelateMultiSetwithEPQ*
-
-\begin {proof}
-
-For the $\textit{only if}$ direction, given an execution $e_m \in \mathcal{I}_m$ of multi-set, such that $e_m$ is accepted by some automaton in $\textit{Auts}_{\textit{MS}}$.
-If $e_m$ is accepted by $\mathcal{A}_{\textit{MS}}^1$, $\mathcal{A}_{\textit{MS}}^2$, $\mathcal{A}_{\textit{MS}}^3$ or $\mathcal{A}_{\textit{MS}}^4$, then it is easy to see that $\textit{MStoEPQ}(e_m)$ is accepted by $\mathcal{A}_{\textit{SinPri}}^2$, $\mathcal{A}_{\textit{SinPri}}^3$, $\mathcal{A}_{\textit{SinPri}}^1$ or $\mathcal{A}_{\seqPQ}^3$, respectively.
-
-For the $\textit{if}$ direction, given an execution $e_{\textit{epq}}$ of extended priority queue, such that $e_{\textit{epq}} = \textit{MStoEPQ}(e_m)$ for some $e_m$ of multi-set, and $e_{\textit{epq}}$ is accepted by some automaton in $\textit{Auts}_{\seqPQ}$.
-
-According to definition of $\textit{MStoEPQ}$, we can see that (1) in $e_{\textit{epq}}$ there does not exist two items with comparable priorities, and (2) in $e_{\textit{epq}}$, there does not exists two items with a same priority.
-Therefore, the register automata for $\seqPQ_1$ and the register automaton $\mathcal{A}_{\textit{SinPri}}^4$ can be safely ignored.
-Then, if $e_{\textit{epq}}$ is accepted by $\mathcal{A}_{\textit{SinPri}}^1$, $\mathcal{A}_{\textit{SinPri}}^2$, $\mathcal{A}_{\textit{SinPri}}^3$ or $\mathcal{A}_{\seqPQ}^3$, then it is easy to see that $e_m$ is accepted by $\mathcal{A}_{\textit{MS}}^3$, $\mathcal{A}_{\textit{MS}}^1$, $\mathcal{A}_{\textit{MS}}^2$ and $\mathcal{A}_{\textit{MS}}^4$, respectively.
-\qed
-\end {proof}
-
-\subsection{Proofs and Definitions in Subsection \ref{subsec:relate stack with extended priority queue}}
-\label{subsec:appendix proof and definition in section relate stack with extended priority queue}
-}

### 2017-04-22 appendix-EPQ.tex
modify

@@
 
-Let $\textit{Items}(e,p)$ be the set of items with priority $p$ in execution $e$. The following lemma states a method to build the linearization of $e$ w.r.t $\mathsf{MatchedMaxPriority}^{=}$.
+Let $\textit{Items}(e,p)$ be the set of values with priority $p$ in execution $e$. The following lemma states a method to build the linearization of $e$ w.r.t $\mathsf{MatchedMaxPriority}^{=}$.
 
@@
 
-\item[-] For item $z$ with priority smaller than $p$.
+\item[-] For value $z$ with priority smaller than $p$.
 If both $\textit{call}(\textit{put},z,\_)$ and $\textit{call}(\textit{rm},z)$ is before $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just after their call actions.
@@
 Then it is easy to see that $\textit{put}(x,p)$ is after $\textit{put}(y,p)$ in $l$.
-Since $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e)$ holds, all other items in $\textit{Items}(e,p)$ has matched $\textit{put}$ and $\textit{rm}$, and it is easy to see that their $\textit{put}$ and $\textit{rm}$ (except for that of $x$) are all before $\textit{rm}(x)$ in $l$.
-
-For item $z$ with priority smaller than $p$, we can see that there are only two possibilities: (1) $\textit{put}(z,\_)$ and $\textit{rm}(z)$ are both before $\textit{rm}(x)$ in $l$, and (2) $\textit{put}(z,\_)$ and $\textit{rm}(z)$ (if exists) are after before $\textit{rm}(x)$ in $l$.
+Since $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e)$ holds, all other values in $\textit{Items}(e,p)$ has matched $\textit{put}$ and $\textit{rm}$, and it is easy to see that their $\textit{put}$ and $\textit{rm}$ (except for that of $x$) are all before $\textit{rm}(x)$ in $l$.
+
+For value $z$ with priority smaller than $p$, we can see that there are only two possibilities: (1) $\textit{put}(z,\_)$ and $\textit{rm}(z)$ are both before $\textit{rm}(x)$ in $l$, and (2) $\textit{put}(z,\_)$ and $\textit{rm}(z)$ (if exists) are after before $\textit{rm}(x)$ in $l$.
 Therefore, before $\textit{rm}(x)$ in $l$, the $\textit{put}$ and $\textit{rm}$ of $z$ are matched.
@@
 According to the definition of $<_{\textit{ob}}^*$, if $a <_{\textit{pb}}^* b$, then there exists $a_1,\ldots,a_m$, such that $a <_{\textit{pb}} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}} a_m <_{\textit{pb}} b$.
-The following lemma states that, the number of intermediate items $a_i$ is in fact bounded.
+The following lemma states that, the number of intermediate values $a_i$ is in fact bounded.
 
@@
 Given a data-differentiated $p$-execution $e$ where $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e)$ holds.
-Let $a$ and $b$ be items with maximal priority $p$.
+Let $a$ and $b$ be values with maximal priority $p$.
 Assume that $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$.
@@
 Given a data-differentiated execution $e$ with only one maximal priority, two actions $\textit{act}_1$, $\textit{act}_2$ of maximal priority in $e$, and assume that $\textit{act}_1$ is before $\textit{act}_2$ in $e$.
-we say that $\textit{act}_1$, $\textit{act}_2$ is covered by items $d_1,\ldots,d_m$ in $e$, if the priorities of $d_1,\ldots,d_m$ is smaller than that of $\textit{act}_1$ and $\textit{act}_2$, and
+we say that $\textit{act}_1$, $\textit{act}_2$ is covered by values $d_1,\ldots,d_m$ in $e$, if the priorities of $d_1,\ldots,d_m$ is smaller than that of $\textit{act}_1$ and $\textit{act}_2$, and
 
@@
 
-Let us begin to represent several register automata that is used to capture the existence of a data-differentiated execution $e$, $e$ has a projection $e'$, $e'$ has only one maximal priority, $\mathsf{Has\text{-}MatchedMax-}$ $\mathsf{Priority}(e')$ holds, there exists items $a$ and $b$ with maximal priority in $e'$, $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,\_)$ or $\textit{call}(\textit{rm},a)$.
+Let us begin to represent several register automata that is used to capture the existence of a data-differentiated execution $e$, $e$ has a projection $e'$, $e'$ has only one maximal priority, $\mathsf{Has\text{-}MatchedMax-}$ $\mathsf{Priority}(e')$ holds, there exists values $a$ and $b$ with maximal priority in $e'$, $a <_{\textit{pb}}^* b$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,\_)$ or $\textit{call}(\textit{rm},a)$.
 
@@
 \noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$.
-$\mathcal{A}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$ if and only if there exists $e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$, $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e')$ holds, $x$ is the value with maximal priority in $e'$, $e'$ has only one maximal priority, $a$ and $b$ are two items with maximal priority $p$ in $e'$, $a <_{\textit{pb}}^* b$ in $e'$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$ in $e'$.
+$\mathcal{A}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$ if and only if there exists $e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$, $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e')$ holds, $x$ is the value with maximal priority in $e'$, $e'$ has only one maximal priority, $a$ and $b$ are two values with maximal priority $p$ in $e'$, $a <_{\textit{pb}}^* b$ in $e'$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$ in $e'$.
 
@@
 By data-independence, there exists data-differentiated execution $e_2 \in \mathcal{I}$ and a renaming function $r$, such that $e_1=r(e_2)$.
-Since $e_1$ is accepted by some register automata in  $\mathcal{A}_{\textit{1-eq}}$, let $x$, $y$ and $z$ (if exists) be the items that are renamed into $b$, $a$ and $a_1$ (if exists) by $r$, respectively, and let $d_1,\ldots,d_m$ be the items that are renamed into $d$ by $r$.
+Since $e_1$ is accepted by some register automata in  $\mathcal{A}_{\textit{1-eq}}$, let $x$, $y$ and $z$ (if exists) be the values that are renamed into $b$, $a$ and $a_1$ (if exists) by $r$, respectively, and let $d_1,\ldots,d_m$ be the values that are renamed into $d$ by $r$.
 
@@
 
-\noindent The $\textit{if}$ direction: Assume that there exists $e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e')$ holds $e'$ has only one maximal priority, $a'$ and $b'$ are two items with maximal priority $p$ in $e'$, $a' <_{\textit{pb}}^* b'$ in $e'$, and the rightmost gap-point of $b'$ is before $\textit{call}(\textit{put},a',p)$ or $\textit{call}(\textit{rm},a')$ in $e'$.
-By data-independence, we can obtain execution $e_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the items $d_1,\ldots,d_m$ that constitute the rightmost gap of $b'$, we rename them into $d$, (3) if $a' <_{\textit{pb}}^A a'_1 <_{\textit{pb}}^B b$, we rename $a'_1$ into $a_1$, and (4) rename the other items into $\top$.
+\noindent The $\textit{if}$ direction: Assume that there exists $e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e')$ holds $e'$ has only one maximal priority, $a'$ and $b'$ are two values with maximal priority $p$ in $e'$, $a' <_{\textit{pb}}^* b'$ in $e'$, and the rightmost gap-point of $b'$ is before $\textit{call}(\textit{put},a',p)$ or $\textit{call}(\textit{rm},a')$ in $e'$.
+By data-independence, we can obtain execution $e_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the values $d_1,\ldots,d_m$ that constitute the rightmost gap of $b'$, we rename them into $d$, (3) if $a' <_{\textit{pb}}^A a'_1 <_{\textit{pb}}^B b$, we rename $a'_1$ into $a_1$, and (4) rename the other values into $\top$.
 It is easy to see that $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e_1)$ holds, $a$ and $b$ has maximal priority in $e_1$, $a <_{\textit{pb}}^* b$ in $e_1$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$ in $e_1$.
@@
 
-Given a data-differentiated execution $e$, we say that $o = \textit{rm}(\textit{empty})$ in $e$ is covered by items $d_1,\ldots,d_m$, if
+Given a data-differentiated execution $e$, we say that $o = \textit{rm}(\textit{empty})$ in $e$ is covered by values $d_1,\ldots,d_m$, if
 
@@
 
-According to the definition of left-right constraint for $o$, in a data-differentiated execution $e$, there is a cycle going through $o$, if and only if there exists items $d_1,\ldots,d_m$, such that $o$ is covered by $d_1,\ldots,d_m$.
+According to the definition of left-right constraint for $o$, in a data-differentiated execution $e$, there is a cycle going through $o$, if and only if there exists values $d_1,\ldots,d_m$, such that $o$ is covered by $d_1,\ldots,d_m$.
 
@@
 \noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$.
-$\mathcal{A}_{\seqPQ}^3 \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, $\mathsf{Has\text{-}EmptyRemove}(e')$ holds, $o = \textit{rm}(\textit{empty})$ is in $e'$, and $o$ is covered by some items $d_1,\ldots,d_m$ in $e'$.
+$\mathcal{A}_{\seqPQ}^3 \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, $\mathsf{Has\text{-}EmptyRemove}(e')$ holds, $o = \textit{rm}(\textit{empty})$ is in $e'$, and $o$ is covered by some values $d_1,\ldots,d_m$ in $e'$.
 
@@
 By data-independence, there exists data-differentiated execution $e_2 \in \mathcal{I}$ and a renaming function $r$, such that $e_1=r(e_2)$.
-Let $d_1,\ldots,d_m$ be the items in $e_2$ such that $r(d_i)=b$ for each $1 \leq i \leq m$.
+Let $d_1,\ldots,d_m$ be the values in $e_2$ such that $r(d_i)=b$ for each $1 \leq i \leq m$.
 Let $e_3 = e_2 \vert_{ \{ o, d_1, \ldots, d_m \} }$.
@@
 \noindent The $\textit{if}$ direction: Assume that there exists such $e$, $e'$, $o$ and $d_1,\ldots,d_m$.
-Then, let $e_1$ be obtained from $e$ by renaming $d_1,\ldots,d_m$ into $b$ and renaming other items into $\top$.
+Then, let $e_1$ be obtained from $e$ by renaming $d_1,\ldots,d_m$ into $b$ and renaming other values into $\top$.
 By data-independence, $e_1 \in \mathcal{I}$.

### 2017-06-29 appendix-EPQ.tex
modify

@@
 
-\begin{restatable}{lemma}{LinEqualsConstraintforEPQOneLar}
-\label{lemma:Lin Equals Constraint for EPQ1Lar}
-Given a data-differentiated execution $e$ where $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, let $p$ be its maximal priority and $\textit{put}(x,p),\textit{rm}(x)$ are only operations of priority $p$ in $e$.
+%\begin{restatable}{lemma}{LinEqualsConstraintforEPQOneLar}
+%\label{lemma:Lin Equals Constraint for EPQ1Lar}
+%Given a data-differentiated execution $e$ where $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, let $p$ be its maximal priority and $\textit{put}(x,p),\textit{rm}(x)$ are only operations of priority $p$ in $e$.
 Let $G$ be the graph representing the left-right constraint of $x$.
 $e$ is $\mathsf{MatchedMaxPriority}$-linearizable, if and only if $G$ has no cycle going through $x$.
-\end{restatable}
+%\end{restatable}
+
+$\newline$
+{\noindent \bf Lemma \ref{lemma:Lin Equals Constraint for EPQ1Lar}}: Given a data-differentiated execution $e$ where $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, let $p$ be its maximal priority and $\textit{put}(x,p),\textit{rm}(x)$ are only operations of priority $p$ in $e$.
+Let $G$ be the graph representing the left-right constraint of $x$.
+$e$ is $\mathsf{MatchedMaxPriority}$-linearizable, if and only if $G$ has no cycle going through $x$.
 
@@
 %Given a data-differentiated $p$-execution $e$ where $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)$ holds.
-Then, $e$ is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ iff $e$ contains two values $x$ and $y$ of maximal priority $p$ such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is strictly smaller than the index of both $\textit{call}(\textit{put},y,p)$ and $\textit{call}(\textit{rm},y)$.
+Then, $e$ is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ iff $e$ contains two values $x$ and $y$ of maximal priority $p$ such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is strictly smaller than the index of $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$.
 
@@
 \end {proof}
+
+\subsection{Proofs and Definitions in Section \ref{subsec:combine step-by-step linearizability and co-regular}}
+\label{subsec:appendix proof and definition in subsection decidability result}
+
+In this subsection, we show how to prove that the verification of linearizability of priority queue to be PSPACE-complete for a fixed number of threads, and EXPSPACE-complete for unbounded number of processes.
+
+When considering decidability and complexity, we consider only finite variables with finite data domain.
+In such case, the priority queue implementations can only model priority queue with bounded capacity.
+Then, at some time, the priority queue may become full and then put can not succeed until some item is removed.
+However, we assume that each method can always make progress in isolation, which means that $\textit{put}$ must return some value when the priority queue is full.
+Therefore, we introduce a specific value $\textit{full}$.
+A $\textit{put}$ method returns $\textit{full}$ indicates that the priority queue is full now.
+To fit our results of priority queue in former section, when checking linearizability of priority queue, we only consider executions without any $\textit{put}(\_,\_,\textit{full})$.
+Since does not mean we consider only executions with bounded number of operations, since an execution can contain unbounded number of pairs of put and remove operations by linearizable w.r.t pairs of put and remove operations.
+
+\noindent {\bf Vector Addition Systems with States (VASS):} A VASS \cite{conf/esop/BouajjaniEEH13} $\mathcal{A}=(Q_\mathcal{A},\rightarrow_\mathcal{A})$ contains a finite set $Q_\mathcal{A}$ of states and a finite set $\rightarrow_\mathcal{A}$ of transitions.
+Each transition is chosen from $Q_\mathcal{A} \times \mathbb{Z}^s \times Q_\mathcal{A}$, where $s$ is a positive integer.
+A vector that has a single non-zero component in $\{1,-1\}$ is called a unit vector.
+We assume, w.l.o.g., that all vectors of transitions are unit vectors.
+
+The operational semantics of a VASS is defined as a LTS.
+A configuration $(q,\vec n)$ of VASS $\mathcal{A}$ contains a state $q \in Q_\mathcal{A}$ with a vector $\vec n \in \mathbb{N}^s$.
+$\vec n$ can be considered as values of several counters.
+We use $\vec n(i)$ to denote the $i$-th component of $\vec n$.
+The transitions between configurations of VASS is defined as follows: $(q_1, \vec n_1) \rightarrow_{\mathcal{A}} (q_2, \vec n_2)$, if there exists vector $\vec n$, such that (1) $(q_1,\vec n,q_2) \in \rightarrow_\mathcal{A}$, (2) if $\vec n(j)=1$ for some $j$, then $\vec n_2$ is obtain from $\vec n_1$ by increasing the $i$-th component by $1$, and (3) if $\vec n(j)=-1$ for some $j$, then $\vec n_2$ is obtain from $\vec n_1$ by decreasing the $i$-th component by $1$.
+
+The state-reachability problem of VASS is to determine whether some configuration with state $q$ is reachable from a given configuration $(q_0,\vec 0)$.
+Here $\vec 0$ is a specific vector for which all components are $0$ and $q_0$ is the initial state of VASS.
+
+\noindent {\bf Reducing priority Queue into PSPACE and EXPSPACE problems:} We describe a class $\mathcal{C}$ of data-independent implementations for which linearizability w.r.t.
+$\seqPQ$ is decidable.
+The implementations in $\mathcal{C}$ allow an unbounded number of values but a bounded number of priorities.
+Each method manipulates a finite number of local variables which store Boolean values, or data values from $\mathbb{D}$.
+Methods communicate through a finite number of shared variables that also store Boolean values, or data values from $\mathbb{D}$.
+Data values may be assigned, but never used in program predicates (e.g., in the conditions of if-then-else and while statements) so as to ensure data independence.
+This class captures typical implementations, or finite-state abstractions thereof, e.g., obtained via predicate abstraction.
+Since the $\Gamma$-complete automata $A(\Gamma)$ uses a fixed set $D=\{a,b,a_1,d,e,\top\}$ of values, we have that $\mathcal{C}\cap A(\Gamma)\neq\emptyset$ for some $\Gamma$ iff $\mathcal{C}_D\cap A(\Gamma)\neq\emptyset$ where $\mathcal{C}_D$ is the subset of $\mathcal{C}$ that uses only values in $D$.
+
+We model the set $\mathcal{C}_D$ as the executions of a Vector Addition System with States (VASS) similarly to Bouajjani et al.
+\cite{conf/esop/BouajjaniEEH13}.
+A VASS has a finite set of states, and manipulates a finite set of counters holding non-negative values.
+The idea is that the states of the VASS represent the global variables of $\mathcal{C}_D$.
+This set is finite as we have bounded the data values.
+Each counter of the VASS then represents the number of threads which are at a particular control location within a method, with a certain valuation of the local variables.
+Here again, as the data values are bounded, there is a finite number of valuations.
+When a thread moves from a control location to another, or updates its local variables, we decrement, resp., increment, the counter corresponding the old, resp., the new, control location or valuation.
+
+For a fixed set of priorities $P$, the register automata $A(\Gamma)$ can be transformed to finite-state automata since the number of possible valuations of the registers is bounded.
+In this way, we reduce checking linearizability of priority queue into the the EXPSPACE-complete problem of checking state-reachability in a VASS.
+When we consider only finite number of threads, the value of counters are bounded and then this problem is PSPACE \cite{DBLP:conf/ac/Esparza96}.
+
+\noindent {\bf Reducing PSPACE and EXPSPACE problems into linearizability of priority queue for bounded number of threads:} We show how to reduce the state-reachability problem of VASS into checking linearizability of priority queue.
+Our construction is based on the reduction of state-reachability problem of counter machine and VASS into linearizability in \cite{conf/esop/BouajjaniEEH13}.
+
+There are already several linearizable lock-free implementations of priority queue \cite{DBLP:conf/ipps/SundellT03,DBLP:conf/ppopp/LiuS12}.
+It is not hard to modify them to fit finite variables with finite data domain and it is obvious that they are still linearizable then.
+Let $\textit{Imp-PQ}_{\textit{lf}}$ be one such implementations.
+It is safe to assume that $\textit{Imp-PQ}_{\textit{lf}}$ is given in the form of pseudo-code.
+We use $\textit{put}_{\textit{lf}}$ and $\textit{rm}_{\textit{lf}}$ to explicitly denote the pseudo-code of put and remove method of $\textit{Imp-PQ}_{\textit{lf}}$.
+
+A counter machine is a VASS that can additional detect if the value of some counter is $0$.
+In \cite{conf/esop/BouajjaniEEH13}, they construct a library $\mathcal{L}_{\mathcal{A}}$ that simulate executions of counter machine $\mathcal{A}$.
+This library use finite variables with finite data domain.
+It contains the following six methods:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] $M_{(q,\vec n,q')}$, which represents a transition resulting of $(q,\vec n,q')$.
+
+\item[-] $M_{(q,i,q')}$, which represents a transition that from state $q$ to state $q'$ while detecting the value of counter $i$ to be $0$.
+
+\item[-] $M_{q_f}$. This method returns if the simulating reaches some state with state $q_f$.
+
+\item[-] $M_{\textit{inc-i}}$, which simulates increasing counter $i$ with $1$.
+
+\item[-] $M_{\textit{dec-i}}$, which simulates decreasing counter $i$ with $1$.
+
+\item[-] $M_{\textit{zero-i}}$, which simulates testing whether the value of counter $i$ to be $0$.
+\end{itemize}
+
+The first three methods do their work with the help of the latter three methods.
+For example, a $M_{(q,\vec n,q')}$ method (that increases counter $j$) will activate a $M_{inc-j}$ method and a $M_{dec-j}$ method.
+
+Let $D' = \{ (q,\vec n,q') \vert q,q' \in Q_{\mathcal{A}} - \{q_f\}, \vec n \in \mathbb{Z}^s \} \cup \{ q_f \} \cup \{\textit{inc-i},\textit{dec-i} \vert 1 \leq i \leq s\}$.
+Let $\textit{RAs}$ be the register automata obtained as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] Given $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$, we modify the $\Gamma$-complete automata while the original value of $\top$ is replaced by values $\top$ and all values in $D'$.
+
+\item[-] For register automata that checks FIFO of sub-executions of operations with a same priority, if the original value under consideration is $\{ b \}$ (the first three automata) or $\{a,b\}$ (the forth automaton), then we replace them with $\{ b \} \cup D'$ and $\{a,b\} \cup D'$, respectively.
+\end{itemize}
+
+Given a VASS $\mathcal{A}$, we construct a priority queue implementation $\textit{PQ}_{\mathcal{A}}$ as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] A new memory location $\textit{reachQf}$ is introduced. Its initial value is set to false.
+
+\item[-] The $\textit{rm}$ method first check the value of $\textit{reachQf}$.
+If it is true, then it returns a random value.
+Otherwise, it works as $\textit{rm}_{\textit{lf}}$.
+The pseudo-code of $\textit{rm}$ is shown in Algorithm \ref{alg:new rm}.
+\end{itemize}
+
+\begin{algorithm}[t]
+%\KwIn {}
+\KwOut{A removed value}
+
+\If {$\textit{reachQf} = \textit{true}$}
+{\KwRet a random value from $D'$\;}
+\Else
+{\KwRet $\textit{rm}_{\textit{lf}}$()\;}
+
+\caption{$\textit{rm}$ of $\textit{PQ}_{\mathcal{A}}$}
+\label{alg:new rm}
+\end{algorithm}
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] The $\textit{put}$ method uses items from $D' \cup \{ a,b,a_1,d,e,\top \}$.
+It first uses $\textit{put}_{\textit{lf}}$ to do the work of priority queue, and then choose a role and begin to simulate transitions of $\mathcal{A}$, until $q_f$ is reached.
+Then, it set the flag $\textit{reachQf}$ to be $\textit{true}$.
+The pseudo-code of $\textit{put}$ is shown in Algorithm \ref{alg:new put}.
+\end{itemize}
+
+In \cite{conf/esop/BouajjaniEEH13}, to use a library to simulate counter machine $\mathcal{A}$, their library $\mathcal{L}_{\mathcal{A}}$ contains the following variables: (1) variable $q$ that uses value from $Q_{\mathcal{A}}$, (2) vector $\textit{req}[]$ and $\textit{ack}[]$, the index is chosen from $1$ to the number of unit vectors of length $s$, (3) $\textit{dec}[]$, where the index is chosen form $1$ to $s$.
+
+Let us explain how $\textit{put}$ method do to act according to its chosen role.
+We use the case of $\textit{trans}(q,\vec n,q')$ as an example while other cases can be similarly obtained.
+For the role $\textit{trans}(q,\vec n,q')$, it works as $M_{(q,\vec n,q')}$ in $\mathcal{L}_{\mathcal{A}}$.
+Or we can say, it do the following work:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] atomically do $\textit{wait}(q)$ and $\textit{signal}(\textit{req}[\vec n])$,
+\item[-] atomically do $\textit{wait}(\textit{ack}[\vec n])$ and $\textit{signal}(q')$,
+\end{itemize}
+
+Note that we want to obtain a implementation where each method can always make progress in isolation.
+To obtain such an implementation, we need to slightly modify the $\textit{wait}$ command as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If a wait command find itself blocked, then it can still proceed.
+However, in this case the $\textit{put}$ method will return a specific value $\textit{fail}$.
+Moreover, when deciding linearizability, we do not consider executions with return value $\textit{fail}$.
+\end{itemize}
+
+\begin{algorithm}[t]
+\KwIn {An item $\textit{itm}$ and its priority $p$}
+\KwOut{$\textit{full}$ if the priority queue is full}
+
+$x = \textit{put}_{\textit{lf}}(\textit{itm},p)$\;
+
+random select a role from $\{ \textit{trans}(q,\vec n,q'), \textit{reach-}q_f, \textit{inc-i},\textit{dec-i} \}$\;
+
+works according to chosen role\;
+
+\If {$q_f$ is reached}
+{$\textit{reachQf} := \textit{true}$\;}
+
+\If {$x = \textit{full}$}
+{\KwRet $x$\;}
+\Else
+{\KwRet\;}
+\caption{$\textit{put}$ of $\textit{PQ}_{\mathcal{A}}$}
+\label{alg:new put}
+\end{algorithm}
+
+Follows the proof in \cite{conf/esop/BouajjaniEEH13}, we can prove that $(q_f,\_)$ is reachable in $\mathcal{A}$, if and only if $\textit{PQ}_{\mathcal{A}}$ is not linearizable w.r.t $\seqPQ$.
+In this way, we reduce the state-reachability problem of VASS into checking linearizablity of priority queue.
+
+With above discussions, we can now prove the following theorem:
+
+$\newline$
+{\noindent \bf Theorem \ref{theorem:complexity of priority queue}}: Verifying whether an implementation in $\mathcal{C}$ is linearizable w.r.t.
+$\seqPQ$ is PSPACE-complete for a fixed number of threads, and EXPSPACE-complete otherwise.

### 2017-06-29 checking-executions.tex
some stuff
modify

@@
 
-Checking whether a data-differentiated sequential execution belongs to $\seqPQ$ could be implemented by checking membership into the language accepted by the LTS $PQ$.
+Checking whether a data-differentiated sequential execution belongs to $\seqPQ$ could be implemented by checking membership into {\color {blue}the language accepted by the LTS $PQ$}{\color {red}the set of traces of the LTS $PQ$}.
 The recursive procedure $\textit{Check-PQ-Seq}$ outlined in Algorithm~\ref{alg:seq_check} is an alternative to this membership test.
@@
 We explain how the procedure works on the following execution:
+{\color {red}
 \begin{align}
-\hspace{-5mm}\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{rm}(\textit{empty}) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)\label{eq:ex_rec1}
+\hspace{-5mm}\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{rm}(\textit{empty}) \cdot \textit{put}(d,p_2) \cdot \textit{put}(f,p_3) \cdot \textit{rm}(f) \cdot \textit{put}(b,p_1)\label{eq:ex_rec1}
 \end{align}
+}
 where $p_1$, $p_2$, $p_3$ are priorities such that $p_1 \prec p_2$ and $p_1 \prec p_3$, and $p_2$ and $p_3$ are incomparable.
-Since the $\textit{rm}(\textit{empty})$ operations\footnote{An $m(a)$-operation in an execution $e$ is an operation identifier $o$ s.t.
-$e$ contains the actions $\textit{call}_o(m,a)$, $\textit{ret}_o(m,a)$.} are read-only (they don't affect the state of the queue), they are selected first.
+Since the $\textit{rm}(\textit{empty})$ operations %\footnote{An $m(a)$-operation in an execution $e$ is an operation identifier $o$ s.t.
+$e$ contains the actions $\textit{call}_o(m,a)$, $\textit{ret}_o(m,a)$.}
+are read-only (they don't affect the state of the queue), they are selected first.
 Ensuring that an operation $o=\textit{rm}(\textit{empty})$ is correct boils down to checking that every $\textit{put}(x,p)$ operation before $o$ is matched to a $\textit{rm}(x)$ operation which also occurs before $o$.
@@
 Therefore, the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
+{\color {red}
 \begin{align*}
-\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)
+\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(f,p_3) \cdot \textit{rm}(f) \cdot \textit{put}(b,p_1)
 \end{align*}
+}
 When there are no more $\textit{rm}(\textit{empty})$ operations, the procedure selects a $\textit{put}$ operation adding a value with maximal priority which is not removed, and then a pair of $\textit{put}$ and $\textit{rm}$ operations adding and removing the same maximal priority value.
@@
 {This operation is correct if $d$ is the last value with priority $p_2$, and the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
+{\color {red}
 \begin{align*}
-\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)
+\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(f,p_3) \cdot \textit{rm}(f) \cdot \textit{put}(b,p_1)
 \end{align*}
+}
 Since there is no other value of maximal priority which is not removed, the procedure selects a pair of $\textit{put}$/$\textit{rm}$ operations with an argument of maximal priority $p_2$,} for instance, $\textit{put}(c,p_2)$ and $\textit{rm}(c)$.
@@
 \end{align*}}
-where $\textit{priorities}(l)$, resp., $\textit{unmatched-priorities}(l)$, is the set of priorities occurring in $\textit{put}$ operations of $e$, resp., in $\textit{put}$ operations of $e$ for which there is no $\textit{rm}$ operation removing the same value.
+{\color{blue} where $\textit{priorities}(l)$, resp., $\textit{unmatched-priorities}(l)$}{\color{red} where $\textit{priorities}(e)$, resp., $\textit{unmatched-priorities}(e)$}, is the set of priorities occurring in $\textit{put}$ operations of $e$, resp., in $\textit{put}$ operations of $e$ for which there is no $\textit{rm}$ operation removing the same value.
 We call the latter \emph{unmatched} put operations.
@@
 For simplicity, we consider the following syntactic sugar $\mathsf{Has\text{-}MatchedMaxPriority}(e)=\neg \mathsf{Has\text{-}EmptyRemoves}(e)\land \neg \mathsf{Has\text{-}UnmatchedMaxPriority}(e)$.
-By an abuse of notation, we also assume that  $\mathsf{Has\text{-}UnmatchedMaxPriority}(e) \Rightarrow \neg \mathsf{Has\text{-}EmptyRemoves}(e)$ (this is sound by the order of the conditionals in $\textit{Check-PQ-Seq}$.
+By an abuse of notation, we also assume that  $\mathsf{Has\text{-}UnmatchedMaxPriority}(e) \Rightarrow \neg \mathsf{Has\text{-}EmptyRemoves}(e)$ (this is sound by the order of the conditionals in $\textit{Check-PQ-Seq}$).
 
@@
 Assume that there exist two sequential executions $l$ and $l'$ such that $e \sqsubseteq l=u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds, and $e \setminus x \sqsubseteq l' \in \seqPQ$.
-Let $u=\epsilon$, $w$ be any sequence containing the set of operations $\textit{put}(z_2,p_2)$, $\textit{put}(x_3,p)$ and $\textit{rm}(z_1)$ (we distinguish them by adding the suffix $-w$ to their name, e.g., $\textit{rm}(z_1)-w$), and $v$ be any sequence containing the remaining operations.
+{\color {red}Let $u=\epsilon$, $w$ be any sequence containing the set of operations $\textit{put}(z_2,p_2)$ and $\textit{rm}(z_1)$} (we distinguish them by adding the suffix $-w$ to their name, e.g., $\textit{rm}(z_1)-w$), and $v$ be any sequence containing the remaining operations.
 In general, the linearization $l'$ can be defined by choosing for each operation, a point in time between its call and return actions, called \emph{linearization point}.
@@
 The $p$-comparable operations in $W'$ are the same as in $w$.
-To identify the $p$-incomparable operations in $W'$, we search for a $p$-incomparable operation $o$ which either happens before some $p$-comparable operation in $w$, or whose linearization point occurs after $\textit{rm}(x)$.
+To identify the $p$-incomparable operations in $W'$, we search for a $p$-incomparable operation $o$ which either happens before some $p$-comparable operation in $w$, {\color {red}or whose linearization point occurs after $\textit{ret}(\textit{rm},x)$}.
 We add to $W'$ the operation $o$ and all the $p$-incomparable operations occurring after $o$ in $l'$.
 In this example, $o$ is $\textit{rm}(z_1)$ and the only $p$-incomparable operation occurring after $o$ in $l'$ is $\textit{rm}(z_2)$ (they are surrounded by boxes in the figure).
-In this process, whether a $p$-incomparable operation is in $W'$ or not only relies on whether it is before or after $o$ in $l'$.
+In this process, whether a $p$-incomparable operation is in $W'$ or not only relies on whether it is before or after such an $o$ in $l'$.
 
@@
 
-\item[-] In conclusion, we have that $l''_1 = \textit{put}(z_1,p_2) \cdot \textit{put}(x_2,p)$, $l''_2 = \textit{put}(z_2,p_2) \cdot \textit{rm}(x_2) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(y_1)$, and $l''_3 = \textit{put}(x_3,p) \cdot \textit{rm}(z_1) \cdot \textit{rm}(z_2)$.
-\figurename~\ref{fig:concurrent execution for EPQ1}(b) draws linearization points for each operation in $e$ defining the linearization $l''$.
+\item[-] In conclusion, we have that $l''_1 = \textit{put}(z_1,p_2) \cdot \textit{put}(x_2,p)$, $l''_2 = \textit{put}(z_2,p_2) \cdot \textit{rm}(x_2) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(y_1)$, {\color {red}and $l''_3 = \textit{rm}(z_1) \cdot \textit{rm}(z_2)$.} \figurename~\ref{fig:concurrent execution for EPQ1}(b) draws linearization points for each operation in $e$ defining the linearization $l''$.
 \end{itemize}
@@
   \centering
-  \includegraphics[width=1 \textwidth]{figures/PIC-HIS-EPQ1-TwoHis.pdf}
+  \includegraphics[width=1 \textwidth]{figures/PIC-HIS-EPQ1-TwoHis-2.pdf}
 %\vspace{-10pt}
@@
 
-Section~\ref{sec:co-regular of extended priority queues} introduces a characterization of concurrent priority queue violations using a set of \emph{non-recursive} automata (i.e., whose states consist of a fixed number of registers) whose product is equivalent to $\textit{Check-PQ-Conc}$ (modulo renaming of values which is possible by data-independence).
+Section~\ref{sec:co-regular of extended priority queues} introduces a characterization of concurrent priority queue violations using a set of \emph{non-recursive} automata (i.e., whose states consist of a fixed number of registers) whose {\color{red}standard synchronized} product is equivalent to $\textit{Check-PQ-Conc}$ (modulo renaming of values which is possible by data-independence).
 Since $\seqPQ$ is closed under projection (Lemma~\ref{lem:closure_proj}), the recursion in $\textit{Check-PQ-Conc}$ can be eliminated by checking that each projection of a given execution $e$ passes a non-recursive version of $\textit{Check-PQ-Conc}$ where every recursive call is replaced by $\mathsf{true}$.

### 2017-06-29 co-regular-EPQ.tex
modify

@@
 We represent each operation as a time interval whose left, resp., right, bound corresponds to the call, resp., return action.
-Operations adding and removing the same value are aligned vertically.}
+{\color {blue}Operations adding and removing the same value are aligned vertically.}}
   \label{fig:introduce gap for EPQ1Lar}
@@
 The execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} is not $\mathsf{MatchedMaxPriority}^>$-linearizable because the left-right constraint of the maximal priority value $b$ contains a cycle: $f \rightarrow e \rightarrow d \rightarrow b \rightarrow f$.
-The presence of such a cycle is equivalent to the execution not being $\mathsf{MatchedMaxPriority}^>$-linearizable (see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}).
+The presence of such a cycle is equivalent to the execution not being $\mathsf{MatchedMaxPriority}^>$-linearizable (see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}), {\color {red}as indicated by the following lemma:
+
+\begin{restatable}{lemma}{LinEqualsConstraintforEPQOneLar}
+\label{lemma:Lin Equals Constraint for EPQ1Lar}
+Given a data-differentiated execution $e$ where $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, let $p$ be its maximal priority and $\textit{put}(x,p),\textit{rm}(x)$ are only operations of priority $p$ in $e$.
+Let $G$ be the graph representing the left-right constraint of $x$.
+$e$ is $\mathsf{MatchedMaxPriority}$-linearizable, if and only if $G$ has no cycle going through $x$.
+\end{restatable}
+}
+
 %The existence of cycle though value with maximal priority in left-right constraint formally modelled the phenomenon in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
@@
 For example, the case where the put happens-before the remove (as in \figurename~\ref{fig:introduce gap for EPQ1Lar}) is pictured in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
-This automaton captures the three possible ways of ordering the first action $\textit{ret}(\textit{put},a,\_)$ w.r.t.
-the actions with value $b$, which are pictured in \figurename~\ref{fig:executions APQ1Lar-1 in paper} (this action cannot occur after $\textit{call}(\textit{rm},b,\_)$ since $b$ must be covered by the $a$-s).
+{\color {red}This automaton captures the three possible ways of ordering the first action $\textit{ret}(\textit{put},a,\_)$ w.r.t.
+the actions with value $b$, which are pictured in \figurename~\ref{fig:executions APQ1Lar-1 in paper} (a)} (this action cannot occur after $\textit{call}(\textit{rm},b,\_)$ since $b$ must be covered by the $a$-s).
 The paths corresponding to these three possible orders are: $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_{10}$ and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$.
-Automata for the other cases can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
+{\color {red}In \figurename~\ref{fig:executions APQ1Lar-1 in paper}, we show all the four orders of the call/return actions of adding and removing $b$, and also possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$.
+In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, three register automata is constructed according to the cases of \figurename~\ref{fig:executions APQ1Lar-1 in paper} (b), (c) and (d), respectively.}
+{\color {blue}Automata for the other cases can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.}
 
@@
   \centering
-  \includegraphics[width=1 \textwidth]{figures/PIC_HIS_PQ1Lar-pprr-in-paper.pdf}
+  %\includegraphics[width=1 \textwidth]{figures/PIC_HIS_PQ1Lar-pprr-in-paper.pdf}
+  \includegraphics[width=1 \textwidth]{figures/PIC_HIS_PQ1Lar-fouCase.pdf}
 %\vspace{-10pt}
-  \caption{Three possible positions of the first $\textit{call}(\textit{put},a,\_)$ in captured execution}
+  %\caption{Three possible positions of the first $\textit{call}(\textit{put},a,\_)$ in captured execution}
+  \caption{{\color {red}Four cases of ordering actions with value $b$, and possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$}}
   \label{fig:executions APQ1Lar-1 in paper}
@@
 
-To recognize the scenarios in \figurename~\ref{fig:introduce pb order}, we introduce an order $<_{\textit{pb}}$ between values which intuitively, can be thought of as ``a value $a$ is putted before another value $b$''.
+To recognize the scenarios in \figurename~\ref{fig:introduce pb order}, we introduce an order $<_{\textit{pb}}$ between values which intuitively, can be thought of as ``a value $a$ is {\color {red}put} before another value $b$''.
 %Let us introduce put-before order $<_{\textit{pb}}$ to formally state that ``an value is putted before another value''.
@@
 Then, to model the time intervals in which a remove operation like $\textit{rm}(b)$ in \figurename~\ref{fig:introduce pb order}, can be linearized (outside of intervals of smaller priority values) we introduce the notion of gap-point.
-This notion relies on a standard indexing of actions in an execution, starting with 0.
+{\color {blue}This notion relies on a standard indexing of actions in an execution, starting with 0.} {\color {red}Here we assume that the index of actions of an execution starts from $0$.}
 
@@
 %Given a data-differentiated $p$-execution $e$ where $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)$ holds.
-Then, $e$ is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ iff $e$ contains two values $x$ and $y$ of maximal priority $p$ such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is strictly smaller than the index of both $\textit{call}(\textit{put},y,p)$ and $\textit{call}(\textit{rm},y)$.
+Then, $e$ is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ iff $e$ contains two values $x$ and $y$ of maximal priority $p$ such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is strictly smaller than the index of $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$.
 \end{restatable}
@@
 $\mathsf{MatchedMaxPriority}^{=}$, one has to consider all the possible orders between call/return actions of the operations on values $a$, $b$, and $a_i$ in Lemma \ref{lemma:ob order has bounded length}, and the right-most gap point of $b$.
-Excluding the inconsistent cases, we are left with 5 possible orders that are listed in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}.
+Excluding the inconsistent cases, {\color {blue}we are left with 5 possible orders that are listed in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}}{\color {red}we are left with 5 possible orders that are shown in \figurename~\ref{fig:five enumerations}, where $o$ denotes the rightmost gap-point of $b$}.
 For each case, we define an automaton recognizing the induced set of violations.
-For instance, one such case and the corresponding register automaton is shown in \figurename~\ref{fig:an enumeration and its witness automaton} ($o$ denotes the rightmost gap-point of $b$).
-In this case, the conditions in Lemma~\ref{lemma:EPQ1Equal as pb order and gap-point} are equivalent to the fact that intuitively, the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$ is covered by lower priority values (and thus, there is no gap-point of $b$ which occurs after $\textit{call}(\textit{rm},a)$).
+{\color {blue}For instance, one such case and the corresponding register automaton is shown in \figurename~\ref{fig:an enumeration and its witness automaton} ($o$ denotes the rightmost gap-point of $b$).} {\color {red}For instance, the register automata for the case of \figurename~\ref{fig:five enumerations}(a) is shown in \figurename~\ref{fig:an enumeration and its witness automaton}.} In this case, the conditions in Lemma~\ref{lemma:EPQ1Equal as pb order and gap-point} are equivalent to the fact that intuitively, the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$ is covered by lower priority values (and thus, there is no gap-point of $b$ which occurs after $\textit{call}(\textit{rm},a)$).
 Using again the data-independence property, these lower priority values can all be renamed to a fixed value $d$, and the other values to a fixed value $\top$.
@@
   \centering
-  \includegraphics[width=0.8 \textwidth]{figures/PIC-Enumeration-WitnessAutomata.pdf}
+  \includegraphics[width=0.8 \textwidth]{figures/PIC-HIS-FiveEnumerations.pdf}
+%\vspace{-10pt}
+  \caption{{\color {red} Five cases that need to be considered.}}
+  \label{fig:five enumerations}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  %\includegraphics[width=0.8 \textwidth]{figures/PIC-Enumeration-WitnessAutomata.pdf}
+  \includegraphics[width=0.8 \textwidth]{figures/PIC-WitnessAutomata-For1.pdf}
 %\vspace{-10pt}
   \caption{A case in deriving a $\mathsf{MatchedMaxPriority}^{=}$-complete automaton.
-We use the following notations: $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true})$, $\textit{call}(\textit{rm},\top), \textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_1 = C \cup \textit{call}(\textit{put},d,<r) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},d) \}$, $C_4 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},d) \}$.}
+We use the following notations: $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true})$, $\textit{call}(\textit{rm},\top), \textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_1 = C \cup \{ \textit{call}(\textit{put},d,<r) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},d) \}$, $C_4 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},d) \}$.}
   \label{fig:an enumeration and its witness automaton}
@@
 This class captures typical implementations, or finite-state abstractions thereof, e.g., obtained via predicate abstraction.
-Since the $\Gamma$-complete automata $A(\Gamma)$ we define use a fixed set $D=\{a,b,d,e,\top\}$ of values, we have that $\mathcal{I}\cap A(\Gamma)\neq\emptyset$ for some $\Gamma$ iff $\mathcal{I}_D\cap A(\Gamma)\neq\emptyset$ where $\mathcal{I}_D$ is the subset of $\mathcal{I}$ that uses only values in $D$.
-The set of executions $\mathcal{I}_D$ can be represented by a Petri Net since bounding the set of values allows us to represent each thread with a finite-state machine.
-For a fixed set of priorities $P$, the register automata $A(\Gamma)$ can be transformed to finite-state automata since the number of possible valuations of the registers is bounded.
+{\color{blue}Since the $\Gamma$-complete automata $A(\Gamma)$ we define use}{\color {red}Since the $\Gamma$-complete automata $A(\Gamma)$ uses} a fixed set {\color {red}$D=\{a,b,a_1,d,e,\top\}$ of values}, we have that {\color {red}$\mathcal{C}\cap A(\Gamma)\neq\emptyset$ for some $\Gamma$ iff $\mathcal{C}_D\cap A(\Gamma)\neq\emptyset$ where $\mathcal{C}_D$ is the subset of $\mathcal{C}$ that uses only values in $D$.} {\color{blue}The set of executions $\mathcal{C}_D$ can be represented by a Petri Net since bounding the set of values allows us to represent each thread with a finite-state machine.
+For a fixed set of priorities $P$, the register automata $A(\Gamma)$ can be transformed to finite-state automata since the number of possible valuations of the registers is bounded.}
+
+{\color{red} The set of executions $\mathcal{C}_D$ can be represented by a Vector Addition Systems with States (VASS), since both values and priorities are finite, which implies that each thread and register automata can be transformed into finite-state automata.
+To obtain this transformation, the states of the VASS represent the global variables of $\mathcal{C}_D$, while each counter of the VASS then represents the number of threads which are at a particular control location within a method, with a certain valuation of the local variables.
+In this way we transform linearizability problem of priority queue into state-reachability problem of VASS, which is EXPSPACE-complete.
+When we consider only finite number of threads, the value of counters are bounded and then this problem is PSPACE \cite{DBLP:conf/ac/Esparza96}.
+
+On the other hand, we can mimic transitions of VASS by a priority queue implementation.
+To obtain this, we keep $\textit{rm}$ unchanged, while in each $\textit{put}$ method, we first put a value into priority queue, and then either simulate one step of transitions or record one increase/decrease of a counter, similarly as that in \cite{conf/esop/BouajjaniEEH13}.
+In this way we transform the state-reachability problem of VASS into linearizability problem of priority queue.
+Based on above discussion, we have the following complexity result.
+The detailed proof can be found in Appendix \ref{subsec:appendix proof and definition in subsection decidability result}.}
+
+%\begin{restatable}{theorem}{ComplexityOfPriorityQueue}
+%\label{theorem:complexity of priority queue}
+%Verifying whether an implementation in $\mathcal{C}$ is linearizable w.r.t. $\seqPQ$ is PSPACE-complete for a fixed number of %threads, and EXPSPACE-complete otherwise.
+%\end{theorem}
 
 \begin{theorem}
+\label{theorem:complexity of priority queue}
 Verifying whether an implementation in $\mathcal{C}$ is linearizable w.r.t. $\seqPQ$ is PSPACE-complete for a fixed number of threads, and EXPSPACE-complete otherwise.

### 2017-06-29 priority-queues.tex
example
modify

@@
 Each value is assigned with a priority when being added to the data structure (using a call to  $\textit{put}$) and the remove method $\textit{rm}$ removes a value with a minimal priority.
-For generality, we assume that the set of priorities is partially-ordered and that incomparable priorities can be removed in any order.
+For generality, we assume that the set of priorities is partially-ordered.
+Incomparable priorities can be removed in any order.
 In case multiple values are assigned with the same priority, $\textit{rm}$ returns the least recent value (according to a FIFO semantics).
@@
 An LTS is a tuple $A=(Q,\Sigma,\rightarrow,q_0)$, where $Q$ is a set of states, $\Sigma$ is an alphabet of transition labels, $\rightarrow\subseteq Q\times\Sigma\times Q$ is a transition relation and $q_0$ is the initial state.
-We model priority queue as an LTS $\textit{PQ}$ defined as follows: each state of $\textit{PQ}$ is a mapping associating priorities in $\mathbb{P}$ with sequences of values in $\mathbb{D}$ representing a snapshot of the priority queue (for each priority, the values are ordered as they were inserted); the transition labels are $\textit{put}(a,p)$ and $\textit{rm}(a)$ operations; Each transition modifies the state as expected.
+We model priority queue as an LTS $\textit{PQ}$ defined as follows: {\color {blue}each state of $\textit{PQ}$}{\color {red}each state of the LTS $\textit{PQ}$} is a mapping associating priorities in $\mathbb{P}$ with sequences of values in $\mathbb{D}$ representing a snapshot of the priority queue (for each priority, the values are ordered as they were inserted); the transition labels are $\textit{put}(a,p)$ and $\textit{rm}(a)$ operations; Each transition modifies the state as expected.
 For example, $q_1 \xrightarrow{\textit{rm}(\textit{empty})} q_2$ if $q_1 = q_2$, and $q_1$ and $q_2$ map each priority to the empty sequence $\epsilon$.
-Then, $\seqPQ$ is the set of words accepted by $\textit{PQ}$.
+Then, {\color{blue}$\seqPQ$ is the set of words accepted by $\textit{PQ}$}{\color {red}$\seqPQ$ is the set of traces of $\textit{PQ}$}.
 The detailed definition of $\textit{PQ}$ can be found in Appendix \ref{sec:appendix definition of seqPQ and proof of Lemma EQP rules and semantics}.
@@
 
-Given a (concurrent) execution $e$ and a sequential execution $s$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if there is a bijection $f: O_1 \rightarrow O_2$, where $O_1$ and $O_2$ are the set of operations of $e$ and $s$, respectively, such that (1) $o$ and $f(o)$ have the same call and return actions, and (2) if $o_1 <_{\textit{hb}} o_2$, then $f(o_1) <_{\textit{hb}} f(o_2)$.
+Given a (concurrent) execution $e$ and a sequential execution $s$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if there is a bijection $f: O_1 \rightarrow O_2$, where $O_1$ and $O_2$ are the set of operations of $e$ and $s$, respectively, such that (1) {\color{blue}$o$ and $f(o)$ have the same call and return actions}{\color {red}$o$ and $f(o)$ is the same operation}\footnote{An $m(a)$-operation in an execution $e$ is an operation identifier $o$ s.t.
+$e$ contains the actions $\textit{call}_o(m,a)$, $\textit{ret}_o(m,a)$.}, and (2) if $o_1 <_{\textit{hb}} o_2$, then $f(o_1) <_{\textit{hb}} f(o_2)$.
 A (concurrent) execution $e$ is linearizable w.r.t a set $S$ of sequential executions, denoted $e \sqsubseteq S$, if there exists $s \in S$ such that $e \sqsubseteq s$.

### 2017-06-29 co-regular-EPQ.tex
modify

@@
 %Given a data-differentiated $p$-execution $e$ where $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)$ holds.
-Then, $e$ is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ iff $e$ contains two values $x$ and $y$ of maximal priority $p$ such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is strictly smaller than the index of $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$.
+Then, $e$ is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ iff $e$ contains two values $x$ and $y$ of maximal priority $p$ such that $y <_{\textit{pb}}^* x$, {\color {red}and the rightmost gap-point of $x$ is strictly smaller than the index of $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$}.
 \end{restatable}

### 2017-06-29 co-regular-EPQ.tex
modify

@@
 $\seqPQ$).
-This property can be checked separately using the techniques introduced for regular FIFO queues in~\cite{DBLP:conf/icalp/BouajjaniEEH15} (see Appendix~\ref{sec:appendix lemma and register automata for FIFO of single-priority executions} for more details).
+{\color {blue}This property can be checked separately using the techniques introduced for regular FIFO queues in~\cite{DBLP:conf/icalp/BouajjaniEEH15}} {\color {red}This property can be checked separately using register automata obtained from the finite automata in~\cite{DBLP:conf/icalp/BouajjaniEEH15} for FIFO queue} (see Appendix~\ref{sec:appendix lemma and register automata for FIFO of single-priority executions} for more details).
 Note that this assumption excludes some obvious violations, such as a $\textit{rm}(a)$ operation happens before a $\textit{put}(a,p)$ operation, for some $p$.

### 2017-07-01 checking-executions-arxiv.tex
modify

--- 2017-06-29 checking-executions.tex
@@
 
-Checking whether a data-differentiated sequential execution belongs to $\seqPQ$ could be implemented by checking membership into {\color {blue}the language accepted by the LTS $PQ$}{\color {red}the set of traces of the LTS $PQ$}.
+Checking whether a data-differentiated sequential execution belongs to $\seqPQ$ could be implemented by checking membership into the set of traces of the LTS $PQ$.
 The recursive procedure $\textit{Check-PQ-Seq}$ outlined in Algorithm~\ref{alg:seq_check} is an alternative to this membership test.
@@
 We explain how the procedure works on the following execution:
-{\color {red}
+{
 \begin{align}
@@
 Therefore, the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
-{\color {red}
+{
 \begin{align*}
@@
 {This operation is correct if $d$ is the last value with priority $p_2$, and the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
-{\color {red}
+{
 \begin{align*}
@@
 \end{align*}}
-{\color{blue} where $\textit{priorities}(l)$, resp., $\textit{unmatched-priorities}(l)$}{\color{red} where $\textit{priorities}(e)$, resp., $\textit{unmatched-priorities}(e)$}, is the set of priorities occurring in $\textit{put}$ operations of $e$, resp., in $\textit{put}$ operations of $e$ for which there is no $\textit{rm}$ operation removing the same value.
+where $\textit{priorities}(e)$, resp., $\textit{unmatched-priorities}(e)$, is the set of priorities occurring in $\textit{put}$ operations of $e$, resp., in $\textit{put}$ operations of $e$ for which there is no $\textit{rm}$ operation removing the same value.
 We call the latter \emph{unmatched} put operations.
@@
 \noindent
-where $p\prec \textit{priorities}(e)$ when $p\prec p'$ for some $p'\in \textit{priorities}(e)$ (and similarly for $p\prec \textit{unmatched-priorities}(e)$ or $p\preceq \textit{unmatched-priorities}(e)$),
+where $p\prec \textit{priorities}(e)$ when $p\prec p'$ for some $p'\in \textit{priorities}(e)$ (and similarly for $p\prec \textit{unmatched-p}$ $\textit{riorities}(e)$ or $p\preceq \textit{unmatched-priorities}(e)$),
 $\textit{matched}_\prec(e,p)$ holds when each value with priority strictly smaller than $p$ is removed in $e$, and $\textit{matched}(e)$ holds when $\textit{matched}_\prec(e,p)$ holds for each $p\in\mathbb{P}$.
@@
 Assume that there exist two sequential executions $l$ and $l'$ such that $e \sqsubseteq l=u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds, and $e \setminus x \sqsubseteq l' \in \seqPQ$.
-{\color {red}Let $u=\epsilon$, $w$ be any sequence containing the set of operations $\textit{put}(z_2,p_2)$ and $\textit{rm}(z_1)$} (we distinguish them by adding the suffix $-w$ to their name, e.g., $\textit{rm}(z_1)-w$), and $v$ be any sequence containing the remaining operations.
+Let $u=\epsilon$, $w$ be any sequence containing the set of operations $\textit{put}(z_2,p_2)$ and $\textit{rm}(z_1)$ (we distinguish them by adding the suffix $-w$ to their name, e.g., $\textit{rm}(z_1)-w$), and $v$ be any sequence containing the remaining operations.
 In general, the linearization $l'$ can be defined by choosing for each operation, a point in time between its call and return actions, called \emph{linearization point}.
@@
 The $p$-comparable operations in $W'$ are the same as in $w$.
-To identify the $p$-incomparable operations in $W'$, we search for a $p$-incomparable operation $o$ which either happens before some $p$-comparable operation in $w$, {\color {red}or whose linearization point occurs after $\textit{ret}(\textit{rm},x)$}.
+To identify the $p$-incomparable operations in $W'$, we search for a $p$-incomparable operation $o$ which either happens before some $p$-comparable operation in $w$, or whose linearization point occurs after $\textit{ret}(\textit{rm},x)$.
 We add to $W'$ the operation $o$ and all the $p$-incomparable operations occurring after $o$ in $l'$.
@@
 
-\item[-] In conclusion, we have that $l''_1 = \textit{put}(z_1,p_2) \cdot \textit{put}(x_2,p)$, $l''_2 = \textit{put}(z_2,p_2) \cdot \textit{rm}(x_2) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(y_1)$, {\color {red}and $l''_3 = \textit{rm}(z_1) \cdot \textit{rm}(z_2)$.} \figurename~\ref{fig:concurrent execution for EPQ1}(b) draws linearization points for each operation in $e$ defining the linearization $l''$.
+\item[-] In conclusion, we have that $l''_1 = \textit{put}(z_1,p_2) \cdot \textit{put}(x_2,p)$, $l''_2 = \textit{put}(z_2,p_2) \cdot \textit{rm}(x_2) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(y_1)$, and $l''_3 = \textit{rm}(z_1) \cdot \textit{rm}(z_2)$.
+\figurename~\ref{fig:concurrent execution for EPQ1}(b) draws linearization points for each operation in $e$ defining the linearization $l''$.
 \end{itemize}
@@
 
-Section~\ref{sec:co-regular of extended priority queues} introduces a characterization of concurrent priority queue violations using a set of \emph{non-recursive} automata (i.e., whose states consist of a fixed number of registers) whose {\color{red}standard synchronized} product is equivalent to $\textit{Check-PQ-Conc}$ (modulo renaming of values which is possible by data-independence).
+Section~\ref{sec:co-regular of extended priority queues} introduces a characterization of concurrent priority queue violations using a set of \emph{non-recursive} automata (i.e., whose states consist of a fixed number of registers) whose standard synchronized product is equivalent to $\textit{Check-PQ-Conc}$ (modulo renaming of values which is possible by data-independence).
 Since $\seqPQ$ is closed under projection (Lemma~\ref{lem:closure_proj}), the recursion in $\textit{Check-PQ-Conc}$ can be eliminated by checking that each projection of a given execution $e$ passes a non-recursive version of $\textit{Check-PQ-Conc}$ where every recursive call is replaced by $\mathsf{true}$.
 More precisely, every occurrence of $\text{{\bf return}}\ \textit{Check-PQ-Conc}$ is replaced by  $\text{{\bf return}}\ \mathsf{true}$.
-Let $\textit{Check-PQ-Conc-NonRec}$ be the thus obtained procedure.
+Let $\textit{Check-PQ-Conc-}$ $\textit{NonRec}$ be the thus obtained procedure.
 

### 2017-07-01 co-regular-EPQ-arxiv.tex
modify

--- 2017-06-29 co-regular-EPQ.tex
@@
 \begin{lemma}
-There exists a $\Gamma$-complete automaton for each $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$.
+There exists a $\Gamma$-complete automaton for each $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMax}$ $\mathsf{Priority}$, $\mathsf{MatchedMaxPriority}\}$.
 \end{lemma}
@@
 $\seqPQ$).
-This property can be checked separately using the techniques introduced for regular FIFO queues in~\cite{DBLP:conf/icalp/BouajjaniEEH15} (see Appendix~\ref{sec:appendix lemma and register automata for FIFO of single-priority executions} for more details).
+This property can be checked separately using register automata obtained from the finite automata in~\cite{DBLP:conf/icalp/BouajjaniEEH15} for FIFO queue (see Appendix~\ref{sec:appendix lemma and register automata for FIFO of single-priority executions} for more details).
 Note that this assumption excludes some obvious violations, such as a $\textit{rm}(a)$ operation happens before a $\textit{put}(a,p)$ operation, for some $p$.
@@
 $l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,x)$.
-Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,x)$ imposes no restriction on the operations in $u$, $v$ and $w$ with priorities incomparable to $p$, erasing all these operations results in a sequential execution which still satisfies this property.
+Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}$ $(s,x)$ imposes no restriction on the operations in $u$, $v$ and $w$ with priorities incomparable to $p$, erasing all these operations results in a sequential execution which still satisfies this property.
 Similarly, for $\Gamma=\mathsf{UnmatchedMaxPriority}$.
@@
 We represent each operation as a time interval whose left, resp., right, bound corresponds to the call, resp., return action.
-{\color {blue}Operations adding and removing the same value are aligned vertically.}}
+Operations adding and removing the same value are aligned vertically.}
   \label{fig:introduce gap for EPQ1Lar}
@@
 The execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} is not $\mathsf{MatchedMaxPriority}^>$-linearizable because the left-right constraint of the maximal priority value $b$ contains a cycle: $f \rightarrow e \rightarrow d \rightarrow b \rightarrow f$.
-The presence of such a cycle is equivalent to the execution not being $\mathsf{MatchedMaxPriority}^>$-linearizable (see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}), {\color {red}as indicated by the following lemma:
+The presence of such a cycle is equivalent to the execution not being $\mathsf{MatchedMaxPriority}^>$-linearizable (see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}), as indicated by the following lemma:
 
@@
 \end{restatable}
-}
 
@@
 For example, the case where the put happens-before the remove (as in \figurename~\ref{fig:introduce gap for EPQ1Lar}) is pictured in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
-{\color {red}This automaton captures the three possible ways of ordering the first action $\textit{ret}(\textit{put},a,\_)$ w.r.t.
-the actions with value $b$, which are pictured in \figurename~\ref{fig:executions APQ1Lar-1 in paper} (a)} (this action cannot occur after $\textit{call}(\textit{rm},b,\_)$ since $b$ must be covered by the $a$-s).
+This automaton captures the three possible ways of ordering the first action $\textit{ret}(\textit{put},a,\_)$ w.r.t.
+the actions with value $b$, which are pictured in \figurename~\ref{fig:executions APQ1Lar-1 in paper} (a) (this action cannot occur after $\textit{call}(\textit{rm},b,\_)$ since $b$ must be covered by the $a$-s).
 The paths corresponding to these three possible orders are: $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_{10}$ and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$.
-{\color {red}In \figurename~\ref{fig:executions APQ1Lar-1 in paper}, we show all the four orders of the call/return actions of adding and removing $b$, and also possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$.
-In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, three register automata is constructed according to the cases of \figurename~\ref{fig:executions APQ1Lar-1 in paper} (b), (c) and (d), respectively.}
-{\color {blue}Automata for the other cases can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.}
+In \figurename~\ref{fig:executions APQ1Lar-1 in paper}, we show all the four orders of the call/return actions of adding and removing $b$, and also possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$.
+In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, three register automata is constructed according to the cases of \figurename~\ref{fig:executions APQ1Lar-1 in paper} (b), (c) and (d), respectively.
 
@@
   %\caption{Three possible positions of the first $\textit{call}(\textit{put},a,\_)$ in captured execution}
-  \caption{{\color {red}Four cases of ordering actions with value $b$, and possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$}}
+  \caption{Four cases of ordering actions with value $b$, and possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$}
   \label{fig:executions APQ1Lar-1 in paper}
@@
 
-To recognize the scenarios in \figurename~\ref{fig:introduce pb order}, we introduce an order $<_{\textit{pb}}$ between values which intuitively, can be thought of as ``a value $a$ is {\color {red}put} before another value $b$''.
+To recognize the scenarios in \figurename~\ref{fig:introduce pb order}, we introduce an order $<_{\textit{pb}}$ between values which intuitively, can be thought of as ``a value $a$ is put before another value $b$''.
 %Let us introduce put-before order $<_{\textit{pb}}$ to formally state that ``an value is putted before another value''.
@@
 Then, to model the time intervals in which a remove operation like $\textit{rm}(b)$ in \figurename~\ref{fig:introduce pb order}, can be linearized (outside of intervals of smaller priority values) we introduce the notion of gap-point.
-{\color {blue}This notion relies on a standard indexing of actions in an execution, starting with 0.} {\color {red}Here we assume that the index of actions of an execution starts from $0$.}
+Here we assume that the index of actions of an execution starts from $0$.
 
@@
 $\mathsf{MatchedMaxPriority}^{=}$, one has to consider all the possible orders between call/return actions of the operations on values $a$, $b$, and $a_i$ in Lemma \ref{lemma:ob order has bounded length}, and the right-most gap point of $b$.
-Excluding the inconsistent cases, {\color {blue}we are left with 5 possible orders that are listed in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}}{\color {red}we are left with 5 possible orders that are shown in \figurename~\ref{fig:five enumerations}, where $o$ denotes the rightmost gap-point of $b$}.
+Excluding the inconsistent cases, we are left with 5 possible orders that are shown in \figurename~\ref{fig:five enumerations}, where $o$ denotes the rightmost gap-point of $b$.
 For each case, we define an automaton recognizing the induced set of violations.
-{\color {blue}For instance, one such case and the corresponding register automaton is shown in \figurename~\ref{fig:an enumeration and its witness automaton} ($o$ denotes the rightmost gap-point of $b$).} {\color {red}For instance, the register automata for the case of \figurename~\ref{fig:five enumerations}(a) is shown in \figurename~\ref{fig:an enumeration and its witness automaton}.} In this case, the conditions in Lemma~\ref{lemma:EPQ1Equal as pb order and gap-point} are equivalent to the fact that intuitively, the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$ is covered by lower priority values (and thus, there is no gap-point of $b$ which occurs after $\textit{call}(\textit{rm},a)$).
+For instance, the register automata for the case of \figurename~\ref{fig:five enumerations}(a) is shown in \figurename~\ref{fig:an enumeration and its witness automaton}.
+In this case, the conditions in Lemma~\ref{lemma:EPQ1Equal as pb order and gap-point} are equivalent to the fact that intuitively, the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$ is covered by lower priority values (and thus, there is no gap-point of $b$ which occurs after $\textit{call}(\textit{rm},a)$).
 Using again the data-independence property, these lower priority values can all be renamed to a fixed value $d$, and the other values to a fixed value $\top$.
@@
 %\vspace{-10pt}
-  \caption{{\color {red} Five cases that need to be considered.}}
+  \caption{Five cases that need to be considered.}
   \label{fig:five enumerations}
@@
 This class captures typical implementations, or finite-state abstractions thereof, e.g., obtained via predicate abstraction.
-{\color{blue}Since the $\Gamma$-complete automata $A(\Gamma)$ we define use}{\color {red}Since the $\Gamma$-complete automata $A(\Gamma)$ uses} a fixed set {\color {red}$D=\{a,b,a_1,d,e,\top\}$ of values}, we have that {\color {red}$\mathcal{C}\cap A(\Gamma)\neq\emptyset$ for some $\Gamma$ iff $\mathcal{C}_D\cap A(\Gamma)\neq\emptyset$ where $\mathcal{C}_D$ is the subset of $\mathcal{C}$ that uses only values in $D$.} {\color{blue}The set of executions $\mathcal{C}_D$ can be represented by a Petri Net since bounding the set of values allows us to represent each thread with a finite-state machine.
-For a fixed set of priorities $P$, the register automata $A(\Gamma)$ can be transformed to finite-state automata since the number of possible valuations of the registers is bounded.}
+Since the $\Gamma$-complete automata $A(\Gamma)$ uses a fixed set $D=\{a,b,a_1,d,e,\top\}$ of values, we have that $\mathcal{C}\cap A(\Gamma)\neq\emptyset$ for some $\Gamma$ iff $\mathcal{C}_D\cap A(\Gamma)\neq\emptyset$ where $\mathcal{C}_D$ is the subset of $\mathcal{C}$ that uses only values in $D$.
 
-{\color{red} The set of executions $\mathcal{C}_D$ can be represented by a Vector Addition Systems with States (VASS), since both values and priorities are finite, which implies that each thread and register automata can be transformed into finite-state automata.
+The set of executions $\mathcal{C}_D$ can be represented by a Vector Addition Systems with States (VASS), since both values and priorities are finite, which implies that each thread and register automata can be transformed into finite-state automata.
 To obtain this transformation, the states of the VASS represent the global variables of $\mathcal{C}_D$, while each counter of the VASS then represents the number of threads which are at a particular control location within a method, with a certain valuation of the local variables.
@@
 Based on above discussion, we have the following complexity result.
-The detailed proof can be found in Appendix \ref{subsec:appendix proof and definition in subsection decidability result}.}
+The detailed proof can be found in Appendix \ref{subsec:appendix proof and definition in subsection decidability result}.
 

### 2017-07-01 priority-queues-arxiv.tex
modify

--- 2017-06-29 priority-queues.tex
@@
 An LTS is a tuple $A=(Q,\Sigma,\rightarrow,q_0)$, where $Q$ is a set of states, $\Sigma$ is an alphabet of transition labels, $\rightarrow\subseteq Q\times\Sigma\times Q$ is a transition relation and $q_0$ is the initial state.
-We model priority queue as an LTS $\textit{PQ}$ defined as follows: {\color {blue}each state of $\textit{PQ}$}{\color {red}each state of the LTS $\textit{PQ}$} is a mapping associating priorities in $\mathbb{P}$ with sequences of values in $\mathbb{D}$ representing a snapshot of the priority queue (for each priority, the values are ordered as they were inserted); the transition labels are $\textit{put}(a,p)$ and $\textit{rm}(a)$ operations; Each transition modifies the state as expected.
+We model priority queue as an LTS $\textit{PQ}$ defined as follows: each state of the LTS $\textit{PQ}$ is a mapping associating priorities in $\mathbb{P}$ with sequences of values in $\mathbb{D}$ representing a snapshot of the priority queue (for each priority, the values are ordered as they were inserted); the transition labels are $\textit{put}(a,p)$ and $\textit{rm}(a)$ operations; Each transition modifies the state as expected.
 For example, $q_1 \xrightarrow{\textit{rm}(\textit{empty})} q_2$ if $q_1 = q_2$, and $q_1$ and $q_2$ map each priority to the empty sequence $\epsilon$.
-Then, {\color{blue}$\seqPQ$ is the set of words accepted by $\textit{PQ}$}{\color {red}$\seqPQ$ is the set of traces of $\textit{PQ}$}.
+Then, $\seqPQ$ is the set of traces of $\textit{PQ}$.
 The detailed definition of $\textit{PQ}$ can be found in Appendix \ref{sec:appendix definition of seqPQ and proof of Lemma EQP rules and semantics}.
@@
 
-Given a (concurrent) execution $e$ and a sequential execution $s$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if there is a bijection $f: O_1 \rightarrow O_2$, where $O_1$ and $O_2$ are the set of operations of $e$ and $s$, respectively, such that (1) {\color{blue}$o$ and $f(o)$ have the same call and return actions}{\color {red}$o$ and $f(o)$ is the same operation}\footnote{An $m(a)$-operation in an execution $e$ is an operation identifier $o$ s.t.
+Given a (concurrent) execution $e$ and a sequential execution $s$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if there is a bijection $f: O_1 \rightarrow O_2$, where $O_1$ and $O_2$ are the set of operations of $e$ and $s$, respectively, such that (1) $o$ and $f(o)$ is the same operation\footnote{An $m(a)$-operation in an execution $e$ is an operation identifier $o$ s.t.
 $e$ contains the actions $\textit{call}_o(m,a)$, $\textit{ret}_o(m,a)$.}, and (2) if $o_1 <_{\textit{hb}} o_2$, then $f(o_1) <_{\textit{hb}} f(o_2)$.

### 2017-07-01 checking-executions.tex
modify

@@
 
-The following lemma states the correctness of $\textit{Check-PQ-Seq}$ (see Appendix~\ref{sec:appendix definition of seqPQ and proof of Lemma EQP rules and semantics} for the proof).
+The following lemma states the correctness of $\textit{Check-PQ-Seq}$. %(see Appendix~\ref{sec:appendix definition of seqPQ and proof of Lemma EQP rules and semantics} for the proof).
 
@@
 
-Proving soundness is highly non-trivial and one of the main technical contributions of this paper (see Appendix~\ref{sec:appendix subsection proof of lemma con-check-EPQ is correct} for a complete proof).
+Proving soundness is highly non-trivial and one of the main technical contributions of this paper.
+%(see Appendix~\ref{sec:appendix subsection proof of lemma con-check-EPQ is correct} for a complete proof).
 The main technical difficulty is showing that for any execution $e$, any linearization of $e\setminus x$ for some maximal priority value $x$ can be extended to a linearization of $e$ provided that $\mathsf{UnmatchedMaxPriority}$ or $\mathsf{MatchedMaxPriority}$ holds (depending on whether there are values with the same priority as $x$ in $e$ which are not removed).

### 2017-07-01 co-regular-EPQ.tex
modify

--- 2017-06-29 co-regular-EPQ.tex
@@
 
-Section~\ref{ssec:aut} describes a $\mathsf{MatchedMaxPriority}$-complete automaton, the other automata being defined in Appendix~\ref{sec:appendix lemma and register automata for FIFO of single-priority executions} %Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar},
+Section~\ref{ssec:aut} describes a $\mathsf{MatchedMaxPriority}$-complete automaton, {\color {red}the other automata can be found in the long version \cite{CONCUR2017Ahmed}.
+Therefore, the following holds.}
+
+%the other automata being defined in Appendix~\ref{sec:appendix lemma and register automata for FIFO of single-priority executions} %Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar},
 %Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}
-and Appendix \ref{subsec:co-regular of EPQ3}. Therefore, the following holds.
+%and Appendix \ref{subsec:co-regular of EPQ3}. Therefore, the following holds.
 
@@
 $\seqPQ$).
-This property can be checked separately using the techniques introduced for regular FIFO queues in~\cite{DBLP:conf/icalp/BouajjaniEEH15} (see Appendix~\ref{sec:appendix lemma and register automata for FIFO of single-priority executions} for more details).
+{\color {blue}This property can be checked separately using the techniques introduced for regular FIFO queues in~\cite{DBLP:conf/icalp/BouajjaniEEH15}} {\color {red}This property can be checked separately using register automata (can be found in the long version \cite{CONCUR2017Ahmed}) obtained from the finite automata in~\cite{DBLP:conf/icalp/BouajjaniEEH15} for FIFO queue}.
+%(see Appendix~\ref{sec:appendix lemma and register automata for FIFO of single-priority executions} for more details).
 Note that this assumption excludes some obvious violations, such as a $\textit{rm}(a)$ operation happens before a $\textit{put}(a,p)$ operation, for some $p$.
@@
 
-The $\textit{if}$ direction follows from the fact that if the projection of an execution to a set of operations $O_1$ has a linearization $l_1$ and the projection of the same execution to the remaining set of operations has a linearization $l_2$, then the execution has a linearization which is defined as an interleaving of $l_1$ and $l_2$ (see Appendix~\ref{sec:appendix proof of Lemma pri execution is enough} for more details).
+The $\textit{if}$ direction follows from the fact that if the projection of an execution to a set of operations $O_1$ has a linearization $l_1$ and the projection of the same execution to the remaining set of operations has a linearization $l_2$, then the execution has a linearization which is defined as an interleaving of $l_1$ and $l_2$.
+%(see Appendix~\ref{sec:appendix proof of Lemma pri execution is enough} for more details).
 
@@
 The paths corresponding to these three possible orders are: $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_{10}$ and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$.
-{\color {red}In \figurename~\ref{fig:executions APQ1Lar-1 in paper}, we show all the four orders of the call/return actions of adding and removing $b$, and also possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$.
-In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, three register automata is constructed according to the cases of \figurename~\ref{fig:executions APQ1Lar-1 in paper} (b), (c) and (d), respectively.}
+{\color {red}In \figurename~\ref{fig:executions APQ1Lar-1 in paper}, we show all the four orders of the call/return actions of adding and removing $b$, and also possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$.} {\color {red}In the long version \cite{CONCUR2017Ahmed}, three register automata is constructed according to the cases of \figurename~\ref{fig:executions APQ1Lar-1 in paper} (b), (c) and (d), respectively.}
+%In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, three register automata is constructed according to the cases of \figurename~\ref{fig:executions APQ1Lar-1 in paper} (b), (c) and (d), respectively.}
 {\color {blue}Automata for the other cases can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.}
@@
 The case of \figurename~\ref{fig:introduce pb order} can be formally described as follows: $a <_{\textit{pb}}^* b$ while the right-most gap-point of $b$ is before $\textit{call}(\textit{rm},a)$ or $\textit{call}(\textit{put},a,p_4)$.
-The following lemma states that these conditions are enough to characterize non-linearizability w.r.t $\mathsf{MatchedMaxPriority}^{=}$ (see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}).
+The following lemma states that these conditions are enough to characterize non-linearizability w.r.t $\mathsf{MatchedMaxPriority}^{=}$.
+%(see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}).
 
@@
 Using again the data-independence property, these lower priority values can all be renamed to a fixed value $d$, and the other values to a fixed value $\top$.
+{\color {red}We show how to generate register automata for other cases in the long version \cite{CONCUR2017Ahmed}.}
 
@@
 In this way we transform the state-reachability problem of VASS into linearizability problem of priority queue.
-Based on above discussion, we have the following complexity result.
-The detailed proof can be found in Appendix \ref{subsec:appendix proof and definition in subsection decidability result}.}
+Based on above discussion, we have the following complexity result.} %The detailed proof can be found in Appendix \ref{subsec:appendix proof and definition in subsection decidability result}.
 

### 2017-07-01 priority-queues.tex
modify

@@
 Then, {\color{blue}$\seqPQ$ is the set of words accepted by $\textit{PQ}$}{\color {red}$\seqPQ$ is the set of traces of $\textit{PQ}$}.
-The detailed definition of $\textit{PQ}$ can be found in Appendix \ref{sec:appendix definition of seqPQ and proof of Lemma EQP rules and semantics}.
+{\color {red}The detailed definition of $\textit{PQ}$ can be found in the long version \cite{CONCUR2017Ahmed}.}
+%{The detailed definition of $\textit{PQ}$ can be found in Appendix \ref{sec:appendix definition of seqPQ and proof of Lemma EQP rules and semantics}.}
 
@@
 Let $\textit{proj}(e)$ be the set of all projections of $e$ to a set of values.
-The proof of the following lemma can be found in Appendix \ref{sec:appendix proofs in section priority queue and data-independence}.
+{\color {red}The following lemma states that $\seqPQ$ is closed under projection.}
+
+%The proof of the following lemma can be found in Appendix \ref{sec:appendix proofs in section priority queue and data-independence}.
 
@@
 
-The following lemma states that by data-independence, it is enough to consider only data-differentiated executions when checking linearizability (see Appendix \ref{sec:appendix proofs in section priority queue and data-independence}).
+The following lemma states that by data-independence, it is enough to consider only data-differentiated executions when checking linearizability.
+%(see Appendix \ref{sec:appendix proofs in section priority queue and data-independence}).
 This is similar to that in \cite{conf/tacas/AbdullaHHJR13,DBLP:conf/icalp/BouajjaniEEH15}, where they use the notion of data-independence in \cite{conf/popl/Wolper86}.

### 2017-07-10 final/abstract.tex
half of the paper

%!TEX root = draft.tex
\begin{abstract}
Efficient implementations of concurrent objects such as atomic collections are essential to modern computing. 
Unfortunately their correctness criteria — linearizability with respect to given ADT specifications — are hard to verify.
%Programming such objects is error prone: in minimizing the synchronization overhead between concurrent object invocations, one risks the conformance to sequential specifications -- or in formal terms, one risks violating linearizability. Unfortunately, 
Verifying linearizability is undecidable in general, even on classes of implementations where the usual control-state reachability is decidable. In this work we consider concurrent priority queues which are fundamental to many multi-threaded applications like task scheduling or discrete event simulation, and show that verifying linearizability of such implementations is reducible to control-state reachability. This reduction entails the first decidability results for verifying concurrent priority queues with an unbounded number of threads, and it enables the application of existing safety-verification tools for establishing their correctness.
\end{abstract}

\forget{
\noindent Keywords: weak memory model, $\textit{linearizability}$,
$\textit{TSO-to-TSO linearizability}$
}

### 2017-07-10 final/checking-executions.tex
half of the paper

--- 2017-07-01 checking-executions.tex
@@
 %While checking linearizability of a given execution is known to be NP-complete in general~\cite{journals/siamcomp/GibbonsK97} we show in this section that it is polynomial-time for data-differentiated executions of the priority queue.
-We define a recursive procedure for checking linearizability of an execution w.r.t. $\seqPQ$.
+We define a recursive procedure for checking linearizability of a data-differentiated execution w.r.t. $\seqPQ$.
 To ease the exposition, Section~\ref{ssec:seq_exec} introduces a recursive procedure for checking whether a data-differentiated \emph{sequential} execution is admitted by the priority queue which is then extended to the concurrent case in Section~\ref{ssec:conc_exec}.
@@
 
-Checking whether a data-differentiated sequential execution belongs to $\seqPQ$ could be implemented by checking membership into {\color {blue}the language accepted by the LTS $PQ$}{\color {red}the set of traces of the LTS $PQ$}.
-The recursive procedure $\textit{Check-PQ-Seq}$ outlined in Algorithm~\ref{alg:seq_check} is an alternative to this membership test.
+%Checking whether a data-differentiated sequential execution belongs to $\seqPQ$ could be implemented by checking membership into {\color {blue}the language accepted by the LTS $PQ$}{\color {red}the set of traces of the LTS $PQ$}.
+The recursive procedure $\textit{Check-PQ-Seq}$ outlined in Algorithm~\ref{alg:seq_check} checks whether a data-differentiated sequential execution belongs to $\seqPQ$ (i.e., if it is accepted by the LTS $PQ$).
+%is an alternative to this membership test. 
  Roughly, it selects one or two operations in the input execution, checks whether their return values are correct by ignoring the order between the other operations other than how they are ordered w.r.t.
@@
 We explain how the procedure works on the following execution:
-{\color {red}
 \begin{align}
-\hspace{-5mm}\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{rm}(\textit{empty}) \cdot \textit{put}(d,p_2) \cdot \textit{put}(f,p_3) \cdot \textit{rm}(f) \cdot \textit{put}(b,p_1)\label{eq:ex_rec1}
+\hspace{-5mm}\textit{put}(c,p_2)\hspace{-.5mm} \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{rm}(\textit{empty}) \cdot \textit{put}(d,p_2) \cdot \textit{put}(\mathit{f},p_3) \cdot \textit{rm}(\mathit{f}) \cdot \textit{put}(b,p_1)\label{eq:ex_rec1}
 \end{align}
-}
 where $p_1$, $p_2$, $p_3$ are priorities such that $p_1 \prec p_2$ and $p_1 \prec p_3$, and $p_2$ and $p_3$ are incomparable.
@@
 $e$ contains the actions $\textit{call}_o(m,a)$, $\textit{ret}_o(m,a)$.}
-are read-only (they don't affect the state of the queue), they are selected first.
-Ensuring that an operation $o=\textit{rm}(\textit{empty})$ is correct boils down to checking that every $\textit{put}(x,p)$ operation before $o$ is matched to a $\textit{rm}(x)$ operation which also occurs before $o$.
+are read-only (they don't affect the queue's state), they are selected first.
+An $\textit{rm}(\textit{empty})$-operation $o$ is correct when every $\textit{put}(x,p)$ operation before $o$ is matched to a $\textit{rm}(x)$ operation which also occurs before $o$.
 This is true in this case for $x\in \{a,c\}$.
-Therefore, the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
-{\color {red}
-\begin{align*}
-\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(f,p_3) \cdot \textit{rm}(f) \cdot \textit{put}(b,p_1)
-\end{align*}
-}
-When there are no more $\textit{rm}(\textit{empty})$ operations, the procedure selects a $\textit{put}$ operation adding a value with maximal priority which is not removed, and then a pair of $\textit{put}$ and $\textit{rm}$ operations adding and removing the same maximal priority value.
-For instance, since $p_2$ is a maximal priority, it selects the operation $\textit{put}(d,p_2)$.
-{This operation is correct if $d$ is the last value with priority $p_2$, and the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
-{\color {red}
-\begin{align*}
-\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(f,p_3) \cdot \textit{rm}(f) \cdot \textit{put}(b,p_1)
-\end{align*}
-}
-Since there is no other value of maximal priority which is not removed, the procedure selects a pair of $\textit{put}$/$\textit{rm}$ operations with an argument of maximal priority $p_2$,} for instance, $\textit{put}(c,p_2)$ and $\textit{rm}(c)$.
-The value returned by $\textit{rm}$ is correct if all the values of priority smaller than $p_2$ added before $\textit{rm}(c)$ are also removed before $\textit{rm}(c)$.
+Thus, the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
+\begin{align}
+\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(f,p_3) \cdot \textit{rm}(f) \cdot \textit{put}(b,p_1)\label{eq:ex_rec2}
+\end{align}
+When the execution contains no $\textit{rm}(\textit{empty})$-operation, the procedure selects a $\textit{put}$ operation adding a value that is not removed and that has a maximal priority.
+For (\ref{eq:ex_rec2}), it selects $\textit{put}(d,p_2)$ because $p_2$ is a maximal priority.
+This operation is correct since $d$ is the last value with priority $p_2$ in the execution, and the correctness of (\ref{eq:ex_rec2}) reduces to the correctness of
+\begin{align}
+\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(f,p_3) \cdot \textit{rm}(f) \cdot \textit{put}(b,p_1)\label{eq:ex_rec3}
+\end{align}
+If no operations like above can be found, $\textit{Check-PQ-Seq}$ selects a pair of $\textit{put}$ and $\textit{rm}$ operations adding and removing the same maximal priority value.
+For (\ref{eq:ex_rec2}), it can select
+%Since there is no other value of maximal priority which is not removed, the procedure selects a pair of $\textit{put}$/$\textit{rm}$ operations with an argument of maximal priority $p_2$, for instance,
+$\textit{put}(c,p_2)$ and $\textit{rm}(c)$.
+The value returned by $\textit{rm}(c)$ is correct if all the values of priority smaller than $p_2$ added before $\textit{rm}(c)$ are also removed before $\textit{rm}(c)$.
 In this case, $a$ is the only value of priority smaller than $p_2$ and it satisfies this property.
@@
 \begin{algorithm}[t]
+\footnotesize{
 \KwIn {A data-differentiated sequential execution $e$}
@@
     \Else {\KwRet $\mathsf{false}$\;}
-}
+}}
 \caption{$\textit{Check-PQ-Seq}$}
@@
 
-In formal terms, the operations which are selected depend on the following set of predicates on executions:
-{\small
+Formally, the selected operations depend on the following set of predicates on executions:
+%{\small
 \begin{align*}
-\mathsf{Has\text{-}EmptyRemoves}(e)=\mathsf{true} & \mbox{ iff  $e$ contains a $\textit{rm}(\textit{empty})$ operation} \\
-\mathsf{Has\text{-}UnmatchedMaxPriority}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{unmatched-priorities}(e)$ for a maximal priority} \\
-&\hspace{4mm}\mbox{$p\in priorities(e)$}
-\end{align*}}
-{\color{blue} where $\textit{priorities}(l)$, resp., $\textit{unmatched-priorities}(l)$}{\color{red} where $\textit{priorities}(e)$, resp., $\textit{unmatched-priorities}(e)$}, is the set of priorities occurring in $\textit{put}$ operations of $e$, resp., in $\textit{put}$ operations of $e$ for which there is no $\textit{rm}$ operation removing the same value.
+& \mathsf{Has\text{-}EmptyRemoves}(e)=\mathsf{true} \mbox{ iff  $e$ contains a $\textit{rm}(\textit{empty})$-operation} \hspace{1cm}\\
+& \mathsf{Has\text{-}UnmatchedMaxPriority}(e)=\mathsf{true} \mbox{ iff $p\in \textit{unmatched-priorities}(e)$ for a maximal $p$} 
+%&\hspace{4mm}\mbox{$p\in priorities(e)$}
+\end{align*}%}
+where $\textit{priorities}(e)$, resp., $\textit{unmatched-priorities}(e)$, is the set of priorities occurring in $\textit{put}$ operations of $e$, resp., in $\textit{put}$ operations of $e$ for which there is no $\textit{rm}$ operation removing the same value.
 We call the latter \emph{unmatched} put operations.
@@
 For simplicity, we consider the following syntactic sugar $\mathsf{Has\text{-}MatchedMaxPriority}(e)=\neg \mathsf{Has\text{-}EmptyRemoves}(e)\land \neg \mathsf{Has\text{-}UnmatchedMaxPriority}(e)$.
-By an abuse of notation, we also assume that  $\mathsf{Has\text{-}UnmatchedMaxPriority}(e) \Rightarrow \neg \mathsf{Has\text{-}EmptyRemoves}(e)$ (this is sound by the order of the conditionals in $\textit{Check-PQ-Seq}$).
+By an abuse of notation, we assume  $\mathsf{Has\text{-}UnmatchedMaxPriority}(e) \Rightarrow \neg \mathsf{Has\text{-}EmptyRemoves}(e)$ (this is sound by the order of the conditionals in $\textit{Check-PQ-Seq}$).
 
 The predicates defining the correctness of the selected operations are defined as follows:
-{\small
+%{\small
 \begin{align*}
-\mathsf{EmptyRemove\text{-}Seq}(e,o)=\mathsf{true} & \mbox{ iff  $e= u\cdot o\cdot v$ and $\textit{matched}(u)$} \\
-\mathsf{UnmatchedMaxPriority\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v$, $p\not\prec \textit{priorities}(u\cdot v)$, and $p\not\in \textit{priorities}(v)$} \\
-\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v\cdot \textit{rm}(x)\cdot w$, $p\not\prec \textit{priorities}(u\cdot v\cdot w)$,} \\
-&\hspace{4mm}\mbox{$p\not\preceq \textit{unmatched-priorities}(u\cdot v\cdot w)$, $\textit{matched}_\prec(u\cdot v,p)$,} \\
-&\hspace{4mm}\mbox{and $p\not\in \textit{priorities}(v\cdot w)$}
+&\mathsf{EmptyRemove\text{-}Seq}(e,o)=\mathsf{true} \mbox{ iff  $e= u\cdot o\cdot v$ and $\textit{matched}(u)$} \\
+&\mathsf{UnmatchedMaxPriority\text{-}Seq}(e,x)=\mathsf{true}  \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v$, $p\not\prec \textit{priorities}(u\cdot v)$,} \\
+&\hspace{6.4cm}p\not\in \textit{priorities}(v) \\
+&\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)=\mathsf{true} \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v\cdot \textit{rm}(x)\cdot w$, $\textit{matched}_\prec(u\cdot v,p)$,} \\
+&\hspace{4.3cm}\mbox{$p\not\preceq \textit{unmatched-priorities}(u\cdot v\cdot w)$, $p\not\prec \textit{priorities}(u\cdot v\cdot w)$,} \\
+&\hspace{4.3cm}\mbox{and  $p\not\in \textit{priorities}(v\cdot w)$}
 \end{align*}
-}
-
-\vspace{-3mm}
-\noindent
+%}
 where $p\prec \textit{priorities}(e)$ when $p\prec p'$ for some $p'\in \textit{priorities}(e)$ (and similarly for $p\prec \textit{unmatched-priorities}(e)$ or $p\preceq \textit{unmatched-priorities}(e)$),
@@
 
-When $o$ is a $\textit{rm}(\textit{empty})$ operation, we write $e\setminus o$ for the maximal subsequence of $e$ which doesn't contain $o$.
-For an execution $e$, $\textit{values}(e)$ is the set of values occurring in call/return actions of $e$.
+When $o$ is an $\textit{rm}(\textit{empty})$-operation, $e\setminus o$ is the maximal subsequence of $e$ which doesn't contain $o$.
+For an execution $e$, $\textit{values}(e)$ is the set of values  in call/return actions of $e$.
 
@@
 \begin{align*}
-\Gamma\mathsf{\text{-}Conc}(e,\alpha) = \mathsf{true}\mbox{ iff there exists a sequential execution $s$ such that $e\sqsubseteq s$ and $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$}
+\Gamma\mathsf{\text{-}Conc}(e,\alpha) = \mathsf{true}\mbox{ iff there is a sequential execution $s$ such that $e\sqsubseteq s$ and $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$}
 \end{align*}
-for each $\Gamma\in \{\mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$.
-Let $\textit{Check-PQ-Conc}$ denote the thus obtained procedure (we assume recursive calls are modified accordingly).
+for each $\Gamma\in \{\mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$. The obtained procedure is denoted by 
+$\textit{Check-PQ-Conc}$ (recursive calls are modified accordingly).
 
@@
 Completeness follows easily from the properties of $\seqPQ$.
-Thus, if $\textit{Check-PQ-Conc}(e) = \textit{false}$, then there exists a set $D$ of values s.t.
+If $\textit{Check-PQ-Conc}(e) = \textit{false}$, then there exists a set $D$ of values s.t.
 either  $\mathsf{EmptyRemove\text{-}Conc}(e \vert D)$ is false, or $\mathsf{UnmatchedMaxPriority\text{-}Conc}(e \vert D,x)$ is false for all the values $x$ of maximal priority that are not removed (and there exists at least one such value), or $\mathsf{MatchedMaxPriority\text{-}Conc}(e \vert D,x)$ is false for all the values $x$ of maximal priority (and these values are all removed in $e \vert D$).
@@
 \label{lemma:con-check-EPQ is correct}
-$\textit{Check-PQ-Conc}(e)=\mathsf{true}$ iff $e \sqsubseteq \seqPQ$, for every data-differentiated execution $e$.
+$\textit{Check-PQ-Conc}(e)=\mathsf{true}$ iff $e \sqsubseteq \seqPQ$, for every data-differentiated $e$.
 \end{restatable}
@@
   \centering
-  \includegraphics[width=1 \textwidth]{figures/PIC-HIS-EPQ1-TwoHis-2.pdf}
+  \includegraphics[width=1 \textwidth]{PIC-HIS-EPQ1-TwoHis-2.pdf}
 %\vspace{-10pt}

### 2017-07-10 final/introduction.tex
half of the paper

--- 2017-04-21 introduction.tex
@@
 
-Modern computer software is increasingly concurrent. Interactive applications
-and services necessitate reactive asynchronous operations to handle requests
-immediately as they happen, rather than waiting for long-running operations to
-complete. Furthermore, as processor manufacturers approach clock-speed limits,
-performance improvements are more-often achieved by parallelizing operations
-across multiple processor cores.
+%Modern computer software is increasingly concurrent. Interactive applications
+%and services necessitate reactive asynchronous operations to handle requests
+%immediately as they happen, rather than waiting for long-running operations to
+%complete. Furthermore, as processor manufacturers approach clock-speed limits,
+%performance improvements are more-often achieved by parallelizing operations
+%across multiple processor cores.
 
@@
   despite being optimized to avoid blocking and exploit parallelism, e.g.,~by
-  using atomic machine instructions like compare-and-swap.
+  using machine instructions like compare-and-swap.
 Intuitively, linearizability asks that every individual operation appears to take place instantaneously at some point between its invocation and its return.
@@
 % (even for finite-state implementations).
-However, recent work~\cite{DBLP:conf/icalp/BouajjaniEEH15} has shown that for particular classes of objects, 
-i.e., registers, mutexes, queues, and stacks, the problem of verifying linearizability becomes decidable (for finite-state implementations).
+However, recent work~\cite{DBLP:conf/icalp/BouajjaniEEH15} has shown that for particular objects,
+e.g., registers, mutexes, queues, and stacks, the problem of verifying linearizability becomes decidable (for finite-state implementations).
 
@@
 Priority queues are collections providing $\textit{put}$ and $\textit{rm}$ methods for adding and removing values.
-Every value added is associated to a priority and a remove operation returns a minimal priority value.
+Every added value is associated to a priority and a remove operation returns a minimal priority value.
 For generality, we consider a partially-ordered set of priorities.
@@
 
-Compared to previously studied collections like stacks and queues, the main challenge in dealing with priority queues is that the order in which values are removed is not fixed by the happens-before between add/remove operations (e.g., in the case of a queue, the values are removed in the order in which they were inserted), but by parameters of the $\textit{put}$ operations (the priorities) which come from an unbounded domain.
+Compared to previously studied collections like stacks and queues, the main challenge in dealing with priority queues is that the order in which values are removed is not fixed by the happens-before between add/remove operations (e.g., in the case of queues, values are removed in the order in which they were inserted), but by parameters of the $\textit{put}$ operations (the priorities) which come from an unbounded domain.
 For instance, the sequential behavior $\textit{put}(a,p_1)\cdot \textit{put}(b,p_3)\cdot \textit{put}(c,p_2)\cdot \textit{rm}(a,p_1)\cdot \textit{rm}(c,p_2)$ where the priority $p_1$ is less than $p_2$ which is less than $p_3$, is not admitted neither by the regular queue nor the stack.
@@
 
-Following the approach in~\cite{DBLP:conf/icalp/BouajjaniEEH15}, we give a characterization of concurrent priority queue behaviors violating linearizability in terms of automata.
+We give a characterization of concurrent priority queue behaviors violating linearizability in terms of automata.
 This characterization enables a reduction of checking linearizability for arbitrary implementations to reachability or invariant checking, and implies decidability for checking linearizability of finite-state implementations.
-However, differently from the case of stacks and queues where finite-state automata are sufficient, we show that the case of priority queues needs \emph{register automata} where registers are used to store and compare priorities.
+While linearizability violations for stacks and queues can be described using finite-state automata~\cite{DBLP:conf/icalp/BouajjaniEEH15}, the case of priority queues requires \emph{register automata} where registers are used to store and compare priorities.
 
 This characterization is obtained in several steps.
-We first define a recursive procedure which recognizes valid sequential executions which is then extended to recognize linearizable concurrent executions.
-Intuitively, for an input execution $e$, this procedure handles values occurring in $e$ one by one, starting with values of maximal priority (which are to be removed the latest).
+We define a recursive procedure that recognizes valid sequential executions, which is then extended to recognize linearizable concurrent executions.
+Intuitively, for an execution $e$, this procedure deals with values occurring in $e$ one by one, starting with values of maximal priority (to be removed the latest).
 For each value $x$, it checks whether $e$ satisfies some property ``local'' to that value, i.e., which is agnostic to how the operations adding or removing other values are ordered between them (w.r.t.
@@
 the operations on $x$.
-When this property holds, the same is done for the rest of the execution, without the operations on $x$.
+When this property holds, the procedure is applied recursively on the rest of the execution, without the operations on $x$.
 This procedure works only for executions where a value is added at most once, but this is not a limitation for \emph{data-independent} implementations whose behavior doesn't depend on the values that are added or removed.
@@
 The remainder of this article is organized as follows.
-Section~\ref{sec:priority queue and data-independence} describes the priority queue ADT, lists several semantic properties like data-independence that are satisfied by implementations of this ADT, and formalizes the notion of linearizability.
-Sections~\ref{sec:checking inclusion by recursive procedure} defines a recursive procedure for checking linearizability of concurrent priority queue behaviors.
+Section~\ref{sec:priority queue and data-independence} describes the priority queue ADT, lists several semantic properties like data-independence, and recalls the notion of linearizability.
+Section~\ref{sec:checking inclusion by recursive procedure} defines a recursive procedure for checking linearizability of concurrent priority queue behaviors.
 Section~\ref{sec:co-regular of extended priority queues} gives an automata characterization of the violations to linearizability, and
-Section~\ref{sec:related} discusses related work.
+Section~\ref{sec:related} discusses related work. Detailed proofs and constructions can be found in the extended version~\cite{CONCUR2017Ahmed}. 
 %, and

### 2017-07-10 final/priority-queues.tex
half of the paper

--- 2017-07-01 priority-queues.tex
@@
 We consider priority queues whose interface contains two methods $\textit{put}$ and $\textit{rm}$ for adding and respectively, removing a value.
-Each value is assigned with a priority when being added to the data structure (using a call to  $\textit{put}$) and the remove method $\textit{rm}$ removes a value with a minimal priority.
+Each value is assigned with a priority when being added to the data structure (by calling $\textit{put}$) and the remove method $\textit{rm}$ removes a value with a minimal priority.
 For generality, we assume that the set of priorities is partially-ordered.
 Incomparable priorities can be removed in any order.
-In case multiple values are assigned with the same priority, $\textit{rm}$ returns the least recent value (according to a FIFO semantics).
+When multiple values are assigned with the same priority, $\textit{rm}$ returns the least recent value.
 Also, when the set of values stored in the priority queue is empty, $\textit{rm}$ returns the distinguished value $\textit{empty}$.
-Concurrent implementations of the priority queue allow the methods $\textit{put}$ and $\textit{rm}$ to be called concurrently from different threads.
+In this section, we formalize (concurrent) executions and implementations, introduce a set of properties satisfied by all the implementations we are aware of, and recall the standard correctness criterion for concurrent implementations of ADTs known as \emph{linearizability}~\cite{journals/toplas/HerlihyW90}.
+%
+%recall the notion of \emph{linearizability}, the standard correctness criterion for concurrent implementations of priority queues (
+%
+%Concurrent implementations of the priority queue allow the methods $\textit{put}$ and $\textit{rm}$ to be called concurrently from different threads.
 However, every method invocation should give the illusion that it takes place instantaneously at some point between its invocation and its return.
@@
 The return value of a remove is transformed to an argument value for uniformity~\footnote{Method return values are guessed nondeterministically, and validated at return points.
-This can be handled using the {\tt assume} statements of typical formal specification languages, which only admit executions satisfying a given predicate.}.
+This can be handled using {\tt assume} statements, which only admit executions satisfying a given predicate.}.
 The return actions are denoted in a similar way as $\textit{ret}_o(\textit{put},a,p)$ and respectively, $\textit{ret}_o(\textit{rm},a')$.
 
-An \emph{execution} $e$ is a sequence of call and return actions which satisfy the following well-formedness properties: each return is preceded by a matching call (i.e., having the same operation identifier), and each operation identifier is used in at most one call/return.
+An \emph{execution} $e$ is a sequence of call and return actions which satisfy the following well-formedness properties: each return is preceded by a matching call (having the same operation identifier), and each operation identifier is used in at most one call/return.
 We assume every set of executions is closed under isomorphic renaming of operation identifiers.
+An $m(a)$-operation in an execution $e$ is an operation identifier $o$ s.t.
+$e$ contains the actions $\textit{call}_o(m,a)$ and $\textit{ret}_o(m,a)$.
 An execution is called \emph{sequential} when no two operations overlap, i.e., each call action is immediately followed by its matching return action, and \emph{concurrent} otherwise.
-To ease the reading, we write a sequential execution as a sequence of $\textit{put}(a,p)$ and $\textit{rm}(a)$ symbols representing a pair of actions $\textit{call}_o(\textit{put},a,p)\cdot \textit{ret}_o(\textit{put},a,p)$ and $\textit{call}_o(\textit{rm},a)\cdot \textit{ret}_o(\textit{rm},a)$, respectively (where $o\in\mathbb{O}$).
+For readability, we write a sequential execution as a sequence of $\textit{put}(a,p)$ and $\textit{rm}(a)$ symbols representing a pair of actions $\textit{call}_o(\textit{put},a,p)\cdot \textit{ret}_o(\textit{put},a,p)$ and $\textit{call}_o(\textit{rm},a)\cdot \textit{ret}_o(\textit{rm},a)$, respectively ($o\in\mathbb{O}$).
 For example, given two priorities $p_1 \prec p_2$, $\textit{put}(a,p_2) \cdot \textit{put}(b,p_1) \cdot \textit{rm}(b)$ is a sequential execution of the priority queue ($\textit{rm}$ returns $b$ because it has smaller priority).
 
-We define $\seqPQ$, the set of sequential priority queue executions, semantically via labelled transition system ($LTS$).
-An LTS is a tuple $A=(Q,\Sigma,\rightarrow,q_0)$, where $Q$ is a set of states, $\Sigma$ is an alphabet of transition labels, $\rightarrow\subseteq Q\times\Sigma\times Q$ is a transition relation and $q_0$ is the initial state.
-We model priority queue as an LTS $\textit{PQ}$ defined as follows: {\color {blue}each state of $\textit{PQ}$}{\color {red}each state of the LTS $\textit{PQ}$} is a mapping associating priorities in $\mathbb{P}$ with sequences of values in $\mathbb{D}$ representing a snapshot of the priority queue (for each priority, the values are ordered as they were inserted); the transition labels are $\textit{put}(a,p)$ and $\textit{rm}(a)$ operations; Each transition modifies the state as expected.
+We define $\seqPQ$, the set of sequential priority queue executions, semantically via a labelled transition system (LTS, for short).
+An LTS is a tuple $A=(Q,\Sigma,\rightarrow,q_0)$, where $Q$ is a set of states, $\Sigma$ is an alphabet of transition labels, $\rightarrow\subseteq Q\times\Sigma\times Q$ is a transition relation, and $q_0$ is the initial state.
+We model the priority queue as an LTS $\textit{PQ}$ where states are mappings associating priorities in $\mathbb{P}$ with sequences of values in $\mathbb{D}$, representing a snapshot of the priority queue (for each priority, the values are ordered as they were inserted), and the transition labels are $\textit{put}(a,p)$ and $\textit{rm}(a)$.
+Each transition modifies the state as expected.
 For example, $q_1 \xrightarrow{\textit{rm}(\textit{empty})} q_2$ if $q_1 = q_2$, and $q_1$ and $q_2$ map each priority to the empty sequence $\epsilon$.
-Then, {\color{blue}$\seqPQ$ is the set of words accepted by $\textit{PQ}$}{\color {red}$\seqPQ$ is the set of traces of $\textit{PQ}$}.
+Then, $\seqPQ$ is the set of traces (words) accepted by $\textit{PQ}$.
+%}{\color {red}$\seqPQ$ is the set of traces of $\textit{PQ}$}.
 {\color {red}The detailed definition of $\textit{PQ}$ can be found in the long version \cite{CONCUR2017Ahmed}.}
@@
 Implementations represent libraries whose methods are called by external programs.
-In the remainder of this work, we consider only completed executions, where each call action has a corresponding return action.
-This simplification is sound when implementation methods can always make progress in isolation: formally, for any execution $e$ with pending operations, there exists an execution $e'$ obtained by extending $e$ only with the return actions of the pending operations of $e$.
-Intuitively this means that methods can always return without any help from outside threads, avoiding deadlock.
+In the remainder of this work, we consider only \emph{completed} executions, where each call action has a corresponding return action.
+This simplification is sound when the method invocations can always make progress in isolation.
+%formally, for any execution $e$ with pending operations, there exists an execution $e'$ obtained by extending $e$ only with the return actions of the pending operations of $e$. 
+%Intuitively this means that methods can always return without any help from outside threads, avoiding deadlock.
 
@@
 
-We define two properties which are satisfied by priority queue implementations and which are important for the results that follow: (1) \emph{data independence}~\cite{conf/popl/Wolper86,conf/tacas/AbdullaHHJR13} states that priority queue behaviors do not depend on the actual values which are added to the queue, and (2) \emph{closure under projection}~\cite{DBLP:conf/icalp/BouajjaniEEH15} states that intuitively, remove operations can return the same values no matter how many other different values are in the queue, assuming they don't have more important priorities.
+We define two properties which are %satisfied by priority queue implementations and which are 
+important for our results: (1) \emph{data independence}~\cite{conf/popl/Wolper86,conf/tacas/AbdullaHHJR13} states that priority queue behaviors do not depend on the actual values which are added to the queue, and (2) \emph{closure under projection}~\cite{DBLP:conf/icalp/BouajjaniEEH15} states that executions remain valid by removing all the operations adding or removing certain values.
+%remove operations can return the same values even if certain values are projected out from the priority queue state.
+%no matter how many other different values are in the queue, provided that they don't have more important priorities.
 
@@
 Given an execution $e$, we denote by $r(e)$ the execution obtained from $e$ by replacing every data value $x$ by $r(x)$.
-Note that $r$ renames only the values and keep the priorities unchanged.
+Note that $r$ renames only the values and keeps the priorities unchanged.
 Intuitively, renaming values has no influence on the behavior of the priority queue, contrary to renaming priorities.
@@
 
-Beyond the sequential executions, every (concurrent) implementation of the priority queue that we are aware of is data-independent.
-Therefore, from now on, we consider only data-independent implementations.
-This assumption enables a reduction from checking the correctness of an implementation $\mathcal{I}$ to checking the correctness of only its data-differentiated executions in $\mathcal{I}_{\neq}$.
+Beyond sequential executions, every concurrent priority queue implementation that we are aware of is data-independent.
+From now on, we consider only data-independent implementations.
+This assumption enables a reduction from checking the correctness of an implementation $\mathcal{I}$ to checking the correctness of its data-differentiated executions in $\mathcal{I}_{\neq}$.
 
-Besides data independence, the sequential behaviors of the priority queue satisfy the following closure property: a behavior remains valid when removing all the operations with an argument in some set of values $D \subseteq \mathbb{D}$ and any $\textit{rm}(\textit{empty})$ operation (since they are read-only and they don't affect the queue's state).
-In order to distinguish between different $\textit{rm}(\textit{empty})$ operations while simplifying the technical exposition, we assume that they receive as argument a value, i.e., call actions are of the form $\textit{call}_o(\textit{rm},\textit{empty},a)$ for some $a\in \mathbb{D}$.
+Besides data independence, the sequential executions of the priority queue satisfy the following closure property: an execution remains valid when removing all the operations with an argument in some set of values $D \subseteq \mathbb{D}$ and any $\textit{rm}(\textit{empty})$ operation (since they are read-only and they don't affect the queue's state).
+To distinguish between different $\textit{rm}(\textit{empty})$ operations while simplifying the technical exposition, we assume that they receive as argument a value, i.e., call actions are of the form $\textit{call}_o(\textit{rm},\textit{empty},a)$ for some $a\in \mathbb{D}$.
 We will make explicit this argument only when needed in our technical development.
-The projection $e \vert D$ of an execution $e$ to a set of values $D \subseteq \mathbb{D}$ is obtained from $e$ by erasing all call/return actions with an argument not in $D$.
+The projection $e \vert D$ of an execution $e$ to a set of values $D \subseteq \mathbb{D}$ is obtained from $e$ by erasing all the call/return actions with an argument not in $D$.
 We write $e \setminus x$ for the projection $e \vert_{ \mathbb{D} \setminus \{ x \} }$.
-Let $\textit{proj}(e)$ be the set of all projections of $e$ to a set of values.
-{\color {red}The following lemma states that $\seqPQ$ is closed under projection.}
+Let $\textit{proj}(e)$ be the set of all projections of $e$ to a set of values $D \subseteq \mathbb{D}$.
+%{\color {red}The following lemma states that $\seqPQ$ is closed under projection.}
 
@@
 
-Given a (concurrent) execution $e$ and a sequential execution $s$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if there is a bijection $f: O_1 \rightarrow O_2$, where $O_1$ and $O_2$ are the set of operations of $e$ and $s$, respectively, such that (1) {\color{blue}$o$ and $f(o)$ have the same call and return actions}{\color {red}$o$ and $f(o)$ is the same operation}\footnote{An $m(a)$-operation in an execution $e$ is an operation identifier $o$ s.t.
-$e$ contains the actions $\textit{call}_o(m,a)$, $\textit{ret}_o(m,a)$.}, and (2) if $o_1 <_{\textit{hb}} o_2$, then $f(o_1) <_{\textit{hb}} f(o_2)$.
-A (concurrent) execution $e$ is linearizable w.r.t a set $S$ of sequential executions, denoted $e \sqsubseteq S$, if there exists $s \in S$ such that $e \sqsubseteq s$.
-A set of concurrent executions $E$ is linearizable w.r.t $S$, denoted $E \sqsubseteq S$, if $e \sqsubseteq S$ for all $e \in E$.
+Given a (concurrent) execution $e$ and a sequential execution $s$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if there is a bijection $f: O_1 \rightarrow O_2$, where $O_1$ and $O_2$ are the set of operations of $e$ and $s$, respectively, such that (1) the call and return actions with identifier $o$ and $f(o)$, respectively, are the same
+%{\color{blue}$o$ and $f(o)$ have the same call and return actions}{\color {red}$o$ and $f(o)$ is the same operation}\footnote{}, 
+and (2) if $o_1 <_{\textit{hb}} o_2$, then $f(o_1) <_{\textit{hb}} f(o_2)$.
+A (concurrent) execution $e$ is linearizable w.r.t.
+a set $S$ of sequential executions, denoted $e \sqsubseteq S$, if there exists $s \in S$ such that $e \sqsubseteq s$.
+A set of concurrent executions $E$ is linearizable w.r.t.
+$S$, denoted $E \sqsubseteq S$, if $e \sqsubseteq S$ for all $e \in E$.
 
@@
 %(see Appendix \ref{sec:appendix proofs in section priority queue and data-independence}).
-This is similar to that in \cite{conf/tacas/AbdullaHHJR13,DBLP:conf/icalp/BouajjaniEEH15}, where they use the notion of data-independence in \cite{conf/popl/Wolper86}.
+%This is similar to that in \cite{conf/tacas/AbdullaHHJR13,DBLP:conf/icalp/BouajjaniEEH15}, where they use the notion of data-independence in \cite{conf/popl/Wolper86}.
 Section~\ref{sec:checking inclusion by recursive procedure} will focus on characterizing linearizability for data-differentiated executions.
@@
 \label{lemma:data differentiated is enough for PQ}
-A data-independent implementation $\mathcal{I}$ is linearizable w.r.t a data-independent set $S$ of sequential executions, if and only if $\mathcal{I}_{\neq}$ is linearizable w.r.t. $S_{\neq}$.
+A data-independent implementation $\mathcal{I}$ is linearizable w.r.t. a data-independent set $S$ of sequential executions, if and only if $\mathcal{I}_{\neq}$ is linearizable w.r.t. $S_{\neq}$.
 \end{restatable}

### 2017-07-10 final/checking-executions.tex
half of the paper

@@
 Assume that there exist two sequential executions $l$ and $l'$ such that $e \sqsubseteq l=u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds, and $e \setminus x \sqsubseteq l' \in \seqPQ$.
-{\color {red}Let $u=\epsilon$, $w$ be any sequence containing the set of operations $\textit{put}(z_2,p_2)$ and $\textit{rm}(z_1)$} (we distinguish them by adding the suffix $-w$ to their name, e.g., $\textit{rm}(z_1)-w$), and $v$ be any sequence containing the remaining operations.
+Let $u=\epsilon$, $w$ be any sequence formed of $\textit{put}(z_2,p_2)$ and $\textit{rm}(z_1)$ (we distinguish them by adding the suffix $-w$ to their name, e.g., $\textit{rm}(z_1)-w$), and $v$ be any sequence containing the remaining operations.
 In general, the linearization $l'$ can be defined by choosing for each operation, a point in time between its call and return actions, called \emph{linearization point}.

### 2017-07-10 final/co-regular-EPQ.tex
modify

--- 2017-07-01 co-regular-EPQ.tex
@@
   \centering
-  \includegraphics[width=0.4 \textwidth]{figures/PIC-HIS-INTRO-GAP-EPQ1L.pdf}
+  \includegraphics[width=0.4 \textwidth]{PIC-HIS-INTRO-GAP-EPQ1L.pdf}
 %\vspace{-10pt}
@@
 The execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} is not $\mathsf{MatchedMaxPriority}^>$-linearizable because the left-right constraint of the maximal priority value $b$ contains a cycle: $f \rightarrow e \rightarrow d \rightarrow b \rightarrow f$.
-The presence of such a cycle is equivalent to the execution not being $\mathsf{MatchedMaxPriority}^>$-linearizable (see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}), {\color {red}as indicated by the following lemma:
+The presence of such a cycle is equivalent to the execution not being $\mathsf{MatchedMaxPriority}^>$-linearizable, % (see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}),
+{\color {red}as indicated by the following lemma:
 
@@
 To define a $\mathsf{MatchedMaxPriority}^>$-complete automaton we need to consider several cases depending on the order between the call/return actions of the $\textit{put}$/$\textit{rm}$ operations that add and respectively, remove the value $b$.
-For example, the case where the put happens-before the remove (as in \figurename~\ref{fig:introduce gap for EPQ1Lar}) is pictured in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
+For example, {\color {red}the automata for the case} where the put happens-before the remove (as in \figurename~\ref{fig:introduce gap for EPQ1Lar}) is pictured in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
 {\color {red}This automaton captures the three possible ways of ordering the first action $\textit{ret}(\textit{put},a,\_)$ w.r.t.
 the actions with value $b$, which are pictured in \figurename~\ref{fig:executions APQ1Lar-1 in paper} (a)} (this action cannot occur after $\textit{call}(\textit{rm},b,\_)$ since $b$ must be covered by the $a$-s).
-The paths corresponding to these three possible orders are: $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_{10}$ and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$.
+The paths corresponding to these three possible orders are: $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_{10} \ldots \rightarrow q_7$ and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$.
 {\color {red}In \figurename~\ref{fig:executions APQ1Lar-1 in paper}, we show all the four orders of the call/return actions of adding and removing $b$, and also possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$.} {\color {red}In the long version \cite{CONCUR2017Ahmed}, three register automata is constructed according to the cases of \figurename~\ref{fig:executions APQ1Lar-1 in paper} (b), (c) and (d), respectively.}
@@
   \centering
-  \includegraphics[width=1 \textwidth]{figures/PIC_AUTO_PQ1Lar-pprr.pdf}
+  \includegraphics[width=1 \textwidth]{PIC_AUTO_PQ1Lar-pprr.pdf}
 %\vspace{-10pt}
   \caption{Register automaton $\mathcal{A}_{\textit{l-lar}}^1$.
-We use the following notations: $C_1 = C \cup \{ \textit{ret}(\textit{rm},a) \}$, $C_2 = C \cup \{ \textit{call}(\textit{put},a,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.}
+We use the following notations: $A_1 = A \cup \{ \textit{ret}(\textit{rm},a) \}$, $A_2 = A \cup \{ \textit{call}(\textit{put},a,=r) \}$, $A_3 = A_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $A = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.}
   \label{fig:automata APQ1Lar-1 in paper}
@@
   %\includegraphics[width=1 \textwidth]{figures/PIC_HIS_PQ1Lar-pprr-in-paper.pdf}
-  \includegraphics[width=1 \textwidth]{figures/PIC_HIS_PQ1Lar-fouCase.pdf}
+  \includegraphics[width=1 \textwidth]{PIC_HIS_PQ1Lar-fouCase.pdf}
 %\vspace{-10pt}
@@
   \centering
-  \includegraphics[width=0.6 \textwidth]{figures/PIC-HIS-INTRO-PB-ORDER-EPQ.pdf}
+  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-PB-ORDER-EPQ.pdf}
 %\vspace{-10pt}
@@
   \centering
-  \includegraphics[width=0.8 \textwidth]{figures/PIC-HIS-FiveEnumerations.pdf}
+  \includegraphics[width=0.8 \textwidth]{PIC-HIS-FiveEnumerations.pdf}
 %\vspace{-10pt}
@@
   %\includegraphics[width=0.8 \textwidth]{figures/PIC-Enumeration-WitnessAutomata.pdf}
-  \includegraphics[width=0.8 \textwidth]{figures/PIC-WitnessAutomata-For1.pdf}
+  \includegraphics[width=0.8 \textwidth]{PIC-WitnessAutomata-For1.pdf}
 %\vspace{-10pt}

### 2017-07-11 co-regular-EPQ-arxiv.tex
modify

@@
   \caption{Register automaton $\mathcal{A}_{\textit{l-lar}}^1$.
-We use the following notations: $C_1 = C \cup \{ \textit{ret}(\textit{rm},a) \}$, $C_2 = C \cup \{ \textit{call}(\textit{put},a,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.}
+We use the following notations: {\color {red}$A_1 = A \cup \{ \textit{ret}(\textit{rm},a) \}$, $A_2 = A \cup \{ \textit{call}(\textit{put},a,=r) \}$, $A_3 = A_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $A = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},\top)$, $\textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$}.}
   \label{fig:automata APQ1Lar-1 in paper}
@@
   \caption{A case in deriving a $\mathsf{MatchedMaxPriority}^{=}$-complete automaton.
-We use the following notations: $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true})$, $\textit{call}(\textit{rm},\top), \textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_1 = C \cup \{ \textit{call}(\textit{put},d,<r) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},d) \}$, $C_4 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},d) \}$.}
+We use the following notations: {\color {red}$A = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true})$, $\textit{call}(\textit{rm},\top), \textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $A_1 = A \cup \{ \textit{call}(\textit{put},d,<r) \}$, $A_2 = A_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $A_3 = A_2 \cup \{ \textit{ret}(\textit{rm},d) \}$, $A_4 = A \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},d) \}$.}}
   \label{fig:an enumeration and its witness automaton}

### 2017-07-11 final/checking-executions.tex
dsad

@@
 
-\begin{restatable}{lemma}{EPQRulesAndSemantics}
+\begin{lemma}
 \label{lemma:EPQ rules and semantics}
 $\textit{Check-PQ-Seq}(e)=\mathsf{true}$ iff $e\in \seqPQ$, for every data-differentiated sequential execution $e$.
-\end{restatable}
+\end{lemma}
 
@@
 
-\begin{restatable}{lemma}{ConCheckEPQIsCorrect}
-\label{lemma:con-check-EPQ is correct}
+\begin{lemma}\label{lemma:con-check-EPQ is correct}
 $\textit{Check-PQ-Conc}(e)=\mathsf{true}$ iff $e \sqsubseteq \seqPQ$, for every data-differentiated $e$.
-\end{restatable}
+\end{lemma}
 
@@
 Assume that there exist two sequential executions $l$ and $l'$ such that $e \sqsubseteq l=u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds, and $e \setminus x \sqsubseteq l' \in \seqPQ$.
-Let $u=\epsilon$, $w$ be any sequence formed of $\textit{put}(z_2,p_2)$ and $\textit{rm}(z_1)$ (we distinguish them by adding the suffix $-w$ to their name, e.g., $\textit{rm}(z_1)-w$), and $v$ be any sequence containing the remaining operations.
-In general, the linearization $l'$ can be defined by choosing for each operation, a point in time between its call and return actions, called \emph{linearization point}.
+Let $u=\epsilon$, $w$ be any sequence formed of $\textit{put}(z_2,p_2)$ and $\textit{rm}(z_1)$ (we distinguish them by adding the suffix ``$-w$'' to their name, e.g., $\textit{rm}(z_1)-w$), and $v$ be any sequence containing the remaining operations.
+In general, the linearization $l'$ can be defined by choosing for each operation, a point in time between its call and return, called \emph{linearization point}.
 The order between the linearization points defines the sequence $l'$.
-\figurename~\ref{fig:concurrent execution for EPQ1}(a) draws linearization points for the operations in $e \setminus x$ defining the linearization $l'$~\footnote{In general, there may exist multiple ways of choosing linearization points to define the same linearization.
+\figurename~\ref{fig:concurrent execution for EPQ1}(a) draws linearization points for the operations in $e \setminus x$ which define $l'$~\footnote{In general, there may exist multiple ways of choosing linearization points to define the same linearization.
 Our construction is agnostic to this choice.}.
 %(the order of these linearization points defines the sequence $l'$).
-We show how to construct a sequence $l''= l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3\in\seqPQ$ such that $e \sqsubseteq l''$.
+We show how to construct a sequence $l''= l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3\in\seqPQ$ s.t. $e \sqsubseteq l''$.
 %In this example, .
@@
 \begin{itemize}
-\setlength{\itemsep}{0.5pt}
+%\setlength{\itemsep}{0.5pt}
 \item[-] An operation is called $p$-comparable (resp., $p$-incomparable) when it receives as argument a value of priority comparable to $p$ (resp., incomparable to $p$).
-We could try to define $l''_1$, $l''_2$ and $l''_3$ as the projection of $l'$ to the operations in $u$, $v$ and $w$, respectively.
-However, this is incorrect, since $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ imposes no restriction on  $p$-incomparable operations in $u \cdot v$, and thus, there is no guarantee that the projection of $l'$ to $p$-incomparable operations in $u \cdot v$ is correct.
-In this example, this projection is $\textit{put}(z_1,p_2) \cdot \textit{rm}(z_2)$ which is incorrect.
+Defining $l''_1$, $l''_2$, and $l''_3$ as the projection of $l'$ to the set of operations in $u$, $v$ and $w$, respectively, leads to a sequence $l''\not\in\seqPQ$.
+This is because $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ imposes no restriction on  $p$-incomparable operations in $u \cdot v$, and the projection of $l'$ to $p$-incomparable operations in $u \cdot v$ is not in $\seqPQ$.
+In this example, this projection is $\textit{put}(z_1,p_2) \cdot \textit{rm}(z_2)$.
 
 \item[-] We define the sets of operations $U'$, $V'$ and $W'$ such that $l''_1$, $l''_2$ and $l''_3$ are the projections of $l'$ to $U'$, $V'$, and $W'$, respectively. This is done in two steps:
-
-\item[-] The first step is to define $W'$.
+\begin{enumerate}
+\item The first step is to define $W'$.
 The $p$-comparable operations in $W'$ are the same as in $w$.
-To identify the $p$-incomparable operations in $W'$, we search for a $p$-incomparable operation $o$ which either happens before some $p$-comparable operation in $w$, {\color {red}or whose linearization point occurs after $\textit{ret}(\textit{rm},x)$}.
+To identify the $p$-incomparable operations in $W'$, we search for a $p$-incomparable operation $o$ which either happens before some $p$-comparable operation in $w$, or whose linearization point occurs after $\textit{ret}(\textit{rm},x)$.
 We add to $W'$ the operation $o$ and all the $p$-incomparable operations occurring after $o$ in $l'$.
@@
 In this process, whether a $p$-incomparable operation is in $W'$ or not only relies on whether it is before or after such an $o$ in $l'$.
-
-\item[-] The second step is to define $U'$ and $V'$.
+\item The second step is to define $U'$ and $V'$.
 $U'$ contains two kinds of operations: (1) operations whose linearization points are before $\textit{ret}(\textit{put},x,p)$, and (2) other $\textit{put}$ operations with priority $p$.
@@
 In this example, $U'$ contains $\textit{put}(z_1,p_2)$ and $\textit{put}(x_2,p)$.
-
-\item[-] In conclusion, we have that $l''_1 = \textit{put}(z_1,p_2) \cdot \textit{put}(x_2,p)$, $l''_2 = \textit{put}(z_2,p_2) \cdot \textit{rm}(x_2) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(y_1)$, {\color {red}and $l''_3 = \textit{rm}(z_1) \cdot \textit{rm}(z_2)$.} \figurename~\ref{fig:concurrent execution for EPQ1}(b) draws linearization points for each operation in $e$ defining the linearization $l''$.
+\end{enumerate}
+\item[-] In conclusion, we have that $l''_1 = \textit{put}(z_1,p_2) \cdot \textit{put}(x_2,p)$, $l''_2 = \textit{put}(z_2,p_2) \cdot \textit{rm}(x_2) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(y_1)$, and $l''_3 = \textit{rm}(z_1) \cdot \textit{rm}(z_2)$.
+\figurename~\ref{fig:concurrent execution for EPQ1}(b) draws linearization points for each operation in $e$ defining the linearization $l''$.
 \end{itemize}
@@
   \centering
-  \includegraphics[width=1 \textwidth]{PIC-HIS-EPQ1-TwoHis-2.pdf}
+  \includegraphics[width=.8\textwidth]{PIC-HIS-EPQ1-TwoHis-2.pdf}
 %\vspace{-10pt}
-  \caption{The process of obtaining linearization of $e$}
+  \caption{A concurrent execution $e$ exemplifying the soundness of $\textit{Check-PQ-Conc}$.}
   \label{fig:concurrent execution for EPQ1}
@@
 
-Section~\ref{sec:co-regular of extended priority queues} introduces a characterization of concurrent priority queue violations using a set of \emph{non-recursive} automata (i.e., whose states consist of a fixed number of registers) whose {\color{red}standard synchronized} product is equivalent to $\textit{Check-PQ-Conc}$ (modulo renaming of values which is possible by data-independence).
-Since $\seqPQ$ is closed under projection (Lemma~\ref{lem:closure_proj}), the recursion in $\textit{Check-PQ-Conc}$ can be eliminated by checking that each projection of a given execution $e$ passes a non-recursive version of $\textit{Check-PQ-Conc}$ where every recursive call is replaced by $\mathsf{true}$.
-More precisely, every occurrence of $\text{{\bf return}}\ \textit{Check-PQ-Conc}$ is replaced by  $\text{{\bf return}}\ \mathsf{true}$.
-Let $\textit{Check-PQ-Conc-NonRec}$ be the thus obtained procedure.
+Section~\ref{sec:co-regular of extended priority queues} introduces a characterization of concurrent priority queue violations using a set of \emph{non-recursive} automata (whose states consist of a fixed number of registers), whose standard synchronized product is equivalent to $\textit{Check-PQ-Conc}$ (modulo a renaming of values which is possible by data-independence).
+Since $\seqPQ$ is closed under projection (Lemma~\ref{lem:closure_proj}), the recursion in $\textit{Check-PQ-Conc}$ can be eliminated by checking that each projection of a given execution $e$ passes a non-recursive version of $\textit{Check-PQ-Conc}$ where every recursive call $\text{{\bf return}}\ \textit{Check-PQ-Conc}(\ldots)$ is
+%is replaced by $\mathsf{true}$. More precisely, every occurrence of $\text{{\bf return}}\ \textit{Check-PQ-Conc}$ is 
+replaced by  $\text{{\bf return}}\ \mathsf{true}$. Let $\textit{Check-PQ-Conc-NonRec}$ be the thus obtained procedure.
 
-\begin{restatable}{lemma}{EPQasMultiInMRpriforHistory}
+\begin{lemma}
 \label{lemma:EPQ as multi in MRpri for history}
 Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$ if and only if for each $e' \in \textit{proj}(e)$, $\textit{Check-PQ-Conc-NonRec}(e')$ returns $\mathsf{true}$.
-\end{restatable}
+\end{lemma}
 

### 2017-07-11 final/co-regular-EPQ.tex
dsad

--- 2017-07-01 co-regular-EPQ.tex
@@
 
-We show that the set of executions for which some projection fails the test $\textit{Check-PQ-Conc-NonRec}$ can be characterized using a set of register automata, modulo a value renaming.
-The possibility of renaming values (which is complete for checking data independent implementations) allows to simplify the reasoning about projections.
-Thus, we assume that all the operations which are not in the projection failing this test use the same distinguished value $\top$, different from those used in the projection.
-Then, it is enough to find an automata characterization for the set of executions $e$ for which $\textit{Check-PQ-Conc-NonRec}$ fails, or equivalently, for which one
+We show that the set of executions for which $\textit{Check-PQ-Conc-NonRec}$ fails on some projection can be described using register automata, modulo a value renaming.
+Renaming values (which is complete under data independence) allows to simplify the reasoning about projections.
+W.l.o.g., we assume that all the operations which are not in the projection failing this test use the same distinguished value $\top$, different from those in the projection.
+Then, it is enough to find an automata characterization of the executions $e$ for which $\textit{Check-PQ-Conc-NonRec}(e)$ is $\mathsf{false}$, i.e., for which
 %a projection $e'$ fails $\textit{Check-PQ-Conc-NonRec}$ when one
-of the following three formulas is false:
-\begin{align*}
-\hspace{-5mm}
-\Gamma(e) := \mathsf{Has\text{-}}\Gamma(e) \Rightarrow \exists \alpha.\ \Gamma\mathsf{\text{-}Conc}(e,\alpha)\mbox{ with $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$}
-\end{align*}
+%of the following three formulas is false:
+%\begin{align*}
+%\hspace{-5mm}
+$\Gamma(e) := \mathsf{Has\text{-}}\Gamma(e) \Rightarrow \exists \alpha.\ \Gamma\mathsf{\text{-}Conc}(e,\alpha)$ is false, for some $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$.
 Intuitively, $\Gamma(e)$ states that $e$ is linearizable w.r.t.
@@
 $\Gamma$}, or \emph{$\Gamma$-linearizable}.
-Extending the automaton characterizing executions which are not $\Gamma$-linearizable, with self-loops that allow any operation with parameter $\top$ results in an automaton satisfying the following property called \emph{$\Gamma$-completeness}.
+Extending the automaton characterizing non $\Gamma$-linearizable executions with self-loops that allow any operation with argument $\top$ results in an automaton satisfying the following property called \emph{$\Gamma$-completeness}.
 
@@
 
-$A \cap \mathcal{I} \neq \emptyset$ if and only if there exists $ e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$ such that $e'$ is not $\Gamma$-linearizable.
+$A \cap \mathcal{I} \neq \emptyset$ iff there exists $e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$ such that $e'$ is not $\Gamma$-linearizable.
 \end{definition}
 
-Section~\ref{ssec:aut} describes a $\mathsf{MatchedMaxPriority}$-complete automaton, {\color {red}the other automata can be found in the long version \cite{CONCUR2017Ahmed}.
+%Section~\ref{ssec:aut} describes a $\mathsf{MatchedMaxPriority}$-complete automaton, {\color {red}the other automata can be found in the long version \cite{CONCUR2017Ahmed}.
 Therefore, the following holds.}
@@
 
-\begin{lemma}
-There exists a $\Gamma$-complete automaton for each $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$.
-\end{lemma}
-
+We can show that for any $\Gamma\in \{\mathsf{EmptyRemove}$,$\mathsf{UnmatchedMaxPriority}$,$\mathsf{MatchedMaxPriority}\}$ there exists a $\Gamma$-complete automaton.
+For lack of space, we only consider the case $\Gamma=\mathsf{MatchedMaxPriority}$ in
+Section~\ref{ssec:aut}.
+%
+%\begin{lemma}
+%There exists a $\Gamma$-complete automaton for each $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$.
+%\end{lemma}
+%
 When defining $\Gamma$-complete automata, we assume that every implementation $\mathcal{I}$ behaves correctly, i.e., as a FIFO queue, when only values with the same priority are observed.
@@
 $\seqPQ$).
-{\color {blue}This property can be checked separately using the techniques introduced for regular FIFO queues in~\cite{DBLP:conf/icalp/BouajjaniEEH15}} {\color {red}This property can be checked separately using register automata (can be found in the long version \cite{CONCUR2017Ahmed}) obtained from the finite automata in~\cite{DBLP:conf/icalp/BouajjaniEEH15} for FIFO queue}.
+This property can be checked separately using register automata similar to the automata in~\cite{DBLP:conf/icalp/BouajjaniEEH15} describing FIFO queue violations.
 %(see Appendix~\ref{sec:appendix lemma and register automata for FIFO of single-priority executions} for more details).
-Note that this assumption excludes some obvious violations, such as a $\textit{rm}(a)$ operation happens before a $\textit{put}(a,p)$ operation, for some $p$.
+This assumption excludes some obvious violations, such as an $\textit{rm}(a)$ operation happening before a $\textit{put}(a,p)$ operation, for some $p$.
 
-Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions which contain only one maximal priority.
-This is possible because any data-differentiated execution for which $\Gamma(e)$ is false has such a projection.
+For $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, we consider $\Gamma$-complete automata recognizing executions which contain only one maximal priority.
+This is w.l.o.g.
+because any data-differentiated execution for which $\Gamma(e)$ is false has such a projection.
 %Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions where any two occurring priorities are comparable (w.r.t.
@@
 
-\begin{restatable}{lemma}{priExecutionIsEnough}
+\begin{lemma}
 \label{lemma:pri execution is enough}
-Let $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$ and $e$ a data-differentiated execution.
-Then, $e$ is $\Gamma$-linearizable iff $e\vert_{\preceq p}$ is $\Gamma$-linearizable for some maximal priority $p$ in $e$.
-\end{restatable}
+For $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, a data-differentiated execution $e$ 
+%and $e$ a . Then, $e$ 
+ is $\Gamma$-linearizable iff $e\vert_{\preceq p}$ is $\Gamma$-linearizable for some maximal priority $p$ in $e$.
+\end{lemma}
 \begin {proof} (Sketch)
 %{\color {blue}We prove an equivalent property: $e$ is $\Gamma$-linearizable if and only if $e\vert_{\preceq p}$ is $\Gamma$-linearizable for some maximal priority $p$ in $e$.} %We consider the case of $\mathsf{MatchedMaxPriority}$ and the case of $\mathsf{UnmatchedMaxPriority}$ is similar.
-To prove the $\textit{only if}$ direction, let $e$ be a data-differentiated execution linearizable w.r.t.
-$l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,x)$.
-Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,x)$ imposes no restriction on the operations in $u$, $v$ and $w$ with priorities incomparable to $p$, erasing all these operations results in a sequential execution which still satisfies this property.
+For the ``only-if'' direction, let $e$ be a data-differentiated execution linearizable w.r.t.
+$l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$ s.t.
+$\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l,x)$ holds.
+Since the predicate $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l,x)$ imposes no restriction on the operations in $u$, $v$, and $w$ with priorities incomparable to $p$, erasing all these operations results in a sequential execution which still satisfies this predicate.
 Similarly, for $\Gamma=\mathsf{UnmatchedMaxPriority}$.
 
-The $\textit{if}$ direction follows from the fact that if the projection of an execution to a set of operations $O_1$ has a linearization $l_1$ and the projection of the same execution to the remaining set of operations has a linearization $l_2$, then the execution has a linearization which is defined as an interleaving of $l_1$ and $l_2$.
+The ``if'' direction follows from the fact that if the projection of an execution to a set of operations $O_1$ has a linearization $l_1$ and the projection of the same execution to the remaining set of operations has a linearization $l_2$, then the execution has a linearization which is defined as an interleaving of $l_1$ and $l_2$.
 %(see Appendix~\ref{sec:appendix proof of Lemma pri execution is enough} for more details).
@@
 Thus, let $e$ be an execution such that $e\vert_{\preceq p}$ is linearizable w.r.t.
-$l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,x)$.
+$l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$ where $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l,x)$ holds.
 By the property above, we know that $e$ has a linearization $l' = u' \cdot \textit{put}(x,p) \cdot v' \cdot \textit{rm}(x) \cdot w'$, such that the projection of $l'$ to values of priority comparable to $p$ is $l$.
 %We can see that $u$, $v$ and $w$ are subsequences of $u'$, $v'$ an $w'$, respectively.
-Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,x)$ does not have a condition on values of priority incomparable to $p$, we obtain that $l' \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
+Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l,x)$ doesn't constrain the values of priority incomparable to $p$, we obtain that $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l',\alpha)$ also holds.
 \end {proof}
 
-The following shows that $\Gamma$-complete automata enable an effective reduction of checking linearizability of concurrent priority queue implementations to state reachability.
-It is a direct consequence of the above definitions.
+The existence of $\Gamma$-complete automata enable an effective reduction of checking linearizability of concurrent priority queue implementations to state reachability. 
+%It is a direct consequence of the above definitions. 
 Section~\ref{subsec:combine step-by-step linearizability and co-regular} discusses decidability results implied by this reduction.
 
-\begin{restatable}{theorem}{ReduceEPQIntoStateReachability}
+\begin{theorem}
 \label{lemma:reduce EPQ into state reachability}
-Let $\mathcal{I}$ be a data-independent implementation, and $A(\Gamma)$ be a $\Gamma$-complete automaton for each $\Gamma$. Then,
-$\mathcal{I} \sqsubseteq \seqPQ$ if and only if $\mathcal{I} \cap A(\Gamma) = \emptyset$ for all $\Gamma$.
-\end{restatable}
+Let $\mathcal{I}$ be a data-independent implementation. Then, there is a $\Gamma$-complete automaton $A(\Gamma)$ for each $\Gamma$. Moreover,
+$\mathcal{I} \sqsubseteq \seqPQ$ iff $\mathcal{I} \cap A(\Gamma) = \emptyset$ for all $\Gamma$.
+\end{theorem}
 
@@
 Instead of formalizing the full class of register automata, we consider a simpler class which suffices our needs.
-More precisely, we consider a class of labeled transition systems whose states consist of a finite control part and a register $r$ interpreted to elements of $\mathbb{P}$.
-The transition labels can be one of the following:
+Thus, we consider a class of labeled transition systems whose states consist of a finite control part and a register $r$ interpreted to elements of $\mathbb{P}$.
+The transition labels are:
 \begin{itemize}
@@
 	\item $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$ for reading call/return actions of a remove,
-	\item $\textit{call}(\textit{put},d,g)$ where $g\in\{=r,\prec r,true\}$ is a guard, for reading a call action $\textit{call}(\textit{put},d,p)$ of a put and checking whether $p$ is either equal to or smaller than the value stored in $r$, or arbitrary,
+	\item $\textit{call}(\textit{put},d,g)$ where $g\in\{=r,\prec r,true\}$ is a guard, for reading a call action $\textit{call}(\textit{put},d,p)$ and checking if $p$ is either equal to or smaller than the value stored in $r$, or arbitrary,
 	\item $\textit{ret}(\textit{put},d,true)$ for reading a return action $\textit{ret}(\textit{put},d,p)$ for any $p$.
 \end{itemize}
-The set of words accepted by such a transition system can be defined as usual.
+The set of sequences (executions) accepted by such a transition system is defined as usual.
 
@@
 \label{subsec:co-regular of EPQ1Lar}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.4 \textwidth]{PIC-HIS-INTRO-GAP-EPQ1L.pdf}
+%\vspace{-10pt}
+  \caption{An execution that is not $\mathsf{MatchedMaxPriority}^{>}$-linearizable.
+We represent each operation as a time interval whose left, resp., right, bound corresponds to the call, resp., return action.
+{\color {blue}Operations adding and removing the same value are aligned vertically.}}
+  \label{fig:introduce gap for EPQ1Lar}
+\end{figure}
 
@@
 %\end{align*}}
-We give a typical example of an execution $e$ which is not $\mathsf{MatchedMaxPriority}^>$-linearizable in
+Figure~\ref{fig:introduce gap for EPQ1Lar} contains a typical example of an execution $e$ which is not $\mathsf{MatchedMaxPriority}^>$-linearizable,
 %Given a data-differentiated execution $e$ such that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Condition}(e)$ holds, Lemma \ref{lemma:automata for extended priority queue with single priority} is not enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$.
@@
 One such example is shown in
-\figurename~\ref{fig:introduce gap for EPQ1Lar}.
+where $p_1 \prec p_4$, $p_2 \prec p_4$, and $p_3 \prec p_4$.
 Intuitively, this is a violation because during the whole execution of $\textit{rm}(b)$, the priority queue stores a smaller priority value (which should be removed before $b$).
 To be more precise, we define \emph{the interval of a value $x$} as the time interval from the return of a put $\textit{ret}(\textit{put},x,p)$ to the call of the matching remove $\textit{call}(rm,x)$, or to the end of the execution if such a call action doesn't exist.
-Intuitively, it represents the time interval in which a value is guaranteed to be stored into the concurrent priority queue.
+This represents the time interval in which a value is guaranteed to be stored into the priority queue.
 Concretely, for a standard indexing of actions in an execution, a time interval is a closed interval between the indexes of two actions in the execution.
-In \figurename~\ref{fig:introduce gap for EPQ1Lar}, we draw the interval of each value by dashed line.
-Here we assume that $p_1 \prec p_4$, $p_2 \prec p_4$, and $p_3 \prec p_4$.
-We can not find a sequence $s$ where $e \sqsubseteq s$ and $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,b)$ holds, since each time point from $\textit{call}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is included in the interval of some smaller priority value, and $\textit{rm}(b)$ can't take effect in the interval of a smaller priority value.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.4 \textwidth]{figures/PIC-HIS-INTRO-GAP-EPQ1L.pdf}
-%\vspace{-10pt}
-  \caption{An execution that is not $\mathsf{MatchedMaxPriority}^{>}$-linearizable.
-We represent each operation as a time interval whose left, resp., right, bound corresponds to the call, resp., return action.
-{\color {blue}Operations adding and removing the same value are aligned vertically.}}
-  \label{fig:introduce gap for EPQ1Lar}
-\end{figure}
-
-To formalize the scenario in \figurename~\ref{fig:introduce gap for EPQ1Lar} we use the notion of \emph{left-right constraint} defined below.
+In \figurename~\ref{fig:introduce gap for EPQ1Lar}, the interval of each value of priority smaller than $p_4$ is pictured as a dashed line.
+There is no sequence $l$ s.t.
+$e \sqsubseteq l$ and $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l,b)$ hold, since each time point from $\textit{call}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is included in the interval of a smaller priority value,
+and $\textit{rm}(b)$ can't take effect in the interval of a smaller priority value.
+To formalize this scenario we use the notion of \emph{left-right constraint} defined below.
 
@@
 
-The execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} is not $\mathsf{MatchedMaxPriority}^>$-linearizable because the left-right constraint of the maximal priority value $b$ contains a cycle: $f \rightarrow e \rightarrow d \rightarrow b \rightarrow f$.
-The presence of such a cycle is equivalent to the execution not being $\mathsf{MatchedMaxPriority}^>$-linearizable (see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}), {\color {red}as indicated by the following lemma:
+The execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} is not $\mathsf{MatchedMaxPriority}^>$-linearizable because the left-right constraint of the maximal priority value $b$ contains the cycle {\color{red} $f \rightarrow e \rightarrow d \rightarrow b \rightarrow f$}.
+The presence of such a cycle is equivalent to \emph{non} $\mathsf{MatchedMaxPriority}^>$-linearizability:
+%(see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}), {\color {red}as indicated by the following lemma:
 
@@
 \label{lemma:Lin Equals Constraint for EPQ1Lar}
-Given a data-differentiated execution $e$ where $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, let $p$ be its maximal priority and $\textit{put}(x,p),\textit{rm}(x)$ are only operations of priority $p$ in $e$.
-Let $G$ be the graph representing the left-right constraint of $x$.
-$e$ is $\mathsf{MatchedMaxPriority}$-linearizable, if and only if $G$ has no cycle going through $x$.
+Let $e$ be a data-differentiated execution such that 
+%Given a data-differentiated execution $e$ where 
+$\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, $p$ is the maximal priority in $e$, and $\textit{put}(x,p)$ and $\textit{rm}(x)$ are only operations with arguments of priority $p$ in $e$. 
+%Let $G$ be the graph representing the left-right constraint of $x$. 
+Then, $e$ is $\mathsf{MatchedMaxPriority}$-linearizable iff the left-right constraint of $x$ contains no cycle going through $x$.
 \end{restatable}
-}
+%}
 
@@
 We then prove that getting rid of cycle though value with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$.
-When the left-right constraint of the maximal priority value $x$ contains a cycle of the form $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ for some $d_1$,$\ldots$,$d_n\in \mathbb{D}$, we say that $x$ is \emph{covered} by $d_1,\ldots,d_m$.
-The shape of such a cycle (i.e., the alternation between call/return actions of $\textit{put}$/$\textit{rm}$ operations) can be detected using our class of automata, the only complication being the unbounded number of values $d_1$,$\ldots$,$d_n$.
+When the left-right constraint contains a cycle $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$, for some $d_1$,$\ldots$,$d_n\in \mathbb{D}$, we say that $x$ is \emph{covered} by $d_1,\ldots,d_m$.
+The shape of an execution witnessing such a cycle (i.e., the alternation between call/return actions) can be identified using our class of automata, the only complication being the unbounded number of values $d_1$,$\ldots$,$d_n$.
 However, by data independence, whenever an implementation contains such an execution it also contains an execution where all the values $d_1$,$\ldots$,$d_n$ are renamed to the same value $a$, and $x$ is renamed to $b$.
-Therefore, our automata can be defined over a fixed set of values $a$, $b$, and $\top$ (recall that $\top$ is used for the operations outside of the non-linearizable projection).
+Therefore, our automata can be defined over a fixed set of values $a$, $b$, and $\top$ (recall that $\top$ is used for operations outside of the non-linearizable projection).
 
@@
   \centering
-  \includegraphics[width=1 \textwidth]{figures/PIC_AUTO_PQ1Lar-pprr.pdf}
+  \includegraphics[width=1 \textwidth]{PIC_AUTO_PQ1Lar-pprr.pdf}
 %\vspace{-10pt}

### 2017-07-11 final/co-regular-EPQ.tex
besides decidability and related work

@@
   \caption{An execution that is not $\mathsf{MatchedMaxPriority}^{>}$-linearizable.
-We represent each operation as a time interval whose left, resp., right, bound corresponds to the call, resp., return action.
-{\color {blue}Operations adding and removing the same value are aligned vertically.}}
+We represent each operation as a time interval whose left, resp., right, bound corresponds to the call, resp., return action.} % {\color {blue}Operations adding and removing the same value are aligned vertically.}}
   \label{fig:introduce gap for EPQ1Lar}
@@
 
-The execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} is not $\mathsf{MatchedMaxPriority}^>$-linearizable because the left-right constraint of the maximal priority value $b$ contains the cycle {\color{red} $f \rightarrow e \rightarrow d \rightarrow b \rightarrow f$}.
+The execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} is not $\mathsf{MatchedMaxPriority}^>$-linearizable because the left-right constraint of the maximal priority value $b$ contains the cycle $f \rightarrow d \rightarrow c \rightarrow b \rightarrow f$.
 The presence of such a cycle is equivalent to \emph{non} $\mathsf{MatchedMaxPriority}^>$-linearizability:
@@
 
-\begin{restatable}{lemma}{LinEqualsConstraintforEPQOneLar}
+\begin{lemma}
 \label{lemma:Lin Equals Constraint for EPQ1Lar}
@@
 Then, $e$ is $\mathsf{MatchedMaxPriority}$-linearizable iff the left-right constraint of $x$ contains no cycle going through $x$.
-\end{restatable}
+\end{lemma}
 %}
@@
 For example, the case where the put happens-before the remove (as in \figurename~\ref{fig:introduce gap for EPQ1Lar}) is pictured in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
-{\color {red}This automaton captures the three possible ways of ordering the first action $\textit{ret}(\textit{put},a,\_)$ w.r.t.
-the actions with value $b$, which are pictured in \figurename~\ref{fig:executions APQ1Lar-1 in paper} (a)} (this action cannot occur after $\textit{call}(\textit{rm},b,\_)$ since $b$ must be covered by the $a$-s).
-The paths corresponding to these three possible orders are: $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_{10}$ and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$.
-{\color {red}In \figurename~\ref{fig:executions APQ1Lar-1 in paper}, we show all the four orders of the call/return actions of adding and removing $b$, and also possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$.} {\color {red}In the long version \cite{CONCUR2017Ahmed}, three register automata is constructed according to the cases of \figurename~\ref{fig:executions APQ1Lar-1 in paper} (b), (c) and (d), respectively.}
-%In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, three register automata is constructed according to the cases of \figurename~\ref{fig:executions APQ1Lar-1 in paper} (b), (c) and (d), respectively.}
-{\color {blue}Automata for the other cases can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.}
+This automaton captures the three possible ways of ordering the first action $\textit{ret}(\textit{put},a,\_)$ w.r.t.
+the actions with value $b$, which are pictured in \figurename~\ref{fig:executions APQ1Lar-1 in paper}(a) (this action cannot occur after $\textit{call}(\textit{rm},b,\_)$ since $b$ must be covered by the $a$-s).
+The paths corresponding to these three possible orders are: $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_{10}$, and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$.
+\figurename~\ref{fig:executions APQ1Lar-1 in paper} lists the four possible orderings of the call/return actions of adding and removing $b$, and also possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$.
+Each such ordering corresponds to an automaton similar to the one in \figurename~\ref{fig:automata APQ1Lar-1 in paper}, their union defining a $\mathsf{MatchedMaxPriority}^>$-complete automaton.
 
@@
 
-\begin{figure}[htbp]
+\begin{figure}[t]
   \centering
-  \includegraphics[width=1 \textwidth]{PIC_AUTO_PQ1Lar-pprr.pdf}
+  %\includegraphics[width=1 \textwidth]{figures/PIC_HIS_PQ1Lar-pprr-in-paper.pdf}
+  \includegraphics[width=.8\textwidth]{PIC_HIS_PQ1Lar-fouCase.pdf}
 %\vspace{-10pt}
-  \caption{Register automaton $\mathcal{A}_{\textit{l-lar}}^1$.
-We use the following notations: $C_1 = C \cup \{ \textit{ret}(\textit{rm},a) \}$, $C_2 = C \cup \{ \textit{call}(\textit{put},a,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.}
-  \label{fig:automata APQ1Lar-1 in paper}
+  %\caption{Three possible positions of the first $\textit{call}(\textit{put},a,\_)$ in captured execution}
+  \caption{Orderings to be considered when defining a $\mathsf{MatchedMaxPriority}^>$-complete automaton.}
+  \label{fig:executions APQ1Lar-1 in paper}
 \end{figure}
 
-\begin{figure}[htbp]
+\begin{figure}[t]
   \centering
-  %\includegraphics[width=1 \textwidth]{figures/PIC_HIS_PQ1Lar-pprr-in-paper.pdf}
-  \includegraphics[width=1 \textwidth]{PIC_HIS_PQ1Lar-fouCase.pdf}
+  \includegraphics[width=.9\textwidth]{PIC_AUTO_PQ1Lar-pprr.pdf}
 %\vspace{-10pt}
-  %\caption{Three possible positions of the first $\textit{call}(\textit{put},a,\_)$ in captured execution}
-  \caption{{\color {red}Four cases of ordering actions with value $b$, and possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$}}
-  \label{fig:executions APQ1Lar-1 in paper}
+  \caption{A register automaton capturing the scenario in Figure~\ref{fig:executions APQ1Lar-1 in paper}(a).
+We use the following notations: $A_1 = A \cup \{ \textit{ret}(\textit{rm},a) \}$, $A_2 = A \cup \{ \textit{call}(\textit{put},a,=r) \}$, $A_3 = A_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $A = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.}
+  \label{fig:automata APQ1Lar-1 in paper}
 \end{figure}
@@
 Intuitively, there may exist a value $a$ which is added before another value $b$ such that all the possible linearization points of $\textit{rm}(b)$ are disabled by the position of $\textit{rm}(a)$ in the happens-before.
-We give an example of such an execution $e$ in \figurename~\ref{fig:introduce pb order}.
+We give an example of such an execution $e$ in \figurename~\ref{fig:introduce pb order}, where $p_1\prec p_4$.
 This execution  is not linearizable w.r.t.
@@
 %According to the previous section, which says that linearizaton point of larger priority can be happen in the interval of values with smaller priority.
-Since $\textit{rm}(b)$ cannot take effect during the interval of a smaller priority value, it could be linearized only in one of the two time intervals pictured with dotted lines in \figurename~\ref{fig:introduce pb order}.
-However, each of these time intervals ends before $\textit{call}(\textit{rm},a)$, and thus $\textit{rm}(a)$ cannot be linearized before $\textit{rm}(b)$.
+Since $\textit{rm}(b)$ cannot take effect during the interval of a smaller priority value, it could be only linearized in one of the two time intervals pictured with dotted lines in \figurename~\ref{fig:introduce pb order}.
+However, each of those time intervals ends before $\textit{call}(\textit{rm},a)$, and thus $\textit{rm}(a)$ cannot be linearized before $\textit{rm}(b)$.
 
-\begin{figure}[htbp]
+\begin{figure}[t]
   \centering
@@
 %\vspace{-10pt}
-  \caption{An execution that is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^=$.}
+  \caption{An execution that is not $\mathsf{MatchedMaxPriority}^=$-linearizable.}
   \label{fig:introduce pb order}
@@
 
-To recognize the scenarios in \figurename~\ref{fig:introduce pb order}, we introduce an order $<_{\textit{pb}}$ between values which intuitively, can be thought of as ``a value $a$ is {\color {red}put} before another value $b$''.
+To recognize the scenarios in \figurename~\ref{fig:introduce pb order}, we introduce an order $<_{\textit{pb}}$ between values which intuitively, can be thought of as ``a value $a$ is put before another value $b$''.
 %Let us introduce put-before order $<_{\textit{pb}}$ to formally state that ``an value is putted before another value''.
-Thus, given a data-differentiated execution $e$ and two values $a$ and $b$ of maximal priority, $a <_{\textit{pb}} b$ if one of the following holds: (1) $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$, (2) $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(b)$, or (3) $\textit{rm}(a) <_{\textit{hb}} \textit{put}(b,\_)$.
-Sometimes we use $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$ and $a <_{\textit{pb}}^C b$ to explicitly distinguish between these three cases.
+More precisely, given a data-differentiated execution $e$ and two values $a$ and $b$ of maximal priority, $a <_{\textit{pb}} b$ if one of the following holds: (1) $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$, (2) $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(b)$, or (3) $\textit{rm}(a) <_{\textit{hb}} \textit{put}(b,\_)$.
+Sometimes we use $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$, and $a <_{\textit{pb}}^C b$ to explicitly distinguish between these three cases.
 Let $<_{\textit{pb}}^*$ be the transitive closure of $<_{\textit{pb}}$.
@@
 
-Then, to model the time intervals in which a remove operation like $\textit{rm}(b)$ in \figurename~\ref{fig:introduce pb order}, can be linearized (outside of intervals of smaller priority values) we introduce the notion of gap-point.
-{\color {blue}This notion relies on a standard indexing of actions in an execution, starting with 0.} {\color {red}Here we assume that the index of actions of an execution starts from $0$.}
+To define the time intervals in which a remove like $\textit{rm}(b)$ in \figurename~\ref{fig:introduce pb order}, can be linearized (outside of intervals of smaller priority values) we use the notion of gap-point.
+As before, defining time intervals relies on an indexing of actions in an execution, starting with 0.
 
 \begin{definition}\label{def:gap-point for matched put and rm operations}
-Let $e$ be a data-differentiated execution which contains only one maximal priority $p$, and $\textit{put}(x,p)$ and $\textit{rm}(x)$ two operations in $e$.
-An index $i\in [0,|e|-1]$ is a \emph{gap-point of $x$} if $i$ is greater than or equal to the index of both $\textit{call}(\textit{put},x,p)$ and $\textit{call}(\textit{rm},x)$, smaller than the index of $\textit{ret}(\textit{rm},x)$, and it is not included in the interval of some value with priority smaller than $p$.
+Let $e$ be a data-differentiated execution with only one maximal priority $p$, and $\textit{put}(x,p)$ and $\textit{rm}(x)$ two operations in $e$.
+An index $i\in [0,|e|-1]$ is a \emph{gap-point of $x$} if $i$ is greater than or equal to the index of both $\textit{call}(\textit{put},x,p)$ and $\textit{call}(\textit{rm},x)$, smaller than the index of $\textit{ret}(\textit{rm},x)$, and not included in the interval of a value with priority smaller than $p$.
 \end{definition}
@@
 The case of \figurename~\ref{fig:introduce pb order} can be formally described as follows: $a <_{\textit{pb}}^* b$ while the right-most gap-point of $b$ is before $\textit{call}(\textit{rm},a)$ or $\textit{call}(\textit{put},a,p_4)$.
-The following lemma states that these conditions are enough to characterize non-linearizability w.r.t $\mathsf{MatchedMaxPriority}^{=}$.
+The following lemma states that these conditions are enough to characterize non-linearizability w.r.t.
+$\mathsf{MatchedMaxPriority}^{=}$.
 %(see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}).
 
-\begin{restatable}{lemma}{EPQOneEqualAsPBandGP}
+\begin{lemma}
 \label{lemma:EPQ1Equal as pb order and gap-point}
-Let $e$ be a data-differentiated execution which contains only one maximal priority $p$ such that $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds.
+Let $e$ be a data-differentiated execution with only one maximal priority $p$ such that $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds.
 %Given a data-differentiated $p$-execution $e$ where $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)$ holds.
-Then, $e$ is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ iff $e$ contains two values $x$ and $y$ of maximal priority $p$ such that $y <_{\textit{pb}}^* x$, {\color {red}and the rightmost gap-point of $x$ is strictly smaller than the index of $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$}.
-\end{restatable}
+Then, $e$ is not $\mathsf{MatchedMaxPriority}^{=}$-linearizable iff $e$ contains two values $x$ and $y$ of maximal priority $p$ such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is strictly smaller than the index of $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$.
+\end{lemma}
 
@@
 To characterize violations to linearizability w.r.t.
-$\mathsf{MatchedMaxPriority}^{=}$ using an automaton that tracks a bounded number of values, we show that the number of values needed to witness that $y <_{\textit{pb}}^* x$ for some $x$ and $y$ is bounded.
+$\mathsf{MatchedMaxPriority}^{=}$ using an automaton that tracks a bounded number of values, we show that the number of values needed to witness that $y <_{\textit{pb}}^* x$, for some $x$ and $y$, is bounded.
 %
@@
 
-\begin{restatable}{lemma}{OBOrderHasBoundedLength}
+\begin{lemma}
 \label{lemma:ob order has bounded length}
@@
 \begin{itemize}
-\setlength{\itemsep}{0.5pt}
 \item[-] $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$, or $a <_{\textit{pb}}^C b$,
@@
 \end{itemize}
-\end{restatable}
+\end{lemma}
 
-\smallskip
-To characterize violations to linearizability w.r.t.
-$\mathsf{MatchedMaxPriority}^{=}$, one has to consider all the possible orders between call/return actions of the operations on values $a$, $b$, and $a_i$ in Lemma \ref{lemma:ob order has bounded length}, and the right-most gap point of $b$.
-Excluding the inconsistent cases, {\color {blue}we are left with 5 possible orders that are listed in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}}{\color {red}we are left with 5 possible orders that are shown in \figurename~\ref{fig:five enumerations}, where $o$ denotes the rightmost gap-point of $b$}.
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.8 \textwidth]{PIC-HIS-FiveEnumerations.pdf}
+%\vspace{-10pt}
+  \caption{Orderings to be considered when defining a $\mathsf{MatchedMaxPriority}^=$-complete automaton.}
+  \label{fig:five enumerations}
+\end{figure}
+
+To characterize violations to $\mathsf{MatchedMaxPriority}^{=}$-linearizability, one has to consider all the possible orders between call/return actions of the operations on values $a$, $b$, and $a_i$ in Lemma \ref{lemma:ob order has bounded length}, and the right-most gap point of $b$.
+Excluding the inconsistent cases, we are left with 5 possible orders that are shown in \figurename~\ref{fig:five enumerations}, where $o$ denotes the rightmost gap-point of $b$.
 For each case, we define an automaton recognizing the induced set of violations.
-{\color {blue}For instance, one such case and the corresponding register automaton is shown in \figurename~\ref{fig:an enumeration and its witness automaton} ($o$ denotes the rightmost gap-point of $b$).} {\color {red}For instance, the register automata for the case of \figurename~\ref{fig:five enumerations}(a) is shown in \figurename~\ref{fig:an enumeration and its witness automaton}.} In this case, the conditions in Lemma~\ref{lemma:EPQ1Equal as pb order and gap-point} are equivalent to the fact that intuitively, the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$ is covered by lower priority values (and thus, there is no gap-point of $b$ which occurs after $\textit{call}(\textit{rm},a)$).
-Using again the data-independence property, these lower priority values can all be renamed to a fixed value $d$, and the other values to a fixed value $\top$.
-{\color {red}We show how to generate register automata for other cases in the long version \cite{CONCUR2017Ahmed}.}
+For instance, the register automaton for the case in \figurename~\ref{fig:five enumerations}(a) is shown in \figurename~\ref{fig:an enumeration and its witness automaton}.
+In this case, the conditions in Lemma~\ref{lemma:EPQ1Equal as pb order and gap-point} are equivalent to the fact that intuitively, the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$ is covered by lower priority values (and thus, there is no gap-point of $b$ which occurs after $\textit{call}(\textit{rm},a)$).
+By data-independence, these lower priority values can all be renamed to a fixed value $c$.
+%and the other values to a fixed value $\top$. {\color {red}We show how to generate register automata for other cases in the long version \cite{CONCUR2017Ahmed}.}
 
@@
 
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8 \textwidth]{PIC-HIS-FiveEnumerations.pdf}
-%\vspace{-10pt}
-  \caption{{\color {red} Five cases that need to be considered.}}
-  \label{fig:five enumerations}
-\end{figure}
-
-\begin{figure}[htbp]
+\begin{figure}[t]
   \centering
@@
 %\vspace{-10pt}
-  \caption{A case in deriving a $\mathsf{MatchedMaxPriority}^{=}$-complete automaton.
-We use the following notations: $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true})$, $\textit{call}(\textit{rm},\top), \textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_1 = C \cup \{ \textit{call}(\textit{put},d,<r) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},d) \}$, $C_4 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},d) \}$.}
+  \caption{A register automaton capturing the scenario in Figure~\ref{fig:five enumerations}(a).
+We use the following notations: $A_1 = A \cup \{ \textit{call}(\textit{put},c,<r)\}$, $A_2 = A_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $A_3 = A_2 \cup \{ \textit{ret}(\textit{rm},c) \}$, $A_4 = A \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},c) \}$, where $A = \{ \textit{call}(\textit{put},\top,\textit{true})$,$\textit{ret}(\textit{put},\top,\textit{true})$,$\textit{call}(\textit{rm},\top)$,$\textit{ret}(\textit{rm},\top)$,$\textit{call}(\textit{rm},\textit{empty})$,$\textit{ret}(\textit{rm},\textit{empty})\}$.}
   \label{fig:an enumeration and its witness automaton}

### 2017-07-12 final/introduction.tex
5 values -> 4 values

@@
 Actually, only two registers are needed: one register $r_1$ for storing a priority guessed at the initial state, and one register $r_2$ for reading priorities as they occur in the execution and comparing them with the one stored in $r_1$.
-We show that registers storing values added to or removed from the priority queue are not needed, since any data-independent implementation admits a violation to linearizability whenever it admits a violation where the number of values is constant, and at most 5 (the number of priorities can still be unbounded).
+We show that registers storing values added to or removed from the priority queue are not needed, since any data-independent implementation admits a violation to linearizability whenever it admits a violation where the number of values is constant, and at most 4 (the number of priorities can still be unbounded).
 

### 2017-07-12 final/co-regular-EPQ.tex
decidability

@@
   %\includegraphics[width=0.8 \textwidth]{figures/PIC-Enumeration-WitnessAutomata.pdf}
-  \includegraphics[width=0.8 \textwidth]{PIC-WitnessAutomata-For1.pdf}
+  \includegraphics[width=0.7 \textwidth]{PIC-WitnessAutomata-For1.pdf}
 %\vspace{-10pt}
-  \caption{A register automaton capturing the scenario in Figure~\ref{fig:five enumerations}(a).
-We use the following notations: $A_1 = A \cup \{ \textit{call}(\textit{put},c,<r)\}$, $A_2 = A_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $A_3 = A_2 \cup \{ \textit{ret}(\textit{rm},c) \}$, $A_4 = A \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},c) \}$, where $A = \{ \textit{call}(\textit{put},\top,\textit{true})$,$\textit{ret}(\textit{put},\top,\textit{true})$,$\textit{call}(\textit{rm},\top)$,$\textit{ret}(\textit{rm},\top)$,$\textit{call}(\textit{rm},\textit{empty})$,$\textit{ret}(\textit{rm},\textit{empty})\}$.}
+  \caption{A register automaton for the case in Figure~\ref{fig:five enumerations}(a), where $A_1 = A \cup \{ \textit{call}(\textit{put},c,<r)\}$, $A_2 = A_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $A_3 = A_2 \cup \{ \textit{ret}(\textit{rm},c) \}$, $A_4 = A \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},c) \}$, where $A = \{ \textit{call}(\textit{put},\top,\textit{true})$,$\textit{ret}(\textit{put},\top,\textit{true})$,$\textit{call}(\textit{rm},\top)$,$\textit{ret}(\textit{rm},\top)$,$\textit{call}(\textit{rm},\textit{empty})$,$\textit{ret}(\textit{rm},\textit{empty})\}$.}
   \label{fig:an enumeration and its witness automaton}
@@
 The implementations in $\mathcal{C}$ allow an unbounded number of values but a bounded number of priorities.
-Each method manipulates a finite number of local variables which store Boolean values, or data values from $\mathbb{D}$.
-Methods communicate through a finite number of shared variables that also store Boolean values, or data values from $\mathbb{D}$.
-Data values may be assigned, but never used in program predicates (e.g., in the conditions of if-then-else and while statements) so as to ensure data independence.
+Each method has a finite set of local variables storing Boolean values or values from $\mathbb{D}$.
+Methods communicate through a finite number of shared variables interpreted also as Booleans or values from $\mathbb{D}$.
+To ensure data independence, values in $\mathbb{D}$ may be assigned, but never used in Boolean expressions (e.g., of if-then-else statements).
 This class captures typical implementations, or finite-state abstractions thereof, e.g., obtained via predicate abstraction.
-{\color{blue}Since the $\Gamma$-complete automata $A(\Gamma)$ we define use}{\color {red}Since the $\Gamma$-complete automata $A(\Gamma)$ uses} a fixed set {\color {red}$D=\{a,b,a_1,d,e,\top\}$ of values}, we have that {\color {red}$\mathcal{C}\cap A(\Gamma)\neq\emptyset$ for some $\Gamma$ iff $\mathcal{C}_D\cap A(\Gamma)\neq\emptyset$ where $\mathcal{C}_D$ is the subset of $\mathcal{C}$ that uses only values in $D$.} {\color{blue}The set of executions $\mathcal{C}_D$ can be represented by a Petri Net since bounding the set of values allows us to represent each thread with a finite-state machine.
-For a fixed set of priorities $P$, the register automata $A(\Gamma)$ can be transformed to finite-state automata since the number of possible valuations of the registers is bounded.}
+The $\Gamma$-complete automata we define use a fixed set $D=\{a,b,c,a_1,\top\}$ of values ($a_1$ is needed to deal with the second item in Lemma~\ref{lemma:ob order has bounded length}).
+Therefore, for any $\Gamma$, $\mathcal{C}\cap A(\Gamma)\neq\emptyset$ iff $\mathcal{C}_D\cap A(\Gamma)\neq\emptyset$, where $\mathcal{C}_D$ is the subset of $\mathcal{C}$ that uses only values in $D$.
 
-{\color{red} The set of executions $\mathcal{C}_D$ can be represented by a Vector Addition Systems with States (VASS), since both values and priorities are finite, which implies that each thread and register automata can be transformed into finite-state automata.
+The set of executions $\mathcal{C}_D$ can be represented by a Vector Addition System with States (VASS).
+Since the set of values and priorities is bounded, each method invocation can be represented by a finite-state automaton (see~\cite{conf/esop/BouajjaniEEH13}).
+For a fixed set of priorities $P\subseteq \mathbb{P}$, the register automata $A(\Gamma)$ can be transformed to finite-state automata (the number of valuations of the registers is bounded).
+Thus, checking linearizability of an implementation in $\mathcal{C}$ is PSPACE when the number of threads is bounded, and EXPSPACE, otherwise.
+Moreover, reachability in VASSs can be reduced to checking linearizability of such an implementation.
+Essentially, given an instance of the VASS reachability problem, one can define a priority queue implementation where the $\textit{put}$ methods behave correctly and additionally, they include the code of the VASS simulation defined in~\cite{conf/esop/BouajjaniEEH13}, and the $\textit{rm}$ methods behave correctly, except for the moment where the target state is reached, in which case they trigger a linearizability violation by returning an arbitrary value.
+
+%{\color{red} The set of executions $\mathcal{C}_D$ can be represented by a Vector Addition Systems with States (VASS), since both values and priorities are finite, which implies that each thread and register automata can be transformed into finite-state automata.
 To obtain this transformation, the states of the VASS represent the global variables of $\mathcal{C}_D$, while each counter of the VASS then represents the number of threads which are at a particular control location within a method, with a certain valuation of the local variables.
@@
 When we consider only finite number of threads, the value of counters are bounded and then this problem is PSPACE \cite{DBLP:conf/ac/Esparza96}.
-
-On the other hand, we can mimic transitions of VASS by a priority queue implementation.
+%
+%On the other hand, we can mimic transitions of VASS by a priority queue implementation.
 To obtain this, we keep $\textit{rm}$ unchanged, while in each $\textit{put}$ method, we first put a value into priority queue, and then either simulate one step of transitions or record one increase/decrease of a counter, similarly as that in \cite{conf/esop/BouajjaniEEH13}.
@@
 Based on above discussion, we have the following complexity result.} %The detailed proof can be found in Appendix \ref{subsec:appendix proof and definition in subsection decidability result}.
-
-%\begin{restatable}{theorem}{ComplexityOfPriorityQueue}
-%\label{theorem:complexity of priority queue}
-%Verifying whether an implementation in $\mathcal{C}$ is linearizable w.r.t. $\seqPQ$ is PSPACE-complete for a fixed number of %threads, and EXPSPACE-complete otherwise.
-%\end{theorem}
+%
+%%\begin{restatable}{theorem}{ComplexityOfPriorityQueue}
+%%\label{theorem:complexity of priority queue}
+%%Verifying whether an implementation in $\mathcal{C}$ is linearizable w.r.t. $\seqPQ$ is PSPACE-complete for a fixed number of %threads, and EXPSPACE-complete otherwise.
+%%\end{theorem}
 

### 2017-07-12 final/co-regular-EPQ.tex
14 pages

@@
 USE BIG CAPS FOR SETS OF TRANSITION LABELS.
-To define a $\mathsf{MatchedMaxPriority}^>$-complete automaton we need to consider several cases depending on the order between the call/return actions of the $\textit{put}$/$\textit{rm}$ operations that add and respectively, remove the value $b$.
-For example, the case where the put happens-before the remove (as in \figurename~\ref{fig:introduce gap for EPQ1Lar}) is pictured in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
+To define a $\mathsf{MatchedMaxPriority}^>$-complete automaton, we need to consider all the possible orders between the call/return actions of the $\textit{put}$/$\textit{rm}$ operations that add and respectively, remove the value $b$.
+The case where the put happens-before the remove (as in \figurename~\ref{fig:introduce gap for EPQ1Lar}) is pictured in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
 This automaton captures the three possible ways of ordering the first action $\textit{ret}(\textit{put},a,\_)$ w.r.t.
@@
 
-To characterize violations to linearizability w.r.t.
-$\mathsf{MatchedMaxPriority}^{=}$ using an automaton that tracks a bounded number of values, we show that the number of values needed to witness that $y <_{\textit{pb}}^* x$, for some $x$ and $y$, is bounded.
+%To characterize violations to linearizability w.r.t. $\mathsf{MatchedMaxPriority}^{=}$ using an automaton that tracks a bounded number of values, we show 
+The following shows that the number of values needed to witness that $y <_{\textit{pb}}^* x$, for some $x$ and $y$, is bounded.
 %
@@
 To characterize violations to $\mathsf{MatchedMaxPriority}^{=}$-linearizability, one has to consider all the possible orders between call/return actions of the operations on values $a$, $b$, and $a_i$ in Lemma \ref{lemma:ob order has bounded length}, and the right-most gap point of $b$.
-Excluding the inconsistent cases, we are left with 5 possible orders that are shown in \figurename~\ref{fig:five enumerations}, where $o$ denotes the rightmost gap-point of $b$.
+Excluding the inconsistent cases, we are left with the five orders in \figurename~\ref{fig:five enumerations}, where $o$ denotes the rightmost gap-point of $b$.
 For each case, we define an automaton recognizing the induced set of violations.
-For instance, the register automaton for the case in \figurename~\ref{fig:five enumerations}(a) is shown in \figurename~\ref{fig:an enumeration and its witness automaton}.
-In this case, the conditions in Lemma~\ref{lemma:EPQ1Equal as pb order and gap-point} are equivalent to the fact that intuitively, the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$ is covered by lower priority values (and thus, there is no gap-point of $b$ which occurs after $\textit{call}(\textit{rm},a)$).
-By data-independence, these lower priority values can all be renamed to a fixed value $c$.
+The register automaton for the case in \figurename~\ref{fig:five enumerations}(a) is shown in \figurename~\ref{fig:an enumeration and its witness automaton}.
+In this case, Lemma~\ref{lemma:EPQ1Equal as pb order and gap-point} is equivalent to the fact that intuitively, the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$ is covered by lower priority values (thus, there is no gap-point of $b$ which occurs after $\textit{call}(\textit{rm},a)$).
+By data-independence, these lower priority values can be renamed to a fixed value $c$.
 %and the other values to a fixed value $\top$. {\color {red}We show how to generate register automata for other cases in the long version \cite{CONCUR2017Ahmed}.}

### 2017-07-12 final/related.tex
14 pages

--- 2017-04-21 related.tex
@@
 Existing automated methods for proving linearizability of a concurrent object
-implementation are also based on reductions to safety
+ are also based on reductions to safety
 verification, e.g.,~\cite{conf/tacas/AbdullaHHJR13, conf/concur/HenzingerSV13,
@@
 proofs~\cite{conf/concur/HenzingerSV13} reduce linearizability to the
-verification of four simpler safety properties. However, this approach has only
+verification of four simpler safety properties. This approach has only
 been applied to queues, and has not produced a fully automated

### 2017-07-12 appendix-EPQ.tex
modify

@@
 We generate register automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}.
-Here $C_1 = \{ \textit{call}(\textit{put},a,\textit{true})$, $\textit{ret}(\textit{put},a,\textit{true}), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},b),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2$ $= C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{put},b,\textit{true}) \}$.
+Here $A_1 = \{ \textit{call}(\textit{put},\top,\textit{true})$, $\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},\top),\textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},b),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $A_2$ $= A_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{put},b,\textit{true}) \}$.
 
@@
 We generate register automata $\mathcal{A}_{\textit{SinPri}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata for FIFO-2}.
-Here $C_1 = \{ \textit{call}(\textit{put},a,\textit{true}),\textit{ret}(\textit{put},a,\textit{true}), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{call}(\textit{rm},b) + \textit{ret}(\textit{rm},b) \}$.
+Here $A_1 = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},\top),\textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $A_2 = A_1 \cup \{ \textit{call}(\textit{rm},b), \textit{ret}(\textit{rm},b) \}$.
 
@@
 We generate register automata $\mathcal{A}_{\textit{SinPri}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata for FIFO-3}.
-Here $C_1 = \{ \textit{call}(\textit{put},a,\textit{true}),\textit{ret}(\textit{put},a,\textit{true}), \textit{call}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,\textit{true}) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_4 = C_3 \cup \{ \textit{call}(\textit{rm},b) \}$, $C_5 = C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $C_6 = C_5 \cup \{ \textit{call}(\textit{rm},b) \}$.
+Here $A_1 = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},\top),\textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $A_2 = A_1 \cup \{ \textit{ret}(\textit{put},b,\textit{true}) \}$, $A_3 = A_2 \cup \{ \textit{ret}(\textit{rm},b) \}$, $A_4 = A_3 \cup \{ \textit{call}(\textit{rm},b) \}$, $A_5 = A_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, $A_6 = A_5 \cup \{ \textit{call}(\textit{rm},b) \}$.
 
@@
 We generate register automata $\mathcal{A}_{\textit{SinPri}}^4$ for the forth case, and it is shown in \figurename~\ref{fig:automata for FIFO-4}.
-Here $C_1 = C \cup \{ \textit{call}(\textit{rm},b) \}$, and $C_2 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{call}(\textit{rm},a), \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},d,\textit{true}),\textit{ret}(\textit{put},d,\textit{true}), \textit{call}(\textit{rm},d),\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.
+Here $A_1 = A \cup \{ \textit{call}(\textit{rm},b) \}$, and $A_2 = A \cup \{ \textit{ret}(\textit{put},b,\textit{true}), \textit{call}(\textit{rm},a), \textit{ret}(\textit{rm},a) \}$, where $A = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},\top),\textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.
 
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] If this is because case $1$, case $2$ or case $3$: $r$ maps $v_b$ into $b$ and maps all other values into $a$.
-
-\item[-] If this is because case $4$: $r$ maps $v_a$ and $v_b$ into $a$ and $b$, respectively, and maps all other values into $d$.
+\item[-] If this is because case $1$, case $2$ or case $3$: $r$ maps $v_b$ into $b$ and maps all other values into $\top$.
+
+\item[-] If this is because case $4$: $r$ maps $v_a$ and $v_b$ into $a$ and $b$, respectively, and maps all other values into $\top$.
 \end{itemize}

### 2017-07-13 final/main.tex
the final version sent

\documentclass[a4paper,UKenglish]{lipics-v2016}

\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{color}
\usepackage{amsfonts,amssymb}
\usepackage{verbatim}
\usepackage{ulem}    
\normalem
\usepackage{enumerate}

\usepackage[linesnumbered,ruled,procnumbered,noend]{algorithm2e}
\usepackage{stmaryrd}
\usepackage{thmtools}

\usepackage[version=0.96]{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,snakes,automata,backgrounds,petri}

\DeclareSymbolFont{largesymbolsA}{U}{txexa}{m}{n}
\SetSymbolFont{largesymbolsA}{bold}{U}{txexa}{bx}{n}
\DeclareFontSubstitution{U}{txexa}{m}{n}
\DeclareMathSymbol{\bigsqcupplus}{\mathop}{largesymbolsA}{"02}

\newcommand{\nondsum}{\bigsqcupplus}
\newcommand{\probplus}[1]{\oplus_{#1}}
\newcommand{\bang}{!\,}
\newcommand{\nondplus}{{\textstyle\bigsqcupplus}}
\newcommand{\partmap}{\rightharpoonup}
\newcommand{\map}{\rightarrow}
\newcommand{\exc}{\alpha}
\newcommand{\exec}{\mathit{exec}}
\newcommand{\execp}{\mathit{execp}}
\newcommand{\Act}{\mathit{Act}}
\newcommand{\Sec}{\mathit{Sec}}
\newcommand{\Obs}{\mathit{Obs}}
\newcommand{\etree}{\mathit{etree}}
\newcommand{\lstate}{\mathit{lst}}
\newcommand{\fstate}{\mathit{fst}}
\newcommand{\trans}{\mathcal{T}}
\newcommand{\Aut}{\mathcal{M}}
\newcommand{\init}{\mathit{init}}
\newcommand{\perr}{\mathcal{P}}

\newcommand{\calo}{\mathcal{O}}
\newcommand{\cals}{\mathcal{S}}
\newcommand{\sseq}{\vec s}
\newcommand{\oseq}{\vec o}
\newcommand{\ccsp}{CCS$_p$}

\newcommand{\bigfrac}[2]{\frac{\raisebox{1ex}{$#1$}}{\raisebox{-1.5ex}{$#2$}}}
\newcommand{\nondarr}[1]{\overset{#1}{\longrightarrow}}
\newcommand{\Nondarr}[1]{\overset{#1}{\Longrightarrow}}
\newcommand{\vectorArrow}[1]{\stackrel{\longrightarrow}{\mbox{#1}}}
\newcommand{\probarr}[1]{\overset{#1}{\dashrightarrow}}
\newcommand{\paral}{\,|\,}
\newcommand{\outp}[1]{\overline{#1}}
\renewcommand{\Pr}{{\rm Pr}}

\newcommand{\TSO}{\textrm{TSO}}
\newcommand{\PSO}{\textrm{PSO}}

\newcommand{\lin}{\textrm{linearizability}}
\newcommand{\slin}{\textrm{static linearizability}}
\newcommand{\qlin}{\textrm{quasi linearizability}}
\newcommand{\TTlin}{\textrm{TSO-to-TSO linearizability}}

\newcommand{\pair}[2]{\langle #1 , #2 \rangle}
\newcommand{\setof}[2]{\{ \, #1 \mid #2 \, \}}
\newcommand{\set}[1]{\{ {#1}  \}  }
\newcommand{\den}[1]{[\![#1]\!]}
\newcommand{\mean}[1]{|\!|#1|\!|}
\newcommand{\forget}[1]{}

\newcommand{\itbox}[1]{{\it #1\/}}
\newcommand{\un}[1]{\uline{#1}}
\newcommand{\ov}[1]{\overline{#1}}
\newcommand{\smallspace}{\vspace{10mm}}
\newcommand{\is}{\mbox{$\Longleftarrow\ $}}
\newcommand{\pright}[1]{\hfill{#1}}
\newcommand{\bnfor}{\;\;\mid\;\;}

\newcommand{\ar}[1]{\stackrel{\scriptstyle #1}{\longrightarrow}}

\newcommand{\calB}{{\cal B}}
\newcommand{\calF}{{\cal F}}
\newcommand{\calP}{{\cal P}}
\newcommand{\order}{{\cal O}}
\newcommand{\size}[1]{|#1|}

\newcommand{\LTS}{\textit{LTS}}
\newcommand{\bedt}[1]{{\color{blue}#1}}
\newcommand{\redt}[1]{{\color{red}#1}}

\title{Checking Linearizability of Concurrent Priority Queues\footnote{This work is supported in part by the European Research Council (ERC) under the
European Union's Horizon 2020 research and innovation programme (grant agreement No 678177).}}
\titlerunning{Checking Linearizability of Concurrent Priority Queues}

\author[1]{Ahmed Bouajjani}
\author[1]{Constantin Enea}
\author[1]{Chao Wang}
\affil[1]{IRIF, University Paris Diderot, Paris, France \\ \texttt{\{abou,cenea,wangch\}@irif.fr}}
\authorrunning{A. Bouajjani, C. Enea, and C. Wang}

\EventEditors{Roland Meyer and Uwe Nestmann}
\EventNoEds{2}
\EventLongTitle{28th International Conference on Concurrency Theory (CONCUR 2017)}
\EventShortTitle{CONCUR 2017}
\EventAcronym{CONCUR}
\EventYear{2017}
\EventDate{September 5--8, 2017}
\EventLocation{Berlin, Germany}
\EventLogo{}
\SeriesVolume{85}
\ArticleNo{12}

\subjclass{D.2.4 Software/Program Verification}

\keywords{Concurrency, Linearizability, Model Checking}

\begin{document}

\maketitle

\begin{abstract}
Efficient implementations of concurrent objects such as atomic collections are essential to modern computing. Unfortunately their correctness criteria — linearizability with respect to given ADT specifications — are hard to verify. Verifying linearizability is undecidable in general, even on classes of implementations where the usual control-state reachability is decidable. In this work we consider concurrent priority queues which are fundamental to many multi-threaded applications like task scheduling or discrete event simulation, and show that verifying linearizability of such implementations is reducible to control-state reachability. This reduction entails the first decidability results for verifying concurrent priority queues with an unbounded number of threads, and it enables the application of existing safety-verification tools for establishing their correctness.
\end{abstract}

\forget{
\noindent Keywords: weak memory model, $\textit{linearizability}$,
$\textit{TSO-to-TSO linearizability}$
}

\section{Introduction}
\label{sec:introduction}

 Multithreaded software is typically built with specialized “concurrent
  objects” like atomic integers, queues, maps, priority queues. These objects’ methods are
  designed to confom to better established sequential specifications, a property known as \emph{linearizability}~\cite{journals/toplas/HerlihyW90},
  despite being optimized to avoid blocking and exploit parallelism, e.g.,~by
  using machine instructions like compare-and-swap. Intuitively, linearizability asks that every individual operation appears to take place instantaneously at some point between its invocation and its return. Verifying linearizability is intrinsically hard, and undecidable in general~\cite{conf/esop/BouajjaniEEH13}. 
However, recent work~\cite{DBLP:conf/icalp/BouajjaniEEH15} has shown that for particular objects,
e.g., registers, mutexes, queues, and stacks, the problem of verifying linearizability becomes decidable (for finite-state implementations).

In this paper, we consider another important object, namely the priority queue, which is essential for applications such as task scheduling and discrete event simulation. Numerous implementations have been proposed in the research literature, e.g.,~\cite{DBLP:conf/ppopp/AlistarhKLS15,DBLP:conf/wdag/CalciuMH14,DBLP:conf/opodis/LindenJ13,DBLP:conf/podc/ShavitZ99,DBLP:conf/ipps/ShavitL00}, and concrete implementations exist in many modern languages like C++ or Java.
Priority queues are collections providing $\textit{put}$ and $\textit{rm}$ methods for adding and removing values. Every added value is associated to a priority and a remove operation returns a minimal priority value. For generality, we consider a partially-ordered set of priorities. Values with incomparable priorities can be removed in any order, and values having the same priority are removed in the FIFO order. Implementations like the PriorityBlockingQueue in Java where same priority values are removed in an arbitrary order can be modeled in our framework by renaming equal priorities to incomparable priorities (while preserving the order constraints).

Compared to previously studied collections like stacks and queues, the main challenge in dealing with priority queues is that the order in which values are removed is not fixed by the happens-before between add/remove operations (e.g., in the case of queues, values are removed in the order in which they were inserted), but by parameters of the $\textit{put}$ operations (the priorities) which come from an unbounded domain. For instance, the sequential behavior $\textit{put}(a,p_1)\cdot \textit{put}(b,p_3)\cdot \textit{put}(c,p_2)\cdot \textit{rm}(a,p_1)\cdot \textit{rm}(c,p_2)$ where the priority $p_1$ is less than $p_2$ which is less than $p_3$, is not admitted neither by the regular queue nor the stack.

We give a characterization of concurrent priority queue behaviors violating linearizability in terms of automata. This characterization enables a reduction of checking linearizability for arbitrary implementations to reachability or invariant checking, and implies decidability for checking linearizability of finite-state implementations. While linearizability violations for stacks and queues can be described using finite-state automata~\cite{DBLP:conf/icalp/BouajjaniEEH15}, the case of priority queues requires \emph{register automata} where registers are used to store and compare priorities.

This characterization is obtained in several steps. We define a recursive procedure that recognizes valid sequential executions, which is then extended to recognize linearizable concurrent executions. Intuitively, for an execution $e$, this procedure deals with values occurring in $e$ one by one, starting with values of maximal priority (to be removed the latest). For each value $x$, it checks whether $e$ satisfies some property ``local'' to that value, i.e., which is agnostic to how the operations adding or removing other values are ordered between them (w.r.t. the happens-before), other than how they are ordered w.r.t. the operations on $x$. When this property holds, the procedure is applied recursively on the rest of the execution, without the operations on $x$. This procedure works only for executions where a value is added at most once, but this is not a limitation for \emph{data-independent} implementations whose behavior doesn't depend on the values that are added or removed. In fact, all the implementations that we are aware of are data-independent.

Next, we show that checking whether an execution violates this ``local'' property for a value $x$ can be done using a class of register automata~\cite{DBLP:journals/tcs/KaminskiF94,DBLP:conf/icalp/Cerans94,DBLP:conf/stacs/SegoufinT11} (transition systems where the states consist of a fixed set of registers that can receive values and be compared). Actually, only two registers are needed: one register $r_1$ for storing a priority guessed at the initial state, and one register $r_2$ for reading priorities as they occur in the execution and comparing them with the one stored in $r_1$. We show that registers storing values added to or removed from the priority queue are not needed, since any data-independent implementation admits a violation to linearizability whenever it admits a violation where the number of values is constant, and at most 4 (the number of priorities can still be unbounded).

The remainder of this article is organized as follows.
Section~\ref{sec:priority queue and data-independence} describes the priority queue ADT, lists several semantic properties like data-independence, and recalls the notion of linearizability.
Section~\ref{sec:checking inclusion by recursive procedure} defines a recursive procedure for checking linearizability of concurrent priority queue behaviors.
Section~\ref{sec:co-regular of extended priority queues} gives an automata characterization of the violations to linearizability, and
Section~\ref{sec:related} discusses related work. Detailed proofs and constructions can be found in the extended version~\cite{CONCUR2017Ahmed}.

\newcommand{\seqPQ}{\mathsf{SeqPQ}}

\section{The Priority Queue ADT}
\label{sec:priority queue and data-independence}

We consider priority queues whose interface contains two methods $\textit{put}$ and $\textit{rm}$ for adding and respectively, removing a value. Each value is assigned with a priority when being added to the data structure (by calling $\textit{put}$) and the remove method $\textit{rm}$ removes a value with a minimal priority. For generality, we assume that the set of priorities is partially-ordered. Incomparable priorities can be removed in any order. When multiple values are assigned with the same priority, $\textit{rm}$ returns the least recent value. Also, when the set of values stored in the priority queue is empty, $\textit{rm}$ returns the distinguished value $\textit{empty}$. In this section, we formalize (concurrent) executions and implementations, introduce a set of properties satisfied by all the implementations we are aware of, and recall the standard correctness criterion for concurrent implementations of ADTs known as \emph{linearizability}~\cite{journals/toplas/HerlihyW90}.

\subsection{Executions}\label{ssec:exec}

We fix a (possibly infinite) set $\mathbb{D}$ of data values, a (possibly infinite) set $\mathbb{P}$ of priorities, a partial order $\prec$ among elements in $\mathbb{P}$, and an infinite set $\mathbb{O}$ of operation identifiers.
The latter are used to match call and return actions of the same invocation. Call actions $\textit{call}_o(\textit{put},a,p)$ and $\textit{call}_o(\textit{rm},a')$ with $a\in \mathbb{D}$, $a'\in \mathbb{D}\cup\{\textit{empty}\}$, $p \in \mathbb{P}$, and $o \in \mathbb{O}$, combine a method name and a set of arguments with an operation identifier. The return value of a remove is transformed to an argument value for uniformity~\footnote{Method return values are guessed nondeterministically, and validated at return points.
This can be handled using {\tt assume} statements, which only admit executions satisfying a given predicate.}.
The return actions are denoted in a similar way as $\textit{ret}_o(\textit{put},a,p)$ and respectively, $\textit{ret}_o(\textit{rm},a')$.

An \emph{execution} $e$ is a sequence of call and return actions which satisfy the following well-formedness properties: each return is preceded by a matching call (having the same operation identifier), and each operation identifier is used in at most one call/return. We assume every set of executions is closed under isomorphic renaming of operation identifiers. An $m(a)$-operation in an execution $e$ is an operation identifier $o$ s.t. $e$ contains the actions $\textit{call}_o(m,a)$ and $\textit{ret}_o(m,a)$.
An execution is called \emph{sequential} when no two operations overlap, i.e., each call action is immediately followed by its matching return action, and \emph{concurrent} otherwise. For readability, we write a sequential execution as a sequence of $\textit{put}(a,p)$ and $\textit{rm}(a)$ symbols representing a pair of actions $\textit{call}_o(\textit{put},a,p)\cdot \textit{ret}_o(\textit{put},a,p)$ and $\textit{call}_o(\textit{rm},a)\cdot \textit{ret}_o(\textit{rm},a)$, respectively ($o\in\mathbb{O}$). For example, given two priorities $p_1 \prec p_2$, $\textit{put}(a,p_2) \cdot \textit{put}(b,p_1) \cdot \textit{rm}(b)$ is a sequential execution of the priority queue ($\textit{rm}$ returns $b$ because it has smaller priority).

We define $\seqPQ$, the set of sequential priority queue executions, semantically via a labelled transition system (LTS, for short). An LTS is a tuple $A=(Q,\Sigma,\rightarrow,q_0)$, where $Q$ is a set of states, $\Sigma$ is an alphabet of transition labels, $\rightarrow\subseteq Q\times\Sigma\times Q$ is a transition relation, and $q_0$ is the initial state. We model the priority queue as an LTS $\textit{PQ}$ where states are mappings associating priorities in $\mathbb{P}$ with sequences of values in $\mathbb{D}$, representing a snapshot of the priority queue (for each priority, the values are ordered as they were inserted), and the transition labels are $\textit{put}(a,p)$ and $\textit{rm}(a)$. Each transition modifies the state as expected. For example, $q_1 \xrightarrow{\textit{rm}(\textit{empty})} q_2$ if $q_1 = q_2$, and $q_1$ and $q_2$ map each priority to the empty sequence $\epsilon$. Then, $\seqPQ$ is the set of traces (words) accepted by $\textit{PQ}$.

An implementation $\mathcal{I}$ is a set of executions. Implementations represent libraries whose methods are called by external programs. In the remainder of this work, we consider only \emph{completed} executions, where each call action has a corresponding return action. This simplification is sound when the method invocations can always make progress in isolation.

\subsection{Semantic Properties of Priority Queues}\label{ssec:semantic_prop}

We define two properties which are 
important for our results: (1) \emph{data independence}~\cite{conf/popl/Wolper86,conf/tacas/AbdullaHHJR13} states that priority queue behaviors do not depend on the actual values which are added to the queue, and (2) \emph{closure under projection}~\cite{DBLP:conf/icalp/BouajjaniEEH15} states that executions remain valid by removing all the operations adding or removing certain values.

An execution $e$ is \emph{data-differentiated} if every value is added at most once, i.e., for each $d \in \mathbb{D}$, $e$ contains at most one action $\textit{call}_o(\textit{put},d,p)$ with $o\in\mathbb{O}$ and $p\in \mathbb{P}$. Note that this property concerns only values, a data-differentiated execution $e$ may contain more than one value with the same priority. The subset of data-differentiated executions of a set of executions $E$ is denoted by $E_{\neq}$.

A renaming function $r$ is a function from $\mathbb{D}$ to $\mathbb{D}$. Given an execution $e$, we denote by $r(e)$ the execution obtained from $e$ by replacing every data value $x$ by $r(x)$. Note that $r$ renames only the values and keeps the priorities unchanged. Intuitively, renaming values has no influence on the behavior of the priority queue, contrary to renaming priorities.

\begin{definition}\label{def:priority-value data-independence}
A set of executions $E$ is \emph{data independent} iff
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] for all $e \in E$, there exists $e' \in E_{\neq}$ and a renaming function $r$, such that $e=r(e')$,

\item[-] for all $e \in E$ and for all renamings $r$, $r(e) \in E$.
\end{itemize}
\end{definition}

The following lemma is a direct consequence of definitions.

\begin{lemma}
$\seqPQ$ is data independent.
\end{lemma}

Beyond sequential executions, every concurrent priority queue implementation that we are aware of is data-independent. From now on, we consider only data-independent implementations. This assumption enables a reduction from checking the correctness of an implementation $\mathcal{I}$ to checking the correctness of its data-differentiated executions in $\mathcal{I}_{\neq}$.

Besides data independence, the sequential executions of the priority queue satisfy the following closure property: an execution remains valid when removing all the operations with an argument in some set of values $D \subseteq \mathbb{D}$ and any $\textit{rm}(\textit{empty})$ operation (since they are read-only and they don't affect the queue's state).
To distinguish between different $\textit{rm}(\textit{empty})$ operations while simplifying the technical exposition, we assume that they receive as argument a value, i.e., call actions are of the form $\textit{call}_o(\textit{rm},\textit{empty},a)$ for some $a\in \mathbb{D}$. We will make explicit this argument only when needed in our technical development. The projection $e \vert D$ of an execution $e$ to a set of values $D \subseteq \mathbb{D}$ is obtained from $e$ by erasing all the call/return actions with an argument not in $D$. We write $e \setminus x$ for the projection $e \vert_{ \mathbb{D} \setminus \{ x \} }$. Let $\textit{proj}(e)$ be the set of all projections of $e$ to a set of values $D \subseteq \mathbb{D}$.

\begin{lemma}
\label{lem:closure_proj}
$\seqPQ$ is closed under projection, i.e., $\textit{proj}(e)\subseteq \seqPQ$ for each $e\in \seqPQ$.
\end{lemma}

\subsection{Linearizability}\label{ssec:lin}

We recall the notion of \emph{linearizability}~\cite{journals/toplas/HerlihyW90} which is the \emph{de facto} standard correctness condition for concurrent data structures.
Given an execution $e$, the happen-before relation $<_{\textit{hb}}$ between operations~\footnote{In general, we refer to operations using their identifiers.} is defined as follows: $o_1 <_{\textit{hb}} o_2$, if the return action of $o_1$ occurs before the call action of $o_2$ in $e$. The happens-before relation is an interval order \cite{DBLP:conf/popl/BouajjaniEEH15}: for distinct $o_1,o_2,o_3,o_4$, if $o_1 <_{\textit{hb}} o_2$ and $o_3 <_{\textit{hb}} o_4$, then either $o_1 <_{\textit{hb}} o_4$, or $o_3 <_{\textit{hb}} o_2$. Intuitively, this comes from the fact that concurrent threads share a notion of global time.

Given a (concurrent) execution $e$ and a sequential execution $s$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if there is a bijection $f: O_1 \rightarrow O_2$, where $O_1$ and $O_2$ are the set of operations of $e$ and $s$, respectively, such that (1) the call and return actions with identifier $o$ and $f(o)$, respectively, are the same
and (2) if $o_1 <_{\textit{hb}} o_2$, then $f(o_1) <_{\textit{hb}} f(o_2)$.
A (concurrent) execution $e$ is linearizable w.r.t. a set $S$ of sequential executions, denoted $e \sqsubseteq S$, if there exists $s \in S$ such that $e \sqsubseteq s$. A set of concurrent executions $E$ is linearizable w.r.t. $S$, denoted $E \sqsubseteq S$, if $e \sqsubseteq S$ for all $e \in E$.

The following lemma states that by data-independence, it is enough to consider only data-differentiated executions when checking linearizability. 
Section~\ref{sec:checking inclusion by recursive procedure} will focus on characterizing linearizability for data-differentiated executions.

\begin{lemma}\label{lemma:data differentiated is enough for PQ}
A data-independent implementation $\mathcal{I}$ is linearizable w.r.t. a data-independent set $S$ of sequential executions, if and only if $\mathcal{I}_{\neq}$ is linearizable w.r.t. $S_{\neq}$.
\end{lemma}

\section{Checking Linearizability of Priority Queue Executions}
\label{sec:checking inclusion by recursive procedure}

We define a recursive procedure for checking linearizability of a data-differentiated execution w.r.t. $\seqPQ$.
To ease the exposition, Section~\ref{ssec:seq_exec} introduces a recursive procedure for checking whether a data-differentiated \emph{sequential} execution is admitted by the priority queue which is then extended to the concurrent case in Section~\ref{ssec:conc_exec}.

\subsection{Characterizing Data-Differentiated Sequential Executions}\label{ssec:seq_exec}

The recursive procedure $\textit{Check-PQ-Seq}$ outlined in Algorithm~\ref{alg:seq_check} checks whether a data-differentiated sequential execution belongs to $\seqPQ$ (i.e., if it is accepted by the LTS $PQ$).
 Roughly, it selects one or two operations in the input execution, checks whether their return values are correct by ignoring the order between the other operations other than how they are ordered w.r.t. the selected ones, and calls itself recursively on the execution without the selected operations.

We explain how the procedure works on the following execution:
\begin{align}
\hspace{-5mm}\textit{put}(c,p_2)\hspace{-.5mm} \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{rm}(\textit{empty}) \cdot \textit{put}(d,p_2) \cdot \textit{put}(\mathit{f},p_3) \cdot \textit{rm}(\mathit{f}) \cdot \textit{put}(b,p_1)\label{eq:ex_rec1}
\end{align}
where $p_1$, $p_2$, $p_3$ are priorities such that $p_1 \prec p_2$ and $p_1 \prec p_3$, and $p_2$ and $p_3$ are incomparable. Since the $\textit{rm}(\textit{empty})$ operations 
are read-only (they don't affect the queue's state), they are selected first. An $\textit{rm}(\textit{empty})$-operation $o$ is correct when every $\textit{put}(x,p)$ operation before $o$ is matched to a $\textit{rm}(x)$ operation which also occurs before $o$. This is true in this case for $x\in \{a,c\}$. Thus, the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
\begin{align}
\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(f,p_3) \cdot \textit{rm}(f) \cdot \textit{put}(b,p_1)\label{eq:ex_rec2}
\end{align}
When the execution contains no $\textit{rm}(\textit{empty})$-operation, the procedure selects a $\textit{put}$ operation adding a value that is not removed and that has a maximal priority. For (\ref{eq:ex_rec2}), it selects $\textit{put}(d,p_2)$ because $p_2$ is a maximal priority. This operation is correct since $d$ is the last value with priority $p_2$ in the execution, and the correctness of (\ref{eq:ex_rec2}) reduces to the correctness of
\begin{align}
\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(f,p_3) \cdot \textit{rm}(f) \cdot \textit{put}(b,p_1)\label{eq:ex_rec3}
\end{align}
If no operations like above can be found, $\textit{Check-PQ-Seq}$ selects a pair of $\textit{put}$ and $\textit{rm}$ operations adding and removing the same maximal priority value. For (\ref{eq:ex_rec2}), it can select 
$\textit{put}(c,p_2)$ and $\textit{rm}(c)$. The value returned by $\textit{rm}(c)$ is correct if all the values of priority smaller than $p_2$ added before $\textit{rm}(c)$ are also removed before $\textit{rm}(c)$. In this case, $a$ is the only value of priority smaller than $p_2$ and it satisfies this property. Applying a similar reasoning for all the remaining values, it can be proved that this execution is correct.

\begin{algorithm}[t]
\footnotesize{
\KwIn {A data-differentiated sequential execution $e$}
\KwOut{$\mathsf{true}$ iff $e\in \seqPQ$}

\If {$e = \epsilon$}
{\Return $\mathsf{true}$\;}

\If {$\mathsf{Has\text{-}EmptyRemoves}(e)$}
{
    \If {$\exists\ o=\textit{rm}(\textit{empty})\in e$ such that $\mathsf{EmptyRemove\text{-}Seq}(e,o)$ holds}
    {
        \KwRet $\textit{Check-PQ-Seq}(e \setminus o)$\;
    }
}
\ElseIf{$\mathsf{Has\text{-}UnmatchedMaxPriority}(e)$}
{
    \If {$\exists\ x \in \textit{values}(e)$ such that $\mathsf{UnmatchedMaxPriority\text{-}Seq}(e,x)$ holds}
    {
        \KwRet $\textit{Check-PQ-Seq}(e \setminus x)$\;
    }
}

\Else
{
    \If {$\exists\ x \in \textit{values}(e)$ such that $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ holds}
    {
        \KwRet $\textit{Check-PQ-Seq}(e \setminus x)$\;
    }
    \Else {\KwRet $\mathsf{false}$\;}
}}
\caption{$\textit{Check-PQ-Seq}$}
\label{alg:seq_check}
\end{algorithm}

Formally, the selected operations depend on the following set of predicates on executions:
\begin{align*}
& \mathsf{Has\text{-}EmptyRemoves}(e)=\mathsf{true} \mbox{ iff  $e$ contains a $\textit{rm}(\textit{empty})$-operation} \hspace{1cm}\\
& \mathsf{Has\text{-}UnmatchedMaxPriority}(e)=\mathsf{true} \mbox{ iff $p\in \textit{unmatched-priorities}(e)$ for a maximal $p$} 
\end{align*}
where $\textit{priorities}(e)$, resp., $\textit{unmatched-priorities}(e)$, is the set of priorities occurring in $\textit{put}$ operations of $e$, resp., in $\textit{put}$ operations of $e$ for which there is no $\textit{rm}$ operation removing the same value. We call the latter \emph{unmatched} put operations. A put operation which is not unmatched is called \emph{matched}. For simplicity, we consider the following syntactic sugar $\mathsf{Has\text{-}MatchedMaxPriority}(e)=\neg \mathsf{Has\text{-}EmptyRemoves}(e)\land \neg \mathsf{Has\text{-}UnmatchedMaxPriority}(e)$. By an abuse of notation, we assume  $\mathsf{Has\text{-}UnmatchedMaxPriority}(e) \Rightarrow \neg \mathsf{Has\text{-}EmptyRemoves}(e)$ (this is sound by the order of the conditionals in $\textit{Check-PQ-Seq}$).

The predicates defining the correctness of the selected operations are defined as follows:
\begin{align*}
&\mathsf{EmptyRemove\text{-}Seq}(e,o)=\mathsf{true} \mbox{ iff  $e= u\cdot o\cdot v$ and $\textit{matched}(u)$} \\
&\mathsf{UnmatchedMaxPriority\text{-}Seq}(e,x)=\mathsf{true}  \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v$, $p\not\prec \textit{priorities}(u\cdot v)$,} \\
&\hspace{6.4cm}p\not\in \textit{priorities}(v) \\
&\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)=\mathsf{true} \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v\cdot \textit{rm}(x)\cdot w$, $\textit{matched}_\prec(u\cdot v,p)$,} \\
&\hspace{4.3cm}\mbox{$p\not\preceq \textit{unmatched-priorities}(u\cdot v\cdot w)$, $p\not\prec \textit{priorities}(u\cdot v\cdot w)$,} \\
&\hspace{4.3cm}\mbox{and  $p\not\in \textit{priorities}(v\cdot w)$}
\end{align*}
where $p\prec \textit{priorities}(e)$ when $p\prec p'$ for some $p'\in \textit{priorities}(e)$ (and similarly for $p\prec \textit{unmatched-priorities}(e)$ or $p\preceq \textit{unmatched-priorities}(e)$),
$\textit{matched}_\prec(e,p)$ holds when each value with priority strictly smaller than $p$ is removed in $e$, and $\textit{matched}(e)$ holds when $\textit{matched}_\prec(e,p)$ holds for each $p\in\mathbb{P}$. Compared to the example presented at the beginning of the section, these predicates take into consideration that multiple values with the same priority are removed in FIFO order: the predicate $\mathsf{MatchedMaxPrioritySeq}(e,x)$ holds when $x$ is the last value with priority $p$ added in $e$.

When $o$ is an $\textit{rm}(\textit{empty})$-operation, $e\setminus o$ is the maximal subsequence of $e$ which doesn't contain $o$. For an execution $e$, $\textit{values}(e)$ is the set of values  in call/return actions of $e$.

The following lemma states the correctness of $\textit{Check-PQ-Seq}$.

\begin{lemma}
\label{lemma:EPQ rules and semantics}
$\textit{Check-PQ-Seq}(e)=\mathsf{true}$ iff $e\in \seqPQ$, for every data-differentiated sequential execution $e$.
\end{lemma}

\subsection{Checking Linearizability of Data-Differentiated Concurrent Executions}\label{ssec:conc_exec}

The extension of $\textit{Check-PQ-Seq}$ to concurrent executions, checking whether they are linearizable w.r.t. $\seqPQ$, is obtained by replacing every predicate $\Gamma\mathsf{\text{-}Seq}$ with
\begin{align*}
\Gamma\mathsf{\text{-}Conc}(e,\alpha) = \mathsf{true}\mbox{ iff there is a sequential execution $s$ such that $e\sqsubseteq s$ and $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$}
\end{align*}
for each $\Gamma\in \{\mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$. The obtained procedure is denoted by 
$\textit{Check-PQ-Conc}$ (recursive calls are modified accordingly).

The following lemma states the correctness of $\textit{Check-PQ-Conc}$. Completeness follows easily from the properties of $\seqPQ$. If $\textit{Check-PQ-Conc}(e) = \textit{false}$, then there exists a set $D$ of values s.t. either  $\mathsf{EmptyRemove\text{-}Conc}(e \vert D)$ is false, or $\mathsf{UnmatchedMaxPriority\text{-}Conc}(e \vert D,x)$ is false for all the values $x$ of maximal priority that are not removed (and there exists at least one such value), or $\mathsf{MatchedMaxPriority\text{-}Conc}(e \vert D,x)$ is false for all the values $x$ of maximal priority (and these values are all removed in $e \vert D$). It can be easily seen that we get $e \vert D\not\sqsubseteq \seqPQ$ in all cases, which by the closure under projection of $\seqPQ$ implies, $e \not\sqsubseteq \seqPQ$ (since every linearization of $e$ includes as a subsequence a linearization of $e \vert D$).

\begin{lemma}\label{lemma:con-check-EPQ is correct}
$\textit{Check-PQ-Conc}(e)=\mathsf{true}$ iff $e \sqsubseteq \seqPQ$, for every data-differentiated $e$.
\end{lemma}

Proving soundness is highly non-trivial and one of the main technical contributions of this paper. 
The main technical difficulty is showing that for any execution $e$, any linearization of $e\setminus x$ for some maximal priority value $x$ can be extended to a linearization of $e$ provided that $\mathsf{UnmatchedMaxPriority}$ or $\mathsf{MatchedMaxPriority}$ holds (depending on whether there are values with the same priority as $x$ in $e$ which are not removed).

We explain the proof of this property on the execution $e$ in \figurename~\ref{fig:concurrent execution for EPQ1}(a) where $p_1 \prec p$, $p_1 \prec p_2$, and the predicate $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds. Assume that there exist two sequential executions $l$ and $l'$ such that $e \sqsubseteq l=u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds, and $e \setminus x \sqsubseteq l' \in \seqPQ$. Let $u=\epsilon$, $w$ be any sequence formed of $\textit{put}(z_2,p_2)$ and $\textit{rm}(z_1)$ (we distinguish them by adding the suffix ``$-w$'' to their name, e.g., $\textit{rm}(z_1)-w$), and $v$ be any sequence containing the remaining operations. In general, the linearization $l'$ can be defined by choosing for each operation, a point in time between its call and return, called \emph{linearization point}. The order between the linearization points defines the sequence $l'$. \figurename~\ref{fig:concurrent execution for EPQ1}(a) draws linearization points for the operations in $e \setminus x$ which define $l'$~\footnote{In general, there may exist multiple ways of choosing linearization points to define the same linearization. Our construction is agnostic to this choice.}.
We show how to construct a sequence $l''= l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3\in\seqPQ$ s.t. $e \sqsubseteq l''$.
\begin{itemize}
\item[-] An operation is called $p$-comparable (resp., $p$-incomparable) when it receives as argument a value of priority comparable to $p$ (resp., incomparable to $p$). Defining $l''_1$, $l''_2$, and $l''_3$ as the projection of $l'$ to the set of operations in $u$, $v$ and $w$, respectively, leads to a sequence $l''\not\in\seqPQ$. This is because $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ imposes no restriction on  $p$-incomparable operations in $u \cdot v$, and the projection of $l'$ to $p$-incomparable operations in $u \cdot v$ is not in $\seqPQ$. In this example, this projection is $\textit{put}(z_1,p_2) \cdot \textit{rm}(z_2)$.

\item[-] We define the sets of operations $U'$, $V'$ and $W'$ such that $l''_1$, $l''_2$ and $l''_3$ are the projections of $l'$ to $U'$, $V'$, and $W'$, respectively. This is done in two steps:
\begin{enumerate}
\item The first step is to define $W'$. The $p$-comparable operations in $W'$ are the same as in $w$. To identify the $p$-incomparable operations in $W'$, we search for a $p$-incomparable operation $o$ which either happens before some $p$-comparable operation in $w$, or whose linearization point occurs after $\textit{ret}(\textit{rm},x)$. We add to $W'$ the operation $o$ and all the $p$-incomparable operations occurring after $o$ in $l'$. In this example, $o$ is $\textit{rm}(z_1)$ and the only $p$-incomparable operation occurring after $o$ in $l'$ is $\textit{rm}(z_2)$ (they are surrounded by boxes in the figure). In this process, whether a $p$-incomparable operation is in $W'$ or not only relies on whether it is before or after such an $o$ in $l'$.
\item The second step is to define $U'$ and $V'$. $U'$ contains two kinds of operations: (1) operations whose linearization points are before $\textit{ret}(\textit{put},x,p)$, and (2) other $\textit{put}$ operations with priority $p$. $V'$ contains the remaining operations. In this example, $U'$ contains $\textit{put}(z_1,p_2)$ and $\textit{put}(x_2,p)$.
\end{enumerate}
\item[-] In conclusion, we have that $l''_1 = \textit{put}(z_1,p_2) \cdot \textit{put}(x_2,p)$, $l''_2 = \textit{put}(z_2,p_2) \cdot \textit{rm}(x_2) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(y_1)$, and $l''_3 = \textit{rm}(z_1) \cdot \textit{rm}(z_2)$. \figurename~\ref{fig:concurrent execution for EPQ1}(b) draws linearization points for each operation in $e$ defining the linearization $l''$.
\end{itemize}

\begin{figure}[t]
  \centering
  \includegraphics[width=.8\textwidth]{PIC-HIS-EPQ1-TwoHis-2.pdf}
  \caption{A concurrent execution $e$ exemplifying the soundness of $\textit{Check-PQ-Conc}$.}
  \label{fig:concurrent execution for EPQ1}
\end{figure}

Section~\ref{sec:co-regular of extended priority queues} introduces a characterization of concurrent priority queue violations using a set of \emph{non-recursive} automata (whose states consist of a fixed number of registers), whose standard synchronized product is equivalent to $\textit{Check-PQ-Conc}$ (modulo a renaming of values which is possible by data-independence). Since $\seqPQ$ is closed under projection (Lemma~\ref{lem:closure_proj}), the recursion in $\textit{Check-PQ-Conc}$ can be eliminated by checking that each projection of a given execution $e$ passes a non-recursive version of $\textit{Check-PQ-Conc}$ where every recursive call $\text{{\bf return}}\ \textit{Check-PQ-Conc}(\ldots)$ is 
replaced by  $\text{{\bf return}}\ \mathsf{true}$. Let $\textit{Check-PQ-Conc-NonRec}$ be the thus obtained procedure.

\begin{lemma}
\label{lemma:EPQ as multi in MRpri for history}
Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$ if and only if for each $e' \in \textit{proj}(e)$, $\textit{Check-PQ-Conc-NonRec}(e')$ returns $\mathsf{true}$.
\end{lemma}

\section{Reducing Linearizability of Priority Queues to Reachability}
\label{sec:co-regular of extended priority queues}

We show that the set of executions for which $\textit{Check-PQ-Conc-NonRec}$ fails on some projection can be described using register automata, modulo a value renaming. Renaming values (which is complete under data independence) allows to simplify the reasoning about projections. W.l.o.g., we assume that all the operations which are not in the projection failing this test use the same distinguished value $\top$, different from those in the projection. Then, it is enough to find an automata characterization of the executions $e$ for which $\textit{Check-PQ-Conc-NonRec}(e)$ is $\mathsf{false}$, i.e., for which 
$\Gamma(e) := \mathsf{Has\text{-}}\Gamma(e) \Rightarrow \exists \alpha.\ \Gamma\mathsf{\text{-}Conc}(e,\alpha)$ is false, for some $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$.
Intuitively, $\Gamma(e)$ states that $e$ is linearizable w.r.t. the set of sequential executions described by $\Gamma\mathsf{\text{-}Seq}$ (provided that $\mathsf{Has\text{-}}\Gamma(e)$ holds). Therefore, by an abuse of terminology, an execution $e$ satisfying $\Gamma(e)$ is called \emph{linearizable w.r.t. $\Gamma$}, or \emph{$\Gamma$-linearizable}.
Extending the automaton characterizing non $\Gamma$-linearizable executions with self-loops that allow any operation with argument $\top$ results in an automaton satisfying the following property called \emph{$\Gamma$-completeness}.

\begin{definition}
For $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$, an automaton $A$ is called \emph{$\Gamma$-complete} when for each data-independent implementation $\mathcal{I}$:

$A \cap \mathcal{I} \neq \emptyset$ iff there exists $e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$ such that $e'$ is not $\Gamma$-linearizable.
\end{definition}

We can show that for any $\Gamma\in \{\mathsf{EmptyRemove}$,$\mathsf{UnmatchedMaxPriority}$,$\mathsf{MatchedMaxPriority}\}$ there exists a $\Gamma$-complete automaton. For lack of space, we only consider the case $\Gamma=\mathsf{MatchedMaxPriority}$ in 
Section~\ref{ssec:aut}.
When defining $\Gamma$-complete automata, we assume that every implementation $\mathcal{I}$ behaves correctly, i.e., as a FIFO queue, when only values with the same priority are observed. More precisely, we assume that for every execution $e\in\mathcal{I}$ and every priority $p\in\mathbb{P}$, the projection of $e$ to values with priority $p$ is linearizable (w.r.t. $\seqPQ$). This property can be checked separately using register automata similar to the automata in~\cite{DBLP:conf/icalp/BouajjaniEEH15} describing FIFO queue violations. 
This assumption excludes some obvious violations, such as an $\textit{rm}(a)$ operation happening before a $\textit{put}(a,p)$ operation, for some $p$.

For $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, we consider $\Gamma$-complete automata recognizing executions which contain only one maximal priority. This is w.l.o.g. because any data-differentiated execution for which $\Gamma(e)$ is false has such a projection.
Formally, given a data-differentiated execution $e$ and $p$ a maximal priority in $e$, $e\vert_{\preceq p}$ is the projection of $e$ to the set of values with priorities smaller or equal to $p$. Then,

\begin{lemma}
\label{lemma:pri execution is enough}
For $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, a data-differentiated execution $e$ 
 is $\Gamma$-linearizable iff $e\vert_{\preceq p}$ is $\Gamma$-linearizable for some maximal priority $p$ in $e$.
\end{lemma}
\begin {proof} (Sketch)
For the ``only-if'' direction, let $e$ be a data-differentiated execution linearizable w.r.t. $l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$ s.t. $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l,x)$ holds. Since the predicate $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l,x)$ imposes no restriction on the operations in $u$, $v$, and $w$ with priorities incomparable to $p$, erasing all these operations results in a sequential execution which still satisfies this predicate. Similarly, for $\Gamma=\mathsf{UnmatchedMaxPriority}$.

The ``if'' direction follows from the fact that if the projection of an execution to a set of operations $O_1$ has a linearization $l_1$ and the projection of the same execution to the remaining set of operations has a linearization $l_2$, then the execution has a linearization which is defined as an interleaving of $l_1$ and $l_2$.

Thus, let $e$ be an execution such that $e\vert_{\preceq p}$ is linearizable w.r.t. $l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$ where $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l,x)$ holds. By the property above, we know that $e$ has a linearization $l' = u' \cdot \textit{put}(x,p) \cdot v' \cdot \textit{rm}(x) \cdot w'$, such that the projection of $l'$ to values of priority comparable to $p$ is $l$.
Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l,x)$ doesn't constrain the values of priority incomparable to $p$, we obtain that $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l',\alpha)$ also holds.
\end {proof}

The existence of $\Gamma$-complete automata enable an effective reduction of checking linearizability of concurrent priority queue implementations to state reachability. 
Section~\ref{subsec:combine step-by-step linearizability and co-regular} discusses decidability results implied by this reduction.

\begin{theorem}
\label{lemma:reduce EPQ into state reachability}
Let $\mathcal{I}$ be a data-independent implementation. Then, there is a $\Gamma$-complete automaton $A(\Gamma)$ for each $\Gamma$. Moreover,
$\mathcal{I} \sqsubseteq \seqPQ$ iff $\mathcal{I} \cap A(\Gamma) = \emptyset$ for all $\Gamma$.
\end{theorem}

\subsection{A $\mathsf{MatchedMaxPriority}$-complete automaton}\label{ssec:aut}

A differentiated execution $e$ is not $\mathsf{MatchedMaxPriority}$-linearizable when all the $\textit{put}$ operations in $e$ using the maximal priority $p$ are matched, and $e$ is not linearizable w.r.t. the set of sequential executions satisfying $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ for each value $x$ of priority $p$. We consider two cases depending on whether $e$ contains exactly one value with priority $p$ or at least two values. We denote by $\mathsf{MatchedMaxPriority}^{>}$ the strengthening of $\mathsf{MatchedMaxPriority}$ with the condition that all the values other than $x$ have a priority strictly smaller than $p$ (corresponding to the first case), and by $\mathsf{MatchedMaxPriority}^{=}$ the strengthening of the same formula with the negation of this condition (corresponding to the second case).
We use particular instances of register automata~\cite{DBLP:journals/tcs/KaminskiF94,DBLP:conf/icalp/Cerans94,DBLP:conf/stacs/SegoufinT11} whose states include only two registers, one for storing a priority guessed at the initial state, and one for storing the priority of the current action in the execution. The transitions can check equality or the order relation $\prec$ between the values stored in the two registers. Instead of formalizing the full class of register automata, we consider a simpler class which suffices our needs. Thus, we consider a class of labeled transition systems whose states consist of a finite control part and a register $r$ interpreted to elements of $\mathbb{P}$. The transition labels are:
\begin{itemize}
	\item $r=*$ for storing an arbitrary value to $r$,
	\item $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$ for reading call/return actions of a remove,
	\item $\textit{call}(\textit{put},d,g)$ where $g\in\{=r,\prec r,true\}$ is a guard, for reading a call action $\textit{call}(\textit{put},d,p)$ and checking if $p$ is either equal to or smaller than the value stored in $r$, or arbitrary,
	\item $\textit{ret}(\textit{put},d,true)$ for reading a return action $\textit{ret}(\textit{put},d,p)$ for any $p$.
\end{itemize}
The set of sequences (executions) accepted by such a transition system is defined as usual.

\subsubsection{A $\mathsf{MatchedMaxPriority}^>$-complete automaton}
\label{subsec:co-regular of EPQ1Lar}

\begin{figure}[t]
  \centering
  \includegraphics[width=0.4 \textwidth]{PIC-HIS-INTRO-GAP-EPQ1L.pdf}
  \caption{An execution that is not $\mathsf{MatchedMaxPriority}^{>}$-linearizable. We represent each operation as a time interval whose left, resp., right, bound corresponds to the call, resp., return action.} 
  \label{fig:introduce gap for EPQ1Lar}
\end{figure}

Figure~\ref{fig:introduce gap for EPQ1Lar} contains a typical example of an execution $e$ which is not $\mathsf{MatchedMaxPriority}^>$-linearizable,
where $p_1 \prec p_4$, $p_2 \prec p_4$, and $p_3 \prec p_4$.
Intuitively, this is a violation because during the whole execution of $\textit{rm}(b)$, the priority queue stores a smaller priority value (which should be removed before $b$). To be more precise, we define \emph{the interval of a value $x$} as the time interval from the return of a put $\textit{ret}(\textit{put},x,p)$ to the call of the matching remove $\textit{call}(rm,x)$, or to the end of the execution if such a call action doesn't exist. This represents the time interval in which a value is guaranteed to be stored into the priority queue. Concretely, for a standard indexing of actions in an execution, a time interval is a closed interval between the indexes of two actions in the execution.
In \figurename~\ref{fig:introduce gap for EPQ1Lar}, the interval of each value of priority smaller than $p_4$ is pictured as a dashed line. There is no sequence $l$ s.t. $e \sqsubseteq l$ and $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l,b)$ hold, since each time point from $\textit{call}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is included in the interval of a smaller priority value, 
and $\textit{rm}(b)$ can't take effect in the interval of a smaller priority value.
To formalize this scenario we use the notion of \emph{left-right constraint} defined below.

\begin{definition}\label{def:left-right constraint for matched put and rm operations}
Let $e$ be a data-differentiated execution which contains only one maximal priority $p$, and only one value $x$ of priority $p$ (and no $\textit{rm}(\textit{empty})$ operations).
The \emph{left-right constraint of $x$} is the graph $G$ where: 
\begin{itemize}
\item the nodes are the values occurring in $e$, 
\item there is an edge from $d_1$ to $x$, if $\textit{put}(d_1,\_) <_{\textit{hb}} \textit{put}(x,p)$ or $\textit{put}(d_1,\_) <_{\textit{hb}} \textit{rm}(x)$,
\item there is an edge from $x$ to $d_1$, if $\textit{rm}(x)<_{\textit{hb}}\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists,
\item there is an edge from $d_1$ to $d_2$, if $\textit{put}(d_1,\_) <_{\textit{hb}} \textit{rm}(d_2,\_)$.
\end{itemize}
\end{definition}

The execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} is not $\mathsf{MatchedMaxPriority}^>$-linearizable because the left-right constraint of the maximal priority value $b$ contains the cycle $f \rightarrow d \rightarrow c \rightarrow b \rightarrow f$. The presence of such a cycle is equivalent to \emph{non} $\mathsf{MatchedMaxPriority}^>$-linearizability:

\begin{lemma}
\label{lemma:Lin Equals Constraint for EPQ1Lar}
Let $e$ be a data-differentiated execution such that 
$\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, $p$ is the maximal priority in $e$, and $\textit{put}(x,p)$ and $\textit{rm}(x)$ are only operations with arguments of priority $p$ in $e$. 
Then, $e$ is $\mathsf{MatchedMaxPriority}$-linearizable iff the left-right constraint of $x$ contains no cycle going through $x$.
\end{lemma}

When the left-right constraint contains a cycle $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$, for some $d_1$,$\ldots$,$d_n\in \mathbb{D}$, we say that $x$ is \emph{covered} by $d_1,\ldots,d_m$. The shape of an execution witnessing such a cycle (i.e., the alternation between call/return actions) can be identified using our class of automata, the only complication being the unbounded number of values $d_1$,$\ldots$,$d_n$. However, by data independence, whenever an implementation contains such an execution it also contains an execution where all the values $d_1$,$\ldots$,$d_n$ are renamed to the same value $a$, and $x$ is renamed to $b$. Therefore, our automata can be defined over a fixed set of values $a$, $b$, and $\top$ (recall that $\top$ is used for operations outside of the non-linearizable projection).

To define a $\mathsf{MatchedMaxPriority}^>$-complete automaton, we need to consider all the possible orders between the call/return actions of the $\textit{put}$/$\textit{rm}$ operations that add and respectively, remove the value $b$. The case where the put happens-before the remove (as in \figurename~\ref{fig:introduce gap for EPQ1Lar}) is pictured in \figurename~\ref{fig:automata APQ1Lar-1 in paper}. This automaton captures the three possible ways of ordering the first action $\textit{ret}(\textit{put},a,\_)$ w.r.t. the actions with value $b$, which are pictured in \figurename~\ref{fig:executions APQ1Lar-1 in paper}(a) (this action cannot occur after $\textit{call}(\textit{rm},b,\_)$ since $b$ must be covered by the $a$-s). The paths corresponding to these three possible orders are: $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_{10}$, and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$. \figurename~\ref{fig:executions APQ1Lar-1 in paper} lists the four possible orderings of the call/return actions of adding and removing $b$, and also possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$. Each such ordering corresponds to an automaton similar to the one in \figurename~\ref{fig:automata APQ1Lar-1 in paper}, their union defining a $\mathsf{MatchedMaxPriority}^>$-complete automaton.

\begin{figure}[t]
  \centering
  \includegraphics[width=.8\textwidth]{PIC_HIS_PQ1Lar-fouCase.pdf}
  \caption{Orderings to be considered when defining a $\mathsf{MatchedMaxPriority}^>$-complete automaton.}
  \label{fig:executions APQ1Lar-1 in paper}
\end{figure}

\begin{figure}[t]
  \centering
  \includegraphics[width=.9\textwidth]{PIC_AUTO_PQ1Lar-pprr.pdf}
  \caption{A register automaton capturing the scenario in Figure~\ref{fig:executions APQ1Lar-1 in paper}(a). We use the following notations: $A_1 = A \cup \{ \textit{ret}(\textit{rm},a) \}$, $A_2 = A \cup \{ \textit{call}(\textit{put},a,=r) \}$, $A_3 = A_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $A = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},\top)$, $\textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.}
  \label{fig:automata APQ1Lar-1 in paper}
\end{figure}

\subsubsection{A $\mathsf{MatchedMaxPriority}^=$-complete automaton}
\label{subsec:co-regular of EPQ1Equal}

When an execution contains at least two values of maximal priority, the acyclicity of the left-right constraints (for all the maximal priority values) is not enough to conclude that the execution is $\mathsf{MatchedMaxPriority}$-linearizable.
Intuitively, there may exist a value $a$ which is added before another value $b$ such that all the possible linearization points of $\textit{rm}(b)$ are disabled by the position of $\textit{rm}(a)$ in the happens-before. We give an example of such an execution $e$ in \figurename~\ref{fig:introduce pb order}, where $p_1\prec p_4$. This execution  is not linearizable w.r.t. $\mathsf{MatchedMaxPriority}$ (or $\mathsf{MatchedMaxPriority}^{=}$) even if
neither $a$ nor $b$ are covered by values with smaller priority. 
Since $\textit{put}(a,p_4) <_{\textit{hb}} \textit{put}(b,p_4)$ and values of the same priority are removed in FIFO order, $\textit{rm}(a)$ should be linearized before $\textit{rm}(b)$ (i.e., this execution should be linearizable w.r.t. a sequence where $\textit{rm}(a)$ occurs before $\textit{rm}(b)$).
Since $\textit{rm}(b)$ cannot take effect during the interval of a smaller priority value, it could be only linearized in one of the two time intervals pictured with dotted lines in \figurename~\ref{fig:introduce pb order}. However, each of those time intervals ends before $\textit{call}(\textit{rm},a)$, and thus $\textit{rm}(a)$ cannot be linearized before $\textit{rm}(b)$.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.5 \textwidth]{PIC-HIS-INTRO-PB-ORDER-EPQ.pdf}
  \caption{An execution that is not $\mathsf{MatchedMaxPriority}^=$-linearizable.}
  \label{fig:introduce pb order}
\end{figure}

To recognize the scenarios in \figurename~\ref{fig:introduce pb order}, we introduce an order $<_{\textit{pb}}$ between values which intuitively, can be thought of as ``a value $a$ is put before another value $b$''. 
More precisely, given a data-differentiated execution $e$ and two values $a$ and $b$ of maximal priority, $a <_{\textit{pb}} b$ if one of the following holds: (1) $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$, (2) $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(b)$, or (3) $\textit{rm}(a) <_{\textit{hb}} \textit{put}(b,\_)$. Sometimes we use $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$, and $a <_{\textit{pb}}^C b$ to explicitly distinguish between these three cases. Let $<_{\textit{pb}}^*$ be the transitive closure of $<_{\textit{pb}}$.

To define the time intervals in which a remove like $\textit{rm}(b)$ in \figurename~\ref{fig:introduce pb order}, can be linearized (outside of intervals of smaller priority values) we use the notion of gap-point. As before, defining time intervals relies on an indexing of actions in an execution, starting with 0.

\begin{definition}\label{def:gap-point for matched put and rm operations}
Let $e$ be a data-differentiated execution with only one maximal priority $p$, and $\textit{put}(x,p)$ and $\textit{rm}(x)$ two operations in $e$. An index $i\in [0,|e|-1]$ is a \emph{gap-point of $x$} if $i$ is greater than or equal to the index of both $\textit{call}(\textit{put},x,p)$ and $\textit{call}(\textit{rm},x)$, smaller than the index of $\textit{ret}(\textit{rm},x)$, and not included in the interval of a value with priority smaller than $p$.
\end{definition}

The case of \figurename~\ref{fig:introduce pb order} can be formally described as follows: $a <_{\textit{pb}}^* b$ while the right-most gap-point of $b$ is before $\textit{call}(\textit{rm},a)$ or $\textit{call}(\textit{put},a,p_4)$. The following lemma states that these conditions are enough to characterize non-linearizability w.r.t. $\mathsf{MatchedMaxPriority}^{=}$.

\begin{lemma}
\label{lemma:EPQ1Equal as pb order and gap-point}
Let $e$ be a data-differentiated execution with only one maximal priority $p$ such that $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds.
Then, $e$ is not $\mathsf{MatchedMaxPriority}^{=}$-linearizable iff $e$ contains two values $x$ and $y$ of maximal priority $p$ such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is strictly smaller than the index of $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$.
\end{lemma}

The following shows that the number of values needed to witness that $y <_{\textit{pb}}^* x$, for some $x$ and $y$, is bounded.

\begin{lemma}
\label{lemma:ob order has bounded length}
Let $e$ be a data-differentiated execution such that $a <_{\textit{pb}} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}} a_m <_{\textit{pb}} b$ holds for some set of values $a$, $a_1$,$\ldots$,$a_m$, $b$. Then, one of the following holds:
\begin{itemize}
\item[-] $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$, or $a <_{\textit{pb}}^C b$,

\item[-] $a <_{\textit{pb}}^A a_i <_{\textit{pb}}^B b$ or $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A b$, for some $i$.
\end{itemize}
\end{lemma}

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8 \textwidth]{PIC-HIS-FiveEnumerations.pdf}
  \caption{Orderings to be considered when defining a $\mathsf{MatchedMaxPriority}^=$-complete automaton. We omit the operations which can be ordered arbitrarily, e.g., $\textit{put}(b)$ in the cases (d) and (e).}
  \label{fig:five enumerations}
\end{figure}

To characterize violations to $\mathsf{MatchedMaxPriority}^{=}$-linearizability, one has to consider all the possible orders between call/return actions of the operations on values $a$, $b$, and $a_i$ in Lemma \ref{lemma:ob order has bounded length}, and the right-most gap point of $b$. Excluding the inconsistent cases, we are left with the five orders in \figurename~\ref{fig:five enumerations}, where $o$ denotes the rightmost gap-point of $b$.
For each case, we define an automaton recognizing the induced set of violations. The register automaton for the case in \figurename~\ref{fig:five enumerations}(a) is shown in \figurename~\ref{fig:an enumeration and its witness automaton}. In this case, Lemma~\ref{lemma:EPQ1Equal as pb order and gap-point} is equivalent to the fact that intuitively, the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$ is covered by lower priority values (thus, there is no gap-point of $b$ which occurs after $\textit{call}(\textit{rm},a)$). By data-independence, these lower priority values can be renamed to a fixed value $c$.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.65 \textwidth]{PIC-WitnessAutomata-For1.pdf}
  \caption{A register automaton for the case in Figure~\ref{fig:five enumerations}(a), where $A_1 = A \cup \{ \textit{call}(\textit{put},c,<r)\}$, $A_2 = A_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $A_3 = A_2 \cup \{ \textit{ret}(\textit{rm},c) \}$, $A_4 = A \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},c) \}$, where $A = \{ \textit{call}(\textit{put},\top,\textit{true})$,$\textit{ret}(\textit{put},\top,\textit{true})$,$\textit{call}(\textit{rm},\top)$,$\textit{ret}(\textit{rm},\top)$,$\textit{call}(\textit{rm},\textit{empty})$,$\textit{ret}(\textit{rm},\textit{empty})\}$.}
  \label{fig:an enumeration and its witness automaton}
\end{figure}

\subsection{Decidability Result}
\label{subsec:combine step-by-step linearizability and co-regular}

We describe a class $\mathcal{C}$ of data-independent implementations for which linearizability w.r.t. $\seqPQ$ is decidable. The implementations in $\mathcal{C}$ allow an unbounded number of values but a bounded number of priorities. Each method has a finite set of local variables storing Boolean values or values from $\mathbb{D}$. Methods communicate through a finite number of shared variables interpreted also as Booleans or values from $\mathbb{D}$. To ensure data independence, values in $\mathbb{D}$ may be assigned, but never used in Boolean expressions (e.g., of if-then-else statements). This class captures typical implementations, or finite-state abstractions thereof, e.g., obtained via predicate abstraction. The $\Gamma$-complete automata we define use a fixed set $D=\{a,b,c,a_1,\top\}$ of values ($a_1$ is needed to deal with the second item in Lemma~\ref{lemma:ob order has bounded length}). Therefore, for any $\Gamma$, $\mathcal{C}\cap A(\Gamma)\neq\emptyset$ iff $\mathcal{C}_D\cap A(\Gamma)\neq\emptyset$, where $\mathcal{C}_D$ is the subset of $\mathcal{C}$ that uses only values in $D$.

The set of executions $\mathcal{C}_D$ can be represented by a Vector Addition System with States (VASS). Since the set of values and priorities is bounded, each method invocation can be represented by a finite-state automaton (see~\cite{conf/esop/BouajjaniEEH13}). For a fixed set of priorities $P\subseteq \mathbb{P}$, the register automata $A(\Gamma)$ can be transformed to finite-state automata (the number of valuations of the registers is bounded). Thus, checking linearizability of an implementation in $\mathcal{C}$ is PSPACE when the number of threads is bounded, and EXPSPACE, otherwise. Moreover, reachability in VASSs can be reduced to checking linearizability of such an implementation. Essentially, given an instance of the VASS reachability problem, one can define a priority queue implementation where the $\textit{put}$ methods behave correctly and additionally, they include the code of the VASS simulation defined in~\cite{conf/esop/BouajjaniEEH13}, and the $\textit{rm}$ methods behave correctly, except for the moment where the target state is reached, in which case they trigger a linearizability violation by returning an arbitrary value.

\begin{theorem}
\label{theorem:complexity of priority queue}
Verifying whether an implementation in $\mathcal{C}$ is linearizable w.r.t. $\seqPQ$ is PSPACE-complete for a fixed number of threads, and EXPSPACE-complete otherwise.
\end{theorem}

\section{Related work}\label{sec:related}

The theoretical limits of checking linearizability have been investigated in previous works. 
Checking linearizability of a single execution w.r.t. an arbitrary ADT is NP-complete~\cite{journals/siamcomp/GibbonsK97} while checking linearizability of all the executions  
of a finite-state implementation w.r.t. an arbitrary ADT 
specification (given as a regular language) is EXPSPACE-complete when the number of program 
threads is bounded~\cite{journals/iandc/AlurMP00,netys-lin}, and
undecidable otherwise~\cite{conf/esop/BouajjaniEEH13}.

Existing automated methods for proving linearizability of a concurrent object
 are also based on reductions to safety
verification, e.g.,~\cite{conf/tacas/AbdullaHHJR13, conf/concur/HenzingerSV13,
conf/cav/Vafeiadis10}. The approach in~\cite{conf/cav/Vafeiadis10} considers
implementations where 
operations' \emph{linearization points}
are 
manually specified.
Essentially, this approach instruments the
implementation with ghost variables simulating the ADT specification at
linearization points. This approach is incomplete since not all implementations
have fixed linearization points. Aspect-oriented
proofs~\cite{conf/concur/HenzingerSV13} reduce linearizability to the
verification of four simpler safety properties. This approach has only
been applied to queues, and has not produced a fully automated
and complete proof technique. The work in~\cite{Dodds:2015:SCT:2676726.2676963} proves 
linearizability of stack implementations with an automated proof assistant. 
Their approach does not lead to full automation however, e.g.,~by reduction to 
safety verification.

Our previous work~\cite{DBLP:conf/icalp/BouajjaniEEH15}
shows that checking linearizability of finite-state implementations of concurrent queues and stacks is decidable.
Roughly, we follow the same schema: the recursive procedure in Section~\ref{ssec:seq_exec} is similar to the inductive rules in~\cite{DBLP:conf/icalp/BouajjaniEEH15}, and its extension to concurrent executions in Section~\ref{ssec:conc_exec} corresponds to the notion of step-by-step linearizability in~\cite{DBLP:conf/icalp/BouajjaniEEH15}. Although similar in nature, defining these procedures and establishing their correctness require proof techniques which are specific to the priority queue semantics. The order in which values are removed from a priority queue is encoded in their priorities which come from an unbounded domain, and not in the happens-before order as in the case of stacks and queues. Therefore, the results we introduce in this paper cannot be inferred from those in~\cite{DBLP:conf/icalp/BouajjaniEEH15}. At a technical level, characterizing the priority queue violations requires a more expressive class of automata (with registers) than the finite-state automata in~\cite{DBLP:conf/icalp/BouajjaniEEH15}.

\newpage

  \bibliography{violin}
\bibliographystyle{plainurl}

\newpage

\appendix

\end{document}

### 2017-07-13 appendix-EPQ.tex

@@
 We also require that $\mathsf{Has\text{-}MatchedMaxPriority}(e')$ holds and there exists a cycle going through the value with maximal priority in $e'$.
-By data-independence, we can obtain $e_r$ from $e$ by renaming function, which maps such value to be $b$, maps values that cover it to be $a$, and maps other values into $d$.
+By data-independence, we can obtain $e_r$ from $e$ by renaming function, which maps such value to be $b$, maps values that cover it to be $a$, and maps other values into $\top$.
 There are four possible enumeration of call and return actions of $\textit{put}(b,p)$ and $\textit{rm}(b)$.
@@
 
-For the case when $e_r \vert_{b} = \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{put},b,p) \cdot \textit{call}(\textit{rm},b) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^1$, as shown in \figurename~\ref{fig:automata APQ1Lar-1}.
-Here $C_1 = C \cup \{ \textit{ret}(\textit{rm},a) \}$, $C_2 = C \cup \{ \textit{call}(\textit{put},a,<r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},\top)$, $\textit{ret}(\textit{rm},\top),$ $\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.
+For the case when $e_r \vert_{b} = \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{put},b,\textit{true}) \cdot \textit{call}(\textit{rm},b) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^1$, as shown in \figurename~\ref{fig:automata APQ1Lar-1}.
+Here $A_1 = A \cup \{ \textit{ret}(\textit{rm},a) \}$, $A_2 = A \cup \{ \textit{call}(\textit{put},a,<r) \}$, $A_3 = A_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $A = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},\top)$, $\textit{ret}(\textit{rm},\top),$ $\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.
 The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^1$ comes from the positions of the first $\textit{ret}(\textit{put},a,\_)$.
@@
 
-For the case when $e_r \vert_{b} = \textit{call}(\textit{put},b,p) \cdot \textit{call}(\textit{rm},b) \cdot \textit{ret}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^2$, as shown in \figurename~\ref{fig:automata APQ1Lar-2}.
-Here $C_1,C_2,C_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$.
+For the case when $e_r \vert_{b} = \textit{call}(\textit{put},b,p) \cdot \textit{call}(\textit{rm},b) \cdot \textit{ret}(\textit{put},b,\textit{true}) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^2$, as shown in \figurename~\ref{fig:automata APQ1Lar-2}.
+Here $A_1,A_2,A_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$.
 The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^2$ comes from the positions of the first $\textit{ret}(\textit{put},a,\_)$.
@@
 
-For the case when $e_r \vert_{b} = \textit{call}(\textit{rm},b) \cdot \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^3$, as shown in \figurename~\ref{fig:automata APQ1Lar-3}.
-Here $C_1,C_2,C_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$.
+For the case when $e_r \vert_{b} = \textit{call}(\textit{rm},b) \cdot \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{put},b,\textit{true}) \cdot \textit{ret}(\textit{rm},b)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^3$, as shown in \figurename~\ref{fig:automata APQ1Lar-3}.
+Here $A_1,A_2,A_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$.
 The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^3$ comes from the positions of the first $\textit{ret}(\textit{put},a,\_)$.
@@
 
-For the case when $e_r \vert_{b} = \textit{call}(\textit{rm},b) \cdot \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b) \cdot \textit{ret}(\textit{put},b,p)$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^4$, as shown in \figurename~\ref{fig:automata APQ1Lar-4}.
-Here $C_1,C_2,C_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$, and $C_4 = C_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$.
+For the case when $e_r \vert_{b} = \textit{call}(\textit{rm},b) \cdot \textit{call}(\textit{put},b,p) \cdot \textit{ret}(\textit{rm},b) \cdot \textit{ret}(\textit{put},b,\textit{true})$, we generate register automaton $\mathcal{A}_{\textit{l-lar}}^4$, as shown in \figurename~\ref{fig:automata APQ1Lar-4}.
+Here $A_1,A_2,A_3$ is the same as that in $\mathcal{A}_{\textit{l-lar}}^1$, and $A_4 = A_1 \cup \{ \textit{ret}(\textit{put},b,\textit{true}) \}$.
 The differentiated branch in $\mathcal{A}_{\textit{l-lar}}^4$ comes from the positions of the first $\textit{ret}(\textit{put},a,\_)$.
@@
 \noindent The $\textit{if}$ direction: Given such $e$, $e'$ and $x$.
-Let renaming function $r$ maps $x$ into $b$, maps values cover $x$ into $a$, and maps other values into $d$.
+Let renaming function $r$ maps $x$ into $b$, maps values cover $x$ into $a$, and maps other values into $\top$.
 By data-independence, $r(e) \in \mathcal{I}$.
@@
 An automaton $\mathcal{A}_{\textit{l-eq}}^1$ is given in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}, and it is constructed for the first enumeration in \figurename~\ref{fig:history enumeration 1 for PQ1Equal}.
-Here we rename the values that cover $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ into $d$, and rename the remanning values into $\top$.
-In this figure, $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true})$, $\textit{call}(\textit{rm},\top), \textit{ret}(\textit{rm},\top),$ $\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_1 = C \cup \{ \textit{call}(\textit{put},d,<r) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $C_3 = C_2 \cup \{ \textit{call}(\textit{put},d,<r),\textit{ret}(\textit{rm},d) \}$, $C_4 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},d) \}$.
+Here we rename the values that cover $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ into $c$, and rename the remanning values into $\top$.
+In this figure, $A = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true})$, $\textit{call}(\textit{rm},\top), \textit{ret}(\textit{rm},\top),$ $\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $A_1 = A \cup \{ \textit{call}(\textit{put},c,<r) \}$, $A_2 = A_1 \cup \{ \textit{ret}(\textit{put},b,\textit{true}) \}$, $A_3 = A_2 \cup \{ \textit{ret}(\textit{rm},c) \}$, $A_4 = A \cup \{ \textit{ret}(\textit{put},b,\textit{true}), \textit{ret}(\textit{rm},c) \}$.
 
@@
 An automaton $\mathcal{A}_{\textit{l-eq}}^2$ is given in \figurename~\ref{fig:automata for second enumeration of PQ1Equal}, and it is constructed for the second enumeration in \figurename~\ref{fig:history enumeration 2 for PQ1Equal}.
-In \figurename~\ref{fig:automata for second enumeration of PQ1Equal}, $C_1$, $C_2$, $C_3$ and $C_4$ is same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}.
+In \figurename~\ref{fig:automata for second enumeration of PQ1Equal}, $A_1$, $A_2$, $A_3$ and $A_4$ is same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}.
 
@@
 For example, for the case $\textit{ca}_1$ when $\textit{call}(\textit{put},a,p)$ is before $\textit{call}(\textit{rm},a_1)$, $\textit{call}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$, and $\textit{call}(\textit{put},b,p)$ is before $\textit{call}(\textit{rm},b)$, we construct a finite automaton $\mathcal{A}_{\textit{l-eq}}^{\textit{3-1}}$ in \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}.
-In \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}, let $C$ and $C_1$ the same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}.
-Let $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},a_1,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $C_4 = C_3 \cup \{ \textit{call}(\textit{put},d,<r), \textit{ret}(\textit{rm},d) \}$, and $C_5 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{put},a_1,=r), \textit{ret}(\textit{rm},d) \}$.
+In \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}, let $A$ and $A_1$ the same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}.
+Let $A_2 = A_1 \cup \{ \textit{ret}(\textit{put},a_1,\textit{true}) \}$, $A_3 = A_2 \cup \{ \textit{ret}(\textit{put},b,\textit{true}) \}$, $A_4 = A_3 \cup \{ \textit{ret}(\textit{rm},c) \}$, and $A_5 = A \cup \{ \textit{ret}(\textit{put},b,\textit{true}), \textit{ret}(\textit{put},a_1,\textit{true}), \textit{ret}(\textit{rm},c) \}$.
 Other register automata can be similarly constructed.
@@
 By data-independence, there exists data-differentiated execution $e_2 \in \mathcal{I}$ and a renaming function $r$, such that $e_1=r(e_2)$.
-Since $e_1$ is accepted by some register automata in  $\mathcal{A}_{\textit{1-eq}}$, let $x$, $y$ and $z$ (if exists) be the values that are renamed into $b$, $a$ and $a_1$ (if exists) by $r$, respectively, and let $d_1,\ldots,d_m$ be the values that are renamed into $d$ by $r$.
+Since $e_1$ is accepted by some register automata in  $\mathcal{A}_{\textit{1-eq}}$, let $x$, $y$ and $z$ (if exists) be the values that are renamed into $b$, $a$ and $a_1$ (if exists) by $r$, respectively, and let $d_1,\ldots,d_m$ be the values that are renamed into $c$ by $r$.
 
@@
 \noindent The $\textit{if}$ direction: Assume that there exists $e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e')$ holds $e'$ has only one maximal priority, $a'$ and $b'$ are two values with maximal priority $p$ in $e'$, $a' <_{\textit{pb}}^* b'$ in $e'$, and the rightmost gap-point of $b'$ is before $\textit{call}(\textit{put},a',p)$ or $\textit{call}(\textit{rm},a')$ in $e'$.
-By data-independence, we can obtain execution $e_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the values $d_1,\ldots,d_m$ that constitute the rightmost gap of $b'$, we rename them into $d$, (3) if $a' <_{\textit{pb}}^A a'_1 <_{\textit{pb}}^B b$, we rename $a'_1$ into $a_1$, and (4) rename the other values into $\top$.
+By data-independence, we can obtain execution $e_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the values $d_1,\ldots,d_m$ that constitute the rightmost gap of $b'$, we rename them into $c$, (3) if $a' <_{\textit{pb}}^A a'_1 <_{\textit{pb}}^B b'$, we rename $a'_1$ into $a_1$, and (4) rename the other values into $\top$.
 It is easy to see that $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e_1)$ holds, $a$ and $b$ has maximal priority in $e_1$, $a <_{\textit{pb}}^* b$ in $e_1$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$ in $e_1$.
@@
 The automaton is $\mathcal{A}_{\seqPQ}^3$, which is given in \figurename~\ref{fig:automata for PQ3}.
-In \figurename~\ref{fig:automata for PQ3}, let $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}($ $\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},\top), \textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_1 = C \cup \{ \textit{call}(\textit{put},b,\textit{true}) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, and $C_3 = C \cup \{ \textit{ret}(\textit{rm},b) \}$.
+In \figurename~\ref{fig:automata for PQ3}, let $A = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}($ $\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},\top), \textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $A_1 = A \cup \{ \textit{call}(\textit{put},b,\textit{true}) \}$, $C_2 = A_1 \cup \{ \textit{ret}(\textit{rm},b) \}$, and $A_3 = A \cup \{ \textit{ret}(\textit{rm},b) \}$.
 
@@
 \noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$.
-$\mathcal{A}_{\seqPQ}^3 \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, $\mathsf{Has\text{-}EmptyRemove}(e')$ holds, $o = \textit{rm}(\textit{empty})$ is in $e'$, and $o$ is covered by some values $d_1,\ldots,d_m$ in $e'$.
+$\mathcal{A}_{\seqPQ}^3 \cap \mathcal{I} \neq \emptyset$, if and only if there exists $o = \textit{rm}(\textit{empty})$ in $e \in \mathcal{I}$, $G$ has a cycle going through $o$, where $G$ is the graph representing the left-right constraint of $o$.
 
@@
 By data-independence, there exists data-differentiated execution $e_2 \in \mathcal{I}$ and a renaming function $r$, such that $e_1=r(e_2)$.
-Let $d_1,\ldots,d_m$ be the values in $e_2$ such that $r(d_i)=b$ for each $1 \leq i \leq m$.
-Let $e_3 = e_2 \vert_{ \{ o, d_1, \ldots, d_m \} }$.
-It is obvious that $e_3 \in \textit{proj}(e_2)$ and $\textit{last}(e_3) = \seqPQ_3$.
-It is easy to see that $o$ is covered by $d_1,\ldots,d_m$.
-
-\noindent The $\textit{if}$ direction: Assume that there exists such $e$, $e'$, $o$ and $d_1,\ldots,d_m$.
-Then, let $e_1$ be obtained from $e$ by renaming $d_1,\ldots,d_m$ into $b$ and renaming other values into $\top$.
-By data-independence, $e_1 \in \mathcal{I}$.
-It is easy to see that $e_1$ is accepted by $\mathcal{A}_{\seqPQ}^3$.
+Since $e_1$ is accepted by $\mathcal{A}_{\seqPQ}^3$, there exists an operation $o=\textit{rm}(\textit{empty})$ in $e_1$ and $e_2$, and let $d_1,\ldots,d_m$ be the values that are renamed into $b$ by $r$.
+Let $e'' = e_2 \vert_{ \{ o,d_1,\ldots,d_m \} }$.
+It is not hard to see that $G$ has a cycle going through $o$, where $G$ is the graph representing the left-right constraint of $o$.
+
+\noindent The $\textit{if}$ direction: Given such $e$ and $o$.
+By data-independence, there exists data-differentiated execution $e_1 \in \mathcal{I}$ and a renaming function $r_1$, such that $e=r_1(e_1)$.
+It is not hard to see that in $e_1$, $o$ is covered by values $d_1,\ldots,d_m$.
+Let renaming function $r_2$ be a renaming function which maps $d_1,\ldots,d_m$ into $b$ and all other values into $\top$.
+Then it is easy to see that $r_2(e_1) \in \mathcal{I}$ is accepted by $\mathcal{A}_{\seqPQ}^3$.
 
@@
 When considering decidability and complexity, we consider only finite variables with finite data domain.
-In such case, the priority queue implementations can only model priority queue with bounded capacity.
+In such case, it seems that the priority queue implementations can only model priority queue with bounded capacity.
 Then, at some time, the priority queue may become full and then put can not succeed until some item is removed.
@@
 To fit our results of priority queue in former section, when checking linearizability of priority queue, we only consider executions without any $\textit{put}(\_,\_,\textit{full})$.
-Since does not mean we consider only executions with bounded number of operations, since an execution can contain unbounded number of pairs of put and remove operations by linearizable w.r.t pairs of put and remove operations.
+This does not mean we consider only executions with bounded number of operations, since an execution can contain unbounded number of pairs of put and remove operations by linearizable w.r.t pairs of put and remove operations.
 
@@
 
-\noindent {\bf Reducing priority Queue into PSPACE and EXPSPACE problems:} We describe a class $\mathcal{C}$ of data-independent implementations for which linearizability w.r.t.
+\noindent {\bf Reducing Linearizability Problem of Priority Queue into State-Reachability Problem of VASS:} We describe a class $\mathcal{C}$ of data-independent implementations for which linearizability w.r.t.
 $\seqPQ$ is decidable.
 The implementations in $\mathcal{C}$ allow an unbounded number of values but a bounded number of priorities.
-Each method manipulates a finite number of local variables which store Boolean values, or data values from $\mathbb{D}$.
-Methods communicate through a finite number of shared variables that also store Boolean values, or data values from $\mathbb{D}$.
-Data values may be assigned, but never used in program predicates (e.g., in the conditions of if-then-else and while statements) so as to ensure data independence.
+Each method has a finite set of local variables storing Boolean values or values from $\mathbb{D}$.
+Methods communicate through a finite number of shared variables interpreted also as Booleans or values from $\mathbb{D}$.
+To ensure data independence, values in $\mathbb{D}$ may be assigned, but never used in Boolean expressions (e.g., of if-then-else statements).
 This class captures typical implementations, or finite-state abstractions thereof, e.g., obtained via predicate abstraction.
-Since the $\Gamma$-complete automata $A(\Gamma)$ uses a fixed set $D=\{a,b,a_1,d,e,\top\}$ of values, we have that $\mathcal{C}\cap A(\Gamma)\neq\emptyset$ for some $\Gamma$ iff $\mathcal{C}_D\cap A(\Gamma)\neq\emptyset$ where $\mathcal{C}_D$ is the subset of $\mathcal{C}$ that uses only values in $D$.
-
-We model the set $\mathcal{C}_D$ as the executions of a Vector Addition System with States (VASS) similarly to Bouajjani et al.
+The $\Gamma$-complete automata we define use a fixed set $D=\{a,b,c,a_1,\top\}$ of values ($a_1$ is needed to deal with the second item in Lemma~\ref{lemma:ob order has bounded length}).
+Therefore, for any $\Gamma$, $\mathcal{C}\cap A(\Gamma)\neq\emptyset$ iff $\mathcal{C}_D\cap A(\Gamma)\neq\emptyset$, where $\mathcal{C}_D$ is the subset of $\mathcal{C}$ that uses only values in $D$.
+
+The set of executions $\mathcal{C}_D$ can be represented by a Vector Addition System with States (VASS).
+Since the set of values and priorities is bounded, each method invocation can be represented by a finite-state automaton (see~\cite{conf/esop/BouajjaniEEH13}).
+For a fixed set of priorities $P\subseteq \mathbb{P}$, the register automata $A(\Gamma)$ can be transformed to finite-state automata (the number of valuations of the registers is bounded).
+Thus, checking linearizability of an implementation in $\mathcal{C}$ is PSPACE when the number of threads is bounded, and EXPSPACE, otherwise.
+
+%The set of executions $\mathcal{C}_D$ can be represented by a Vector Addition System with States (VASS).
+Since the set of values and priorities is bounded, each method invocation can be represented by a finite-state automaton (see~\cite{conf/esop/BouajjaniEEH13}).
+For a fixed set of priorities $P\subseteq \mathbb{P}$, the register automata $A(\Gamma)$ can be transformed to finite-state automata (the number of valuations of the registers is bounded).
+A VASS has a finite set of states, and manipulates a finite set of counters holding non-negative values.
+The idea is that the states of the VASS is a product of the valuation of global variables of $\mathcal{C}_D$ and state of $A(\Gamma)$.
+This set is finite.
+Each counter of the VASS then represents the number of threads which are at a particular control location within a method, with a certain valuation of the local variables.
+Here again, as the data values are bounded, there is a finite number of valuations.
+When a thread moves from a control location to another, we decrement, resp., increment, the counter corresponding the old, resp., the new control location.
+In this way, we model $\mathcal{C}_D$ as VASS.
+Thus, checking linearizability of an implementation in $\mathcal{C}$ is PSPACE when the number of threads is bounded, and EXPSPACE, otherwise.
+
+%We model the set $\mathcal{C}_D$ as the executions of a Vector Addition System with States (VASS) similarly to Bouajjani et al.
 \cite{conf/esop/BouajjaniEEH13}.
@@
 
-For a fixed set of priorities $P$, the register automata $A(\Gamma)$ can be transformed to finite-state automata since the number of possible valuations of the registers is bounded.
+%For a fixed set of priorities $P$, the register automata $A(\Gamma)$ can be transformed to finite-state automata since the number of possible valuations of the registers is bounded.
 In this way, we reduce checking linearizability of priority queue into the the EXPSPACE-complete problem of checking state-reachability in a VASS.
@@
 
-\noindent {\bf Reducing PSPACE and EXPSPACE problems into linearizability of priority queue for bounded number of threads:} We show how to reduce the state-reachability problem of VASS into checking linearizability of priority queue.
-Our construction is based on the reduction of state-reachability problem of counter machine and VASS into linearizability in \cite{conf/esop/BouajjaniEEH13}.
-
-There are already several linearizable lock-free implementations of priority queue \cite{DBLP:conf/ipps/SundellT03,DBLP:conf/ppopp/LiuS12}.
+\noindent {\bf Reducing State-Reachability Problem of VASS into Linearizability Problem of Priority Queue:} We show how to reduce the state-reachability problem of VASS into checking linearizability of priority queue.
+Our construction is based on the reduction of state-reachability problem of counter machine and VASS into linearizability problem in \cite{conf/esop/BouajjaniEEH13}.
+
+%There are already several linearizable lock-free implementations of priority queue \cite{DBLP:conf/ipps/SundellT03,DBLP:conf/ppopp/LiuS12}.
 It is not hard to modify them to fit finite variables with finite data domain and it is obvious that they are still linearizable then.
@@
 
+Let $\textit{Imp-PQ}_{\textit{lf}}$ be a correct priority queue implementation where (1) each method execute in a sequentially way (can be achieved by using lock and unlock) and (2) uses finite variables with finite data domain.
+It is quite easy to construct $\textit{Imp-PQ}_{\textit{lf}}$ and we omit its detailed definition here.
+By condition (1) we know that each method does not to wait for other method to complete its execution.
+It is safe to assume that $\textit{Imp-PQ}_{\textit{lf}}$ is given in the form of pseudo-code.
+We use $\textit{put}_{\textit{lf}}$ and $\textit{rm}_{\textit{lf}}$ to explicitly denote the pseudo-code of put and remove method of $\textit{Imp-PQ}_{\textit{lf}}$.
+
 A counter machine is a VASS that can additional detect if the value of some counter is $0$.
@@
 
-Let $D' = \{ (q,\vec n,q') \vert q,q' \in Q_{\mathcal{A}} - \{q_f\}, \vec n \in \mathbb{Z}^s \} \cup \{ q_f \} \cup \{\textit{inc-i},\textit{dec-i} \vert 1 \leq i \leq s\}$.
+%Let $D' = \{ (q,\vec n,q') \vert q,q' \in Q_{\mathcal{A}} - \{q_f\}, \vec n \in \mathbb{Z}^s \} \cup \{ q_f \} \cup \{\textit{inc-i},\textit{dec-i} \vert 1 \leq i \leq s\}$.
 Let $\textit{RAs}$ be the register automata obtained as follows:
 
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] Given $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$, we modify the $\Gamma$-complete automata while the original value of $\top$ is replaced by values $\top$ and all values in $D'$.
-
-\item[-] For register automata that checks FIFO of sub-executions of operations with a same priority, if the original value under consideration is $\{ b \}$ (the first three automata) or $\{a,b\}$ (the forth automaton), then we replace them with $\{ b \} \cup D'$ and $\{a,b\} \cup D'$, respectively.
-\end{itemize}
+%\begin{itemize}
+%\setlength{\itemsep}{0.5pt}
+%\item[-] Given $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$, we modify the $\Gamma$-complete automata while the original value of $\top$ is replaced by values $\top$ and all values in $D'$.
+
+%\item[-] For register automata that checks FIFO of sub-executions of operations with a same priority, if the original value under consideration is $\{ b \}$ (the first three automata) or $\{a,b\}$ (the forth automaton), then we replace them with $\{ b \} \cup D'$ and $\{a,b\} \cup D'$, respectively.
+%\end{itemize}
 
@@
 \setlength{\itemsep}{0.5pt}
-\item[-] The $\textit{put}$ method uses items from $D' \cup \{ a,b,a_1,d,e,\top \}$.
+\item[-] The $\textit{put}$ method first uses $\textit{put}_{\textit{lf}}$ to do the work of priority queue, and then choose a role in $\{ \textit{trans}(q,\vec n,q'), \textit{reach-}q_f, \textit{inc-i},\textit{dec-i} \}$ and work according to it.
+If $q_f$ is reached, it then set the flag $\textit{reachQf}$ to be $\textit{true}$.
+The pseudo-code of $\textit{put}$ is shown in Algorithm \ref{alg:new put}.
+\end{itemize}
+
+%\begin{itemize}
+%\setlength{\itemsep}{0.5pt}
+%\item[-] The $\textit{put}$ method uses items from $D' \cup \{ a,b,a_1,d,e,\top \}$.
 It first uses $\textit{put}_{\textit{lf}}$ to do the work of priority queue, and then choose a role and begin to simulate transitions of $\mathcal{A}$, until $q_f$ is reached.
@@
 The pseudo-code of $\textit{put}$ is shown in Algorithm \ref{alg:new put}.
-\end{itemize}
-
-In \cite{conf/esop/BouajjaniEEH13}, to use a library to simulate counter machine $\mathcal{A}$, their library $\mathcal{L}_{\mathcal{A}}$ contains the following variables: (1) variable $q$ that uses value from $Q_{\mathcal{A}}$, (2) vector $\textit{req}[]$ and $\textit{ack}[]$, the index is chosen from $1$ to the number of unit vectors of length $s$, (3) $\textit{dec}[]$, where the index is chosen form $1$ to $s$.
-
-Let us explain how $\textit{put}$ method do to act according to its chosen role.
+%\end{itemize}
+
+\cite{conf/esop/BouajjaniEEH13} uses wait and signal command.
+To obtain an implementation where each method can always make progress in isolation, we need to slightly modify the $\textit{wait}$ command as follows:
+
+\begin{itemize}
+\setlength{\itemsep}{0.5pt}
+\item[-] If a wait command find itself blocked, then it can still proceed.
+However, in this case the $\textit{put}$ method will return a specific value $\textit{fail}$.
+Moreover, when deciding linearizability, we do not consider executions with return value $\textit{fail}$.
+\end{itemize}
+
+%In \cite{conf/esop/BouajjaniEEH13}, to use a library to simulate counter machine $\mathcal{A}$, their library $\mathcal{L}_{\mathcal{A}}$ contains the following variables: (1) variable $q$ that uses value from $Q_{\mathcal{A}}$, (2) vector $\textit{req}[]$ and $\textit{ack}[]$, the index is chosen from $1$ to the number of unit vectors of length $s$, (3) $\textit{dec}[]$, where the index is chosen form $1$ to $s$.
+
+%Let us explain how $\textit{put}$ method do to act according to its chosen role.
 We use the case of $\textit{trans}(q,\vec n,q')$ as an example while other cases can be similarly obtained.
@@
 
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] atomically do $\textit{wait}(q)$ and $\textit{signal}(\textit{req}[\vec n])$,
-\item[-] atomically do $\textit{wait}(\textit{ack}[\vec n])$ and $\textit{signal}(q')$,
-\end{itemize}
-
-Note that we want to obtain a implementation where each method can always make progress in isolation.
+%\begin{itemize}
+%\setlength{\itemsep}{0.5pt}
+%\item[-] atomically do $\textit{wait}(q)$ and $\textit{signal}(\textit{req}[\vec n])$,
+%\item[-] atomically do $\textit{wait}(\textit{ack}[\vec n])$ and $\textit{signal}(q')$,
+%\end{itemize}
+
+%Note that we want to obtain a implementation where each method can always make progress in isolation.
 To obtain such an implementation, we need to slightly modify the $\textit{wait}$ command as follows:
 
-\begin{itemize}
-\setlength{\itemsep}{0.5pt}
-\item[-] If a wait command find itself blocked, then it can still proceed.
+%\begin{itemize}
+%\setlength{\itemsep}{0.5pt}
+%\item[-] If a wait command find itself blocked, then it can still proceed.
 However, in this case the $\textit{put}$ method will return a specific value $\textit{fail}$.
 Moreover, when deciding linearizability, we do not consider executions with return value $\textit{fail}$.
-\end{itemize}
+%\end{itemize}
 

### 2017-07-13 priority-queues-arxiv.tex
arxiv

@@
 The return value of a remove is transformed to an argument value for uniformity~\footnote{Method return values are guessed nondeterministically, and validated at return points.
-This can be handled using the {\tt assume} statements of typical formal specification languages, which only admit executions satisfying a given predicate.}.
+This can be handled using {\tt assume} statements, which only admit executions satisfying a given predicate.}.
 The return actions are denoted in a similar way as $\textit{ret}_o(\textit{put},a,p)$ and respectively, $\textit{ret}_o(\textit{rm},a')$.
 
-An \emph{execution} $e$ is a sequence of call and return actions which satisfy the following well-formedness properties: each return is preceded by a matching call (i.e., having the same operation identifier), and each operation identifier is used in at most one call/return.
+An \emph{execution} $e$ is a sequence of call and return actions which satisfy the following well-formedness properties: each return is preceded by a matching call (having the same operation identifier), and each operation identifier is used in at most one call/return.
 We assume every set of executions is closed under isomorphic renaming of operation identifiers.
+An $m(a)$-operation in an execution $e$ is an operation identifier $o$ s.t.
+$e$ contains the actions $\textit{call}_o(m,a)$ and $\textit{ret}_o(m,a)$.
 An execution is called \emph{sequential} when no two operations overlap, i.e., each call action is immediately followed by its matching return action, and \emph{concurrent} otherwise.
-To ease the reading, we write a sequential execution as a sequence of $\textit{put}(a,p)$ and $\textit{rm}(a)$ symbols representing a pair of actions $\textit{call}_o(\textit{put},a,p)\cdot \textit{ret}_o(\textit{put},a,p)$ and $\textit{call}_o(\textit{rm},a)\cdot \textit{ret}_o(\textit{rm},a)$, respectively (where $o\in\mathbb{O}$).
+For readability, we write a sequential execution as a sequence of $\textit{put}(a,p)$ and $\textit{rm}(a)$ symbols representing a pair of actions $\textit{call}_o(\textit{put},a,p)\cdot \textit{ret}_o(\textit{put},a,p)$ and $\textit{call}_o(\textit{rm},a)\cdot \textit{ret}_o(\textit{rm},a)$, respectively ($o\in\mathbb{O}$).
 For example, given two priorities $p_1 \prec p_2$, $\textit{put}(a,p_2) \cdot \textit{put}(b,p_1) \cdot \textit{rm}(b)$ is a sequential execution of the priority queue ($\textit{rm}$ returns $b$ because it has smaller priority).
 
-We define $\seqPQ$, the set of sequential priority queue executions, semantically via labelled transition system ($LTS$).
-An LTS is a tuple $A=(Q,\Sigma,\rightarrow,q_0)$, where $Q$ is a set of states, $\Sigma$ is an alphabet of transition labels, $\rightarrow\subseteq Q\times\Sigma\times Q$ is a transition relation and $q_0$ is the initial state.
-We model priority queue as an LTS $\textit{PQ}$ defined as follows: each state of the LTS $\textit{PQ}$ is a mapping associating priorities in $\mathbb{P}$ with sequences of values in $\mathbb{D}$ representing a snapshot of the priority queue (for each priority, the values are ordered as they were inserted); the transition labels are $\textit{put}(a,p)$ and $\textit{rm}(a)$ operations; Each transition modifies the state as expected.
+We define $\seqPQ$, the set of sequential priority queue executions, semantically via a labelled transition system (LTS, for short).
+An LTS is a tuple $A=(Q,\Sigma,\rightarrow,q_0)$, where $Q$ is a set of states, $\Sigma$ is an alphabet of transition labels, $\rightarrow\subseteq Q\times\Sigma\times Q$ is a transition relation, and $q_0$ is the initial state.
+We model the priority queue as an LTS $\textit{PQ}$ where states are mappings associating priorities in $\mathbb{P}$ with sequences of values in $\mathbb{D}$, representing a snapshot of the priority queue (for each priority, the values are ordered as they were inserted), and the transition labels are $\textit{put}(a,p)$ and $\textit{rm}(a)$.
+Each transition modifies the state as expected.
 For example, $q_1 \xrightarrow{\textit{rm}(\textit{empty})} q_2$ if $q_1 = q_2$, and $q_1$ and $q_2$ map each priority to the empty sequence $\epsilon$.
-Then, $\seqPQ$ is the set of traces of $\textit{PQ}$.
+Then, $\seqPQ$ is the set of traces (words) accepted by $\textit{PQ}$.
+%}{\color {red}$\seqPQ$ is the set of traces of $\textit{PQ}$}.
+{\color {red}The detailed definition of $\textit{PQ}$ can be found in the long version \cite{CONCUR2017Ahmed}.}
 The detailed definition of $\textit{PQ}$ can be found in Appendix \ref{sec:appendix definition of seqPQ and proof of Lemma EQP rules and semantics}.
@@
 Implementations represent libraries whose methods are called by external programs.
-In the remainder of this work, we consider only completed executions, where each call action has a corresponding return action.
-This simplification is sound when implementation methods can always make progress in isolation: formally, for any execution $e$ with pending operations, there exists an execution $e'$ obtained by extending $e$ only with the return actions of the pending operations of $e$.
-Intuitively this means that methods can always return without any help from outside threads, avoiding deadlock.
+In the remainder of this work, we consider only \emph{completed} executions, where each call action has a corresponding return action.
+This simplification is sound when the method invocations can always make progress in isolation.
+%formally, for any execution $e$ with pending operations, there exists an execution $e'$ obtained by extending $e$ only with the return actions of the pending operations of $e$.
+%Intuitively this means that methods can always return without any help from outside threads, avoiding deadlock.
 
@@
 
-We define two properties which are satisfied by priority queue implementations and which are important for the results that follow: (1) \emph{data independence}~\cite{conf/popl/Wolper86,conf/tacas/AbdullaHHJR13} states that priority queue behaviors do not depend on the actual values which are added to the queue, and (2) \emph{closure under projection}~\cite{DBLP:conf/icalp/BouajjaniEEH15} states that intuitively, remove operations can return the same values no matter how many other different values are in the queue, assuming they don't have more important priorities.
+We define two properties which are %satisfied by priority queue implementations and which are
+important for our results: (1) \emph{data independence}~\cite{conf/popl/Wolper86,conf/tacas/AbdullaHHJR13} states that priority queue behaviors do not depend on the actual values which are added to the queue, and (2) \emph{closure under projection}~\cite{DBLP:conf/icalp/BouajjaniEEH15} states that executions remain valid by removing all the operations adding or removing certain values.
+%remove operations can return the same values even if certain values are projected out from the priority queue state.
+%no matter how many other different values are in the queue, provided that they don't have more important priorities.
 
@@
 
-Beyond the sequential executions, every (concurrent) implementation of the priority queue that we are aware of is data-independent.
-Therefore, from now on, we consider only data-independent implementations.
-This assumption enables a reduction from checking the correctness of an implementation $\mathcal{I}$ to checking the correctness of only its data-differentiated executions in $\mathcal{I}_{\neq}$.
+Beyond sequential executions, every concurrent priority queue implementation that we are aware of is data-independent.
+From now on, we consider only data-independent implementations.
+This assumption enables a reduction from checking the correctness of an implementation $\mathcal{I}$ to checking the correctness of its data-differentiated executions in $\mathcal{I}_{\neq}$.
 
-Besides data independence, the sequential behaviors of the priority queue satisfy the following closure property: a behavior remains valid when removing all the operations with an argument in some set of values $D \subseteq \mathbb{D}$ and any $\textit{rm}(\textit{empty})$ operation (since they are read-only and they don't affect the queue's state).
-In order to distinguish between different $\textit{rm}(\textit{empty})$ operations while simplifying the technical exposition, we assume that they receive as argument a value, i.e., call actions are of the form $\textit{call}_o(\textit{rm},\textit{empty},a)$ for some $a\in \mathbb{D}$.
+Besides data independence, the sequential executions of the priority queue satisfy the following closure property: an execution remains valid when removing all the operations with an argument in some set of values $D \subseteq \mathbb{D}$ and any $\textit{rm}(\textit{empty})$ operation (since they are read-only and they don't affect the queue's state).
+To distinguish between different $\textit{rm}(\textit{empty})$ operations while simplifying the technical exposition, we assume that they receive as argument a value, i.e., call actions are of the form $\textit{call}_o(\textit{rm},\textit{empty},a)$ for some $a\in \mathbb{D}$.
 We will make explicit this argument only when needed in our technical development.
-The projection $e \vert D$ of an execution $e$ to a set of values $D \subseteq \mathbb{D}$ is obtained from $e$ by erasing all call/return actions with an argument not in $D$.
+The projection $e \vert D$ of an execution $e$ to a set of values $D \subseteq \mathbb{D}$ is obtained from $e$ by erasing all the call/return actions with an argument not in $D$.
 We write $e \setminus x$ for the projection $e \vert_{ \mathbb{D} \setminus \{ x \} }$.
-Let $\textit{proj}(e)$ be the set of all projections of $e$ to a set of values.
+Let $\textit{proj}(e)$ be the set of all projections of $e$ to a set of values $D \subseteq \mathbb{D}$.
+%{\color {red}The following lemma states that $\seqPQ$ is closed under projection.}
+
 The proof of the following lemma can be found in Appendix \ref{sec:appendix proofs in section priority queue and data-independence}.
@@
 %\end{lemma}
-\begin{restatable}{lemma}{EPQClosedUnderProjection}
+\begin{lemma}
 \label{lem:closure_proj}
 $\seqPQ$ is closed under projection, i.e., $\textit{proj}(e)\subseteq \seqPQ$ for each $e\in \seqPQ$.
-\end{restatable}
+\end{lemma}
 
@@
 
-Given a (concurrent) execution $e$ and a sequential execution $s$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if there is a bijection $f: O_1 \rightarrow O_2$, where $O_1$ and $O_2$ are the set of operations of $e$ and $s$, respectively, such that (1) $o$ and $f(o)$ is the same operation\footnote{An $m(a)$-operation in an execution $e$ is an operation identifier $o$ s.t.
-$e$ contains the actions $\textit{call}_o(m,a)$, $\textit{ret}_o(m,a)$.}, and (2) if $o_1 <_{\textit{hb}} o_2$, then $f(o_1) <_{\textit{hb}} f(o_2)$.
-A (concurrent) execution $e$ is linearizable w.r.t a set $S$ of sequential executions, denoted $e \sqsubseteq S$, if there exists $s \in S$ such that $e \sqsubseteq s$.
-A set of concurrent executions $E$ is linearizable w.r.t $S$, denoted $E \sqsubseteq S$, if $e \sqsubseteq S$ for all $e \in E$.
+Given a (concurrent) execution $e$ and a sequential execution $s$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if there is a bijection $f: O_1 \rightarrow O_2$, where $O_1$ and $O_2$ are the set of operations of $e$ and $s$, respectively, such that (1) the call and return actions with identifier $o$ and $f(o)$, respectively, are the same
+%{\color{blue}$o$ and $f(o)$ have the same call and return actions}{\color {red}$o$ and $f(o)$ is the same operation}\footnote{},
+and (2) if $o_1 <_{\textit{hb}} o_2$, then $f(o_1) <_{\textit{hb}} f(o_2)$.
+A (concurrent) execution $e$ is linearizable w.r.t.
+a set $S$ of sequential executions, denoted $e \sqsubseteq S$, if there exists $s \in S$ such that $e \sqsubseteq s$.
+A set of concurrent executions $E$ is linearizable w.r.t.
+$S$, denoted $E \sqsubseteq S$, if $e \sqsubseteq S$ for all $e \in E$.
 
 The following lemma states that by data-independence, it is enough to consider only data-differentiated executions when checking linearizability (see Appendix \ref{sec:appendix proofs in section priority queue and data-independence}).
-This is similar to that in \cite{conf/tacas/AbdullaHHJR13,DBLP:conf/icalp/BouajjaniEEH15}, where they use the notion of data-independence in \cite{conf/popl/Wolper86}.
+%This is similar to that in \cite{conf/tacas/AbdullaHHJR13,DBLP:conf/icalp/BouajjaniEEH15}, where they use the notion of data-independence in \cite{conf/popl/Wolper86}.
 Section~\ref{sec:checking inclusion by recursive procedure} will focus on characterizing linearizability for data-differentiated executions.
@@
 
-\begin{restatable}{lemma}{DataDifferentiatedisEnoughforPQ}
-\label{lemma:data differentiated is enough for PQ}
-A data-independent implementation $\mathcal{I}$ is linearizable w.r.t a data-independent set $S$ of sequential executions, if and only if $\mathcal{I}_{\neq}$ is linearizable w.r.t. $S_{\neq}$.
-\end{restatable}
+\begin{lemma}\label{lemma:data differentiated is enough for PQ}
+A data-independent implementation $\mathcal{I}$ is linearizable w.r.t. a data-independent set $S$ of sequential executions, if and only if $\mathcal{I}_{\neq}$ is linearizable w.r.t. $S_{\neq}$.
+\end{lemma}

### 2017-07-13 checking-executions-arxiv.tex
arxiv

@@
 
-%TODO REPHRASE THIS
-%While checking linearizability of a given execution is known to be NP-complete in general~\cite{journals/siamcomp/GibbonsK97} we show in this section that it is polynomial-time for data-differentiated executions of the priority queue.
-We define a recursive procedure for checking linearizability of an execution w.r.t. $\seqPQ$.
+We define a recursive procedure for checking linearizability of a data-differentiated execution w.r.t. $\seqPQ$.
 To ease the exposition, Section~\ref{ssec:seq_exec} introduces a recursive procedure for checking whether a data-differentiated \emph{sequential} execution is admitted by the priority queue which is then extended to the concurrent case in Section~\ref{ssec:conc_exec}.
@@
 
-Checking whether a data-differentiated sequential execution belongs to $\seqPQ$ could be implemented by checking membership into the set of traces of the LTS $PQ$.
-The recursive procedure $\textit{Check-PQ-Seq}$ outlined in Algorithm~\ref{alg:seq_check} is an alternative to this membership test.
+The recursive procedure $\textit{Check-PQ-Seq}$ outlined in Algorithm~\ref{alg:seq_check} checks whether a data-differentiated sequential execution belongs to $\seqPQ$ (i.e., if it is accepted by the LTS $PQ$).
  Roughly, it selects one or two operations in the input execution, checks whether their return values are correct by ignoring the order between the other operations other than how they are ordered w.r.t.
 the selected ones, and calls itself recursively on the execution without the selected operations.
-%{\color {blue} Here we implicitly mix operation $o$ (when $\textit{call}_o(m,a)$, $\textit{ret}_o(m,b)$ in $e$) with $m(a,b)$.}
 
 We explain how the procedure works on the following execution:
-{
 \begin{align}
-\hspace{-5mm}\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{rm}(\textit{empty}) \cdot \textit{put}(d,p_2) \cdot \textit{put}(f,p_3) \cdot \textit{rm}(f) \cdot \textit{put}(b,p_1)\label{eq:ex_rec1}
+\hspace{-5mm}\textit{put}(c,p_2)\hspace{-.5mm} \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{rm}(\textit{empty}) \cdot \textit{put}(d,p_2) \cdot \textit{put}(\mathit{f},p_3) \cdot \textit{rm}(\mathit{f}) \cdot \textit{put}(b,p_1)\label{eq:ex_rec1}
 \end{align}
-}
-where $p_1$, $p_2$, $p_3$ are priorities such that $p_1 \prec p_2$ and $p_1 \prec p_3$, and $p_2$ and $p_3$ are incomparable.
-Since the $\textit{rm}(\textit{empty})$ operations %\footnote{An $m(a)$-operation in an execution $e$ is an operation identifier $o$ s.t.
-$e$ contains the actions $\textit{call}_o(m,a)$, $\textit{ret}_o(m,a)$.}
-are read-only (they don't affect the state of the queue), they are selected first.
-Ensuring that an operation $o=\textit{rm}(\textit{empty})$ is correct boils down to checking that every $\textit{put}(x,p)$ operation before $o$ is matched to a $\textit{rm}(x)$ operation which also occurs before $o$.
+where $p_1$, $p_2$, $p_3$ are priorities such that $p_1 \prec p_2$ and $p_1 \prec p_3$, and $p_2$ and $p_3$ are incomparable. Since the $\textit{rm}(\textit{empty})$ operations
+are read-only (they don't affect the queue's state), they are selected first.
+An $\textit{rm}(\textit{empty})$-operation $o$ is correct when every $\textit{put}(x,p)$ operation before $o$ is matched to a $\textit{rm}(x)$ operation which also occurs before $o$.
 This is true in this case for $x\in \{a,c\}$.
-Therefore, the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
-{
-\begin{align*}
-\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(f,p_3) \cdot \textit{rm}(f) \cdot \textit{put}(b,p_1)
-\end{align*}
-}
-When there are no more $\textit{rm}(\textit{empty})$ operations, the procedure selects a $\textit{put}$ operation adding a value with maximal priority which is not removed, and then a pair of $\textit{put}$ and $\textit{rm}$ operations adding and removing the same maximal priority value.
-For instance, since $p_2$ is a maximal priority, it selects the operation $\textit{put}(d,p_2)$.
-{This operation is correct if $d$ is the last value with priority $p_2$, and the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
-{
-\begin{align*}
-\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(f,p_3) \cdot \textit{rm}(f) \cdot \textit{put}(b,p_1)
-\end{align*}
-}
-Since there is no other value of maximal priority which is not removed, the procedure selects a pair of $\textit{put}$/$\textit{rm}$ operations with an argument of maximal priority $p_2$,} for instance, $\textit{put}(c,p_2)$ and $\textit{rm}(c)$.
-The value returned by $\textit{rm}$ is correct if all the values of priority smaller than $p_2$ added before $\textit{rm}(c)$ are also removed before $\textit{rm}(c)$.
+Thus, the correctness of (\ref{eq:ex_rec1}) reduces to the correctness of
+\begin{align}
+\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(f,p_3) \cdot \textit{rm}(f) \cdot \textit{put}(b,p_1)\label{eq:ex_rec2}
+\end{align}
+When the execution contains no $\textit{rm}(\textit{empty})$-operation, the procedure selects a $\textit{put}$ operation adding a value that is not removed and that has a maximal priority.
+For (\ref{eq:ex_rec2}), it selects $\textit{put}(d,p_2)$ because $p_2$ is a maximal priority.
+This operation is correct since $d$ is the last value with priority $p_2$ in the execution, and the correctness of (\ref{eq:ex_rec2}) reduces to the correctness of
+\begin{align}
+\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(f,p_3) \cdot \textit{rm}(f) \cdot \textit{put}(b,p_1)\label{eq:ex_rec3}
+\end{align}
+If no operations like above can be found, $\textit{Check-PQ-Seq}$ selects a pair of $\textit{put}$ and $\textit{rm}$ operations adding and removing the same maximal priority value.
+For (\ref{eq:ex_rec2}), it can select
+$\textit{put}(c,p_2)$ and $\textit{rm}(c)$.
+The value returned by $\textit{rm}(c)$ is correct if all the values of priority smaller than $p_2$ added before $\textit{rm}(c)$ are also removed before $\textit{rm}(c)$.
 In this case, $a$ is the only value of priority smaller than $p_2$ and it satisfies this property.
@@
     }
-    %\Else {\KwRet $\mathsf{false}$\;}
 }
@@
 
-In formal terms, the operations which are selected depend on the following set of predicates on executions:
-{\small
+Formally, the selected operations depend on the following set of predicates on executions:
 \begin{align*}
-\mathsf{Has\text{-}EmptyRemoves}(e)=\mathsf{true} & \mbox{ iff  $e$ contains a $\textit{rm}(\textit{empty})$ operation} \\
-\mathsf{Has\text{-}UnmatchedMaxPriority}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{unmatched-priorities}(e)$ for a maximal priority} \\
-&\hspace{4mm}\mbox{$p\in priorities(e)$}
-\end{align*}}
+& \mathsf{Has\text{-}EmptyRemoves}(e)=\mathsf{true} \mbox{ iff  $e$ contains a $\textit{rm}(\textit{empty})$-operation} \hspace{1cm}\\
+& \mathsf{Has\text{-}UnmatchedMaxPriority}(e)=\mathsf{true} \mbox{ iff $p\in \textit{unmatched-priorities}(e)$ for a maximal $p$}
+\end{align*}
 where $\textit{priorities}(e)$, resp., $\textit{unmatched-priorities}(e)$, is the set of priorities occurring in $\textit{put}$ operations of $e$, resp., in $\textit{put}$ operations of $e$ for which there is no $\textit{rm}$ operation removing the same value.
@@
 For simplicity, we consider the following syntactic sugar $\mathsf{Has\text{-}MatchedMaxPriority}(e)=\neg \mathsf{Has\text{-}EmptyRemoves}(e)\land \neg \mathsf{Has\text{-}UnmatchedMaxPriority}(e)$.
-By an abuse of notation, we also assume that  $\mathsf{Has\text{-}UnmatchedMaxPriority}(e) \Rightarrow \neg \mathsf{Has\text{-}EmptyRemoves}(e)$ (this is sound by the order of the conditionals in $\textit{Check-PQ-Seq}$).
+By an abuse of notation, we assume  $\mathsf{Has\text{-}UnmatchedMaxPriority}(e) \Rightarrow \neg \mathsf{Has\text{-}EmptyRemoves}(e)$ (this is sound by the order of the conditionals in $\textit{Check-PQ-Seq}$).
 
 The predicates defining the correctness of the selected operations are defined as follows:
-{\small
 \begin{align*}
-\mathsf{EmptyRemove\text{-}Seq}(e,o)=\mathsf{true} & \mbox{ iff  $e= u\cdot o\cdot v$ and $\textit{matched}(u)$} \\
-\mathsf{UnmatchedMaxPriority\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v$, $p\not\prec \textit{priorities}(u\cdot v)$, and $p\not\in \textit{priorities}(v)$} \\
-\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v\cdot \textit{rm}(x)\cdot w$, $p\not\prec \textit{priorities}(u\cdot v\cdot w)$,} \\
-&\hspace{4mm}\mbox{$p\not\preceq \textit{unmatched-priorities}(u\cdot v\cdot w)$, $\textit{matched}_\prec(u\cdot v,p)$,} \\
-&\hspace{4mm}\mbox{and $p\not\in \textit{priorities}(v\cdot w)$}
+&\mathsf{EmptyRemove\text{-}Seq}(e,o)=\mathsf{true} \mbox{ iff  $e= u\cdot o\cdot v$ and $\textit{matched}(u)$} \\
+&\mathsf{UnmatchedMaxPriority\text{-}Seq}(e,x)=\mathsf{true}  \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v$, $p\not\prec \textit{priorities}(u\cdot v)$,} \\
+&\hspace{6.4cm}p\not\in \textit{priorities}(v) \\
+&\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)=\mathsf{true} \mbox{ iff  $e= u\cdot \textit{put}(x,p)\cdot v\cdot \textit{rm}(x)\cdot w$, $\textit{matched}_\prec(u\cdot v,p)$,} \\
+&\hspace{4.3cm}\mbox{$p\not\preceq \textit{unmatched-priorities}(u\cdot v\cdot w)$, $p\not\prec \textit{priorities}(u\cdot v\cdot w)$,} \\
+&\hspace{4.3cm}\mbox{and  $p\not\in \textit{priorities}(v\cdot w)$}
 \end{align*}
-}
-
-\vspace{-3mm}
-\noindent
-where $p\prec \textit{priorities}(e)$ when $p\prec p'$ for some $p'\in \textit{priorities}(e)$ (and similarly for $p\prec \textit{unmatched-p}$ $\textit{riorities}(e)$ or $p\preceq \textit{unmatched-priorities}(e)$),
+where $p\prec \textit{priorities}(e)$ when $p\prec p'$ for some $p'\in \textit{priorities}(e)$ (and similarly for $p\prec \textit{unmatched-priorities}(e)$ or $p\preceq \textit{unmatched-priorities}(e)$),
 $\textit{matched}_\prec(e,p)$ holds when each value with priority strictly smaller than $p$ is removed in $e$, and $\textit{matched}(e)$ holds when $\textit{matched}_\prec(e,p)$ holds for each $p\in\mathbb{P}$.
@@
 
-When $o$ is a $\textit{rm}(\textit{empty})$ operation, we write $e\setminus o$ for the maximal subsequence of $e$ which doesn't contain $o$.
-For an execution $e$, $\textit{values}(e)$ is the set of values occurring in call/return actions of $e$.
-
-% and items of unmatched put in $l$, respectively.
-$p \prec \textit{priorities}(l)$ (resp., $p \preceq \textit{priorities}(l)$), if $p \prec p'$ (resp., $p \preceq p'$)for some $p' \in \textit{priorities}(l)$.
-Similarly we can define $p \preceq \textit{unmatched-priorities}(l)$.
-%
-%
-%Essentially, this procedure chooses either a $\textit{rm}(\textit{empty})$ operation or a set of operations adding/removing a maximal priority in the input execution and checks whether they satisfy certain properties
-%
-%
-% proceeds by
-%
-% The projection $e \vert{O}$ of $e$ into $O$ is obtained from $e$ by erasing all call and return actions of non-$O$ operations.
-We write $e \setminus o$ for the projection $e \vert_{ O_e \setminus \{ o \} }$, where $O_e$ is the set of operations of $e$.
-%
-%Given a data-differentiated sequential execution $e$, it is costly to check whether $e$ belongs to priority queue, since for each element of $e$, we should keep the current content of priority queue and try to modify the priority queue according to this element.
-Such process is too complex.
-In this section, we propose another method for checking inclusion of priority queue by using a recursive procedure $\textit{check-PQ}$.
-For checking a sequence, every time we only check a much simpler property, and then recursively check the remanning sequences obtained by erasing one or two elements.
-Each step of our method is easier to deal with.
-%
-%$\textit{put}(a,p)$ matches $\textit{rm}(b)$, if $a = b$. To introduce $\textit{check-PQ}$, let us introduce several predicates. Given sequential execution $l$ and priority $p$:
-%
-%\begin{itemize}
-%\setlength{\itemsep}{0.5pt}
-%\item[-] $\textit{rm}(\textit{empty}) \notin l$ is satisfied when $l$ does not contain $\textit{rm}(\textit{empty})$.
-%
-%\item[-] Let $\textit{priorities}(l)$ and $\textit{unmatched-priorities}(l)$ be the set of priorities of putted items and items of unmatched put in $l$, respectively.
-$p \prec \textit{priorities}(l)$ (resp., $p \preceq \textit{priorities}(l)$), if $p \prec p'$ (resp., $p \preceq p'$)for some $p' \in \textit{priorities}(l)$.
-Similarly we can define $p \preceq \textit{unmatched-priorities}(l)$.
-%
-%\item[-] $\textit{matched}_{\prec}(l,p)$ is satisfied, if for each item of $l$ with priorities smaller than $p$, it has matched $\textit{put}$ and $\textit{rm}$.
-Similarly, we can define $\textit{matched}(l)$, where we consider all items in $l$ instead of items with certain priorities.
-%\end{itemize}
-%
-%Let us introduce some notations: given sequential executions $u,v,w$ and priority $p$,
-%
-%\begin{itemize}
-%\setlength{\itemsep}{0.5pt}
-%\item[-] $\textit{PQ}_1(u,v,w,p) \equiv
-%(\textit{rm}(\textit{empty}) \notin u \cdot v \cdot w) \wedge
-%(p \not\prec \textit{priorities}(u \cdot v \cdot w)) \wedge
-%(p \not\preceq \textit{unmatched-priorities}(u \cdot v \cdot w)) \wedge
-%(\textit{matched}_{\prec}(u \cdot v,p) ) \wedge
-%(p \notin \textit{priorities}(v \cdot w))$
-%
-%\item[-] $\textit{PQ}_2(u,v,p) \equiv
-%(\textit{rm}(\textit{empty}) \notin u \cdot v) \wedge
-%(p \not\prec \textit{priorities}(u \cdot v \cdot w)) \wedge
-%(p \notin \textit{priorities}(v \cdot w))$
-%
-%\item[-] $\textit{PQ}_3(u,v) \equiv
-%\textit{matched}(u \cdot v)$
-%\end{itemize}
-%
-%
-%Let us defined set $\textit{MS}(R)$ ($R \in \textit{PQ}_1,\textit{PQ}_2,\textit{PQ}_3$), which is the set of data-differentiated sequential executions that respect $R$.
-%
-%\begin{itemize}
-%\setlength{\itemsep}{0.5pt}
-%\item[-] $e \in \textit{MS}(\textit{PQ}_1)$, if $e = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$ and $\textit{PQ}_1(u,v,w,p)$ holds for some item $x$ and a maximal priority $p$,
-%
-%\item[-] $e \in \textit{MS}(\textit{PQ}_2)$, if $e = u \cdot \textit{put}(x,p) \cdot v$ and $\textit{PQ}_2(u,v,p)$ holds for some item $x$ and a maximal priority $p$,
-%
-%\item[-] $e \in \textit{MS}(\textit{PQ}_3)$, if $e = u \cdot \textit{rm}(\textit{empty}) \cdot v$ and $\textit{PQ}_3(u,v)$ holds.
-%\end{itemize}
-%
-%We call such $x$ and $\textit{rm}(\textit{empty})$ (if exists) the witness of $e$.
-Note that sequences in $\textit{MS}(R)$ does not guaranteed to be a priority queue execution.
-Let us introduce the notion $\textit{last}(e)$, which is a set and used as the branch condition of $\textit{check-PQ}$.
-Here $e$ is a data-differentiated sequential execution.
-%
-%\begin{itemize}
-%\setlength{\itemsep}{0.5pt}
-%\item[-] If $e$ contains $\textit{rm}(\textit{empty})$, then $\textit{last}(e) = \{ \textit{PQ}_3 \}$.
-%
-%\item[-] Else, if for a maximal priority $p$ of $e$, $\textit{unmatched-priorities}(e) \neq \emptyset$, then $\textit{PQ}_2 \in \textit{last}(e)$.
-%
-%\item[-] Else, if for a maximal priority $p$ of $e$, $\textit{unmatched-priorities}(e) = \emptyset$, then $\textit{PQ}_1 \in \textit{last}(e)$.
-%\end{itemize}
-%
-%When $\textit{last}(e)$ contains only one element $R$, we write $\textit{last}(w)=R$ for simplicity.
-%
-%The projection $e \vert{\mathcal{D}}$ of $e$ into $\mathcal{D} \subseteq \mathbb{D}$ is obtained from $e$ by erasing all elements with a data value not in $\mathcal{D}$.
-We write $e \setminus x$ for the projection $e \vert_{ \mathbb{D} \setminus \{ x \} }$.
-The projection $e \vert{O}$ of $e$ into $O$ is obtained from $e$ by erasing all call and return actions of non-$O$ operations.
-We write $e \setminus o$ for the projection $e \vert_{ O_e \setminus \{ o \} }$, where $O_e$ is the set of operations of $e$.
-%
-%The recursive procedure $\textit{con-check-EPQ}$ is given as follows. Here $P_i(e,o)$ and $P_i(e,x)$ is the predicates of $e \in \textit{MS}(\textit{PQ}_i)$ withe witness $o$ or $x$, respectively.
-%
-%\begin{minipage}{.5\textwidth}
-%\begin{algorithm}[H]
-%\KwIn {a data-differentiated sequential execution $e$}
-%
-%\If {$e = \epsilon$}
-%{\KwRet $\textit{true}$;}
-%
-%\If {$\textit{last}(e) = \textit{PQ}_3$}
-%{
-%    \If {$\exists o=\textit{rm}(\textit{empty})$, $P_3(e,o)$ holds}
-%    {
-%        \KwRet $\textit{check-PQ}(e \setminus o)$;
-%    }
-%    \KwRet $\textit{false}$;
-%}
-%
-%\If {$\textit{PQ}_i \in \textit{last}(e)$ for $i \in \{1,2\}$}
-%{
-%    \If {$\exists l$, $P_i(e,x)$ holds}
-%    {
-%        \KwRet $\textit{check-PQ}(e \setminus x)$;
-%    }
-%    \KwRet $\textit{false}$;
-%}
-%\caption{$\textit{check-PQ}$}
-%\label{Method-check-PQ}
-%\end{algorithm}
-%\end{minipage}
-%\hfill
-
-%TODO MODIFY THE FOLLOWING EXAMPLE
-%
-%Let us use an example to show how $\textit{check-PQ}$ works.
-Here we use $w_1 \xrightarrow{R} w_2$ to denote that in $\textit{check-PQ}(w_1)$, we choose the branch of $R$ and recursively call $\textit{check-PQ}(w_2)$.
-%
-%\begin{example}\label{example:generate extended priority queue executions}
-%Given priorities $p_1,p_2,p_3$ with orders $p_1 \prec p_2$ and $p_1 \prec p_3$,
-%
-%\noindent $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{rm}(\textit{empty}) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)$
-%
-%$\xrightarrow{\textit{EPQ}_3}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(e,p_3) \cdot \textit{rm}(e) \cdot \textit{put}(b,p_1)$
-%
-%$\xrightarrow{\textit{EPQ}_1}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(d,p_2) \cdot \textit{put}(b,p_1)$
-%
-%$\xrightarrow{\textit{EPQ}_2}$ $\textit{put}(c,p_2) \cdot \textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{rm}(c) \cdot \textit{put}(b,p_1)$
-%
-%$\xrightarrow{\textit{EPQ}_1}$ $\textit{put}(a,p_1) \cdot \textit{rm}(a) \cdot \textit{put}(b,p_1)$
-%
-%$\xrightarrow{\textit{EPQ}_2}$ $\textit{put}(a,p_1) \cdot \textit{rm}(a)$
-%
-%$\xrightarrow{\textit{EPQ}_1}$ $\epsilon$
-%\end{example}
+When $o$ is an $\textit{rm}(\textit{empty})$-operation, $e\setminus o$ is the maximal subsequence of $e$ which doesn't contain $o$.
+For an execution $e$, $\textit{values}(e)$ is the set of values  in call/return actions of $e$.
 
@@
 
-\begin{restatable}{lemma}{EPQRulesAndSemantics}
+\begin{lemma}
 \label{lemma:EPQ rules and semantics}
 $\textit{Check-PQ-Seq}(e)=\mathsf{true}$ iff $e\in \seqPQ$, for every data-differentiated sequential execution $e$.
-\end{restatable}
-
-%
-%\begin{lemma}
-%Algorithm~\ref{alg:seq_check} returns $\mathsf{true}$ on an execution $e$ iff $e\in \seqPQ$.
-%\end{lemma}
-
-%Let $\textit{PQ}$ be set of sequences obtained by renaming some data-differentiated sequential execution that is accepted by $\textit{check-PQ}$.
-To persuade readers that $\textit{PQ}$ is indeed the set of behaviors of priority queue, in Appendix \ref{sec:appendix in section inductive rules of extended priority queue}, we give a semantical version definition $\textit{PQ}_s$ of priority queue, and shows that $\textit{PQ} = \textit{PQ}_s$.
-To model the possible behaviors of priority queue, we model it as an labelled transition system (shortened as LTS) $\textit{LTS}_e$.
-Each state of $\textit{LTS}_e$ is a function from $\mathbb{P}$ into sequences over $\mathbb{D}$, and represents a snapshot of contents of priority queue.
-$\textit{PQ}_s$ is the set of traces of $\textit{LTS}_e$.
-The definition of $\textit{PQ}_s$ and the proof of the following lemma can be found in Appendix \ref{sec:appendix in section inductive rules of extended priority queue}.
-%
-%\begin{restatable}{lemma}{EPQRulesAndSemantics}
-%\label{lemma:EPQ rules and semantics}
-%$\textit{PQ} = \textit{PQ}_s$.
-%\end{restatable}
-%
-%By Example \ref{example:generate extended priority queue executions} we can see how to generate a sequence $e \in \textit{EP}$ from $\epsilon$ as follows:
-%
-%\begin{itemize}
-%\setlength{\itemsep}{0.5pt}
-%\item[-] First we add non-$\textit{rm}(\textit{empty})$ elements with a loop: Let $P$ be the set of priorities of items in $e$.
-In the first round of the loop, we choose a minimal priorities $p \in P$, we add items by using $\textit{PQ}_1$ and $\textit{PQ}_2$, let $P = P \setminus \{p\}$, and begins the next round of loop.
-%
-%\item[-] Then, we add $\textit{rm}(\textit{empty})$ by using $\textit{PQ}_3$.
-%\end{itemize}
+\end{lemma}
 
@@
 \begin{align*}
-\Gamma\mathsf{\text{-}Conc}(e,\alpha) = \mathsf{true}\mbox{ iff there exists a sequential execution $s$ such that $e\sqsubseteq s$ and $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$}
+\Gamma\mathsf{\text{-}Conc}(e,\alpha) = \mathsf{true}\mbox{ iff there is a sequential execution $s$ such that $e\sqsubseteq s$ and $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$}
 \end{align*}
-for each $\Gamma\in \{\mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$.
-Let $\textit{Check-PQ-Conc}$ denote the thus obtained procedure (we assume recursive calls are modified accordingly).
-
-% with linearizability to deal with linearizability w.r.t $\textit{PQ}$ as follows:
-%
-%\begin{itemize}
-%\setlength{\itemsep}{0.5pt}
-%\item[-] The input $e$ is a concurrent execution.
-%
-%\item[-] $\textit{last}(e)$ is the union of $\textit{last}(u)$ for any sequential execution $u$ such that $e \sqsubseteq u$.
-$e \sqsubseteq \textit{MS}(R)$ with witness $x$ (resp., $o$), if $e \sqsubseteq u \in \textit{MS}(R)$ and $x$ (resp., $o$) is the witness of $u$.
-%
-%\item[-] $P_i(e,o)$ and $P_i(e,x)$ holds, if $e \sqsubseteq \textit{MS}(\textit{PQ}_i)$ withe witness $o$ or $x$, respectively.
-%\end{itemize}
+for each $\Gamma\in \{\mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$. The obtained procedure is denoted by
+$\textit{Check-PQ-Conc}$ (recursive calls are modified accordingly).
 
@@
 
-\begin{restatable}{lemma}{ConCheckEPQIsCorrect}
-\label{lemma:con-check-EPQ is correct}
-$\textit{Check-PQ-Conc}(e)=\mathsf{true}$ iff $e \sqsubseteq \seqPQ$, for every data-differentiated execution $e$.
-\end{restatable}
+\begin{lemma}\label{lemma:con-check-EPQ is correct}
+$\textit{Check-PQ-Conc}(e)=\mathsf{true}$ iff $e \sqsubseteq \seqPQ$, for every data-differentiated $e$.
+\end{lemma}
 
@@
 The main technical difficulty is showing that for any execution $e$, any linearization of $e\setminus x$ for some maximal priority value $x$ can be extended to a linearization of $e$ provided that $\mathsf{UnmatchedMaxPriority}$ or $\mathsf{MatchedMaxPriority}$ holds (depending on whether there are values with the same priority as $x$ in $e$ which are not removed).
-%
-%
-%To ensure that $\textit{check-PQ}(e) = \textit{true} \Rightarrow e \sqsubseteq \textit{PQ}$, we require $\textit{PQ}$ to satisfy the following property.
-%
-%for any data-differentiated execution $e$,
-%\begin{itemize}
-%\setlength{\itemsep}{0.5pt}
-%\item[-] if $e \sqsubseteq \textit{MS}(R)$ ($R \in \{ \textit{PQ}_1, \textit{PQ}_2 \}$) with witness $x$, we have: $e \setminus x \sqsubseteq \textit{PQ} \Rightarrow e \sqsubseteq \textit{PQ}$.
-%
-%\item[-] if $e \sqsubseteq \textit{MS}(\textit{PQ}_3)$ and $o$ is a $\textit{rm}(\textit{empty})$, we have: $e \setminus o \sqsubseteq \textit{PQ} \Rightarrow e \sqsubseteq \textit{PQ}$.
-%\end{itemize}
-%
-%With this property, we can build a linearization of a whole execution by increasingly construct linearization of sub-execution from $\epsilon$, and this convince us that $\textit{check-PQ}(e) = \textit{true} \Rightarrow e \sqsubseteq \textit{PQ}$.
-%
 
@@
 Assume that there exist two sequential executions $l$ and $l'$ such that $e \sqsubseteq l=u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds, and $e \setminus x \sqsubseteq l' \in \seqPQ$.
-Let $u=\epsilon$, $w$ be any sequence containing the set of operations $\textit{put}(z_2,p_2)$ and $\textit{rm}(z_1)$ (we distinguish them by adding the suffix $-w$ to their name, e.g., $\textit{rm}(z_1)-w$), and $v$ be any sequence containing the remaining operations.
-In general, the linearization $l'$ can be defined by choosing for each operation, a point in time between its call and return actions, called \emph{linearization point}.
+Let $u=\epsilon$, $w$ be any sequence formed of $\textit{put}(z_2,p_2)$ and $\textit{rm}(z_1)$ (we distinguish them by adding the suffix ``$-w$'' to their name, e.g., $\textit{rm}(z_1)-w$), and $v$ be any sequence containing the remaining operations.
+In general, the linearization $l'$ can be defined by choosing for each operation, a point in time between its call and return, called \emph{linearization point}.
 The order between the linearization points defines the sequence $l'$.
-\figurename~\ref{fig:concurrent execution for EPQ1}(a) draws linearization points for the operations in $e \setminus x$ defining the linearization $l'$~\footnote{In general, there may exist multiple ways of choosing linearization points to define the same linearization.
+\figurename~\ref{fig:concurrent execution for EPQ1}(a) draws linearization points for the operations in $e \setminus x$ which define $l'$~\footnote{In general, there may exist multiple ways of choosing linearization points to define the same linearization.
 Our construction is agnostic to this choice.}.
-%(the order of these linearization points defines the sequence $l'$).
-We show how to construct a sequence $l''= l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3\in\seqPQ$ such that $e \sqsubseteq l''$.
-%In this example, .
-We also explicitly draw the linearization points according to $l'$.
-We construct $l''$ as follows, and this method does not rely on the locations of linearization points of $l'$ and operations of $x$.
+We show how to construct a sequence $l''= l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3\in\seqPQ$ s.t. $e \sqsubseteq l''$.
 \begin{itemize}
-\setlength{\itemsep}{0.5pt}
 \item[-] An operation is called $p$-comparable (resp., $p$-incomparable) when it receives as argument a value of priority comparable to $p$ (resp., incomparable to $p$).
-We could try to define $l''_1$, $l''_2$ and $l''_3$ as the projection of $l'$ to the operations in $u$, $v$ and $w$, respectively.
-However, this is incorrect, since $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ imposes no restriction on  $p$-incomparable operations in $u \cdot v$, and thus, there is no guarantee that the projection of $l'$ to $p$-incomparable operations in $u \cdot v$ is correct.
-In this example, this projection is $\textit{put}(z_1,p_2) \cdot \textit{rm}(z_2)$ which is incorrect.
+Defining $l''_1$, $l''_2$, and $l''_3$ as the projection of $l'$ to the set of operations in $u$, $v$ and $w$, respectively, leads to a sequence $l''\not\in\seqPQ$.
+This is because $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ imposes no restriction on  $p$-incomparable operations in $u \cdot v$, and the projection of $l'$ to $p$-incomparable operations in $u \cdot v$ is not in $\seqPQ$.
+In this example, this projection is $\textit{put}(z_1,p_2) \cdot \textit{rm}(z_2)$.
 
 \item[-] We define the sets of operations $U'$, $V'$ and $W'$ such that $l''_1$, $l''_2$ and $l''_3$ are the projections of $l'$ to $U'$, $V'$, and $W'$, respectively. This is done in two steps:
-
-\item[-] The first step is to define $W'$.
+\begin{enumerate}
+\item The first step is to define $W'$.
 The $p$-comparable operations in $W'$ are the same as in $w$.
@@
 In this example, $U'$ contains $\textit{put}(z_1,p_2)$ and $\textit{put}(x_2,p)$.
-
+\end{enumerate}
 \item[-] In conclusion, we have that $l''_1 = \textit{put}(z_1,p_2) \cdot \textit{put}(x_2,p)$, $l''_2 = \textit{put}(z_2,p_2) \cdot \textit{rm}(x_2) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(y_1)$, and $l''_3 = \textit{rm}(z_1) \cdot \textit{rm}(z_2)$.
@@
 
-%\begin {proof} (Sketch)
-%It is not hard to see that $\textit{check-PQ}(e) = \textit{false} \Rightarrow e \not\sqsubseteq \textit{PQ}$, since $\textit{PQ}_{\neq} \subset \textit{MS}(R)_{\neq}$.
-
-%This completes the proof of this lemma.
-%
-%
-%\qed
-%\end {proof}
 \begin{figure}[t]
   \centering
-  \includegraphics[width=1 \textwidth]{figures/PIC-HIS-EPQ1-TwoHis-2.pdf}
-%\vspace{-10pt}
-  \caption{The process of obtaining linearization of $e$}
+  \includegraphics[width=.8\textwidth]{figures/PIC-HIS-EPQ1-TwoHis-2.pdf}
+  \caption{A concurrent execution $e$ exemplifying the soundness of $\textit{Check-PQ-Conc}$.}
   \label{fig:concurrent execution for EPQ1}
@@
 
-Section~\ref{sec:co-regular of extended priority queues} introduces a characterization of concurrent priority queue violations using a set of \emph{non-recursive} automata (i.e., whose states consist of a fixed number of registers) whose standard synchronized product is equivalent to $\textit{Check-PQ-Conc}$ (modulo renaming of values which is possible by data-independence).
-Since $\seqPQ$ is closed under projection (Lemma~\ref{lem:closure_proj}), the recursion in $\textit{Check-PQ-Conc}$ can be eliminated by checking that each projection of a given execution $e$ passes a non-recursive version of $\textit{Check-PQ-Conc}$ where every recursive call is replaced by $\mathsf{true}$.
-More precisely, every occurrence of $\text{{\bf return}}\ \textit{Check-PQ-Conc}$ is replaced by  $\text{{\bf return}}\ \mathsf{true}$.
-Let $\textit{Check-PQ-Conc-}$ $\textit{NonRec}$ be the thus obtained procedure.
+Section~\ref{sec:co-regular of extended priority queues} introduces a characterization of concurrent priority queue violations using a set of \emph{non-recursive} automata (whose states consist of a fixed number of registers), whose standard synchronized product is equivalent to $\textit{Check-PQ-Conc}$ (modulo a renaming of values which is possible by data-independence).
+Since $\seqPQ$ is closed under projection (Lemma~\ref{lem:closure_proj}), the recursion in $\textit{Check-PQ-Conc}$ can be eliminated by checking that each projection of a given execution $e$ passes a non-recursive version of $\textit{Check-PQ-Conc}$ where every recursive call $\text{{\bf return}}\ \textit{Check-PQ-Conc}(\ldots)$ is
+replaced by  $\text{{\bf return}}\ \mathsf{true}$. Let $\textit{Check-PQ-Conc-NonRec}$ be the thus obtained procedure.
 
-\begin{restatable}{lemma}{EPQasMultiInMRpriforHistory}
+\begin{lemma}
 \label{lemma:EPQ as multi in MRpri for history}
 Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$ if and only if for each $e' \in \textit{proj}(e)$, $\textit{Check-PQ-Conc-NonRec}(e')$ returns $\mathsf{true}$.
-\end{restatable}
-
-%The following section shows that the linearizability check in the above lemma can be implemented using a set of register automata.
-
-%In this section, we divide checking linearizability w.r.t $\seqPQ$ into checking linearizablity w.r.t $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$, each of which can be done in polynomial-time.
-Due to data-independence, violations to each $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$ can be captured by register automata.
-This reduces checking linearizability w.r.t $\seqPQ$ into state reachability problem.
-
-%\begin{restatable}{lemma}{EPQasMultiInMRpriforHistory}
-%\label{lemma:EPQ as multi in MRpri for history}
-%Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$ if and only if for each $e' \in \textit{proj}(e)$, $\mathsf{Has\text{-}}\Gamma(e')$ implies $\Gamma\mathsf{\text{-}Conc}(e')$.
-%
-%$\forall \Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$,  $\Gamma\mathsf{\text{-}Condition}(e)$ $\Rightarrow$ $e$ is linearizable w.r.t $\Gamma$.
-%\end{restatable}
-%
-%
-%
-%\subsection{Co-Regular}
-%\label{subsec:definition of co-regular}
-%
-%Let $\Gamma\mathsf{\text{-}Condition}(e)$ be $\mathsf{Has\text{-}EmptyRemoves}(e)$, $\neg \mathsf{Has\text{-}EmptyRemoves}(e) \wedge \mathsf{HasUnmatchedMaxPriority}(e)$, and $\mathsf{HasMatchedMaxPriority}(e)$, for $\Gamma = \mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}$, respectively.
-$\textit{Check-PQ-Conc}$ inspires us that, when $\Gamma\mathsf{\text{-}Condition}(e)$ holds, there should exist some $\alpha$, such that  $\Gamma\mathsf{\text{-}Conc}(e,\alpha)$ holds.
-The following lemma states this formally.
-Let $\textit{proj}(e)$ be the set of projections of $e$ into set of values.
-When refer to $\textit{proj}(e)$, we implicitly assume that each $\textit{rm}(\textit{empty})$ in $e$ has a ghost argument that is unique.
-We say that execution $l$ is linearizable w.r.t $\Gamma$, if we can find a sequence $s$ where $e \sqsubseteq s$ and $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$ holds for some $\alpha$.
-%
-%\begin{restatable}{lemma}{EPQasMultiInMRpriforHistory}
-%\label{lemma:EPQ as multi in MRpri for history}
-%Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$, if and only if, $\forall e' \in \textit{proj}(e)$, $\forall \Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$,  $\Gamma\mathsf{\text{-}Condition}(e)$ $\Rightarrow$ $e$ is linearizable w.r.t $\Gamma$.
-%\end{restatable}
+\end{lemma}

### 2017-07-13 co-regular-EPQ-arxiv.tex
arxiv

--- 2017-07-01 co-regular-EPQ-arxiv.tex
@@
 
-We show that the set of executions for which some projection fails the test $\textit{Check-PQ-Conc-NonRec}$ can be characterized using a set of register automata, modulo a value renaming.
-The possibility of renaming values (which is complete for checking data independent implementations) allows to simplify the reasoning about projections.
-Thus, we assume that all the operations which are not in the projection failing this test use the same distinguished value $\top$, different from those used in the projection.
-Then, it is enough to find an automata characterization for the set of executions $e$ for which $\textit{Check-PQ-Conc-NonRec}$ fails, or equivalently, for which one
-%a projection $e'$ fails $\textit{Check-PQ-Conc-NonRec}$ when one
-of the following three formulas is false:
-\begin{align*}
-\hspace{-5mm}
-\Gamma(e) := \mathsf{Has\text{-}}\Gamma(e) \Rightarrow \exists \alpha.\ \Gamma\mathsf{\text{-}Conc}(e,\alpha)\mbox{ with $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$}
-\end{align*}
+We show that the set of executions for which $\textit{Check-PQ-Conc-NonRec}$ fails on some projection can be described using register automata, modulo a value renaming.
+Renaming values (which is complete under data independence) allows to simplify the reasoning about projections.
+W.l.o.g., we assume that all the operations which are not in the projection failing this test use the same distinguished value $\top$, different from those in the projection.
+Then, it is enough to find an automata characterization of the executions $e$ for which $\textit{Check-PQ-Conc-NonRec}(e)$ is $\mathsf{false}$, i.e., for which
+$\Gamma(e) := \mathsf{Has\text{-}}\Gamma(e) \Rightarrow \exists \alpha.\ \Gamma\mathsf{\text{-}Conc}(e,\alpha)$ is false, for some $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$.
 Intuitively, $\Gamma(e)$ states that $e$ is linearizable w.r.t.
@@
 
-$A \cap \mathcal{I} \neq \emptyset$ if and only if there exists $ e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$ such that $e'$ is not $\Gamma$-linearizable.
+$A \cap \mathcal{I} \neq \emptyset$ iff there exists $e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$ such that $e'$ is not $\Gamma$-linearizable.
 \end{definition}
 
-Section~\ref{ssec:aut} describes a $\mathsf{MatchedMaxPriority}$-complete automaton, the other automata being defined in Appendix~\ref{sec:appendix lemma and register automata for FIFO of single-priority executions} %Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar},
-%Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}
-and Appendix \ref{subsec:co-regular of EPQ3}. Therefore, the following holds.
-
-\begin{lemma}
-There exists a $\Gamma$-complete automaton for each $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMax}$ $\mathsf{Priority}$, $\mathsf{MatchedMaxPriority}\}$.
-\end{lemma}
-
+We can show that for any $\Gamma\in \{\mathsf{EmptyRemove}$,$\mathsf{UnmatchedMaxPriority}$,$\mathsf{MatchedMaxPriority}\}$ there exists a $\Gamma$-complete automaton.
+For lack of space, we only consider the case $\Gamma=\mathsf{MatchedMaxPriority}$ in
+Section~\ref{ssec:aut}.
 When defining $\Gamma$-complete automata, we assume that every implementation $\mathcal{I}$ behaves correctly, i.e., as a FIFO queue, when only values with the same priority are observed.
@@
 $\seqPQ$).
-This property can be checked separately using register automata obtained from the finite automata in~\cite{DBLP:conf/icalp/BouajjaniEEH15} for FIFO queue (see Appendix~\ref{sec:appendix lemma and register automata for FIFO of single-priority executions} for more details).
-Note that this assumption excludes some obvious violations, such as a $\textit{rm}(a)$ operation happens before a $\textit{put}(a,p)$ operation, for some $p$.
+This property can be checked separately using register automata similar to the automata in~\cite{DBLP:conf/icalp/BouajjaniEEH15} describing FIFO queue violations (see Appendix~\ref{sec:appendix lemma and register automata for FIFO of single-priority executions} for more details).
+This assumption excludes some obvious violations, such as an $\textit{rm}(a)$ operation happening before a $\textit{put}(a,p)$ operation, for some $p$.
 
-Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions which contain only one maximal priority.
-This is possible because any data-differentiated execution for which $\Gamma(e)$ is false has such a projection.
-%Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions where any two occurring priorities are comparable (w.r.t.
-$\prec$).
-This is possible because any data-differentiated execution which is not $\Gamma$-linearizable has such a projection.
-Formally, given a data-differentiated execution $e$ and $p$ a maximal priority in $e$, $e\vert_{\preceq p}$ is the projection of $e$ to the set of values with priorities smaller than $p$. Then,
+For $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, we consider $\Gamma$-complete automata recognizing executions which contain only one maximal priority.
+This is w.l.o.g.
+because any data-differentiated execution for which $\Gamma(e)$ is false has such a projection.
+Formally, given a data-differentiated execution $e$ and $p$ a maximal priority in $e$, $e\vert_{\preceq p}$ is the projection of $e$ to the set of values with priorities smaller or equal to $p$. Then,
 
-\begin{restatable}{lemma}{priExecutionIsEnough}
+\begin{lemma}
 \label{lemma:pri execution is enough}
-Let $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$ and $e$ a data-differentiated execution.
-Then, $e$ is $\Gamma$-linearizable iff $e\vert_{\preceq p}$ is $\Gamma$-linearizable for some maximal priority $p$ in $e$.
-\end{restatable}
+For $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, a data-differentiated execution $e$
+ is $\Gamma$-linearizable iff $e\vert_{\preceq p}$ is $\Gamma$-linearizable for some maximal priority $p$ in $e$.
+\end{lemma}
 \begin {proof} (Sketch)
-%{\color {blue}We prove an equivalent property: $e$ is $\Gamma$-linearizable if and only if $e\vert_{\preceq p}$ is $\Gamma$-linearizable for some maximal priority $p$ in $e$.} %We consider the case of $\mathsf{MatchedMaxPriority}$ and the case of $\mathsf{UnmatchedMaxPriority}$ is similar.
-To prove the $\textit{only if}$ direction, let $e$ be a data-differentiated execution linearizable w.r.t.
-$l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,x)$.
-Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}$ $(s,x)$ imposes no restriction on the operations in $u$, $v$ and $w$ with priorities incomparable to $p$, erasing all these operations results in a sequential execution which still satisfies this property.
+For the ``only-if'' direction, let $e$ be a data-differentiated execution linearizable w.r.t.
+$l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$ s.t.
+$\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l,x)$ holds.
+Since the predicate $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l,x)$ imposes no restriction on the operations in $u$, $v$, and $w$ with priorities incomparable to $p$, erasing all these operations results in a sequential execution which still satisfies this predicate.
 Similarly, for $\Gamma=\mathsf{UnmatchedMaxPriority}$.
 
-The $\textit{if}$ direction follows from the fact that if the projection of an execution to a set of operations $O_1$ has a linearization $l_1$ and the projection of the same execution to the remaining set of operations has a linearization $l_2$, then the execution has a linearization which is defined as an interleaving of $l_1$ and $l_2$ (see Appendix~\ref{sec:appendix proof of Lemma pri execution is enough} for more details).
+The ``if'' direction follows from the fact that if the projection of an execution to a set of operations $O_1$ has a linearization $l_1$ and the projection of the same execution to the remaining set of operations has a linearization $l_2$, then the execution has a linearization which is defined as an interleaving of $l_1$ and $l_2$ (see Appendix~\ref{sec:appendix proof of Lemma pri execution is enough} for more details).
 
 Thus, let $e$ be an execution such that $e\vert_{\preceq p}$ is linearizable w.r.t.
-$l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,x)$.
+$l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$ where $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l,x)$ holds.
 By the property above, we know that $e$ has a linearization $l' = u' \cdot \textit{put}(x,p) \cdot v' \cdot \textit{rm}(x) \cdot w'$, such that the projection of $l'$ to values of priority comparable to $p$ is $l$.
-%We can see that $u$, $v$ and $w$ are subsequences of $u'$, $v'$ an $w'$, respectively.
-Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,x)$ does not have a condition on values of priority incomparable to $p$, we obtain that $l' \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$.
+Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l,x)$ doesn't constrain the values of priority incomparable to $p$, we obtain that $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l',\alpha)$ also holds.
 \end {proof}
 
-The following shows that $\Gamma$-complete automata enable an effective reduction of checking linearizability of concurrent priority queue implementations to state reachability.
-It is a direct consequence of the above definitions.
+We shows in Appendix \ref{subsec:appendix co-regular of EPQ2Lar} and Appendix \ref{subsec:appendix co-regular of EPQ2Equal} that it is safe to ignore constructing $\mathsf{UnmatchedMaxPriority}$-complete register automata, and we give the $\mathsf{EmptyRemove}$-complete register automata in Appendix \ref{subsec:co-regular of EPQ3}.
+The existence of $\Gamma$-complete automata enable an effective reduction of checking linearizability of concurrent priority queue implementations to state reachability.
 Section~\ref{subsec:combine step-by-step linearizability and co-regular} discusses decidability results implied by this reduction.
 
-\begin{restatable}{theorem}{ReduceEPQIntoStateReachability}
+\begin{theorem}
 \label{lemma:reduce EPQ into state reachability}
-Let $\mathcal{I}$ be a data-independent implementation, and $A(\Gamma)$ be a $\Gamma$-complete automaton for each $\Gamma$. Then,
-$\mathcal{I} \sqsubseteq \seqPQ$ if and only if $\mathcal{I} \cap A(\Gamma) = \emptyset$ for all $\Gamma$.
-\end{restatable}
-
-%Section~\ref{ssec:aut} discusses
-
-%When the data domain and priority domain are both fixed to be finite set, the register automata can be simulated with finite automata, and by Lemma \ref{lemma:reduce EPQ into state reachability}, checking linearizable w.r.t $\seqPQ$ is decidable.
-
-%
-%
-% and
-%
-%Let $A(\Gamma)$ be a $\Gamma$-complete automaton for each $\Gamma$. The following theorem
-%
-%
-%We will describe this automaton only in the case of $\Gamma=\mathsf{MatchedMaxPriority}$. The rest can be found in Appendix~\ref{}. These automata satisfy the following property
-%
-%the main technical challenge is to show that each of the following sets of executions can be characterized in terms of automata (modulo renaming of values):
-%
-%
-% define a set of register automata whose product with a given implementation $\mathcal{I}$ accepts a linearizability violation whenever $\mathcal{I}$ is not linearizable.
-%
-%are precise enough to recognize every non-linearizable implementation.
-%
-%In this section, we divide checking linearizability w.r.t $\seqPQ$ into checking linearizablity w.r.t $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$, each of which can be done in polynomial-time.
-Due to data-independence, violations to each $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$ can be captured by register automata.
-This reduces checking linearizability w.r.t $\seqPQ$ into state reachability problem.
-%
-%
-%
-%\subsection{Co-Regular}
-%\label{subsec:definition of co-regular}
-%
-
-%Let $\Gamma\mathsf{\text{-}Condition}(e)$ be $\mathsf{HasEmptyRemoves}(e)$, $\neg \mathsf{HasEmptyRemoves}(e) \wedge \mathsf{HasUnmatchedMaxPriority}(e)$, and $\mathsf{HasMatchedMaxPriority}(e)$, for $\Gamma = \mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}$, respectively.
-$\textit{Check-PQ-Conc}$ inspires us that, when $\Gamma\mathsf{\text{-}Condition}(e)$ holds, there should exist some $\alpha$, such that  $\Gamma\mathsf{\text{-}Conc}(e,\alpha)$ holds.
-The following lemma states this formally.
-Let $\textit{proj}(e)$ be the set of projections of $e$ into set of values.
-When refer to $\textit{proj}(e)$, we implicitly assume that each $\textit{rm}(\textit{empty})$ in $e$ has a ghost argument that is unique.
-We say that execution $l$ is linearizable w.r.t $\Gamma$, if we can find a sequence $s$ where $e \sqsubseteq s$ and $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$ holds for some $\alpha$.
-
-%\begin{restatable}{lemma}{EPQasMultiInMRpriforHistory}
-%\label{lemma:EPQ as multi in MRpri for history}
-%Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$, if and only if, $\forall e' \in \textit{proj}(e)$, $\forall \Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$,  $\Gamma\mathsf{\text{-}Condition}(e)$ $\Rightarrow$ $e$ is linearizable w.r.t $\Gamma$.
-%\end{restatable}
-
-%Note that due to data-independence, register automata are used to detect some renamed violations, instead of store and checking directly.
-Let us introduce the notion of co-regular, which reduce checking linearizable w.r.t $\Gamma\mathsf{\text{-}Conc}(e,\alpha)$ into checking set of witness automata.
-%
-%%\vspace{-6pt}
-%\begin{definition}\label{def:co-regular of rules of extended priority queues}
-%Given $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$.
-$\Gamma$ is co-regular, if there is a finite set $\textit{Auts}_{\Gamma}$ of register automata, such that for each data-independent implementation $\mathcal{I}$, we have that:
-%
-%$\textit{Auts}_{\Gamma} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $\Gamma\mathsf{\text{-}Condition}(e')$ holds but $e$ does not linearizable w.r.t $\Gamma$.
-%
-%$\seqPQ$ is co-regular, if each $\Gamma$ is co-regular.
-%\end{definition}
+Let $\mathcal{I}$ be a data-independent implementation. Then, there is a $\Gamma$-complete automaton $A(\Gamma)$ for each $\Gamma$. Moreover,
+$\mathcal{I} \sqsubseteq \seqPQ$ iff $\mathcal{I} \cap A(\Gamma) = \emptyset$ for all $\Gamma$.
+\end{theorem}
 
 \subsection{A $\mathsf{MatchedMaxPriority}$-complete automaton}\label{ssec:aut}
-
-%{\color {blue}When $\mathsf{MatchedMaxPriority}$ is false for some execution $e$, then there are two possibilities:
-%
-%\begin{itemize}
-%\setlength{\itemsep}{0.5pt}
-%\item[-] There exists value $x$, let $e_x$ be obtained from $e$ when all the values of maximal priorities in $e$ but $x$ are removed. Then $\mathsf{MatchedMaxPriority\text{-}Conc}(e_x,x)$ is false.
-%
-%\item[-] There exists set $X$ of values, let $e'$ be obtained from $e$ when all the values of maximal priorities in $e$ but values in $X$ are removed.
-Then, although $\mathsf{MatchedMaxPriority\text{-}Conc}(e'_x,x)$ is true for each $x \in X$, there is no value $y$ to make $\mathsf{MatchedMaxPriority\text{-}Conc}(e',y)$ to be true.
-%\end{itemize}
-%
-%It is surprised to see that these two possibilities can be captured by two sub-cases of $\mathsf{MatchedMaxPriority}$, depending on whether is only one value $x$ with maximal priority $p$.
-We denote by $\mathsf{MatchedMaxPriority}^{>}$ the strengthening of $\mathsf{MatchedMaxPriority}$ with this condition (i.e., all the values other than $x$ have a priority strictly smaller than $p$) and $\mathsf{MatchedMaxPriority}^{=}$ for the strengthening of the same formula with the negation of this condition (i.e., there exists another value $y$ with priority $p$).
-%}
 
@@
 	\item $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$ for reading call/return actions of a remove,
-	\item $\textit{call}(\textit{put},d,g)$ where $g\in\{=r,\prec r,true\}$ is a guard, for reading a call action $\textit{call}(\textit{put},d,p)$ of a put and checking whether $p$ is either equal to or smaller than the value stored in $r$, or arbitrary,
+	\item $\textit{call}(\textit{put},d,g)$ where $g\in\{=r,\prec r,true\}$ is a guard, for reading a call action $\textit{call}(\textit{put},d,p)$ and checking if $p$ is either equal to or smaller than the value stored in $r$, or arbitrary,
 	\item $\textit{ret}(\textit{put},d,true)$ for reading a return action $\textit{ret}(\textit{put},d,p)$ for any $p$.
 \end{itemize}
-The set of words accepted by such a transition system can be defined as usual.
-
-%$\textit{call}(\textit{put},d,g)$ $\textit{ret}(\textit{put},d,g),$ $\textit{call}(\textit{rm},d),\textit{ret}(\textit{rm},d)$, where $d \in \mathbb{D} \cup \{ \textit{empty} \}$, and $g$ and $\top$ are guards
-%
-% which are complete for each case. Since we don't require the full generality of register automata,
-%
-%
-%We define $\Gamma$-complete automata for both cases and their union will give a $\mathsf{MatchedMaxPriority}$-complete automaton
-%
-%
-%
-%Register automata is a kind of finite automata with registers used to deal with infinite data domain, where register can be assigned values and do simple operations.
-In this paper we use a restricted kind of register automata, which contains only one register, whose value is obtained once by guessing and never changed.
-Our register machine contains a register $r$.
-Its transition labels are $r=*,\textit{call}(\textit{put},d,g), \textit{ret}(\textit{put},d,g),$ $\textit{call}(\textit{rm},d),\textit{ret}(\textit{rm},d)$, where $d \in \mathbb{D} \cup \{ \textit{empty} \}$, and $g$ and $\top$ are guards.
-The first transition from initial state is a $r=*$ transition, and is used to assign an arbitrary value to $r$.
-A guard is a predicate chosen from $\{=r,< r,\top \}$, which represents equals the value of $r$, less than the value of $r$, and always holds, respectively.
-Given an execution $e = \alpha_1 \cdots \alpha_k$ of priority queue, register automata $\mathcal{A}$ accepts $e$, if there exists value $d_r\in \mathbb{D}$ and transitions $q_0 \xrightarrow{\beta_1} q_1 \ldots \xrightarrow{\beta_k} q_k$ of $\mathcal{A}$ from a initial state to a final state, such that for each $i$
-%
-%\begin{itemize}
-%\setlength{\itemsep}{0.5pt}
-%\item[-] If $\alpha_i = \textit{call}(\textit{put},a,p)$, then $\beta_i= \textit{call}(\textit{put},a,g)$ and $g$ holds with $p$ and $d_r$. Similarly for $\textit{ret}(\textit{put},a,p)$.
-%
-%\item[-] Else, $\alpha_i = \beta_i$.
-%\end{itemize}
+The set of sequences (executions) accepted by such a transition system is defined as usual.
 
@@
 
-%We use two point to simply our proof of co-regular.
-The first point is to use the results in \cite{Bouajjani:2015} to ensure FIFO (first in first out) property of each single-priority projections.
-A execution where only values of one priority is putted is called a single-priority execution.
-It is obvious that each single-priority projection of sequences in $\seqPQ$ satisfy FIFO property.
-Let $\textit{toQueue}(e)$ be an execution generated from $e$ by transforming $\textit{put}$ and $\textit{rm}$ into $\textit{enq}$ and $\textit{deq}$, respectively, and then discarding priorities.
-In Appendix \ref{sec:appendix proof and definition in section definition of co-regular}, according to the result of queue in \cite{Bouajjani:2015}, we construct a set $\textit{Auts}_{\textit{sinPri}}$ of witness automata, and shows that they are enough to ensure that for each data-differentiated executions, each of its single-priority projection without $\textit{rm}(\textit{empty})$ to have ``FIFO'' property, as shown by the following lemma.
-%
-%\begin{restatable}{lemma}{AutoForEPQwithSignlePri}
-%\label{lemma:automata for extended priority queue with single priority}
-%
-%Given a data-independent implementations $\mathcal{I}$ of extended priority queue, $\mathcal{I} \cap \textit{Auts}_{\textit{sinPri}} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{toQueue}(e')$ does not linearizable to queue.
-%\end{restatable}
-%
-%According to Lemma \ref{lemma:automata for extended priority queue with single priority}, from now on, it is safe to assume that, for each data-differentiated execution, any of its single-priority projection without $\textit{rm}(\textit{empty})$ has ``FIFO'' property.
-This effectively exclude some obvious violations, such as $\textit{rm}(a)$ happens before $\textit{put}(a,\_)$.
-
-%Lemma \ref{lemma:pri execution is enough} makes our co-regular proof more clear.
-One consequence of it is that for given $e\vert_{\preceq p}$, $\Gamma\mathsf{\text{-}Condition}(e')$ holds only for one $\Gamma$.
-%
-%
-%
-%
-%
-%
-%\subsection{Co-Regular of $\mathsf{MatchedMaxPriority}^{>}$}
-%\label{subsec:co-regular of EPQ1Lar}
-
-%To facilitate the proof of co-Regular of $\mathsf{MatchedMaxPriority}$, we separate $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ into the following two sub-predicates, where the first one represents that $x$ is the only value with priority $p$ in $e$ , and the second one represents that there are more than one value with priority $p$ in $e$.
-Here $\textit{occurOnce}(p,l)$ holds if there is only one value with priority $p$ in $l$.
-%
-%{\small
-%\begin{align*}
-%\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(e,x)=\mathsf{true}$,} \\
-%&\hspace{4mm}\mbox{$\textit{occurOnce}(p,e)$ where $p$ is the priority of $x$} \\
-%\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(e,x)=\mathsf{true}$,} \\
-%&\hspace{4mm}\mbox{$\neg\textit{occurOnce}(p,e)$ where $p$ is the priority of $x$} \\
-%\end{align*}
-%}
-%
-%$\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ is obtained from $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Conc}(e,x)$ by using $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(e,x)$ instead of $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$.
-And similarly for $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Conc}(e,x)$.
-%
-%We can also define two conditions, where $\textit{unmatched-priorities}(l)$ is the set of priorities occurring in $\textit{put}$ operations of $e$ for which there is $\textit{rm}$ operation removing the same value:
-%
-%{\small
-%\begin{align*}
-%\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Condition}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{matched-priorities}(e)$ for a maximal priority} \\
-%&\hspace{4mm}\mbox{$p\in priorities(e)$, $\textit{occurOnce}(p,e)$} \\
-%\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{matched-priorities}(e)$ for a maximal priority} \\
-%&\hspace{4mm}\mbox{$p\in priorities(e)$, $\neg\textit{occurOnce}(p,e)$}
-%\end{align*}}
-We give a typical example of an execution $e$ which is not $\mathsf{MatchedMaxPriority}^>$-linearizable in
-%Given a data-differentiated execution $e$ such that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Condition}(e)$ holds, Lemma \ref{lemma:automata for extended priority queue with single priority} is not enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$.
-This is because that it is possible that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ does not holds because of interaction between actions of multiple priorities.
-One such example is shown in
-\figurename~\ref{fig:introduce gap for EPQ1Lar}.
-Intuitively, this is a violation because during the whole execution of $\textit{rm}(b)$, the priority queue stores a smaller priority value (which should be removed before $b$).
-To be more precise, we define \emph{the interval of a value $x$} as the time interval from the return of a put $\textit{ret}(\textit{put},x,p)$ to the call of the matching remove $\textit{call}(rm,x)$, or to the end of the execution if such a call action doesn't exist.
-Intuitively, it represents the time interval in which a value is guaranteed to be stored into the concurrent priority queue.
-Concretely, for a standard indexing of actions in an execution, a time interval is a closed interval between the indexes of two actions in the execution.
-In \figurename~\ref{fig:introduce gap for EPQ1Lar}, we draw the interval of each value by dashed line.
-Here we assume that $p_1 \prec p_4$, $p_2 \prec p_4$, and $p_3 \prec p_4$.
-We can not find a sequence $s$ where $e \sqsubseteq s$ and $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,b)$ holds, since each time point from $\textit{call}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is included in the interval of some smaller priority value, and $\textit{rm}(b)$ can't take effect in the interval of a smaller priority value.
-
-\begin{figure}[htbp]
+\begin{figure}[t]
   \centering
   \includegraphics[width=0.4 \textwidth]{figures/PIC-HIS-INTRO-GAP-EPQ1L.pdf}
-%\vspace{-10pt}
   \caption{An execution that is not $\mathsf{MatchedMaxPriority}^{>}$-linearizable.
-We represent each operation as a time interval whose left, resp., right, bound corresponds to the call, resp., return action.
-Operations adding and removing the same value are aligned vertically.}
+We represent each operation as a time interval whose left, resp., right, bound corresponds to the call, resp., return action.}
   \label{fig:introduce gap for EPQ1Lar}
@@
 
-To formalize the scenario in \figurename~\ref{fig:introduce gap for EPQ1Lar} we use the notion of \emph{left-right constraint} defined below.
+Figure~\ref{fig:introduce gap for EPQ1Lar} contains a typical example of an execution $e$ which is not $\mathsf{MatchedMaxPriority}^>$-linearizable,
+where $p_1 \prec p_4$, $p_2 \prec p_4$, and $p_3 \prec p_4$.
+Intuitively, this is a violation because during the whole execution of $\textit{rm}(b)$, the priority queue stores a smaller priority value (which should be removed before $b$).
+To be more precise, we define \emph{the interval of a value $x$} as the time interval from the return of a put $\textit{ret}(\textit{put},x,p)$ to the call of the matching remove $\textit{call}(rm,x)$, or to the end of the execution if such a call action doesn't exist.
+This represents the time interval in which a value is guaranteed to be stored into the priority queue.
+Concretely, for a standard indexing of actions in an execution, a time interval is a closed interval between the indexes of two actions in the execution.
+In \figurename~\ref{fig:introduce gap for EPQ1Lar}, the interval of each value of priority smaller than $p_4$ is pictured as a dashed line.
+There is no sequence $l$ s.t.
+$e \sqsubseteq l$ and $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(l,b)$ hold, since each time point from $\textit{call}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is included in the interval of a smaller priority value,
+and $\textit{rm}(b)$ can't take effect in the interval of a smaller priority value.
+To formalize this scenario we use the notion of \emph{left-right constraint} defined below.
 
@@
 Let $e$ be a data-differentiated execution which contains only one maximal priority $p$, and only one value $x$ of priority $p$ (and no $\textit{rm}(\textit{empty})$ operations).
-% and let $\textit{put}(x,p)$ and $\textit{rm}(x)$ be the only operations with priority $p$.
-The \emph{left-right constraint of $x$} is the graph $G$ where: %$\textit{put}(x,p)$ and $\textit{rm}(x)$ is the graph $G$ where:
+The \emph{left-right constraint of $x$} is the graph $G$ where:
 \begin{itemize}
-\item the nodes are the values occurring in $e$, %, to which we add a node,
+\item the nodes are the values occurring in $e$,
 \item there is an edge from $d_1$ to $x$, if $\textit{put}(d_1,\_) <_{\textit{hb}} \textit{put}(x,p)$ or $\textit{put}(d_1,\_) <_{\textit{hb}} \textit{rm}(x)$,
@@
 
-The execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} is not $\mathsf{MatchedMaxPriority}^>$-linearizable because the left-right constraint of the maximal priority value $b$ contains a cycle: $f \rightarrow e \rightarrow d \rightarrow b \rightarrow f$.
-The presence of such a cycle is equivalent to the execution not being $\mathsf{MatchedMaxPriority}^>$-linearizable (see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}), as indicated by the following lemma:
+The execution in \figurename~\ref{fig:introduce gap for EPQ1Lar} is not $\mathsf{MatchedMaxPriority}^>$-linearizable because the left-right constraint of the maximal priority value $b$ contains the cycle $f \rightarrow d \rightarrow c \rightarrow b \rightarrow f$.
+The following lemma states that the presence of such a cycle is equivalent to \emph{non} $\mathsf{MatchedMaxPriority}^>$-linearizability, and it is proved in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}:
 
-\begin{restatable}{lemma}{LinEqualsConstraintforEPQOneLar}
+\begin{lemma}
 \label{lemma:Lin Equals Constraint for EPQ1Lar}
-Given a data-differentiated execution $e$ where $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, let $p$ be its maximal priority and $\textit{put}(x,p),\textit{rm}(x)$ are only operations of priority $p$ in $e$.
-Let $G$ be the graph representing the left-right constraint of $x$.
-$e$ is $\mathsf{MatchedMaxPriority}$-linearizable, if and only if $G$ has no cycle going through $x$.
-\end{restatable}
+Let $e$ be a data-differentiated execution such that
+$\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, $p$ is the maximal priority in $e$, and $\textit{put}(x,p)$ and $\textit{rm}(x)$ are only operations with arguments of priority $p$ in $e$.
+Then, $e$ is $\mathsf{MatchedMaxPriority}$-linearizable iff the left-right constraint of $x$ contains no cycle going through $x$.
+\end{lemma}
 
-%The existence of cycle though value with maximal priority in left-right constraint formally modelled the phenomenon in \figurename~\ref{fig:introduce gap for EPQ1Lar}.
-For example, in \figurename~\ref{fig:introduce gap for EPQ1Lar}, $f \rightarrow e \rightarrow d \rightarrow b \rightarrow f$ is a cycle in left-right constraint.
-We then prove that getting rid of cycle though value with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$.
-When the left-right constraint of the maximal priority value $x$ contains a cycle of the form $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ for some $d_1$,$\ldots$,$d_n\in \mathbb{D}$, we say that $x$ is \emph{covered} by $d_1,\ldots,d_m$.
-The shape of such a cycle (i.e., the alternation between call/return actions of $\textit{put}$/$\textit{rm}$ operations) can be detected using our class of automata, the only complication being the unbounded number of values $d_1$,$\ldots$,$d_n$.
+When the left-right constraint contains a cycle $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$, for some $d_1$,$\ldots$,$d_n\in \mathbb{D}$, we say that $x$ is \emph{covered} by $d_1,\ldots,d_m$.
+The shape of an execution witnessing such a cycle (i.e., the alternation between call/return actions) can be identified using our class of automata, the only complication being the unbounded number of values $d_1$,$\ldots$,$d_n$.
 However, by data independence, whenever an implementation contains such an execution it also contains an execution where all the values $d_1$,$\ldots$,$d_n$ are renamed to the same value $a$, and $x$ is renamed to $b$.
-Therefore, our automata can be defined over a fixed set of values $a$, $b$, and $\top$ (recall that $\top$ is used for the operations outside of the non-linearizable projection).
+Therefore, our automata can be defined over a fixed set of values $a$, $b$, and $\top$ (recall that $\top$ is used for operations outside of the non-linearizable projection).
 
-%TODO WHY IT IS A SET OF AUTOMATA ?
-SAY WHAT ARE THE CASES THAT NEED TO BE COVERED.
-SAY WHAT IS THE CASE WE GIVE IN THE FIGURE.
-USE $\top$ AS MENTIONED AT THE BEGINNING OF THE SECTION INSTEAD OF 	$c$.
-USE BIG CAPS FOR SETS OF TRANSITION LABELS.
-To define a $\mathsf{MatchedMaxPriority}^>$-complete automaton we need to consider several cases depending on the order between the call/return actions of the $\textit{put}$/$\textit{rm}$ operations that add and respectively, remove the value $b$.
-For example, the case where the put happens-before the remove (as in \figurename~\ref{fig:introduce gap for EPQ1Lar}) is pictured in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
+To define a $\mathsf{MatchedMaxPriority}^>$-complete automaton, we need to consider all the possible orders between the call/return actions of the $\textit{put}$/$\textit{rm}$ operations that add and respectively, remove the value $b$.
+The case where the put happens-before the remove (as in \figurename~\ref{fig:introduce gap for EPQ1Lar}) is pictured in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
 This automaton captures the three possible ways of ordering the first action $\textit{ret}(\textit{put},a,\_)$ w.r.t.
-the actions with value $b$, which are pictured in \figurename~\ref{fig:executions APQ1Lar-1 in paper} (a) (this action cannot occur after $\textit{call}(\textit{rm},b,\_)$ since $b$ must be covered by the $a$-s).
-The paths corresponding to these three possible orders are: $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_{10}$ and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$.
-In \figurename~\ref{fig:executions APQ1Lar-1 in paper}, we show all the four orders of the call/return actions of adding and removing $b$, and also possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$.
+the actions with value $b$, which are pictured in \figurename~\ref{fig:executions APQ1Lar-1 in paper}(a) (this action cannot occur after $\textit{call}(\textit{rm},b,\_)$ since $b$ must be covered by the $a$-s).
+The paths corresponding to these three possible orders are: $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_{10}$, and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$.
+\figurename~\ref{fig:executions APQ1Lar-1 in paper} lists the four possible orderings of the call/return actions of adding and removing $b$, and also possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$.
+Each such ordering corresponds to an automaton similar to the one in \figurename~\ref{fig:automata APQ1Lar-1 in paper}, their union defining a $\mathsf{MatchedMaxPriority}^>$-complete automaton.
 In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, three register automata is constructed according to the cases of \figurename~\ref{fig:executions APQ1Lar-1 in paper} (b), (c) and (d), respectively.
 
-%For each case, we define an automaton recognizing the induced set of violations.
-%For example, the automaton corresponding to the case ??
-is pictured in in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
-More details can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
-%
-%
-%{\color {blue}
-%Defining a $\mathsf{MatchedMaxPriority}^>$-complete automaton needs to capture the executions when a value $x$ is covered by values with smaller priority.
-By data-independence, we rename $x$ into $b$, rename the items covers it into $a$ and rename all other value into $\top$.
-There are four possibility of projection of captured execution into $b$, and for clarity we construct a register automaton for each of these case.
-For example, one such case is as in \figurename~\ref{fig:introduce gap for EPQ1Lar} (consider $b$ as $x$), and the corresponding register automaton is shown in \figurename~\ref{fig:automata APQ1Lar-1 in paper}.
-There are three possible place of the first $\textit{call}(\textit{put},a,\_)$ in the captured execution, as shown in \figurename~\ref{fig:executions APQ1Lar-1 in paper}.
-These three possible places correspond to three paths in $\mathcal{A}_{\textit{l-lar}}^1$: $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_7$, $q_1 \rightarrow q_2 \rightarrow q_3 \ldots \rightarrow q_{10}$ and $q_1 \rightarrow q_9 \rightarrow q_{10} \ldots \rightarrow q_7$.
-More details can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
-%}
-
-\begin{figure}[htbp]
+\begin{figure}[t]
   \centering
-  \includegraphics[width=1 \textwidth]{figures/PIC_AUTO_PQ1Lar-pprr.pdf}
-%\vspace{-10pt}
-  \caption{Register automaton $\mathcal{A}_{\textit{l-lar}}^1$.
-We use the following notations: $C_1 = C \cup \{ \textit{ret}(\textit{rm},a) \}$, $C_2 = C \cup \{ \textit{call}(\textit{put},a,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},d)$, $\textit{ret}(\textit{rm},d),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.}
-  \label{fig:automata APQ1Lar-1 in paper}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  %\includegraphics[width=1 \textwidth]{figures/PIC_HIS_PQ1Lar-pprr-in-paper.pdf}
-  \includegraphics[width=1 \textwidth]{figures/PIC_HIS_PQ1Lar-fouCase.pdf}
-%\vspace{-10pt}
-  %\caption{Three possible positions of the first $\textit{call}(\textit{put},a,\_)$ in captured execution}
-  \caption{Four cases of ordering actions with value $b$, and possible orders of the first $\textit{ret}(\textit{put},a,\_)$ w.r.t the actions with value $b$}
+  \includegraphics[width=.8\textwidth]{figures/PIC_HIS_PQ1Lar-fouCase.pdf}
+  \caption{Orderings to be considered when defining a $\mathsf{MatchedMaxPriority}^>$-complete automaton.}
   \label{fig:executions APQ1Lar-1 in paper}
@@
 
-%{\color {blue}
-%Let $\mathcal{A}_{\textit{1-lar}}$ be the union of the four register automata mentioned above, then we prove that $\mathcal{A}_{\textit{1-lar}}$ is $\mathsf{MatchedMaxPriority}^{>}$-complete.
-The detailed construction and proof can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}.
-%}
-%
-%\begin{restatable}{lemma}{EPQOneLarisCoRegular}
-%\label{lemma:EPQ1Lar is co-regular}
-%$\textit{Auts}_{\textit{1-lar}}$ is $\mathsf{MatchedMaxPriority}^{>}$-complete.
-%\end{restatable}
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/PIC_AUTO_PQ1Lar-pprr.pdf}
+  \caption{A register automaton capturing the scenario in Figure~\ref{fig:executions APQ1Lar-1 in paper}(a).
+We use the following notations: $A_1 = A \cup \{ \textit{ret}(\textit{rm},a) \}$, $A_2 = A \cup \{ \textit{call}(\textit{put},a,=r) \}$, $A_3 = A_2 \cup \{ \textit{ret}(\textit{rm},a) \}$, where $A = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true}), \textit{call}(\textit{rm},\top)$, $\textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$.}
+  \label{fig:automata APQ1Lar-1 in paper}
+\end{figure}
 
@@
 
-%In this subsection, we briefly introduce the idea for proving co-regular of $\mathsf{MatchedMaxPriority}^{=}$.
-The proof of this subsection can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}.
-%Given a data-differentiated $\_$-execution $e$ such that $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)$ holds, Lemma \ref{lemma:automata for extended priority queue with single priority} and $\textit{Auts}_{\textit{1-lar}}$ are sill not enough for ensure that $e$ linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$.
 When an execution contains at least two values of maximal priority, the acyclicity of the left-right constraints (for all the maximal priority values) is not enough to conclude that the execution is $\mathsf{MatchedMaxPriority}$-linearizable.
 Intuitively, there may exist a value $a$ which is added before another value $b$ such that all the possible linearization points of $\textit{rm}(b)$ are disabled by the position of $\textit{rm}(a)$ in the happens-before.
-We give an example of such an execution $e$ in \figurename~\ref{fig:introduce pb order}.
+We give an example of such an execution $e$ in \figurename~\ref{fig:introduce pb order}, where $p_1\prec p_4$.
 This execution  is not linearizable w.r.t.
 $\mathsf{MatchedMaxPriority}$ (or $\mathsf{MatchedMaxPriority}^{=}$) even if
-%$h \vert_{\prec p_1}$ and $h \vert_{p_4}$ both linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$, and
-neither $a$ nor $b$ are covered by values with smaller priority. %The explanation of $e$ not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ is as follows:
+neither $a$ nor $b$ are covered by values with smaller priority.
 Since $\textit{put}(a,p_4) <_{\textit{hb}} \textit{put}(b,p_4)$ and values of the same priority are removed in FIFO order, $\textit{rm}(a)$ should be linearized before $\textit{rm}(b)$ (i.e., this execution should be linearizable w.r.t.
 a sequence where $\textit{rm}(a)$ occurs before $\textit{rm}(b)$).
-%According to the previous section, which says that linearizaton point of larger priority can be happen in the interval of values with smaller priority.
-Since $\textit{rm}(b)$ cannot take effect during the interval of a smaller priority value, it could be linearized only in one of the two time intervals pictured with dotted lines in \figurename~\ref{fig:introduce pb order}.
-However, each of these time intervals ends before $\textit{call}(\textit{rm},a)$, and thus $\textit{rm}(a)$ cannot be linearized before $\textit{rm}(b)$.
+Since $\textit{rm}(b)$ cannot take effect during the interval of a smaller priority value, it could be only linearized in one of the two time intervals pictured with dotted lines in \figurename~\ref{fig:introduce pb order}.
+However, each of those time intervals ends before $\textit{call}(\textit{rm},a)$, and thus $\textit{rm}(a)$ cannot be linearized before $\textit{rm}(b)$.
 
-\begin{figure}[htbp]
+\begin{figure}[t]
   \centering
-  \includegraphics[width=0.6 \textwidth]{figures/PIC-HIS-INTRO-PB-ORDER-EPQ.pdf}
-%\vspace{-10pt}
-  \caption{An execution that is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^=$.}
+  \includegraphics[width=0.5 \textwidth]{figures/PIC-HIS-INTRO-PB-ORDER-EPQ.pdf}
+  \caption{An execution that is not $\mathsf{MatchedMaxPriority}^=$-linearizable.}
   \label{fig:introduce pb order}
@@
 To recognize the scenarios in \figurename~\ref{fig:introduce pb order}, we introduce an order $<_{\textit{pb}}$ between values which intuitively, can be thought of as ``a value $a$ is put before another value $b$''.
-%Let us introduce put-before order $<_{\textit{pb}}$ to formally state that ``an value is putted before another value''.
-Thus, given a data-differentiated execution $e$ and two values $a$ and $b$ of maximal priority, $a <_{\textit{pb}} b$ if one of the following holds: (1) $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$, (2) $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(b)$, or (3) $\textit{rm}(a) <_{\textit{hb}} \textit{put}(b,\_)$.
-Sometimes we use $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$ and $a <_{\textit{pb}}^C b$ to explicitly distinguish between these three cases.
+More precisely, given a data-differentiated execution $e$ and two values $a$ and $b$ of maximal priority, $a <_{\textit{pb}} b$ if one of the following holds: (1) $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$, (2) $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(b)$, or (3) $\textit{rm}(a) <_{\textit{hb}} \textit{put}(b,\_)$.
+Sometimes we use $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$, and $a <_{\textit{pb}}^C b$ to explicitly distinguish between these three cases.
 Let $<_{\textit{pb}}^*$ be the transitive closure of $<_{\textit{pb}}$.
-%Intuitively, $a <_{\textit{pb}}^* b$ means that we can infer that $a$ is putted before $b$ with the help of several intermediate values.
 
-Then, to model the time intervals in which a remove operation like $\textit{rm}(b)$ in \figurename~\ref{fig:introduce pb order}, can be linearized (outside of intervals of smaller priority values) we introduce the notion of gap-point.
-Here we assume that the index of actions of an execution starts from $0$.
+To define the time intervals in which a remove like $\textit{rm}(b)$ in \figurename~\ref{fig:introduce pb order}, can be linearized (outside of intervals of smaller priority values) we use the notion of gap-point.
+As before, defining time intervals relies on an indexing of actions in an execution, starting with 0.
 
 \begin{definition}\label{def:gap-point for matched put and rm operations}
-Let $e$ be a data-differentiated execution which contains only one maximal priority $p$, and $\textit{put}(x,p)$ and $\textit{rm}(x)$ two operations in $e$.
-An index $i\in [0,|e|-1]$ is a \emph{gap-point of $x$} if $i$ is greater than or equal to the index of both $\textit{call}(\textit{put},x,p)$ and $\textit{call}(\textit{rm},x)$, smaller than the index of $\textit{ret}(\textit{rm},x)$, and it is not included in the interval of some value with priority smaller than $p$.
+Let $e$ be a data-differentiated execution with only one maximal priority $p$, and $\textit{put}(x,p)$ and $\textit{rm}(x)$ two operations in $e$.
+An index $i\in [0,|e|-1]$ is a \emph{gap-point of $x$} if $i$ is greater than or equal to the index of both $\textit{call}(\textit{put},x,p)$ and $\textit{call}(\textit{rm},x)$, smaller than the index of $\textit{ret}(\textit{rm},x)$, and not included in the interval of a value with priority smaller than $p$.
 \end{definition}
@@
 The case of \figurename~\ref{fig:introduce pb order} can be formally described as follows: $a <_{\textit{pb}}^* b$ while the right-most gap-point of $b$ is before $\textit{call}(\textit{rm},a)$ or $\textit{call}(\textit{put},a,p_4)$.
-The following lemma states that these conditions are enough to characterize non-linearizability w.r.t $\mathsf{MatchedMaxPriority}^{=}$ (see Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}).
+The following lemma states that these conditions are enough to characterize non-linearizability w.r.t.
+$\mathsf{MatchedMaxPriority}^{=}$.
+It is proved in Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}.
 
-\begin{restatable}{lemma}{EPQOneEqualAsPBandGP}
+\begin{lemma}
 \label{lemma:EPQ1Equal as pb order and gap-point}
-Let $e$ be a data-differentiated execution which contains only one maximal priority $p$ such that $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds.
-%Given a data-differentiated $p$-execution $e$ where $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)$ holds.
-Then, $e$ is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ iff $e$ contains two values $x$ and $y$ of maximal priority $p$ such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is strictly smaller than the index of $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$.
-\end{restatable}
+Let $e$ be a data-differentiated execution with only one maximal priority $p$ such that $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds.
+Then, $e$ is not $\mathsf{MatchedMaxPriority}^{=}$-linearizable iff $e$ contains two values $x$ and $y$ of maximal priority $p$ such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is strictly smaller than the index of $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$.
+\end{lemma}
 
-%\begin {proof} (Sketch)
-%We have already gave an intuitive explanation for the $\textit{if}$ direction.
-To prove the $\textit{only if}$ direction, we prove its contrapositive.
-%, where we already know that $\forall x, y$ with maximal priority, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{cal}(\textit{put},y,\textit{pri})$ and $\textit{cal}(\textit{rm},x)$.
-We need to prove that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$.
-%Or we can say, we need to explicitly construct linearization of $e$.
-%
-%Let $\textit{values}(e,p)$ be the set of values with priority $p$ in $e$.
-We introduce another lemma (Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal} in Appendix), which states that if we can find a value $\exists g_1 \in \textit{values}(e,p)$, such that $\forall g_2 \in \textit{values}(e,p)$, (1) $g_1 \nless_{\textit{pb}} g_2$, and (2) the right-most gap-point of $g_1$ is after $\textit{call}(\textit{put},g_2,p)$ and $\textit{call}(\textit{rm},g_2)$.
-Then $e$ is linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$.
-%The remaining of our proof is a loop for searching such $g_1$ in $e$.
-Let $e_p$ be the projection of $e$ into values of priority $p$, and let $l_p$ be the linearization of $e$ which satisfy the requirements of queue.
-Let the initial value of $a_1$ be the last inserted value in $l_p$.
-%
-%\begin{itemize}
-%\setlength{\itemsep}{0.5pt}
-%\item[-] check if $a_1$ satisfy the conditions of Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}. If it does, then this loop terminates.
-%
-%\item[-] Otherwise, $\exists a_2 \in \textit{values}(e,p)$, such that the rightmost gap-point of $a_1$ is before $\textit{cal}(\textit{put},a_2,p)$ or $\textit{cal}(\textit{rm},a_2)$ in $e$.
-By assumption, we know that $a_2 \nless_{\textit{pb}} a_1$, and each gap-point of $a_2$ is after the rightmost gap-point of $a_1$.
-%
-%    \begin{itemize}
-%    \setlength{\itemsep}{0.5pt}
-%    \item[-] If $\forall b \in \textit{values}(e,p)$, $a_2\not<_{\textit{pb}}b$. Then we go to next round of the loop and make $a_2$ to be the ``new-round $a_1$''.
-%    \item[-] Otherwise, there exists $a_3 \in \textit{values}(e,p)$ such that $a_2 <_{\textit{pb}}^* a_3$.
-It is not hard to see that there is no cycle in $<_{\textit{pb}}$, and it is safe to assume that $a_3$ is maximal w.r.t $<_{\textit{pb}}^*$.
-%
-%        By assumption,we know that the rightmost gap-point of $a_3$ is after $\textit{call}(\textit{put},a_2,p)$ and $\textit{call}(\textit{rm},a_2)$.
-Therefore, we can see that the rightmost gap-point of $a_3$ is after the rightmost gap-point of $a_1$.
-Then we go to next round of the loop and make $a_3$ to be the ``new-round $a_1$''.
-%    \end{itemize}
-%\end{itemize}
-%
-%Let $a^i$ be the $a_1$ in the $\textit{i-th}$ round of loop.
-We can see that the rightmost gap-point of $a^j$ is after the rightmost gap-point of $a^i$ for each $i<j$.
-Therefore, the loop finally terminates at some $a^f$ that satisfies the condition of Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}.
-This implies that $e$ is linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ and completes the proof of the $\textit{only if}$ direction.
-%\end {proof}
+The following shows that the number of values needed to witness that $y <_{\textit{pb}}^* x$, for some $x$ and $y$, is bounded.
 
-To characterize violations to linearizability w.r.t.
-$\mathsf{MatchedMaxPriority}^{=}$ using an automaton that tracks a bounded number of values, we show that the number of values needed to witness that $y <_{\textit{pb}}^* x$ for some $x$ and $y$ is bounded.
-%
-%While Lemma \ref{lemma:EPQ1Equal as pb order and gap-point} doesn't imply a bound for
-%the number of values needs to witness that
-%
-%The result of Lemma \ref{lemma:EPQ1Equal as pb order and gap-point} still needs to be impove, since obtaining $y <_{\textit{pb}}^* x$ may requires arbitrary many intermediate values, and it is hard to store unbounded values by automata.
-Fortunately, we prove that the number of intermediate values is essentially bounded, as stated by the following lemma.
-
-\begin{restatable}{lemma}{OBOrderHasBoundedLength}
+\begin{lemma}
 \label{lemma:ob order has bounded length}
@@
 \item[-] $a <_{\textit{pb}}^A a_i <_{\textit{pb}}^B b$ or $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A b$, for some $i$.
-%\item[-] $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A a_j <_{\textit{pb}}^B b$, for some $i$ and $j$,
 \end{itemize}
-\end{restatable}
+\end{lemma}
 
-\smallskip
-To characterize violations to linearizability w.r.t.
-$\mathsf{MatchedMaxPriority}^{=}$, one has to consider all the possible orders between call/return actions of the operations on values $a$, $b$, and $a_i$ in Lemma \ref{lemma:ob order has bounded length}, and the right-most gap point of $b$.
-Excluding the inconsistent cases, we are left with 5 possible orders that are shown in \figurename~\ref{fig:five enumerations}, where $o$ denotes the rightmost gap-point of $b$.
-For each case, we define an automaton recognizing the induced set of violations.
-For instance, the register automata for the case of \figurename~\ref{fig:five enumerations}(a) is shown in \figurename~\ref{fig:an enumeration and its witness automaton}.
-In this case, the conditions in Lemma~\ref{lemma:EPQ1Equal as pb order and gap-point} are equivalent to the fact that intuitively, the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$ is covered by lower priority values (and thus, there is no gap-point of $b$ which occurs after $\textit{call}(\textit{rm},a)$).
-Using again the data-independence property, these lower priority values can all be renamed to a fixed value $d$, and the other values to a fixed value $\top$.
-
-%We denote by $o$ the rightmost gap-point of $b$, we rename the values that ``cover the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$'' (see Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}) into $d$, and rename the renaming values into $e$.
-
-\begin{figure}[htbp]
+\begin{figure}[t]
   \centering
   \includegraphics[width=0.8 \textwidth]{figures/PIC-HIS-FiveEnumerations.pdf}
-%\vspace{-10pt}
-  \caption{Five cases that need to be considered.}
+  \caption{Orderings to be considered when defining a $\mathsf{MatchedMaxPriority}^=$-complete automaton.
+We omit the operations which can be ordered arbitrarily, e.g., $\textit{put}(b)$ in the cases (d) and (e).}
   \label{fig:five enumerations}
@@
 
-\begin{figure}[htbp]
+To characterize violations to $\mathsf{MatchedMaxPriority}^{=}$-linearizability, one has to consider all the possible orders between call/return actions of the operations on values $a$, $b$, and $a_i$ in Lemma \ref{lemma:ob order has bounded length}, and the right-most gap point of $b$.
+Excluding the inconsistent cases, we are left with the five orders in \figurename~\ref{fig:five enumerations}, where $o$ denotes the rightmost gap-point of $b$ (see Lemma \ref{lemma:five enumeration is enough for EPQ1Equal} in Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}).
+For each case, we define an automaton recognizing the induced set of violations.
+The register automaton for the case in \figurename~\ref{fig:five enumerations}(a) is shown in \figurename~\ref{fig:an enumeration and its witness automaton}.
+In this case, Lemma~\ref{lemma:EPQ1Equal as pb order and gap-point} is equivalent to the fact that intuitively, the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$ is covered by lower priority values (thus, there is no gap-point of $b$ which occurs after $\textit{call}(\textit{rm},a)$).
+By data-independence, these lower priority values can be renamed to a fixed value $c$.
+In Appendix~\ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}, we give the detailed construction for register automata for all five orders in \figurename~\ref{fig:five enumerations}.
+
+\begin{figure}[t]
   \centering
-  %\includegraphics[width=0.8 \textwidth]{figures/PIC-Enumeration-WitnessAutomata.pdf}
-  \includegraphics[width=0.8 \textwidth]{figures/PIC-WitnessAutomata-For1.pdf}
-%\vspace{-10pt}
-  \caption{A case in deriving a $\mathsf{MatchedMaxPriority}^{=}$-complete automaton.
-We use the following notations: $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true})$, $\textit{call}(\textit{rm},\top), \textit{ret}(\textit{rm},\top),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_1 = C \cup \{ \textit{call}(\textit{put},d,<r) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{rm},d) \}$, $C_4 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},d) \}$.}
+  \includegraphics[width=0.65 \textwidth]{figures/PIC-WitnessAutomata-For1.pdf}
+  \caption{A register automaton for the case in Figure~\ref{fig:five enumerations}(a), where $A_1 = A \cup \{ \textit{call}(\textit{put},c,<r)\}$, $A_2 = A_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $A_3 = A_2 \cup \{ \textit{ret}(\textit{rm},c) \}$, $A_4 = A \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},c) \}$, where $A = \{ \textit{call}(\textit{put},\top,\textit{true})$,$\textit{ret}(\textit{put},\top,\textit{true})$,$\textit{call}(\textit{rm},\top)$,$\textit{ret}(\textit{rm},\top)$,$\textit{call}(\textit{rm},\textit{empty})$,$\textit{ret}(\textit{rm},\textit{empty})\}$.}
   \label{fig:an enumeration and its witness automaton}
 \end{figure}
-
-%we can use register automata to detect $a <_{\textit{pb}}^* b$ by enumerating all possible enumerations of operations of $a$, $b$ and possibly $a_1$.
-Since some combinations of $<_{\textit{pb}}^*$ and gap-points are conflict, we finally reduce the number of potential enumerations of operations of $a$, $b$ and possibly $a_1$ into only five in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}.
-
-%In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}, we construct a set $\textit{Auts}_{\textit{1-eq}}$ of register automata, and prove that $\mathsf{MatchedMaxPriority}^{=}$ is co-regular, as stated by the following lemma.
-%
-%\begin{restatable}{lemma}{EPQOneEqualIsCoRegular}
-%\label{lemma:EPQ1Equal is co-regular}
-%$\mathsf{MatchedMaxPriority}^{=}$ is co-regular.
-%\end{restatable}
 
@@
 The implementations in $\mathcal{C}$ allow an unbounded number of values but a bounded number of priorities.
-Each method manipulates a finite number of local variables which store Boolean values, or data values from $\mathbb{D}$.
-Methods communicate through a finite number of shared variables that also store Boolean values, or data values from $\mathbb{D}$.
-Data values may be assigned, but never used in program predicates (e.g., in the conditions of if-then-else and while statements) so as to ensure data independence.
+Each method has a finite set of local variables storing Boolean values or values from $\mathbb{D}$.
+Methods communicate through a finite number of shared variables interpreted also as Booleans or values from $\mathbb{D}$.
+To ensure data independence, values in $\mathbb{D}$ may be assigned, but never used in Boolean expressions (e.g., of if-then-else statements).
 This class captures typical implementations, or finite-state abstractions thereof, e.g., obtained via predicate abstraction.
-Since the $\Gamma$-complete automata $A(\Gamma)$ uses a fixed set $D=\{a,b,a_1,d,e,\top\}$ of values, we have that $\mathcal{C}\cap A(\Gamma)\neq\emptyset$ for some $\Gamma$ iff $\mathcal{C}_D\cap A(\Gamma)\neq\emptyset$ where $\mathcal{C}_D$ is the subset of $\mathcal{C}$ that uses only values in $D$.
+The $\Gamma$-complete automata we define use a fixed set $D=\{a,b,c,a_1,\top\}$ of values ($a_1$ is needed to deal with the second item in Lemma~\ref{lemma:ob order has bounded length}).
+Therefore, for any $\Gamma$, $\mathcal{C}\cap A(\Gamma)\neq\emptyset$ iff $\mathcal{C}_D\cap A(\Gamma)\neq\emptyset$, where $\mathcal{C}_D$ is the subset of $\mathcal{C}$ that uses only values in $D$.
 
-The set of executions $\mathcal{C}_D$ can be represented by a Vector Addition Systems with States (VASS), since both values and priorities are finite, which implies that each thread and register automata can be transformed into finite-state automata.
-To obtain this transformation, the states of the VASS represent the global variables of $\mathcal{C}_D$, while each counter of the VASS then represents the number of threads which are at a particular control location within a method, with a certain valuation of the local variables.
-In this way we transform linearizability problem of priority queue into state-reachability problem of VASS, which is EXPSPACE-complete.
-When we consider only finite number of threads, the value of counters are bounded and then this problem is PSPACE \cite{DBLP:conf/ac/Esparza96}.
+The set of executions $\mathcal{C}_D$ can be represented by a Vector Addition System with States (VASS).
+Since the set of values and priorities is bounded, each method invocation can be represented by a finite-state automaton (see~\cite{conf/esop/BouajjaniEEH13}).
+For a fixed set of priorities $P\subseteq \mathbb{P}$, the register automata $A(\Gamma)$ can be transformed to finite-state automata (the number of valuations of the registers is bounded).
+Thus, checking linearizability of an implementation in $\mathcal{C}$ is PSPACE when the number of threads is bounded, and EXPSPACE, otherwise.
+Moreover, reachability in VASSs can be reduced to checking linearizability of such an implementation.
+Essentially, given an instance of the VASS reachability problem, one can define a priority queue implementation where the $\textit{put}$ methods behave correctly and additionally, they include the code of the VASS simulation defined in~\cite{conf/esop/BouajjaniEEH13}, and the $\textit{rm}$ methods behave correctly, except for the moment where the target state is reached, in which case they trigger a linearizability violation by returning an arbitrary value.
 
-On the other hand, we can mimic transitions of VASS by a priority queue implementation.
-To obtain this, we keep $\textit{rm}$ unchanged, while in each $\textit{put}$ method, we first put a value into priority queue, and then either simulate one step of transitions or record one increase/decrease of a counter, similarly as that in \cite{conf/esop/BouajjaniEEH13}.
-In this way we transform the state-reachability problem of VASS into linearizability problem of priority queue.
-Based on above discussion, we have the following complexity result.
-The detailed proof can be found in Appendix \ref{subsec:appendix proof and definition in subsection decidability result}.
+Based on above discussion, we have the following complexity result. The detailed proof can be found in Appendix \ref{subsec:appendix proof and definition in subsection decidability result}.
 
@@
 \end{theorem}
-
-%We also prove that $\mathsf{UnmatchedMaxPriority}$ are co-regular in Appendix \ref{subsec:appendix co-regular of EPQ2Lar}, Appendix \ref{subsec:appendix co-regular of EPQ2Equal}, and prove that $\mathsf{EmptyRemove}$ are co-regular in Appendix \ref{subsec:co-regular of EPQ3}.
-Then we can see that $\seqPQ$ is co-regular.
-%
-%\begin{restatable}{lemma}{EPQIsCoRegular}
-%\label{lemma:EPQ is co-regular}
-%$\seqPQ$ is co-regular.
-%\end{restatable}
-%
-%By Lemma \ref{lemma:EPQ as multi in MRpri for history} and Lemma \ref{lemma:EPQ is co-regular}, we can finally reduce the verification of linearizability w.r.t $\seqPQ$ into a reachability problem, as illustrated by the following theorem, where $\textit{Auts}_{\textit{EPQ}}$ is the set of witness automata of this section.
-%
-%\begin{restatable}{theorem}{ReduceEPQIntoStateReachability}
-%\label{lemma:reduce EPQ into state reachability}
-%Given a data-independence implementation $\mathcal{I}$. $\mathcal{I} \sqsubseteq \seqPQ$, if and only if, $\mathcal{I} \cap \textit{Auts}_{\textit{EPQ}} = \emptyset$.
-%\end{restatable}
-%
-%When the data domain and priority domain are both fixed to be finite set, the register automata can be simulated with finite automata, and by Lemma \ref{lemma:reduce EPQ into state reachability}, checking linearizable w.r.t $\seqPQ$ is decidable.

### 2017-07-13 priority-queues-arxiv.tex
arxiv

@@
 In this section, we formalize (concurrent) executions and implementations, introduce a set of properties satisfied by all the implementations we are aware of, and recall the standard correctness criterion for concurrent implementations of ADTs known as \emph{linearizability}~\cite{journals/toplas/HerlihyW90}.
-%
-%recall the notion of \emph{linearizability}, the standard correctness criterion for concurrent implementations of priority queues (
-%
-%Concurrent implementations of the priority queue allow the methods $\textit{put}$ and $\textit{rm}$ to be called concurrently from different threads.
-However, every method invocation should give the illusion that it takes place instantaneously at some point between its invocation and its return.
-We formalize (concurrent) executions and implementations in Section~\ref{ssec:exec}, Section~\ref{ssec:semantic_prop} introduces a set of properties satisfied by all the  implementations we are aware of, and Section~\ref{ssec:lin} defines the standard correctness criterion for concurrent implementations of ADTs known as \emph{linearizability}~\cite{journals/toplas/HerlihyW90}.
 
@@
 Then, $\seqPQ$ is the set of traces (words) accepted by $\textit{PQ}$.
-%}{\color {red}$\seqPQ$ is the set of traces of $\textit{PQ}$}.
-{\color {red}The detailed definition of $\textit{PQ}$ can be found in the long version \cite{CONCUR2017Ahmed}.}
 The detailed definition of $\textit{PQ}$ can be found in Appendix \ref{sec:appendix definition of seqPQ and proof of Lemma EQP rules and semantics}.
@@
 This simplification is sound when the method invocations can always make progress in isolation.
-%formally, for any execution $e$ with pending operations, there exists an execution $e'$ obtained by extending $e$ only with the return actions of the pending operations of $e$.
-%Intuitively this means that methods can always return without any help from outside threads, avoiding deadlock.
-
-%a finite set $\mathbb{M}$ of methods, and an infinite set $\mathbb{O}$ of operation (identifiers).
-Given $m \in \mathbb{M}$, $x \in \mathbb{D}$, and $o \in \mathbb{O}$, a call action $\textit{call}_o (m,x)$ represents an invocation to method $m$ with argument $x$, while a return action $\textit{ret}_o (m,x)$ represents a response from method $m$ with return value $x$.
-\footnote{Call actions with more than one arguments are similarly defined.} Here $o$ is used to match return actions to their call actions: $\textit{cal}_o (m,x)$ matches $\textit{ret}_{o'} (m',x')$, if $o=o' \wedge m=m'$.
-A sequential execution is a sequence of call and return actions, while each call action is immediately followed by its matching return action.
-Let $\cdot$ be the concatenation of sequences.
-To ease the reading, given a sequential execution $e=\textit{call}_{o_1}(m_1,a_1) \cdot \textit{rm}_{o_1}(m_1,b_1) \cdot \ldots \cdot \textit{call}_{o_n}(m_n,a_n) \cdot \textit{rm}_{o_n}(m_n,b_n)$, when the context is clear, we can write it as $e=m_1(a_1,b_1) \cdot \ldots \cdot m_n(a_n,b_n)$.
-We write $m(a,b)$ as $m(a)$ or $m(b)$, when there is no return value or there is no argument, respectively.
-
-%A priority queue with partially-ordered priorities (priority queue for short) contains two method: $\textit{put}$ and $\textit{rm}$.
-A $\textit{put}$ method has two arguments, while the first argument is an item and the second argument is its priority.
-A $\textit{put}$ method is used to put an item into the priority queue with certain priority.
-Here the item is chosen from a specific (possibly infinite) data domain $\mathbb{D}$ and priority is chosen from a (possibly infinite) set $\mathbb{P}$.
-Moreover, there is a strict partial-order $\prec$ among elements in $\mathbb{P}$.
-A $\textit{rm}$ method intends to remove the item with minimal priority (w.r.t $\prec$) in priority queue and then returns it.
-It works as follows:
-
-%\begin{itemize}
-%\setlength{\itemsep}{0.5pt}
-%\item[-] If the priority queue is empty, then $\textit{rm}$ returns $\textit{empty}$.
-%
-%\item[-] Else, $\textit{rm}$ choose one of minimal priority of items in priority queue, and returns the earliest putted item of this priority.
-Note that if there are more than one incomparable and minimal candidate priorities, then the chosen of priority is arbitrary.
-%\end{itemize}
-
-%Let us introduce the notion of sequential executions, which modelled the specification of concurrent libraries.
-We fix a (possibly infinite) set $\mathbb{D}$ of data values, a finite set $\mathbb{M}$ of methods, and an infinite set $\mathbb{O}$ of operation (identifiers).
-Given $m \in \mathbb{M}$, $x \in \mathbb{D}$, and $o \in \mathbb{O}$, a call action $\textit{call}_o (m,x)$ represents an invocation to method $m$ with argument $x$, while a return action $\textit{ret}_o (m,x)$ represents a response from method $m$ with return value $x$.
-\footnote{Call actions with more than one arguments are similarly defined.} Here $o$ is used to match return actions to their call actions: $\textit{cal}_o (m,x)$ matches $\textit{ret}_{o'} (m',x')$, if $o=o' \wedge m=m'$.
-A sequential execution is a sequence of call and return actions, while each call action is immediately followed by its matching return action.
-Let $\cdot$ be the concatenation of sequences.
-To ease the reading, given a sequential execution $e=\textit{call}_{o_1}(m_1,a_1) \cdot \textit{rm}_{o_1}(m_1,b_1) \cdot \ldots \cdot \textit{call}_{o_n}(m_n,a_n) \cdot \textit{rm}_{o_n}(m_n,b_n)$, when the context is clear, we can write it as $e=m_1(a_1,b_1) \cdot \ldots \cdot m_n(a_n,b_n)$.
-We write $m(a,b)$ as $m(a)$ or $m(b)$, when there is no return value or there is no argument, respectively.
 
@@
 
-We define two properties which are %satisfied by priority queue implementations and which are
+We define two properties which are
 important for our results: (1) \emph{data independence}~\cite{conf/popl/Wolper86,conf/tacas/AbdullaHHJR13} states that priority queue behaviors do not depend on the actual values which are added to the queue, and (2) \emph{closure under projection}~\cite{DBLP:conf/icalp/BouajjaniEEH15} states that executions remain valid by removing all the operations adding or removing certain values.
-%remove operations can return the same values even if certain values are projected out from the priority queue state.
-%no matter how many other different values are in the queue, provided that they don't have more important priorities.
 
-%Data-independence \cite{Wolper:1986} can be used to effectively handle unbounded data domain.
-In this paper, we slightly modify the notion of data-independence in \cite{Wolper:1986} and propose data-independence for priority queues.
-Let $\_$ denote an element, of which the value is irrelevant.
 An execution $e$ is \emph{data-differentiated} if every value is added at most once, i.e., for each $d \in \mathbb{D}$, $e$ contains at most one action $\textit{call}_o(\textit{put},d,p)$ with $o\in\mathbb{O}$ and $p\in \mathbb{P}$.
@@
 
-%\vspace{-6pt}
 \begin{definition}\label{def:priority-value data-independence}
@@
 Let $\textit{proj}(e)$ be the set of all projections of $e$ to a set of values $D \subseteq \mathbb{D}$.
-%{\color {red}The following lemma states that $\seqPQ$ is closed under projection.}
 
@@
 
-%\begin{lemma}\label{lem:closure_proj}
-%$\seqPQ$ is closed under projection, i.e., $\textit{proj}(e)\subseteq \seqPQ$ for each $e\in \seqPQ$.
-%\end{lemma}
 \begin{lemma}
@@
 
-%A (concurrent) execution $e$ is a sequence of call and return actions which satisfy a well-formedness property: every return has a matching call action before it in $e$, and an operation $o$ can be used only twice in $e$, once in a call action, and once in a return action.
-The definition of data-differentiated, renaming function, data-independence and projection extends to (concurrent) executions.
-For instance, an execution is data-differentiated if, for all $d \in \mathbb{D}$, there is at most one $\textit{cal}_{\_}(\textit{put},d,\_)$.
-%
-%An implementation $\mathcal{I}$ is a set of concurrent executions.
-Implementations represent libraries whose methods are called by external programs.
-In the remainder of this work, we consider only completed executions, where each call action has a corresponding return action.
-This simplification is sound when implementation methods can always make progress in isolation \cite{Henzinger:2013}: formally, for any execution $e$ with pending operations, there exists an execution $e'$ obtained by extending $e$ only with the return actions of the pending operations of $e$.
-Intuitively this means that methods can always return without any help from outside threads, avoiding deadlock.
 We recall the notion of \emph{linearizability}~\cite{journals/toplas/HerlihyW90} which is the \emph{de facto} standard correctness condition for concurrent data structures.
@@
 Given a (concurrent) execution $e$ and a sequential execution $s$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if there is a bijection $f: O_1 \rightarrow O_2$, where $O_1$ and $O_2$ are the set of operations of $e$ and $s$, respectively, such that (1) the call and return actions with identifier $o$ and $f(o)$, respectively, are the same
-%{\color{blue}$o$ and $f(o)$ have the same call and return actions}{\color {red}$o$ and $f(o)$ is the same operation}\footnote{},
 and (2) if $o_1 <_{\textit{hb}} o_2$, then $f(o_1) <_{\textit{hb}} f(o_2)$.
@@
 The following lemma states that by data-independence, it is enough to consider only data-differentiated executions when checking linearizability (see Appendix \ref{sec:appendix proofs in section priority queue and data-independence}).
-%This is similar to that in \cite{conf/tacas/AbdullaHHJR13,DBLP:conf/icalp/BouajjaniEEH15}, where they use the notion of data-independence in \cite{conf/popl/Wolper86}.
 Section~\ref{sec:checking inclusion by recursive procedure} will focus on characterizing linearizability for data-differentiated executions.
-%The proof of this lemma can be found in Appendix \ref{sec:appendix in section data-independence of EPQ}.
 

### 2017-07-13 related.tex
arxiv

@@
 are
-% fixed to particular source-code locations.
 manually specified.
@@
 At a technical level, characterizing the priority queue violations requires a more expressive class of automata (with registers) than the finite-state automata in~\cite{DBLP:conf/icalp/BouajjaniEEH15}.
-
-%
-%these results require establishing key results which are specific to priority queues and which are not implied by those for stacks and queues (say that more details are given in the related work)
-%
-%\noindent The most related work of our paper is ~\cite{DBLP:conf/icalp/BouajjaniEEH15}.
-Compared to ~\cite{DBLP:conf/icalp/BouajjaniEEH15}, our work are specific to priority queue and thus have points not occur in queues and stacks.
-Due to the partial-order priorities, to prove the correctness of recursive procedure, we need to alter the positions of linearization points of incomparable items while keeps the linearization points of comparable items unchanged.
-{\color {blue}In the proof process we use the notion of left-right constraint, which is inspired by left-right constraint of queue \cite{Bouajjani:2015}.} For checking violations of one local property, we just need to find a value, such that at every possible time point to locate its remove operation, there is already values with smaller priorities in priority queue.
-For checking violations of another local property, we just need to find two values with maximal priority, such that every possible time point to locate remove operation of one value is disabled by remove operation of another value.
-Such phenomenons are unique for priority queue.
